### 2017-09-21 proposal/draft.tex
start taking notes

    \documentclass[12pt,
        usenames, % allows access to some tikz colors
        dvipsnames % more colors: https://en.wikibooks.org/wiki/LaTeX/Colors
    ]{article}
    \usepackage{
        amsmath,
        amssymb,
        fancyhdr, % page styling
        lastpage, % footer fanciness
        hyperref, % various links
        setspace, % line spacing
        amsthm, % newtheorem and proof environment
        mathtools, % \Aboxed for boxing inside aligns, among others
        float, % Allow [H] figure env alignment
        enumerate, % Allow custom enumerate numbering
        graphicx, % allow includegraphics with more filetypes
        wasysym, % \smiley!
        upgreek, % \upmu for \mum macro
        listings, % writing TrueType fonts and including code prettily
        tikz, % drawing things
        booktabs % \bottomrule instead of hline apparently
    }
    \usepackage[margin=1in]{geometry} % page geometry
    \usepackage[
        labelfont=bf, % caption names are labeled in bold
        font=scriptsize % smaller font for captions
    ]{caption}
    \usepackage[font=scriptsize]{subcaption} % subfigures

    \newcommand*{\scinot}[2]{#1\times10^{#2}}
    \newcommand*{\bra}[1]{\left<#1\right|}
    \newcommand*{\ket}[1]{\left|#1\right>}
    \newcommand*{\dotp}[2]{\left<#1\,\middle|\,#2\right>}
    \newcommand*{\rd}[2]{\frac{\mathrm{d}#1}{\mathrm{d}#2}}
    \newcommand*{\pd}[2]{\frac{\partial#1}{\partial#2}}
    \newcommand*{\rtd}[2]{\frac{\mathrm{d}^2#1}{\mathrm{d}#2^2}}
    \newcommand*{\ptd}[2]{\frac{\partial^2 #1}{\partial#2^2}}
    \newcommand*{\md}[2]{\frac{\mathrm{D}#1}{\mathrm{D}#2}}
    \newcommand*{\norm}[1]{\left|\left|#1\right|\right|}
    \newcommand*{\abs}[1]{\left|#1\right|}
    \newcommand*{\pvec}[1]{\vec{#1}^{\,\prime}}
    \newcommand*{\svec}[1]{\vec{#1}\;\!}
    \newcommand*{\bm}[1]{\boldsymbol{\mathbf{#1}}}
    \newcommand*{\expvalue}[1]{\left<#1\right>}
    \newcommand*{\ang}[0]{\text{\AA}}
    \newcommand*{\mum}[0]{\upmu \mathrm{m}}

    \newtheorem{theorem}{Theorem}[section]

    \let\Re\undefined
    \let\Im\undefined
    \DeclareMathOperator{\Res}{Res}
    \DeclareMathOperator{\Re}{Re}
    \DeclareMathOperator{\Im}{Im}
    \DeclareMathOperator{\Log}{Log}
    \DeclareMathOperator{\Arg}{Arg}
    \DeclareMathOperator{\Tr}{Tr}
    \DeclareMathOperator{\E}{E}
    \DeclareMathOperator{\Var}{Var}
    \DeclareMathOperator*{\argmin}{argmin}
    \DeclareMathOperator*{\argmax}{argmax}
    \DeclareMathOperator{\sgn}{sgn}
    \DeclareMathOperator{\diag}{diag}

    \everymath{\displaystyle} % biggify limits of inline sums and integrals
    \tikzstyle{circ} % usage: \node[circ, placement] (label) {text};
        = [draw, circle, fill=white, node distance=3cm, minimum height=2em]
    \definecolor{commentgreen}{rgb}{0,0.6,0}
    \lstset{
        basicstyle=\ttfamily\footnotesize,
        frame=single,
        numbers=left,
        showstringspaces=false,
        keywordstyle=\color{blue},
        stringstyle=\color{purple},
        commentstyle=\color{commentgreen},
        morecomment=[l][\color{magenta}]{\#}
    }

\begin{document}

\def\Snospace~{\S{}} % hack to remove the space left after autorefs
\renewcommand*{\sectionautorefname}{\Snospace}
\renewcommand*{\appendixautorefname}{\Snospace}
\renewcommand*{\figureautorefname}{Fig.}
\renewcommand*{\equationautorefname}{Eq.}
\renewcommand*{\tableautorefname}{Tab.}

\onehalfspacing

\pagestyle{fancy}
\rhead{Yubo Su}
\cfoot{\thepage/\pageref{LastPage}}

% helpful links:
% http://www.alexhunterlang.com/nsf-fellowship#TOC-Examples-of-Successful-Essays

\begin{itemize}
    \item \url{http://journals.ametsoc.org/doi/pdf/10.1175/1520-%
        0469\%281983\%29040\%3C2497\%3ATIOGWB\%3E2.0.CO\%3B2} Wave breaking is
        an essential part of circulation % holten et al 1983

    \item \url{https://arxiv.org/pdf/1211.0624.pdf}, % fuller 2012, vick 2016
        \url{https://arxiv.org/abs/1612.07316} Show necessary to get heating as
        function of depth

    \item \url{http://journals.ametsoc.org/doi/abs/10.1175/JAS-D-11%
        -097.1} has already studied to some extent in atmospheres, weakly
        nonlinear regime
        % sutherland 2011
\end{itemize}

\end{document}

### 2017-09-22 proposal/draft.tex
shvenv

@@
         % sutherland 2011
+
+    \item LIGO seems to have detected neutron star binaries
+        (\url{https://www.sciencenews.org/article/rumors-swirl-ligo-snagged-%
+        gravitational-waves-neutron-star-collision},
+        \url{http://news.nationalgeographic.com/2017/08/new-%
+        gravitational-waves-neutron-stars-ligo-space-science/}), white dwarf
+        waveforms crucial!
+
+    \item Do WD GW papers account for this?
 \end{itemize}

### 2017-11-27 writups/notes.tex
writeup

    \documentclass[11pt,
        usenames, % allows access to some tikz colors
        dvipsnames % more colors: https://en.wikibooks.org/wiki/LaTeX/Colors
    ]{report}
    \usepackage{
        amsmath,
        amssymb,
        fouriernc, % fourier font w/ new century book
        fancyhdr, % page styling
        lastpage, % footer fanciness
        hyperref, % various links
        setspace, % line spacing
        amsthm, % newtheorem and proof environment
        mathtools, % \Aboxed for boxing inside aligns, among others
        float, % Allow [H] figure env alignment
        enumerate, % Allow custom enumerate numbering
        graphicx, % allow includegraphics with more filetypes
        wasysym, % \smiley!
        upgreek, % \upmu for \mum macro
        listings, % writing TrueType fonts and including code prettily
        tikz, % drawing things
        booktabs, % \bottomrule instead of hline apparently
        cancel % can cancel things out!
    }
    \usepackage[margin=1in]{geometry} % page geometry
    \usepackage[
        labelfont=bf, % caption names are labeled in bold
        font=scriptsize % smaller font for captions
    ]{caption}
    \usepackage[font=scriptsize]{subcaption} % subfigures

    \newcommand*{\scinot}[2]{#1\times10^{#2}}
    \newcommand*{\bra}[1]{\left<#1\right|}
    \newcommand*{\ket}[1]{\left|#1\right>}
    \newcommand*{\dotp}[2]{\left<#1\,\middle|\,#2\right>}
    \newcommand*{\rd}[2]{\frac{\mathrm{d}#1}{\mathrm{d}#2}}
    \newcommand*{\pd}[2]{\frac{\partial#1}{\partial#2}}
    \newcommand*{\rtd}[2]{\frac{\mathrm{d}^2#1}{\mathrm{d}#2^2}}
    \newcommand*{\ptd}[2]{\frac{\partial^2 #1}{\partial#2^2}}
    \newcommand*{\md}[2]{\frac{\mathrm{D}#1}{\mathrm{D}#2}}
    \newcommand*{\norm}[1]{\left|\left|#1\right|\right|}
    \newcommand*{\abs}[1]{\left|#1\right|}
    \newcommand*{\pvec}[1]{\vec{#1}^{\,\prime}}
    \newcommand*{\svec}[1]{\vec{#1}\;\!}
    \newcommand*{\bm}[1]{\boldsymbol{\mathbf{#1}}}
    \newcommand*{\expvalue}[1]{\left<#1\right>}
    \newcommand*{\ang}[0]{\text{\AA}}
    \newcommand*{\mum}[0]{\upmu \mathrm{m}}
    \newcommand*{\at}[1]{\left.#1\right|}

    \newtheorem{theorem}{Theorem}[section]

    \let\Re\undefined
    \let\Im\undefined
    \DeclareMathOperator{\Res}{Res}
    \DeclareMathOperator{\Re}{Re}
    \DeclareMathOperator{\Im}{Im}
    \DeclareMathOperator{\Log}{Log}
    \DeclareMathOperator{\Arg}{Arg}
    \DeclareMathOperator{\Tr}{Tr}
    \DeclareMathOperator{\E}{E}
    \DeclareMathOperator{\Var}{Var}
    \DeclareMathOperator*{\argmin}{argmin}
    \DeclareMathOperator*{\argmax}{argmax}
    \DeclareMathOperator{\sgn}{sgn}
    \DeclareMathOperator{\diag}{diag\;}

    \DeclarePairedDelimiter\p{\lparen}{\rparen}
    \DeclarePairedDelimiter\s{\lbrack}{\rbrack}
    \DeclarePairedDelimiter\z{\lbrace}{\rbrace}

    % \everymath{\displaystyle} % biggify limits of inline sums and integrals
    \tikzstyle{circ} % usage: \node[circ, placement] (label) {text};
        = [draw, circle, fill=white, node distance=3cm, minimum height=2em]
    \definecolor{commentgreen}{rgb}{0,0.6,0}
    \lstset{
        basicstyle=\ttfamily\footnotesize,
        frame=single,
        numbers=left,
        showstringspaces=false,
        keywordstyle=\color{blue},
        stringstyle=\color{purple},
        commentstyle=\color{commentgreen},
        morecomment=[l][\color{magenta}]{\#}
    }

\begin{document}

\def\Snospace~{\S{}} % hack to remove the space left after autorefs
\renewcommand*{\sectionautorefname}{\Snospace}
\renewcommand*{\appendixautorefname}{\Snospace}
\renewcommand*{\figureautorefname}{Fig.}
\renewcommand*{\equationautorefname}{Eq.}
\renewcommand*{\tableautorefname}{Tab.}

\onehalfspacing

\pagestyle{fancy}
\rfoot{Yubo Su}
\cfoot{\thepage/\pageref{LastPage}}

\title{Research Notes}
\author{Yubo Su}

\maketitle

\tableofcontents

\newpage

\chapter{2D Wave Breaking in Atmospheres}

The goal of this will be to lay out a formalism that can reproduce Sutherland et
al.~2011\footnote{DOI:10.1175/JAS-D-11--097.1} and also investigate driven
oscillations versus the breaking of a single wave packet. This represents wave
breaking in the atmosphere.

\section{Dynamical Setup}

We adopt notation where $q_0$ is the background quantity and $q_1$ is the
perturbed quantity from the propagating wave.

\subsection{Nonturbulent Equations of Motion}

The equations of motion are the usual fluid equations
\begin{subequations}\label{se:fl_eq}
    \begin{align}
        \pd{\rho}{t} + \vec{\nabla} \cdot \rho \vec{u} &= 0,
            \label{eq:fl_eq.cont}\\
        \rd{\vec{u}}{t} &= -\frac{\vec{\nabla}P}{\rho} - g\hat{z},
            \label{eq:fl_eq.mom}
    \end{align}
\end{subequations}
where $\rd{}{t} = \pd{}{t} + \vec{u} \cdot \vec{\nabla}$ denotes the Lagrangian
derivative. We take $\vec{u}_0 = 0$ no background flow, $P_0(x, z) = P_0(z)
\propto e^{-z/H}$ where $H$ is the scale height of the atmosphere. We take the
fluid to be incompressible $\vec{\nabla} \cdot \vec{u} = 0$, which turns the
continuity equation \autoref{eq:fl_eq.cont} into $\rd{\rho}{t} = 0$. Finally, we
take the atmosphere to be isothermal, implying $\rho_0 \propto P_0 \propto
e^{-z/H}$, and more importantly $\frac{P_0}{\rho_0} = \frac{P_1}{\rho_1} =
\frac{k_BT}{\mu m_p}$. We denote $P_0 = P_0(z=0), \rho_0 = \rho_0(z=0)$.

We write out the new fluid equations under these assumptions; note that we do
not assume perturbed quantities are small and must keep all terms
\begin{subequations}\label{se:lin_eq}
    \begin{align}
        \pd{\rho_1}{t} &= \frac{u_{1z}}{H}\rho_0e^{-z/H} - \vec{u}_1 \cdot
            \vec{\nabla} \rho_1, \label{eq:lin_eq.cont}\\
        \pd{\vec{u}_1}{t} &=
            -\p*{\vec{u}_1 \cdot \vec{\nabla}}\vec{u}_1
            -\frac{-\frac{\hat{z}}{H}P_0e^{-z/H} + \frac{k_BT}{\mu m_p}
                \vec{\nabla} \rho_1}{\rho_0e^{-z/H} + \rho_1}
                - g\hat{z}, \label{eq:lin_eq.mom}.
    \end{align}
\end{subequations}
This consists of four equations in four unknowns, $\rho_1$ and the three
components of $\vec{u}_1$.

\subsection{Turbulent Dissipation}

In order to mimic the effect of turbulent dissipation, we introduce a viscosity
term into the equations of motion. Per the incompressible Navier-Stokes
equation, we quantify viscosity via $\nu \nabla^2 \vec{u}_1$ where $\nu$ is the
turbulent viscosity. For this problem, we simply use $\nu \neq
0.1\;\mathrm{cm^2/s}$ for air.

We identify the kinetic energy density carried by the perturbation to be
$\mathcal{U}_T = \frac{1}{2}\rho \p*{\vec{u}_1 \cdot \vec{u}_1}$. We require
$\rd{\mathcal{U}}{t} = 0$ under the dissipation-free fluid equations
\autoref{se:fl_eq} where we denote $\mathcal{U} \equiv \mathcal{U}_T +
\mathcal{U}_V$ the sum of the kinetic and potential energy densities; we assume
incompressibility as per our present problem:
\begin{align}
    \rd{\p*{\mathcal{U}_T + \mathcal{U}_V}}{t} = 0
        &= \frac{1}{2}u_1^2 \cancelto{0}{\rd{\rho}{t}}
        + \rho \vec{u}_1 \cdot \rd{\vec{u}_1}{t} + \rd{\mathcal{U}_V}{t}
        ,\nonumber\\
        &= \rho \p*{
                -\frac{\vec{u}_1 \cdot \vec{\nabla}P}{\rho} - u_{1z}g
            } + \rd{\mathcal{U}_V}{t},\nonumber\\
        &= -\cancelto{0}{\rd{P}{t}} + \cancelto{0}{\pd{P_0}{t}}
            + \pd{P_1}{t} - \rho u_{1z}g + \rd{\mathcal{U}_V}{t}.
\end{align}

%TODO figure this out from L&L, figure out the energy dissipation upon adding
%Navier-Stokes term.

\subsection{Boundary Conditions}

We must bound this to a finite domain. We choose periodic boundary conditions in
$x$ with total length $L_x$, $x \in \s*{-\frac{L}{2}, \frac{L}{2}}$. We choose
$z$ dimension to have total length $L_z$, $z \in [0, L_z]$.

To set up the boundary condition at $z = 0$, we recall that gravity waves in an
atmosphere with $e^{-z/H}$ profile have form
\begin{equation}
    u_{1z} \propto e^{\frac{z}{2H}}e^{i\p*{k_xx + k_zz - \omega t}},
        \label{eq:bc_below}
\end{equation}
where ($N^2 \equiv \frac{g}{H}$ is the Brunt-V\"ais\"al\"a frequency in an
incompressible, stratified atmosphere)
\begin{equation}
    \omega^2 = \frac{N^2 k_x^2}{k_x^2 + k_z^2 + \frac{1}{4H^2}}.
        \label{eq:omega}
\end{equation}
Thus, at constant $z = 0$ we must have
\begin{equation}
    \vec{u}_{1z}(z = 0) = e^{i(k_xx - \omega t)}\label{eq:bc_u1z}.
\end{equation}

The boundary condition at $z = L_z$ is much harder to determine. It is clear
that $\vec{u}_1(z = \infty) = \rho_1(z = \infty) = 0$, and for $L_z \gg H$ this
would be a reasonable approximation, if simply because we expect the majority of
the wave to dissipate via turbulent dissipation as $z$ reaches many $H$. We will
simply choose the BC to be many multiples of $H$. We can solve with both a
Dirichlet and Neumann BC and compare the two solutions; if the solutions
differ significantly then we must choose a larger $L_z$. These are the only two
solutions that can be implemented where we do not need the phase of the linear
wave, which we lose during the nonlinear breaking region, to relate the function
and derivative at the boundary.

For the boundary conditions on $\rho$, we note that in the linear regime it
should just have a phase offset from $u_{1z}$, thus we choose $\rho(z = 0)
\propto e^{\frac{z}{2H}}ie^{i\p*{k_xx + k_zz - \omega t}}$ and a similar
treatment at $z = L_z$, taking both a Dirichlet and Neumann BC\@. This should be
sufficient information to constrain our problem, as satisfying incompressibility
along the boundaries constrains $u_{1x}$.

\subsubsection{Old Stuff---$u_{1z}(z = L_z)$}

\textbf{Note:} The reason the below is difficult to make work is that it tries
to restore a linear treatment to the wave. However, since we must capture
nonlinear dynamics, we lose the concept of the ``phase'' of the wave in the wave
breaking region, and it proves difficult to write down a BC that is
phase-independent.

However, at any finite boundary there is nonzero reflection thanks to the
continually varying impedance. \emph{The correct BC to use when performing a
finite truncation must match propagation characteristics with an untruncated
medium.}

We assume the perturbations have once again become small at $L_z$, i.e.\
dissipation between $[0, L_z]$ is sufficient that only a small amplitude
perturbation persists to $z = L_z$. This is a necessary criterion for choosing
$L_z$ else we do not have a representative portrait of the energy dissipation
profile of the propagating wave with our simulation. Then we recall that in the
linear regime, the solution given by \autoref{eq:bc_below} is the propagating
solution. We assume that $k_x, \omega$ are the same as the values at the $z = 0$
boundary, then
\begin{equation}
    \frac{\rho_{,z}}{\rho} = \frac{u_{1z, z}}{u_{1z}} = \frac{1}{2H} + ik_z.
        \label{eq:bc_above}
\end{equation}
Recall that since our equations of motion are nonlinear, we cannot use complex
notation except in very limited, time-averaged cases. Thus, how to implement
this BC is tricky\dots

\subsection{Simulation}

We begin our simulation with $\rho_1 = 0, \vec{u}_1 = 0$ strictly within the
domain of simulation. We will borrow some values from Sutherland's paper and use
$k_z = 2\;\mathrm{km^{-1}}$ then define $k_x = -0.4k_z, H = 10 / k_z, A = 0.05 /
k_z, L_z = 300 / k_z, L_x = 20 / k_z$, We also use $\mu \approx 29, T =
273\;\mathrm{K}, \rho_0 = 1\;\mathrm{kg/m^3}, P_0 = \frac{\rho_0 k_BT}{\mu m_p},
g = 10\;\mathrm{m/s^2}$.

\end{document}

### 2017-12-02 proposal/ford/ford_proposal.tex
rename

    \documentclass[12pt,
        usenames, % allows access to some tikz colors
        dvipsnames % more colors: https://en.wikibooks.org/wiki/LaTeX/Colors
    ]{article}
    \usepackage{
        amsmath,
        amssymb,
        fancyhdr, % page styling
        lastpage, % footer fanciness
        hyperref, % various links
        setspace, % line spacing
        amsthm, % newtheorem and proof environment
        mathtools, % \Aboxed for boxing inside aligns, among others
        float, % Allow [H] figure env alignment
        enumerate, % Allow custom enumerate numbering
        graphicx, % allow includegraphics with more filetypes
        wasysym, % \smiley!
        upgreek, % \upmu for \mum macro
        listings, % writing TrueType fonts and including code prettily
        tikz, % drawing things
        times,
        booktabs % \bottomrule instead of hline apparently
    }
    \usepackage[margin=0.5in, top=1in, bottom=1in]{geometry} % page geometry
    \usepackage[
        labelfont=bf, % caption names are labeled in bold
        font=scriptsize % smaller font for captions
    ]{caption}
    \usepackage[font=scriptsize]{subcaption} % subfigures

    \newcommand*{\scinot}[2]{#1\times10^{#2}}
    \newcommand*{\bra}[1]{\left<#1\right|}
    \newcommand*{\ket}[1]{\left|#1\right>}
    \newcommand*{\dotp}[2]{\left<#1\,\middle|\,#2\right>}
    \newcommand*{\rd}[2]{\frac{\mathrm{d}#1}{\mathrm{d}#2}}
    \newcommand*{\pd}[2]{\frac{\partial#1}{\partial#2}}
    \newcommand*{\rtd}[2]{\frac{\mathrm{d}^2#1}{\mathrm{d}#2^2}}
    \newcommand*{\ptd}[2]{\frac{\partial^2 #1}{\partial#2^2}}
    \newcommand*{\md}[2]{\frac{\mathrm{D}#1}{\mathrm{D}#2}}
    \newcommand*{\norm}[1]{\left|\left|#1\right|\right|}
    \newcommand*{\abs}[1]{\left|#1\right|}
    \newcommand*{\pvec}[1]{\vec{#1}^{\,\prime}}
    \newcommand*{\svec}[1]{\vec{#1}\;\!}
    \newcommand*{\bm}[1]{\boldsymbol{\mathbf{#1}}}
    \newcommand*{\expvalue}[1]{\left<#1\right>}
    \newcommand*{\ang}[0]{\text{\AA}}
    \newcommand*{\mum}[0]{\upmu \mathrm{m}}

    \newtheorem{theorem}{Theorem}[section]

    \let\Re\undefined
    \let\Im\undefined
    \DeclareMathOperator{\Res}{Res}
    \DeclareMathOperator{\Re}{Re}
    \DeclareMathOperator{\Im}{Im}
    \DeclareMathOperator{\Log}{Log}
    \DeclareMathOperator{\Arg}{Arg}
    \DeclareMathOperator{\Tr}{Tr}
    \DeclareMathOperator{\E}{E}
    \DeclareMathOperator{\Var}{Var}
    \DeclareMathOperator*{\argmin}{argmin}
    \DeclareMathOperator*{\argmax}{argmax}
    \DeclareMathOperator{\sgn}{sgn}
    \DeclareMathOperator{\diag}{diag}

    \everymath{\displaystyle} % biggify limits of inline sums and integrals
    \tikzstyle{circ} % usage: \node[circ, placement] (label) {text};
        = [draw, circle, fill=white, node distance=3cm, minimum height=2em]
    \definecolor{commentgreen}{rgb}{0,0.6,0}
    \lstset{
        basicstyle=\ttfamily\footnotesize,
        frame=single,
        numbers=left,
        showstringspaces=false,
        keywordstyle=\color{blue},
        stringstyle=\color{purple},
        commentstyle=\color{commentgreen},
        morecomment=[l][\color{magenta}]{\#}
    }

\begin{document}

\def\Snospace~{\S{}} % hack to remove the space left after autorefs
\renewcommand*{\sectionautorefname}{\Snospace}
\renewcommand*{\appendixautorefname}{\Snospace}
\renewcommand*{\figureautorefname}{Fig.}
\renewcommand*{\equationautorefname}{Eq.}
\renewcommand*{\tableautorefname}{Tab.}

\doublespacing

% helpful links:
% http://www.alexhunterlang.com/nsf-fellowship#TOC-Examples-of-Successful-Essays

\section{Proposed Research}

In the last decade, astrophysical systems with one white dwarf orbiting another
body e.g.\ a black hole or another white dwarf (referred to collectively as WD
binaries) have become increasingly important for several topics in astrophysics.
As gravitational wave (GW) astronomy takes off with the recent triumphs of the
Laser Interferometer Gravitational-Wave Observatory (LIGO), understanding GW
profiles of WD systems is paramount. Mergers of WD-WD binaries create a variety
of astrophysical systems, the most important of which are type Ia supernovae,
the standard candle of observational astronomy. WD-BH (black hole) systems are
expected to produce detectable flares that can be observed to yield constraints
on currently unknown physics.
% (https://arxiv.org/pdf/1701.08162.pdf)

\emph{Tidal interactions}, the effect of an uneven gravitational field produced
by one body on the other, are the key component in such systems. In WD binaries,
tidal interactions are expected to excite internal gravity waves in the WD\@.
These then propagate towards the surface and dissipate by wave breaking, as ocean
waves do on a beach, among other mechanisms.
Prior work shows that this wave breaking can play a dominant role in determining
the internal structure of the WD\@. Nonetheless, previous treatments of WD
binaries adopt incomplete, semi-analytical treatments to estimate the location
and strength of wave-breaking induced dissipation. For such fluid systems,
numerical simulation is the only way to obtain quantitatively accurate results.
% (Fuller 2012d)

\textbf{We propose to study nonlinear wave breaking of internal gravity waves in
WDs using numerical hydrodynamic codes.} There are already a wide assortment of
suitable hydrodynamical codes
% (Dedalus\footnote{
% http://iopscience.iop.org/article/10.1088/0004-%
% 637X/796/1/17}, Athena\footnote{e.g.\
% http://iopscience.iop.org/article/10.1088/0004-%
% 637X/741/1/57})
capable of solving the fully nonlinear hydrodynamical equations, we expect to be
able to apply such codes with slight modifications to solve the proposed
problem. Should existing codes prove inapplicable, the applicant is well
equipped to develop suitable software from scratch (\autoref{s:plan}). We aim
to extend our findings to observational signatures for a variety of compositions
and uncertain material properties.

The results of our study as GW astronomy continues to improve; inaccuracies are
especially important to mitigate with the high sensitivity requirements
involved. Conversely, a gravitational wave detection could yield significant
information on WD dynamics that, thanks to their ubiquity, could have
ramifications throughout astronomy. Equally importantly, our approach would be
highly transferable to other problems such as tidal heating during exoplanet
formation, another area of crucial and open research.

A tentative timeline for research follows:
\begin{itemize}
    \item 2--4 months: Set up hydrodynamic code, calibrate against simplified,
        well-understood fiducial models.
    \item 4--10 months: Apply hydrodynamic code to systems found in literature.
    \item 10--14 months: Generate observational counterparts to predicted
        phenomena.
    \item 14--24 months: Explore related interesting astrophysical systems, seek
        analytical descriptions of any discovered phenomena.
\end{itemize}

\section{Plan of Study}\label{s:plan}

Cornell University is the ideal institution to pursue the proposed research.
Professor Dong Lai, the applicant's adviser, has authored numerous papers on WD
tidal dissipation and is a foremost expert in the subject. Other faculty and
researchers at Cornell are equally well-suited for the project. Notably,
Professor Saul Teukolsky is a leading researcher in numerical techniques in
astrophysics for the past five decades.

The applicant earned two B.S.\ from the California Institute of Technology in
Physics and Computer science and is well equipped for the proposed problem with
teaching proficiency in differential equations and completed coursework in
computational physics, astrophysical fluid dynamics, GPU programming and others.
Future relevant coursework includes astrophysical processes, celestial
mechanics and physics of compact objects, all of which are scheduled to be
completed within six months of recipt of the fellowship funding. These courses
will enable both a complete treatment of relevant physics and accurate, fast
programming of numerical codes.

The proposed study is integral to the long-term career goals of the applicant,
a professorship in astrophysics at a research institution. Proficiency in
numerical techniques and software development is extremely important in modern
astrophysical studies. White dwarfs both are important to observational
astronomy and involve many astrophysical processes, so such a project would be
relevant to many members of the academic community and open avenues for
collaboration.

% \begin{itemize}
%     \item Lots of WKB treatments
%     % TODO radiative diffusion vs turbulent dissipation?
%     %   - if in stars can power energic outbursts, same for WD?
%     % Planet formation, WD inspirals + waveforms (important b/c Fuller et al P4)
%     %   - heating should have visible signatures in flares
%     \item

%     \item \url{http://journals.ametsoc.org/doi/pdf/10.1175/1520-%
%         0469\%281983\%29040\%3C2497\%3ATIOGWB\%3E2.0.CO\%3B2} Wave breaking is
%         an essential part of circulation % holten et al 1983

%     \item \url{https://arxiv.org/pdf/1211.0624.pdf}, % fuller 2012, vick 2016
%         \url{https://arxiv.org/abs/1612.07316} Show necessary to get heating as
%         function of depth

%     \item \url{http://journals.ametsoc.org/doi/abs/10.1175/JAS-D-11%
%         -097.1} has already studied to some extent in atmospheres, weakly
%         nonlinear regime
%         % sutherland 2011

%     \item Tidal heating in planet formation too!
% \end{itemize}

\end{document}

### 2017-12-05 proposal/proposal.tex
Stash

--- 2017-12-02 proposal/ford/ford_proposal.tex
@@
 
-In the last decade, astrophysical systems with one white dwarf orbiting another
-body e.g.\ a black hole or another white dwarf (referred to collectively as WD
-binaries) have become increasingly important for several topics in astrophysics.
-As gravitational wave (GW) astronomy takes off with the recent triumphs of the
-Laser Interferometer Gravitational-Wave Observatory (LIGO), understanding GW
-profiles of WD systems is paramount. Mergers of WD-WD binaries create a variety
-of astrophysical systems, the most important of which are type Ia supernovae,
-the standard candle of observational astronomy. WD-BH (black hole) systems are
-expected to produce detectable flares that can be observed to yield constraints
-on currently unknown physics.
-% (https://arxiv.org/pdf/1701.08162.pdf)
+In the last decade, astrophysical systems with one white dwarf and another
+body (e.g.\ a black hole or another white dwarf), referred to collectively as WD
+binaries, have become increasingly important for several topics in astrophysics.
+Mergers of WD-WD binaries create a variety of astrophysical systems, the most
+important of which are type Ia supernovae, the standard candle of observational
+astronomy. WD-BH (black hole) systems are expected to produce detectable flares
+that can be observed to yield constraints on currently unknown physics
+(https://arxiv.org/pdf/1701.08162.pdf).
 

### 2017-12-11 proposal/proposal.tex
outline
try proposal

@@
-    \documentclass[12pt,
+    \documentclass[11pt,
         usenames, % allows access to some tikz colors
@@
     ]{article}
+    \usepackage[numbers]{natbib}
     \usepackage{
@@
         amssymb,
+        fouriernc,
         fancyhdr, % page styling
@@
     }
-    \usepackage[margin=0.5in, top=1in, bottom=1in]{geometry} % page geometry
+    \usepackage[margin=0.8in, top=1in, bottom=1in]{geometry} % page geometry
     \usepackage[
@@
 
-    \everymath{\displaystyle} % biggify limits of inline sums and integrals
     \tikzstyle{circ} % usage: \node[circ, placement] (label) {text};
@@
 
-\doublespacing
+\onehalfspacing
 
@@
 
+\title{Nonlinear Tidal Dissipation in White Dwarfs}
+\author{Yubo Su}
+\date{}
+
+\maketitle
+
+\section{Introduction: White Dwarf Binaries}
+
+White dwarfs (WDs) are luminous, long-lasting products of late-stage stellar
+evolution that are held up by electron degeneracy pressure, a quantum mechanical
+effect arising from subjecting electrons to immense pressures. They are some of
+the densest objects in the known universe, behind only neutron stars (NS) and
+black holes (BH), fitting a solar mass into an Earth-sized sphere. Furthermore,
+they exhibit various chemical compositions, each of which is subject to
+different physics, and are found in a variety of interesting systems. As such,
+WDs are a unique and diverse window into matter under extreme conditions.
+
+WDs are commonly found in binary orbits with companion objects, in which the two
+orbit their center of mass under their mutual gravitational attraction. The
+companion object ranges from another WD to a supermassive BH (SMBH),
+hypothesized to be at the center of galaxies and have mass $10^5$--$10^6$ solar
+masses. All of these binary systems are very important to astrophysics. WD-WD
+binaries are most important for being thought to generate \emph{Type Ia
+supernovae} in which one of the two WDs reaches a critical mass through
+accreting surrounding matter and collapses under its own gravity, producing a
+shock wave that tears the WD apart and releases immense amounts of energy. Type
+Ia supernovae are unique in astrophysics for their near-constancy throughout
+space and time, consistently occuring at the WD critical mass, and have been
+used to discover numerous cosmological phenomena such as dark energy.
+
+WD-BH systems are also interesting subjects of study. Studies indicate that
+as WDs orbit the extreme gravity surrounding BHs, they will produce observable
+flares induced by gravitational tidal forces\cite{flares}. A WD orbiting a SMBH
+would produce gravitational waves, waves carried by distortions in space-time
+and generated according to Einstein's general relativity by accelerating massive
+objects, that are expected to be detected by the space-based \emph{Laser
+Interferometer Space Antenna (LISA)} when deployed\cite{lisa}. Gravitational
+wave astronomy is an increasingly exciting field as the \emph{Laser
+Interferometer Gravitational-Wave Observatory (LIGO)} continues to make
+detections since its first in late 2015. As gravitational wave astronomy relies
+on accurate predictions of the expected signals, it is important to build as
+accurate models as possible before observation runs begin.
+
+\subsection{Tidal Dissipation}
+
+The excitation of \emph{internal gravity waves} in the WD by the tidal forces of
+the companion is an effect seen in all of the aforementioned systems. Internal
+gravity waves, not to be confused with the gravitational waves discussed above,
+are internal displacements in the WD that oscillate and propagate due to a
+restoring buoyancy force. As these waves propagate outward from where they are
+excited, they are expected to grow in amplitude until they break, as do ocean
+waves on a shore, and deposit both energy and angular momentum inside the WD\@.
+
+Previous work predicts that this dissipation mechanism can generate
+significantly more energy than thermal radiation from the WD surface alone and
+are thus a significant contribution to the WD energy budget\cite{fullerII}. The
+exact radial dissipation profile is of enormous interest since it is both
+sensitive to WD properties and can produce drastically different observable
+outcomes. One proposed outcome is a \emph{tidal nova}, in which heating in the
+WD's degenerate hydrogen layer is significant enough to trigger runaway nuclear
+fusion and a significant, observable explosion\cite{tidal_novae}. Understanding
+whether such phenomena occur requires understanding how energy is distributed
+internally inside a WD\@.
+
+Internal gravity wave breaking is a nonlinear hydrodynamic phenomenon. Such
+phenomena are known to require numerical simulation to study. It is therefore
+paramount to begin numerical study to build dissipation models inside WDs to
+characterise what phenomena can be observed in which WD models and
+characteristics.
+
 \section{Proposed Research}
 
-In the last decade, astrophysical systems with one white dwarf and another
-body (e.g.\ a black hole or another white dwarf), referred to collectively as WD
-binaries, have become increasingly important for several topics in astrophysics.
-Mergers of WD-WD binaries create a variety of astrophysical systems, the most
-important of which are type Ia supernovae, the standard candle of observational
-astronomy. WD-BH (black hole) systems are expected to produce detectable flares
-that can be observed to yield constraints on currently unknown physics
-(https://arxiv.org/pdf/1701.08162.pdf).
+\textbf{We propose to study tidal dissipation via nonlinear gravity wave
+breaking in white dwarfs via numerical simulation.}
 
-\emph{Tidal interactions}, the effect of an uneven gravitational field produced
-by one body on the other, are the key component in such systems. In WD binaries,
-tidal interactions are expected to excite internal gravity waves in the WD\@.
-These then propagate towards the surface and dissipate by wave breaking, as ocean
-waves do on a beach, among other mechanisms.
-Prior work shows that this wave breaking can play a dominant role in determining
-the internal structure of the WD\@. Nonetheless, previous treatments of WD
-binaries adopt incomplete, semi-analytical treatments to estimate the location
-and strength of wave-breaking induced dissipation. For such fluid systems,
-numerical simulation is the only way to obtain quantitatively accurate results.
-% (Fuller 2012d)
+Our research will consist of computing for various WD models and compositions
+the energy and angular momentum dissipation profiles inside WDs. Once such
+profiles are obtained, we intend to add these profiles to existing stellar
+evolution codes to study the dynamical effects of tidal dissipation. We will
+attempt to find a compact or even analytical representation of our numerical
+work to greatly simplify and accelerate such integration. Finally, we will make
+observational predictions with our results and compare our predictions with
+existing observed WDs. Any software and results would be made in accordance with
+best practice and publicly available for the entire community.
 
-\textbf{We propose to study nonlinear wave breaking of internal gravity waves in
-WDs using numerical hydrodynamic codes.} There are already a wide assortment of
-suitable hydrodynamical codes
-% (Dedalus\footnote{
-% http://iopscience.iop.org/article/10.1088/0004-%
-% 637X/796/1/17}, Athena\footnote{e.g.\
-% http://iopscience.iop.org/article/10.1088/0004-%
-% 637X/741/1/57})
-capable of solving the fully nonlinear hydrodynamical equations, we expect to be
-able to apply such codes with slight modifications to solve the proposed
-problem. Should existing codes prove inapplicable, the applicant is well
-equipped to develop suitable software from scratch (\autoref{s:plan}). We aim
-to extend our findings to observational signatures for a variety of compositions
-and uncertain material properties.
+\subsection{Qualifications}
 
-The results of our study as GW astronomy continues to improve; inaccuracies are
-especially important to mitigate with the high sensitivity requirements
-involved. Conversely, a gravitational wave detection could yield significant
-information on WD dynamics that, thanks to their ubiquity, could have
-ramifications throughout astronomy. Equally importantly, our approach would be
-highly transferable to other problems such as tidal heating during exoplanet
-formation, another area of crucial and open research.
+The proposing researcher (YS) and his adviser (DL) are uniquely qualified to
+pursue such studies. YS has ample research experience in numerical simulation
+and is continuing to pursue academic study in hydrodynamics, turbulence and
+computation. He also completed a double Bachelors degree in physics and computer
+science from the California Institute of Technology. Finally, he has worked in
+the software industry for a year and continues to follow discussions of best
+practice far ahead of those in academia and apply them to his own work.
 
-A tentative timeline for research follows:
-\begin{itemize}
-    \item 2--4 months: Set up hydrodynamic code, calibrate against simplified,
-        well-understood fiducial models.
-    \item 4--10 months: Apply hydrodynamic code to systems found in literature.
-    \item 10--14 months: Generate observational counterparts to predicted
-        phenomena.
-    \item 14--24 months: Explore related interesting astrophysical systems, seek
-        analytical descriptions of any discovered phenomena.
-\end{itemize}
+DL is a co-author on much of the latest literature on WD binaries and is an
+expert on many related subjects such as fluid dynamics in extreme matter and
+studies of other compact object systems. Cornell University also houses Saul
+Teukolsky's research group, one of the pre-eminent numerical relativity groups
+in the world, which will be frequently consulted for numerical best practices.
 
-\section{Plan of Study}\label{s:plan}
+In terms of software, YS proposes to only implement what is not currently
+available in the community. Current literature indicates that the spectral
+hydrodynamic code Dedalus is able to handle instability-driven turbulence with
+high accuracy and speed, and our preliminary work seems to indicate that Dedalus
+will be sufficiently accurate to simulate wave breaking
+dissipation\cite{dedalus}. Modules for Experiments in Stellar Astrophysics
+(MESA) is a proven stellar evolution code that will likely need only be extended
+to incorporate the discovered dissipation models\cite{MESA}. Nevertheless, both
+codes will require adaptation to be applied to the present problem (YS has
+already contributed an improvement to Dedalus), a task for which YS is
+well-equiped.
 
-Cornell University is the ideal institution to pursue the proposed research.
-Professor Dong Lai, the applicant's adviser, has authored numerous papers on WD
-tidal dissipation and is a foremost expert in the subject. Other faculty and
-researchers at Cornell are equally well-suited for the project. Notably,
-Professor Saul Teukolsky is a leading researcher in numerical techniques in
-astrophysics for the past five decades.
+Should no existing software prove adequate, we are more than able and willing to
+develop a new code and contribute it to the community. Notably, no
+GPU-accelerated spectral hydrodynamic code has yet been released, which would be
+of great value to researchers, and YS is well-versed in GPU programming.
+Nevertheless, in the interest of the proposed science objectives codes will be
+developed on an as-needed basis.
 
-The applicant earned two B.S.\ from the California Institute of Technology in
-Physics and Computer science and is well equipped for the proposed problem with
-teaching proficiency in differential equations and completed coursework in
-computational physics, astrophysical fluid dynamics, GPU programming and others.
-Future relevant coursework includes astrophysical processes, celestial
-mechanics and physics of compact objects, all of which are scheduled to be
-completed within six months of recipt of the fellowship funding. These courses
-will enable both a complete treatment of relevant physics and accurate, fast
-programming of numerical codes.
+\subsection{Applications, Interest to the DoD}
 
-The proposed study is integral to the long-term career goals of the applicant,
-a professorship in astrophysics at a research institution. Proficiency in
-numerical techniques and software development is extremely important in modern
-astrophysical studies. White dwarfs both are important to observational
-astronomy and involve many astrophysical processes, so such a project would be
-relevant to many members of the academic community and open avenues for
-collaboration.
+A robust pipeline coupling nonlinear wave breaking to system evolution
+would be of interest spanning all fields of physics. Barring BHs, all
+astrophysical objects exhibit internal waves that can break as they propagate
+towards the surface. Notable present interests in the astrophysical community
+include tidal disruption, planet formation and stars. Even in planetary and
+atmospheric sciences, wave breaking in atmospheres would share many common
+techniques with the proposed research.
 
-% \begin{itemize}
-%     \item Lots of WKB treatments
-%     % TODO radiative diffusion vs turbulent dissipation?
-%     %   - if in stars can power energic outbursts, same for WD?
-%     % Planet formation, WD inspirals + waveforms (important b/c Fuller et al P4)
-%     %   - heating should have visible signatures in flares
-%     \item
+The proposed research is extremely relevant to Department of Defense (DoD)
+fields of interest. WDs and its energetic phenomena we propose to study e.g.\
+novae of all sorts are natural laboratories of extreme physics that have
+direct consequences for understanding of plasmas and other technologies integral
+to propulsion in aeronautics. A theory of nonlinear wave breaking would have
+far-reaching consequences in turbulence study and atmospheric sciences. Finally,
+the observation of our predictions could shed light on atmospheric optics,
+whether as a use for calibration or understanding contamination effects.
 
-%     \item \url{http://journals.ametsoc.org/doi/pdf/10.1175/1520-%
-%         0469\%281983\%29040\%3C2497\%3ATIOGWB\%3E2.0.CO\%3B2} Wave breaking is
-%         an essential part of circulation % holten et al 1983
-
-%     \item \url{https://arxiv.org/pdf/1211.0624.pdf}, % fuller 2012, vick 2016
-%         \url{https://arxiv.org/abs/1612.07316} Show necessary to get heating as
-%         function of depth
-
-%     \item \url{http://journals.ametsoc.org/doi/abs/10.1175/JAS-D-11%
-%         -097.1} has already studied to some extent in atmospheres, weakly
-%         nonlinear regime
-%         % sutherland 2011
-
-%     \item Tidal heating in planet formation too!
-% \end{itemize}
+\bibliographystyle{plainnat}
+\renewcommand{\bibname}{References}
+{\scriptsize \bibliography{proposal}}
 

### 2017-12-28 proposal/NDSEG_personal.tex
target NDSEG

--- 2017-11-27 writups/notes.tex
@@
         dvipsnames % more colors: https://en.wikibooks.org/wiki/LaTeX/Colors
-    ]{report}
+    ]{article}
     \usepackage{
@@
 
-\onehalfspacing
+\singlespacing
 
-\pagestyle{fancy}
-\rfoot{Yubo Su}
-\cfoot{\thepage/\pageref{LastPage}}
+As an aspiring theoretical astrophysics professor, my goals consist of
+developing both a strong grasp of pertinent numerical and analytical techniques
+and a broad understanding of all science that can be observed in astrophysical
+systems. Having this extra understanding will allow me to contribute to many
+fields using the natural laboratories that astronomical objects supply, probing
+physics inacessible to traditional experiment. To this end, the NDSEG fellowship
+along with its associated benefits are a great fit for my goals, and I would
+look forward to contributing to the NDSEG fellow community and the scientific
+community at large.
 
-\title{Research Notes}
-\author{Yubo Su}
+My immediate priority in becoming a professor is to seek highly versatile,
+interdisciplinary research to best prepare myself for future work. Such research
+will build up a core skillset that can be applied to various fields. My proposed
+research, numerical studies of nonlinear wave breaking in white dwarfs, is an
+example of such research that combines high energy physics, fluid dynamics and
+numerics. I intend to continue studying energetic phenomena in astrophysics...
+% maybe one more goal
 
-\maketitle
+Ultimately, I seek to become a research professor developing simple, tractable
+mobels of behaviors in astrophysical systems inacessible to man-made
+laboratories. I believe such models are invaluable for turning astronomical
+observations into constraints useful for other fields of study. My proposed work
+accomplishes this by turning observed white dwarf luminosity and orbital
+parameters into constraints on the behavior of matter under extreme conditions.
+Furthermore, such models are
 
-\tableofcontents
-
-\newpage
-
-\chapter{2D Wave Breaking in Atmospheres}
-
-The goal of this will be to lay out a formalism that can reproduce Sutherland et
-al.~2011\footnote{DOI:10.1175/JAS-D-11--097.1} and also investigate driven
-oscillations versus the breaking of a single wave packet. This represents wave
-breaking in the atmosphere.
-
-\section{Dynamical Setup}
-
-We adopt notation where $q_0$ is the background quantity and $q_1$ is the
-perturbed quantity from the propagating wave.
-
-\subsection{Nonturbulent Equations of Motion}
-
-The equations of motion are the usual fluid equations
-\begin{subequations}\label{se:fl_eq}
-    \begin{align}
-        \pd{\rho}{t} + \vec{\nabla} \cdot \rho \vec{u} &= 0,
-            \label{eq:fl_eq.cont}\\
-        \rd{\vec{u}}{t} &= -\frac{\vec{\nabla}P}{\rho} - g\hat{z},
-            \label{eq:fl_eq.mom}
-    \end{align}
-\end{subequations}
-where $\rd{}{t} = \pd{}{t} + \vec{u} \cdot \vec{\nabla}$ denotes the Lagrangian
-derivative. We take $\vec{u}_0 = 0$ no background flow, $P_0(x, z) = P_0(z)
-\propto e^{-z/H}$ where $H$ is the scale height of the atmosphere. We take the
-fluid to be incompressible $\vec{\nabla} \cdot \vec{u} = 0$, which turns the
-continuity equation \autoref{eq:fl_eq.cont} into $\rd{\rho}{t} = 0$. Finally, we
-take the atmosphere to be isothermal, implying $\rho_0 \propto P_0 \propto
-e^{-z/H}$, and more importantly $\frac{P_0}{\rho_0} = \frac{P_1}{\rho_1} =
-\frac{k_BT}{\mu m_p}$. We denote $P_0 = P_0(z=0), \rho_0 = \rho_0(z=0)$.
-
-We write out the new fluid equations under these assumptions; note that we do
-not assume perturbed quantities are small and must keep all terms
-\begin{subequations}\label{se:lin_eq}
-    \begin{align}
-        \pd{\rho_1}{t} &= \frac{u_{1z}}{H}\rho_0e^{-z/H} - \vec{u}_1 \cdot
-            \vec{\nabla} \rho_1, \label{eq:lin_eq.cont}\\
-        \pd{\vec{u}_1}{t} &=
-            -\p*{\vec{u}_1 \cdot \vec{\nabla}}\vec{u}_1
-            -\frac{-\frac{\hat{z}}{H}P_0e^{-z/H} + \frac{k_BT}{\mu m_p}
-                \vec{\nabla} \rho_1}{\rho_0e^{-z/H} + \rho_1}
-                - g\hat{z}, \label{eq:lin_eq.mom}.
-    \end{align}
-\end{subequations}
-This consists of four equations in four unknowns, $\rho_1$ and the three
-components of $\vec{u}_1$.
-
-\subsection{Turbulent Dissipation}
-
-In order to mimic the effect of turbulent dissipation, we introduce a viscosity
-term into the equations of motion. Per the incompressible Navier-Stokes
-equation, we quantify viscosity via $\nu \nabla^2 \vec{u}_1$ where $\nu$ is the
-turbulent viscosity. For this problem, we simply use $\nu \neq
-0.1\;\mathrm{cm^2/s}$ for air.
-
-We identify the kinetic energy density carried by the perturbation to be
-$\mathcal{U}_T = \frac{1}{2}\rho \p*{\vec{u}_1 \cdot \vec{u}_1}$. We require
-$\rd{\mathcal{U}}{t} = 0$ under the dissipation-free fluid equations
-\autoref{se:fl_eq} where we denote $\mathcal{U} \equiv \mathcal{U}_T +
-\mathcal{U}_V$ the sum of the kinetic and potential energy densities; we assume
-incompressibility as per our present problem:
-\begin{align}
-    \rd{\p*{\mathcal{U}_T + \mathcal{U}_V}}{t} = 0
-        &= \frac{1}{2}u_1^2 \cancelto{0}{\rd{\rho}{t}}
-        + \rho \vec{u}_1 \cdot \rd{\vec{u}_1}{t} + \rd{\mathcal{U}_V}{t}
-        ,\nonumber\\
-        &= \rho \p*{
-                -\frac{\vec{u}_1 \cdot \vec{\nabla}P}{\rho} - u_{1z}g
-            } + \rd{\mathcal{U}_V}{t},\nonumber\\
-        &= -\cancelto{0}{\rd{P}{t}} + \cancelto{0}{\pd{P_0}{t}}
-            + \pd{P_1}{t} - \rho u_{1z}g + \rd{\mathcal{U}_V}{t}.
-\end{align}
-
-%TODO figure this out from L&L, figure out the energy dissipation upon adding
-%Navier-Stokes term.
-
-\subsection{Boundary Conditions}
-
-We must bound this to a finite domain. We choose periodic boundary conditions in
-$x$ with total length $L_x$, $x \in \s*{-\frac{L}{2}, \frac{L}{2}}$. We choose
-$z$ dimension to have total length $L_z$, $z \in [0, L_z]$.
-
-To set up the boundary condition at $z = 0$, we recall that gravity waves in an
-atmosphere with $e^{-z/H}$ profile have form
-\begin{equation}
-    u_{1z} \propto e^{\frac{z}{2H}}e^{i\p*{k_xx + k_zz - \omega t}},
-        \label{eq:bc_below}
-\end{equation}
-where ($N^2 \equiv \frac{g}{H}$ is the Brunt-V\"ais\"al\"a frequency in an
-incompressible, stratified atmosphere)
-\begin{equation}
-    \omega^2 = \frac{N^2 k_x^2}{k_x^2 + k_z^2 + \frac{1}{4H^2}}.
-        \label{eq:omega}
-\end{equation}
-Thus, at constant $z = 0$ we must have
-\begin{equation}
-    \vec{u}_{1z}(z = 0) = e^{i(k_xx - \omega t)}\label{eq:bc_u1z}.
-\end{equation}
-
-The boundary condition at $z = L_z$ is much harder to determine. It is clear
-that $\vec{u}_1(z = \infty) = \rho_1(z = \infty) = 0$, and for $L_z \gg H$ this
-would be a reasonable approximation, if simply because we expect the majority of
-the wave to dissipate via turbulent dissipation as $z$ reaches many $H$. We will
-simply choose the BC to be many multiples of $H$. We can solve with both a
-Dirichlet and Neumann BC and compare the two solutions; if the solutions
-differ significantly then we must choose a larger $L_z$. These are the only two
-solutions that can be implemented where we do not need the phase of the linear
-wave, which we lose during the nonlinear breaking region, to relate the function
-and derivative at the boundary.
-
-For the boundary conditions on $\rho$, we note that in the linear regime it
-should just have a phase offset from $u_{1z}$, thus we choose $\rho(z = 0)
-\propto e^{\frac{z}{2H}}ie^{i\p*{k_xx + k_zz - \omega t}}$ and a similar
-treatment at $z = L_z$, taking both a Dirichlet and Neumann BC\@. This should be
-sufficient information to constrain our problem, as satisfying incompressibility
-along the boundaries constrains $u_{1x}$.
-
-\subsubsection{Old Stuff---$u_{1z}(z = L_z)$}
-
-\textbf{Note:} The reason the below is difficult to make work is that it tries
-to restore a linear treatment to the wave. However, since we must capture
-nonlinear dynamics, we lose the concept of the ``phase'' of the wave in the wave
-breaking region, and it proves difficult to write down a BC that is
-phase-independent.
-
-However, at any finite boundary there is nonzero reflection thanks to the
-continually varying impedance. \emph{The correct BC to use when performing a
-finite truncation must match propagation characteristics with an untruncated
-medium.}
-
-We assume the perturbations have once again become small at $L_z$, i.e.\
-dissipation between $[0, L_z]$ is sufficient that only a small amplitude
-perturbation persists to $z = L_z$. This is a necessary criterion for choosing
-$L_z$ else we do not have a representative portrait of the energy dissipation
-profile of the propagating wave with our simulation. Then we recall that in the
-linear regime, the solution given by \autoref{eq:bc_below} is the propagating
-solution. We assume that $k_x, \omega$ are the same as the values at the $z = 0$
-boundary, then
-\begin{equation}
-    \frac{\rho_{,z}}{\rho} = \frac{u_{1z, z}}{u_{1z}} = \frac{1}{2H} + ik_z.
-        \label{eq:bc_above}
-\end{equation}
-Recall that since our equations of motion are nonlinear, we cannot use complex
-notation except in very limited, time-averaged cases. Thus, how to implement
-this BC is tricky\dots
-
-\subsection{Simulation}
-
-We begin our simulation with $\rho_1 = 0, \vec{u}_1 = 0$ strictly within the
-domain of simulation. We will borrow some values from Sutherland's paper and use
-$k_z = 2\;\mathrm{km^{-1}}$ then define $k_x = -0.4k_z, H = 10 / k_z, A = 0.05 /
-k_z, L_z = 300 / k_z, L_x = 20 / k_z$, We also use $\mu \approx 29, T =
-273\;\mathrm{K}, \rho_0 = 1\;\mathrm{kg/m^3}, P_0 = \frac{\rho_0 k_BT}{\mu m_p},
-g = 10\;\mathrm{m/s^2}$.
+% short term goals: perform highly interdisciplinary research that will enable
+% me to contribute to a variety of fields and bring novel tools across
+% specialties.
+%   - build core "swiss army knife" to apply to variety of problems
+%   - practice learning new ideas, i.e. breadth ~= depth, cross-pollination
+% long term goals: astrophysics research professor studying them as natural
+% laboratories for simple, tractable models of behavior inaccessible to
+% experiment
+%   - strong grasp of wide breadth of literature to understand what mechasims
+%   can be at play
+%   - passion for teaching, information dissemination
+% goals developed from desire to study complicated problems with easily grasped
+% explanations
+%   - implications beyond target problem
+%   - demonstration of principle of prescient analysis, outreach inspiration
+% I have already begun to lay foundation for this
+%   - research with multidisciplinary Dong Lai, strong emphasis on simplified
+%   problems
+%   - bringing new tools to table; double major, software best practice
+% fellowship fits into goals of understanding a wide breadth of material at a
+% strong enough level for application to unrelated problems
 

### 2018-01-20 proposal/NDSEG_personal.tex
proposals + check in review

@@
 
-\singlespacing
+\onehalfspacing
+
+\textbf{Yubo Su Personal Statement}\\[\parskip]
 
@@
 and a broad understanding of all science that can be observed in astrophysical
-systems. Having this extra understanding will allow me to contribute to many
+systems. Having this extra understanding will allow me to contribute to various
 fields using the natural laboratories that astronomical objects supply, probing
@@
 along with its associated benefits are a great fit for my goals, and I would
-look forward to contributing to the NDSEG fellow community and the scientific
-community at large.
+look forward to contributing to the NDSEG community and the scientific community
+at large.
 
@@
 example of such research that combines high energy physics, fluid dynamics and
-numerics. I intend to continue studying energetic phenomena in astrophysics...
-% maybe one more goal
+numerics. A second core component of building a transferrable skillset is
+collaborating with scientists with diverse areas of expertise. The NDSEG
+practicum will be a great opportunity both to meet research leaders in relevant
+but different fields and to work alongside fellow growing scientists.
 
 Ultimately, I seek to become a research professor developing simple, tractable
-mobels of behaviors in astrophysical systems inacessible to man-made
-laboratories. I believe such models are invaluable for turning astronomical
-observations into constraints useful for other fields of study. My proposed work
-accomplishes this by turning observed white dwarf luminosity and orbital
-parameters into constraints on the behavior of matter under extreme conditions.
-Furthermore, such models are
+mobels of behaviors in astrophysical systems as probes of conditions
+beyond man-made laboratories. I believe such models are invaluable for turning
+astronomical observations into constraints useful for other fields of study. My
+proposed work accomplishes this by turning observed white dwarf luminosity and
+orbital parameters into constraints on the behavior of matter under extreme
+conditions. My preference for simpler models is rooted in my excitement to
+teach; intuitive models are more pedagogically suitable. Many of my best-taught
+classes were transformative experiences for me, and I hope as part of my job to
+instill the same clarity as did my professors.
+
+My academic career to date has revolved around my goal of becoming an
+interdisciplinary research and teaching faculty member developing simplified
+models of intractable behavior. At Caltech, I pursued a double major and a
+broad courseload that prepared me to apply many skills flexibly in my research.
+I then worked for a year at a software startup, where I brought in many of my
+problem solving skills from my physics background, and am now bringing industry
+best practice to my research efforts at Cornell. I will therefore be well
+prepared to bilaterally share knowledge between a DoD practicum and Cornell
+research groups. Now, I have begun research with the versatile Professor Dong
+Lai whose research centers on the same principle of distilling intractable
+problems into intuitive, simplified models. My present track lines up well with
+my desired goals, and I am primed to both contribute to and learn from an NDSEG
+fellowship.
 

### 2018-01-30 writups/notes.tex
01/28

@@
 
-\subsection{Nonturbulent Equations of Motion}
-
-The equations of motion are the usual fluid equations
+The fluid equations are
 \begin{subequations}\label{se:fl_eq}
@@
             \label{eq:fl_eq.cont}\\
-        \rd{\vec{u}}{t} &= -\frac{\vec{\nabla}P}{\rho} - g\hat{z},
+        \rd{\vec{u}}{t} &= -\vec{\nabla}\frac{P}{\rho} - g\hat{z},
             \label{eq:fl_eq.mom}
@@
 \end{subequations}
-where $\rd{}{t} = \pd{}{t} + \vec{u} \cdot \vec{\nabla}$ denotes the Lagrangian
-derivative. We take $\vec{u}_0 = 0$ no background flow, $P_0(x, z) = P_0(z)
-\propto e^{-z/H}$ where $H$ is the scale height of the atmosphere. We take the
-fluid to be incompressible $\vec{\nabla} \cdot \vec{u} = 0$, which turns the
-continuity equation \autoref{eq:fl_eq.cont} into $\rd{\rho}{t} = 0$. Finally, we
-take the atmosphere to be isothermal, implying $\rho_0 \propto P_0 \propto
-e^{-z/H}$, and more importantly $\frac{P_0}{\rho_0} = \frac{P_1}{\rho_1} =
-\frac{k_BT}{\mu m_p}$. We denote $P_0 = P_0(z=0), \rho_0 = \rho_0(z=0)$.
+where we will take gravity to be uniform throughout the domain of interest. We
+will study for no background flow $\vec{u}_0 = 0$ and in vertically stratified
+atmosphere $\rho_0 \propto e^{-z/H}$. In the absence of any perturbations it is
+then easy to show that $P_0 = \rho_0 g H$.
 
-We write out the new fluid equations under these assumptions; note that we do
-not assume perturbed quantities are small and must keep all terms
-\begin{subequations}\label{se:lin_eq}
+\subsection{Linear Regime, Incompressible}
+
+First, we solve the incompressible case $c_s^2 \to \infty, \vec{\nabla} \cdot
+\vec{u} = 0$ in the linear regime. The fluid equations to first order reduce to
+\begin{equation}
+    \begin{split}
+        \pd{\rho_1}{t} + \p*{\vec{u}_1 \cdot \vec{\nabla}}\rho_0 &= 0,\\
+        \vec{\nabla} \cdot \vec{u}_1 &= 0,\\
+        \pd{\vec{u}_1}{t} &= -\frac{\vec{\nabla}P_1}{\rho_0}
+            - P_1 \vec{\nabla}\frac{1}{\rho_0}
+            + \rho_1 gH \vec{\nabla}\frac{1}{\rho_0}
+            + \frac{gH}{\rho_0} \vec{\nabla}\rho_1.
+    \end{split}\label{eq:lin_incomp}
+\end{equation}
+Substituting in plane waves of form $e^{i\p*{\vec{k} \cdot \vec{r} - \omega t}}$
+and the given vertically stratified $e^{-z/H}$ density and pressure profiles
+gives system of equations
 \begin{align}
-        \pd{\rho_1}{t} &= \frac{u_{1z}}{H}\rho_0e^{-z/H} - \vec{u}_1 \cdot
-            \vec{\nabla} \rho_1, \label{eq:lin_eq.cont}\\
-        \pd{\vec{u}_1}{t} &=
-            -\p*{\vec{u}_1 \cdot \vec{\nabla}}\vec{u}_1
-            -\frac{-\frac{\hat{z}}{H}P_0e^{-z/H} + \frac{k_BT}{\mu m_p}
-                \vec{\nabla} \rho_1}{\rho_0e^{-z/H} + \rho_1}
-                - g\hat{z}, \label{eq:lin_eq.mom}.
+    \begin{bmatrix}
+        -i\omega & 0 & -\frac{\rho_0}{H} & 0\\
+        0 & ik_x & ik_z & 0\\
+        0 & -i\omega & 0 & \frac{ik_x}{\rho_0}\\
+        \frac{P_0}{\rho_0^2H} & 0 & -i\omega & \frac{ik_z}{\rho_0}
+    \end{bmatrix}\begin{bmatrix}
+        \rho_0 \\ u_{1x} \\ u_{1z} \\ P_1
+    \end{bmatrix} &= 0,\nonumber\\
+    \omega^2 k^2 + \frac{P_0}{\rho_0^2H}k_x^2 = \omega^2k^2 - N^2k_x^2 &= 0.
 \end{align}
-\end{subequations}
-This consists of four equations in four unknowns, $\rho_1$ and the three
-components of $\vec{u}_1$.
+We identify $\frac{P_0}{\rho_0^2H} = \frac{g}{H} = N^2$ the Brunt-V\"ais\"al\"a
+frequency. This agrees with the usual incompressible dispersion relation
+$\omega^2 = N^2\cos^2\theta$ where $\theta$ is the angle formed between
+$\vec{k}$ and $\hat{x}$.
 
-\subsection{Turbulent Dissipation}
-
-In order to mimic the effect of turbulent dissipation, we introduce a viscosity
-term into the equations of motion. Per the incompressible Navier-Stokes
-equation, we quantify viscosity via $\nu \nabla^2 \vec{u}_1$ where $\nu$ is the
-turbulent viscosity. For this problem, we simply use $\nu \neq
-0.1\;\mathrm{cm^2/s}$ for air.
-
-We identify the kinetic energy density carried by the perturbation to be
-$\mathcal{U}_T = \frac{1}{2}\rho \p*{\vec{u}_1 \cdot \vec{u}_1}$. We require
-$\rd{\mathcal{U}}{t} = 0$ under the dissipation-free fluid equations
-\autoref{se:fl_eq} where we denote $\mathcal{U} \equiv \mathcal{U}_T +
-\mathcal{U}_V$ the sum of the kinetic and potential energy densities; we assume
-incompressibility as per our present problem:
-\begin{align}
-    \rd{\p*{\mathcal{U}_T + \mathcal{U}_V}}{t} = 0
-        &= \frac{1}{2}u_1^2 \cancelto{0}{\rd{\rho}{t}}
-        + \rho \vec{u}_1 \cdot \rd{\vec{u}_1}{t} + \rd{\mathcal{U}_V}{t}
-        ,\nonumber\\
-        &= \rho \p*{
-                -\frac{\vec{u}_1 \cdot \vec{\nabla}P}{\rho} - u_{1z}g
-            } + \rd{\mathcal{U}_V}{t},\nonumber\\
-        &= -\cancelto{0}{\rd{P}{t}} + \cancelto{0}{\pd{P_0}{t}}
-            + \pd{P_1}{t} - \rho u_{1z}g + \rd{\mathcal{U}_V}{t}.
-\end{align}
-
-%TODO figure this out from L&L, figure out the energy dissipation upon adding
-%Navier-Stokes term.
+\subsection{Linear Regime, Compressible}
 
@@
 \begin{equation}
-    \vec{u}_{1z}(z = 0) = e^{i(k_xx - \omega t)}\label{eq:bc_u1z}.
+    \vec{u}_{1z}(z = 0) \propto e^{i(k_xx - \omega t)}\label{eq:bc_u1z}.
 \end{equation}
@@
 \propto e^{\frac{z}{2H}}ie^{i\p*{k_xx + k_zz - \omega t}}$ and a similar
-treatment at $z = L_z$, taking both a Dirichlet and Neumann BC\@. This should be
-sufficient information to constrain our problem, as satisfying incompressibility
-along the boundaries constrains $u_{1x}$.
-
-\subsubsection{Old Stuff---$u_{1z}(z = L_z)$}
-
-\textbf{Note:} The reason the below is difficult to make work is that it tries
-to restore a linear treatment to the wave. However, since we must capture
-nonlinear dynamics, we lose the concept of the ``phase'' of the wave in the wave
-breaking region, and it proves difficult to write down a BC that is
-phase-independent.
-
-However, at any finite boundary there is nonzero reflection thanks to the
-continually varying impedance. \emph{The correct BC to use when performing a
-finite truncation must match propagation characteristics with an untruncated
-medium.}
-
-We assume the perturbations have once again become small at $L_z$, i.e.\
-dissipation between $[0, L_z]$ is sufficient that only a small amplitude
-perturbation persists to $z = L_z$. This is a necessary criterion for choosing
-$L_z$ else we do not have a representative portrait of the energy dissipation
-profile of the propagating wave with our simulation. Then we recall that in the
-linear regime, the solution given by \autoref{eq:bc_below} is the propagating
-solution. We assume that $k_x, \omega$ are the same as the values at the $z = 0$
-boundary, then
-\begin{equation}
-    \frac{\rho_{,z}}{\rho} = \frac{u_{1z, z}}{u_{1z}} = \frac{1}{2H} + ik_z.
-        \label{eq:bc_above}
-\end{equation}
-Recall that since our equations of motion are nonlinear, we cannot use complex
-notation except in very limited, time-averaged cases. Thus, how to implement
-this BC is tricky\dots
+treatment at $z = L_z$, taking both a Dirichlet and Neumann BC\@.
 

### 2018-01-30 writups/notes.tex
ugh
ugh wrote up a hw problem

@@
 where we will take gravity to be uniform throughout the domain of interest. We
-will study for no background flow $\vec{u}_0 = 0$ and in vertically stratified
-atmosphere $\rho_0 \propto e^{-z/H}$. In the absence of any perturbations it is
-then easy to show that $P_0 = \rho_0 g H$.
+will study for no background flow $\vec{u}_0 = 0$ and in the presence of
+stratification $\rho_0 \propto e^{-z/H}$. In the absence of any perturbations it
+is then easy to show that $\rd{P_0}{z} = -\rho_0 g$.
 
-\subsection{Linear Regime, Incompressible}
+\subsection{Linear, Incompressible}
+
+\subsubsection{Solution for Arbitrary Stratification}
 
 First, we solve the incompressible case $c_s^2 \to \infty, \vec{\nabla} \cdot
-\vec{u} = 0$ in the linear regime. The fluid equations to first order reduce to
+\vec{u} = 0$ in the linear regime. For funsies, we solve for arbitrary
+stratification first. The fluid equations to first order reduce to
 \begin{equation}
@@
             - P_1 \vec{\nabla}\frac{1}{\rho_0}
-            + \rho_1 gH \vec{\nabla}\frac{1}{\rho_0}
-            + \frac{gH}{\rho_0} \vec{\nabla}\rho_1.
     \end{split}\label{eq:lin_incomp}
 \end{equation}
-Substituting in plane waves of form $e^{i\p*{\vec{k} \cdot \vec{r} - \omega t}}$
-and the given vertically stratified $e^{-z/H}$ density and pressure profiles
-gives system of equations
+We expect there to be some $z$ dependence in the amplitude, so we substitute
+variables of form $e^{i(kx - \omega t)}$ and do not specify the $z$ dependence.
+This gives us
+\begin{equation}
+    \begin{split}
+        -i\omega \rho_1 - u_{1z}\rd{\rho_0}{z} &= 0,\\
+        iku_{1x} + \rd{u_{1z}}{z} &= 0,\\
+        -iw u_{1x} + \frac{ik_x P_1}{\rho_0} &= 0,\\
+        -iw u_{1z} + \frac{1}{\rho_0}\pd{P_1}{z} +
+            \frac{\rho_1}{\rho_0^2}\rd{P_0}{z} &= 0.
+    \end{split}
+\end{equation}
+We substitute $N^2 = -\frac{g}{\rho_0}\rd{\rho_0}{z}$ and $\rd{P_0}{z} = -\rho
+g$ to obtain
+\begin{subequations}
     \begin{align}
-    \begin{bmatrix}
-        -i\omega & 0 & -\frac{\rho_0}{H} & 0\\
-        0 & ik_x & ik_z & 0\\
-        0 & -i\omega & 0 & \frac{ik_x}{\rho_0}\\
-        \frac{P_0}{\rho_0^2H} & 0 & -i\omega & \frac{ik_z}{\rho_0}
-    \end{bmatrix}\begin{bmatrix}
-        \rho_0 \\ u_{1x} \\ u_{1z} \\ P_1
-    \end{bmatrix} &= 0,\nonumber\\
-    \omega^2 k^2 + \frac{P_0}{\rho_0^2H}k_x^2 = \omega^2k^2 - N^2k_x^2 &= 0.
+        -i\omega \rho_1 - u_{1z}\frac{\rho_0N^2}{g} &= 0,\label{eq:lin.cont}\\
+        iku_{1x} + \rd{u_{1z}}{z} &= 0,\label{eq:lin.div}\\
+        -iw u_{1x} + \frac{ik_x P_1}{\rho_0} &= 0,\label{eq:lin.mom_x}\\
+        -iw u_{1z} + \frac{1}{\rho_0}\pd{P_1}{z} +
+            \frac{\rho_1g}{\rho_0} &= 0.\label{eq:lin.mom_z}
     \end{align}
-We identify $\frac{P_0}{\rho_0^2H} = \frac{g}{H} = N^2$ the Brunt-V\"ais\"al\"a
-frequency. This agrees with the usual incompressible dispersion relation
-$\omega^2 = N^2\cos^2\theta$ where $\theta$ is the angle formed between
-$\vec{k}$ and $\hat{x}$.
+\end{subequations}
+Eliminating $u_{1x}$ by substituting~\eqref{eq:lin.div}
+into~\eqref{eq:lin.mom_x} and $\rho_1$ by substituting~\eqref{eq:lin.cont}
+into~\eqref{eq:lin.mom_z} give
+\begin{subequations}
+    \begin{align}
+        i\omega \rd{u_{1z}}{z} + \frac{k_x^2 P_1}{\rho_0} &= 0,
+            \label{eq:lin.eq1}\\
+        \p*{\omega^2 - N^2 }u_{1z} + \frac{i\omega}{\rho_0}\rd{P_1}{z} &= 0.
+            \label{eq:lin.eq2}
+    \end{align}
+\end{subequations}
+Finally, we multiply~\eqref{eq:lin.eq1} with $\rho_0$ and differentiate
+$\mathrm{d}z$ and combine with~\eqref{eq:lin.eq2} to give
+\begin{equation}
+    \rtd{u_{1z}}{z} + \frac{1}{\rho_0}\rd{\rho_0}{z}\rd{u_{1z}}{z}
+        + k_x^2\p*{\frac{N^2}{\omega^2} - 1}u_{1z} = 0.\label{eq:lin.incomp.gen}
+\end{equation}
+
+\subsubsection{Introduce Stratification}
+
+With stratification $\rho \propto e^{-z/H}$ \autoref{eq:lin.incomp.gen} clearly
+has exponential solutions $e^{\kappa z}$ for
+\begin{equation}
+    \kappa^2 - \frac{\kappa}{H} + k_x^2\p*{\frac{N^2}{\omega^2} - 1} = 0.
+\end{equation}
+
+We permit complex $\kappa = -\frac{1}{2H} + ik_z$, and from the above clearly
+\begin{align}
+    k_z^2 &= -\frac{1}{4H^2} + k_x^2\p*{\frac{N^2}{\omega^2} - 1},\nonumber\\
+    \omega^2 &= \frac{N^2k_x^2}{k_x^2 + k_z^2 + \frac{1}{4H^2}}.
+\end{align}
 

### 2018-02-05 writups/notes.tex

@@
 
-\subsection{Linear Regime, Compressible}
+\subsection{Developing the Anelastic/Boussinesq Approximations}
+
+Let's relax the incompressibility constraint (we will expand the continuity
+equation to first order, but the momentum equation will merit a separate
+treatment):
+\begin{equation}
+    \begin{split}
+        \pd{\rho_1}{t} + \vec{\nabla} \cdot \p*{\rho_0 \vec{u}_1} &= 0,\\
+        \pd{\vec{u}_1}{t} &= -\frac{\vec{\nabla}P}{\rho} - \vec{g}.
+    \end{split}
+\end{equation}
+Suppose we are interested in phenomena with characteristic length scale $L$ and
+time scale $\tau$. Let's first examine the relative magnitudes of the terms in
+the continuity equation
+\begin{equation*}
+    \frac{\rho_1}{\tau} + \frac{\rho_0 \abs{u_1}}{L} = 0.
+\end{equation*}
+Thus, if we are interested in time scales $\tau \gg
+\frac{\rho_1}{\rho_0}\frac{L}{\abs{u_1}}$ then we neglect the first term, the
+time derivative. This corresponds to making the perturbation incompressible;
+note that $\pd{\rho_1}{t} \approx \rd{\rho_1}{t}$ to first order, so we drop the
+high frequency restoring forces in the perturbation.
+
+For the momentum equation, we instead first manipulate to first order
+\begin{align}
+    -\frac{\vec{\nabla}P}{\rho} - \vec{g}
+        &= -\frac{\vec{\nabla}P_0}{\rho} - \frac{\vec{\nabla}P_1}{\rho_0}
+            - \vec{g},\nonumber\\
+        &= - \frac{\vec{\nabla}P_1}{\rho_0} +
+            \p*{\frac{\rho_0}{\rho} - 1}\vec{g},\nonumber\\
+        &= -\vec{\nabla}\p*{\frac{P_1}{\rho_0}}
+            - \frac{P_1}{\rho_0^2}\vec{\nabla}\rho_0
+            - \frac{\rho_1}{\rho_0}\vec{g}.
+\end{align}
+
+We now have three equations for four variables, $\vec{u}_1, \rho_1, P_1$. We
+must introduce a fourth equation, a thermodynamic equation. For an adiabatic
+process $P\rho^{-\gamma} \propto P^{1-\gamma}T^\gamma$ is constant. We thus
+introduce the concept of the \emph{potential temperature}
+\begin{equation}
+    \theta = T\p*{\frac{P_0}{P}}^\kappa.
+        \label{eq:pot_temp}
+\end{equation}
+For an adiabatic process, $\rd{\theta}{t} = 0$. Motivated by this, we use
+\begin{equation}
+    \begin{split}
+        \pd{1}{\rho_0}\pd{\rho_0}{z} &= \frac{1}{\gamma P_0}\pd{P_0}{z}
+            - \frac{1}{\theta_0}\pd{\theta_0}{z},\\
+        \frac{\rho_1}{\rho_0} &= \frac{1}{\gamma} \frac{P_1}{P_0}
+            - \frac{\theta_1}{\theta_0},
+    \end{split}
+\end{equation}
+to give the momentum equation form
+\begin{equation}
+    \rd{\vec{u}_1}{t} = -\vec{\nabla}\p*{\frac{P_1}{\rho_0}}
+        + \frac{P_1}{\rho_0}\p*{\frac{1}{\theta_0}\vec{\nabla}\theta_0}
+        + \vec{g} \frac{\theta_1}{\theta_0}.\label{eq:mom_unsimp}
+\end{equation}
+
+We also recognize $N^2 = \frac{g}{\theta_0}\pd{\theta_0}{z}$. We now do the same
+trick where we consider dynamics on length scale $D$ and compare the first and
+second terms in \autoref{eq:mom_unsimp}. Their ratio is $\frac{N^2 D}{g}$, and
+so as $N^2 \ll \frac{g}{D}$ the freefall time we neglect the second term.
+
+The anelastic fluid equations thus read
+\begin{equation}
+    \begin{split}
+        \vec{\nabla} \cdot \p*{\rho_0\vec{u}} &= 0,\\
+        \pd{\vec{u}_1}{t} + \vec{\nabla}\p*{\frac{P_1}{\rho_0}}
+            - \vec{g}\frac{\theta_1}{\theta_0} &= 0,\\
+        \pd{\theta_1}{t} + \p*{\vec{u} \cdot \vec{\nabla}}\theta_0 &= 0.
+    \end{split}\label{eq:lin_anelastic}
+\end{equation}
+
+The Boussinesq equations are obtained from these in the limit where $H \gg D$
+the relevant length scale, thus we allow $\rho_0$ to be approximately constant.
 

### 2018-02-06 writups/notes.tex
stash

@@
     \begin{split}
-        -i\omega \rho_1 - u_{1z}\rd{\rho_0}{z} &= 0,\\
-        iku_{1x} + \rd{u_{1z}}{z} &= 0,\\
+        -i\omega \rho_1 - u_{1z}\pd{\rho_0}{z} &= 0,\\
+        iku_{1x} + \pd{u_{1z}}{z} &= 0,\\
         -iw u_{1x} + \frac{ik_x P_1}{\rho_0} &= 0,\\
         -iw u_{1z} + \frac{1}{\rho_0}\pd{P_1}{z} +
-            \frac{\rho_1}{\rho_0^2}\rd{P_0}{z} &= 0.
+            \frac{\rho_1}{\rho_0^2}\pd{P_0}{z} &= 0.
     \end{split}
 \end{equation}
-We substitute $N^2 = -\frac{g}{\rho_0}\rd{\rho_0}{z}$ and $\rd{P_0}{z} = -\rho
+We substitute $N^2 = -\frac{g}{\rho_0}\pd{\rho_0}{z}$ and $\pd{P_0}{z} = -\rho
 g$ to obtain
@@
         -i\omega \rho_1 - u_{1z}\frac{\rho_0N^2}{g} &= 0,\label{eq:lin.cont}\\
-        iku_{1x} + \rd{u_{1z}}{z} &= 0,\label{eq:lin.div}\\
+        iku_{1x} + \pd{u_{1z}}{z} &= 0,\label{eq:lin.div}\\
         -iw u_{1x} + \frac{ik_x P_1}{\rho_0} &= 0,\label{eq:lin.mom_x}\\
@@
     \begin{align}
-        i\omega \rd{u_{1z}}{z} + \frac{k_x^2 P_1}{\rho_0} &= 0,
+        i\omega \pd{u_{1z}}{z} + \frac{k_x^2 P_1}{\rho_0} &= 0,
             \label{eq:lin.eq1}\\
-        \p*{\omega^2 - N^2 }u_{1z} + \frac{i\omega}{\rho_0}\rd{P_1}{z} &= 0.
+        \p*{\omega^2 - N^2 }u_{1z} + \frac{i\omega}{\rho_0}\pd{P_1}{z} &= 0.
             \label{eq:lin.eq2}
@@
 \begin{equation}
-    \rtd{u_{1z}}{z} + \frac{1}{\rho_0}\rd{\rho_0}{z}\rd{u_{1z}}{z}
+    \rtd{u_{1z}}{z} + \frac{1}{\rho_0}\pd{\rho_0}{z}\pd{u_{1z}}{z}
         + k_x^2\p*{\frac{N^2}{\omega^2} - 1}u_{1z} = 0.\label{eq:lin.incomp.gen}
@@
 
-\subsection{Boundary Conditions}
+\subsection{Anelastic Solution}
+
+We simply substitute $e^{i(\vec{k} \cdot \vec{r} - \omega t)}$ into
+\autoref{eq:lin_anelastic} with $\rho_0 \propto e^{-z/H}$ and obtain
+\begin{equation}
+    \begin{bmatrix}
+        0 & 0 & ik_x\rho_0 & ik_z \rho_0 - \frac{\rho_0}{H}\\
+        0 & -i\omega & 0 & \frac{N^2\theta_0}{g}\\
+        \frac{ik_x}{\rho_0} & 0 & -i\omega & 0\\
+        \frac{ik_z}{\rho_0} + \frac{1}{\rho_0 H} & -\frac{g}{\theta_0}
+            & 0 & -i\omega
+    \end{bmatrix} \begin{bmatrix}
+        P_1 \\ \theta_1 \\ u_{1x} \\ u_{1z}
+    \end{bmatrix} = 0.
+\end{equation}
+Taking the determinant of this matrix produces
+\begin{align}
+    -k_x^2\p*{-N^2 + \omega^2} + \p*{ik_z - \frac{1}{H}}
+        \p*{ik_z + \frac{1}{H}}\omega^2 &= 0,\nonumber\\
+    \frac{N^2k_x^2}{k_x^2 + k_z^2 + \frac{1}{4H^2}} &= \omega^2.
+\end{align}
+
+\section{Boundary Conditions}
+
+\subsection{Incompressible}
 
@@
 \begin{equation}
-    u_{1z} \propto e^{\frac{z}{2H}}e^{i\p*{k_xx + k_zz - \omega t}},
+    u_{1z} \propto e^{\frac{z}{2H}}e^{i\p*{k_xx + k_zz - \omega t}}.
         \label{eq:bc_below}
 \end{equation}
-where ($N^2 \equiv \frac{g}{H}$ is the Brunt-V\"ais\"al\"a frequency in an
-incompressible, stratified atmosphere)
-\begin{equation}
-    \omega^2 = \frac{N^2 k_x^2}{k_x^2 + k_z^2 + \frac{1}{4H^2}}.
-        \label{eq:omega}
-\end{equation}
-Thus, at constant $z = 0$ we must have
-\begin{equation}
-    \vec{u}_{1z}(z = 0) \propto e^{i(k_xx - \omega t)}\label{eq:bc_u1z}.
-\end{equation}
+We will take Boussinesq/WKB approximation $H \to 0$, reasonable in the linear
+regime, to simplify the BCs. Then $\rho_1 = 0$ (reasonable in the incompressible
+limit), $u_{1x} = -\frac{k_z}{k_x}u_{1z}$ and $P_1 = -\frac{\omega
+\rho_0}{k_z}u_{1z}$, from the linearized incompressible equations.
 
@@
 
-For the boundary conditions on $\rho$, we note that in the linear regime it
-should just have a phase offset from $u_{1z}$, thus we choose $\rho(z = 0)
-\propto e^{\frac{z}{2H}}ie^{i\p*{k_xx + k_zz - \omega t}}$ and a similar
-treatment at $z = L_z$, taking both a Dirichlet and Neumann BC\@.
-
-\subsection{Simulation}
+\section{Simulation}
 

### 2018-02-18 writups/notes.tex
no_g case

@@
 \newpage
+
+\chapter{Preliminary Problems}
+
+To get an intuition for how Dedalus and fluid mechanics works, we will solve
+some toy problems. Recall fluid equations in the presence of a uniform
+gravitational field $\vec{g} = -g\hat{z}$:
+\begin{equation}
+    \begin{split}
+        \rd{\rho}{t} + \vec{\nabla} \cdot \vec{u} &= 0,\\
+        \rd{\vec{u}}{t} + \frac{\vec{\nabla}P}{\rho} - \vec{g} &= 0.
+    \end{split}\label{eq:fluid_eq}
+\end{equation}
+
+In the incompressible limit, $\rd{\rho}{t} = 0$, which implies $\vec{\nabla}
+\cdot \vec{u} = 0$. We use subscripts to indicate perturbed quantities, $Q_0$ is
+background and $Q_1$ is perturbed. We will generally use $\vec{u}_0 = 0$ unless
+otherwise noted. We will also generally assume symmetry along all axes except
+$z$ the vertical axis.
+
+In the incompressible limit, the fluid equations become
+\begin{equation}
+    \begin{split}
+        \vec{\nabla} \cdot \vec{u}_1 &= 0,\\
+        \pd{\rho_1}{t} + u_{1z}\pd{\rho_0}{z} &= 0,\\
+        \pd{\vec{u}_1}{t} + \frac{1}{\rho_0}\vec{\nabla}P_1
+            + \frac{\rho_1 g \hat{z}}{\rho_0} &= 0.
+    \end{split}\label{eq:lin.incomp}
+\end{equation}
+We have used $\vec{\nabla}P_0 = -\rho_0 g \hat{z}$ in the absence of
+perturbations.
+
+\section{Incompressible, No Gravity}
+
+We note that in the no gravity limit that $\rho_1$ does not have an effect on
+other dynamical variables, so the equations of motion we must solve are
+\begin{equation}
+    \begin{split}
+        \vec{\nabla} \cdot \vec{u}_1 &= 0,\\
+        \pd{\vec{u}_1}{t} + \frac{\vec{\nabla}P_1}{\rho_0} &= 0.
+    \end{split}\label{eq:lin.no_g_eom}
+\end{equation}
+We can take the divergence of the momentum equation and substitute the
+continuity equation to get $\nabla^2 P = 0$. This is a Laplace equation, which
+we've solved countless times. Imposing periodic boundary conditions in the $x$
+direction and $P_1(z = L) = 0, P_1(z = 0) = \mathcal{P}(x, t)$, we obtain
+eigenfunctions
+\begin{equation}
+    \begin{split}
+        P_{1, n}(x, z, t) &=
+            \frac{\mathcal{P}_n(t)}{\sinh(k_nL)}e^{ik_nx}\sinh\p*{k_n(L - z)},\\
+        u_{1x, n}(x, z, t) &=
+            \int\limits^t -\frac{1}{\rho_0}\pd{P_{1, n}}{x}\;\mathrm{d}t,\\
+        u_{1z, n}(x, z, t) &=
+            \int\limits^t -\frac{1}{\rho_0}\pd{P_{1, n}}{z}\;\mathrm{d}t.
+    \end{split}\label{eq:incomp_nog}
+\end{equation}
+We define $k_n = \frac{2\pi n}{L}, n \in [0, L - 1]$ and $\mathcal{P}(x, t) =
+\sum\limits_n \mathcal{P}_n(t)e^{ik_nx}$.
+
+Thus, if we impose BCs $\mathcal{P}(x, t) = \sin \frac{2\pi x}{L}$ and start
+with initial conditions such that all quantities are zero, we would expect after
+transients die out that
+\begin{equation}
+    \begin{split}
+        P(x, z, t) &= \frac{\sin \frac{2\pi x}{L}}{\sinh 2\pi}
+            \sinh \p*{2\pi \frac{L - z}{L}},\\
+        u_{1x}(x, z, t) &= -\frac{2\pi t}{L\rho_0}
+            \frac{\cos \frac{2\pi x}{L}}{\sinh 2\pi}\sinh \p*{2\pi \frac{L -
+                z}{L}},\\
+        u_{1z}(x, z, t) &= +\frac{2\pi t}{L\rho_0}
+            \frac{\sin \frac{2\pi z}{L}}{\sinh 2\pi}\cosh \p*{2\pi \frac{L -
+                z}{L}}.
+    \end{split}\label{eq:incomp_nog_sol}
+\end{equation}
+This is in good agreement with the results, presented in \autoref{fig:no_g}.
+Note that $P$ is constant while $\vec{u}$ increases linearly in time, and we
+observe the expected $\sim \sin x \sinh \frac{L - z}{z}$ dependence.
+
+\begin{figure}[!h]
+    \centering
+    \begin{subfigure}{0.3\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{../sims/prelim/no_g/no_g_P_t50.png}
+    \end{subfigure}
+    \begin{subfigure}{0.3\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{../sims/prelim/no_g/no_g_ux_t50.png}
+    \end{subfigure}
+    \begin{subfigure}{0.3\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{../sims/prelim/no_g/no_g_uz_t50.png}
+    \end{subfigure}
+
+    \begin{subfigure}{0.3\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{../sims/prelim/no_g/no_g_P_t100.png}
+    \end{subfigure}
+    \begin{subfigure}{0.3\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{../sims/prelim/no_g/no_g_ux_t100.png}
+    \end{subfigure}
+    \begin{subfigure}{0.3\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{../sims/prelim/no_g/no_g_uz_t100.png}
+    \end{subfigure}
+    \caption{$P, u_x, u_z$ at $t = 0.5$ and $t = 1$ for $\rho_0 = 1$.
+    }\label{fig:no_g}
+\end{figure}
+
+It is worth noting that, since our \autoref{eq:lin.no_g_eom} reduced to a Laplace
+equation, we needed two $z$ BCs and two $x$ BCs (periodic BCs amount to equating
+the value and derivative of the function). This is in agreement with the
+observation that the original \autoref{eq:lin.no_g_eom} had two derivatives in
+$x, z$ apiece, so we needed two BCs each.
 

### 2018-02-18 dedalus_progs/2d_incomp/2d_travel.py

@@ -1,153 +1,0 @@
-'''
-Incompressible fluid equations w/o delta rho, velocity perturbation just travels
-'''
-
-import logging
-import numpy as np
-import matplotlib.pyplot as plt
-
-from dedalus import public as de
-from dedalus.extras.plot_tools import quad_mesh, pad_limits
-
-logger = logging.getLogger(__name__)
-XMAX = 10
-ZMAX = 100
-N_X = 64
-N_Z = 256
-DX = XMAX / N_X
-DZ = ZMAX / N_Z
-T_F = 20
-DT = 1e-2
-
-rho0 = 1 # kg/m^3
-P0 = 100 # Pa
-q0 = 1e5 # K
-k = 2
-S = 10 / k
-H = 10 / k
-A = 0.05 / k
-m = -0.4 * k
-g = P0 / (H * rho0)
-nu = 2e-3
-
-# Bases and domain
-x_basis = de.Fourier('x', N_X, interval=(0, XMAX), dealias=3/2)
-z_basis = de.Chebyshev('z', N_Z, interval=(0, ZMAX), dealias=3/2)
-domain = de.Domain([x_basis, z_basis], np.float64)
-
-problem = de.IVP(domain, variables=['P', 'u', 'w', 'uz', 'wz'])
-problem.meta['P','u','w']['z']['dirichlet'] = True
-problem.parameters['nu'] = nu
-problem.parameters['H'] = H
-problem.parameters['g'] = g
-problem.parameters['rho0'] = rho0
-problem.add_equation("dx(u) + wz = 0")
-# problem.add_equation('dt(rho) = -u * dx(rho) - w * dz(rho)')
-problem.add_equation(
-    "dt(u) - nu*(dx(dx(u)) + dz(uz)) + dx(P)/rho0 = -(u*dx(u) + w*uz)")
-problem.add_equation(
-    "dt(w) - nu*(dx(dx(w)) + dz(wz)) + dz(P)/rho0 = -(u*dx(w) + w*wz) - g")
-problem.add_equation("uz - dz(u) = 0")
-problem.add_equation("wz - dz(w) = 0")
-problem.add_bc("left(u) = 0")
-problem.add_bc("left(w) = 0", condition="(nx != 0)")
-problem.add_bc("left(P) = rho0 * g * 100", condition="(nx == 0)")
-problem.add_bc("right(u) = 0")
-problem.add_bc("right(w) = 0", condition="(nx != 0)")
[…]

### 2018-02-18 sims/prelim/no_g/no_g.py

@@ -1,0 +1,76 @@
+'''
+Incompressible fluid equations w/
+'''
+
+import logging
+import numpy as np
+import matplotlib.pyplot as plt
+
+from dedalus import public as de
+from dedalus.extras.plot_tools import quad_mesh, pad_limits
+
+logger = logging.getLogger(__name__)
+XMAX = 10
+ZMAX = 10
+N_X = 64
+N_Z = 64
+DX = XMAX / N_X
+DZ = ZMAX / N_Z
+T_F = 20
+DT = 1e-2
+
+# Bases and domain
+x_basis = de.Fourier('x', N_X, interval=(0, XMAX), dealias=3/2)
+z_basis = de.Chebyshev('z', N_Z, interval=(0, ZMAX), dealias=3/2)
+domain = de.Domain([x_basis, z_basis], np.float64)
+
+problem = de.IVP(domain, variables=['P', 'ux', 'uz'])
+problem.meta['P']['z']['dirichlet'] = True
+problem.parameters['rho0'] = 1
+problem.add_equation("dx(ux) + dz(uz) = 0")
+problem.add_equation("dt(ux) + dx(P)/rho0 = 0")
+problem.add_equation("dt(uz) + dz(P)/rho0 = 0")
+problem.add_bc("right(P) = 0")
+problem.add_bc("left(P) = sin(6.28 * x / 10)")
+
+# Build solver
+solver = problem.build_solver(de.timesteppers.RK222)
+solver.stop_sim_time = T_F
+solver.stop_wall_time = np.inf
+solver.stop_iteration = 100
+
+# Initial conditions
+x = domain.grid(0)
+z = domain.grid(1)
+ux = solver.state['ux']
+uz = solver.state['uz']
+P = solver.state['P']
+
+# need to reverse for some reason?
+gshape = domain.dist.grid_layout.global_shape(scales=1)
+P['g'] = np.zeros(gshape)
+ux['g'] = np.zeros(gshape)
+uz['g'] = np.zeros(gshape)
+
+# Main loop
+for stop_iter in [50, 100]:
+    solver.stop_iteration = stop_iter
+    while solver.ok:
+        solver.step(DT)
+
+    ux.set_scales(1, keep_data=True)
+    uz.set_scales(1, keep_data=True)
+    P.set_scales(1, keep_data=True)
+
+    # plot vars
+    xmesh, zmesh = quad_mesh(x=x[:,0], y=z[0])
[…]

### 2018-02-18 writups/notes.tex
stash

@@
 \author{Yubo Su}
+\rhead{}
 
@@
 We can take the divergence of the momentum equation and substitute the
-continuity equation to get $\nabla^2 P = 0$. This is a Laplace equation, which
-we've solved countless times. Imposing periodic boundary conditions in the $x$
-direction and $P_1(z = L) = 0, P_1(z = 0) = \mathcal{P}(x, t)$, we obtain
-eigenfunctions
+continuity equation to get $\nabla^2 P = 0$.
+
+\subsection{Dirichlet BCs}
+
+This is a Laplace equation, which we've solved countless times. Imposing
+periodic boundary conditions in the $x$ direction and $P_1(z = L) = 0, P_1(z =
+0) = \mathcal{P}(x, t)$, we obtain eigenfunctions
 \begin{equation}
@@
 \end{equation}
-We define $k_n = \frac{2\pi n}{L}, n \in [0, L - 1]$ and $\mathcal{P}(x, t) =
+We define $k_n = \frac{2\pi n}{L}, n \geq 0$ and $\mathcal{P}(x, t) =
 \sum\limits_n \mathcal{P}_n(t)e^{ik_nx}$.
@@
 Note that $P$ is constant while $\vec{u}$ increases linearly in time, and we
-observe the expected $\sim \sin x \sinh \frac{L - z}{z}$ dependence.
+observe the expected $\sim \sin x \sinh \frac{L - z}{z}$ dependence. In fact,
+$u_{1x}, u_{1z}$ are exactly $\frac{2\pi}{10}$ at $t = 1$.
 
@@
     \end{subfigure}
-    \caption{$P, u_x, u_z$ at $t = 0.5$ and $t = 1$ for $\rho_0 = 1$.
-    }\label{fig:no_g}
+    \caption{$P, u_x, u_z$ at $t = 0.5$ and $t = 1$ for $\rho_0 = 1$. We choose
+    a $L = 10$ square domain.}\label{fig:no_g}
 \end{figure}
@@
 
-\chapter{2D Wave Breaking in Atmospheres}
+\subsection{Sommerfield/Radiative + Driving BCs}
 
-The goal of this will be to lay out a formalism that can reproduce Sutherland et
-al.~2011\footnote{DOI:10.1175/JAS-D-11--097.1} and also investigate driven
-oscillations versus the breaking of a single wave packet. This represents wave
-breaking in the atmosphere.
+This is the more interesting case. Let's go back to the Laplace equation
+$\nabla^2 P = 0$ but instead implement a driving term at $z = 0$ and radiative
+BCs at $z = L$. This
 
-\section{Dynamical Setup}
+\section{Incompressible, Stratified w/ Gravity}
 
-We adopt notation where $q_0$ is the background quantity and $q_1$ is the
-perturbed quantity from the propagating wave.
+\subsection{Eigenfunctions}
 
-The fluid equations are
-\begin{subequations}\label{se:fl_eq}
-    \begin{align}
-        \pd{\rho}{t} + \vec{\nabla} \cdot \rho \vec{u} &= 0,
-            \label{eq:fl_eq.cont}\\
-        \rd{\vec{u}}{t} &= -\vec{\nabla}\frac{P}{\rho} - g\hat{z},
-            \label{eq:fl_eq.mom}
-    \end{align}
-\end{subequations}
-where we will take gravity to be uniform throughout the domain of interest. We
-will study for no background flow $\vec{u}_0 = 0$ and in the presence of
-stratification $\rho_0 \propto e^{-z/H}$. In the absence of any perturbations it
-is then easy to show that $\rd{P_0}{z} = -\rho_0 g$.
-
-\subsection{Linear, Incompressible}
-
-\subsubsection{Solution for Arbitrary Stratification}
-
-First, we solve the incompressible case $c_s^2 \to \infty, \vec{\nabla} \cdot
-\vec{u} = 0$ in the linear regime. For funsies, we solve for arbitrary
-stratification first. The fluid equations to first order reduce to
+Let's restore the $\rho_1 g$ term now. For funsies, we begin by solving for
+arbitrary stratification $\rho_0(z)$ first. The fluid equations to first order
+reduce to
 \begin{equation}
@@
         \pd{\vec{u}_1}{t} &= -\frac{\vec{\nabla}P_1}{\rho_0}
-            - P_1 \vec{\nabla}\frac{1}{\rho_0}
+            - \frac{\rho_1 g}{\rho_0}
     \end{split}\label{eq:lin_incomp}
@@
         -iw u_{1z} + \frac{1}{\rho_0}\pd{P_1}{z} +
-            \frac{\rho_1}{\rho_0^2}\pd{P_0}{z} &= 0.
+            \frac{\rho_1 g}{\rho_g} &= 0.
     \end{split}
 \end{equation}
-We substitute $N^2 = -\frac{g}{\rho_0}\pd{\rho_0}{z}$ and $\pd{P_0}{z} = -\rho
-g$ to obtain
+We substitute $N^2 = -\frac{g}{\rho_0}\pd{\rho_0}{z}$ to obtain
 \begin{subequations}
@@
 
-\subsubsection{Introduce Stratification}
-
-With stratification $\rho \propto e^{-z/H}$ \autoref{eq:lin.incomp.gen} clearly
-has exponential solutions $e^{\kappa z}$ for
+Let's now pick stratification $\rho \propto e^{-z/H}$
+\autoref{eq:lin.incomp.gen} clearly has exponential solutions $e^{\kappa z}$ for
 \begin{equation}
@@
 \end{equation}
-
-We permit complex $\kappa = -\frac{1}{2H} + ik_z$, and from the above clearly
+We permit complex $\kappa = \frac{1}{2H} + ik_z$, and from the above clearly
 \begin{align}
@@
 \end{align}
+Thus the eigenfunctions are
+\begin{equation}
+    \begin{split}
+        u_{1z} &= e^{z/2H} e^{i(k_zz + k_xx - \omega t)},\\
+        u_{1x} &= -\frac{k_z + i/2H}{k_x} u_{1z},\\
+        \rho_1 &= \frac{i \rho_0}{H\omega} u_{1z},\\
+        P_1 &= -\frac{\rho_0 \omega}{k_x^2}\p*{k_z + i/2H}u_{1z}.
+    \end{split}\label{eq:strat_eigens}
+\end{equation}
+
+\subsection{Solving an IVP}
+
+\section{(Algebra) The Anelastic/Boussinesq Approximations}
 
@@
 
-\subsection{Anelastic Solution}
+\subsection{Anelastic Solution to Stratified Atmosphere}
 
@@
 
+\chapter{2D Wave Breaking in Atmospheres}
+
+\section{Dynamical Setup}
+
+TODO (fluid equations, driven on bottom, parameters)
+
 \section{Boundary Conditions}
 
-\subsection{Incompressible}
-
-We must bound this to a finite domain. We choose periodic boundary conditions in
-$x$ with total length $L_x$, $x \in \s*{-\frac{L}{2}, \frac{L}{2}}$. We choose
-$z$ dimension to have total length $L_z$, $z \in [0, L_z]$.
-
-To set up the boundary condition at $z = 0$, we recall that gravity waves in an
-atmosphere with $e^{-z/H}$ profile have form
-\begin{equation}
-    u_{1z} \propto e^{\frac{z}{2H}}e^{i\p*{k_xx + k_zz - \omega t}}.
-        \label{eq:bc_below}
-\end{equation}
-We will take Boussinesq/WKB approximation $H \to 0$, reasonable in the linear
-regime, to simplify the BCs. Then $\rho_1 = 0$ (reasonable in the incompressible
-limit), $u_{1x} = -\frac{k_z}{k_x}u_{1z}$ and $P_1 = -\frac{\omega
-\rho_0}{k_z}u_{1z}$, from the linearized incompressible equations.
-
-The boundary condition at $z = L_z$ is much harder to determine. It is clear
-that $\vec{u}_1(z = \infty) = \rho_1(z = \infty) = 0$, and for $L_z \gg H$ this
-would be a reasonable approximation, if simply because we expect the majority of
-the wave to dissipate via turbulent dissipation as $z$ reaches many $H$. We will
-simply choose the BC to be many multiples of $H$. We can solve with both a
-Dirichlet and Neumann BC and compare the two solutions; if the solutions
-differ significantly then we must choose a larger $L_z$. These are the only two
-solutions that can be implemented where we do not need the phase of the linear
-wave, which we lose during the nonlinear breaking region, to relate the function
-and derivative at the boundary.
+TODO (periodic in x, show that right number of BCs in $z$, discuss whether gauge
+choice).
 
@@
 
-We begin our simulation with $\rho_1 = 0, \vec{u}_1 = 0$ strictly within the
-domain of simulation. We will borrow some values from Sutherland's paper and use
-$k_z = 2\;\mathrm{km^{-1}}$ then define $k_x = -0.4k_z, H = 10 / k_z, A = 0.05 /
-k_z, L_z = 300 / k_z, L_x = 20 / k_z$, We also use $\mu \approx 29, T =
-273\;\mathrm{K}, \rho_0 = 1\;\mathrm{kg/m^3}, P_0 = \frac{\rho_0 k_BT}{\mu m_p},
-g = 10\;\mathrm{m/s^2}$.
+TODO (CFL conditions etc.)
 

### 2018-02-19 writups/notes.tex
get some work done on stratified

@@
 $u_{1x}, u_{1z}$ are exactly $\frac{2\pi}{10}$ at $t = 1$.
-
 \begin{figure}[!h]
@@
 
+\subsection{Things to Note}
+
 It is worth noting that, since our \autoref{eq:lin.no_g_eom} reduced to a Laplace
@@
 
-\subsection{Sommerfield/Radiative + Driving BCs}
+It is also worth seeing from our solution that $P$ immediately goes to the
+equilibrium solution. This is not surprising since in the incompressible limit,
+sound speed goes to infinity which is the timescale on which the pressure field
+adjusts to net forces. Thus, the dynamics solely arise from the static pressure
+field pushing the velocity field to equilibrium.
 
-This is the more interesting case. Let's go back to the Laplace equation
-$\nabla^2 P = 0$ but instead implement a driving term at $z = 0$ and radiative
-BCs at $z = L$. This
+We chose time-independent $\mathcal{P}(x, t)$, but it is clear that whatever
+$\mathcal{P}(x, t)$ we choose, the time dependence propagates to the velocities
+by way of an integral. If we had instead chosen to take Fourier transform $t \to
+\omega$, we would have had to integrate the boundary condition against the
+eigenfunctions for each of the $\omega$, which is still easily computable, to
+get the full $u_{1x}(x, z, t)$. We consider this in preparation for when we can
+only solve for a set of $\vec{k}, \omega$ in the next problem.
+
+The next thing we would have wanted to do is solve a problem with radiative BCs,
+but we need to have wave solutions, which are missing in the absence of gravity.
+We thus move on to the next configuration.
 
@@
     k_z^2 &= -\frac{1}{4H^2} + k_x^2\p*{\frac{N^2}{\omega^2} - 1},\nonumber\\
-    \omega^2 &= \frac{N^2k_x^2}{k_x^2 + k_z^2 + \frac{1}{4H^2}}.
+    \omega^2 &= \frac{N^2k_x^2}{k_x^2 + k_z^2 +
+        \frac{1}{4H^2}}.\label{eq:strat_disp}
 \end{align}
@@
 
-\subsection{Solving an IVP}
+\subsection{Analytically Solving an IVP, Dirichlet + Driving BCs}
+
+We will analyze everything in terms of $u_{1z}$ since it has the simplest form;
+note that when actually choosing the BCs we will have to consider the gauge
+freedom of $P$ and some considerations we defer to the computational section.
+
+Currently, we have a set of eigenfunctions
+\begin{equation}
+    u_{1z}\p*{x, z, t \middle| \vec{k}, \omega}
+        = e^{z/2H} e^{i(k_zz + k_xx - \omega t)}.
+        \label{eq:strat.uz_eigen}
+\end{equation}
+Note that $u_{1z}$ is really only a function of two parameters rather than the
+three $(k_x, k_z, \omega)$, since the three are related by dispersion relation
+\autoref{eq:strat_disp}.
+
+Now, we implement BCs. Consider domain $x, z \in [0, L]$. We will use periodic
+BCs again in $x$, so then $k_{x, n} = \frac{2\pi n}{L}, n \geq 0$. Then, we will
+require $u_{1z, n}(x, L, t) = 0$, a Dirichlet condition at the top boundary,
+which restricts us to eigenfunctions of form
+\begin{equation}
+    u_{1z, n}\p*{x, z, t \middle| \vec{k}, \omega}
+        = e^{z/2H}e^{i(k_xx - \omega t)}\sin\p*{k_z\p*{L - z}}.
+        \label{eq:strat.uz_eigen_l-z}
+\end{equation}
+
+Finally, we must choose a BC at $z = 0$. We will choose a general function
+$u_{1z}(x, 0, t) = F(x, t)$ where we can decompose
+\begin{equation}
+    F(x, t) = \int\limits \sum\limits_n \mathcal{F}(k_{x, n}, \omega)
+        e^{i(k_{x, n}x - \omega t)}\;\mathrm{d}\omega.
+\end{equation}
+Matching BCs then gives us general solution for $u_{1z}$ given an arbitrary
+driving function
+\begin{equation}
+    u_{1z}\p*{x, z, t \middle| \vec{k}, \omega}
+        = \int\limits \sum\limits_n \mathcal{F}\p*{k_{x, n}, \omega}
+            \frac{e^{z/2H}e^{i\p*{k_{x, n}x - \omega t}}\sin \p*{k_z(L - z)}}{
+            \sin k_zL}\;\mathrm{d}\omega.
+            \label{eq:strat.full_sol}
+\end{equation}
+
+For ease of computation, let's pick $F(x, t) = \cos \p*{\frac{2 \pi x}{L} -
+\omega_0 t}$, so our full expected solution is (note $A + \epsilon =
+Ae^{\epsilon/A} + \mathcal{O}(\epsilon^2)$)
+\begin{equation}
+    \begin{split}
+        u_{1z}\p*{x, z, t \middle | \vec{k}, \omega}
+            &= e^{z/2H}\frac{\cos\p*{\frac{2\pi x}{L} - \omega_0 t}
+                \sin \p*{k_z(L - z)}}{\sin k_zL},\\
+        u_{1x}\p*{x, z, t \middle | \vec{k}, \omega}
+            &\approx -\frac{k_z}{k_x} e^{z/2H}\frac{
+                \cos\p*{\frac{2\pi x}{L} - \omega_0 t + \frac{1}{2Hk_z}}
+                \sin \p*{k_z(L - z)}
+            }{\sin k_zL},\\
+        \rho_1\p*{x, z, t \middle | \vec{k}, \omega}
+            &\approx \frac{\rho_0}{H\omega} e^{z/2H}\frac{
+                -\sin\p*{\frac{2\pi x}{L} - \omega_0 t}
+                \sin \p*{k_z(L - z)}
+            }{\sin k_zL},\\
+        P_1\p*{x, z, t \middle | \vec{k}, \omega}
+            &\approx -\frac{\rho_0 \omega k_z}{k_x^2} e^{z/2H}\frac{
+                -\cos\p*{\frac{2\pi x}{L} - \omega_0 t + \frac{1}{2Hk_z}}
+                \sin \p*{k_z(L - z)}
+            }{\sin k_zL},\\
+    \end{split}\label{eq:strat.part_sol}
+\end{equation}
+where $k_z: \omega(k_x, k_z) = \omega_0$ by the dispersion relation
+\autoref{eq:strat_disp}. Note that $H$ contributes both to the overall
+exponential profile and to the phase lag of $u_{1z}$.
+
+\subsection{Computationally Solving an IVP, Dirichlet + Driving BCs}
+
+To solve this computationally with the aforementioned BCs, periodic in $x$,
+Dirichlet $0$ at $z = L$ and $\cos \p*{\frac{2\pi x}{L} - \omega_0 t}$ at $z =
+0$, we must address the gauge freedom in $P$. This arises because for $k_x = 0$,
+the divergence-free condition $\vec{\nabla} \cdot \vec{u}_1 = \pd{u_z}{z} = 0$
+specificies $u_{1z}$ up to a constant already, so the bottom BC will fix the
+value of $u_z$ when $k_x = 0$.
+
+A different way of phrasing the same argument is as follows. Consider the
+discrete $N \times N$ (square for notational simplicity) grid. At the boundary,
+there is a list of values $f\p*{\z*{x_i}, z_0}$ that lives in an $N$ dimensional
+space. We can thus pick a spanning set of $N$ basis vectors, and for each of
+these $N$ vectors, by enforcing $\vec{\nabla} \cdot \vec{u} = 0$ at the boundary
+we fix the allowed $f\p*{\z*{x_i}, z_{-1}}$ boundary conditions we can implement.
+But there exists a choice of basis vectors for which one of the basis vectors is
+constant $f_i\p*{\z*{x_i}, z_0} = C$. For this basis vector, the boundary
+condition is fully determined, so we only have $N - 1$ dimensions from which to
+choose the BCs $f\p*{\z*{x_i}, z_{-1}}$. Since the dimensionality of a space
+cannot depend on the choice of basis vector, the divergence free condition
+actually yields an extra degree of freedom.
+
+We can use this extra degreee of freedom to specify $P(z = L) = P_L$ some
+constant. Then we can just simulate away! Let's choose $L = 10, k_x = k_z = 1, H
+= 10, g = 10$ which constrains $N^2 \equiv g/H = 1, \omega^2 = \frac{1}{2 +
+1/400}$.
+
+\subsection{Sommerfield/Radiative BCs}
 

### 2018-02-19 sims/prelim/strat/strat.py

@@ -1,11 +1,5 @@
 '''
-linearized fluid equations w/o gravity. Fluid equations reduce to
-- div u = 0
-- dt(u) + grad(P) / rho0 = 0
-
-Domain is L x L
-ICs are zero everywhere
-BCs are periodic x, Dirichlet 0 at z = L, sin(2*pi*x/L) at z = 0.
+Incompressible fluid equations w/
 '''
 
 import logging
@@ -36,8 +30,8 @@ problem.parameters['rho0'] = 1
 problem.add_equation("dx(ux) + dz(uz) = 0")
 problem.add_equation("dt(ux) + dx(P)/rho0 = 0")
 problem.add_equation("dt(uz) + dz(P)/rho0 = 0")
-problem.add_bc("right(P) = 0")
-problem.add_bc("left(P) = sin(6.28 * x / 10)")
+problem.add_bc("right(dt(P) - dz(P)) = 0")
+problem.add_bc("left(P) = sin(6.28 * (x - t) / 10)")
 
 # Build solver
 solver = problem.build_solver(de.timesteppers.RK222)

### 2018-02-19 writups/notes.tex
stuff is working

@@
 
-We can use this extra degreee of freedom to specify $P(z = L) = P_L$ some
-constant. Then we can just simulate away! Let's choose $L = 10, k_x = k_z = 1, H
-= 10, g = 10$ which constrains $N^2 \equiv g/H = 1, \omega^2 = \frac{1}{2 +
-1/400}$.
+We can use this extra degreee of freedom to specify $P(z = L) = 0$ so the
+oscillations should have zero mean. Then we can just simulate away!
+
+Since we're much more interested in the $z$ behavior, we will set the $z$ axis
+to be further away. Let's choose $\rho_0 = 1, L_x = 10, L_z = 20, k_x = 2, k_z =
+1, H = 5, g = 10$ which constrains $N^2 \equiv g/H = 1, \omega^2 = \frac{4}{5 +
+1/100}$. Some considerations:
+\begin{itemize}
+    \item $\frac{1}{k_xH}, \frac{1}{k_zH}$ are $1/10, 1/5$ respectively,
+        small-ish but $L_Z = 8H$ so we should be able to see the $e^{z/2H}$
+        behavior well.
+
+    \item The vertical phase velocity is $\frac{\omega}{k_z} \sim 1$, so we
+        should simulate at least out to $t = 20$ for the wave to fully propagate
+        to the boundary.
+\end{itemize}
 

### 2018-02-19 sims/prelim/no_g/no_g.py

@@ -8,14 +8,12 @@ ICs are zero everywhere
 BCs are periodic x, Dirichlet 0 at z = L, sin(2*pi*x/L) at z = 0.
 '''
 
-import logging
 import numpy as np
 import matplotlib.pyplot as plt
 
 from dedalus import public as de
 from dedalus.extras.plot_tools import quad_mesh, pad_limits
 
-logger = logging.getLogger(__name__)
 XMAX = 10
 ZMAX = 10
 N_X = 64

### 2018-02-19 sims/prelim/strat/strat.py

@@ -1,76 +1,102 @@
 '''
-Incompressible fluid equations w/
+Incompressible fluid equations w/ vertical stratification
+- div u1 = 0
+- dt (rho1) - rho0 * u1z / H = 0
+- dt(u1) = -grad(P1) / rho0 - rho1 * g / rho0
 '''
-
 import logging
 import numpy as np
 import matplotlib.pyplot as plt
 
+logger = logging.getLogger(__name__)
 from dedalus import public as de
 from dedalus.extras.plot_tools import quad_mesh, pad_limits
 
-logger = logging.getLogger(__name__)
 XMAX = 10
-ZMAX = 10
+ZMAX = 20
 N_X = 64
-N_Z = 64
+N_Z = 256
 DX = XMAX / N_X
 DZ = ZMAX / N_Z
-T_F = 20
+
+T_F = 100
 DT = 1e-2
+KX = 2
+KZ = 1
+H = 5
+G = 10
+N = G / H
+RHO0 = 1
 
-# Bases and domain
-x_basis = de.Fourier('x', N_X, interval=(0, XMAX), dealias=3/2)
-z_basis = de.Chebyshev('z', N_Z, interval=(0, ZMAX), dealias=3/2)
-domain = de.Domain([x_basis, z_basis], np.float64)
+NUM_SNAPSHOTS = 40
 
-problem = de.IVP(domain, variables=['P', 'ux', 'uz'])
-problem.meta['P']['z']['dirichlet'] = True
-problem.parameters['rho0'] = 1
-problem.add_equation("dx(ux) + dz(uz) = 0")
-problem.add_equation("dt(ux) + dx(P)/rho0 = 0")
-problem.add_equation("dt(uz) + dz(P)/rho0 = 0")
-problem.add_bc("right(dt(P) - dz(P)) = 0")
-problem.add_bc("left(P) = sin(6.28 * (x - t) / 10)")
+if __name__ == '__main__':
+    # Bases and domain
+    x_basis = de.Fourier('x', N_X, interval=(0, XMAX), dealias=3/2)
+    z_basis = de.Chebyshev('z', N_Z, interval=(0, ZMAX), dealias=3/2)
+    domain = de.Domain([x_basis, z_basis], np.float64)
 
-# Build solver
-solver = problem.build_solver(de.timesteppers.RK222)
-solver.stop_sim_time = T_F
-solver.stop_wall_time = np.inf
-solver.stop_iteration = 100
[…]

### 2018-02-26 writups/notes.tex
fork out helper

@@
         \centering
-        \includegraphics[width=\textwidth]{../sims/prelim/no_g/no_g_P_t50.png}
+        \includegraphics[width=\textwidth]{../sims/2d_no_g/no_g_P_t50.png}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{../sims/prelim/no_g/no_g_ux_t50.png}
+        \includegraphics[width=\textwidth]{../sims/2d_no_g/no_g_ux_t50.png}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{../sims/prelim/no_g/no_g_uz_t50.png}
+        \includegraphics[width=\textwidth]{../sims/2d_no_g/no_g_uz_t50.png}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{../sims/prelim/no_g/no_g_P_t100.png}
+        \includegraphics[width=\textwidth]{../sims/2d_no_g/no_g_P_t100.png}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{../sims/prelim/no_g/no_g_ux_t100.png}
+        \includegraphics[width=\textwidth]{../sims/2d_no_g/no_g_ux_t100.png}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{../sims/prelim/no_g/no_g_uz_t100.png}
+        \includegraphics[width=\textwidth]{../sims/2d_no_g/no_g_uz_t100.png}
     \end{subfigure}
@@
 \autoref{eq:strat_disp}. Note that $H$ contributes both to the overall
-exponential profile and to the phase lag of $u_{1z}$.
+exponential profile and to the phase lag of $u_{1x}$.
 

### 2018-02-26 sims/2d_strat/strat.py

@@ -6,110 +6,64 @@ Incompressible fluid equations w/ vertical stratification
 
 Periodic BC in x, Dirichlet/Neumann 0 at z=L, Driving term at z=0
 '''
-import logging
 import numpy as np
-import matplotlib.pyplot as plt
-
-logger = logging.getLogger(__name__)
-from dedalus import public as de
-from dedalus.extras.plot_tools import quad_mesh
-from dedalus.extras.flow_tools import CFL
-
-XMAX = 10
-ZMAX = 20
-N_X = 64
-N_Z = 256
-DX = XMAX / N_X
-DZ = ZMAX / N_Z
-
-T_F = 80
-DT = 2e-2
-KX = 2
-KZ = 1
-H = ZMAX / 3
-G = 10
-N = G / H
-RHO0 = 1
-
-NUM_SNAPSHOTS = 80
+from strat_helper import run_strat_sim
 
 if __name__ == '__main__':
-    # Bases and domain
-    x_basis = de.Fourier('x', N_X, interval=(0, XMAX), dealias=3/2)
-    z_basis = de.Chebyshev('z', N_Z, interval=(0, ZMAX), dealias=3/2)
-    domain = de.Domain([x_basis, z_basis], np.float64)
-    x = domain.grid(0)
-    z = domain.grid(1)
+    params = {'H_FACT': 3,
+              'ZMAX': 20,
+              'XMAX': 10,
+              'KX': 2 * np.pi / 10, # 2pi/L_X
+              'KZ': 1,
+              'RHO0': 1,
+              'G': 10}
 
-    problem = de.IVP(domain, variables=['P', 'rho', 'ux', 'uz'])
-    problem.meta['uz']['z']['dirichlet'] = True
-    problem.parameters['L'] = XMAX
-    problem.parameters['g'] = G
-    problem.parameters['H'] = H
-    problem.parameters['omega'] = np.sqrt(
-        N**2 * KX**2 / (KX**2 + KZ**2 + 0.25 / H**2))
+    def dirichlet_bc(problem):
+        problem.add_bc("right(uz) = 0", condition="nx != 0")
 
-    # rho0 stratification
-    rho0 = domain.new_field()
-    rho0.meta['x']['constant'] = True
-    rho0['g'] = RHO0 * np.exp(-z / H)
-    xmesh, zmesh = quad_mesh(x=x[:,0], y=z[0])
-    problem.parameters['rho0'] = rho0
[…]

### 2018-02-26 sims/2d_strat/strat_helper.py

@@ -1,0 +1,101 @@
+#!/usr/bin/env python
+'''
+helper function to run the shared stratification scenario. user just has to
+specify BCs and ICs
+'''
+import logging
+import numpy as np
+import matplotlib.pyplot as plt
+
+from dedalus import public as de
+from dedalus.extras.plot_tools import quad_mesh
+# from dedalus.extras.flow_tools import CFL
+
+logger = logging.getLogger(__name__)
+
+def run_strat_sim(set_bc,
+                  set_ICs,
+                  XMAX=10,
+                  ZMAX=20,
+                  N_X=64,
+                  N_Z=256,
+                  T_F=40,
+                  DT=2e-2,
+                  KX=2,
+                  KZ=1,
+                  H_FACT=3,
+                  RHO0=1,
+                  G=10):
+    H = ZMAX / H_FACT
+
+    NUM_SNAPSHOTS = 80
+
+    # Bases and domain
+    x_basis = de.Fourier('x', N_X, interval=(0, XMAX), dealias=3/2)
+    z_basis = de.Chebyshev('z', N_Z, interval=(0, ZMAX), dealias=3/2)
+    domain = de.Domain([x_basis, z_basis], np.float64)
+    x = domain.grid(0)
+    z = domain.grid(1)
+
+    problem = de.IVP(domain, variables=['P', 'rho', 'ux', 'uz'])
+    problem.meta['uz']['z']['dirichlet'] = True
+    problem.parameters['L'] = XMAX
+    problem.parameters['g'] = G
+    problem.parameters['H'] = H
+    problem.parameters['KX'] = KX
+    problem.parameters['omega'] = np.sqrt(
+        (G / H) * KX**2 / (KX**2 + KZ**2 + 0.25 / H**2))
+
+    # rho0 stratification
+    rho0 = domain.new_field()
+    rho0.meta['x']['constant'] = True
+    rho0['g'] = RHO0 * np.exp(-z / H)
+    xmesh, zmesh = quad_mesh(x=x[:, 0], y=z[0])
+    problem.parameters['rho0'] = rho0
+
+    plt.pcolormesh(xmesh, zmesh, np.transpose(rho0['g']))
+    plt.xlabel('x')
+    plt.ylabel('z')
+    plt.title('Background rho0')
+    plt.colorbar()
+    plt.savefig('strat_rho0.png')
+
+    problem.add_equation("dx(ux) + dz(uz) = 0")
[…]

### 2018-02-27 writups/notes.tex
stash

@@
 
-Since we're much more interested in the $z$ behavior, we will set the $z$ axis
-to be further away. Let's choose $\rho_0 = 1, L_x = 10, L_z = 20, k_x = 2, k_z =
-1, H = 5, g = 10$ which constrains $N^2 \equiv g/H = 1, \omega^2 = \frac{4}{5 +
-1/100}$. Some considerations:
-\begin{itemize}
-    \item $\frac{1}{k_xH}, \frac{1}{k_zH}$ are $1/10, 1/5$ respectively,
-        small-ish but $L_Z = 8H$ so we should be able to see the $e^{z/2H}$
-        behavior well.
+From here we can just tweak the parameters until we get something useful, can
+look at the videos. Note that we should simulate until at least $T >
+\frac{2L_z}{v_{p, z}}$ the $z$ phase space velocity, so we can capture any
+reflections off the boundary. Turns out both Dirichlet and Neumann BCs give
+strong reflections that produce standing waves in the $z$ and traveling waves in
+the $x$.
 
-    \item The vertical phase velocity is $\frac{\omega}{k_z} \sim 1$, so we
-        should simulate at least out to $t = 20$ for the wave to fully propagate
-        to the boundary.
-\end{itemize}
+\subsection{Supressing Reflection}
 
-\subsection{Sommerfield/Radiative BCs}
+Sourced from \url{https://people.maths.ox.ac.uk/trefethen/6all.pdf}. There are
+largely two ways to do this. The first is to add a damping zone, a region in
+which $\dot{q} = -q/\tau_d(z)$ where $\tau_d(0) = 0$ and increases to some
+non-small number for some $z_0$ beyond which we want damping.
 
-\section{(Algebra) The Anelastic/Boussinesq Approximations}
+This requires we compute a ton of extra points though. Ideally, we want to use
+BCs. Based on our explicit solution \autoref{eq:strat.full_sol} we might expect
+to write something like $\pd{u_{1z}}{z} = \frac{1/2H +
+ik_z}{\mp i\omega}\pd{u_{1z}}{z} \approx \pm \frac{k_z}{\omega}\pd{u_{1z}}{z}$.
+For an outgoing wave we expect to choose the negative sign.
+
+\section{(Algebra/Optional) The Anelastic/Boussinesq Approximations}
 
@@
 
-\chapter{2D Wave Breaking in Atmospheres}
-
-\section{Dynamical Setup}
-
-TODO (fluid equations, driven on bottom, parameters)
-
-\section{Boundary Conditions}
-
-TODO (periodic in x, show that right number of BCs in $z$, discuss whether gauge
-choice).
-
-\section{Simulation}
-
-TODO (CFL conditions etc.)
-
 \end{document}

### 2018-02-27 sims/2d_strat/strat_helper.py

@@ -14,7 +14,7 @@ from dedalus.extras.plot_tools import quad_mesh
 
 SNAPSHOTS_DIR = 'snapshots'
 
-def run_strat_sim(set_bc,
+def run_strat_sim(setup_problem,
                   set_ICs,
                   XMAX,
                   ZMAX,
@@ -63,13 +63,7 @@ def run_strat_sim(set_bc,
     plt.colorbar()
     plt.savefig('strat_rho0.png')
 
-    problem.add_equation("dx(ux) + dz(uz) = 0")
-    problem.add_equation("dt(rho) - rho0 * uz / H = 0")
-    problem.add_equation("dt(ux) + dx(P) / rho0 = 0")
-    problem.add_equation("dt(uz) + dz(P) / rho0 + rho * g / rho0 = 0")
-    set_bc(problem)
-    problem.add_bc("left(P) = 0", condition="nx == 0")
-    problem.add_bc("left(uz) = cos(KX * x - omega * t)")
+    setup_problem(problem)
 
     # Build solver
     solver = problem.build_solver(de.timesteppers.RK222)
@@ -103,3 +97,11 @@ def run_strat_sim(set_bc,
                         solver.sim_time,
                         np.mean(timesteps))
             timesteps = []
+
+def default_problem(problem):
+    problem.add_equation("dx(ux) + dz(uz) = 0")
+    problem.add_equation("dt(rho) - rho0 * uz / H = 0")
+    problem.add_equation("dt(ux) + dx(P) / rho0 = 0")
+    problem.add_equation("dt(uz) + dz(P) / rho0 + rho * g / rho0 = 0")
+    problem.add_bc("left(P) = 0", condition="nx == 0")
+    problem.add_bc("left(uz) = cos(KX * x - omega * t)")

### 2018-02-27 sims/2d_strat/strat_sommer.py

@@ -1,90 +1,0 @@
-'''
-Almost the same as strat.py, just different BC
-
-Incompressible fluid equations w/ vertical stratification
-- div u1 = 0
-- dt (rho1) - rho0 * u1z / H = 0
-- dt(u1) = -grad(P1) / rho0 - rho1 * g / rho0
-
-Periodic BC in x, radiative at z=L, Driving term at z=0
-'''
-import logging
-import numpy as np
-import matplotlib.pyplot as plt
-
-logger = logging.getLogger(__name__)
-from dedalus import public as de
-from dedalus.extras.plot_tools import quad_mesh, pad_limits
-
-XMAX = 10
-ZMAX = 20
-N_X = 64
-N_Z = 256
-DX = XMAX / N_X
-DZ = ZMAX / N_Z
-
-T_F = 200
-DT = 2e-2
-KX = 2
-KZ = 1
-H = 5
-G = 10
-N = G / H
-RHO0 = 1
-
-NUM_SNAPSHOTS = 800
-
-if __name__ == '__main__':
-    # Bases and domain
-    x_basis = de.Fourier('x', N_X, interval=(0, XMAX), dealias=3/2)
-    z_basis = de.Chebyshev('z', N_Z, interval=(0, ZMAX), dealias=3/2)
-    domain = de.Domain([x_basis, z_basis], np.float64)
-
-    problem = de.IVP(domain, variables=['P', 'rho', 'ux', 'uz'])
-    problem.meta['uz']['z']['dirichlet'] = True
-    problem.parameters['rho0'] = RHO0
-    problem.parameters['L'] = XMAX
-    problem.parameters['g'] = G
-    problem.parameters['KX'] = KX
-    problem.parameters['H'] = H
-    problem.parameters['N'] = N
-    problem.parameters['omega'] = np.sqrt(
-        N**2 * KX**2 / (KX**2 + KZ**2 + 0.25 / H**2))
-    problem.add_equation("dx(ux) + dz(uz) = 0")
-    problem.add_equation("dt(rho) - rho0 * uz / H = 0")
-    problem.add_equation("dt(ux) + dx(P)/rho0 = 0")
-    problem.add_equation("dt(uz) + dz(P)/rho0 + rho * g / rho0= 0")
-    problem.add_bc("right(dt(uz) * N * KX / omega ** 2 - dz(uz)) = 0",
-                   condition="nx != 0")
[…]

### 2018-03-05 writups/notes.tex
regenerate png, add some notes

@@
 
+\subsection{Extracting Information from the Waveform}
+
+Let's make sure we can extract information from the linearized problem. Notable
+things we should be interested in: phase/group velocity, energy density and
+energy flux.
+
+Let's first consider the phase velocity. We traditionally think about the phase
+velocity in 1D $v_{ph} = \frac{\omega(k)}{k}$, but in general it is the function
+such that $\vec{k} \cdot \vec{r} - \omega t = \vec{k} \cdot \p*{\vec{r} -
+\vec{v}_{ph}t}$ is constant (the phase of the wave). Thus, it must satisfy
+$\vec{k} \cdot \vec{v}_{ph} = \omega$, and so one sensible choice is
+$\vec{v}_{ph} = \frac{\omega \hat{k}}{\abs{\vec{k}}} = \frac{\omega
+\vec{k}}{\abs{\vec{k}}^2}$. Thus, if we have some waveform $f(x, z, t)$, we can
+estimate its phase velocity $\vec{v}_{ph}(x, z, t)$ by
+\begin{equation}
+    \vec{v}_{ph}(x, z, t) = \frac{\partial_t f
+        \vec{\nabla}f}{\abs{\vec{\nabla}f}^2}.\label{eq:v_ph}
+\end{equation}
+This can be validated in the $f(x, z, t) \propto e^{i(\vec{k} \cdot \vec{x} -
+\omega t)}$ case to give $\vec{v}_{ph} = \frac{\omega \vec{k}}{k^2}$. Note the
+sign in \autoref{eq:v_ph}
+
 \subsection{Supressing Reflection}
@@
 which $\dot{q} = -q/\tau_d(z)$ where $\tau_d(0) = 0$ and increases to some
-non-small number for some $z_0$ beyond which we want damping.
+non-small number for some $z_0$ beyond which we want damping. One form, chosen
+in Ryan and Dong's paper, is to use multiplicative factor $f(z) = \max\s*{0, 1 -
+\frac{\p*{z - z_b}^2}{\p*{z_d - z_b}^2}}$ where $z_b$ is where one begins
+supression and $z_b$ is the boundary. Then a dynamical variable $q$ can be
+supressed via $\dot{q} = (\dots) - f(z)\frac{q - q_0}{\tau}$ where $\tau$ is the
+dynamical timescale on which to damp and $q_0$ is the value to damp to.
 

### 2018-03-06 writups/notes.tex
stash

@@
 
+\subsection{Waveform Properties, Analytical}
+
+Let's figure out the analytical forms for the phase, group velocity and the
+energy density/power flux. We will try to recover these from waveforms.
+
+For our stratified atmosphere problem, $\omega^2 = \frac{N^2 k_x^2}{k^2 +
+\frac{1}{4H^2}}$. Plugging in our parameters $k_z = -k_x = 2pi, N^2 = 10$ we can
+find $\omega \approx 2.23\;\mathrm{Hz}$, so $\vec{v}_{ph} \approx
+0.1777\p*{-\hat{x} + \hat{z}}\;\mathrm{s^{-1}}$.
+
+Since we use a height of $5$, we should expect the initial disturbance to reach
+the top boundary at $t \sim 30\;\mathrm{s}$.
+
+On the other hand the group velocity is given $\vec{v}_g =
+\pd{\omega}{k_i}\hat{\imath}$. In our problem, $v_{g, z} =
+-\frac{N^2k_x^2k_z}{\p*{k_x^2 + k_z^2 + 1/4H^2}^{3/2}}$ while $v_{g, x} =
+\frac{2N^2k_x}{\sqrt{k_x^2 + k_z^2 + 1/4H^2}} - \frac{N^2k_x^3}{\p*{k_x^2 +
+k_z^2 + 1/4H^2}^{3/2}}$.
+
+We should note that the kinetic energy density is just $E = \frac{1}{2}\rho
+v^2$, the kinetic energy. Using then $\rd{E}{t} = 0$ shows that the
+corresponding power flux is $\vec{P} = \frac{\vec{v}}{2}\rho v^2$.
+
 \subsection{Extracting Information from the Waveform}
@@
 This requires we compute a ton of extra points though. Ideally, we want to use
-BCs. Based on our explicit solution \autoref{eq:strat.full_sol} we might expect
-to write something like $\pd{u_{1z}}{z} = \frac{1/2H +
-ik_z}{\mp i\omega}\pd{u_{1z}}{z} \approx \pm \frac{k_z}{\omega}\pd{u_{1z}}{z}$.
-For an outgoing wave we expect to choose the negative sign.
+BCs. TODO
 

### 2018-03-06 writups/notes.tex

@@
 
+Let's first consider the phase velocity. We traditionally think about the phase
+velocity in 1D $v_{ph} = \frac{\omega(k)}{k}$, but in general it is the function
+such that $\vec{k} \cdot \vec{r} - \omega t = \vec{k} \cdot \p*{\vec{r} -
+\vec{v}_{ph}t}$ is constant (the phase of the wave). Thus, it must satisfy
+$\vec{k} \cdot \vec{v}_{ph} = \omega$, and so one sensible choice is
+\begin{equation}
+    \vec{v}_{ph} = \frac{\omega \hat{k}}{\abs{\vec{k}}}
+        = \frac{\omega \vec{k}}{\abs{\vec{k}}^2}.
+        \label{eq:v_ph}
+\end{equation}
+
 For our stratified atmosphere problem, $\omega^2 = \frac{N^2 k_x^2}{k^2 +
@@
 
-Let's first consider the phase velocity. We traditionally think about the phase
-velocity in 1D $v_{ph} = \frac{\omega(k)}{k}$, but in general it is the function
-such that $\vec{k} \cdot \vec{r} - \omega t = \vec{k} \cdot \p*{\vec{r} -
-\vec{v}_{ph}t}$ is constant (the phase of the wave). Thus, it must satisfy
-$\vec{k} \cdot \vec{v}_{ph} = \omega$, and so one sensible choice is
-$\vec{v}_{ph} = \frac{\omega \hat{k}}{\abs{\vec{k}}} = \frac{\omega
-\vec{k}}{\abs{\vec{k}}^2}$. Thus, if we have some waveform $f(x, z, t)$, we can
-estimate its phase velocity $\vec{v}_{ph}(x, z, t)$ by
+Based on our phase velocity expression \autoref{eq:v_ph}, if we have some
+waveform $f(x, z, t)$, we can estimate its phase velocity $\vec{v}_{ph}(x, z,
+t)$ by
 \begin{equation}

### 2018-03-21 sims/2d_strat/strat.py
check in agree_plots

@@ -19,7 +19,7 @@ num_timesteps = 4e4
 XMAX = H
 ZMAX = 5 * H
 KX = -2 * np.pi / H
-KZ = 2 * np.pi / H
+KZ = (np.pi / 2) * np.pi / H
 G = 10
 OMEGA = strat_helper.get_omega(G, H, KX, KZ)
 VPH_X, VPH_Z = strat_helper.get_vph(G, H, KX, KZ)
@@ -53,11 +53,8 @@ def dirichlet_bc(problem, *_):
     strat_helper.default_problem(problem)
     problem.add_bc('right(uz) = 0', condition='nx != 0')
 
-def neumann_bc(problem, *_):
-    strat_helper.default_problem(problem)
-    problem.add_bc('right(dz(uz)) = 0', condition='nx != 0')
-
 def rad_bc(problem, *_):
+    # TODO incorrect
     strat_helper.default_problem(problem)
     problem.add_bc('right(dt(uz) + omega / KZ * dz(uz)) = 0',
                    condition='nx != 0')
@@ -108,9 +105,8 @@ def run(bc, ic, name, params_dict):
 
 if __name__ == '__main__':
     tasks = [
-        (dirichlet_bc, zero_ic, 'd0', build_interp_params(16, 2)),
-        (neumann_bc, zero_ic, 'n0', build_interp_params(16, 4)),
-        (sponge, zero_ic, 'sponge2', build_interp_params(16, 4)),
+        (dirichlet_bc, zero_ic, 'd0', build_interp_params(8, 2)),
+        (sponge, zero_ic, 'sponge2', build_interp_params(8, 4)),
     ]
 
     with Pool(processes=N_PARALLEL) as p:

### 2018-03-21 sims/2d_strat/strat_helper.py

@@ -146,9 +146,9 @@ def load(setup_problem,
          RHO0,
          INTERP_X,
          INTERP_Z,
-         dyn_vars,
          name=None,
          **_):
+    dyn_vars = ['uz', 'ux', 'rho', 'P']
     snapshots_dir = SNAPSHOTS_DIR % name
     filename = '{s}/{s}_s1/{s}_s1_p0.h5'.format(s=snapshots_dir)
 
@@ -181,8 +181,6 @@ def load(setup_problem,
     rho0 = RHO0 * np.exp(-z / H)
     state_vars['E'] = ((rho0 + state_vars['rho']) *
                        (state_vars['ux']**2 + state_vars['uz']**2)) / 2
-    state_vars['dE_t'] = np.gradient(state_vars['E'])[0] / DT
-    state_vars['P_x'] = state_vars['E'] * state_vars['ux']
     state_vars['P_z'] = state_vars['E'] * state_vars['uz']
     return sim_times, domain, state_vars
 
@@ -207,11 +205,15 @@ def plot(setup_problem,
     snapshots_dir = SNAPSHOTS_DIR % name
     path = '{s}/{s}_s1'.format(s=snapshots_dir)
     matplotlib.rcParams.update({'font.size': 6})
-    dyn_vars = ['uz', 'ux', 'rho', 'P']
-    z_vars = ['E', 'dE_t', 'P_x', 'P_z'] # sum these over x
+    plot_vars = ['uz', 'ux', 'rho', 'P', 'P_z']
+    z_vars = ['E'] # sum these over x
     n_cols = 3
-    n_rows = 3
-    plot_stride = 2
+    n_rows = 2
+    plot_stride = 1
+
+    if os.path.exists('%s.mp4' % name):
+        print('%s.mp4 already exists, not regenerating' % name)
+        return
 
     sim_times, domain, state_vars = load(setup_problem,
                                          XMAX=XMAX,
@@ -228,7 +230,6 @@ def plot(setup_problem,
                                          RHO0=RHO0,
                                          INTERP_X=INTERP_X,
                                          INTERP_Z=INTERP_Z,
-                                         dyn_vars=dyn_vars,
                                          name=name)
 
     x = domain.grid(0, scales=INTERP_X)
@@ -242,8 +243,8 @@ def plot(setup_problem,
         fig = plt.figure(dpi=200)
 
         idx = 1
-        for var in dyn_vars:
[…]

### 2018-03-23 presentations/Su_GM_032318.tex
check in some agree plots, non-compiling presentation

--- 2018-03-05 writups/notes.tex
@@
-    \documentclass[11pt,
-        usenames, % allows access to some tikz colors
-        dvipsnames % more colors: https://en.wikibooks.org/wiki/LaTeX/Colors
-    ]{report}
+    \documentclass[dvipsnames]{beamer}
+    \usetheme{Madrid}
+    \usefonttheme{professionalfonts}
     \usepackage{
@@
     }
-    \usepackage[margin=1in]{geometry} % page geometry
     \usepackage[
@@
     \newcommand*{\at}[1]{\left.#1\right|}
-
-    \newtheorem{theorem}{Theorem}[section]
 
@@
 
-\def\Snospace~{\S{}} % hack to remove the space left after autorefs
-\renewcommand*{\sectionautorefname}{\Snospace}
-\renewcommand*{\appendixautorefname}{\Snospace}
-\renewcommand*{\figureautorefname}{Fig.}
-\renewcommand*{\equationautorefname}{Eq.}
-\renewcommand*{\tableautorefname}{Tab.}
-
-\onehalfspacing
-
 \pagestyle{fancy}
@@
 
-\title{Research Notes}
-\author{Yubo Su}
-\rhead{}
+\title[Exciting Internal Gravity Waves]{Exciting Internal Gravity Waves, Toy
+Problem}
+\subtitle[]{Mar 23, 2018 Group Meeting}
+\author[Yubo Su]{Yubo Su}
+\date{Mar 23, 2018}
 
-\maketitle
+\frame{\titlepage}
 
-\tableofcontents
+\begin{frame}
+    \frametitle{Problem Setup}
+    \framesubtitle{Problem Statement}
+    \begin{itemize}
+        \item 2D hydrodynamics with driving boundary condition, produce gravity
+            waves.
+        \item Begin with linear incompressible case, equations:
+        \begin{subequations}\label{se:eom}
+            \begin{align}
+                \pd{\rho_1}{t} &= 0,\\
+                \vec{\nabla} \cdot \vec{u}_1 &= 0,\\
+                \pd{\vec{u}_1}{t} &= -\frac{\vec{\nabla}P_1}{\rho_0} + \vec{g}.
+            \end{align}
+        \end{subequations}
 
-\newpage
+        \item Consider stratified atmosphere $\rho_0(x, z, t) = \rho_0
+            e^{-z/H}$.
+    \end{itemize}
+\end{frame}
 
-\chapter{Preliminary Problems}
+\begin{frame}
+    \frametitle{Problem Setup}
+    \framesubtitle{Boundary Conditions}
 
-To get an intuition for how Dedalus and fluid mechanics works, we will solve
-some toy problems. Recall fluid equations in the presence of a uniform
-gravitational field $\vec{g} = -g\hat{z}$:
+    \begin{itemize}
+        \item Four variables ($\rho_1, P_1, u_{1x}, u_{1z}$), all first order in
+            equations of motion, four boundary conditions.
+
+        \item Periodic BCs in $x$ (2).
+
+        \item Wavelike solutions $u_{1z} \propto e^{z/2H}e^{i\p*{k_xx + k_zz -
+            \omega t}}$, $u_{1x}, P_1, \rho_1$ similar up to phase and terms
+            $\sim \mathcal{O}\p*{H^{-1}}$.
+
+        \item Driven bottom BC\@: $u_{1z}(x, 0, t) = A\cos \p*{k_xx + \omega t}$
+            (1).
+    \end{itemize}
+\end{frame}
+
+\begin{frame}
+    \frametitle{Problem Setup}
+    \framesubtitle{Top Boundary Condition}
+
+    \begin{itemize}
+        \item Top BC can be chosen Dirichlet $u_{1z}(x, L_z, t) = 0$,
+            analytically tractable.
+
+        \item Driving term will pump energy into the system, without dissipation
+            becomes like driven undamped SHO\@. In full nonlinear problem not an
+            issue\dots
+
+        \item More realistically, two solutions: radiative boundary conditions
+            and damping zone.
+
+        \item We use damping zone, add term $\rd{}{t}\vec{q} = L\vec{q}
+            - f(z)\p*{\vec{q} - \vec{q}_0}$ where
             \begin{equation}
-    \begin{split}
-        \rd{\rho}{t} + \vec{\nabla} \cdot \vec{u} &= 0,\\
-        \rd{\vec{u}}{t} + \frac{\vec{\nabla}P}{\rho} - \vec{g} &= 0.
-    \end{split}\label{eq:fluid_eq}
+                f(z) =
+                \begin{cases}
+                    \Gamma \s*{
+                        1 - \frac{\p*{z - L_z}^2}
+                            {\p*{z_{damp} - L_z}^2}
+                    }
+                        & z > z_{damp}\\
+                    0 & z < z_{damp}
+                \end{cases}.
             \end{equation}
+    \end{itemize}
+\end{frame}
 
-In the incompressible limit, $\rd{\rho}{t} = 0$, which implies $\vec{\nabla}
-\cdot \vec{u} = 0$. We use subscripts to indicate perturbed quantities, $Q_0$ is
-background and $Q_1$ is perturbed. We will generally use $\vec{u}_0 = 0$ unless
-otherwise noted. We will also generally assume symmetry along all axes except
-$z$ the vertical axis.
+\begin{frame}
+    \frametitle{Results}
+    \framesubtitle{Problem Statement Summary}
 
-In the incompressible limit, the fluid equations become
-\begin{equation}
-    \begin{split}
-        \vec{\nabla} \cdot \vec{u}_1 &= 0,\\
-        \pd{\rho_1}{t} + u_{1z}\pd{\rho_0}{z} &= 0,\\
-        \pd{\vec{u}_1}{t} + \frac{1}{\rho_0}\vec{\nabla}P_1
-            + \frac{\rho_1 g \hat{z}}{\rho_0} &= 0.
-    \end{split}\label{eq:lin.incomp}
-\end{equation}
-We have used $\vec{\nabla}P_0 = -\rho_0 g \hat{z}$ in the absence of
-perturbations.
+    \begin{itemize}
+        \item Linear incompressible hydrodynamics with periodic BCs in $x$,
+            driving term at $z = 0$ and either Dirichlet or reflection-free BCs
+            at $z = L_z$.
 
-\section{Incompressible, No Gravity}
+        \item Spectral code Dedalus, built-in Runge-Kutta timestepper (simplest
+            to characterize).
+    \end{itemize}
+\end{frame}
 
-We note that in the no gravity limit that $\rho_1$ does not have an effect on
-other dynamical variables, so the equations of motion we must solve are
-\begin{equation}
-    \begin{split}
-        \vec{\nabla} \cdot \vec{u}_1 &= 0,\\
-        \pd{\vec{u}_1}{t} + \frac{\vec{\nabla}P_1}{\rho_0} &= 0.
-    \end{split}\label{eq:lin.no_g_eom}
-\end{equation}
-We can take the divergence of the momentum equation and substitute the
-continuity equation to get $\nabla^2 P = 0$.
+\begin{frame}
+    \frametitle{Results}
+    \framesubtitle{Waveform Agreement}
 
-\subsection{Dirichlet BCs}
+    \begin{figure}[!h]
+        \centering
+        \begin{subfigure}{0.4\textwidth}
+            \centering
+            \includegraphics[width=\textwidth]{../sims/2d_strat/agree_plots/sponge_0.png}
+        \end{subfigure}
+        \begin{subfigure}{0.4\textwidth}
+            \centering
+            \includegraphics[width=\textwidth]{../sims/2d_strat/agree_plots/sponge_1.png}
+        \end{subfigure}
 
-This is a Laplace equation, which we've solved countless times. Imposing
-periodic boundary conditions in the $x$ direction and $P_1(z = L) = 0, P_1(z =
-0) = \mathcal{P}(x, t)$, we obtain eigenfunctions
-\begin{equation}
-    \begin{split}
-        P_{1, n}(x, z, t) &=
-            \frac{\mathcal{P}_n(t)}{\sinh(k_nL)}e^{ik_nx}\sinh\p*{k_n(L - z)},\\
-        u_{1x, n}(x, z, t) &=
-            \int\limits^t -\frac{1}{\rho_0}\pd{P_{1, n}}{x}\;\mathrm{d}t,\\
-        u_{1z, n}(x, z, t) &=
-            \int\limits^t -\frac{1}{\rho_0}\pd{P_{1, n}}{z}\;\mathrm{d}t.
-    \end{split}\label{eq:incomp_nog}
-\end{equation}
-We define $k_n = \frac{2\pi n}{L}, n \geq 0$ and $\mathcal{P}(x, t) =
-\sum\limits_n \mathcal{P}_n(t)e^{ik_nx}$.
+        \begin{subfigure}{0.4\textwidth}
+            \centering
+            \includegraphics[width=\textwidth]{../sims/2d_strat/agree_plots/sponge_2.png}
+        \end{subfigure}
+        \begin{subfigure}{0.4\textwidth}
+            \centering
+            \includegraphics[width=\textwidth]{../sims/2d_strat/agree_plots/sponge_3.png}
+        \end{subfigure}
+        \caption{Agreement of numerical and analytical solution at four adjacent
+        snapshots.}\label{fig:sponge_agree}
+    \end{figure}
+\end{frame}
 
-Thus, if we impose BCs $\mathcal{P}(x, t) = \sin \frac{2\pi x}{L}$ and start
-with initial conditions such that all quantities are zero, we would expect after
-transients die out that
-\begin{equation}
-    \begin{split}
-        P(x, z, t) &= \frac{\sin \frac{2\pi x}{L}}{\sinh 2\pi}
-            \sinh \p*{2\pi \frac{L - z}{L}},\\
-        u_{1x}(x, z, t) &= -\frac{2\pi t}{L\rho_0}
-            \frac{\cos \frac{2\pi x}{L}}{\sinh 2\pi}\sinh \p*{2\pi \frac{L -
-                z}{L}},\\
-        u_{1z}(x, z, t) &= +\frac{2\pi t}{L\rho_0}
-            \frac{\sin \frac{2\pi z}{L}}{\sinh 2\pi}\cosh \p*{2\pi \frac{L -
-                z}{L}}.
-    \end{split}\label{eq:incomp_nog_sol}
-\end{equation}
-This is in good agreement with the results, presented in \autoref{fig:no_g}.
-Note that $P$ is constant while $\vec{u}$ increases linearly in time, and we
-observe the expected $\sim \sin x \sinh \frac{L - z}{z}$ dependence. In fact,
-$u_{1x}, u_{1z}$ are exactly $\frac{2\pi}{10}$ at $t = 1$.
+\begin{frame}
+    \frametitle{Results}
+    \framesubtitle{Convergence Tests}
+
+    \begin{itemize}
+        \item $RMS = \frac{\sqrt{\sum\limits_{x, z, t} \p*{u_{1z}^{(N)}(x, z, t)
+            - u_{1z}^{(2N)}(x, z, t)}^2}}{\sqrt{
+            \sum\limits_{x, z, t} \p*{u_{1z}^{(N)}(x, z, t)}^2}}$
+    \end{itemize}
     \begin{figure}[!h]
@@
             \centering
-        \includegraphics[width=\textwidth]{../sims/2d_no_g/no_g_P_t50.png}
+            \includegraphics[width=\textwidth]{../sims/2d_strat_conv/t_conv.png}
         \end{subfigure}
@@
             \centering
-        \includegraphics[width=\textwidth]{../sims/2d_no_g/no_g_ux_t50.png}
-    \end{subfigure}
-    \begin{subfigure}{0.3\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{../sims/2d_no_g/no_g_uz_t50.png}
+            \includegraphics[width=\textwidth]{../sims/2d_strat_conv/x_conv.png}
         \end{subfigure}
@@
             \centering
-        \includegraphics[width=\textwidth]{../sims/2d_no_g/no_g_P_t100.png}
+            \includegraphics[width=\textwidth]{../sims/2d_strat_conv/z_conv.png}
         \end{subfigure}
-    \begin{subfigure}{0.3\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{../sims/2d_no_g/no_g_ux_t100.png}
-    \end{subfigure}
-    \begin{subfigure}{0.3\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{../sims/2d_no_g/no_g_uz_t100.png}
-    \end{subfigure}
-    \caption{$P, u_x, u_z$ at $t = 0.5$ and $t = 1$ for $\rho_0 = 1$. We choose
-    a $L = 10$ square domain.}\label{fig:no_g}
+        \caption{Convergences of time stepsize and grid resolution. Note the
+        power law convergence in timestep but exponential in $z$ points (since
+        $x$ is always monochromatic, increasing resolution does little).}\label{fig:convsv}
     \end{figure}
-
-\subsection{Things to Note}
-
-It is worth noting that, since our \autoref{eq:lin.no_g_eom} reduced to a Laplace
-equation, we needed two $z$ BCs and two $x$ BCs (periodic BCs amount to equating
-the value and derivative of the function). This is in agreement with the
-observation that the original \autoref{eq:lin.no_g_eom} had two derivatives in
-$x, z$ apiece, so we needed two BCs each.
-
-It is also worth seeing from our solution that $P$ immediately goes to the
-equilibrium solution. This is not surprising since in the incompressible limit,
-sound speed goes to infinity which is the timescale on which the pressure field
-adjusts to net forces. Thus, the dynamics solely arise from the static pressure
-field pushing the velocity field to equilibrium.
-
-We chose time-independent $\mathcal{P}(x, t)$, but it is clear that whatever
-$\mathcal{P}(x, t)$ we choose, the time dependence propagates to the velocities
-by way of an integral. If we had instead chosen to take Fourier transform $t \to
-\omega$, we would have had to integrate the boundary condition against the
-eigenfunctions for each of the $\omega$, which is still easily computable, to
-get the full $u_{1x}(x, z, t)$. We consider this in preparation for when we can
-only solve for a set of $\vec{k}, \omega$ in the next problem.
-
-The next thing we would have wanted to do is solve a problem with radiative BCs,
-but we need to have wave solutions, which are missing in the absence of gravity.
-We thus move on to the next configuration.
-
-\section{Incompressible, Stratified w/ Gravity}
-
-\subsection{Eigenfunctions}
-
-Let's restore the $\rho_1 g$ term now. For funsies, we begin by solving for
-arbitrary stratification $\rho_0(z)$ first. The fluid equations to first order
-reduce to
-\begin{equation}
-    \begin{split}
-        \pd{\rho_1}{t} + \p*{\vec{u}_1 \cdot \vec{\nabla}}\rho_0 &= 0,\\
-        \vec{\nabla} \cdot \vec{u}_1 &= 0,\\
-        \pd{\vec{u}_1}{t} &= -\frac{\vec{\nabla}P_1}{\rho_0}
-            - \frac{\rho_1 g}{\rho_0}
-    \end{split}\label{eq:lin_incomp}
-\end{equation}
-We expect there to be some $z$ dependence in the amplitude, so we substitute
-variables of form $e^{i(kx - \omega t)}$ and do not specify the $z$ dependence.
-This gives us
-\begin{equation}
-    \begin{split}
-        -i\omega \rho_1 - u_{1z}\pd{\rho_0}{z} &= 0,\\
-        iku_{1x} + \pd{u_{1z}}{z} &= 0,\\
-        -iw u_{1x} + \frac{ik_x P_1}{\rho_0} &= 0,\\
-        -iw u_{1z} + \frac{1}{\rho_0}\pd{P_1}{z} +
-            \frac{\rho_1 g}{\rho_g} &= 0.
-    \end{split}
-\end{equation}
-We substitute $N^2 = -\frac{g}{\rho_0}\pd{\rho_0}{z}$ to obtain
-\begin{subequations}
-    \begin{align}
-        -i\omega \rho_1 - u_{1z}\frac{\rho_0N^2}{g} &= 0,\label{eq:lin.cont}\\
-        iku_{1x} + \pd{u_{1z}}{z} &= 0,\label{eq:lin.div}\\
-        -iw u_{1x} + \frac{ik_x P_1}{\rho_0} &= 0,\label{eq:lin.mom_x}\\
-        -iw u_{1z} + \frac{1}{\rho_0}\pd{P_1}{z} +
-            \frac{\rho_1g}{\rho_0} &= 0.\label{eq:lin.mom_z}
-    \end{align}
-\end{subequations}
-Eliminating $u_{1x}$ by substituting~\eqref{eq:lin.div}
-into~\eqref{eq:lin.mom_x} and $\rho_1$ by substituting~\eqref{eq:lin.cont}
-into~\eqref{eq:lin.mom_z} give
-\begin{subequations}
-    \begin{align}
-        i\omega \pd{u_{1z}}{z} + \frac{k_x^2 P_1}{\rho_0} &= 0,
-            \label{eq:lin.eq1}\\
-        \p*{\omega^2 - N^2 }u_{1z} + \frac{i\omega}{\rho_0}\pd{P_1}{z} &= 0.
-            \label{eq:lin.eq2}
-    \end{align}
-\end{subequations}
-Finally, we multiply~\eqref{eq:lin.eq1} with $\rho_0$ and differentiate
-$\mathrm{d}z$ and combine with~\eqref{eq:lin.eq2} to give
-\begin{equation}
-    \rtd{u_{1z}}{z} + \frac{1}{\rho_0}\pd{\rho_0}{z}\pd{u_{1z}}{z}
-        + k_x^2\p*{\frac{N^2}{\omega^2} - 1}u_{1z} = 0.\label{eq:lin.incomp.gen}
-\end{equation}
-
-Let's now pick stratification $\rho \propto e^{-z/H}$
-\autoref{eq:lin.incomp.gen} clearly has exponential solutions $e^{\kappa z}$ for
-\begin{equation}
-    \kappa^2 - \frac{\kappa}{H} + k_x^2\p*{\frac{N^2}{\omega^2} - 1} = 0.
-\end{equation}
-We permit complex $\kappa = \frac{1}{2H} + ik_z$, and from the above clearly
-\begin{align}
-    k_z^2 &= -\frac{1}{4H^2} + k_x^2\p*{\frac{N^2}{\omega^2} - 1},\nonumber\\
-    \omega^2 &= \frac{N^2k_x^2}{k_x^2 + k_z^2 +
-        \frac{1}{4H^2}}.\label{eq:strat_disp}
-\end{align}
-Thus the eigenfunctions are
-\begin{equation}
-    \begin{split}
-        u_{1z} &= e^{z/2H} e^{i(k_zz + k_xx - \omega t)},\\
-        u_{1x} &= -\frac{k_z + i/2H}{k_x} u_{1z},\\
-        \rho_1 &= \frac{i \rho_0}{H\omega} u_{1z},\\
-        P_1 &= -\frac{\rho_0 \omega}{k_x^2}\p*{k_z + i/2H}u_{1z}.
-    \end{split}\label{eq:strat_eigens}
-\end{equation}
-
-\subsection{Analytically Solving an IVP, Dirichlet + Driving BCs}
-
-We will analyze everything in terms of $u_{1z}$ since it has the simplest form;
-note that when actually choosing the BCs we will have to consider the gauge
-freedom of $P$ and some considerations we defer to the computational section.
-
-Currently, we have a set of eigenfunctions
-\begin{equation}
-    u_{1z}\p*{x, z, t \middle| \vec{k}, \omega}
-        = e^{z/2H} e^{i(k_zz + k_xx - \omega t)}.
-        \label{eq:strat.uz_eigen}
-\end{equation}
-Note that $u_{1z}$ is really only a function of two parameters rather than the
-three $(k_x, k_z, \omega)$, since the three are related by dispersion relation
-\autoref{eq:strat_disp}.
-
-Now, we implement BCs. Consider domain $x, z \in [0, L]$. We will use periodic
-BCs again in $x$, so then $k_{x, n} = \frac{2\pi n}{L}, n \geq 0$. Then, we will
-require $u_{1z, n}(x, L, t) = 0$, a Dirichlet condition at the top boundary,
-which restricts us to eigenfunctions of form
-\begin{equation}
-    u_{1z, n}\p*{x, z, t \middle| \vec{k}, \omega}
-        = e^{z/2H}e^{i(k_xx - \omega t)}\sin\p*{k_z\p*{L - z}}.
-        \label{eq:strat.uz_eigen_l-z}
-\end{equation}
-
-Finally, we must choose a BC at $z = 0$. We will choose a general function
-$u_{1z}(x, 0, t) = F(x, t)$ where we can decompose
-\begin{equation}
-    F(x, t) = \int\limits \sum\limits_n \mathcal{F}(k_{x, n}, \omega)
-        e^{i(k_{x, n}x - \omega t)}\;\mathrm{d}\omega.
-\end{equation}
-Matching BCs then gives us general solution for $u_{1z}$ given an arbitrary
-driving function
-\begin{equation}
-    u_{1z}\p*{x, z, t \middle| \vec{k}, \omega}
-        = \int\limits \sum\limits_n \mathcal{F}\p*{k_{x, n}, \omega}
-            \frac{e^{z/2H}e^{i\p*{k_{x, n}x - \omega t}}\sin \p*{k_z(L - z)}}{
-            \sin k_zL}\;\mathrm{d}\omega.
-            \label{eq:strat.full_sol}
-\end{equation}
-
-For ease of computation, let's pick $F(x, t) = \cos \p*{\frac{2 \pi x}{L} -
-\omega_0 t}$, so our full expected solution is (note $A + \epsilon =
-Ae^{\epsilon/A} + \mathcal{O}(\epsilon^2)$)
-\begin{equation}
-    \begin{split}
-        u_{1z}\p*{x, z, t \middle | \vec{k}, \omega}
-            &= e^{z/2H}\frac{\cos\p*{\frac{2\pi x}{L} - \omega_0 t}
-                \sin \p*{k_z(L - z)}}{\sin k_zL},\\
-        u_{1x}\p*{x, z, t \middle | \vec{k}, \omega}
-            &\approx -\frac{k_z}{k_x} e^{z/2H}\frac{
-                \cos\p*{\frac{2\pi x}{L} - \omega_0 t + \frac{1}{2Hk_z}}
-                \sin \p*{k_z(L - z)}
-            }{\sin k_zL},\\
-        \rho_1\p*{x, z, t \middle | \vec{k}, \omega}
-            &\approx \frac{\rho_0}{H\omega} e^{z/2H}\frac{
-                -\sin\p*{\frac{2\pi x}{L} - \omega_0 t}
-                \sin \p*{k_z(L - z)}
-            }{\sin k_zL},\\
-        P_1\p*{x, z, t \middle | \vec{k}, \omega}
-            &\approx -\frac{\rho_0 \omega k_z}{k_x^2} e^{z/2H}\frac{
-                -\cos\p*{\frac{2\pi x}{L} - \omega_0 t + \frac{1}{2Hk_z}}
-                \sin \p*{k_z(L - z)}
-            }{\sin k_zL},\\
-    \end{split}\label{eq:strat.part_sol}
-\end{equation}
-where $k_z: \omega(k_x, k_z) = \omega_0$ by the dispersion relation
-\autoref{eq:strat_disp}. Note that $H$ contributes both to the overall
-exponential profile and to the phase lag of $u_{1x}$.
-
-\subsection{Computationally Solving an IVP, Dirichlet + Driving BCs}
-
-To solve this computationally with the aforementioned BCs, periodic in $x$,
-Dirichlet $0$ at $z = L$ and $\cos \p*{\frac{2\pi x}{L} - \omega_0 t}$ at $z =
-0$, we must address the gauge freedom in $P$. This arises because for $k_x = 0$,
-the divergence-free condition $\vec{\nabla} \cdot \vec{u}_1 = \pd{u_z}{z} = 0$
-specificies $u_{1z}$ up to a constant already, so the bottom BC will fix the
-value of $u_z$ when $k_x = 0$.
-
-A different way of phrasing the same argument is as follows. Consider the
-discrete $N \times N$ (square for notational simplicity) grid. At the boundary,
-there is a list of values $f\p*{\z*{x_i}, z_0}$ that lives in an $N$ dimensional
-space. We can thus pick a spanning set of $N$ basis vectors, and for each of
-these $N$ vectors, by enforcing $\vec{\nabla} \cdot \vec{u} = 0$ at the boundary
-we fix the allowed $f\p*{\z*{x_i}, z_{-1}}$ boundary conditions we can implement.
-But there exists a choice of basis vectors for which one of the basis vectors is
-constant $f_i\p*{\z*{x_i}, z_0} = C$. For this basis vector, the boundary
-condition is fully determined, so we only have $N - 1$ dimensions from which to
-choose the BCs $f\p*{\z*{x_i}, z_{-1}}$. Since the dimensionality of a space
-cannot depend on the choice of basis vector, the divergence free condition
-actually yields an extra degree of freedom.
-
-We can use this extra degreee of freedom to specify $P(z = L) = 0$ so the
-oscillations should have zero mean. Then we can just simulate away!
-
-From here we can just tweak the parameters until we get something useful, can
-look at the videos. Note that we should simulate until at least $T >
-\frac{2L_z}{v_{p, z}}$ the $z$ phase space velocity, so we can capture any
-reflections off the boundary. Turns out both Dirichlet and Neumann BCs give
-strong reflections that produce standing waves in the $z$ and traveling waves in
-the $x$.
-
-\subsection{Extracting Information from the Waveform}
-
-Let's make sure we can extract information from the linearized problem. Notable
-things we should be interested in: phase/group velocity, energy density and
-energy flux.
-
-Let's first consider the phase velocity. We traditionally think about the phase
-velocity in 1D $v_{ph} = \frac{\omega(k)}{k}$, but in general it is the function
-such that $\vec{k} \cdot \vec{r} - \omega t = \vec{k} \cdot \p*{\vec{r} -
-\vec{v}_{ph}t}$ is constant (the phase of the wave). Thus, it must satisfy
-$\vec{k} \cdot \vec{v}_{ph} = \omega$, and so one sensible choice is
-$\vec{v}_{ph} = \frac{\omega \hat{k}}{\abs{\vec{k}}} = \frac{\omega
-\vec{k}}{\abs{\vec{k}}^2}$. Thus, if we have some waveform $f(x, z, t)$, we can
-estimate its phase velocity $\vec{v}_{ph}(x, z, t)$ by
-\begin{equation}
-    \vec{v}_{ph}(x, z, t) = \frac{\partial_t f
-        \vec{\nabla}f}{\abs{\vec{\nabla}f}^2}.\label{eq:v_ph}
-\end{equation}
-This can be validated in the $f(x, z, t) \propto e^{i(\vec{k} \cdot \vec{x} -
-\omega t)}$ case to give $\vec{v}_{ph} = \frac{\omega \vec{k}}{k^2}$. Note the
-sign in \autoref{eq:v_ph}
-
-\subsection{Supressing Reflection}
-
-Sourced from \url{https://people.maths.ox.ac.uk/trefethen/6all.pdf}. There are
-largely two ways to do this. The first is to add a damping zone, a region in
-which $\dot{q} = -q/\tau_d(z)$ where $\tau_d(0) = 0$ and increases to some
-non-small number for some $z_0$ beyond which we want damping. One form, chosen
-in Ryan and Dong's paper, is to use multiplicative factor $f(z) = \max\s*{0, 1 -
-\frac{\p*{z - z_b}^2}{\p*{z_d - z_b}^2}}$ where $z_b$ is where one begins
-supression and $z_b$ is the boundary. Then a dynamical variable $q$ can be
-supressed via $\dot{q} = (\dots) - f(z)\frac{q - q_0}{\tau}$ where $\tau$ is the
-dynamical timescale on which to damp and $q_0$ is the value to damp to.
-
-This requires we compute a ton of extra points though. Ideally, we want to use
-BCs. Based on our explicit solution \autoref{eq:strat.full_sol} we might expect
-to write something like $\pd{u_{1z}}{z} = \frac{1/2H +
-ik_z}{\mp i\omega}\pd{u_{1z}}{z} \approx \pm \frac{k_z}{\omega}\pd{u_{1z}}{z}$.
-For an outgoing wave we expect to choose the negative sign.
-
-\section{(Algebra/Optional) The Anelastic/Boussinesq Approximations}
-
-\subsection{Developing the Anelastic/Boussinesq Approximations}
-
-Let's relax the incompressibility constraint (we will expand the continuity
-equation to first order, but the momentum equation will merit a separate
-treatment):
-\begin{equation}
-    \begin{split}
-        \pd{\rho_1}{t} + \vec{\nabla} \cdot \p*{\rho_0 \vec{u}_1} &= 0,\\
-        \pd{\vec{u}_1}{t} &= -\frac{\vec{\nabla}P}{\rho} - \vec{g}.
-    \end{split}
-\end{equation}
-Suppose we are interested in phenomena with characteristic length scale $L$ and
-time scale $\tau$. Let's first examine the relative magnitudes of the terms in
-the continuity equation
-\begin{equation*}
-    \frac{\rho_1}{\tau} + \frac{\rho_0 \abs{u_1}}{L} = 0.
-\end{equation*}
-Thus, if we are interested in time scales $\tau \gg
-\frac{\rho_1}{\rho_0}\frac{L}{\abs{u_1}}$ then we neglect the first term, the
-time derivative. This corresponds to making the perturbation incompressible;
-note that $\pd{\rho_1}{t} \approx \rd{\rho_1}{t}$ to first order, so we drop the
-high frequency restoring forces in the perturbation.
-
-For the momentum equation, we instead first manipulate to first order
-\begin{align}
-    -\frac{\vec{\nabla}P}{\rho} - \vec{g}
-        &= -\frac{\vec{\nabla}P_0}{\rho} - \frac{\vec{\nabla}P_1}{\rho_0}
-            - \vec{g},\nonumber\\
-        &= - \frac{\vec{\nabla}P_1}{\rho_0} +
-            \p*{\frac{\rho_0}{\rho} - 1}\vec{g},\nonumber\\
-        &= -\vec{\nabla}\p*{\frac{P_1}{\rho_0}}
-            - \frac{P_1}{\rho_0^2}\vec{\nabla}\rho_0
-            - \frac{\rho_1}{\rho_0}\vec{g}.
-\end{align}
-
-We now have three equations for four variables, $\vec{u}_1, \rho_1, P_1$. We
-must introduce a fourth equation, a thermodynamic equation. For an adiabatic
-process $P\rho^{-\gamma} \propto P^{1-\gamma}T^\gamma$ is constant. We thus
-introduce the concept of the \emph{potential temperature}
-\begin{equation}
-    \theta = T\p*{\frac{P_0}{P}}^\kappa.
-        \label{eq:pot_temp}
-\end{equation}
-For an adiabatic process, $\rd{\theta}{t} = 0$. Motivated by this, we use
-\begin{equation}
-    \begin{split}
-        \pd{1}{\rho_0}\pd{\rho_0}{z} &= \frac{1}{\gamma P_0}\pd{P_0}{z}
-            - \frac{1}{\theta_0}\pd{\theta_0}{z},\\
-        \frac{\rho_1}{\rho_0} &= \frac{1}{\gamma} \frac{P_1}{P_0}
-            - \frac{\theta_1}{\theta_0},
-    \end{split}
-\end{equation}
-to give the momentum equation form
-\begin{equation}
-    \rd{\vec{u}_1}{t} = -\vec{\nabla}\p*{\frac{P_1}{\rho_0}}
-        + \frac{P_1}{\rho_0}\p*{\frac{1}{\theta_0}\vec{\nabla}\theta_0}
-        + \vec{g} \frac{\theta_1}{\theta_0}.\label{eq:mom_unsimp}
-\end{equation}
-
-We also recognize $N^2 = \frac{g}{\theta_0}\pd{\theta_0}{z}$. We now do the same
-trick where we consider dynamics on length scale $D$ and compare the first and
-second terms in \autoref{eq:mom_unsimp}. Their ratio is $\frac{N^2 D}{g}$, and
-so as $N^2 \ll \frac{g}{D}$ the freefall time we neglect the second term.
-
-The anelastic fluid equations thus read
-\begin{equation}
-    \begin{split}
-        \vec{\nabla} \cdot \p*{\rho_0\vec{u}} &= 0,\\
-        \pd{\vec{u}_1}{t} + \vec{\nabla}\p*{\frac{P_1}{\rho_0}}
-            - \vec{g}\frac{\theta_1}{\theta_0} &= 0,\\
-        \pd{\theta_1}{t} + \p*{\vec{u} \cdot \vec{\nabla}}\theta_0 &= 0.
-    \end{split}\label{eq:lin_anelastic}
-\end{equation}
-
-The Boussinesq equations are obtained from these in the limit where $H \gg D$
-the relevant length scale, thus we allow $\rho_0$ to be approximately constant.
-
-\subsection{Anelastic Solution to Stratified Atmosphere}
-
-We simply substitute $e^{i(\vec{k} \cdot \vec{r} - \omega t)}$ into
-\autoref{eq:lin_anelastic} with $\rho_0 \propto e^{-z/H}$ and obtain
-\begin{equation}
-    \begin{bmatrix}
-        0 & 0 & ik_x\rho_0 & ik_z \rho_0 - \frac{\rho_0}{H}\\
-        0 & -i\omega & 0 & \frac{N^2\theta_0}{g}\\
-        \frac{ik_x}{\rho_0} & 0 & -i\omega & 0\\
-        \frac{ik_z}{\rho_0} + \frac{1}{\rho_0 H} & -\frac{g}{\theta_0}
-            & 0 & -i\omega
-    \end{bmatrix} \begin{bmatrix}
-        P_1 \\ \theta_1 \\ u_{1x} \\ u_{1z}
-    \end{bmatrix} = 0.
-\end{equation}
-Taking the determinant of this matrix produces
-\begin{align}
-    -k_x^2\p*{-N^2 + \omega^2} + \p*{ik_z - \frac{1}{H}}
-        \p*{ik_z + \frac{1}{H}}\omega^2 &= 0,\nonumber\\
-    \frac{N^2k_x^2}{k_x^2 + k_z^2 + \frac{1}{4H^2}} &= \omega^2.
-\end{align}
+\end{frame}
 

### 2018-03-23 sims/2d_strat/agree_plots.py

@@ -1,6 +1,7 @@
 #!/usr/bin/env python
 # plot agreement of simulation/analytical at various (x, t) slices over z.
 
+import os
 import h5py
 import matplotlib.pyplot as plt
 import matplotlib
@@ -101,17 +102,20 @@ def agree_d0(plot_idx, t_idx=-10):
     return plot_idx + 1
 
 if __name__ == '__main__':
+    try:
+        os.makedirs('agree_plots')
+    except FileExistsError:
+        pass
+
     matplotlib.rcParams.update({'font.size': 6})
     plot_idx = 0
 
-    plot_idx = agree_sponge(plot_idx, -5)
     plot_idx = agree_sponge(plot_idx, -4)
     plot_idx = agree_sponge(plot_idx, -3)
     plot_idx = agree_sponge(plot_idx, -2)
     plot_idx = agree_sponge(plot_idx, -1)
 
     plot_idx = 0
-    plot_idx = agree_d0(plot_idx, -5)
     plot_idx = agree_d0(plot_idx, -4)
     plot_idx = agree_d0(plot_idx, -3)
     plot_idx = agree_d0(plot_idx, -2)

### 2018-03-23 presentations/Su_GM_032318.tex
check in final presentation

@@
     \begin{itemize}
-        \item 2D hydrodynamics with driving boundary condition, produce gravity
-            waves.
+        \item Eventual objective: simulate IGW breaking in WDs, compute
+            energy/angular momentum dissipation profiles.
+        \item Toy problem: 2D hydrodynamics with driving boundary condition.
         \item Begin with linear incompressible case, equations:
@@
             \omega t}}$, $u_{1x}, P_1, \rho_1$ similar up to phase and terms
-            $\sim \mathcal{O}\p*{H^{-1}}$.
+            $\sim \mathcal{O}\p*{H^{-1}} \ll k_z$.
+
+        \item Obeys dispersion relation
+            \begin{equation}
+                \omega^2 = \frac{N^2k_x^2}{k_x^2 + k_z^2 + \frac{1}{4H^2}}.
+            \end{equation}
 
@@
     \begin{itemize}
-        \item Linear incompressible hydrodynamics with periodic BCs in $x$,
+        \item Linear incompressible 2D hydrodynamics with periodic BCs in $x$,
             driving term at $z = 0$ and either Dirichlet or reflection-free BCs
@@
 
-        \item Spectral code Dedalus, built-in Runge-Kutta timestepper (simplest
-            to characterize).
+        \item Spectral code Dedalus, exponential convergence for shock-free
+            hydrodynamics.
     \end{itemize}
@@
         \centering
-        \begin{subfigure}{0.4\textwidth}
+        \begin{subfigure}{0.37\textwidth}
             \centering
@@
         \end{subfigure}
-        \begin{subfigure}{0.4\textwidth}
+        \begin{subfigure}{0.37\textwidth}
             \centering
@@
 
-        \begin{subfigure}{0.4\textwidth}
+        \begin{subfigure}{0.37\textwidth}
             \centering
@@
     \begin{itemize}
-        \item $RMS = \frac{\sqrt{\sum\limits_{x, z, t} \p*{u_{1z}^{(N)}(x, z, t)
-            - u_{1z}^{(2N)}(x, z, t)}^2}}{\sqrt{
-            \sum\limits_{x, z, t} \p*{u_{1z}^{(N)}(x, z, t)}^2}}$
+        \item $RMS_i = \frac{\sqrt{\sum\limits_{x, z, t} \p*{u_{1z}^{(i)}(x, z,
+            t) - u_{1z}^{(N)}(x, z, t)}^2}}{\sqrt{ \sum\limits_{x, z, t}
+            \p*{u_{1z}^{(N)}(x, z, t)}^2}}$
     \end{itemize}
@@
         \centering
-        \begin{subfigure}{0.3\textwidth}
-            \centering
-            \includegraphics[width=\textwidth]{../sims/2d_strat_conv/t_conv.png}
-        \end{subfigure}
-        \begin{subfigure}{0.3\textwidth}
+        \begin{subfigure}{0.37\textwidth}
             \centering
@@
         \end{subfigure}
-
-        \begin{subfigure}{0.3\textwidth}
+        \begin{subfigure}{0.37\textwidth}
             \centering
@@
         \end{subfigure}
-        \caption{Convergences of time stepsize and grid resolution. Note the
-        power law convergence in timestep but exponential in $z$ points (since
-        $x$ is always monochromatic, increasing resolution does little).}\label{fig:convsv}
+        \caption{Convergences of RMS over grid resolution. Note the exponential
+        convergence in $z$ to machine precision (since $x$ is always
+        monochromatic, increasing resolution does little).}\label{fig:convsv}
     \end{figure}

### 2018-03-23 sims/2d_strat_conv/strat.py

@@ -29,7 +29,7 @@ DT = T_F / num_timesteps
 PARAMS_RAW = {'XMAX': XMAX,
               'ZMAX': ZMAX,
               'N_X': 64,
-              'N_Z': 256,
+              'N_Z': 512,
               'T_F': T_F,
               'DT': DT,
               'KX': KX,
@@ -110,22 +110,26 @@ def run(name, args):
     return '%s completed' % name
 
 def rms_diff(arr1, arr2):
-    return np.sqrt(np.sum((arr1 - arr2)**2)) / np.sqrt(np.sum(arr1**2))
+    num_times = len(arr1)
+    return np.sqrt(np.sum((arr1[num_times // 2: ] - arr2[num_times // 2: ])**2))\
+        / np.sqrt(np.sum(arr1[num_times //2 : ]**2))
 
 if __name__ == '__main__':
     tasks = {
-        'd0_dt0': (dirichlet_bc, zero_ic, build_interp_params(16, 4, dt=DT * 2)),
-        'd0_dt2': (dirichlet_bc, zero_ic, build_interp_params(16, 4, dt=DT / 2)),
-        'd0_dt3': (dirichlet_bc, zero_ic, build_interp_params(16, 4, dt=DT / 4)),
-        'd0_dt4': (dirichlet_bc, zero_ic, build_interp_params(16, 4, dt=DT / 8)),
+        'd0_dt0': (dirichlet_bc, zero_ic, build_interp_params(16, 8, dt=DT * 2)),
+        'd0_dt2': (dirichlet_bc, zero_ic, build_interp_params(16, 8, dt=DT / 2)),
+        'd0_dt3': (dirichlet_bc, zero_ic, build_interp_params(16, 8, dt=DT / 4)),
+        'd0_dt4': (dirichlet_bc, zero_ic, build_interp_params(16, 8, dt=DT / 8)),
+        'd0_dt5': (dirichlet_bc, zero_ic, build_interp_params(16, 8, dt=DT / 16)),
         'd0_16_1': (dirichlet_bc, zero_ic, build_interp_params(16, 1, dt=DT)),
         'd0_16_2': (dirichlet_bc, zero_ic, build_interp_params(16, 2, dt=DT)),
         'd0_16_4': (dirichlet_bc, zero_ic, build_interp_params(16, 4, dt=DT)),
         'd0_16_8': (dirichlet_bc, zero_ic, build_interp_params(16, 8, dt=DT)),
[…]

### 2018-03-23 presentations/Su_GM_032318.tex
small changes to presentation

@@
         \item Top BC can be chosen Dirichlet $u_{1z}(x, L_z, t) = 0$,
-            analytically tractable.
+            analytically tractable as linear combination of $\pm k_z$ solutions.
 
@@
 
-        \item We use damping zone, add term $\rd{}{t}\vec{q} = L\vec{q}
+        \item We use damping zone, add term $\pd{}{t}\vec{q} = L\vec{q}
             - f(z)\p*{\vec{q} - \vec{q}_0}$ where
@@
         \end{subfigure}
-        \caption{Agreement of numerical and analytical solution at four adjacent
-        snapshots.}\label{fig:sponge_agree}
+        \caption{Numerical and analytical solutions agree. Numerics chooses
+        $v_{gz} > 0$.}\label{fig:sponge_agree}
     \end{figure}

### 2018-04-07 writups/notes.tex
check in agree_plots
write up some scratch work

@@
         u_{1x}\p*{x, z, t \middle | \vec{k}, \omega}
-            &\approx -\frac{k_z}{k_x} e^{z/2H}\frac{
+            &\approx \frac{k_z}{k_x} e^{z/2H}\frac{
                 \cos\p*{\frac{2\pi x}{L} - \omega_0 t + \frac{1}{2Hk_z}}
@@
             &\approx -\frac{\rho_0 \omega k_z}{k_x^2} e^{z/2H}\frac{
-                -\cos\p*{\frac{2\pi x}{L} - \omega_0 t + \frac{1}{2Hk_z}}
+                \cos\p*{\frac{2\pi x}{L} - \omega_0 t + \frac{1}{2Hk_z}}
                 \sin \p*{k_z(L - z)}
@@
 
-\subsection{Waveform Properties, Analytical}
+\subsection{Phase/Group Velocity}
 
 Let's figure out the analytical forms for the phase, group velocity and the
-energy density/power flux. We will try to recover these from waveforms.
+energy density/power flux.
 
@@
 For our stratified atmosphere problem, $\omega^2 = \frac{N^2 k_x^2}{k^2 +
-\frac{1}{4H^2}}$. Plugging in our parameters $k_z = -k_x = 2pi, N^2 = 10$ we can
-find $\omega \approx 2.23\;\mathrm{Hz}$, so $\vec{v}_{ph} \approx
-0.1777\p*{-\hat{x} + \hat{z}}\;\mathrm{s^{-1}}$.
-
-Since we use a height of $5$, we should expect the initial disturbance to reach
-the top boundary at $t \sim 30\;\mathrm{s}$.
+\frac{1}{4H^2}}$. This corresponds to $\vec{v}_{ph} = \frac{Nk_x}{\sqrt{k_x^2 +
+k_z^2 + 1/4H^2}} \vec{k}$.
 
@@
 \pd{\omega}{k_i}\hat{\imath}$. In our problem, $v_{g, z} =
--\frac{N^2k_x^2k_z}{\p*{k_x^2 + k_z^2 + 1/4H^2}^{3/2}}$ while $v_{g, x} =
-\frac{2N^2k_x}{\sqrt{k_x^2 + k_z^2 + 1/4H^2}} - \frac{N^2k_x^3}{\p*{k_x^2 +
+-\frac{Nk_xk_z}{\p*{k_x^2 + k_z^2 + 1/4H^2}^{3/2}}$ while $v_{g, x} =
+\frac{N}{\sqrt{k_x^2 + k_z^2 + 1/4H^2}} - \frac{Nk_x^2}{\p*{k_x^2 +
+k_z^2 + 1/4H^2}^{3/2}} = \frac{N\p*{k_z^2 + 1/4H^2}}{\p*{k_x^2 +
 k_z^2 + 1/4H^2}^{3/2}}$.
 
-We should note that the kinetic energy density is just $E = \frac{1}{2}\rho
-v^2$, the kinetic energy. Using then $\rd{E}{t} = 0$ shows that the
-corresponding power flux is $\vec{P} = \frac{\vec{v}}{2}\rho v^2$.
+\subsection{Energy/Power Flux}
 
-\subsection{Extracting Information from the Waveform}
+To compute the energy and power flux of the wave, we recall that for a general
+fluid the energy conservation equation reads
+\begin{equation}
+    \pd{}{t}\p*{\frac{1}{2}\rho v^2 + \rho \epsilon}
+        = \vec{\nabla} \cdot \p*{
+            \rho \vec{v}\p*{v^2 + \epsilon + \frac{P}{\rho}}
+        },
+\end{equation}
+where $\epsilon$ is the internal energy. But since $P = (\gamma -
+1)\rho\epsilon$ and we take $\gamma \to \infty$ incompressible limit, $\epsilon
+= 0$, so the energy of the wave is just $\frac{1}{2}\p*{\rho_0 + \rho_1} u_1^2$
+and the power flux is $\p*{\rho_0 + \rho_1}\vec{u}_1u_1^2 + P_1\vec{u}_1$.
 
-Let's make sure we can extract information from the linearized problem. Notable
-things we should be interested in: phase/group velocity, energy density and
-energy flux.
-
-Based on our phase velocity expression \autoref{eq:v_ph}, if we have some
-waveform $f(x, z, t)$, we can estimate its phase velocity $\vec{v}_{ph}(x, z,
-t)$ by
-\begin{equation}
-    \vec{v}_{ph}(x, z, t) = \frac{\partial_t f
-        \vec{\nabla}f}{\abs{\vec{\nabla}f}^2}.\label{eq:v_ph}
-\end{equation}
-This can be validated in the $f(x, z, t) \propto e^{i(\vec{k} \cdot \vec{x} -
-\omega t)}$ case to give $\vec{v}_{ph} = \frac{\omega \vec{k}}{k^2}$. Note the
-sign in \autoref{eq:v_ph}
+Another formula given by Sutherland 2011 is $\pd{\expvalue{E}}{t} = v_{g, z}
+\expvalue{E}$ where we average over $x$ one wavelength. I wasn't able to show
+that these two agree in general; Sutherland's formula seems to be derived for a
+traveling wavepacket whereas we have no such thing. We do not use this
+expression since the Landau \& Lifshitz expression works well.
 
@@
 
-This requires we compute a ton of extra points though. Ideally, we want to use
-BCs. TODO
+In order to compute the reflecting boundary condition, we must find a
+sequence of differential operators that well approxmiates the system of PDEs. We
+can get this from the dispersion relation; since our radiating boundary is along
+$z$, we should try to solve for $k_z$ in the dispersion relation to get a
+pseudodifferential operator form for $\pd{}{z}u_z$ (we take $k_xH \ll 1$ as well
+to simplify):
+\begin{align}
+    k_z^2 &= \frac{N^2}{\omega^2}k_x^2 - k_x^2 - \frac{1}{4H^2},\nonumber\\
+        &\approx k_x^2\p*{\frac{N^2}{\omega^2} - 1},\\
+    \pd{u_z}{z} &\approx -\pd{u_z}{x}\sqrt{\frac{N^2}{\omega^2} - 1}.
+\end{align}
+We pick the negative sign in accordance with an outgoing wave, per the group
+velocity formulae for $k_z > k_x$.
 
-\section{(Algebra/Optional) The Anelastic/Boussinesq Approximations}
+\clearpage
+
+\appendix
+
+\chapter{Deriving Fluids Results}
+
+\section{Equation of Energy Conservation}
+
+We follow Landau \& Lifshitz's derivation for this expression. Consider that the
+total energy stored in a wave must be the sum of its kinetic and internal energy
+$\frac{1}{2}\rho v^2 + \rho \epsilon$ where $\epsilon$ is the internal energy
+density. To obtain an equation of energy conservation, we must take the time
+derivative of this expression. First, we consider
+\begin{align}
+    \pd{}{t}\p*{\frac{1}{2}\rho v^2} &= \frac{1}{2}v^2\pd{\rho}{t}
+        + \rho \vec{v} \cdot \pd{\vec{v}}{t},\nonumber\\
+        &= -\frac{1}{2}v^2 \p*{\vec{\nabla} \cdot \rho \vec{v}}
+            - \rho \vec{v} \cdot \p*{
+                \p*{\vec{v} \cdot \vec{\nabla}}\vec{v} +
+                    \frac{\vec{\nabla}P}{\rho} - \vec{g}
+            }_{T, V},\nonumber\\
+        &= -\frac{1}{2} v^2 \p*{\vec{\nabla} \cdot \rho \vec{v}}
+            - \frac{1}{2}\rho \vec{v} \cdot \p*{\vec{\nabla} v^2}
+            - \vec{v} \cdot \p*{
+                \rho \vec{\nabla} w - \rho T \vec{\nabla}s
+            } + \rho \vec{v} \cdot \vec{g}.
+\end{align}
+We have denoted $s$ the specific internal entropy density of the fluid and
+$\mathrm{d}w = T\mathrm{d}s + \frac{\mathrm{d}P}{\rho} = T\mathrm{d}s +
+\epsilon$ the specific internal enthalpy density of the fluid.
+Recall enthalpy $\epsilon = w - Ts$ is the usual thermodynamic definition.
+
+At the same time, consider
+\begin{align}
+    \pd{}{t}\p*{\rho \epsilon} &= \epsilon \pd{\rho}{t}
+        + \rho \pd{}{t}\p*{Ts - \frac{P}{\rho}}_{T, P},\nonumber\\
+        &= \epsilon \pd{\rho}{t} + \rho T \pd{s}{t} + \frac{P}{\rho}\pd{\rho}{t}
+            ,\nonumber\\
+        &= w \pd{\rho}{t} + \rho T \pd{s}{t},\nonumber\\
+        &= -w\vec{\nabla}\p*{\rho \vec{v}} - \rho T \vec{v} \cdot \p*{
+            \vec{\nabla}s
+        }.
+\end{align}
+Summing the two, we find
+\begin{equation}
+    \pd{}{t}\p*{\frac{\rho v^2}{2} + \rho \epsilon}
+        = -\vec{\nabla} \cdot \s*{
+            \rho \vec{v}\p*{v^2 + w}
+        } = -\vec{\nabla} \cdot \s*{
+            \rho \vec{v}\p*{v^2 + \epsilon + \frac{P}{\rho}}
+        }.
+\end{equation}
+
+\section{The Anelastic/Boussinesq Approximations}
 

### 2018-04-07 writups/notes.tex
implement radiative conditions, writeup first of nonlinear

@@
     \begin{split}
-        \rd{\rho}{t} + \vec{\nabla} \cdot \vec{u} &= 0,\\
+        \pd{\rho}{t} + \vec{\nabla} \cdot \p*{\rho \vec{u}} &= 0,\\
         \rd{\vec{u}}{t} + \frac{\vec{\nabla}P}{\rho} - \vec{g} &= 0.
@@
     \begin{split}
-        \pd{\rho_1}{t} + \p*{\vec{u}_1 \cdot \vec{\nabla}}\rho_0 &= 0,\\
+        \pd{\rho_1}{t} + \vec{u}_1 \cdot \p*{\vec{\nabla}\rho_0} &= 0,\\
         \vec{\nabla} \cdot \vec{u}_1 &= 0,\\
@@
 supressed via $\dot{q} = (\dots) - f(z)\frac{q - q_0}{\tau}$ where $\tau$ is the
-dynamical timescale on which to damp and $q_0$ is the value to damp to.
+dynamical timescale on which to damp and $q_0$ is the value to damp to. We will
+apply this only to the velocity variables with an eye to extending this approach
+to the nonlinear regime, where the velocity variables should still be damped to
+zero but $\rho, P$ must capture the stratification and so cannot easily be
+artificially damped (maybe? Revisit if this does poorly compared to damping all
+linear variables, and damp the nonlinear variables to their
+equilibrium/stratified values).
 
@@
 velocity formulae for $k_z > k_x$.
+
+From the numerical simulations, it is clear that reflections are well supressed
+initially, but the the simulation blows up. This is because our supression is
+imperfect (differential approximation to a pseudodifferential operator) and
+since our system is dissipation free, reflection grows. Moreover, since any
+reflected components seem to have different $k_x, k_z, \omega$, the second time
+they're incident on the boundary, our above condition becomes wildly inaccurate.
+
+\section{Incompressible, Stratified w/ Gravity, Nonlinear}
+
+We consider the problem where $\rho_1 \sim \rho_0$, where we cannot make such a
+perturbative expansion. We write down thus nonlinear fluid equations
+\begin{equation}
+    \begin{split}
+        \pd{\rho}{t} + \vec{u} \cdot \p*{\vec{\nabla}\rho} &= 0,\\
+        \vec{\nabla} \cdot \vec{u} &= 0,\\
+        \pd{\vec{u}}{t} &= -\frac{\vec{\nabla}P}{\rho} - g\hat{z}.
+    \end{split}\label{eq:nonlin_incomp}
+\end{equation}
+Note that we cannot even subtract off hydrostatic equilibrium anymore since
+$\rho$ can deviate greatly from the normal $\rho_0e^{-z/H}$!
+
+Since Dedalus cannot handle a field quantity in the denominator, we must get rid
+of $\rho$ somehow. We choose to do this by recasting the equations in terms of
+$\rho^{-1}$, which is everywhere nonzero and finite since $\rho$ is everywhere
+nonzero and finite. But then $\pd{\rho}{t} = -\rho^2\pd{\rho^{-1}}{t}$ while
+$\vec{\nabla}\rho = -\rho^2\vec{\nabla}\rho^{-1}$, and so factoring out the
+$\rho^2$ we can simply rewrite
+\begin{equation}
+    \begin{split}
+        \pd{\rho^{-1}}{t} + \vec{u} \cdot \p*{\vec{\nabla}\rho^{-1}} &= 0,\\
+        \vec{\nabla} \cdot \vec{u} &= 0,\\
+        \pd{\vec{u}}{t} &= -\rho^{-1}\vec{\nabla}P - g\hat{z}.
+    \end{split}\label{eq:nonlin_incomp}
+\end{equation}
+
+We must also consider how we implement the outgoing boundary condition in this
+regime. If we use the radiative boundary condition formula, thankfully we have
+little to change, though it also doesn't do a great job of supressing
+reflections. If we instead use a sponge zone, we should be careful about how we
+damp $\rho^{-1}$, since we obviously should not try damping it to zero! This is
+why we initially chose to only apply our sponge zone to the velocity variables,
+even in the linear problem.
 

### 2018-04-07 sims/2d_strat/strat.py

@@ -53,9 +53,9 @@ def dirichlet_bc(problem, *_):
     problem.add_bc('right(uz) = 0', condition='nx != 0')
 
 def rad_bc(problem, *_):
-    # TODO incorrect
+    problem.parameters['BC_PARAM'] = np.sqrt(G / (H * OMEGA ** 2) - 1)
     strat_helper.default_problem(problem)
-    problem.add_bc('right(dt(uz) + omega / KZ * dz(uz)) = 0',
+    problem.add_bc('right(dz(uz) - dx(uz) * BC_PARAM) = 0',
                    condition='nx != 0')
 
 def sponge(problem, domain):
@@ -106,8 +106,9 @@ def run(bc, ic, name, params_dict):
 
 if __name__ == '__main__':
     tasks = [
-        (dirichlet_bc, zero_ic, 'd0', build_interp_params(8, 1)),
-        (sponge, zero_ic, 'sponge', build_interp_params(4, 4)),
+        (dirichlet_bc, zero_ic, 'd0', build_interp_params(8, 4)),
+        (sponge, zero_ic, 'sponge', build_interp_params(8, 8)),
+        (rad_bc, zero_ic, 'rad', build_interp_params(8, 8)),
     ]
 
     with Pool(processes=N_PARALLEL) as p:

### 2018-04-07 sims/2d_strat/strat_helper.py

@@ -51,15 +51,6 @@ def get_solver(setup_problem,
     rho0['g'] = RHO0 * np.exp(-z / H)
     problem.parameters['rho0'] = rho0
 
-    # plot rho0
-    # xmesh, zmesh = quad_mesh(x=x[:, 0], y=z[0])
-    # plt.pcolormesh(xmesh, zmesh, rho0['g'].T)
-    # plt.xlabel('x')
-    # plt.ylabel('z')
-    # plt.title('Background rho0')
-    # plt.colorbar()
-    # plt.savefig('strat_rho0.png')
-
     setup_problem(problem, domain)
 
     # Build solver
@@ -134,7 +125,7 @@ def get_omega(g, h, kx, kz):
     return np.sqrt((g / h) * kx**2 / (kx**2 + kz**2 + 0.25 / h**2))
 
 def get_vph(g, h, kx, kz):
-    norm = get_omega(g, h, kx, kz) / (kx**2 + kz**2)
+    norm = abs(get_omega(g, h, kx, kz)) / (kx**2 + kz**2)
     return norm * kz, norm * kz
 
 def load(setup_problem,
@@ -187,6 +178,9 @@ def load(setup_problem,
     rho0 = RHO0 * np.exp(-z / H)
     state_vars['E'] = ((rho0 + state_vars['rho']) *
                        (state_vars['ux']**2 + state_vars['uz']**2)) / 2
+    state_vars['F_z'] =  state_vars['uz'] * (
+        (rho0 + state_vars['rho']) * (state_vars['ux']**2 + state_vars['uz']**2)
+        + state_vars['P'])
     state_vars['rho0'] = rho0 * np.ones(np.shape(state_vars['E']))
     return sim_times, domain, state_vars
 
@@ -212,9 +206,9 @@ def plot(setup_problem,
     path = '{s}/{s}_s1'.format(s=snapshots_dir)
     matplotlib.rcParams.update({'font.size': 6})
     plot_vars = ['uz', 'ux', 'rho', 'P', 'rho0']
-    z_vars = ['E'] # sum these over x
+    z_vars = ['F_z', 'E'] # sum these over x
     n_cols = 3
-    n_rows = 2
+    n_rows = 3
     plot_stride = 1
 
     if os.path.exists('%s.mp4' % name):

### 2018-04-12 writups/notes.tex
nonlin is working

@@
         \centering
-        \includegraphics[width=\textwidth]{../sims/2d_no_g/no_g_P_t50.png}
+        \includegraphics[width=\textwidth]{../sims/2d_0_no_g/no_g_P_t50.png}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{../sims/2d_no_g/no_g_ux_t50.png}
+        \includegraphics[width=\textwidth]{../sims/2d_0_no_g/no_g_ux_t50.png}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{../sims/2d_no_g/no_g_uz_t50.png}
+        \includegraphics[width=\textwidth]{../sims/2d_0_no_g/no_g_uz_t50.png}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{../sims/2d_no_g/no_g_P_t100.png}
+        \includegraphics[width=\textwidth]{../sims/2d_0_no_g/no_g_P_t100.png}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{../sims/2d_no_g/no_g_ux_t100.png}
+        \includegraphics[width=\textwidth]{../sims/2d_0_no_g/no_g_ux_t100.png}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{../sims/2d_no_g/no_g_uz_t100.png}
+        \includegraphics[width=\textwidth]{../sims/2d_0_no_g/no_g_uz_t100.png}
     \end{subfigure}
@@
         \vec{\nabla} \cdot \vec{u} &= 0,\\
-        \pd{\vec{u}}{t} &= -\frac{\vec{\nabla}P}{\rho} - g\hat{z}.
+        \pd{\vec{u}}{t} &= -\p*{\vec{u} \cdot \vec{\nabla}} \vec{u}
+            - \frac{\vec{\nabla}P}{\rho} - g\hat{z}.
     \end{split}\label{eq:nonlin_incomp}
@@
 $\rho$ can deviate greatly from the normal $\rho_0e^{-z/H}$!
-
-Since Dedalus cannot handle a field quantity in the denominator, we must get rid
-of $\rho$ somehow. We choose to do this by recasting the equations in terms of
-$\rho^{-1}$, which is everywhere nonzero and finite since $\rho$ is everywhere
-nonzero and finite. But then $\pd{\rho}{t} = -\rho^2\pd{\rho^{-1}}{t}$ while
-$\vec{\nabla}\rho = -\rho^2\vec{\nabla}\rho^{-1}$, and so factoring out the
-$\rho^2$ we can simply rewrite
-\begin{equation}
-    \begin{split}
-        \pd{\rho^{-1}}{t} + \vec{u} \cdot \p*{\vec{\nabla}\rho^{-1}} &= 0,\\
-        \vec{\nabla} \cdot \vec{u} &= 0,\\
-        \pd{\vec{u}}{t} &= -\rho^{-1}\vec{\nabla}P - g\hat{z}.
-    \end{split}\label{eq:nonlin_incomp}
-\end{equation}
-
-We must also consider how we implement the outgoing boundary condition in this
-regime. If we use the radiative boundary condition formula, thankfully we have
-little to change, though it also doesn't do a great job of supressing
-reflections. If we instead use a sponge zone, we should be careful about how we
-damp $\rho^{-1}$, since we obviously should not try damping it to zero! This is
-why we initially chose to only apply our sponge zone to the velocity variables,
-even in the linear problem.
 
@@
 \end{equation}
+
+\section{Conservative Fluid Equations}
+
+This isn't particularly useful for our work since spectral methods don't benefit
+from having the dynamical equations in conservative form, but form a useful
+exercise for the ever-so-stupid writer.
+
+The Reynolds transport theorem tells us that
+\begin{equation}
+    \rd{}{t}\int\limits_V Q\;\mathrm{d}V
+        = \int\limits_V \pd{Q}{t}\;\mathrm{d}V
+            + \int\limits_{\partial V}
+                \p*{\vec{v}\p*{\mathrm{d}A} \cdot \hat{n}} Q\;\mathrm{d}A.
+\end{equation}
+We have notated $\vec{v}\p*{\mathrm{d}A}$ the velocity of the surface element.
+In fluid mechanics, this is just the velocity field. Thus, for conserved
+quantities where $\rd{}{t}\int\limits_V Q\;\mathrm{d}V = 0$, we can apply
+divergence theorem to obtain the simple result
+\begin{equation}
+    \pd{Q}{t} + \vec{\nabla} \cdot \p*{Q\vec{v}} = 0.\label{eq:cons_eq}
+\end{equation}
+Of course, if there are any net forces etc.\ on the system, they are simply
+inserted as fileds on the right hand side of \autoref{eq:cons_eq}.
+
+The application of this to the three major conserved quantities: mass, momentum
+and energy, yield:
+\begin{description}
+    \item[Conservation of Mass] This produces simply $\pd{\rho}{t} +
+        \vec{\nabla} \cdot \p*{\rho \vec{v}} = 0$ the continuity equation.
+
+    \item[Conservation of Momentum] This produces $\pd{\rho v_i}{t} +
+        \vec{\nabla} \cdot \p*{\rho v_i \vec{v}} = 0$. To obtain the traditional
+        momentum equation, it is perhaps most clear in index notation where
+        repeated indicies denote summation:
+        \begin{align}
+            \pd{\p*{\rho v_i}}{t} + \partial_j \rho v_iv_j &= 0,\nonumber\\
+            v_i \pd{\rho}{t} + \rho \pd{v_i}{t} + v_I \partial_j \p*{\rho v_j}
+                + \rho v_j \partial_j v_i &= 0,\nonumber\\
+            v_i\p*{\pd{\rho}{t} + \partial_j \p*{\rho v_j}}
+                + \rho\p*{
+                    \pd{v_i}{t} + v_j \partial_t v_i
+                } = 0.
+        \end{align}
+        The first parenthetical term we recognize is just the continuity
+        equation though. The remainder is the traditional source-free momentum
+        equation.
+
+    \item[Conservation of Energy] This just reads $\pd{e}{t} + \vec{\nabla}
+        \cdot \p*{e \vec{v}} = 0$. It's a pretty uninteresting equation without
+        pressure and thermodynamic effects, which would arise from a more
+        careful treatment of the source terms.
+
+        In fact, pressure contributes a source term $P\vec{\nabla} \cdot
+        \vec{v}$, so the full energy equation is often written
+        \begin{equation}
+            \rd{e}{t} + \p*{e + P}\vec{\nabla} \cdot \vec{v} = 0.
+        \end{equation}
+\end{description}
 

### 2018-04-12 sims/2d_1_strat/strat.py

@@ -22,7 +22,7 @@ KZ = (np.pi / 2) * np.pi / H
 G = 10
 OMEGA = strat_helper.get_omega(G, H, KX, KZ)
 VPH_X, VPH_Z = strat_helper.get_vph(G, H, KX, KZ)
-T_F = (ZMAX / VPH_Z) * 2
+T_F = (ZMAX / VPH_Z) * 4
 DT = T_F / num_timesteps
 
 PARAMS_RAW = {'XMAX': XMAX,
@@ -134,9 +134,9 @@ def run(bc, ic, name, params_dict):
 if __name__ == '__main__':
     tasks = [
         (dirichlet_bc, zero_ic, 'd0', build_interp_params(8, 4)),
-        (sponge, zero_ic, 'sponge', build_interp_params(8, 8)),
-        (sponge_part, zero_ic, 'sponge_part', build_interp_params(8, 8)),
-        (rad_bc, zero_ic, 'rad', build_interp_params(8, 8)),
+        (sponge, zero_ic, 'sponge', build_interp_params(8, 4)),
+        (sponge_part, zero_ic, 'sponge_part', build_interp_params(8, 4)),
+        (rad_bc, zero_ic, 'rad', build_interp_params(8, 4)),
     ]
 
     with Pool(processes=N_PARALLEL) as p:

### 2018-04-12 sims/2d_2_strat_nonlin/strat_helper.py

@@ -1,0 +1,286 @@
+#!/usr/bin/env python
+'''
+helper function to run the shared stratification scenario. user just has to
+specify BCs and ICs
+'''
+import logging
+import os
+from collections import defaultdict
+
+import h5py
+import matplotlib
+import matplotlib.pyplot as plt
+import numpy as np
+
+from dedalus import public as de
+from dedalus.extras.flow_tools import CFL
+from dedalus.extras.plot_tools import quad_mesh, pad_limits
+
+SNAPSHOTS_DIR = 'snapshots_%s'
+def get_solver(setup_problem,
+               XMAX,
+               ZMAX,
+               N_X,
+               N_Z,
+               T_F,
+               KX,
+               KZ,
+               H,
+               RHO0,
+               G,
+               A,
+               NU):
+    # Bases and domain
+    x_basis = de.Fourier('x', N_X, interval=(0, XMAX), dealias=3/2)
+    z_basis = de.Chebyshev('z', N_Z, interval=(0, ZMAX), dealias=3/2)
+    domain = de.Domain([x_basis, z_basis], np.float64)
+    z = domain.grid(1)
+
+    problem = de.IVP(domain, variables=['P', 'rho', 'ux', 'uz', 'ux_z', 'uz_z'])
+    problem.parameters['L'] = XMAX
+    problem.parameters['g'] = G
+    problem.parameters['H'] = H
+    problem.parameters['A'] = A
+    problem.parameters['KX'] = KX
+    problem.parameters['KZ'] = KZ
+    problem.parameters['NU'] = NU
+    problem.parameters['RHO0'] = RHO0
+    problem.parameters['omega'] = get_omega(G, H, KX, KZ)
+
+    # rho0 stratification
+    rho0 = domain.new_field()
+    rho0.meta['x']['constant'] = True
+    rho0['g'] = RHO0 * np.exp(-z / H)
+    problem.parameters['rho0'] = rho0
+
+    setup_problem(problem, domain)
+
+    # Build solver
+    solver = problem.build_solver(de.timesteppers.RK443)
+    solver.stop_sim_time = T_F
+    solver.stop_wall_time = np.inf
+    solver.stop_iteration = np.inf
+    return solver, domain
+
+def run_strat_sim(setup_problem,
+                  set_ICs,
+                  XMAX,
+                  ZMAX,
+                  N_X,
+                  N_Z,
+                  T_F,
+                  DT,
[…]

### 2018-04-22 writups/notes.tex
update writeup with recent musings

@@
 
-\subsection{Supressing Reflection}
+\section{Supressing Reflection}
 
@@
 equilibrium/stratified values).
+
+\subsection{First Order}
 
@@
 
-\section{Incompressible, Stratified w/ Gravity, Nonlinear}
+\section{The Difficulty of the Incompressible, Nonlinear Problem}
 
-We consider the problem where $\rho_1 \sim \rho_0$, where we cannot make such a
-perturbative expansion. We write down thus nonlinear fluid equations
+We consider the problem where both advective terms are kept and $\rho_1 \sim
+\rho_0$. We write down thus nonlinear fluid equations
 \begin{equation}
@@
 $\rho$ can deviate greatly from the normal $\rho_0e^{-z/H}$!
+
+\subsection{Rearrangement for Dedalus}
+
+In order to simulate this at all in a spectral program, recall the way that
+nonlinear PDEs are decomposed in a spectral code. Given a phase space $Q$,
+consider first a system of PDEs expressible in form
+\begin{equation}
+    \dot{Q} + \bm{L}Q + f(Q) = g
+\end{equation}
+where $\dot{Q}$ is the time derivative, $\bm{L}$ is some linear operator, $f(Q)$
+are nonlinear terms and $g$ is an arbitrary driving function. Spectral codes
+work by re-casting the operators to be purely algebraic in an appropriate
+spectral basis consisting of some $\varphi_i$ trial and $\xi_i$ test functions.
+The \emph{tau spectral method} that Dedalus uses considers using the trial
+functions as the test functions, so the effective decomposition that is
+performed is
+\begin{equation}
+    \bra*{\varphi_i}\dot{Q} + \bm{L}Q \ket*{\varphi_j}
+        = \dot{Q}_i + L_{ij}Q_q = \bra*{\varphi_i}-f(Q)  + g \ket*{\varphi_i}.
+\end{equation}
+It is then clear that $-f(Q) + g$ can just be treated as inhomogeneous source
+terms, and over a short $\Delta t$ it can be approximated via $-f(Q_0) + g$.
+This first-order inhomegeneous ODE then admits closed form exact solutions
+$e^{-L_{ij}t}\p*{\dots}$.
+
+One can imagine, for this matter, that $\rd{}{t}$ can be wrapped in the $L_{ij}$
+operator and the timestepping computed implicitly; I believe this is what
+Dedalus does, implicitly timestepping the left hand side and explicitly the
+inhomogeneous terms.
+
+Both of these procedures require $\bm{L}$ linear operator or $\rd{}{t} + \bm{L}$
+to be full rank, or invertible. This is enforced analytically by requiring the
+same number of BCs as derivatives, and numerically this is little different.
+This is why it is possible to have components of $Q$ that are not differentiated
+$\partial_t$ but still have well-defined time evolutions; the only condition
+required is that the linear terms of the PDE make up a full-rank operator.
+Hence, in our linear incompressible equations, we permitted use of $\vec{\nabla}
+\cdot \vec{u}_1 = 0$ in lieu of a $\pd{P}{t}$ explicit equation of motion; so
+long as the gauge for $P$ (which only appears in derivatives) is properly set,
+the operator is full rank and numerically well-defined.
+
+In the case of the full nonlinear equations though, it bears noting that $P$
+only appears in nonlinear terms! Thus, the linear terms clearly cannot have full
+rank since they fail to reference $P$ at all. We remedy this by re-casting the
+momentum equation as
+\begin{equation}
+    \pd{\vec{u}}{t} + \frac{\vec{\nabla}P}{\rho_0}
+        = -\vec{\nabla}P\p*{\frac{1}{\rho} - \frac{1}{\rho_0}}
+            + \vec{g}.
+\end{equation}
+This ensures that, along with suitable gauge choice for $P$, the linear
+operators on the left hand side of the equation have full rank. The remaining
+equations
+\begin{subequations}
+    \begin{align}
+        \pd{\rho}{t} &= -\vec{u} \cdot \vec{\nabla}\rho,\\
+        \vec{\nabla} \cdot \vec{u} &= 0,
+    \end{align}
+\end{subequations}
+remain unchanged.
+
+\subsection{Pathology with $u_z$ BCs}
+
+Our problem turns out to be ill-conditioned, the pathology arising from our
+driving term in conjunction with the incompressibility constraint. Since we want
+to model the propagation of waves from where they are excited, rather than
+exciting them directly in our problem, we must drive our region of interest via
+boundary conditions rather than forcing terms.
+
+By driving via boundary conditions, we can only specify boundary conditions on
+the variables being differentiated $\partial_z$ in our equations of motion,
+since we enforce fully periodic BCs in $x$ via using a Fourier spectral
+decomposition. Thus, we are restricted to enforcing BCs on either $\vec{u}$ or
+$P$.
+
+It turns out that by using BCs on $\vec{u}$, we run into a problem: the BCs we
+specify on $u_z$ or its derivatives will in general affect grid points
+$\partial_z u_{z, 1x}$ while those specified on $u_x$ can only affect
+$\partial_x u_{x, 0x}$. This means that $\partial_x u_x$ will pick up
+sharp jerks by being coupled to the BC, producing high frequency components in
+$u_x$ as can be seen in \autoref{fig:agree_plots}. These come from enforcing the
+divergence-free condition too literally at the boundary. Not that the figure was
+produced with only a Dirichlet BC on $u_z$, no BCs on $u_x$, in a linear system
+with absorbing BCs.
+\begin{figure}[!h]
+    \centering
+    \includegraphics[width=0.7\textwidth]{../sims/2d_1_strat/agree_plots/sponge_2.png}
+    \caption{Note the jaggedness in $u_x$ induced by a Dirichlet BC on
+    $u_z$.}\label{fig:agree_plots}
+\end{figure}
+
+This problem seems to be remedied somewhat by instead enforcing Neumann BCs on
+$u_z$, or even when allowing dissipation enforcing Neumann BCs on $\partial_z
+u_z$ (the extra derivatives granted by $\nu \nabla^2 \vec{u}$ let us specify
+higher order BCs). However, in the fully nonlinear system we still observe
+divergences near $z=0$ regardless of how smooth of BCs we use on $u_z$.
+Moreover, since the discontinuities come from the nature of the BC and not its
+strength, we run into these instabilities regardless of how gradually we
+introduce the BC (e.g.\ we could consider multiplying by $1 - e^{-t/\tau}$ to
+gradually introduce the BC, but this turns out also to be unstable).
+
+There are thus three possible solutions to this:
+\begin{itemize}
+    \item Consider BCs only on $P$, the only other choice of dynamical variable.
+
+    \item Consider using a $\nabla^2 P$ equation instead of a $\vec{\nabla}
+        \cdot \vec{u} = 0$ constraint. The correct choice of BCs in this case is
+        discussed in Nordstrom et al, 2006.
+
+    \item Change the system of equations to no longer consider an incompressible
+        system; the anelastic system is an attractive alternative. This boasts
+        the advantage of not having an algebraic constraint equation.
+\end{itemize}
+
+\section{Limits of the Incompressibility and Anelasticity Assumptions}
+
+To get a better handle on the regimes of validity of the incompressible and
+anelastic approximations, we return to the full fluid equations
+\begin{equation}
+    \begin{split}
+        \pd{\rho}{t} + \vec{\nabla} \cdot \p*{\rho \vec{u}} &= 0,\\
+        \pd{\vec{u}}{t} + \p*{\vec{u} \cdot \vec{\nabla}}\vec{u} +
+            \frac{\vec{\nabla}P}{\rho} - \vec{g} &= 0.
+    \end{split}
+\end{equation}
+Furthermore, we assume small perturbations $\delta P = c_s^2 \delta \rho$ where
+$c_s^2$ is the sound speed, also the velocity at which an adiabatic perturbation
+is restored. We drop the $-\vec{g}$ from the analysis assuming that it is
+largely negated by an appropriate background $P$.
+
+Given these considerations, we consider in what limit we may presume the fluid
+to be incompressible. Consider if $\abs*{\vec{u}} \ll c_s$, then for some
+characteristic frequency $\omega$ and wavenumber $k$ the momentum equation
+scales as $\omega u + ku^2 + c_s^2 \delta \rho/\rho = 0$. Comparing scales it is
+thus clear that $\delta \rho \ll \rho$ in this limit. Then, considering the
+mass continuity equation, we recognize that the scalings are $\omega \delta \rho
++ \rho \vec{\nabla} \cdot \vec{u} + k\delta \rho u = 0$. Thus, $\vec{\nabla}
+\cdot \vec{u} = 0$ necessarily, since $\rho$ is much larger than the other two
+terms, and we obtain the incompressible mass continuity equation from the other
+two terms. While this is obviously not a very careful treatment, it should
+suffice to convince us that $u \ll c, \delta \rho \ll \rho, \vec{\nabla} \cdot
+\vec{u} = 0$ are all roughly equivalent conditions.
+
+On the other hand, when may we drop the total derivative terms? This seems to be
+a slightly different criterion, and indeed we may compare $\pd{\vec{u}}{t} +
+\p*{\vec{u} \cdot \vec{\nabla}} \vec{u} \sim \omega \vec{u} + ku^2$. Thus, the
+second term is dominated by the first term when $\vec{u} \ll \vec{v}_{ph}$ where
+$\vec{v}_{ph} \equiv \frac{\omega}{k}\hat{k}$ is the phase velocity. Recall that
+for our stratified waves that $\omega \leq N = \sqrt{g/H}$ as a point of
+reference. So we require $u \ll v_{ph}$ to be able to drop the total derivative
+terms.
+
+Since in general $v_{ph} \ll c_s$, we thus identify that there are three
+regimes. The first, $u \ll v_{ph}$, we solve the completely linear problem.
+Then, for $v_{ph} \lesssim u \ll c_s$, we must begin to consider the advective
+terms but are still allowed to assume $\rho_1 \ll \rho_0$, an assumption that is
+only violated when $u \sim c_s$.
+
+If we look ahead to the anelastic system derivation, we can identify that the
+anelastic approximation corresponds to assuming $c_s^2 \gg uv_{ph}$, making no
+assertion of the scale between $u, v_{ph}$. Thus, we can also consider the
+anelastic approximation with advective terms just fine in teh regime $v_{ph}
+\lesssim u$, and if this is sufficient to induce wave breaking then we may never
+need to consider the $\rho_1 \sim \rho_0$ enormous perturbations.
 

### 2018-06-06 correspondence/180606-lecoanet/brief.tex
finished brief

    \documentclass[11pt,
        usenames, % allows access to some tikz colors
        dvipsnames % more colors: https://en.wikibooks.org/wiki/LaTeX/Colors
    ]{article}
    \usepackage{
        amsmath,
        amssymb,
        fouriernc, % fourier font w/ new century book
        hyperref, % various links
        setspace, % line spacing
        amsthm, % newtheorem and proof environment
        mathtools, % \Aboxed for boxing inside aligns, among others
        float, % Allow [H] figure env alignment
        enumerate, % Allow custom enumerate numbering
        graphicx, % allow includegraphics with more filetypes
        wasysym, % \smiley!
        upgreek, % \upmu for \mum macro
        listings, % writing TrueType fonts and including code prettily
        tikz, % drawing things
        booktabs, % \bottomrule instead of hline apparently
        cancel % can cancel things out!
    }
    \usepackage[margin=1in]{geometry} % page geometry
    \usepackage[
        labelfont=bf, % caption names are labeled in bold
        font=scriptsize % smaller font for captions
    ]{caption}
    \usepackage[font=scriptsize]{subcaption} % subfigures

    \newcommand*{\scinot}[2]{#1\times10^{#2}}
    \newcommand*{\dotp}[2]{\left<#1\,\middle|\,#2\right>}
    \newcommand*{\rd}[2]{\frac{\mathrm{d}#1}{\mathrm{d}#2}}
    \newcommand*{\pd}[2]{\frac{\partial#1}{\partial#2}}
    \newcommand*{\rtd}[2]{\frac{\mathrm{d}^2#1}{\mathrm{d}#2^2}}
    \newcommand*{\ptd}[2]{\frac{\partial^2 #1}{\partial#2^2}}
    \newcommand*{\md}[2]{\frac{\mathrm{D}#1}{\mathrm{D}#2}}
    \newcommand*{\pvec}[1]{\vec{#1}^{\,\prime}}
    \newcommand*{\svec}[1]{\vec{#1}\;\!}
    \newcommand*{\bm}[1]{\boldsymbol{\mathbf{#1}}}
    \newcommand*{\ang}[0]{\;\text{\AA}}
    \newcommand*{\mum}[0]{\;\upmu \mathrm{m}}
    \newcommand*{\at}[1]{\left.#1\right|}

    \newtheorem{theorem}{Theorem}[section]

    \let\Re\undefined
    \let\Im\undefined
    \DeclareMathOperator{\Res}{Res}
    \DeclareMathOperator{\Re}{Re}
    \DeclareMathOperator{\Im}{Im}
    \DeclareMathOperator{\Log}{Log}
    \DeclareMathOperator{\Arg}{Arg}
    \DeclareMathOperator{\Tr}{Tr}
    \DeclareMathOperator{\E}{E}
    \DeclareMathOperator{\Var}{Var}
    \DeclareMathOperator*{\argmin}{argmin}
    \DeclareMathOperator*{\argmax}{argmax}
    \DeclareMathOperator{\sgn}{sgn}
    \DeclareMathOperator{\diag}{diag\;}

    \DeclarePairedDelimiter\bra{\langle}{\rvert}
    \DeclarePairedDelimiter\ket{\lvert}{\rangle}
    \DeclarePairedDelimiter\abs{\lvert}{\rvert}
    \DeclarePairedDelimiter\ev{\langle}{\rangle}
    \DeclarePairedDelimiter\p{\lparen}{\rparen}
    \DeclarePairedDelimiter\s{\lbrack}{\rbrack}
    \DeclarePairedDelimiter\z{\lbrace}{\rbrace}

    % \everymath{\displaystyle} % biggify limits of inline sums and integrals
    \tikzstyle{circ} % usage: \node[circ, placement] (label) {text};
        = [draw, circle, fill=white, node distance=3cm, minimum height=2em]
    \definecolor{commentgreen}{rgb}{0,0.6,0}
    \lstset{
        basicstyle=\ttfamily\footnotesize,
        frame=single,
        numbers=left,
        showstringspaces=false,
        keywordstyle=\color{blue},
        stringstyle=\color{purple},
        commentstyle=\color{commentgreen},
        morecomment=[l][\color{magenta}]{\#}
    }

\begin{document}

\def\Snospace~{\S{}} % hack to remove the space left after autorefs
\renewcommand*{\sectionautorefname}{\Snospace}
\renewcommand*{\appendixautorefname}{\Snospace}
\renewcommand*{\figureautorefname}{Fig.}
\renewcommand*{\equationautorefname}{Eq.}
\renewcommand*{\tableautorefname}{Tab.}

\onehalfspacing

\section{Problem}

To the end of simulating breaking of internal gravity waves (IGW) in white
dwarfs (WDs), we begin with the toy problem of IGW breaking in a uniformly
stratified atmosphere in 2D. To begin with the simplest treatment, we consider
an incompressible fluid characterized by four dynamical variables $\rho, P, u_x,
u_z$.

\subsection{Physical Description}

We consider perturbations, not necessarily small, in a medium that at
equilibrium is at rest and has uniform density stratification
\begin{equation}
    \rho_0(x, z) \propto e^{-z/H}.
\end{equation}

For an incompressible fluid, the density of the fluid $\rho\p*{x, z, t}$
decomposes into this background stratification $\rho_0(x, z)$ and a small
perturbation $\rho_1\p*{x, z, t} \ll \rho_0$ ($\rho_1 \ll \rho_0$ is a necessary
condition for an incompressible treatment). Permitting a similar decomposition for
$P(x, z, t) = P_0(x, z) + P_1(x, z, t)$ and using hydrostatic equilibrium
$\rd{P_0(x, z)}{z} = -\rho g$ gives fluid equations
\begin{subequations}\label{eq:nonlin_feq}
    \begin{align}
        \vec{\nabla} \cdot \vec{u}_1 &= 0,\\
        \pd{\rho_1}{t} - u_{1z}\frac{\rho_0}{H} &=
            -\p*{\vec{u}_1 \cdot \vec{\nabla}} \rho_1,\\
        \pd{\vec{u}_1}{t} + \frac{\vec{\nabla}P_1}{\rho_0}
            + \frac{\rho_1}{\rho_0}\vec{g} &= -\p*{\vec{u}_1 \cdot \vec{\nabla}}
            \vec{u}_1.
    \end{align}
\end{subequations}
I notate $\vec{u} \equiv \vec{u_1}$ to identify that it is a perturbation
variable even though there is no distinction between the two without a
background flow.

\subsection{Linear Description}

If the perturbation is sufficiently small $u_1 \ll v_{ph}$ the phase velocity,
then we drop terms nonlinear in the perturbation quantities. The linearized
equations admit wave-like solutions described by
\begin{subequations}\label{eq:lin_sol}
    \begin{align}
        u_{1z} &= Ae^{z/2H}e^{i\p*{\vec{k} \cdot \vec{r} - \omega t}},\\
        \pd{u_{1x}}{x} &= -\pd{u_{1z}}{z},\\
        \pd{\rho_1}{t} &= u_{1z}\frac{\rho_0}{H},\\
        \pd{P}{x} &= -\rho_0 \pd{u_{1x}}{t},\\
        \omega^2 &= \frac{N^2k_x^2}{k_x^2 + k_z^2 + \frac{1}{4H^2}}.
    \end{align}
\end{subequations}
Note $N^2 = g/H$ the buoyancy frequency here.

\section{Linear Numerical Simulation using \texttt{Dedalus}}

In the WD problem, the IGW is generated deep within the WD and propagates
outwards. Since we are only concerned with its breaking behavior, we initially
seek to excite waves in a way agnostic of how the waves are generated. Put
another way, we don't want to generate the wave, we just want to see how the
wave evolves as it nears nonlinear amplitudes. We begin with the linear problem
to see whether our mechanism produces expected results.

The code described in this section is attached as \texttt{strat.py}, a minimum
working example that saves plots to \texttt{plots/} on the fly.

\subsection{Numerical Setup}

To this end, we simulate the linearized fluid equations
\begin{subequations}\label{eq:lin_feq}
    \begin{align}
        \vec{\nabla} \cdot \vec{u}_1 &= 0,\\
        \pd{\rho_1}{t} - u_{1z}\frac{\rho_0}{H} &= 0,\\
        \pd{\vec{u}_1}{t} + \frac{\vec{\nabla}P_1}{\rho_0}
            + \frac{\rho_1}{\rho_0}\vec{g} &= 0.
    \end{align}
\end{subequations}

The domain of our simulation is $x \in [0, H], z \in [0, 4H]$. We use a Fourier
basis in the $x$ and a Chebyshev in the $z$. We try with just $16$ modes in the
$x$ and $64$ modes in the $z$. Since we choose parameters of the problem such
that $\frac{2\pi}{k_x} = H, \frac{2\pi}{k_z} = -\frac{4H}{\pi}$\footnote{As
seen later, $k_z$ never explicitly enters into the problem description; instead,
we specify $\omega(k_x, k_z)$ using the desired $k_x, k_z$ values.}, this should
be plenty of resolution to resolve the IGW (I chose $k_z$ as an irrational value
to ensure any undesired reflections are far from any normal modes of the box).
The timescale is set such that $N = 1$.

\subsection{Reflection Suppression}

To simulate a purely outgoing wave, we adopt a damping zone near the top
boundary to suppress reflection. Specifically, we use
\begin{subequations}\label{eq:lin_feq_sponge}
    \begin{align}
        \vec{\nabla} \cdot \vec{u}_1 &= 0,\\
        \pd{\rho_1}{t} + f(z)\rho_1 - u_{1z}\frac{\rho_0}{H} &= 0,\\
        \pd{\vec{u}_1}{t} + f(z)\rho_1 + \frac{\vec{\nabla}P_1}{\rho_0}
            + \frac{\rho_1}{\rho_0}\vec{g} &= 0,
    \end{align}
\end{subequations}
where
\begin{equation}
    f(z) = f_0 \frac{\max\p*{z - z_0, 0}^2}{\p*{z_{\max} - z_0}^2}.
\end{equation}
We use $f_0 = 2, z_0 = 0.7 z_{\max}$, so the damping turns on quadratically
starting at $z_0 = 0.7z_{\max}$.

\subsection{Boundary Conditions}

To excite the wave, we adopt bottom boundary condition (BC)
\begin{equation}
    \at{\pd{u_{1z}}{z}}_{z = 0} = -Ak_z\cos\p*{k_xx - \omega t + \frac{1}{2H}}.
\end{equation}
The motivation for this comes from differentiating $\pd{u_{1z}}{z}$ in
\autoref{eq:lin_sol}. The reason our BC isn't just $u_{1z}(z = 0)$ is because the
initial condition $u_x(t = 0) = u_{1z}(t = 0) = 0$ is not divergence free.

Since we have the reflection suppression term above, it turns out we need fix
the $u_{1z}$ gauge, otherwise the $k_x = 0$ mode is numerically ill-conditioned
and causes \texttt{Dedalus} to blow up. Thus, we enforce $u_{1z}(k_x = 0) = 0$.

The remaining boundary condition (there are two $\partial_z$ terms in
\autoref{eq:lin_feq}) need only fix the pressure gauge, so we choose $P_1(k_x =
0) = 0$ and a throwaway $u_{1x}(k_x \neq 0) = 0$.

\subsection{Initial Conditions}

We initialize everything $\rho_1 = P_1 = u_{1x} = u_{1z} = 0$ everywhere.

\subsection{Results}

A snapshot of the simulation produced by \texttt{strat.py} is shown below in
\autoref{fig:lin_plot}. We see reasonable agreement between the analytical
solution and the simulation data as the wave propagates. More plots can be found
in the \texttt{plots/} folder.
\begin{figure}[t]
    \centering
    \begin{subfigure}{0.7\textwidth}
        \centering
        \includegraphics[width=\textwidth]{plots/t_20.png}
    \end{subfigure}

    \begin{subfigure}{0.7\textwidth}
        \centering
        \includegraphics[width=\textwidth]{plots/t_70.png}
    \end{subfigure}
    \caption{For line plots, red line indicates beginning of damping zone, blue
    the analytical solution and orange the simulation data. Variables are
    $u_{1z} = uz, u_{1x} = ux$, $F_z$ the energy flux, $E$ the energy, $P_1 = P,
    \rho_1 = rho$. Line plots of dynamical variables are slices through $x = 0$.
    }\label{fig:lin_plot}
\end{figure}

\section{Nonlinear Numerical Simulation}

\subsection{Low Amplitude}

To move to the nonlinear problem, we first introduce the nonlinear terms as in
\autoref{eq:nonlin_feq} but use small $A$ such that $u_{1z} \sim
Ae^{z_{\max}/2H} \ll v_{ph}$. All other problem setup is kept the same. The code
is in \texttt{strat\_nonlin\_lowA.py}. We expect agreement with the analytical
solution to the linear problem.

\subsection{Non-negligible amplitude: Instability}

We now increase the amplitude, and a numerical instability seems to develop. We
explore this with \texttt{strat\_nonlin.py}. An instability seems to grow near
$z = 0$.

\section{Attempted Solutions}

It seems that the observed instability has to do with the interface driving. A
few possible solutions have been considered:
\begin{itemize}
    \item I've considered other types of boundary conditions, such as
        $\ptd{u_{1z}}{z}$ or $\pd{P}{z}$, but all of these seem to produce
        similar behavior, instability near $z = 0$.

    \item Start the simulation with a small amplitude and use $z_{\max}$ very
        large, relying on the $u_{1z} \sim e^{-z/2H}$ to bring the wave to
        breaking amplitudes. This will require a much larger number of spatial
        modes to resolve the vertical wavelengths though and will be very slow.

    \item Bulk forcing has also been considered, where we use a driving term in
        the momentum equation $\pd{\vec{u}_1}{t}$ to generate IGW, then use
        reflection-suppressing damping zones on both sides of the $z$ domain.
        This seems to draw focus away from the focus on the breaking portion.

        I was able to get bulk forcing to work, however, and the results seem
        reasonable. It would be a less desirable entry point though, since we're
        trying to study just the breaking irrespective of the forcing mechanism.
\end{itemize}

\end{document}

### 2018-06-06 correspondence/180606-lecoanet/strat.py

@@ -21,20 +21,20 @@ XMAX = H
 ZMAX = 5 * H
 KX = 2 * np.pi / H
 KZ = -(np.pi / 2) * np.pi / H
-G = 10
+G = 1
 
 OMEGA = get_omega(G, H, KX, KZ)
 VPH_X, VPH_Z = get_vph(G, H, KX, KZ)
 
-num_timesteps = 2e3
+num_timesteps = 4e2
 DAMP_START = ZMAX * 0.7 # start damping zone
-T_F = -ZMAX / VPH_Z # VPH_Z < 0
+T_F = -ZMAX / VPH_Z * 2 # VPH_Z < 0
 DT = T_F / num_timesteps
 N_X = 16
 N_Z = 64
 RHO0 = 1
 A = 0.05
-NUM_SNAPSHOTS = 1e3
+NUM_SNAPSHOTS = 100
 
 def get_analytical_sponge(name, z_pts, t):
     uz_anal = A * np.exp(z_pts / (2 * H)) *\
@@ -76,7 +76,7 @@ if __name__ == '__main__':
     rho0['g'] = RHO0 * np.exp(-z / H)
     problem.parameters['rho0'] = rho0
 
-    sponge_strength = 10
+    sponge_strength = 0.3
     z = domain.grid(1)
 
     # sponge field
@@ -88,12 +88,10 @@ if __name__ == '__main__':
 
     problem.parameters['sponge'] = sponge
     problem.add_equation('dx(ux) + dz(uz) = 0')
-    problem.add_equation('dt(rho) - rho0 * uz / H + sponge * rho'
-                         + '= -ux * dx(rho) - uz * dz(rho)')
-    problem.add_equation('dt(ux) + dx(P) / rho0 + sponge * ux'
-                         + '= -ux * dx(ux) - uz * dz(ux)')
-    problem.add_equation('dt(uz) + dz(P) / rho0 + rho * g / rho0 + sponge * uz'
-                         + '= -ux * dx(uz) - uz * dz(uz)')
+    problem.add_equation('dt(rho) - rho0 * uz / H + sponge * rho = 0')
+    problem.add_equation('dt(ux) + dx(P) / rho0 + sponge * ux = 0')
+    problem.add_equation(
+        'dt(uz) + dz(P) / rho0 + rho * g / rho0 + sponge * uz = 0')
 
     problem.add_bc('left(P) = 0', condition='nx == 0')
[…]

### 2018-06-06 correspondence/180606-lecoanet/strat_nonlin.py

@@ -1,0 +1,215 @@
+from collections import defaultdict
+import logging
+import matplotlib
+import matplotlib.pyplot as plt
+import numpy as np
+import os
+
+from dedalus import public as de
+from dedalus.extras.flow_tools import CFL
+from dedalus.extras.plot_tools import quad_mesh, pad_limits
+
+def get_omega(g, h, kx, kz):
+    return np.sqrt((g / h) * kx**2 / (kx**2 + kz**2 + 0.25 / h**2))
+
+def get_vph(g, h, kx, kz):
+    norm = abs(get_omega(g, h, kx, kz)) / (kx**2 + kz**2)
+    return norm * kz, norm * kz
+
+H = 1
+XMAX = H
+ZMAX = 5 * H
+KX = 2 * np.pi / H
+KZ = -(np.pi / 2) * np.pi / H
+G = 1
+
+OMEGA = get_omega(G, H, KX, KZ)
+VPH_X, VPH_Z = get_vph(G, H, KX, KZ)
+
+num_timesteps = 2e3
+DAMP_START = ZMAX * 0.7 # start damping zone
+T_F = -ZMAX / VPH_Z # VPH_Z < 0
+DT = T_F / num_timesteps
+N_X = 16
+N_Z = 64
+RHO0 = 1
+A = 0.05
+NUM_SNAPSHOTS = 1e3
+
+def get_analytical_sponge(name, z_pts, t):
+    uz_anal = A * np.exp(z_pts / (2 * H)) *\
+        np.cos(KZ * z_pts - OMEGA * t)
+    rho0 = RHO0 * np.exp(-z_pts / H)
+    analyticals = {
+        'uz': uz_anal,
+        'ux': -KZ / KX * uz_anal,
+        'rho': -rho0 * A / (H * OMEGA) *\
+            np.exp(z_pts / (2 * H)) *\
+            np.sin(KZ * z_pts - OMEGA * t),
+        'P': -rho0 * OMEGA / KX**2 * KZ *\
+            uz_anal,
+    }
+    return analyticals[name]
+
+if __name__ == '__main__':
+    logger = logging.getLogger(__name__)
+
+    x_basis = de.Fourier('x', N_X, interval=(0, XMAX), dealias=3/2)
+    z_basis = de.Chebyshev('z', N_Z, interval=(0, ZMAX), dealias=3/2)
+    domain = de.Domain([x_basis, z_basis], np.float64)
+    x = domain.grid(0)
+    z = domain.grid(1)
[…]

### 2018-06-06 correspondence/180606-lecoanet/strat_nonlin_lowA.py

@@ -21,7 +21,7 @@ XMAX = H
 ZMAX = 5 * H
 KX = 2 * np.pi / H
 KZ = -(np.pi / 2) * np.pi / H
-G = 10
+G = 1
 
 OMEGA = get_omega(G, H, KX, KZ)
 VPH_X, VPH_Z = get_vph(G, H, KX, KZ)
@@ -33,7 +33,7 @@ DT = T_F / num_timesteps
 N_X = 16
 N_Z = 64
 RHO0 = 1
-A = 0.05
+A = 0.00005
 NUM_SNAPSHOTS = 100
 
 def get_analytical_sponge(name, z_pts, t):
@@ -76,22 +76,23 @@ if __name__ == '__main__':
     rho0['g'] = RHO0 * np.exp(-z / H)
     problem.parameters['rho0'] = rho0
 
-    sponge_strength = 10
+    sponge_strength = 2
     z = domain.grid(1)
 
     # sponge field
     sponge = domain.new_field()
     sponge.meta['x']['constant'] = True
-    sponge['g'] = sponge_strength * np.maximum(
-        1 - (z - ZMAX)**2 / (DAMP_START - ZMAX)**2,
-        np.zeros(np.shape(z)))
+    sponge['g'] = sponge_strength * (
+        np.maximum(z - DAMP_START, 0) ** 2 / (ZMAX - DAMP_START) ** 2)
 
     problem.parameters['sponge'] = sponge
     problem.add_equation('dx(ux) + dz(uz) = 0')
-    problem.add_equation('dt(rho) - rho0 * uz / H + sponge * rho = 0')
-    problem.add_equation('dt(ux) + dx(P) / rho0 + sponge * ux = 0')
-    problem.add_equation(
-        'dt(uz) + dz(P) / rho0 + rho * g / rho0 + sponge * uz = 0')
+    problem.add_equation('dt(rho) - rho0 * uz / H + sponge * rho'
+                         + '= -ux * dx(rho) - uz * dz(rho)')
+    problem.add_equation('dt(ux) + dx(P) / rho0 + sponge * ux'
+                         + '= -ux * dx(ux) - uz * dz(ux)')
+    problem.add_equation('dt(uz) + dz(P) / rho0 + rho * g / rho0 + sponge * uz'
+                         + '= -ux * dx(uz) - uz * dz(uz)')
 
     problem.add_bc('left(P) = 0', condition='nx == 0')
     problem.add_bc('right(uz) = 0', condition='nx != 0')
[…]

### 2018-06-06 correspondence/180606-lecoanet/brief.tex
final version of code

@@
 we specify $\omega(k_x, k_z)$ using the desired $k_x, k_z$ values.}, this should
-be plenty of resolution to resolve the IGW (I chose $k_z$ as an irrational value
-to ensure any undesired reflections are far from any normal modes of the box).
-The timescale is set such that $N = 1$.
+be plenty of resolution to resolve the IGW\footnote{I chose $k_z$ as an
+irrational value to ensure any undesired reflections are far from any normal
+modes of the box.}. The timescale is set by $N = 1$.
 
@@
 \end{equation}
-We use $f_0 = 2, z_0 = 0.7 z_{\max}$, so the damping turns on quadratically
+We use $f_0 = 1, z_0 = 0.7 z_{\max}$, so the damping turns on quadratically
 starting at $z_0 = 0.7z_{\max}$.
@@
 \end{equation}
-The motivation for this comes from differentiating $\pd{u_{1z}}{z}$ in
-\autoref{eq:lin_sol}. The reason our BC isn't just $u_{1z}(z = 0)$ is because the
-initial condition $u_x(t = 0) = u_{1z}(t = 0) = 0$ is not divergence free.
+We use $A = 0.05$. The reason our BC isn't just $u_{1z}(z = 0)$ is because the
+initial condition $u_x(t = 0) = u_{1z}(t = 0) = 0$ is not divergence free and
+produces high-$k$ modes in $u_x$. We just differentiate $\pd{u_{1z}}{z}$ in
+\autoref{eq:lin_sol} and take real part.
 
@@
     \centering
-    \begin{subfigure}{0.7\textwidth}
+    \begin{subfigure}{0.75\textwidth}
         \centering
@@
     \end{subfigure}
-
-    \begin{subfigure}{0.7\textwidth}
+    \begin{subfigure}{0.75\textwidth}
         \centering
@@
 \autoref{eq:nonlin_feq} but use small $A$ such that $u_{1z} \sim
-Ae^{z_{\max}/2H} \ll v_{ph}$. All other problem setup is kept the same. The code
-is in \texttt{strat\_nonlin\_lowA.py}. We expect agreement with the analytical
-solution to the linear problem.
+Ae^{z_{\max}/2H} \ll v_{ph}$; we use $A = \scinot{5}{-5}$. All other problem
+setup is kept the same. The code is in \texttt{strat\_nonlin\_lowA.py}. We
+expect agreement with the analytical solution to the linear problem, and this is
+shown well in \autoref{fig:nonlin_plot_lowA}. Further plots are available in
+\texttt{plots\_nonlin\_lowA/}.
+\begin{figure}[t]
+    \centering
+    \begin{subfigure}{0.75\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{plots_nonlin_lowA/t_20.png}
+    \end{subfigure}
+    \begin{subfigure}{0.75\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{plots_nonlin_lowA/t_70.png}
+    \end{subfigure}
+    \caption{Same plots as \autoref{fig:lin_plot} but with nonlinear terms added
+    back to the fluid equations and very small $A = \scinot{5}{-5}$.
+    }\label{fig:nonlin_plot_lowA}
+\end{figure}
 
@@
 
-We now increase the amplitude, and a numerical instability seems to develop. We
-explore this with \texttt{strat\_nonlin.py}. An instability seems to grow near
-$z = 0$.
+We now increase the amplitude back to $A = 0.05$, and a numerical instability
+seems to develop. We explore this with \texttt{strat\_nonlin.py}. An instability
+seems to grow near $z = 0$, and two plots of its evolution can be seen in
+\autoref{fig:nonlin_plot}. Further plots can be found in
+\texttt{plots\_nonlin/}.
+\begin{figure}[t]
+    \centering
+    \begin{subfigure}{0.75\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{plots_nonlin/t_70.png}
+    \end{subfigure}
+    \begin{subfigure}{0.75\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{plots_nonlin/t_77.png}
+    \end{subfigure}
+    \caption{Nonlinear equations with $A = 0.05$ as it diverges near $z = 0$.
+    }\label{fig:nonlin_plot}
+\end{figure}
 
@@
 \begin{itemize}
+    \item Doubling the temporal resolution doesn't seem to change when the
+        instability sets in. Doubling the spatial resolution makes it set in
+        earlier, and halving makes it set in later.
+
     \item I've considered other types of boundary conditions, such as

### 2018-06-06 correspondence/180606-lecoanet/strat_nonlin_lowA.py

@@ -76,7 +76,7 @@ if __name__ == '__main__':
     rho0['g'] = RHO0 * np.exp(-z / H)
     problem.parameters['rho0'] = rho0
 
-    sponge_strength = 2
+    sponge_strength = 1
     z = domain.grid(1)
 
     # sponge field

### 2018-06-11 writups/notes.tex
update notes w/ TODO, results from P
stash chebyshev notes

@@
     \item Consider BCs only on $P$, the only other choice of dynamical variable.
+        This seems to also exhibit instabilities near $z = 0$.
 
@@
         \cdot \vec{u} = 0$ constraint. The correct choice of BCs in this case is
-        discussed in Nordstrom et al, 2006.
+        discussed in Nordstrom et al, 2006. We choose not to investigate this.
 
@@
 \end{itemize}
+
+TODO Can taking $\rho_1 \ll \rho_0$, giving us a full rank operator w/o doing
+the janky double add thing, solve in the incompressible limit? Otherwise, go
+anelastic.
 
@@
 
+\chapter{Numerical Results}
+
+\section{Chebyshev Polynomials}
+
+Note that, unlike a Fourier basis, a Chebyshev basis has $N^2$ grid spacing near
+its boundaries. By this, we mean that for the Chebyshev polynomials $T_n(x)$,
+the spacing between zeros of $T_n(x)$ near $x = \pm 1$ scales like $n^{-2}$.
+This contrasts with Fourier series for which $\psi_n(x) = \cos n\pi x, \sin n\pi
+x$ for which the zeros are spaced $\Delta x = \frac{1}{n}$. This is important
+because the CFL condition for PDEs requires that $u\frac{\Delta t}{\Delta x} <
+C$ for some constant $C$, otherwise the simulation will move fluid elements more
+than $\Delta x$ in a single timestep, an illegal operation for the basis
+function.
+
+If we consider instead $T_n(x)$ satisfying $T_n(\cos \theta) = \cos n\theta$,
+then within some spacing of the edge of the domain $\cos\theta = 1 - \Delta x$
+then $\Delta x \approx \theta^2/2$ then the number of zeros seen is the number
+of solutions $\cos \phi = 0, \phi = n\theta = n\sqrt{2\Delta x}$ which requires
+$\sim k\pi \in [0, n\sqrt{2\Delta x}]$. Thus, for any interval $\Delta x$, the
+number of zeros contained scales with $n^2$!
+
 \end{document}

### 2018-06-11 sims/2d_3_bulk_force/strat.py

@@ -7,7 +7,7 @@ from strat_helper import *
 
 N_PARALLEL = 20
 H = 1
-num_timesteps = 4e3
+num_timesteps = 5e2
 
 XMAX = H
 ZMAX = 3 * H
@@ -16,9 +16,9 @@ KZ = -(np.pi) * np.pi / H
 G = (KX**2 + KZ**2 + 1 / (4 * H**2)) / KX**2 * (2 * np.pi)**2 * H # omega = 2pi
 OMEGA = get_omega(G, H, KX, KZ)
 _, VPH_Z = get_vph(G, H, KX, KZ)
-T_F = abs(ZMAX / VPH_Z) * 8
+T_F = abs(ZMAX / VPH_Z) * 2
 DT = T_F / num_timesteps
-NUM_SNAPSHOTS = 400
+NUM_SNAPSHOTS = 100
 
 PARAMS_RAW = {'XMAX': XMAX,
               'ZMAX': ZMAX,
@@ -33,10 +33,10 @@ PARAMS_RAW = {'XMAX': XMAX,
               'RHO0': 1,
               'G': G,
               'A': 0.005,
-              'F': 0.1,
-              'SPONGE_STRENGTH': 10,
-              'SPONGE_START_HIGH': 0.7 * ZMAX,
-              'SPONGE_START_LOW': 0.3 * ZMAX,
+              'F': 0.01,
+              'SPONGE_STRENGTH': 30,
+              'SPONGE_START_HIGH': 0.8 * ZMAX,
+              'SPONGE_START_LOW': 0.2 * ZMAX,
               'NUM_SNAPSHOTS': NUM_SNAPSHOTS}
 
 def build_interp_params(interp_x, interp_z, dt=DT, overrides=None):
@@ -61,21 +61,21 @@ if __name__ == '__main__':
         (get_solver, setup_problem, zero_ic,
          'F1',
          build_interp_params(4, 4)),
-        (get_solver, setup_problem, zero_ic,
-         'F2',
-         build_interp_params(4, 4, overrides={'F': 0.3})),
-        (get_solver, setup_problem, zero_ic,
-         'F3',
-         build_interp_params(4, 4, overrides={'F': 0.6})),
-        (get_solver, setup_problem, zero_ic,
-         'F4',
-         build_interp_params(4, 4, overrides={'F': 1.0})),
-        (get_solver, setup_problem, zero_ic,
-         'F5',
[…]

### 2018-06-11 sims/2d_3_bulk_force/strat_helper.py

@@ -77,19 +77,18 @@ def setup_problem(problem, domain, params):
         'dt(ux) + sponge * ux + dx(P) / rho0' +
         '= - ux * dx(ux) - uz * dz(ux) +'
         # '='
-        'F * KX / KZ * exp(-(z - 0.15 * ZMAX)**2 / (2 * 0.1**2)) *' +
+        'F * KX * exp(-(z - 0.3 * ZMAX)**2 / (2 * 0.1**2)) *' +
             'sin(KX * x - omega * t)')
     problem.add_equation(
         'dt(uz) + sponge * uz + dz(P) / rho0 + rho * g / rho0' +
         '= - ux * dx(uz) - uz * dz(uz) +' +
         # '='
-        'F * exp(-(z - 0.15 * ZMAX)**2 / (2 * 0.1**2)) *' +
-            'sin(KX * x - omega * t)')
+        'F * (z - 0.3 * ZMAX) / 0.1**1 * exp(-(z - 0.3 * ZMAX)**2 / (2 * 0.1**2)) *' +
+            'cos(KX * x - omega * t)')
 
     z = domain.grid(1)
     x = domain.grid(0)
-    problem.add_bc('left(uz) = 0', condition='nx != 0')
-    problem.add_bc('left(P) = 0', condition='nx == 0')
+    problem.add_bc('left(P) = 0')
     problem.add_bc('right(uz) = 0')
 
 ###
@@ -243,7 +242,7 @@ def plot(get_solver, setup_problem, name, params):
                   for i in ['uz', 'ux', 'rho1', 'P1']]
     n_cols = 3
     n_rows = 3
-    plot_stride = 1
+    plot_stride = 2
 
     if os.path.exists('%s.mp4' % name):
         print('%s.mp4 already exists, not regenerating' % name)

### 2018-06-17 writups/notes.tex
check in partial notes

@@
 
+\section{Volumetric Forcing}
+
+Due to the structure of \lstinline{Dedalus}, we must use Chebyshev polynomials
+in the vertical direction (to be able to enforce BCs). But as discussed in
+\autoref{s:chebyshev}, the grid spacing of the Chebyshev polynomials near
+boundaries goes as $1/N^2$, which means that interface forcing necessitates an
+extremely strict Courant condition to capture evolution near the interface; this
+is the explanation for the divergences we were seeing in earlier sections.
+
+Instead, we employ volumetric forcing. As a toy problem of seeing how this
+works, we will study the restricted 1D wave equation.
+
+\subsection{1D Wave Equation}
+
+Consider the forced 1D wave equation (we set $c = 1$)
+\begin{equation}
+    \ptd{u(x, t)}{t} - \ptd{u(x, t)}{x} = f(x, t).
+\end{equation}
+Note that the only allowed propagation modes here are $u \sim e^{iC\p*{x \pm
+t}}$ for some constant $C$.
+
+We wish to choose $f(x, t)$ such that $u\p*{x \to -\infty} = 0$ while $u\p*{x
+\to +\infty} \sim e^{i\p*{x - t}}$. It is clear that if we consider the Fourier
+transform $\tilde{f}(k, t)$, the response in $\tilde{u}(k, t)$ will also contain
+components for each of the nonzero $\tilde{f}(k, t)$ components. Thus, if we
+want a pure $e^{i(x - t)}$ outgoing wave, we should have a pure $\tilde{f}(k, t)
+\propto \delta\p*{k - 1}$. However, it is clear that driving the oscillator
+exactly on resonance produces divergent oscillation amplitudes, and moreover it
+is impossible to drive with an $f(x, t)$ that doesn't vanish as $x \to \pm
+\infty$.
+
+It is nonetheless clear that $\tilde{f}(k, t)$ should be strongly peaked at $k =
+1$ and small elsewhere. We can thus consider a form $f(x, t) \propto
+e^{-x^2/2\sigma^2} e^{i(x - t)}$ such that $\sigma \gg 1$, i.e.\ the envelope
+contains many wavelengths. The Fourier transform $\tilde{f}(k, t)$ is then a
+Gaussian with width $1/\sigma \ll 1$ centered at $k = 1$ (convolution).
+
+More precisely, consider $f(x, t) = f_0e^{-x^2/2\sigma^2} e^{i(x - t)}$ and $u(x,
+t) = u_0(x)e^{i(x - t)}$ such that $u_0(x \to \infty) = 0$. Substituting this
+into the inhomogeneous PDE gives
+\begin{equation}
+    i\pd{u_0(x)}{x} - \ptd{u_0(x)}{x} = f_0e^{-x^2/2\sigma^2}.
+\end{equation}
+We note that if $\sigma \gg 1$ then roughly $\pd{}{x} \ll 1$ or
+\begin{equation}
+    \pd{u_0(x)}{x} \approx -if_0e^{-x^2/2\sigma^2}.
+\end{equation}
+Enforcing $u_0(x \to -\infty) = 0$, then $\abs*{u_0(x \to +\infty)} =
+f_0\sigma\sqrt{2\pi}$.
+
 \clearpage
@@
 
-\section{Chebyshev Polynomials}
+\section{Chebyshev Polynomials}\label{s:chebyshev}
 

### 2018-07-16 writups/notes.tex
update refs in writeup

@@
         \centering
-        \includegraphics[width=\textwidth]{../sims/2d_0_no_g/no_g_P_t50.png}
+        \includegraphics[width=\textwidth]{../sims_old/2d_0_no_g/no_g_P_t50.png}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{../sims/2d_0_no_g/no_g_ux_t50.png}
+        \includegraphics[width=\textwidth]{../sims_old/2d_0_no_g/no_g_ux_t50.png}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{../sims/2d_0_no_g/no_g_uz_t50.png}
+        \includegraphics[width=\textwidth]{../sims_old/2d_0_no_g/no_g_uz_t50.png}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{../sims/2d_0_no_g/no_g_P_t100.png}
+        \includegraphics[width=\textwidth]{../sims_old/2d_0_no_g/no_g_P_t100.png}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{../sims/2d_0_no_g/no_g_ux_t100.png}
+        \includegraphics[width=\textwidth]{../sims_old/2d_0_no_g/no_g_ux_t100.png}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{../sims/2d_0_no_g/no_g_uz_t100.png}
+        \includegraphics[width=\textwidth]{../sims_old/2d_0_no_g/no_g_uz_t100.png}
     \end{subfigure}
@@
     \centering
-    \includegraphics[width=0.7\textwidth]{../sims/2d_1_strat/agree_plots/sponge_2.png}
+    \includegraphics[width=0.7\textwidth]{../sims_old/2d_1_strat/agree_plots/sponge_2.png}
     \caption{Note the jaggedness in $u_x$ induced by a Dirichlet BC on

### 2018-07-17 writups/notes.tex
update writeups, clean up dictionary merging

@@
 
+\chapter{Simulation Setup}
+
+We describe now some concerns that arise in setting up the simulation.
+
 \section{Volumetric Forcing}
@@
 extremely strict Courant condition to capture evolution near the interface; this
-is the explanation for the divergences we were seeing in earlier sections.
+is the explanation for the divergences we were seeing in earlier sections. In
+fact, in the literature, the spectral CFL condition meets a $\Delta t_{\max}
+\sim 1/N^2$ scaling thanks to poor bounds near the boundary.
 
@@
 
-\subsection{1D Wave Equation}
+\subsection{1D Wave Equation (Deprecated)}
 
@@
 f_0\sigma\sqrt{2\pi}$.
+
+\subsection{Narrow-band forcing}
+
+Instead of doing this on-frequency forcing, we choose instead to use a narrow
+forcing zone $f(x, t) \sim N_{\sigma}(z)e^{it}$ where $\sigma$ is not large.
+Specifically, we need $\sigma$ to be well-resolved spectrally, but we need
+$\tilde{f}(k_z, t)$ not to be small. Thus, since $1/k_z$ must be resolved by the
+simulation, choosing $\sigma = 1/k_z$ is a natural choice.
+
+Note that this implies that waves will be excited going both upwards and
+downwards, as $\tilde{f}(k_z) = \tilde{f}(-k_z)$. We can however place the
+forcing zone near the bottom boundary so that downwards-propagating waves will
+be attenuated by damping zones.
+
+\section{Timestepping}
+
+\subsection{Signatures of Insufficient Temporal Resolution}
+
+When running these simulations, it can occasionally be seen that the waves do
+not grow in amplitude like $e^{z/2H}$, such that the energy and flux are not
+constant (or the waves will grow too quickly; this is dependent on the exact
+timestepping used). This is caused by using an overly-long timestep, as can be
+seen with the below illustration.
+
+We study the 1D advection equation $\psi_t = c\psi_x$ corresponding to
+rightwards propagating waves. Consider first order implicit Euler timestepping
+for wave mode $\psi_x = ik\psi$, then
+\begin{equation}
+    \psi_{i + 1} - \psi_i = ick\Delta t\psi_{i + 1}.
+\end{equation}
+Solving yields
+\begin{align}
+    \psi_{i + 1} &= \p*{1 - ick\Delta t}^{-1}\psi_i,\nonumber\\
+        &= \p*{1 - i\omega \Delta t}^{-1}\psi_i.
+\end{align}
+If we then want to step forwards for some total time $T$, with $N \Delta t = T$
+the number of timesteps taken, then we approximate
+\begin{align}
+    \psi(t_0 + T) \approx \p*{1 - i\omega \frac{T}{N}}^{-N}\psi(t_0).
+\end{align}
+It is worth noting that in the $N \to \infty$ limit we recover $\psi(t_0 + T) =
+e^{i\omega T}\psi(t_0)$.
+
+However, consider if $\omega \Delta t \lesssim 1$, then we should see the wave
+amplitude decrease over time like
+\begin{equation}
+    \abs*{\psi(t_0 + T)} \approx \p*{1 + \omega^2 \Delta t^2}^{-N/2}\psi(t_0).
+\end{equation}
+Thus, since we have an implicit first order scheme, the wave amplitude
+\emph{decreases} over time spuriously. An explicit method would produce an
+increase, but \lstinline{Dedalus} uses implicit timestepping. The exact rate of
+decrease depends on the timestepper used, but it is clear that $\omega \Delta t
+\ll 1$ is necessary to avoid this spurious decrease (in numerics language, ``the
+oscillation must be well-resolved'').
 

### 2018-07-17 sims/2d_0_lin/strat.py

@@ -10,17 +10,17 @@ XMAX = H
 ZMAX = 10 * H
 KX = 8 * np.pi / H
 KZ = -20 / H
-G = (KX**2 + KZ**2 + 1 / (4 * H**2)) / KX**2 * (2 * np.pi)**2 * H # omega = 2pi
-OMEGA = get_omega(G, H, KX, KZ)
-_, VG_Z = get_vg(G, H, KX, KZ)
+g = (KX**2 + KZ**2 + 1 / (4 * H**2)) / KX**2 * (2 * np.pi)**2 * H # omega = 2pi
+OMEGA = get_omega(g, H, KX, KZ)
+_, VG_Z = get_vg(g, H, KX, KZ)
 
 T_F = abs(ZMAX / VG_Z) * 1.2
-DT = 0.06
-NUM_SNAPSHOTS = 100
+DT = 0.02
+NUM_SNAPSHOTS = 200
 
 PARAMS_RAW = {'XMAX': XMAX,
               'ZMAX': ZMAX,
-              'N_X': 64,
+              'N_X': 128,
               'N_Z': 512,
               'T_F': T_F,
               'DT': DT,
@@ -29,12 +29,14 @@ PARAMS_RAW = {'XMAX': XMAX,
               'KZ': KZ,
               'H': H,
               'RHO0': 1,
-              'G': G,
+              'g': g,
               'A': 0.005,
               'F': 0.005,
+              'S': 1 / KZ,
+              'Z0': 0.2 * ZMAX,
               'SPONGE_STRENGTH': 1,
-              'SPONGE_START_HIGH': 0.9 * ZMAX,
-              'SPONGE_START_LOW': 0.1 * ZMAX,
+              'SPONGE_HIGH': 0.9 * ZMAX,
+              'SPONGE_LOW': 0.1 * ZMAX,
               'NUM_SNAPSHOTS': NUM_SNAPSHOTS}
 
 def build_interp_params(interp_x, interp_z, dt=DT, overrides=None):
@@ -56,7 +58,7 @@ def run(ic, name, params_dict):
 if __name__ == '__main__':
     tasks = [
         (zero_ic, 'F0',
-         build_interp_params(1, 1, overrides={'F': 0.00003})),
+         build_interp_params(1, 1, overrides={'F': 0.00002})),
     ]
     if '-plot' not in sys.argv:
         for task in tasks:

### 2018-07-17 sims/2d_0_nonlin/strat.py

@@ -7,16 +7,16 @@ from strat_helper import *
 
 H = 1
 XMAX = H
-ZMAX = 15 * H
+ZMAX = 10 * H
 KX = 8 * np.pi / H
 KZ = -20 / H
 G = (KX**2 + KZ**2 + 1 / (4 * H**2)) / KX**2 * (2 * np.pi)**2 * H # omega = 2pi
 OMEGA = get_omega(G, H, KX, KZ)
-_, VPH_Z = get_vph(G, H, KX, KZ)
+_, VG_Z = get_vg(G, H, KX, KZ)
 
-T_F = abs(ZMAX / VPH_Z)
-DT = 0.1
-NUM_SNAPSHOTS = 100
+T_F = abs(ZMAX / VG_Z) * 1.2
+DT = 0.02
+NUM_SNAPSHOTS = 200
 
 PARAMS_RAW = {'XMAX': XMAX,
               'ZMAX': ZMAX,

### 2018-07-17 sims/2d_0_nonlin/strat_helper.py

@@ -22,8 +22,8 @@ SNAPSHOTS_DIR = 'snapshots_%s'
 def get_omega(g, h, kx, kz):
     return np.sqrt((g / h) * kx**2 / (kx**2 + kz**2 + 0.25 / h**2))
 
-def get_vph(g, h, kx, kz):
-    norm = get_omega(g, h, kx, kz) / (kx**2 + kz**2)
+def get_vg(g, h, kx, kz):
+    norm = get_omega(g, h, kx, kz) / (kx**2 + kz**2) * (kz / kx)
     return norm * kx, norm * kz
 
 def zero_ic(solver, domain, params):
@@ -51,7 +51,6 @@ def get_solver(params):
     problem.parameters['KX'] = params['KX']
     problem.parameters['KZ'] = params['KZ']
     problem.parameters['NU'] = params['NU']
-    problem.parameters['RHO0'] = params['RHO0']
     problem.parameters['omega'] = params['OMEGA']
     problem.parameters['ZMAX'] = params['ZMAX']
     problem.parameters['SPONGE_STRENGTH'] = params['SPONGE_STRENGTH']
@@ -75,7 +74,7 @@ def get_solver(params):
         'dt(rho) - rho0 * uz / H' +
         '= - sponge * rho - ux * dx(rho) - uz * dz(rho) +' +
         'F * exp(-(z - Z0)**2 / (2 * S**2)) *' +
-            'cos(KX * x + KZ * z - omega * t)')
+            'cos(KX * x - omega * t)')
     problem.add_equation(
         'dt(ux) + dx(P) / rho0' +
         '= - sponge * ux - ux * dx(ux) - uz * dz(ux)')
@@ -88,15 +87,11 @@ def get_solver(params):
     problem.add_bc('right(P) = 0', condition = 'nx == 0')
 
     # Build solver
-    solver = problem.build_solver(de.timesteppers.RK222)
+    solver = problem.build_solver(de.timesteppers.RK443)
     solver.stop_sim_time = params['T_F']
     solver.stop_wall_time = np.inf
     solver.stop_iteration = np.inf
     return solver, domain
-
-###
-### ENTRY POINTS
-###
 
 def run_strat_sim(set_ICs, name, params):
     snapshots_dir = SNAPSHOTS_DIR % name
@@ -170,7 +165,7 @@ def load(name, params):
         params['RHO0'] * (np.exp(-z / params['H']) - 1) *\
         params['G'] * params['H']
 
-    state_vars['E'] = state_vars['rho'] * \
[…]

### 2018-07-24 writeups/notes.tex
use analytical estimate for F:Uz1 to plot, init nonlin sims

--- 2018-07-17 writups/notes.tex
@@
     \newcommand*{\scinot}[2]{#1\times10^{#2}}
-    \newcommand*{\bra}[1]{\left<#1\right|}
-    \newcommand*{\ket}[1]{\left|#1\right>}
     \newcommand*{\dotp}[2]{\left<#1\,\middle|\,#2\right>}
@@
     \newcommand*{\md}[2]{\frac{\mathrm{D}#1}{\mathrm{D}#2}}
-    \newcommand*{\norm}[1]{\left|\left|#1\right|\right|}
-    \newcommand*{\abs}[1]{\left|#1\right|}
     \newcommand*{\pvec}[1]{\vec{#1}^{\,\prime}}
@@
     \newcommand*{\bm}[1]{\boldsymbol{\mathbf{#1}}}
-    \newcommand*{\expvalue}[1]{\left<#1\right>}
-    \newcommand*{\ang}[0]{\text{\AA}}
-    \newcommand*{\mum}[0]{\upmu \mathrm{m}}
+    \newcommand*{\ang}[0]{\;\text{\AA}}
+    \newcommand*{\mum}[0]{\;\upmu \mathrm{m}}
     \newcommand*{\at}[1]{\left.#1\right|}
@@
 
+    \DeclarePairedDelimiter\bra{\langle}{\rvert}
+    \DeclarePairedDelimiter\ket{\lvert}{\rangle}
+    \DeclarePairedDelimiter\abs{\lvert}{\rvert}
+    \DeclarePairedDelimiter\ev{\langle}{\rangle}
     \DeclarePairedDelimiter\p{\lparen}{\rparen}
@@
 
-Another formula given by Sutherland 2011 is $\pd{\expvalue{E}}{t} = v_{g, z}
-\expvalue{E}$ where we average over $x$ one wavelength. I wasn't able to show
+Another formula given by Sutherland 2011 is $\pd{\ev*{E}}{t} = v_{g, z}
+\ev*{E}$ where we average over $x$ one wavelength. I wasn't able to show
 that these two agree in general; Sutherland's formula seems to be derived for a
@@
 
-One can imagine, for this matter, that $\rd{}{t}$ can be wrapped in the $L_{ij}$
+One can imagine, for this matter, that $\pd{}{t}$ can be wrapped in the $L_{ij}$
 operator and the timestepping computed implicitly; I believe this is what
@@
 
-Both of these procedures require $\bm{L}$ linear operator or $\rd{}{t} + \bm{L}$
+Both of these procedures require $\bm{L}$ linear operator or $\pd{}{t} + \bm{L}$
 to be full rank, or invertible. This is enforced analytically by requiring the
@@
 
+\section{Damping Zone}
+
+Since our code is a spectral code, instead of using the quadratic damping zone
+presented from before (whose second derivative is discontinuous over the entire
+domain), we prefer instead an $\arctan$ profile, such that
+\begin{equation}
+    \Gamma(z) \propto 0.5 * \p*{
+        2 + \tanh C\p*{\frac{z - z_t}{L_z - z_t}}
+            + \tanh C\p*{\frac{z - z_b}{z_b}}}.
+\end{equation}
+This is a profile that mostly vanishes $z \in [z_b, z_t]$ but is unity for $z
+\lesssim z_b, z \gtrsim z_t$, with steepness of transition governed by factor
+$C$. While this profile may have reflection issues if $C$ is chosen to be too
+steep, it is smooth and so has exponentially convergent spectral expansion and
+empirically seems to do alright.
+
+It is moreover prudent to note that, since implicit calculations are expensive,
+as long as the damping timescale $\Gamma^{-1}$ is everywhere resolved it is
+cheaper to put $\Gamma \rho_1, \Gamma \vec{u}_1$ on the right hand sides of
+their equations in Dedalus, which is timestepped explicitly.
+
 \section{Volumetric Forcing}
@@
 be attenuated by damping zones.
+
+Let us derive the resultant equations of motion. In the interest of maintaining
+a divergence-free velocity field, we place the driving term on the density
+equation. Since the forcing is purely to excite waves and not physical (for the
+time being), it should not worry us that driving the density equation is
+ill-justifiable physically.
+
+Consider the linearized system of equations (to examine the excited perturbation
+in the linear legime)
+\begin{subequations}\label{se:vol_drive}
+    \begin{align}
+        \vec{\nabla} \cdot \vec{u} &= 0,\\
+        \pd{\vec{u}_1}{t} + \frac{\vec{\nabla}P}{\rho_0}
+            + \frac{\rho_1 \vec{g}}{\rho_0}
+            &= 0,\\
+        \pd{\rho_1}{t} - \frac{u_{1z} \rho_0}{H}
+            &= F(x, z, t),\\
+        F(x, z, t) &= \Re Fe^{-\frac{(z - z_0)^2}{2\sigma^2}}
+            e^{i\p*{k_xx - \omega t}}.
+    \end{align}
+\end{subequations}
+For the linearized system, we can immediately substitute $\pd{}{t} \to -i\omega,
+\pd{}{x} \to ik_x$, giving
+\begin{align*}
+    \pd{u_{1z}}{z} + ik_xu_x &= 0,\\
+    -i\omega u_x + \frac{ik_xP}{\rho_0} &= 0,\\
+    -i\omega u_{1z} + \frac{1}{\rho_0}\pd{P}{z} + \frac{\rho_1 g}{\rho_0} &= 0,\\
+    -i\omega \rho_1 - \frac{u_{1z}\rho_0}{H} &= F.
+\end{align*}
+This can be recast solely in terms of $u_{1z}$ as
+\begin{align*}
+    -i\omega u_{1z} + \frac{\omega}{k_x}\p*{-\frac{1}{ik_x}\ptd{u_{1z}}{z}
+        + \frac{1}{ik_xH}\pd{u_{1z}}{z}} - \frac{u_{1z}g}{Hi\omega}
+        - \frac{Fg}{i\omega \rho_0} = 0,\\
+    \omega^2\p*{
+        k_x^2u_{1z} + \frac{1}{H}\pd{u_{1z}}{z} - \ptd{u_{1z}}{z}
+    } - u_{1z}N^2k_x^2 - \frac{Fgk_x^2}{\rho_0} = 0.
+\end{align*}
+Note that if $\pd{}{z} \to 1/2H, F = 0$ then this simplifies to the expected
+dispersion relation $\omega^2 = \frac{N^2k_x^2}{k_x^2 + k_z^2 +
+\frac{1}{4H^2}}$.
+
+If we approximate $F(x, z, t) \approx F_0\delta\p*{z - z_0} e^{i(k_xx - \omega t)}$
+for the time being, then clearly the only term that can be equally singluar is
+the $\ptd{}{z}$ term, leaving us with
+\begin{equation}
+    -\omega^2 \ptd{u_{1z}}{z} - \frac{Fgk_x^2}{\rho_0} = 0.
+\end{equation}
+Since $u_{1z}$ both above and below $z_0$ satisfy the homogeneous PDE, which has
+plane wave solutions with an $e^{z/2H}$ envelope where $k_zH \gg 1$, we can
+integrate about a small domain around $z_0$ and obtain
+\begin{equation}
+    \at{\pd{u_{1z}}{z}}_{z > z_0} - \at{\pd{u_{1z}}{z}}_{z < z_0}
+        = -F_0\frac{gk_x^2}{\rho_0\omega^2}e^{i(k_xx - \omega t)}.
+\end{equation}
+It should then be apparent that the $u_{1z}$ solution is
+\begin{equation}
+    u_{1z} = \frac{F_0gk_x^2}{\rho_0(z = z_0)\omega^2}\frac{1}{2ik_z} \times
+    \begin{cases}
+        -e^{(z - z_0)/2H}e^{i(k_xx + k_z(z - z_0) - \omega t)} & z > z_0\\
+        e^{(z - z_0)/2H}e^{i(k_xx + k_z(z - z_0) - \omega t)} & z < z_0\\
+    \end{cases}.
+\end{equation}
+
+Given that we are actually using a Gaussian $F(x, z, t) \propto F_0e^{-\frac{(z
+- z_0)^2}{2\sigma^2}}$, the integral under $F(z)$ is actually
+$F_0\sqrt{2\pi}\sigma$ rather than just $F_0$. Finally, since a delta function
+is flat in spatial frequency space while a Gaussian tapers off with width
+$1/\sigma$ ($\sigma$ is the spatial width, so $1/\sigma$ is the spatial
+frequency width), we should expect an overall multiplicative factor of
+$e^{-k_z^2\sigma^2 / 2}$ in the $k_z$ spatial component of $F(x, z, t)$ compared
+to the delta function case. Thus, we finally obtain
+\begin{equation}
+    \abs*{u_{1z}(z = z_0)} = \frac{F_0\sqrt{2\pi}\sigma gk_x^2}{
+        2\rho_0\p*{z = z_0}\omega^2 k_z}e^{-k_z^2\sigma^2/2}.
+\end{equation}
 
@@
 
+Note: There exist papers on the Courant stability condition of Chebyshev
+spectral methods which cite exactly the $N^{-2}$ minimum timestep scaling we can
+infer from the above. But if our field is small near the boundaries and is only
+interesting in a region isolated from the boundary (e.g.\ with our damping
+zones) then the bound used in the papers scales more like $N^{-1}$, which is
+much more desirable.
+
 \end{document}

### 2018-07-24 sims/2d_0_lin/strat_helper.py

@@ -29,6 +29,12 @@ def get_vg(g, h, kx, kz):
 def zero_ic(solver, domain, params):
     pass
 
+def get_uz_f_ratio(params):
+    return (np.sqrt(2 * np.pi) * params['S'] * params['g'] *
+            params['KX']**2) * np.exp(-1/2) / (
+                2 * params['RHO0'] * np.exp(-params['Z0'] / params['H'])
+                * params['OMEGA']**2 * params['KZ'])
+
 def get_solver(params):
     ''' sets up solver '''
     x_basis = de.Fourier('x',
@@ -188,13 +194,7 @@ def plot(name, params):
     for var in slice_vars:
         state_vars[var] = state_vars[var.replace(slice_suffix, '')][:, 0, :]
 
-    # we estimate the analytical envelope by picking a z value, finding the
-    # median over time of the max amplitude over x, then extrapolating a
-    # exp(z/2H) profile.
-    z_est_loc = params['Z0'] + 20 * params['S']
-    z_est_idx = int(z_est_loc / params['ZMAX'] * z.size)
-    uz_maxes = np.max(state_vars['uz'][:, :, z_est_idx], 1)
-    uz_est = np.max(uz_maxes)
+    uz_est = params['F'] * get_uz_f_ratio(params)
 
     for t_idx, sim_time in list(enumerate(sim_times))[::plot_stride]:
         fig = plt.figure(dpi=200)
@@ -223,8 +223,14 @@ def plot(name, params):
                           linewidth=0.5)
             if var == 'uz%s' % slice_suffix:
                 p = axes.plot(
-                    uz_est * np.exp((z_pts - z_est_loc) / (2 * params['H'])),
+                    uz_est * np.exp((z_pts - params['Z0']) / (2 * params['H'])),
                     z_pts,
+                    'orange',
+                    linewidth=0.5)
+                p = axes.plot(
+                    -uz_est * np.exp((z_pts - params['Z0']) / (2 * params['H'])),
+                    z_pts,
+                    'orange',
                     linewidth=0.5)
 
             plt.xticks(rotation=30)

### 2018-07-26 sims/2d_0_ns/strat_helper.py
NS solver

@@ -22,9 +22,8 @@ SNAPSHOTS_DIR = 'snapshots_%s'
 def get_omega(g, h, kx, kz):
     return np.sqrt((g / h) * kx**2 / (kx**2 + kz**2 + 0.25 / h**2))
 
-def get_vg(g, h, kx, kz):
-    norm = get_omega(g, h, kx, kz) / (kx**2 + kz**2) * (kz / kx)
-    return norm * kx, norm * kz
+def get_vgz(g, h, kx, kz):
+    return get_omega(g, h, kx, kz) / (kx**2 + kz**2) * kz
 
 def zero_ic(solver, domain, params):
     pass
@@ -269,7 +268,7 @@ def plot(name, params):
                                    idx,
                                    title='%s (kx=kx_d)' % var)
             var_dat = state_vars[var]
-            kx_idx = round(params['KX'] / (2 * np.pi / params['H']))
+            kx_idx = round(params['KX'] / (2 * np.pi / params['XMAX']))
             p = axes.semilogx(var_dat[t_idx][kx_idx],
                               range(len(var_dat[t_idx][kx_idx])),
                               linewidth=0.5)

### 2018-07-30 writeups/notes.tex
NS solver
check in modulational instability writeup

@@
 
+\section{Algebraic Manipulations}
+
+\subsection{Factoring out the exponential dependence}
+
+I've had to do this algebra too many times now, so let's just consider if we
+factor out the $e^{z/2H}$ exponential scaling on the dynamical variables and the
+$e^{-z/H}$ scaling on $\rho_0$. This leaves us with:
+\begin{subequations}\label{se:nonlin_factored}
+    \begin{align}
+        \vec{\nabla} \cdot \vec{u}_1 + \frac{u_{1z}}{2H} &= 0,\\
+        \pd{\rho_1}{t} - \frac{u_{1z} \rho_0}{H} &= -e^{z/2H}\s*{
+            \p*{\vec{u}_1 \cdot \vec{\nabla} \rho_1} + \frac{u_{1z}\rho_1}{2H}
+        },\\
+        \pd{\vec{u}_1}{t} + \frac{\vec{\nabla}P}{\rho_0}
+            + \frac{P}{2H\rho_0}\hat{z} + \frac{\rho_1\vec{g}}{\rho_0}
+            &= -e^{z/2H}\s*{
+            \p*{\vec{u} \cdot \vec{\nabla} \vec{u}_1}
+                + \frac{u_{1z}\vec{u}_1}{2H}
+        }.
+    \end{align}
+\end{subequations}
+I'm too lazy to re-notate my variables, so all of them are constant amplitude
+over $z$ now.
+
+\chapter{Stability Analysis}
+
+\section{Modulational Instability of NLSE}
+
+We study the nonlinear Schr\"odinger Equation (NLSE)
+\begin{equation}
+    i\psi_t + \psi_{xx} + \nu \abs*{\psi}^2\psi = 0.
+\end{equation}
+Sutherland's paper demonstrates an interaction between the wave and the
+wave-induced mean flow ($\propto \abs*{\psi}^2$) that reduces the IGW system to
+the NLSE\@. Here we follow Whitham's textbook in rederiving the modulational
+instability criterion.
+
+Consider for some fixed value $k_0$, then the linear SE (LSE) admits a plane
+wave solution $\psi = e^{i(k_0x - \omega(k_0)t)}$ for some dispersion relation
+$\omega(k)$. If instead we permit $\psi$ to be distributed in frequency space
+near $k_0$, then we may notate
+\begin{equation}
+    \psi = \psi_0 \varphi = e^{i(k_0x - \omega(k_0) t)} \int\limits
+        \exp\s*{i\p*{k' x - k'\pd{\omega}{k}t -
+            \frac{1}{2}k'^2 \ptd{\omega}{k}t}}\;\mathrm{d}k',
+\end{equation}
+where $k' = k - k_0$ denotes the deviation from the central $k_0$
+wavenumber. Defining $\omega' = \omega - \omega_0$, we may observe
+\begin{equation}
+    \omega' = k' \pd{\omega}{k} + \frac{1}{2}k'^2 \ptd{\omega}{k}.
+\end{equation}
+Making correspondence $-i\omega' \to \partial_t, ik' \to \partial_x$, we see
+that $\varphi$ obeys dispersion relation
+\begin{equation}
+    i\p*{\varphi_t + \pd{\omega}{k}\varphi_x}
+        + \frac{1}{2}\ptd{\omega}{k}\varphi_{xx} = 0.
+\end{equation}
+Thus, the modulation $\varphi$ translates at the group velocity, and we find
+that $\varphi$ obeys a modified LSE in a comoving frame.
+
+If we reintroduce the nonlinearity, we observe that $\nu \abs*{\psi}^2\psi = \nu
+\abs*{\varphi}^2 \psi_0\varphi$, and so the nonlinearity modifies the LSE as
+\begin{align}
+    i\p*{\varphi_t + \pd{\omega}{k}\varphi_x}
+        + \frac{1}{2}\ptd{\omega}{k}\varphi_{xx}
+        + \nu \abs*{\varphi^2}\varphi &= 0,\\
+    \omega' = k' \pd{\omega}{k} + \frac{1}{2}k'^2 \ptd{\omega}{k}
+        - \nu a^2.\label{eq:nlse_disp}
+\end{align}
+
+Armed with this dispersion relation, we can apply the method of characteristics
+to the system of equations describing any wavetrain
+\begin{subequations}\label{eq:wavetrain}
+    \begin{align}
+        \pd{k}{t} + \pd{\omega}{x} &= 0,\label{eq:wavetrain_1}\\
+        \pd{A^2}{t} + \pd{}{x}\p*{c_gA^2} &= 0.\label{eq:wavetrain_2}
+    \end{align}
+\end{subequations}
+We denote $A$ the wave amplitude and $c_g$ the group velocity. In a nonlinear
+system, $\omega = \omega(A, k)$, and to leading order in $A$ we expect $\omega =
+\omega_0 + \omega_2A^2$, i.e.\ the dispersion should not depend on the sign of
+$A$. Thus we may rewrite \autoref{eq:wavetrain} as
+\begin{subequations}\label{eq:wavetrain_mod}
+    \begin{align}
+        \pd{k}{t} + \p*{\pd{\omega_{0}}{k} + \pd{\omega_2}{k}A^2}\pd{k}{x}
+            + \omega_2 \pd{A^2}{x} &= 0,\\
+        \pd{A^2}{t} + \pd{A^2}{x}\pd{\omega_0}{k} + A^2\ptd{\omega_0}{k}\pd{k}{x}
+            &= 0.
+    \end{align}
+\end{subequations}
+We first note that, in $A \to 0$, the equations decouple and both equations
+describe propagation at the group velocity (they are already in characteristic
+form $Q_{,t} + cQ_{,x} = 0$). With some foresight, we anticipate that the
+$\pd{A^2}{x}$ term in the first equation will provide a $\mathcal{O}(A)$
+correction to the group velocity, while the $\pd{\omega_2}{k}A^2\pd{k}{x}$ term
+provides an $\mathcal{O}(A^2)$ correction, so we drop it.
+
+The method of characteristics then requires we seek linear combinations of the
+two equations such that all equations are in characteristic form. The Whitham
+prescription says that if the original system is of form $A_{ij}q_{j, t} +
+B_{ij}q_{j, x} = 0$, then the characteristic curves $X(T)$ satisfy $A_{ij}X +
+B_{ij}T = 0$. Here, this yields
+\begin{equation}
+    X(T) = \p*{\pd{\omega_0}{x} \pm A\sqrt{\omega_2 \ptd{\omega_0}{k}}}T.
+\end{equation}
+
+Comparing to the NLSE dispersion relation \autoref{eq:nlse_disp}, we identify
+that $\sqrt{-q \ptd{\omega_0}{k}}A$ is the correction to the group velocity.
+Thus, if $q\ptd{\omega_0}{k} > 0$, we obtain a \emph{complex} group velocity,
+which is identified with \emph{modulational instability}. Naively, this might
+imply a steepening $k = c_{g, 0}t - i\sqrt{q\omega_0''}t$ with timescale
+$\Gamma \sim A\p*{q\omega_0''}^{-1/2}$.
+
 \clearpage

### 2018-08-01 writeups/notes.tex
update ns as well

@@
 \begin{equation}
-    X(T) = \p*{\pd{\omega_0}{x} \pm A\sqrt{\omega_2 \ptd{\omega_0}{k}}}T.
+    X(T) = \p*{\pd{\omega_0}{k} \pm A\sqrt{\omega_2 \ptd{\omega_0}{k}}}T.
 \end{equation}
@@
 Comparing to the NLSE dispersion relation \autoref{eq:nlse_disp}, we identify
-that $\sqrt{-q \ptd{\omega_0}{k}}A$ is the correction to the group velocity.
-Thus, if $q\ptd{\omega_0}{k} > 0$, we obtain a \emph{complex} group velocity,
+that $\sqrt{-\nu \ptd{\omega_0}{k}}A$ is the correction to the group velocity.
+Thus, if $\nu\ptd{\omega_0}{k} > 0$, we obtain a \emph{complex} group velocity,
 which is identified with \emph{modulational instability}. Naively, this might
-imply a steepening $k = c_{g, 0}t - i\sqrt{q\omega_0''}t$ with timescale
-$\Gamma \sim A\p*{q\omega_0''}^{-1/2}$.
+imply a steepening $k = c_{g, 0}t - i\sqrt{\nu\omega_0''}At$ with timescale
+$\Gamma \sim \p*{A^2\nu\omega_0''}^{-1/2}$.
 

### 2018-08-01 sims/2d_0_ns/strat.py

@@ -11,7 +11,7 @@ ZMAX = 10 * H
 
 NUM_TIMESTEPS = 2e3
 NUM_SNAPSHOTS = 200
-TARGET_UZ = 0.02 # target uz at forcing zone
+TARGET_UZ = 0.01 # target uz at forcing zone
 
 PARAMS_RAW = {'XMAX': XMAX,
               'ZMAX': ZMAX,
@@ -22,7 +22,7 @@ PARAMS_RAW = {'XMAX': XMAX,
               'H': H,
               'RHO0': 1,
               'Z0': 0.15 * ZMAX,
-              'SPONGE_STRENGTH': 2,
+              'SPONGE_STRENGTH': 1,
               'SPONGE_HIGH': 0.9 * ZMAX,
               'SPONGE_LOW': 0.1 * ZMAX,
               'NUM_SNAPSHOTS': NUM_SNAPSHOTS}
@@ -31,7 +31,8 @@ def build_interp_params(interp_x, interp_z, overrides=None):
     params = {**PARAMS_RAW, **(overrides or {})}
     KX = params['KX']
     KZ = params['KZ']
-    g = (KX**2 + KZ**2 + 1 / (4 * H**2)) / KX**2 * (2 * np.pi)**2 * H # omega = 2pi
+    # g = (KX**2 + KZ**2 + 1 / (4 * H**2)) / KX**2 * (2 * np.pi)**2 * H # omega = 2pi
+    g = H # N^2 = 1
 
     OMEGA = get_omega(g, H, KX, KZ)
     VG_Z = get_vgz(g, H, KX, KZ)
@@ -45,10 +46,12 @@ def build_interp_params(interp_x, interp_z, overrides=None):
     params['INTERP_Z'] = interp_z
     params['N_X'] //= interp_x
     params['N_Z'] //= interp_z
-    params['DT'] = min(params['T_F'] / NUM_TIMESTEPS, 0.001) # omega * DT << 1
+    # omega * DT << 1 is required
+    params['DT'] = min(params['T_F'] / NUM_TIMESTEPS, 0.01 / OMEGA)
     if not params.get('F'): # default value
         params['F'] = TARGET_UZ / get_uz_f_ratio(params)
-    params['NU'] = params['ZMAX'] / params['N_Z'] # L_min * V / nu ~ 1
+    # NU / (kmax/2)^2 ~ omega
+    params['NU'] = OMEGA * (params['ZMAX'] / (np.pi * params['N_Z']))**2
     print(params['NU'])
     return params
 

### 2018-08-02 writeups/notes.tex
writeup typo fix, shorter timestepper for all

@@
 \end{align*}
-Note that if $\pd{}{z} \to 1/2H, F = 0$ then this simplifies to the expected
+Note that if $\pd{}{z} \to 1/2H + ik_z, F = 0$ then this simplifies to the expected
 dispersion relation $\omega^2 = \frac{N^2k_x^2}{k_x^2 + k_z^2 +
@@
 If we approximate $F(x, z, t) \approx F_0\delta\p*{z - z_0} e^{i(k_xx - \omega t)}$
-for the time being, then clearly the only term that can be equally singluar is
+for the time being, then clearly the only term that can be equally singular is
 the $\ptd{}{z}$ term, leaving us with

### 2018-08-02 sims/2d_0_nonlin/strat.py

@@ -15,7 +15,7 @@ TARGET_UZ = 0.01 # target uz at forcing zone
 
 PARAMS_RAW = {'XMAX': XMAX,
               'ZMAX': ZMAX,
-              'N_X': 128,
+              'N_X': 256,
               'N_Z': 1024,
               'KX': 2 * np.pi / XMAX,
               'KZ': -20 / H,
@@ -46,8 +46,8 @@ def build_interp_params(interp_x, interp_z, overrides=None):
     params['INTERP_Z'] = interp_z
     params['N_X'] //= interp_x
     params['N_Z'] //= interp_z
-    # omega * DT << 1 is required
-    params['DT'] = min(params['T_F'] / NUM_TIMESTEPS, 0.1 / OMEGA)
+    # omega * DT << 1 is required, as is DT << 1/N = 1
+    params['DT'] = min(params['T_F'] / NUM_TIMESTEPS, 0.1 / OMEGA, 0.02)
     if not params.get('F'): # default value
         params['F'] = TARGET_UZ / get_uz_f_ratio(params)
     return params

### 2018-08-02 sims/2d_0_ns/strat.py

@@ -15,7 +15,7 @@ TARGET_UZ = 0.01 # target uz at forcing zone
 
 PARAMS_RAW = {'XMAX': XMAX,
               'ZMAX': ZMAX,
-              'N_X': 128,
+              'N_X': 256,
               'N_Z': 1024,
               'KX': 2 * np.pi / XMAX,
               'KZ': -20 / H,
@@ -47,7 +47,7 @@ def build_interp_params(interp_x, interp_z, overrides=None):
     params['N_X'] //= interp_x
     params['N_Z'] //= interp_z
     # omega * DT << 1 is required, as is DT << 1/N = 1
-    params['DT'] = min(params['T_F'] / NUM_TIMESTEPS, 0.1 / OMEGA, 0.05)
+    params['DT'] = min(params['T_F'] / NUM_TIMESTEPS, 0.1 / OMEGA, 0.02)
     if not params.get('F'): # default value
         params['F'] = TARGET_UZ / get_uz_f_ratio(params)
     # NU / (kmax/2)^2 ~ omega
@@ -64,8 +64,8 @@ def run(ic, name, params_dict):
 
 if __name__ == '__main__':
     tasks = [
-        (zero_ic, 'linear_ns',
-         build_interp_params(1, 1, overrides={'F': 0.00001})),
+        # (zero_ic, 'linear_ns',
+        #  build_interp_params(1, 1, overrides={'F': 0.00001})),
         (zero_ic, 'nonlinear_ns',
          build_interp_params(1, 1, overrides={'KX': 4 * np.pi / H})),
     ]

### 2018-08-02 sims/2d_0_ns/strat_helper.py

@@ -131,6 +131,7 @@ def load(name, params):
     dyn_vars = ['uz', 'ux', 'rho', 'P']
     snapshots_dir = SNAPSHOTS_DIR % name
     filename = '{s}/{s}_s1.h5'.format(s=snapshots_dir)
+    start_idx = 0
 
     post.merge_analysis(snapshots_dir, cleanup=False)
 
@@ -138,13 +139,13 @@ def load(name, params):
     z = domain.grid(1, scales=params['INTERP_Z'])
 
     with h5py.File(filename, mode='r') as dat:
-        sim_times = np.array(dat['scales']['sim_time'])
+        sim_times = np.array(dat['scales']['sim_time'])[start_idx: ]
     # we let the file close before trying to reopen it again in load
 
     # load into state_vars
     state_vars = defaultdict(list)
     for idx in range(len(sim_times)):
-        solver.load_state(filename, idx)
+        solver.load_state(filename, idx + start_idx)
 
         for varname in dyn_vars:
             values = solver.state[varname]

### 2018-08-08 writeups/q_writeup/q_writeup.tex
q writeup

    \documentclass[11pt,
        usenames, % allows access to some tikz colors
        dvipsnames % more colors: https://en.wikibooks.org/wiki/LaTeX/Colors
    ]{article}
    \usepackage{
        amsmath,
        amssymb,
        fouriernc, % fourier font w/ new century book
        fancyhdr, % page styling
        lastpage, % footer fanciness
        hyperref, % various links
        setspace, % line spacing
        amsthm, % newtheorem and proof environment
        mathtools, % \Aboxed for boxing inside aligns, among others
        float, % Allow [H] figure env alignment
        enumerate, % Allow custom enumerate numbering
        graphicx, % allow includegraphics with more filetypes
        wasysym, % \smiley!
        upgreek, % \upmu for \mum macro
        listings, % writing TrueType fonts and including code prettily
        tikz, % drawing things
        booktabs, % \bottomrule instead of hline apparently
        cancel % can cancel things out!
    }
    \usepackage[margin=0.8in]{geometry} % page geometry
    \usepackage[
        labelfont=bf, % caption names are labeled in bold
        font=scriptsize % smaller font for captions
    ]{caption}
    \usepackage[font=scriptsize]{subcaption} % subfigures
    \usepackage[numbers]{natbib}

    \newcommand*{\scinot}[2]{#1\times10^{#2}}
    \newcommand*{\dotp}[2]{\left<#1\,\middle|\,#2\right>}
    \newcommand*{\rd}[2]{\frac{\mathrm{d}#1}{\mathrm{d}#2}}
    \newcommand*{\pd}[2]{\frac{\partial#1}{\partial#2}}
    \newcommand*{\rtd}[2]{\frac{\mathrm{d}^2#1}{\mathrm{d}#2^2}}
    \newcommand*{\ptd}[2]{\frac{\partial^2 #1}{\partial#2^2}}
    \newcommand*{\md}[2]{\frac{\mathrm{D}#1}{\mathrm{D}#2}}
    \newcommand*{\pvec}[1]{\vec{#1}^{\,\prime}}
    \newcommand*{\svec}[1]{\vec{#1}\;\!}
    \newcommand*{\bm}[1]{\boldsymbol{\mathbf{#1}}}
    \newcommand*{\ang}[0]{\;\text{\AA}}
    \newcommand*{\mum}[0]{\;\upmu \mathrm{m}}
    \newcommand*{\at}[1]{\left.#1\right|}

    \newtheorem{theorem}{Theorem}[section]

    \let\Re\undefined
    \let\Im\undefined
    \DeclareMathOperator{\Res}{Res}
    \DeclareMathOperator{\Re}{Re}
    \DeclareMathOperator{\Im}{Im}
    \DeclareMathOperator{\Log}{Log}
    \DeclareMathOperator{\Arg}{Arg}
    \DeclareMathOperator{\Tr}{Tr}
    \DeclareMathOperator{\E}{E}
    \DeclareMathOperator{\Var}{Var}
    \DeclareMathOperator*{\argmin}{argmin}
    \DeclareMathOperator*{\argmax}{argmax}
    \DeclareMathOperator{\sgn}{sgn}
    \DeclareMathOperator{\diag}{diag\;}

    \DeclarePairedDelimiter\bra{\langle}{\rvert}
    \DeclarePairedDelimiter\ket{\lvert}{\rangle}
    \DeclarePairedDelimiter\abs{\lvert}{\rvert}
    \DeclarePairedDelimiter\ev{\langle}{\rangle}
    \DeclarePairedDelimiter\p{\lparen}{\rparen}
    \DeclarePairedDelimiter\s{\lbrack}{\rbrack}
    \DeclarePairedDelimiter\z{\lbrace}{\rbrace}

    % \everymath{\displaystyle} % biggify limits of inline sums and integrals
    \tikzstyle{circ} % usage: \node[circ, placement] (label) {text};
        = [draw, circle, fill=white, node distance=3cm, minimum height=2em]
    \definecolor{commentgreen}{rgb}{0,0.6,0}
    \lstset{
        basicstyle=\ttfamily\footnotesize,
        frame=single,
        numbers=left,
        showstringspaces=false,
        keywordstyle=\color{blue},
        stringstyle=\color{purple},
        commentstyle=\color{commentgreen},
        morecomment=[l][\color{magenta}]{\#}
    }

\begin{document}

\def\Snospace~{\S{}} % hack to remove the space left after autorefs
\renewcommand*{\sectionautorefname}{\Snospace}
\renewcommand*{\appendixautorefname}{\Snospace}
\renewcommand*{\figureautorefname}{Fig.}
\renewcommand*{\equationautorefname}{Eq.}
\renewcommand*{\tableautorefname}{Tab.}

\singlespacing

\pagestyle{fancy}
% \rfoot{Yubo Su}
\rhead{}
\cfoot{\thepage/\pageref{LastPage}}

\title{Q Exam Written Report}
\author{Yubo Su\\
Adviser: Dong Lai}

\maketitle

\section{Introduction: White Dwarf Binaries}

White Dwarfs (WDs) are commonly found in binaries in which two objects orbit
their center of mass under mutual gravitational attraction. The companion object
ranges from another WD to a supermassive BH (SMBH). Understanding the evolution
of such systems is important to astrophysics. WD-WD binaries are most important
for being thought to generate Type Ia supernovae which have been used as a
``standard candle'' to probe the expansion rate of the
universe\cite{darkEnergy}. WD-BH systems are also interesting subjects of study.
Recent works indicate that as WDs plunge close to BHs, they will produce
observable flares induced by gravitational tidal forces\cite{flares}. A WD
orbiting a SMBH would also produce gravitational waves that are expected to be
detected by the space-based \emph{Laser Interferometer Space Antenna (LISA)}
when deployed\cite{lisa}. Gravitational wave astronomy is an increasingly
exciting field as the \emph{Laser Interferometer Gravitational-Wave Observatory
(LIGO)} continues to make progress since its first detection in late 2015. As
gravitational wave astronomy relies on accurate predictions of the expected
signals, it is important to build as accurate models as possible before
observation runs begin.

\subsection{Tidal Dissipation}

The excitation of \emph{internal gravity waves} (IGW) in the WD by the tidal
forces of the companion is an effect exhibited in the aforementioned systems.
IGWs, not to be confused with the gravitational waves above, are internal
displacements in the WD fluid that oscillate and propagate due to a restoring
buoyancy force. Previous work\citep{fullerI} indicates that IGWs can be excited
in WD binaries that propagate outwards and grow exponentially due to density
rarefaction, reaching nonlinear amplitudes well before the WD boundary. These
waves are then expected to break and dissipate via nonlinear processes.

Previous work predicts that this dissipation mechanism can generate
significantly more energy than thermal radiation from the WD surface alone and
are thus a significant contribution to the WD energy budget\cite{fullerII}. The
exact radial dissipation profile is of interest since it both is sensitive to WD
properties and can produce vastly different observable outcomes. One proposed
outcome is a \emph{tidal nova}, in which heating in the WD's degenerate hydrogen
layer is sufficient to trigger runaway nuclear fusion and an observable surface
explosion\cite{tidal_novae}. Understanding whether such phenomena occur requires
understanding how energy and angular momentum are redistributed internally
inside the WD\@.

Internal gravity wave breaking is a nonlinear hydrodynamic phenomenon. Such
phenomena are known to require numerical simulation to study in detail.

\section{Work Accomplished}

To begin our numerical study of IGW breaking, we first considered 2D IGW
breaking in a plane-parallel, uniformly stratified, incompressible atmosphere.
Even in this reduced problem, we encountered numerous difficulties and efforts
to surmount them are ongoing. Thus the subsequent two sections on the theory and
numerics of IGW breaking largely focus on the toy problem with only brief
references to extensions to the full 3D, compressible problem we wish to
eventually study.

\subsection{IGW Theory}

\subsubsection{IGW System and Linear Theory}

We briefly recap the studied IGW system in detail and its solution in linear
theory. An incompressible 2D hydrodynamical system can be described by just four
Eulerian dynamical variables $\rho, P, u_x, u_z$, the density, pressure and $x,
z$ components of Eulerian velocity respectively, each of which is a function of
$x, z, t$. We notate $\rho = \rho_0 + \rho_1$ where $\rho_0$ is the background
density satisfying hydrostatic equilibrium in the absence of any IGW and
$\rho_1$ is the deviation that represents the IGW flow, and likewise for the
remaining variables. In the linear theory, we assume $\rho_1 \ll \rho_0, P_1 \ll
P_0$, and to further simplify the problem we take $u_{0x} = u_{0z} = 0$. The
resulting system of equations describing IGW propagation are:
\begin{subequations}\label{se:igw_lin}
    \begin{align}
        \vec{\nabla} \cdot \vec{u}_1 &= 0,\\
        \rd{\rho_1}{t} + u_{1z}\pd{\rho_0}{z} &= 0,\\
        \rd{\vec{u}_1}{t} + \frac{\vec{\nabla}P_1}{\rho_0}
            - \rho_1 \frac{\vec{\nabla}P_0}{\rho_0^2} &= 0.
    \end{align}
\end{subequations}
In the linearized system, $\rd{}{t} = \pd{}{t} + \vec{u} \cdot \vec{\nabla}
\approx \pd{}{t}$. Uniform stratification takes the form $-\pd{\ln \rho_0}{z} =
-\pd{\ln P_0}{z} = H$ where $H$ is some characteristic scale height, then it can
be shown the IGW solution takes on form
\begin{align}
    u_{1z}(x, z, t) &= Ae^{z/2H}e^{i\p*{k_xx + k_zz - \omega(k_x, k_z) t}},
        \label{eq:plane}\\
    \omega(k_x, k_z) &= \frac{Nk_x}{\sqrt{k_x^2 + k_z^2 + \frac{1}{4H^2}}},
\end{align}
where $N^2 = \frac{g}{H}$ the Brunt-V\"ais\"al\"a frequency is constant. Note
the $u_{1z} \propto e^{z/2H}$ growth induced by the density stratification, and
so we expect nonlinear effects to become important as $Ae^{z/2H} \sim 1$.

\subsubsection{Nonlinear IGW Theory}

The leading-order nonlinearities in \autoref{se:igw_lin} causing the plane wave
solution \autoref{eq:plane} to break down are thoroughly studied in atmospheric
sciences e.g.\cite{Sutherland1}\cite{Sutherland2}, though a verdict on the exact
breaking phenomenon is still unclear. The primary two mechanisms considered by
the community are:
\begin{description}
    \item[Modulational Instability] By arguments similar to those underlying
        Stokes Flow (the mean motion of particles in a plane wave flow), it can
        be shown that the plane wave solution \autoref{eq:plane} induces a mean
        flow $\mathcal{O}(A^2)$. Since, in the incompressible system, $\vec{k}
        \cdot \vec{u} = 0$ (where $\vec{k} = k_x\hat{x} + k_z\hat{z}$ is the
        wave vector), many of the advective terms $\vec{u} \cdot \vec{\nabla}$
        vanish to second order $\vec{u}_1 \cdot \vec{\nabla}\vec{u}_1 = 0$, and
        so the first nonlinear term to appear must be the interaction of the
        plane wave with its own induced mean flow.

        With some algebraic manipulations, \autoref{se:igw_lin} can be recast in
        the form of a nonlinear Schr\"odinger Equation, whose form is $iA_t =
        \alpha A_{xx} + \beta \abs*{A}^2A$. $A$ happens here to be the complex
        amplitude of the Eulerian displacement, and $\abs*{A}^2$ is the mean
        flow term.

        The weakly nonlinear Schr\"odinger equation can be examined for
        \emph{modulational instability} by considering the evolution of modes
        near the propagating plane wave, i.e.\ a wave train centered on
        $\vec{k}_0$. A brief derivation of the criterion is provided
        in~\cite{Whitham} and applied in~\cite{Sutherland1}. The calculations
        show that for certain values of $\vec{k}$, the wave propagates with a
        weakly split group velocity that acts to reduce the amplitude and delay
        breaking, while for other values the group velocity is complex and
        exhibits breaking at lower amplitudes.

    \item[Parametric Subharmonic Instability] If a triad of modes with
        wavevectors $\vec{k}_1, \vec{k}_2, \vec{k}_3$ satisfies the triad
        resonance condition
        \begin{align}
            \vec{k}_1 &= \vec{k}_2 + \vec{k}_3 &
            \omega(\vec{k}_1) &= \omega(\vec{k}_2) + \omega(\vec{k}_3),
            \label{eq:triad}
        \end{align}
        then it is possible in the nonlinear advective term $\vec{u} \cdot
        \vec{\nabla}$ for two modes to have a frequency component that drives
        the third. A careful albeit tedious derivation for the growth rate of
        two infinitesimal modes in the presence of a third mode such that the
        three satisfy \autoref{eq:triad} is given in~\cite{psi}. Further Floquet
        analysis, as done in~\cite{DrazinFloquet}, illustrates that this
        produces cascades to integer multiples of the plane wave wavevector and
        fractions of the temporal frequency.

        It bears noting that, as this is a parametric resonance, the growth of
        the subharmonics grows exponentially from a nonzero initial value. An
        example of the development of these higher modes is given in
        \autoref{fig:igw_psi}.
        \begin{figure}[t]
            \centering
            \includegraphics[width=0.8\textwidth]{figs/igw_psi.png}
            \caption{Breakdown of plane wave via parametric subharmonic
            instability. Doubly periodic boundary conditions, courtesy
            of~\cite{SutherlandBook}.}\label{fig:igw_psi}
        \end{figure}
\end{description}

While a thorough understanding of the nonlinear theory is not strictly
necessary to run simulations, it is important to be able to identify whether
deviations from the linear solution in simulations are numerical artifacts or
real physical effects.

\subsection{Numerical Simulations}

\subsubsection{Choice of Software}

For numerical simulations we use the spectral PDE solver
\lstinline{Dedalus}\cite{dedalus}. A number of concerns motivated our choice of
software:
\begin{itemize}
    \item Though we expect the problem to develop fine scale structure at later
        times, no shocks develop. Spectral methods exhibit exponential
        convergence when dynamical variables vary smoothly\cite{NR}, and so we
        are able to capture better numerical precision for the same
        computational cost.

    \item As the key focus of our study is how energy cascades into smaller
        length scales but we have limited spatial resolution, we must introduce
        some nonphysical regularization at small length scales to prevent them
        from accumulating energy. This usually comes in the form of viscosity,
        so having precise control over the viscosity is necessary to understand
        the regime of validity of our results. Spectral methods have no
        numerical viscosity\cite{NR} and so allow for much more precise control
        over regularization.

    \item \lstinline{Dedalus} is an exceptionally easy to use code and remains
        under active development. It also comes parallelized and has been
        extended to 3D geometries as well (although still in development).
\end{itemize}
Note however that \lstinline{Dedalus}, being a fully spectral solver, does not
support numerical techniques such as adaptive mesh refinement and so must use a
uniform resolution throughout the simulation domain. It is thus prudent to
choose a simulation subject that maximizes usage of the domain with interesting
behavior.

\subsubsection{Setup}

Note that the objective only requires simulating IGWs near the end of their
propagation, after an infinitesimal displacement has already grown to
near-nonlinear amplitudes. Focusing on the end of their evolution is a
tremendous computational savings both in temporal evolution and spatial
resolution as discussed above.

We discuss the simulation setup and its various considerations below:
\begin{description}
    \item[Domain] We choose a 2D domain with periodic boundary conditions in the
        $x$ direction and aperiodic boundary conditions in the $z$ direction to
        permit density stratification. In conjunction with the damping zones
        below, this mimics a 2D infinite plane geometry. The choice of spectral
        basis functions is then Fourier in the $x$ and Chebyshev in the $z$.

        We want a $z$ domain of a many scale heights to induce growth to
        nonlinear amplitudes, so we use $z \in [0, 10H]$. To use a similarly
        sized domain, we use $x \in [0, 4H]$.

        We are able to use $1024$ Chebyshev modes and $256$ Fourier modes
        (memory and computation resource limited), meaning we can resolve length
        scales $\sim H/100$, plenty for our target $k_z = \frac{20}{H}$.

    \item[Damping Zones] In order to mimic an unbounded domain in the $z$, we
        must ensure that the boundaries do not produce reflections. Such
        radiative boundary conditions (also outgoing/Sommerfield/absorbing
        boundary conditions) can only be approximate when applied to many
        $\vec{k}$ incoming waves\cite{RadBC}\cite{Wagatha1983}.

        Instead, we implement \emph{damping zones} modeled after those used
        in~\cite{LecoanetDamp}. We introduce a field $\Gamma(z)$ such that
        $\pd{}{t} \to \pd{}{t} + \Gamma(z)$, so that dynamical variables
        dissipate where $\Gamma$ is nonzero. Choosing then $\Gamma$ to vanish
        everywhere but near the boundaries suppresses reflection. In order to
        ensure $\Gamma(z)$ has an exponentially convergent spectral
        representation, we must choose smooth $\Gamma(z)$, and a traditional
        choice is (let $z \in [0, L_z]$ be the simulation domain)
        \begin{equation}
            \Gamma(z) = \frac{\Gamma_0}{2}\s*{
                2 + \tanh \gamma\frac{z - z_t}{L_z - z_t} +
                    \tanh \gamma\frac{z_b - z}{z_b}
            }.
        \end{equation}
        This function is approximately zero $z \in [z_b, z_t]$ and approximately
        $1$ in the damping zones. Furthermore, its steepness can be tuned via
        the parameter $\gamma$ (it cannot be so steep as to run up against
        spatial resolution, however), and the damping timescale can be tuned by
        $\Gamma_0$.

        In our simulation, $z_b = 0.07L_z, z_t = 0.93L_z, \gamma = 3$ and we
        choose $\Gamma_0 = N$ such that, in the damping zone oscillations damp
        on the timescale of $N^{-1}$.

    \item[Forcing] Since we only wish to model the later stages of IGW
        evolution, we need only excite an IGW in the linear regime with any
        mechanism rather than specifically the tidal mechanism describe in the
        introduction.

        Because Chebyshev polynomials (which satisfy $T_n(x) = \cos
        \p*{n \arccos x}$) have spacing between zeros scaling as $n^{-1}$ on the
        interior of the domain but cluster as $n^{-2}$ near the edge of the
        domain, the effective Courant-Friedrichs-Levy condition is much stricter
        when spectral features exist near the edge of the
        domain\cite{GottliebCFL}.

        Instead, we again follow~\cite{LecoanetDamp} in using a volumetric
        forcing term to generate a plane wave IGW\@. We choose a forcing profile
        of form
        \begin{equation}
            F(x, z, t) = F_0
                e^{-\frac{\p*{z - z_0}^2}{2\sigma^2}}e^{i(k_xx - \omega_0 t)}.
        \end{equation}
        If $\sigma$ is chosen to be sufficiently small, then $F(x, z, t)$ will
        have a reasonably large component along the $k_z$ wavevector satisfying
        $\omega(k_x, k_z) = \omega_0$ where $\omega(k_x, k_z)$ is the IGW
        dispersion relation. Thus we are able to drive on resonance and generate
        IGW propagating in both directions from $z_0$. Lastly, if we choose
        $z_0$ near the bottom of the domain, the upward propagating wave is
        exactly what we wish to study and the downward propagating wave is
        damped by the damping zone.

        Since our equations of motion require the velocity to be
        divergence-free, we instead attach the driving term to the \emph{density
        equation}. While this is ill-justifiable physically, corresponding to
        violating the conservation of mass periodically, it suffices for the
        application at hand. Thus, we use continuity equation
        \begin{equation}
            \rd{\rho_1}{t} - \rho_0 \frac{u_{1z}}{H} + \Gamma(z) \rho_1
                = F_0 e^{-\frac{\p*{z - z_0}^2}{2\sigma^2}}
                    \cos(k_xx - \omega_0 t).
        \end{equation}

        It is not so hard to compute the excited $u_{1z}(z = z_0)$ for such a
        forcing profile: solving for a forcing $F(x, z, t) = F_0\delta\p*{z -
        z_0}e^{i(k_xx - \omega_0t)}$ is a simple matter of matching homogeneous
        solutions, then an approximation that the Gaussian excites
        a $e^{-k_z^2\sigma^2/2}$ smaller wave is a serviceable approximation.
        This gives
        \begin{equation}
            \abs*{u_{1z}(z = z_0)} = \frac{F_0\sqrt{2\pi}\sigma gk_x^2}{
                2\rho_0\p*{z = z_0}\omega^2 k_z}e^{-k_z^2\sigma^2/2}.
        \end{equation}

        In order for the above approximation to hold, $\sigma$ must be small, so
        that the forcing is broad in spatial frequencies. We choose $\sigma k_z
        = 1$. We also use $z_0 = 0.25L_z$, so $0.1L_z$ separated from the
        damping zone, and we choose $F_0$ depending on the $u_{1z}(z = z_0)$
        desired (often $0.01$).

    \item[Timestepping] The primary timestep that must be resolved for IGW is
        $N^{-1}$ the buoyancy timescale (in IGW, $k_x \ll k_z$ and so $\omega
        \ll N$), so we have been testing with timesteps $\Delta t = 0.02N^{-1}$.
        Other simulations use a variety of timesteps $\Delta t \in
        [0.0025N^{-1}, 0.08N^{-1}]$
        e.g.~\cite{LecoanetDamp},~\cite{Sutherland1}. Because Dedalus uses
        implicit timestepping for linear terms, an overly large timestep induces
        amplitude loss during propagation; the proper diagnostic for timestep
        size is ensuring $e^{z/2H}$ growth in the linear regime.

    \item[Viscosity] This is the current stage of our work and is not yet
        well-characterized. Viscosity serves to regularize the system and
        drain energy preferentially from high $k_z$ modes. In order to maintain
        exponential convergence, we need to ensure that any mode that excites
        modes beyond the scope of our spectral expansion must be viscously
        damped. Thus, our regularization should damp the upper half of spectral
        modes.

        There are two choices for implementing such viscous regularization that
        we are currently investigating:

        \begin{description}
            \item[Navier-Stokes Viscosity] This is the traditional Navier-Stokes
                viscosity, which modifies the momentum equation of
                \autoref{se:igw_lin} as
                \begin{equation}
                    \rd{\vec{u}_1}{t} + \frac{\vec{\nabla}P_1}{\rho_0}
                        + \frac{\rho_1\vec{g}}{\rho_0} + \Gamma \vec{u}_1
                        - \nu \nabla^2 \vec{u}_1 = 0.
                \end{equation}

                Using our (rather heuristic) argument that half of modes should
                be damped in the nonlinear regime $u_1 \sim \frac{\omega}{k_z}$
                then we need $\nu \sim \omega \p*{\frac{L_z}{\pi N_Z}}^2$ where
                $N_Z$ is the number of $z$ modes. Note that $k_x\ll k_z$ does
                not factor into the $\nu \nabla^2$ size calculation.

            \item[Hyperviscosity] Much more of a numerical trick,
                \emph{hyperviscosity} refers to a family of
                stronger-than-Navier-Stokes viscosities. The one we consider is
                reaching directly into the $k_z$ spectral coefficients and
                artificially decreasing those beyond $N_Z/2$. Conventionally,
                hyperviscosity sets these directly to zero but this seems to be
                too strong by our tests. Because results are still pending, the
                exact hyperviscosity procedure is yet to be determined.
        \end{description}
\end{description}
$N^{-1}$ is the unit of time and $H$ the unit of distance in our computations,
both of which are set to $1$.

\clearpage
\bibliographystyle{plainnat}
\renewcommand{\bibname}{References}
{\scriptsize \bibliography{q_writeup}}

\end{document}

### 2018-08-09 writeups/q_writeup/q_writeup.tex
final q writeup, my piece of trash thing

@@
 
-\section{Work Accomplished}
+\section{Work Accomplished: Theory and Numerics}
 
@@
         breaking, while for other values the group velocity is complex and
-        exhibits breaking at lower amplitudes.
+        exhibits breaking at lower amplitudes. An exmaple of a modulationally
+        unstable Gaussian wavepacket is shown in \autoref{fig:modinstab}.
+        \begin{figure}[t]
+            \centering
+            \includegraphics[width=\textwidth]{figs/sutherland_modinstab.png}
+            \caption{Modulationally unstable wavepacket breaking before the
+            height predicted by linear theory (thick black
+            line).}\label{fig:modinstab}
+        \end{figure}
 
@@
         \end{description}
+
+    \item[Gradual Forcing] One last trick that we are considering is to turn on
+        the forcing term gradually; \autoref{fig:plots} shows many examples
+        where ringing from turning the forcing term on immediately excites
+        ringing that generates spurious nonlinear interactions. We choose to
+        effect this by $F(t) \tanh 10\omega_0t$, so over the course of about
+        twenty periods the forcing will reach full amplitude. Results are
+        pending.
 \end{description}
+
 $N^{-1}$ is the unit of time and $H$ the unit of distance in our computations,
-both of which are set to $1$.
+both of which are set to $1$. Some snapshots from our simulation can be seen in
+\autoref{fig:plots}. Note the smooth
+\begin{figure}[t]
+    \centering
+    \begin{subfigure}{0.47\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{figs/linear.png}
+        \caption{Linear simulation. Note the close agreement of the analytically
+        predicted excitement amplitude and the actual waveform. The spectrum is
+        broader than a pure Fourier spectrum due to the waveforms being
+        sinusoidal and the basis being Chebyshev.}
+    \end{subfigure}\hfill
+    \begin{subfigure}{0.47\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{figs/nonlin_psi.png}
+        \caption{Nonlinear simulation without regularization. The mismatch of
+        the analytical envelope is not significant. Note the significant
+        amplitude in higher Chebyshev modes. The simulation breaks soon after
+        this point, seemingly due to parametric subharmonic instability.}
+    \end{subfigure}
+
+    \begin{subfigure}{0.47\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{figs/nonlin_ns.png}
+        \caption{Nonlinear simulation with Navier-Stokes regularization. The
+        mismatch of the analytical envelope is again not significant. The
+        supposed PSI from the non-regularized system is clearly just ringing
+        from starting the forcing in discontinuous fashion.}
+    \end{subfigure}\hfill
+    \begin{subfigure}{0.47\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{figs/nonlin_driving_blowup.png}
+        \caption{Another consequence of starting the forcing in discontinuous
+        fashion, nonlinear terms amplify the waveform in the driving zone and
+        produces nonphysical shocks.}\label{fig:blowup}
+    \end{subfigure}
+    \caption{The four simulation snapshots above include the following notation:
+        the left plot is a simple 2D plot of the $u_{1z}(x, z)$ field, the
+        middle plot is a slice $u_{1z}(x=0, z)$ and the right plot plots the
+        Chebyshev spectral expansion at the driving $k_x$. In the middle plots,
+        the two dashed red lines indicate the damping zones, the dashed blue
+        lines indicate the forcing zone $\pm 3\sigma$ and the orange solid lines
+        indicate the analytically predicted envelope.}\label{fig:plots}
+\end{figure}
+
+\section{Future Work}
+
+Mercifully omitted from this writeup are all of the misguided efforts that were
+slowly revised out over time. Nevertheless, against all odds we find ourselves
+very near conquering the 2D toy problem, with a few remaining difficulties to be
+elaborated below.
+
+\subsection{Remaining 2D Work}
+
+We are very close to having a full working simulation of 2D wave breaking
+without numerical singularities. The largest further obstacle that stands in our
+way is computational expense, which is being rectified presently within the
+group. Compounding our difficulties are that, in the $k_x \ll k_z$ limit that
+most astrophysical IGWs find themselves in, the group velocity is extremely
+slow, and so requiring resolving $N^{-1}$ timescales results in a very slow
+simulation. Even so, we should soon understand where the dissipation in these
+excited IGWs occurs as they break via nonlinear processes.
+
+\subsection{3D Work}
+
+\lstinline{Dedalus} has been adapted for 3D work, using spherical harmonics and
+Chebyshev polynomials, so the software package is sufficiently up-to-date for
+our problem. Nevertheless, 3D calculations will be even more expensive and will
+likely carry with them another host of instabilities to sort out. Armed with our
+understanding of these 2D difficulties, however, it should be much easier to
+tackle.
+
+We also intend to extend our 3D calculations to realstic WD equations of state
+and density profiles\cite{fullerI}, such as those in \autoref{fig:eos}.
+These will all have to be adapted somewhat to fit the requirements of
+\lstinline{Dedalus}.
+\begin{figure}[t]
+    \centering
+    \includegraphics[width=\textwidth]{figs/fuller_eos.png}
+    \caption{Dynamical variables for a realistic WD model. The thick line is the
+    density and the thin line is $N^2$. The dashed line is the square of the
+    Lamb frequency, not discussed in this paper but important to 3D
+    waves.}\label{fig:eos}
+\end{figure}.
+
+\section{Conclusion and Acknowledgements}
+
+In conclusion, we are very close to a numerical solution to the 2D plane
+parallel atmosphere IGW breaking problem, with only a few more kinks to sort
+out. With this, we should be able to interpret our results within the framework
+of existing theory and, after having understood this thoroughly, generalize our
+approach to the full realistic 3D problem.
+
+I would like to thank Professor Dong Lai for advising me both as a lab member
+and as a first year graduate student during the past year. I would also like to
+thank Daniel Lecoanet for explaining his unending back of numerical tricks for
+using \lstinline{Dedalus} to me.
 

### 2018-08-14 writeups/notes.tex

@@
     \begin{cases}
-        -e^{(z - z_0)/2H}e^{i(k_xx + k_z(z - z_0) - \omega t)} & z > z_0\\
-        e^{(z - z_0)/2H}e^{i(k_xx + k_z(z - z_0) - \omega t)} & z < z_0\\
+        e^{(z - z_0)/2H}e^{i(k_xx + k_z(z - z_0) - \omega t)} & z > z_0\\
+        e^{(z - z_0)/2H}e^{i(k_xx - k_z(z - z_0) - \omega t)} & z < z_0\\
     \end{cases}.
@@
 $\Gamma \sim \p*{A^2\nu\omega_0''}^{-1/2}$.
+
+\chapter{Stratified Flow}
+
+\section{Linear Breaking Criterion}
+
+Since we're having numerical difficulties, we will simultaneously consider
+\emph{corotation resonances} as a pathway to instability. This corresponds in
+the 2D case to a shear flow $\vec{u}_0 = u_0(z)\hat{x}$. We can do the math for
+this out, but the end result produces dispersion relation
+\begin{equation}
+    \p*{\omega - u_0k_x}^2 = \frac{k_x^2}{k_x^2 + k_z^2 + \frac{1}{4H^2}}
+        \p*{N^2 - \omega \pd{u_0}{z}}.
+\end{equation}
+Since our fluid is generally not stratified strongly and $\omega \ll N$ in the
+IGW regime, we can drop the $\pd{u_0}{z}$ term and the shear flow becomes a
+simple Doppler shift of $\omega$. Denote $\omega_{cr}(z) = \omega - u_0(z)k_x$
+the corotating frequency. A singularity is obvious at $\omega_{cr} = 0$, exactly
+the \emph{corotation resonance}, and one expects that as
+\begin{equation}
+    k_z(z) \approx \frac{k_xN}{\omega_{cr}(z)} \sim \frac{\omega_{cr}}{u_{1z}},
+        \label{eq:corot_nonlin_cond}
+\end{equation}
+that nonlinear effects should kick in.
+
+One advantage of a shear flow is that, instead of requiring $2H$ of vertical
+height for amplitude growth, we can just choose $u_0 =
+\frac{\omega}{k_x}\frac{z}{z_{\max}}$ to meet the corotation citerion at $z =
+z_{\max}$. As $\pd{u_0}{z} \sim \frac{\omega}{k_xz_{\max}} \approx
+\frac{N}{k_zz_{\max}}$, to meet the $N^2 \gg \omega \pd{u_0}{z}$ criterion we
+require $N \gg \frac{\omega}{k_zz_{\max}}$ which is always true. We can even
+induce the corotation resonance well below $z_{\max}$ and still be perfectly
+within the regime of validity of our assumption.
+
+A second advantage of a shear flow is that we can force at very small amplitudes
+and still observe resonances, since the resonance is driven by the convergence
+of $\omega_{cr} \to 0$ (purely coordinate driven) instead of $u_{1z} \sim
+v_{ph}$ (dependent on wave amplitude).
+
+The width of the dissipation zone can be estimated by noticing $u_0k_x = \omega
+- \sqrt{k_xNu_{1z}}$ from the nonlinearity criterion
+\autoref{eq:corot_nonlin_cond}. Thus, the width is roughly of order
+$2\sqrt{k_xNu_{1z}}$, which has the right scaling behavior (a stronger
+perturbation should be resonant farther away). Note however that, since $\omega$
+is small, this width can easily be large compared to $\omega$!
 

### 2018-08-27 writeups/notes.tex

--- 2018-08-02 writeups/notes.tex
@@
 $\Gamma \sim \p*{A^2\nu\omega_0''}^{-1/2}$.
+
+\chapter{Stratified Flow}
+
+\section{Linear Equations}
+
+We largely follow Booker \& Bretherton from 1966 though do not adopt the
+Boussinesq approximation. Consider background shear flow $\vec{u}_0 =
+u_0(z)\hat{x}$ and vertical stratification $\rho_0 \propto e^{-z/H}$, then
+\begin{subequations}\label{se:shear_eq}
+    \begin{align}
+        \vec{\nabla} \cdot \vec{u}_1 &= 0,\\
+        \pd{\rho_1}{t} + \p*{\vec{u} \cdot \vec{\nabla}}\rho_1
+            - u_{1z}\frac{\rho_0}{H} &= 0,\\
+        \pd{u_{1x}}{t} + \p*{\vec{u} \cdot \vec{\nabla}}\p*{u_{1x} + u_0}
+            + \frac{1}{\rho_0}\pd{P}{x} &= 0,\\
+        \pd{u_{1z}}{t} + \p*{\vec{u} \cdot \vec{\nabla}}u_{1z}
+            + \frac{1}{\rho_0}\pd{P}{z} + \frac{\rho_1 g}{\rho_0} &= 0.
+    \end{align}
+\end{subequations}
+Note that $\vec{u} \cdot \vec{\nabla} = \p*{\vec{u}_0 + \vec{u}_1} \cdot
+\vec{\nabla}$.
+
+We again assume plane waves in $x, t$ such that $u_{1z}(x, z, t) =
+u_{1z}(z)e^{i(k_xx - \omega t)}$. We bracket second-order terms below and set
+them to zero in the ensuing calculations
+\begin{subequations}
+    \begin{align}
+        \pd{u_{1z}}{z} + ik_xu_{1x} &= 0,\\
+        -i\omega \rho_1 + u_0ik_x\rho_1
+            - u_{1z}\frac{\rho_0}{H}
+            + \s*{\p*{\vec{u}_1 \cdot \vec{\nabla}} \rho_1} &= 0,\\
+        -i\omega u_{1x} + u_0ik_x u_{1x}
+            + \frac{1}{\rho_0}ik_xP + u_{1z}\pd{u_0}{z}
+            + \s*{\vec{u}_1 \cdot \vec{\nabla} u_{1x}} &= 0,\\
+        -i\omega u_{1z} + u_0ik_x u_{1z}
+            + \frac{1}{\rho_0}\pd{P}{z} + \frac{\rho_1 g}{\rho_0}
+            + \s*{\vec{u}_1 \cdot \vec{\nabla} u_{1z}} &= 0.\\
+    \end{align}
+\end{subequations}
+Denote $\omega_{cr}(z) = \omega - u_0(z)k_x$ the corotating frequency. Using
+then that $u_{1x} = \frac{i}{k_x}\pd{u_{1z}}{z}$, $\rho_1 = \frac{i
+\rho_0}{H\omega_{cr}} u_{1z}$ and $P = \s*{i\omega_{cr}u_{1x} - u_{1z
+\pd{u_0}{z}}} \frac{\rho_0}{ik_x}$, we can finally arrive at linearized equation
+\begin{equation}
+    \ptd{u_{1z}}{z} - \frac{1}{H}\pd{u_{1z}}{z} + \s*{
+        \frac{k_x^2N^2}{\omega_{cr}^2}
+            - k_x^2
+            - \frac{k_x}{\omega_{cr}H}\pd{u_0}{z}
+            + \frac{k_x}{\omega_{cr}} \ptd{u_0}{z}
+    } u_{1z} = 0. \label{eq:shear_lin_uz}
+\end{equation}
+Note that if we make correspondences $ik_x \to \partial_x, i\omega_{cr} \to
+\partial_t + u_0\partial_x$ and take $H \to \infty$ while holding $N^2$
+constant, we recover Equation 2.1 of Booker \& Bretherton. We can also note that
+if we take constant $u_0$ we recover the expected dispersion relation
+\autoref{eq:strat_disp} for $\omega = \omega_{cr}$.
+
+\section{Linear Amplification}
+
+One can solve \autoref{eq:shear_lin_uz}
+
+\section{Breaking Criterion}
+
+If the fluid is not strongly stratified $\ptd{u_0}{z}, \pd{u_0}{z}\frac{1}{H}
+\ll \frac{k_xN^2}{\omega_{cr}}$ then the singularity is largely induced by
+$\omega_{cr} \to 0$ (note that the condition becomes easier to satisfy as we
+approach the singularity). $\omega_{cr} = 0$, is exactly the \emph{corotation
+resonance}, and one expects that as (note that if $k_z \gg k_x$ then the
+dispersion relation is $\omega_{cr} \approx Nk_x/k_z$):
+\begin{equation}
+    k_z(z) \approx \frac{Nk_x}{\omega_{cr}(z)}
+        \sim \frac{\omega_{cr}(z)}{\abs*{u_{1z}}},
+        \label{eq:corot_nonlin_cond}
+\end{equation}
+that nonlinear effects should kick in.
+
+One advantage of a shear flow is that, instead of requiring $2H$ of vertical
+height for amplitude growth, we can just choose $u_0 =
+\frac{\omega}{k_x}\frac{z}{z_{\max}}$ to meet the corotation citerion at $z =
+z_{\max}$. Consider $\pd{u_0}{z} \sim \frac{\omega}{k_xz_{\max}} \approx
+\frac{N}{k_zz_{\max}}$. This both satisfies $\pd{u_0}{z} \frac{1}{H} =
+\frac{N}{k_zz_{\max}H} \ll \frac{k_xN^2}{\omega_{cr}}$ (as $k_z^2z_{\max}H \ll
+1$), and $\ptd{u_0}{z} = 0 \ll \frac{k_xN^2}{\omega_{cr}}$. We can even induce
+the corotation resonance well below $z_{\max}$ and still be perfectly within the
+regime of validity of our assumption.
+
+A second advantage of a shear flow is that we can force at very small amplitudes
+and still observe resonances, since the resonance is driven by the convergence
+of $\omega_{cr} \to 0$ (purely coordinate driven) instead of $u_{1z} \sim
+v_{ph}$ (dependent on wave amplitude).
+
+The width of the dissipation zone can be estimated by noticing $u_0k_x = \omega
+\pm \sqrt{k_xNu_{1z}}$ from the nonlinearity criterion
+\autoref{eq:corot_nonlin_cond}. Thus, the width of $u_0$ that incurs nonlinear
+effects is roughly of order $2\sqrt{Nu_{1z} / k_x}$, which has the right scaling
+behavior (a stronger perturbation should be resonant farther away).
 

### 2018-09-09 writeups/notes.tex
check in writeup progress

@@
 
-\section{Linear Amplification}
-
-One can solve \autoref{eq:shear_lin_uz}
-
-\section{Breaking Criterion}
-
-If the fluid is not strongly stratified $\ptd{u_0}{z}, \pd{u_0}{z}\frac{1}{H}
-\ll \frac{k_xN^2}{\omega_{cr}}$ then the singularity is largely induced by
-$\omega_{cr} \to 0$ (note that the condition becomes easier to satisfy as we
-approach the singularity). $\omega_{cr} = 0$, is exactly the \emph{corotation
-resonance}, and one expects that as (note that if $k_z \gg k_x$ then the
-dispersion relation is $\omega_{cr} \approx Nk_x/k_z$):
-\begin{equation}
-    k_z(z) \approx \frac{Nk_x}{\omega_{cr}(z)}
-        \sim \frac{\omega_{cr}(z)}{\abs*{u_{1z}}},
-        \label{eq:corot_nonlin_cond}
-\end{equation}
-that nonlinear effects should kick in.
-
-One advantage of a shear flow is that, instead of requiring $2H$ of vertical
-height for amplitude growth, we can just choose $u_0 =
-\frac{\omega}{k_x}\frac{z}{z_{\max}}$ to meet the corotation citerion at $z =
-z_{\max}$. Consider $\pd{u_0}{z} \sim \frac{\omega}{k_xz_{\max}} \approx
-\frac{N}{k_zz_{\max}}$. This both satisfies $\pd{u_0}{z} \frac{1}{H} =
-\frac{N}{k_zz_{\max}H} \ll \frac{k_xN^2}{\omega_{cr}}$ (as $k_z^2z_{\max}H \ll
-1$), and $\ptd{u_0}{z} = 0 \ll \frac{k_xN^2}{\omega_{cr}}$. We can even induce
-the corotation resonance well below $z_{\max}$ and still be perfectly within the
-regime of validity of our assumption.
-
-A second advantage of a shear flow is that we can force at very small amplitudes
-and still observe resonances, since the resonance is driven by the convergence
-of $\omega_{cr} \to 0$ (purely coordinate driven) instead of $u_{1z} \sim
-v_{ph}$ (dependent on wave amplitude).
-
-The width of the dissipation zone can be estimated by noticing $u_0k_x = \omega
-\pm \sqrt{k_xNu_{1z}}$ from the nonlinearity criterion
-\autoref{eq:corot_nonlin_cond}. Thus, the width of $u_0$ that incurs nonlinear
-effects is roughly of order $2\sqrt{Nu_{1z} / k_x}$, which has the right scaling
-behavior (a stronger perturbation should be resonant farther away).
+\section{Reflection/Transmission in Linear Theory}
+
+One can solve \autoref{eq:shear_lin_uz} via \emph{Frobenius Theory}, but if we
+consider only $\omega_{cr} \to 0$, we obtain a very simple power series
+solution. In $\omega_{cr} \to 0$, we have only leading order contributions
+\begin{equation}
+    \ptd{u_{1z}}{z} + \frac{k_x^2N^2}{\omega_{cr}^2}u_{1z} = 0.
+\end{equation}
+Consider a power series solution $u_{1z} \propto \omega_{cr}^\alpha$, where
+$\omega_{cr}(z) = \omega - k_xu_0(z)$. Note that very near the critical level,
+$\pd{u_0}{z}\delta z k_x$ where $\delta z = z - z_c, \omega_{cr}(z_c) = 0$. We
+thus have
+\begin{align}
+    \p*{\pd{\omega_{cr}}{z}}^2 \alpha\p*{\alpha - 1} + k_x^2N^2 &= 0,\\
+    \alpha &= \frac{1}{2}\pm i\sqrt{\frac{N^2}{\p*{\pd{u_0}{z}}^2} -
+        \frac{1}{4}}.
+\end{align}
+The classic Booker \& Bretherton paper notates this result $u_{1z} \propto
+\p*{\omega_{cr}(z)}^{\frac{1}{2} \pm i\mu} \propto (\delta z)^{\frac{1}{2} \pm
+i\mu}$.
+
+Now, consistent with a perturbation that vanishes as $t \to -\infty$
+(alternatively, this is just the Landau prescription for deforming below a
+singularity on the real line), we must interpret $\omega_{cr}$ as having a
+slight positive imaginary component. Given this, the expression for $u_{1z}(z)$
+is multivalued, and we must specify a branch cut. For $\Arg\p*{z - z_c} = 0$, or
+intuitively when we are above the critical layer, choose branch such that
+\begin{equation}
+    \p*{z - z_c}^{\frac{1}{2} \pm i\mu} \equiv
+        \abs*{z - z_c}^{1/2\pm i\mu}.
+\end{equation}
+It is then clear that for $\Arg\p*{z - z_c} = -\pi$, intuitively corresponding
+to being below the critical layer (note the arg is $-\pi$ since $z_c$ is
+infinitesimally above the real axis), our branch cut enforces
+\begin{equation}
+    \p*{z - z_c}^{\frac{1}{2} \pm i\mu} =
+        \abs*{z - z_c}^{\frac{1}{2} \pm i\mu}e^{-\frac{i\pi}{2}}e^{\mp \mu \pi}.
+\end{equation}
+Lastly, we interpret $\abs*{z - z_c}^{\pm i\mu} = e^{\pm i\mu \ln \abs*{z -
+z_c}}$ to have the form of a positive/negative $k_z$ value. Interestingly, this
+means that $e^{-i\mu \ln \abs*{z - z_c}}$ corresponds to the \emph{positive
+group velocity}, which is $e^{\mu \pi}$ magnified compared to the $\Arg\p*{z -
+z_c} = 0$ case, and $e^{2\mu \pi}$ magnified over the negative group velocity
+component below $z_c$.
+
+The interpretation we choose for this is that an incoming wavepacket moving with
+positive group velocity inbound on $z_c$ is transmitted with amplitude $e^{-\mu
+\pi}$ and reflected with amplitude $e^{-2\mu \pi}$. The remainder is interpreted
+as being \emph{absorbed} by the critical layer.
+
+\section{Validity of the Linear Approximation}
+
+The real physics is expected to deviate from our computation near the critical
+layer owing primarily to two effects: nonlinear advective terms and diffusivity.
+Hazel 1967 found numerically that including diffusive tersm in the equations of
+motion does not change the attenuation factor; even in the presence of
+dissipation no reflection/transimission are observed. This makes sense;
+examining the equations of motion, the effect of an extra $\pd{}{t} \to \pd{}{t}
+- \nu \nabla^2$ is to take $\omega_{cr} \to \omega_{cr} + i\nu k_x^2 - i\nu
+\ptd{}{z}$, giving the singularity at $\omega_{cr} + i\nu k_x^2 = 0$ the
+positive imaginary component we before asserted.
+
+On the other hand, other studies e.g.\ Brown \& Stewartson 1980, Winters \&
+d'Asaro 1994, find that once the wave amplitude becomes nonnegligible, sizable
+reflection can be found. In contrast, Ogilvie \& Barker 2010 find that in
+geometrically focused breaking, no reflection is observed, despite the nonlinear
+effects being stronger than the dissipitive ones. The study of how nonlinear and
+diffusive effects compete is then nontrivial.
 

### 2018-10-07 writeups/notes.tex
check in new meeting notes
stash

@@
 
-\section{Algebraic Manipulations}
-
-\subsection{Factoring out the exponential dependence}
-
-I've had to do this algebra too many times now, so let's just consider if we
-factor out the $e^{z/2H}$ exponential scaling on the dynamical variables and the
-$e^{-z/H}$ scaling on $\rho_0$. This leaves us with:
-\begin{subequations}\label{se:nonlin_factored}
-    \begin{align}
-        \vec{\nabla} \cdot \vec{u}_1 + \frac{u_{1z}}{2H} &= 0,\\
-        \pd{\rho_1}{t} - \frac{u_{1z} \rho_0}{H} &= -e^{z/2H}\s*{
-            \p*{\vec{u}_1 \cdot \vec{\nabla} \rho_1} + \frac{u_{1z}\rho_1}{2H}
-        },\\
-        \pd{\vec{u}_1}{t} + \frac{\vec{\nabla}P}{\rho_0}
-            + \frac{P}{2H\rho_0}\hat{z} + \frac{\rho_1\vec{g}}{\rho_0}
-            &= -e^{z/2H}\s*{
-            \p*{\vec{u} \cdot \vec{\nabla} \vec{u}_1}
-                + \frac{u_{1z}\vec{u}_1}{2H}
-        }.
-    \end{align}
-\end{subequations}
-I'm too lazy to re-notate my variables, so all of them are constant amplitude
-over $z$ now.
+\section{Navier-Stokes Regularization}
+
+This was a major tripping point for quite a while. As the wave breaks to higher
+wavenumbers, it is expected that energy will accumulate in the higher $k$
+spectral modes, and so some numerical viscosity must be added. This is not a
+physical viscosity but a numerical one (since spectral codes have no numerical
+viscosity of their own), and so the viscous term is added to all dynamical
+equations $\rd{}{t} \to \pd{}{t} + \vec{u} \cdot \vec{\nabla} - \nu \nabla^2$.
+This makes the equations resemble the Navier-Stokes equations, but since $\nu$
+is not grounded in any physical process, it is not helpful to think of it as a
+Navier-Stokes equation.
+
+We choose the $\nu \nabla^2$ term, which scales $\sim k^2$, such that it becomes
+strong in the same regime that nonlinearity $\p*{\vec{u} \cdot \vec{\nabla}}$
+becomes important. Consider some $k_{turb}$ turbulent wavenumber, then we will
+want $\nu k_{turb}^2 \sim uk_{turb}$. We can get an idea of what $u$ will be
+when turbulence sets in using the wave breaking criterion, that breaking sets in
+when $u \sim \omega/k$ where $k$ is the base wavelength of the wave. This
+substitution suggests we should use $\nu = \frac{\omega}{k_{turb}k}$. We can
+choose $k_{turb} = \frac{2\pi L}{N}$ the grid spacing, so that at this turbulent
+wavelength the regularization kicks in. If we then take $k, \omega$ to be the
+parent wave's properties, we have a prescription for $\nu$.
+
+This turns out to be insufficient for regularization, we need roughly $\nu \sim
+\frac{3\omega_0}{k_zk_{turb}}$.
+
+This is because we performed the naive substitution $k, \omega$ from the parent
+wave. Instead, it is advisable to scale $k, \omega$ to the same $k_{turb}$
+length scale. We make ansatz that the cascade is roughly Kolmogorov, and so
+(TODO work this out) we obtain that we should choose $\frac{\omega_0}{k_z^2 \nu}
+\sim \p*{\frac{k_{turb}}{k_z}}^{4/3}$; this comes from scaling $\omega, k$ per
+the Kolmogorov prescription $\omega_l u_l^2$ is approximately constant over
+length scale $l$.
+
+\subsection{Gotcha: $\nu_x, \nu_z$ and Grid Spacing}
+
+If the grid resolution is uneven in the $x, z$ directions, it might be tempting
+to use different $\nu_x, \nu_z$ per the above criteria, since their $k_{turb} =
+k_{\max}$ differ. This is dangerous however, since $\nu \nabla^2 \vec{u}_1$ has
+to be divergenceless! Thus, it is more advisable to choose an equal grid spacing
+in the two directions.
 
@@
 $\omega_{cr}(z) = \omega - k_xu_0(z)$. Note that very near the critical level,
-$\pd{u_0}{z}\delta z k_x$ where $\delta z = z - z_c, \omega_{cr}(z_c) = 0$. We
-thus have
+$\pd{u_0}{z}\delta z k_x = \omega_{cr}$ where $\delta z = z - z_c,
+\omega_{cr}(z_c) = 0$. We thus have
 \begin{align}
@@
 
+\section{Algebraic Manipulations}
+
+\subsection{Factoring out the exponential dependence}
+
+I've had to do this algebra too many times now, so let's just consider if we
+factor out the $e^{z/2H}$ exponential scaling on the dynamical variables and the
+$e^{-z/H}$ scaling on $\rho_0$. This leaves us with:
+\begin{subequations}\label{se:nonlin_factored}
+    \begin{align}
+        \vec{\nabla} \cdot \vec{u}_1 + \frac{u_{1z}}{2H} &= 0,\\
+        \pd{\rho_1}{t} - \frac{u_{1z} \rho_0}{H} &= -e^{z/2H}\s*{
+            \p*{\vec{u}_1 \cdot \vec{\nabla} \rho_1} + \frac{u_{1z}\rho_1}{2H}
+        },\\
+        \pd{\vec{u}_1}{t} + \frac{\vec{\nabla}P}{\rho_0}
+            + \frac{P}{2H\rho_0}\hat{z} + \frac{\rho_1\vec{g}}{\rho_0}
+            &= -e^{z/2H}\s*{
+            \p*{\vec{u} \cdot \vec{\nabla} \vec{u}_1}
+                + \frac{u_{1z}\vec{u}_1}{2H}
+        }.
+    \end{align}
+\end{subequations}
+I'm too lazy to re-notate my variables, so all of them are constant amplitude
+over $z$ now.
+
 \end{document}

### 2018-10-15 writeups/notes.tex
check in 2d_1_gradual param scan, mostly working

@@
 wave. Instead, it is advisable to scale $k, \omega$ to the same $k_{turb}$
-length scale. We make ansatz that the cascade is roughly Kolmogorov, and so
-(TODO work this out) we obtain that we should choose $\frac{\omega_0}{k_z^2 \nu}
-\sim \p*{\frac{k_{turb}}{k_z}}^{4/3}$; this comes from scaling $\omega, k$ per
-the Kolmogorov prescription $\omega_l u_l^2$ is approximately constant over
-length scale $l$.
+length scale. We first make ansatz that the cascade is roughly Kolmogorov, in
+which case we do the hand-wavy argument that energy $u_l^2$ at a certain length
+scale $l$ is dissipated on timescale $\omega_l = u_l / l$, and so the total
+energy dissipated from the original flow $\epsilon = u_l^3/l$. Thus, $\nu k_l^2
+\sim u_lk_l$ gives $\frac{\omega_0}{\nu k_0^2}\p*{\frac{k_0}{k_{turb}}}^{4/3} =
+1$. This isn't quite right though, as per comparison.
+
+My hypothesis is currently two fold: first, the turbulent cascade is not
+instantaneous, so we can solve a diffusion equation in $k$ space to figure out
+how far the wave can propagate (and grow in amplitude) before the Kolmogorov
+spectrum is obtained. Second, the turbulence causes some energy to be reflected
+in higher $k$ modes, which interact nonlinearly with the flow at lower
+amplitudes.
 
@@
 
-\section{Validity of the Linear Approximation}
-
-The real physics is expected to deviate from our computation near the critical
-layer owing primarily to two effects: nonlinear advective terms and diffusivity.
-Hazel 1967 found numerically that including diffusive tersm in the equations of
-motion does not change the attenuation factor; even in the presence of
-dissipation no reflection/transimission are observed. This makes sense;
-examining the equations of motion, the effect of an extra $\pd{}{t} \to \pd{}{t}
-- \nu \nabla^2$ is to take $\omega_{cr} \to \omega_{cr} + i\nu k_x^2 - i\nu
-\ptd{}{z}$, giving the singularity at $\omega_{cr} + i\nu k_x^2 = 0$ the
-positive imaginary component we before asserted.
-
-On the other hand, other studies e.g.\ Brown \& Stewartson 1980, Winters \&
-d'Asaro 1994, find that once the wave amplitude becomes nonnegligible, sizable
-reflection can be found. In contrast, Ogilvie \& Barker 2010 find that in
-geometrically focused breaking, no reflection is observed, despite the nonlinear
-effects being stronger than the dissipitive ones. The study of how nonlinear and
-diffusive effects compete is then nontrivial.
+\section{Competition between Viscous and Nonlinear Forces}
+
+As we zoom in on the critical layer, we expect to be able to take Boussinesq
+approximation $H \to \infty$ within the immediate vicinity of the critical
+layer.
+
+In this limit, the group velocity is just $v_{g,z} = -\frac{Nk_xk_z}{\p*{k_x^2 +
+k_z^2}^{3/2}} = -\omega\frac{k_z}{k_x^2 + k_z^2}$. As we approach corotation,
+$\omega \to 0, \omega \approx \frac{Nk_x}{k_z}, k_z \to \infty$, and so $v_{g,
+z} \to 0$. Then considering an energy flux incident on the critical layer
+clearly shows that energy must accumulate near the critical layer, and must be
+dissipated somehow. This must be either via viscous forces or nonlinear
+cascades.
+
+The characteristic strength of the viscous force $\nu \nabla^2$ is $\nu k_z^2$,
+while the nonlinear advective term has characteristic strength $u_{1z}k_z$. Note
+that in the Boussinesq approximation $\abs*{u_{1z}}$ is constant during
+propagation. Thus, the ratio of forces $R = \frac{f_{\nu}}{f_{adv}} \sim
+\frac{\nu k_z}{u_{1z}}$. Note that this ratio varies spatially. Qualitatively,
+$R$ is important since it dictates whether the wave is damped viscously before
+it needs to be treated in nonlinear fashion or whether a full
+nonlinear/hydrodynamic treatment is necessary.
+
+We want to understand for what viscosity we expect linear behavior to dominate.
+Thus, at critical $z_{c}$ where nonlinear effects become important, we want $R
+\gtrsim 1$. Shifting to coordinate system $\delta z$ where $\delta z = 0$ is the
+corotation resonance, we can write $\abs*{\omega_{cr}} = k_xu_0' \delta z$ where
+$u_0' = \at{\pd{u_0}{z}}_{\delta z = 0}$, and $k_z = \frac{N}{u_0' \delta z}$.
+Then the breaking criterion is $\frac{u_{1z}k_z(z)}{\omega_{cr}(z)} \sim 1$, or
+$\delta z_{c} \sim \sqrt{\frac{u_{1z}N}{k_xu_0'^2}}$. We can then evaluate
+\begin{equation}
+    R\p*{z_{c}} = \frac{\nu k_z}{u_{1z}} = \nu \sqrt{\frac{Nk_x}{u_{1z}^3}}.
+\end{equation}
+
+It is helpful also to write this down for any arbitrary hyperviscosity $f_{visc}
+= \nu^{(n)} \nabla^{(n)}$, which is cited without proof
+\begin{equation}
+    R\p*{z_c} = \nu^{(n)} \sqrt{\frac{N^{n - 1}k_x^{n - 1}}{u_{1z}^{n + 1}}}.
+\end{equation}
 

### 2018-10-15 sims/2d_1_gradual/strat.py

@@ -81,20 +81,32 @@ if __name__ == '__main__':
                                               'USE_CFL': True})),
         (zero_ic, 'nl3_lowres',
          build_interp_params(2, 2, overrides={'F_MULT': 2,
+                                              'NU_MULT': 3,
+                                              'USE_CFL': True})),
+        (zero_ic, 'nl4_lowres',
+         build_interp_params(2, 2, overrides={'F_MULT': 2,
                                               'NU_MULT': 4,
                                               'USE_CFL': True})),
-        # (zero_ic, 'nl1',
-        #  build_interp_params(1, 1, overrides={'F_MULT': 2,
-        #                                       'NU_MULT': 1,
-        #                                       'USE_CFL': True})),
-        # (zero_ic, 'nl2',
-        #  build_interp_params(1, 1, overrides={'F_MULT': 2,
-        #                                       'NU_MULT': 1.4,
-        #                                       'USE_CFL': True})),
-        # (zero_ic, 'nl3',
-        #  build_interp_params(1, 1, overrides={'F_MULT': 2,
-        #                                       'NU_MULT': 2,
-        #                                       'USE_CFL': True})),
+        (zero_ic, 'nl0',
+         build_interp_params(1, 1, overrides={'F_MULT': 4,
+                                              'NU_MULT': 1,
+                                              'USE_CFL': True})),
+        (zero_ic, 'nl1',
+         build_interp_params(1, 1, overrides={'F_MULT': 4,
+                                              'NU_MULT': 2,
+                                              'USE_CFL': True})),
+        (zero_ic, 'nl2',
+         build_interp_params(1, 1, overrides={'F_MULT': 4,
+                                              'NU_MULT': 3,
+                                              'USE_CFL': True})),
+        (zero_ic, 'nl3',
+         build_interp_params(1, 1, overrides={'F_MULT': 4,
+                                              'NU_MULT': 4,
+                                              'USE_CFL': True})),
+        (zero_ic, 'nl4',
+         build_interp_params(1, 1, overrides={'F_MULT': 4,
+                                              'NU_MULT': 6,
+                                              'USE_CFL': True})),
     ]
     if '-plot' not in sys.argv:
         for task in tasks:

### 2018-10-22 writeups/notes.tex
check in new notes

@@
 \end{subequations}
-Denote $\omega_{cr}(z) = \omega - u_0(z)k_x$ the corotating frequency. Using
+Denote $\tilde{\omega}(z) = \omega - u_0(z)k_x$ the corotating frequency. Using
 then that $u_{1x} = \frac{i}{k_x}\pd{u_{1z}}{z}$, $\rho_1 = \frac{i
-\rho_0}{H\omega_{cr}} u_{1z}$ and $P = \s*{i\omega_{cr}u_{1x} - u_{1z
-\pd{u_0}{z}}} \frac{\rho_0}{ik_x}$, we can finally arrive at linearized equation
+\rho_0}{H\tilde{\omega}} u_{1z}$ and $P = \s*{i\tilde{\omega}u_{1x} - u_{1z}
+\pd{u_0}{z}} \frac{\rho_0}{ik_x}$, we can finally arrive at linearized equation
 \begin{equation}
     \ptd{u_{1z}}{z} - \frac{1}{H}\pd{u_{1z}}{z} + \s*{
-        \frac{k_x^2N^2}{\omega_{cr}^2}
+        \frac{k_x^2N^2}{\tilde{\omega}^2}
             - k_x^2
-            - \frac{k_x}{\omega_{cr}H}\pd{u_0}{z}
-            + \frac{k_x}{\omega_{cr}} \ptd{u_0}{z}
+            - \frac{k_x}{\tilde{\omega}H}\pd{u_0}{z}
+            + \frac{k_x}{\tilde{\omega}} \ptd{u_0}{z}
     } u_{1z} = 0. \label{eq:shear_lin_uz}
 \end{equation}
-Note that if we make correspondences $ik_x \to \partial_x, i\omega_{cr} \to
+Note that if we make correspondences $ik_x \to \partial_x, i\tilde{\omega} \to
 \partial_t + u_0\partial_x$ and take $H \to \infty$ while holding $N^2$
@@
 if we take constant $u_0$ we recover the expected dispersion relation
-\autoref{eq:strat_disp} for $\omega = \omega_{cr}$.
+\autoref{eq:strat_disp} for $\omega = \tilde{\omega}$.
+
+As a brief exercise, we can take the $H \to \infty$ limit and compute the
+resultant shear-induced splitting; this is akin to rotational mode splitting,
+and is given without proof from $-k_z^2 - k_x^2 + \frac{k_x^2N^2}{\omega^2} +
+\frac{k_x}{\omega}\ptd{u_0}{z} = 0$ as
+\begin{equation}
+    \omega^2 = \frac{k_x^2 N^2}{k_x^2 + k_z^2}\s*{1 + \frac{1}{\sqrt{k_x^2 +
+        k_z^2}} \ptd{u_0}{z}}.
+\end{equation}
 
@@
 One can solve \autoref{eq:shear_lin_uz} via \emph{Frobenius Theory}, but if we
-consider only $\omega_{cr} \to 0$, we obtain a very simple power series
-solution. In $\omega_{cr} \to 0$, we have only leading order contributions
-\begin{equation}
-    \ptd{u_{1z}}{z} + \frac{k_x^2N^2}{\omega_{cr}^2}u_{1z} = 0.
-\end{equation}
-Consider a power series solution $u_{1z} \propto \omega_{cr}^\alpha$, where
-$\omega_{cr}(z) = \omega - k_xu_0(z)$. Note that very near the critical level,
-$\pd{u_0}{z}\delta z k_x = \omega_{cr}$ where $\delta z = z - z_c,
-\omega_{cr}(z_c) = 0$. We thus have
-\begin{align}
-    \p*{\pd{\omega_{cr}}{z}}^2 \alpha\p*{\alpha - 1} + k_x^2N^2 &= 0,\\
+consider only $\tilde{\omega} \to 0$, we obtain a very simple power series
+solution. In $\tilde{\omega} \to 0$, we have only leading order contributions
+\begin{equation}
+    \ptd{u_{1z}}{z} + \frac{k_x^2N^2}{\tilde{\omega}^2}u_{1z} = 0.
+\end{equation}
+Consider a power series solution $u_{1z} \propto \tilde{\omega}^\alpha$, where
+$\tilde{\omega}(z) = \omega - k_xu_0(z)$. Note that very near the critical level,
+$\pd{u_0}{z}\delta z k_x = \tilde{\omega}$ where $\delta z = z - z_c,
+\tilde{\omega}(z_c) = 0$. We thus have
+\begin{align}
+    \p*{\pd{\tilde{\omega}}{z}}^2 \alpha\p*{\alpha - 1} + k_x^2N^2 &= 0,\\
     \alpha &= \frac{1}{2}\pm i\sqrt{\frac{N^2}{\p*{\pd{u_0}{z}}^2} -
@@
 The classic Booker \& Bretherton paper notates this result $u_{1z} \propto
-\p*{\omega_{cr}(z)}^{\frac{1}{2} \pm i\mu} \propto (\delta z)^{\frac{1}{2} \pm
+\p*{\tilde{\omega}(z)}^{\frac{1}{2} \pm i\mu} \propto (\delta z)^{\frac{1}{2} \pm
 i\mu}$.
@@
 (alternatively, this is just the Landau prescription for deforming below a
-singularity on the real line), we must interpret $\omega_{cr}$ as having a
+singularity on the real line), we must interpret $\tilde{\omega}$ as having a
 slight positive imaginary component. Given this, the expression for $u_{1z}(z)$
@@
 \gtrsim 1$. Shifting to coordinate system $\delta z$ where $\delta z = 0$ is the
-corotation resonance, we can write $\abs*{\omega_{cr}} = k_xu_0' \delta z$ where
+corotation resonance, we can write $\abs*{\tilde{\omega}} = k_xu_0' \delta z$ where
 $u_0' = \at{\pd{u_0}{z}}_{\delta z = 0}$, and $k_z = \frac{N}{u_0' \delta z}$.
-Then the breaking criterion is $\frac{u_{1z}k_z(z)}{\omega_{cr}(z)} \sim 1$, or
+Then the breaking criterion is $\frac{u_{1z}k_z(z)}{\tilde{\omega}(z)} \sim 1$, or
 $\delta z_{c} \sim \sqrt{\frac{u_{1z}N}{k_xu_0'^2}}$. We can then evaluate
@@
 \end{equation}
+
+\chapter{Cylindrical Stratified Flow}
+
+We fearlessly tackle the cylindrical coordinate problem, using Tsang \& Lai 2008
+non-barotropic flows as a guide but reproducing all algebra. We show that in an
+appropriate limit it recovers \autoref{eq:shear_lin_uz}.
+
+First, we take $\vec{\nabla} \cdot \vec{u}_0 = 0$, which we can define $u_0 =
+u_0(r)\hat{\phi} = r\Omega(r) \hat{\phi}$. Furthermore, under hydrostatic
+equilibrium if we assume a thin atmosphere (such that $\vec{\nabla}\Phi$ does
+not vary much) then $\frac{\vec{\nabla}P_0}{\rho_0} = -g\hat{r}$. We assume
+stratification $\rho \propto e^{-r/a_0}$, which gives buoyancy frequency $N^2 =
+g/a_0$ in the incompressible fluid limit. Finally, we make plane wave
+substitution $\propto e^{i\p*{m\phi - \omega t}}$ and begin
+\begin{align}
+    0 ={}& \frac{1}{r}\pd{\p*{u_{1r}}}{r} + \frac{im}{r}u_{1\phi},\\
+    0 ={}& -i\omega \rho_1 - u_{1r}\frac{\rho_0}{a_0} + im\Omega\rho_1
+        \z*{+ u_{1r}\pd{\rho_1}{r} + \frac{u_{1\phi}r}{r}\pd{\rho_1}{\phi}}
+            ,\\
+    0 ={}& -i\omega \vec{u}_1
+            + \s*{\Omega im u_{1r} - 2\Omega u_{1\phi}} \hat{r}
+            +\s*{u_{1r}\rd{\p*{r\Omega}}{r} + \Omega imu_{1\phi}
+                + \Omega u_{1r}}\hat{\phi}
+            + \frac{1}{\rho_0}\s*{\pd{P_1}{r}\hat{r} + \frac{imP_1}{r}
+                \hat{\phi}} + \frac{\rho_1 g}{\rho_0}\hat{r}\nonumber\\
+        &+ \z*{\s*{u_{1r}\pd{u_{1r}}{r} + \frac{u_{1\phi}}{r}
+            \pd{u_{1r}}{\phi} - \frac{u_{1\phi}^2}{r}}\hat{r}
+            + \s*{u_{1r}\pd{u_{1\phi}}{r}
+                + \frac{u_{1\phi}}{r}\pd{u_{1\phi}}{\phi}
+                + \frac{u_{1\phi} u_{1r}}{r}}\hat{\phi}}.
+\end{align}
+We have bracketed nonlinear terms in curly braces. The natural definition
+$\tilde{\omega} = \omega - m\Omega$ and common definition $\frac{\kappa^2}{2\Omega}
+= \Omega + \rd{(r\Omega)}{r}$ may be used, which agrees with Tsang \& Lai.
+Dropping bracketed nonlinear terms allows us to manipulate
+\begin{align}
+    u_{1\theta} ={}& \frac{i}{m} \pd{ru_{1r}}{r},\\
+    \rho_1 ={}& u_{1r}\frac{\rho_0}{a_0}\frac{1}{-i\tilde{\omega}},\\
+    P_1 ={}& \frac{i\rho_0 r}{m}\s*{-i\tilde{\omega}u_{1\theta}
+        + u_{1r}\frac{\kappa^2}{2\Omega}},\\
+    0 ={}& -i\tilde{\omega}u_{1r} - 2\Omega\p*{\frac{i}{m}\pd{(ru_{1r})}{r}}
+        + \frac{1}{\rho_0}\pd{}{r}\s*{\frac{i\rho_0 r}{m}\s*{
+            -i\tilde{\omega} \frac{i}{m}\pd{(ru_{1r})}{r}
+            +u_{1r}\frac{\kappa^2}{2\Omega}}}
+        + \frac{iu_{1r}g}{a_0\tilde{\omega}}.
+\end{align}
+Expanding the rather ugly derivative gives
+\begin{align}
+    0 ={}& -i\tilde{\omega}u_{1r} - \frac{2i\Omega}{m}\s*{r\pd{u_{1r}}{r}
+        + u_{1r}} + \p*{\frac{i}{m} - \frac{ir}{a_0m}}
+            \s*{\frac{\tilde{\omega}}{m}
+            \p*{r\pd{u_{1r}}{r} + u_{1r}} + u_{1r}\frac{\kappa^2}{2\Omega}}
+            \nonumber\\
+        &+ \frac{ir}{m}\s*{\frac{\tilde{\omega}}{m}\p*{r\ptd{u_{1r}}{r}
+            + 2\pd{u_{1r}}{r}} + \frac{\kappa^2}{2\Omega}\pd{u_{1r}}{r}
+            + u_{1r}\pd{}{r}\p*{\frac{\kappa^2}{2\Omega}}}
+            \nonumber\\
+        &+ \frac{r}{m}\pd{\tilde{\omega}}{r}\frac{i}{m}\p*{r\pd{u_{1r}}{r}
+            + u_{1r}} + \frac{iu_{1r}g}{a_0\tilde{\omega}}.
+\end{align}
+Let's group terms to find
+\begin{align}
+    \frac{\tilde{\omega}r^2}{m^2}\ptd{u_{1r}}{r} + &
+        \pd{u_{1r}}{r}\s*{-\frac{2\Omega r}{m}
+            + \frac{r\tilde{\omega}}{m} \p*{\frac{1}{m} - \frac{r}{a_0m}}
+            + \frac{2r\tilde{\omega}}{m^2}
+            + \frac{r}{m}\frac{\kappa^2}{2\Omega}
+            + \frac{r^2}{m^2}\pd{\tilde{\omega}}{r}}\nonumber\\
+        & +u_{1r}\s*{-\frac{2\Omega}{m}
+            + \frac{\tilde{\omega}}{m} \p*{\frac{1}{m} - \frac{r}{a_0m}}
+            + \p*{\frac{1}{m} - \frac{r}{a_0m}}\frac{\kappa^2}{2\Omega}
+            + \frac{r}{m}\pd{}{r}\p*{\frac{\kappa^2}{2\Omega}}
+            + \frac{r}{m^2}\pd{\tilde{\omega}}{r}
+            + \frac{N^2}{\tilde{\omega}} - \tilde{\omega}} = 0.
+\end{align}
+One small simplification can be made: observe that $\frac{\kappa^2}{2\Omega} =
+\Omega + \pd{(r\Omega)}{r} = 2\Omega + r\pd{\Omega}{r}$, so
+\begin{equation}
+    \frac{\tilde{\omega}r^2}{m^2}\ptd{u_{1r}}{r}
+        + \pd{u_{1r}}{r}
+            \frac{r\tilde{\omega}}{m^2} \p*{3 - \frac{r}{a_0}}
+        + u_{1r}\s*{
+            \frac{\tilde{\omega}}{m} \p*{\frac{1}{m} - \frac{r}{a_0m}}
+            - \frac{r}{a_0m}\frac{\kappa^2}{2\Omega}
+            + \frac{r}{m}\pd{}{r}\p*{\frac{\kappa^2}{2\Omega}}
+            + \frac{N^2}{\tilde{\omega}} - \tilde{\omega}} = 0.
+\end{equation}
+I think that the $3$ actually makes sense, it's basically
+$\pd{}{r}r\pd{(ru_r)}{r} = \dots + 2r\pd{u_r}{r}$ and $\pd{(ru_r)}{r} =
+r\pd{u_r}{r} + u_r$, which also explains the $1$.
+
+\section{Comparison to Plane Parallel}
+
+\subsection{Reduction in $r \to \infty$}
+
+First, let's consider the limit where $r \to \infty$, so we should have a plane
+parallel waveform. In this limit, $\frac{m}{r} \to k_x$ can be identified (and
+so $m$ must also be taken to be infinitely large to have finite wavelengths),
+and of course $\Omega \to \frac{u_0}{r}$ in line with our plane parallel
+definitions. We need one last identification, that $\pd{\tilde{\omega}}{r} =
+-m\rd{\Omega}{r} \approx -k_x\rd{u_0}{r}$ also per the plane parallel result.
+These identifications allow us to write, approximating $\frac{\kappa^2}{2\Omega}
+\approx \rd{u_0}{r}$:
+\begin{equation}
+    \ptd{u_{1r}}{r} - \pd{u_{1r}}{r}\frac{1}{a_0} +
+        u_{1r}\s*{\frac{N^2k_x^2}{\tilde{\omega}^2}
+            + \frac{k_x}{\omega} \rtd{u_0}{r} - \frac{k_x}{a_0\tilde{\omega}}
+            \rd{u_0}{r} - k_x^2}u_{1r} = 0.
+\end{equation}
+This can be seen to agree with \autoref{eq:shear_lin_uz}, which is no small
+miracle for my usual algebra skills.
+
+\subsection{Boundary Matching at $\tilde{\omega} = 0$}
+
+Again, there is nothing more singular than a $\tilde{\omega}^{-2}$ coefficient
+on $u_{1r}$, after dividing out the leading coefficient, and so we may simply
+analyze
+\begin{equation}
+    \ptd{u_{1r}}{r} + u_{1r}\frac{N^2m^2}{\tilde{\omega}^2r^2} = 0.
+\end{equation}
+Using the usual $u_{1r} = (r - r_c)^\alpha = \delta r^\alpha$ and
+$\tilde{\omega} = -m\at{\pd{\Omega}{r}}_{r_c}\delta r \equiv -m\Omega'_c \delta
+r$ gives
+\begin{equation}
+    \alpha(\alpha - 1) + \frac{N^2}{(\Omega'_c)^2r^2} = 0.\label{eq:cylin_crit}
+\end{equation}
+This solves to $\alpha = \frac{1}{2} \pm
+i\sqrt{\frac{N^2}{(\Omega'_c)^2r^2} - \frac{1}{4}}$. This agrees with
+the exponent derived in the Tsang \& Lai paper. Thus, in the linear theory,
+transmission/reflection are attenuated by factor
+$\exp\s*{-\pi\sqrt{\frac{N^2}{(\Omega'_c)^2 r^2} - \frac{1}{4}}}$.
+
+It can be observed in passing that if $N \sim \Omega'_c r$ then the scenario
+reduces to that considered in Tsang \& Lai, where some variable such as $\zeta$
+can induce superreflection. We don't compute it here, but it bears mentioning
+that the corresponding analysis for Tsang \& Lai in plane parallel gives $\nu =
+-4\rd{\ln u_0'}{r}$, and so $u_0'$ plays the role of $\zeta$.
+
+Note that the H\o iland Criteria, such as given in Balbus \& Hawley 1998, states
+that disks are hydrodynamically stable if
+\begin{equation}
+    N^2 + \frac{1}{R^3}\pd{R^4 \Omega^2}{R} > 0.
+\end{equation}
+We may compare this to our transmission/reflection criterion by computing at
+critical $\Omega_c$, where
+\begin{equation}
+    R^2 \p*{\Omega_c'}^2 = \p*{\frac{N^2 + 4\Omega^2}{2\Omega}}^2.
+\end{equation}
+This was never going to agree with \autoref{eq:cylin_crit}, since it doesn't
+depend simply on $\Omega_c'$. This reinforces Booker \& Bretherton's conclusion
+that the agreement of their calculation with convective stability was pure
+coincidence.
+
+\section{Wave Properties}
+
+Let's compute dispersion relation. I'm too lazy to sort out $u_{1r} \propto
+e^{r/2a_0}$ nonsense, so we will take Boussinesq approximation $a_0 \to \infty$,
+and so
+\begin{equation}
+    \ptd{u_{1r}}{r} + \pd{u_{1r}}{r}\p*{\frac{3}{r} - \frac{1}{a_0}}
+        + u_{1r}\s*{\p*{\frac{1}{r^2} - \frac{1}{a_0r}}
+            + \frac{m}{r\tilde{\omega}} \pd{}{r}\p*{\frac{\kappa^2}{2\Omega}}
+            + \frac{N^2m^2}{r^2\tilde{\omega}^2} - \frac{m^2}{r^2}} = 0.
+\end{equation}
+We've been a bit hand-wavy about what terms to drop, but we can identify now
+that some of the terms can come together to form (we have to drop the
+$\pd{u_{1r}}{r}$ coefficient, otherwise we'll have $ik_r$ terms\dots ugh)
+\begin{equation}
+    \tilde{\omega}^2 \simeq \frac{N^2\p*{m^2/r^2}}{k_r^2 + \frac{m^2}{r^2}}.
+\end{equation}
+Thus, in a nutshell, we should expect the group velocity to largely vanish in
+cylindrical coordinates once the WKB regime is hit (where we're reasonably far
+away that we don't have to worry about the non-wave-like distortion of the
+geometry; is this the same as the plane wave limit though?).
 

### 2018-10-22 writeups/notes.tex
hypothesis abound

@@
 
+\section{Traveling Breaking Front}
+
+Motivated by simulations, we posit the existence of a hydraulic jump, similar to
+shocks in nature. We thus need to develop jump conditions, similar to the
+Rankine-Huginot jump conditions for shocks.
+
+The R-H jump conditions take the conservation equations and integrate over a
+control volume. Since in a hydrodynamic shock, all fluid quantities are
+advected, all conservation equations are necessary. A simpler treatment is
+merited for IGW breaking, since very few quantities are advected (notably,
+vertical momentum and density are not).
+
+Let's consider first the advection of horizontal momentum $\rho_0 u_{1x}$ to
+leading order. This satisfies conservation equation
+\begin{equation}
+    \pd{P_x}{t} + \vec{\nabla} \cdot \p*{\rho_0 u_{1x}\vec{u}_1} = 0.
+\end{equation}
+Integrating over control volume $V = \s*{-\infty, \infty} \times \s*{z < z_b, z
+> z_b}$ the full horizontal domain times the breaking zone, we require the total
+horizontal momentum of the zone be constant in time, if we permit $z_b$ to move
+with velocity $u_b$ the breaking front velocity. This equates to
+\begin{equation}
+    \Delta P_x - \s*{\rho_0 u_{1x}u_b}_{z > z_b}
+        + \s*{\rho_0 u_{1x}\p*{u_{1z} - u_b}}_{z < z_b} = 0.
+\end{equation}
+We've called $\Delta P_x$ the change in horizontal momentum in the breaking zone
+(so the boundary integral over the endcaps of the domain, effectively). Now,
+based on simulations, we identify that $u_{1x}\p*{z > z_c} = B\cos \kappa z$
+seems to form a standing wave pattern. Note that since this is a pure shear
+flow, it is consistent with an incompressible fluid and $u_{1z}\p*{z > z_c} =
+0$.
+
+Permitting such a flow above the breaking zone, then we substitute our known
+linear solution $u_{1x} \approx \frac{k_z}{k_x} u_{1z} \propto Ae^{z / 2H}\cos
+\omega_0 t$ and $\rho_0 = \tilde{\rho}e^{-z/H}$ and evaluate harmonics of $\cos
+\omega_0 t$ to obtain
+\begin{align}
+    \Delta P_x\p*{\omega = 0} &= \tilde{\rho}\frac{A^2}{2}\frac{k_z}{k_x},\\
+    \Delta P_x\p*{\omega = \omega_0}
+        &= u_b Ae^{-z/2H}\tilde{\rho}\cos \omega_0 t,\\
+    \Delta P_k(\omega = 2\omega_0) &=
+        -u_b \cos (\kappa u_bt)
+            + \frac{k_z}{k_x}\frac{A^2}{2B}e^{z/H} \cos 2 \omega_0 t.
+\end{align}
+We can easily interpret $\Delta P_x(0)$ to be the deposition of the horizontal
+momentum of the wave, and $\Delta P_x\p*{\omega = \omega_0}$ to be just the wave
+flowing through the deposition zone; it time averages to zero anyways, so it's
+not super interesting.
+
+Instead, if we dare make identification $u_b = \frac{2\omega_0}{\kappa}$ (the
+only way $u_b \neq 0$), then the deposition to
+\begin{equation}
+    \Delta P_k\p*{\omega = 2\omega_0} = \p*{\frac{A^2}{2B}e^{z/H} -
+        \frac{2\omega_0}{\kappa}}\cos 2\omega_0 t,
+\end{equation}
+i.e.\ the standing wave pattern above can sap some $2\omega_0$ frequency $P_x$
+out of the mean flow via a breaking front.
+
+\subsection{Finding and Explaining the Breaking Front}
+
+It is a bit tricky to interpret this, so I'll just jot down some ideas lest I
+forget them in my sleep-deprived state. If we require that the breaking front
+only sap momentum from the wave (and not from the mean flow), then the breaking
+front shuts of when $\frac{A^2}{2B}e^{z/H} \sim \frac{2\omega_0}{\kappa}$. It's
+also unclear how $\kappa$ is chosen as well. But based on preliminary
+simulations, this is actually pretty accurate for observed $\kappa, u_b$!
+
+Encouragingly, we can guess at where $\kappa$ comes from. Since the waveform
+above the breaking zone has to lose its $k_x$ component, we may guess that the
+incoming wave $\cos \p*{k_xx + k_zz}$ must interact nonlinearly  with another
+wave to knock out the $k_xx$ contribution. One possible candidate would have a
+$\cos \p*{-k_xx + k_zz}$ such that their product $\cos 2k_zz$, in which case
+$\kappa = 2k_z$. This also seems to be reasonable from a few of our simulations,
+but this requires thus that $2k_z$ be much longer than the viscous length scale;
+this would explain why we don't see it on our low-resolution simulations at $z
+\in [0, 12H], N_Z = 512$ but only on $N_Z = 1024$. The desired $\cos \p*{-k_xx +
+k_zz}$ is just a reflected wave in the $x$ direction, so I guess it can come
+from the viscous subscales pushing against the horizontal momentum/mean flow?
+In any case it's not inconceivable it would exist, and it should be the dominant
+$-k_xx$ term compared e.g.\ to $-k_xx + 0z, -k_xx + 2k_zz$.
+
+TODO We need simulations next of the front breaking down, to understand where it
+would stop, since we don't have a criterion for it yet.
+
 \chapter{Cylindrical Stratified Flow}

### 2018-10-30 writeups/notes.tex
huh

--- 2018-10-22 writeups/notes.tex
@@
 
+\section{Hydraulic Jumps in Surface Waves}
+
+We originally got started on this since it seems like a dimensionally-restricted
+version of what we're seeing in simulations, but presently unclear whether
+really the case. Also, Faber's book for fluids \emph{Fluid Dynamics for
+Physicists} cites a critical breaking jump height of $1.3$ which is markedly
+false; papers such as Chanson's 2009 hydraulic jumps review in Euro.\ J.\ Mech\
+B/Fluids shows that undular jumps (instead of turbulent jumps) can be found
+easily up to $3.5/4$.
+
+\subsection{1D Hydraulic Jump}
+
+This is one of the OG papers; B\'elanger solved this as early as in 1838 but
+Rayleigh's work in 1916 is also pretty famous. Consider the 1D nonlinear Euler
+equations for a body of water whose surface can move $h(x)$, then give it
+velocity $u(x, z)$. The problem is really 2D, but we can integrate along the $z$
+direction, in which case $\pd{P(x, z)}{x} = \frac{\rho g h(x, z)}{x}$ simply
+integrates to $\frac{\rho g h(x)^2}{2} = \frac{g\Sigma(x)^2}{2\rho}$. Assuming
+also the velocity is uniform $u(x, z) = u(x)$ gives equations of motion
+\begin{align}
+    \pd{\Sigma}{t} + \pd{(\Sigma u)}{x} &= 0,\\
+    \pd{u}{t} + u\pd{u}{x} &=
+        -\frac{1}{\Sigma}\pd{}{x}\p*{\frac{g\Sigma^2}{2\rho}},\nonumber\\
+        &= -\frac{g}{\rho}\pd{\Sigma}{x}.
+\end{align}
+It proves easier to divide the density equation by $\rho$ so that we can replace
+$\frac{\Sigma}{\rho} = h(x)$ in the above. A specific momentum conservation
+equation can be obtained
+\begin{equation}
+    \pd{(hu)}{t} + \pd{}{x}\p*{hu^2 + \frac{gh^2}{2}} = 0.
+\end{equation}
+
+Now, consider a hydraulic jump $u_1, h_1 \to u_2, h_2$ such that the
+discontinuity is fixed (we've shifted to the jump front frame), so that
+$\s*{uh}, \s*{hu^2 + gh^2/2}$ bracketed quantities are continuous across the jump.
+This results in
+\begin{align}
+    h_1u_1^2 - h_2u_2^2 = h_1u_1^2\p*{1 - \frac{u_2}{u_1}}
+        &= h_1u_1^2\p*{1 - \frac{h_1}{h_2}},\nonumber\\
+        &= \frac{g\p*{h_2^2 - h_1^2}}{2},\nonumber\\
+    u_1^2 &= \frac{h_2}{h_1}\frac{g\p*{h_1 + h_2}}{2}.
+\end{align}
+Customarily one defines $a_1^2 = gh_1, \mathrm{Fr} =
+\frac{u_1}{a_1}$ and $\gamma = \frac{h_2}{h_1}$ which gives simplification
+\begin{align}
+    2\mathrm{Fr}^2 &= \gamma(1 + \gamma)\nonumber\\
+    \gamma = \frac{-1 + \sqrt{1 + 8\mathrm{Fr}^2}}{2}.
+\end{align}
+Note now that $\mathrm{Fr}^2 > 1$ produces a $\gamma > 1$, so $h_2 > h_1$.
+Noting moreover that $\frac{u_1}{u_2} = \gamma$ and so $u_1 > u_2$. If we have a
+picture of a jet incident on a medium at rest, $u_2$ is just the velocity of the
+jump.
+
+\subsection{1D Viscous Theory}
+
+We follow Bohr's paper from J.\ Fluid Mech.\ in 1993 (also done by Tani in 1949
+to some extent) but in 1D first so that the equations are a bit easier to
+intuit. It's completely unnecessary but illustrates some features of the
+approach.
+
+Consider the same 2D equations as before (before integrating along $z$), but now
+permit a viscosity term. In the boundary layer approximation, the usual
+assumptions are that flow largely varies horizontally but variations are
+stronger vertically (i.e.\ we only consider the $x$ velocity equation but only
+consider the $z$ viscosity). We make one small modification, it seems that we
+permit an incompressibility assumption since $\rd{h(x, z, t)}{t} = 0$; this is
+just $\Delta P = \rho g h = $ the Lagrangian pressure boundary condition. Thus,
+the equations of motion are
+\begin{align}
+    \pd{u_x}{x} + \pd{u_z}{z} &= 0,\\
+    u_x \pd{u_x}{x} + u_z \pd{u_x}{z} &= -g\pd{h}{x} + \nu \ptd{u_x}{z}.
+\end{align}
+Integration along $z$ of the momentum equation can be done with appropriate BCs:
+no-slip/penetration at the bottom $u_x(z = 0) = u_z(z = 0) = 0$, while no shear
+at the surface $\pd{u_x}{z}(z = h) = 0$. We handle the integral of $u_z u_{x,
+z}$ by integrating by parts, dropping the boundary term and continuity equation,
+giving
+\begin{equation}
+    2u_x \int\limits_0^h u_x\rd{u_x}{x}\;\mathrm{d}z
+        = -g\pd{h}{x}h - \nu\at{\pd{u_x}{z}}_{z = 0}.
+\end{equation}
+We make a few handwavy substitutions for representative values $\partial_z u_x(z
+= 0) \sim u_x / h$ and $\partial_x u_x^2 \sim \partial_x \bar{u}_x^2$ average
+value. This averaging procedure and dropping the factor of $2$ (the two papers
+seem to disagree on this coefficient, I think Bohr's paper has an error) lets
+us examine
+\begin{equation}
+    \bar{u}_x \pd{\bar{u}_x}{x} + g\pd{h}{x} \sim -\frac{\bar{u}_x}{h^2}.
+\end{equation}
+Next, let's identify $\bar{u}_xh \equiv q$ volume flux constant in $x$, which
+leaves us with
+\begin{equation}
+    \pd{\bar{u}_x}{x}\p*{\bar{u}_x - \frac{gq}{\bar{u}_x^2}}
+        \sim -\nu\frac{\bar{u}_x^3}{q^2}.
+\end{equation}
+Let's study in terms of $\mathrm{Fr}^2 = \frac{\bar{u}_x^2}{gh} =
+\frac{\bar{u}_x^3}{gq}$ again, albeit a bit messier, which takes a bit of work
+but eventually lets us rearrange
+\begin{equation}
+    \pd{\mathrm{Fr}}{x} = \frac{\p*{gq}^{1/3}\nu}{q^2}\frac{
+        \mathrm{Fr}^{5/3}}{1 - \mathrm{Fr}^2}.
+\end{equation}
+Note that the units work here, as $\frac{\p*{gq}^{1/3}\nu}{q^2} \equiv L_\nu$
+has units of length.
+
+Let's interpret the equation we have now: we have exactly how the inbound
+$\mathrm{Fr}\p*{x = -\infty}$ evolves over $x$, which should describe (extremely
+coarsely, owing to our approximations) the boundary layer of the jump.
+
+Qualitatively, we can use a parameterization $\mathrm{Fr}(s), x(s)$ such that
+$\dot{\mathrm{Fr}}(s) = -\mathrm{Fr}^{5/3}, \dot{x}(s) = L_\nu \p*{\mathrm{Fr}^2
+- 1}$ where we define $L_\nu \equiv a_0h^2/\nu$ and we've absorbed the negative
+sign into $\dot{\mathrm{Fr}}(s)$. I'm too lazy to plot this, but if we start at
+a point $\mathrm{Fr} > 1$, it's clear that $\mathrm{Fr}$ increases and $x$
+decreases until $\mathrm{Fr} = 1$. Since the sign convention on the parametric
+is arbitrary, we can flip the sign and see that the parameterized curve formally
+follows decreasing $\dot{\mathrm{Fr}}(s) \propto -1/s$ while $\dot{x}(s) \propto
+-L$. This is double-valued with the solution $\mathrm{Fr} > 1$ though, so with
+some physical intuition we discard this and consider what values of
+$\mathrm{Fr}$ can take on beyond its intersection with $\mathrm{Fr} = 1$.
+
+It must discontinuously jump to another family of curves. But any nonzero value
+of $\mathrm{Fr}$ will live on another one of these such trajectories and
+increase without bound; indeed the only reasonable asymptotic solution is
+$\mathrm{Fr} = 0$. This completes the picture: for an inbound $\mathrm{Fr} \gg
+1$, it follows roughly $\pd{\mathrm{Fr}}{x} \gtrsim -\mathrm{Fr}^{-1/3}$ until
+it reaches $\mathrm{Fr} = 1$, then it discontinously disappears. This is the
+upstream flow, and so by mass conservation the location of the discontinuity
+must travel.
+
+This is exactly the picture we derived above: for an incoming $\mathrm{Fr}
+\gtrsim 1$, the upstream flow has zero velocity and the jump travels.
+
+\subsection{2D Hydraulic Jump}
+
+This is the case worked out in detail in Bohr's paper. Following much the same
+approach but in cylindrical coordinates we may write down equations of motion
+\begin{align}
+    u_r \pd{u_r}{r} + u_z \pd{u_r}{z} &= -g\pd{h}{r} + \nu
+        \ptd{u_r}{z},\nonumber\\
+    \frac{1}{r}\pd{(ru_r)}{r} + \pd{u_z}{z} &= 0.
+\end{align}
+Note the slight difference here due to the $\vec{\nabla} \cdot$ in cylindrical
+coordinates. Interestingly, as a consequence of this $\pd{P_r}{t} +
+\frac{1}{r}\partial_r \p*{hu_r^2 + gh^2/2} \neq 0$, since one term is a
+$\vec{\nabla} \cdot (hu_r)$ and the other is a $\vec{\nabla}gh^2/2$! Radial
+momentum is not conserved as the flow spreads out. Averaging using the same
+techniques as above gives
+\begin{align}
+    \bar{u}_r\pd{\bar{u}_r}{r} + g\pd{h}{r} &= -\nu \frac{\bar{u}_r}{h^2},\nonumber\\
+    \bar{u}_rhr &= q.
+\end{align}
+Combining the two gives
+\begin{align}
+    \pd{\bar{u}_r}{r}\p*{\bar{u}_r - \frac{gq}{\bar{u}_r^2r}}
+        &= \frac{gq}{\bar{u}_rr^2} - \nu\frac{\bar{u}_r^3r^2}{q^2}.
+        \label{eq:2d_hjump}
+\end{align}
+It's not very worth converting into dimensionless $\mathrm{Fr}$ number since it
+would carry an $r$ dependence (if we are to hold $q$ constant) and we need to
+study the $r$ dependence.
+
+It may be observed that \autoref{eq:2d_hjump} has only a single fixed point,
+$u_r^3r = gq, u_r^4r^4 = gq^3/\nu$. Call this point $r_c$ the critical radius,
+then it turns out that the two parametric solutions (as we analyzed in the
+previous section) both spiral around this point. This has a few important
+properties:
+\begin{itemize}
+    \item No asymptotic solutions: the solution for $r > r_c$ always diverges
+        at some finite $r_s < \infty$. We can try an approximate solution in the
+        form $u_r = y / r$; this is done in Bohr's paper.
+
+    \item For a fixed $r_s \gg r_c, < \infty$, we can require the usual jump
+        condition which takes the slightly modified form
+\end{itemize}
+
+Edit: Looks like Bhagat 2018 J.\ Fluid.\ Mech.\ does this more correctly, they
+find that surface tension is the proper term to invoke after performing this
+boundary layer analysis. I'm not going to finish reworking the rest of Bohr's
+paper here, since it's up through the boundary layer equations that is reusable.
 \end{document}

### 2018-11-02 writeups/notes.tex
check in new notes + higher k sims

@@
     2\mathrm{Fr}^2 &= \gamma(1 + \gamma)\nonumber\\
-    \gamma = \frac{-1 + \sqrt{1 + 8\mathrm{Fr}^2}}{2}.
+    \gamma &= \frac{-1 + \sqrt{1 + 8\mathrm{Fr}^2}}{2}.
 \end{align}
@@
 
-This is the case worked out in detail in Bohr's paper. Following much the same
-approach but in cylindrical coordinates we may write down equations of motion
+This is the case worked out in detail in Bohr's paper. First, perhaps the
+easiest thing to do is to enforce jump conditions $[rhu], [hu^2 + gh^2/2]$
+subject to $q = 2\pi rh_iu_{ri}$ conserved total flow. This gives us
+\begin{align}
+    h_1u_{r1}^2 - h_2u_{r2}^2 = \frac{q}{2\pi r}(u_{r1} - u_{r2})
+        &= \frac{gh_1^2}{2} - \frac{gh_2^2}{2},\\
+        &= \frac{gq^2}{8\pi r^2u_{r1}^2} - \frac{gq^2}{8\pi r^2u_{r2}^2}
+            ,\nonumber\\
+    2u_{r1}^2 u_{r2}^2 &= \frac{gq}{2\pi r}\p*{u_{r2} + u_{r1}},\nonumber\\
+    u_{r1} &= \frac{gq}{8\pi u_{r2}^2r} \p*{1 +
+        \sqrt{1 + 16\frac{\pi u_{r2}^3r}{gq}}},\\
+    \frac{u_{r1}}{u_{r2}} &= \frac{1}{4\mathrm{Fr}^2}\p*{1 +
+        \sqrt{1 + 8\mathrm{Fr}^2}}.
+\end{align}
+We've defined $\mathrm{Fr}^2 \equiv \frac{2\pi r u_r^3}{gq} = \frac{u_r^2}{gh}$.
+Note that this is in agreement with $\gamma = \frac{h_1}{h_2} = \frac{u_2}{u_1}$
+from the plane parallel theory. The major difference is that now $\mathrm{Fr}^2
+= \mathrm{Fr}^2(r)$, and so the location of the hydraulic jump can be pinned.
+One might expect that once $\mathrm{Fr}(r) > 1$ that the jump occurs, but this
+turns out not to be the case.
+
+We now perform the boundary layer analysis. Following much the same approach but
+in cylindrical coordinates we may write down equations of motion
 \begin{align}

### 2018-11-02 sims/2d_1_gradual/strat_helper.py

@@ -124,8 +124,8 @@ def run_strat_sim(set_ICs, name, params):
     logger.info('Starting sim...')
     while solver.ok:
         cfl_dt = cfl.compute_dt() if params.get('USE_CFL') else params['DT']
-        if cfl_dt < params['DT'] / 4: # small step sizes if strongly cfl limited
-            cfl_dt /= 4
+        if cfl_dt < params['DT'] / 8: # small step sizes if strongly cfl limited
+            cfl_dt /= 2
         solver.step(cfl_dt)
         curr_iter = solver.iteration
 

### 2018-11-05 presentations/Su_DL_110618.tex
stash Daniel Visit ppt

--- 2018-10-30 writeups/notes.tex
@@
     \newcommand*{\ang}[0]{\;\text{\AA}}
-    \newcommand*{\mum}[0]{\;\upmu \mathrm{m}}
+    \newcommand*{\mum}[0]{\\;upmu \mathrm{m}}
     \newcommand*{\at}[1]{\left.#1\right|}
-
-    \newtheorem{theorem}{Theorem}[section]
 
@@
 
-\def\Snospace~{\S{}} % hack to remove the space left after autorefs
-\renewcommand*{\sectionautorefname}{\Snospace}
-\renewcommand*{\appendixautorefname}{\Snospace}
-\renewcommand*{\figureautorefname}{Fig.}
-\renewcommand*{\equationautorefname}{Eq.}
-\renewcommand*{\tableautorefname}{Tab.}
-
-\onehalfspacing
-
-\pagestyle{fancy}
-\rfoot{Yubo Su}
-\cfoot{\thepage/\pageref{LastPage}}
-
-\title{Research Notes}
+\title{Quick Research Review}
 \author{Yubo Su}
-\rhead{}
+\date{Nov. 06, 2018}
 
@@
 
-\tableofcontents
+\begin{frame}
+    \frametitle{Objectives}
 
-\newpage
+    \begin{itemize}
+        \item Continuous train of IGW excited in stratified WD atmosphere by
+            binary companion, grows due to stratification, reaches nonlinear
+            amplitudes $\xi_z k_z \gtrsim 1$ before the peak of the envelope.
 
-\chapter{Preliminary Problems}
+        \item Seek long-term behavior of nonlinear dissipation, whether
+            steady-state or periodic.
+    \end{itemize}
+\end{frame}
 
-To get an intuition for how Dedalus and fluid mechanics works, we will solve
-some toy problems. Recall fluid equations in the presence of a uniform
-gravitational field $\vec{g} = -g\hat{z}$:
-\begin{equation}
-    \begin{split}
-        \pd{\rho}{t} + \vec{\nabla} \cdot \p*{\rho \vec{u}} &= 0,\\
-        \rd{\vec{u}}{t} + \frac{\vec{\nabla}P}{\rho} - \vec{g} &= 0.
-    \end{split}\label{eq:fluid_eq}
-\end{equation}
+\begin{frame}
+    \frametitle{Numerical Simulations, No Shear}
+    \framesubtitle{Setup}
 
-In the incompressible limit, $\rd{\rho}{t} = 0$, which implies $\vec{\nabla}
-\cdot \vec{u} = 0$. We use subscripts to indicate perturbed quantities, $Q_0$ is
-background and $Q_1$ is perturbed. We will generally use $\vec{u}_0 = 0$ unless
-otherwise noted. We will also generally assume symmetry along all axes except
-$z$ the vertical axis.
+    \begin{itemize}
+        \item 2D, Fourier $\times$ Chebyshev, damping zones, ran at both $128
+            \times 512$ and $256 \times 1024$ over $3H \times 12H$ grid, where
+            $H$ is scale height.
 
-In the incompressible limit, the fluid equations become
-\begin{equation}
-    \begin{split}
-        \vec{\nabla} \cdot \vec{u}_1 &= 0,\\
-        \pd{\rho_1}{t} + u_{1z}\pd{\rho_0}{z} &= 0,\\
-        \pd{\vec{u}_1}{t} + \frac{1}{\rho_0}\vec{\nabla}P_1
-            + \frac{\rho_1 g \hat{z}}{\rho_0} &= 0.
-    \end{split}\label{eq:lin.incomp}
-\end{equation}
-We have used $\vec{\nabla}P_0 = -\rho_0 g \hat{z}$ in the absence of
-perturbations.
+        \item Target waves $k_x = \frac{2\pi}{3H}, k_z = \frac{2\pi}{H}, \omega
+            \approx 0.32N$, $N$ B-V frequency.
 
-\section{Incompressible, No Gravity}
+        \item Wave is excited by volumetric forcing
+            $e^{-z^2/2\sigma^2}\cos\p*{k_xx - \omega t}$ with $\sigma =
+            \frac{3H}{32} \lesssim \frac{1}{k_z}$.
 
-We note that in the no gravity limit that $\rho_1$ does not have an effect on
-other dynamical variables, so the equations of motion we must solve are
-\begin{equation}
-    \begin{split}
-        \vec{\nabla} \cdot \vec{u}_1 &= 0,\\
-        \pd{\vec{u}_1}{t} + \frac{\vec{\nabla}P_1}{\rho_0} &= 0.
-    \end{split}\label{eq:lin.no_g_eom}
-\end{equation}
-We can take the divergence of the momentum equation and substitute the
-continuity equation to get $\nabla^2 P = 0$.
+        \item Regularized by numerical viscosity $\nu \gtrsim
+            \frac{\omega}{k_{\max}k_z}$.
+    \end{itemize}
+\end{frame}
 
-\subsection{Dirichlet BCs}
+\begin{frame}
+    \frametitle{Numerical Simulations, No Shear}
+    \framesubtitle{Results}
 
-This is a Laplace equation, which we've solved countless times. Imposing
-periodic boundary conditions in the $x$ direction and $P_1(z = L) = 0, P_1(z =
-0) = \mathcal{P}(x, t)$, we obtain eigenfunctions
-\begin{equation}
-    \begin{split}
-        P_{1, n}(x, z, t) &=
-            \frac{\mathcal{P}_n(t)}{\sinh(k_nL)}e^{ik_nx}\sinh\p*{k_n(L - z)},\\
-        u_{1x, n}(x, z, t) &=
-            \int\limits^t -\frac{1}{\rho_0}\pd{P_{1, n}}{x}\;\mathrm{d}t,\\
-        u_{1z, n}(x, z, t) &=
-            \int\limits^t -\frac{1}{\rho_0}\pd{P_{1, n}}{z}\;\mathrm{d}t.
-    \end{split}\label{eq:incomp_nog}
-\end{equation}
-We define $k_n = \frac{2\pi n}{L}, n \geq 0$ and $\mathcal{P}(x, t) =
-\sum\limits_n \mathcal{P}_n(t)e^{ik_nx}$.
+    \begin{itemize}
+        \item High resolution ($\nu = 4\frac{\omega}{k_{\max}k_z}$):
+            \lstinline{nl3.mp4}. Note that mean flow starts to build up at the
+            outgoing damping zone, and once it hits criticality it starts to
+            move down while continuing to steepen. Eventually, it stops moving
+            and produces reflection below, while leaving behind a sinusoidal
+            mean flow above.
 
-Thus, if we impose BCs $\mathcal{P}(x, t) = \sin \frac{2\pi x}{L}$ and start
-with initial conditions such that all quantities are zero, we would expect after
-transients die out that
-\begin{equation}
-    \begin{split}
-        P(x, z, t) &= \frac{\sin \frac{2\pi x}{L}}{\sinh 2\pi}
-            \sinh \p*{2\pi \frac{L - z}{L}},\\
-        u_{1x}(x, z, t) &= -\frac{2\pi t}{L\rho_0}
-            \frac{\cos \frac{2\pi x}{L}}{\sinh 2\pi}\sinh \p*{2\pi \frac{L -
-                z}{L}},\\
-        u_{1z}(x, z, t) &= +\frac{2\pi t}{L\rho_0}
-            \frac{\sin \frac{2\pi z}{L}}{\sinh 2\pi}\cosh \p*{2\pi \frac{L -
-                z}{L}}.
-    \end{split}\label{eq:incomp_nog_sol}
-\end{equation}
-This is in good agreement with the results, presented in \autoref{fig:no_g}.
-Note that $P$ is constant while $\vec{u}$ increases linearly in time, and we
-observe the expected $\sim \sin x \sinh \frac{L - z}{z}$ dependence. In fact,
-$u_{1x}, u_{1z}$ are exactly $\frac{2\pi}{10}$ at $t = 1$.
-\begin{figure}[!h]
-    \centering
-    \begin{subfigure}{0.3\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{../sims_old/2d_0_no_g/no_g_P_t50.png}
-    \end{subfigure}
-    \begin{subfigure}{0.3\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{../sims_old/2d_0_no_g/no_g_ux_t50.png}
-    \end{subfigure}
-    \begin{subfigure}{0.3\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{../sims_old/2d_0_no_g/no_g_uz_t50.png}
-    \end{subfigure}
+        \item Low resolution ($\nu = 3\frac{\omega}{k_{\max}k_z}$):
+            \lstinline{nl3_lowres.mp4}. Qualitatively similar, but the mean flow
+            forms later.
 
-    \begin{subfigure}{0.3\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{../sims_old/2d_0_no_g/no_g_P_t100.png}
-    \end{subfigure}
-    \begin{subfigure}{0.3\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{../sims_old/2d_0_no_g/no_g_ux_t100.png}
-    \end{subfigure}
-    \begin{subfigure}{0.3\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{../sims_old/2d_0_no_g/no_g_uz_t100.png}
-    \end{subfigure}
-    \caption{$P, u_x, u_z$ at $t = 0.5$ and $t = 1$ for $\rho_0 = 1$. We choose
-    a $L = 10$ square domain.}\label{fig:no_g}
-\end{figure}
+        \item Two hypotheses:
+            \begin{itemize}
+                \item Damping layer is causing mean flow to begin building up
+                    like Daniel said; how can we fix? Seems like low $k_z$ modes
+                    (which have higher $v_{ph}$).
 
-\subsection{Things to Note}
+                \item Further growth/propagation is critical layer absorption?
+                    Eventual reflection might build up when shear flow is too
+                    steep $\mathrm{Ri} \gtrsim 1/2$ or WKB breaks down
+                    $\pd{k_z}{z} \sim \frac{k_z}{\lambda_z}$.
+            \end{itemize}
+    \end{itemize}
+\end{frame}
 
-It is worth noting that, since our \autoref{eq:lin.no_g_eom} reduced to a Laplace
-equation, we needed two $z$ BCs and two $x$ BCs (periodic BCs amount to equating
-the value and derivative of the function). This is in agreement with the
-observation that the original \autoref{eq:lin.no_g_eom} had two derivatives in
-$x, z$ apiece, so we needed two BCs each.
+\begin{frame}
+    \frametitle{Numerical Simulations, Shear Flow}
+    \framesubtitle{Setup}
 
-It is also worth seeing from our solution that $P$ immediately goes to the
-equilibrium solution. This is not surprising since in the incompressible limit,
-sound speed goes to infinity which is the timescale on which the pressure field
-adjusts to net forces. Thus, the dynamics solely arise from the static pressure
-field pushing the velocity field to equilibrium.
+    \begin{itemize}
+        \item 2D, Fourier $\times$ Fourier, damping zones, running at $64 \times
+            256$. No height stratification, domain is $ H \times H$ (though $H$
+            is now physically meaningless).
 
-We chose time-independent $\mathcal{P}(x, t)$, but it is clear that whatever
-$\mathcal{P}(x, t)$ we choose, the time dependence propagates to the velocities
-by way of an integral. If we had instead chosen to take Fourier transform $t \to
-\omega$, we would have had to integrate the boundary condition against the
-eigenfunctions for each of the $\omega$, which is still easily computable, to
-get the full $u_{1x}(x, z, t)$. We consider this in preparation for when we can
-only solve for a set of $\vec{k}, \omega$ in the next problem.
+        \item Target waves $k_x = \frac{2\pi}{H}, k_z = \frac{20\pi}{H}, \omega
+            \approx 0.1N$, $N$ B-V frequency. Volumetric forcing.
 
-The next thing we would have wanted to do is solve a problem with radiative BCs,
-but we need to have wave solutions, which are missing in the absence of gravity.
-We thus move on to the next configuration.
+        \item Regularized by $\nabla^6$ hyperviscosity.
+    \end{itemize}
+\end{frame}
 
-\section{Incompressible, Stratified w/ Gravity}
+\begin{frame}
+    \frametitle{Numerical Simulations, Shear Flow}
+    \framesubtitle{Results}
 
-\subsection{Eigenfunctions}
+    \begin{itemize}
+        \item Initialize w/ shear flow and let evolve. Try both narrow and wide
+            profiles (\lstinline{vstrat*.mp4}).
 
-Let's restore the $\rho_1 g$ term now. For funsies, we begin by solving for
-arbitrary stratification $\rho_0(z)$ first. The fluid equations to first order
-reduce to
-\begin{equation}
-    \begin{split}
-        \pd{\rho_1}{t} + \vec{u}_1 \cdot \p*{\vec{\nabla}\rho_0} &= 0,\\
-        \vec{\nabla} \cdot \vec{u}_1 &= 0,\\
-        \pd{\vec{u}_1}{t} &= -\frac{\vec{\nabla}P_1}{\rho_0}
-            - \frac{\rho_1 g}{\rho_0}
-    \end{split}\label{eq:lin_incomp}
-\end{equation}
-We expect there to be some $z$ dependence in the amplitude, so we substitute
-variables of form $e^{i(kx - \omega t)}$ and do not specify the $z$ dependence.
-This gives us
-\begin{equation}
-    \begin{split}
-        -i\omega \rho_1 - u_{1z}\pd{\rho_0}{z} &= 0,\\
-        iku_{1x} + \pd{u_{1z}}{z} &= 0,\\
-        -iw u_{1x} + \frac{ik_x P_1}{\rho_0} &= 0,\\
-        -iw u_{1z} + \frac{1}{\rho_0}\pd{P_1}{z} +
-            \frac{\rho_1 g}{\rho_g} &= 0.
-    \end{split}
-\end{equation}
-We substitute $N^2 = -\frac{g}{\rho_0}\pd{\rho_0}{z}$ to obtain
-\begin{subequations}
+        \item Conclusion: attenuated transmission when thin, barely-critical
+            layer, smooth WKB-like wavelength shortening when broad, critical
+            flow, but steep $\pd{U_0}{z}$ gives reflection!
+
+        \item Recall Booker \& Bretherton result
             \begin{align}
-        -i\omega \rho_1 - u_{1z}\frac{\rho_0N^2}{g} &= 0,\label{eq:lin.cont}\\
-        iku_{1x} + \pd{u_{1z}}{z} &= 0,\label{eq:lin.div}\\
-        -iw u_{1x} + \frac{ik_x P_1}{\rho_0} &= 0,\label{eq:lin.mom_x}\\
-        -iw u_{1z} + \frac{1}{\rho_0}\pd{P_1}{z} +
-            \frac{\rho_1g}{\rho_0} &= 0.\label{eq:lin.mom_z}
+                T &\sim \exp\s*{-\pi \sqrt{\frac{N^2}{(U_0')^2} - \frac{1}{4}}}
+                & R &\sim
+                    \exp\s*{-2\pi \sqrt{\frac{N^2}{(U_0')^2} - \frac{1}{4}}}.
             \end{align}
-\end{subequations}
-Eliminating $u_{1x}$ by substituting~\eqref{eq:lin.div}
-into~\eqref{eq:lin.mom_x} and $\rho_1$ by substituting~\eqref{eq:lin.cont}
-into~\eqref{eq:lin.mom_z} give
-\begin{subequations}
-    \begin{align}
-        i\omega \pd{u_{1z}}{z} + \frac{k_x^2 P_1}{\rho_0} &= 0,
-            \label{eq:lin.eq1}\\
-        \p*{\omega^2 - N^2 }u_{1z} + \frac{i\omega}{\rho_0}\pd{P_1}{z} &= 0.
-            \label{eq:lin.eq2}
-    \end{align}
-\end{subequations}
-Finally, we multiply~\eqref{eq:lin.eq1} with $\rho_0$ and differentiate
-$\mathrm{d}z$ and combine with~\eqref{eq:lin.eq2} to give
-\begin{equation}
-    \rtd{u_{1z}}{z} + \frac{1}{\rho_0}\pd{\rho_0}{z}\pd{u_{1z}}{z}
-        + k_x^2\p*{\frac{N^2}{\omega^2} - 1}u_{1z} = 0.\label{eq:lin.incomp.gen}
-\end{equation}
+            WKB criterion $\Rightarrow \frac{k_z^2}{N}U_0' \frac{\lambda^2}{2}
+            \ll 1$.
 
-Let's now pick stratification $\rho \propto e^{-z/H}$
-\autoref{eq:lin.incomp.gen} clearly has exponential solutions $e^{\kappa z}$ for
-\begin{equation}
-    \kappa^2 - \frac{\kappa}{H} + k_x^2\p*{\frac{N^2}{\omega^2} - 1} = 0.
-\end{equation}
-We permit complex $\kappa = \frac{1}{2H} + ik_z$, and from the above clearly
-\begin{align}
-    k_z^2 &= -\frac{1}{4H^2} + k_x^2\p*{\frac{N^2}{\omega^2} - 1},\nonumber\\
-    \omega^2 &= \frac{N^2k_x^2}{k_x^2 + k_z^2 +
-        \frac{1}{4H^2}}.\label{eq:strat_disp}
-\end{align}
-Thus the eigenfunctions are
-\begin{equation}
-    \begin{split}
-        u_{1z} &= e^{z/2H} e^{i(k_zz + k_xx - \omega t)},\\
-        u_{1x} &= -\frac{k_z + i/2H}{k_x} u_{1z},\\
-        \rho_1 &= \frac{i \rho_0}{H\omega} u_{1z},\\
-        P_1 &= -\frac{\rho_0 \omega}{k_x^2}\p*{k_z + i/2H}u_{1z}.
-    \end{split}\label{eq:strat_eigens}
-\end{equation}
+        \item Winters d'Asaro had a local $\mathrm{Ri} \sim 0.5$.
+    \end{itemize}
+\end{frame}
 
-\subsection{Analytically Solving an IVP, Dirichlet + Driving BCs}
-
-We will analyze everything in terms of $u_{1z}$ since it has the simplest form;
-note that when actually choosing the BCs we will have to consider the gauge
-freedom of $P$ and some considerations we defer to the computational section.
-
-Currently, we have a set of eigenfunctions
-\begin{equation}
-    u_{1z}\p*{x, z, t \middle| \vec{k}, \omega}
-        = e^{z/2H} e^{i(k_zz + k_xx - \omega t)}.
-        \label{eq:strat.uz_eigen}
-\end{equation}
-Note that $u_{1z}$ is really only a function of two parameters rather than the
-three $(k_x, k_z, \omega)$, since the three are related by dispersion relation
-\autoref{eq:strat_disp}.
-
-Now, we implement BCs. Consider domain $x, z \in [0, L]$. We will use periodic
-BCs again in $x$, so then $k_{x, n} = \frac{2\pi n}{L}, n \geq 0$. Then, we will
-require $u_{1z, n}(x, L, t) = 0$, a Dirichlet condition at the top boundary,
-which restricts us to eigenfunctions of form
-\begin{equation}
-    u_{1z, n}\p*{x, z, t \middle| \vec{k}, \omega}
-        = e^{z/2H}e^{i(k_xx - \omega t)}\sin\p*{k_z\p*{L - z}}.
-        \label{eq:strat.uz_eigen_l-z}
-\end{equation}
-
-Finally, we must choose a BC at $z = 0$. We will choose a general function
-$u_{1z}(x, 0, t) = F(x, t)$ where we can decompose
-\begin{equation}
-    F(x, t) = \int\limits \sum\limits_n \mathcal{F}(k_{x, n}, \omega)
-        e^{i(k_{x, n}x - \omega t)}\;\mathrm{d}\omega.
-\end{equation}
-Matching BCs then gives us general solution for $u_{1z}$ given an arbitrary
-driving function
-\begin{equation}
-    u_{1z}\p*{x, z, t \middle| \vec{k}, \omega}
-        = \int\limits \sum\limits_n \mathcal{F}\p*{k_{x, n}, \omega}
-            \frac{e^{z/2H}e^{i\p*{k_{x, n}x - \omega t}}\sin \p*{k_z(L - z)}}{
-            \sin k_zL}\;\mathrm{d}\omega.
-            \label{eq:strat.full_sol}
-\end{equation}
-
-For ease of computation, let's pick $F(x, t) = \cos \p*{\frac{2 \pi x}{L} -
-\omega_0 t}$, so our full expected solution is (note $A + \epsilon =
-Ae^{\epsilon/A} + \mathcal{O}(\epsilon^2)$)
-\begin{equation}
-    \begin{split}
-        u_{1z}\p*{x, z, t \middle | \vec{k}, \omega}
-            &= e^{z/2H}\frac{\cos\p*{\frac{2\pi x}{L} - \omega_0 t}
-                \sin \p*{k_z(L - z)}}{\sin k_zL},\\
-        u_{1x}\p*{x, z, t \middle | \vec{k}, \omega}
-            &\approx \frac{k_z}{k_x} e^{z/2H}\frac{
-                \cos\p*{\frac{2\pi x}{L} - \omega_0 t + \frac{1}{2Hk_z}}
-                \sin \p*{k_z(L - z)}
-            }{\sin k_zL},\\
-        \rho_1\p*{x, z, t \middle | \vec{k}, \omega}
-            &\approx \frac{\rho_0}{H\omega} e^{z/2H}\frac{
-                -\sin\p*{\frac{2\pi x}{L} - \omega_0 t}
-                \sin \p*{k_z(L - z)}
-            }{\sin k_zL},\\
-        P_1\p*{x, z, t \middle | \vec{k}, \omega}
-            &\approx -\frac{\rho_0 \omega k_z}{k_x^2} e^{z/2H}\frac{
-                \cos\p*{\frac{2\pi x}{L} - \omega_0 t + \frac{1}{2Hk_z}}
-                \sin \p*{k_z(L - z)}
-            }{\sin k_zL},\\
-    \end{split}\label{eq:strat.part_sol}
-\end{equation}
-where $k_z: \omega(k_x, k_z) = \omega_0$ by the dispersion relation
-\autoref{eq:strat_disp}. Note that $H$ contributes both to the overall
-exponential profile and to the phase lag of $u_{1x}$.
-
-\subsection{Computationally Solving an IVP, Dirichlet + Driving BCs}
-
-To solve this computationally with the aforementioned BCs, periodic in $x$,
-Dirichlet $0$ at $z = L$ and $\cos \p*{\frac{2\pi x}{L} - \omega_0 t}$ at $z =
-0$, we must address the gauge freedom in $P$. This arises because for $k_x = 0$,
-the divergence-free condition $\vec{\nabla} \cdot \vec{u}_1 = \pd{u_z}{z} = 0$
-specificies $u_{1z}$ up to a constant already, so the bottom BC will fix the
-value of $u_z$ when $k_x = 0$.
-
-A different way of phrasing the same argument is as follows. Consider the
-discrete $N \times N$ (square for notational simplicity) grid. At the boundary,
-there is a list of values $f\p*{\z*{x_i}, z_0}$ that lives in an $N$ dimensional
-space. We can thus pick a spanning set of $N$ basis vectors, and for each of
-these $N$ vectors, by enforcing $\vec{\nabla} \cdot \vec{u} = 0$ at the boundary
-we fix the allowed $f\p*{\z*{x_i}, z_{-1}}$ boundary conditions we can implement.
-But there exists a choice of basis vectors for which one of the basis vectors is
-constant $f_i\p*{\z*{x_i}, z_0} = C$. For this basis vector, the boundary
-condition is fully determined, so we only have $N - 1$ dimensions from which to
-choose the BCs $f\p*{\z*{x_i}, z_{-1}}$. Since the dimensionality of a space
-cannot depend on the choice of basis vector, the divergence free condition
-actually yields an extra degree of freedom.
-
-We can use this extra degreee of freedom to specify $P(z = L) = 0$ so the
-oscillations should have zero mean. Then we can just simulate away!
-
-From here we can just tweak the parameters until we get something useful, can
-look at the videos. Note that we should simulate until at least $T >
-\frac{2L_z}{v_{p, z}}$ the $z$ phase space velocity, so we can capture any
-reflections off the boundary. Turns out both Dirichlet and Neumann BCs give
-strong reflections that produce standing waves in the $z$ and traveling waves in
-the $x$.
-
-\subsection{Phase/Group Velocity}
-
-Let's figure out the analytical forms for the phase, group velocity and the
-energy density/power flux.
-
-Let's first consider the phase velocity. We traditionally think about the phase
-velocity in 1D $v_{ph} = \frac{\omega(k)}{k}$, but in general it is the function
-such that $\vec{k} \cdot \vec{r} - \omega t = \vec{k} \cdot \p*{\vec{r} -
-\vec{v}_{ph}t}$ is constant (the phase of the wave). Thus, it must satisfy
-$\vec{k} \cdot \vec{v}_{ph} = \omega$, and so one sensible choice is
-\begin{equation}
-    \vec{v}_{ph} = \frac{\omega \hat{k}}{\abs{\vec{k}}}
-        = \frac{\omega \vec{k}}{\abs{\vec{k}}^2}.
-        \label{eq:v_ph}
-\end{equation}
-
-For our stratified atmosphere problem, $\omega^2 = \frac{N^2 k_x^2}{k^2 +
-\frac{1}{4H^2}}$. This corresponds to $\vec{v}_{ph} = \frac{Nk_x}{\sqrt{k_x^2 +
-k_z^2 + 1/4H^2}} \vec{k}$.
-
-On the other hand the group velocity is given $\vec{v}_g =
-\pd{\omega}{k_i}\hat{\imath}$. In our problem, $v_{g, z} =
--\frac{Nk_xk_z}{\p*{k_x^2 + k_z^2 + 1/4H^2}^{3/2}}$ while $v_{g, x} =
-\frac{N}{\sqrt{k_x^2 + k_z^2 + 1/4H^2}} - \frac{Nk_x^2}{\p*{k_x^2 +
-k_z^2 + 1/4H^2}^{3/2}} = \frac{N\p*{k_z^2 + 1/4H^2}}{\p*{k_x^2 +
-k_z^2 + 1/4H^2}^{3/2}}$.
-
-\subsection{Energy/Power Flux}
-
-To compute the energy and power flux of the wave, we recall that for a general
-fluid the energy conservation equation reads
-\begin{equation}
-    \pd{}{t}\p*{\frac{1}{2}\rho v^2 + \rho \epsilon}
-        = \vec{\nabla} \cdot \p*{
-            \rho \vec{v}\p*{v^2 + \epsilon + \frac{P}{\rho}}
-        },
-\end{equation}
-where $\epsilon$ is the internal energy. But since $P = (\gamma -
-1)\rho\epsilon$ and we take $\gamma \to \infty$ incompressible limit, $\epsilon
-= 0$, so the energy of the wave is just $\frac{1}{2}\p*{\rho_0 + \rho_1} u_1^2$
-and the power flux is $\p*{\rho_0 + \rho_1}\vec{u}_1u_1^2 + P_1\vec{u}_1$.
-
-Another formula given by Sutherland 2011 is $\pd{\ev*{E}}{t} = v_{g, z}
-\ev*{E}$ where we average over $x$ one wavelength. I wasn't able to show
-that these two agree in general; Sutherland's formula seems to be derived for a
-traveling wavepacket whereas we have no such thing. We do not use this
-expression since the Landau \& Lifshitz expression works well.
-
-\section{Supressing Reflection}
-
-Sourced from \url{https://people.maths.ox.ac.uk/trefethen/6all.pdf}. There are
-largely two ways to do this. The first is to add a damping zone, a region in
-which $\dot{q} = -q/\tau_d(z)$ where $\tau_d(0) = 0$ and increases to some
-non-small number for some $z_0$ beyond which we want damping. One form, chosen
-in Ryan and Dong's paper, is to use multiplicative factor $f(z) = \max\s*{0, 1 -
-\frac{\p*{z - z_b}^2}{\p*{z_d - z_b}^2}}$ where $z_b$ is where one begins
-supression and $z_b$ is the boundary. Then a dynamical variable $q$ can be
-supressed via $\dot{q} = (\dots) - f(z)\frac{q - q_0}{\tau}$ where $\tau$ is the
-dynamical timescale on which to damp and $q_0$ is the value to damp to. We will
-apply this only to the velocity variables with an eye to extending this approach
-to the nonlinear regime, where the velocity variables should still be damped to
-zero but $\rho, P$ must capture the stratification and so cannot easily be
-artificially damped (maybe? Revisit if this does poorly compared to damping all
-linear variables, and damp the nonlinear variables to their
-equilibrium/stratified values).
-
-\subsection{First Order}
-
-In order to compute the reflecting boundary condition, we must find a
-sequence of differential operators that well approxmiates the system of PDEs. We
-can get this from the dispersion relation; since our radiating boundary is along
-$z$, we should try to solve for $k_z$ in the dispersion relation to get a
-pseudodifferential operator form for $\pd{}{z}u_z$ (we take $k_xH \ll 1$ as well
-to simplify):
-\begin{align}
-    k_z^2 &= \frac{N^2}{\omega^2}k_x^2 - k_x^2 - \frac{1}{4H^2},\nonumber\\
-        &\approx k_x^2\p*{\frac{N^2}{\omega^2} - 1},\\
-    \pd{u_z}{z} &\approx -\pd{u_z}{x}\sqrt{\frac{N^2}{\omega^2} - 1}.
-\end{align}
-We pick the negative sign in accordance with an outgoing wave, per the group
-velocity formulae for $k_z > k_x$.
-
-From the numerical simulations, it is clear that reflections are well supressed
-initially, but the the simulation blows up. This is because our supression is
-imperfect (differential approximation to a pseudodifferential operator) and
-since our system is dissipation free, reflection grows. Moreover, since any
-reflected components seem to have different $k_x, k_z, \omega$, the second time
-they're incident on the boundary, our above condition becomes wildly inaccurate.
-
-\section{The Difficulty of the Incompressible, Nonlinear Problem}
-
-We consider the problem where both advective terms are kept and $\rho_1 \sim
-\rho_0$. We write down thus nonlinear fluid equations
-\begin{equation}
-    \begin{split}
-        \pd{\rho}{t} + \vec{u} \cdot \p*{\vec{\nabla}\rho} &= 0,\\
-        \vec{\nabla} \cdot \vec{u} &= 0,\\
-        \pd{\vec{u}}{t} &= -\p*{\vec{u} \cdot \vec{\nabla}} \vec{u}
-            - \frac{\vec{\nabla}P}{\rho} - g\hat{z}.
-    \end{split}\label{eq:nonlin_incomp}
-\end{equation}
-Note that we cannot even subtract off hydrostatic equilibrium anymore since
-$\rho$ can deviate greatly from the normal $\rho_0e^{-z/H}$!
-
-\subsection{Rearrangement for Dedalus}
-
-In order to simulate this at all in a spectral program, recall the way that
-nonlinear PDEs are decomposed in a spectral code. Given a phase space $Q$,
-consider first a system of PDEs expressible in form
-\begin{equation}
-    \dot{Q} + \bm{L}Q + f(Q) = g
-\end{equation}
-where $\dot{Q}$ is the time derivative, $\bm{L}$ is some linear operator, $f(Q)$
-are nonlinear terms and $g$ is an arbitrary driving function. Spectral codes
-work by re-casting the operators to be purely algebraic in an appropriate
-spectral basis consisting of some $\varphi_i$ trial and $\xi_i$ test functions.
-The \emph{tau spectral method} that Dedalus uses considers using the trial
-functions as the test functions, so the effective decomposition that is
-performed is
-\begin{equation}
-    \bra*{\varphi_i}\dot{Q} + \bm{L}Q \ket*{\varphi_j}
-        = \dot{Q}_i + L_{ij}Q_q = \bra*{\varphi_i}-f(Q)  + g \ket*{\varphi_i}.
-\end{equation}
-It is then clear that $-f(Q) + g$ can just be treated as inhomogeneous source
-terms, and over a short $\Delta t$ it can be approximated via $-f(Q_0) + g$.
-This first-order inhomegeneous ODE then admits closed form exact solutions
-$e^{-L_{ij}t}\p*{\dots}$.
-
-One can imagine, for this matter, that $\pd{}{t}$ can be wrapped in the $L_{ij}$
-operator and the timestepping computed implicitly; I believe this is what
-Dedalus does, implicitly timestepping the left hand side and explicitly the
-inhomogeneous terms.
-
-Both of these procedures require $\bm{L}$ linear operator or $\pd{}{t} + \bm{L}$
-to be full rank, or invertible. This is enforced analytically by requiring the
-same number of BCs as derivatives, and numerically this is little different.
-This is why it is possible to have components of $Q$ that are not differentiated
-$\partial_t$ but still have well-defined time evolutions; the only condition
-required is that the linear terms of the PDE make up a full-rank operator.
-Hence, in our linear incompressible equations, we permitted use of $\vec{\nabla}
-\cdot \vec{u}_1 = 0$ in lieu of a $\pd{P}{t}$ explicit equation of motion; so
-long as the gauge for $P$ (which only appears in derivatives) is properly set,
-the operator is full rank and numerically well-defined.
-
-In the case of the full nonlinear equations though, it bears noting that $P$
-only appears in nonlinear terms! Thus, the linear terms clearly cannot have full
-rank since they fail to reference $P$ at all. We remedy this by re-casting the
-momentum equation as
-\begin{equation}
-    \pd{\vec{u}}{t} + \frac{\vec{\nabla}P}{\rho_0}
-        = -\vec{\nabla}P\p*{\frac{1}{\rho} - \frac{1}{\rho_0}}
-            + \vec{g}.
-\end{equation}
-This ensures that, along with suitable gauge choice for $P$, the linear
-operators on the left hand side of the equation have full rank. The remaining
-equations
-\begin{subequations}
-    \begin{align}
-        \pd{\rho}{t} &= -\vec{u} \cdot \vec{\nabla}\rho,\\
-        \vec{\nabla} \cdot \vec{u} &= 0,
-    \end{align}
-\end{subequations}
-remain unchanged.
-
-\subsection{Pathology with $u_z$ BCs}
-
-Our problem turns out to be ill-conditioned, the pathology arising from our
-driving term in conjunction with the incompressibility constraint. Since we want
-to model the propagation of waves from where they are excited, rather than
-exciting them directly in our problem, we must drive our region of interest via
-boundary conditions rather than forcing terms.
-
-By driving via boundary conditions, we can only specify boundary conditions on
-the variables being differentiated $\partial_z$ in our equations of motion,
-since we enforce fully periodic BCs in $x$ via using a Fourier spectral
-decomposition. Thus, we are restricted to enforcing BCs on either $\vec{u}$ or
-$P$.
-
-It turns out that by using BCs on $\vec{u}$, we run into a problem: the BCs we
-specify on $u_z$ or its derivatives will in general affect grid points
-$\partial_z u_{z, 1x}$ while those specified on $u_x$ can only affect
-$\partial_x u_{x, 0x}$. This means that $\partial_x u_x$ will pick up
-sharp jerks by being coupled to the BC, producing high frequency components in
-$u_x$ as can be seen in \autoref{fig:agree_plots}. These come from enforcing the
-divergence-free condition too literally at the boundary. Not that the figure was
-produced with only a Dirichlet BC on $u_z$, no BCs on $u_x$, in a linear system
-with absorbing BCs.
-\begin{figure}[!h]
-    \centering
-    \includegraphics[width=0.7\textwidth]{../sims_old/2d_1_strat/agree_plots/sponge_2.png}
-    \caption{Note the jaggedness in $u_x$ induced by a Dirichlet BC on
-    $u_z$.}\label{fig:agree_plots}
-\end{figure}
-
-This problem seems to be remedied somewhat by instead enforcing Neumann BCs on
-$u_z$, or even when allowing dissipation enforcing Neumann BCs on $\partial_z
-u_z$ (the extra derivatives granted by $\nu \nabla^2 \vec{u}$ let us specify
-higher order BCs). However, in the fully nonlinear system we still observe
-divergences near $z=0$ regardless of how smooth of BCs we use on $u_z$.
-Moreover, since the discontinuities come from the nature of the BC and not its
-strength, we run into these instabilities regardless of how gradually we
-introduce the BC (e.g.\ we could consider multiplying by $1 - e^{-t/\tau}$ to
-gradually introduce the BC, but this turns out also to be unstable).
-
-There are thus three possible solutions to this:
-\begin{itemize}
-    \item Consider BCs only on $P$, the only other choice of dynamical variable.
-        This seems to also exhibit instabilities near $z = 0$.
-
-    \item Consider using a $\nabla^2 P$ equation instead of a $\vec{\nabla}
-        \cdot \vec{u} = 0$ constraint. The correct choice of BCs in this case is
-        discussed in Nordstrom et al, 2006. We choose not to investigate this.
-
-    \item Change the system of equations to no longer consider an incompressible
-        system; the anelastic system is an attractive alternative. This boasts
-        the advantage of not having an algebraic constraint equation.
-\end{itemize}
-
-TODO Can taking $\rho_1 \ll \rho_0$, giving us a full rank operator w/o doing
-the janky double add thing, solve in the incompressible limit? Otherwise, go
-anelastic.
-
-\section{Limits of the Incompressibility and Anelasticity Assumptions}
-
-To get a better handle on the regimes of validity of the incompressible and
-anelastic approximations, we return to the full fluid equations
-\begin{equation}
-    \begin{split}
-        \pd{\rho}{t} + \vec{\nabla} \cdot \p*{\rho \vec{u}} &= 0,\\
-        \pd{\vec{u}}{t} + \p*{\vec{u} \cdot \vec{\nabla}}\vec{u} +
-            \frac{\vec{\nabla}P}{\rho} - \vec{g} &= 0.
-    \end{split}
-\end{equation}
-Furthermore, we assume small perturbations $\delta P = c_s^2 \delta \rho$ where
-$c_s^2$ is the sound speed, also the velocity at which an adiabatic perturbation
-is restored. We drop the $-\vec{g}$ from the analysis assuming that it is
-largely negated by an appropriate background $P$.
-
-Given these considerations, we consider in what limit we may presume the fluid
-to be incompressible. Consider if $\abs*{\vec{u}} \ll c_s$, then for some
-characteristic frequency $\omega$ and wavenumber $k$ the momentum equation
-scales as $\omega u + ku^2 + c_s^2 \delta \rho/\rho = 0$. Comparing scales it is
-thus clear that $\delta \rho \ll \rho$ in this limit. Then, considering the
-mass continuity equation, we recognize that the scalings are $\omega \delta \rho
-+ \rho \vec{\nabla} \cdot \vec{u} + k\delta \rho u = 0$. Thus, $\vec{\nabla}
-\cdot \vec{u} = 0$ necessarily, since $\rho$ is much larger than the other two
-terms, and we obtain the incompressible mass continuity equation from the other
-two terms. While this is obviously not a very careful treatment, it should
-suffice to convince us that $u \ll c, \delta \rho \ll \rho, \vec{\nabla} \cdot
-\vec{u} = 0$ are all roughly equivalent conditions.
-
-On the other hand, when may we drop the total derivative terms? This seems to be
-a slightly different criterion, and indeed we may compare $\pd{\vec{u}}{t} +
-\p*{\vec{u} \cdot \vec{\nabla}} \vec{u} \sim \omega \vec{u} + ku^2$. Thus, the
-second term is dominated by the first term when $\vec{u} \ll \vec{v}_{ph}$ where
-$\vec{v}_{ph} \equiv \frac{\omega}{k}\hat{k}$ is the phase velocity. Recall that
-for our stratified waves that $\omega \leq N = \sqrt{g/H}$ as a point of
-reference. So we require $u \ll v_{ph}$ to be able to drop the total derivative
-terms.
-
-Since in general $v_{ph} \ll c_s$, we thus identify that there are three
-regimes. The first, $u \ll v_{ph}$, we solve the completely linear problem.
-Then, for $v_{ph} \lesssim u \ll c_s$, we must begin to consider the advective
-terms but are still allowed to assume $\rho_1 \ll \rho_0$, an assumption that is
-only violated when $u \sim c_s$.
-
-If we look ahead to the anelastic system derivation, we can identify that the
-anelastic approximation corresponds to assuming $c_s^2 \gg uv_{ph}$, making no
-assertion of the scale between $u, v_{ph}$. Thus, we can also consider the
-anelastic approximation with advective terms just fine in teh regime $v_{ph}
-\lesssim u$, and if this is sufficient to induce wave breaking then we may never
-need to consider the $\rho_1 \sim \rho_0$ enormous perturbations.
-
-\chapter{Simulation Setup}
-
-We describe now some concerns that arise in setting up the simulation.
-
-\section{Damping Zone}
-
-Since our code is a spectral code, instead of using the quadratic damping zone
-presented from before (whose second derivative is discontinuous over the entire
-domain), we prefer instead an $\arctan$ profile, such that
-\begin{equation}
-    \Gamma(z) \propto 0.5 * \p*{
-        2 + \tanh C\p*{\frac{z - z_t}{L_z - z_t}}
-            + \tanh C\p*{\frac{z - z_b}{z_b}}}.
-\end{equation}
-This is a profile that mostly vanishes $z \in [z_b, z_t]$ but is unity for $z
-\lesssim z_b, z \gtrsim z_t$, with steepness of transition governed by factor
-$C$. While this profile may have reflection issues if $C$ is chosen to be too
-steep, it is smooth and so has exponentially convergent spectral expansion and
-empirically seems to do alright.
-
-It is moreover prudent to note that, since implicit calculations are expensive,
-as long as the damping timescale $\Gamma^{-1}$ is everywhere resolved it is
-cheaper to put $\Gamma \rho_1, \Gamma \vec{u}_1$ on the right hand sides of
-their equations in Dedalus, which is timestepped explicitly.
-
-\section{Volumetric Forcing}
-
-Due to the structure of \lstinline{Dedalus}, we must use Chebyshev polynomials
-in the vertical direction (to be able to enforce BCs). But as discussed in
-\autoref{s:chebyshev}, the grid spacing of the Chebyshev polynomials near
-boundaries goes as $1/N^2$, which means that interface forcing necessitates an
-extremely strict Courant condition to capture evolution near the interface; this
-is the explanation for the divergences we were seeing in earlier sections. In
-fact, in the literature, the spectral CFL condition meets a $\Delta t_{\max}
-\sim 1/N^2$ scaling thanks to poor bounds near the boundary.
-
-Instead, we employ volumetric forcing. As a toy problem of seeing how this
-works, we will study the restricted 1D wave equation.
-
-\subsection{1D Wave Equation (Deprecated)}
-
-Consider the forced 1D wave equation (we set $c = 1$)
-\begin{equation}
-    \ptd{u(x, t)}{t} - \ptd{u(x, t)}{x} = f(x, t).
-\end{equation}
-Note that the only allowed propagation modes here are $u \sim e^{iC\p*{x \pm
-t}}$ for some constant $C$.
-
-We wish to choose $f(x, t)$ such that $u\p*{x \to -\infty} = 0$ while $u\p*{x
-\to +\infty} \sim e^{i\p*{x - t}}$. It is clear that if we consider the Fourier
-transform $\tilde{f}(k, t)$, the response in $\tilde{u}(k, t)$ will also contain
-components for each of the nonzero $\tilde{f}(k, t)$ components. Thus, if we
-want a pure $e^{i(x - t)}$ outgoing wave, we should have a pure $\tilde{f}(k, t)
-\propto \delta\p*{k - 1}$. However, it is clear that driving the oscillator
-exactly on resonance produces divergent oscillation amplitudes, and moreover it
-is impossible to drive with an $f(x, t)$ that doesn't vanish as $x \to \pm
-\infty$.
-
-It is nonetheless clear that $\tilde{f}(k, t)$ should be strongly peaked at $k =
-1$ and small elsewhere. We can thus consider a form $f(x, t) \propto
-e^{-x^2/2\sigma^2} e^{i(x - t)}$ such that $\sigma \gg 1$, i.e.\ the envelope
-contains many wavelengths. The Fourier transform $\tilde{f}(k, t)$ is then a
-Gaussian with width $1/\sigma \ll 1$ centered at $k = 1$ (convolution).
-
-More precisely, consider $f(x, t) = f_0e^{-x^2/2\sigma^2} e^{i(x - t)}$ and $u(x,
-t) = u_0(x)e^{i(x - t)}$ such that $u_0(x \to \infty) = 0$. Substituting this
-into the inhomogeneous PDE gives
-\begin{equation}
-    i\pd{u_0(x)}{x} - \ptd{u_0(x)}{x} = f_0e^{-x^2/2\sigma^2}.
-\end{equation}
-We note that if $\sigma \gg 1$ then roughly $\pd{}{x} \ll 1$ or
-\begin{equation}
-    \pd{u_0(x)}{x} \approx -if_0e^{-x^2/2\sigma^2}.
-\end{equation}
-Enforcing $u_0(x \to -\infty) = 0$, then $\abs*{u_0(x \to +\infty)} =
-f_0\sigma\sqrt{2\pi}$.
-
-\subsection{Narrow-band forcing}
-
-Instead of doing this on-frequency forcing, we choose instead to use a narrow
-forcing zone $f(x, t) \sim N_{\sigma}(z)e^{it}$ where $\sigma$ is not large.
-Specifically, we need $\sigma$ to be well-resolved spectrally, but we need
-$\tilde{f}(k_z, t)$ not to be small. Thus, since $1/k_z$ must be resolved by the
-simulation, choosing $\sigma = 1/k_z$ is a natural choice.
-
-Note that this implies that waves will be excited going both upwards and
-downwards, as $\tilde{f}(k_z) = \tilde{f}(-k_z)$. We can however place the
-forcing zone near the bottom boundary so that downwards-propagating waves will
-be attenuated by damping zones.
-
-Let us derive the resultant equations of motion. In the interest of maintaining
-a divergence-free velocity field, we place the driving term on the density
-equation. Since the forcing is purely to excite waves and not physical (for the
-time being), it should not worry us that driving the density equation is
-ill-justifiable physically.
-
-Consider the linearized system of equations (to examine the excited perturbation
-in the linear legime)
-\begin{subequations}\label{se:vol_drive}
-    \begin{align}
-        \vec{\nabla} \cdot \vec{u} &= 0,\\
-        \pd{\vec{u}_1}{t} + \frac{\vec{\nabla}P}{\rho_0}
-            + \frac{\rho_1 \vec{g}}{\rho_0}
-            &= 0,\\
-        \pd{\rho_1}{t} - \frac{u_{1z} \rho_0}{H}
-            &= F(x, z, t),\\
-        F(x, z, t) &= \Re Fe^{-\frac{(z - z_0)^2}{2\sigma^2}}
-            e^{i\p*{k_xx - \omega t}}.
-    \end{align}
-\end{subequations}
-For the linearized system, we can immediately substitute $\pd{}{t} \to -i\omega,
-\pd{}{x} \to ik_x$, giving
-\begin{align*}
-    \pd{u_{1z}}{z} + ik_xu_x &= 0,\\
-    -i\omega u_x + \frac{ik_xP}{\rho_0} &= 0,\\
-    -i\omega u_{1z} + \frac{1}{\rho_0}\pd{P}{z} + \frac{\rho_1 g}{\rho_0} &= 0,\\
-    -i\omega \rho_1 - \frac{u_{1z}\rho_0}{H} &= F.
-\end{align*}
-This can be recast solely in terms of $u_{1z}$ as
-\begin{align*}
-    -i\omega u_{1z} + \frac{\omega}{k_x}\p*{-\frac{1}{ik_x}\ptd{u_{1z}}{z}
-        + \frac{1}{ik_xH}\pd{u_{1z}}{z}} - \frac{u_{1z}g}{Hi\omega}
-        - \frac{Fg}{i\omega \rho_0} = 0,\\
-    \omega^2\p*{
-        k_x^2u_{1z} + \frac{1}{H}\pd{u_{1z}}{z} - \ptd{u_{1z}}{z}
-    } - u_{1z}N^2k_x^2 - \frac{Fgk_x^2}{\rho_0} = 0.
-\end{align*}
-Note that if $\pd{}{z} \to 1/2H + ik_z, F = 0$ then this simplifies to the expected
-dispersion relation $\omega^2 = \frac{N^2k_x^2}{k_x^2 + k_z^2 +
-\frac{1}{4H^2}}$.
-
-If we approximate $F(x, z, t) \approx F_0\delta\p*{z - z_0} e^{i(k_xx - \omega t)}$
-for the time being, then clearly the only term that can be equally singular is
-the $\ptd{}{z}$ term, leaving us with
-\begin{equation}
-    -\omega^2 \ptd{u_{1z}}{z} - \frac{Fgk_x^2}{\rho_0} = 0.
-\end{equation}
-Since $u_{1z}$ both above and below $z_0$ satisfy the homogeneous PDE, which has
-plane wave solutions with an $e^{z/2H}$ envelope where $k_zH \gg 1$, we can
-integrate about a small domain around $z_0$ and obtain
-\begin{equation}
-    \at{\pd{u_{1z}}{z}}_{z > z_0} - \at{\pd{u_{1z}}{z}}_{z < z_0}
-        = -F_0\frac{gk_x^2}{\rho_0\omega^2}e^{i(k_xx - \omega t)}.
-\end{equation}
-It should then be apparent that the $u_{1z}$ solution is
-\begin{equation}
-    u_{1z} = \frac{F_0gk_x^2}{\rho_0(z = z_0)\omega^2}\frac{1}{2ik_z} \times
-    \begin{cases}
-        e^{(z - z_0)/2H}e^{i(k_xx + k_z(z - z_0) - \omega t)} & z > z_0\\
-        e^{(z - z_0)/2H}e^{i(k_xx - k_z(z - z_0) - \omega t)} & z < z_0\\
-    \end{cases}.
-\end{equation}
-
-Given that we are actually using a Gaussian $F(x, z, t) \propto F_0e^{-\frac{(z
-- z_0)^2}{2\sigma^2}}$, the integral under $F(z)$ is actually
-$F_0\sqrt{2\pi}\sigma$ rather than just $F_0$. Finally, since a delta function
-is flat in spatial frequency space while a Gaussian tapers off with width
-$1/\sigma$ ($\sigma$ is the spatial width, so $1/\sigma$ is the spatial
-frequency width), we should expect an overall multiplicative factor of
-$e^{-k_z^2\sigma^2 / 2}$ in the $k_z$ spatial component of $F(x, z, t)$ compared
-to the delta function case. Thus, we finally obtain
-\begin{equation}
-    \abs*{u_{1z}(z = z_0)} = \frac{F_0\sqrt{2\pi}\sigma gk_x^2}{
-        2\rho_0\p*{z = z_0}\omega^2 k_z}e^{-k_z^2\sigma^2/2}.
-\end{equation}
-
-\section{Timestepping}
-
-\subsection{Signatures of Insufficient Temporal Resolution}
-
-When running these simulations, it can occasionally be seen that the waves do
-not grow in amplitude like $e^{z/2H}$, such that the energy and flux are not
-constant (or the waves will grow too quickly; this is dependent on the exact
-timestepping used). This is caused by using an overly-long timestep, as can be
-seen with the below illustration.
-
-We study the 1D advection equation $\psi_t = c\psi_x$ corresponding to
-rightwards propagating waves. Consider first order implicit Euler timestepping
-for wave mode $\psi_x = ik\psi$, then
-\begin{equation}
-    \psi_{i + 1} - \psi_i = ick\Delta t\psi_{i + 1}.
-\end{equation}
-Solving yields
-\begin{align}
-    \psi_{i + 1} &= \p*{1 - ick\Delta t}^{-1}\psi_i,\nonumber\\
-        &= \p*{1 - i\omega \Delta t}^{-1}\psi_i.
-\end{align}
-If we then want to step forwards for some total time $T$, with $N \Delta t = T$
-the number of timesteps taken, then we approximate
-\begin{align}
-    \psi(t_0 + T) \approx \p*{1 - i\omega \frac{T}{N}}^{-N}\psi(t_0).
-\end{align}
-It is worth noting that in the $N \to \infty$ limit we recover $\psi(t_0 + T) =
-e^{i\omega T}\psi(t_0)$.
-
-However, consider if $\omega \Delta t \lesssim 1$, then we should see the wave
-amplitude decrease over time like
-\begin{equation}
-    \abs*{\psi(t_0 + T)} \approx \p*{1 + \omega^2 \Delta t^2}^{-N/2}\psi(t_0).
-\end{equation}
-Thus, since we have an implicit first order scheme, the wave amplitude
-\emph{decreases} over time spuriously. An explicit method would produce an
-increase, but \lstinline{Dedalus} uses implicit timestepping. The exact rate of
-decrease depends on the timestepper used, but it is clear that $\omega \Delta t
-\ll 1$ is necessary to avoid this spurious decrease (in numerics language, ``the
-oscillation must be well-resolved'').
-
-\section{Navier-Stokes Regularization}
-
-This was a major tripping point for quite a while. As the wave breaks to higher
-wavenumbers, it is expected that energy will accumulate in the higher $k$
-spectral modes, and so some numerical viscosity must be added. This is not a
-physical viscosity but a numerical one (since spectral codes have no numerical
-viscosity of their own), and so the viscous term is added to all dynamical
-equations $\rd{}{t} \to \pd{}{t} + \vec{u} \cdot \vec{\nabla} - \nu \nabla^2$.
-This makes the equations resemble the Navier-Stokes equations, but since $\nu$
-is not grounded in any physical process, it is not helpful to think of it as a
-Navier-Stokes equation.
-
-We choose the $\nu \nabla^2$ term, which scales $\sim k^2$, such that it becomes
-strong in the same regime that nonlinearity $\p*{\vec{u} \cdot \vec{\nabla}}$
-becomes important. Consider some $k_{turb}$ turbulent wavenumber, then we will
-want $\nu k_{turb}^2 \sim uk_{turb}$. We can get an idea of what $u$ will be
-when turbulence sets in using the wave breaking criterion, that breaking sets in
-when $u \sim \omega/k$ where $k$ is the base wavelength of the wave. This
-substitution suggests we should use $\nu = \frac{\omega}{k_{turb}k}$. We can
-choose $k_{turb} = \frac{2\pi L}{N}$ the grid spacing, so that at this turbulent
-wavelength the regularization kicks in. If we then take $k, \omega$ to be the
-parent wave's properties, we have a prescription for $\nu$.
-
-This turns out to be insufficient for regularization, we need roughly $\nu \sim
-\frac{3\omega_0}{k_zk_{turb}}$.
-
-This is because we performed the naive substitution $k, \omega$ from the parent
-wave. Instead, it is advisable to scale $k, \omega$ to the same $k_{turb}$
-length scale. We first make ansatz that the cascade is roughly Kolmogorov, in
-which case we do the hand-wavy argument that energy $u_l^2$ at a certain length
-scale $l$ is dissipated on timescale $\omega_l = u_l / l$, and so the total
-energy dissipated from the original flow $\epsilon = u_l^3/l$. Thus, $\nu k_l^2
-\sim u_lk_l$ gives $\frac{\omega_0}{\nu k_0^2}\p*{\frac{k_0}{k_{turb}}}^{4/3} =
-1$. This isn't quite right though, as per comparison.
-
-My hypothesis is currently two fold: first, the turbulent cascade is not
-instantaneous, so we can solve a diffusion equation in $k$ space to figure out
-how far the wave can propagate (and grow in amplitude) before the Kolmogorov
-spectrum is obtained. Second, the turbulence causes some energy to be reflected
-in higher $k$ modes, which interact nonlinearly with the flow at lower
-amplitudes.
-
-\subsection{Gotcha: $\nu_x, \nu_z$ and Grid Spacing}
-
-If the grid resolution is uneven in the $x, z$ directions, it might be tempting
-to use different $\nu_x, \nu_z$ per the above criteria, since their $k_{turb} =
-k_{\max}$ differ. This is dangerous however, since $\nu \nabla^2 \vec{u}_1$ has
-to be divergenceless! Thus, it is more advisable to choose an equal grid spacing
-in the two directions.
-
-\chapter{Stability Analysis}
-
-\section{Modulational Instability of NLSE}
-
-We study the nonlinear Schr\"odinger Equation (NLSE)
-\begin{equation}
-    i\psi_t + \psi_{xx} + \nu \abs*{\psi}^2\psi = 0.
-\end{equation}
-Sutherland's paper demonstrates an interaction between the wave and the
-wave-induced mean flow ($\propto \abs*{\psi}^2$) that reduces the IGW system to
-the NLSE\@. Here we follow Whitham's textbook in rederiving the modulational
-instability criterion.
-
-Consider for some fixed value $k_0$, then the linear SE (LSE) admits a plane
-wave solution $\psi = e^{i(k_0x - \omega(k_0)t)}$ for some dispersion relation
-$\omega(k)$. If instead we permit $\psi$ to be distributed in frequency space
-near $k_0$, then we may notate
-\begin{equation}
-    \psi = \psi_0 \varphi = e^{i(k_0x - \omega(k_0) t)} \int\limits
-        \exp\s*{i\p*{k' x - k'\pd{\omega}{k}t -
-            \frac{1}{2}k'^2 \ptd{\omega}{k}t}}\;\mathrm{d}k',
-\end{equation}
-where $k' = k - k_0$ denotes the deviation from the central $k_0$
-wavenumber. Defining $\omega' = \omega - \omega_0$, we may observe
-\begin{equation}
-    \omega' = k' \pd{\omega}{k} + \frac{1}{2}k'^2 \ptd{\omega}{k}.
-\end{equation}
-Making correspondence $-i\omega' \to \partial_t, ik' \to \partial_x$, we see
-that $\varphi$ obeys dispersion relation
-\begin{equation}
-    i\p*{\varphi_t + \pd{\omega}{k}\varphi_x}
-        + \frac{1}{2}\ptd{\omega}{k}\varphi_{xx} = 0.
-\end{equation}
-Thus, the modulation $\varphi$ translates at the group velocity, and we find
-that $\varphi$ obeys a modified LSE in a comoving frame.
-
-If we reintroduce the nonlinearity, we observe that $\nu \abs*{\psi}^2\psi = \nu
-\abs*{\varphi}^2 \psi_0\varphi$, and so the nonlinearity modifies the LSE as
-\begin{align}
-    i\p*{\varphi_t + \pd{\omega}{k}\varphi_x}
-        + \frac{1}{2}\ptd{\omega}{k}\varphi_{xx}
-        + \nu \abs*{\varphi^2}\varphi &= 0,\\
-    \omega' = k' \pd{\omega}{k} + \frac{1}{2}k'^2 \ptd{\omega}{k}
-        - \nu a^2.\label{eq:nlse_disp}
-\end{align}
-
-Armed with this dispersion relation, we can apply the method of characteristics
-to the system of equations describing any wavetrain
-\begin{subequations}\label{eq:wavetrain}
-    \begin{align}
-        \pd{k}{t} + \pd{\omega}{x} &= 0,\label{eq:wavetrain_1}\\
-        \pd{A^2}{t} + \pd{}{x}\p*{c_gA^2} &= 0.\label{eq:wavetrain_2}
-    \end{align}
-\end{subequations}
-We denote $A$ the wave amplitude and $c_g$ the group velocity. In a nonlinear
-system, $\omega = \omega(A, k)$, and to leading order in $A$ we expect $\omega =
-\omega_0 + \omega_2A^2$, i.e.\ the dispersion should not depend on the sign of
-$A$. Thus we may rewrite \autoref{eq:wavetrain} as
-\begin{subequations}\label{eq:wavetrain_mod}
-    \begin{align}
-        \pd{k}{t} + \p*{\pd{\omega_{0}}{k} + \pd{\omega_2}{k}A^2}\pd{k}{x}
-            + \omega_2 \pd{A^2}{x} &= 0,\\
-        \pd{A^2}{t} + \pd{A^2}{x}\pd{\omega_0}{k} + A^2\ptd{\omega_0}{k}\pd{k}{x}
-            &= 0.
-    \end{align}
-\end{subequations}
-We first note that, in $A \to 0$, the equations decouple and both equations
-describe propagation at the group velocity (they are already in characteristic
-form $Q_{,t} + cQ_{,x} = 0$). With some foresight, we anticipate that the
-$\pd{A^2}{x}$ term in the first equation will provide a $\mathcal{O}(A)$
-correction to the group velocity, while the $\pd{\omega_2}{k}A^2\pd{k}{x}$ term
-provides an $\mathcal{O}(A^2)$ correction, so we drop it.
-
-The method of characteristics then requires we seek linear combinations of the
-two equations such that all equations are in characteristic form. The Whitham
-prescription says that if the original system is of form $A_{ij}q_{j, t} +
-B_{ij}q_{j, x} = 0$, then the characteristic curves $X(T)$ satisfy $A_{ij}X +
-B_{ij}T = 0$. Here, this yields
-\begin{equation}
-    X(T) = \p*{\pd{\omega_0}{k} \pm A\sqrt{\omega_2 \ptd{\omega_0}{k}}}T.
-\end{equation}
-
-Comparing to the NLSE dispersion relation \autoref{eq:nlse_disp}, we identify
-that $\sqrt{-\nu \ptd{\omega_0}{k}}A$ is the correction to the group velocity.
-Thus, if $\nu\ptd{\omega_0}{k} > 0$, we obtain a \emph{complex} group velocity,
-which is identified with \emph{modulational instability}. Naively, this might
-imply a steepening $k = c_{g, 0}t - i\sqrt{\nu\omega_0''}At$ with timescale
-$\Gamma \sim \p*{A^2\nu\omega_0''}^{-1/2}$.
-
-\chapter{Stratified Flow}
-
-\section{Linear Equations}
-
-We largely follow Booker \& Bretherton from 1966 though do not adopt the
-Boussinesq approximation. Consider background shear flow $\vec{u}_0 =
-u_0(z)\hat{x}$ and vertical stratification $\rho_0 \propto e^{-z/H}$, then
-\begin{subequations}\label{se:shear_eq}
-    \begin{align}
-        \vec{\nabla} \cdot \vec{u}_1 &= 0,\\
-        \pd{\rho_1}{t} + \p*{\vec{u} \cdot \vec{\nabla}}\rho_1
-            - u_{1z}\frac{\rho_0}{H} &= 0,\\
-        \pd{u_{1x}}{t} + \p*{\vec{u} \cdot \vec{\nabla}}\p*{u_{1x} + u_0}
-            + \frac{1}{\rho_0}\pd{P}{x} &= 0,\\
-        \pd{u_{1z}}{t} + \p*{\vec{u} \cdot \vec{\nabla}}u_{1z}
-            + \frac{1}{\rho_0}\pd{P}{z} + \frac{\rho_1 g}{\rho_0} &= 0.
-    \end{align}
-\end{subequations}
-Note that $\vec{u} \cdot \vec{\nabla} = \p*{\vec{u}_0 + \vec{u}_1} \cdot
-\vec{\nabla}$.
-
-We again assume plane waves in $x, t$ such that $u_{1z}(x, z, t) =
-u_{1z}(z)e^{i(k_xx - \omega t)}$. We bracket second-order terms below and set
-them to zero in the ensuing calculations
-\begin{subequations}
-    \begin{align}
-        \pd{u_{1z}}{z} + ik_xu_{1x} &= 0,\\
-        -i\omega \rho_1 + u_0ik_x\rho_1
-            - u_{1z}\frac{\rho_0}{H}
-            + \s*{\p*{\vec{u}_1 \cdot \vec{\nabla}} \rho_1} &= 0,\\
-        -i\omega u_{1x} + u_0ik_x u_{1x}
-            + \frac{1}{\rho_0}ik_xP + u_{1z}\pd{u_0}{z}
-            + \s*{\vec{u}_1 \cdot \vec{\nabla} u_{1x}} &= 0,\\
-        -i\omega u_{1z} + u_0ik_x u_{1z}
-            + \frac{1}{\rho_0}\pd{P}{z} + \frac{\rho_1 g}{\rho_0}
-            + \s*{\vec{u}_1 \cdot \vec{\nabla} u_{1z}} &= 0.\\
-    \end{align}
-\end{subequations}
-Denote $\tilde{\omega}(z) = \omega - u_0(z)k_x$ the corotating frequency. Using
-then that $u_{1x} = \frac{i}{k_x}\pd{u_{1z}}{z}$, $\rho_1 = \frac{i
-\rho_0}{H\tilde{\omega}} u_{1z}$ and $P = \s*{i\tilde{\omega}u_{1x} - u_{1z}
-\pd{u_0}{z}} \frac{\rho_0}{ik_x}$, we can finally arrive at linearized equation
-\begin{equation}
-    \ptd{u_{1z}}{z} - \frac{1}{H}\pd{u_{1z}}{z} + \s*{
-        \frac{k_x^2N^2}{\tilde{\omega}^2}
-            - k_x^2
-            - \frac{k_x}{\tilde{\omega}H}\pd{u_0}{z}
-            + \frac{k_x}{\tilde{\omega}} \ptd{u_0}{z}
-    } u_{1z} = 0. \label{eq:shear_lin_uz}
-\end{equation}
-Note that if we make correspondences $ik_x \to \partial_x, i\tilde{\omega} \to
-\partial_t + u_0\partial_x$ and take $H \to \infty$ while holding $N^2$
-constant, we recover Equation 2.1 of Booker \& Bretherton. We can also note that
-if we take constant $u_0$ we recover the expected dispersion relation
-\autoref{eq:strat_disp} for $\omega = \tilde{\omega}$.
-
-As a brief exercise, we can take the $H \to \infty$ limit and compute the
-resultant shear-induced splitting; this is akin to rotational mode splitting,
-and is given without proof from $-k_z^2 - k_x^2 + \frac{k_x^2N^2}{\omega^2} +
-\frac{k_x}{\omega}\ptd{u_0}{z} = 0$ as
-\begin{equation}
-    \omega^2 = \frac{k_x^2 N^2}{k_x^2 + k_z^2}\s*{1 + \frac{1}{\sqrt{k_x^2 +
-        k_z^2}} \ptd{u_0}{z}}.
-\end{equation}
-
-\section{Reflection/Transmission in Linear Theory}
-
-One can solve \autoref{eq:shear_lin_uz} via \emph{Frobenius Theory}, but if we
-consider only $\tilde{\omega} \to 0$, we obtain a very simple power series
-solution. In $\tilde{\omega} \to 0$, we have only leading order contributions
-\begin{equation}
-    \ptd{u_{1z}}{z} + \frac{k_x^2N^2}{\tilde{\omega}^2}u_{1z} = 0.
-\end{equation}
-Consider a power series solution $u_{1z} \propto \tilde{\omega}^\alpha$, where
-$\tilde{\omega}(z) = \omega - k_xu_0(z)$. Note that very near the critical level,
-$\pd{u_0}{z}\delta z k_x = \tilde{\omega}$ where $\delta z = z - z_c,
-\tilde{\omega}(z_c) = 0$. We thus have
-\begin{align}
-    \p*{\pd{\tilde{\omega}}{z}}^2 \alpha\p*{\alpha - 1} + k_x^2N^2 &= 0,\\
-    \alpha &= \frac{1}{2}\pm i\sqrt{\frac{N^2}{\p*{\pd{u_0}{z}}^2} -
-        \frac{1}{4}}.
-\end{align}
-The classic Booker \& Bretherton paper notates this result $u_{1z} \propto
-\p*{\tilde{\omega}(z)}^{\frac{1}{2} \pm i\mu} \propto (\delta z)^{\frac{1}{2} \pm
-i\mu}$.
-
-Now, consistent with a perturbation that vanishes as $t \to -\infty$
-(alternatively, this is just the Landau prescription for deforming below a
-singularity on the real line), we must interpret $\tilde{\omega}$ as having a
-slight positive imaginary component. Given this, the expression for $u_{1z}(z)$
-is multivalued, and we must specify a branch cut. For $\Arg\p*{z - z_c} = 0$, or
-intuitively when we are above the critical layer, choose branch such that
-\begin{equation}
-    \p*{z - z_c}^{\frac{1}{2} \pm i\mu} \equiv
-        \abs*{z - z_c}^{1/2\pm i\mu}.
-\end{equation}
-It is then clear that for $\Arg\p*{z - z_c} = -\pi$, intuitively corresponding
-to being below the critical layer (note the arg is $-\pi$ since $z_c$ is
-infinitesimally above the real axis), our branch cut enforces
-\begin{equation}
-    \p*{z - z_c}^{\frac{1}{2} \pm i\mu} =
-        \abs*{z - z_c}^{\frac{1}{2} \pm i\mu}e^{-\frac{i\pi}{2}}e^{\mp \mu \pi}.
-\end{equation}
-Lastly, we interpret $\abs*{z - z_c}^{\pm i\mu} = e^{\pm i\mu \ln \abs*{z -
-z_c}}$ to have the form of a positive/negative $k_z$ value. Interestingly, this
-means that $e^{-i\mu \ln \abs*{z - z_c}}$ corresponds to the \emph{positive
-group velocity}, which is $e^{\mu \pi}$ magnified compared to the $\Arg\p*{z -
-z_c} = 0$ case, and $e^{2\mu \pi}$ magnified over the negative group velocity
-component below $z_c$.
-
-The interpretation we choose for this is that an incoming wavepacket moving with
-positive group velocity inbound on $z_c$ is transmitted with amplitude $e^{-\mu
-\pi}$ and reflected with amplitude $e^{-2\mu \pi}$. The remainder is interpreted
-as being \emph{absorbed} by the critical layer.
-
-\section{Competition between Viscous and Nonlinear Forces}
-
-As we zoom in on the critical layer, we expect to be able to take Boussinesq
-approximation $H \to \infty$ within the immediate vicinity of the critical
-layer.
-
-In this limit, the group velocity is just $v_{g,z} = -\frac{Nk_xk_z}{\p*{k_x^2 +
-k_z^2}^{3/2}} = -\omega\frac{k_z}{k_x^2 + k_z^2}$. As we approach corotation,
-$\omega \to 0, \omega \approx \frac{Nk_x}{k_z}, k_z \to \infty$, and so $v_{g,
-z} \to 0$. Then considering an energy flux incident on the critical layer
-clearly shows that energy must accumulate near the critical layer, and must be
-dissipated somehow. This must be either via viscous forces or nonlinear
-cascades.
-
-The characteristic strength of the viscous force $\nu \nabla^2$ is $\nu k_z^2$,
-while the nonlinear advective term has characteristic strength $u_{1z}k_z$. Note
-that in the Boussinesq approximation $\abs*{u_{1z}}$ is constant during
-propagation. Thus, the ratio of forces $R = \frac{f_{\nu}}{f_{adv}} \sim
-\frac{\nu k_z}{u_{1z}}$. Note that this ratio varies spatially. Qualitatively,
-$R$ is important since it dictates whether the wave is damped viscously before
-it needs to be treated in nonlinear fashion or whether a full
-nonlinear/hydrodynamic treatment is necessary.
-
-We want to understand for what viscosity we expect linear behavior to dominate.
-Thus, at critical $z_{c}$ where nonlinear effects become important, we want $R
-\gtrsim 1$. Shifting to coordinate system $\delta z$ where $\delta z = 0$ is the
-corotation resonance, we can write $\abs*{\tilde{\omega}} = k_xu_0' \delta z$ where
-$u_0' = \at{\pd{u_0}{z}}_{\delta z = 0}$, and $k_z = \frac{N}{u_0' \delta z}$.
-Then the breaking criterion is $\frac{u_{1z}k_z(z)}{\tilde{\omega}(z)} \sim 1$, or
-$\delta z_{c} \sim \sqrt{\frac{u_{1z}N}{k_xu_0'^2}}$. We can then evaluate
-\begin{equation}
-    R\p*{z_{c}} = \frac{\nu k_z}{u_{1z}} = \nu \sqrt{\frac{Nk_x}{u_{1z}^3}}.
-\end{equation}
-
-It is helpful also to write this down for any arbitrary hyperviscosity $f_{visc}
-= \nu^{(n)} \nabla^{(n)}$, which is cited without proof
-\begin{equation}
-    R\p*{z_c} = \nu^{(n)} \sqrt{\frac{N^{n - 1}k_x^{n - 1}}{u_{1z}^{n + 1}}}.
-\end{equation}
-
-\chapter{Cylindrical Stratified Flow}
-
-We fearlessly tackle the cylindrical coordinate problem, using Tsang \& Lai 2008
-non-barotropic flows as a guide but reproducing all algebra. We show that in an
-appropriate limit it recovers \autoref{eq:shear_lin_uz}.
-
-First, we take $\vec{\nabla} \cdot \vec{u}_0 = 0$, which we can define $u_0 =
-u_0(r)\hat{\phi} = r\Omega(r) \hat{\phi}$. Furthermore, under hydrostatic
-equilibrium if we assume a thin atmosphere (such that $\vec{\nabla}\Phi$ does
-not vary much) then $\frac{\vec{\nabla}P_0}{\rho_0} = -g\hat{r}$. We assume
-stratification $\rho \propto e^{-r/a_0}$, which gives buoyancy frequency $N^2 =
-g/a_0$ in the incompressible fluid limit. Finally, we make plane wave
-substitution $\propto e^{i\p*{m\phi - \omega t}}$ and begin
-\begin{align}
-    0 ={}& \frac{1}{r}\pd{\p*{u_{1r}}}{r} + \frac{im}{r}u_{1\phi},\\
-    0 ={}& -i\omega \rho_1 - u_{1r}\frac{\rho_0}{a_0} + im\Omega\rho_1
-        \z*{+ u_{1r}\pd{\rho_1}{r} + \frac{u_{1\phi}r}{r}\pd{\rho_1}{\phi}}
-            ,\\
-    0 ={}& -i\omega \vec{u}_1
-            + \s*{\Omega im u_{1r} - 2\Omega u_{1\phi}} \hat{r}
-            +\s*{u_{1r}\rd{\p*{r\Omega}}{r} + \Omega imu_{1\phi}
-                + \Omega u_{1r}}\hat{\phi}
-            + \frac{1}{\rho_0}\s*{\pd{P_1}{r}\hat{r} + \frac{imP_1}{r}
-                \hat{\phi}} + \frac{\rho_1 g}{\rho_0}\hat{r}\nonumber\\
-        &+ \z*{\s*{u_{1r}\pd{u_{1r}}{r} + \frac{u_{1\phi}}{r}
-            \pd{u_{1r}}{\phi} - \frac{u_{1\phi}^2}{r}}\hat{r}
-            + \s*{u_{1r}\pd{u_{1\phi}}{r}
-                + \frac{u_{1\phi}}{r}\pd{u_{1\phi}}{\phi}
-                + \frac{u_{1\phi} u_{1r}}{r}}\hat{\phi}}.
-\end{align}
-We have bracketed nonlinear terms in curly braces. The natural definition
-$\tilde{\omega} = \omega - m\Omega$ and common definition $\frac{\kappa^2}{2\Omega}
-= \Omega + \rd{(r\Omega)}{r}$ may be used, which agrees with Tsang \& Lai.
-Dropping bracketed nonlinear terms allows us to manipulate
-\begin{align}
-    u_{1\theta} ={}& \frac{i}{m} \pd{ru_{1r}}{r},\\
-    \rho_1 ={}& u_{1r}\frac{\rho_0}{a_0}\frac{1}{-i\tilde{\omega}},\\
-    P_1 ={}& \frac{i\rho_0 r}{m}\s*{-i\tilde{\omega}u_{1\theta}
-        + u_{1r}\frac{\kappa^2}{2\Omega}},\\
-    0 ={}& -i\tilde{\omega}u_{1r} - 2\Omega\p*{\frac{i}{m}\pd{(ru_{1r})}{r}}
-        + \frac{1}{\rho_0}\pd{}{r}\s*{\frac{i\rho_0 r}{m}\s*{
-            -i\tilde{\omega} \frac{i}{m}\pd{(ru_{1r})}{r}
-            +u_{1r}\frac{\kappa^2}{2\Omega}}}
-        + \frac{iu_{1r}g}{a_0\tilde{\omega}}.
-\end{align}
-Expanding the rather ugly derivative gives
-\begin{align}
-    0 ={}& -i\tilde{\omega}u_{1r} - \frac{2i\Omega}{m}\s*{r\pd{u_{1r}}{r}
-        + u_{1r}} + \p*{\frac{i}{m} - \frac{ir}{a_0m}}
-            \s*{\frac{\tilde{\omega}}{m}
-            \p*{r\pd{u_{1r}}{r} + u_{1r}} + u_{1r}\frac{\kappa^2}{2\Omega}}
-            \nonumber\\
-        &+ \frac{ir}{m}\s*{\frac{\tilde{\omega}}{m}\p*{r\ptd{u_{1r}}{r}
-            + 2\pd{u_{1r}}{r}} + \frac{\kappa^2}{2\Omega}\pd{u_{1r}}{r}
-            + u_{1r}\pd{}{r}\p*{\frac{\kappa^2}{2\Omega}}}
-            \nonumber\\
-        &+ \frac{r}{m}\pd{\tilde{\omega}}{r}\frac{i}{m}\p*{r\pd{u_{1r}}{r}
-            + u_{1r}} + \frac{iu_{1r}g}{a_0\tilde{\omega}}.
-\end{align}
-Let's group terms to find
-\begin{align}
-    \frac{\tilde{\omega}r^2}{m^2}\ptd{u_{1r}}{r} + &
-        \pd{u_{1r}}{r}\s*{-\frac{2\Omega r}{m}
-            + \frac{r\tilde{\omega}}{m} \p*{\frac{1}{m} - \frac{r}{a_0m}}
-            + \frac{2r\tilde{\omega}}{m^2}
-            + \frac{r}{m}\frac{\kappa^2}{2\Omega}
-            + \frac{r^2}{m^2}\pd{\tilde{\omega}}{r}}\nonumber\\
-        & +u_{1r}\s*{-\frac{2\Omega}{m}
-            + \frac{\tilde{\omega}}{m} \p*{\frac{1}{m} - \frac{r}{a_0m}}
-            + \p*{\frac{1}{m} - \frac{r}{a_0m}}\frac{\kappa^2}{2\Omega}
-            + \frac{r}{m}\pd{}{r}\p*{\frac{\kappa^2}{2\Omega}}
-            + \frac{r}{m^2}\pd{\tilde{\omega}}{r}
-            + \frac{N^2}{\tilde{\omega}} - \tilde{\omega}} = 0.
-\end{align}
-One small simplification can be made: observe that $\frac{\kappa^2}{2\Omega} =
-\Omega + \pd{(r\Omega)}{r} = 2\Omega + r\pd{\Omega}{r}$, so
-\begin{equation}
-    \frac{\tilde{\omega}r^2}{m^2}\ptd{u_{1r}}{r}
-        + \pd{u_{1r}}{r}
-            \frac{r\tilde{\omega}}{m^2} \p*{3 - \frac{r}{a_0}}
-        + u_{1r}\s*{
-            \frac{\tilde{\omega}}{m} \p*{\frac{1}{m} - \frac{r}{a_0m}}
-            - \frac{r}{a_0m}\frac{\kappa^2}{2\Omega}
-            + \frac{r}{m}\pd{}{r}\p*{\frac{\kappa^2}{2\Omega}}
-            + \frac{N^2}{\tilde{\omega}} - \tilde{\omega}} = 0.
-\end{equation}
-I think that the $3$ actually makes sense, it's basically
-$\pd{}{r}r\pd{(ru_r)}{r} = \dots + 2r\pd{u_r}{r}$ and $\pd{(ru_r)}{r} =
-r\pd{u_r}{r} + u_r$, which also explains the $1$.
-
-\section{Comparison to Plane Parallel}
-
-\subsection{Reduction in $r \to \infty$}
-
-First, let's consider the limit where $r \to \infty$, so we should have a plane
-parallel waveform. In this limit, $\frac{m}{r} \to k_x$ can be identified (and
-so $m$ must also be taken to be infinitely large to have finite wavelengths),
-and of course $\Omega \to \frac{u_0}{r}$ in line with our plane parallel
-definitions. We need one last identification, that $\pd{\tilde{\omega}}{r} =
--m\rd{\Omega}{r} \approx -k_x\rd{u_0}{r}$ also per the plane parallel result.
-These identifications allow us to write, approximating $\frac{\kappa^2}{2\Omega}
-\approx \rd{u_0}{r}$:
-\begin{equation}
-    \ptd{u_{1r}}{r} - \pd{u_{1r}}{r}\frac{1}{a_0} +
-        u_{1r}\s*{\frac{N^2k_x^2}{\tilde{\omega}^2}
-            + \frac{k_x}{\omega} \rtd{u_0}{r} - \frac{k_x}{a_0\tilde{\omega}}
-            \rd{u_0}{r} - k_x^2}u_{1r} = 0.
-\end{equation}
-This can be seen to agree with \autoref{eq:shear_lin_uz}, which is no small
-miracle for my usual algebra skills.
-
-\subsection{Boundary Matching at $\tilde{\omega} = 0$}
-
-Again, there is nothing more singular than a $\tilde{\omega}^{-2}$ coefficient
-on $u_{1r}$, after dividing out the leading coefficient, and so we may simply
-analyze
-\begin{equation}
-    \ptd{u_{1r}}{r} + u_{1r}\frac{N^2m^2}{\tilde{\omega}^2r^2} = 0.
-\end{equation}
-Using the usual $u_{1r} = (r - r_c)^\alpha = \delta r^\alpha$ and
-$\tilde{\omega} = -m\at{\pd{\Omega}{r}}_{r_c}\delta r \equiv -m\Omega'_c \delta
-r$ gives
-\begin{equation}
-    \alpha(\alpha - 1) + \frac{N^2}{(\Omega'_c)^2r^2} = 0.\label{eq:cylin_crit}
-\end{equation}
-This solves to $\alpha = \frac{1}{2} \pm
-i\sqrt{\frac{N^2}{(\Omega'_c)^2r^2} - \frac{1}{4}}$. This agrees with
-the exponent derived in the Tsang \& Lai paper. Thus, in the linear theory,
-transmission/reflection are attenuated by factor
-$\exp\s*{-\pi\sqrt{\frac{N^2}{(\Omega'_c)^2 r^2} - \frac{1}{4}}}$.
-
-It can be observed in passing that if $N \sim \Omega'_c r$ then the scenario
-reduces to that considered in Tsang \& Lai, where some variable such as $\zeta$
-can induce superreflection. We don't compute it here, but it bears mentioning
-that the corresponding analysis for Tsang \& Lai in plane parallel gives $\nu =
--4\rd{\ln u_0'}{r}$, and so $u_0'$ plays the role of $\zeta$.
-
-Note that the H\o iland Criteria, such as given in Balbus \& Hawley 1998, states
-that disks are hydrodynamically stable if
-\begin{equation}
-    N^2 + \frac{1}{R^3}\pd{R^4 \Omega^2}{R} > 0.
-\end{equation}
-We may compare this to our transmission/reflection criterion by computing at
-critical $\Omega_c$, where
-\begin{equation}
-    R^2 \p*{\Omega_c'}^2 = \p*{\frac{N^2 + 4\Omega^2}{2\Omega}}^2.
-\end{equation}
-This was never going to agree with \autoref{eq:cylin_crit}, since it doesn't
-depend simply on $\Omega_c'$. This reinforces Booker \& Bretherton's conclusion
-that the agreement of their calculation with convective stability was pure
-coincidence.
-
-\section{Wave Properties}
-
-Let's compute dispersion relation. I'm too lazy to sort out $u_{1r} \propto
-e^{r/2a_0}$ nonsense, so we will take Boussinesq approximation $a_0 \to \infty$,
-and so
-\begin{equation}
-    \ptd{u_{1r}}{r} + \pd{u_{1r}}{r}\p*{\frac{3}{r} - \frac{1}{a_0}}
-        + u_{1r}\s*{\p*{\frac{1}{r^2} - \frac{1}{a_0r}}
-            + \frac{m}{r\tilde{\omega}} \pd{}{r}\p*{\frac{\kappa^2}{2\Omega}}
-            + \frac{N^2m^2}{r^2\tilde{\omega}^2} - \frac{m^2}{r^2}} = 0.
-\end{equation}
-We've been a bit hand-wavy about what terms to drop, but we can identify now
-that some of the terms can come together to form (we have to drop the
-$\pd{u_{1r}}{r}$ coefficient, otherwise we'll have $ik_r$ terms\dots ugh)
-\begin{equation}
-    \tilde{\omega}^2 \simeq \frac{N^2\p*{m^2/r^2}}{k_r^2 + \frac{m^2}{r^2}}.
-\end{equation}
-Thus, in a nutshell, we should expect the group velocity to largely vanish in
-cylindrical coordinates once the WKB regime is hit (where we're reasonably far
-away that we don't have to worry about the non-wave-like distortion of the
-geometry; is this the same as the plane wave limit though?).
-
-\clearpage
-
-\appendix
-
-\chapter{Deriving Fluids Results}
-
-\section{Equation of Energy Conservation}
-
-We follow Landau \& Lifshitz's derivation for this expression. Consider that the
-total energy stored in a wave must be the sum of its kinetic and internal energy
-$\frac{1}{2}\rho v^2 + \rho \epsilon$ where $\epsilon$ is the internal energy
-density. To obtain an equation of energy conservation, we must take the time
-derivative of this expression. First, we consider
-\begin{align}
-    \pd{}{t}\p*{\frac{1}{2}\rho v^2} &= \frac{1}{2}v^2\pd{\rho}{t}
-        + \rho \vec{v} \cdot \pd{\vec{v}}{t},\nonumber\\
-        &= -\frac{1}{2}v^2 \p*{\vec{\nabla} \cdot \rho \vec{v}}
-            - \rho \vec{v} \cdot \p*{
-                \p*{\vec{v} \cdot \vec{\nabla}}\vec{v} +
-                    \frac{\vec{\nabla}P}{\rho} - \vec{g}
-            }_{T, V},\nonumber\\
-        &= -\frac{1}{2} v^2 \p*{\vec{\nabla} \cdot \rho \vec{v}}
-            - \frac{1}{2}\rho \vec{v} \cdot \p*{\vec{\nabla} v^2}
-            - \vec{v} \cdot \p*{
-                \rho \vec{\nabla} w - \rho T \vec{\nabla}s
-            } + \rho \vec{v} \cdot \vec{g}.
-\end{align}
-We have denoted $s$ the specific internal entropy density of the fluid and
-$\mathrm{d}w = T\mathrm{d}s + \frac{\mathrm{d}P}{\rho} = T\mathrm{d}s +
-\epsilon$ the specific internal enthalpy density of the fluid.
-Recall enthalpy $\epsilon = w - Ts$ is the usual thermodynamic definition.
-
-At the same time, consider
-\begin{align}
-    \pd{}{t}\p*{\rho \epsilon} &= \epsilon \pd{\rho}{t}
-        + \rho \pd{}{t}\p*{Ts - \frac{P}{\rho}}_{T, P},\nonumber\\
-        &= \epsilon \pd{\rho}{t} + \rho T \pd{s}{t} + \frac{P}{\rho}\pd{\rho}{t}
-            ,\nonumber\\
-        &= w \pd{\rho}{t} + \rho T \pd{s}{t},\nonumber\\
-        &= -w\vec{\nabla}\p*{\rho \vec{v}} - \rho T \vec{v} \cdot \p*{
-            \vec{\nabla}s
-        }.
-\end{align}
-Summing the two, we find
-\begin{equation}
-    \pd{}{t}\p*{\frac{\rho v^2}{2} + \rho \epsilon}
-        = -\vec{\nabla} \cdot \s*{
-            \rho \vec{v}\p*{v^2 + w}
-        } = -\vec{\nabla} \cdot \s*{
-            \rho \vec{v}\p*{v^2 + \epsilon + \frac{P}{\rho}}
-        }.
-\end{equation}
-
-\section{Conservative Fluid Equations}
-
-This isn't particularly useful for our work since spectral methods don't benefit
-from having the dynamical equations in conservative form, but form a useful
-exercise for the ever-so-stupid writer.
-
-The Reynolds transport theorem tells us that
-\begin{equation}
-    \rd{}{t}\int\limits_V Q\;\mathrm{d}V
-        = \int\limits_V \pd{Q}{t}\;\mathrm{d}V
-            + \int\limits_{\partial V}
-                \p*{\vec{v}\p*{\mathrm{d}A} \cdot \hat{n}} Q\;\mathrm{d}A.
-\end{equation}
-We have notated $\vec{v}\p*{\mathrm{d}A}$ the velocity of the surface element.
-In fluid mechanics, this is just the velocity field. Thus, for conserved
-quantities where $\rd{}{t}\int\limits_V Q\;\mathrm{d}V = 0$, we can apply
-divergence theorem to obtain the simple result
-\begin{equation}
-    \pd{Q}{t} + \vec{\nabla} \cdot \p*{Q\vec{v}} = 0.\label{eq:cons_eq}
-\end{equation}
-Of course, if there are any net forces etc.\ on the system, they are simply
-inserted as fileds on the right hand side of \autoref{eq:cons_eq}.
-
-The application of this to the three major conserved quantities: mass, momentum
-and energy, yield:
-\begin{description}
-    \item[Conservation of Mass] This produces simply $\pd{\rho}{t} +
-        \vec{\nabla} \cdot \p*{\rho \vec{v}} = 0$ the continuity equation.
-
-    \item[Conservation of Momentum] This produces $\pd{\rho v_i}{t} +
-        \vec{\nabla} \cdot \p*{\rho v_i \vec{v}} = 0$. To obtain the traditional
-        momentum equation, it is perhaps most clear in index notation where
-        repeated indicies denote summation:
-        \begin{align}
-            \pd{\p*{\rho v_i}}{t} + \partial_j \rho v_iv_j &= 0,\nonumber\\
-            v_i \pd{\rho}{t} + \rho \pd{v_i}{t} + v_I \partial_j \p*{\rho v_j}
-                + \rho v_j \partial_j v_i &= 0,\nonumber\\
-            v_i\p*{\pd{\rho}{t} + \partial_j \p*{\rho v_j}}
-                + \rho\p*{
-                    \pd{v_i}{t} + v_j \partial_t v_i
-                } = 0.
-        \end{align}
-        The first parenthetical term we recognize is just the continuity
-        equation though. The remainder is the traditional source-free momentum
-        equation.
-
-    \item[Conservation of Energy] This just reads $\pd{e}{t} + \vec{\nabla}
-        \cdot \p*{e \vec{v}} = 0$. It's a pretty uninteresting equation without
-        pressure and thermodynamic effects, which would arise from a more
-        careful treatment of the source terms.
-
-        In fact, pressure contributes a source term $P\vec{\nabla} \cdot
-        \vec{v}$, so the full energy equation is often written
-        \begin{equation}
-            \rd{e}{t} + \p*{e + P}\vec{\nabla} \cdot \vec{v} = 0.
-        \end{equation}
-\end{description}
-
-\section{The Anelastic/Boussinesq Approximations}
-
-\subsection{Developing the Anelastic/Boussinesq Approximations}
-
-Let's relax the incompressibility constraint (we will expand the continuity
-equation to first order, but the momentum equation will merit a separate
-treatment):
-\begin{equation}
-    \begin{split}
-        \pd{\rho_1}{t} + \vec{\nabla} \cdot \p*{\rho_0 \vec{u}_1} &= 0,\\
-        \pd{\vec{u}_1}{t} &= -\frac{\vec{\nabla}P}{\rho} - \vec{g}.
-    \end{split}
-\end{equation}
-Suppose we are interested in phenomena with characteristic length scale $L$ and
-time scale $\tau$. Let's first examine the relative magnitudes of the terms in
-the continuity equation
-\begin{equation*}
-    \frac{\rho_1}{\tau} + \frac{\rho_0 \abs{u_1}}{L} = 0.
-\end{equation*}
-Thus, if we are interested in time scales $\tau \gg
-\frac{\rho_1}{\rho_0}\frac{L}{\abs{u_1}}$ then we neglect the first term, the
-time derivative. This corresponds to making the perturbation incompressible;
-note that $\pd{\rho_1}{t} \approx \rd{\rho_1}{t}$ to first order, so we drop the
-high frequency restoring forces in the perturbation.
-
-For the momentum equation, we instead first manipulate to first order
-\begin{align}
-    -\frac{\vec{\nabla}P}{\rho} - \vec{g}
-        &= -\frac{\vec{\nabla}P_0}{\rho} - \frac{\vec{\nabla}P_1}{\rho_0}
-            - \vec{g},\nonumber\\
-        &= - \frac{\vec{\nabla}P_1}{\rho_0} +
-            \p*{\frac{\rho_0}{\rho} - 1}\vec{g},\nonumber\\
-        &= -\vec{\nabla}\p*{\frac{P_1}{\rho_0}}
-            - \frac{P_1}{\rho_0^2}\vec{\nabla}\rho_0
-            - \frac{\rho_1}{\rho_0}\vec{g}.
-\end{align}
-
-We now have three equations for four variables, $\vec{u}_1, \rho_1, P_1$. We
-must introduce a fourth equation, a thermodynamic equation. For an adiabatic
-process $P\rho^{-\gamma} \propto P^{1-\gamma}T^\gamma$ is constant. We thus
-introduce the concept of the \emph{potential temperature}
-\begin{equation}
-    \theta = T\p*{\frac{P_0}{P}}^\kappa.
-        \label{eq:pot_temp}
-\end{equation}
-For an adiabatic process, $\rd{\theta}{t} = 0$. Motivated by this, we use
-\begin{equation}
-    \begin{split}
-        \pd{1}{\rho_0}\pd{\rho_0}{z} &= \frac{1}{\gamma P_0}\pd{P_0}{z}
-            - \frac{1}{\theta_0}\pd{\theta_0}{z},\\
-        \frac{\rho_1}{\rho_0} &= \frac{1}{\gamma} \frac{P_1}{P_0}
-            - \frac{\theta_1}{\theta_0},
-    \end{split}
-\end{equation}
-to give the momentum equation form
-\begin{equation}
-    \rd{\vec{u}_1}{t} = -\vec{\nabla}\p*{\frac{P_1}{\rho_0}}
-        + \frac{P_1}{\rho_0}\p*{\frac{1}{\theta_0}\vec{\nabla}\theta_0}
-        + \vec{g} \frac{\theta_1}{\theta_0}.\label{eq:mom_unsimp}
-\end{equation}
-
-We also recognize $N^2 = \frac{g}{\theta_0}\pd{\theta_0}{z}$. We now do the same
-trick where we consider dynamics on length scale $D$ and compare the first and
-second terms in \autoref{eq:mom_unsimp}. Their ratio is $\frac{N^2 D}{g}$, and
-so as $N^2 \ll \frac{g}{D}$ the freefall time we neglect the second term.
-
-The anelastic fluid equations thus read
-\begin{equation}
-    \begin{split}
-        \vec{\nabla} \cdot \p*{\rho_0\vec{u}} &= 0,\\
-        \pd{\vec{u}_1}{t} + \vec{\nabla}\p*{\frac{P_1}{\rho_0}}
-            - \vec{g}\frac{\theta_1}{\theta_0} &= 0,\\
-        \pd{\theta_1}{t} + \p*{\vec{u} \cdot \vec{\nabla}}\theta_0 &= 0.
-    \end{split}\label{eq:lin_anelastic}
-\end{equation}
-
-The Boussinesq equations are obtained from these in the limit where $H \gg D$
-the relevant length scale, thus we allow $\rho_0$ to be approximately constant.
-
-\subsection{Anelastic Solution to Stratified Atmosphere}
-
-We simply substitute $e^{i(\vec{k} \cdot \vec{r} - \omega t)}$ into
-\autoref{eq:lin_anelastic} with $\rho_0 \propto e^{-z/H}$ and obtain
-\begin{equation}
-    \begin{bmatrix}
-        0 & 0 & ik_x\rho_0 & ik_z \rho_0 - \frac{\rho_0}{H}\\
-        0 & -i\omega & 0 & \frac{N^2\theta_0}{g}\\
-        \frac{ik_x}{\rho_0} & 0 & -i\omega & 0\\
-        \frac{ik_z}{\rho_0} + \frac{1}{\rho_0 H} & -\frac{g}{\theta_0}
-            & 0 & -i\omega
-    \end{bmatrix} \begin{bmatrix}
-        P_1 \\ \theta_1 \\ u_{1x} \\ u_{1z}
-    \end{bmatrix} = 0.
-\end{equation}
-Taking the determinant of this matrix produces
-\begin{align}
-    -k_x^2\p*{-N^2 + \omega^2} + \p*{ik_z - \frac{1}{H}}
-        \p*{ik_z + \frac{1}{H}}\omega^2 &= 0,\nonumber\\
-    \frac{N^2k_x^2}{k_x^2 + k_z^2 + \frac{1}{4H^2}} &= \omega^2.
-\end{align}
-
-\chapter{Numerical Results}
-
-\section{Chebyshev Polynomials}\label{s:chebyshev}
-
-Note that, unlike a Fourier basis, a Chebyshev basis has $N^2$ grid spacing near
-its boundaries. By this, we mean that for the Chebyshev polynomials $T_n(x)$,
-the spacing between zeros of $T_n(x)$ near $x = \pm 1$ scales like $n^{-2}$.
-This contrasts with Fourier series for which $\psi_n(x) = \cos n\pi x, \sin n\pi
-x$ for which the zeros are spaced $\Delta x = \frac{1}{n}$. This is important
-because the CFL condition for PDEs requires that $u\frac{\Delta t}{\Delta x} <
-C$ for some constant $C$, otherwise the simulation will move fluid elements more
-than $\Delta x$ in a single timestep, an illegal operation for the basis
-function.
-
-If we consider instead $T_n(x)$ satisfying $T_n(\cos \theta) = \cos n\theta$,
-then within some spacing of the edge of the domain $\cos\theta = 1 - \Delta x$
-then $\Delta x \approx \theta^2/2$ then the number of zeros seen is the number
-of solutions $\cos \phi = 0, \phi = n\theta = n\sqrt{2\Delta x}$ which requires
-$\sim k\pi \in [0, n\sqrt{2\Delta x}]$. Thus, for any interval $\Delta x$, the
-number of zeros contained scales with $n^2$!
-
-Note: There exist papers on the Courant stability condition of Chebyshev
-spectral methods which cite exactly the $N^{-2}$ minimum timestep scaling we can
-infer from the above. But if our field is small near the boundaries and is only
-interesting in a region isolated from the boundary (e.g.\ with our damping
-zones) then the bound used in the papers scales more like $N^{-1}$, which is
-much more desirable.
-
-\section{Algebraic Manipulations}
-
-\subsection{Factoring out the exponential dependence}
-
-I've had to do this algebra too many times now, so let's just consider if we
-factor out the $e^{z/2H}$ exponential scaling on the dynamical variables and the
-$e^{-z/H}$ scaling on $\rho_0$. This leaves us with:
-\begin{subequations}\label{se:nonlin_factored}
-    \begin{align}
-        \vec{\nabla} \cdot \vec{u}_1 + \frac{u_{1z}}{2H} &= 0,\\
-        \pd{\rho_1}{t} - \frac{u_{1z} \rho_0}{H} &= -e^{z/2H}\s*{
-            \p*{\vec{u}_1 \cdot \vec{\nabla} \rho_1} + \frac{u_{1z}\rho_1}{2H}
-        },\\
-        \pd{\vec{u}_1}{t} + \frac{\vec{\nabla}P}{\rho_0}
-            + \frac{P}{2H\rho_0}\hat{z} + \frac{\rho_1\vec{g}}{\rho_0}
-            &= -e^{z/2H}\s*{
-            \p*{\vec{u} \cdot \vec{\nabla} \vec{u}_1}
-                + \frac{u_{1z}\vec{u}_1}{2H}
-        }.
-    \end{align}
-\end{subequations}
-I'm too lazy to re-notate my variables, so all of them are constant amplitude
-over $z$ now.
-
-\section{Hydraulic Jumps in Surface Waves}
-
-We originally got started on this since it seems like a dimensionally-restricted
-version of what we're seeing in simulations, but presently unclear whether
-really the case. Also, Faber's book for fluids \emph{Fluid Dynamics for
-Physicists} cites a critical breaking jump height of $1.3$ which is markedly
-false; papers such as Chanson's 2009 hydraulic jumps review in Euro.\ J.\ Mech\
-B/Fluids shows that undular jumps (instead of turbulent jumps) can be found
-easily up to $3.5/4$.
-
-\subsection{1D Hydraulic Jump}
-
-This is one of the OG papers; B\'elanger solved this as early as in 1838 but
-Rayleigh's work in 1916 is also pretty famous. Consider the 1D nonlinear Euler
-equations for a body of water whose surface can move $h(x)$, then give it
-velocity $u(x, z)$. The problem is really 2D, but we can integrate along the $z$
-direction, in which case $\pd{P(x, z)}{x} = \frac{\rho g h(x, z)}{x}$ simply
-integrates to $\frac{\rho g h(x)^2}{2} = \frac{g\Sigma(x)^2}{2\rho}$. Assuming
-also the velocity is uniform $u(x, z) = u(x)$ gives equations of motion
-\begin{align}
-    \pd{\Sigma}{t} + \pd{(\Sigma u)}{x} &= 0,\\
-    \pd{u}{t} + u\pd{u}{x} &=
-        -\frac{1}{\Sigma}\pd{}{x}\p*{\frac{g\Sigma^2}{2\rho}},\nonumber\\
-        &= -\frac{g}{\rho}\pd{\Sigma}{x}.
-\end{align}
-It proves easier to divide the density equation by $\rho$ so that we can replace
-$\frac{\Sigma}{\rho} = h(x)$ in the above. A specific momentum conservation
-equation can be obtained
-\begin{equation}
-    \pd{(hu)}{t} + \pd{}{x}\p*{hu^2 + \frac{gh^2}{2}} = 0.
-\end{equation}
-
-Now, consider a hydraulic jump $u_1, h_1 \to u_2, h_2$ such that the
-discontinuity is fixed (we've shifted to the jump front frame), so that
-$\s*{uh}, \s*{hu^2 + gh^2/2}$ bracketed quantities are continuous across the jump.
-This results in
-\begin{align}
-    h_1u_1^2 - h_2u_2^2 = h_1u_1^2\p*{1 - \frac{u_2}{u_1}}
-        &= h_1u_1^2\p*{1 - \frac{h_1}{h_2}},\nonumber\\
-        &= \frac{g\p*{h_2^2 - h_1^2}}{2},\nonumber\\
-    u_1^2 &= \frac{h_2}{h_1}\frac{g\p*{h_1 + h_2}}{2}.
-\end{align}
-Customarily one defines $a_1^2 = gh_1, \mathrm{Fr} =
-\frac{u_1}{a_1}$ and $\gamma = \frac{h_2}{h_1}$ which gives simplification
-\begin{align}
-    2\mathrm{Fr}^2 &= \gamma(1 + \gamma)\nonumber\\
-    \gamma = \frac{-1 + \sqrt{1 + 8\mathrm{Fr}^2}}{2}.
-\end{align}
-Note now that $\mathrm{Fr}^2 > 1$ produces a $\gamma > 1$, so $h_2 > h_1$.
-Noting moreover that $\frac{u_1}{u_2} = \gamma$ and so $u_1 > u_2$. If we have a
-picture of a jet incident on a medium at rest, $u_2$ is just the velocity of the
-jump.
-
-\subsection{1D Viscous Theory}
-
-We follow Bohr's paper from J.\ Fluid Mech.\ in 1993 (also done by Tani in 1949
-to some extent) but in 1D first so that the equations are a bit easier to
-intuit. It's completely unnecessary but illustrates some features of the
-approach.
-
-Consider the same 2D equations as before (before integrating along $z$), but now
-permit a viscosity term. In the boundary layer approximation, the usual
-assumptions are that flow largely varies horizontally but variations are
-stronger vertically (i.e.\ we only consider the $x$ velocity equation but only
-consider the $z$ viscosity). We make one small modification, it seems that we
-permit an incompressibility assumption since $\rd{h(x, z, t)}{t} = 0$; this is
-just $\Delta P = \rho g h = $ the Lagrangian pressure boundary condition. Thus,
-the equations of motion are
-\begin{align}
-    \pd{u_x}{x} + \pd{u_z}{z} &= 0,\\
-    u_x \pd{u_x}{x} + u_z \pd{u_x}{z} &= -g\pd{h}{x} + \nu \ptd{u_x}{z}.
-\end{align}
-Integration along $z$ of the momentum equation can be done with appropriate BCs:
-no-slip/penetration at the bottom $u_x(z = 0) = u_z(z = 0) = 0$, while no shear
-at the surface $\pd{u_x}{z}(z = h) = 0$. We handle the integral of $u_z u_{x,
-z}$ by integrating by parts, dropping the boundary term and continuity equation,
-giving
-\begin{equation}
-    2u_x \int\limits_0^h u_x\rd{u_x}{x}\;\mathrm{d}z
-        = -g\pd{h}{x}h - \nu\at{\pd{u_x}{z}}_{z = 0}.
-\end{equation}
-We make a few handwavy substitutions for representative values $\partial_z u_x(z
-= 0) \sim u_x / h$ and $\partial_x u_x^2 \sim \partial_x \bar{u}_x^2$ average
-value. This averaging procedure and dropping the factor of $2$ (the two papers
-seem to disagree on this coefficient, I think Bohr's paper has an error) lets
-us examine
-\begin{equation}
-    \bar{u}_x \pd{\bar{u}_x}{x} + g\pd{h}{x} \sim -\frac{\bar{u}_x}{h^2}.
-\end{equation}
-Next, let's identify $\bar{u}_xh \equiv q$ volume flux constant in $x$, which
-leaves us with
-\begin{equation}
-    \pd{\bar{u}_x}{x}\p*{\bar{u}_x - \frac{gq}{\bar{u}_x^2}}
-        \sim -\nu\frac{\bar{u}_x^3}{q^2}.
-\end{equation}
-Let's study in terms of $\mathrm{Fr}^2 = \frac{\bar{u}_x^2}{gh} =
-\frac{\bar{u}_x^3}{gq}$ again, albeit a bit messier, which takes a bit of work
-but eventually lets us rearrange
-\begin{equation}
-    \pd{\mathrm{Fr}}{x} = \frac{\p*{gq}^{1/3}\nu}{q^2}\frac{
-        \mathrm{Fr}^{5/3}}{1 - \mathrm{Fr}^2}.
-\end{equation}
-Note that the units work here, as $\frac{\p*{gq}^{1/3}\nu}{q^2} \equiv L_\nu$
-has units of length.
-
-Let's interpret the equation we have now: we have exactly how the inbound
-$\mathrm{Fr}\p*{x = -\infty}$ evolves over $x$, which should describe (extremely
-coarsely, owing to our approximations) the boundary layer of the jump.
-
-Qualitatively, we can use a parameterization $\mathrm{Fr}(s), x(s)$ such that
-$\dot{\mathrm{Fr}}(s) = -\mathrm{Fr}^{5/3}, \dot{x}(s) = L_\nu \p*{\mathrm{Fr}^2
-- 1}$ where we define $L_\nu \equiv a_0h^2/\nu$ and we've absorbed the negative
-sign into $\dot{\mathrm{Fr}}(s)$. I'm too lazy to plot this, but if we start at
-a point $\mathrm{Fr} > 1$, it's clear that $\mathrm{Fr}$ increases and $x$
-decreases until $\mathrm{Fr} = 1$. Since the sign convention on the parametric
-is arbitrary, we can flip the sign and see that the parameterized curve formally
-follows decreasing $\dot{\mathrm{Fr}}(s) \propto -1/s$ while $\dot{x}(s) \propto
--L$. This is double-valued with the solution $\mathrm{Fr} > 1$ though, so with
-some physical intuition we discard this and consider what values of
-$\mathrm{Fr}$ can take on beyond its intersection with $\mathrm{Fr} = 1$.
-
-It must discontinuously jump to another family of curves. But any nonzero value
-of $\mathrm{Fr}$ will live on another one of these such trajectories and
-increase without bound; indeed the only reasonable asymptotic solution is
-$\mathrm{Fr} = 0$. This completes the picture: for an inbound $\mathrm{Fr} \gg
-1$, it follows roughly $\pd{\mathrm{Fr}}{x} \gtrsim -\mathrm{Fr}^{-1/3}$ until
-it reaches $\mathrm{Fr} = 1$, then it discontinously disappears. This is the
-upstream flow, and so by mass conservation the location of the discontinuity
-must travel.
-
-This is exactly the picture we derived above: for an incoming $\mathrm{Fr}
-\gtrsim 1$, the upstream flow has zero velocity and the jump travels.
-
-\subsection{2D Hydraulic Jump}
-
-This is the case worked out in detail in Bohr's paper. Following much the same
-approach but in cylindrical coordinates we may write down equations of motion
-\begin{align}
-    u_r \pd{u_r}{r} + u_z \pd{u_r}{z} &= -g\pd{h}{r} + \nu
-        \ptd{u_r}{z},\nonumber\\
-    \frac{1}{r}\pd{(ru_r)}{r} + \pd{u_z}{z} &= 0.
-\end{align}
-Note the slight difference here due to the $\vec{\nabla} \cdot$ in cylindrical
-coordinates. Interestingly, as a consequence of this $\pd{P_r}{t} +
-\frac{1}{r}\partial_r \p*{hu_r^2 + gh^2/2} \neq 0$, since one term is a
-$\vec{\nabla} \cdot (hu_r)$ and the other is a $\vec{\nabla}gh^2/2$! Radial
-momentum is not conserved as the flow spreads out. Averaging using the same
-techniques as above gives
-\begin{align}
-    \bar{u}_r\pd{\bar{u}_r}{r} + g\pd{h}{r} &= -\nu \frac{\bar{u}_r}{h^2},\nonumber\\
-    \bar{u}_rhr &= q.
-\end{align}
-Combining the two gives
-\begin{align}
-    \pd{\bar{u}_r}{r}\p*{\bar{u}_r - \frac{gq}{\bar{u}_r^2r}}
-        &= \frac{gq}{\bar{u}_rr^2} - \nu\frac{\bar{u}_r^3r^2}{q^2}.
-        \label{eq:2d_hjump}
-\end{align}
-It's not very worth converting into dimensionless $\mathrm{Fr}$ number since it
-would carry an $r$ dependence (if we are to hold $q$ constant) and we need to
-study the $r$ dependence.
-
-It may be observed that \autoref{eq:2d_hjump} has only a single fixed point,
-$u_r^3r = gq, u_r^4r^4 = gq^3/\nu$. Call this point $r_c$ the critical radius,
-then it turns out that the two parametric solutions (as we analyzed in the
-previous section) both spiral around this point. This has a few important
-properties:
-\begin{itemize}
-    \item No asymptotic solutions: the solution for $r > r_c$ always diverges
-        at some finite $r_s < \infty$. We can try an approximate solution in the
-        form $u_r = y / r$; this is done in Bohr's paper.
-
-    \item For a fixed $r_s \gg r_c, < \infty$, we can require the usual jump
-        condition which takes the slightly modified form
-\end{itemize}
-
-Edit: Looks like Bhagat 2018 J.\ Fluid.\ Mech.\ does this more correctly, they
-find that surface tension is the proper term to invoke after performing this
-boundary layer analysis. I'm not going to finish reworking the rest of Bohr's
-paper here, since it's up through the boundary layer equations that is reusable.
 \end{document}

### 2018-11-12 writeups/notes.tex
typo
check in new notes, sync

@@
 For our stratified atmosphere problem, $\omega^2 = \frac{N^2 k_x^2}{k^2 +
-\frac{1}{4H^2}}$. This corresponds to $\vec{v}_{ph} = \frac{Nk_x}{\sqrt{k_x^2 +
-k_z^2 + 1/4H^2}} \vec{k}$.
+\frac{1}{4H^2}}$. This corresponds to $\vec{v}_{ph} = \frac{Nk_x}{\p*{k_x^2 +
+k_z^2 + 1/4H^2}^{3/2}} \vec{k}$.
 
@@
 paper here, since it's up through the boundary layer equations that is reusable.
+
+\clearpage
+
+\section{Paper Notes}
+
+\begin{itemize}
+    \item \emph{Lindzen 1968} A Theory of the QBO\@: IGW are excited from below
+        with a distribution of phase velocities $c \in \s*{-c_0, c_0}$. Then, if
+        you allow a shear layer to absorb a particular phase velocity, the mean
+        flow will propate downwards until it creates a discontinuity (because
+        there's no more inbound IGW at the correct $c > c_0$). Then, once the
+        ledge leaves the driveable zone, perhaps the flow flips (very roughly).
+
+        Note that follow-up paper Holton \& Lindzen 1972 attributes the mean
+        flow deposition to radiative cooling instead of critical layer
+        absorption, TODO investigate.
+
+    \item \emph{Jones 1970} Coupling IGW mean flow: numerical study: as IGW
+        steepens shelf, more transmission than linear steady state theory.
+
+    \item \emph{Couson \& Lecoanet 2018} Slowly-reversing Mean Flows: Convection
+        drives spectrum of waves that become IGW at transition, produces QBOs.
+
+        Key note: momentum deposition by viscous dissipation dominates critical
+        layer absorption in the WKB model?
+\end{itemize}
 \end{document}

### 2018-11-12 sims/2d_2_anel/strat_helper.py

@@ -72,44 +72,38 @@ def get_solver(params):
     problem.substitutions['sponge'] = 'SPONGE_STRENGTH * 0.5 * ' +\
         '(2 + tanh((z - SPONGE_HIGH) / (SPONGE_WIDTH * (ZMAX - SPONGE_HIGH))) - ' +\
         'tanh((z - SPONGE_LOW) / (SPONGE_WIDTH * (SPONGE_LOW))))'
-    # problem.substitutions['mask'] = '0.5 * ' +\
-    #     '(-tanh((z - 9.5 * H) /' + \
-    #         '(SPONGE_WIDTH * (ZMAX - 9.5 * H))) + ' +\
-    #     'tanh((z - (SPONGE_LOW + SPONGE_WIDTH)) /' + \
-    #         '(SPONGE_WIDTH * (SPONGE_LOW + SPONGE_WIDTH))))'
-    problem.substitutions['mask'] = '1'
     problem.substitutions['rho0'] = 'RHO0 * exp(-z / H)'
-    problem.substitutions['t_s'] = 'T_F / 10'
-    problem.add_equation('dx(ux) + uz_z = 0')
+    problem.add_equation('rho0 * dx(ux) + rho0 * uz_z - uz * rho0 / H = 0')
     problem.add_equation(
         'dt(rho) - rho0 * uz / H' +
         '- (NU * dx(dx(rho)) + NU * dz(rho_z))' +
-        ' = - sponge * rho - mask * (ux * dx(rho) + uz * dz(rho)) +' +
+        ' = - sponge * rho -' +
+        '(ux * dx(rho) + uz * dz(rho) + rho * (dx(ux) + uz_z)) +' +
         'F * exp(-(z - Z0)**2 / (2 * S**2)) *' +
             'cos(KX * x - OMEGA * t)')
     problem.add_equation(
         'dt(ux) + dx(P) / rho0' +
         '- (NU * dx(dx(ux)) + NU * dz(ux_z))' +
-        '= - sponge * ux - mask * (ux * dx(ux) + uz * dz(ux))')
+        '= - sponge * ux - (ux * dx(ux) + uz * dz(ux))')
     problem.add_equation(
         'dt(uz) + dz(P) / rho0 + rho * g / rho0' +
         '- (NU * dx(dx(uz)) + NU * dz(uz_z))' +
-        '= - sponge * uz - mask * (ux * dx(uz) + uz * dz(uz))')
[…]

### 2018-11-12 writeups/notes.tex
writeup simplified anelastic

@@
 
-\section{Traveling Breaking Front}
-
-Motivated by simulations, we posit the existence of a hydraulic jump, similar to
-shocks in nature. We thus need to develop jump conditions, similar to the
-Rankine-Huginot jump conditions for shocks.
-
-The R-H jump conditions take the conservation equations and integrate over a
-control volume. Since in a hydrodynamic shock, all fluid quantities are
-advected, all conservation equations are necessary. A simpler treatment is
-merited for IGW breaking, since very few quantities are advected (notably,
-vertical momentum and density are not).
-
-Let's consider first the advection of horizontal momentum $\rho_0 u_{1x}$ to
-leading order. This satisfies conservation equation
-\begin{equation}
-    \pd{P_x}{t} + \vec{\nabla} \cdot \p*{\rho_0 u_{1x}\vec{u}_1} = 0.
-\end{equation}
-Integrating over control volume $V = \s*{-\infty, \infty} \times \s*{z < z_b, z
-> z_b}$ the full horizontal domain times the breaking zone, we require the total
-horizontal momentum of the zone be constant in time, if we permit $z_b$ to move
-with velocity $u_b$ the breaking front velocity. This equates to
-\begin{equation}
-    \Delta P_x - \s*{\rho_0 u_{1x}u_b}_{z > z_b}
-        + \s*{\rho_0 u_{1x}\p*{u_{1z} - u_b}}_{z < z_b} = 0.
-\end{equation}
-We've called $\Delta P_x$ the change in horizontal momentum in the breaking zone
-(so the boundary integral over the endcaps of the domain, effectively). Now,
-based on simulations, we identify that $u_{1x}\p*{z > z_c} = B\cos \kappa z$
-seems to form a standing wave pattern. Note that since this is a pure shear
-flow, it is consistent with an incompressible fluid and $u_{1z}\p*{z > z_c} =
-0$.
-
-Permitting such a flow above the breaking zone, then we substitute our known
-linear solution $u_{1x} \approx \frac{k_z}{k_x} u_{1z} \propto Ae^{z / 2H}\cos
-\omega_0 t$ and $\rho_0 = \tilde{\rho}e^{-z/H}$ and evaluate harmonics of $\cos
-\omega_0 t$ to obtain
-\begin{align}
-    \Delta P_x\p*{\omega = 0} &= \tilde{\rho}\frac{A^2}{2}\frac{k_z}{k_x},\\
-    \Delta P_x\p*{\omega = \omega_0}
-        &= u_b Ae^{-z/2H}\tilde{\rho}\cos \omega_0 t,\\
-    \Delta P_k(\omega = 2\omega_0) &=
-        -u_b \cos (\kappa u_bt)
-            + \frac{k_z}{k_x}\frac{A^2}{2B}e^{z/H} \cos 2 \omega_0 t.
-\end{align}
-We can easily interpret $\Delta P_x(0)$ to be the deposition of the horizontal
-momentum of the wave, and $\Delta P_x\p*{\omega = \omega_0}$ to be just the wave
-flowing through the deposition zone; it time averages to zero anyways, so it's
-not super interesting.
-
-Instead, if we dare make identification $u_b = \frac{2\omega_0}{\kappa}$ (the
-only way $u_b \neq 0$), then the deposition to
-\begin{equation}
-    \Delta P_k\p*{\omega = 2\omega_0} = \p*{\frac{A^2}{2B}e^{z/H} -
-        \frac{2\omega_0}{\kappa}}\cos 2\omega_0 t,
-\end{equation}
-i.e.\ the standing wave pattern above can sap some $2\omega_0$ frequency $P_x$
-out of the mean flow via a breaking front.
-
-\subsection{Finding and Explaining the Breaking Front}
-
-It is a bit tricky to interpret this, so I'll just jot down some ideas lest I
-forget them in my sleep-deprived state. If we require that the breaking front
-only sap momentum from the wave (and not from the mean flow), then the breaking
-front shuts of when $\frac{A^2}{2B}e^{z/H} \sim \frac{2\omega_0}{\kappa}$. It's
-also unclear how $\kappa$ is chosen as well. But based on preliminary
-simulations, this is actually pretty accurate for observed $\kappa, u_b$!
-
-Encouragingly, we can guess at where $\kappa$ comes from. Since the waveform
-above the breaking zone has to lose its $k_x$ component, we may guess that the
-incoming wave $\cos \p*{k_xx + k_zz}$ must interact nonlinearly  with another
-wave to knock out the $k_xx$ contribution. One possible candidate would have a
-$\cos \p*{-k_xx + k_zz}$ such that their product $\cos 2k_zz$, in which case
-$\kappa = 2k_z$. This also seems to be reasonable from a few of our simulations,
-but this requires thus that $2k_z$ be much longer than the viscous length scale;
-this would explain why we don't see it on our low-resolution simulations at $z
-\in [0, 12H], N_Z = 512$ but only on $N_Z = 1024$. The desired $\cos \p*{-k_xx +
-k_zz}$ is just a reflected wave in the $x$ direction, so I guess it can come
-from the viscous subscales pushing against the horizontal momentum/mean flow?
-In any case it's not inconceivable it would exist, and it should be the dominant
-$-k_xx$ term compared e.g.\ to $-k_xx + 0z, -k_xx + 2k_zz$.
-
-TODO We need simulations next of the front breaking down, to understand where it
-would stop, since we don't have a criterion for it yet.
-
 \chapter{Cylindrical Stratified Flow}
@@
 
+\chapter{Anelastic Fluid Equations}
+
+\section{Inadequacy of Incompressible Stratified Equations}
+
+To perform simulations up to $u \sim c_p \ll c_s$ and include the advective
+terms but no sound-speed-order terms, we need to make sure our equations
+qualitatively match the fully nonlinear (including sound speed) equations.
+In a stratified atmosphere, the incompressible equations are qualitatively
+wrong. The full fluid equations consist of conservation equations plus an
+equation of state, so our reduced equations must conserve the same quantities.
+The equations which we simulated above, the incompressible nonlinear equations,
+must themselves also conserve the same quantities.
+
+Our system, which drops its energy conservation equation, consists of momentum
+conservation, mass conservation and an incompressible equation of state
+$\rd{\rho}{t} = 0$. Note however that the horizontally averaged horizontal
+momentum $\ev*{\rho u_{1x}} = \ev*{\rho_0 u_{1x}}$ is \emph{not} conserved by
+our above equations:
+\begin{equation}
+    \rho_0 \pd{u_{1x}}{t} + \rho_0 \p*{\vec{u}_1 \cdot \vec{\nabla}} u_{1x}
+        + \vec{\nabla}P_1 = 0.
+\end{equation}
+Since $\pd{\rho_0}{t} = 0$, we may try to put this into flux conservative form
+via
+\begin{equation}
+    \pd{(\rho_0 u_{1x})}{t} + \vec{\nabla} \cdot
+        \p*{\rho_0 \vec{u}_1 u_{1x} + \bm{P}_1}
+        = u_{1x} \vec{\nabla} \cdot \p*{\rho_0 \vec{u}_1}.
+\end{equation}
+We combine some scalar gradients into tensors, but it's clear that there is an
+extra term that \emph{is not} cancelled by the usual incompressibility criterion
+$\vec{\nabla} \cdot \vec{u}_1 = 0$. It bears noting this term only cancels
+\emph{in the Boussinesq approximation} where $\rho_0$ is held constant.
+
+\section{Anelastic Fluid Equations}
+
+The anelastic fluid equations replace incompressibility with the obvious
+$\vec{\nabla} \cdot (\rho_0\vec{u}_1) = 0$ to conserve momentum $\rho_0 u_{1x}$
+exactly. There are however many flavors of anelastic approximations however, the
+large majority of which reintroduce a temperature/entropy variable. This
+probably comes about because most uses of the anelastic equations examine
+convective systems, for which thermal diffusivity is important. For us, where
+all fluid motion is fairly local, it seems reasonable to only permit fluid
+advection and simplify our model, so we can stay with only two state variables
+$P, \rho$ along with the flow $\vec{u}$. We still use numerical viscosities, but
+these are more so a subgrid model and not physical, so we will just reintroduce
+them before simulating.
+
+From here, it's a simple matter of taking limits of existing anelastic
+approximations. We will start with the full system from Glatzmaier's book:
+\begin{align}
+    0 &= \vec{\nabla} \cdot \p*{\rho_0 \vec{u}_1},\\
+    \rho_1 &= \frac{\rho_0}{\gamma P_0}P_1 - \frac{\rho_0}{c_p}S_1
+        = \at{\pd{\rho_0}{P_0}}_S P_1 + \at{\pd{\rho_0}{S_0}}_P S_1,\\
+    N^2 &= \frac{g}{c_p}\rd{S_0}{z},\\
+    0 &= \pd{\vec{u}_1}{t} + \p*{\vec{u}_1 \cdot \vec{\nabla}}\vec{u}_1
+        + \vec{\nabla}\p*{\frac{P_1}{\rho_0} + \Phi_1}
+        + \at{\pd{\rho_0}{S_0}}P \frac{g}{\rho_0}S\hat{z}
+        - \nu \p*{\nabla^2 \vec{u}_1 + \frac{1}{3}\vec{\nabla}\p*{\vec{\nabla}
+            \cdot \vec{u}_1}},\\
+    0 &= \pd{S_1}{t} + \p*{\vec{u}_1 \cdot \vec{\nabla}}S_1
+        + u_{1z}\rd{S_0}{z}
+        - \frac{1}{\rho_0 T_0} \vec{\nabla} \cdot \p*{\kappa \rho_0 T_0
+            \vec{\nabla} S_1}.
+\end{align}
+We obviously take $\Phi_1 = 0$ the Cowling approximation. Of course, in our
+problem, we may drop all the viscous terms from the momentum equation and energy
+equation, which immediately leaves us with
+\begin{align}
+    \vec{\nabla} \cdot \p*{\rho_0 \vec{u}_1} &= 0,\\
+    \pd{\vec{u}_1}{t} + \p*{\vec{u}_1 \cdot \vec{\nabla}}\vec{u}_1
+        + \vec{\nabla} \p*{\frac{P_1}{\rho_0}} - \frac{g}{c_p}S_1 \hat{z}
+            &= 0,\\
+    \pd{S_1}{t} + \p*{\vec{u}_1 \cdot \vec{\nabla}}S_1
+        &= 0.
+\end{align}
+Note that when accounting for $N^2 = \frac{g}{c_p}\rd{S_0}{z}$ so $c_p = S_0$
+immediately gives us Sutherland's equations from his 2011 paper.
+
+We now make one further approximation: we have already taken the low Mach number
+approximation, so we will approximate $\rho_1 = \frac{\rho_0}{\gamma P_0}P_1 -
+\frac{\rho_0}{S_0}S_1 \approx -\frac{\rho_0}{S_0}S_1$, which allows us to
+eliminate $S$. We must be careful in taking this approximation (and indeed, it
+may be unjustified), since the difference term $\vec{\nabla}\frac{P_1}{\rho_0} -
+\frac{\vec{\nabla}P_1}{\rho_0} = -\frac{P_1 \vec{\nabla}\rho_0}{\rho_0^2}$ is
+now small as well. What to do with it is dictated by the requirement that
+horizontal momentum be conserved. With this in mind, the momentum equation
+must be
+\begin{equation}
+    \pd{\vec{u}_1}{t} + \p*{\vec{u}_1 \cdot \vec{\nabla}}\vec{u}_1
+        + \frac{\vec{\nabla}P_1}{\rho_0} + \frac{\rho_1}{\rho_0}g\hat{z} = 0.
+\end{equation}
+We may verify that horizontal momentum is conserved by multiplying through with
+$\rho_0$ again.
+
+The entropy equation takes a bit more work, but we may show that
+\begin{align}
+    -\frac{c_p}{\rho_0}\pd{\rho_1}{t} - c_p \vec{u}_1 \cdot
+        \vec{\nabla}\p*{\frac{\rho_1}{\rho_0}} - \frac{u_1 c_p N^2}{g} &= 0,\\
+    \pd{\rho_1}{t} + \vec{u}_1 \cdot \vec{\nabla}\rho_1
+        - \frac{\rho_1\vec{u}_1 \cdot \vec{\nabla}\rho_0}{\rho_0}
+        - \frac{u_{1z}\rho_0}{H} &= 0,\nonumber\\
+    \pd{\rho_1}{t} + \vec{\nabla}\p*{\rho_1 \vec{u}_1} - \frac{u_{1z}\rho_0}{H}
+        &= 0.
+\end{align}
+We have used $\vec{\nabla} \cdot \p*{\rho_0 \vec{u}_1} = 0$.
+
 \clearpage
@@
 
-\section{The Anelastic/Boussinesq Approximations}
+\section{The Anelastic/Boussinesq Approximations (old)}
 
@@
 
+This seems to be largely dated now, we believe that the sharp jump in the
+nonlinear simulations we are seeing are mean flow interactions, rather than
+hydraulic jumps.
+
+\subsection{Traveling Breaking Front}
+
+Motivated by simulations, we posit the existence of a hydraulic jump, similar to
+shocks in nature. We thus need to develop jump conditions, similar to the
+Rankine-Huginot jump conditions for shocks.
+
+The R-H jump conditions take the conservation equations and integrate over a
+control volume. Since in a hydrodynamic shock, all fluid quantities are
+advected, all conservation equations are necessary. A simpler treatment is
+merited for IGW breaking, since very few quantities are advected (notably,
+vertical momentum and density are not).
+
+Let's consider first the advection of horizontal momentum $\rho_0 u_{1x}$ to
+leading order. This satisfies conservation equation
+\begin{equation}
+    \pd{P_x}{t} + \vec{\nabla} \cdot \p*{\rho_0 u_{1x}\vec{u}_1} = 0.
+\end{equation}
+Integrating over control volume $V = \s*{-\infty, \infty} \times \s*{z < z_b, z
+> z_b}$ the full horizontal domain times the breaking zone, we require the total
+horizontal momentum of the zone be constant in time, if we permit $z_b$ to move
+with velocity $u_b$ the breaking front velocity. This equates to
+\begin{equation}
+    \Delta P_x - \s*{\rho_0 u_{1x}u_b}_{z > z_b}
+        + \s*{\rho_0 u_{1x}\p*{u_{1z} - u_b}}_{z < z_b} = 0.
+\end{equation}
+We've called $\Delta P_x$ the change in horizontal momentum in the breaking zone
+(so the boundary integral over the endcaps of the domain, effectively). Now,
+based on simulations, we identify that $u_{1x}\p*{z > z_c} = B\cos \kappa z$
+seems to form a standing wave pattern. Note that since this is a pure shear
+flow, it is consistent with an incompressible fluid and $u_{1z}\p*{z > z_c} =
+0$.
+
+Permitting such a flow above the breaking zone, then we substitute our known
+linear solution $u_{1x} \approx \frac{k_z}{k_x} u_{1z} \propto Ae^{z / 2H}\cos
+\omega_0 t$ and $\rho_0 = \tilde{\rho}e^{-z/H}$ and evaluate harmonics of $\cos
+\omega_0 t$ to obtain
+\begin{align}
+    \Delta P_x\p*{\omega = 0} &= \tilde{\rho}\frac{A^2}{2}\frac{k_z}{k_x},\\
+    \Delta P_x\p*{\omega = \omega_0}
+        &= u_b Ae^{-z/2H}\tilde{\rho}\cos \omega_0 t,\\
+    \Delta P_k(\omega = 2\omega_0) &=
+        -u_b \cos (\kappa u_bt)
+            + \frac{k_z}{k_x}\frac{A^2}{2B}e^{z/H} \cos 2 \omega_0 t.
+\end{align}
+We can easily interpret $\Delta P_x(0)$ to be the deposition of the horizontal
+momentum of the wave, and $\Delta P_x\p*{\omega = \omega_0}$ to be just the wave
+flowing through the deposition zone; it time averages to zero anyways, so it's
+not super interesting.
+
+Instead, if we dare make identification $u_b = \frac{2\omega_0}{\kappa}$ (the
+only way $u_b \neq 0$), then the deposition to
+\begin{equation}
+    \Delta P_k\p*{\omega = 2\omega_0} = \p*{\frac{A^2}{2B}e^{z/H} -
+        \frac{2\omega_0}{\kappa}}\cos 2\omega_0 t,
+\end{equation}
+i.e.\ the standing wave pattern above can sap some $2\omega_0$ frequency $P_x$
+out of the mean flow via a breaking front.
+
+\subsection{Finding and Explaining the Breaking Front}
+
+It is a bit tricky to interpret this, so I'll just jot down some ideas lest I
+forget them in my sleep-deprived state. If we require that the breaking front
+only sap momentum from the wave (and not from the mean flow), then the breaking
+front shuts of when $\frac{A^2}{2B}e^{z/H} \sim \frac{2\omega_0}{\kappa}$. It's
+also unclear how $\kappa$ is chosen as well. But based on preliminary
+simulations, this is actually pretty accurate for observed $\kappa, u_b$!
+
+Encouragingly, we can guess at where $\kappa$ comes from. Since the waveform
+above the breaking zone has to lose its $k_x$ component, we may guess that the
+incoming wave $\cos \p*{k_xx + k_zz}$ must interact nonlinearly  with another
+wave to knock out the $k_xx$ contribution. One possible candidate would have a
+$\cos \p*{-k_xx + k_zz}$ such that their product $\cos 2k_zz$, in which case
+$\kappa = 2k_z$. This also seems to be reasonable from a few of our simulations,
+but this requires thus that $2k_z$ be much longer than the viscous length scale;
+this would explain why we don't see it on our low-resolution simulations at $z
+\in [0, 12H], N_Z = 512$ but only on $N_Z = 1024$. The desired $\cos \p*{-k_xx +
+k_zz}$ is just a reflected wave in the $x$ direction, so I guess it can come
+from the viscous subscales pushing against the horizontal momentum/mean flow?
+In any case it's not inconceivable it would exist, and it should be the dominant
+$-k_xx$ term compared e.g.\ to $-k_xx + 0z, -k_xx + 2k_zz$.
+
+TODO We need simulations next of the front breaking down, to understand where it
+would stop, since we don't have a criterion for it yet.
+
 We originally got started on this since it seems like a dimensionally-restricted

### 2018-11-12 writeups/notes.tex
WIP

@@
 \begin{equation}
-    \rtd{u_{1z}}{z} + \frac{1}{\rho_0}\pd{\rho_0}{z}\pd{u_{1z}}{z}
+    \ptd{u_{1z}}{z} + \frac{1}{\rho_0}\pd{\rho_0}{z}\pd{u_{1z}}{z}
         + k_x^2\p*{\frac{N^2}{\omega^2} - 1}u_{1z} = 0.\label{eq:lin.incomp.gen}
@@
 
-Let's now pick stratification $\rho \propto e^{-z/H}$
+Let's now pick stratification $\rho_0 \propto e^{-z/H}$
 \autoref{eq:lin.incomp.gen} clearly has exponential solutions $e^{\kappa z}$ for
@@
             &= 0,\\
-    \pd{S_1}{t} + \p*{\vec{u}_1 \cdot \vec{\nabla}}S_1
+    \pd{S_1}{t} + \p*{\vec{u}_1 \cdot \vec{\nabla}}(S_0 + S_1)
         &= 0.
@@
 \frac{\rho_0}{S_0}S_1 \approx -\frac{\rho_0}{S_0}S_1$, which allows us to
-eliminate $S$. We must be careful in taking this approximation (and indeed, it
-may be unjustified), since the difference term $\vec{\nabla}\frac{P_1}{\rho_0} -
-\frac{\vec{\nabla}P_1}{\rho_0} = -\frac{P_1 \vec{\nabla}\rho_0}{\rho_0^2}$ is
-now small as well. What to do with it is dictated by the requirement that
-horizontal momentum be conserved. With this in mind, the momentum equation
-must be
+eliminate $S$. With this in mind, the momentum equation can be rewritten
 \begin{equation}
     \pd{\vec{u}_1}{t} + \p*{\vec{u}_1 \cdot \vec{\nabla}}\vec{u}_1
-        + \frac{\vec{\nabla}P_1}{\rho_0} + \frac{\rho_1}{\rho_0}g\hat{z} = 0.
+        + \vec{\nabla}\p*{\frac{P_1}{\rho_0}}
+        + \frac{\rho_1}{\rho_0}g\hat{z} = 0.
 \end{equation}
@@
 
-\section{The Anelastic/Boussinesq Approximations (old)}
+\section{The Anelastic/Boussinesq Approximations}
 
 \subsection{Developing the Anelastic/Boussinesq Approximations}
+
+This follows a step by step derivation of the usual anelastic approximation that
+is used in atmospheric sciences. It's not the most careful treatment but should
+give some intuition as to how the equation of state is handled.
 
@@
 process $P\rho^{-\gamma} \propto P^{1-\gamma}T^\gamma$ is constant. We thus
-introduce the concept of the \emph{potential temperature}
+introduce the concept of the \emph{potential temperature} (just the entropy)
 \begin{equation}
@@
     \begin{split}
-        \pd{1}{\rho_0}\pd{\rho_0}{z} &= \frac{1}{\gamma P_0}\pd{P_0}{z}
+        \frac{1}{\rho_0}\pd{\rho_0}{z} &= \frac{1}{\gamma P_0}\pd{P_0}{z}
             - \frac{1}{\theta_0}\pd{\theta_0}{z},\\
@@
     \begin{split}
-        \vec{\nabla} \cdot \p*{\rho_0\vec{u}} &= 0,\\
+        \vec{\nabla} \cdot \p*{\rho_0\vec{u}_1} &= 0,\\
         \pd{\vec{u}_1}{t} + \vec{\nabla}\p*{\frac{P_1}{\rho_0}}

### 2018-11-18 presentations/Su_DL_110618/Su_DL_110618.tex
reorg presentations dir

--- 2018-11-05 presentations/Su_DL_110618.tex
@@
 
-        \item Two hypotheses:
+        \item Tried masking nonlinear terms near damping zones,
+            \lstinline{nl3_lowres_test.mp4}. Still mean flow interaction!
+    \end{itemize}
+\end{frame}
+
+\begin{frame}
+    \frametitle{Numerical Simulations, No Shear}
+    \framesubtitle{Hypotheses}
     \begin{itemize}
         \item Damping layer is causing mean flow to begin building up
-                    like Daniel said; how can we fix? Seems like low $k_z$ modes
-                    (which have higher $v_{ph}$).
+            like Daniel said; is removing nonlinear terms near damping layer a
+            good idea?
 
@@
             $\pd{k_z}{z} \sim \frac{k_z}{\lambda_z}$.
-            \end{itemize}
+
+            The sign of the flow is
+            wrong though, $\sim -\frac{1}{c_{gz}}\ev*{u_xu_z}_x$!
     \end{itemize}

### 2018-11-18 writeups/notes.tex
Stash

@@
 
+\chapter{State of the Art}
+
+While the succeeding content is largely chronological, this chapter will be
+devoted to a description of the current state-of-the-art.
+
+\section{2D Cartesian, Pseudo-Anelastic Atmosphere}
+
+Our equations are:
+\begin{subequations}\label{se:base.pseudo-anel}
+    \begin{align}
+        \vec{\nabla} \cdot \vec{u} &= 0,\\
+        \pd{\rho_1}{t} - \frac{u_{1z} \rho_0}{H}
+            + \vec{u} \cdot \vec{\nabla}\rho_1 &= 0,\\
+        \pd{\vec{u}_1}{t} + \p*{\vec{u} \cdot \vec{\nabla}}\vec{u}_1
+            + \frac{\vec{\nabla}P_1}{\rho_0} + \frac{\rho_1 \vec{g}}{\rho_0}
+            - \frac{\vec{\nabla}P_1}{\rho_0}\rho_1 &= 0.
+    \end{align}
+\end{subequations}
+Note that we permit background stratification $-\pd{\ln \rho_0}{z} = H$ but no
+background shear flow $\vec{u}_0$, so we write $\vec{u}_1 = \vec{u}$ without
+ambiguity.
+
+\section{2D Cartesian, Boussinesq Local Simulations}
+
+Our equations are the same:
+\begin{subequations}\label{se:base.bouss}
+    \begin{align}
+        \vec{\nabla} \cdot \vec{u} &= 0,\\
+        \pd{\rho_1}{t} - \frac{u_{1z} \rho_0}{H}
+            + \vec{u} \cdot \vec{\nabla}\rho_1 &= 0,\\
+        \pd{\vec{u}_1}{t} + \p*{\vec{u} \cdot \vec{\nabla}}\vec{u}_1
+            + \frac{\vec{\nabla}P_1}{\rho_0} + \frac{\rho_1 \vec{g}}{\rho_0}
+            - \frac{\vec{\nabla}P_1}{\rho_0}\rho_1 &= 0.
+    \end{align}
+\end{subequations}
+The only difference is that we take $\rho_0(z) = \rho_0$, effectively assuming
+$H$ is much larger than the vertical extent of the domain of simulation.
+
+\clearpage
+
 \chapter{Preliminary Problems}

### 2018-11-18 sims/2d_2_strat/strat_helper.py

@@ -143,6 +143,8 @@ def run_strat_sim(set_ICs, name, params):
     logger.info('Starting sim...')
     while solver.ok:
         cfl_dt = cfl.compute_dt() if params.get('USE_CFL') else params['DT']
+        # if cfl_dt < params['DT'] / 8: # small step sizes if strongly cfl limited
+        #     cfl_dt /= 2
         solver.step(cfl_dt)
         curr_iter = solver.iteration
 
@@ -303,7 +305,6 @@ def plot(name, params):
     z = domain.grid(1, scales=params['INTERP_Z'])[: , z_b:]
     xmesh, zmesh = quad_mesh(x=x[:, 0], y=z[0])
     x2mesh, z2mesh = quad_mesh(x=np.arange(params['N_X'] // 2), y=z[0])
-    z_pts = np.array((zmesh[1:, 0] + zmesh[:-1, 0]) / 2)
 
     # preprocess
     for var in dyn_vars + ['F_px']:
@@ -365,6 +366,7 @@ def plot(name, params):
             for var in z_vars + slice_vars:
                 axes = fig.add_subplot(n_rows, n_cols, idx, title=var)
                 var_dat = state_vars[var][:, z_b:]
+                z_pts = (zmesh[1:, 0] + zmesh[:-1, 0]) / 2
                 p = axes.plot(var_dat[t_idx],
                               z_pts,
                               linewidth=0.5)

### 2018-11-23 writeups/running_writeup.tex
updated writeups

    \documentclass[11pt,
        usenames, % allows access to some tikz colors
        dvipsnames % more colors: https://en.wikibooks.org/wiki/LaTeX/Colors
    ]{article}
    \usepackage{
        amsmath,
        amssymb,
        fouriernc, % fourier font w/ new century book
        fancyhdr, % page styling
        lastpage, % footer fanciness
        hyperref, % various links
        setspace, % line spacing
        amsthm, % newtheorem and proof environment
        mathtools, % \Aboxed for boxing inside aligns, among others
        float, % Allow [H] figure env alignment
        enumerate, % Allow custom enumerate numbering
        graphicx, % allow includegraphics with more filetypes
        wasysym, % \smiley!
        upgreek, % \upmu for \mum macro
        listings, % writing TrueType fonts and including code prettily
        tikz, % drawing things
        booktabs, % \bottomrule instead of hline apparently
        cancel % can cancel things out!
    }
    \usepackage[margin=1in]{geometry} % page geometry
    \usepackage[
        labelfont=bf, % caption names are labeled in bold
        font=scriptsize % smaller font for captions
    ]{caption}
    \usepackage[font=scriptsize]{subcaption} % subfigures

    \newcommand*{\scinot}[2]{#1\times10^{#2}}
    \newcommand*{\dotp}[2]{\left<#1\,\middle|\,#2\right>}
    \newcommand*{\rd}[2]{\frac{\mathrm{d}#1}{\mathrm{d}#2}}
    \newcommand*{\pd}[2]{\frac{\partial#1}{\partial#2}}
    \newcommand*{\rtd}[2]{\frac{\mathrm{d}^2#1}{\mathrm{d}#2^2}}
    \newcommand*{\ptd}[2]{\frac{\partial^2 #1}{\partial#2^2}}
    \newcommand*{\md}[2]{\frac{\mathrm{D}#1}{\mathrm{D}#2}}
    \newcommand*{\pvec}[1]{\vec{#1}^{\,\prime}}
    \newcommand*{\svec}[1]{\vec{#1}\;\!}
    \newcommand*{\bm}[1]{\boldsymbol{\mathbf{#1}}}
    \newcommand*{\ang}[0]{\;\text{\AA}}
    \newcommand*{\mum}[0]{\;\upmu \mathrm{m}}
    \newcommand*{\at}[1]{\left.#1\right|}

    \newtheorem{theorem}{Theorem}[section]

    \let\Re\undefined
    \let\Im\undefined
    \DeclareMathOperator{\Res}{Res}
    \DeclareMathOperator{\Re}{Re}
    \DeclareMathOperator{\Im}{Im}
    \DeclareMathOperator{\Log}{Log}
    \DeclareMathOperator{\Arg}{Arg}
    \DeclareMathOperator{\Tr}{Tr}
    \DeclareMathOperator{\E}{E}
    \DeclareMathOperator{\Var}{Var}
    \DeclareMathOperator*{\argmin}{argmin}
    \DeclareMathOperator*{\argmax}{argmax}
    \DeclareMathOperator{\sgn}{sgn}
    \DeclareMathOperator{\diag}{diag\;}

    \DeclarePairedDelimiter\bra{\langle}{\rvert}
    \DeclarePairedDelimiter\ket{\lvert}{\rangle}
    \DeclarePairedDelimiter\abs{\lvert}{\rvert}
    \DeclarePairedDelimiter\ev{\langle}{\rangle}
    \DeclarePairedDelimiter\p{\lparen}{\rparen}
    \DeclarePairedDelimiter\s{\lbrack}{\rbrack}
    \DeclarePairedDelimiter\z{\lbrace}{\rbrace}

    % \everymath{\displaystyle} % biggify limits of inline sums and integrals
    \tikzstyle{circ} % usage: \node[circ, placement] (label) {text};
        = [draw, circle, fill=white, node distance=3cm, minimum height=2em]
    \definecolor{commentgreen}{rgb}{0,0.6,0}
    \lstset{
        basicstyle=\ttfamily\footnotesize,
        frame=single,
        numbers=left,
        showstringspaces=false,
        keywordstyle=\color{blue},
        stringstyle=\color{purple},
        commentstyle=\color{commentgreen},
        morecomment=[l][\color{magenta}]{\#}
    }

\begin{document}

\def\Snospace~{\S{}} % hack to remove the space left after autorefs
\renewcommand*{\sectionautorefname}{\Snospace}
\renewcommand*{\appendixautorefname}{\Snospace}
\renewcommand*{\figureautorefname}{Fig.}
\renewcommand*{\equationautorefname}{Eq.}
\renewcommand*{\tableautorefname}{Tab.}

\onehalfspacing

\pagestyle{fancy}
\rfoot{Yubo Su}
\rhead{}
\cfoot{\thepage/\pageref{LastPage}}

\section{Equations}

We begin with the fully compressible fluid equations and general equation of
state
\begin{subequations}\label{se:comp}
    \begin{align}
        \rd{\rho}{t} + \rho \p*{\vec{\nabla} \cdot \vec{u}} &= 0,
            \label{eq:density}\\
        \rd{S}{t} &= 0,\label{eq:entropy}\\
        \rd{\vec{u}}{t} + \frac{\vec{\nabla}P}{\rho} + \vec{g} &= 0,
            \label{eq:velocity}\\
        P &= P(\rho, S).\label{eq:eos}
    \end{align}
\end{subequations}
We notate $\rd{}{t} = \pd{}{t} + \p*{\vec{u} \cdot \vec{\nabla}}$. Note this is
$5$ equations (in 2D) for 5 variables $(\rho, P, S, \vec{u})$ and so is properly
closed.

\subsection{Current Implementation}

We assume $P(\rho, S) = P(\rho)$ a barotropic equation of state, such that $S$
is entirely decoupled from the dynamical equations. We thus drop
\autoref{eq:entropy} of the full fluid equations \autoref{se:comp}. Then, we
will decompose dynamical variables into background and fluctuation quantities $P
= P_0 + P_1, \rho = \rho_0 + \rho_1$, where background quantities are
time-independent. Assuming no background velocities, we write $\vec{u}$ to
denote the fluctuation velocity with no ambiguity.

Furthermore, we will implement incompressibility by mandating
$\at{\pd{P(\rho)}{\rho}}_{ad} \to \infty$ adiabatic derivative. This forces
$\Delta P \gg \Delta \rho$ within a displaced fluid parcel, or $\rd{\rho}{t} =
0$ comoving derivative. This allows us to rewrite \autoref{eq:density}, and so
we arrive at the system of equations
\begin{subequations}\label{se:pseudo_mid}
    \begin{align}
        \vec{\nabla} \cdot \vec{u} &= 0, \label{eq:mid_density}\\
        \rd{\vec{u}}{t} + \frac{\vec{\nabla}P}{\rho} + \vec{g} &= 0,
            \label{eq:mid_velocity}\\
        \rd{\rho}{t} = \rd{\rho_1}{t} + u_z \rd{\rho_0}{z} &= 0.
            \label{eq:mid_eos}
    \end{align}
\end{subequations}

We now assume $P_0 = \rho_0 c_s^2 \propto e^{-z/H}$ isothermal exponential
stratification. At hydrostatic equilibrium, \autoref{eq:mid_velocity} forces
$\vec{\nabla}P_0 = -\rho_0 \vec{g}$. We next claim that $\frac{\rho_1}{\rho_0}
\sim \p*{\frac{u}{c_s}}^2 = \mathrm{Ma}^2 \ll 1$\footnote{I guess this doesn't
have to be true if $\frac{P_0}{\rho_0} = c_s^2 \neq \at{\pd{P}{\rho}}_{ad}$
which is the velocity we are sending to infinity.}. This may be verified to be
true by looking briefly at the momentum equation: $\omega u +
k\p*{\frac{P}{\rho} - \frac{P_0}{\rho_0}} = 0$, which then dividing through by
$P_0 / \rho_0 = c_s^2$ shows that $\frac{P/P_0}{\rho/\rho_0} - 1 \sim
\mathrm{Ma}^2$.

Given this, we expand \autoref{eq:mid_velocity} by $\frac{\vec{\nabla}P}{\rho}
\approx \frac{\vec{\nabla}P_0}{\rho_0} + \frac{\vec{\nabla}P_1}{\rho_0} -
\frac{\rho_1 \vec{\nabla}P_0}{\rho_0^2} + \dots$ and obtain the full system of
equations I have been using
\begin{subequations}\label{se:pseudo_fin}
    \begin{align}
        \vec{\nabla} \cdot \vec{u} &= 0, \label{eq:pseudo_density}\\
        \pd{\vec{u}}{t} + \p*{\vec{u} \cdot \vec{\nabla}}\vec{u}
            + \frac{\vec{\nabla}P_1}{\rho_0}
            + \frac{\rho_1 \vec{g}}{\vec{_0}} - \frac{\rho_1 \vec{\nabla}P_1}{
                \rho_0^2} &= 0,
            \label{eq:pseudo_velocity}\\
        \pd{\rho_1}{t} + \p*{\vec{u} \cdot \vec{\nabla}}\rho_1
            - \frac{u_z \rho_0}{H} &= 0.
            \label{eq:pseudo_eos}
    \end{align}
\end{subequations}
These are complete up to numerical and forcing terms.

\subsection{Anelastic}

We can argue for this in a very formal/nondimensionalized way, but we can also
see this comparatively simply (albeit less rigorously). We will start from
\autoref{se:comp}. Expand all variables in terms of $\epsilon \equiv \mathrm{Ma}
= \frac{u}{c_s}$ where $u$ is the characteristic velocity of the flows we want
to study, then
\begin{align*}
    P &= P_0 + \epsilon P_1 + \dots,& \rho &= \rho_0 + \epsilon \rho_1 + \dots,&
        S &= S_0 + \epsilon S_1 + \dots,& u &\sim \mathcal{O}(\epsilon^1),\\
    \vec{\nabla} &\sim k \sim \mathcal{O}(\epsilon^0) ,&
        \partial_t &\approx uk \sim \mathcal{O}\p*{\epsilon^1}.
\end{align*}
Note that we pick a particular length scale with $k$ and with typical flow
velocity $u$, then $\partial_t = uk$ is the correct order to examine
at\footnote{More formally, we can probably do $\partial_t = \partial_{t_0} +
\partial_{t_1/\epsilon} + \dots$}.

With this dictionary, we may consider the incompressible equations at:
\begin{itemize}
    \item $\mathcal{O}(\epsilon^0)$, which is just
        $\frac{\vec{\nabla}P_0}{\rho_0} + \vec{g} = 0$ hydrostatic equilibrium.

        At this point, we may make the same argument as in the preceeding
        section: $\rd{}{t} \sim \mathcal{O}(\epsilon^2)$, so
        $\frac{\vec{\nabla}P}{\rho} + \vec{g} = \frac{\vec{\nabla}P}{\rho} -
        \frac{\vec{\nabla}P_0}{\rho_0} \sim \mathcal{O}(\epsilon^2)$ as well.
        Thus, $P_1, \rho_1 = 0$ necessarily, otherwise
        $\frac{\vec{\nabla}P}{\rho} - \frac{\vec{\nabla}P_0}{\rho_0}$ must have
        $\mathcal{O}(\epsilon^1)$ terms that are not cancelled by
        $\rd{\vec{u}}{t}$ which is prohibited.

    \item $\mathcal{O}(\epsilon^1)$. Since $P_1 = \rho_1 = 0$, the only terms at
        this order are $\vec{\nabla} \cdot \p*{\rho_0 \vec{u}} = 0$ and
        $\p*{\vec{u}_1 \cdot \vec{\nabla}}S_0 = 0$.

    \item $\mathcal{O}(\epsilon^2)$. This is where the momentum equation gets
        most of its terms, it becomes the familiar
        \begin{equation}
            \rd{\vec{u}}{t} + \frac{\vec{\nabla}P_2}{\rho_0} + \frac{\rho_2
                \vec{g}}{\rho_0} = 0.
        \end{equation}
        Now we haven't argued for $S_1 \neq 0$, so we must permit it as well, so
        the entropy equation becomes
        \begin{equation}
            \rd{S_1}{t} + \p*{\vec{u} \cdot \vec{\nabla}}S_0 = 0.
        \end{equation}

    \item Higher order expansions are not hard to find, and relax $\vec{\nabla}
        \cdot \p*{\rho_0 \vec{u}} = 0$.
\end{itemize}
Thus, to order $\mathrm{Ma}^2$, our equations are
\begin{subequations}\label{se:anel_0}
    \begin{align}
        \vec{\nabla} \cdot \p*{\rho_0 \vec{u}} &= 0,\\
        \rd{S_1}{t} + \vec{u} \cdot \vec{\nabla}S_0 &= 0,\\
        \rd{\vec{u}}{t} + \frac{\vec{\nabla}P_2}{\rho_0}
            + \frac{\rho_2 \vec{g}}{\rho_0} &= 0,\\
        P &= P(\rho, S).
    \end{align}
\end{subequations}
The last step is just to remove $\rho_2$ from the equations using the equation
of state so that the first three equations of \autoref{se:anel_0} above can form
a complete system. We simply follow Glatzmaier and rewrite
\begin{subequations}\label{se:anel_f}
    \begin{align}
        \vec{\nabla} \cdot \p*{\rho_0 \vec{u}} &= 0,\\
        \pd{S_1}{t} + \p*{\vec{u} \cdot \vec{\nabla}}S_1 + \vec{u} \cdot
            \vec{\nabla}S_0 &= 0,\\
        \pd{\vec{u}}{t} + \p*{\vec{u} \cdot \vec{\nabla}}\vec{u} +
            \vec{\nabla}\frac{P_2}{\rho_0} - \frac{S_1\vec{g}}{S_0} &= 0.
    \end{align}
\end{subequations}
Solving this system gives buoyancy waves with frequency $N^2 =
\frac{g}{H_S} = \frac{g}{H} \at{\pd{\ln S_0}{\ln \rho_0}}_P \ll \frac{g}{H}$ as
we obtain $\pd{\ln S_0}{z} = -\frac{1}{H_S}$ stratification using the EoS.

\end{document}

### 2018-11-24 writeups/notes.tex
small updates, rm sims_old

--- 2018-11-12 writeups/notes.tex
@@
         \centering
-        \includegraphics[width=\textwidth]{../sims_old/2d_0_no_g/no_g_P_t50.png}
+        \includegraphics[width=\textwidth]{plots/no_g_P_t50.png}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{../sims_old/2d_0_no_g/no_g_ux_t50.png}
+        \includegraphics[width=\textwidth]{plots/no_g_ux_t50.png}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{../sims_old/2d_0_no_g/no_g_uz_t50.png}
+        \includegraphics[width=\textwidth]{plots/no_g_uz_t50.png}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{../sims_old/2d_0_no_g/no_g_P_t100.png}
+        \includegraphics[width=\textwidth]{plots/no_g_P_t100.png}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{../sims_old/2d_0_no_g/no_g_ux_t100.png}
+        \includegraphics[width=\textwidth]{plots/no_g_ux_t100.png}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{../sims_old/2d_0_no_g/no_g_uz_t100.png}
+        \includegraphics[width=\textwidth]{plots/no_g_uz_t100.png}
     \end{subfigure}
@@
     \centering
-    \includegraphics[width=0.7\textwidth]{../sims_old/2d_1_strat/agree_plots/sponge_2.png}
+    \includegraphics[width=0.7\textwidth]{plots/sponge_2.png}
     \caption{Note the jaggedness in $u_x$ induced by a Dirichlet BC on

### 2018-11-24 sims_old/1d_uniform/1d_uniform.py

@@ -1,93 +1,0 @@
-"""
-Simplest 1D fluid equation in uniform medium
-
-rho_1/rho = 0.001 seems to look linear, rho_1/rho = 0.1 produces steepening!
-u << c, rho_1 << rho = linear
-"""
-
-import logging
-import numpy as np
-import matplotlib.pyplot as plt
-
-from dedalus import public as de
-from dedalus.extras.plot_tools import quad_mesh, pad_limits
-
-logger = logging.getLogger(__name__)
-XMAX = 10
-N_X = 2048
-DX = XMAX / N_X
-DT = 1e-5
-C = 5 # speed
-T_F = 0.5 * XMAX / C
-
-SIGMA = 0.3 # for the gaussian wave packet
-WIDTH_GAUSS = 3 # how many sigma to go on either side of center of wave packet
-RHO_0 = 10 # background rho
-RAT = 0.5 # ratio u_1/C, rho_1/RHO_0
-
-# Bases and domain
-x_basis = de.Fourier('x', N_X, interval=(0, XMAX), dealias=3/2)
-domain = de.Domain([x_basis], np.float64)
-
-# Problem
-problem = de.IVP(domain, variables=['u', 'r'])
-problem.parameters['c'] = C
-problem.add_equation("dt(r) = -r * dx(u) - u * dx(r)")
-problem.add_equation("dt(u) = -c * c/r * dx(r) - u * dx(u)")
-
-# Build solver
-solver = problem.build_solver(de.timesteppers.SBDF2)
-solver.stop_sim_time = T_F
-solver.stop_wall_time = 100000 # should never get hit
-solver.stop_iteration = 100000 # should never get hit
-
-# Initial conditions
-x = domain.grid(0)
-u = solver.state['u']
-r = solver.state['r']
-
-u['g'] = [0] * len(x)
-r['g'] = [10] * len(x)
-for i in range(int(2 * WIDTH_GAUSS * SIGMA / DX)):
-    center = WIDTH_GAUSS * SIGMA / DX
-    u['g'][i] += C * RAT * np.exp(-(center - i) ** 2 / (2 * (SIGMA / DX) ** 2))
-    r['g'][i] += RHO_0 * RAT * \
-        np.exp(-(center - i) ** 2 / (2 * (SIGMA / DX) ** 2))
-
-# Store data for final plot
-u.set_scales(1, keep_data=True)
-u_list = [np.copy(u['g'])]
-r_list = [np.copy(r['g'])]
-t_list = [solver.sim_time]
-
[…]

### 2018-11-24 sims_old/2d_0_no_g/no_g.py

@@ -1,80 +1,0 @@
-'''
-linearized fluid equations w/o gravity. Fluid equations reduce to
-- div u = 0
-- dt(u) + grad(P) / rho0 = 0
-
-Domain is L x L
-ICs are zero everywhere
-BCs are periodic x, Dirichlet 0 at z = L, sin(2*pi*x/L) at z = 0.
-'''
-
-import numpy as np
-import matplotlib.pyplot as plt
-
-from dedalus import public as de
-from dedalus.extras.plot_tools import quad_mesh, pad_limits
-
-XMAX = 10
-ZMAX = 10
-N_X = 64
-N_Z = 64
-DX = XMAX / N_X
-DZ = ZMAX / N_Z
-T_F = 20
-DT = 1e-2
-
-# Bases and domain
-x_basis = de.Fourier('x', N_X, interval=(0, XMAX), dealias=3/2)
-z_basis = de.Chebyshev('z', N_Z, interval=(0, ZMAX), dealias=3/2)
-domain = de.Domain([x_basis, z_basis], np.float64)
-
-problem = de.IVP(domain, variables=['P', 'ux', 'uz'])
-problem.meta['P']['z']['dirichlet'] = True
-problem.parameters['rho0'] = 1
-problem.add_equation("dx(ux) + dz(uz) = 0")
-problem.add_equation("dt(ux) + dx(P)/rho0 = 0")
-problem.add_equation("dt(uz) + dz(P)/rho0 = 0")
-problem.add_bc("right(P) = 0")
-problem.add_bc("left(P) = sin(6.28 * x / 10)")
-
-# Build solver
-solver = problem.build_solver(de.timesteppers.RK222)
-solver.stop_sim_time = T_F
-solver.stop_wall_time = np.inf
-solver.stop_iteration = 100
-
-# Initial conditions
-x = domain.grid(0)
-z = domain.grid(1)
-ux = solver.state['ux']
-uz = solver.state['uz']
-P = solver.state['P']
-
-# need to reverse for some reason?
-gshape = domain.dist.grid_layout.global_shape(scales=1)
-P['g'] = np.zeros(gshape)
-ux['g'] = np.zeros(gshape)
-uz['g'] = np.zeros(gshape)
-
-# Main loop
-for stop_iter in [50, 100]:
-    solver.stop_iteration = stop_iter
-    while solver.ok:
-        solver.step(DT)
-
-    ux.set_scales(1, keep_data=True)
[…]

### 2018-11-24 sims_old/2d_1_strat/agree_plots.py

@@ -1,122 +1,0 @@
-#!/usr/bin/env python
-# plot agreement of simulation/analytical at various (x, t) slices over z.
-
-import os
-import h5py
-import matplotlib.pyplot as plt
-import matplotlib
-import numpy as np
-import strat_helper
-
-A = 0.05
-H = 1
-XMAX = H
-ZMAX = 5 * H
-KX = 2 * np.pi / H
-_KZ = (np.pi / 2) * np.pi / H
-G = 10
-RHO0 = 1
-OMEGA = strat_helper.get_omega(G, H, KX, _KZ)
-
-def agree_sponge(plot_idx, t_idx=-30):
-    KZ = -_KZ
-    dat = h5py.File('snapshots_sponge/snapshots_sponge_s1/snapshots_sponge_s1_p0.h5', mode='r')
-    tmesh = np.array(dat['tasks']['uz'].dims[0][0])
-    xmesh = np.array(dat['tasks']['uz'].dims[1][0])
-    zmesh = np.array(dat['tasks']['uz'].dims[2][0])
-
-    x_idx = len(xmesh) // 2
-
-    uz_anal = A * np.exp(zmesh / (2 * H)) \
-        * np.cos(KX * xmesh[x_idx] + KZ * zmesh - OMEGA * tmesh[t_idx])
-    rho0 = RHO0 * np.exp(-zmesh / H)
-    analyticals = {
-        'uz': uz_anal,
-        'ux': KZ / KX * uz_anal,
-        'rho': -rho0 * A / (H * OMEGA)* np.exp(zmesh / (2 * H)) \
-            * np.sin(KX * xmesh[x_idx] + KZ * zmesh - OMEGA * tmesh[t_idx]),
-        'P': -rho0 * OMEGA / KX**2 * KZ * uz_anal,
-    }
-
-    fig = plt.figure(dpi=200)
-    for idx, var in enumerate(['uz', 'ux', 'rho', 'P']):
-        axes = fig.add_subplot(2, 2, idx + 1, title=var)
-        vardat = dat['tasks'][var]
-        sliced = vardat[t_idx, x_idx, :]
-        axes.plot(zmesh, sliced, label='Numerical')
-
-        axes.plot(zmesh,
-                 analyticals[var],
-                 label='Analytical'
-                )
-        axes.legend()
-    fig.suptitle('Sponge (t=%.3f, x=%.3f) (damping zone z>=3.875)' %
[…]

### 2018-11-24 sims_old/2d_1_strat/strat_helper.py

@@ -1,274 +1,0 @@
-#!/usr/bin/env python
-'''
-helper function to run the shared stratification scenario. user just has to
-specify BCs and ICs
-'''
-import logging
-import os
-from collections import defaultdict
-
-import h5py
-import matplotlib
-import matplotlib.pyplot as plt
-import numpy as np
-
-from dedalus import public as de
-from dedalus.extras.flow_tools import CFL
-from dedalus.extras.plot_tools import quad_mesh, pad_limits
-
-SNAPSHOTS_DIR = 'snapshots_%s'
-def get_solver(setup_problem,
-               XMAX,
-               ZMAX,
-               N_X,
-               N_Z,
-               T_F,
-               KX,
-               KZ,
-               H,
-               RHO0,
-               G,
-               A):
-    # Bases and domain
-    x_basis = de.Fourier('x', N_X, interval=(0, XMAX), dealias=3/2)
-    z_basis = de.Chebyshev('z', N_Z, interval=(0, ZMAX), dealias=3/2)
-    domain = de.Domain([x_basis, z_basis], np.float64)
-    z = domain.grid(1)
-
-    problem = de.IVP(domain, variables=['P', 'rho', 'ux', 'uz'])
-    problem.meta['uz']['z']['dirichlet'] = True
-    problem.parameters['L'] = XMAX
-    problem.parameters['g'] = G
-    problem.parameters['H'] = H
-    problem.parameters['A'] = A
-    problem.parameters['KX'] = KX
-    problem.parameters['KZ'] = KZ
-    problem.parameters['omega'] = get_omega(G, H, KX, KZ)
-
-    # rho0 stratification
-    rho0 = domain.new_field()
-    rho0.meta['x']['constant'] = True
-    rho0['g'] = RHO0 * np.exp(-z / H)
-    problem.parameters['rho0'] = rho0
-
-    setup_problem(problem, domain)
-
-    # Build solver
-    solver = problem.build_solver(de.timesteppers.RK443)
-    solver.stop_sim_time = T_F
-    solver.stop_wall_time = np.inf
-    solver.stop_iteration = np.inf
-    return solver, domain
-
-def run_strat_sim(setup_problem,
-                  set_ICs,
-                  XMAX,
-                  ZMAX,
-                  N_X,
-                  N_Z,
-                  T_F,
-                  DT,
-                  KX,
-                  KZ,
-                  H,
[…]

### 2018-11-24 sims_old/2d_2_strat_drop_terms/strat_helper.py

@@ -1,496 +1,0 @@
-#!/usr/bin/env python
-'''
-helper function to run the shared stratification scenario. user just has to
-specify BCs and ICs
-'''
-import logging
-import os
-from collections import defaultdict
-
-import h5py
-import matplotlib
-import matplotlib.pyplot as plt
-import numpy as np
-
-from dedalus import public as de
-from dedalus.extras.flow_tools import CFL
-from dedalus.extras.plot_tools import quad_mesh, pad_limits
-
-SNAPSHOTS_DIR = 'snapshots_%s'
-
-###
-### UTILS
-###
-
-def get_omega(g, h, kx, kz):
-    return np.sqrt((g / h) * kx**2 / (kx**2 + kz**2 + 0.25 / h**2))
-
-def get_vph(g, h, kx, kz):
-    norm = get_omega(g, h, kx, kz) / (kx**2 + kz**2)
-    return norm * kx, norm * kz
-
-def get_analytical_sponge(name, z_pts, t, params):
-    """ gets the analytical form of the variables for radiative BCs """
-    uz_anal = params['A'] * np.exp(z_pts / (2 * params['H'])) *\
-        np.cos(params['KZ'] * z_pts - params['OMEGA'] * t)
-    rho0 = params['RHO0'] * np.exp(-z_pts / params['H'])
-    analyticals = {
-        'uz': uz_anal,
-        'ux': -params['KZ'] / params['KX'] * uz_anal,
-        'rho1': -rho0 * params['A'] / (params['H'] * params['OMEGA']) *\
-            np.exp(z_pts / (2 * params['H'])) *\
-            np.sin(params['KZ'] * z_pts - params['OMEGA'] * t),
-        'P1': -rho0 * params['OMEGA'] / params['KX']**2 * params['KZ'] *\
-            uz_anal,
-    }
-    return analyticals[name]
-
-def get_sponge(domain, params):
-    sponge_strength = params['SPONGE_STRENGTH']
-    zmax = params['ZMAX']
-    damp_start = params['SPONGE_START']
-    z = domain.grid(1)
-
-    # sponge field
-    sponge = domain.new_field()
-    sponge.meta['x']['constant'] = True
-    sponge['g'] = sponge_strength * np.maximum(
-        1 - (z - zmax)**2 / (damp_start - zmax)**2,
-        np.zeros(np.shape(z)))
-    return sponge
-
-###
-### IC
-###
-
[…]

### 2018-11-24 sims_old/2d_2_strat_nonlin/strat.py

@@ -1,127 +1,0 @@
-'''
-2d_1_strat + Navier-Stokes dissipation
-'''
-from multiprocessing import Pool
-import numpy as np
-from strat_helper import *
-
-N_PARALLEL = 20
-H = 1
-num_timesteps = 3e5
-
-XMAX = H
-ZMAX = 3 * H
-KX = 2 * np.pi / H
-KZ = -(np.pi) * np.pi / H
-G = (KX**2 + KZ**2 + 1 / (4 * H**2)) / KX**2 * (2 * np.pi)**2 * H # omega = 2pi
-OMEGA = get_omega(G, H, KX, KZ)
-_, VPH_Z = get_vph(G, H, KX, KZ)
-T_F = abs(ZMAX / VPH_Z) * 4
-DT = T_F / num_timesteps
-NUM_SNAPSHOTS = 400
-
-PARAMS_RAW = {'XMAX': XMAX,
-              'ZMAX': ZMAX,
-              'N_X': 64,
-              'N_Z': 256,
-              'T_F': T_F,
-              'DT': DT,
-              'OMEGA': OMEGA,
-              'KX': KX,
-              'KZ': KZ,
-              'H': H,
-              'RHO0': 1,
-              'G': G,
-              'A': 0.005,
-              'SPONGE_STRENGTH': 6,
-              'SPONGE_START': 0.7 * ZMAX,
-              'NUM_SNAPSHOTS': NUM_SNAPSHOTS}
-
-def build_interp_params(interp_x, interp_z, dt=DT, overrides=None):
-    params = {**PARAMS_RAW, **(overrides or {})}
-    params['INTERP_X'] = interp_x
-    params['INTERP_Z'] = interp_z
-    params['N_X'] //= interp_x
-    params['N_Z'] //= interp_z
-    params['DT'] = dt
-    params['NU'] = 0.1 * (ZMAX / params['N_Z'])**2 / np.pi**2 # smallest wavenumber
-    return params
-
-def run(get_solver, bc, ic, name, params_dict):
-    assert 'INTERP_X' in params_dict and 'INTERP_Z' in params_dict,\
-        'params need INTERP'
-    run_strat_sim(get_solver, bc, ic, name, params_dict)
-    return '%s completed' % name
-
-if __name__ == '__main__':
-    tasks = [
-        # linear, neumann works to drop ux waviness
-        (get_solver, sponge_lin, zero_ic,
-         '0sponge_lin',
-         build_interp_params(8, 8)),
-
-        # nonlin, low amplitude
[…]

### 2018-11-24 sims_old/2d_2_strat_nonlin/strat_helper.py

@@ -1,496 +1,0 @@
-#!/usr/bin/env python
-'''
-helper function to run the shared stratification scenario. user just has to
-specify BCs and ICs
-'''
-import logging
-import os
-from collections import defaultdict
-
-import h5py
-import matplotlib
-import matplotlib.pyplot as plt
-import numpy as np
-
-from dedalus import public as de
-from dedalus.extras.flow_tools import CFL
-from dedalus.extras.plot_tools import quad_mesh, pad_limits
-
-SNAPSHOTS_DIR = 'snapshots_%s'
-
-###
-### UTILS
-###
-
-def get_omega(g, h, kx, kz):
-    return np.sqrt((g / h) * kx**2 / (kx**2 + kz**2 + 0.25 / h**2))
-
-def get_vph(g, h, kx, kz):
-    norm = get_omega(g, h, kx, kz) / (kx**2 + kz**2)
-    return norm * kx, norm * kz
-
-def get_analytical_sponge(name, z_pts, t, params):
-    """ gets the analytical form of the variables for radiative BCs """
-    uz_anal = params['A'] * np.exp(z_pts / (2 * params['H'])) *\
-        np.cos(params['KZ'] * z_pts - params['OMEGA'] * t)
-    rho0 = params['RHO0'] * np.exp(-z_pts / params['H'])
-    analyticals = {
-        'uz': uz_anal,
-        'ux': -params['KZ'] / params['KX'] * uz_anal,
-        'rho1': -rho0 * params['A'] / (params['H'] * params['OMEGA']) *\
-            np.exp(z_pts / (2 * params['H'])) *\
-            np.sin(params['KZ'] * z_pts - params['OMEGA'] * t),
-        'P1': -rho0 * params['OMEGA'] / params['KX']**2 * params['KZ'] *\
-            uz_anal,
-    }
-    return analyticals[name]
-
-def get_sponge(domain, params):
-    sponge_strength = params['SPONGE_STRENGTH']
-    zmax = params['ZMAX']
-    damp_start = params['SPONGE_START']
-    z = domain.grid(1)
-
-    # sponge field
-    sponge = domain.new_field()
-    sponge.meta['x']['constant'] = True
-    sponge['g'] = sponge_strength * np.maximum(
-        1 - (z - zmax)**2 / (damp_start - zmax)**2,
-        np.zeros(np.shape(z)))
-    return sponge
-
-###
[…]

### 2018-11-24 sims_old/2d_3_factored/strat.py

@@ -1,78 +1,0 @@
-'''
-2d_1_strat + Navier-Stokes dissipation
-'''
-from multiprocessing import Pool
-import numpy as np
-from strat_helper import *
-
-N_PARALLEL = 20
-XMAX = 1
-KX = 2 * np.pi / XMAX
-ZMAX = 30 * (2 * np.pi / KX)
-H = 10 / KX
-KZ = -0.4 * KX
-
-G = 1
-OMEGA = get_omega(G, H, KX, KZ)
-_, VPH_Z = get_vph(G, H, KX, KZ)
-
-num_timesteps = 1e3
-T_F = abs(ZMAX / VPH_Z)
-DT = T_F / num_timesteps
-NUM_SNAPSHOTS = 200
-
-PARAMS_RAW = {'XMAX': XMAX,
-              'ZMAX': ZMAX,
-              'N_X': 128,
-              'N_Z': 256,
-              'T_F': T_F,
-              'DT': DT,
-              'OMEGA': OMEGA,
-              'KX': KX,
-              'KZ': KZ,
-              'H': H,
-              'RHO0': 1,
-              'G': G,
-              'A': 0.005,
-              'F': 0.1,
-              'SPONGE_STRENGTH': 100,
-              'SPONGE_START_HIGH': 0.9 * ZMAX,
-              'SPONGE_START_LOW': 0.1 * ZMAX,
-              'NUM_SNAPSHOTS': NUM_SNAPSHOTS}
-
-def build_interp_params(interp_x, interp_z, dt=DT, overrides=None):
-    params = {**PARAMS_RAW, **(overrides or {})}
-    params['INTERP_X'] = interp_x
-    params['INTERP_Z'] = interp_z
-    params['N_X'] //= interp_x
-    params['N_Z'] //= interp_z
-    params['DT'] = dt
-    params['NU'] = (ZMAX / params['N_Z']) / np.pi # ~ 1/k_max
-    return params
-
-def run(get_solver, bc, ic, name, params_dict):
-    assert 'INTERP_X' in params_dict and 'INTERP_Z' in params_dict,\
-        'params need INTERP'
-    run_strat_sim(get_solver, bc, ic, name, params_dict)
-    return '%s completed' % name
-
-if __name__ == '__main__':
-    tasks = [
-        (get_solver, setup_problem_unforced, wavepacket_ic,
-         'RKSMR',
-         build_interp_params(4, 2, overrides={'TIMESTEPPER':
[…]

### 2018-11-26 writeups/running_writeup.tex
check in updated writeup w/ change of variables

@@
 
-We begin with the fully compressible fluid equations and general equation of
-state
-\begin{subequations}\label{se:comp}
+\subsection{Boussinesq Equations}
+
+The physical equations in the Boussinesq approximation are:
+\begin{subequations}\label{se:bouss}
     \begin{align}
-        \rd{\rho}{t} + \rho \p*{\vec{\nabla} \cdot \vec{u}} &= 0,
-            \label{eq:density}\\
-        \rd{S}{t} &= 0,\label{eq:entropy}\\
-        \rd{\vec{u}}{t} + \frac{\vec{\nabla}P}{\rho} + \vec{g} &= 0,
-            \label{eq:velocity}\\
-        P &= P(\rho, S).\label{eq:eos}
+        \vec{\nabla} \cdot \vec{u} &= 0,\\
+        \pd{\rho_1}{t} + \p*{\vec{u} \cdot \vec{\nabla}}\rho_1
+            - \frac{\rho_0 u_z}{H} &= 0,\\
+        \pd{\vec{u}}{t} + \p*{\vec{u} \cdot \vec{\nabla}}\vec{u}
+            + \frac{\vec{\nabla}P_1}{\rho_0}
+            + \frac{\rho_1 \vec{g}}{\rho_0} &= 0.
     \end{align}
 \end{subequations}
-We notate $\rd{}{t} = \pd{}{t} + \p*{\vec{u} \cdot \vec{\nabla}}$. Note this is
-$5$ equations (in 2D) for 5 variables $(\rho, P, S, \vec{u})$ and so is properly
-closed.
+In these equations, we take $\rho_0(z) = \rho_0$ a constant reference density.
 
-\subsection{Current Implementation}
+Including numerical terms and driving terms, my full equations are
+\begin{subequations}\label{se:bouss_num}
+    \begin{align}
+        \vec{\nabla} \cdot \vec{u}_1 &= 0,\\
+        \pd{\rho_1}{t} - \frac{\rho_0 u_z}{H}
+            - \nu \nabla^6 \rho_1
+            &= -\Gamma(z) \rho_1
+                - \p*{\vec{u} \cdot \vec{\nabla}}\rho_1
+                + Fe^{-\frac{(z - z_0)^2}{2\sigma^2}}
+                    \cos \p*{k_xx - \omega t},\\
+        \pd{u_x}{t} + \frac{\partial_x P}{\rho_0}
+            - \nu \nabla^6 u_x
+            &= -\Gamma(z) u_x
+                - \p*{\vec{u} \cdot \vec{\nabla}}u_x,\\
+        \pd{u_z}{t} + \frac{\partial_z P}{\rho_0}
+            + \frac{\rho_1 g}{\rho_0}
+            - \nu \nabla^6 u_z
+            &= -\Gamma(z) u_z
+                - \p*{\vec{u} \cdot \vec{\nabla}}u_z,\\
+        \Gamma(z) &= 5\s*{2 + \tanh \frac{z - z_T}{(z_{\max} - z_T) / 2}
+            + \tanh \frac{z_B - z}{z_B / 2}},
+    \end{align}
+\end{subequations}
+where $z_B = 0.07z_{\max}, z_T = 0.93z_{\max}$ are the boundaries of the damping
+zones, and my domain is $z \in [0, z_{\max}]$. The forcing term $F$ is chosen to
+be weakly nonlinear, $\omega$ is chosen by inverting $\omega(k_x, k_z)$
+dispersion relation for fixed $k_x = 2\pi / x_{\max}$ and some desired $k_z \ll
+H, z_{\max}$ (I've chosen $k_z \approx -\frac{2\pi}{z_{\max} / 5}$ here), and
+$\sigma \lesssim \frac{1}{k_z}$ is used to excite a broad band of modes
+including the desired $k_z$ mode.
 
-We assume $P(\rho, S) = P(\rho)$ a barotropic equation of state, such that $S$
-is entirely decoupled from the dynamical equations. We thus drop
-\autoref{eq:entropy} of the full fluid equations \autoref{se:comp}. Then, we
-will decompose dynamical variables into background and fluctuation quantities $P
-= P_0 + P_1, \rho = \rho_0 + \rho_1$, where background quantities are
-time-independent. Assuming no background velocities, we write $\vec{u}$ to
-denote the fluctuation velocity with no ambiguity.
+\subsection{Stratification}
 
-Furthermore, we will implement incompressibility by mandating
-$\at{\pd{P(\rho)}{\rho}}_{ad} \to \infty$ adiabatic derivative. This forces
-$\Delta P \gg \Delta \rho$ within a displaced fluid parcel, or $\rd{\rho}{t} =
-0$ comoving derivative. This allows us to rewrite \autoref{eq:density}, and so
-we arrive at the system of equations
-\begin{subequations}\label{se:pseudo_mid}
+We simulate an incompressible, isothermal fluid, representatitive of degenerate
+matter in WD bulks. We assume a barotropic equation of state to simplify for the
+time being. The physical equations for an incompressible, barotropic fluid are:
+\begin{subequations}\label{se:fc_orig}
     \begin{align}
-        \vec{\nabla} \cdot \vec{u} &= 0, \label{eq:mid_density}\\
-        \rd{\vec{u}}{t} + \frac{\vec{\nabla}P}{\rho} + \vec{g} &= 0,
-            \label{eq:mid_velocity}\\
-        \rd{\rho}{t} = \rd{\rho_1}{t} + u_z \rd{\rho_0}{z} &= 0.
-            \label{eq:mid_eos}
+        \vec{\nabla} \cdot \vec{u} &= 0,\\
+        \pd{\rho}{t} + \vec{u} \cdot \vec{\nabla}\rho &= 0,\\
+        \pd{\vec{u}}{t} + \p*{\vec{u} \cdot \vec{\nabla}}\vec{u}
+            + \frac{\vec{\nabla}P}{\rho}
+            + g\hat{z} &= 0.
+    \end{align}
+\end{subequations}
+We will introduce variables $T = P/\rho$. We then mandate $\rho_0, T_0$
+backgrounds satisfy hydrostatic equilibrium $\vec{\nabla}T_0 + T_0
+\vec{\nabla}\rho_0 + \vec{g} = 0$. Taking isothermal stratification, we find
+$T_0 = gH$. Making then substitution of variables $\Upsilon = \ln \rho - \ln
+\rho_0$ and $T_1 = T - T_0$ deviations from the background state, the exact
+fluid equations in the new variables are:
+\begin{subequations}\label{se:fc_var}
+    \begin{align}
+        \vec{\nabla} \cdot \vec{u} &= 0,\\
+        \pd{\Upsilon}{t} - \frac{u_z}{H} &= 0,\\
+        \pd{u_x}{t} + \p*{\vec{u} \cdot \vec{\nabla}}u_x
+            + \pd{T}{x} + gH\pd{\Upsilon}{x}
+            + T_1 \pd{\Upsilon}{x} &= 0,\\
+        \pd{u_z}{t} + \p*{\vec{u} \cdot \vec{\nabla}}u_z
+            + \pd{T}{z} + gH\pd{\Upsilon}{z}
+            + T_1 \pd{\Upsilon}{z} - \frac{T_1}{H} &= 0.
     \end{align}
@@
 
-We now assume $P_0 = \rho_0 c_s^2 \propto e^{-z/H}$ isothermal exponential
-stratification. At hydrostatic equilibrium, \autoref{eq:mid_velocity} forces
-$\vec{\nabla}P_0 = -\rho_0 \vec{g}$. We next claim that $\frac{\rho_1}{\rho_0}
-\sim \p*{\frac{u}{c_s}}^2 = \mathrm{Ma}^2 \ll 1$\footnote{I guess this doesn't
-have to be true if $\frac{P_0}{\rho_0} = c_s^2 \neq \at{\pd{P}{\rho}}_{ad}$
-which is the velocity we are sending to infinity.}. This may be verified to be
-true by looking briefly at the momentum equation: $\omega u +
-k\p*{\frac{P}{\rho} - \frac{P_0}{\rho_0}} = 0$, which then dividing through by
-$P_0 / \rho_0 = c_s^2$ shows that $\frac{P/P_0}{\rho/\rho_0} - 1 \sim
-\mathrm{Ma}^2$.
-
-Given this, we expand \autoref{eq:mid_velocity} by $\frac{\vec{\nabla}P}{\rho}
-\approx \frac{\vec{\nabla}P_0}{\rho_0} + \frac{\vec{\nabla}P_1}{\rho_0} -
-\frac{\rho_1 \vec{\nabla}P_0}{\rho_0^2} + \dots$ and obtain the full system of
-equations I have been using
-\begin{subequations}\label{se:pseudo_fin}
+These are implemented as:
+\begin{subequations}\label{se:curr_num}
     \begin{align}
-        \vec{\nabla} \cdot \vec{u} &= 0, \label{eq:pseudo_density}\\
-        \pd{\vec{u}}{t} + \p*{\vec{u} \cdot \vec{\nabla}}\vec{u}
-            + \frac{\vec{\nabla}P_1}{\rho_0}
-            + \frac{\rho_1 \vec{g}}{\vec{_0}} - \frac{\rho_1 \vec{\nabla}P_1}{
-                \rho_0^2} &= 0,
-            \label{eq:pseudo_velocity}\\
-        \pd{\rho_1}{t} + \p*{\vec{u} \cdot \vec{\nabla}}\rho_1
-            - \frac{u_z \rho_0}{H} &= 0.
-            \label{eq:pseudo_eos}
+        \vec{\nabla} \cdot \vec{u} &= 0,\\
+        \pd{\Upsilon}{t} - \frac{u_z}{H}
+            - \nu \nabla^2 \Upsilon &= -\Gamma(z) \Upsilon
+                - \p*{\vec{u} \cdot \vec{\nabla}}\Upsilon
+                + \frac{F}{\rho_0(z)}e^{-\frac{(z - z_0)^2}{2\sigma^2}}
+                    \cos \p*{k_xx - \omega t},\\
+        \pd{u_x}{t} + \pd{T}{x} + gH\pd{\Upsilon}{x}
+            - \nu \nabla^2 u_x
+            &= -\Gamma(z) u_x
+                - \p*{\vec{u} \cdot \vec{\nabla}}u_x
+                - T_1 \pd{\Upsilon}{x},\\
+        \pd{u_z}{t} + \pd{T}{z} + gH\pd{\Upsilon}{z} - \frac{T_1}{H}
+            - \nu \nabla^2 u_z &= -\Gamma(z) u_z
+                - \p*{\vec{u} \cdot \vec{\nabla}}u_z
+                - T_1 \pd{\Upsilon}{z}.
     \end{align}
 \end{subequations}
-These are complete up to numerical and forcing terms.
-
-\subsection{Anelastic}
-
-We can argue for this in a very formal/nondimensionalized way, but we can also
-see this comparatively simply (albeit less rigorously). We will start from
-\autoref{se:comp}. Expand all variables in terms of $\epsilon \equiv \mathrm{Ma}
-= \frac{u}{c_s}$ where $u$ is the characteristic velocity of the flows we want
-to study, then
-\begin{align*}
-    P &= P_0 + \epsilon P_1 + \dots,& \rho &= \rho_0 + \epsilon \rho_1 + \dots,&
-        S &= S_0 + \epsilon S_1 + \dots,& u &\sim \mathcal{O}(\epsilon^1),\\
-    \vec{\nabla} &\sim k \sim \mathcal{O}(\epsilon^0) ,&
-        \partial_t &\approx uk \sim \mathcal{O}\p*{\epsilon^1}.
-\end{align*}
-Note that we pick a particular length scale with $k$ and with typical flow
-velocity $u$, then $\partial_t = uk$ is the correct order to examine
-at\footnote{More formally, we can probably do $\partial_t = \partial_{t_0} +
-\partial_{t_1/\epsilon} + \dots$}.
-
-With this dictionary, we may consider the incompressible equations at:
-\begin{itemize}
-    \item $\mathcal{O}(\epsilon^0)$, which is just
-        $\frac{\vec{\nabla}P_0}{\rho_0} + \vec{g} = 0$ hydrostatic equilibrium.
-
-        At this point, we may make the same argument as in the preceeding
-        section: $\rd{}{t} \sim \mathcal{O}(\epsilon^2)$, so
-        $\frac{\vec{\nabla}P}{\rho} + \vec{g} = \frac{\vec{\nabla}P}{\rho} -
-        \frac{\vec{\nabla}P_0}{\rho_0} \sim \mathcal{O}(\epsilon^2)$ as well.
-        Thus, $P_1, \rho_1 = 0$ necessarily, otherwise
-        $\frac{\vec{\nabla}P}{\rho} - \frac{\vec{\nabla}P_0}{\rho_0}$ must have
-        $\mathcal{O}(\epsilon^1)$ terms that are not cancelled by
-        $\rd{\vec{u}}{t}$ which is prohibited.
-
-    \item $\mathcal{O}(\epsilon^1)$. Since $P_1 = \rho_1 = 0$, the only terms at
-        this order are $\vec{\nabla} \cdot \p*{\rho_0 \vec{u}} = 0$ and
-        $\p*{\vec{u}_1 \cdot \vec{\nabla}}S_0 = 0$.
-
-    \item $\mathcal{O}(\epsilon^2)$. This is where the momentum equation gets
-        most of its terms, it becomes the familiar
-        \begin{equation}
-            \rd{\vec{u}}{t} + \frac{\vec{\nabla}P_2}{\rho_0} + \frac{\rho_2
-                \vec{g}}{\rho_0} = 0.
-        \end{equation}
-        Now we haven't argued for $S_1 \neq 0$, so we must permit it as well, so
-        the entropy equation becomes
-        \begin{equation}
-            \rd{S_1}{t} + \p*{\vec{u} \cdot \vec{\nabla}}S_0 = 0.
-        \end{equation}
-
-    \item Higher order expansions are not hard to find, and relax $\vec{\nabla}
-        \cdot \p*{\rho_0 \vec{u}} = 0$.
-\end{itemize}
-Thus, to order $\mathrm{Ma}^2$, our equations are
-\begin{subequations}\label{se:anel_0}
-    \begin{align}
-        \vec{\nabla} \cdot \p*{\rho_0 \vec{u}} &= 0,\\
-        \rd{S_1}{t} + \vec{u} \cdot \vec{\nabla}S_0 &= 0,\\
-        \rd{\vec{u}}{t} + \frac{\vec{\nabla}P_2}{\rho_0}
-            + \frac{\rho_2 \vec{g}}{\rho_0} &= 0,\\
-        P &= P(\rho, S).
-    \end{align}
-\end{subequations}
-The last step is just to remove $\rho_2$ from the equations using the equation
-of state so that the first three equations of \autoref{se:anel_0} above can form
-a complete system. We simply follow Glatzmaier and rewrite
-\begin{subequations}\label{se:anel_f}
-    \begin{align}
-        \vec{\nabla} \cdot \p*{\rho_0 \vec{u}} &= 0,\\
-        \pd{S_1}{t} + \p*{\vec{u} \cdot \vec{\nabla}}S_1 + \vec{u} \cdot
-            \vec{\nabla}S_0 &= 0,\\
-        \pd{\vec{u}}{t} + \p*{\vec{u} \cdot \vec{\nabla}}\vec{u} +
-            \vec{\nabla}\frac{P_2}{\rho_0} - \frac{S_1\vec{g}}{S_0} &= 0.
-    \end{align}
-\end{subequations}
-Solving this system gives buoyancy waves with frequency $N^2 =
-\frac{g}{H_S} = \frac{g}{H} \at{\pd{\ln S_0}{\ln \rho_0}}_P \ll \frac{g}{H}$ as
-we obtain $\pd{\ln S_0}{z} = -\frac{1}{H_S}$ stratification using the EoS.
+$\Gamma(z)$ is as before, but we use $k_z = -2\pi/H$ here, and $\omega$ in the
+forcing term accordingly.
 

### 2018-12-06 writeups/running_writeup.tex
Stash a bunch of new simulations + writeup

@@
-    \documentclass[11pt,
+    \documentclass[twocolumn,
         usenames, % allows access to some tikz colors
         dvipsnames % more colors: https://en.wikibooks.org/wiki/LaTeX/Colors
-    ]{article}
+    ]{revtex4-1}% chktex 8
     \usepackage{
@@
         amssymb,
-        fouriernc, % fourier font w/ new century book
         fancyhdr, % page styling
@@
     \newcommand*{\svec}[1]{\vec{#1}\;\!}
-    \newcommand*{\bm}[1]{\boldsymbol{\mathbf{#1}}}
     \newcommand*{\ang}[0]{\;\text{\AA}}
@@
 
-\onehalfspacing
+\title{IGWs and WDs Ongoing Writeup}
+\author{Yubo Su}
+\maketitle
 
-\pagestyle{fancy}
-\rfoot{Yubo Su}
-\rhead{}
-\cfoot{\thepage/\pageref{LastPage}}
+\section{Introduction}
 
-\section{Equations}
+WIP ongoing writeup, last updated \today.
 
-\subsection{Boussinesq Equations}
+% Define white dwarfs (WD), internal gravity waves (IGW), citations
+
+Basic narrative is: the prevailing understanding is that mean flow absorption
+slowly spins up a WD from the outside in as the critical layer propagates
+inwards. Our numerical simulations validate this prediction and give an
+analytical prediction for the motion of the critical layer. However, it appears
+our numerical simulations are viscosity limited. We present a brief argument for
+why the viscosity limitations are insignificant, but also present some
+exploratory, inconclusive work towards direct numerical simulation resolving
+critical layer behavior without being viscosity limited.
+
+In \autoref{s:eq}, we will cover the equations and numerical setup. In
+\autoref{s:lin}, we will discuss the agreement of our simulation in the
+low-amplitude limit with analytical linear theory. In \autoref{s:nonlin} we will
+present a few reference simulations illustrating critical layer absorption
+depositing horizontal angular momentum into the fluid. Finally, in
+\autoref{s:bouss} we will consider a speculative series of simulations
+attempting to rectify the seemingly viscosity-limited simulations of the
+previous section.
+
+\section{Equations}\label{s:eq}
+
+We simulate an incompressible, isothermal fluid, representatitive of degenerate
+matter in WD bulks. We assume a barotropic equation of state as a first
+approximation. Furthermore, we model the background, hydrostatic density
+stratification as $\rho_0(x, z) = \rho_0(z) \propto e^{-z/H}$.
+
+The physical equations for an incompressible, barotropic fluid in a uniform
+gravitational field are
+\begin{subequations}\label{se:fc_orig}
+    \begin{align}
+        \vec{\nabla} \cdot \vec{u} &= 0,\\
+        \pd{\rho}{t} + \vec{u} \cdot \vec{\nabla}\rho &= 0
+            ,\label{eq:fc_density}\\
+        \pd{\vec{u}}{t} + \p*{\vec{u} \cdot \vec{\nabla}}\vec{u}
+            + \frac{\vec{\nabla}P}{\rho}
+            + g\hat{z} &= 0.
+    \end{align}
+\end{subequations}
+$\vec{u}, \rho, P$ are the fluid dynamical variables and $-g\hat{z}$ is constant
+gravitational acceleration. Note that at hydrostatic equilibrium $\partial_t =
+0$ we have $\vec{u} = 0, \vec{\nabla}P_0 = -\rho_0 g\hat{z}$ and so $P_0 =
+\rho_0 gH$.
+
+We nondimensionalize by taking $H = N = \rho_{ref} = 1$ where
+\begin{equation}
+    N^2 \equiv g^2\p*{\rd{\rho}{P} - \at{\pd{\rho}{P}}_{ad}}
+        = \frac{g}{H},
+\end{equation}
+is the Brunt-V\"ais\"al\"a frequency and $\rho_{ref}$ is the background density
+stratification at $z = 0$. The full numerical details of the simulation are
+described in \autoref{ss:strat_impl}.
+
+\subsection{Wave Generation}
+
+To model a continuous IGW wave train excited deep in the WD interior propagating
+towards the surface without unnecessary computational cost, we excite an
+IGW wave train from the bottom of the simulation domain. As Chebyshev grid is
+a factor of $\propto N_z$ denser near the edges of the simulation domain, we
+use a volumetric forcing to excite IGW at some $z_0 > 0$ interior to the
+simulation domain. Our forcing excites both IGW propagating upwards, the
+physically interesting solution, and downwards, which are damped away by damping
+layers (see \autoref{ss:strat_impl}) before inducing strict Courant-Friedrichs-Lewy
+(CFL) timestepping constraints (cite).
+
+As not to interfere with the incompressibility constraint, we force the system
+on the density equation, replacing \autoref{eq:fc_density} with
+\begin{equation}
+    \pd{\rho}{t} + \vec{u} \cdot \vec{\nabla}\rho =
+        Fe^{-\frac{(z - z_0)^2}{2\sigma^2}}
+            \cos \p*{k_{0, x}x - \omega_0 t}.
+\end{equation}
+Define $k_{0, z}: \omega(k_{0, x}, k_{0, z}) = \omega_0$. If $k_{0, z}\sigma \ll
+1$, the forcing term has the effect of picking out $k_{0, z}$ driven by
+$\cos\p*{k_{0, x}x - \omega_0 t}$ from the broad $k_z$ wavenumber spectrum
+permitted by the narrow Gaussian profile.
+
+\section{Linear Regime}\label{s:lin}
+
+We first test our numerical simulation with a ``linear'' simulation, where
+all flow quantities are small: we assume  $\rho_1 \equiv \rho - \rho_0 \ll
+\rho_0$ and $\vec{u} \cdot \vec{\nabla} \ll \partial_t$. We clarify that our
+``linear'' simulations solve the full nonlinear equations, just in the weak
+perturbation limit.
+
+\subsection{Analytical Solution}
+
+In the small perturbation limit, we linearize \autoref{se:fc_orig} with
+inhomegeneous driving term and obtain:
+\begin{subequations}\label{se:vol_drive}
+    \begin{align}
+        \vec{\nabla} \cdot \vec{u} &= 0,\\
+        \pd{\vec{u}_1}{t} + \frac{\vec{\nabla}P}{\rho_0}
+            + \frac{\rho_1 \vec{g}}{\rho_0}
+            &= 0,\\
+        \pd{\rho_1}{t} - \frac{u_{1z} \rho_0}{H}
+            &= Fe^{-\frac{(z - z_0)^2}{2\sigma^2}}
+                e^{i(k_xx - \omega t)}.
+    \end{align}
+\end{subequations}
+We have substituted $\cos\p*{k_xx - \omega_0t}$ for a complex exponential for
+easier manipulation; taking real parts of all dynamical variables above recovers
+the physical solution.
+
+It is well known that the solutions to the $F = 0$ homogeneous
+\autoref{se:vol_drive} are of form (cite)
+\begin{subequations}\label{se:lin_sol}
+    \begin{align}
+        u_z\p*{x, z, t} &= Ae^{z/2H}e^{i(k_{0x}x + k_{0z}z - \omega_0 t)},\\
+        \omega_0^2 &= \frac{N^2k_{0x}^2}{k_{0x}^2 + k_{0z}^2 + \frac{1}{4H^2}}.
+    \end{align}
+\end{subequations}
+
+To solve the inhomogeneous system, we may first approximate the driving term
+using $e^{\frac{-(z - z_0)^2}{2\sigma^2}} \approx \sqrt{2\pi \sigma^2}\delta(z -
+z_0)$ if $\sigma \to 0$, then relax to $k_{0z}\sigma \lesssim 1$. This may be
+solved to obtain the solution
+\begin{align}
+    A(F) &= \frac{Fgk_{0x}^2}{\rho_0(z_0)\omega_0^2}
+        \frac{1}{2ik_z}\frac{e^{\frac{-(k_{0z}\sigma)^2}{2}}}
+        {\sqrt{2\pi\sigma^2}},\\
+    u_{1z} &= A(F) \times\begin{cases}
+        e^{\frac{z - z_0}{2H}}e^{i(k_{0x}x + k_{0z}(z - z_0) - \omega_0 t)}
+            & z > z_0\\
+        e^{\frac{z - z_0}{2H}}e^{i(k_{0x}x - k_{0z}(z - z_0) - \omega_0 t)}
+            & z < z_0\\
+    \end{cases}.\label{eq:uz_lin}
+\end{align}
+The extra factor of $e^{-\frac{(k_{0z}\sigma)^2}{2}}$ compared to the
+$\delta$-function solution comes from evaluating the Fourier Transform of
+$e^{-\frac{(z - z_0)^2}{2\sigma^2}}$ at $k_z = k_{0z}$.
+
+Furthermore, in the low-amplitude, dissipation free limit, the time-averaged
+total $x$-momentum flux in the $\hat{z}$ direction can be computed. Since all
+dynamical variables $\propto e^{i(k_{0x}x - \omega_0 t)}$, $x$ averaging and
+time averaging are equivalent, so we may write
+\begin{equation}
+    F_{px} \equiv \frac{1}{L_x}\int_0^{L_x}\limits \rho u_x u_z\;\mathrm{d}x
+\end{equation}
+We will denote this $F_{px} = \ev*{\rho u_xu_z}_x$ for simplicity. For a given
+forcing amplitude then:
+\begin{equation}
+    F_{px}(F) \approx -\frac{A(F)^2}{2}\rho_0(z_0) \frac{k_z}{k_x}.
+        \label{eq:fpx_lin}
+\end{equation}
+We've approximated $k_zH \ll 1$ and noticed $\ev*{\rho_1 u_xu_z}_x = 0$ in the
+above for clarity.
+
+Finally, it may be noted that horizontally averaged horizontal momentum
+$p_x = \ev*{\rho u_x}_x$ obeys $\pd{p_x}{t} + \pd{F_{px}}{z} = 0$, and since
+flux is transported on the group velocity timescale, this simplifies to the
+known expression for the wave-induced horizontal mean flow (cite)
+\begin{equation}
+    \bar{U}_0 \equiv \ev*{u_x}_x = \frac{\ev*{u_xu_z}_x}{c_{gz}}
+\end{equation}
+We've denoted $c_{gz} = \at{\pd{\omega}{k_{z}}}_{k_{z0}}$ the vertical group
+velocity. In terms of our forcing amplitude,
+\begin{equation}
+    \bar{U}_0(F) = -\frac{A(F)^2}{2}e^{\frac{z - z_0}{H}} \frac{k_z}{k_x}c_{gz},
+        \label{eq:u0_lin}
+\end{equation}
+where $c_{gz} = -\frac{Nk_{0x}k_{0z}}{\p*{k_{0x}^2 + k_{0z}^2 +
+\frac{1}{4H^2}}^{3/2}}$.
+
+$\omega$ is chosen by inverting $\omega(k_x, k_z)$ dispersion relation for fixed
+$k_x = 2\pi / L_x$ and some desired $k_z = -2\pi/H$, and $\sigma \lesssim
+\frac{1}{k_z}$ is used to excite a broad band of modes including the desired
+$k_z$ mode.
+
+\subsection{Numerical Simulation}\label{ss:lin_ns}
+
+We simulate using Dedalus (cite) a spectral numerical method. We use a Fourier
+basis in the $x$ direction and a Chebyshev in the $z$ direction. Our simulation
+starts with all perturbation quantities at zero, an initial condition at rest.
+Our parameters are chosen to be as similar as possible to the later nonlinear
+simulations (\autoref{s:nonlin}). We choose as follows:
+\begin{itemize}
+    \item $k_{0x}$: Astrophysical IGWs have $k_x \ll k_z$, so we choose $k_{0x}
+        = \frac{2\pi}{L_x}$ the smallest permitted wavenumber permitted by
+        periodic boundary conditions.
+
+    \item We choose $\omega_0$ by choosing it to produce a desired $k_{0z}$.
+        Astrophysical systems generally also exhibit $k_zH \gg 1$, where the
+        stratification height is significantly larger than the vertical
+        wavelength. However, in order for the simulation to both be well-damped
+        at grid resolutions and be neglibily damped at leading order nonlinear
+        wavelengths, we require $k_{0z} \ll \frac{1}{L_{NL}} \ll \frac{2\pi
+        N_z}{L_z}$, where $L_{NL}$ denotes some characteristic length scale of
+        nonlinear features. Finally, we need $L_z$ to be many $H$ to capture
+        significant $e^{z/2H}$ growth and separate the linear and nonlinear
+        amplitudes of $u_z$.
+
+        Fixing $L_z = 10H$ to give $\sim e^4$ amplitude growth between the
+        damping zones, this implies we require separation of scales $1 \ll
+        k_{0z}H \ll \frac{H}{L_{NL}} \ll \frac{2\pi N_z}{10}$.
+
+        Since $N_z$ is fixed by computational cost (we tried both $N_z = 512,
+        N_z = 1024$), we choose $k_{0z} \sim \frac{2\pi}{H}$. This is less
+        physically representative but gives more wavenumber space for the
+        nonlinear cascade. We then invert $\omega_0 = \omega(k_{0x}, k_{0z})$.
+
+    \item We choose $\nu$ $1/10$ that of the nonlinear simulation, to verify
+        that the undamped linear solution is recovered.
+
+    \item We choose $F$ forcing strength $1/20$ that of the nonlinear
+        simulation, which keeps the flow amplitude sufficiently small to be
+        treated in the linear approximation at all points in the simulation
+        domain.
+\end{itemize}
+
+A representative snapshot of this simulation
+is provided in \autoref{fig:linear_nonu}. Key features to note:
+\begin{itemize}
+    \item $u_z$ grows with \autoref{eq:uz_lin} to reasonable accuracy.
+
+    \item $F_{px}$ is seen to be constant, as would be expected by
+        \autoref{eq:fpx_lin}.
+
+    \item $\bar{U}_0$ agrees with \autoref{eq:u0_lin} at lower $z$ but deviates
+        significantly at higher $z$. This deviation is caused by the top damping
+        zone that does not conserve horizontal momentum.
+\end{itemize}
+We also include a qualitatively similar snapshot of the same simulation run
+using the $\nu$ from the later nonlinear simulations in \autoref{fig:linear_nu}.
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.5\textwidth]{plots/linear_nonu.png}
+    \caption{Snapshot of linear simulation after reaching steady state.
+        Analytical predictions of $\abs*{u_z}, \bar{U}_0, F_{px}$ are shown in
+        orange. Red dotted lines indicate the onset of the damping region and
+        blue dotted lines denote the driving region.}\label{fig:linear_nonu}
+\end{figure}
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.5\textwidth]{plots/linear_nu.png}
+    \caption{Same as \autoref{fig:linear_nonu} but with $0.3\times $viscosity
+    used for nonlinear simulations (YUBONOTE should be the same but that
+    simulation mysteriously stopped without me noticing; restarting worked, so
+    currently rerunning). Note the slight resulting dissipation in $F_{px}$.
+    }\label{fig:linear_nu}
+\end{figure}
+
+\section{Nonlinear Regime}\label{s:nonlin}
+
+\subsection{Mean Flow Critical Layer Absorption}
+
+In past studies of IGWs in WDs (cite Fuller \& Lai), $\xi_z \equiv
+\frac{u_z}{\omega_0}$ the Lagrangian fluid displacement was often used towards
+wave breaking criterion $k_{0z}\xi_z \gtrsim 1$. We argue that the wave's
+self-interaction via its generated mean flow $\bar{U}_0$ induces total
+absorption when the mean flow exceeds critical value
+\begin{equation}
+    \bar{U}_c = \frac{\omega_0}{k_{0x}}.
+\end{equation}
+This is consistent with the picture put forth in e.g.\ Goldreich and Nicholson
+(cite).
+
+A purely horizontal shear flow $\bar{U}_0(z) \hat{x}$ can be seen in
+\autoref{se:fc_orig} to have the effect of modifying time derivatives
+$\partial_t$ to their frequency in the comoving frame of the fluid $\partial_t -
+\bar{U}_0(z)\partial_x$. For a critical value $\omega_0 - \bar{U}_c k_{0x} = 0$,
+the frequency of the linear wave in the fluid's frame of reference vanishes and
+critical behavior is observed. In a linear theory or a theory where small scales
+are viscosity rather than advection dominated, the incident wave has amplitude
+reflection and transmission coefficients
+\begin{align}
+    \mathcal{R} &= e^{-2\pi \sqrt{\mathrm{Ri} - \frac{1}{4}}}, &
+    \mathcal{T} &= e^{-\pi \sqrt{\mathrm{Ri} - \frac{1}{4}}},
+    \label{eq:crit_coeffs}
+\end{align}
+where we have defined Richardson number $\mathrm{Ri} \equiv
+\at{\frac{N^2}{\p*{\pd{\bar{U}_0}{z}}^2}}_{z_c}$ at the critical layer $z_c:
+\bar{U}_0(z_c) = \frac{\omega_0}{k_{0x}}$. For most shear flows, $\mathrm{Ri}
+\gg 1$ and so $\mathcal{R}, \mathcal{T} \ll 1$ and the incident wave is
+absorbed.
+
+When the fluid absorbs the incident wave, it absorbs the incident horizontal
+momentum flux as well, which is converted into additional horizontal momentum of
+the shear flow. Since the shear flow cannot exceed $\bar{U}_c$ the horizontal
+phase velocity of the incident wave, the critical layer must thus propagate
+downwards (towards the wave source) to accommodate the incident momentum. In
+other words, the total horizontal momentum of the shear flow obeys conservation
+equation
+\begin{equation}
+    \pd{}{t}\int\limits_0^{L_z} \rho(z) \bar{U}_0(z, t)\;\mathrm{d}z
+        - F_{px} = 0.
+\end{equation}
+Treating $\bar{U}_0(z > z_c) = \bar{U}_c, \bar{U}(z < z_c) = 0$ gives us exactly
+\begin{equation}
+    \rho \bar{U}_c\pd{z_c}{t} = -F_{px}.\label{eq:zc_anal}
+\end{equation}
+
+\subsection{Numerical Simulation}
+
+We use the same $k_{0x}, \omega_0$ as \autoref{ss:lin_ns}. Our other parameters
+are:
+\begin{itemize}
+    \item We choose $\nu = 0.35 \frac{\omega_0}{k_{0z}k_{z, \max}}$, where
+        $k_{z, \max} = \frac{2\pi N_z}{L_z}$. Note that $\nu =
+        \frac{\omega_0}{k_{0z} k_{z, \max}}$ corresponds to the advective term
+        $\vec{u} \cdot \vec{\nabla}$ being of the same order as the time
+        derivative $\partial_t$ for flow velocities $\vec{u} \sim
+        \frac{\omega_0}{k_{0z}}$ at the grid spacing.
+
+    \item We choose $F$ such that $\bar{U}_0(z)$ predicted by
+        \autoref{eq:u0_lin} exceeds $\bar{U}_c$ critical velocity for $z <
+        L_z$, i.e.\ the wave-induced mean flow is sufficiently large within the
+        simulation domain to induce critical layer absorption.
+\end{itemize}
+
+Two representative snapshots from our simulation are provided in
+\autoref{fig:nl} after the critical layer has had time to form. We may note that
+the critical layer, where $F_{px}$ is absorbed and $\bar{U}_0 = \bar{U}_c$,
+travels downwards as predicted.
+\begin{figure}[h]
+    \centering
+    \begin{subfigure}{0.5\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{plots/nl35_1.png}
+        \caption{Earlier time snapshot. Legend is the same as
+        \autoref{fig:linear_nonu} except $\bar{U}_c$ is marked in red on the
+        third panel.}
+    \end{subfigure}
+
+    \begin{subfigure}{0.5\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{plots/nl35_2.png}
+        \caption{Later snapshot illustrating propagation of $z_c$.}
+    \end{subfigure}
+    \caption{Nonlinear numerical simulation.}\label{fig:nl}
+\end{figure}
+
+\subsection{Propagating Critical Layer}
+
+For the simulation in \autoref{fig:nl}, we may define $z_c = \argmax_z
+\pd{\bar{U}_0}{z}$. Computing $\pd{z_c}{t}$ using the analytical flux
+\autoref{eq:fpx_lin} allows comparison to \autoref{eq:zc_anal}, which we exhibit
+in \autoref{fig:zc_anal}. We see overall good agreement, though some small
+deviation is expected since $F_{px}$ is not perfectly conserved owing to
+numerical viscosity (and YUBONOTE $F_{px}$ is misestimated?).
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.5\textwidth]{plots/nl35_front_v.png}
+    \caption{Comparison of simulated $\pd{z_c}{t}$ to analytical
+    \autoref{eq:zc_anal}. Plot begins after formation of critical layer in
+    simulation.}\label{fig:zc_anal}
+\end{figure}
+
+However, recalling \autoref{eq:crit_coeffs}, we only expect complete critical
+layer absorption when $\mathrm{Ri} \gg \frac{1}{4}$. We may plot $\mathrm{Ri}$
+over the same time interval in \autoref{fig:nl35_f_ri}. We may observe that the
+Richardson number is initially decreasing but eventually levels out and
+begins to increase again. This corresponds to an increasingly sharp critical
+layer transition then subsequently a decreasingly sharp critical layer
+transition.
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.5\textwidth]{plots/nl35_f_ri.png}
+    \caption{Plot of $\mathrm{Ri}_{\max}(t) = \max_z
+    \frac{N^2}{\bar{U}_0'}(t)$ over the simulation time.}\label{fig:nl35_f_ri}
+\end{figure}
+
+We argue that the Richardson number is bounded from below by viscosity. A
+repeated simulation with a larger viscosity is shown in \autoref{fig:nl1_f_ri},
+where the Richardson number does not go nearly as low.
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.5\textwidth]{plots/nl1_f_ri.png}
+    \caption{Same as \autoref{fig:nl35_f_ri} but with $\sim 1.3\times$
+    viscosity.}\label{fig:nl1_f_ri}
+\end{figure}
+
+Additionally, a lower-viscosity simulation shows significant features resembling
+Kelvin-Helmholtz instabilities (KHI), \autoref{fig:nl_full_01}. As we have only
+plotted $\bar{U}_0(z)$, the local Richardson number $N^2 / \pd{u_x}{z}$ may
+exceed the plotted value and incur KHIs.
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.5\textwidth]{plots/nl_full_01.png}
+    \caption{Low viscosity simulation showing seemingly KHI-like
+    features.}\label{fig:nl_full_01}
+\end{figure}
+
+This seems to suggest that our reference simulation is viscosity limited. This
+is perhaps physical: a Richardson number $\sim 0.25$ incurs the KHI which saps
+energy and angular momentum from the flow. It is conceivable then that our
+numerical viscosity is physically motivated as a effective KHI viscosity.
+
+\section{Boussinesq Local Simulation (Speculative)}\label{s:bouss}
+
+Alternatively, we can attempt direct numerical simulation probing lower
+Richardson numbers. We will move to the Boussinesq approximation and set up a
+mean flow as an initial condition. This relaxes the constraint that the domain
+must be sufficiently large to permit substantial $e^{z/H}$ growth of the mean
+flow $\bar{U}_0$ and allows us to probe much finer vertical resolution.
 
@@
         \pd{\rho_1}{t} + \p*{\vec{u} \cdot \vec{\nabla}}\rho_1
-            - \frac{\rho_0 u_z}{H} &= 0,\\
+            - N^2\frac{\rho_0 u_z}{g} &= 0,\\
         \pd{\vec{u}}{t} + \p*{\vec{u} \cdot \vec{\nabla}}\vec{u}
             + \frac{\vec{\nabla}P_1}{\rho_0}
-            + \frac{\rho_1 \vec{g}}{\rho_0} &= 0.
+            + \frac{\rho_1 g\hat{z}}{\rho_0} &= 0.
     \end{align}
 \end{subequations}
-In these equations, we take $\rho_0(z) = \rho_0$ a constant reference density.
+In these equations, we take $\rho_0(z) = \rho_0 = 1$ a constant reference
+density. We nondimensionalize $\rho_0 = N = g = 1$. By using a subgrid model
+(details \autoref{ss:bouss_impl}), we are able to probe significantly smaller
+Richardson numbers.
 
-Including numerical terms and driving terms, my full equations are
-\begin{subequations}\label{se:bouss_num}
-    \begin{align}
-        \vec{\nabla} \cdot \vec{u}_1 &= 0,\\
-        \pd{\rho_1}{t} - \frac{\rho_0 u_z}{H}
-            - \nu \nabla^6 \rho_1
-            &= -\Gamma(z) \rho_1
-                - \p*{\vec{u} \cdot \vec{\nabla}}\rho_1
-                + Fe^{-\frac{(z - z_0)^2}{2\sigma^2}}
-                    \cos \p*{k_xx - \omega t},\\
-        \pd{u_x}{t} + \frac{\partial_x P}{\rho_0}
-            - \nu \nabla^6 u_x
-            &= -\Gamma(z) u_x
-                - \p*{\vec{u} \cdot \vec{\nabla}}u_x,\\
-        \pd{u_z}{t} + \frac{\partial_z P}{\rho_0}
-            + \frac{\rho_1 g}{\rho_0}
-            - \nu \nabla^6 u_z
-            &= -\Gamma(z) u_z
-                - \p*{\vec{u} \cdot \vec{\nabla}}u_z,\\
-        \Gamma(z) &= 5\s*{2 + \tanh \frac{z - z_T}{(z_{\max} - z_T) / 2}
-            + \tanh \frac{z_B - z}{z_B / 2}},
-    \end{align}
-\end{subequations}
-where $z_B = 0.07z_{\max}, z_T = 0.93z_{\max}$ are the boundaries of the damping
-zones, and my domain is $z \in [0, z_{\max}]$. The forcing term $F$ is chosen to
-be weakly nonlinear, $\omega$ is chosen by inverting $\omega(k_x, k_z)$
-dispersion relation for fixed $k_x = 2\pi / x_{\max}$ and some desired $k_z \ll
-H, z_{\max}$ (I've chosen $k_z \approx -\frac{2\pi}{z_{\max} / 5}$ here), and
-$\sigma \lesssim \frac{1}{k_z}$ is used to excite a broad band of modes
-including the desired $k_z$ mode.
+\subsection{Low-Resolution Simulation}
 
-\subsection{Stratification}
+The parameters of the simulation differ slightly; since we do not have
+exponential density stratification to induce $\bar{U}_0 \propto e^{z/H}$ growth,
+we initialize the system with an initial $\bar{U}_0(z) \leq \bar{U}_c$,
+simulating a later phase of the stratified simulation where the critical layer
+has propagated deep into the domain.
 
-We simulate an incompressible, isothermal fluid, representatitive of degenerate
-matter in WD bulks. We assume a barotropic equation of state to simplify for the
-time being. The physical equations for an incompressible, barotropic fluid are:
-\begin{subequations}\label{se:fc_orig}
-    \begin{align}
-        \vec{\nabla} \cdot \vec{u} &= 0,\\
-        \pd{\rho}{t} + \vec{u} \cdot \vec{\nabla}\rho &= 0,\\
-        \pd{\vec{u}}{t} + \p*{\vec{u} \cdot \vec{\nabla}}\vec{u}
-            + \frac{\vec{\nabla}P}{\rho}
-            + g\hat{z} &= 0.
-    \end{align}
-\end{subequations}
-We will introduce variables $T = P/\rho$. We then mandate $\rho_0, T_0$
-backgrounds satisfy hydrostatic equilibrium $\vec{\nabla}T_0 + T_0
-\vec{\nabla}\rho_0 + \vec{g} = 0$. Taking isothermal stratification, we find
-$T_0 = gH$. Making then substitution of variables $\Upsilon = \ln \rho - \ln
-\rho_0$ and $T_1 = T - T_0$ deviations from the background state, the exact
-fluid equations in the new variables are:
+Sample snapshots from a reference simulation are \autoref{fig:bouss_1}.
+\begin{figure}[h]
+    \centering
+    \begin{subfigure}{0.5\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{plots/vstrat_1.png}
+        \caption{Boussinesq simulation initial condition.}
+    \end{subfigure}
+
+    \begin{subfigure}{0.5\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{plots/vstrat_2.png}
+        \caption{At later time, when critical layer absorption has begun.}
+    \end{subfigure}
+
+    \begin{subfigure}{0.5\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{plots/vstrat_3.png}
+        \caption{At an even later time, a strong reflective character seems to
+        emerge as the Richardson number $\to 0.5$.}
+    \end{subfigure}
+    \caption{Boussinesq simulation snapshots.}\label{fig:bouss_1}
+\end{figure}
+The resulting front velocity and Richardson number are depicted in
+\autoref{fig:bouss}. Most important to note are the significantly diminished
+critical layer velocity compared to the prediction (owing to critical layer
+reflection) and the Richardson number which approaches the critical $1/4$ value
+for KHI-instability.
+\begin{figure}[t]
+    \centering
+    \begin{subfigure}{0.5\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{plots/vstrat_ri.png}
+        \caption{Richardson number evolution over time.}
+    \end{subfigure}
+
+    \begin{subfigure}{0.5\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{plots/vstrat_front_v.png}
+        \caption{Front velocity compared to analytical \autoref{eq:u0_lin}
+        prediction.}
+    \end{subfigure}
+    \caption{Boussinesq data analysis plots.}\label{fig:bouss}
+\end{figure}
+
+YUBONOTE\@: The next step would be to run simulations at sufficient resolution to
+resolve the KHI (only possible in the Boussinesq system, where we are not
+required to have a sufficiently large vertical domain to capture exponential
+growth and can choose ever smaller vertical domains); I have started to try this
+but haven't resolved the resultant numerical issues yet.
+
+\clearpage
+\onecolumngrid
+\appendix
+
+\section{Equation Implementations}
+
+We denote $x \in [0, L_x], z \in [0, L_z]$ the simulation domain and $N_x, N_z$
+the number of spectral modes in the respective dimensions.
+
+\subsection{Stratified Fluid}\label{ss:strat_impl}
+
+We introduce variable $T = P/\rho$. We then mandate $\rho_0, T_0$ background
+fields satisfy hydrostatic equilibrium $\vec{\nabla}T_0 + T_0 \vec{\nabla}\rho_0
++ \vec{g} = 0$. Taking isothermal stratification, we find $T_0 = gH$. Making
+then substitution of variables $\Upsilon = \ln \rho - \ln \rho_0$ and $T_1 = T -
+T_0$ deviations from the background state, the exact fluid equations in the new
+variables are:
 \begin{subequations}\label{se:fc_var}
@@
                 - \p*{\vec{u} \cdot \vec{\nabla}}u_z
-                - T_1 \pd{\Upsilon}{z}.
+                - T_1 \pd{\Upsilon}{z},\\
+    \Gamma(z) &= 7.5\s*{2 + \tanh \frac{z - z_T}{(L_z - z_T) / 2}
+        + \tanh \frac{z_B - z}{z_B / 2}},\label{eq:Gamma}
     \end{align}
 \end{subequations}
-$\Gamma(z)$ is as before, but we use $k_z = -2\pi/H$ here, and $\omega$ in the
-forcing term accordingly.
+where $z_B = 0.05L_z, z_T = 0.95L_z$ are the boundaries of the damping zones. A
+Navier-Stokes numerical viscosity $\nu$ is used to damp high wavenumbers and
+regularize the nonlinear cascade at near grid resolution: $\nu \sim 0.2
+\frac{\omega}{\abs*{k_z}}\frac{L_z}{2\pi N_z}$ was found to be suitable for $N_z
+= 1024$, where $k_z$ is the wavenumber of the excited linear mode.
+
+\subsection{Boussinesq Fluid}\label{ss:bouss_impl}
+
+In the Boussinesq system, there is no need to transform to $T, \Upsilon$
+variables as $\rho_0$ is constant. Including numerical terms and driving terms,
+our full equations are
+\begin{subequations}\label{se:bouss_num}
+    \begin{align}
+        \vec{\nabla} \cdot \vec{u}_1 &= 0,\\
+        \pd{\rho_1}{t} - \frac{\rho_0 u_z}{H}
+            - \nu \nabla^6 \rho_1
+            &= -\Gamma(z) \rho_1
+                - \p*{\vec{u} \cdot \vec{\nabla}}\rho_1
+                + Fe^{-\frac{(z - z_0)^2}{2\sigma^2}}
+                    \cos \p*{k_xx - \omega t},\\
+        \pd{u_x}{t} + \frac{\partial_x P}{\rho_0}
+            - \nu \nabla^6 u_x
+            &= -\Gamma(z) u_x
+                - \p*{\vec{u} \cdot \vec{\nabla}}u_x,\\
+        \pd{u_z}{t} + \frac{\partial_z P}{\rho_0}
+            + \frac{\rho_1 g}{\rho_0}
+            - \nu \nabla^6 u_z
+            &= -\Gamma(z) u_z
+                - \p*{\vec{u} \cdot \vec{\nabla}}u_z.
+    \end{align}
+\end{subequations}
+$\Gamma(z)$ is as \autoref{eq:Gamma}. We use subgrid model $\nabla^6$ (cite) to
+decrease the effect of viscosity on scales above the grid resolution.
 

### 2018-12-06 sims/2d_1_vstrat/strat_helper.py

@@ -448,28 +448,42 @@ def plot_front(name, params):
     else:
         print('data.log found, loading')
         with open(logfile) as data:
-            sim_times = eval(data.readline())
-            front_pos = eval(data.readline())
-            ri_inv = eval(data.readline())
-            fluxes = eval(data.readline())
+            sim_times = eval(data.readline())[30: ]
+            front_pos = eval(data.readline())[30: ]
+            ri_inv = eval(data.readline())[30: ]
+            fluxes = eval(data.readline())[30: ]
+    flux_anal = (params['F'] * get_uz_f_ratio(params))**2 / 2 \
+         * np.ones(np.shape(fluxes)) * abs(params['KZ'] / params['KX'])
     plt.plot(sim_times, front_pos)
-    plt.ylabel('Front Position')
+    plt.ylabel('Critical Layer Position')
     plt.xlabel('Time')
     plt.title(name)
     plt.savefig('%s/front.png' % snapshots_dir, dpi=200)
     plt.clf()
 
-    plt.plot(sim_times, ri_inv)
-    plt.ylabel('1/Ri')
+    front_v = np.gradient(front_pos) / np.gradient(sim_times)
+    anal_v = -flux_anal / \
+        (params['RHO0'] * np.exp(-np.array(front_pos) / params['H'])) * \
+        params['KX'] / params['OMEGA']
+    plt.plot(sim_times, front_v, label='Data')
+    plt.plot(sim_times, anal_v, label='Analytic')
+    plt.ylabel('Critical Layer Velocity')
+    plt.xlabel('Time')
+    plt.ylim([min(min(front_v), min(anal_v)), 0])
+    plt.legend()
+    plt.title(name)
+    plt.savefig('%s/front_v.png' % snapshots_dir, dpi=200)
+    plt.clf()
+
+    plt.plot(sim_times, 1 / np.array(ri_inv)**2)
+    plt.ylabel('Ri')
     plt.xlabel('Time')
     plt.title(name)
     plt.savefig('%s/f_ri.png' % snapshots_dir, dpi=200)
     plt.clf()
 
     plt.plot(sim_times, np.array(fluxes) * 1e6)
-    plt.plot(sim_times, (params['F'] * get_uz_f_ratio(params))**2 / 2
-             * np.ones(np.shape(fluxes)) * abs(params['KZ'] / params['KX'])
-             * 1e6)
+    plt.plot(sim_times,flux_anal * 1e6)
[…]

### 2018-12-06 sims/2d_2_fc/strat2.py

@@ -68,10 +68,10 @@ def run(ic, name, params_dict):
 
 if __name__ == '__main__':
     tasks = [
-       (set_ic, 'nl1',
-        build_interp_params(2, 2, overrides={'F_MULT': 1,
-                                             'Re': 0.5,
-                                             'USE_CFL': True})),
+       # (set_ic, 'nl1',
+       #  build_interp_params(2, 2, overrides={'F_MULT': 1,
+       #                                       'Re': 0.5,
+       #                                       'USE_CFL': True})),
        # (set_ic, 'linear1',
        #  build_interp_params(2, 2, overrides={'F_MULT': 0.05,
        #                                       'Re': 0.5})),
@@ -81,7 +81,7 @@ if __name__ == '__main__':
        #                                       'USE_CFL': True})),
        # (set_ic, 'linear0',
        #  build_interp_params(2, 2, overrides={'F_MULT': 0.05,
-       #                                       'Re': 0.01})),
+       #                                       'Re': 0.0001})),
        # (set_ic, 'nl_full',
        #  build_interp_params(1, 1, overrides={'F_MULT': 1,
        #                                       'Re': 0.7,
@@ -113,14 +113,14 @@ if __name__ == '__main__':
        # (set_ic, 'linear_01',
        #  build_interp_params(2, 2, overrides={'F_MULT': 0.05,
        #                                       'Re': 0.1})),
-       # (set_ic, 'nl_full_01',
-       #  build_interp_params(1, 1, overrides={'F_MULT': 1,
-       #                                       'Re': 0.2,
-       #                                       'USE_CFL': True})),
-       # (set_ic, 'nl_full_02',
-       #  build_interp_params(1, 1, overrides={'F_MULT': 1,
-       #                                       'Re': 0.4,
-       #                                       'USE_CFL': True})),
+       (set_ic, 'nl_full_01',
+        build_interp_params(1, 1, overrides={'F_MULT': 1,
+                                             'Re': 0.2,
+                                             'USE_CFL': True})),
+       (set_ic, 'nl_full_02',
+        build_interp_params(1, 1, overrides={'F_MULT': 1,
+                                             'Re': 0.4,
+                                             'USE_CFL': True})),
     ]
     if '-plot' in sys.argv:
         for _, name, params_dict in tasks:

### 2018-12-06 sims/2d_2_fc/strat_helper.py

@@ -280,15 +280,20 @@ def plot(name, params):
         return ret
 
     plot_cfgs = [
+        # {
+        #     'save_fmt_str': 'p_%03i.png',
+        #     'z_vars': ['F_px', 'ux'],
+        #     'slice_vars': ['uz'],
+        #     # 'sub_vars': ['ux'],
+        # },
+        # {
+        #     'save_fmt_str': 'm_%03i.png',
+        #     'plot_vars': ['ux', 'uz', 'rho', 'P'],
+        # },
         {
-            'save_fmt_str': 'p_%03i.png',
-            'z_vars': ['ux', 'F_px', 'ux_z'],
+            'save_fmt_str': 't_%03i.png',
+            'z_vars': ['F_px', 'ux'],
             'slice_vars': ['uz'],
-            'sub_vars': ['ux'],
-        },
-        {
-            'save_fmt_str': 'm_%03i.png',
-            'plot_vars': ['ux', 'uz', 'rho', 'P'],
         },
     ]
 
@@ -377,6 +382,17 @@ def plot(name, params):
                         'orange',
                         linewidth=0.5)
 
+                if var == 'F_px%s' % sum_suffix:
+                    p = axes.plot(
+                        (params['F'] * get_uz_f_ratio(params))**2 / 2
+                            * abs(params['KZ'] / params['KX'])
+                            * params['RHO0']
+                                * np.exp(-params['Z0'] / params['H'])
+                            * np.ones(np.shape(z_pts)),
+                        z_pts,
+                        'orange',
+                        linewidth=0.5)
+
                 if var == 'ux%s' % sum_suffix:
                     # mean flow = E[ux * uz] / V_GZ
                     p = axes.plot(
@@ -385,6 +401,13 @@ def plot(name, params):
                         z_pts,
                         'orange',
                         linewidth=0.5)
+                    # critical = omega / kx
+                    p = axes.plot(
+                        params['OMEGA'] / params['KX']
+                            * np.ones(np.shape(z_pts)),
+                        z_pts,
+                        'red',
+                        linewidth=0.5)
 
                 plt.xticks(rotation=30)
                 plt.yticks(rotation=30)
@@ -392,11 +415,13 @@ def plot(name, params):
                 axes.set_xlim(*xlims)
                 axes.set_ylim(z_pts.min(), z_pts.max())
[…]

### 2018-12-30 writeups/paper.tex
stash

--- 2018-12-06 writeups/running_writeup.tex
@@
     \documentclass[twocolumn,
+        nofootinbib,
         usenames, % allows access to some tikz colors
@@
 
-\title{IGWs and WDs Ongoing Writeup}
-\author{Yubo Su}
+\begin{abstract}
+    In white dwarf binaries tidal torques are expected to raise a train of
+    internal gravity waves in the white dwarf interior. These
+    outward-propagating waves can then grow to be nonlinear and break via
+    hydrodynamical instabilities, transfering energy and angular momentum from
+    the binary orbit to the white dwarf. We perform 2D numerical simulations of
+    nonlinear wave breaking of outwards-propagating internal gravity waves in an
+    incompressible, isothermal atmosphere with an exponentially decaying density
+    stratification. We find that tidal synchronization, after an initial
+    transient phase, proceeds inwards from the surface in thin layers. We argue
+    the thickness of these layers is limited by the Kelvin-Helmholtz
+    Instability. We provide simple analytical formulae for the location and
+    thickness of tidal synchronization over time that are in good agreement with
+    our simulations.
+\end{abstract}
+
+\title{Tidal Heating via Internal Gravity Waves in White Dwarf Binaries}
+\author{Authors}
+\date{\today}
 \maketitle
@@
 
-WIP ongoing writeup, last updated \today.
+% Usual WD binaries introduction, from proposals
 
-% Define white dwarfs (WD), internal gravity waves (IGW), citations
+% IGW tidal dissipation has been thoroughly studied in many astrophysical
+% binary systems (Barker Ogilvie, Goldreich Nicholson, Zahn?). Its effect in WDs
+% (Fuller Lai) resembles early solar type stars but has not been studied
+% numerically. While the linear IGW solution is an exact solution of the
+% nonlinear system (Drazin) but is globally unstable (Drazin), the IVP must be
+% studied numerically to determine how the linear solution breaks down and
+% deposits energy/angular momentum.
 
-Basic narrative is: the prevailing understanding is that mean flow absorption
-slowly spins up a WD from the outside in as the critical layer propagates
-inwards. Our numerical simulations validate this prediction and give an
-analytical prediction for the motion of the critical layer. However, it appears
-our numerical simulations are viscosity limited. We present a brief argument for
-why the viscosity limitations are insignificant, but also present some
-exploratory, inconclusive work towards direct numerical simulation resolving
-critical layer behavior without being viscosity limited.
+% From numerical studies, IGWs are known to be anti-diffusive (Lecoanet mean
+% flow papers, Lindzen) and sharpen mean flows (m = 0). Mean flows produce
+% reflection, known both analytically (Booker & Bretherton) and numerically
+% (Winters d'Asaro). IGW self-interaction via its wave-induced mean flow is
+% known to play a crucial role in IGW wavebreaking (Sutherland x2).
 
-In \autoref{s:eq}, we will cover the equations and numerical setup. In
-\autoref{s:lin}, we will discuss the agreement of our simulation in the
-low-amplitude limit with analytical linear theory. In \autoref{s:nonlin} we will
-present a few reference simulations illustrating critical layer absorption
-depositing horizontal angular momentum into the fluid. Finally, in
-\autoref{s:bouss} we will consider a speculative series of simulations
-attempting to rectify the seemingly viscosity-limited simulations of the
-previous section.
+% We study IGW breaking in the plane-parallel approximation in 2D. Known to
+% describe both wave steepening (Kloyermeister 1991) and captures mean flow
+% absorption (Booker & Bretherton). While the details of breaking are presently
+% understood to be 3D (Winters & d'Asaro), we argue in (section) that our
+% results should be qualitatively similar to 3D (both 2D and 3D exhibit KHI, and
+% since we are able to resolve KHI, the spinup in 3D should be similar?).
 
-\section{Equations}\label{s:eq}
+In \autoref{s:num_setup}, we will cover the numerical setup. In \autoref{s:lin}, we
+will discuss the agreement of our simulation in the low-amplitude limit with
+analytical linear theory. In \autoref{s:nonlin} we will present a few reference
+simulations illustrating critical layer absorption depositing horizontal angular
+momentum into the fluid.
 
-We simulate an incompressible, isothermal fluid, representatitive of degenerate
-matter in WD bulks. We assume a barotropic equation of state as a first
-approximation. Furthermore, we model the background, hydrostatic density
-stratification as $\rho_0(x, z) = \rho_0(z) \propto e^{-z/H}$.
+\section{Problem Description}\label{s:num_setup}
 
-The physical equations for an incompressible, barotropic fluid in a uniform
+\subsection{Fluid Equations}
+
+We consider an incompressible, isothermal fluid, representative of degenerate
+matter in WD bulks, in 2D. We assume a barotropic equation of state as a first
+approximation. As we are interested in the behavior of the wave far from the
+center of the WD, we approximate the gravitational field as uniform. We model
+the background density stratification as $\rho_0(x, z) = \rho_0(z) = \rho_{ref}
+e^{-z/H}$ for some reference density $\rho_{ref} = \at{\rho_0(z)}_{z = 0}$.
+
+The Euler equations for an incompressible, barotropic fluid in a uniform
 gravitational field are
@@
         \vec{\nabla} \cdot \vec{u} &= 0,\\
-        \pd{\rho}{t} + \vec{u} \cdot \vec{\nabla}\rho &= 0
-            ,\label{eq:fc_density}\\
-        \pd{\vec{u}}{t} + \p*{\vec{u} \cdot \vec{\nabla}}\vec{u}
-            + \frac{\vec{\nabla}P}{\rho}
-            + g\hat{z} &= 0.
+        \md{\rho}{t} &= 0 ,\label{eq:fc_density}\\
+        \md{\vec{u}}{t} + \frac{\vec{\nabla}P}{\rho} + g\hat{z} &= 0.
     \end{align}
 \end{subequations}
-$\vec{u}, \rho, P$ are the fluid dynamical variables and $-g\hat{z}$ is constant
-gravitational acceleration. Note that at hydrostatic equilibrium $\partial_t =
-0$ we have $\vec{u} = 0, \vec{\nabla}P_0 = -\rho_0 g\hat{z}$ and so $P_0 =
-\rho_0 gH$.
+$\md{}{t} = \pd{}{t} + \p*{\vec{u} \cdot \vec{\nabla}}$ is the material
+derivative. $\vec{u}, \rho, P$ are the fluid dynamical variables and $-g\hat{z}$
+is constant gravitational acceleration. Note that at hydrostatic equilibrium
+$\pd{}{t} = 0$ we have $\vec{u} = 0, \vec{\nabla}P_0 = -\rho_0 g\hat{z}$ and
+so $P_0 = \rho_0 gH$.
 
@@
 \end{equation}
-is the Brunt-V\"ais\"al\"a frequency and $\rho_{ref}$ is the background density
-stratification at $z = 0$. The full numerical details of the simulation are
-described in \autoref{ss:strat_impl}.
+is the Brunt-V\"ais\"al\"a frequency. The full numerical details of the
+simulation are described in \autoref{ss:strat_impl}.
 
@@
 towards the surface without unnecessary computational cost, we excite an
-IGW wave train from the bottom of the simulation domain. As Chebyshev grid is
-a factor of $\propto N_z$ denser near the edges of the simulation domain, we
-use a volumetric forcing to excite IGW at some $z_0 > 0$ interior to the
-simulation domain. Our forcing excites both IGW propagating upwards, the
-physically interesting solution, and downwards, which are damped away by damping
-layers (see \autoref{ss:strat_impl}) before inducing strict Courant-Friedrichs-Lewy
-(CFL) timestepping constraints (cite).
+IGW wave train from the bottom of the simulation domain. We use a volumetric
+forcing term to excite IGW at some $z_0 > 0$ interior to the simulation
+domain\footnote{Interfacial forcing at the bottom boundary incurs severe CFL
+timestepping constraints, as we use a Chebyshev polynomial basis along the $z$
+axis in our spectral method which has very small grid spacing at the
+boundaries.}. Our forcing excites both IGWs propagating upwards, imitating a
+wave tidally excited deeper in the WD, and downwards, which are damped away by
+damping layers (see \autoref{ss:numerics}) before inducing strict
+Courant-Friedrichs-Lewy (CFL) timestepping constraints.
 
@@
 \begin{equation}
-    \pd{\rho}{t} + \vec{u} \cdot \vec{\nabla}\rho =
-        Fe^{-\frac{(z - z_0)^2}{2\sigma^2}}
-            \cos \p*{k_{0, x}x - \omega_0 t}.
+    \md{\rho}{t} = Fe^{-\frac{(z - z_0)^2}{2\sigma^2}}
+            \cos \p*{k_{0x}x - \omega_0 t}.
 \end{equation}
-Define $k_{0, z}: \omega(k_{0, x}, k_{0, z}) = \omega_0$. If $k_{0, z}\sigma \ll
-1$, the forcing term has the effect of picking out $k_{0, z}$ driven by
-$\cos\p*{k_{0, x}x - \omega_0 t}$ from the broad $k_z$ wavenumber spectrum
-permitted by the narrow Gaussian profile.
+Using a narrow Gaussian profile excites a broad $k_z$ wavenumber spectrum, and
+only the $k_{0z}$ satisfying the IGW's dispersion relation for the given
+$k_{0x}, \omega_0$ will propagate.
+% Define $k_{0z}: \omega(k_{0x}, k_{0z}) = \omega_0$ the wavenumber picked
+% out by $k_{0x}, \omega_0$. If $k_{0z}\sigma \ll 1$, the $\cos\p*{k_{0x}x -
+% \omega_0 t}$ term permits only the $k_{0z}$ mode to propagate among the broad
+% $k_z$ wavenumber spectrum permitted by the narrow Gaussian profile.
 
-\section{Linear Regime}\label{s:lin}
-
-We first test our numerical simulation with a ``linear'' simulation, where
-all flow quantities are small: we assume  $\rho_1 \equiv \rho - \rho_0 \ll
-\rho_0$ and $\vec{u} \cdot \vec{\nabla} \ll \partial_t$. We clarify that our
-``linear'' simulations solve the full nonlinear equations, just in the weak
-perturbation limit.
-
-\subsection{Analytical Solution}
+\subsection{Internal Gravity Waves}
 
@@
 We have substituted $\cos\p*{k_xx - \omega_0t}$ for a complex exponential for
-easier manipulation; taking real parts of all dynamical variables above recovers
-the physical solution.
+easier manipulation.
 
@@
 using $e^{\frac{-(z - z_0)^2}{2\sigma^2}} \approx \sqrt{2\pi \sigma^2}\delta(z -
-z_0)$ if $\sigma \to 0$, then relax to $k_{0z}\sigma \lesssim 1$. This may be
-solved to obtain the solution
+z_0)$ if $\sigma \to 0$. This system is solved exactly by matching the two
+homogeneous solutions above and below $z_0$, We may then approximately relax the
+solution to $k_{0z}\sigma \lesssim 1$. This may be solved to obtain the solution
 \begin{align}
@@
 \end{align}
-The extra factor of $e^{-\frac{(k_{0z}\sigma)^2}{2}}$ compared to the
-$\delta$-function solution comes from evaluating the Fourier Transform of
+An extra factor of $e^{-\frac{(k_{0z}\sigma)^2}{2}}$ arises compared to the
+$\delta$-function solution when evaluating the Fourier Transform of
 $e^{-\frac{(z - z_0)^2}{2\sigma^2}}$ at $k_z = k_{0z}$.
+
+% Linear solution: e^z/2H, dispersion relation, phase/group velocity. Observe is
+% exact solution of nonlinear.
+% Linear solution is globally unstable to RTI though, thus not stable
+% equilibrium of IVP
+%
+% Nonlinear: Critical layer absorption (viscous = Hazel, nonlin = Winters),
+% wave-induced mean flow, horizontal momentum flux
 
@@
 \frac{1}{4H^2}}^{3/2}}$.
+
+\subsection{Numerical Parameters}\label{ss:numerics}
+
+% Description of spectral method, basis functions, initial conditions
+% Grid Size
+% Viscosity + diffusivity, damping layers (to simulate radiative BC)
+% timestepping? RK443, (CFL limiter??)
+
+\section{Linear Regime}\label{s:lin}
+
+We first test our numerical simulation with a ``linear'' simulation, where
+all flow quantities are small: we assume  $\rho_1 \equiv \rho - \rho_0 \ll
+\rho_0$ and $\vec{u} \cdot \vec{\nabla} \ll \partial_t$. We emphasize that our
+``linear'' simulations solve the full nonlinear equations, just in the weak
+perturbation limit.
 
@@
 
-\section{Boussinesq Local Simulation (Speculative)}\label{s:bouss}
-
-Alternatively, we can attempt direct numerical simulation probing lower
-Richardson numbers. We will move to the Boussinesq approximation and set up a
-mean flow as an initial condition. This relaxes the constraint that the domain
-must be sufficiently large to permit substantial $e^{z/H}$ growth of the mean
-flow $\bar{U}_0$ and allows us to probe much finer vertical resolution.
-
-The physical equations in the Boussinesq approximation are:
-\begin{subequations}\label{se:bouss}
-    \begin{align}
-        \vec{\nabla} \cdot \vec{u} &= 0,\\
-        \pd{\rho_1}{t} + \p*{\vec{u} \cdot \vec{\nabla}}\rho_1
-            - N^2\frac{\rho_0 u_z}{g} &= 0,\\
-        \pd{\vec{u}}{t} + \p*{\vec{u} \cdot \vec{\nabla}}\vec{u}
-            + \frac{\vec{\nabla}P_1}{\rho_0}
-            + \frac{\rho_1 g\hat{z}}{\rho_0} &= 0.
-    \end{align}
-\end{subequations}
-In these equations, we take $\rho_0(z) = \rho_0 = 1$ a constant reference
-density. We nondimensionalize $\rho_0 = N = g = 1$. By using a subgrid model
-(details \autoref{ss:bouss_impl}), we are able to probe significantly smaller
-Richardson numbers.
-
-\subsection{Low-Resolution Simulation}
-
-The parameters of the simulation differ slightly; since we do not have
-exponential density stratification to induce $\bar{U}_0 \propto e^{z/H}$ growth,
-we initialize the system with an initial $\bar{U}_0(z) \leq \bar{U}_c$,
-simulating a later phase of the stratified simulation where the critical layer
-has propagated deep into the domain.
-
-Sample snapshots from a reference simulation are \autoref{fig:bouss_1}.
-\begin{figure}[h]
-    \centering
-    \begin{subfigure}{0.5\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{plots/vstrat_1.png}
-        \caption{Boussinesq simulation initial condition.}
-    \end{subfigure}
-
-    \begin{subfigure}{0.5\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{plots/vstrat_2.png}
-        \caption{At later time, when critical layer absorption has begun.}
-    \end{subfigure}
-
-    \begin{subfigure}{0.5\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{plots/vstrat_3.png}
-        \caption{At an even later time, a strong reflective character seems to
-        emerge as the Richardson number $\to 0.5$.}
-    \end{subfigure}
-    \caption{Boussinesq simulation snapshots.}\label{fig:bouss_1}
-\end{figure}
-The resulting front velocity and Richardson number are depicted in
-\autoref{fig:bouss}. Most important to note are the significantly diminished
-critical layer velocity compared to the prediction (owing to critical layer
-reflection) and the Richardson number which approaches the critical $1/4$ value
-for KHI-instability.
-\begin{figure}[t]
-    \centering
-    \begin{subfigure}{0.5\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{plots/vstrat_ri.png}
-        \caption{Richardson number evolution over time.}
-    \end{subfigure}
-
-    \begin{subfigure}{0.5\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{plots/vstrat_front_v.png}
-        \caption{Front velocity compared to analytical \autoref{eq:u0_lin}
-        prediction.}
-    \end{subfigure}
-    \caption{Boussinesq data analysis plots.}\label{fig:bouss}
-\end{figure}
-
-YUBONOTE\@: The next step would be to run simulations at sufficient resolution to
-resolve the KHI (only possible in the Boussinesq system, where we are not
-required to have a sufficiently large vertical domain to capture exponential
-growth and can choose ever smaller vertical domains); I have started to try this
-but haven't resolved the resultant numerical issues yet.
-
 \clearpage
@@
 
-\subsection{Boussinesq Fluid}\label{ss:bouss_impl}
-
-In the Boussinesq system, there is no need to transform to $T, \Upsilon$
-variables as $\rho_0$ is constant. Including numerical terms and driving terms,
-our full equations are
-\begin{subequations}\label{se:bouss_num}
-    \begin{align}
-        \vec{\nabla} \cdot \vec{u}_1 &= 0,\\
-        \pd{\rho_1}{t} - \frac{\rho_0 u_z}{H}
-            - \nu \nabla^6 \rho_1
-            &= -\Gamma(z) \rho_1
-                - \p*{\vec{u} \cdot \vec{\nabla}}\rho_1
-                + Fe^{-\frac{(z - z_0)^2}{2\sigma^2}}
-                    \cos \p*{k_xx - \omega t},\\
-        \pd{u_x}{t} + \frac{\partial_x P}{\rho_0}
-            - \nu \nabla^6 u_x
-            &= -\Gamma(z) u_x
-                - \p*{\vec{u} \cdot \vec{\nabla}}u_x,\\
-        \pd{u_z}{t} + \frac{\partial_z P}{\rho_0}
-            + \frac{\rho_1 g}{\rho_0}
-            - \nu \nabla^6 u_z
-            &= -\Gamma(z) u_z
-                - \p*{\vec{u} \cdot \vec{\nabla}}u_z.
-    \end{align}
-\end{subequations}
-$\Gamma(z)$ is as \autoref{eq:Gamma}. We use subgrid model $\nabla^6$ (cite) to
-decrease the effect of viscosity on scales above the grid resolution.
-
 \end{document}

### 2018-12-30 sims/2d_3_final/strat.py

@@ -48,7 +48,7 @@ def build_interp_params(interp_x, interp_z, overrides=None):
     params['N_X'] //= interp_x
     params['N_Z'] //= interp_z
     # omega * DT << 1 is required, as is DT << 1/N = 1
-    params['DT'] = min(0.1 / OMEGA, 0.1)
+    params['DT'] = params.get('DT', min(0.1 / OMEGA, 0.1))
     params['F'] = params.get('F_MULT', 1) * \
         (TARGET_DISP_RAT * OMEGA / KZ) / get_uz_f_ratio(params) \
         * np.exp(-params['Z0'] / (2 * H))
@@ -68,22 +68,23 @@ def run(ic, name, params_dict):
 
 if __name__ == '__main__':
     tasks = [
-       (set_ic, 'lin_0',
-        build_interp_params(4, 4, overrides={'F_MULT': 0.05,
-                                             'Re': 0.01,
-                                             'T_F': 500})),
-       (set_ic, 'lin_1',
-        build_interp_params(4, 4, overrides={'F_MULT': 0.05,
-                                             'Re': 0.2,
-                                             'T_F': 500})),
+       # (set_ic, 'lin_0',
+       #  build_interp_params(4, 4, overrides={'F_MULT': 0.05,
+       #                                       'Re': 0.01,
+       #                                       'T_F': 500})),
+       # (set_ic, 'lin_1',
+       #  build_interp_params(4, 4, overrides={'F_MULT': 0.05,
+       #                                       'Re': 0.2,
+       #                                       'DT': 0.3,
+       #                                       'T_F': 500})),
        # (set_ic, 'nl_low_1',
        #  build_interp_params(2, 2, overrides={'F_MULT': 1,
        #                                       'Re': 0.5,
        #                                       'USE_CFL': True})),
-       # (set_ic, 'nl_1',
-       #  build_interp_params(1, 1, overrides={'F_MULT': 1,
-       #                                       'Re': 0.7,
-       #                                       'USE_CFL': True})),
+       (set_ic, 'nl_1',
+        build_interp_params(1, 1, overrides={'F_MULT': 1,
+                                             'Re': 0.7,
+                                             'USE_CFL': True})),
        # (set_ic, 'nl_2',
        #  build_interp_params(1, 1, overrides={'F_MULT': 1,
        #                                       'Re': 0.4,

### 2018-12-30 sims/2d_3_final/strat_helper.py

@@ -34,7 +34,7 @@ def get_vgz(g, h, kx, kz):
 
 def set_ic(name, solver, domain, params):
     snapshots_dir = SNAPSHOTS_DIR % name
-    filename = '{s}/{s}_s1.h5'.format(s=snapshots_dir)
+    filename = '{s}/{s}_s2.h5'.format(s=snapshots_dir)
 
     if not os.path.exists(snapshots_dir):
         print('No snapshots found, no IC loaded')
@@ -77,9 +77,9 @@ def get_solver(params):
     problem.add_equation('dx(ux) + uz_z = 0')
     problem.add_equation(
         'dt(U) - uz / H' +
-        '- (NU * dx(dx(U)) + NU * dz(U_z) - 2 * U_z / H)' +
+        '- NU * (dx(dx(U)) + dz(U_z) - 2 * U_z / H)' +
         '= - sponge * U' +
-        '- (ux * dx(U) + uz * dz(U)) +' +
+        '- (ux * dx(U) + uz * dz(U))' +
         '+ NU * (dx(U) * dx(U) + U_z * U_z)' +
         '+ F * exp(-(z - Z0)**2 / (2 * S**2) + Z0 / H) *' +
             'cos(KX * x - OMEGA * t)')
@@ -92,7 +92,8 @@ def get_solver(params):
         '+ NU * ux * (dx(dx(U)) + dz(U_z))' +
         '+ NU * ux * (dx(U) * dx(U) + U_z * U_z)' +
         '- 2 * NU * ux * U_z / H' +
-        '+ NU * ux * (1 - exp(-U)) / H**2' +
+        # '+ NU * ux * (1 - exp(-U)) / H**2' +
+        '+ NU * ux * U / H**2' +
         '- W * dx(U)')
     problem.add_equation(
         'dt(uz) + dz(W) + (g * H) * dz(U) - W/H' +
@@ -103,7 +104,8 @@ def get_solver(params):
         '+ NU * uz * (dx(dx(U)) + dz(U_z))' +
         '+ NU * uz * (dx(U) * dx(U) + U_z * U_z)' +
         '- 2 * NU * uz * U_z / H' +
-        '+ NU * uz * (1 - exp(-U)) / H**2' +
+        # '+ NU * uz * (1 - exp(-U)) / H**2' +
+        '+ NU * uz * U / H**2' +
         '- W * dz(U)')
     problem.add_equation('dz(ux) - ux_z = 0')
[…]

### 2018-12-30 writeups/notes.tex

@@
 
-\subsection{Gotcha: $\nu_x, \nu_z$ and Grid Spacing}
-
-If the grid resolution is uneven in the $x, z$ directions, it might be tempting
-to use different $\nu_x, \nu_z$ per the above criteria, since their $k_{turb} =
-k_{\max}$ differ. This is dangerous however, since $\nu \nabla^2 \vec{u}_1$ has
-to be divergenceless! Thus, it is more advisable to choose an equal grid spacing
-in the two directions.
+\subsection{Choices of $k_x, k_z$ and Scales}
+
+Lots of scales to consider:
+\begin{itemize}
+    \item $k_zH \lesssim 1$ is required for the density stratification to be
+        considered weak.
+
+    \item $k_z \ll k_{z, \max} \equiv \frac{2\pi N_z}{L_z}$ is required for
+        there to be significant separation of scales between the linear mode,
+        the nonlinear cascade and the grid resolution.
+
+    \item $L_Z \gtrsim 10H$ or a few, to get a decent amplitude growth going
+        from linear to nonlinear amplitudes.
+
+    \item Viscosity induced decay length scale: $\partial_t \sim \omega + i\nu
+        k^2$ where $k^2 = k_x^2 + k_z^2$, so the characteristic length scale on
+        which the wave decays is $L_\nu \sim \frac{v_{ph, z}}{\nu k_z^2}$. Call
+        $\tilde{\nu} = \nu / \frac{\omega}{k_zk_{z, \max}}$, \lstinline{NU_MULT}
+        in our simulations, then we may simplify
+        \begin{equation}
+            L_\nu \sim \frac{k_z^2k_{z, \max}}{\tilde{\nu}k^4}.
+        \end{equation}
+        Approximating $k_z \approx k$ and requiring $L_\nu \gtrsim L_Z$ enforces
+        \begin{equation}
+            \frac{L_\nu}{L_Z} \sim \frac{N_z}{\tilde{\nu}200\pi} \gtrsim 1.
+        \end{equation}
+\end{itemize}
+
+However, in order for the simulation to both be well-damped
+at grid resolutions and be neglibily damped at leading order nonlinear
+wavelengths, we require $k_{0z} \ll \frac{1}{L_{NL}} \ll \frac{2\pi
+N_z}{L_z}$, where $L_{NL}$ denotes some characteristic length scale of
+nonlinear features. Finally, we need $L_z$ to be many $H$ to capture
+significant $e^{z/2H}$ growth and separate the linear and nonlinear
+amplitudes of $u_z$.
+
+Fixing $L_z = 10H$ to give $\sim e^4$ amplitude growth between the
+damping zones, this implies we require separation of scales $1 \ll
+k_{0z}H \ll \frac{H}{L_{NL}} \ll \frac{2\pi N_z}{10}$.
+
+Since $N_z$ is fixed by computational cost (we tried both $N_z = 512,
+N_z = 1024$), we choose $k_{0z} \sim \frac{2\pi}{H}$. This is less
+physically representative but gives more wavenumber space for the
+nonlinear cascade.
 

### 2019-01-15 writeups/paper.tex
stash paper

@@
 is constant gravitational acceleration. Note that at hydrostatic equilibrium
-$\pd{}{t} = 0$ we have $\vec{u} = 0, \vec{\nabla}P_0 = -\rho_0 g\hat{z}$ and
-so $P_0 = \rho_0 gH$.
+$\pd{}{t} = 0$ we have $\vec{\nabla}P_0 = -\rho_0 g\hat{z}$ and so $P_0 = \rho_0
+gH$. A shear flow $u_x(x, z, t) = u_x(z)$ is permitted in hydrostatic
+equilibrium, but we will initially consider no shear flow in equilibrium.
 
-We nondimensionalize by taking $H = N = \rho_{ref} = 1$ where
+\subsection{Internal Gravity Waves: Linear Theory}
+
+In the small perturbation limit, where flow velocities are small compared to the
+characteristic space and time scales $\pd{}{t} \gg \vec{u} \cdot \vec{\nabla}$,
+we linearize \autoref{se:fc_orig} by ignoring the advective components of the
+material derivative, defining $\rho_1 \equiv \rho - \rho_0 \ll \rho_0$ and $P_1
+\equiv P - P_0$ deviations from equilibrium and obtain:
+\begin{subequations}\label{se:lin_homo}
+    \begin{align}
+        \vec{\nabla} \cdot \vec{u} &= 0,\\
+        \pd{\rho_1}{t} - \frac{u_{z} \rho_0}{H} &= 0,\\
+        \pd{\vec{u}}{t} + \frac{\vec{\nabla}P_1}{\rho_0}
+            + \frac{\rho_1 g\hat{z}}{\rho_0}
+            &= 0.
+    \end{align}
+\end{subequations}
+
+It is well known that the solutions to the $F = 0$ homogeneous
+\autoref{se:lin_homo} are of form (cite)
+\begin{subequations}\label{se:lin_sol}
+    \begin{align}
+        u_z\p*{x, z, t} &= Ae^{z/2H}e^{i(k_{0x}x + k_{0z}z - \omega_0 t)},\\
+        \omega_0^2 &= \frac{N^2k_{0x}^2}{k_{0x}^2 + k_{0z}^2 + \frac{1}{4H^2}},
+    \end{align}
+\end{subequations}
+up to undetermined amplitude $A$, where
 \begin{equation}
@@
 \end{equation}
-is the Brunt-V\"ais\"al\"a frequency. The full numerical details of the
-simulation are described in \autoref{ss:strat_impl}.
+is the Brunt-V\"ais\"al\"a, or buoyancy, frequency.
+
+In the weak stratification limit $k_zH \gg 1$, the solution exhibits the
+following characteristics:
+\begin{itemize}
+    \item The amplitude of the perturbations grows like $e^{z/2H}$. Thus, the
+        linear approximation is violated for sufficiently large $z$.
+
+    \item The phase and group velocities can be computed respectively:
+        \begin{align}
+            \vec{c}_{ph} &= \frac{Nk_x}{k_x^2 + k_z^2 + 1/4H^2} \hat{k},\\
+            \vec{c}_{g} &= \frac{N\p*{k_xk_z\hat{x}
+                - \p*{k_z^2 + \frac{1}{4H^2}}\hat{z}}}
+                {\p*{k_x^2 + k_z^2 + \frac{1}{4H^2}}^{3/2}} \nonumber\\
+                &\approx \frac{Nk_z}{k_x^2 + k_z^2}\p*{k_x\hat{x} - k_z\hat{z}}.
+        \end{align}
+        We recover the usual result $\vec{c}_{ph} \cdot \vec{c}_g = 0$ (CITE).
+        Note that an IGW transporting energy and momentum upwards has $k_z < 0$.
+
+%     \item The linear solution is an exact solution to the full nonlinear
+%         equations \autoref{se:fc_orig}, thanks to the advective terms vanishing
+%         identically when $\vec{\nabla} \cdot \vec{u} = 0$ (CITE). However, this
+%         solution is globally unstable to resonant triad interactions (CITE).
+
+    \item The time-averaged total $x$-momentum flux in the $\hat{z}$ direction
+        can be computed. Since the fluid variables have separable as $f(x, z, t)
+        = f(z)e^{i(k_{0x}x - \omega_0 t)}$, $x$ averaging and time averaging are
+        equivalent, so we may write ($\ev*{\dots}_x$ denotes $x$-averaging)
+        \begin{align}
+            F_{px, z} &\equiv \ev*{\rho u_x u_z}_x =
+                \frac{1}{L_x}\int_0^{L_x}\limits \rho u_x u_z\;\mathrm{d}x
+                    \nonumber\\
+                &\approx -\frac{A^2}{2}\rho_{ref}\frac{k_z}{k_x}.
+                    \label{eq:fpx_lin}
+        \end{align}
+
+    \item Finally, it may be noted that horizontally averaged horizontal
+        momentum $p_x = \ev*{\rho u_x}_x$ obeys $\pd{p_x}{t} + \pd{F_{px}}{z} =
+        0$. Since flux is transported on the group velocity timescale (CITE),
+        this simplifies to the known expression for the wave-induced horizontal
+        mean flow (cite)
+        \begin{equation}
+            \bar{U}_0 \equiv \ev*{u_x}_x = \frac{\ev*{u_xu_z}_x}{c_{gz}}
+        \end{equation}
+        We've denoted $c_{gz} = \at{\pd{\omega}{k_{z}}}_{k_{z0}}$ the vertical group
+        velocity. In terms of our forcing amplitude,
+        \begin{equation}
+            \bar{U}_0(F) = -\frac{A(F)^2}{2}e^{\frac{z - z_0}{H}} \frac{k_z}{k_x}c_{gz},
+                \label{eq:u0_lin}
+        \end{equation}
+        where $c_{gz} = -\frac{Nk_{0x}k_{0z}}{\p*{k_{0x}^2 + k_{0z}^2 +
+        \frac{1}{4H^2}}^{3/2}}$.
+\end{itemize}
+% Linear solution: e^z/2H, dispersion relation, phase/group velocity. Observe is
+% exact solution of nonlinear.
+% Linear solution is globally unstable to RTI though, thus not stable
+% equilibrium of IVP
+%
+% Nonlinear: Critical layer absorption (viscous = Hazel, nonlin = Winters),
+% wave-induced mean flow, horizontal momentum flux
 
@@
 only the $k_{0z}$ satisfying the IGW's dispersion relation for the given
-$k_{0x}, \omega_0$ will propagate.
-% Define $k_{0z}: \omega(k_{0x}, k_{0z}) = \omega_0$ the wavenumber picked
-% out by $k_{0x}, \omega_0$. If $k_{0z}\sigma \ll 1$, the $\cos\p*{k_{0x}x -
-% \omega_0 t}$ term permits only the $k_{0z}$ mode to propagate among the broad
-% $k_z$ wavenumber spectrum permitted by the narrow Gaussian profile.
-
-\subsection{Internal Gravity Waves}
-
-In the small perturbation limit, we linearize \autoref{se:fc_orig} with
-inhomegeneous driving term and obtain:
+$k_{0x}, \omega_0(k_{0x}, k_{0z})$ will propagate. The fluid equations then take
+form
 \begin{subequations}\label{se:vol_drive}
@@
         \vec{\nabla} \cdot \vec{u} &= 0,\\
-        \pd{\vec{u}_1}{t} + \frac{\vec{\nabla}P}{\rho_0}
-            + \frac{\rho_1 \vec{g}}{\rho_0}
-            &= 0,\\
-        \pd{\rho_1}{t} - \frac{u_{1z} \rho_0}{H}
-            &= Fe^{-\frac{(z - z_0)^2}{2\sigma^2}}
-                e^{i(k_xx - \omega t)}.
-    \end{align}
-\end{subequations}
-We have substituted $\cos\p*{k_xx - \omega_0t}$ for a complex exponential for
-easier manipulation.
-
-It is well known that the solutions to the $F = 0$ homogeneous
-\autoref{se:vol_drive} are of form (cite)
-\begin{subequations}\label{se:lin_sol}
-    \begin{align}
-        u_z\p*{x, z, t} &= Ae^{z/2H}e^{i(k_{0x}x + k_{0z}z - \omega_0 t)},\\
-        \omega_0^2 &= \frac{N^2k_{0x}^2}{k_{0x}^2 + k_{0z}^2 + \frac{1}{4H^2}}.
+        \md{\rho}{t} &= Fe^{-\frac{(z - z_0)^2}{2\sigma^2}}
+                e^{i(k_xx - \omega t)},\\
+        \md{\vec{u}}{t} + \frac{\vec{\nabla}P}{\rho}
+            + g\hat{z} &= 0.
     \end{align}
@@
 
-To solve the inhomogeneous system, we may first approximate the driving term
-using $e^{\frac{-(z - z_0)^2}{2\sigma^2}} \approx \sqrt{2\pi \sigma^2}\delta(z -
-z_0)$ if $\sigma \to 0$. This system is solved exactly by matching the two
-homogeneous solutions above and below $z_0$, We may then approximately relax the
-solution to $k_{0z}\sigma \lesssim 1$. This may be solved to obtain the solution
+The linearized \autoref{se:vol_drive}, obtained by dropping the advective term
+in the material derivative, can be solved analytically up to good accuracy. We
+may first approximate the driving term using $e^{\frac{-(z - z_0)^2}{2\sigma^2}}
+\approx \sqrt{2\pi \sigma^2}\delta(z - z_0)$ if $\sigma \to 0$\footnote{$\sigma$
+must be large enough to be numerically resolved by the spectral code.}. This
+system is solved exactly by matching the two homogeneous solutions above and
+below $z_0$, We may then approximately relax the solution to $k_{0z}\sigma
+\lesssim 1$: an extra factor of $e^{-\frac{(k_{0z}\sigma)^2}{2}}$ arises
+compared to the $\delta$-function solution (evaluating the Fourier Transform
+of $e^{-\frac{(z - z_0)^2}{2\sigma^2}}$ at $k_z = k_{0z}$), and we obtain
 \begin{align}
-    A(F) &= \frac{Fgk_{0x}^2}{\rho_0(z_0)\omega_0^2}
+    A(F) ={}& \frac{Fgk_{0x}^2}{\rho_0(z_0)\omega_0^2}
         \frac{1}{2ik_z}\frac{e^{\frac{-(k_{0z}\sigma)^2}{2}}}
         {\sqrt{2\pi\sigma^2}},\\
-    u_{1z} &= A(F) \times\begin{cases}
+    u_{1z}(x, z, t) ={}& A(F) \times\nonumber\\
+        &{}\begin{cases}
         e^{\frac{z - z_0}{2H}}e^{i(k_{0x}x + k_{0z}(z - z_0) - \omega_0 t)}
@@
 \end{align}
-An extra factor of $e^{-\frac{(k_{0z}\sigma)^2}{2}}$ arises compared to the
-$\delta$-function solution when evaluating the Fourier Transform of
-$e^{-\frac{(z - z_0)^2}{2\sigma^2}}$ at $k_z = k_{0z}$.
-
-% Linear solution: e^z/2H, dispersion relation, phase/group velocity. Observe is
-% exact solution of nonlinear.
-% Linear solution is globally unstable to RTI though, thus not stable
-% equilibrium of IVP
-%
-% Nonlinear: Critical layer absorption (viscous = Hazel, nonlin = Winters),
-% wave-induced mean flow, horizontal momentum flux
-
-Furthermore, in the low-amplitude, dissipation free limit, the time-averaged
-total $x$-momentum flux in the $\hat{z}$ direction can be computed. Since all
-dynamical variables $\propto e^{i(k_{0x}x - \omega_0 t)}$, $x$ averaging and
-time averaging are equivalent, so we may write
-\begin{equation}
-    F_{px} \equiv \frac{1}{L_x}\int_0^{L_x}\limits \rho u_x u_z\;\mathrm{d}x
-\end{equation}
-We will denote this $F_{px} = \ev*{\rho u_xu_z}_x$ for simplicity. For a given
-forcing amplitude then:
-\begin{equation}
-    F_{px}(F) \approx -\frac{A(F)^2}{2}\rho_0(z_0) \frac{k_z}{k_x}.
-        \label{eq:fpx_lin}
-\end{equation}
-We've approximated $k_zH \ll 1$ and noticed $\ev*{\rho_1 u_xu_z}_x = 0$ in the
-above for clarity.
-
-Finally, it may be noted that horizontally averaged horizontal momentum
-$p_x = \ev*{\rho u_x}_x$ obeys $\pd{p_x}{t} + \pd{F_{px}}{z} = 0$, and since
-flux is transported on the group velocity timescale, this simplifies to the
-known expression for the wave-induced horizontal mean flow (cite)
-\begin{equation}
-    \bar{U}_0 \equiv \ev*{u_x}_x = \frac{\ev*{u_xu_z}_x}{c_{gz}}
-\end{equation}
-We've denoted $c_{gz} = \at{\pd{\omega}{k_{z}}}_{k_{z0}}$ the vertical group
-velocity. In terms of our forcing amplitude,
-\begin{equation}
-    \bar{U}_0(F) = -\frac{A(F)^2}{2}e^{\frac{z - z_0}{H}} \frac{k_z}{k_x}c_{gz},
-        \label{eq:u0_lin}
-\end{equation}
-where $c_{gz} = -\frac{Nk_{0x}k_{0z}}{\p*{k_{0x}^2 + k_{0z}^2 +
-\frac{1}{4H^2}}^{3/2}}$.
 
 \subsection{Numerical Parameters}\label{ss:numerics}
+
+We nondimensionalize by taking $H = N = \rho_{ref} = 1$
+
+%%%
+These are implemented as:
+\begin{subequations}\label{se:curr_num}
+    \begin{align}
+        \vec{\nabla} \cdot \vec{u} &= 0,\\
+        \pd{\Upsilon}{t} - \frac{u_z}{H}
+            - \nu \nabla^2 \Upsilon &= -\Gamma(z) \Upsilon
+                - \p*{\vec{u} \cdot \vec{\nabla}}\Upsilon
+                + \frac{F}{\rho_0(z)}e^{-\frac{(z - z_0)^2}{2\sigma^2}}
+                    \cos \p*{k_xx - \omega t},\\
+        \pd{u_x}{t} + \pd{T}{x} + gH\pd{\Upsilon}{x}
+            - \nu \nabla^2 u_x
+            &= -\Gamma(z) u_x
+                - \p*{\vec{u} \cdot \vec{\nabla}}u_x
+                - T_1 \pd{\Upsilon}{x},\\
+        \pd{u_z}{t} + \pd{T}{z} + gH\pd{\Upsilon}{z} - \frac{T_1}{H}
+            - \nu \nabla^2 u_z &= -\Gamma(z) u_z
+                - \p*{\vec{u} \cdot \vec{\nabla}}u_z
+                - T_1 \pd{\Upsilon}{z},\\
+    \Gamma(z) &= 7.5\s*{2 + \tanh \frac{z - z_T}{(L_z - z_T) / 2}
+        + \tanh \frac{z_B - z}{z_B / 2}},\label{eq:Gamma}
+    \end{align}
+\end{subequations}
+where $z_B = 0.05L_z, z_T = 0.95L_z$ are the boundaries of the damping zones. A
+Navier-Stokes numerical viscosity $\nu$ is used to damp high wavenumbers and
+regularize the nonlinear cascade at near grid resolution: $\nu \sim 0.2
+\frac{\omega}{\abs*{k_z}}\frac{L_z}{2\pi N_z}$ was found to be suitable for $N_z
+= 1024$, where $k_z$ is the wavenumber of the excited linear mode.
+%%%
 
@@
 % timestepping? RK443, (CFL limiter??)
+
+The exact implementation of these equations in Dedalus is described in
+\autoref{ss:strat_impl}.
 
@@
 
-\section{Equation Implementations}
+\section{Equation Implementations}\label{ss:strat_impl}
 
 We denote $x \in [0, L_x], z \in [0, L_z]$ the simulation domain and $N_x, N_z$
-the number of spectral modes in the respective dimensions.
+the number of spectral modes in the respective dimensions. We perform direct
+numerical simulation of \autoref{se:vol_drive} with the open-source
+pseudo-spectral code Dedalus (CITE).
 
-\subsection{Stratified Fluid}\label{ss:strat_impl}
-
-We introduce variable $T = P/\rho$. We then mandate $\rho_0, T_0$ background
-fields satisfy hydrostatic equilibrium $\vec{\nabla}T_0 + T_0 \vec{\nabla}\rho_0
-+ \vec{g} = 0$. Taking isothermal stratification, we find $T_0 = gH$. Making
-then substitution of variables $\Upsilon = \ln \rho - \ln \rho_0$ and $T_1 = T -
-T_0$ deviations from the background state, the exact fluid equations in the new
-variables are:
+Numerically, the nonlinear $\frac{\vec{\nabla}P}{\rho}$ term is problematic: we
+desire a system where the fluid fields are not divided by one another. We
+introduce $\varpi = \frac{P}{\rho}$ instead, then mandate $\rho_0, \varpi_0$
+background fields satisfy hydrostatic equilibrium $\vec{\nabla}\varpi_0 +
+\varpi_0 \vec{\nabla}\rho_0 + g\hat{z} = 0$. Taking isothermal stratification,
+we find $\varpi_0 = gH$. We further change variables to $\Upsilon = \ln \rho -
+\ln \rho_0$ and $\varpi_1 = \varpi - \varpi_0$ deviations from the background
+state to obtain a system of equations at most quadratic in fluid fields:
 \begin{subequations}\label{se:fc_var}
@@
         \vec{\nabla} \cdot \vec{u} &= 0,\\
-        \pd{\Upsilon}{t} - \frac{u_z}{H} &= 0,\\
+        \pd{\Upsilon}{t} + \p*{\vec{u} \cdot \vec{\nabla}} \Upsilon
+            - \frac{u_z}{H} &= 0,\\
         \pd{u_x}{t} + \p*{\vec{u} \cdot \vec{\nabla}}u_x
-            + \pd{T}{x} + gH\pd{\Upsilon}{x}
-            + T_1 \pd{\Upsilon}{x} &= 0,\\
+            + \pd{\varpi_1}{x} + gH\pd{\Upsilon}{x}
+            + \varpi_1 \pd{\Upsilon}{x} &= 0,\\
         \pd{u_z}{t} + \p*{\vec{u} \cdot \vec{\nabla}}u_z
-            + \pd{T}{z} + gH\pd{\Upsilon}{z}
-            + T_1 \pd{\Upsilon}{z} - \frac{T_1}{H} &= 0.
+            + \pd{\varpi_1}{z} + gH\pd{\Upsilon}{z}
+            + \varpi_1 \pd{\Upsilon}{z} - \frac{\varpi_1}{H} &= 0.
     \end{align}
 \end{subequations}
+It bears noting that these equations are exactly equivalent to the original
+Euler equations and hence conserve horizontal momentum.
 
-These are implemented as:
-\begin{subequations}\label{se:curr_num}
+\subsection{Artificial Dissipation}
+
+The nonlinear terms in the above equations will transfer energy from lower
+wavenumbers to higher wavenumbers. Since spectral codes have no numerical
+dissipation, artificial dissipation must be added. To ensure the dissipitive
+system conserves horizontal momentum, we begin by adding dissipitive terms to
+the flux-conservative form of the Euler fluid equations \autoref{se:fc_orig} (we
+use stress tensor $\tau_{ij} = P\delta_{ij}$):
+\begin{subequations}
     \begin{align}
         \vec{\nabla} \cdot \vec{u} &= 0,\\
-        \pd{\Upsilon}{t} - \frac{u_z}{H}
-            - \nu \nabla^2 \Upsilon &= -\Gamma(z) \Upsilon
-                - \p*{\vec{u} \cdot \vec{\nabla}}\Upsilon
-                + \frac{F}{\rho_0(z)}e^{-\frac{(z - z_0)^2}{2\sigma^2}}
-                    \cos \p*{k_xx - \omega t},\\
-        \pd{u_x}{t} + \pd{T}{x} + gH\pd{\Upsilon}{x}
-            - \nu \nabla^2 u_x
-            &= -\Gamma(z) u_x
-                - \p*{\vec{u} \cdot \vec{\nabla}}u_x
-                - T_1 \pd{\Upsilon}{x},\\
-        \pd{u_z}{t} + \pd{T}{z} + gH\pd{\Upsilon}{z} - \frac{T_1}{H}
-            - \nu \nabla^2 u_z &= -\Gamma(z) u_z
-                - \p*{\vec{u} \cdot \vec{\nabla}}u_z
-                - T_1 \pd{\Upsilon}{z},\\
-    \Gamma(z) &= 7.5\s*{2 + \tanh \frac{z - z_T}{(L_z - z_T) / 2}
-        + \tanh \frac{z_B - z}{z_B / 2}},\label{eq:Gamma}
+        \partial_t \rho + \vec{\nabla} \cdot (\rho \vec{u} - \nu
+            \vec{\nabla}(\rho - \rho_0)) &= 0,\label{eq:visc_cons_mom}\\
+        \partial_t (\rho \vec{u}) + \vec{\nabla} \cdot (\rho \vec{u} \vec{u} +
+            \mathrm{diag}(\rho \varpi) - \nu \rho \vec{\nabla}\vec{u}) + \rho g
+            \hat{z} &= 0.
     \end{align}
 \end{subequations}
-where $z_B = 0.05L_z, z_T = 0.95L_z$ are the boundaries of the damping zones. A
-Navier-Stokes numerical viscosity $\nu$ is used to damp high wavenumbers and
-regularize the nonlinear cascade at near grid resolution: $\nu \sim 0.2
-\frac{\omega}{\abs*{k_z}}\frac{L_z}{2\pi N_z}$ was found to be suitable for $N_z
-= 1024$, where $k_z$ is the wavenumber of the excited linear mode.
+The same $\nu$ is used for both the diffusive and viscous term, though this is
+not required. Since the dissipation is not physical and is purely used for
+numerical stability, we choose it such that hydrostatic equilibrium is not
+modified. Some algebraic manipulation to re-cast it in the form of
+\autoref{se:fc_var} gives
+\begin{subequations}
+    \begin{align}
+        \vec{\nabla} \cdot \vec{u} &= 0,\\
+        \partial_t \Upsilon + \p*{\vec{u} \cdot \vec{\nabla}} \Upsilon -
+            \frac{u_z}{H} - \nu\p*{\nabla^2 \Upsilon + \p*{\vec{\nabla}
+            \Upsilon} \cdot \p*{\vec{\nabla}\Upsilon} - \frac{2}{H}\partial_z
+            \Upsilon + \frac{1 - e^{-\Upsilon}}{H^2}} &= 0,\\
+        \partial_t \vec{u} + \p*{\vec{u} \cdot \vec{\nabla}}\vec{u} +
+            \vec{\nabla} \varpi + \varpi \vec{\nabla} \Upsilon - \nu \nabla^2
+            \vec{u} + \vec{u} \nu\p*{\nabla^2 \Upsilon + \p*{\vec{\nabla}
+            \Upsilon} \cdot \p*{\vec{\nabla}\Upsilon} - \frac{2}{H}\partial_z
+            \Upsilon + \frac{1 - e^{-\Upsilon}}{H^2}}&{}\nonumber\\
+        - 2\nu \p*{\p*{\p*{\vec{\nabla}\Upsilon} \cdot \vec{\nabla}}\vec{u} -
+            \frac{1}{H}\partial_z \vec{u}} - \frac{\varpi_1}{H} &= 0.
+    \end{align}
+\end{subequations}
+Hydrostatic equilibrium is still $\vec{\nabla} \varpi_0 + g\hat{z} = 0$ where
+$\rho = \rho_0, \vec{u} = 0$. Including the damping layers and forcing terms as
+described in \autoref{ss:numerics}, we finally obtain the full system of
+equations as simulated in Dedalus:
 

### 2019-01-18 proposal/GRFP/Su_GRFP.tex
reorg proposals

    \documentclass[11pt,
        usenames, % allows access to some tikz colors
        dvipsnames % more colors: https://en.wikibooks.org/wiki/LaTeX/Colors
    ]{article}
    \usepackage{
        amsmath,
        amssymb,
        fouriernc, % fourier font w/ new century book
        fancyhdr, % page styling
        lastpage, % footer fanciness
        hyperref, % various links
        setspace, % line spacing
        amsthm, % newtheorem and proof environment
        mathtools, % \Aboxed for boxing inside aligns, among others
        float, % Allow [H] figure env alignment
        enumerate, % Allow custom enumerate numbering
        graphicx, % allow includegraphics with more filetypes
        wasysym, % \smiley!
        upgreek, % \upmu for \mum macro
        listings, % writing TrueType fonts and including code prettily
        tikz, % drawing things
        booktabs, % \bottomrule instead of hline apparently
        cancel % can cancel things out!
    }
    \usepackage[margin=1in]{geometry} % page geometry
    \usepackage[
        labelfont=bf, % caption names are labeled in bold
        font=scriptsize % smaller font for captions
    ]{caption}
    \usepackage[font=scriptsize]{subcaption} % subfigures

    \newcommand*{\scinot}[2]{#1\times10^{#2}}
    \newcommand*{\dotp}[2]{\left<#1\,\middle|\,#2\right>}
    \newcommand*{\rd}[2]{\frac{\mathrm{d}#1}{\mathrm{d}#2}}
    \newcommand*{\pd}[2]{\frac{\partial#1}{\partial#2}}
    \newcommand*{\rtd}[2]{\frac{\mathrm{d}^2#1}{\mathrm{d}#2^2}}
    \newcommand*{\ptd}[2]{\frac{\partial^2 #1}{\partial#2^2}}
    \newcommand*{\md}[2]{\frac{\mathrm{D}#1}{\mathrm{D}#2}}
    \newcommand*{\pvec}[1]{\vec{#1}^{\,\prime}}
    \newcommand*{\svec}[1]{\vec{#1}\;\!}
    \newcommand*{\bm}[1]{\boldsymbol{\mathbf{#1}}}
    \newcommand*{\ang}[0]{\;\text{\AA}}
    \newcommand*{\mum}[0]{\;\upmu \mathrm{m}}
    \newcommand*{\at}[1]{\left.#1\right|}

    \newtheorem{theorem}{Theorem}[section]

    \let\Re\undefined
    \let\Im\undefined
    \DeclareMathOperator{\Res}{Res}
    \DeclareMathOperator{\Re}{Re}
    \DeclareMathOperator{\Im}{Im}
    \DeclareMathOperator{\Log}{Log}
    \DeclareMathOperator{\Arg}{Arg}
    \DeclareMathOperator{\Tr}{Tr}
    \DeclareMathOperator{\E}{E}
    \DeclareMathOperator{\Var}{Var}
    \DeclareMathOperator*{\argmin}{argmin}
    \DeclareMathOperator*{\argmax}{argmax}
    \DeclareMathOperator{\sgn}{sgn}
    \DeclareMathOperator{\diag}{diag\;}

    \DeclarePairedDelimiter\bra{\langle}{\rvert}
    \DeclarePairedDelimiter\ket{\lvert}{\rangle}
    \DeclarePairedDelimiter\abs{\lvert}{\rvert}
    \DeclarePairedDelimiter\ev{\langle}{\rangle}
    \DeclarePairedDelimiter\p{\lparen}{\rparen}
    \DeclarePairedDelimiter\s{\lbrack}{\rbrack}
    \DeclarePairedDelimiter\z{\lbrace}{\rbrace}

    % \everymath{\displaystyle} % biggify limits of inline sums and integrals
    \tikzstyle{circ} % usage: \node[circ, placement] (label) {text};
        = [draw, circle, fill=white, node distance=3cm, minimum height=2em]
    \definecolor{commentgreen}{rgb}{0,0.6,0}
    \lstset{
        basicstyle=\ttfamily\footnotesize,
        frame=single,
        numbers=left,
        showstringspaces=false,
        keywordstyle=\color{blue},
        stringstyle=\color{purple},
        commentstyle=\color{commentgreen},
        morecomment=[l][\color{magenta}]{\#}
    }
    \usepackage[
        backend=bibtex8,
        style=nature]{biblatex}

\bibliography{Su_GRFP}
% \AtEveryBibitem{\clearfield{title}}
\begin{document}

% NOTE: this is not the version I submitted, I wrote it on gdocs

\def\Snospace~{\S{}} % hack to remove the space left after autorefs
\renewcommand*{\sectionautorefname}{\Snospace}
\renewcommand*{\appendixautorefname}{\Snospace}
\renewcommand*{\figureautorefname}{Fig.}
\renewcommand*{\equationautorefname}{Eq.}
\renewcommand*{\tableautorefname}{Tab.}

\onehalfspacing

\pagestyle{fancy}
\rfoot{Yubo Su}
\rhead{}
\cfoot{\thepage/\pageref{LastPage}}

\title{Nonlinear Tidal Dissipation in Binary White Dwarfs}
\author{Yubo Su}
\date{\today}

\maketitle

\section{White Dwarf Binaries}

White dwarfs (WDs) are remnants of stellar evolution for stars with mass
$\lesssim 8M_{\odot}$. They are some of the densest objects in the universe,
fitting a solar mass into an Earth-sized sphere. They are supported against
gravity by degeneracy pressure, a quantum mechanical effect arising from
subjecting matter to immense pressure. As such, WDs are a unique window into
matter under extreme conditions.

WDs are commonly found in binary systems in which two objects orbit their center
of mass under mutual gravitational attraction. The companion object ranges from
another WD to a supermassive black hole, and all of these binary systems are
very important to astrophysics. In particular, merging WD-WD binaries are
thought to generate Type Ia supernovae (SNe Ia). Because SNe Ia are highly
luminous and can be seen at large distances, they have been used as ``standard
candles'' to probe the expansion rate of the universe (e.g.\ in 1998 SNe Ia
observations provided the first evidence for dark energy\cite{DE}). WD binaries
also produce gravitational waves (GW), periodic warping of space-time predicted
by Einstein's theory of general relativity. The first detection of GWs by the
Laser Interferometer Gravitational-Wave Observatory (LIGO) in 2015 heralded a
new era of GW astrophysics. Although LIGO cannot detect the low-frequency GWs
expected from WD binaries, the space-based Laser Interferometer Space Antenna
(LISA) will detect such GWs in the coming decades\cite{lisa}. As GW astronomy
relies on accurate predictions of the expected gravitational waveforms, it is
important to build highly precise models of the WD binary inspiral before LISA
observation runs begin (scheduled for $\sim 2030$).

\subsection{Tidal Dissipation and Internal Gravity Waves}

Tidal dissipation in binary WDs can affect the pre-merger state of the WD and
have a major impact on the gravitational waves produced by the binary. Such
dissipation arises from the excitation of internal gravity waves (IGWs) in the
deep envelope of the WD by the gravitational force of its companion and their
subsequent dissipation in the outer envelope\cite{fullerII}. IGWs, not to be
confused with the gravitational waves above, are internal displacements in the
WD fluid that oscillate and propagate due to a restoring buoyancy force. These
waves are analogous to ocean tides on Earth raised by the Moon and the Sun,
except that, since WDs do not have sharp surfaces, these waves are internal to
the WD\@. As these waves propagate outwards from where they are excited, they
grow in amplitude until they break, as do ocean waves on a shore, and deposit
both energy and angular momentum from the binary orbit in the outer envelope of
the WD\@.

Previous works have shown that this dissipation mechanism can generate
significantly more energy than thermal radiation from the WD surface alone and
is thus a major contributor to the WD energy budget\cite{fullerII,fullerIV}.
However, these works treated nonlinear wave dissipation by a simple
parameterization. Related works in other fields find that properly including
nonlinear effects produces drastically different IGW dissipation
behavior\cite{winters1994}. Such fully nonlinear studies have not been performed
for astrophysical binary systems but are important for the thermal and orbital
evolution of WD binaries undergoing tidal heating. Characterizing this nonlinear
tidal dissipation will require extensive numerical simulation to capture the
turbulent cascade to small scales that drives IGW dissipation. This is the goal
of my proposed research.

\section{Proposed Research}

I propose to study the dynamical effects of tidal dissipation via nonlinear IGW
breaking in binary WDs using both analytical techniques and numerical
simulations. As WDs vary widely in composition and effective temperature, I will
study select WD models to capture a wide range of possible phenomena. My
research will proceed in the following stages:

I will perform simulations of nonlinear IGW breaking for various WD models. In
the last year, I have adapted the spectral hydrodynamic code
Dedalus\cite{dedalus} to simulate the simplest WD models. I have excited IGWs in
the deep envelope and observed nonlinear dissipation as the wave propagates into
the outer envelope. I am in the process of characterizing this dissipation and
extending my simulation to other, more realistic WD models. From these
simulations, I will compute energy and angular momentum dissipation as a
function of time and depth in the envelope for each WD model.

I will apply my dissipation profiles to study tidally heated WDs. Modules for
Experiments in Stellar Astrophysics (MESA) is a proven stellar evolution code
that is readily extensible to include tidal heating\cite{MESA}. Packaging my dissipation
profiles for each WD model as a MESA module would enable simulations of the
binary WDs’ internal structures evolving under tidal heating. From these
simulations, I will extract the WD luminosity over time.

I will compare the simulated binary WD luminosity to observational data to
constrain WD properties. For instance, the WD binary SDSS JJ065133+284423
(period 13 minutes) may undergo strong tidal heating and exhibit energetic
behavior such as tidal novae\cite{12min, tidal_novae}. The Large Synoptic Survey
Telescope (LSST, expected 2021) is expected to detect a few thousand more WD
binaries\cite{lsst_wd}. I will forecast properties of these samples due to tidal
heating.

Tidal dissipation can also affect the orbital evolution of the WD binary and
hence its GW emission during inspiral. I will calculate the signature of tidal
dissipation on these GWs for each WD model. I will publish my corrected GW
waveforms for use by LISA and the GW community.

Intellectual Merit: As discussed, WD binaries are important for being potential
progenitors of Type Ia supernovae, for being future LISA GW sources and for
producing various transients that would be detectable. My work will characterize
a key ingredient of WD binary evolution that may explain observed properties of
such binaries. Additionally, the nonlinear IGW dissipation model I will build
has applications beyond WDs binaries. Within astrophysics, IGWs are a key
angular momentum transport mechanism in stellar interiors: they are believed to
fill a crucial gap in stellar evolution theory\cite{l_trans_rev}. Beyond
astrophysics, IGWs are also studied in terrestrial sciences such as oceanography
and atmospheric sciences. A robust understanding of nonlinear IGW dissipation
would thus have consequences for problems ranging from planetary and stellar
evolution to weather patterns and biosphere dynamics on Earth.

Broader Impacts: With the support of the NSF GRFP, I will grow as an effective
researcher. I will present my results at conferences and to collaborators in the
US and abroad. I will generate videos for use in future Cornell outreach
efforts. As hydrodynamical simulation produces visually spectacular videos, my
research will be an effective tool in inspiring the public.

\printbibliography

\end{document}

### 2019-01-18 proposal/NDSEG_2017/NDSEG_proposal.tex
reorg proposals

--- 2017-12-11 proposal/proposal.tex
@@
 \renewcommand*{\sectionautorefname}{\Snospace}
+\renewcommand*{\subsectionautorefname}{\Snospace}
 \renewcommand*{\appendixautorefname}{\Snospace}
@@
 
-\onehalfspacing
+\setstretch{1.2}
 
@@
 
-White dwarfs (WDs) are luminous, long-lasting products of late-stage stellar
-evolution that are held up by electron degeneracy pressure, a quantum mechanical
-effect arising from subjecting electrons to immense pressures. They are some of
-the densest objects in the known universe, behind only neutron stars (NS) and
-black holes (BH), fitting a solar mass into an Earth-sized sphere. Furthermore,
-they exhibit various chemical compositions, each of which is subject to
-different physics, and are found in a variety of interesting systems. As such,
-WDs are a unique and diverse window into matter under extreme conditions.
+White dwarfs (WDs) are remnants of stellar evolution for stars with mass
+$\lesssim 8 M_{sun}$. They are some of the densest objects in the universe,
+fitting a solar mass into an Earth-sized sphere. They are supported against
+gravity by electron degeneracy pressure, a quantum mechanical effect arising from
+subjecting electrons to immense pressures Furthermore, they exhibit various
+chemical compositions, each of which is subject to different physics, and are
+found in a variety of interesting systems. As such, WDs are a unique and diverse
+window into matter under extreme conditions.
 
-WDs are commonly found in binary orbits with companion objects, in which the two
-orbit their center of mass under their mutual gravitational attraction. The
-companion object ranges from another WD to a supermassive BH (SMBH),
-hypothesized to be at the center of galaxies and have mass $10^5$--$10^6$ solar
-masses. All of these binary systems are very important to astrophysics. WD-WD
-binaries are most important for being thought to generate \emph{Type Ia
-supernovae} in which one of the two WDs reaches a critical mass through
-accreting surrounding matter and collapses under its own gravity, producing a
-shock wave that tears the WD apart and releases immense amounts of energy. Type
-Ia supernovae are unique in astrophysics for their near-constancy throughout
-space and time, consistently occuring at the WD critical mass, and have been
-used to discover numerous cosmological phenomena such as dark energy.
+WDs are commonly found in binaries in which two objects orbit their center of
+mass under mutual gravitational attraction. The companion object ranges
+from another WD to a supermassive BH (SMBH), hypothesized to be at the center of
+galaxies and have mass $\gtrsim 10^5$ solar masses. All of these binary systems
+are very important to astrophysics. WD-WD binaries are most important for being
+thought to generate \emph{Type Ia supernovae} in which two WDs merge and
+initiate runaway thermonuclear fusion that powers an explosion, releasing
+immense amounts of energy. Because Type Ia supernovae (SNe Ia) are highly
+luminous, can be seen at large distances and are very consistent, they have been
+used as a ``standard candle'' to probe the expansion rate of the universe (e.g.\
+observations of SNe Ia provided the first evidence of the existence of dark
+energy in the universe).
 
-WD-BH systems are also interesting subjects of study. Studies indicate that
-as WDs orbit the extreme gravity surrounding BHs, they will produce observable
-flares induced by gravitational tidal forces\cite{flares}. A WD orbiting a SMBH
-would produce gravitational waves, waves carried by distortions in space-time
-and generated according to Einstein's general relativity by accelerating massive
-objects, that are expected to be detected by the space-based \emph{Laser
-Interferometer Space Antenna (LISA)} when deployed\cite{lisa}. Gravitational
-wave astronomy is an increasingly exciting field as the \emph{Laser
-Interferometer Gravitational-Wave Observatory (LIGO)} continues to make
-detections since its first in late 2015. As gravitational wave astronomy relies
-on accurate predictions of the expected signals, it is important to build as
-accurate models as possible before observation runs begin.
+WD-BH systems are also interesting subjects of study. Recent works indicate that
+as WDs plunge close to BHs, they will produce observable flares induced by
+gravitational tidal forces\cite{flares}. A WD orbiting a SMBH would also produce
+gravitational waves, waves of space-time warping as predicted by Einstein's
+theory of general relativity, that are expected to be detected by the
+space-based \emph{Laser Interferometer Space Antenna (LISA)} when
+deployed\cite{lisa}. Gravitational wave astronomy is an increasingly exciting
+field as the \emph{Laser Interferometer Gravitational-Wave Observatory (LIGO)}
+continues to make progress since its first detection in late 2015. As
+gravitational wave astronomy relies on accurate predictions of the expected
+signals, it is important to build as accurate models as possible before
+observation runs begin.
 
@@
 The excitation of \emph{internal gravity waves} in the WD by the tidal forces of
-the companion is an effect seen in all of the aforementioned systems. Internal
-gravity waves, not to be confused with the gravitational waves discussed above,
-are internal displacements in the WD that oscillate and propagate due to a
-restoring buoyancy force. As these waves propagate outward from where they are
-excited, they are expected to grow in amplitude until they break, as do ocean
-waves on a shore, and deposit both energy and angular momentum inside the WD\@.
+the companion is an effect exhibited in all of the aforementioned systems.
+Internal gravity waves, not to be confused with the gravitational waves above,
+are internal displacements in the WD fluid that oscillate and propagate due to a
+restoring buoyancy force. These tidally excited gravity waves are analogous to
+tides on Earth raised by the Moon and the Sun, except that since WDs do not have
+sharp surfaces these waves are internal to the WD\@. As these waves propagate
+outward from where they are excited, they are expected to grow in amplitude
+until they break, as do ocean waves on a shore, and deposit both energy and
+angular momentum in the WD envelope.
 
@@
 are thus a significant contribution to the WD energy budget\cite{fullerII}. The
-exact radial dissipation profile is of enormous interest since it is both
-sensitive to WD properties and can produce drastically different observable
-outcomes. One proposed outcome is a \emph{tidal nova}, in which heating in the
-WD's degenerate hydrogen layer is significant enough to trigger runaway nuclear
-fusion and a significant, observable explosion\cite{tidal_novae}. Understanding
-whether such phenomena occur requires understanding how energy is distributed
-internally inside a WD\@.
+exact radial dissipation profile is of interest since it both is sensitive to WD
+properties and can produce vastly different observable outcomes. One proposed
+outcome is a \emph{tidal nova}, in which heating in the WD's degenerate hydrogen
+layer is sufficient to trigger runaway nuclear fusion and an observable
+surface explosion\cite{tidal_novae}. Understanding whether such phenomena occur
+requires understanding how energy is distributed internally inside a WD\@.
 
@@
 paramount to begin numerical study to build dissipation models inside WDs to
-characterise what phenomena can be observed in which WD models and
-characteristics.
+characterize what phenomena can be observed in different WD models.
 
@@
 \textbf{We propose to study tidal dissipation via nonlinear gravity wave
-breaking in white dwarfs via numerical simulation.}
+breaking in white dwarfs using numerical simulation.}
 
-Our research will consist of computing for various WD models and compositions
-the energy and angular momentum dissipation profiles inside WDs. Once such
-profiles are obtained, we intend to add these profiles to existing stellar
-evolution codes to study the dynamical effects of tidal dissipation. We will
-attempt to find a compact or even analytical representation of our numerical
-work to greatly simplify and accelerate such integration. Finally, we will make
-observational predictions with our results and compare our predictions with
-existing observed WDs. Any software and results would be made in accordance with
-best practice and publicly available for the entire community.
+Our research will initially consist of numerically computing for various WD
+models and compositions the energy and angular momentum dissipation profiles
+inside WDs. Once such profiles are obtained, we intend to add these profiles to
+existing stellar evolution codes to study the dynamical effects of tidal
+dissipation. We will attempt to find a compact or even analytical representation
+of our numerical work to greatly simplify and accelerate such integration.
+Finally, we will make observational predictions with our results and compare our
+predictions with existing observed WDs in binaries. Any software and results
+would be made in accordance with best practice and publicly available for the
+scientific community.
 
-\subsection{Qualifications}
+\subsection{Numerical Approach}
+
+Much of our proposed research will rely on numerical solutions to fluid
+dynamical equations. Guided by current literature, we have begun with the
+Dedalus numerical solver, a modern spectral solver that integrates the fluid
+equations both quickly and to high accuracy\cite{dedalus}. Dedalus has been
+shown in current literature to describe turbulence and fluid instabilities well,
+and our preliminary work supports this conclusion.
+
+For our problem, we will first adapt Dedalus to study gravity waves in the
+stratified atmosphere/envelope of WDs and later consider global (spherical) WD
+models. Should Dedalus prove inadequate, we are able (\autoref{ss:quals}) and
+willing to contribute a new hydrodynamic code to the community. Notably, no
+GPU-accelerated spectral hydrodynamic code has yet been released, which would be
+of great value to researchers. Nevertheless, in the interest of the proposed
+science objectives codes will be developed on an as-needed basis.
+
+After characterizing tidal dissipation, we intend to apply our models to WD
+evolution under tidal heating. Modules for Experiments in Stellar Astrophysics
+(MESA) is a proven stellar evolution code that would require comparatively
+little adaptation to incorporate tidal heating in studying WD
+evolution\cite{MESA}. A released extension to MESA containing our work would be
+the most effective way to release our work to prospective users.
+
+\subsection{Qualifications}\label{ss:quals}
 
 The proposing researcher (YS) and his adviser (DL) are uniquely qualified to
-pursue such studies. YS has ample research experience in numerical simulation
-and is continuing to pursue academic study in hydrodynamics, turbulence and
-computation. He also completed a double Bachelors degree in physics and computer
-science from the California Institute of Technology. Finally, he has worked in
+pursue such studies. YS has completed a double Bachelors degree in physics and
+computer science from the California Institute of Technology. He also has ample
+research experience in numerical simulation and is continuing to pursue academic
+study in hydrodynamics, turbulence and computation. Finally, he has worked in
 the software industry for a year and continues to follow discussions of best
@@
 
-DL is a co-author on much of the latest literature on WD binaries and is an
+DL is a co-author of a number of recent papers on WD binaries and is an
 expert on many related subjects such as fluid dynamics in extreme matter and
@@
 
-In terms of software, YS proposes to only implement what is not currently
-available in the community. Current literature indicates that the spectral
-hydrodynamic code Dedalus is able to handle instability-driven turbulence with
-high accuracy and speed, and our preliminary work seems to indicate that Dedalus
-will be sufficiently accurate to simulate wave breaking
-dissipation\cite{dedalus}. Modules for Experiments in Stellar Astrophysics
-(MESA) is a proven stellar evolution code that will likely need only be extended
-to incorporate the discovered dissipation models\cite{MESA}. Nevertheless, both
-codes will require adaptation to be applied to the present problem (YS has
-already contributed an improvement to Dedalus), a task for which YS is
-well-equiped.
-
-Should no existing software prove adequate, we are more than able and willing to
-develop a new code and contribute it to the community. Notably, no
-GPU-accelerated spectral hydrodynamic code has yet been released, which would be
-of great value to researchers, and YS is well-versed in GPU programming.
-Nevertheless, in the interest of the proposed science objectives codes will be
-developed on an as-needed basis.
-
 \subsection{Applications, Interest to the DoD}
 
-A robust pipeline coupling nonlinear wave breaking to system evolution
-would be of interest spanning all fields of physics. Barring BHs, all
-astrophysical objects exhibit internal waves that can break as they propagate
-towards the surface. Notable present interests in the astrophysical community
-include tidal disruption, planet formation and stars. Even in planetary and
-atmospheric sciences, wave breaking in atmospheres would share many common
-techniques with the proposed research.
+Although our proposed research deals with astronomical objects and phenomena
+(WDs and BHs), the tools we use and develop in our research
+have applications in many areas of physical sciences and engineering. As
+discussed above, through our research we will develop extensive expertise in
+nonlinear hydrodynamics (internal gravity waves), radiation physics (radiative
+transfer and diffusion of heat inside WD), nuclear physics (tidal heating
+leading to runaway fusion) and numerical computation. Such a research project
+thus contributes to many science and engineering fields and equips me to tackle
+an exceptionally large variety of problems.
 
 The proposed research is extremely relevant to Department of Defense (DoD)
-fields of interest. WDs and its energetic phenomena we propose to study e.g.\
-novae of all sorts are natural laboratories of extreme physics that have
+fields of interest. WDs and its energetic phenomena we propose to study (e.g.\
+novae) of all sorts are natural laboratories of extreme physics that have
 direct consequences for understanding of plasmas and other technologies integral

### 2019-01-18 proposal/NDSEG_2017/outline.tex
reorg proposals

    \documentclass[11pt,
        usenames, % allows access to some tikz colors
        dvipsnames % more colors: https://en.wikibooks.org/wiki/LaTeX/Colors
    ]{article}
    \usepackage{
        amsmath,
        amssymb,
        fouriernc, % fourier font w/ new century book
        fancyhdr, % page styling
        lastpage, % footer fanciness
        hyperref, % various links
        setspace, % line spacing
        amsthm, % newtheorem and proof environment
        mathtools, % \Aboxed for boxing inside aligns, among others
        float, % Allow [H] figure env alignment
        enumerate, % Allow custom enumerate numbering
        graphicx, % allow includegraphics with more filetypes
        wasysym, % \smiley!
        upgreek, % \upmu for \mum macro
        listings, % writing TrueType fonts and including code prettily
        tikz, % drawing things
        booktabs, % \bottomrule instead of hline apparently
        cancel % can cancel things out!
    }
    \usepackage[margin=1in]{geometry} % page geometry
    \usepackage[
        labelfont=bf, % caption names are labeled in bold
        font=scriptsize % smaller font for captions
    ]{caption}
    \usepackage[font=scriptsize]{subcaption} % subfigures

    \newcommand*{\scinot}[2]{#1\times10^{#2}}
    \newcommand*{\bra}[1]{\left<#1\right|}
    \newcommand*{\ket}[1]{\left|#1\right>}
    \newcommand*{\dotp}[2]{\left<#1\,\middle|\,#2\right>}
    \newcommand*{\rd}[2]{\frac{\mathrm{d}#1}{\mathrm{d}#2}}
    \newcommand*{\pd}[2]{\frac{\partial#1}{\partial#2}}
    \newcommand*{\rtd}[2]{\frac{\mathrm{d}^2#1}{\mathrm{d}#2^2}}
    \newcommand*{\ptd}[2]{\frac{\partial^2 #1}{\partial#2^2}}
    \newcommand*{\md}[2]{\frac{\mathrm{D}#1}{\mathrm{D}#2}}
    \newcommand*{\norm}[1]{\left|\left|#1\right|\right|}
    \newcommand*{\abs}[1]{\left|#1\right|}
    \newcommand*{\pvec}[1]{\vec{#1}^{\,\prime}}
    \newcommand*{\svec}[1]{\vec{#1}\;\!}
    \newcommand*{\bm}[1]{\boldsymbol{\mathbf{#1}}}
    \newcommand*{\expvalue}[1]{\left<#1\right>}
    \newcommand*{\ang}[0]{\text{\AA}}
    \newcommand*{\mum}[0]{\upmu \mathrm{m}}
    \newcommand*{\at}[1]{\left.#1\right|}

    \newtheorem{theorem}{Theorem}[section]

    \let\Re\undefined
    \let\Im\undefined
    \DeclareMathOperator{\Res}{Res}
    \DeclareMathOperator{\Re}{Re}
    \DeclareMathOperator{\Im}{Im}
    \DeclareMathOperator{\Log}{Log}
    \DeclareMathOperator{\Arg}{Arg}
    \DeclareMathOperator{\Tr}{Tr}
    \DeclareMathOperator{\E}{E}
    \DeclareMathOperator{\Var}{Var}
    \DeclareMathOperator*{\argmin}{argmin}
    \DeclareMathOperator*{\argmax}{argmax}
    \DeclareMathOperator{\sgn}{sgn}
    \DeclareMathOperator{\diag}{diag\;}

    \DeclarePairedDelimiter\p{\lparen}{\rparen}
    \DeclarePairedDelimiter\s{\lbrack}{\rbrack}
    \DeclarePairedDelimiter\z{\lbrace}{\rbrace}

    % \everymath{\displaystyle} % biggify limits of inline sums and integrals
    \tikzstyle{circ} % usage: \node[circ, placement] (label) {text};
        = [draw, circle, fill=white, node distance=3cm, minimum height=2em]
    \definecolor{commentgreen}{rgb}{0,0.6,0}
    \lstset{
        basicstyle=\ttfamily\footnotesize,
        frame=single,
        numbers=left,
        showstringspaces=false,
        keywordstyle=\color{blue},
        stringstyle=\color{purple},
        commentstyle=\color{commentgreen},
        morecomment=[l][\color{magenta}]{\#}
    }

\begin{document}

\def\Snospace~{\S{}} % hack to remove the space left after autorefs
\renewcommand*{\sectionautorefname}{\Snospace}
\renewcommand*{\appendixautorefname}{\Snospace}
\renewcommand*{\figureautorefname}{Fig.}
\renewcommand*{\equationautorefname}{Eq.}
\renewcommand*{\tableautorefname}{Tab.}

\section*{Timelines}

\begin{itemize}
    \item 12/31/17---NDSEG \url{http://www.ndsegfellowships.org/application}
        \begin{itemize}
            \item \emph{Please submit a proposal relevant to the DoD Agency
                (AIRFORCE, ARMY, ONR)/field that interest you.} (Pitch as
                ``science of extreme materials, results that can be useful in
                understanding turbulence''?)
            \item 3 page proposal.
        \end{itemize}

    \item 01/10/17---NNSA SSGF \url{
        https://www.krellinst.org/ssgf/about-doe-nnsa-ssgf/fields-study}
        \begin{itemize}
            \item Promote stewardship science (National Nuclear Security
                Administration).
            \item Fields of study: \textbf{High Energy Density Physics,
                Materials under Extreme Conditions and Hydrodynamics}.
            \item Ask to propose \emph{program of study}, I picked:
                \begin{itemize}
                    \item Advanced Plasmas
                    \item Physics of BHs, WDs, stars
                    \item Theory stellar structure/evolution
                    \item Applications of Parallel Computers
                    \item Computational fluid dynamics
                    \item Turbulence/turbulent flows (not sure this one)
                \end{itemize}
            \item Essays:
                \begin{itemize}
                    \item 500w area of research
                    \item 300w intellectual excitement
                    \item 300w ``how your research will directly compile to
                        stockpile stewardship in your area'' Advice from
                        theoretical astrophysicist: ``\emph{I pretty much argued
                        that SNe are labs for many different types of extreme
                        physics and can help inform ground-based experiments.
                        Also many of the same principles in my simulations could
                        be applied to other things. Depending on what you study
                        it shouldn't be too hard to justify (especially if
                        you're into plasmas, high energy particles, fluid
                        dynamics, etc.)}''
                \end{itemize}
        \end{itemize}

    \item 01/17/17---DOE CSGF \url{
        https://www.krellinst.org/csgf/about-doe-csgf/fields-study}
        \begin{itemize}
            \item Math + Computers to study fields e.g.\ astrophysics, etc.,
                promote interdisciplinary, require broad program of study.
            \item Possible proposal: \textbf{Propose to develop new software? Or
                just say that we'll be ready to, just in case.}
            \item Program of Study:
                \begin{itemize}
                    \item (Science/Eng) Physics BH, WD, NS
                    \item (Science/Eng) Computational Fluid Dynamics
                    \item (Math/Stats) Applied Dynamical Systems
                    \item (Math/Stats) Functional Analysis
                    \item (Computer Science) Applications of Parallel Computers
                    \item (Computer Science) Advanced Programming Languages
                \end{itemize}
            \item Essays:
                \begin{itemize}
                    \item 2250c Field of Interest, role of CS
                    \item 2250c use of CS + math in proposed research
                    \item 2250c justify program of study.
                \end{itemize}
        \end{itemize}
    \item 02/01/17---NSPIRES NESSF \url{
        https://nspires.nasaprs.com/external/solicitations/summary.do?method=init&solId={1A7F0A8C-1126-EA39-6584-10DBD500C8AC}&path=open
        }
        \begin{itemize}
            \item \textbf{Adviser starts}, check instructions at \href{
                https://nspires.nasaprs.com/external/viewrepositorydocument/cmdocumentid=539448/solicitationId=\%7B1A7F0A8C-1126-EA39-6584-10DBD500C8AC\%7D/viewSolicitationDocument=1/NESSF\%202017-2018\%20Proposal\%20Submission\%20Instructions.pdf}{here}
            \item Proposal limit 6 pages
        \end{itemize}
\end{itemize}

\newpage

\title{Nonlinear Tidal Dissipation in White Dwarfs}
\author{Yubo Su}
\date{}

\maketitle

\section{Introduction}

\subsection{Background}

\begin{itemize}
    \item \emph{Steal a bunch from Fuller et al}
    \item WD-WD systems are important, produce sdO stars, giant stars,
        Ia supernovae (Fuller Paper II citations)
    \item WD-BH systems, tidal disruption produces observable flares
        (https://arxiv.org/pdf/1701.08162.pdf)
    \item WD binaries are in LISA sensitivity band, GW astronomy is powerful
        tool evidenced by the recent LIGO success (FullerII)
    \item Tidal dissipation can contribute significantly to the brightnesses of
        these binaries (FullerII)
\end{itemize}

\subsection{Previous Work}

\begin{itemize}
    \item Fuller + Lai determined in WD binaries, interval gravity waves are
        excited that propagate towards the surface and nonlinear dissipate in
        some WDs.
    \item Predicted dissipation luminosity dominates black body dissipation,
        significant contribution to energy dynamics. (FullerII)
    \item State of the art is WKB (FullerIV)
    \item Sensitive probe of WD conditions/models, e.g.\ discern whether certain
        white dwarfs are undergoing strong tidal heating (FullerIV).
    \item Require numerical simulations to determine dissipation profile.
\end{itemize}

\section{Proposed Research}

\textbf{Propose to study nonlinear wave breaking in white dwarf binaries via
numerical simulation}.

\begin{itemize}
    \item Research Plan
        \begin{itemize}
            \item For various WD models + compositions, run simulations to
                determine $E, \vec{L}$ dissipation distribution.
            \item Start with spectral code (Dedalus), existing literature shows
                that handles instability-driven turbulence with high accuracy
                and speed.
            \item Have the computational expertise to modify existing
                codes/build new code if existing ones are insufficient.
            \item Determine whether observed events e.g.\ Brown et al 2011 match
                our predictions for any models, other observational signatures.
            \item Study model output for simpler models to develop analytical
                tools.
        \end{itemize}
    \item Further applications
        \begin{itemize}
            \item Forming hot Jupiters (Michelle paper)
            \item \textbf{LIGO detecting kilonovae means many energetic
                phenomena to study, any chance to apply similar approach?}
        \end{itemize}
\end{itemize}

\end{document}

### 2019-01-24 proposal/FINESST/Su_FINESST_2019.tex
FINESST proposal draft v1

    \documentclass[12pt,
        usenames, % allows access to some tikz colors
        dvipsnames % more colors: https://en.wikibooks.org/wiki/LaTeX/Colors
    ]{article}
    \usepackage{
        amsmath,
        amssymb,
        times,
        fancyhdr, % page styling
        lastpage, % footer fanciness
        hyperref, % various links
        setspace, % line spacing
        amsthm, % newtheorem and proof environment
        mathtools, % \Aboxed for boxing inside aligns, among others
        float, % Allow [H] figure env alignment
        enumerate, % Allow custom enumerate numbering
        graphicx, % allow includegraphics with more filetypes
        wasysym, % \smiley!
        upgreek, % \upmu for \mum macro
        listings, % writing TrueType fonts and including code prettily
        tikz, % drawing things
        booktabs, % \bottomrule instead of hline apparently
        cancel % can cancel things out!
    }
    \usepackage[margin=1in]{geometry} % page geometry
    \usepackage[
        labelfont=bf, % caption names are labeled in bold
        font=scriptsize % smaller font for captions
    ]{caption}
    \usepackage[font=scriptsize]{subcaption} % subfigures

    \newcommand*{\scinot}[2]{#1\times10^{#2}}
    \newcommand*{\dotp}[2]{\left<#1\,\middle|\,#2\right>}
    \newcommand*{\rd}[2]{\frac{\mathrm{d}#1}{\mathrm{d}#2}}
    \newcommand*{\pd}[2]{\frac{\partial#1}{\partial#2}}
    \newcommand*{\rtd}[2]{\frac{\mathrm{d}^2#1}{\mathrm{d}#2^2}}
    \newcommand*{\ptd}[2]{\frac{\partial^2 #1}{\partial#2^2}}
    \newcommand*{\md}[2]{\frac{\mathrm{D}#1}{\mathrm{D}#2}}
    \newcommand*{\pvec}[1]{\vec{#1}^{\,\prime}}
    \newcommand*{\svec}[1]{\vec{#1}\;\!}
    \newcommand*{\bm}[1]{\boldsymbol{\mathbf{#1}}}
    \newcommand*{\ang}[0]{\;\text{\AA}}
    \newcommand*{\mum}[0]{\;\upmu \mathrm{m}}
    \newcommand*{\at}[1]{\left.#1\right|}

    \newtheorem{theorem}{Theorem}[section]

    \let\Re\undefined
    \let\Im\undefined
    \DeclareMathOperator{\Res}{Res}
    \DeclareMathOperator{\Re}{Re}
    \DeclareMathOperator{\Im}{Im}
    \DeclareMathOperator{\Log}{Log}
    \DeclareMathOperator{\Arg}{Arg}
    \DeclareMathOperator{\Tr}{Tr}
    \DeclareMathOperator{\E}{E}
    \DeclareMathOperator{\Var}{Var}
    \DeclareMathOperator*{\argmin}{argmin}
    \DeclareMathOperator*{\argmax}{argmax}
    \DeclareMathOperator{\sgn}{sgn}
    \DeclareMathOperator{\diag}{diag\;}

    \DeclarePairedDelimiter\bra{\langle}{\rvert}
    \DeclarePairedDelimiter\ket{\lvert}{\rangle}
    \DeclarePairedDelimiter\abs{\lvert}{\rvert}
    \DeclarePairedDelimiter\ev{\langle}{\rangle}
    \DeclarePairedDelimiter\p{\lparen}{\rparen}
    \DeclarePairedDelimiter\s{\lbrack}{\rbrack}
    \DeclarePairedDelimiter\z{\lbrace}{\rbrace}

    % \everymath{\displaystyle} % biggify limits of inline sums and integrals
    \tikzstyle{circ} % usage: \node[circ, placement] (label) {text};
        = [draw, circle, fill=white, node distance=3cm, minimum height=2em]
    \definecolor{commentgreen}{rgb}{0,0.6,0}
    \lstset{
        basicstyle=\ttfamily\footnotesize,
        frame=single,
        numbers=left,
        showstringspaces=false,
        keywordstyle=\color{blue},
        stringstyle=\color{purple},
        commentstyle=\color{commentgreen},
        morecomment=[l][\color{magenta}]{\#}
    }
    \usepackage[
        backend=bibtex8,
        style=nature]{biblatex}
\defbibenvironment{bibliography}
  {\noindent}
  {\unspace}
  {\printtext[labelnumberwidth]{%
     \printfield{labelprefix}%
     \printfield{labelnumber}}%
   \addspace}
\renewbibmacro*{finentry}{\finentry\addspace}
\bibliography{Su_FINESST_2019}
\AtEveryBibitem{\clearfield{title}}
\begin{document}

\def\Snospace~{\S{}} % hack to remove the space left after autorefs
\renewcommand*{\sectionautorefname}{\Snospace}
\renewcommand*{\appendixautorefname}{\Snospace}
\renewcommand*{\figureautorefname}{Fig.}
\renewcommand*{\equationautorefname}{Eq.}
\renewcommand*{\tableautorefname}{Tab.}

\singlespacing

\pagestyle{fancy}
\rfoot{Yubo Su}
\rhead{}
\cfoot{\thepage/\pageref{LastPage}}

\title{Nonlinear Tidal Dissipation in Binary White Dwarfs}
\author{Yubo Su}
\date{}

\maketitle

\section{White Dwarf Binaries}

Compact white dwarf (WD) binary systems, with orbital periods in the range of
minutes to hours, are important for several topics in astrophysics. They are
expected to be important sources of gravitational wave (GW) radiation for the
Laser Interferometric Space Antenna (LISA)\cite{lisa}. They are also thought to
be likely progenitors of type Ia supernovae (e.g.\ \cite{Ia0,webbink} or more
recently\cite{Ia1,Ia2}). While presently only a few tens of compact WD binaries are
known, \emph{Gaia} (currently gathering data) is expected to expand the catalog
to a few hundred, and the Large Synoptic Survey Telescope (LSST, first light
scheduled for 2020) is expected to detect a few thousand more\cite{lsst_wd}.
These observations are expected to significantly advance understanding of WD
binaries: results based on \emph{Gaia}'s second data release have already begun
to appear\cite{gaiaDD,gaiaDD2}. The proposed research is well-timed to take
advantage of these new resources.

In spite of the broad importance of WD binaries, the evolution of these systems
pre-merger is sparsely studied. In particular, compact WD binaries are expected
to exhibit major tidal interactions. In these interactions, tidal forces from
the companion WD excite internal gravity waves (IGW), displacements in the WD
fluid experiencing a buoyant restoring force due to density stratification.
As these waves propagate outwards from where they are excited, they grow in
amplitude until they break, as do ocean waves on a shore, and deposit both
energy and angular momentum from the binary orbit in the outer envelope of the
WD\@.

Previous works have shown that this tidal dissipation mechanism can generate
significantly more energy than thermal radiation from the WD surface alone and
is thus a major contributor to the WD energy budget\cite{fullerII,fullerIV}.
However, these works treated the nonlinear wave breaking in a parameterized
fashion. Related works in other fields find that properly including nonlinear
phenomena in hydrodynamical wave breaking produces drastically different IGW
dissipation behavior\cite{winters1994,barker_ogilvie}. Such fully nonlinear
studies have not been performed for WD binaries but are important for the
thermal and orbital evolution of WDs undergoing tidal heating. Characterizing
this nonlinear tidal dissipation will require numerical simulation to capture
the turbulent cascade to small scales that drives IGW dissipation.

\textbf{I propose to study the dynamical effects of tidal dissipation via
nonlinear IGW breaking in binary WDs.} To accomplish this, I aim to address the
following three aims:
\begin{itemize}
    \item Characterize the location, spatial extent, and other properties of
        wave breaking in realistic WD models via direct numerical simulation.
        The location and spatial extent will furnish a simple yet effective
        parameterization of tidal dissipation that can substitute full
        hydrodynamic simulation.

    \item Predict signatures of tidal dissipation over a wide range of possible
        WD systems. In particular, I will describe the impact of tidal heating
        on the apparent temperature of a binary WD and maybe even the production
        of observable flares.

    \item Compute modified GW templates for LISA that account for changes in the
        phase evolution of the orbit due to tidal dissipation.
\end{itemize}

\section{Nonlinear Tidal Dissipation}

\subsection{Background and Preliminary Work}

The current understanding of tidal synchronization in WD binaries is laid out
in~\cite{fullerII}: a tidally-excited train of IGW undergoes wave breaking
in the outer envelope of the WD, locally depositing angular momentum and
synchronizing the WD spin to the orbit. Such a process is broadly consistent
with canonical models of tidal synchronization in stars\cite{zahn75,gn89}, the
only major difference being in the specifics of wave excitation\footnote{While
IGWs in stars are excited at their radiative-convective boundaries, IGWs in WDs
are excited at sharp changes in composition.}. While this tidal synchronization
mechanism is generally accepted, direct numerical simulation of the wave
breaking process is sparse. Since wave breaking is a strongly nonlinear
phenomenon, where a larger wave breaks down into many smaller-scale waves,
numerical simulation is paramount to an accurate understanding of IGW breaking.

IGW breaking is well studied in atmospheric sciences. The wave breaking process
proceeds as follows: initially, the universal instability of an IGW to the
parametric subharmonic instability is responsible for transfering energy and
angular momentum from the unstable wave to the mean flow of the
fluid\cite{drazin}. Subsequently, after the mean flow velocity reaches the
horizontal phase velocity of the IGW, a well-known calculation shows that the
IGW is nearly completely absorbed by the mean flow in the linear
approximation and endowing the atmosphere with a mean horizontal
flow\cite{booker_bretherton,hazel}. However, when this mean flow absorption was
numerically studied including full nonlinear interactions, \emph{new phenomena
not described by the linear theory were observed}\cite{jones_num,winters1994}.
This highlights the importance of numerical simulation in capturing the wave
breaking process.

To this end, I have adapted the spectral hydrodynamic code Dedalus\cite{dedalus}
to simulate a plane-parallel IGW wavetrain excited at the bottom of a 2D
incompressible fluid. A spectral code like Dedalus is ideal for simulating
complex hydrodynamical phenomena as spectral methods have no inherent numerical
viscosity and so are more suitable for resolving the nonlinear cascade to small
length scales in wave breaking.

In my simulations, I observe IGW propagating upwards from where they are
excited, then breaking and depositing horizontal momentum in the fluid, causing
the fluid to acquire average horizontal momentum, consistent with
literature\cite{fullerII}. I derive simple formulae for the location and spatial
extent of the dissipation zone where the IGW is absorbed by the fluid. More
interestingly, I observe partial reflection of the IGW at the synchronization
layer\cite{me}, a phenomenon not considered in the current astrophysical
literature but consistent with atmospheric science results\cite{winters1994}. A
sample simulation is presented in \autoref{fig:nl_fluxes}. I am preparing these
results for publication\cite{me}.
\begin{figure}[!h]
    \centering
    \includegraphics[width=0.7\textwidth]{nl_fluxes.png}
    \caption{The evolution of the horizontal momentum flux $S_{px}(z)$ over time
    in a representative simulation. Note that the location where horizontal
    momentum is deposited by wave breaking, the sharp decrease in $S_{px}$,
    moves to lower $z$ at later times; this reflects spinup of the fluid from
    the top down. The incident flux appears to decrease over time as well, but
    this is attributed to partial reflection of the IGW off the critical
    layer. The vertical axis is in units of the total excited horizontal momentum
    flux, while distances and time are in units of the stratification scale
    height and the buoyancy period.}\label{fig:nl_fluxes}
\end{figure}

\subsection{Proposed Work}

It is therefore clear that full numerical modeling of IGW breaking is necessary
to fully characterize tidal dissipation. The first aim of my proposal is then:
\textbf{I will extend my preliminary results to characterize the dynamics of
nonlinear IGW breaking in realistic WDs.} Via numerical simulation, I will
develop simple models for tidal dissipation that can be used to study WD
evolution under tidal heating without performing further simulations. I will
continue my work in the following stages:
\begin{itemize}
    \item I will perform simulations examining the validity of my results in
        more realistic geometries, including polar and spherical geometries. I
        will continue to use Dedalus, which also supports these coordinate
        systems.

    \item I will extend my simulations to realistic WD models and equations of
        state such as those in~\cite{brassard1992}, continuing to track the
        location and spatial extent of the dissipation layer as well as any new
        phenomena. As WDs vary widely in composition and effective temperature,
        studying representative WD models is vital to obtaining a robust
        characterization of tidal dissipation. My selected WD models are those
        used in the current literature\cite{fullerII, fullerIV}.

    \item With these simulations, I will build a model of tidal dissipation
        valid for a broad range of WDs. I will package my model as a module for
        the 1D stellar evolution code Modules for Experiments in Stellar
        Astrophysics (MESA)\cite{MESA}. MESA is a popular choice for stellar
        evolution simulation due to its computational speed and extensibility.
        It has been used in previous studies to study the evolution of WDs under
        tidal heating\cite{fullerIV,tidal_novae}. My module will provide a
        parameterized and numerically-validated model of tidal dissipation that
        can be used to quickly and easily model tidal interactions in binary
        WDs.
\end{itemize}

\section{Tidal Heating Thermodynamics}

\subsection{Background}

As discussed earlier, compact WD binaries exhibit a range of observed and
hypothesized optical transients. Tidal novae\cite{tidal_novae}, underluminous
supernovae\cite{underlum}, and Ca-rich fast transients\cite{carich} are all
hypothesized to arise in WD binary systems. Given that tidal heating can become
a significant contributor to a WD's total energy budget, a realistic model of
tidal dissipation is important to understanding a binary WD's thermodynamic
evolution during inspiral.

In~\cite{tidal_novae} (hereafter FL), the authors use MESA to study the
production of tidal novae in binary WDs. A simple two-zone parameterization
where the tidal heat is deposited throughout the outer zone was used to model
tidal dissipation. They find that cool WDs in sufficiently compact binaries
(orbital period $\lesssim 15$ minutes) may incur a thermonuclear detonation of
the hydrogen envelope. WD binaries with such short orbital periods have been
observed, e.g.\ SDSS J065133+284423 has a period of 12.75 minutes\cite{12min}.
FL fits the observed properties of SDSS J065133+284423 to their tidal heating
models and finds evidence for tidal heating of the secondary WD in SDSS
J065133+284423. From their fits, FL are able to estimate the temperature of the
WD in the absence of tidal heating. As WDs form around a well known temperature
and cool predictably over time, accounting for tidal heating equates to a large
correction to the age of the WD\@. This illustrates the importance of accurately
modeling tidal interactions when inferring physical properties.

\subsection{Proposed Work}

With the tidal dissipation models I will develop, I will be able to perform a
simulation similar to that in FL but with a more realistic tidal interaction.
The second aim of my proposal is accordingly: \textbf{I will use my tidal
dissipation profiles to simulate a binary WD undergoing tidal heating and make
comparisons to observational data.} The location of the tidal heating is a key
ingredient in determining whether deposited energy can be efficiently transfered
away or whether the WD experiences sudden detonations. As such, an improved
understanding of the properties of tidal dissipation is important to proper
forecasting of binary WDs' thermodynamic evolution. Moreover, while only a few
sufficiently compact WD binaries were available at the time of writing of FL,
\emph{Gaia} data releases 2 and 3 will provide many more compact WD binaries to
examine for evidence of tidal heating.

Using MESA, I will evolve various WD models undergoing tidal heating. From these
simulations, I will extract the increased temperature of the WDs and make
comparisons to observational data, in particular to WDs in new \emph{Gaia} data
releases. I will also identify the occurrence rate and observational properties
of any predicted optical transients such as tidal novae and attempt to identify
them among existing detected events. Predictions of the occurence rates of such
phenomena are vital to guiding future observations. Finally, comparison of
observational data from known WD binaries to simulations could yield new
insights to the behavior of degenerate matter.

\section{Tidal Heating and LISA}

As discussed earlier, WD binaries are a primary source of GW radiation for
LISA\@. LISA will attain optimal sensitivity at frequencies
$10^{-4}$--$10^{-1}\;\mathrm{Hz}$\cite{LISA_band}. Exactly in this frequency
range, tidal effects act to synchronize the orbit of the WDs to the binary
orbit and transfer energy from the orbit into the WDs. While the decay
of the binary orbit is still mostly driven by GW radiation, the tidal energy
dissipation rate grows to $\sim10^{-2}$ the GW energy
dissipation\cite{fullerII,fullerIV}. An effect of such a magnitude causes the
phase of the emitted GWs to significantly from the point-mass binary
prediction; the emitted wave may exhibit ``missing cycles'' due to tidal
effects\cite{fullerII}. GW astronomy uses matched filtering, where a library of
template waveforms is matched against instrument data, to identify GW signals.
As such, the accuracy and completeness of the template library is of utmost
importance.

The final aim of my proposal is then: \textbf{I will use my tidal dissipation
model to compute WD binary GW waveforms including tidal dissipation for use in
the LISA detection pipeline}. This aim is much less computationally expensive
than it appears: LISA-band WD binaries can be well described using the Newtonian
dynamics of two co-orbiting point masses\cite{DWD_pointmass}. The resultant GW
emission can then be accurately computed using the weak gravity quadrupole
approximation (see e.g.~\cite{peters,lsst_wd}). Under these two approximations,
the GW waveform can be computed analytically without resorting to numerical
relativity simulations at all, an enormous computational savings. Thus, I will
compute GW waveforms accounting for the additional phase evolution due to tidal
dissipation. I will publish my corrected waveforms for use by LISA and the GW
community.

\section{Project Timeline}

During the first year of work and first half of the second year, I will complete
my tidal dissipation model. I anticipate the extension of my 2D plane-parallel
work to more complex geometries will be complete within the first half year,
while the extension to realistic WD models will take up to a year. I expect that
these two results together will produce one peer-reviewed publication and one
MESA module.

During the following year, I will use my tidal dissipation model to perform MESA
simulations of tidally heated WDs and compare to observational data. I expect my
MESA simulations and extracting appropriate observables to take about half a
year. I then intend to spend another half year analyzing observational data for
potential signatures of tidal heating or tidal novae. I expect both of these
phases will produce peer-reviewed publications.

Finally, in the last six months I will compute WD binary GW templates for use by
the LISA community. This work will also produce one peer-reviewed publication.

\section{Relevance to NASA Objectives}

This project is extremely relevant to the NASA Astrophysics research program.
My work directly relates to (i) the interactions of particles under the extreme
conditions found in astrophysical situations, (ii) how complex systems create
and shape the structure and composition of the universe on all scales, and (iii)
the development of new techniques that can be applied to future major missions.

My work will improve understanding of possible energetic phenomena that can
occur in compact object binaries. My results concerning angular momentum
transfer via IGW breaking will be applicable to astrophysical systems beyond WD
binaries, for instance angular momentum transfer in stars\cite{l_trans_rev}. I
will devise new ways of analyzing astrophysically interesting systems in
\emph{Gaia} and LSST\@. Finally, my GW templates will be important for LISA GW
detection efforts.

My work has direct relevance to NASA missions in the detectability of
astrophysical transients related to WD binaries. For instance, tidal novae are
theorized to have a similar observational signature to dwarf novae which have
been observed with Chandra and the Hubble Space Telescope, among others.

\printbibliography

\end{document}

### 2019-01-27 proposal/FINESST/Su_FINESST_2019.tex
update plots, draft v2 of FINESST proposal

@@
 
-\section{White Dwarf Binaries}
+\section{Background and Proposed Research}
+
+\subsection{White Dwarf Binaries}
 
 Compact white dwarf (WD) binary systems, with orbital periods in the range of
-minutes to hours, are important for several topics in astrophysics. They are
-expected to be important sources of gravitational wave (GW) radiation for the
-Laser Interferometric Space Antenna (LISA)\cite{lisa}. They are also thought to
-be likely progenitors of type Ia supernovae (e.g.\ \cite{Ia0,webbink} or more
-recently\cite{Ia1,Ia2}). While presently only a few tens of compact WD binaries are
-known, \emph{Gaia} (currently gathering data) is expected to expand the catalog
-to a few hundred, and the Large Synoptic Survey Telescope (LSST, first light
-scheduled for 2020) is expected to detect a few thousand more\cite{lsst_wd}.
-These observations are expected to significantly advance understanding of WD
-binaries: results based on \emph{Gaia}'s second data release have already begun
-to appear\cite{gaiaDD,gaiaDD2}. The proposed research is well-timed to take
-advantage of these new resources.
+minutes to hours, are important for a range of astrophysical problems. They are
+the most important sources of gravitational waves (GWs) for the Laser
+Interferometric Space Antenna (LISA)\cite{lisa}. They are also thought to
+produce interesting optical transients such as tidal novae\cite{tidal_novae},
+underluminous supernovae\cite{underlum}, and Ca-rich fast
+transients\cite{carich}. Most importantly, they have been proposed as the likely
+progenitors of type Ia supernovae (e.g.~\cite{Ia0,webbink} or more
+recently\cite{Ia1,Ia2}). While presently only a few tens of compact WD binaries
+are known\cite{lsst_wd}, \emph{Gaia} (currently gathering data) is expected to
+expand the catalog to a few hundred\cite{lsst_wd} (results based on
+\emph{Gaia}'s second data release have already begun to
+appear\cite{gaiaDD,gaiaDD2}), and the Large Synoptic Survey Telescope (LSST,
+first light scheduled for 2020) will likely detect a few thousand
+more\cite{lsst_wd}. These observations will significantly advance the
+understanding of WD binaries. My proposed theoretical and computational research
+is well-timed to take advantage of these new advances.
 
 In spite of the broad importance of WD binaries, the evolution of these systems
-pre-merger is sparsely studied. In particular, compact WD binaries are expected
-to exhibit major tidal interactions. In these interactions, tidal forces from
-the companion WD excite internal gravity waves (IGW), displacements in the WD
-fluid experiencing a buoyant restoring force due to density stratification.
-As these waves propagate outwards from where they are excited, they grow in
-amplitude until they break, as do ocean waves on a shore, and deposit both
-energy and angular momentum from the binary orbit in the outer envelope of the
-WD\@.
+prior to their final merger is not well understood. Much of this uncertainty
+comes from our imprecise understanding of tidal interactions, which play an
+important role during a compact WD binary's inspiral\cite{fullerII}. Studies
+show that these interactions manifest as tidal excitation of internal gravity
+waves (IGW), waves in the WD fluid restored by the buoyancy force due to density
+stratification\cite{fullerI}. As these waves propagate outwards towards the
+surface, they grow in amplitude until they break, as do ocean waves on a shore,
+and transfer both energy and angular momentum from the binary orbit to the outer
+envelope of the WD\cite{fullerI,fullerII}.
 
-Previous works have shown that this tidal dissipation mechanism can generate
+Previous works have found that this tidal dissipation mechanism can generate
 significantly more energy than thermal radiation from the WD surface alone and
 is thus a major contributor to the WD energy budget\cite{fullerII,fullerIV}.
-However, these works treated the nonlinear wave breaking in a parameterized
-fashion. Related works in other fields find that properly including nonlinear
-phenomena in hydrodynamical wave breaking produces drastically different IGW
-dissipation behavior\cite{winters1994,barker_ogilvie}. Such fully nonlinear
-studies have not been performed for WD binaries but are important for the
-thermal and orbital evolution of WDs undergoing tidal heating. Characterizing
-this nonlinear tidal dissipation will require numerical simulation to capture
-the turbulent cascade to small scales that drives IGW dissipation.
+However, these works parameterized the wave breaking process without
+hydrodynamical simulations. The details of tidal dissipation, namely the
+location and spatial extent of the wave breaking, affect the observable outcome:
+dissipation near the surface of the WD can be efficiently radiated away and
+simply brightens the WD, while dissipation deep in the WD causes an energy
+buildup that results in energetic flares\cite{tidal_novae}. Similar works in
+other fields find that numerical simulations capturing the strongly nonlinear
+wave breaking process find new behavior that cannot be described by linear and
+weakly nonlinear theory\cite{winters1994,barker_ogilvie}. Such fully nonlinear
+numerical simulations have not been performed for WDs.
 
-\textbf{I propose to study the dynamical effects of tidal dissipation via
-nonlinear IGW breaking in binary WDs.} To accomplish this, I aim to address the
-following three aims:
+\subsection{Proposed Research}
+
+Characterizing the location and spatial extent of tidal dissipation in WD
+binaries will require numerical simulation to capture the turbulent cascade to
+small scales causing wave breaking. \textbf{I propose to study the dynamical
+effects of tidal dissipation via nonlinear IGW breaking in binary WDs.} To
+accomplish this, I aim to address the following three aims:
 \begin{itemize}
@@
         The location and spatial extent will furnish a simple yet effective
-        parameterization of tidal dissipation that can substitute full
-        hydrodynamic simulation.
+        parameterization of tidal dissipation. \autoref{s:2} describes the
+        steps required for this aim.
 
@@
         on the apparent temperature of a binary WD and maybe even the production
-        of observable flares.
+        of observable flares. \autoref{s:3} details my plan to do so.
 
     \item Compute modified GW templates for LISA that account for changes in the
-        phase evolution of the orbit due to tidal dissipation.
+        phase evolution of the orbit due to tidal dissipation. \autoref{s:4}
+        elaborates on how I will perform this computation.
 \end{itemize}
 
-\section{Nonlinear Tidal Dissipation}
+\section{Nonlinear Tidal Dissipation}\label{s:2}
 
@@
 The current understanding of tidal synchronization in WD binaries is laid out
-in~\cite{fullerII}: a tidally-excited train of IGW undergoes wave breaking
-in the outer envelope of the WD, locally depositing angular momentum and
-synchronizing the WD spin to the orbit. Such a process is broadly consistent
-with canonical models of tidal synchronization in stars\cite{zahn75,gn89}, the
-only major difference being in the specifics of wave excitation\footnote{While
-IGWs in stars are excited at their radiative-convective boundaries, IGWs in WDs
-are excited at sharp changes in composition.}. While this tidal synchronization
-mechanism is generally accepted, direct numerical simulation of the wave
-breaking process is sparse. Since wave breaking is a strongly nonlinear
-phenomenon, where a larger wave breaks down into many smaller-scale waves,
-numerical simulation is paramount to an accurate understanding of IGW breaking.
+in~\cite{fullerII}: a tidally-excited train of IGW undergoes wave breaking in
+the outer envelope of the WD, locally depositing angular momentum and
+synchronizing the WD spin to the binary orbit. A similar process may also
+operate in stellar binaries consisting of early type stars\cite{zahn75,gn89},
+the only major difference being in the specifics of wave
+excitation\footnote{While IGWs in stars are excited at their
+radiative-convective boundaries, IGWs in WDs are excited at sharp changes in
+composition.}. Nevertheless, direct numerical simulation of the wave breaking
+process has not been performed in either of these systems. Since wave breaking
+is a strongly nonlinear phenomenon, where a larger wave breaks down into many
+smaller-scale waves, numerical simulation is paramount to an accurate
+understanding of IGW breaking.
 
-IGW breaking is well studied in atmospheric sciences. The wave breaking process
-proceeds as follows: initially, the universal instability of an IGW to the
-parametric subharmonic instability is responsible for transfering energy and
-angular momentum from the unstable wave to the mean flow of the
+IGW breaking has been studied in detail in atmospheric sciences. The wave
+breaking process proceeds as follows: initially, as the IGW reaches nonlinear
+amplitudes, it breaks down via the parametric subharmonic instability and
+transfers energy and angular momentum from the wave to the mean flow of the
 fluid\cite{drazin}. Subsequently, after the mean flow velocity reaches the
-horizontal phase velocity of the IGW, a well-known calculation shows that the
-IGW is nearly completely absorbed by the mean flow in the linear
-approximation and endowing the atmosphere with a mean horizontal
-flow\cite{booker_bretherton,hazel}. However, when this mean flow absorption was
-numerically studied including full nonlinear interactions, \emph{new phenomena
-not described by the linear theory were observed}\cite{jones_num,winters1994}.
-This highlights the importance of numerical simulation in capturing the wave
-breaking process.
+horizontal phase velocity of the IGW, a critical layer through which the IGW
+cannot propagate forms. A well-known calculation shows that the IGW is nearly
+completely absorbed at this critical layer in the linear approximation and
+endows the atmosphere with a mean horizontal flow\cite{booker_bretherton,hazel}.
+However, when this mean flow absorption was numerically studied including full
+nonlinear interactions, new phenomena not described by the linear theory
+(reflection off the critical layer and sharpening of the mean flow) were
+observed\cite{jones_num,winters1994} that affected the evolution of the
+atmosphere over time. \emph{This highlights the importance of numerical
+simulation in capturing the wave breaking process.}
 
-To this end, I have adapted the spectral hydrodynamic code Dedalus\cite{dedalus}
-to simulate a plane-parallel IGW wavetrain excited at the bottom of a 2D
-incompressible fluid. A spectral code like Dedalus is ideal for simulating
-complex hydrodynamical phenomena as spectral methods have no inherent numerical
-viscosity and so are more suitable for resolving the nonlinear cascade to small
-length scales in wave breaking.
+To gain insight into the tidal dissipation process, I began by using the
+spectral hydrodynamics code Dedalus\cite{dedalus} to study IGW breaking in a 2D
+isothermal, stratified atmosphere. A spectral code like Dedalus is ideal for
+simulating complex hydrodynamical phenomena, as spectral methods have no
+inherent numerical viscosity and so better resolve the nonlinear cascade to
+small length scales in wave breaking.
 
-In my simulations, I observe IGW propagating upwards from where they are
-excited, then breaking and depositing horizontal momentum in the fluid, causing
-the fluid to acquire average horizontal momentum, consistent with
-literature\cite{fullerII}. I derive simple formulae for the location and spatial
+Working with Dr.\ Daniel Lecoanet (Princeton; one of the authors of the Dedalus
+code) and my advisor, Prof.\ Dong Lai, I simulated an upward-propagating IGW
+wavetrain excited at the bottom of the atmosphere. I observe the excited wave
+breaking and depositing horizontal momentum in the fluid, causing the fluid to
+acquire an average horizontal flow, consistent with previous
+studies\cite{fullerII}. I derive simple formulae for the location and spatial
 extent of the dissipation zone where the IGW is absorbed by the fluid. More
@@
 layer\cite{me}, a phenomenon not considered in the current astrophysical
-literature but consistent with atmospheric science results\cite{winters1994}. A
+literature but consistent with the aforementioned results\cite{winters1994}. A
 sample simulation is presented in \autoref{fig:nl_fluxes}. I am preparing these
 results for publication\cite{me}.
+
 \begin{figure}[!h]
@@
     \includegraphics[width=0.7\textwidth]{nl_fluxes.png}
-    \caption{The evolution of the horizontal momentum flux $S_{px}(z)$ over time
-    in a representative simulation. Note that the location where horizontal
-    momentum is deposited by wave breaking, the sharp decrease in $S_{px}$,
-    moves to lower $z$ at later times; this reflects spinup of the fluid from
-    the top down. The incident flux appears to decrease over time as well, but
-    this is attributed to partial reflection of the IGW off the critical
-    layer. The vertical axis is in units of the total excited horizontal momentum
-    flux, while distances and time are in units of the stratification scale
-    height and the buoyancy period.}\label{fig:nl_fluxes}
+    \caption{The evolution of the average horizontal flow velocity of the fluid
+    $U_0$ (in units of $c_{ph,x}$ the horizontal phase velocity of the IGW) and
+    horizontal momentum flux contained in the IGW $S_{px}(z)$ (in units of the
+    total excited flux $S_0$) for one of my simulations. Different colored lines
+    correspond to different times in the simulation. Times are measured in units
+    of $N^{-1}$, the inverse of the Brunt-V\"ais\"al\"a frequency (or the
+    buoyancy period), and heights in units of $H$ the scale height of the
+    density stratification. IGW are excited at $z = 2H$ and propagate to higher
+    $z$ before undergoing wave breaking. The sharp decrease in $S_{px}$ is
+    the location of the critical layer where the IGW breaks and drives $U_0$
+    towards $c_{ph, x}$; it moves to lower $z$ over time. The apparent decrease
+    in $S_{px}$ at later times indicates reflection off the critical
+    layer.}\label{fig:nl_fluxes}
 \end{figure}
@@
         used in the current literature\cite{fullerII, fullerIV}.
-
-    \item With these simulations, I will build a model of tidal dissipation
-        valid for a broad range of WDs. I will package my model as a module for
-        the 1D stellar evolution code Modules for Experiments in Stellar
-        Astrophysics (MESA)\cite{MESA}. MESA is a popular choice for stellar
-        evolution simulation due to its computational speed and extensibility.
-        It has been used in previous studies to study the evolution of WDs under
-        tidal heating\cite{fullerIV,tidal_novae}. My module will provide a
-        parameterized and numerically-validated model of tidal dissipation that
-        can be used to quickly and easily model tidal interactions in binary
-        WDs.
 \end{itemize}
 
-\section{Tidal Heating Thermodynamics}
+\section{Tidal Heating Thermodynamics}\label{s:3}
 
@@
 The second aim of my proposal is accordingly: \textbf{I will use my tidal
+% "obtained in Section 2"
 dissipation profiles to simulate a binary WD undergoing tidal heating and make
@@
 
-\section{Tidal Heating and LISA}
+\section{Tidal Heating and LISA}\label{s:4}
 
@@
 while the extension to realistic WD models will take up to a year. I expect that
-these two results together will produce one peer-reviewed publication and one
-MESA module.
+these two results together will produce one peer-reviewed publication in
+addition to the one currently in preparation.
 

### 2019-01-27 sims/2d_3_final/strat_helper.py

@@ -4,6 +4,7 @@ helper function to run the shared stratification scenario. user just has to
 specify BCs and ICs
 '''
 import logging
+import pickle
 logger = logging.getLogger()
 
 import os
@@ -14,7 +15,9 @@ import matplotlib
 matplotlib.use('Agg')
 import matplotlib.pyplot as plt
 import numpy as np
-plt.style.use('ggplot')
+# plt.style.use('ggplot')
+plt.rc('text', usetex=True)
+plt.rc('font', family='serif')
 
 from dedalus import public as de
 from dedalus.tools import post
@@ -477,7 +480,7 @@ def plot_front(name, params):
     N_Z = params['N_Z'] * params['INTERP_Z']
     dyn_vars = ['uz', 'ux', 'U', 'W', 'ux_z']
     snapshots_dir = SNAPSHOTS_DIR % name
-    logfile = '%s/data.log' % snapshots_dir
+    logfile = '%s/data.pkl' % snapshots_dir
 
     sim_times = []
     front_pos = []
@@ -498,21 +501,15 @@ def plot_front(name, params):
         z0 = z[0]
 
         F_px = np.sum(state_vars['F_px'], axis=1) / N_X
+        u0 = np.sum(state_vars['ux'], axis=1) / N_X
 
-        with open(logfile, 'w') as data:
-            data.write(repr(z0.tolist()))
-            data.write('\n')
-            data.write(repr(sim_times.tolist()))
-            data.write('\n')
-            data.write(repr(F_px.tolist()))
-            data.write('\n')
+        with open(logfile, 'wb') as data:
+            pickle.dump((z0, sim_times, F_px, u0), data)
 
     else:
-        print('data.log found, loading')
-        with open(logfile) as data:
-            z0 = np.array(eval(data.readline()))
-            sim_times = np.array(eval(data.readline()))
-            F_px = np.array(eval(data.readline()))
+        print('data found, loading')
+        with open(logfile, 'rb') as data:
+            z0, sim_times, F_px, u0 = pickle.load(data)
 
     for t_idx, sim_time in enumerate(sim_times):
         max_pos = len(F_px[t_idx]) - 1
@@ -523,6 +520,7 @@ def plot_front(name, params):
 
     start_idx = 10
[…]

### 2019-01-29 presentations/meeting.tex
1/29 meeting

--- 2018-11-05 presentations/Su_DL_110618.tex
@@
     \newcommand*{\ang}[0]{\;\text{\AA}}
-    \newcommand*{\mum}[0]{\\;upmu \mathrm{m}}
+    \newcommand*{\mum}[0]{\;\upmu \mathrm{m}}
     \newcommand*{\at}[1]{\left.#1\right|}
@@
 
-\title{Quick Research Review}
-\author{Yubo Su}
-\date{Nov. 06, 2018}
+\begin{frame}
+    \frametitle{Linear}
 
-\maketitle
-
-\begin{frame}
-    \frametitle{Objectives}
-
-    \begin{itemize}
-        \item Continuous train of IGW excited in stratified WD atmosphere by
-            binary companion, grows due to stratification, reaches nonlinear
-            amplitudes $\xi_z k_z \gtrsim 1$ before the peak of the envelope.
-
-        \item Seek long-term behavior of nonlinear dissipation, whether
-            steady-state or periodic.
-    \end{itemize}
+    \begin{figure}[t]
+        \centering
+        \includegraphics[width=0.7\textwidth]{../sims/2d_3_final/snapshots_lin_0/fluxes.png}
+        \caption{Linear simulation.}
+    \end{figure}
 \end{frame}
@@
 \begin{frame}
-    \frametitle{Numerical Simulations, No Shear}
-    \framesubtitle{Setup}
+    \frametitle{New Plotting for Nonlinear}
 
-    \begin{itemize}
-        \item 2D, Fourier $\times$ Chebyshev, damping zones, ran at both $128
-            \times 512$ and $256 \times 1024$ over $3H \times 12H$ grid, where
-            $H$ is scale height.
-
-        \item Target waves $k_x = \frac{2\pi}{3H}, k_z = \frac{2\pi}{H}, \omega
-            \approx 0.32N$, $N$ B-V frequency.
-
-        \item Wave is excited by volumetric forcing
-            $e^{-z^2/2\sigma^2}\cos\p*{k_xx - \omega t}$ with $\sigma =
-            \frac{3H}{32} \lesssim \frac{1}{k_z}$.
-
-        \item Regularized by numerical viscosity $\nu \gtrsim
-            \frac{\omega}{k_{\max}k_z}$.
-    \end{itemize}
+    \begin{figure}[t]
+        \centering
+        \includegraphics[width=0.7\textwidth]{../sims/2d_3_final/snapshots_nl_3/fluxes.png}
+        \caption{Nonlinear simulation, new plotting methodology. Feedback?}
+    \end{figure}
 \end{frame}
@@
 \begin{frame}
-    \frametitle{Numerical Simulations, No Shear}
-    \framesubtitle{Results}
+    \frametitle{Reflection?}
 
-    \begin{itemize}
-        \item High resolution ($\nu = 4\frac{\omega}{k_{\max}k_z}$):
-            \lstinline{nl3.mp4}. Note that mean flow starts to build up at the
-            outgoing damping zone, and once it hits criticality it starts to
-            move down while continuing to steepen. Eventually, it stops moving
-            and produces reflection below, while leaving behind a sinusoidal
-            mean flow above.
-
-        \item Low resolution ($\nu = 3\frac{\omega}{k_{\max}k_z}$):
-            \lstinline{nl3_lowres.mp4}. Qualitatively similar, but the mean flow
-            forms later.
-
-        \item Two hypotheses:
-            \begin{itemize}
-                \item Damping layer is causing mean flow to begin building up
-                    like Daniel said; how can we fix? Seems like low $k_z$ modes
-                    (which have higher $v_{ph}$).
-
-                \item Further growth/propagation is critical layer absorption?
-                    Eventual reflection might build up when shear flow is too
-                    steep $\mathrm{Ri} \gtrsim 1/2$ or WKB breaks down
-                    $\pd{k_z}{z} \sim \frac{k_z}{\lambda_z}$.
-            \end{itemize}
-    \end{itemize}
+    Get exact solution, up to viscous dissipation. Plot $\delta u_z =
+    \frac{u_{z,sim} - u_{z, anal}}{\abs*{u_{z, anal}}}$, ``fractional
+    deviation'' between driving zone and critical layer. Compute $\mathrm{RMS} =
+    \sqrt{\ev*{\delta u_z^2}}$.
+    \begin{figure}[t]
+        \centering
+        \includegraphics[width=0.7\textwidth]{../sims/2d_3_final/snapshots_lin_1/p_074.png}
+        \caption{Linear simulation with same viscosity as nonlinear. Green is
+        viscosity-free envelope, orange analytical.}
+    \end{figure}
 \end{frame}
@@
 \begin{frame}
-    \frametitle{Numerical Simulations, Shear Flow}
-    \framesubtitle{Setup}
+    \frametitle{Reflection (nl4)}
 
-    \begin{itemize}
-        \item 2D, Fourier $\times$ Fourier, damping zones, running at $64 \times
-            256$. No height stratification, domain is $ H \times H$ (though $H$
-            is now physically meaningless).
-
-        \item Target waves $k_x = \frac{2\pi}{H}, k_z = \frac{20\pi}{H}, \omega
-            \approx 0.1N$, $N$ B-V frequency. Volumetric forcing.
-
-        \item Regularized by $\nabla^6$ hyperviscosity.
-    \end{itemize}
+    Get exact solution, up to viscous dissipation. Plot $\delta u_z =
+    \frac{u_{z,sim} - u_{z, anal}}{\abs*{u_{z, anal}}}$, ``fractional
+    deviation'' between driving zone and critical layer. Compute $\mathrm{RMS} =
+    \sqrt{\ev*{\delta u_z^2}}$.
+    \begin{figure}[t]
+        \centering
+        \includegraphics[width=0.7\textwidth]{../sims/2d_3_final/snapshots_nl_4/p_074.png}
+        \caption{Nonlinear simulation. Green is viscosity-free envelope, orange
+        analytical.}
+    \end{figure}
 \end{frame}
@@
 \begin{frame}
-    \frametitle{Numerical Simulations, Shear Flow}
-    \framesubtitle{Results}
+    \frametitle{Reflection?}
 
-    \begin{itemize}
-        \item Initialize w/ shear flow and let evolve. Try both narrow and wide
-            profiles (\lstinline{vstrat*.mp4}).
+    \begin{table}[t]
+        \centering
+        \begin{tabular}{l|l l }
+            Name & RMS & Notes\\
+            lin0 & 0.077 & Almost-no visc\\
+            lin1 & 0.067 & Same visc as nl4\\
+            nl1 (low) & 0.132 & half-res nl1, higher visc\\
+            nl1 & 0.120 & $0.7\nu$\\
+            nl2 & 0.213 & $0.4\nu$\\
+            nl3 & 0.179 & $0.4\nu$ (double A)\\
+            nl4 & 0.269 & $0.3\nu$\\
+            nl5 & 0.207 & $0.3\nu$ (double $k_z$)\\
+            nl6 & 0.246 & $0.2\nu$\\
+            nl7 & 0.358 & $0.1\nu$
+        \end{tabular}
+        \caption{Table of RMS values.}
+    \end{table}
+\end{frame}
 
-        \item Conclusion: attenuated transmission when thin, barely-critical
-            layer, smooth WKB-like wavelength shortening when broad, critical
-            flow, but steep $\pd{U_0}{z}$ gives reflection!
+\begin{frame}
+    \frametitle{Reflection (nl6)}
 
-        \item Recall Booker \& Bretherton result
-            \begin{align}
-                T &\sim \exp\s*{-\pi \sqrt{\frac{N^2}{(U_0')^2} - \frac{1}{4}}}
-                & R &\sim
-                    \exp\s*{-2\pi \sqrt{\frac{N^2}{(U_0')^2} - \frac{1}{4}}}.
-            \end{align}
-            WKB criterion $\Rightarrow \frac{k_z^2}{N}U_0' \frac{\lambda^2}{2}
-            \ll 1$.
+    Get exact solution, up to viscous dissipation. Plot $\delta u_z =
+    \frac{u_{z,sim} - u_{z, anal}}{\abs*{u_{z, anal}}}$, ``fractional
+    deviation'' between driving zone and critical layer. Compute $\mathrm{RMS} =
+    \sqrt{\ev*{\delta u_z^2}}$.
+    \begin{figure}[t]
+        \centering
+        \includegraphics[width=0.7\textwidth]{../sims/2d_3_final/snapshots_nl_6/p_074.png}
+        \caption{Nonlinear simulation. Green is viscosity-free envelope, orange
+        analytical.}
+    \end{figure}
+\end{frame}
 
-        \item Winters d'Asaro had a local $\mathrm{Ri} \sim 0.5$.
-    \end{itemize}
+\begin{frame}
+    \frametitle{Reflection (nl7)}
+
+    Get exact solution, up to viscous dissipation. Plot $\delta u_z =
+    \frac{u_{z,sim} - u_{z, anal}}{\abs*{u_{z, anal}}}$, ``fractional
+    deviation'' between driving zone and critical layer. Compute $\mathrm{RMS} =
+    \sqrt{\ev*{\delta u_z^2}}$.
+    \begin{figure}[t]
+        \centering
+        \includegraphics[width=0.7\textwidth]{../sims/2d_3_final/snapshots_nl_7/p_074.png}
+        \caption{Nonlinear simulation. Green is viscosity-free envelope, orange
+        analytical.}
+    \end{figure}
+\end{frame}
+
+\begin{frame}
+    \frametitle{Reflection (nl3)}
+
+    Get exact solution, up to viscous dissipation. Plot $\delta u_z =
+    \frac{u_{z,sim} - u_{z, anal}}{\abs*{u_{z, anal}}}$, ``fractional
+    deviation'' between driving zone and critical layer. Compute $\mathrm{RMS} =
+    \sqrt{\ev*{\delta u_z^2}}$.
+    \begin{figure}[t]
+        \centering
+        \includegraphics[width=0.7\textwidth]{../sims/2d_3_final/snapshots_nl_3/p_058.png}
+        \caption{Nonlinear simulation. Green is viscosity-free envelope, orange
+        analytical.}
+    \end{figure}
+\end{frame}
+
+\begin{frame}
+    \frametitle{Reflection (nl5)}
+
+    Get exact solution, up to viscous dissipation. Plot $\delta u_z =
+    \frac{u_{z,sim} - u_{z, anal}}{\abs*{u_{z, anal}}}$, ``fractional
+    deviation'' between driving zone and critical layer. Compute $\mathrm{RMS} =
+    \sqrt{\ev*{\delta u_z^2}}$.
+    \begin{figure}[t]
+        \centering
+        \includegraphics[width=0.7\textwidth]{../sims/2d_3_final/snapshots_nl_5/p_074.png}
+        \caption{Nonlinear simulation. Green is viscosity-free envelope, orange
+        analytical.}
+    \end{figure}
 \end{frame}

### 2019-02-01 writeups/notes.tex
check in some temp paper work

@@
     \begin{cases}
-        e^{(z - z_0)/2H}e^{i(k_xx + k_z(z - z_0) - \omega t)} & z > z_0\\
-        e^{(z - z_0)/2H}e^{i(k_xx - k_z(z - z_0) - \omega t)} & z < z_0\\
+        e^{(z - z_0)/2H}e^{i(k_xx + k_z(z - z_0) - \omega t)
+            + \frac{1}{2k_zH}} & z > z_0\\
+        e^{(z - z_0)/2H}e^{i(k_xx - k_z(z - z_0) - \omega t)
+            + \frac{1}{2k_zH}} & z < z_0\\
     \end{cases}.
 \end{equation}
+The extra phase shift comes from $\pd{}{z} \approx ik_z + \frac{1}{2H} \approx
+k_z\p*{e^{-\frac{1}{2k_zH}}}$.
 
@@
 
+\chapter{Tidal Synchronization}
+
+Chapter containing notes slightly more pertaining to the paper. I've kinda not
+been updating my notes here for the last few months, so I'll just add what I
+can.
+
+\section{Momentum-Conserving Equations}
+
+First, we want to simplate the Euler equations, but $\frac{\vec{\nabla}P}{\rho}$
+is difficult to deal with. Instead, define $\varpi = P / \rho$, then we can
+divide $\rho$ out of the momentum equation and get no quotients of dynamical
+variables. It is then easier to use $\Upsilon: \rho = \rho e^\Upsilon$.
+Proceeding:
+
+We start from the flux-conservative Euler equations and manually insert
+artificial viscosity terms in a momentum-conserving way:
+\begin{subequations}
+    \begin{align}
+        \vec{\nabla} \cdot \vec{u} &= 0,\\
+        \partial_t \rho + \vec{\nabla} \cdot (\rho \vec{u} - \nu
+            \vec{\nabla}(\rho - \rho_0)) &= 0,\label{eq:visc_cons_mom}\\
+        \partial_t (\rho \vec{u}) + \vec{\nabla} \cdot (\rho \vec{u} \vec{u} +
+            \mathrm{diag}(\rho \varpi) - \nu \rho \vec{\nabla}\vec{u}) + \rho g
+            \hat{z} &= 0.
+    \end{align}
+\end{subequations}
+We can do some algebra and obtain
+\begin{subequations}
+    \begin{align}
+        \vec{\nabla} \cdot \vec{u} &= 0,\\
+        \partial_t \Upsilon + \p*{\vec{u} \cdot \vec{\nabla}} \Upsilon -
+            \frac{u_z}{H} - \nu\p*{\nabla^2 \Upsilon + \p*{\vec{\nabla}
+            \Upsilon} \cdot \p*{\vec{\nabla}\Upsilon} - \frac{2}{H}\partial_z
+            \Upsilon + \frac{1 - e^{-\Upsilon}}{H^2}} &= 0,\\
+        \partial_t \vec{u} + \p*{\vec{u} \cdot \vec{\nabla}}\vec{u} +
+            \vec{\nabla} \varpi + \varpi \vec{\nabla} \Upsilon - \nu \nabla^2
+            \vec{u} + \vec{u} \nu\p*{\nabla^2 \Upsilon + \p*{\vec{\nabla}
+            \Upsilon} \cdot \p*{\vec{\nabla}\Upsilon} - \frac{2}{H}\partial_z
+            \Upsilon + \frac{1 - e^{-\Upsilon}}{H^2}}&{}\nonumber\\
+        - 2\nu \p*{\p*{\p*{\vec{\nabla}\Upsilon} \cdot \vec{\nabla}}\vec{u} -
+            \frac{1}{H}\partial_z \vec{u}} - \frac{\varpi_1}{H} &= 0.
+    \end{align}
+\end{subequations}
+Including numerical dissipation, sponge layers and driving gives the full system
+of equations
+\begin{subequations}
+    \begin{align}
+        \vec{\nabla} \cdot \vec{u} ={}& 0,\\
+        \partial_t \Upsilon - \frac{u_z}{H}
+            ={}& \nu\p*{\nabla^2 \Upsilon + \p*{\vec{\nabla}
+            \Upsilon} \cdot \p*{\vec{\nabla}\Upsilon} - \frac{2}{H}\partial_z
+            \Upsilon + \frac{1 - e^{-\Upsilon}}{H^2}},\nonumber\\
+            & - \p*{\vec{u} \cdot \vec{\nabla}}\Upsilon
+                -\Gamma(z) \Upsilon
+                + \frac{F}{\rho_0(z)}e^{-\frac{(z - z_0)^2}{2\sigma^2}}
+                    \cos \p*{k_xx - \omega t},\\
+        \pd{u_x}{t} + \pd{T}{x} + gH\pd{\Upsilon}{x} ={}&
+            \nu \nabla^2 u_x
+            - u_x \nu\p*{\nabla^2 \Upsilon + \p*{\vec{\nabla} \Upsilon} \cdot
+                \p*{\vec{\nabla}\Upsilon} - \frac{2}{H}\partial_z \Upsilon
+                + \frac{1 - e^{-\Upsilon}}{H^2}}\nonumber\\
+            &+ 2\nu \p*{\p*{\p*{\vec{\nabla}\Upsilon} \cdot \vec{\nabla}}u_x
+                - \frac{1}{H}\partial_z u_x}
+            -\Gamma(z) u_x
+                - \p*{\vec{u} \cdot \vec{\nabla}}u_x
+                - T_1 \pd{\Upsilon}{x},\\
+        \pd{u_z}{t} + \pd{T}{z} + gH\pd{\Upsilon}{z} - \frac{T_1}{H} ={}&
+            \nu \nabla^2 u_z
+            - u_z \nu\p*{\nabla^2 \Upsilon + \p*{\vec{\nabla} \Upsilon} \cdot
+                \p*{\vec{\nabla}\Upsilon} - \frac{2}{H}\partial_z \Upsilon
+                + \frac{1 - e^{-\Upsilon}}{H^2}}\nonumber\\
+            &+ 2\nu \p*{\p*{\p*{\vec{\nabla}\Upsilon} \cdot \vec{\nabla}}u_z -
+                \frac{1}{H}\partial_z u_{z}}
+            -\Gamma(z) u_z - \p*{\vec{u} \cdot \vec{\nabla}}u_z
+            - T_1 \pd{\Upsilon}{z},\\
+        \Gamma(z) &= 7.5\s*{2 + \tanh \frac{z - z_T}{(L_z - z_T) / 2}
+            + \tanh \frac{z_B - z}{z_B / 2}},\label{eq:Gamma}
+    \end{align}
+\end{subequations}
+
+\section{Mean Flow Absorption}
+
+The model we're examining now is if part of the flow has synchronized to the
+horizontal phase velocity of the wave. Then as the wavetrain is incident on the
+wave with horizontal momentum flux $S_{px}$, it completely deposits its
+horizonttal momentum. Thus, $S_{px}$ goes from the wave's, which is well known,
+down to zero, being deposited into the mean flow, which forces
+\begin{equation}
+    \rho(z) c_{ph,x} \pd{z_c}{t} = -S_{px}.
+\end{equation}
+where $c_{ph, x}$ is the horizontal phase velocity/critical velocity, $z_c$ is
+the location of the dissipation front, and $S_{px}$ constant is the incident
+momentum flux. This can be analytically solved if we take $\rho(z) \approx
+\rho_0 \propto e^{-z/H}$. Try ansatz $z(t) = -H\p*{\ln t + \ln \tau}$ and we
+find
+\begin{equation}
+    z(t) = -H\ln t - H\ln \frac{H\rho_0(z = 0)c_{ph, x}}{S_{px}}.
+\end{equation}
+
 \clearpage

### 2019-02-01 sims/2d_3_final/strat_helper.py

@@ -27,7 +27,7 @@ CW = MPI.COMM_WORLD
 
 SNAPSHOTS_DIR = 'snapshots_%s'
 FILENAME_EXPR = '{s}/{s}_s{idx}.h5'
-plot_stride = 4
+plot_stride = 75
 
 def get_omega(g, h, kx, kz):
     return np.sqrt((g / h) * kx**2 / (kx**2 + kz**2 + 0.25 / h**2))
@@ -316,17 +316,16 @@ def plot(name, params):
     plot_cfgs = [
         {
             'save_fmt_str': 'p_%03i.png',
-            'mean_vars': ['S_{px}', 'ux'],
+            'plot_vars': ['uz'],
             'slice_vars': ['uz'],
-            'sub_vars': ['ux'],
             'anal_vars': ['uz'],
+            'f2_vars': ['uz'],
         },
-        # {
-        #     'save_fmt_str': 'm_%03i.png',
-        #     'plot_vars': ['uz'],
-        #     'anal_vars': ['uz'],
-        #     'slice_vars': ['uz'],
-        # },
+        {
+            'save_fmt_str': 'm_%03i.png',
+            'plot_vars': ['uz', 'ux'],
+            'mean_vars': ['S_{px}', 'ux'],
+        },
     ]
 
     dyn_vars = ['uz', 'ux', 'U', 'W']
@@ -568,7 +567,6 @@ def plot(name, params):
             logger.info('Saved %s/%s' % (snapshots_dir, savefig))
             plt.close()
 
-
 def plot_front(name, params):
     ''' few plots for front, defined where flux drops below 1/2 of theory '''
     N_X = params['N_X'] * params['INTERP_X']
@@ -588,7 +586,7 @@ def plot_front(name, params):
     if not os.path.exists(logfile):
         print('log file not found, generating')
         sim_times, domain, state_vars = load(
-            name, params, dyn_vars, plot_stride, start=0)
+            name, params, dyn_vars, plot_stride=1, start=0)
         x = domain.grid(0, scales=params['INTERP_X'])
         z = domain.grid(1, scales=params['INTERP_Z'])
         xmesh, zmesh = quad_mesh(x=x[:, 0], y=z[0])

### 2019-02-02 writeups/paper.tex

@@
         \begin{align}
-            F_{px, z} &\equiv \ev*{\rho u_x u_z}_x =
+            S_{px, z} &\equiv \ev*{\rho u_x u_z}_x =
                 \frac{1}{L_x}\int_0^{L_x}\limits \rho u_x u_z\;\mathrm{d}x
@@
     \item Finally, it may be noted that horizontally averaged horizontal
-        momentum $p_x = \ev*{\rho u_x}_x$ obeys $\pd{p_x}{t} + \pd{F_{px}}{z} =
+        momentum $p_x = \ev*{\rho u_x}_x$ obeys $\pd{p_x}{t} + \pd{S_{px}}{z} =
         0$. Since flux is transported on the group velocity timescale (CITE),
@@
 system is solved exactly by matching the two homogeneous solutions above and
-below $z_0$, We may then approximately relax the solution to $k_{0z}\sigma
+below $z_0$. We may then approximately relax the solution to $k_{0z}\sigma
 \lesssim 1$: an extra factor of $e^{-\frac{(k_{0z}\sigma)^2}{2}}$ arises
@@
 \begin{align}
-    A(F) ={}& \frac{Fgk_{0x}^2}{\rho_0(z_0)\omega_0^2}
+    u_{1z}&(x, z, t) ={} \frac{Fgk_{0x}^2}{\rho_0(z_0)\omega_0^2}
         \frac{1}{2ik_z}\frac{e^{\frac{-(k_{0z}\sigma)^2}{2}}}
-        {\sqrt{2\pi\sigma^2}},\\
-    u_{1z}(x, z, t) ={}& A(F) \times\nonumber\\
+        {\sqrt{2\pi\sigma^2}} \times\nonumber\\
         &{}\begin{cases}
-        e^{\frac{z - z_0}{2H}}e^{i(k_{0x}x + k_{0z}(z - z_0) - \omega_0 t)}
+        e^{\frac{z - z_0}{2H}}e^{i(k_{0x}x + k_{0z}(z - z_0) - \omega_0 t)
+            + \frac{1}{2k_zH}}
             & z > z_0\\
-        e^{\frac{z - z_0}{2H}}e^{i(k_{0x}x - k_{0z}(z - z_0) - \omega_0 t)}
+        e^{\frac{z - z_0}{2H}}e^{i(k_{0x}x - k_{0z}(z - z_0) - \omega_0 t)
+            + \frac{1}{2k_zH}}
             & z < z_0\\
@@
 
-%%%
-These are implemented as:
-\begin{subequations}\label{se:curr_num}
-    \begin{align}
-        \vec{\nabla} \cdot \vec{u} &= 0,\\
-        \pd{\Upsilon}{t} - \frac{u_z}{H}
-            - \nu \nabla^2 \Upsilon &= -\Gamma(z) \Upsilon
-                - \p*{\vec{u} \cdot \vec{\nabla}}\Upsilon
-                + \frac{F}{\rho_0(z)}e^{-\frac{(z - z_0)^2}{2\sigma^2}}
-                    \cos \p*{k_xx - \omega t},\\
-        \pd{u_x}{t} + \pd{T}{x} + gH\pd{\Upsilon}{x}
-            - \nu \nabla^2 u_x
-            &= -\Gamma(z) u_x
-                - \p*{\vec{u} \cdot \vec{\nabla}}u_x
-                - T_1 \pd{\Upsilon}{x},\\
-        \pd{u_z}{t} + \pd{T}{z} + gH\pd{\Upsilon}{z} - \frac{T_1}{H}
-            - \nu \nabla^2 u_z &= -\Gamma(z) u_z
-                - \p*{\vec{u} \cdot \vec{\nabla}}u_z
-                - T_1 \pd{\Upsilon}{z},\\
-    \Gamma(z) &= 7.5\s*{2 + \tanh \frac{z - z_T}{(L_z - z_T) / 2}
-        + \tanh \frac{z_B - z}{z_B / 2}},\label{eq:Gamma}
-    \end{align}
-\end{subequations}
 where $z_B = 0.05L_z, z_T = 0.95L_z$ are the boundaries of the damping zones. A
@@
 
-    \item $F_{px}$ is seen to be constant, as would be expected by
+    \item $S_{px}$ is seen to be constant, as would be expected by
         \autoref{eq:fpx_lin}.
@@
     \caption{Snapshot of linear simulation after reaching steady state.
-        Analytical predictions of $\abs*{u_z}, \bar{U}_0, F_{px}$ are shown in
+        Analytical predictions of $\abs*{u_z}, \bar{U}_0, S_{px}$ are shown in
         orange. Red dotted lines indicate the onset of the damping region and
@@
     simulation mysteriously stopped without me noticing; restarting worked, so
-    currently rerunning). Note the slight resulting dissipation in $F_{px}$.
+    currently rerunning). Note the slight resulting dissipation in $S_{px}$.
     }\label{fig:linear_nu}
@@
     \pd{}{t}\int\limits_0^{L_z} \rho(z) \bar{U}_0(z, t)\;\mathrm{d}z
-        - F_{px} = 0.
+        - S_{px} = 0.
 \end{equation}
@@
 \begin{equation}
-    \rho \bar{U}_c\pd{z_c}{t} = -F_{px}.\label{eq:zc_anal}
+    \rho \bar{U}_c\pd{z_c}{t} = -S_{px}.\label{eq:zc_anal}
+\end{equation}
+For constant $S_{px}$ in space and $\rho \approx \rho_0$, this has analytical
+solution
+\begin{equation}
+    z(t) = -H\ln t - H\ln \frac{H\rho_0(z = 0)c_{ph, x}}{S_{px}}.
 \end{equation}
@@
 \autoref{fig:nl} after the critical layer has had time to form. We may note that
-the critical layer, where $F_{px}$ is absorbed and $\bar{U}_0 = \bar{U}_c$,
+the critical layer, where $S_{px}$ is absorbed and $\bar{U}_0 = \bar{U}_c$,
 travels downwards as predicted.
@@
 in \autoref{fig:zc_anal}. We see overall good agreement, though some small
-deviation is expected since $F_{px}$ is not perfectly conserved owing to
-numerical viscosity (and YUBONOTE $F_{px}$ is misestimated?).
+deviation is expected since $S_{px}$ is not perfectly conserved owing to
+numerical viscosity (and YUBONOTE $S_{px}$ is misestimated?).
 \begin{figure}[h]
@@
 equations as simulated in Dedalus:
+\begin{subequations}
+    \begin{align}
+        \vec{\nabla} \cdot \vec{u} ={}& 0,\\
+        \partial_t \Upsilon - \frac{u_z}{H}
+            ={}& \nu\p*{\nabla^2 \Upsilon + \p*{\vec{\nabla}
+            \Upsilon} \cdot \p*{\vec{\nabla}\Upsilon} - \frac{2}{H}\partial_z
+            \Upsilon + \frac{1 - e^{-\Upsilon}}{H^2}},\nonumber\\
+            & - \p*{\vec{u} \cdot \vec{\nabla}}\Upsilon
+                -\Gamma(z) \Upsilon
+                + \frac{F}{\rho_0(z)}e^{-\frac{(z - z_0)^2}{2\sigma^2}}
+                    \cos \p*{k_xx - \omega t},\\
+        \pd{u_x}{t} + \pd{T}{x} + gH\pd{\Upsilon}{x} ={}&
+            \nu \nabla^2 u_x
+            - u_x \nu\p*{\nabla^2 \Upsilon + \p*{\vec{\nabla} \Upsilon} \cdot
+                \p*{\vec{\nabla}\Upsilon} - \frac{2}{H}\partial_z \Upsilon
+                + \frac{1 - e^{-\Upsilon}}{H^2}}\nonumber\\
+            &+ 2\nu \p*{\p*{\p*{\vec{\nabla}\Upsilon} \cdot \vec{\nabla}}u_x
+                - \frac{1}{H}\partial_z u_x}
+            -\Gamma(z) u_x
+                - \p*{\vec{u} \cdot \vec{\nabla}}u_x
+                - T_1 \pd{\Upsilon}{x},\\
+        \pd{u_z}{t} + \pd{T}{z} + gH\pd{\Upsilon}{z} - \frac{T_1}{H} ={}&
+            \nu \nabla^2 u_z
+            - u_z \nu\p*{\nabla^2 \Upsilon + \p*{\vec{\nabla} \Upsilon} \cdot
+                \p*{\vec{\nabla}\Upsilon} - \frac{2}{H}\partial_z \Upsilon
+                + \frac{1 - e^{-\Upsilon}}{H^2}}\nonumber\\
+            &+ 2\nu \p*{\p*{\p*{\vec{\nabla}\Upsilon} \cdot \vec{\nabla}}u_z -
+                \frac{1}{H}\partial_z u_{z}}
+            -\Gamma(z) u_z - \p*{\vec{u} \cdot \vec{\nabla}}u_z
+            - T_1 \pd{\Upsilon}{z},\\
+        \Gamma(z) &= 7.5\s*{2 + \tanh \frac{z - z_T}{(L_z - z_T) / 2}
+            + \tanh \frac{z_B - z}{z_B / 2}},\label{eq:Gamma}
+    \end{align}
+\end{subequations}
 

### 2019-02-03 writeups/notes.tex
update analytical fit function

@@
 momentum flux. This can be analytically solved if we take $\rho(z) \approx
-\rho_0 \propto e^{-z/H}$. Try ansatz $z(t) = -H\p*{\ln t + \ln \tau}$ and we
+\rho_0 \propto e^{-z/H}$. Try ansatz $z(t) = -H\p*{\ln t / \tau}$ and we
 find
-\begin{equation}
-    z(t) = -H\ln t - H\ln \frac{H\rho_0(z = 0)c_{ph, x}}{S_{px}}.
+\begin{align}
+    \rho_0(z = 0) \frac{t}{\tau} c_{ph, x} \p*{-\frac{H}{t}} &= -S_{px}
+        \nonumber\\
+    \tau &= \frac{H\rho_0(z = 0)c_{ph, x}}{S_{px}},\nonumber\\
+    z(t) &= -H\ln \frac{t + t_0}{\tau}.
+\end{align}
+$t_0$ is some constant of integration that is fit by setting $z_c(t = 0) = z_0$
+some initial condition. Let's first notice that $z(-t_0) = +\infty, z(\tau -
+t_0) = 0$, and so $\tau$ corresponds to the time it takes to spin up the entire
+atmosphere above $z = 0$. Indeed, halving $S_{px}$ doubles the amount of time
+necessary.
+
+In our simulations, it seems like $S_{px}$ decreases over time due to reflective
+behavior setting in (similar to Winters \& d'Asaro, though since they had a very
+narrow critical layer they observed transmission too, impossible in our case).
+We can thus consider how two $z(t, S_{px})$ solutions differ based on choice of
+$S_{px}$ given some fixed initial condition. In our case, since the beginning of
+the simulation is dominated by transients, horizontal momentum dissipation and
+formation of the critical layer, our equation is only really valid towards the
+end of the simulation. Thus, we formulate our problem as follows: given two
+$S_{px}, \alpha S_{px}$, compute $z(t)$ for both given initial condition
+$z(t_f) = z_f$. We can solve and obtain
+\begin{equation}
+    t_0 = \tau e^{-z_f/H} - t_f,
+\end{equation}
+and so the exact solution becomes
+\begin{equation}
+    z(t) = -H\ln \frac{t - t_f + \tau e^{-z_f/H}}{\tau}.
 \end{equation}

### 2019-02-04 writeups/paper.tex
stash paper

@@
     \documentclass[twocolumn,
-        nofootinbib,
+        nofootinbib, % unbreaks footnotes from being put in bib
         usenames, % allows access to some tikz colors
+        aps,
+        prd,
         dvipsnames % more colors: https://en.wikibooks.org/wiki/LaTeX/Colors
@@
 
-In \autoref{s:num_setup}, we will cover the numerical setup. In \autoref{s:lin}, we
+In \autoref{s:equations}, we will cover the numerical setup. In \autoref{s:lin}, we
 will discuss the agreement of our simulation in the low-amplitude limit with
@@
 
-\section{Problem Description}\label{s:num_setup}
-
-\subsection{Fluid Equations}
+\section{Equations}\label{s:equations}
 
@@
 center of the WD, we approximate the gravitational field as uniform. We model
-the background density stratification as $\rho_0(x, z) = \rho_0(z) = \rho_{ref}
-e^{-z/H}$ for some reference density $\rho_{ref} = \at{\rho_0(z)}_{z = 0}$.
+the background density stratification as $\rho_0(x, z) = \rho_0(z) = \rho_0(z=0)
+e^{-z/H}$ for some reference density $\rho_0(z=0) = \at{\rho_0(z)}_{z = 0}$.
 
@@
 
-\subsection{Internal Gravity Waves: Linear Theory}
+\section{Internal Gravity Waves: Linear Theory}
+
+\subsection{Analytical Properties}
 
@@
 we linearize \autoref{se:fc_orig} by ignoring the advective components of the
-material derivative, defining $\rho_1 \equiv \rho - \rho_0 \ll \rho_0$ and $P_1
-\equiv P - P_0$ deviations from equilibrium and obtain:
+material derivative, defining $\rho_1 \equiv \rho - \rho_0 \ll \rho_0 \ll
+\rho_0$ and $P_1 \equiv P - P_0 \ll P_0$ deviations from equilibrium and obtain:
 \begin{subequations}\label{se:lin_homo}
@@
 It is well known that the solutions to the $F = 0$ homogeneous
-\autoref{se:lin_homo} are of form (cite)
+\autoref{se:lin_homo} are of form\cite{drazin,sutherland0}:
 \begin{subequations}\label{se:lin_sol}
@@
 \end{subequations}
-up to undetermined amplitude $A$, where
+up to undetermined amplitude $A$, and
 \begin{equation}
@@
 \end{equation}
-is the Brunt-V\"ais\"al\"a, or buoyancy, frequency.
+is the Brunt-V\"ais\"al\"a/buoyancy frequency for our incompressible,
+exponentially stratified atmosphere.
 
-In the weak stratification limit $k_zH \gg 1$, the solution exhibits the
+In the weak stratification limit $k_{0z}H \gg 1$, the solution exhibits the
 following characteristics:
@@
         \begin{align}
-            \vec{c}_{ph} &= \frac{Nk_x}{k_x^2 + k_z^2 + 1/4H^2} \hat{k},\\
-            \vec{c}_{g} &= \frac{N\p*{k_xk_z\hat{x}
-                - \p*{k_z^2 + \frac{1}{4H^2}}\hat{z}}}
-                {\p*{k_x^2 + k_z^2 + \frac{1}{4H^2}}^{3/2}} \nonumber\\
-                &\approx \frac{Nk_z}{k_x^2 + k_z^2}\p*{k_x\hat{x} - k_z\hat{z}}.
+            \vec{c}_{ph} &= \frac{Nk_{0x}}{k_{0x}^2 + k_{0z}^2 + 1/4H^2} \hat{k},\\
+            \vec{c}_{g} &= \frac{N\p*{k_{0x}k_{0z}\hat{x}
+                - \p*{k_{0z}^2 + \frac{1}{4H^2}}\hat{z}}}
+                {\p*{k_{0x}^2 + k_{0z}^2 + \frac{1}{4H^2}}^{3/2}} \nonumber\\
+                &\approx \frac{Nk_{0z}}{k_{0x}^2 + k_{0z}^2}\p*{k_{0x}\hat{x} -
+                k_{0z}\hat{z}}.
         \end{align}
-        We recover the usual result $\vec{c}_{ph} \cdot \vec{c}_g = 0$ (CITE).
-        Note that an IGW transporting energy and momentum upwards has $k_z < 0$.
-
-%     \item The linear solution is an exact solution to the full nonlinear
-%         equations \autoref{se:fc_orig}, thanks to the advective terms vanishing
-%         identically when $\vec{\nabla} \cdot \vec{u} = 0$ (CITE). However, this
-%         solution is globally unstable to resonant triad interactions (CITE).
+        We recover the usual result $\vec{c}_{ph} \cdot \vec{c}_g = 0$
+        (e.g.\ \cite{drazin,sutherland1}). Note that an IGW transporting energy
+        and momentum upwards $c_{g, z} > 0$ has $k_{0z} < 0$.
 
     \item The time-averaged total $x$-momentum flux in the $\hat{z}$ direction
-        can be computed. Since the fluid variables have separable as $f(x, z, t)
-        = f(z)e^{i(k_{0x}x - \omega_0 t)}$, $x$ averaging and time averaging are
+        can be computed. Since the linear solution is separable as $f(x, z, t) =
+        f(z)e^{i(k_{0x}x - \omega_0 t)}$, $x$ averaging and time averaging are
         equivalent, so we may write ($\ev*{\dots}_x$ denotes $x$-averaging)
         \begin{align}
-            S_{px, z} &\equiv \ev*{\rho u_x u_z}_x =
+            S_{px, z} &\equiv \ev*{\rho u_x u_z}_x \equiv
                 \frac{1}{L_x}\int_0^{L_x}\limits \rho u_x u_z\;\mathrm{d}x
                     \nonumber\\
-                &\approx -\frac{A^2}{2}\rho_{ref}\frac{k_z}{k_x}.
+                &\approx -\frac{A^2}{2}\rho_0(z=0)\frac{k_{0z}}{k_{0x}}.
                     \label{eq:fpx_lin}
@@
         momentum $p_x = \ev*{\rho u_x}_x$ obeys $\pd{p_x}{t} + \pd{S_{px}}{z} =
-        0$. Since flux is transported on the group velocity timescale (CITE),
-        this simplifies to the known expression for the wave-induced horizontal
-        mean flow (cite)
+        0$. Since flux is transported on the group velocity
+        timescale\cite{sutherland0}, this simplifies to the known expression for
+        the wave-induced horizontal mean flow\cite{sutherland0,sutherland1}
         \begin{equation}
@@
         We've denoted $c_{gz} = \at{\pd{\omega}{k_{z}}}_{k_{z0}}$ the vertical group
-        velocity. In terms of our forcing amplitude,
+        velocity. This may also be rewritten
         \begin{equation}
-            \bar{U}_0(F) = -\frac{A(F)^2}{2}e^{\frac{z - z_0}{H}} \frac{k_z}{k_x}c_{gz},
-                \label{eq:u0_lin}
+            \bar{U}_0 = -\frac{A^2}{2}e^{\frac{z - z_0}{H}}
+                \frac{k_{0z}}{k_{0x}}c_{gz}, \label{eq:u0_lin}
         \end{equation}
@@
 \end{itemize}
-% Linear solution: e^z/2H, dispersion relation, phase/group velocity. Observe is
-% exact solution of nonlinear.
-% Linear solution is globally unstable to RTI though, thus not stable
-% equilibrium of IVP
-%
-% Nonlinear: Critical layer absorption (viscous = Hazel, nonlin = Winters),
-% wave-induced mean flow, horizontal momentum flux
 
@@
     \md{\rho}{t} = Fe^{-\frac{(z - z_0)^2}{2\sigma^2}}
-            \cos \p*{k_{0x}x - \omega_0 t}.
+            \cos \p*{k_{0x}x - \omega_0 t}.\label{eq:vol_drive}
 \end{equation}
@@
 only the $k_{0z}$ satisfying the IGW's dispersion relation for the given
-$k_{0x}, \omega_0(k_{0x}, k_{0z})$ will propagate. The fluid equations then take
-form
-\begin{subequations}\label{se:vol_drive}
-    \begin{align}
-        \vec{\nabla} \cdot \vec{u} &= 0,\\
-        \md{\rho}{t} &= Fe^{-\frac{(z - z_0)^2}{2\sigma^2}}
-                e^{i(k_xx - \omega t)},\\
-        \md{\vec{u}}{t} + \frac{\vec{\nabla}P}{\rho}
-            + g\hat{z} &= 0.
-    \end{align}
-\end{subequations}
+$k_{0x}, \omega_0(k_{0x}, k_{0z})$ will propagate.
 
-The linearized \autoref{se:vol_drive}, obtained by dropping the advective term
+The linearized \autoref{eq:vol_drive}, obtained by dropping the advective term
 in the material derivative, can be solved analytically up to good accuracy. We
@@
     u_{1z}&(x, z, t) ={} \frac{Fgk_{0x}^2}{\rho_0(z_0)\omega_0^2}
-        \frac{1}{2ik_z}\frac{e^{\frac{-(k_{0z}\sigma)^2}{2}}}
+        \frac{1}{2ik_{0z}}\frac{e^{\frac{-(k_{0z}\sigma)^2}{2}}}
         {\sqrt{2\pi\sigma^2}} \times\nonumber\\
@@
         e^{\frac{z - z_0}{2H}}e^{i(k_{0x}x + k_{0z}(z - z_0) - \omega_0 t)
-            + \frac{1}{2k_zH}}
+            + \frac{1}{2k_{0z}H}}
             & z > z_0\\
         e^{\frac{z - z_0}{2H}}e^{i(k_{0x}x - k_{0z}(z - z_0) - \omega_0 t)
-            + \frac{1}{2k_zH}}
+            + \frac{1}{2k_{0z}H}}
             & z < z_0\\
@@
 \end{align}
+The $z > z_0$ solution models an upwards-propagating IGW wavetrain inbound on
+the simulation domain from below. The solution for $u_{1x}(x, z, t)$ moreover is
+just $u_{1x} \approx \frac{k_{0z} u_{1z}}{k_{0x}}e^{\frac{1}{2k_{0z}H}}$.
+
+\section{Numerical Simulation}
 
@@
 
-We nondimensionalize by taking $H = N = \rho_{ref} = 1$
+We nondimensionalize by taking $H = N = \rho_0(z=0) = 1$ in
+\autoref{se:fc_orig}.
 
@@
 regularize the nonlinear cascade at near grid resolution: $\nu \sim 0.2
-\frac{\omega}{\abs*{k_z}}\frac{L_z}{2\pi N_z}$ was found to be suitable for $N_z
-= 1024$, where $k_z$ is the wavenumber of the excited linear mode.
+\frac{\omega}{\abs*{k_{0z}}}\frac{L_z}{2\pi N_z}$ was found to be suitable for
+$N_z = 1024$, where $k_{0z}$ is the wavenumber of the excited linear mode.
 %%%
@@
 
-$\omega$ is chosen by inverting $\omega(k_x, k_z)$ dispersion relation for fixed
-$k_x = 2\pi / L_x$ and some desired $k_z = -2\pi/H$, and $\sigma \lesssim
-\frac{1}{k_z}$ is used to excite a broad band of modes including the desired
-$k_z$ mode.
+$\omega$ is chosen by inverting $\omega(k_{0x}, k_{0z})$ dispersion relation for fixed
+$k_{0x} = 2\pi / L_x$ and some desired $k_{0z} = -2\pi/H$, and $\sigma \lesssim
+\frac{1}{k_{0z}}$ is used to excite a broad band of $k_z$ modes including the
+desired $k_{0z}$ mode.
 
@@
 
-We simulate using Dedalus (cite) a spectral numerical method. We use a Fourier
-basis in the $x$ direction and a Chebyshev in the $z$ direction. Our simulation
-starts with all perturbation quantities at zero, an initial condition at rest.
-Our parameters are chosen to be as similar as possible to the later nonlinear
-simulations (\autoref{s:nonlin}). We choose as follows:
+% Nonlinear: Critical layer absorption (viscous = Hazel, nonlin = Winters),
+% wave-induced mean flow, horizontal momentum flux
+
+We simulate using Dedalus\cite{dedalus} a spectral numerical method. We use a
+Fourier basis in the $x$ direction and a Chebyshev in the $z$ direction. Our
+simulation starts with all perturbation quantities at zero, an initial condition
+at rest. Our parameters are chosen to be as similar as possible to the later
+nonlinear simulations (\autoref{s:nonlin}). We choose as follows:
 \begin{itemize}
-    \item $k_{0x}$: Astrophysical IGWs have $k_x \ll k_z$, so we choose $k_{0x}
+    \item $k_{0x}$: Astrophysical IGWs have $k_{0x} \ll k_{0z}$, so we choose $k_{0x}
         = \frac{2\pi}{L_x}$ the smallest permitted wavenumber permitted by
@@
     \item We choose $\omega_0$ by choosing it to produce a desired $k_{0z}$.
-        Astrophysical systems generally also exhibit $k_zH \gg 1$, where the
+        Astrophysical systems generally also exhibit $k_{0z}H \gg 1$, where the
         stratification height is significantly larger than the vertical
@@
 using the $\nu$ from the later nonlinear simulations in \autoref{fig:linear_nu}.
-\begin{figure}[h]
+\begin{figure}[t]
     \centering
@@
 \end{figure}
-\begin{figure}[h]
+\begin{figure}[t]
     \centering
@@
 
-\section{Nonlinear Regime}\label{s:nonlin}
+\subsection{Nonlinear Regime}\label{s:nonlin}
 
@@
 travels downwards as predicted.
-\begin{figure}[h]
+\begin{figure}[t]
     \centering
@@
 numerical viscosity (and YUBONOTE $S_{px}$ is misestimated?).
-\begin{figure}[h]
+\begin{figure}[t]
     \centering
@@
 transition.
-\begin{figure}[h]
+\begin{figure}[t]
     \centering
@@
 where the Richardson number does not go nearly as low.
-\begin{figure}[h]
+\begin{figure}[t]
     \centering
@@
 \end{figure}
-
-Additionally, a lower-viscosity simulation shows significant features resembling
-Kelvin-Helmholtz instabilities (KHI), \autoref{fig:nl_full_01}. As we have only
-plotted $\bar{U}_0(z)$, the local Richardson number $N^2 / \pd{u_x}{z}$ may
-exceed the plotted value and incur KHIs.
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.5\textwidth]{plots/nl_full_01.png}
-    \caption{Low viscosity simulation showing seemingly KHI-like
-    features.}\label{fig:nl_full_01}
-\end{figure}
-
-This seems to suggest that our reference simulation is viscosity limited. This
-is perhaps physical: a Richardson number $\sim 0.25$ incurs the KHI which saps
-energy and angular momentum from the flow. It is conceivable then that our
-numerical viscosity is physically motivated as a effective KHI viscosity.
 
@@
 the number of spectral modes in the respective dimensions. We perform direct
-numerical simulation of \autoref{se:vol_drive} with the open-source
+numerical simulation of \autoref{eq:vol_drive} with the open-source
 pseudo-spectral code Dedalus (CITE).
@@
 
+\bibliographystyle{apsrev4-1} % chktex 8
+\bibliography{paper}
 \end{document}

### 2019-02-05 presentations/meeting.tex
2/5/19 meeting?

@@
 \begin{frame}
-    \frametitle{Linear}
+    \frametitle{Review}
 
+    \begin{columns}
+        \begin{column}{0.5\textwidth}
+            \begin{itemize}
+                \item Linear: $\nu = 0$ has $S_{px}(z) = S_{px, 0}$ great
+                    prediction. Fit $\vec{u}$ for $\nu \neq 0$ as well, though
+                    not $S_{px}(z)$ yet (\textbf{Updated!}).
+
+                \item In nonlinear, review: expect deposited flux $\Delta S_{px}
+                    = S_{px, 0}$ incident/generated flux.
+
+                \item Instead, observe $\Delta S_{px} < S_{px, 0}$, and moreover
+                    $S_{px}$ far from critical layer changing over time!
+            \end{itemize}
+        \end{column}
+        \begin{column}{0.5\textwidth}
+            \begin{itemize}
+                \item Last week: defined some
+                    \begin{equation}
+                        \delta u_z = \frac{u_z - u_{z, 0}
+                            - \ev*{u_z}_x}{\abs*{u_{z, 0}}},
+                    \end{equation}
+                    deviation (new: subtracting mean flow too), computed RMS\@.
+                    Found higher RMS for lower $\nu$.
+
+                \item Could it be $\ev*{\dots}_x$ vs $\ev*{\dots}_t$?
+
+                \item Objective: understand ``reflection''?
+            \end{itemize}
+        \end{column}
+    \end{columns}
+\end{frame}
+
+\begin{frame}
+    \frametitle{Overview}
+
+    \begin{itemize}
+        \item Understand exactly how much flux is being absorbed at the critical
+            layer:
+            \begin{itemize}
+                \item Can the decreased flux absorption be explained by
+                    viscous dissipation before the critical alyer alone?
+
+                \item Compare predictions of $\pd{z_c}{t}$ to the observed
+                    $\Delta S_{px}$ at the critical layer.
+            \end{itemize}
+
+        \item Understand nature of ``reflection'', why cannot see in certain
+            simulations.
+    \end{itemize}
+\end{frame}
+
+\begin{frame}
+    \frametitle{Results}
+    \framesubtitle{Linear}
+
+    Had great $u_z$ fits including viscosity (error $\sim 3-5\%$), but did not
+    show great $S_{px}(z; \nu)$ fits last week. Here they are!
     \begin{figure}[t]
         \centering
-        \includegraphics[width=0.7\textwidth]{../sims/2d_3_final/snapshots_lin_0/fluxes.png}
-        \caption{Linear simulation.}
+        \includegraphics[width=0.7\textwidth]{../sims/2d_3_final/snapshots_lin_1/fluxes.png}
+        \caption{Fits linear $S_{px}(z; \nu)$.}
     \end{figure}
@@
 \begin{frame}
-    \frametitle{New Plotting for Nonlinear}
+    \frametitle{Results}
+    \framesubtitle{Mean Flow and Flux}
 
+    \begin{columns}
+        \begin{column}{0.5\textwidth}
             \begin{figure}[t]
                 \centering
-        \includegraphics[width=0.7\textwidth]{../sims/2d_3_final/snapshots_nl_3/fluxes.png}
-        \caption{Nonlinear simulation, new plotting methodology. Feedback?}
+                \includegraphics[width=\textwidth]{../sims/2d_3_final/snapshots_nl_4/fluxes.png}
             \end{figure}
+        \end{column}
+        \begin{column}{0.5\textwidth}
+            \begin{itemize}
+                \item Last week: Evolution of $\ev*{u_x}_x(t),
+                    \ev*{S_{px}}_x(t)$ at different times. Note flux seems to
+                    decrease over time throughout domain.
+
+                \item Bottom panel, solid lines are time slices, dotted lines
+                    indicate averaging over $\sim \frac{2\pi}{\omega}$ one
+                    period (6 times), thick line is linear viscous prediction.
+
+                \item Shaded grey indicates one vertical wavelength.
+            \end{itemize}
+        \end{column}
+    \end{columns}
 \end{frame}
@@
 \begin{frame}
-    \frametitle{Reflection?}
+    \frametitle{Results}
+    \framesubtitle{Mean Flow and Flux}
 
-    Get exact solution, up to viscous dissipation. Plot $\delta u_z =
-    \frac{u_{z,sim} - u_{z, anal}}{\abs*{u_{z, anal}}}$, ``fractional
-    deviation'' between driving zone and critical layer. Compute $\mathrm{RMS} =
-    \sqrt{\ev*{\delta u_z^2}}$.
+    \begin{columns}
+        \begin{column}{0.5\textwidth}
             \begin{figure}[t]
                 \centering
-        \includegraphics[width=0.7\textwidth]{../sims/2d_3_final/snapshots_lin_1/p_074.png}
-        \caption{Linear simulation with same viscosity as nonlinear. Green is
-        viscosity-free envelope, orange analytical.}
+                \includegraphics[width=\textwidth]{../sims/2d_3_final/snapshots_nl_1/fluxes.png}
             \end{figure}
+        \end{column}
+        \begin{column}{0.5\textwidth}
+            \begin{itemize}
+                \item More damped for comparison ($\nu \to 7\nu/3$).
+
+                \item Note that $S_{px}$ does not change much over time, general
+                    trend matches viscous linear decrease.
+            \end{itemize}
+        \end{column}
+    \end{columns}
 \end{frame}
@@
 \begin{frame}
-    \frametitle{Reflection (nl4)}
+    \frametitle{Results}
+    \framesubtitle{Absorbed Flux and Front Propagation}
 
-    Get exact solution, up to viscous dissipation. Plot $\delta u_z =
-    \frac{u_{z,sim} - u_{z, anal}}{\abs*{u_{z, anal}}}$, ``fractional
-    deviation'' between driving zone and critical layer. Compute $\mathrm{RMS} =
-    \sqrt{\ev*{\delta u_z^2}}$.
+    \begin{columns}
+        \begin{column}{0.5\textwidth}
             \begin{figure}[t]
                 \centering
-        \includegraphics[width=0.7\textwidth]{../sims/2d_3_final/snapshots_nl_4/p_074.png}
-        \caption{Nonlinear simulation. Green is viscosity-free envelope, orange
-        analytical.}
+                \includegraphics[width=\textwidth]{../sims/2d_3_final/snapshots_nl_4/front.png}
             \end{figure}
+            \begin{itemize}
+                \item Three different $\Delta S_{px}$:
+                    \begin{itemize}
+                        \item LinVisc extrapolate $S_{px, 0}$.
+
+                        \item $\Delta S_{px; S}$ ($z$-avg).
+
+                        \item $\Delta S_{px; U}$ ($z$-avg too).
+                    \end{itemize}
+            \end{itemize}
+        \end{column}
+        \begin{column}{0.5\textwidth}
+            \begin{itemize}
+
+                \item \textbf{FIXED} $z_c(t)$ for different incident fluxes,
+                    even more models:
+                    \begin{itemize}
+                        \item Solid lines are three ways of extracting from
+                            data.
+
+                        \item Dotted lines use $\pd{z_c}{t}(\Delta S_{px})$
+                            model for three values of $S_{px}$:
+                            \begin{itemize}
+                                \item Average $\Delta S_{px}$.
+                                \item $\nu$-extrapolated $S_{px, 0}$.
+                                \item Full $S_{px, 0}$.
+                            \end{itemize}
+                    \end{itemize}
+                    Takeaway: strongly inconsistent with $S_{px, 0}$ being
+                    completly absorbed! In fact, most are $\sim 0.5S_{px}$ being
+                    absorbed.
+            \end{itemize}
+        \end{column}
+    \end{columns}
 \end{frame}
@@
 \begin{frame}
-    \frametitle{Reflection?}
+    \frametitle{Results}
+    \framesubtitle{Absorbed Flux and Front Propagation}
 
-    \begin{table}[t]
+    \begin{columns}
+        \begin{column}{0.5\textwidth}
+            \begin{figure}[t]
                 \centering
-        \begin{tabular}{l|l l }
-            Name & RMS & Notes\\
-            lin0 & 0.077 & Almost-no visc\\
-            lin1 & 0.067 & Same visc as nl4\\
-            nl1 (low) & 0.132 & half-res nl1, higher visc\\
-            nl1 & 0.120 & $0.7\nu$\\
-            nl2 & 0.213 & $0.4\nu$\\
-            nl3 & 0.179 & $0.4\nu$ (double A)\\
-            nl4 & 0.269 & $0.3\nu$\\
-            nl5 & 0.207 & $0.3\nu$ (double $k_z$)\\
-            nl6 & 0.246 & $0.2\nu$\\
-            nl7 & 0.358 & $0.1\nu$
-        \end{tabular}
-        \caption{Table of RMS values.}
-    \end{table}
+                \includegraphics[width=\textwidth]{../sims/2d_3_final/snapshots_nl_1/front.png}
+            \end{figure}
+        \end{column}
+        \begin{column}{0.5\textwidth}
+            \begin{itemize}
+                \item Overdamped for comparison.
+
+                \item Absorbed flux is in good agreement w/ linear viscous
+                    extrapolation of $S_{px, 0}$, seems to imply full
+                    absorption.
+            \end{itemize}
+        \end{column}
+    \end{columns}
 \end{frame}
@@
 \begin{frame}
-    \frametitle{Reflection (nl6)}
+    \frametitle{Results}
+    \framesubtitle{Residual Flow Properties}
 
-    Get exact solution, up to viscous dissipation. Plot $\delta u_z =
-    \frac{u_{z,sim} - u_{z, anal}}{\abs*{u_{z, anal}}}$, ``fractional
-    deviation'' between driving zone and critical layer. Compute $\mathrm{RMS} =
-    \sqrt{\ev*{\delta u_z^2}}$.
+    \begin{columns}
+        \begin{column}{0.5\textwidth}
             \begin{figure}[t]
                 \centering
-        \includegraphics[width=0.7\textwidth]{../sims/2d_3_final/snapshots_nl_6/p_074.png}
-        \caption{Nonlinear simulation. Green is viscosity-free envelope, orange
-        analytical.}
+                \includegraphics[width=\textwidth]{../sims/2d_3_final/snapshots_nl_4/fft.png}
             \end{figure}
+        \end{column}
+        \begin{column}{0.5\textwidth}
+            \begin{itemize}
+                \item FT of $\delta \vec{u}$ slightly below the critical layer.
+                    Red line denotes where $2\nu k_x^2 = \omega$, ``viscous
+                    wavenumber''?
+
+                \item Note that we've chosen $\nu k_x^2 \lesssim u_xk_x \sim
+                    \frac{\omega}{k_{0x}}k_x$, our previous viscous wavenumber
+                    already.
+            \end{itemize}
+        \end{column}
+    \end{columns}
 \end{frame}
@@
 \begin{frame}
-    \frametitle{Reflection (nl7)}
+    \frametitle{Results}
+    \framesubtitle{Residual Flow Properties}
 
-    Get exact solution, up to viscous dissipation. Plot $\delta u_z =
-    \frac{u_{z,sim} - u_{z, anal}}{\abs*{u_{z, anal}}}$, ``fractional
-    deviation'' between driving zone and critical layer. Compute $\mathrm{RMS} =
-    \sqrt{\ev*{\delta u_z^2}}$.
+    \begin{columns}
+        \begin{column}{0.5\textwidth}
             \begin{figure}[t]
                 \centering
-        \includegraphics[width=0.7\textwidth]{../sims/2d_3_final/snapshots_nl_7/p_074.png}
-        \caption{Nonlinear simulation. Green is viscosity-free envelope, orange
-        analytical.}
+                \includegraphics[width=\textwidth]{../sims/2d_3_final/snapshots_nl_1/fft.png}
+                \caption{Overdamped for comparison, not power law!}
             \end{figure}
+        \end{column}
+        \begin{column}{0.5\textwidth}
+            \begin{figure}[t]
+                \centering
+                \includegraphics[width=\textwidth]{../sims/2d_3_final/snapshots_lin_2/fft.png}
+                \caption{Linear for comparison; taken just below damping zone.}
+            \end{figure}
+        \end{column}
+    \end{columns}
 \end{frame}
@@
 \begin{frame}
-    \frametitle{Reflection (nl3)}
+    \frametitle{Useful Plots}
+    \framesubtitle{Residuals}
 
-    Get exact solution, up to viscous dissipation. Plot $\delta u_z =
-    \frac{u_{z,sim} - u_{z, anal}}{\abs*{u_{z, anal}}}$, ``fractional
-    deviation'' between driving zone and critical layer. Compute $\mathrm{RMS} =
-    \sqrt{\ev*{\delta u_z^2}}$.
+    \begin{columns}
+        \begin{column}{0.5\textwidth}
             \begin{figure}[t]
                 \centering
-        \includegraphics[width=0.7\textwidth]{../sims/2d_3_final/snapshots_nl_3/p_058.png}
-        \caption{Nonlinear simulation. Green is viscosity-free envelope, orange
-        analytical.}
+                \includegraphics[width=\textwidth]{../sims/2d_3_final/snapshots_nl_4/p_009.png}
+                \caption{Residuals for not overdamped. }
             \end{figure}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Reflection (nl5)}
-
-    Get exact solution, up to viscous dissipation. Plot $\delta u_z =
-    \frac{u_{z,sim} - u_{z, anal}}{\abs*{u_{z, anal}}}$, ``fractional
-    deviation'' between driving zone and critical layer. Compute $\mathrm{RMS} =
-    \sqrt{\ev*{\delta u_z^2}}$.
+        \end{column}
+        \begin{column}{0.5\textwidth}
             \begin{figure}[t]
                 \centering
-        \includegraphics[width=0.7\textwidth]{../sims/2d_3_final/snapshots_nl_5/p_074.png}
-        \caption{Nonlinear simulation. Green is viscosity-free envelope, orange
-        analytical.}
+                \includegraphics[width=\textwidth]{../sims/2d_3_final/snapshots_nl_1/p_009.png}
+                \caption{Residuals for overdamped.}
             \end{figure}
+        \end{column}
+    \end{columns}
 \end{frame}

### 2019-02-05 writeups/notes.tex
2/5/19 meeting?

@@
 
+\section{Estimating Various Other Scales}
+
+\subsection{Viscous Damping Length}
+
+This is a rather hacky derivation. Consider that in the fluid equations, if we
+assume $\frac{1}{k_zH} \ll 1$, then viscosity basically just enters as a bunch
+of $\nu k^2$ terms. We can absorb this into the frequency $\partial_t + \nu k^2
+\sim i\omega_0\p*{1 - \frac{i\nu k^2}{\omega_0}}$. Plugging into the dispersion
+relation then shows that ($k_0 = \abs*{\vec{k}_0} = \sqrt{k_{x0}^2 + k_{z0}^2}$)
+\begin{align*}
+    \omega^2 &\approx \frac{N^2k_{x0}^2}{k^2}\p*{1 - \frac{2i\nu k^2}{\omega_0}},\\
+        &\approx \frac{N^2k_{x0}^2}{k_{x0}^2 k_{z0}^2\p*{1 + \frac{2i\nu k^4}{\omega
+            k_{z0}^2}}},\\
+        &\approx \frac{N^2k_{x0}^2}{k_{x0}^2 + \s*{k_{0z}
+            \p*{1 + \frac{i\nu k_0^4}{\omega k_{0z}^2}}}},\\
+        &\approx \frac{N^2k_{x0}^2}{k_{x0}^2 + \s*{k_{0z}
+            \p*{1 + \frac{i\nu k_0^5}{Nk_{x0} k_{0z}^2}}}},\\
+    e^{ik_zz} &\approx \exp\s*{ik_{z0}z - \frac{\nu k_0^5}{Nk_{x0}k_{z0}}z}.
+\end{align*}
+Thus, in the small dissipation limit we can treat the artificial viscosity the
+same as a vertical damping length.
+
 \clearpage

### 2019-02-05 sims/2d_3_final/strat_helper.py

@@ -935,6 +935,9 @@ def plot_front(name, params):
         z_avg = np.mean(uz_ffts[idx - half_avg:idx + half_avg, :], 0)
         ax1.loglog(kx, x_avg, label='t=%.1f' % sim_times[idx], linewidth=0.7)
         ax2.loglog(kx, z_avg, label='t=%.1f' % sim_times[idx], linewidth=0.7)
+    visc_kx = np.sqrt(OMEGA / (2 * params['NU']))
+    ax1.axvline(x=visc_kx, linewidth=1.5, color='red')
+    ax2.axvline(x=visc_kx, linewidth=1.5, color='red')
     ax1.set_ylabel(r'$\tilde{u}_x(k_x)$')
     ax2.set_ylabel(r'$\tilde{u}_z(k_x)$')
     ax2.set_xlabel(r'$k_x$')

### 2019-02-05 sims/2d_3_final/strat_helper.py
slight cleanup

@@ -883,13 +883,13 @@ def plot_front(name, params):
                  '%s-' % PLT_COLORS[color_idx],
                  label='Data (U)',
                  linewidth=0.7)
-        mean_pos = np.max(front_pos_S[len(front_pos_S) // 3: ])
         color_idx += 1
 
         # three multipliers are (i) average incident flux, (ii) estimated
         # incident flux extrapolated from nu and (iii) full flux
         mean_incident = np.mean(-dSpx_S[len(dSpx_S) // 3: ])
         est_generated_flux = -S_px0[len(S_px0) // 5]
+        mean_pos = np.max(front_pos_S[len(front_pos_S) // 3: ])
         est_incident_flux = est_generated_flux *\
             np.exp(-k_damp * 2 * (mean_pos - params['Z0']))
         flux_mults = [mean_incident / flux_th,

### 2019-02-12 proposal/FINESST/Su_FINESST_2019.tex
add convolve for amps

@@
 
-\def\Snospace~{\S{}} % hack to remove the space left after autorefs
-\renewcommand*{\sectionautorefname}{\Snospace}
-\renewcommand*{\appendixautorefname}{\Snospace}
-\renewcommand*{\figureautorefname}{Fig.}
-\renewcommand*{\equationautorefname}{Eq.}
-\renewcommand*{\tableautorefname}{Tab.}
+% \def\Snospace~{\S{}} % hack to remove the space left after autorefs
+\renewcommand*{\sectionautorefname}{Section}
+% \renewcommand*{\appendixautorefname}{\Snospace}
+% \renewcommand*{\figureautorefname}{Fig.}
+% \renewcommand*{\equationautorefname}{Eq.}
+% \renewcommand*{\tableautorefname}{Tab.}
 
@@
 Interferometric Space Antenna (LISA)\cite{lisa}. They are also thought to
-produce interesting optical transients such as tidal novae\cite{tidal_novae},
-underluminous supernovae\cite{underlum}, and Ca-rich fast
-transients\cite{carich}. Most importantly, they have been proposed as the likely
+produce interesting optical transients such as underluminous
+supernovae\cite{underlum}, Ca-rich fast transients\cite{carich}, and tidal
+novae\cite{tidal_novae}. Most importantly, they have been proposed as the likely
 progenitors of type Ia supernovae (e.g.~\cite{Ia0,webbink} or more
@@
 are known\cite{lsst_wd}, \emph{Gaia} (currently gathering data) is expected to
-expand the catalog to a few hundred\cite{lsst_wd} (results based on
+expand the catalog to a few hundreds\cite{lsst_wd} (results based on
 \emph{Gaia}'s second data release have already begun to
@@
 more\cite{lsst_wd}. These observations will significantly advance the
-understanding of WD binaries. My proposed theoretical and computational research
-is well-timed to take advantage of these new advances.
+understanding of WD binaries and their evolution. My proposed theoretical and
+computational research is well-timed to take advantage of these new advances.
 
 In spite of the broad importance of WD binaries, the evolution of these systems
-prior to their final merger is not well understood. Much of this uncertainty
+prior to their final mergers is not well understood. Much of this uncertainty
 comes from our imprecise understanding of tidal interactions, which play an
-important role during a compact WD binary's inspiral\cite{fullerII}. Studies
-show that these interactions manifest as tidal excitation of internal gravity
-waves (IGW), waves in the WD fluid restored by the buoyancy force due to density
-stratification\cite{fullerI}. As these waves propagate outwards towards the
-surface, they grow in amplitude until they break, as do ocean waves on a shore,
-and transfer both energy and angular momentum from the binary orbit to the outer
-envelope of the WD\cite{fullerI,fullerII}.
+important role during a compact WD binary's inspiral\cite{fullerII}. Previous
+studies have shown that these interactions manifest as tidal excitation of
+internal gravity waves (IGW), waves in the WD fluid restored by the buoyancy
+force due to density stratification\cite{fullerI}. As these waves propagate
+outwards towards the WD surface, they grow in amplitude until they break, as do
+ocean waves on a shore, and transfer both energy and angular momentum from the
+binary orbit to the outer envelope of the WD\cite{fullerI,fullerII}.
 
-Previous works have found that this tidal dissipation mechanism can generate
-significantly more energy than thermal radiation from the WD surface alone and
-is thus a major contributor to the WD energy budget\cite{fullerII,fullerIV}.
-However, these works parameterized the wave breaking process without
-hydrodynamical simulations. The details of tidal dissipation, namely the
-location and spatial extent of the wave breaking, affect the observable outcome:
-dissipation near the surface of the WD can be efficiently radiated away and
-simply brightens the WD, while dissipation deep in the WD causes an energy
-buildup that results in energetic flares\cite{tidal_novae}. Similar works in
-other fields find that numerical simulations capturing the strongly nonlinear
-wave breaking process find new behavior that cannot be described by linear and
-weakly nonlinear theory\cite{winters1994,barker_ogilvie}. Such fully nonlinear
-numerical simulations have not been performed for WDs.
+Previous works have found that the dissipation of IGW can generate significantly
+more energy than thermal radiation from the isolated WD surface and is thus a
+major contributor to the WD energy budget\cite{fullerII,fullerIV}. However,
+these works parameterized the wave breaking process in an ad hoc manner. The
+details of dissipation, namely the location and spatial extent of the wave
+breaking, affect the observable outcome: dissipation near the surface of the WD
+can be efficiently radiated away and simply brightens the WD, while dissipation
+deep in the WD envelope causes an energy buildup that results in energetic
+flares\cite{tidal_novae}. Works in other fields based on numerical simulations
+show that strongly nonlinear wave breaking process exhibits new behavior that
+cannot be described by linear and weakly nonlinear
+theory\cite{winters1994,barker_ogilvie}. Such fully nonlinear numerical
+simulations have not been performed for WDs.
 
-\subsection{Proposed Research}
+\subsection{Goals of Proposed Research}
 
@@
 binaries will require numerical simulation to capture the turbulent cascade to
-small scales causing wave breaking. \textbf{I propose to study the dynamical
-effects of tidal dissipation via nonlinear IGW breaking in binary WDs.} To
-accomplish this, I aim to address the following three aims:
+small scales that causes wave breaking. \textbf{I propose to study the dynamical
+effects of tidal dissipation via nonlinear Internal Gravity Wave (IGW) breaking
+in binary WDs.} There are three specific goals of my proposed research:
 \begin{itemize}
@@
         wave breaking in realistic WD models via direct numerical simulation.
-        The location and spatial extent will furnish a simple yet effective
-        parameterization of tidal dissipation. \autoref{s:2} describes the
-        steps required for this aim.
+        The location and spatial extent of wave breaking will furnish a simple
+        yet effective parameterization of tidal dissipation in a range of WD
+        models. \autoref{s:2} describes the steps required to acheive this goal.
 
     \item Predict signatures of tidal dissipation over a wide range of possible
-        WD systems. In particular, I will describe the impact of tidal heating
-        on the apparent temperature of a binary WD and maybe even the production
-        of observable flares. \autoref{s:3} details my plan to do so.
+        WD systems. In particular, I will study the impact of tidal heating
+        on the luminosity of WDs in binaries and explore the possibility of
+        producing of observable flares. \autoref{s:3} details my plan.
 
@@
 The current understanding of tidal synchronization in WD binaries is laid out
-in~\cite{fullerII}: a tidally-excited train of IGW undergoes wave breaking in
+in~\cite{fullerII}: tidal forces from the companion excites IGWs in the deep
+envelope of the WDs. These IGW propagate outwards and undergo wave breaking in
 the outer envelope of the WD, locally depositing angular momentum and
-synchronizing the WD spin to the binary orbit. A similar process may also
-operate in stellar binaries consisting of early type stars\cite{zahn75,gn89},
+synchronizing the WD spin to the binary orbit. A similar process also
+operates in binaries consisting of early type stars\cite{zahn75,gn89},
 the only major difference being in the specifics of wave
-excitation\footnote{While IGWs in stars are excited at their
-radiative-convective boundaries, IGWs in WDs are excited at sharp changes in
-composition.}. Nevertheless, direct numerical simulation of the wave breaking
-process has not been performed in either of these systems. Since wave breaking
-is a strongly nonlinear phenomenon, where a larger wave breaks down into many
-smaller-scale waves, numerical simulation is paramount to an accurate
-understanding of IGW breaking.
+excitation\footnote{While IGWs in massive stars are excited at their
+radiative-convective boundaries, the excitation of IGWs in WDs is more gradual
+and is associated with sharp composition changes in the stellar
+envelope\cite{fullerII}.}. Nevertheless, direct numerical simulation of the wave
+breaking process has not been performed in either of these systems. Since wave
+breaking is a strongly nonlinear phenomenon, where a larger wave breaks down
+into many smaller-scale waves, numerical simulation is paramount to an accurate
+understanding of the tidal dissipation process.
 
-IGW breaking has been studied in detail in atmospheric sciences. The wave
-breaking process proceeds as follows: initially, as the IGW reaches nonlinear
+IGW breaking has been studied in atmospheric sciences. The wave
+breaking process proceeds as follows: Initially, as the IGW reaches nonlinear
 amplitudes, it breaks down via the parametric subharmonic instability and
 transfers energy and angular momentum from the wave to the mean flow of the
-fluid\cite{drazin}. Subsequently, after the mean flow velocity reaches the
-horizontal phase velocity of the IGW, a critical layer through which the IGW
-cannot propagate forms. A well-known calculation shows that the IGW is nearly
-completely absorbed at this critical layer in the linear approximation and
-endows the atmosphere with a mean horizontal flow\cite{booker_bretherton,hazel}.
-However, when this mean flow absorption was numerically studied including full
-nonlinear interactions, new phenomena not described by the linear theory
-(reflection off the critical layer and sharpening of the mean flow) were
-observed\cite{jones_num,winters1994} that affected the evolution of the
-atmosphere over time. \emph{This highlights the importance of numerical
-simulation in capturing the wave breaking process.}
+fluid\cite{drazin}; after the mean flow velocity reaches the
+horizontal phase velocity of the IGW, a critical layer forms. Analytical
+calculations show that the IGW is nearly completely absorbed at this critical
+layer in the linear approximation and endows the atmosphere with a mean
+horizontal flow\cite{booker_bretherton,hazel}. However, when this mean flow
+absorption was numerically studied including full nonlinear interactions, new
+phenomena not described by the linear theory (reflection off the critical layer
+and sharpening of the mean flow) were observed\cite{jones_num,winters1994}; this
+nonlinear behavior significantly affectedthe evolution of the atmosphere over
+time. \emph{This highlights the importance of numerical simulation in capturing
+the wave breaking process.}
 
-To gain insight into the tidal dissipation process, I began by using the
+To gain insight into the tidal dissipation process, I began by adapting the
 spectral hydrodynamics code Dedalus\cite{dedalus} to study IGW breaking in a 2D
@@
 simulating complex hydrodynamical phenomena, as spectral methods have no
-inherent numerical viscosity and so better resolve the nonlinear cascade to
-small length scales in wave breaking.
+inherent numerical viscosity and so can more accurately resolve the nonlinear
+cascade to small length scales in wave breaking.
 
-Working with Dr.\ Daniel Lecoanet (Princeton; one of the authors of the Dedalus
-code) and my advisor, Prof.\ Dong Lai, I simulated an upward-propagating IGW
-wavetrain excited at the bottom of the atmosphere. I observe the excited wave
-breaking and depositing horizontal momentum in the fluid, causing the fluid to
-acquire an average horizontal flow, consistent with previous
-studies\cite{fullerII}. I derive simple formulae for the location and spatial
-extent of the dissipation zone where the IGW is absorbed by the fluid. More
-interestingly, I observe partial reflection of the IGW at the synchronization
-layer\cite{me}, a phenomenon not considered in the current astrophysical
-literature but consistent with the aforementioned results\cite{winters1994}. A
-sample simulation is presented in \autoref{fig:nl_fluxes}. I am preparing these
-results for publication\cite{me}.
+Working with Dr.\ Daniel Lecoanet (a post-doc at Princeton University, and one
+of the authors of the Dedalus code) and my advisor, Prof.\ Dong Lai, I have
+simulated the nonlinear evolution of an upward-propagating IGW wavetrain excited
+at the bottom of the atmosphere. My simulations show the waves breaking and
+depositing horizontal momentum in the fluid, causing the fluid to acquire an
+average horizontal flow, consistent with previous studies\cite{fullerII}. I have
+derived simple formulae for the location and spatial extent of the dissipation
+zone where the IGW is absorbed by the fluid. More interestingly, my simulation
+reveals a partial reflection of the IGW at the critical layer\cite{me}, a
+phenomenon not considered in the current astrophysical literature but consistent
+with the aforementioned results\cite{winters1994}. A sample simulation is
+presented in \autoref{fig:nl_fluxes}. I am preparing these results for
+publication\cite{me}.
 
@@
     \centering
-    \includegraphics[width=0.7\textwidth]{nl_fluxes.png}
+    \includegraphics[width=0.8\textwidth]{nl_fluxes.png}
     \caption{The evolution of the average horizontal flow velocity of the fluid
@@
     horizontal momentum flux contained in the IGW $S_{px}(z)$ (in units of the
-    total excited flux $S_0$) for one of my simulations. Different colored lines
+    total excited flux $S_0$) for one of my simulations. Different lines
     correspond to different times in the simulation. Times are measured in units
     of $N^{-1}$, the inverse of the Brunt-V\"ais\"al\"a frequency (or the
-    buoyancy period), and heights in units of $H$ the scale height of the
-    density stratification. IGW are excited at $z = 2H$ and propagate to higher
-    $z$ before undergoing wave breaking. The sharp decrease in $S_{px}$ is
-    the location of the critical layer where the IGW breaks and drives $U_0$
+    buoyancy period), and heights in units of $H$ (the scale height of the
+    density stratification). IGWs are excited at $z = 2H$ and propagate to
+    higher $z$ before undergoing wave breaking. The sharp decrease in $S_{px}$
+    is the location of the critical layer where the IGW breaks and drives $U_0$
     towards $c_{ph, x}$; it moves to lower $z$ over time. The apparent decrease
-    in $S_{px}$ at later times indicates reflection off the critical
+    in $S_{px}$ at later times indicates wave reflection off the critical
     layer.}\label{fig:nl_fluxes}
@@
 
-It is therefore clear that full numerical modeling of IGW breaking is necessary
-to fully characterize tidal dissipation. The first aim of my proposal is then:
-\textbf{I will extend my preliminary results to characterize the dynamics of
-nonlinear IGW breaking in realistic WDs.} Via numerical simulation, I will
-develop simple models for tidal dissipation that can be used to study WD
-evolution under tidal heating without performing further simulations. I will
-continue my work in the following stages:
+It is clear that full numerical modeling of IGW breaking is necessary to fully
+characterize tidal dissipation. The first aim of my proposal is to
+\textbf{extend my preliminary results concerning IGW breaking in stratified
+atmospheres to characterize the dynamics of nonlinear IGW breaking in realistic
+WD models.} Via numerical simulation, I will develop models for tidal
+dissipation that can be used to study long-term WD evolution under tidal heating
+without expensive hydrodynamical simulations. I will continue my work in the
+following stages:
 \begin{itemize}
     \item I will perform simulations examining the validity of my results in
-        more realistic geometries, including polar and spherical geometries. I
-        will continue to use Dedalus, which also supports these coordinate
-        systems.
+        spherical geometries to capture tidal effects in WD binaries. I will
+        continue to use Dedalus, which supports spherical coordinates. Although
+        3D simulations are necessarily more complex than my 3D simulations of
+        stratified atmospheres, the underlying dynamics of IGW breaking are the
+        same.
 
     \item I will extend my simulations to realistic WD models and equations of
-        state such as those in~\cite{brassard1992}, continuing to track the
-        location and spatial extent of the dissipation layer as well as any new
-        phenomena. As WDs vary widely in composition and effective temperature,
-        studying representative WD models is vital to obtaining a robust
-        characterization of tidal dissipation. My selected WD models are those
-        used in the current literature\cite{fullerII, fullerIV}.
+        state such as those in~\cite{brassard1992} as well as those generated by
+        MESA (see \autoref{s:3} below), continuing to track the location and
+        spatial extent of the dissipation layer as well as any new phenomena. As
+        WDs vary widely in composition and effective temperature, studying
+        representative WD models is vital to obtaining a robust characterization
+        of tidal dissipation.
 \end{itemize}

### 2019-02-12 sims/2d_3_final/strat_helper.py

@@ -668,11 +668,11 @@ def write_front(name, params):
     flux_th = get_flux_th(params)
     flux_threshold = flux_th * 0.3
 
-    # load if exists
+    # generate if does not exist
     if not os.path.exists(logfile):
         print('log file not found, generating')
         sim_times, domain, state_vars = load(
-            name, params, dyn_vars, plot_stride=1, start=0)
+            name, params, dyn_vars, plot_stride=4, start=0)
         x = domain.grid(0, scales=params['INTERP_X'])
         z = domain.grid(1, scales=params['INTERP_Z'])
         xmesh, zmesh = quad_mesh(x=x[:, 0], y=z[0])
@@ -684,6 +684,8 @@ def write_front(name, params):
         Spx01 = []
         Spx10 = []
         Spx11 = []
+        x_amps = []
+        z_amps = []
 
         S_px = horiz_mean(state_vars['S_{px}'], N_X)
         u0 = horiz_mean(state_vars['ux'], N_X)
@@ -691,26 +693,39 @@ def write_front(name, params):
         uz_est = params['F'] * get_uz_f_ratio(params)
         ux_est = uz_est * KZ / KX
         for t_idx, sim_time in enumerate(sim_times):
-            dux = state_vars['ux'][t_idx] - get_anal_ux(params, sim_time, x, z)
-            duz = state_vars['uz'][t_idx] - get_anal_uz(params, sim_time, x, z)
+            # figure out what to subtract by convolution
+            front_idx = get_front_idx(S_px[t_idx], flux_threshold)
+            z_bot = get_z_idx(params['Z0'] + 3 * params['S'], z0)
+
+            front_idx = (front_idx + z_bot) // 2 # convolve only over bottom half
+            anal_ux = get_anal_ux(params, sim_time, x, z)[:, z_bot: front_idx]
+            anal_uz = get_anal_uz(params, sim_time, x, z)[:, z_bot: front_idx]
+            x_amp = np.sum(state_vars['ux'][t_idx, :, z_bot: front_idx]
+                           * anal_ux) / np.sum(anal_ux**2)
+            z_amp = np.sum(state_vars['uz'][t_idx, :, z_bot: front_idx]
+                           * anal_uz) / np.sum(anal_uz**2)
+            x_lin_est = x_amp * get_anal_ux(params, sim_time, x, z)
+            z_lin_est = z_amp * get_anal_uz(params, sim_time, x, z)
+
[…]

### 2019-02-12 presentations/weeklies/meeting_190205.tex
reorg meetings
compute reflectivity(t)
plot Ri(nu), R(nu), explore Ri-R?

--- 2019-02-05 presentations/meeting.tex
@@
         \centering
-        \includegraphics[width=0.7\textwidth]{../sims/2d_3_final/snapshots_lin_1/fluxes.png}
+        \includegraphics[width=0.7\textwidth]{../../sims/2d_3_final/snapshots_lin_1/fluxes.png}
         \caption{Fits linear $S_{px}(z; \nu)$.}
@@
                 \centering
-                \includegraphics[width=\textwidth]{../sims/2d_3_final/snapshots_nl_4/fluxes.png}
+                \includegraphics[width=\textwidth]{../../sims/2d_3_final/snapshots_nl_4/fluxes.png}
             \end{figure}
@@
                 \centering
-                \includegraphics[width=\textwidth]{../sims/2d_3_final/snapshots_nl_1/fluxes.png}
+                \includegraphics[width=\textwidth]{../../sims/2d_3_final/snapshots_nl_1/fluxes.png}
             \end{figure}
@@
                 \centering
-                \includegraphics[width=\textwidth]{../sims/2d_3_final/snapshots_nl_4/front.png}
+                \includegraphics[width=\textwidth]{../../sims/2d_3_final/snapshots_nl_4/front.png}
             \end{figure}
@@
                 \centering
-                \includegraphics[width=\textwidth]{../sims/2d_3_final/snapshots_nl_1/front.png}
+                \includegraphics[width=\textwidth]{../../sims/2d_3_final/snapshots_nl_1/front.png}
             \end{figure}
@@
     \frametitle{Results}
+    \framesubtitle{Absorbed Flux and Front Propagation}
+
+    \begin{table}[t]
+        \centering
+        \begin{tabular}{c c c}
+            $\nu$ & $\Delta S_{px; S}$ & LinVisc Extrapolation\\\midrule
+            $0.7\nu_0$ & $0.55$ & $0.52$\\
+            $0.4\nu_0$ & $0.54$ & $0.66$\\
+            $0.4\nu_0$ (double strength) & $0.42$ & $0.56$\\
+            $0.3\nu_0$ & $0.51$ & $0.71$\\
+            $0.2\nu_0$ & $0.47$ & $0.76$\\
+            $0.1\nu_0$ & $0.45$ & $0.82$
+        \end{tabular}
+        \caption{Fluxes in units of $S_{px, 0}$ excited flux. For linear viscous
+        extrapolation, I took a slightly earlier (but not immediate) flux and
+        extrapolated to an average $z_c$, so rather approximate.}
+    \end{table}
+\end{frame}
+
+\begin{frame}
+    \frametitle{Results}
     \framesubtitle{Residual Flow Properties}
@@
                 \centering
-                \includegraphics[width=\textwidth]{../sims/2d_3_final/snapshots_nl_4/fft.png}
+                \includegraphics[width=\textwidth]{../../sims/2d_3_final/snapshots_nl_4/fft.png}
             \end{figure}
@@
                 \centering
-                \includegraphics[width=\textwidth]{../sims/2d_3_final/snapshots_nl_1/fft.png}
+                \includegraphics[width=\textwidth]{../../sims/2d_3_final/snapshots_nl_1/fft.png}
                 \caption{Overdamped for comparison, not power law!}
@@
                 \centering
-                \includegraphics[width=\textwidth]{../sims/2d_3_final/snapshots_lin_2/fft.png}
+                \includegraphics[width=\textwidth]{../../sims/2d_3_final/snapshots_lin_2/fft.png}
                 \caption{Linear for comparison; taken just below damping zone.}
@@
                 \centering
-                \includegraphics[width=\textwidth]{../sims/2d_3_final/snapshots_nl_4/p_009.png}
+                \includegraphics[width=\textwidth]{../../sims/2d_3_final/snapshots_nl_4/p_009.png}
                 \caption{Residuals for not overdamped. }
@@
                 \centering
-                \includegraphics[width=\textwidth]{../sims/2d_3_final/snapshots_nl_1/p_009.png}
+                \includegraphics[width=\textwidth]{../../sims/2d_3_final/snapshots_nl_1/p_009.png}
                 \caption{Residuals for overdamped.}

### 2019-02-12 presentations/weeklies/meeting_190212.tex
reorg meetings
compute reflectivity(t)
plot Ri(nu), R(nu), explore Ri-R?

--- 2019-01-29 presentations/meeting.tex
@@
 \begin{frame}
-    \frametitle{Linear}
+    \frametitle{Review}
+
+    \begin{itemize}
+        \item Last time: found flux absorbed $\neq$ flux injected w/
+            $\nu$-extrapolation, \textbf{must have reflection.}
+
+        \item Hypothesis:
+            \begin{equation}
+                \vec{u} = A_0\vec{u}_0 + \delta \vec{u},
+            \end{equation}
+            where $A_0$ is set by forcing, $\delta \vec{u}$ is ``turbulent.''
+
+        \item Since $S_{px} = \ev*{\rho u_xu_z}_x$, decided to plot
+            \begin{equation}
+                \delta S_{px} = \ev*{\rho \delta u_x \delta u_z}_x.
+            \end{equation}
+    \end{itemize}
+\end{frame}
+
+\begin{frame}
+    \frametitle{Results}
+    \framesubtitle{Driving amplitude}
+
+    \begin{itemize}
+        \item Actually, discovered $A < A_0$, \textbf{driving amplitude changes
+            over time} (via convolution). Oscillation timescale $\sim
+            \frac{z_0 - z_c}{v_{gz}}$ for higher order modes.
+
+        \item Adjusting for this gives $\ev*{\rho u_{x0} \delta u_z}_x,
+            \ev*{\rho \delta u_x u_{z0}}_x$ centered on zero, no net flux.
 
@@
             \centering
-        \includegraphics[width=0.7\textwidth]{../sims/2d_3_final/snapshots_lin_0/fluxes.png}
-        \caption{Linear simulation.}
+            \begin{subfigure}{0.45\textwidth}
+                \centering
+                \includegraphics[width=\textwidth]{../../sims/2d_3_final/snapshots_lin_1/f_amps.png}
+                \caption{Linear excited $\frac{u_x}{u_{x0}}, \frac{u_z}{u_{z0}}$ over
+                time.}
+            \end{subfigure}
+            \begin{subfigure}{0.45\textwidth}
+                \centering
+                \includegraphics[width=\textwidth]{../../sims/2d_3_final/snapshots_nl_4/f_amps.png}
+                \caption{Nonlinear excited $\frac{u_x}{u_{x0}}, \frac{u_z}{u_{z0}}$ over
+                time.}
+            \end{subfigure}
+        \end{figure}
+    \end{itemize}
+\end{frame}
+
+\begin{frame}
+    \frametitle{Results}
+    \framesubtitle{Driving Amplitude}
+
+    \begin{figure}[t]
+        \centering
+        \includegraphics[width=0.6\textwidth]{../../sims/2d_3_final/snapshots_nl_4/fluxes_time.png}
+        \caption{Decompositions of the flux over time. $01$ means $u_{x0} \delta
+        u_z$. Oscillating $01, 10$ implies reflected $k_z \to -k_z$ mode.}
     \end{figure}
@@
 \begin{frame}
-    \frametitle{New Plotting for Nonlinear}
+    \frametitle{Results}
+    \framesubtitle{Front Position}
 
@@
         \centering
-        \includegraphics[width=0.7\textwidth]{../sims/2d_3_final/snapshots_nl_3/fluxes.png}
-        \caption{Nonlinear simulation, new plotting methodology. Feedback?}
+        \includegraphics[width=0.6\textwidth]{../../sims/2d_3_final/snapshots_nl_4/front.png}
+        \caption{Front evolution, nonlinear.}
     \end{figure}
@@
 \begin{frame}
-    \frametitle{Reflection?}
+    \frametitle{Results}
+    \framesubtitle{Minimum Ri}
 
-    Get exact solution, up to viscous dissipation. Plot $\delta u_z =
-    \frac{u_{z,sim} - u_{z, anal}}{\abs*{u_{z, anal}}}$, ``fractional
-    deviation'' between driving zone and critical layer. Compute $\mathrm{RMS} =
-    \sqrt{\ev*{\delta u_z^2}}$.
     \begin{figure}[t]
         \centering
-        \includegraphics[width=0.7\textwidth]{../sims/2d_3_final/snapshots_lin_1/p_074.png}
-        \caption{Linear simulation with same viscosity as nonlinear. Green is
-        viscosity-free envelope, orange analytical.}
+        \includegraphics[width=0.6\textwidth]{../../sims/2d_3_final/snapshots_nl_4/fluxes.png}
+        \caption{Mean flow and flux evolution, non-linear.}
     \end{figure}
@@
 \begin{frame}
-    \frametitle{Reflection (nl4)}
+    \frametitle{Conclusion}
+    \framesubtitle{All effects}
 
-    Get exact solution, up to viscous dissipation. Plot $\delta u_z =
-    \frac{u_{z,sim} - u_{z, anal}}{\abs*{u_{z, anal}}}$, ``fractional
-    deviation'' between driving zone and critical layer. Compute $\mathrm{RMS} =
-    \sqrt{\ev*{\delta u_z^2}}$.
-    \begin{figure}[t]
-        \centering
-        \includegraphics[width=0.7\textwidth]{../sims/2d_3_final/snapshots_nl_4/p_074.png}
-        \caption{Nonlinear simulation. Green is viscosity-free envelope, orange
-        analytical.}
-    \end{figure}
-\end{frame}
+    \begin{itemize}
+        \item Excite wave with $S_{px}$. Deposit some $\eta S_{px}$ in
+            critical layer (efficiency $\eta$).
 
-\begin{frame}
-    \frametitle{Reflection?}
+        \item Critical layer width $\delta z$ such that $\mathrm{Ri} \gtrsim
+            \frac{1}{4}$.
 
-    \begin{table}[t]
-        \centering
-        \begin{tabular}{l|l l }
-            Name & RMS & Notes\\
-            lin0 & 0.077 & Almost-no visc\\
-            lin1 & 0.067 & Same visc as nl4\\
-            nl1 (low) & 0.132 & half-res nl1, higher visc\\
-            nl1 & 0.120 & $0.7\nu$\\
-            nl2 & 0.213 & $0.4\nu$\\
-            nl3 & 0.179 & $0.4\nu$ (double A)\\
-            nl4 & 0.269 & $0.3\nu$\\
-            nl5 & 0.207 & $0.3\nu$ (double $k_z$)\\
-            nl6 & 0.246 & $0.2\nu$\\
-            nl7 & 0.358 & $0.1\nu$
-        \end{tabular}
-        \caption{Table of RMS values.}
-    \end{table}
-\end{frame}
+        \item Critical layer position obeys $\rho c_{ph, x} \pd{z_c}{t}
+            = -\eta S_{px}$, or ($\tau = \frac{H\rho_0(z = 0)c_{ph,x}}
+            {\eta S_{px}}$)
+            \begin{equation}
+                z_c(t) = -H\ln \frac{t - t_i + \tau e^{-z_i/H}}{\tau},
+            \end{equation}
 
-\begin{frame}
-    \frametitle{Reflection (nl6)}
+        \item Reflect $(1 - \eta)S_{px}$, some in $k_z \to k_z$ linear
+            reflection, some in higher-order modes.
 
-    Get exact solution, up to viscous dissipation. Plot $\delta u_z =
-    \frac{u_{z,sim} - u_{z, anal}}{\abs*{u_{z, anal}}}$, ``fractional
-    deviation'' between driving zone and critical layer. Compute $\mathrm{RMS} =
-    \sqrt{\ev*{\delta u_z^2}}$.
-    \begin{figure}[t]
-        \centering
-        \includegraphics[width=0.7\textwidth]{../sims/2d_3_final/snapshots_nl_6/p_074.png}
-        \caption{Nonlinear simulation. Green is viscosity-free envelope, orange
-        analytical.}
-    \end{figure}
-\end{frame}
+        \item Reflected wave causes inefficient excitation $S_{px,0} \to \alpha
+            S_{px, 0}$, $\alpha \lesssim 1$.
 
-\begin{frame}
-    \frametitle{Reflection (nl7)}
-
-    Get exact solution, up to viscous dissipation. Plot $\delta u_z =
-    \frac{u_{z,sim} - u_{z, anal}}{\abs*{u_{z, anal}}}$, ``fractional
-    deviation'' between driving zone and critical layer. Compute $\mathrm{RMS} =
-    \sqrt{\ev*{\delta u_z^2}}$.
-    \begin{figure}[t]
-        \centering
-        \includegraphics[width=0.7\textwidth]{../sims/2d_3_final/snapshots_nl_7/p_074.png}
-        \caption{Nonlinear simulation. Green is viscosity-free envelope, orange
-        analytical.}
-    \end{figure}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Reflection (nl3)}
-
-    Get exact solution, up to viscous dissipation. Plot $\delta u_z =
-    \frac{u_{z,sim} - u_{z, anal}}{\abs*{u_{z, anal}}}$, ``fractional
-    deviation'' between driving zone and critical layer. Compute $\mathrm{RMS} =
-    \sqrt{\ev*{\delta u_z^2}}$.
-    \begin{figure}[t]
-        \centering
-        \includegraphics[width=0.7\textwidth]{../sims/2d_3_final/snapshots_nl_3/p_058.png}
-        \caption{Nonlinear simulation. Green is viscosity-free envelope, orange
-        analytical.}
-    \end{figure}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Reflection (nl5)}
-
-    Get exact solution, up to viscous dissipation. Plot $\delta u_z =
-    \frac{u_{z,sim} - u_{z, anal}}{\abs*{u_{z, anal}}}$, ``fractional
-    deviation'' between driving zone and critical layer. Compute $\mathrm{RMS} =
-    \sqrt{\ev*{\delta u_z^2}}$.
-    \begin{figure}[t]
-        \centering
-        \includegraphics[width=0.7\textwidth]{../sims/2d_3_final/snapshots_nl_5/p_074.png}
-        \caption{Nonlinear simulation. Green is viscosity-free envelope, orange
-        analytical.}
-    \end{figure}
+        \item Probably can't be any more exact, time delays + noisy data.
+    \end{itemize}
 \end{frame}

### 2019-02-18 writeups/notes.tex
correct mask, justification

@@
             + \frac{1}{2k_zH}} & z < z_0\\
-    \end{cases}.
+    \end{cases}.\label{eq:anal_forced}
 \end{equation}
@@
 
+\section{Masking Advective Terms}
+
+Recall \autoref{eq:anal_forced} tells us the amplitude of the excited wave in
+terms of the forcing function. However, numerically, it appears that the forcing
+zone evolves a local mean flow $\bar{U}_0 < 0$ due to the nonlinear terms. The
+effect of such a term is to replace $\omega_0 \to \omega_0 - k_x\bar{U}_0 >
+\omega_0$ in \autoref{eq:anal_forced}, which has the effect of attenuating the
+forcing amplitude by $\sim \frac{1}{1 + \frac{\bar{U}_0}{c_{ph, x}}} \approx
+\p*{1 - \frac{\bar{U}_0}{c_{ph, x}}}$.
+
+The cause of this term comes from the viscous terms, which $\sim \partial_t -
+\nu \nabla^2$ (our particular implementation differs slightly). In the
+$\delta$-function forcing limit, $u_x, u_z$ are only $\mathcal{C}^0$ at $z_0$.
+The derivative jumps at $z_0$ with height dictated by the forcing strength.
+Thus, $\nabla^2 u_x, \nabla^2 u_z \sim \delta(z - z_0)$, and there is a strong
+extra contribution to the EOM at $z_0$.
+
 \section{Estimating Various Other Scales}

### 2019-02-18 sims/2d_3_final/strat.py

@@ -41,7 +41,7 @@ def build_interp_params(interp_x, interp_z, overrides=None):
 
     params['T_F'] = params.get('T_F', T_F)
     params['NL'] = params.get('NL', True)
-    params['adv_mask'] = params.get('adv_mask', False)
+    params['mask'] = params.get('mask', False)
     params['g'] = g
     params['OMEGA'] = OMEGA
     params['S'] = params['ZMAX'] / 512 * 4
@@ -96,7 +96,7 @@ if __name__ == '__main__':
                                              'Re': 0.3})),
        (set_ic, 'nl_4_masked',
         build_interp_params(1, 1, overrides={'F_MULT': 1,
-                                             'adv_mask': True,
+                                             'mask': True,
                                              'Re': 0.3})),
        (set_ic, 'nl_5',
         build_interp_params(1, 1, overrides={'F_MULT': 1,

### 2019-02-18 sims/2d_3_final/strat_helper.py

@@ -113,53 +113,6 @@ def set_ic(name, solver, domain, params):
     print('Loaded snapshots')
     return write, dt
 
-def add_nl_problem(problem):
-    problem.add_equation('dx(ux) + uz_z = 0')
-    problem.add_equation(
-        'dt(U) - uz / H' +
-        '- NU * (dx(dx(U)) + dz(U_z) - 2 * U_z / H)' +
-        '= - sponge * U' +
-        '- NL * adv_mask * (ux * dx(U) + uz * dz(U))' +
-        '+ NU * (dx(U) * dx(U) + U_z * U_z)' +
-        '+ F * exp(-(z - Z0)**2 / (2 * S**2) + Z0 / H) *' +
-            'cos(KX * x - OMEGA * t)')
-    problem.add_equation(
-        'dt(ux) + dx(W) + (g * H) * dx(U)' +
-        '- (NU * dx(dx(ux)) + NU * dz(ux_z))' +
-        '+ NU * dz(ux) / H'
-        '= - sponge * ux' +
-        '- NL * adv_mask * (ux * dx(ux) + uz * dz(ux))' +
-        '- NU * (dx(U) * dx(ux) + U_z * ux_z)' +
-        '+ NU * ux * (dx(dx(U)) + dz(U_z))' +
-        '+ NU * ux * (dx(U) * dx(U) + U_z * U_z)' +
-        '- 2 * NU * ux * U_z / H' +
-        # '+ NU * ux * (1 - exp(-U)) / H**2' +
-        '+ NU * ux * U / H**2' +
-        '- NL * (W * dx(U))')
-    problem.add_equation(
-        'dt(uz) + dz(W) + (g * H) * dz(U) - W/H' +
-        '- (NU * dx(dx(uz)) + NU * dz(uz_z))' +
-        '+ NU * dz(uz) / H'
-        '= - sponge * uz - NL * (ux * dx(uz) + uz * dz(uz))' +
-        '- NU * adv_mask * (dx(U) * dx(uz) + U_z * uz_z)' +
-        '+ NU * uz * (dx(dx(U)) + dz(U_z))' +
-        '+ NU * uz * (dx(U) * dx(U) + U_z * U_z)' +
-        '- 2 * NU * uz * U_z / H' +
-        # '+ NU * uz * (1 - exp(-U)) / H**2' +
-        '+ NU * uz * U / H**2' +
-        '- NL * (W * dz(U))')
-    problem.add_equation('dz(ux) - ux_z = 0')
[…]

### 2019-02-19 proposal/FINESST/Su_FINESST_2019.tex
just don't mask visc, stash proposal changes

@@
         style=nature]{biblatex}
-\defbibenvironment{bibliography}
-  {\noindent}
-  {\unspace}
-  {\printtext[labelnumberwidth]{%
-     \printfield{labelprefix}%
-     \printfield{labelnumber}}%
-   \addspace}
-\renewbibmacro*{finentry}{\finentry\addspace}
+% \defbibenvironment{bibliography}
+%   {\noindent}
+%   {\unspace}
+%   {\printtext[labelnumberwidth]{%
+%      \printfield{labelprefix}%
+%      \printfield{labelnumber}}%
+%    \addspace}
+% \renewbibmacro*{finentry}{\finentry\addspace}
 \bibliography{Su_FINESST_2019}
@@
 
-\section{Background and Proposed Research}
+\section{Background and Proposed Research}\label{s:1}
 
@@
         state such as those in~\cite{brassard1992} as well as those generated by
-        MESA (see \autoref{s:3} below), continuing to track the location and
-        spatial extent of the dissipation layer as well as any new phenomena. As
-        WDs vary widely in composition and effective temperature, studying
-        representative WD models is vital to obtaining a robust characterization
-        of tidal dissipation.
+        MESA\cite{MESA} (see \autoref{s:3} below), continuing to track the
+        location and spatial extent of the dissipation layer as well as any new
+        phenomena. As WDs vary widely in composition and effective temperature,
+        studying representative WD models is vital to obtaining a robust
+        characterization of tidal dissipation.
 \end{itemize}
 
-\section{Tidal Heating Thermodynamics}\label{s:3}
+\section{Tidal Heating and Binary White Dwarf Evolution}\label{s:3}
 
@@
 
-As discussed earlier, compact WD binaries exhibit a range of observed and
-hypothesized optical transients. Tidal novae\cite{tidal_novae}, underluminous
+As discussed earlier, compact WD binaries may exhibit a range of transient
+phenomena: tidal novae\cite{tidal_novae}, underluminous
 supernovae\cite{underlum}, and Ca-rich fast transients\cite{carich} are all
@@
 a significant contributor to a WD's total energy budget, a realistic model of
-tidal dissipation is important to understanding a binary WD's thermodynamic
-evolution during inspiral.
+tidal dissipation is important to understanding the thermal evolution of WDs
+during their binary inspiral.
 
-In~\cite{tidal_novae} (hereafter FL), the authors use MESA to study the
-production of tidal novae in binary WDs. A simple two-zone parameterization
+In~\cite{tidal_novae} (hereafter FL), the authors used MESA\cite{MESA} to study
+the production of tidal novae in binary WDs. A simple two-zone parameterization
 where the tidal heat is deposited throughout the outer zone was used to model
-tidal dissipation. They find that cool WDs in sufficiently compact binaries
+tidal dissipation. It was found that cool WDs in sufficiently compact binaries
 (orbital period $\lesssim 15$ minutes) may incur a thermonuclear detonation of
@@
 observed, e.g.\ SDSS J065133+284423 has a period of 12.75 minutes\cite{12min}.
-FL fits the observed properties of SDSS J065133+284423 to their tidal heating
+FL fitted the observed properties of SDSS J065133+284423 to their tidal heating
 models and finds evidence for tidal heating of the secondary WD in SDSS
-J065133+284423. From their fits, FL are able to estimate the temperature of the
-WD in the absence of tidal heating. As WDs form around a well known temperature
-and cool predictably over time, accounting for tidal heating equates to a large
-correction to the age of the WD\@. This illustrates the importance of accurately
-modeling tidal interactions when inferring physical properties.
+J065133+284423.
 
@@
 With the tidal dissipation models I will develop, I will be able to perform a
-simulation similar to that in FL but with a more realistic tidal interaction.
-The second aim of my proposal is accordingly: \textbf{I will use my tidal
-% "obtained in Section 2"
-dissipation profiles to simulate a binary WD undergoing tidal heating and make
-comparisons to observational data.} The location of the tidal heating is a key
-ingredient in determining whether deposited energy can be efficiently transfered
-away or whether the WD experiences sudden detonations. As such, an improved
-understanding of the properties of tidal dissipation is important to proper
-forecasting of binary WDs' thermodynamic evolution. Moreover, while only a few
-sufficiently compact WD binaries were available at the time of writing of FL,
-\emph{Gaia} data releases 2 and 3 will provide many more compact WD binaries to
-examine for evidence of tidal heating.
+binary WD evolution study similar to that in FL but with a realistic tidal
+dissipation profile (instead of their parameterized model). The second aim of my
+proposal is to \textbf{use my tidal dissipation profiles to simulate a binary WD
+undergoing tidal heating and make comparisons to observational data.} The
+location of the tidal heating is a key ingredient in determining whether
+the deposited tidal energy can be efficiently transfered away or whether the WD
+experiences sudden detonations. As such, an improved understanding of the
+properties of tidal dissipation is important to proper forecasting of binary
+WDs' thermodynamic evolution. Moreover, while only a few sufficiently compact WD
+binaries were available at the time of writing of FL, \emph{Gaia} data releases
+2 and 3 will provide many more compact WD binaries to examine for observational
+manifestations of tidal heating.
 
 Using MESA, I will evolve various WD models undergoing tidal heating. From these
-simulations, I will extract the increased temperature of the WDs and make
+studies, I will extract the increased temperature of the WDs and make
 comparisons to observational data, in particular to WDs in new \emph{Gaia} data
@@
 phenomena are vital to guiding future observations. Finally, comparison of
-observational data from known WD binaries to simulations could yield new
-insights to the behavior of degenerate matter.
+observational data from known WD binaries to my theoretical predictions could
+yield new constraints on the interior properties of WDs (e.g.\ crystallization).
 
-\section{Tidal Heating and LISA}\label{s:4}
+\section{Tidal Dissipation and LISA}\label{s:4}
 
-As discussed earlier, WD binaries are a primary source of GW radiation for
-LISA\@. LISA will attain optimal sensitivity at frequencies
+As discussed in \autoref{s:1}, WD binaries are an important source of GW
+radiation for LISA\@. LISA will attain optimal sensitivity at frequencies
 $10^{-4}$--$10^{-1}\;\mathrm{Hz}$\cite{LISA_band}. Exactly in this frequency
-range, tidal effects act to synchronize the orbit of the WDs to the binary
-orbit and transfer energy from the orbit into the WDs. While the decay
-of the binary orbit is still mostly driven by GW radiation, the tidal energy
-dissipation rate grows to $\sim10^{-2}$ the GW energy
-dissipation\cite{fullerII,fullerIV}. An effect of such a magnitude causes the
-phase of the emitted GWs to significantly from the point-mass binary
-prediction; the emitted wave may exhibit ``missing cycles'' due to tidal
-effects\cite{fullerII}. GW astronomy uses matched filtering, where a library of
-template waveforms is matched against instrument data, to identify GW signals.
-As such, the accuracy and completeness of the template library is of utmost
-importance.
+range, tidal effects act to synchronize the spin of the WDs to the binary orbit
+and transfer energy from the orbit into the WDs. While the decay of the binary
+orbit is still mostly driven by GW radiation, the tidal energy dissipation rate
+grows to $\sim10^{-2}$ the GW luminosity\cite{fullerII,fullerIV}. An effect of
+such a magnitude causes the phase of the emitted GWs to deviate significantly
+from the point-mass binary prediction; the emitted wave may exhibit ``missing
+cycles'' due to tidal effects\cite{fullerII}. GW astronomy uses matched
+filtering, where a library of template waveforms is matched against instrument
+data, to identify GW signals. As such, the accuracy and completeness of the
+template library is of utmost importance.
 
-The final aim of my proposal is then: \textbf{I will use my tidal dissipation
-model to compute WD binary GW waveforms including tidal dissipation for use in
-the LISA detection pipeline}. This aim is much less computationally expensive
-than it appears: LISA-band WD binaries can be well described using the Newtonian
+The final aim of my proposal is \textbf{to use my tidal dissipation model to
+compute WD binary GW waveforms including tidal dissipation for use in the LISA
+detection pipeline}. This aim is much less computationally expensive than it
+appears: LISA-band WD binaries can be well described using the Newtonian
 dynamics of two co-orbiting point masses\cite{DWD_pointmass}. The resultant GW
@@
 the GW waveform can be computed analytically without resorting to numerical
-relativity simulations at all, an enormous computational savings. Thus, I will
-compute GW waveforms accounting for the additional phase evolution due to tidal
-dissipation. I will publish my corrected waveforms for use by LISA and the GW
-community.
+relativity simulations at all. Thus, I will compute GW waveforms accounting for
+the additional phase evolution due to tidal dissipation. I will publish my
+corrected waveforms for use by LISA and the GW community.
 
@@
 During the first year of work and first half of the second year, I will complete
-my tidal dissipation model. I anticipate the extension of my 2D plane-parallel
-work to more complex geometries will be complete within the first half year,
-while the extension to realistic WD models will take up to a year. I expect that
-these two results together will produce one peer-reviewed publication in
-addition to the one currently in preparation.
+calculations of tidal dissipation models. I anticipate that the extension of my
+2D plane-parallel work to 3D spherical geometries will be complete within the
+first half year, while the extension to realistic WD models will take up to a
+year. I expect that these two results together will produce two peer-reviewed
+publications in addition to the one currently in preparation.
 
@@
 transfer via IGW breaking will be applicable to astrophysical systems beyond WD
-binaries, for instance angular momentum transfer in stars\cite{l_trans_rev}. I
-will devise new ways of analyzing astrophysically interesting systems in
-\emph{Gaia} and LSST\@. Finally, my GW templates will be important for LISA GW
-detection efforts.
+binaries, for instance angular momentum transfer in massive
+stars\cite{l_trans_rev}. I will devise new ways of analyzing astrophysically
+interesting systems in \emph{Gaia} and LSST\@. Finally, my GW templates will be
+important for LISA GW detection efforts.
 
@@
 
+\clearpage
+
 \printbibliography

### 2019-02-19 sims/2d_3_final/strat_helper.py

@@ -138,7 +138,6 @@ def add_nl_problem(problem):
         '= - sponge * U' +
         '+ F * exp(-(z - Z0)**2 / (2 * S**2) + Z0 / H) *' +
             'cos(KX * x - OMEGA * t)' +
-        '- (1 - NL_MASK) * NU * (dx(dx(U)) + dz(U_z))' +
         '+ NL_MASK * (- (ux * dx(U) + uz * dz(U))' +
         '- NU * 2 * U_z / H' +
         '+ NU * (dx(U) * dx(U) + U_z * U_z))')
@@ -146,7 +145,6 @@ def add_nl_problem(problem):
         'dt(ux) + dx(W) + (g * H) * dx(U)' +
         '- NU * (dx(dx(ux)) + dz(ux_z))' +
         '= - sponge * ux' +
-        '- (1 - NL_MASK) * (NU * (dx(dx(ux)) + dz(ux_z)))' +
         '+ NL_MASK * (- (ux * dx(ux) + uz * dz(ux))'
         '- NU * dz(ux) / H' +
         '- NU * (dx(U) * dx(ux) + U_z * ux_z)' +
@@ -159,7 +157,6 @@ def add_nl_problem(problem):
         'dt(uz) + dz(W) + (g * H) * U_z - W/H' +
         '- NU * (dx(dx(uz)) + dz(uz_z))' +
         '= - sponge * uz' +
-        '- (1 - NL_MASK) * (NU * (dx(dx(uz)) + dz(uz_z)))' +
         '+ NL_MASK * (- (ux * dx(uz) + uz * dz(uz))' +
         '- NU * dz(uz) / H'
         '- NU * (dx(U) * dx(uz) + U_z * uz_z)' +

### 2019-02-19 proposal/FINESST/Su_FINESST_2019.tex
typo fixes

@@
 flares\cite{tidal_novae}. Works in other fields based on numerical simulations
-show that strongly nonlinear wave breaking process exhibits new behavior that
-cannot be described by linear and weakly nonlinear
-theory\cite{winters1994,barker_ogilvie}. Such fully nonlinear numerical
-simulations have not been performed for WDs.
+show that strongly nonlinear wave breaking behaves differently than predictions
+based in linear and weakly nonlinear theory\cite{winters1994,barker_ogilvie}.
+Such fully nonlinear numerical simulations have not been performed for WDs.
 
@@
 small scales that causes wave breaking. \textbf{I propose to study the dynamical
-effects of tidal dissipation via nonlinear Internal Gravity Wave (IGW) breaking
+effects of tidal dissipation via nonlinear internal gravity wave (IGW) breaking
 in binary WDs.} There are three specific goals of my proposed research:
@@
         on the luminosity of WDs in binaries and explore the possibility of
-        producing of observable flares. \autoref{s:3} details my plan.
+        producing of observable flares. \autoref{s:3} details my plan to predict
+        observational manifestations of tidal heating.
 
@@
 The current understanding of tidal synchronization in WD binaries is laid out
-in~\cite{fullerII}: tidal forces from the companion excites IGWs in the deep
+in~\cite{fullerII}: Tidal forces from the companion excites IGWs in the deep
 envelope of the WDs. These IGW propagate outwards and undergo wave breaking in
@@
 transfers energy and angular momentum from the wave to the mean flow of the
-fluid\cite{drazin}; after the mean flow velocity reaches the
+fluid\cite{drazin}. After the mean flow velocity reaches the
 horizontal phase velocity of the IGW, a critical layer forms. Analytical
@@
 and sharpening of the mean flow) were observed\cite{jones_num,winters1994}; this
-nonlinear behavior significantly affectedthe evolution of the atmosphere over
+nonlinear behavior significantly affected the evolution of the atmosphere over
 time. \emph{This highlights the importance of numerical simulation in capturing
@@
 It is clear that full numerical modeling of IGW breaking is necessary to fully
-characterize tidal dissipation. The first aim of my proposal is to
-\textbf{extend my preliminary results concerning IGW breaking in stratified
-atmospheres to characterize the dynamics of nonlinear IGW breaking in realistic
-WD models.} Via numerical simulation, I will develop models for tidal
-dissipation that can be used to study long-term WD evolution under tidal heating
-without expensive hydrodynamical simulations. I will continue my work in the
-following stages:
+capture tidal dissipation. The first aim of my proposal is to \textbf{extend my
+preliminary results concerning IGW breaking in stratified atmospheres to
+characterize the dynamics of nonlinear IGW breaking in realistic WD models.} Via
+numerical simulation, I will develop models for tidal dissipation that can be
+used to study long-term WD evolution under tidal heating without expensive
+hydrodynamical simulations. I will continue my work in the following stages:
 \begin{itemize}

### 2019-02-25 writeups/notes.tex
stash

@@
 
+\subsection{Energy Dissipation Thickness}
+
+The wave action $I = \frac{E}{\tilde{\omega}} = \frac{E}{\omega_0 -
+k_x\bar{U}_0}$ is a conserved quantity for an adiabatically varying shear flow
+$\bar{U}_0(z)$. Then, writing $E = I\tilde{\omega}$, we can compute the energy
+dissipation as a function of $z$:
+\begin{align*}
+    \pd{E}{z} &= -Ik_x\pd{\bar{U}_0}{z},\\
+        &= -\frac{E_0}{\omega_0} k_x \pd{\bar{U}_0}{z},\\
+        &= -\frac{E_0}{N} \abs*{\vec{k}_0} \pd{\bar{U}_0}{z},\\
+        &= -E_0 \frac{\abs*{\vec{k}_0}}{\sqrt{\mathrm{Ri}}}.
+\end{align*}
+
+In most astrophysical applications, $k_z \gg k_x$, and so we find that the
+characteristic dissipation lengthscale is $\sim
+\frac{\sqrt{\mathrm{Ri}}}{k_{z0}}$.
+
 \clearpage

### 2019-02-25 writeups/paper.tex
stash

@@
 
+Compact white dwarf (WD) binary systems, with orbital periods in the range of
+minutes to hours, are important for a range of astrophysical problems. They are
+the most important sources of gravitational waves (GWs) for the Laser
+Interferometric Space Antenna (LISA)\cite{lisa}. They are also thought to
+produce interesting optical transients such as underluminous
+supernovae\cite{underlum}, Ca-rich fast transients\cite{carich}, and tidal
+novae\cite{tidal_novae}. Most importantly, they have been proposed as the likely
+progenitors of type Ia supernovae (e.g.~\cite{Ia0,webbink} or more
+recently\cite{Ia1,Ia2}). While presently only a few tens of compact WD binaries
+are known\cite{lsst_wd}, \emph{Gaia} (currently gathering data) is expected to
+expand the catalog to a few hundreds\cite{lsst_wd} (results based on
+\emph{Gaia}'s second data release have already begun to
+appear\cite{gaiaDD,gaiaDD2}), and the Large Synoptic Survey Telescope (LSST,
+first light scheduled for 2020) will likely detect a few thousand
+more\cite{lsst_wd}. These observations will significantly advance the
+understanding of WD binaries and their evolution. My proposed theoretical and
+computational research is well-timed to take advantage of these new advances.
+
+In spite of the broad importance of WD binaries, the evolution of these systems
+prior to their final mergers is not well understood. Much of this uncertainty
+comes from our imprecise understanding of tidal interactions, which play an
+important role during a compact WD binary's inspiral\cite{fullerII}. Previous
+studies have shown that these interactions manifest as tidal excitation of
+internal gravity waves (IGW), waves in the WD fluid restored by the buoyancy
+force due to density stratification\cite{fullerI}. As these waves propagate
+outwards towards the WD surface, they grow in amplitude until they break, as do
+ocean waves on a shore, and transfer both energy and angular momentum from the
+binary orbit to the outer envelope of the WD\cite{fullerI,fullerII}.
+
+Previous works have found that the dissipation of IGW can generate significantly
+more energy than thermal radiation from the isolated WD surface and is thus a
+major contributor to the WD energy budget\cite{fullerII,fullerIV}. However,
+these works parameterized the wave breaking process in an ad hoc manner. The
+details of dissipation, namely the location and spatial extent of the wave
+breaking, affect the observable outcome: dissipation near the surface of the WD
+can be efficiently radiated away and simply brightens the WD, while dissipation
+deep in the WD envelope causes an energy buildup that results in energetic
+flares\cite{tidal_novae}. Works in other fields based on numerical simulations
+show that strongly nonlinear wave breaking behaves differently than predictions
+based in linear and weakly nonlinear theory\cite{winters1994,barker_ogilvie}.
+Such fully nonlinear numerical simulations have not been performed for WDs.
+
 % Usual WD binaries introduction, from proposals
@@
 % We study IGW breaking in the plane-parallel approximation in 2D. Known to
-% describe both wave steepening (Kloyermeister 1991) and captures mean flow
+% describe both wave steepening (klostermeyer 1991) and captures mean flow
 % absorption (Booker & Bretherton). While the details of breaking are presently
@@
 
-In \autoref{s:equations}, we will cover the numerical setup. In \autoref{s:lin}, we
-will discuss the agreement of our simulation in the low-amplitude limit with
-analytical linear theory. In \autoref{s:nonlin} we will present a few reference
-simulations illustrating critical layer absorption depositing horizontal angular
-momentum into the fluid.
+In \autoref{s:equations}, we will describe the system of equations we will
+use to analyze IGW breaking. In \autoref{s:theory}, we discuss relevant
+analytical results. In \autoref{s:sim} we present the results of numerical
+simulations. Finally, in \autoref{s:discussion} we discuss the results of the
+preceeding section.
 
-\section{Equations}\label{s:equations}
+\section{Problem Description}\label{s:equations}
 
 We consider an incompressible, isothermal fluid, representative of degenerate
-matter in WD bulks, in 2D. We assume a barotropic equation of state as a first
-approximation. As we are interested in the behavior of the wave far from the
-center of the WD, we approximate the gravitational field as uniform. We model
-the background density stratification as $\rho_0(x, z) = \rho_0(z) = \rho_0(z=0)
-e^{-z/H}$ for some reference density $\rho_0(z=0) = \at{\rho_0(z)}_{z = 0}$.
+matter in WD bulks, in 2D. We neglect temperature variations and assume a
+barotropic equation of state as a first approximation. As we are interested in
+the behavior of the wave far from the center of the WD, we approximate the
+gravitational field as uniform. We model the background density stratification
+as $\rho_0(x, z) = \rho_0(z) = \rho_0(z=0) e^{-z/H}$ for some reference density
+$\rho_0(z=0) = \at{\rho_0(z)}_{z = 0}$.
 
@@
 gravitational field are
-\begin{subequations}\label{se:fc_orig}
+\begin{subequations}\label{se:nl_orig}
     \begin{align}
         \vec{\nabla} \cdot \vec{u} &= 0,\\
-        \md{\rho}{t} &= 0 ,\label{eq:fc_density}\\
-        \md{\vec{u}}{t} + \frac{\vec{\nabla}P}{\rho} + g\hat{z} &= 0.
+        \md{\rho}{t} &= 0 ,\label{eq:nl_density}\\
+        \md{\vec{u}}{t} + \frac{\vec{\nabla}p}{\rho} + g\hat{z} &=
+            0.\label{eq:nl_mom}.
     \end{align}
 \end{subequations}
-$\md{}{t} = \pd{}{t} + \p*{\vec{u} \cdot \vec{\nabla}}$ is the material
-derivative. $\vec{u}, \rho, P$ are the fluid dynamical variables and $-g\hat{z}$
-is constant gravitational acceleration. Note that at hydrostatic equilibrium
-$\pd{}{t} = 0$ we have $\vec{\nabla}P_0 = -\rho_0 g\hat{z}$ and so $P_0 = \rho_0
-gH$. A shear flow $u_x(x, z, t) = u_x(z)$ is permitted in hydrostatic
-equilibrium, but we will initially consider no shear flow in equilibrium.
+$\md{}{t} = \pd{}{t} + \p*{\vec{u} \cdot \vec{\nabla}}$ is the Lagrangian or
+material derivative. $\vec{u}, \rho, p$ denote the velocity field, density and
+pressure respectively. We denote $-g\hat{z}$ constant gravitational
+acceleration. Note that at hydrostatic equilibrium $\pd{}{t} = 0$ we have
+$\vec{\nabla}P_0 = -\rho_0 g\hat{z}$ and so $P_0 = \rho_0 gH$. A shear flow
+$u_x(x, z, t) = u_x(z)$ is permitted in hydrostatic equilibrium, but we will
+initially consider no shear flow in equilibrium.
 
-\section{Internal Gravity Waves: Linear Theory}
+In practice, it is convenient to introduce coordinate $\Upsilon = \ln
+\frac{\rho}{\rho_0}$. This both identically enforces $\rho > 0$ and avoids
+numerical issues if $\rho$ is small. We also define reduced pressure $P =
+\frac{p}{\rho}$. Then, we may rewrite the second two equations of
+\autoref{se:nl_orig} as
+\begin{subequations}\label{se:nl_upsilon}
+    \begin{align}
+        \md{\Upsilon}{t} + u_z \pd{\ln \rho_0}{z} &= 0,\label{eq:nl_up_density}
+            \\
+        \md{\vec{u}}{t} + \vec{\nabla}P + P\vec{\nabla}\Upsilon + g\hat{z} &= 0.
+    \end{align}
+\end{subequations}
+Note that in the new coordinates, hydrostatic equilibrium corresponds to
+$\Upsilon_0 = 1, P_0 = gH$.
 
-\subsection{Analytical Properties}
+\section{Internal Gravity Wave Theory}\label{s:theory}
+
+\subsection{Analytical Properties: Linear}
 
@@
 characteristic space and time scales $\pd{}{t} \gg \vec{u} \cdot \vec{\nabla}$,
-we linearize \autoref{se:fc_orig} by ignoring the advective components of the
-material derivative, defining $\rho_1 \equiv \rho - \rho_0 \ll \rho_0 \ll
-\rho_0$ and $P_1 \equiv P - P_0 \ll P_0$ deviations from equilibrium and obtain:
+we may linearize \autoref{se:nl_upsilon}. We ignore the advective components of the
+material derivative and consider small deviations $\Upsilon_1 = \Upsilon -
+\Upsilon_0, P_1 = P - P_0$ about hydrostatic equilibrium:
 \begin{subequations}\label{se:lin_homo}
@@
         \vec{\nabla} \cdot \vec{u} &= 0,\\
-        \pd{\rho_1}{t} - \frac{u_{z} \rho_0}{H} &= 0,\\
-        \pd{\vec{u}}{t} + \frac{\vec{\nabla}P_1}{\rho_0}
-            + \frac{\rho_1 g\hat{z}}{\rho_0}
+        \pd{\Upsilon_1}{t} - \frac{u_{z}}{H} &= 0,\\
+        \pd{\vec{u}}{t} + \vec{\nabla}P_1
+            + \p*{1 + \Upsilon_1} g\hat{z}
             &= 0.
@@
 
-It is well known that the solutions to the $F = 0$ homogeneous
-\autoref{se:lin_homo} are of form\cite{drazin,sutherland0}:
+This can be solved to obtain solutions up to undetermined amplitude $A$ the
+well-known result\cite{drazin,sutherland0}:
 \begin{subequations}\label{se:lin_sol}
@@
         \omega_0^2 &= \frac{N^2k_{0x}^2}{k_{0x}^2 + k_{0z}^2 + \frac{1}{4H^2}},
+            \label{eq:disp_rel}
     \end{align}
 \end{subequations}
-up to undetermined amplitude $A$, and
+where
 \begin{equation}
@@
 \end{equation}
-is the Brunt-V\"ais\"al\"a/buoyancy frequency for our incompressible,
-exponentially stratified atmosphere.
+the Brunt-V\"ais\"al\"a frequency is constant
 
@@
         \begin{align}
-            \vec{c}_{ph} &= \frac{Nk_{0x}}{k_{0x}^2 + k_{0z}^2 + 1/4H^2} \hat{k},\\
+            \vec{c}_{ph} &=
+                \p*{\frac{\hat{x}}{k_{0x}} + \frac{\hat{z}}{k_{0z}}}\omega,\\
             \vec{c}_{g} &= \frac{N\p*{k_{0x}k_{0z}\hat{x}
@@
                 {\p*{k_{0x}^2 + k_{0z}^2 + \frac{1}{4H^2}}^{3/2}} \nonumber\\
-                &\approx \frac{Nk_{0z}}{k_{0x}^2 + k_{0z}^2}\p*{k_{0x}\hat{x} -
-                k_{0z}\hat{z}}.
+                &\approx \frac{\omega k_{0z}}{\abs*{\vec{k}_0}^{3/2}}
+                    \p*{k_{0x}\hat{x} - k_{0z}\hat{z}}.
         \end{align}
@@
         f(z)e^{i(k_{0x}x - \omega_0 t)}$, $x$ averaging and time averaging are
-        equivalent, so we may write ($\ev*{\dots}_x$ denotes $x$-averaging)
+        equivalent, so we may write
         \begin{align}
@@
                     \nonumber\\
-                &\approx -\frac{A^2}{2}\rho_0(z=0)\frac{k_{0z}}{k_{0x}}.
+                &\approx -\frac{A^2}{2}\rho_0(z=0)\frac{k_{0z}}{k_{0x}},
                     \label{eq:fpx_lin}
         \end{align}
-
-    \item Finally, it may be noted that horizontally averaged horizontal
-        momentum $p_x = \ev*{\rho u_x}_x$ obeys $\pd{p_x}{t} + \pd{S_{px}}{z} =
-        0$. Since flux is transported on the group velocity
-        timescale\cite{sutherland0}, this simplifies to the known expression for
-        the wave-induced horizontal mean flow\cite{sutherland0,sutherland1}
-        \begin{equation}
-            \bar{U}_0 \equiv \ev*{u_x}_x = \frac{\ev*{u_xu_z}_x}{c_{gz}}
-        \end{equation}
-        We've denoted $c_{gz} = \at{\pd{\omega}{k_{z}}}_{k_{z0}}$ the vertical group
-        velocity. This may also be rewritten
-        \begin{equation}
-            \bar{U}_0 = -\frac{A^2}{2}e^{\frac{z - z_0}{H}}
-                \frac{k_{0z}}{k_{0x}}c_{gz}, \label{eq:u0_lin}
-        \end{equation}
-        where $c_{gz} = -\frac{Nk_{0x}k_{0z}}{\p*{k_{0x}^2 + k_{0z}^2 +
-        \frac{1}{4H^2}}^{3/2}}$.
+        where $\ev*{\dots}_x$ denotes $x$-averaging.
 \end{itemize}
@@
 To model a continuous IGW wave train excited deep in the WD interior propagating
-towards the surface without unnecessary computational cost, we excite an
-IGW wave train from the bottom of the simulation domain. We use a volumetric
-forcing term to excite IGW at some $z_0 > 0$ interior to the simulation
-domain\footnote{Interfacial forcing at the bottom boundary incurs severe CFL
-timestepping constraints, as we use a Chebyshev polynomial basis along the $z$
-axis in our spectral method which has very small grid spacing at the
-boundaries.}. Our forcing excites both IGWs propagating upwards, imitating a
+towards the surface, we use a volumetric forcing term to excite IGW near the
+bottom of the simulation domain\footnote{Interfacial forcing at the bottom
+boundary incurs strong stability limits, as we use a Chebyshev polynomial basis
+along the $z$ axis in our spectral method which has very small grid spacing at
+the boundaries.}. Our forcing excites both IGWs propagating upwards, imitating a
 wave tidally excited deeper in the WD, and downwards, which are damped away by
-damping layers (see \autoref{ss:numerics}) before inducing strict
-Courant-Friedrichs-Lewy (CFL) timestepping constraints.
+the damping layers described in \autoref{ss:damping}.
 
 As not to interfere with the incompressibility constraint, we force the system
-on the density equation, replacing \autoref{eq:fc_density} with
+on the density equation. This constitutes replacing \autoref{eq:nl_up_density}
+with
 \begin{equation}
-    \md{\rho}{t} = Fe^{-\frac{(z - z_0)^2}{2\sigma^2}}
+    \md{\Upsilon}{t} + u_z\pd{\ln \rho_0}{z}
+        = Fe^{-\frac{(z - z_0)^2}{2\sigma^2}}
             \cos \p*{k_{0x}x - \omega_0 t}.\label{eq:vol_drive}
@@
 Using a narrow Gaussian profile excites a broad $k_z$ wavenumber spectrum, and
-only the $k_{0z}$ satisfying the IGW's dispersion relation for the given
-$k_{0x}, \omega_0(k_{0x}, k_{0z})$ will propagate.
+only the $k_{0z}$ satisfying dispersion relation \autoref{eq:disp_rel} for the
+given $k_{0x}, \omega_0(k_{0x}, k_{0z})$ will propagate.
 
-The linearized \autoref{eq:vol_drive}, obtained by dropping the advective term
-in the material derivative, can be solved analytically up to good accuracy. We
-may first approximate the driving term using $e^{\frac{-(z - z_0)^2}{2\sigma^2}}
-\approx \sqrt{2\pi \sigma^2}\delta(z - z_0)$ if $\sigma \to 0$\footnote{$\sigma$
-must be large enough to be numerically resolved by the spectral code.}. This
-system is solved exactly by matching the two homogeneous solutions above and
-below $z_0$. We may then approximately relax the solution to $k_{0z}\sigma
-\lesssim 1$: an extra factor of $e^{-\frac{(k_{0z}\sigma)^2}{2}}$ arises
-compared to the $\delta$-function solution (evaluating the Fourier Transform
-of $e^{-\frac{(z - z_0)^2}{2\sigma^2}}$ at $k_z = k_{0z}$), and we obtain
+In the linearized system, the effect of this forcing can be solved analytically
+up to good accuracy: we first approximate the driving term using
+$e^{\frac{-(z - z_0)^2}{2\sigma^2}} \approx \sqrt{2\pi \sigma^2}\delta(z -
+z_0)$, the $\sigma \to 0$ limit\footnote{In practice, $\sigma$ must be large
+enough to be numerically resolved by the spectral code.}. This system is solved
+exactly by matching the two homogeneous solutions above and below $z_0$. We may
+then approximately relax the solution to nonzero $\sigma$: an extra factor of
+$e^{-\frac{(k_{0z}\sigma)^2}{2}}$ arises compared to the $\delta$-function
+solution (evaluating the Fourier Transform of $e^{-\frac{(z -
+z_0)^2}{2\sigma^2}}$ at $k_z = k_{0z}$), and we obtain
 \begin{align}
-    u_{1z}&(x, z, t) ={} \frac{Fgk_{0x}^2}{\rho_0(z_0)\omega_0^2}
+    u_{1z}&(x, z, t) ={} \frac{Fgk_{0x}^2}{\omega_0^2}
         \frac{1}{2ik_{0z}}\frac{e^{\frac{-(k_{0z}\sigma)^2}{2}}}
@@
         &{}\begin{cases}
-        e^{\frac{z - z_0}{2H}}e^{i(k_{0x}x + k_{0z}(z - z_0) - \omega_0 t)
-            + \frac{1}{2k_{0z}H}}
+        e^{\frac{z - z_0}{2H}}e^{i\p*{k_{0x}x + k_{0z}(z - z_0) - \omega_0 t
+            + \frac{1}{2k_{0z}H}}}
             & z > z_0\\
-        e^{\frac{z - z_0}{2H}}e^{i(k_{0x}x - k_{0z}(z - z_0) - \omega_0 t)
-            + \frac{1}{2k_{0z}H}}
+        e^{\frac{z - z_0}{2H}}e^{i\p*{k_{0x}x - k_{0z}(z - z_0) - \omega_0 t
+            + \frac{1}{2k_{0z}H}}}
             & z < z_0\\
@@
 The $z > z_0$ solution models an upwards-propagating IGW wavetrain inbound on
-the simulation domain from below. The solution for $u_{1x}(x, z, t)$ moreover is
-just $u_{1x} \approx \frac{k_{0z} u_{1z}}{k_{0x}}e^{\frac{1}{2k_{0z}H}}$.
+the simulation domain from below. The solution for $u_{1x}(x, z, t)$ is
+correspondingly $u_{1x} \approx \frac{k_{0z} u_{1z}}{k_{0x}}e^{\frac{i}{2k_{0z}H}}$.
 
-\section{Numerical Simulation}
+\subsection{Wave Breaking}
+
+In developing the results above, we have neglected advective terms $\pd{}{t} \gg
+\vec{u} \cdot \vec{\nabla}$. However, since $\abs*{\vec{u}} \propto e^{z/2H}$,
+any nonzero amplitude IGW will eventually violate the linearity criterion. This
+regime is relevant in practice since the IGW grows to nonlinear amplitudes
+before it reaches the surface of the WD\cite{fullerI,fullerII}. We can coarsely
+estimate the height at which this happens by $\abs*{\vec{u}}(z) \gtrsim
+\frac{\omega}{\abs*{\vec{k}}}$. This corresponds to when the Lagrangian
+displacement $\vec{\xi} \equiv \vec{u} / \omega$ satisfies $\vec{\xi} \cdot
+\vec{k} \gtrsim 1$, the same criterion used in\cite{fullerI}.
+
+The general understanding of nonlinear IGW breaking is that the waves'
+interactions with the mean flow of the fluid transfer horizontal momentum from
+the waves into the fluid; such a process has been conjectured to be responsible
+for tidal synchronization in stellar binaries\cite{zahn75,gn89} as well as the
+quasi-biennial oscillation\cite{lindzen_qbo}. The details of this process have
+been laid out in a few key papers: IGWs are globally unstable to resonant
+three-wave interactions\cite{drazin}, causing energy transfer out of the IGW to
+daughter modes. These steeper daughter modes facilitate wave breaking,
+depositing horizontal momentum in the fluid's mean flow\cite{klostermeyer}. Once
+the mean flow reaches the horizontal phase velocity of the parent IGW, a
+critical layer forms at which the frequency of the IGW is Doppler-shifted to
+zero. The interaction of the parent IGW with the mean flow was first studied in
+the inviscid, linear regime in~\cite{booker_bretherton}, which found nearly
+complete absorption. This result was reproduced with nonzero
+viscosity\cite{hazel}, but weakly nonlinear theory\cite{brown_stewartson} and
+fully nonlinear simulations\cite{winters1994} suggest that nonlinear effects can
+induce reflection.
+
+An estimate of where these mean flow effects become important can be made using
+the wave-induced mean flow formula\cite{eliassen_palm_cite,sutherland0}:
+\begin{equation}
+     \ev*{u_x} = \bar{U}_0 \equiv \frac{1}{L_x}
+        \int\limits u_x\;\mathrm{d}x = \frac{\ev*{u_xu_z}_x}{c_{g,z}},
+\end{equation}
+where $\ev*{\dots}x$ denotes horizontal averaging. Setting $\bar{U}_0 = c_{ph,
+x}$ the critical layer condition, we can solve that $\vec{\xi} \cdot \vec{k}
+= \sqrt{\frac{k_x}{k_z}}$ is where the mean flow attains the horizontal phase
+velocity. In an astrophysical context, $N \gg \omega$ and $k_z \gg k_x$. Thus,
+wave breaking may occur significantly earlier than assumed in current
+astrophysics literature.
+
+\section{Internal Gravity Wave Numerical Simulation}\label{s:sim}
+
+To study IGW breaking as a weak perturbation grows to nonlinear amplitudes, we
+perform direct numerical simulation using Dedalus\cite{dedalus} a spectral
+numerical method. Our region of interest is strictly interior to the WD and so
+our simulation must strive to eliminate boundary effects resulting from
+nonphysical boundaries. To do so, we adopt periodic boundary conditions in the
+$x$ direction and damping layers (described in \autoref{ss:damping}) in the $z$
+direction to damp perturbations exiting the domain. Correspondingly, we use a
+Fourier basis in the $x$ direction and a Chebyshev in the $z$ direction.
 
@@
 We nondimensionalize by taking $H = N = \rho_0(z=0) = 1$ in
-\autoref{se:fc_orig}.
+\autoref{se:nl_orig}.
 
-where $z_B = 0.05L_z, z_T = 0.95L_z$ are the boundaries of the damping zones. A
-Navier-Stokes numerical viscosity $\nu$ is used to damp high wavenumbers and
-regularize the nonlinear cascade at near grid resolution: $\nu \sim 0.2
-\frac{\omega}{\abs*{k_{0z}}}\frac{L_z}{2\pi N_z}$ was found to be suitable for
-$N_z = 1024$, where $k_{0z}$ is the wavenumber of the excited linear mode.
-%%%
-
-% Description of spectral method, basis functions, initial conditions
 % Grid Size
@@
 
-The exact implementation of these equations in Dedalus is described in
-\autoref{ss:strat_impl}.
-
-\section{Linear Regime}\label{s:lin}
-
-We first test our numerical simulation with a ``linear'' simulation, where
-all flow quantities are small: we assume  $\rho_1 \equiv \rho - \rho_0 \ll
-\rho_0$ and $\vec{u} \cdot \vec{\nabla} \ll \partial_t$. We emphasize that our
-``linear'' simulations solve the full nonlinear equations, just in the weak
-perturbation limit.
-
-$\omega$ is chosen by inverting $\omega(k_{0x}, k_{0z})$ dispersion relation for fixed
-$k_{0x} = 2\pi / L_x$ and some desired $k_{0z} = -2\pi/H$, and $\sigma \lesssim
-\frac{1}{k_{0z}}$ is used to excite a broad band of $k_z$ modes including the
-desired $k_{0z}$ mode.
-
-\subsection{Numerical Simulation}\label{ss:lin_ns}
-
-% Nonlinear: Critical layer absorption (viscous = Hazel, nonlin = Winters),
-% wave-induced mean flow, horizontal momentum flux
-
-We simulate using Dedalus\cite{dedalus} a spectral numerical method. We use a
-Fourier basis in the $x$ direction and a Chebyshev in the $z$ direction. Our
-simulation starts with all perturbation quantities at zero, an initial condition
-at rest. Our parameters are chosen to be as similar as possible to the later
-nonlinear simulations (\autoref{s:nonlin}). We choose as follows:
+Our
+simulation starts with all perturbation quantities at zero. We choose as
+follows:
 \begin{itemize}
@@
 
-    \item We choose $\nu$ $1/10$ that of the nonlinear simulation, to verify
-        that the undamped linear solution is recovered.
-
     \item We choose $F$ forcing strength $1/20$ that of the nonlinear
@@
 \end{itemize}
+
+\subsection{Damping Layers}\label{ss:damping}
+
+To imitate an infinite fluid using a finite simulation domain, we use periodic
+boundary conditions in the $x$ direction and damping layers at the top and
+bottom of the $z$ direction. This damps waves that reach the edge of the
+simulation domain without inducing nonphysical reflection. Specifically, these
+are implemented by introducing linear damping terms to \autoref{se:nl_upsilon}:
+\begin{align}
+    \md{\Upsilon}{t} + u_z \pd{\ln \rho_0}{z} &=
+        -\Gamma(z)\Upsilon,\nonumber\\
+    \md{\vec{u}}{t} + \vec{\nabla}P + P\vec{\nabla}\Upsilon + g\hat{z} &=
+        - \Gamma(z)\vec{u},\nonumber\\
+    \frac{1}{2\tau}\s*{2 + \tanh \frac{z - z_T}{\Delta z}
+        + \tanh \frac{z_B - z}{\Delta z}} &= \Gamma(z),\label{eq:Gamma}
+\end{align}
+where $z_B = 0.05L_z, z_T = 0.95L_z$ are the boundaries of the damping zone.
+This strongly damps perturbations below $z_B$ and above $z_T$ with damping time
+$\tau$, negligibly affects dynamics between $z_B, z_T$ and has transition width
+governed by $\Delta z$; we use $\Delta z = 0.025L_z$. This prescription, used in
+similar studies\cite{lecoanet_damp}, has the advantage of being smooth,
+important for spectral methods. As these damping terms are linear, they are
+retained in the linearized equations of motion as well.
+
+The full implementation of the fluid equations in Dedalus is described in
+\autoref{ss:strat_impl}.
+
+\subsection{Linear Regime}\label{ss:lin}
+
+We first test our numerical simulation with a ``linear'' simulation, where
+all flow quantities are small: we assume  $\rho_1 \equiv \rho - \rho_0 \ll
+\rho_0$ and $\vec{u} \cdot \vec{\nabla} \ll \partial_t$. We emphasize that our
+``linear'' simulations solve the full nonlinear equations, just in the weak
+perturbation limit.
+
+$\omega$ is chosen by inverting $\omega(k_{0x}, k_{0z})$ dispersion relation for fixed
+$k_{0x} = 2\pi / L_x$ and some desired $k_{0z} = -2\pi/H$, and $\sigma \lesssim
+\frac{1}{k_{0z}}$ is used to excite a broad band of $k_z$ modes including the
+desired $k_{0z}$ mode.
+
+\subsection{Numerical Simulation}\label{ss:lin_ns}
 
@@
         \autoref{eq:fpx_lin}.
-
-    \item $\bar{U}_0$ agrees with \autoref{eq:u0_lin} at lower $z$ but deviates
-        significantly at higher $z$. This deviation is caused by the top damping
-        zone that does not conserve horizontal momentum.
 \end{itemize}
@@
 
-\subsection{Nonlinear Regime}\label{s:nonlin}
+\subsection{Mean Flow Critical Layer Absorption}
 
-\subsection{Mean Flow Critical Layer Absorption}
+A
+Navier-Stokes numerical viscosity $\nu$ is used to damp high wavenumbers and
+regularize the nonlinear cascade at near grid resolution: $\nu \sim 0.2
+\frac{\omega}{\abs*{k_{0z}}}\frac{L_z}{2\pi N_z}$ was found to be suitable for
+$N_z = 1024$, where $k_{0z}$ is the wavenumber of the excited linear mode.
+%%%
 
@@
 A purely horizontal shear flow $\bar{U}_0(z) \hat{x}$ can be seen in
-\autoref{se:fc_orig} to have the effect of modifying time derivatives
+\autoref{se:nl_orig} to have the effect of modifying time derivatives
 $\partial_t$ to their frequency in the comoving frame of the fluid $\partial_t -
@@
         \frac{\omega_0}{k_{0z}}$ at the grid spacing.
-
-    \item We choose $F$ such that $\bar{U}_0(z)$ predicted by
-        \autoref{eq:u0_lin} exceeds $\bar{U}_c$ critical velocity for $z <
-        L_z$, i.e.\ the wave-induced mean flow is sufficiently large within the
-        simulation domain to induce critical layer absorption.
 \end{itemize}
@@
 
+\section{Discussion}\label{s:discussion}
+
 \clearpage
@@
 state to obtain a system of equations at most quadratic in fluid fields:
-\begin{subequations}\label{se:fc_var}
+\begin{subequations}\label{se:nl_var}
     \begin{align}
@@
 system conserves horizontal momentum, we begin by adding dissipitive terms to
-the flux-conservative form of the Euler fluid equations \autoref{se:fc_orig} (we
+the flux-conservative form of the Euler fluid equations \autoref{se:nl_orig} (we
 use stress tensor $\tau_{ij} = P\delta_{ij}$):
@@
 modified. Some algebraic manipulation to re-cast it in the form of
-\autoref{se:fc_var} gives
+\autoref{se:nl_var} gives
 \begin{subequations}
@@
             -\Gamma(z) u_z - \p*{\vec{u} \cdot \vec{\nabla}}u_z
-            - T_1 \pd{\Upsilon}{z},\\
-        \Gamma(z) &= 7.5\s*{2 + \tanh \frac{z - z_T}{(L_z - z_T) / 2}
-            + \tanh \frac{z_B - z}{z_B / 2}},\label{eq:Gamma}
+            - T_1 \pd{\Upsilon}{z}.
     \end{align}

### 2019-03-03 proposal/FINESST/Su_FINESST_2019.tex
notes + repropose
final proposal

@@
 
-\section{Background and Proposed Research}\label{s:1}
+\section{Introduction and Goals of Proposed Research}\label{s:1}
 
@@
 breaking, affect the observable outcome: dissipation near the surface of the WD
-can be efficiently radiated away and simply brightens the WD, while dissipation
+can be efficiently radiated away and brightens the WD, while dissipation
 deep in the WD envelope causes an energy buildup that results in energetic
@@
 show that strongly nonlinear wave breaking behaves differently than predictions
-based in linear and weakly nonlinear theory\cite{winters1994,barker_ogilvie}.
+based on linear and weakly nonlinear theory\cite{winters1994,barker_ogilvie}.
 Such fully nonlinear numerical simulations have not been performed for WDs.
@@
 effects of tidal dissipation via nonlinear internal gravity wave (IGW) breaking
-in binary WDs.} There are three specific goals of my proposed research:
+in binary WDs.} There are three specific objectives of my proposed research:
 \begin{itemize}
-    \item Characterize the location, spatial extent, and other properties of
-        wave breaking in realistic WD models via direct numerical simulation.
-        The location and spatial extent of wave breaking will furnish a simple
-        yet effective parameterization of tidal dissipation in a range of WD
-        models. \autoref{s:2} describes the steps required to acheive this goal.
+    \item Objective A\@: Characterize the location, spatial extent, and other
+        properties of wave breaking in realistic WD models via direct numerical
+        simulation. The location and spatial extent of wave breaking will
+        furnish a simple yet effective parameterization of tidal dissipation in
+        a range of WD models. \autoref{s:2} describes the steps required to
+        acheive this goal.
 
-    \item Predict signatures of tidal dissipation over a wide range of possible
-        WD systems. In particular, I will study the impact of tidal heating
-        on the luminosity of WDs in binaries and explore the possibility of
-        producing of observable flares. \autoref{s:3} details my plan to predict
-        observational manifestations of tidal heating.
+    \item Objective B\@: Predict signatures of tidal dissipation over a wide
+        range of possible WD systems. In particular, I will study the impact of
+        tidal heating on the luminosity of WDs in binaries and explore the
+        possibility of producing observable flares. \autoref{s:3} details my
+        plan to predict observational manifestations of tidal heating.
 
-    \item Compute modified GW templates for LISA that account for changes in the
-        phase evolution of the orbit due to tidal dissipation. \autoref{s:4}
-        elaborates on how I will perform this computation.
+    \item Objective C\@: Compute modified GW templates for LISA that account for
+        changes in the phase evolution of the orbit due to tidal dissipation.
+        \autoref{s:4} elaborates on how I will perform this computation.
 \end{itemize}
 
-\section{Nonlinear Tidal Dissipation}\label{s:2}
+\section{Objective A\@: Nonlinear Tidal Dissipation}\label{s:2}
 
@@
 the outer envelope of the WD, locally depositing angular momentum and
-synchronizing the WD spin to the binary orbit. A similar process also
-operates in binaries consisting of early type stars\cite{zahn75,gn89},
-the only major difference being in the specifics of wave
-excitation\footnote{While IGWs in massive stars are excited at their
-radiative-convective boundaries, the excitation of IGWs in WDs is more gradual
-and is associated with sharp composition changes in the stellar
-envelope\cite{fullerII}.}. Nevertheless, direct numerical simulation of the wave
-breaking process has not been performed in either of these systems. Since wave
-breaking is a strongly nonlinear phenomenon, where a larger wave breaks down
-into many smaller-scale waves, numerical simulation is paramount to an accurate
-understanding of the tidal dissipation process.
+synchronizing the WD spin to the binary orbit. A similar process also operates
+in binaries consisting of early type stars\cite{zahn75,gn89}, the only major
+difference being in the specifics of wave excitation\footnote{While IGWs in
+massive stars are excited at their radiative-convective boundaries, the
+excitation of IGWs in WDs is more gradual and is associated with sharp
+composition changes in the stellar envelope\cite{fullerII}.}. Nevertheless,
+direct numerical simulation of the wave breaking process has not been performed
+in either of these systems. Since wave breaking is a strongly nonlinear
+phenomenon, where a larger wave breaks down into many smaller-scale waves,
+numerical simulation is paramount to an accurate understanding of the tidal
+dissipation process.
 
-IGW breaking has been studied in atmospheric sciences. The wave
-breaking process proceeds as follows: Initially, as the IGW reaches nonlinear
-amplitudes, it breaks down via the parametric subharmonic instability and
-transfers energy and angular momentum from the wave to the mean flow of the
-fluid\cite{drazin}. After the mean flow velocity reaches the
-horizontal phase velocity of the IGW, a critical layer forms. Analytical
-calculations show that the IGW is nearly completely absorbed at this critical
-layer in the linear approximation and endows the atmosphere with a mean
-horizontal flow\cite{booker_bretherton,hazel}. However, when this mean flow
-absorption was numerically studied including full nonlinear interactions, new
-phenomena not described by the linear theory (reflection off the critical layer
-and sharpening of the mean flow) were observed\cite{jones_num,winters1994}; this
-nonlinear behavior significantly affected the evolution of the atmosphere over
-time. \emph{This highlights the importance of numerical simulation in capturing
-the wave breaking process.}
+IGW breaking has been studied in atmospheric sciences. The wave breaking process
+proceeds as follows: Initially, as the IGW reaches nonlinear amplitudes, it
+breaks down via the parametric instability and transfers energy and angular
+momentum from the wave to the mean flow of the fluid\cite{drazin}. After the
+mean flow velocity reaches the horizontal phase velocity of the IGW, a critical
+layer forms. Analytical calculations show that the IGW is nearly completely
+absorbed at this critical layer in the linear approximation and endows the
+atmosphere with a mean horizontal flow\cite{booker_bretherton,hazel}. However,
+when this mean flow absorption was numerically studied including full nonlinear
+interactions, new phenomena not described by the linear theory (reflection off
+the critical layer and sharpening of the mean flow) were
+observed\cite{jones_num,winters1994}; this nonlinear behavior significantly
+affected the evolution of the atmosphere over time. \emph{This highlights the
+importance of numerical simulation in capturing the wave breaking process.}
 
-To gain insight into the tidal dissipation process, I began by adapting the
-spectral hydrodynamics code Dedalus\cite{dedalus} to study IGW breaking in a 2D
-isothermal, stratified atmosphere. A spectral code like Dedalus is ideal for
-simulating complex hydrodynamical phenomena, as spectral methods have no
-inherent numerical viscosity and so can more accurately resolve the nonlinear
-cascade to small length scales in wave breaking.
+To gain insight into the tidal dissipation process, I have already begun to
+adapt the spectral hydrodynamics code Dedalus\cite{dedalus} to study IGW
+breaking in a 2D isothermal, stratified atmosphere. A spectral code like Dedalus
+is ideal for simulating complex hydrodynamical phenomena, as spectral methods
+have no inherent numerical viscosity and so can more accurately resolve the
+nonlinear cascade to small length scales in wave breaking.
 
@@
 simulated the nonlinear evolution of an upward-propagating IGW wavetrain excited
-at the bottom of the atmosphere. My simulations show the waves breaking and
-depositing horizontal momentum in the fluid, causing the fluid to acquire an
-average horizontal flow, consistent with previous studies\cite{fullerII}. I have
-derived simple formulae for the location and spatial extent of the dissipation
-zone where the IGW is absorbed by the fluid. More interestingly, my simulation
-reveals a partial reflection of the IGW at the critical layer\cite{me}, a
-phenomenon not considered in the current astrophysical literature but consistent
-with the aforementioned results\cite{winters1994}. A sample simulation is
-presented in \autoref{fig:nl_fluxes}. I am preparing these results for
+at the bottom of the atmosphere (see \autoref{fig:nl_fluxes} for an example). My
+simulations show that waves break and deposit horizontal momentum in the fluid,
+causing the fluid to acquire an average horizontal flow, consistent with
+previous studies\cite{fullerII}. I have derived simple formulae for the location
+and spatial extent of the dissipation zone where the IGW is absorbed by the
+fluid. More interestingly, my simulation reveals a partial reflection of the IGW
+at the critical layer\cite{me}, a phenomenon not considered in the current
+astrophysical literature but consistent with the aforementioned
+results\cite{winters1994}. I am preparing these results for
 publication\cite{me}.
@@
     \includegraphics[width=0.8\textwidth]{nl_fluxes.png}
-    \caption{The evolution of the average horizontal flow velocity of the fluid
-    $U_0$ (in units of $c_{ph,x}$ the horizontal phase velocity of the IGW) and
+    \caption{A sample numerical simulation of the evolution of the nonlinear
+    internal gravity wave (IGW) propagating upwards and breaking in a stratified
+    isothermal atmosphere. The upper panel shows the evolution of the average
+    horizontal flow velocity of the fluid $U_0$ (in units of $c_{ph,x}$ the
+    horizontal phase velocity of the IGW), and the lower panel shows the
     horizontal momentum flux contained in the IGW $S_{px}(z)$ (in units of the
-    total excited flux $S_0$) for one of my simulations. Different lines
-    correspond to different times in the simulation. Times are measured in units
-    of $N^{-1}$, the inverse of the Brunt-V\"ais\"al\"a frequency (or the
-    buoyancy period), and heights in units of $H$ (the scale height of the
-    density stratification). IGWs are excited at $z = 2H$ and propagate to
-    higher $z$ before undergoing wave breaking. The sharp decrease in $S_{px}$
-    is the location of the critical layer where the IGW breaks and drives $U_0$
-    towards $c_{ph, x}$; it moves to lower $z$ over time. The apparent decrease
-    in $S_{px}$ at later times indicates wave reflection off the critical
-    layer.}\label{fig:nl_fluxes}
+    total excited flux $S_0$). Different lines correspond to different times in
+    the simulation. Times are measured in units of $N^{-1}$, the inverse of the
+    Brunt-V\"ais\"al\"a frequency (or the buoyancy period), and heights in units
+    of $H$ (the scale height of the density stratification). IGWs are excited at
+    $z = 2H$ and propagate to higher $z$ before undergoing wave breaking. The
+    sharp decrease in $S_{px}$ corresponds to the location of the critical layer
+    where the IGW breaks and drives $U_0$ towards $c_{ph, x}$; it moves to lower
+    $z$ over time. The apparent decrease in $S_{px}$ at later times indicates
+    wave reflection off the critical layer.}\label{fig:nl_fluxes}
 \end{figure}
@@
 numerical simulation, I will develop models for tidal dissipation that can be
-used to study long-term WD evolution under tidal heating without expensive
-hydrodynamical simulations. I will continue my work in the following stages:
+used to study long-term WD evolution under tidal heating. I will continue my
+work in the following stages:
 \begin{itemize}
@@
         continue to use Dedalus, which supports spherical coordinates. Although
-        3D simulations are necessarily more complex than my 3D simulations of
-        stratified atmospheres, the underlying dynamics of IGW breaking are the
-        same.
+        3D simulations are necessarily more complex than my 2D simulations of
+        stratified atmospheres (see above), the underlying dynamics of IGW
+        breaking are the same.
 
@@
 
-\section{Tidal Heating and Binary White Dwarf Evolution}\label{s:3}
+\subsection{Feasibility and Availability of Resources}
+
+In my preliminary 2D simulations, I use $256$ horizontal and $1024$ vertical
+spectral modes. These simulations take about $2$ days to run on $32$ threads using
+Dedalus\cite{dedalus} using adaptive timestepping. Based on these preliminary
+results, I intend to perform 3D simulations at resolution $128 \times 128 \times
+512$. Including some optimizations I am still developing, I anticipate these
+simulations will take about 2 weeks each when performed using 64 threads.
+
+The computational resources available to me at Cornell include a cluster with
+104 cores/208 threads as well as multiple 32 core/64 thread machines. There are
+sufficient to perform many such 3D simulations within the timespan of the
+proposal.
+
+\section{Objective B\@: Tidal Heating and Binary White Dwarf Evolution
+}\label{s:3}
 
@@
 the hydrogen envelope. WD binaries with such short orbital periods have been
-observed, e.g.\ SDSS J065133+284423 has a period of 12.75 minutes\cite{12min}.
-FL fitted the observed properties of SDSS J065133+284423 to their tidal heating
-models and finds evidence for tidal heating of the secondary WD in SDSS
-J065133+284423.
+observed, e.g.\ SDSS J0651 has a period of 12.75 minutes\cite{12min}. FL fitted
+the observed properties of J0651 to their tidal heating models and finds
+evidence for tidal heating of the secondary WD in this system.
 
@@
 
-With the tidal dissipation models I will develop, I will be able to perform a
-binary WD evolution study similar to that in FL but with a realistic tidal
-dissipation profile (instead of their parameterized model). The second aim of my
-proposal is to \textbf{use my tidal dissipation profiles to simulate a binary WD
-undergoing tidal heating and make comparisons to observational data.} The
-location of the tidal heating is a key ingredient in determining whether
-the deposited tidal energy can be efficiently transfered away or whether the WD
-experiences sudden detonations. As such, an improved understanding of the
-properties of tidal dissipation is important to proper forecasting of binary
-WDs' thermodynamic evolution. Moreover, while only a few sufficiently compact WD
-binaries were available at the time of writing of FL, \emph{Gaia} data releases
-2 and 3 will provide many more compact WD binaries to examine for observational
-manifestations of tidal heating.
+With the tidal dissipation models I will develop in Objective A (see
+\autoref{s:2}), I will be able to perform a binary WD evolution study similar to
+that in FL but with a realistic tidal dissipation profile (instead of their
+parameterized model). The second aim of my proposal is to \textbf{use my tidal
+dissipation profiles to simulate a binary WD undergoing tidal heating and make
+comparisons to observational data.} The location of the tidal heating is a key
+ingredient in determining whether the deposited tidal energy can be efficiently
+transfered away or whether the WD experiences sudden detonations. As such, an
+improved understanding of the properties of tidal dissipation is important to
+proper forecasting of binary WDs' thermodynamic evolution. Moreover, while only
+a few sufficiently compact WD binaries were available at the time of writing of
+FL, \emph{Gaia} data releases 2 and 3 will provide many more compact WD binaries
+to examine for observational manifestations of tidal heating.
 
@@
 
-\section{Tidal Dissipation and LISA}\label{s:4}
+As MESA is a 1D stellar evolution code, I will easily have sufficient
+computational resources to carry out this investigation.
+
+\section{Objective C\@: Tidal Dissipation and LISA}\label{s:4}
 
@@
 detection pipeline}. This aim is much less computationally expensive than it
-appears: LISA-band WD binaries can be well described using the Newtonian
-dynamics of two co-orbiting point masses\cite{DWD_pointmass}. The resultant GW
-emission can then be accurately computed using the weak gravity quadrupole
-approximation (see e.g.~\cite{peters,lsst_wd}). Under these two approximations,
-the GW waveform can be computed analytically without resorting to numerical
-relativity simulations at all. Thus, I will compute GW waveforms accounting for
-the additional phase evolution due to tidal dissipation. I will publish my
-corrected waveforms for use by LISA and the GW community.
+appears: LISA-band WD binaries can be well described using leading-order
+post-Newtonian dynamics and including the effect of tidal
+deformations.\cite{DWD_pointmass}. The resultant GW emission can then be
+accurately computed using the weak gravity quadrupole approximation (see
+e.g.~\cite{peters,lsst_wd}). Under these two approximations, the GW waveform can
+be computed analytically without resorting to numerical relativity simulations
+at all. Thus, I will compute GW waveforms accounting for the additional phase
+evolution due to tidal dissipation. I will publish my corrected waveforms for
+use by LISA and the GW community.
 
@@
 \end{document}
+% mesa school?
+% finish FINESST
+% e disp? plot
+
+% Kushnir?
+% Bouss?? maybe

### 2019-03-03 proposal/FINESST/Su_FINESST_2019.tex
update w/ MESA use ref

@@
 
-As MESA is a 1D stellar evolution code, I will easily have sufficient
-computational resources to carry out this investigation.
+This objective is very feasible using the resources available to me at Cornell.
+Many members of my research group have used MESA in their work on tides in WDs
+(e.g.\ \cite{fullerII,fullerIV,vick}) and will be able to advise me.
+Additionally, as MESA is a 1D stellar evolution code, I will have sufficient
+computing resources to carry out this investigation.
 

### 2019-03-06 proposal/FINESST/Su_FINESST_2019.tex
LISA abbrev
clean up objectives

@@
 the most important sources of gravitational waves (GWs) for the Laser
-Interferometric Space Antenna (LISA)\cite{lisa}. They are also thought to
+Interferometer Space Antenna (LISA)\cite{lisa}. They are also thought to
 produce interesting optical transients such as underluminous
@@
 
-This project is extremely relevant to the NASA Astrophysics research program.
-My work directly relates to (i) the interactions of particles under the extreme
-conditions found in astrophysical situations, (ii) how complex systems create
-and shape the structure and composition of the universe on all scales, and (iii)
-the development of new techniques that can be applied to future major missions.
+This project is extremely relevant to the NASA Astrophysics Division, both in
+terms of supporting the division's objectives and enhancing the science output
+from NASA missions. My work directly relates to (i) the interactions of
+particles under the extreme conditions found in astrophysical situations, (ii)
+how complex systems create and shape the structure and composition of the
+universe on all scales, and (iii) the development of new techniques that can be
+applied to future major missions.
 
-My work will improve understanding of possible energetic phenomena that can
-occur in compact object binaries. My results concerning angular momentum
-transfer via IGW breaking will be applicable to astrophysical systems beyond WD
-binaries, for instance angular momentum transfer in massive
-stars\cite{l_trans_rev}. I will devise new ways of analyzing astrophysically
-interesting systems in \emph{Gaia} and LSST\@. Finally, my GW templates will be
-important for LISA GW detection efforts.
-
-My work has direct relevance to NASA missions in the detectability of
+Towards (i), my work will improve understanding of possible energetic phenomena
+that can occur in compact object binaries. Towards (ii), my results concerning
+angular momentum transfer via IGW breaking will be applicable to astrophysical
+systems beyond WD binaries, for instance in massive stars\cite{l_trans_rev}.
+Towards (iii), my work has direct relevance to NASA missions by forecasting
 astrophysical transients related to WD binaries. For instance, tidal novae are
@@
 been observed with Chandra and the Hubble Space Telescope, among others.
+Additionally, my GW templates will also be important for LISA GW detection
+efforts.
 
@@
 \end{document}
-% mesa school?
-% finish FINESST
-% e disp? plot
-
-% Kushnir?
-% Bouss?? maybe

### 2019-03-06 proposal/FINESST/Su_FINESST_2019.tex
word
update proposal one last time

@@
 \pagestyle{fancy}
-\rfoot{Yubo Su}
 \rhead{}
@@
 terms of supporting the division's objectives and enhancing the science output
-from NASA missions. My work directly relates to (i) the interactions of
-particles under the extreme conditions found in astrophysical situations, (ii)
-how complex systems create and shape the structure and composition of the
-universe on all scales, and (iii) the development of new techniques that can be
-applied to future major missions.
+from NASA missions. My work directly relates to NASA objectives: (i) ``the
+interactions of particles under the extreme conditions found in astrophysical
+situations'', (ii) ``how complex systems create and shape the structure and
+composition of the universe on all scales'', and (iii) ``the development of new
+techniques that can be applied to future major missions''.
 
@@
 been observed with Chandra and the Hubble Space Telescope, among others.
-Additionally, my GW templates will also be important for LISA GW detection
-efforts.
+Additionally, my GW templates will be important for LISA GW detection efforts.
 

### 2019-03-07 proposal/FINESST/Su_FINESST_2019.tex
disclaimer

@@
 
-\section{Objective B\@: Tidal Heating and Binary White Dwarf Evolution
-}\label{s:3}
+\section{Objective B\@: Tidal Heating and White Dwarf Evolution}\label{s:3}
 
@@
 
+\emph{Research performed by Cornell University is conducted as fundamental
+research (basic and applied research ordinarily published and shared broadly
+within the scientific community) and is therefore exempt from the requirements
+of the International Traffic in Arms Regulations (ITAR) and the Export
+Administration Regulations (EAR).}
+
 \clearpage

### 2019-03-23 sims/2d_3_final/strat_helper.py
stash

@@ -1197,7 +1197,7 @@ def plot_front(name, params):
         ax2.plot(t, width_med[start_idx: ], 'g', linewidth=0.7)
         ax2.plot(t, width_min[start_idx: ], 'r:', linewidth=0.5)
         ax2.plot(t, width_max[start_idx: ], 'r:', linewidth=0.5)
-        ax2.set_ylim([0, 1])
+        ax2.set_ylim([0, 0.1])
         ax2.set_ylabel(r"$\Delta z$")
 
         ri_width = N**2 * width_med**2 / (0.7 * u_c)**2
@@ -1209,7 +1209,7 @@ def plot_front(name, params):
         ax3.plot(t,
                  ri_min[start_idx: ],
                  'r:',
-                 linewidth=0.5,
+                 linewidth=0.7,
                  label='Global (min)')
         ax3.plot(t, ri_width[start_idx: ], 'g', linewidth=0.7, label='Width')
         ax3.set_ylim([0, 0.6])

### 2019-04-16 writeups/paper.tex
stash
rm all old plots

@@
 
-We consider an incompressible, isothermal fluid, representative of degenerate
-matter in WD bulks, in 2D. We neglect temperature variations and assume a
+We consider a two-dimensional incompressible, isothermal fluid, representative
+of degenerate matter in WDs. We neglect temperature variations and assume a
 barotropic equation of state as a first approximation. As we are interested in
-the behavior of the wave far from the center of the WD, we approximate the
-gravitational field as uniform. We model the background density stratification
-as $\rho_0(x, z) = \rho_0(z) = \rho_0(z=0) e^{-z/H}$ for some reference density
-$\rho_0(z=0) = \at{\rho_0(z)}_{z = 0}$.
+dynamics far from the center of the WD, we approximate the gravitational field
+as uniform. We model the background density stratification as $\rho_0(x, z) =
+\rho_0(z) = \rho_0(z=0) e^{-z/H}$ for some reference density $\rho_0(z=0) =
+\at{\rho_0(z)}_{z = 0}$.
 
@@
 \begin{itemize}
-    \item The amplitude of the perturbations grows like $e^{z/2H}$. Thus, the
+    \item The amplitude of the wave grows like $e^{z/2H}$. Thus, the
         linear approximation is violated for sufficiently large $z$.
@@
 the simulation domain from below. The solution for $u_{1x}(x, z, t)$ is
-correspondingly $u_{1x} \approx \frac{k_{0z} u_{1z}}{k_{0x}}e^{\frac{i}{2k_{0z}H}}$.
+correspondingly $u_{1x} \approx \frac{k_{0z}
+u_{1z}}{k_{0x}}e^{\frac{i}{2k_{0z}H}}$.
 
@@
 any nonzero amplitude IGW will eventually violate the linearity criterion. This
-regime is relevant in practice since the IGW grows to nonlinear amplitudes
+regime is relevant in practice since the IGW may grow to nonlinear amplitudes
 before it reaches the surface of the WD\cite{fullerI,fullerII}. We can coarsely
-estimate the height at which this happens by $\abs*{\vec{u}}(z) \gtrsim
-\frac{\omega}{\abs*{\vec{k}}}$. This corresponds to when the Lagrangian
-displacement $\vec{\xi} \equiv \vec{u} / \omega$ satisfies $\vec{\xi} \cdot
-\vec{k} \gtrsim 1$, the same criterion used in\cite{fullerI}.
+estimate the height at which this happens by when the Lagrangian displacement
+$\vec{\xi}$ of a fluid parcel satisfies
+\begin{equation}
+    \vec{\xi} \cdot \vec{k} \gtrsim 1.\label{eq:nl}
+\end{equation}
 
@@
 An estimate of where these mean flow effects become important can be made using
-the wave-induced mean flow formula\cite{eliassen_palm_cite,sutherland0}:
+the wave-induced mean flow\cite{eliassen_palm_cite,sutherland0}:
 \begin{equation}
@@
 \end{equation}
-where $\ev*{\dots}x$ denotes horizontal averaging. Setting $\bar{U}_0 = c_{ph,
-x}$ the critical layer condition, we can solve that $\vec{\xi} \cdot \vec{k}
-= \sqrt{\frac{k_x}{k_z}}$ is where the mean flow attains the horizontal phase
-velocity. In an astrophysical context, $N \gg \omega$ and $k_z \gg k_x$. Thus,
-wave breaking may occur significantly earlier than assumed in current
-astrophysics literature.
+where $\ev*{\dots}_x$ denotes horizontal averaging. Setting $\bar{U}_0 = c_{ph,
+x}$ the critical layer condition, we recover $\frac{u_xk_x}{\omega} = \xi_x k_x
+\sim 1$, roughly equivalent to the nonlinearity criterion \autoref{eq:nl}.
 
@@
 perform direct numerical simulation using Dedalus\cite{dedalus} a spectral
-numerical method. Our region of interest is strictly interior to the WD and so
-our simulation must strive to eliminate boundary effects resulting from
-nonphysical boundaries. To do so, we adopt periodic boundary conditions in the
-$x$ direction and damping layers (described in \autoref{ss:damping}) in the $z$
-direction to damp perturbations exiting the domain. Correspondingly, we use a
-Fourier basis in the $x$ direction and a Chebyshev in the $z$ direction.
+numerical method. Our simulation domain is smaller than our full physical
+domain, so we adopt reflection-suppressing boundary conditions. We use periodic
+boundary conditions in the $x$ direction and damping layers (described in
+\autoref{ss:damping}) in the $z$ direction to damp perturbations exiting the
+top and bottom of the domain. We use a Fourier basis in the $x$ direction and
+either a Chebyshev or Fourier basis in the $z$ direction; our results were
+robust to either set of basis functions. We varied the number of $x, z$ modes
+(denoted $N_x, N_z$ respectively), and used $3/2$ dealiasing\cite{boyd}.
+
+The time integration uses a split implicit-explicit first-order scheme where
+certain terms are treated implicitly using for- ward Euler, and the remaining
+terms are treated explicitly using backward Euler. A third-order, four-stage
+DIRK-ERK scheme\cite{ascher} is used with adaptive timesteps taken to be
+the advective Courant-Friedrichs-Lewy (CFL) time given by $0.7 \min (\Delta
+x / u_x,\Delta z / u_z)$. The minimum is taken over every grid point in
+the domain and $\Delta x,\Delta z$ are the grid spacings in the $x$ and $z$
+directions respectively.
 
@@
 
-% Grid Size
-% Viscosity + diffusivity, damping layers (to simulate radiative BC)
-% timestepping? RK443, (CFL limiter??)
+Denote $L_x, L_z$ to be the physical dimensions of the simulation domain. The
+physics of our simulation is fixed by six parameters: $L_x, L_z$, and $k_{0x},
+\omega_0, F, \nu$. We describe and choose these parameters below:
+\begin{itemize}
+    \item $L_z$: Since the wave amplitude grows $\sim e^{z/2H}$, we choose $L_z
+        = 10H$ to give $\sim e^4$ apmlitude growth between the damping zones.
 
-Our
-simulation starts with all perturbation quantities at zero. We choose as
-follows:
-\begin{itemize}
-    \item $k_{0x}$: Astrophysical IGWs have $k_{0x} \ll k_{0z}$, so we choose $k_{0x}
-        = \frac{2\pi}{L_x}$ the smallest permitted wavenumber permitted by
-        periodic boundary conditions.
+    \item $L_x$: We want similar grid spacing $\frac{L_x}{N_x} \sim
+        \frac{L_z}{N_z}$. For computational savings, we fix $N_z = 4N_x$.
+        Expecting wave breaking to be mostly in the $z$ direction, we use $L_x =
+        4H$.
 
-    \item We choose $\omega_0$ by choosing it to produce a desired $k_{0z}$.
-        Astrophysical systems generally also exhibit $k_{0z}H \gg 1$, where the
-        stratification height is significantly larger than the vertical
-        wavelength. However, in order for the simulation to both be well-damped
-        at grid resolutions and be neglibily damped at leading order nonlinear
-        wavelengths, we require $k_{0z} \ll \frac{1}{L_{NL}} \ll \frac{2\pi
-        N_z}{L_z}$, where $L_{NL}$ denotes some characteristic length scale of
-        nonlinear features. Finally, we need $L_z$ to be many $H$ to capture
-        significant $e^{z/2H}$ growth and separate the linear and nonlinear
-        amplitudes of $u_z$.
+    \item $k_{0x}$: Astrophysical IGWs in stars are generally excited by the $l
+        = 2$ component of the tidal potential, for which $k_{\perp} \sim
+        \frac{1}{R}$ where $R$ is the radius of the star. To best emulate this,
+        we use $k_{0x} = \frac{2\pi}{L_x}$ the smallest permitted wavenumber
+        permitted by periodic boundary conditions.
 
-        Fixing $L_z = 10H$ to give $\sim e^4$ amplitude growth between the
-        damping zones, this implies we require separation of scales $1 \ll
-        k_{0z}H \ll \frac{H}{L_{NL}} \ll \frac{2\pi N_z}{10}$.
+    \item $\omega_0$: We choose $\omega_0$ by evaluating dispersion relation
+        $\omega_0\p*{k_{0x}, k_{0z}}$ for some desired $k_{0z}$.
 
-        Since $N_z$ is fixed by computational cost (we tried both $N_z = 512,
-        N_z = 1024$), we choose $k_{0z} \sim \frac{2\pi}{H}$. This is less
-        physically representative but gives more wavenumber space for the
-        nonlinear cascade. We then invert $\omega_0 = \omega(k_{0x}, k_{0z})$.
+        Astrophysical IGWs also generally satisfy $\omega \ll N$, or
+        equivalently $\frac{k_r}{k_{\perp}} \sim k_rR \gg 1$. Since $H \lesssim
+        R$, we aim to study $k_{0z}H \gg 1$. However, to satisfy this well is
+        computationally expensive: $L_z \gg H$ is required to give waves ample
+        room to grow within the simulation domain, but we want $k_{0z}$ to be
+        sufficiently separated from the grid spacing $\sim L_z / N_z$ that
+        nonlinear effects can be seen in later simulations (see
+        \autoref{ss:nl_ns}). Thus, $N_z$ sets how well we can satisfy $k_{0z}H
+        \gg 1$; we choose $k_{0z}H = 1$.
 
-    \item We choose $F$ forcing strength $1/20$ that of the nonlinear
-        simulation, which keeps the flow amplitude sufficiently small to be
-        treated in the linear approximation at all points in the simulation
-        domain.
+    \item $F$: We first choose $F$ forcing strength such that $\vec{\xi} \cdot
+        \vec{k} \ll 1$ is satisfied everywhere throughout the domain. This
+        becomes a constraint on $F$ by using \autoref{eq:uz_lin} and
+        $\abs*{\vec{u}} \propto e^{z/2H}$.
+
+    \item $\nu$: Nonlinear effects transfer wave energy from $\vec{k}_{0}$ to
+        smaller scales. Since well-resolved simulations using spectral methods
+        have no inherent numerical viscosity, energy will accumulate at grid
+        scales in the absence of artificial dissipation. We introduce
+        Navier-Stokes viscosity $\nu$ while ensuring our equations of motion
+        continue to conserve horizontal momentum (see \autoref{ss:strat_impl}
+        for details). When $\vec{\xi} \cdot \vec{k} \ll 1$, we may set $\nu =
+        0$.
 \end{itemize}
@@
 similar studies\cite{lecoanet_damp}, has the advantage of being smooth,
-important for spectral methods. As these damping terms are linear, they are
-retained in the linearized equations of motion as well.
+important for spectral methods. The full implementation of the fluid equations
+in Dedalus is described in \autoref{ss:strat_impl}.
 
-The full implementation of the fluid equations in Dedalus is described in
-\autoref{ss:strat_impl}.
-
-\subsection{Linear Regime}\label{ss:lin}
+\subsection{Linear Regime}\label{ss:lin_ns}
 
 We first test our numerical simulation with a ``linear'' simulation, where
-all flow quantities are small: we assume  $\rho_1 \equiv \rho - \rho_0 \ll
-\rho_0$ and $\vec{u} \cdot \vec{\nabla} \ll \partial_t$. We emphasize that our
-``linear'' simulations solve the full nonlinear equations, just in the weak
-perturbation limit.
-
-$\omega$ is chosen by inverting $\omega(k_{0x}, k_{0z})$ dispersion relation for fixed
-$k_{0x} = 2\pi / L_x$ and some desired $k_{0z} = -2\pi/H$, and $\sigma \lesssim
-\frac{1}{k_{0z}}$ is used to excite a broad band of $k_z$ modes including the
-desired $k_{0z}$ mode.
-
-\subsection{Numerical Simulation}\label{ss:lin_ns}
-
-A representative snapshot of this simulation
-is provided in \autoref{fig:linear_nonu}. Key features to note:
-\begin{itemize}
-    \item $u_z$ grows with \autoref{eq:uz_lin} to reasonable accuracy.
-
-    \item $S_{px}$ is seen to be constant, as would be expected by
-        \autoref{eq:fpx_lin}.
-\end{itemize}
-We also include a qualitatively similar snapshot of the same simulation run
-using the $\nu$ from the later nonlinear simulations in \autoref{fig:linear_nu}.
-\begin{figure}[t]
-    \centering
-    \includegraphics[width=0.5\textwidth]{plots/linear_nonu.png}
-    \caption{Snapshot of linear simulation after reaching steady state.
-        Analytical predictions of $\abs*{u_z}, \bar{U}_0, S_{px}$ are shown in
-        orange. Red dotted lines indicate the onset of the damping region and
-        blue dotted lines denote the driving region.}\label{fig:linear_nonu}
-\end{figure}
-\begin{figure}[t]
-    \centering
-    \includegraphics[width=0.5\textwidth]{plots/linear_nu.png}
-    \caption{Same as \autoref{fig:linear_nonu} but with $0.3\times $viscosity
-    used for nonlinear simulations (YUBONOTE should be the same but that
-    simulation mysteriously stopped without me noticing; restarting worked, so
-    currently rerunning). Note the slight resulting dissipation in $S_{px}$.
-    }\label{fig:linear_nu}
-\end{figure}
+$F$ is chosen such that $\vec{\xi} \cdot \vec{k} \ll 1$ everywhere. We emphasize
+that our ``linear'' simulations solve the full nonlinear equations, just in the
+weak perturbation limit.
 
@@
 
-A
-Navier-Stokes numerical viscosity $\nu$ is used to damp high wavenumbers and
-regularize the nonlinear cascade at near grid resolution: $\nu \sim 0.2
-\frac{\omega}{\abs*{k_{0z}}}\frac{L_z}{2\pi N_z}$ was found to be suitable for
-$N_z = 1024$, where $k_{0z}$ is the wavenumber of the excited linear mode.
-%%%
+A Navier-Stokes numerical viscosity $\nu$ is used to damp high wavenumbers and
+regularize the nonlinear cascade at near grid resolution: $\nu \sim 0.1
+\frac{\omega_0}{\abs*{k_{0z}}}\frac{L_z}{2\pi N_z}$ was found to be suitable for
+$N_z = 2048$.
 
@@
 
-\subsection{Numerical Simulation}
+\subsection{Numerical Simulation}\label{ss:nl_ns}
 
@@
 Two representative snapshots from our simulation are provided in
-\autoref{fig:nl} after the critical layer has had time to form. We may note that
+TODO after the critical layer has had time to form. We may note that
 the critical layer, where $S_{px}$ is absorbed and $\bar{U}_0 = \bar{U}_c$,
 travels downwards as predicted.
-\begin{figure}[t]
-    \centering
-    \begin{subfigure}{0.5\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{plots/nl35_1.png}
-        \caption{Earlier time snapshot. Legend is the same as
-        \autoref{fig:linear_nonu} except $\bar{U}_c$ is marked in red on the
-        third panel.}
-    \end{subfigure}
-
-    \begin{subfigure}{0.5\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{plots/nl35_2.png}
-        \caption{Later snapshot illustrating propagation of $z_c$.}
-    \end{subfigure}
-    \caption{Nonlinear numerical simulation.}\label{fig:nl}
-\end{figure}
 
@@
 
-For the simulation in \autoref{fig:nl}, we may define $z_c = \argmax_z
+For the simulation in TODO, we may define $z_c = \argmax_z
 \pd{\bar{U}_0}{z}$. Computing $\pd{z_c}{t}$ using the analytical flux
 \autoref{eq:fpx_lin} allows comparison to \autoref{eq:zc_anal}, which we exhibit
-in \autoref{fig:zc_anal}. We see overall good agreement, though some small
+in TODO\@. We see overall good agreement, though some small
 deviation is expected since $S_{px}$ is not perfectly conserved owing to
 numerical viscosity (and YUBONOTE $S_{px}$ is misestimated?).
-\begin{figure}[t]
-    \centering
-    \includegraphics[width=0.5\textwidth]{plots/nl35_front_v.png}
-    \caption{Comparison of simulated $\pd{z_c}{t}$ to analytical
-    \autoref{eq:zc_anal}. Plot begins after formation of critical layer in
-    simulation.}\label{fig:zc_anal}
-\end{figure}
 
@@
 layer absorption when $\mathrm{Ri} \gg \frac{1}{4}$. We may plot $\mathrm{Ri}$
-over the same time interval in \autoref{fig:nl35_f_ri}. We may observe that the
+over the same time interval in TODO\@. We may observe that the
 Richardson number is initially decreasing but eventually levels out and
@@
 transition.
-\begin{figure}[t]
-    \centering
-    \includegraphics[width=0.5\textwidth]{plots/nl35_f_ri.png}
-    \caption{Plot of $\mathrm{Ri}_{\max}(t) = \max_z
-    \frac{N^2}{\bar{U}_0'}(t)$ over the simulation time.}\label{fig:nl35_f_ri}
-\end{figure}
 
 We argue that the Richardson number is bounded from below by viscosity. A
-repeated simulation with a larger viscosity is shown in \autoref{fig:nl1_f_ri},
+repeated simulation with a larger viscosity is shown in TODO,
 where the Richardson number does not go nearly as low.
-\begin{figure}[t]
-    \centering
-    \includegraphics[width=0.5\textwidth]{plots/nl1_f_ri.png}
-    \caption{Same as \autoref{fig:nl35_f_ri} but with $\sim 1.3\times$
-    viscosity.}\label{fig:nl1_f_ri}
-\end{figure}
 
@@
 \end{document}
+
+% nonlinear wave breaking, tidal synchronization, dynamical tides, internal
+% gravity waves

### 2019-04-16 writeups/paper.tex
slight cleanup of plots, move to MNRAS style file

@@
-    \documentclass[twocolumn,
-        nofootinbib, % unbreaks footnotes from being put in bib
-        usenames, % allows access to some tikz colors
-        aps,
-        prd,
-        dvipsnames % more colors: https://en.wikibooks.org/wiki/LaTeX/Colors
-    ]{revtex4-1}% chktex 8
+    \documentclass[
+        fleqn,
+        usenatbib,
+    ]{mnras}
     \usepackage{
@@
         amssymb,
-        fancyhdr, % page styling
-        lastpage, % footer fanciness
-        hyperref, % various links
-        setspace, % line spacing
-        amsthm, % newtheorem and proof environment
-        mathtools, % \Aboxed for boxing inside aligns, among others
-        float, % Allow [H] figure env alignment
-        enumerate, % Allow custom enumerate numbering
-        graphicx, % allow includegraphics with more filetypes
-        wasysym, % \smiley!
-        upgreek, % \upmu for \mum macro
-        listings, % writing TrueType fonts and including code prettily
-        tikz, % drawing things
-        booktabs, % \bottomrule instead of hline apparently
-        cancel % can cancel things out!
+        newtxtext,
+        newtxmath,
+        ae, aecompl,
+        graphicx
     }
+    \usepackage[T1]{fontenc}
     \usepackage[
@@
     ]{caption}
-    \usepackage[font=scriptsize]{subcaption} % subfigures
 
-    \newcommand*{\scinot}[2]{#1\times10^{#2}}
-    \newcommand*{\dotp}[2]{\left<#1\,\middle|\,#2\right>}
     \newcommand*{\rd}[2]{\frac{\mathrm{d}#1}{\mathrm{d}#2}}
     \newcommand*{\pd}[2]{\frac{\partial#1}{\partial#2}}
-    \newcommand*{\rtd}[2]{\frac{\mathrm{d}^2#1}{\mathrm{d}#2^2}}
-    \newcommand*{\ptd}[2]{\frac{\partial^2 #1}{\partial#2^2}}
     \newcommand*{\md}[2]{\frac{\mathrm{D}#1}{\mathrm{D}#2}}
-    \newcommand*{\pvec}[1]{\vec{#1}^{\,\prime}}
-    \newcommand*{\svec}[1]{\vec{#1}\;\!}
-    \newcommand*{\ang}[0]{\;\text{\AA}}
-    \newcommand*{\mum}[0]{\;\upmu \mathrm{m}}
     \newcommand*{\at}[1]{\left.#1\right|}
-
-    \newtheorem{theorem}{Theorem}[section]
-
-    \let\Re\undefined
-    \let\Im\undefined
-    \DeclareMathOperator{\Res}{Res}
-    \DeclareMathOperator{\Re}{Re}
-    \DeclareMathOperator{\Im}{Im}
-    \DeclareMathOperator{\Log}{Log}
-    \DeclareMathOperator{\Arg}{Arg}
-    \DeclareMathOperator{\Tr}{Tr}
-    \DeclareMathOperator{\E}{E}
-    \DeclareMathOperator{\Var}{Var}
+    \newcommand*{\abs}[1]{\left|#1\right|}
+    \newcommand*{\ev}[1]{\langle#1\rangle}
+    \newcommand*{\p}[1]{\left(#1\right)}
+    \newcommand*{\s}[1]{\left[#1\right]}
+    \newcommand*{\z}[1]{\left\{#1\right\}}
     \DeclareMathOperator*{\argmin}{argmin}
     \DeclareMathOperator*{\argmax}{argmax}
-    \DeclareMathOperator{\sgn}{sgn}
-    \DeclareMathOperator{\diag}{diag\;}
 
-    \DeclarePairedDelimiter\bra{\langle}{\rvert}
-    \DeclarePairedDelimiter\ket{\lvert}{\rangle}
-    \DeclarePairedDelimiter\abs{\lvert}{\rvert}
-    \DeclarePairedDelimiter\ev{\langle}{\rangle}
-    \DeclarePairedDelimiter\p{\lparen}{\rparen}
-    \DeclarePairedDelimiter\s{\lbrack}{\rbrack}
-    \DeclarePairedDelimiter\z{\lbrace}{\rbrace}
-
-    % \everymath{\displaystyle} % biggify limits of inline sums and integrals
-    \tikzstyle{circ} % usage: \node[circ, placement] (label) {text};
-        = [draw, circle, fill=white, node distance=3cm, minimum height=2em]
-    \definecolor{commentgreen}{rgb}{0,0.6,0}
-    \lstset{
-        basicstyle=\ttfamily\footnotesize,
-        frame=single,
-        numbers=left,
-        showstringspaces=false,
-        keywordstyle=\color{blue},
-        stringstyle=\color{purple},
-        commentstyle=\color{commentgreen},
-        morecomment=[l][\color{magenta}]{\#}
+\title[IGW Breaking in WDs]{Tidal Heating via Internal Gravity Waves in White Dwarf Binaries}
+\author[Y. Su et\ al.]{
+Yubo Su,$^1$,
+Daniel Lecoanet,$^2$
+Dong Lai,$^1$
+\\
+$^1,^3$ ABC
+$^2$ DEF
 }
 
-\begin{document}
+\date{Accepted XXX\@. Received YYY\@; in original form ZZZ}
 
-\def\Snospace~{\S{}} % hack to remove the space left after autorefs
-\renewcommand*{\sectionautorefname}{\Snospace}
-\renewcommand*{\appendixautorefname}{\Snospace}
-\renewcommand*{\figureautorefname}{Fig.}
-\renewcommand*{\equationautorefname}{Eq.}
-\renewcommand*{\tableautorefname}{Tab.}
+\pubyear{2019}
+
+\begin{document}\label{firstpage}
+\pagerange{\pageref{firstpage}--\pageref{lastpage}}
+\maketitle
 
@@
 
-\title{Tidal Heating via Internal Gravity Waves in White Dwarf Binaries}
-\author{Authors}
-\date{\today}
-\maketitle
+\begin{keywords}
+Wave Breaking, Dynamical Tides, Internal Gravity Waves
+\end{keywords}
 
@@
 
+Discussion foo
+
+\bibliographystyle{mnras}
+\bibliography{paper}
+
 \clearpage
-\onecolumngrid
+\onecolumn
 \appendix
@@
 \end{subequations}
-
-\bibliographystyle{apsrev4-1} % chktex 8
-\bibliography{paper}
+\label{lastpage} % chktex 24
 \end{document}
-
-% nonlinear wave breaking, tidal synchronization, dynamical tides, internal
-% gravity waves

### 2019-04-16 sims/2d_3_final/strat.py

@@ -36,7 +36,7 @@ PARAMS_DEFAULT = {'XMAX': XMAX,
 
                   'NUM_SNAPSHOTS': NUM_SNAPSHOTS,
                   'NL': True,
-                  'mask': False}
+                  'mask': True}
 
 def get_params(overrides=None):
     params = {**PARAMS_DEFAULT, **(overrides or {})}
@@ -67,9 +67,15 @@ def get_params(overrides=None):
 
 if __name__ == '__main__':
     tasks = [
+       ('lin_0_masked',
+        get_params(overrides={'F_MULT': 0.0005,
+                              'T_F': 4000,
+                              'DT': 0.1,
+                              'N_X': 64,
+                              'N_Z': 256,
+                              'Re_inv': 1e-4})),
        ('lin_1_masked',
         get_params(overrides={'F_MULT': 0.01,
-                              'mask': True,
                               'N_X': 64,
                               'N_Z': 256,
                               'Re_inv': 0.075})),
@@ -83,20 +89,16 @@ if __name__ == '__main__':
                               'Re_inv': 0})),
        ('nl_1_masked',
         get_params(overrides={'F_MULT': 1,
-                              'mask': True,
                               'Re_inv': 0.7})),
        ('nl_3_masked',
         get_params(overrides={'F_MULT': 1,
-                              'mask': True,
                               'Re_inv': 0.5})),
        ('nl_4_masked',
         get_params(overrides={'F_MULT': 1,
-                              'mask': True,
                               'NL': True,
                               'Re_inv': 0.3})),
        ('nl_6_masked',
         get_params(overrides={'F_MULT': 1,
-                              'mask': True,
                               'Re_inv': 0.2})),
     ]
     if '-plot' in sys.argv:

### 2019-04-16 sims/2d_4_fourier/strat_helper.py

@@ -1121,14 +1121,14 @@ def plot_front(name, params):
         ax1.plot(t,
                  smooth(S_px0[start_idx: ] / flux_th),
                  '%s-' % PLT_COLORS[color_idx],
-                 label=r'$\Delta S_{px,0}|_{z=z_0}$',
+                 label=r'$S(z=z_0)$',
                  linewidth=0.7)
         color_idx += 1
-        ax1.plot(t,
-                 smooth(dSpx0),
-                 '%s-' % PLT_COLORS[color_idx],
-                 label=r'$\Delta S_{px,0}|_{z=z_{c}}$',
-                 linewidth=0.7)
+        # ax1.plot(t,
+        #          smooth(dSpx0),
+        #          '%s-' % PLT_COLORS[color_idx],
+        #          label=r'$\Delta S_{0}|_{z=z_{c}}$',
+        #          linewidth=0.7)
         color_idx += 1
         ax1.plot(t,
                  smooth(-dSpx[start_idx: ] / flux_th),
@@ -1136,7 +1136,7 @@ def plot_front(name, params):
                  label=r'$\Delta S(z_{c})$',
                  linewidth=0.7)
         color_idx += 1
-        ax1.set_ylabel(r'$S / S_{px, 0}$')
+        ax1.set_ylabel(r'$S / S_0$')
         ax1.legend(fontsize=6, loc='lower right')
 
         # compare forecasts of front position using two predictors integrated
@@ -1167,7 +1167,7 @@ def plot_front(name, params):
         ax2.plot(t,
                  pos_anal,
                  '%s:' % PLT_COLORS[color_idx],
-                 label='Avg Absorbed $%.2fS_{px,0}$' %
+                 label='Avg Absorbed $%.2fS_{0}$' %
                     (mean_incident / flux_th),
                  linewidth=0.7)
         color_idx += 1

### 2019-04-16 writeups/paper.tex
stash paper writeup

@@
 
-\title[IGW Breaking in WDs]{Tidal Heating via Internal Gravity Waves in White Dwarf Binaries}
+\title[Internal Gravity Wave Breaking in White Dwarf Binaries]{Internal Gravity Wave Breaking in White Dwarf Binaries}
 \author[Y. Su et\ al.]{
@@
 \\
-$^1,^3$ ABC
-$^2$ DEF
+$^1$ Cornell Center for Astrophysics and Planetary Science, Department of
+Astronomy, Cornell University, Ithaca, NY 14853, USA
+\\
+$^2$ Princeton Center for Theoretical Science, Princeton University, Princeton,
+NJ 08544, USA
 }
@@
 \pagerange{\pageref{firstpage}--\pageref{lastpage}}
+\renewcommand*{\sectionautorefname}{Section}
 \maketitle
@@
 \begin{keywords}
-Wave Breaking, Dynamical Tides, Internal Gravity Waves
+white dwarfs -- hydrodynamics -- binaries:close -- waves % chktex 8
 \end{keywords}
@@
 the most important sources of gravitational waves (GWs) for the Laser
-Interferometric Space Antenna (LISA)\cite{lisa}. They are also thought to
+Interferometric Space Antenna (LISA) \citep{lisa}. They are also thought to
 produce interesting optical transients such as underluminous
-supernovae\cite{underlum}, Ca-rich fast transients\cite{carich}, and tidal
-novae\cite{tidal_novae}. Most importantly, they have been proposed as the likely
-progenitors of type Ia supernovae (e.g.~\cite{Ia0,webbink} or more
-recently\cite{Ia1,Ia2}). While presently only a few tens of compact WD binaries
-are known\cite{lsst_wd}, \emph{Gaia} (currently gathering data) is expected to
-expand the catalog to a few hundreds\cite{lsst_wd} (results based on
+supernovae \citep{underlum}, Ca-rich fast transients \citep{carich}, and tidal
+novae \citep{tidal_novae}. Most importantly, they have been proposed as the likely
+progenitors of type Ia supernovae (e.g. \citep{Ia0,webbink} or more
+recently \citep{Ia1,Ia2}). While presently only a few tens of compact WD binaries
+are known \citep{lsst_wd}, \emph{Gaia} (currently gathering data) is expected to
+expand the catalog to a few hundreds \citep{lsst_wd} (results based on
 \emph{Gaia}'s second data release have already begun to
-appear\cite{gaiaDD,gaiaDD2}), and the Large Synoptic Survey Telescope (LSST,
+appear \citep{gaiaDD,gaiaDD2}), and the Large Synoptic Survey Telescope (LSST,
 first light scheduled for 2020) will likely detect a few thousand
-more\cite{lsst_wd}. These observations will significantly advance the
+more \citep{lsst_wd}. These observations will significantly advance the
 understanding of WD binaries and their evolution. My proposed theoretical and
@@
 comes from our imprecise understanding of tidal interactions, which play an
-important role during a compact WD binary's inspiral\cite{fullerII}. Previous
+important role during a compact WD binary's inspiral \citep{fullerII}. Previous
 studies have shown that these interactions manifest as tidal excitation of
 internal gravity waves (IGW), waves in the WD fluid restored by the buoyancy
-force due to density stratification\cite{fullerI}. As these waves propagate
+force due to density stratification \citep{fullerI}. As these waves propagate
 outwards towards the WD surface, they grow in amplitude until they break, as do
 ocean waves on a shore, and transfer both energy and angular momentum from the
-binary orbit to the outer envelope of the WD\cite{fullerI,fullerII}.
+binary orbit to the outer envelope of the WD \citep{fullerI,fullerII}.
 
@@
 more energy than thermal radiation from the isolated WD surface and is thus a
-major contributor to the WD energy budget\cite{fullerII,fullerIV}. However,
+major contributor to the WD energy budget \citep{fullerII,fullerIV}. However,
 these works parameterized the wave breaking process in an ad hoc manner. The
@@
 deep in the WD envelope causes an energy buildup that results in energetic
-flares\cite{tidal_novae}. Works in other fields based on numerical simulations
+flares \citep{tidal_novae}. Works in other fields based on numerical simulations
 show that strongly nonlinear wave breaking behaves differently than predictions
-based in linear and weakly nonlinear theory\cite{winters1994,barker_ogilvie}.
+based in linear and weakly nonlinear theory \citep{winters1994,barker_ogilvie}.
 Such fully nonlinear numerical simulations have not been performed for WDs.
@@
     \begin{align}
-        \vec{\nabla} \cdot \vec{u} &= 0,\\
+        \vec{\nabla} \cdot \vec{u} &= 0,\label{eq:nl_incomp}\\
         \md{\rho}{t} &= 0 ,\label{eq:nl_density}\\
@@
 \end{subequations}
-$\md{}{t} = \pd{}{t} + \p*{\vec{u} \cdot \vec{\nabla}}$ is the Lagrangian or
+$\md{}{t} = \pd{}{t} + \p{\vec{u} \cdot \vec{\nabla}}$ is the Lagrangian or
 material derivative. $\vec{u}, \rho, p$ denote the velocity field, density and
@@
 
-\section{Internal Gravity Wave Theory}\label{s:theory}
+\section{Internal Gravity Waves: Theory}\label{s:theory}
 
@@
         \pd{\vec{u}}{t} + \vec{\nabla}P_1
-            + \p*{1 + \Upsilon_1} g\hat{z}
+            + \p{1 + \Upsilon_1} g\hat{z}
             &= 0.
@@
 This can be solved to obtain solutions up to undetermined amplitude $A$ the
-well-known result\cite{drazin,sutherland0}:
+well-known result \citep{drazin,sutherland0}:
 \begin{subequations}\label{se:lin_sol}
     \begin{align}
-        u_z\p*{x, z, t} &= Ae^{z/2H}e^{i(k_{0x}x + k_{0z}z - \omega_0 t)},\\
+        u_z\p{x, z, t} &= Ae^{z/2H}e^{i(k_{0x}x + k_{0z}z - \omega_0 t)},\\
         \omega_0^2 &= \frac{N^2k_{0x}^2}{k_{0x}^2 + k_{0z}^2 + \frac{1}{4H^2}},
@@
 \begin{equation}
-    N^2 \equiv g^2\p*{\rd{\rho}{P} - \at{\pd{\rho}{P}}_{ad}}
+    N^2 \equiv g^2\p{\rd{\rho}{P} - \at{\pd{\rho}{P}}_{ad}}
         = \frac{g}{H},
@@
             \vec{c}_{ph} &=
-                \p*{\frac{\hat{x}}{k_{0x}} + \frac{\hat{z}}{k_{0z}}}\omega,\\
-            \vec{c}_{g} &= \frac{N\p*{k_{0x}k_{0z}\hat{x}
-                - \p*{k_{0z}^2 + \frac{1}{4H^2}}\hat{z}}}
-                {\p*{k_{0x}^2 + k_{0z}^2 + \frac{1}{4H^2}}^{3/2}} \nonumber\\
-                &\approx \frac{\omega k_{0z}}{\abs*{\vec{k}_0}^{3/2}}
-                    \p*{k_{0x}\hat{x} - k_{0z}\hat{z}}.
+                \p{\frac{\hat{x}}{k_{0x}} + \frac{\hat{z}}{k_{0z}}}\omega,\\
+            \vec{c}_{g} &= \frac{N\p{k_{0x}k_{0z}\hat{x}
+                - \p{k_{0z}^2 + \frac{1}{4H^2}}\hat{z}}}
+                {\p{k_{0x}^2 + k_{0z}^2 + \frac{1}{4H^2}}^{3/2}} \nonumber\\
+                &\approx \frac{\omega k_{0z}}{\abs{\vec{k}_0}^{3/2}}
+                    \p{k_{0x}\hat{x} - k_{0z}\hat{z}}.
         \end{align}
         We recover the usual result $\vec{c}_{ph} \cdot \vec{c}_g = 0$
-        (e.g.\ \cite{drazin,sutherland1}). Note that an IGW transporting energy
+        (e.g.\  \citep{drazin,sutherland1}). Note that an IGW transporting energy
         and momentum upwards $c_{g, z} > 0$ has $k_{0z} < 0$.
@@
         \begin{align}
-            S_{px, z} &\equiv \ev*{\rho u_x u_z}_x \equiv
-                \frac{1}{L_x}\int_0^{L_x}\limits \rho u_x u_z\;\mathrm{d}x
-                    \nonumber\\
+            S &\equiv \ev{\rho u_x u_z}_x \equiv
+                \frac{1}{L_x}\int_0^{L_x}\limits \rho u_x u_z\;\mathrm{d}x,
+                    \label{eq:S_def},\\
                 &\approx -\frac{A^2}{2}\rho_0(z=0)\frac{k_{0z}}{k_{0x}},
-                    \label{eq:fpx_lin}
+                    \label{eq:S_lin}
         \end{align}
-        where $\ev*{\dots}_x$ denotes $x$-averaging.
+        where $\ev{\dots}_x$ denotes $x$-averaging.
 \end{itemize}
@@
         = Fe^{-\frac{(z - z_0)^2}{2\sigma^2}}
-            \cos \p*{k_{0x}x - \omega_0 t}.\label{eq:vol_drive}
+            \cos \p{k_{0x}x - \omega_0 t}.\label{eq:vol_drive}
 \end{equation}
@@
         &{}\begin{cases}
-        e^{\frac{z - z_0}{2H}}e^{i\p*{k_{0x}x + k_{0z}(z - z_0) - \omega_0 t
+        e^{\frac{z - z_0}{2H}}e^{i\p{k_{0x}x + k_{0z}(z - z_0) - \omega_0 t
             + \frac{1}{2k_{0z}H}}}
             & z > z_0\\
-        e^{\frac{z - z_0}{2H}}e^{i\p*{k_{0x}x - k_{0z}(z - z_0) - \omega_0 t
+        e^{\frac{z - z_0}{2H}}e^{i\p{k_{0x}x - k_{0z}(z - z_0) - \omega_0 t
             + \frac{1}{2k_{0z}H}}}
@@
 The $z > z_0$ solution models an upwards-propagating IGW wavetrain inbound on
-the simulation domain from below. The solution for $u_{1x}(x, z, t)$ is
-correspondingly $u_{1x} \approx \frac{k_{0z}
-u_{1z}}{k_{0x}}e^{\frac{i}{2k_{0z}H}}$.
+the simulation domain from below. The solution for $u_{1x}(x, z, t)$ can be
+obtained analytically from \autoref{eq:nl_incomp}.
 
@@
 In developing the results above, we have neglected advective terms $\pd{}{t} \gg
-\vec{u} \cdot \vec{\nabla}$. However, since $\abs*{\vec{u}} \propto e^{z/2H}$,
+\vec{u} \cdot \vec{\nabla}$. However, since $\abs{\vec{u}} \propto e^{z/2H}$,
 any nonzero amplitude IGW will eventually violate the linearity criterion. This
 regime is relevant in practice since the IGW may grow to nonlinear amplitudes
-before it reaches the surface of the WD\cite{fullerI,fullerII}. We can coarsely
+before it reaches the surface of the WD \citep{fullerI,fullerII}. We can coarsely
 estimate the height at which this happens by when the Lagrangian displacement
@@
 the waves into the fluid; such a process has been conjectured to be responsible
-for tidal synchronization in stellar binaries\cite{zahn75,gn89} as well as the
-quasi-biennial oscillation\cite{lindzen_qbo}. The details of this process have
+for tidal synchronization in stellar binaries \citep{zahn75,gn89} as well as the
+quasi-biennial oscillation \citep{lindzen_qbo}. The details of this process have
 been laid out in a few key papers: IGWs are globally unstable to resonant
-three-wave interactions\cite{drazin}, causing energy transfer out of the IGW to
-daughter modes. These steeper daughter modes facilitate wave breaking,
-depositing horizontal momentum in the fluid's mean flow\cite{klostermeyer}. Once
-the mean flow reaches the horizontal phase velocity of the parent IGW, a
+three-wave interactions \citep{drazin}, causing energy transfer out of the IGW
+to daughter modes. These steeper daughter modes facilitate wave breaking,
+depositing horizontal momentum in the fluid's mean flow \citep{klostermeyer}.
+Once the mean flow reaches the horizontal phase velocity of the parent IGW, a
 critical layer forms at which the frequency of the IGW is Doppler-shifted to
 zero. The interaction of the parent IGW with the mean flow was first studied in
-the inviscid, linear regime in~\cite{booker_bretherton}, which found nearly
-complete absorption. This result was reproduced with nonzero
-viscosity\cite{hazel}, but weakly nonlinear theory\cite{brown_stewartson} and
-fully nonlinear simulations\cite{winters1994} suggest that nonlinear effects can
+the inviscid, linear regime in \citep{booker_bretherton}, which found nearly
+complete absorption. This result was reproduced with nonzero viscosity
+\citep{hazel}, but weakly nonlinear theory \citep{brown_stewartson} and fully
+nonlinear simulations \citep{winters1994} suggest that nonlinear effects can
 induce reflection.
@@
 An estimate of where these mean flow effects become important can be made using
-the wave-induced mean flow\cite{eliassen_palm_cite,sutherland0}:
+the wave-induced mean flow \citep{eliassen_palm_cite,sutherland0}:
 \begin{equation}
-     \ev*{u_x} = \bar{U}_0 \equiv \frac{1}{L_x}
-        \int\limits u_x\;\mathrm{d}x = \frac{\ev*{u_xu_z}_x}{c_{g,z}},
+     \ev{u_x} = \bar{U}_0 \equiv \frac{1}{L_x}
+        \int\limits u_x\;\mathrm{d}x = \frac{\ev{u_xu_z}_x}{c_{g,z}},
+        \label{eq:mean_flow}
 \end{equation}
-where $\ev*{\dots}_x$ denotes horizontal averaging. Setting $\bar{U}_0 = c_{ph,
+where $\ev{\dots}_x$ denotes horizontal averaging. Setting $\bar{U}_0 = c_{ph,
 x}$ the critical layer condition, we recover $\frac{u_xk_x}{\omega} = \xi_x k_x
@@
 
-\section{Internal Gravity Wave Numerical Simulation}\label{s:sim}
+\section{Internal Gravity Waves: Linear Numerical Simulation}\label{s:sim}
 
-To study IGW breaking as a weak perturbation grows to nonlinear amplitudes, we
-perform direct numerical simulation using Dedalus\cite{dedalus} a spectral
-numerical method. Our simulation domain is smaller than our full physical
-domain, so we adopt reflection-suppressing boundary conditions. We use periodic
-boundary conditions in the $x$ direction and damping layers (described in
-\autoref{ss:damping}) in the $z$ direction to damp perturbations exiting the
-top and bottom of the domain. We use a Fourier basis in the $x$ direction and
-either a Chebyshev or Fourier basis in the $z$ direction; our results were
-robust to either set of basis functions. We varied the number of $x, z$ modes
-(denoted $N_x, N_z$ respectively), and used $3/2$ dealiasing\cite{boyd}.
+Towards numerical simulation of IGW breaking, we first verify agreement with
+linear theory at weak forcing amplitudes. We perform direct numerical simulation
+using the pseudo-spectral code Dedalus \citep{dedalus}. In \autoref{ss:numerics}
+we discuss choices of numerical parameters, in \autoref{ss:damping} we
+separately discuss use of a damping zone at the top and bottom of the domain,
+and in \autoref{ss:lin_ns} we present the results of our ``linear'' simulations,
+where we solve the full nonlinear fluid equations for weak forcing.
 
-The time integration uses a split implicit-explicit first-order scheme where
-certain terms are treated implicitly using for- ward Euler, and the remaining
-terms are treated explicitly using backward Euler. A third-order, four-stage
-DIRK-ERK scheme\cite{ascher} is used with adaptive timesteps taken to be
-the advective Courant-Friedrichs-Lewy (CFL) time given by $0.7 \min (\Delta
-x / u_x,\Delta z / u_z)$. The minimum is taken over every grid point in
-the domain and $\Delta x,\Delta z$ are the grid spacings in the $x$ and $z$
-directions respectively.
-
-\subsection{Numerical Parameters}\label{ss:numerics}
+\subsection{Numerical Setup}\label{ss:numerics}
 
@@
 
-Denote $L_x, L_z$ to be the physical dimensions of the simulation domain. The
-physics of our simulation is fixed by six parameters: $L_x, L_z$, and $k_{0x},
-\omega_0, F, \nu$. We describe and choose these parameters below:
+Denote $L_x, L_z$ to be the physical dimensions of the simulation domain. Our
+simulation domain is smaller than our full physical domain, so we adopt
+reflection-suppressing boundary conditions. We use periodic boundary conditions
+in the $x$ direction and damping layers (described in \autoref{ss:damping}) in
+the $z$ direction to damp perturbations exiting the top and bottom of the
+domain. We use a Fourier basis in the $x$ direction and either a Chebyshev or
+Fourier basis in the $z$ direction; our results were robust to either set of
+basis functions. We varied the number of $x, z$ modes (denoted $N_x, N_z$
+respectively), and used $3/2$ dealiasing \citep{boyd}.
+
+The time integration uses a split implicit-explicit third-order scheme where
+certain terms are treated implicitly and the remaining terms are treated
+explicitly. A third-order, four-stage DIRK-ERK scheme \citep{ascher} is used
+with adaptive timesteps computed from advective Courant-Friedrichs-Lewy (CFL)
+time. Specifically, we use $\Delta t = 0.7 \min(\Delta x / u_x,\Delta z / u_z)$,
+where the minimum is taken over every grid point in the domain and $\Delta
+x,\Delta z$ are the grid spacings in the $x$ and $z$ directions respectively.
+
+The physics of our simulation is fixed by six parameters: $L_x, L_z$, and
+$k_{0x}, \omega_0, F, \nu$. We describe our choices for these parameters below:
 \begin{itemize}
@@
     \item $\omega_0$: We choose $\omega_0$ by evaluating dispersion relation
-        $\omega_0\p*{k_{0x}, k_{0z}}$ for some desired $k_{0z}$.
+        $\omega_0\p{k_{0x}, k_{0z}}$ for some desired $k_{0z}$.
 
@@
         becomes a constraint on $F$ by using \autoref{eq:uz_lin} and
-        $\abs*{\vec{u}} \propto e^{z/2H}$.
+        $\abs{\vec{u}} \propto e^{z/2H}$.
 
@@
         scales in the absence of artificial dissipation. We introduce
-        Navier-Stokes viscosity $\nu$ while ensuring our equations of motion
-        continue to conserve horizontal momentum (see \autoref{ss:strat_impl}
-        for details). When $\vec{\xi} \cdot \vec{k} \ll 1$, we may set $\nu =
-        0$.
+        dissipation parameter $\nu$ used for both artificial viscosity and artificial
+        diffusivity. We ensure our equations of motion continue to conserve
+        horizontal momentum (see \autoref{ss:strat_impl} for details). Under
+        weak forcing, we may set $\nu = 0$, since the energy cascade is
+        negligible.
 \end{itemize}
@@
         - \Gamma(z)\vec{u},\nonumber\\
-    \frac{1}{2\tau}\s*{2 + \tanh \frac{z - z_T}{\Delta z}
+    \frac{1}{2\tau}\s{2 + \tanh \frac{z - z_T}{\Delta z}
         + \tanh \frac{z_B - z}{\Delta z}} &= \Gamma(z),\label{eq:Gamma}
@@
 governed by $\Delta z$; we use $\Delta z = 0.025L_z$. This prescription, used in
-similar studies\cite{lecoanet_damp}, has the advantage of being smooth,
+similar studies \citep{lecoanet_damp}, has the advantage of being smooth,
 important for spectral methods. The full implementation of the fluid equations
@@
 
-\subsection{Linear Regime}\label{ss:lin_ns}
+\subsection{Simulation Results}\label{ss:lin_ns}
 
-We first test our numerical simulation with a ``linear'' simulation, where
-$F$ is chosen such that $\vec{\xi} \cdot \vec{k} \ll 1$ everywhere. We emphasize
-that our ``linear'' simulations solve the full nonlinear equations, just in the
-weak perturbation limit.
+We present the results of a simulation satisfying the parameter choices in
+\autoref{ss:numerics} such that $\vec{\xi} \cdot \vec{k} \ll 1$ everywhere.
+\begin{figure}[t]
+    \centering
+    \includegraphics[width=\columnwidth]{plots/lin_amps.png}
+    \caption{Amplitude of excited wave over time in weak forcing simulation,
+    in units of the analytical estimate \autoref{eq:uz_lin}. We observe great
+    agreement with the linear calculation.}\label{fig:lin_amps}
+\end{figure}
+\begin{figure}[t]
+    \centering
+    \includegraphics[width=\columnwidth]{plots/lin_fluxes.png}
+    \caption{Horizontal momentum flux $S(z, t)$ in units of the analytical estimator
+    given by \autoref{eq:S_lin} as a function of $z$ at select times $t$. As
+    initial transients die out, the flux converges to be flat below the damping
+    layer. The flux excited in the forcing zone is transported without loss to
+    the top of the domain, where it is dissipated by the damping layer
+    (see \autoref{ss:damping}) without reflection.}\label{fig:lin_fluxes}
+\end{figure}
+
+\section{Internal Gravity Waves: Nonlinear Simulation}\label{s:sim}
 
@@
 regularize the nonlinear cascade at near grid resolution: $\nu \sim 0.1
-\frac{\omega_0}{\abs*{k_{0z}}}\frac{L_z}{2\pi N_z}$ was found to be suitable for
+\frac{\omega_0}{\abs{k_{0z}}}\frac{L_z}{2\pi N_z}$ was found to be suitable for
 $N_z = 2048$.
@@
 where we have defined Richardson number $\mathrm{Ri} \equiv
-\at{\frac{N^2}{\p*{\pd{\bar{U}_0}{z}}^2}}_{z_c}$ at the critical layer $z_c:
+\at{\frac{N^2}{\p{\pd{\bar{U}_0}{z}}^2}}_{z_c}$ at the critical layer $z_c:
 \bar{U}_0(z_c) = \frac{\omega_0}{k_{0x}}$. For most shear flows, $\mathrm{Ri}
@@
     \pd{}{t}\int\limits_0^{L_z} \rho(z) \bar{U}_0(z, t)\;\mathrm{d}z
-        - S_{px} = 0.
+        - \Delta S = 0.
 \end{equation}
-Treating $\bar{U}_0(z > z_c) = \bar{U}_c, \bar{U}(z < z_c) = 0$ gives us exactly
+We define $\Delta S$ to be the change in flux across the critical layer, or
+equivalently the absorbed flux. Treating $\bar{U}_0(z > z_c) = \bar{U}_c,
+\bar{U}(z < z_c) = 0$ gives us exactly
 \begin{equation}
-    \rho \bar{U}_c\pd{z_c}{t} = -S_{px}.\label{eq:zc_anal}
+    \rho \bar{U}_c\pd{z_c}{t} = -\Delta S.\label{eq:zc_anal}
 \end{equation}
-For constant $S_{px}$ in space and $\rho \approx \rho_0$, this has analytical
+For constant $\Delta S$ in space and $\rho \approx \rho_0$, this has analytical
 solution
 \begin{equation}
-    z(t) = -H\ln t - H\ln \frac{H\rho_0(z = 0)c_{ph, x}}{S_{px}}.
+    z(t) = -H\ln t - H\ln \frac{H\rho_0(z = 0)c_{ph, x}}{\Delta S}
+        .\label{eq:zc_sol}
 \end{equation}
@@
 \begin{itemize}
-    \item We choose $\nu = 0.35 \frac{\omega_0}{k_{0z}k_{z, \max}}$, where
+    \item We choose $\nu = 0.1 \frac{\omega_0}{k_{0z}k_{z, \max}}$, where
         $k_{z, \max} = \frac{2\pi N_z}{L_z}$. Note that $\nu =
@@
 \pd{\bar{U}_0}{z}$. Computing $\pd{z_c}{t}$ using the analytical flux
-\autoref{eq:fpx_lin} allows comparison to \autoref{eq:zc_anal}, which we exhibit
+\autoref{eq:S_lin} allows comparison to \autoref{eq:zc_anal}, which we exhibit
 in TODO\@. We see overall good agreement, though some small
@@
 
+\begin{figure}[t]
+    \centering
+    \includegraphics[width=\columnwidth]{plots/nl_f_amps.png}
+    \caption{The top panel measures the incident wave amplitude $A_i(t)$ (green)
+    and the downwards propagating wave amplitude $A_d(t)$ (red) just above the
+    forcing zone, normalized to the analytical estimate \autoref{eq:uz_lin}.
+    $A_d \neq 0$ due to reflection off the critical layer. The bottom panel
+    shows the behavior of three horizontal momentum fluxes over time, in units
+    of the analytical estimate \autoref{eq:S_lin}: (blue) flux incident on the
+    critical layer, (black) flux absorbed by the critical layer, and (green)
+    flux transmitted through the critical layer. A fiducial simulation of
+    $\mathrm{Re} = 1000$ is shown.}\label{fig:nl_amps}
+\end{figure}
+
+\begin{figure}[t]
+    \centering
+    \includegraphics[width=\columnwidth]{plots/nl_fluxes.png}
+    \caption{The top and bottom panels show respectively the mean flow
+    $\bar{U}_0$ (see \autoref{eq:mean_flow}) and horizontal momentum flux $S$
+    (see \autoref{eq:S_def}) as a function of $z$ at certain times $t$. The same
+    color coding is used for the two plots. The mean flow is normalized to its
+    critical value $c_{ph,x}$ while again $S$ is normalized to its analytical
+    estimate \autoref{eq:S_lin}.}\label{fig:nl_fluxes}
+\end{figure}
+
+\begin{figure}[t]
+    \centering
+    \includegraphics[width=\columnwidth]{plots/nl_front.png}
+    \caption{Shown is the position of the critical layer (black) as well as two
+    predictors of the critical layer location using the absorbed flux (see the
+    second panel of \autoref{fig:nl_amps}). The green line is generated by
+    integrating \autoref{eq:zc_anal} using $\Delta S$ from the data, while the
+    blue line is generated by using the average value of $\Delta S$ in
+    \autoref{eq:zc_sol}.}\label{fig:nl_front}
+\end{figure}
+
+\begin{figure}[t]
+    \centering
+    \includegraphics[width=\columnwidth]{plots/nl_f_refl.png}
+    \caption{NL reflectivity}\label{fig:nl_refl}
+\end{figure}
+
 \section{Discussion}\label{s:discussion}
 
+\begin{figure}[t]
+    \centering
+    \includegraphics[width=\columnwidth]{plots/agg.png}
+    \caption{Convergence w/ all runs}\label{fig:agg}
+\end{figure}
+
+\subsection{Applicability to Astrophysical Systems}
+
+% magnetic winding, radiative diffusivity + effective Re. Might be applicable to
+% inwards focusing too
+
 Discussion foo
+
+\section{Acknowledgements}\label{s:ack}
 
@@
         \vec{\nabla} \cdot \vec{u} &= 0,\\
-        \pd{\Upsilon}{t} + \p*{\vec{u} \cdot \vec{\nabla}} \Upsilon
+        \pd{\Upsilon}{t} + \p{\vec{u} \cdot \vec{\nabla}} \Upsilon
             - \frac{u_z}{H} &= 0,\\
-        \pd{u_x}{t} + \p*{\vec{u} \cdot \vec{\nabla}}u_x
+        \pd{u_x}{t} + \p{\vec{u} \cdot \vec{\nabla}}u_x
             + \pd{\varpi_1}{x} + gH\pd{\Upsilon}{x}
             + \varpi_1 \pd{\Upsilon}{x} &= 0,\\
-        \pd{u_z}{t} + \p*{\vec{u} \cdot \vec{\nabla}}u_z
+        \pd{u_z}{t} + \p{\vec{u} \cdot \vec{\nabla}}u_z
             + \pd{\varpi_1}{z} + gH\pd{\Upsilon}{z}
@@
         \vec{\nabla} \cdot \vec{u} &= 0,\\
-        \partial_t \Upsilon + \p*{\vec{u} \cdot \vec{\nabla}} \Upsilon -
-            \frac{u_z}{H} - \nu\p*{\nabla^2 \Upsilon + \p*{\vec{\nabla}
-            \Upsilon} \cdot \p*{\vec{\nabla}\Upsilon} - \frac{2}{H}\partial_z
+        \partial_t \Upsilon + \p{\vec{u} \cdot \vec{\nabla}} \Upsilon -
+            \frac{u_z}{H} - \nu\p{\nabla^2 \Upsilon + \p{\vec{\nabla}
+            \Upsilon} \cdot \p{\vec{\nabla}\Upsilon} - \frac{2}{H}\partial_z
             \Upsilon + \frac{1 - e^{-\Upsilon}}{H^2}} &= 0,\\
-        \partial_t \vec{u} + \p*{\vec{u} \cdot \vec{\nabla}}\vec{u} +
+        \partial_t \vec{u} + \p{\vec{u} \cdot \vec{\nabla}}\vec{u} +
             \vec{\nabla} \varpi + \varpi \vec{\nabla} \Upsilon - \nu \nabla^2
-            \vec{u} + \vec{u} \nu\p*{\nabla^2 \Upsilon + \p*{\vec{\nabla}
-            \Upsilon} \cdot \p*{\vec{\nabla}\Upsilon} - \frac{2}{H}\partial_z
+            \vec{u} + \vec{u} \nu\p{\nabla^2 \Upsilon + \p{\vec{\nabla}
+            \Upsilon} \cdot \p{\vec{\nabla}\Upsilon} - \frac{2}{H}\partial_z
             \Upsilon + \frac{1 - e^{-\Upsilon}}{H^2}}&{}\nonumber\\
-        - 2\nu \p*{\p*{\p*{\vec{\nabla}\Upsilon} \cdot \vec{\nabla}}\vec{u} -
+        - 2\nu \p{\p{\p{\vec{\nabla}\Upsilon} \cdot \vec{\nabla}}\vec{u} -
             \frac{1}{H}\partial_z \vec{u}} - \frac{\varpi_1}{H} &= 0.
@@
         \partial_t \Upsilon - \frac{u_z}{H}
-            ={}& \nu\p*{\nabla^2 \Upsilon + \p*{\vec{\nabla}
-            \Upsilon} \cdot \p*{\vec{\nabla}\Upsilon} - \frac{2}{H}\partial_z
+            ={}& \nu\p{\nabla^2 \Upsilon + \p{\vec{\nabla}
+            \Upsilon} \cdot \p{\vec{\nabla}\Upsilon} - \frac{2}{H}\partial_z
             \Upsilon + \frac{1 - e^{-\Upsilon}}{H^2}},\nonumber\\
-            & - \p*{\vec{u} \cdot \vec{\nabla}}\Upsilon
+            & - \p{\vec{u} \cdot \vec{\nabla}}\Upsilon
                 -\Gamma(z) \Upsilon
                 + \frac{F}{\rho_0(z)}e^{-\frac{(z - z_0)^2}{2\sigma^2}}
-                    \cos \p*{k_xx - \omega t},\\
+                    \cos \p{k_xx - \omega t},\\
         \pd{u_x}{t} + \pd{T}{x} + gH\pd{\Upsilon}{x} ={}&
             \nu \nabla^2 u_x
-            - u_x \nu\p*{\nabla^2 \Upsilon + \p*{\vec{\nabla} \Upsilon} \cdot
-                \p*{\vec{\nabla}\Upsilon} - \frac{2}{H}\partial_z \Upsilon
+            - u_x \nu\p{\nabla^2 \Upsilon + \p{\vec{\nabla} \Upsilon} \cdot
+                \p{\vec{\nabla}\Upsilon} - \frac{2}{H}\partial_z \Upsilon
                 + \frac{1 - e^{-\Upsilon}}{H^2}}\nonumber\\
-            &+ 2\nu \p*{\p*{\p*{\vec{\nabla}\Upsilon} \cdot \vec{\nabla}}u_x
+            &+ 2\nu \p{\p{\p{\vec{\nabla}\Upsilon} \cdot \vec{\nabla}}u_x
                 - \frac{1}{H}\partial_z u_x}
             -\Gamma(z) u_x
-                - \p*{\vec{u} \cdot \vec{\nabla}}u_x
+                - \p{\vec{u} \cdot \vec{\nabla}}u_x
                 - T_1 \pd{\Upsilon}{x},\\
@@
             \nu \nabla^2 u_z
-            - u_z \nu\p*{\nabla^2 \Upsilon + \p*{\vec{\nabla} \Upsilon} \cdot
-                \p*{\vec{\nabla}\Upsilon} - \frac{2}{H}\partial_z \Upsilon
+            - u_z \nu\p{\nabla^2 \Upsilon + \p{\vec{\nabla} \Upsilon} \cdot
+                \p{\vec{\nabla}\Upsilon} - \frac{2}{H}\partial_z \Upsilon
                 + \frac{1 - e^{-\Upsilon}}{H^2}}\nonumber\\
-            &+ 2\nu \p*{\p*{\p*{\vec{\nabla}\Upsilon} \cdot \vec{\nabla}}u_z -
+            &+ 2\nu \p{\p{\p{\vec{\nabla}\Upsilon} \cdot \vec{\nabla}}u_z -
                 \frac{1}{H}\partial_z u_{z}}
-            -\Gamma(z) u_z - \p*{\vec{u} \cdot \vec{\nabla}}u_z
+            -\Gamma(z) u_z - \p{\vec{u} \cdot \vec{\nabla}}u_z
             - T_1 \pd{\Upsilon}{z}.

### 2019-04-16 sims/2d_3_final/strat_helper.py

@@ -976,10 +976,10 @@ def plot_front(name, params):
                      S_px[time, z_b_idx: ] / flux_th,
                      linewidth=0.7,
                      label=r't=%.1f$N^{-1}$' % sim_times[time])
-        plt.plot(z0_cut,
-                 np.exp(-k_damp * 2 * (z0_cut - Z0)),
-                 linewidth=1.5,
-                 label=r'Model')
+        # plt.plot(z0_cut,
+        #          np.exp(-k_damp * 2 * (z0_cut - Z0)),
+        #          linewidth=1.5,
+        #          label=r'Model')
         plt.xlim(z_b, ZMAX)
         plt.ylim(-0.1, 1.1)
         plt.legend(fontsize=6)
@@ -1007,7 +1007,7 @@ def plot_front(name, params):
         #          linewidth=0.7)
         # ax1.legend(fontsize=6)
         ax1.set_xlabel(r'$t (N^{-1})$')
-        ax1.set_ylabel(r'$A$')
+        ax1.set_ylabel(r'$A/A_0$')
         # ax2.plot(t,
         #          np.unwrap(phis_down[start_idx: ]),
         #          'r',
@@ -1169,7 +1169,7 @@ def plot_front(name, params):
                  'r',
                  label=r'$A_d$',
                  linewidth=0.7)
-        ax1.set_ylabel(r'$A$')
+        ax1.set_ylabel(r'$A/A_0$')
         ax1.set_ylim(bottom=0)
         ax1.legend(loc=0, fontsize=6)
 

### 2019-04-16 sims/2d_4_fourier/strat.py

@@ -81,12 +81,12 @@ if __name__ == '__main__':
                                'N_Z': 3072})),
         ('yubo_nu2_hres',
          get_params(overrides={'Re_inv': 0.2})),
-        # ('yubo_nu2_width',
-        #  get_params(overrides={'Re_inv': 0.2,
-        #                        'N_X': 256,
-        #                        'NUM_SNAPSHOTS': 1500,
-        #                        'STRIDE': 2,
-        #                        'N_Z': 1024})),
+        ('yubo_nu2_width',
+         get_params(overrides={'Re_inv': 0.2,
+                               'N_X': 256,
+                               'NUM_SNAPSHOTS': 1500,
+                               'STRIDE': 2,
+                               'N_Z': 1024})),
         ('yubo_nu3_width',
          get_params(overrides={'Re_inv': 0.3,
                                'N_X': 256,

### 2019-04-16 sims/aggregate.py

@@ -19,7 +19,7 @@ DATA = [
     (0.05, ((0.4869, 0.3428, 0.5560), (0.5632, 0.4939, 0.5996), (0.1528, 0.0863, 0.1885), (0.3727, 0.3203, 0.4493))),
     (0.05, ((0.5169, 0.4096, 0.6131), (0.5571, 0.4914, 0.6078), (0.1089, 0.0768, 0.1456), (0.3716, 0.3074, 0.4428))),
     (0.10, ((0.4914, 0.2686, 0.5305), (0.5287, 0.3905, 0.6236), (0.1538, 0.1016, 0.1787), (0.3479, 0.2871, 0.4105))),
-    # (0.10, ((0.3876, 0.3201, 0.4097), (0.4540, 0.4292, 0.4783), (0.0858, 0.0651, 0.1261), (0.3646, 0.2847, 0.4419))),
+    (0.10, ((0.5220, 0.3271, 0.6194), (0.5603, 0.4294, 0.5899), (0.1274, 0.1000, 0.1805), (0.3627, 0.2888, 0.4312))),
     (0.20, ((0.4626, 0.4088, 0.5010), (0.5487, 0.4951, 0.5867), (0.0340, 0.0207, 0.0456), (0.4124, 0.2874, 0.5553))),
     (0.30, ((0.3475, 0.2300, 0.3747), (0.4441, 0.4382, 0.4462), (0.0022, 0.0009, 0.0024), (0.3467, 0.3050, 0.4099))),
 ]

### 2019-04-26 writeups/paper.tex
check in new abstract?

--- 2019-04-16 writeups/paper.tex
@@
 \begin{abstract}
-    In white dwarf binaries tidal torques are expected to raise a train of
-    internal gravity waves in the white dwarf interior. These
-    outward-propagating waves can then grow to be nonlinear and break via
-    hydrodynamical instabilities, transfering energy and angular momentum from
-    the binary orbit to the white dwarf. We perform 2D numerical simulations of
-    nonlinear wave breaking of outwards-propagating internal gravity waves in an
-    incompressible, isothermal atmosphere with an exponentially decaying density
-    stratification. We find that tidal synchronization, after an initial
-    transient phase, proceeds inwards from the surface in thin layers. We argue
-    the thickness of these layers is limited by the Kelvin-Helmholtz
-    Instability. We provide simple analytical formulae for the location and
-    thickness of tidal synchronization over time that are in good agreement with
-    our simulations.
+    In sufficiently compact white dwarf binaries, dynamical tides raise a train
+    of internal gravity waves that propagate towards the surface. We perform 2D
+    numerical simulations of these waves undergoing nonlinear wave breaking in
+    an incompressible, isothermal atmosphere. After an initial transient phase,
+    we find that these waves induce a sharp transition between a non-rotating
+    core and synchronously rotating envelope. We find evidence that the width of
+    this transition layer is bound from below by the Kelvin-Helmholtz
+    Instability. We provide analytical formulae for absorption and reflection of
+    incident waves off the critical layer up to prefactors of order unity. These
+    prefactors converge to constant values when artificial dissipation is
+    decreased. We provide dimensionless criteria necessary to resolving momentum
+    transfer within the critical layer. Finally, we speculate on the application
+    of our model to tidal synchronization and heating in astrophysical systems.
 \end{abstract}
 
-\title{Tidal Heating via Internal Gravity Waves in White Dwarf Binaries}
+\title{Nonlinear Dynamical Tides in White Dwarf Binaries}
 \author{Authors}
@@
 
-We consider a two-dimensional incompressible, isothermal fluid, representative
-of degenerate matter in WDs. We neglect temperature variations and assume a
+We consider an incompressible, isothermal fluid, representative of degenerate
+matter in WD bulks, in 2D. We neglect temperature variations and assume a
 barotropic equation of state as a first approximation. As we are interested in
-dynamics far from the center of the WD, we approximate the gravitational field
-as uniform. We model the background density stratification as $\rho_0(x, z) =
-\rho_0(z) = \rho_0(z=0) e^{-z/H}$ for some reference density $\rho_0(z=0) =
-\at{\rho_0(z)}_{z = 0}$.
+the behavior of the wave far from the center of the WD, we approximate the
+gravitational field as uniform. We model the background density stratification
+as $\rho_0(x, z) = \rho_0(z) = \rho_0(z=0) e^{-z/H}$ for some reference density
+$\rho_0(z=0) = \at{\rho_0(z)}_{z = 0}$.
 
@@
 \begin{itemize}
-    \item The amplitude of the wave grows like $e^{z/2H}$. Thus, the
+    \item The amplitude of the perturbations grows like $e^{z/2H}$. Thus, the
         linear approximation is violated for sufficiently large $z$.
@@
 the simulation domain from below. The solution for $u_{1x}(x, z, t)$ is
-correspondingly $u_{1x} \approx \frac{k_{0z}
-u_{1z}}{k_{0x}}e^{\frac{i}{2k_{0z}H}}$.
+correspondingly $u_{1x} \approx \frac{k_{0z} u_{1z}}{k_{0x}}e^{\frac{i}{2k_{0z}H}}$.
 
@@
 any nonzero amplitude IGW will eventually violate the linearity criterion. This
-regime is relevant in practice since the IGW may grow to nonlinear amplitudes
+regime is relevant in practice since the IGW grows to nonlinear amplitudes
 before it reaches the surface of the WD\cite{fullerI,fullerII}. We can coarsely
-estimate the height at which this happens by when the Lagrangian displacement
-$\vec{\xi}$ of a fluid parcel satisfies
-\begin{equation}
-    \vec{\xi} \cdot \vec{k} \gtrsim 1.\label{eq:nl}
-\end{equation}
+estimate the height at which this happens by $\abs*{\vec{u}}(z) \gtrsim
+\frac{\omega}{\abs*{\vec{k}}}$. This corresponds to when the Lagrangian
+displacement $\vec{\xi} \equiv \vec{u} / \omega$ satisfies $\vec{\xi} \cdot
+\vec{k} \gtrsim 1$, the same criterion used in\cite{fullerI}.
 
@@
 An estimate of where these mean flow effects become important can be made using
-the wave-induced mean flow\cite{eliassen_palm_cite,sutherland0}:
+the wave-induced mean flow formula\cite{eliassen_palm_cite,sutherland0}:
 \begin{equation}
@@
 \end{equation}
-where $\ev*{\dots}_x$ denotes horizontal averaging. Setting $\bar{U}_0 = c_{ph,
-x}$ the critical layer condition, we recover $\frac{u_xk_x}{\omega} = \xi_x k_x
-\sim 1$, roughly equivalent to the nonlinearity criterion \autoref{eq:nl}.
+where $\ev*{\dots}x$ denotes horizontal averaging. Setting $\bar{U}_0 = c_{ph,
+x}$ the critical layer condition, we can solve that $\vec{\xi} \cdot \vec{k}
+= \sqrt{\frac{k_x}{k_z}}$ is where the mean flow attains the horizontal phase
+velocity. In an astrophysical context, $N \gg \omega$ and $k_z \gg k_x$. Thus,
+wave breaking may occur significantly earlier than assumed in current
+astrophysics literature.
 
@@
 perform direct numerical simulation using Dedalus\cite{dedalus} a spectral
-numerical method. Our simulation domain is smaller than our full physical
-domain, so we adopt reflection-suppressing boundary conditions. We use periodic
-boundary conditions in the $x$ direction and damping layers (described in
-\autoref{ss:damping}) in the $z$ direction to damp perturbations exiting the
-top and bottom of the domain. We use a Fourier basis in the $x$ direction and
-either a Chebyshev or Fourier basis in the $z$ direction; our results were
-robust to either set of basis functions. We varied the number of $x, z$ modes
-(denoted $N_x, N_z$ respectively), and used $3/2$ dealiasing\cite{boyd}.
-
-The time integration uses a split implicit-explicit first-order scheme where
-certain terms are treated implicitly using for- ward Euler, and the remaining
-terms are treated explicitly using backward Euler. A third-order, four-stage
-DIRK-ERK scheme\cite{ascher} is used with adaptive timesteps taken to be
-the advective Courant-Friedrichs-Lewy (CFL) time given by $0.7 \min (\Delta
-x / u_x,\Delta z / u_z)$. The minimum is taken over every grid point in
-the domain and $\Delta x,\Delta z$ are the grid spacings in the $x$ and $z$
-directions respectively.
+numerical method. Our region of interest is strictly interior to the WD and so
+our simulation must strive to eliminate boundary effects resulting from
+nonphysical boundaries. To do so, we adopt periodic boundary conditions in the
+$x$ direction and damping layers (described in \autoref{ss:damping}) in the $z$
+direction to damp perturbations exiting the domain. Correspondingly, we use a
+Fourier basis in the $x$ direction and a Chebyshev in the $z$ direction.
 
@@
 
-Denote $L_x, L_z$ to be the physical dimensions of the simulation domain. The
-physics of our simulation is fixed by six parameters: $L_x, L_z$, and $k_{0x},
-\omega_0, F, \nu$. We describe and choose these parameters below:
+% Grid Size
+% Viscosity + diffusivity, damping layers (to simulate radiative BC)
+% timestepping? RK443, (CFL limiter??)
+
+Our
+simulation starts with all perturbation quantities at zero. We choose as
+follows:
 \begin{itemize}
-    \item $L_z$: Since the wave amplitude grows $\sim e^{z/2H}$, we choose $L_z
-        = 10H$ to give $\sim e^4$ apmlitude growth between the damping zones.
+    \item $k_{0x}$: Astrophysical IGWs have $k_{0x} \ll k_{0z}$, so we choose $k_{0x}
+        = \frac{2\pi}{L_x}$ the smallest permitted wavenumber permitted by
+        periodic boundary conditions.
 
-    \item $L_x$: We want similar grid spacing $\frac{L_x}{N_x} \sim
-        \frac{L_z}{N_z}$. For computational savings, we fix $N_z = 4N_x$.
-        Expecting wave breaking to be mostly in the $z$ direction, we use $L_x =
-        4H$.
+    \item We choose $\omega_0$ by choosing it to produce a desired $k_{0z}$.
+        Astrophysical systems generally also exhibit $k_{0z}H \gg 1$, where the
+        stratification height is significantly larger than the vertical
+        wavelength. However, in order for the simulation to both be well-damped
+        at grid resolutions and be neglibily damped at leading order nonlinear
+        wavelengths, we require $k_{0z} \ll \frac{1}{L_{NL}} \ll \frac{2\pi
+        N_z}{L_z}$, where $L_{NL}$ denotes some characteristic length scale of
+        nonlinear features. Finally, we need $L_z$ to be many $H$ to capture
+        significant $e^{z/2H}$ growth and separate the linear and nonlinear
+        amplitudes of $u_z$.
 
-    \item $k_{0x}$: Astrophysical IGWs in stars are generally excited by the $l
-        = 2$ component of the tidal potential, for which $k_{\perp} \sim
-        \frac{1}{R}$ where $R$ is the radius of the star. To best emulate this,
-        we use $k_{0x} = \frac{2\pi}{L_x}$ the smallest permitted wavenumber
-        permitted by periodic boundary conditions.
+        Fixing $L_z = 10H$ to give $\sim e^4$ amplitude growth between the
+        damping zones, this implies we require separation of scales $1 \ll
+        k_{0z}H \ll \frac{H}{L_{NL}} \ll \frac{2\pi N_z}{10}$.
 
-    \item $\omega_0$: We choose $\omega_0$ by evaluating dispersion relation
-        $\omega_0\p*{k_{0x}, k_{0z}}$ for some desired $k_{0z}$.
+        Since $N_z$ is fixed by computational cost (we tried both $N_z = 512,
+        N_z = 1024$), we choose $k_{0z} \sim \frac{2\pi}{H}$. This is less
+        physically representative but gives more wavenumber space for the
+        nonlinear cascade. We then invert $\omega_0 = \omega(k_{0x}, k_{0z})$.
 
-        Astrophysical IGWs also generally satisfy $\omega \ll N$, or
-        equivalently $\frac{k_r}{k_{\perp}} \sim k_rR \gg 1$. Since $H \lesssim
-        R$, we aim to study $k_{0z}H \gg 1$. However, to satisfy this well is
-        computationally expensive: $L_z \gg H$ is required to give waves ample
-        room to grow within the simulation domain, but we want $k_{0z}$ to be
-        sufficiently separated from the grid spacing $\sim L_z / N_z$ that
-        nonlinear effects can be seen in later simulations (see
-        \autoref{ss:nl_ns}). Thus, $N_z$ sets how well we can satisfy $k_{0z}H
-        \gg 1$; we choose $k_{0z}H = 1$.
-
-    \item $F$: We first choose $F$ forcing strength such that $\vec{\xi} \cdot
-        \vec{k} \ll 1$ is satisfied everywhere throughout the domain. This
-        becomes a constraint on $F$ by using \autoref{eq:uz_lin} and
-        $\abs*{\vec{u}} \propto e^{z/2H}$.
-
-    \item $\nu$: Nonlinear effects transfer wave energy from $\vec{k}_{0}$ to
-        smaller scales. Since well-resolved simulations using spectral methods
-        have no inherent numerical viscosity, energy will accumulate at grid
-        scales in the absence of artificial dissipation. We introduce
-        Navier-Stokes viscosity $\nu$ while ensuring our equations of motion
-        continue to conserve horizontal momentum (see \autoref{ss:strat_impl}
-        for details). When $\vec{\xi} \cdot \vec{k} \ll 1$, we may set $\nu =
-        0$.
+    \item We choose $F$ forcing strength $1/20$ that of the nonlinear
+        simulation, which keeps the flow amplitude sufficiently small to be
+        treated in the linear approximation at all points in the simulation
+        domain.
 \end{itemize}
@@
 similar studies\cite{lecoanet_damp}, has the advantage of being smooth,
-important for spectral methods. The full implementation of the fluid equations
-in Dedalus is described in \autoref{ss:strat_impl}.
+important for spectral methods. As these damping terms are linear, they are
+retained in the linearized equations of motion as well.
 
-\subsection{Linear Regime}\label{ss:lin_ns}
+The full implementation of the fluid equations in Dedalus is described in
+\autoref{ss:strat_impl}.
+
+\subsection{Linear Regime}\label{ss:lin}
 
 We first test our numerical simulation with a ``linear'' simulation, where
-$F$ is chosen such that $\vec{\xi} \cdot \vec{k} \ll 1$ everywhere. We emphasize
-that our ``linear'' simulations solve the full nonlinear equations, just in the
-weak perturbation limit.
+all flow quantities are small: we assume  $\rho_1 \equiv \rho - \rho_0 \ll
+\rho_0$ and $\vec{u} \cdot \vec{\nabla} \ll \partial_t$. We emphasize that our
+``linear'' simulations solve the full nonlinear equations, just in the weak
+perturbation limit.
+
+$\omega$ is chosen by inverting $\omega(k_{0x}, k_{0z})$ dispersion relation for fixed
+$k_{0x} = 2\pi / L_x$ and some desired $k_{0z} = -2\pi/H$, and $\sigma \lesssim
+\frac{1}{k_{0z}}$ is used to excite a broad band of $k_z$ modes including the
+desired $k_{0z}$ mode.
+
+\subsection{Numerical Simulation}\label{ss:lin_ns}
+
+A representative snapshot of this simulation
+is provided in \autoref{fig:linear_nonu}. Key features to note:
+\begin{itemize}
+    \item $u_z$ grows with \autoref{eq:uz_lin} to reasonable accuracy.
+
+    \item $S_{px}$ is seen to be constant, as would be expected by
+        \autoref{eq:fpx_lin}.
+\end{itemize}
+We also include a qualitatively similar snapshot of the same simulation run
+using the $\nu$ from the later nonlinear simulations in \autoref{fig:linear_nu}.
+\begin{figure}[t]
+    \centering
+    \includegraphics[width=0.5\textwidth]{plots/linear_nonu.png}
+    \caption{Snapshot of linear simulation after reaching steady state.
+        Analytical predictions of $\abs*{u_z}, \bar{U}_0, S_{px}$ are shown in
+        orange. Red dotted lines indicate the onset of the damping region and
+        blue dotted lines denote the driving region.}\label{fig:linear_nonu}
+\end{figure}
+\begin{figure}[t]
+    \centering
+    \includegraphics[width=0.5\textwidth]{plots/linear_nu.png}
+    \caption{Same as \autoref{fig:linear_nonu} but with $0.3\times $viscosity
+    used for nonlinear simulations (YUBONOTE should be the same but that
+    simulation mysteriously stopped without me noticing; restarting worked, so
+    currently rerunning). Note the slight resulting dissipation in $S_{px}$.
+    }\label{fig:linear_nu}
+\end{figure}
 
@@
 
-A Navier-Stokes numerical viscosity $\nu$ is used to damp high wavenumbers and
-regularize the nonlinear cascade at near grid resolution: $\nu \sim 0.1
-\frac{\omega_0}{\abs*{k_{0z}}}\frac{L_z}{2\pi N_z}$ was found to be suitable for
-$N_z = 2048$.
+A
+Navier-Stokes numerical viscosity $\nu$ is used to damp high wavenumbers and
+regularize the nonlinear cascade at near grid resolution: $\nu \sim 0.2
+\frac{\omega}{\abs*{k_{0z}}}\frac{L_z}{2\pi N_z}$ was found to be suitable for
+$N_z = 1024$, where $k_{0z}$ is the wavenumber of the excited linear mode.
+%%%
 
@@
 
-\subsection{Numerical Simulation}\label{ss:nl_ns}
+\subsection{Numerical Simulation}
 
@@
 Two representative snapshots from our simulation are provided in
-TODO after the critical layer has had time to form. We may note that
+\autoref{fig:nl} after the critical layer has had time to form. We may note that
 the critical layer, where $S_{px}$ is absorbed and $\bar{U}_0 = \bar{U}_c$,
 travels downwards as predicted.
+\begin{figure}[t]
+    \centering
+    \begin{subfigure}{0.5\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{plots/nl35_1.png}
+        \caption{Earlier time snapshot. Legend is the same as
+        \autoref{fig:linear_nonu} except $\bar{U}_c$ is marked in red on the
+        third panel.}
+    \end{subfigure}
+
+    \begin{subfigure}{0.5\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{plots/nl35_2.png}
+        \caption{Later snapshot illustrating propagation of $z_c$.}
+    \end{subfigure}
+    \caption{Nonlinear numerical simulation.}\label{fig:nl}
+\end{figure}
 
@@
 
-For the simulation in TODO, we may define $z_c = \argmax_z
+For the simulation in \autoref{fig:nl}, we may define $z_c = \argmax_z
 \pd{\bar{U}_0}{z}$. Computing $\pd{z_c}{t}$ using the analytical flux
 \autoref{eq:fpx_lin} allows comparison to \autoref{eq:zc_anal}, which we exhibit
-in TODO\@. We see overall good agreement, though some small
+in \autoref{fig:zc_anal}. We see overall good agreement, though some small
 deviation is expected since $S_{px}$ is not perfectly conserved owing to
 numerical viscosity (and YUBONOTE $S_{px}$ is misestimated?).
+\begin{figure}[t]
+    \centering
+    \includegraphics[width=0.5\textwidth]{plots/nl35_front_v.png}
+    \caption{Comparison of simulated $\pd{z_c}{t}$ to analytical
+    \autoref{eq:zc_anal}. Plot begins after formation of critical layer in
+    simulation.}\label{fig:zc_anal}
+\end{figure}
 
@@
 layer absorption when $\mathrm{Ri} \gg \frac{1}{4}$. We may plot $\mathrm{Ri}$
-over the same time interval in TODO\@. We may observe that the
+over the same time interval in \autoref{fig:nl35_f_ri}. We may observe that the
 Richardson number is initially decreasing but eventually levels out and
@@
 transition.
+\begin{figure}[t]
+    \centering
+    \includegraphics[width=0.5\textwidth]{plots/nl35_f_ri.png}
+    \caption{Plot of $\mathrm{Ri}_{\max}(t) = \max_z
+    \frac{N^2}{\bar{U}_0'}(t)$ over the simulation time.}\label{fig:nl35_f_ri}
+\end{figure}
 
 We argue that the Richardson number is bounded from below by viscosity. A
-repeated simulation with a larger viscosity is shown in TODO,
+repeated simulation with a larger viscosity is shown in \autoref{fig:nl1_f_ri},
 where the Richardson number does not go nearly as low.
+\begin{figure}[t]
+    \centering
+    \includegraphics[width=0.5\textwidth]{plots/nl1_f_ri.png}
+    \caption{Same as \autoref{fig:nl35_f_ri} but with $\sim 1.3\times$
+    viscosity.}\label{fig:nl1_f_ri}
+\end{figure}
 

### 2019-04-29 writeups/paper.tex

--- 2019-04-16 writeups/paper.tex
@@
 \begin{abstract}
-    In white dwarf binaries tidal torques are expected to raise a train of
-    internal gravity waves in the white dwarf interior. These
-    outward-propagating waves can then grow to be nonlinear and break via
-    hydrodynamical instabilities, transfering energy and angular momentum from
-    the binary orbit to the white dwarf. We perform 2D numerical simulations of
-    nonlinear wave breaking of outwards-propagating internal gravity waves in an
-    incompressible, isothermal atmosphere with an exponentially decaying density
-    stratification. We find that tidal synchronization, after an initial
-    transient phase, proceeds inwards from the surface in thin layers. We argue
-    the thickness of these layers is limited by the Kelvin-Helmholtz
-    Instability. We provide simple analytical formulae for the location and
-    thickness of tidal synchronization over time that are in good agreement with
-    our simulations.
+    In sufficiently compact white dwarf binaries, dynamical tides raise a train
+    of internal gravity waves that propagate towards the surface. We perform 2D
+    numerical simulations of these waves undergoing nonlinear wave breaking in
+    an incompressible, isothermal atmosphere. After an initial transient phase,
+    we find that these waves induce a sharp transition between a non-rotating
+    core and synchronously rotating envelope. We find evidence that the width of
+    this transition layer is bound from below by the Kelvin-Helmholtz
+    Instability. We provide analytical formulae for absorption and reflection of
+    incident waves off the critical layer up to prefactors of order unity. These
+    prefactors converge to constant values when artificial dissipation is
+    decreased. We provide dimensionless criteria necessary to resolving momentum
+    transfer within the critical layer. Finally, we speculate on the application
+    of our model to tidal synchronization and heating in astrophysical systems.
 \end{abstract}

### 2019-04-30 writeups/paper.tex
stash WIP

@@
         ae, aecompl,
-        graphicx
+        graphicx,
+        booktabs,
     }
@@
     \DeclareMathOperator*{\argmax}{argmax}
+    \DeclareMathOperator*{\med}{med}
 
@@
 \section{Introduction}
+
+[Below copied from proposal, will rewrite]
 
@@
 We consider a two-dimensional incompressible, isothermal fluid, representative
-of degenerate matter in WDs. We neglect temperature variations and assume a
-barotropic equation of state as a first approximation. As we are interested in
-dynamics far from the center of the WD, we approximate the gravitational field
-as uniform. We model the background density stratification as $\rho_0(x, z) =
-\rho_0(z) = \rho_0(z=0) e^{-z/H}$ for some reference density $\rho_0(z=0) =
-\at{\rho_0(z)}_{z = 0}$.
+of degenerate matter in WDs. We furthermore neglect temperature variations and
+assume a barotropic equation of state $P(\rho, T) = P(\rho)$ as a first
+approximation. As we are interested in dynamics far from the center of the WD,
+we approximate the gravitational field as uniform. We model the background
+density stratification as $\rho_0(x, z) = \rho_0(z) = \rho_0(z=0) e^{-z/H}$ for
+some reference density $\rho_0(z=0) = \at{\rho_0(z)}_{z = 0}$.
 
@@
     \begin{align}
-        \vec{\nabla} \cdot \vec{u} &= 0,\label{eq:nl_incomp}\\
+        \vec{\nabla} \cdot \vec{u}_1 &= 0,\label{eq:nl_incomp}\\
         \md{\rho}{t} &= 0 ,\label{eq:nl_density}\\
-        \md{\vec{u}}{t} + \frac{\vec{\nabla}p}{\rho} + g\hat{z} &=
+        \md{\vec{u}_1}{t} + \frac{\vec{\nabla}p}{\rho} + g\hat{z} &=
             0.\label{eq:nl_mom}.
@@
 \end{subequations}
-$\md{}{t} = \pd{}{t} + \p{\vec{u} \cdot \vec{\nabla}}$ is the Lagrangian or
-material derivative. $\vec{u}, \rho, p$ denote the velocity field, density and
-pressure respectively. We denote $-g\hat{z}$ constant gravitational
+$\md{}{t} = \pd{}{t} + \p{\vec{u}_1 \cdot \vec{\nabla}}$ is the Lagrangian or
+material derivative, and $\vec{u}_1, \rho, p$ denote the velocity field, density
+and pressure respectively. We denote $-g\hat{z}$ constant gravitational
 acceleration. Note that at hydrostatic equilibrium $\pd{}{t} = 0$ we have
-$\vec{\nabla}P_0 = -\rho_0 g\hat{z}$ and so $P_0 = \rho_0 gH$. A shear flow
-$u_x(x, z, t) = u_x(z)$ is permitted in hydrostatic equilibrium, but we will
-initially consider no shear flow in equilibrium.
+$\vec{\nabla}P_0 = -\rho_0 g\hat{z}$ and so $P_0 = \rho_0 gH$. A shear flow of
+form $\vec{u}_1 = u_{1x}(x, z, t)\hat{x} = u_{1x}(z)\hat{x}$ is permitted in hydrostatic
+equilibrium, but we will assume no background shear flow. Physically, this
+assumption corresponds to a non-rotating star, or going to the corotating frame
+of a rigidly rotating star.
 
@@
 \frac{\rho}{\rho_0}$. This both identically enforces $\rho > 0$ and avoids
-numerical issues if $\rho$ is small. We also define reduced pressure $P =
-\frac{p}{\rho}$. Then, we may rewrite the second two equations of
-\autoref{se:nl_orig} as
+numerical issues in the $\frac{\vec{\nabla}p}{\rho}$ term if $\rho$ is small. We
+also define reduced pressure $P = \frac{p}{\rho}$. Then, we may rewrite the
+second two equations of \autoref{se:nl_orig} as
 \begin{subequations}\label{se:nl_upsilon}
     \begin{align}
-        \md{\Upsilon}{t} + u_z \pd{\ln \rho_0}{z} &= 0,\label{eq:nl_up_density}
-            \\
-        \md{\vec{u}}{t} + \vec{\nabla}P + P\vec{\nabla}\Upsilon + g\hat{z} &= 0.
+        \md{\Upsilon}{t} + u_{1z} \pd{\ln \rho_0}{z} &= 0
+            ,\label{eq:nl_up_density} \\
+        \md{\vec{u}_1}{t} + \vec{\nabla}P + P\vec{\nabla}\Upsilon + g\hat{z} &= 0.
     \end{align}
@@
 In the small perturbation limit, where flow velocities are small compared to the
-characteristic space and time scales $\pd{}{t} \gg \vec{u} \cdot \vec{\nabla}$,
+characteristic space and time scales $\pd{}{t} \gg \vec{u}_1 \cdot \vec{\nabla}$,
 we may linearize \autoref{se:nl_upsilon}. We ignore the advective components of the
@@
     \begin{align}
-        \vec{\nabla} \cdot \vec{u} &= 0,\\
+        \vec{\nabla} \cdot \vec{u}_1 &= 0,\\
         \pd{\Upsilon_1}{t} - \frac{u_{z}}{H} &= 0,\\
-        \pd{\vec{u}}{t} + \vec{\nabla}P_1
+        \pd{\vec{u}_1}{t} + \vec{\nabla}P_1
             + \p{1 + \Upsilon_1} g\hat{z}
@@
 well-known result \citep{drazin,sutherland0}:
-\begin{subequations}\label{se:lin_sol}
-    \begin{align}
-        u_z\p{x, z, t} &= Ae^{z/2H}e^{i(k_{0x}x + k_{0z}z - \omega_0 t)},\\
-        \omega_0^2 &= \frac{N^2k_{0x}^2}{k_{0x}^2 + k_{0z}^2 + \frac{1}{4H^2}},
+\begin{equation}
+    u_{1z}\p{x, z, t} = Ae^{z/2H}e^{i\p{k_{1x}x - k_{1z}z - \omega_1 t}},
+        \label{eq:k0z_sign}
+\end{equation}
+where $\omega_1$ satisfies dispersion relation
+\begin{equation}
+    \omega_1^2 = \frac{N^2k_{1x}^2}{k_{1x}^2 + k_{1z}^2 + \frac{1}{4H^2}},
         \label{eq:disp_rel}
-    \end{align}
-\end{subequations}
+\end{equation}
 where
 \begin{equation}
-    N^2 \equiv g^2\p{\rd{\rho}{P} - \at{\pd{\rho}{P}}_{ad}}
-        = \frac{g}{H},
+    N^2 \equiv g^2\p{\rd{\rho}{P} - \at{\pd{\rho}{P}}_{ad}} = \frac{g}{H},
 \end{equation}
-the Brunt-V\"ais\"al\"a frequency is constant
+the Brunt-V\"ais\"al\"a frequency is constant. Our sign convention is such that
+a wave with $k_{1x}, k_{1z} > 0$ has positive vertical group velocity (see
+\autoref{eq:vg}).
 
-In the weak stratification limit $k_{0z}H \gg 1$, the solution exhibits the
+In the weak stratification limit $k_{1z}H \gg 1$, the solution exhibits the
 following characteristics:
 \begin{itemize}
-    \item The amplitude of the wave grows like $e^{z/2H}$. Thus, the
-        linear approximation is violated for sufficiently large $z$.
+    \item The amplitude of the wave grows with $z$ as $e^{z/2H}$. Thus, the
+        linear approximation is always violated for sufficiently large $z$.
 
@@
             \vec{c}_{ph} &=
-                \p{\frac{\hat{x}}{k_{0x}} + \frac{\hat{z}}{k_{0z}}}\omega,\\
-            \vec{c}_{g} &= \frac{N\p{k_{0x}k_{0z}\hat{x}
-                - \p{k_{0z}^2 + \frac{1}{4H^2}}\hat{z}}}
-                {\p{k_{0x}^2 + k_{0z}^2 + \frac{1}{4H^2}}^{3/2}} \nonumber\\
-                &\approx \frac{\omega k_{0z}}{\abs{\vec{k}_0}^{3/2}}
-                    \p{k_{0x}\hat{x} - k_{0z}\hat{z}}.
+                \p{k_{1x}\hat{x} - k_{1z}\hat{z}}\frac{\omega_1}
+                {k_{1x}^2 + k_{1z}^2 + \frac{1}{4H^2}},\\
+            \vec{c}_{g} &= N\frac{\p{k_{1z}^2 + \frac{1}{4H^2}}\hat{x}
+                + \p{k_{1x}k_{1z}\hat{z}}}
+                {\p{k_{1x}^2 + k_{1z}^2 + \frac{1}{4H^2}}^{3/2}}.\label{eq:vg}
         \end{align}
-        We recover the usual result $\vec{c}_{ph} \cdot \vec{c}_g = 0$
-        (e.g.\  \citep{drazin,sutherland1}). Note that an IGW transporting energy
-        and momentum upwards $c_{g, z} > 0$ has $k_{0z} < 0$.
+        We note $\vec{c}_{ph} \cdot \vec{c}_g = +
+        \mathcal{O}\p{\frac{1}{\p{k_{1z}H}^{2}}} \approx 0$, the usual result
+        (e.g.\  \citep{drazin,sutherland1}). Note again that $c_{gz} > 0$ for
+        $k_{1x}, k_{1z} > 0$, a consequence of our sign choice in
+        \autoref{eq:k0z_sign}.
 
-    \item The time-averaged total $x$-momentum flux in the $\hat{z}$ direction
-        can be computed. Since the linear solution is separable as $f(x, z, t) =
-        f(z)e^{i(k_{0x}x - \omega_0 t)}$, $x$ averaging and time averaging are
-        equivalent, so we may write
+    \item The time-averaged total $x$-momentum flux carried in the $\hat{z}$
+        direction can be computed. Since the linear solution is separable as
+        $f(x, z, t) = f(z)e^{i(k_{1x}x - \omega_1 t)}$, $x$ averaging and time
+        averaging are equivalent, so we may write
         \begin{align}
-            S &\equiv \ev{\rho u_x u_z}_x \equiv
-                \frac{1}{L_x}\int_0^{L_x}\limits \rho u_x u_z\;\mathrm{d}x,
+            S &\equiv \ev{\rho u_{1x} u_{1z}}_x \equiv
+                \frac{1}{L_x}\int_0^{L_x}\limits \rho u_{1x} u_{1z}\;\mathrm{d}x,
                     \label{eq:S_def},\\
-                &\approx -\frac{A^2}{2}\rho_0(z=0)\frac{k_{0z}}{k_{0x}},
+                &\approx -\frac{A^2}{2}\rho_0(z=0)\frac{k_{1z}}{k_{1x}},
                     \label{eq:S_lin}
@@
 
-\subsection{Wave Generation}
+\subsection{Wave Generation: Analytical Analysis}
 
-To model a continuous IGW wave train excited deep in the WD interior propagating
+Since we are only concerned with the behavior of the wave at high $z$ where wave
+breaking occurs, we generate waves in the most numerically convenient way. As
+long as physically represesentative IGW are generated and wave dynamics are only
+examined far from the forcing region, the exact excitation mechanism is not
+important for our conclusions.
+
+To model continuous excitation of IGWs deep in the WD interior propagating
 towards the surface, we use a volumetric forcing term to excite IGW near the
-bottom of the simulation domain\footnote{Interfacial forcing at the bottom
-boundary incurs strong stability limits, as we use a Chebyshev polynomial basis
-along the $z$ axis in our spectral method which has very small grid spacing at
-the boundaries.}. Our forcing excites both IGWs propagating upwards, imitating a
-wave tidally excited deeper in the WD, and downwards, which are damped away by
-the damping layers described in \autoref{ss:damping}.
+bottom of the simulation domain\footnote{Interfacial forcing is numerically
+unstable when using a Chebyshev polynomial basis in the $z$ direction and
+impossible when using a Fourier basis.}. Our forcing excites both IGWs
+propagating upwards, imitating a wave tidally excited deeper in the WD, and
+downwards, which are damped away by the damping layers described in
+\autoref{ss:damping}.
 
@@
 \begin{equation}
-    \md{\Upsilon}{t} + u_z\pd{\ln \rho_0}{z}
+    \md{\Upsilon}{t} + u_{1z}\pd{\ln \rho_0}{z}
         = Fe^{-\frac{(z - z_0)^2}{2\sigma^2}}
-            \cos \p{k_{0x}x - \omega_0 t}.\label{eq:vol_drive}
+            \cos \p{k_{1x}x - \omega_1 t}.\label{eq:vol_drive}
 \end{equation}
 Using a narrow Gaussian profile excites a broad $k_z$ wavenumber spectrum, and
-only the $k_{0z}$ satisfying dispersion relation \autoref{eq:disp_rel} for the
-given $k_{0x}, \omega_0(k_{0x}, k_{0z})$ will propagate.
+only the $k_{1z}$ satisfying dispersion relation \autoref{eq:disp_rel} for the
+given $k_{1x}, \omega_1(k_{1x}, k_{1z})$ will propagate.
 
@@
 then approximately relax the solution to nonzero $\sigma$: an extra factor of
-$e^{-\frac{(k_{0z}\sigma)^2}{2}}$ arises compared to the $\delta$-function
+$e^{-\frac{(k_{1z}\sigma)^2}{2}}$ arises compared to the $\delta$-function
 solution (evaluating the Fourier Transform of $e^{-\frac{(z -
-z_0)^2}{2\sigma^2}}$ at $k_z = k_{0z}$), and we obtain
+z_0)^2}{2\sigma^2}}$ at $k_z = k_{1z}$), and we obtain
 \begin{align}
-    u_{1z}&(x, z, t) ={} \frac{Fgk_{0x}^2}{\omega_0^2}
-        \frac{1}{2ik_{0z}}\frac{e^{\frac{-(k_{0z}\sigma)^2}{2}}}
+    u_{z}&(x, z, t) ={} \frac{Fgk_{1x}^2}{\omega_1^2}
+        \frac{1}{2ik_{1z}}\frac{e^{\frac{-(k_{1z}\sigma)^2}{2}}}
         {\sqrt{2\pi\sigma^2}} \times\nonumber\\
         &{}\begin{cases}
-        e^{\frac{z - z_0}{2H}}e^{i\p{k_{0x}x + k_{0z}(z - z_0) - \omega_0 t
-            + \frac{1}{2k_{0z}H}}}
+        e^{\frac{z - z_0}{2H}}e^{i\p{k_{1x}x + k_{1z}(z - z_0) - \omega_1 t
+            + \frac{1}{2k_{1z}H}}}
             & z > z_0\\
-        e^{\frac{z - z_0}{2H}}e^{i\p{k_{0x}x - k_{0z}(z - z_0) - \omega_0 t
-            + \frac{1}{2k_{0z}H}}}
+        e^{\frac{z - z_0}{2H}}e^{i\p{k_{1x}x - k_{1z}(z - z_0) - \omega_1 t
+            + \frac{1}{2k_{1z}H}}}
             & z < z_0\\
@@
 the simulation domain from below. The solution for $u_{1x}(x, z, t)$ can be
-obtained analytically from \autoref{eq:nl_incomp}.
+obtained analytically from incompressibility \autoref{eq:nl_incomp}.
 
-\subsection{Wave Breaking}
+\subsection{Wave Breaking Location}\label{ss:wave_breaking}
 
 In developing the results above, we have neglected advective terms $\pd{}{t} \gg
-\vec{u} \cdot \vec{\nabla}$. However, since $\abs{\vec{u}} \propto e^{z/2H}$,
-any nonzero amplitude IGW will eventually violate the linearity criterion. This
-regime is relevant in practice since the IGW may grow to nonlinear amplitudes
-before it reaches the surface of the WD \citep{fullerI,fullerII}. We can coarsely
-estimate the height at which this happens by when the Lagrangian displacement
-$\vec{\xi}$ of a fluid parcel satisfies
+\vec{u}_1 \cdot \vec{\nabla}$. However, since $\abs{\vec{u}_1} \propto
+e^{z/2H}$, any nonzero amplitude IGW will eventually violate the linearity
+criterion. This regime is relevant in practice since the IGW may grow to
+nonlinear amplitudes before it reaches the surface of the WD
+\citep{fullerI,fullerII}. We can coarsely estimate the height at which this
+happens by when the Lagrangian displacement $\vec{\xi} =
+\frac{\vec{u}_1}{-i\omega}$ of a fluid parcel satisfies
 \begin{equation}
-    \vec{\xi} \cdot \vec{k} \gtrsim 1.\label{eq:nl}
+    \vec{\xi} \cdot \vec{k}_1 \gtrsim 1.\label{eq:nl}
 \end{equation}
+
+Another useful result quantifies the wave-induced mean flow
+\citep{eliassen_palm_cite,sutherland0}:
+\begin{equation}
+     \ev{u_{1x}} \equiv \bar{U} = \frac{1}{L_x}
+        \int\limits u_{1x}\;\mathrm{d}x = \frac{\ev{u_{1x}u_{1z}}_x}{c_{g,z}},
+        \label{eq:mean_flow}
+\end{equation}
+This formula is analogous to Stokes' drift for surface waves and can be derived
+by considering the propagation of $S$ (\autoref{eq:S_def}) into a medium at
+rest. Solving for $\bar{U} = c_{ph, x}$ the critical layer condition (see
+\autoref{ss:crit_layer} immediately below), we recover
+$\frac{u_{1x}k_{1x}}{\omega} = \xi_x k_{1x} \sim 1$, roughly equivalent to the
+nonlinearity criterion \autoref{eq:nl}. This condition predicts where the wave's
+self-acceleration alone is sufficient to induce critical layer absorption.
+
+\subsection{Formation and Dynamics of Critical Layer}\label{ss:crit_layer}
+
+[TODO rewrite more clearly, it's just copy pasted together right now]
 
@@
 depositing horizontal momentum in the fluid's mean flow \citep{klostermeyer}.
-Once the mean flow reaches the horizontal phase velocity of the parent IGW, a
-critical layer forms at which the frequency of the IGW is Doppler-shifted to
-zero. The interaction of the parent IGW with the mean flow was first studied in
-the inviscid, linear regime in \citep{booker_bretherton}, which found nearly
+
+In accordance with literature and the simulation setup, we will assume momentum
+redistribution is inefficient (see \autoref{ss:disp} for analysis of this
+assumption). The mean flow will then grow in place until it reaches the
+horizontal phase velocity of the parent IGW\@. Then, a critical layer forms at
+which the frequency of the incident IGW is Doppler-shifted to zero. The
+interaction of the parent IGW with the mean flow was first studied in the
+inviscid, linear regime in \citep{booker_bretherton}, which found nearly
 complete absorption. This result was reproduced with nonzero viscosity
@@
 
-An estimate of where these mean flow effects become important can be made using
-the wave-induced mean flow \citep{eliassen_palm_cite,sutherland0}:
+A purely horizontal shear flow $\bar{U}(z) \hat{x}$ can be seen in
+\autoref{se:nl_orig} to have the effect of modifying time derivatives
+$\partial_t$ to their frequency in the comoving frame of the fluid $\partial_t -
+\bar{U}(z)\partial_x$. For a critical value $\omega_1 - \bar{U}_c k_{1x} = 0$,
+the frequency of the linear wave in the fluid's frame of reference vanishes and
+critical behavior is observed. In a linear \citep{booker_bretherton} or viscous
+\citep{hazel} theory, the incident wave has amplitude reflection and
+transmission coefficients
+\begin{align}
+    \mathrm{Ri} &\equiv \at{\frac{N^2}{\p{\pd{\bar{U}}{z}}^2}}_{z_c},
+        \label{eq:ri_def}\\
+    \mathcal{R} &= e^{-2\pi \sqrt{\mathrm{Ri} - \frac{1}{4}}}, &
+    \mathcal{T} &= e^{-\pi \sqrt{\mathrm{Ri} - \frac{1}{4}}},
+        \label{eq:crit_coeffs}
+\end{align}
+where we have defined local Richardson number Ri at the critical layer $z_c:
+\bar{U}(z_c) = \frac{\omega_1}{k_{1x}}$. In the $\mathrm{Ri} \gg 1$ limit,
+$\mathcal{R}, \mathcal{T} \ll 1$ and the incident wave is completely absorbed to
+good approximation.
+
+When the fluid absorbs the incident wave, it absorbs the incident horizontal
+momentum flux as well, which is converted into additional horizontal momentum of
+the shear flow. Since the shear flow cannot exceed $\bar{U}_c$ the horizontal
+phase velocity of the incident wave, the critical layer must instead propagate
+downwards to accommodate the incident momentum. The total horizontal momentum of
+the shear flow then obeys conservation equation
 \begin{equation}
-     \ev{u_x} = \bar{U}_0 \equiv \frac{1}{L_x}
-        \int\limits u_x\;\mathrm{d}x = \frac{\ev{u_xu_z}_x}{c_{g,z}},
-        \label{eq:mean_flow}
+    \pd{}{t}\int\limits_0^{L_z} \rho(z) \bar{U}(z, t)\;\mathrm{d}z
+        \Delta S = 0.
 \end{equation}
-where $\ev{\dots}_x$ denotes horizontal averaging. Setting $\bar{U}_0 = c_{ph,
-x}$ the critical layer condition, we recover $\frac{u_xk_x}{\omega} = \xi_x k_x
-\sim 1$, roughly equivalent to the nonlinearity criterion \autoref{eq:nl}.
+We define $\Delta S$ to be the change in flux across the critical layer, or
+equivalently the negative of the absorbed flux. Treating $\bar{U}(z > z_c) =
+\bar{U}_c, \bar{U}(z < z_c) = 0$ gives us exactly
+\begin{equation}
+    \rho(z_c) \bar{U}_c\pd{z_c}{t} = \Delta S(z_c(t)).\label{eq:zc_anal}
+\end{equation}
+$\Delta S(z_c)$ denotes the change in $S$ across $z_c$. For constant $\Delta S$
+in time and $\rho \approx \rho_0$, this has analytical solution
+\begin{equation}
+    z_c(t) = -H\ln t - H\ln \frac{H\rho_0(z = 0)c_{ph, x}}{-\Delta S}
+        .\label{eq:zc_sol}
+\end{equation}
+Note that constant $-\Delta S$ corresponds physically to the picture where
+constant $S$ carried by the IGWs, given by \autoref{eq:S_lin}, is fully absorbed
+at the critical layer, in which case $\Delta S = -S$.
 
@@
 
+The geometry of our simulation domain is fixed by three parameters: $L_x, L_z$,
+and $z_0$ the forcing location. We use $z_0 = 0.2L_z$ to force sufficiently far
+from the lower damping zone and permit sufficient room for the upwards-moving
+wave to grow at $\propto e^{z/2H}$. We choose $L_z = 10H$ to give $\sim e^3$
+apmlitude growth between the damping zones. Finally, we want similar grid
+spacing $\frac{L_x}{N_x} \sim \frac{L_z}{N_z}$. For computational savings, we
+fix $N_z = 4N_x$. We use $L_x = 4H = \frac{2}{5}L_z$.
+
 The time integration uses a split implicit-explicit third-order scheme where
@@
 with adaptive timesteps computed from advective Courant-Friedrichs-Lewy (CFL)
-time. Specifically, we use $\Delta t = 0.7 \min(\Delta x / u_x,\Delta z / u_z)$,
+time. Specifically, we use $\Delta t = 0.7 \min(\Delta x / u_x,\Delta z / u_{1z})$,
 where the minimum is taken over every grid point in the domain and $\Delta
@@
 
-The physics of our simulation is fixed by six parameters: $L_x, L_z$, and
-$k_{0x}, \omega_0, F, \nu$. We describe our choices for these parameters below:
+The physics of our simulation is fixed by four parameters: $k_{1x}, \omega_1, F,
+\nu$. We describe our choices for these parameters below:
 \begin{itemize}
-    \item $L_z$: Since the wave amplitude grows $\sim e^{z/2H}$, we choose $L_z
-        = 10H$ to give $\sim e^4$ apmlitude growth between the damping zones.
-
-    \item $L_x$: We want similar grid spacing $\frac{L_x}{N_x} \sim
-        \frac{L_z}{N_z}$. For computational savings, we fix $N_z = 4N_x$.
-        Expecting wave breaking to be mostly in the $z$ direction, we use $L_x =
-        4H$.
-
-    \item $k_{0x}$: Astrophysical IGWs in stars are generally excited by the $l
+    \item $k_{1x}$: Astrophysical IGWs in stars are generally excited by the $l
         = 2$ component of the tidal potential, for which $k_{\perp} \sim
         \frac{1}{R}$ where $R$ is the radius of the star. To best emulate this,
-        we use $k_{0x} = \frac{2\pi}{L_x}$ the smallest permitted wavenumber
+        we use $k_{1x} = \frac{2\pi}{L_x}$ the smallest permitted wavenumber
         permitted by periodic boundary conditions.
 
-    \item $\omega_0$: We choose $\omega_0$ by evaluating dispersion relation
-        $\omega_0\p{k_{0x}, k_{0z}}$ for some desired $k_{0z}$.
+    \item $\omega_1$: We choose $\omega_1$ by evaluating dispersion relation
+        $\omega_1\p{k_{1x}, k_{1z}}$ for some desired $k_{1z}$.
 
@@
         equivalently $\frac{k_r}{k_{\perp}} \sim k_rR \gg 1$. Since $H \lesssim
-        R$, we aim to study $k_{0z}H \gg 1$. However, to satisfy this well is
+        R$, we aim to study $k_{1z}H \gg 1$. However, to satisfy this well is
         computationally expensive: $L_z \gg H$ is required to give waves ample
-        room to grow within the simulation domain, but we want $k_{0z}$ to be
+        room to grow within the simulation domain, but we want $k_{1z}$ to be
         sufficiently separated from the grid spacing $\sim L_z / N_z$ that
         nonlinear effects can be seen in later simulations (see
-        \autoref{ss:nl_ns}). Thus, $N_z$ sets how well we can satisfy $k_{0z}H
-        \gg 1$; we choose $k_{0z}H = 1$.
+        \autoref{ss:nl_ns}). Thus, $N_z$ sets how well we can satisfy $k_{1z}H
+        \gg 1$; we choose $k_{1z}H = 1$.
 
     \item $F$: We first choose $F$ forcing strength such that $\vec{\xi} \cdot
-        \vec{k} \ll 1$ is satisfied everywhere throughout the domain. This
-        becomes a constraint on $F$ by using \autoref{eq:uz_lin} and
-        $\abs{\vec{u}} \propto e^{z/2H}$.
+        \vec{k}_1 \ll 1$ is satisfied everywhere throughout the domain. This
+        constraints $F$ by \autoref{eq:uz_lin}.
 
-    \item $\nu$: Nonlinear effects transfer wave energy from $\vec{k}_{0}$ to
+    \item $\nu$: Nonlinear effects transfer wave energy from $\vec{k}_{1}$ to
         smaller scales. Since well-resolved simulations using spectral methods
@@
         horizontal momentum (see \autoref{ss:strat_impl} for details). Under
-        weak forcing, we may set $\nu = 0$, since the energy cascade is
-        negligible.
+        weak forcing, we may set $\nu = 0$, since energy transfer out of the
+        parent $\vec{k}_1$ mode is negligible.
 \end{itemize}
+
+Finally, we use initial conditions $\vec{u}(0) = \Upsilon(0) = 0, \varpi(0) =
+1$ corresponding to hydrostatic equilibrium and no initial fluid motion.
 
@@
 \begin{align}
-    \md{\Upsilon}{t} + u_z \pd{\ln \rho_0}{z} &=
+    \md{\Upsilon}{t} + u_{1z} \pd{\ln \rho_0}{z} &=
         -\Gamma(z)\Upsilon,\nonumber\\
-    \md{\vec{u}}{t} + \vec{\nabla}P + P\vec{\nabla}\Upsilon + g\hat{z} &=
-        - \Gamma(z)\vec{u},\nonumber\\
+    \md{\vec{u}_1}{t} + \vec{\nabla}P + P\vec{\nabla}\Upsilon + g\hat{z} &=
+        - \Gamma(z)\vec{u}_1,\nonumber\\
     \frac{1}{2\tau}\s{2 + \tanh \frac{z - z_T}{\Delta z}
@@
 similar studies \citep{lecoanet_damp}, has the advantage of being smooth,
-important for spectral methods. The full implementation of the fluid equations
-in Dedalus is described in \autoref{ss:strat_impl}.
+important for spectral methods. Further details of our implementation of the
+fluid equations in Dedalus are described in \autoref{ss:strat_impl}.
 
@@
 
-We present the results of a simulation satisfying the parameter choices in
-\autoref{ss:numerics} such that $\vec{\xi} \cdot \vec{k} \ll 1$ everywhere.
+We describe the results of a simulation satisfying the parameter choices in
+\autoref{ss:numerics} such that $\vec{\xi} \cdot \vec{k}_1 \ll 1$ everywhere.
+
+During the simulation, we expect IGW of form \autoref{eq:uz_lin} to be
+excited. Incompressibility allows us to compute a $\vec{u}_1(x, z, t)$ profile
+analytically. The amplitude of the observed IGW in $\vec{u}$ relative to
+analytical solution $\vec{u}_1$  can be estimated with estimator $\hat{A}(t)$
+\begin{equation}
+    \hat{A}(t) = \frac{\int\limits_{z_b}^{z_t}\int\limits_0^{L_x}
+        \rho_0\p{k_{1x}u_xu_{1x} - k_{1z}u_{1z}u_{1z}}\;\mathrm{d}x\mathrm{d}z}
+        {\int\limits_{z_b}^{z_t}\int\limits_0^{L_x}
+        \rho_0\p{k_{1x}u_{1x}^2 - k_{1z}u_{1z}^2}\;\mathrm{d}x\mathrm{d}z}
+        \label{eq:ahat_def}
+\end{equation}
+If $\vec{u} = \vec{u}_1$ over $z \in [z_b, z_t], \forall t$, then it's clear
+that $\hat{A}(t) = 1$. The choice of multiplication by $\rho_0$ is such that the
+integrand is of roughly equal magnitude throughout the domain of integration (as
+$\vec{u} \propto e^{z/2H}$), while including a weighting by $k_{1x}, k_{1z}$
+ensures that reflected components with wavevector $\vec{k} = k_{1x}\hat{x} -
+k_{1z}\hat{z}$ identically vanish in the $\hat{A}$ estimator. Both of these
+features become important in later sections (see \autoref{ss:reflectivity}).
+
+For a simulation satisfying $\vec{\xi} \cdot \vec{k}_1 \ll 1$ everywhere, we
+expect $\hat{A}(t) = 1$ so long as we consider only the domain between the
+forcing and damping zones, i.e.\  $z_b \gtrsim z_0, z_t \lesssim z_T$ ($z_0,
+z_T$ are defined in \autoref{eq:vol_drive} and \autoref{eq:Gamma} respectively).
+For consistency with the nonlinear case later, we choose $z_b = z_0 + 3\sigma,
+z_t = z_b + H$, though instead using $z_t = z_T - \Delta z$ does not change the
+results appreciably. The resulting measurement of $\hat{A}(t)$ is shown in
+\autoref{fig:lin_amps}.
 \begin{figure}[t]
@@
     \caption{Amplitude of excited wave over time in weak forcing simulation,
-    in units of the analytical estimate \autoref{eq:uz_lin}. We observe great
-    agreement with the linear calculation.}\label{fig:lin_amps}
+    computed using \autoref{eq:ahat_def}. $\hat{A}(t) = 1$ corresponds to
+    perfect agreement with the analytical estimate. After an initial transient
+    phase, we observe great agreement with \autoref{eq:uz_lin}, and in
+    particular $\hat{A}(t)$ asymptotes to a constant value, implying continuous
+    excitation of identical IGW.}\label{fig:lin_amps}
 \end{figure}
+
+A second prediction of the analytical theory is that $S(z, t)$ should be
+spatially flat between the forcing and damping zone(\autoref{eq:S_def}). We make
+may a slightly more accurate estimate of $S_1$ the flux carried in the IGW by
+directly substituting $\vec{u}_1$ into \autoref{eq:S_def}. Calling this estimate
+$S_1$, we may then measure the agreement of our simulation with analytical
+expectation by computing
+\begin{equation}
+    \hat{S}(z, t) \equiv \frac{\int\limits_0^{L_x}\rho u_xu_z\;\mathrm{d}x}{
+        \int\limits_{0}^{L_x} \rho_0 u_{1x}u_{1z}\;\mathrm{d}x}.
+        \label{eq:shat_def}
+\end{equation}
+For our weak forcing simulation, $\hat{S} = 1$ is expected between $z_0, z_T$,
+and indeed we observe agreement with this in \autoref{fig:lin_fluxes}.
 \begin{figure}[t]
@@
     \includegraphics[width=\columnwidth]{plots/lin_fluxes.png}
-    \caption{Horizontal momentum flux $S(z, t)$ in units of the analytical estimator
-    given by \autoref{eq:S_lin} as a function of $z$ at select times $t$. As
-    initial transients die out, the flux converges to be flat below the damping
-    layer. The flux excited in the forcing zone is transported without loss to
-    the top of the domain, where it is dissipated by the damping layer
-    (see \autoref{ss:damping}) without reflection.}\label{fig:lin_fluxes}
+    \caption{\autoref{eq:shat_def} as a function of $z$ at select times $t$. As
+    initial transients die out, $\hat{S}(z, t) = 1$ to very good agreement below
+    $z \lesssim z_T = 9.5H$. The flux excited in the forcing zone is transported
+    without loss to the top of the domain, where it is dissipated by the damping
+    layer (see \autoref{ss:damping}) without reflection.}\label{fig:lin_fluxes}
 \end{figure}
@@
 
-\subsection{Mean Flow Critical Layer Absorption}
+To perform simulations of wave breaking phenomena, we use the same values as
+\autoref{ss:numerics} except for $F, \nu$. In particular, we choose $F$ such
+that $\vec{k}_1 \cdot \vec{\xi} = 0.1$ in the forcing zone, which implies
+$\vec{k}_1 \cdot \vec{\xi}$ is exceeded before $z_T$ the upper damping zone.
+The dissipation parameter $\nu$ was varied across the various simulations. To
+quantify $\nu$, we define Reynolds number
+\begin{equation}
+    \mathrm{Re} = \frac{LU}{\nu} \equiv \frac{\omega_1}{\nu k_{1z}^2}.
+        \label{eq:re_def}
+\end{equation}
+A table of our simulation viscosities can be found in \autoref{tab:params}.
+\begin{table*}
+    \centering
+    \begin{tabular}{l c c c}
+        Name & Resolution & $\mathrm{Re}$ & $z$ spectral basis\\\bottomrule
+        \texttt{0p5\_hres} & $512 \times 2048$ & $2048$ & Fourier\\
+        \texttt{0p5\_shres} & $1024 \times 4096$ & $2048$ & Fourier\\
+        \texttt{1\_hres} & $512 \times 2048$ & $1024$ & Fourier\\
+        \texttt{1\_vhres} & $768 \times 3072$ & $1024$ & Fourier\\
+        \texttt{2\_hres} & $512 \times 2048$ & $512$ & Fourier\\
+        \texttt{3\_width} & $256 \times 1024$ & $341$ & Fourier\\
+        \texttt{6\_masked} & $256 \times 1024$ & $512$ & Chebyshev\\
+        \texttt{4\_masked} & $256 \times 1024$ & $341$ & Chebyshev\\
+        \texttt{3\_masked} & $256 \times 1024$ & $205$ & Chebyshev\\
+        \texttt{1\_masked} & $256 \times 1024$ & $146$ & Chebyshev\\
+    \end{tabular}
+    \caption{Table of simulation parameters. Note that $\mathrm{Re}$ is defined
+    as in \autoref{eq:re_def}.}\label{tab:params}
+\end{table*}
 
-A Navier-Stokes numerical viscosity $\nu$ is used to damp high wavenumbers and
-regularize the nonlinear cascade at near grid resolution: $\nu \sim 0.1
-\frac{\omega_0}{\abs{k_{0z}}}\frac{L_z}{2\pi N_z}$ was found to be suitable for
-$N_z = 2048$.
+\subsection{Numerical Simulation Results}\label{ss:nl_ns}
 
-In past studies of IGWs in WDs (cite Fuller \& Lai), $\xi_z \equiv
-\frac{u_z}{\omega_0}$ the Lagrangian fluid displacement was often used towards
-wave breaking criterion $k_{0z}\xi_z \gtrsim 1$. We argue that the wave's
-self-interaction via its generated mean flow $\bar{U}_0$ induces total
-absorption when the mean flow exceeds critical value
-\begin{equation}
-    \bar{U}_c = \frac{\omega_0}{k_{0x}}.
-\end{equation}
-This is consistent with the picture put forth in e.g.\ Goldreich and Nicholson
-(cite).
-
-A purely horizontal shear flow $\bar{U}_0(z) \hat{x}$ can be seen in
-\autoref{se:nl_orig} to have the effect of modifying time derivatives
-$\partial_t$ to their frequency in the comoving frame of the fluid $\partial_t -
-\bar{U}_0(z)\partial_x$. For a critical value $\omega_0 - \bar{U}_c k_{0x} = 0$,
-the frequency of the linear wave in the fluid's frame of reference vanishes and
-critical behavior is observed. In a linear theory or a theory where small scales
-are viscosity rather than advection dominated, the incident wave has amplitude
-reflection and transmission coefficients
-\begin{align}
-    \mathcal{R} &= e^{-2\pi \sqrt{\mathrm{Ri} - \frac{1}{4}}}, &
-    \mathcal{T} &= e^{-\pi \sqrt{\mathrm{Ri} - \frac{1}{4}}},
-    \label{eq:crit_coeffs}
-\end{align}
-where we have defined Richardson number $\mathrm{Ri} \equiv
-\at{\frac{N^2}{\p{\pd{\bar{U}_0}{z}}^2}}_{z_c}$ at the critical layer $z_c:
-\bar{U}_0(z_c) = \frac{\omega_0}{k_{0x}}$. For most shear flows, $\mathrm{Ri}
-\gg 1$ and so $\mathcal{R}, \mathcal{T} \ll 1$ and the incident wave is
-absorbed.
-
-When the fluid absorbs the incident wave, it absorbs the incident horizontal
-momentum flux as well, which is converted into additional horizontal momentum of
-the shear flow. Since the shear flow cannot exceed $\bar{U}_c$ the horizontal
-phase velocity of the incident wave, the critical layer must thus propagate
-downwards (towards the wave source) to accommodate the incident momentum. In
-other words, the total horizontal momentum of the shear flow obeys conservation
-equation
-\begin{equation}
-    \pd{}{t}\int\limits_0^{L_z} \rho(z) \bar{U}_0(z, t)\;\mathrm{d}z
-        - \Delta S = 0.
-\end{equation}
-We define $\Delta S$ to be the change in flux across the critical layer, or
-equivalently the absorbed flux. Treating $\bar{U}_0(z > z_c) = \bar{U}_c,
-\bar{U}(z < z_c) = 0$ gives us exactly
-\begin{equation}
-    \rho \bar{U}_c\pd{z_c}{t} = -\Delta S.\label{eq:zc_anal}
-\end{equation}
-For constant $\Delta S$ in space and $\rho \approx \rho_0$, this has analytical
-solution
-\begin{equation}
-    z(t) = -H\ln t - H\ln \frac{H\rho_0(z = 0)c_{ph, x}}{\Delta S}
-        .\label{eq:zc_sol}
-\end{equation}
-
-\subsection{Numerical Simulation}\label{ss:nl_ns}
-
-We use the same $k_{0x}, \omega_0$ as \autoref{ss:lin_ns}. Our other parameters
-are:
-\begin{itemize}
-    \item We choose $\nu = 0.1 \frac{\omega_0}{k_{0z}k_{z, \max}}$, where
-        $k_{z, \max} = \frac{2\pi N_z}{L_z}$. Note that $\nu =
-        \frac{\omega_0}{k_{0z} k_{z, \max}}$ corresponds to the advective term
-        $\vec{u} \cdot \vec{\nabla}$ being of the same order as the time
-        derivative $\partial_t$ for flow velocities $\vec{u} \sim
-        \frac{\omega_0}{k_{0z}}$ at the grid spacing.
-\end{itemize}
-
-Two representative snapshots from our simulation are provided in
-TODO after the critical layer has had time to form. We may note that
-the critical layer, where $S_{px}$ is absorbed and $\bar{U}_0 = \bar{U}_c$,
-travels downwards as predicted.
+A full video of our higher-resolution run at $N_x = 1024, N_z = 4096$ can be
+found at\footnote{ http://www.princeton.edu/~lecoanet/data/breaking\_wave.mov}.
+Slices of $\bar{U}, \hat{S}$ across $z$ at various times $t$ are shown in
+\autoref{fig:nl_fluxes}. While the behavior of $\bar{U}$ seems to conform
+qualitatively with the predictions of \autoref{ss:crit_layer}, the behavior of
+$\hat{S}$ exhibits two salient features: (i) the incident flux seems to fluctuate
+greatly with time, and (ii) there seems to be a small transmitted feature at many
+of the later times. We will discuss further these features in
+\autoref{ss:reflectivity}, after first analyzing the propagation of the critical
+layer.
+\begin{figure}[t]
+    \centering
+    \includegraphics[width=\columnwidth]{plots/nl_fluxes.png}
+    \caption{Plot of $\bar{U}(z, t)$ and $\hat{S}(z, t)$ respectively at various
+    times over $z$. $\bar{U}, \hat{S}$ follow their definitions in
+    \autoref{eq:mean_flow} and \autoref{eq:shat_def}. The values are extracted
+    from run \texttt{0p5\_shres}. $\bar{U}$ is given
+    in units of $c_{ph, x} = \frac{\omega_1}{k_{1x}}$ the critical layer flow
+    velocity (see \autoref{ss:crit_layer}). The propagation of the critical
+    layer towards lower $z$ and sharp deposition of $S$ at the critical layer
+    are evident.}\label{fig:nl_fluxes}
+\end{figure}
 
@@
 
-For the simulation in TODO, we may define $z_c = \argmax_z
-\pd{\bar{U}_0}{z}$. Computing $\pd{z_c}{t}$ using the analytical flux
-\autoref{eq:S_lin} allows comparison to \autoref{eq:zc_anal}, which we exhibit
-in TODO\@. We see overall good agreement, though some small
-deviation is expected since $S_{px}$ is not perfectly conserved owing to
-numerical viscosity (and YUBONOTE $S_{px}$ is misestimated?).
+For the simulation shown in \autoref{fig:nl_fluxes}, we may analyze the location
+of the critical layer. As the simulations are very noisy, we define the location
+of the critical layer to be
+\begin{align}
+    z_{c, \min} &= \argmin_z \z{z: S(z) < 0.3S_0},\nonumber\\
+    z_{c, \max} &= \argmax_z \z{z: S(z) > 0.3S_0},\nonumber\\
+    z_c &\equiv \frac{z_{c, \min} + z_{c, \max}}{2}.\label{eq:zc_def}
+\end{align}
+This was found to a relatively stable estimator of the critical layer location.
+Other estimators were used and do not significantly change the results of the
+analysis.
 
-However, recalling \autoref{eq:crit_coeffs}, we only expect complete critical
-layer absorption when $\mathrm{Ri} \gg \frac{1}{4}$. We may plot $\mathrm{Ri}$
-over the same time interval in TODO\@. We may observe that the
-Richardson number is initially decreasing but eventually levels out and
-begins to increase again. This corresponds to an increasingly sharp critical
-layer transition then subsequently a decreasingly sharp critical layer
-transition.
+The evolution of $z_c$ is described by ODE \autoref{eq:zc_anal}, and in the
+limit $\Delta S(t) = \Delta S$ we obtained closed-form solution
+\autoref{eq:zc_sol}. To estimate $\Delta S(t)$ from the data, we defined:
+\begin{align}
+    S_{<}(t) &= \ev{S(z)}_{z \in [z_c - \Delta z - H, z_c - \Delta z]}
+        ,\label{eq:sbelow_def}\\
+    S_>(t) &= \ev{\z{S(z): S(z) < 0}}_{z \in [z_c, z_c + \Delta z]},
+        \label{eq:sabove_def},\\
+    \Delta S(t) &\equiv S_>(t) - S_{<}(t).\label{eq:ds_def}
+\end{align}
+$\ev{\dots}_{z \in [z_a, z_b]}$ denotes averaging over interval $[z_a, z_b]$.
+We average over an interval of length $H = \frac{2\pi}{k_{1z}}$ a full vertical
+wavelength. The offset $\Delta z$ is set by the observation that the critical
+layer's width is limited by $\mathrm{Ri} \lesssim 1$, translating to a critical
+layer width of $\sim \frac{1}{k_{1z}}$. Here, we found $\Delta z =
+\frac{3}{k_z}$ was necessary to be sufficiently offset from strong fluctuations
+near the critical layer.
 
-We argue that the Richardson number is bounded from below by viscosity. A
-repeated simulation with a larger viscosity is shown in TODO,
-where the Richardson number does not go nearly as low.
+Since the transmission feature can be seen in \autoref{fig:nl_fluxes} to
+attenuate rather quickly and is indeed rather weak, it is difficult to establish
+the correct interval to average over; we use the observation that $S(z) < 0$
+only for a small interval for $z$ to select the interval length implicitly.
+
+Finally, we may plot the observed $z_c$ from data against two simple predictors:
+(i) integration of \autoref{eq:zc_anal} using the measured $\Delta S(t)$, and
+(ii) substituting the time-averaged $\ev{\Delta S}_t$ into \autoref{eq:zc_sol}.
+Since $z_c(t)$ is both stable and less well-defined at early times (when the
+critical layer is thick and transient behavior is still strong), we instead
+integrate backwards from the end of the simulation, using $z_c(t_f)$ as initial
+condition. The resulting predictors are depicted in \autoref{fig:nl_front}. The
+good agreement between the evolution of $z_c(t)$ and its estimate via $\Delta
+S(t)$ and \autoref{eq:zc_anal} are noteworthy. Further of interest is the
+discrepancy between the time-averaged predictor and the data. The general
+agreement clearly demonstrates $\Delta S < S_1$ and that incomplete absorption
+is observed. Moreover, its significant underprediction of critical layer motion
+at early times contrasts with its comparatively satisfactory prediction at later
+times to indicate that $\Delta S(t)$ must vary significantly over time. Both of
+these observations indicate a careful characterization of reflectivity is
+necessary.
+
+\begin{figure}[t]
+    \centering
+    \includegraphics[width=\columnwidth]{plots/nl_front.png}
+    \caption{Propagation of the critical layer over time. Shown are (black)
+    $z_c(t)$ from simulation data, (green) predictor of $z_c(t)$ using direct
+    integration of \autoref{eq:zc_anal} for $\Delta S(t)$ measured from
+    simulation data (described in \autoref{eq:ds_def}) and (blue) direct
+    substitution of time-averaged $\ev{\Delta S(t)}_t$ into
+    \autoref{eq:zc_sol}. Predictors use the end of the simulation as initial
+    conditions and integrate backwards, as $z_c$ is less well-defined at early
+    times. The agreement of the directly-integrated predictor with the data
+    shows \autoref{eq:zc_anal} is a good description of the evolution of $z_c$.
+    The the poorer but qualitatively correct agreement of the time-averaged
+    predictor with the data shows both that $\Delta S(t) \neq S_1$ and that
+    $\Delta S(t)$ likely has some very real variation with
+    time.}\label{fig:nl_front}
+\end{figure}
+
+\subsection{Kelvin-Helmholtz Instability and Critical Layer Width}
+
+The observation of reflected/transmitted waves is in accordange with previous
+studies as discussed in \autoref{ss:wave_breaking}. Many studies have shown that
+a local Richardson number $\mathrm{Ri} \sim 1/4$ corresponds to the onset of
+reflectivity. This agrees with the observed Kelvin-Helmholtz instability in our
+simulations [TODO image]. To verify that the critical layer's width is bound
+from below by Kelvin-Helmholtz instabilities, we decided to compute a
+$\mathrm{Ri}$ within the critical layer.
+
+Since fluid instabilities are local, $\mathrm{Ri}$ must be measured in the
+immediate vicinity of a fluid parcel. Thus, we first assign an $\mathrm{Ri}$ for
+every $x$ in the critical layer, then take the median as $\mathrm{Ri}$ for the
+entire layer. To avoid noisiness, the local $\mathrm{Ri}$ is computed using the
+vertical distance over which the local $u_x$ increases from $0.3 \times$ its
+critical value to its critical value. This can be written
+\begin{align}
+    z_{u, \min}(x, t) &= \argmin_\zeta
+        \z{z: u_x(x, \zeta, t) > 0.3\bar{U}_c},\nonumber\\
+    z_{u, \max}(x, t) &= \argmax
+        \zeta \z{z: u_x(x, \zeta, t) < \bar{U}_c},\nonumber\\
+    \mathrm{Ri}(t) &\equiv
+        \med_x\p{\frac{N^2 \p{z_{u, \max} - z_{u, \min}}^2}{(0.7
+            \bar{U}_c)^2}}.\label{ss:ri_med_def}
+\end{align}
+To understand the variation in $\mathrm{Ri}$ over $x$, we can also plot using
+the minimum over $x$; the maximum is significantly noisier. Both of these are
+shown in \autoref{fig:nl_f_ri}. The quick evolution of $\mathrm{Ri}$ to its
+saturated value is reflective of the fact that IGW are \emph{anti-diffusive}.
+This property is a simple consequence of the IGW dispersion relation
+\autoref{eq:disp_rel}, which is approximately $\omega k_z \approx Nk_x$ such
+that as an IGW propagates into a shear flow, $\omega$ decreases and $k_z$
+increases, enhancing dissipation. Dissipation in a momentum-conserving fluid
+then corresponds to further mean flow acceleration.
+\begin{figure}[t]
+    \centering
+    \includegraphics[width=\columnwidth]{plots/nl_f_ri.png}
+    \caption{Local Richardson number of the flow at the critical layer over time
+    as defined in \autoref{ss:ri_med_def}. Dotted red lines demarcate the
+    minimum and maximum of $\mathrm{Ri}$ over $x$ instead of the median. These
+    numbers effectively measure the mean and spread in width of the critical
+    layer over $x$. Note that $\mathrm{Ri} \lesssim \frac{1}{4}$ corresponds
+    to Kelvin-Helmholtz instability, so this plot suggests the critical layer
+    width roughly saturates the lower bound from the Kelvin-Helmholtz
+    instability after an initial transient phase.}\label{fig:nl_f_ri}
+\end{figure}
+
+\subsection{Non-absorption at Critical Layer}\label{ss:reflectivity}
+
+We identify two instances of non-absorptive behavior: (i), the presence of a
+reflected wave with wave vector $\vec{k} = k_{1x}\hat{x} + k_{1z}\hat{z}$, and
+(ii) the amount of horizontal momentum flux $S$ reflected off the critical
+layer.
+
+To measure the reflected wave amplitude, we use almost the same definition
+\autoref{eq:ahat_def} except using $-k_{1z} \to +k_{1z}$; call this estimated
+$\hat{A}_r(t)$ the amplitude of the downwards-propagating reflected wave, and
+call $\hat{A}_i(t)$ the upwards-propagating incident wave. To compute
+$\hat{A}_r(t)$, we furthermore permit an arbitrary phase offset $\phi_r(t)$ at
+each time $t$, since the phase of the reflected wave is unknown, unlike that of
+the incident wave. We observe that the evolution of $\phi_r(t)$ is consistent
+with the interpretation of a reflected wave, as it is well approximated by
+$\pd{\phi_r}{t} \approx 2\pd{(k_{1z}z_c)}{t}$, consistent with a wave reflecting
+off a moving boundary at $z_c$.
+
+Since reflectivity depends sensitively on accurate measurements of $\hat{A}_i,
+\hat{A}_r$, it is important \autoref{eq:ahat_def} identically not pick up
+contributions between the two $k_{1x}\hat{x} \pm k_{1z}\hat{z}$ modes, as we
+argued earlier. The integral for $\hat{A}_r$ is also performed over $z \in [z_0
++ 3\sigma, z_0 + 3\sigma + H]$. Since in general $\hat{A}_i(t), \hat{A}_r(t)$
+are functions of time, we must measure reflectivity accounting for propagation
+time between the measurement zone and the critical layer where the wave
+reflects. For some
+\begin{equation}
+    \delta t(t) \approx \frac{z_c(t) - z_0 + 3\sigma + H/2}{c_{ph, z}},
+        \label{eq:dt_def}
+\end{equation}
+an incident wave measured at $t - \delta t$ reflects at $t$ and is re-measured
+as a reflected wave at $t + \delta t$. The choice of vertical phase velocity
+instead of group velocity is necessary to good results, and a possible
+explanation is provided in \autoref{ss:modes}. We can then define the amplitude
+reflectivity
+\begin{equation}
+    \mathcal{R}_A(t) \equiv \frac{\hat{A}_r(t - \delta t)}{\hat{A}_i(t +
+        \delta t)}.\label{eq:Ra_def}
+\end{equation}
+Since in general $\delta t$ can take on arbitrary values while $\hat{A}_i,
+\hat{A}_r$ are sampled at discrete timesteps set by our simulation output, we
+evaluate $\hat{A}_i, \hat{A}_r$ using a simple linear interpolation to
+intermediate times.
+
+To measure the reflected horizontal momentum flux, we recall from our linear
+simulations described in \autoref{ss:lin_ns} that we are able to accurately
+predict the incident flux from the incident wave amplitude. Thus, we write for
+incident flux
+\begin{equation}
+    S_i(t) \equiv \ev{\rho u_{1x} u_{1z}}_x\hat{A}_i^2(t).
+\end{equation}
+But then, since we already have $\Delta S(t), S_>(t)$ jump and transmitted flux
+across the critical layer at time $t$ respectively, we can immediately write
+down the reflected flux
+\begin{equation}
+    S_d(t) = S_i(t - \delta t) + \Delta S(t) - S_>(t).
+\end{equation}
+$\delta t$ is still \autoref{eq:dt_def}; propagation time between $z_c$ and $z_c
+- 3/k_z - H/2$ the center averaging zone was ignored at the level of accuracy of
+the study. Then we can define flux reflectivity and transmissivity coefficients
+\begin{align}
+    \mathcal{R}_S(t) &\equiv -\frac{S_d(t)}{S_i(t - \delta t)},&
+    \mathcal{T}_S(t) &\equiv -\frac{S_>(t)}{S_i(t - \delta t)}.
+        \label{eq:srefl_def}
+\end{align}
+
+The measurements of $\hat{A}_i, \hat{A}_r, S_i, \Delta S, S_d, S_>$ are given in
+\autoref{fig:nl_f_amps}. The oscillations in $\hat{A}_i, S_i$ are evident and
+neatly correspond to oscillations in $A_d, S_d$ respectively. We further analyze
+these oscillations in \autoref{ss:modes}.
 
@@
     \includegraphics[width=\columnwidth]{plots/nl_f_amps.png}
-    \caption{The top panel measures the incident wave amplitude $A_i(t)$ (green)
-    and the downwards propagating wave amplitude $A_d(t)$ (red) just above the
-    forcing zone, normalized to the analytical estimate \autoref{eq:uz_lin}.
-    $A_d \neq 0$ due to reflection off the critical layer. The bottom panel
-    shows the behavior of three horizontal momentum fluxes over time, in units
-    of the analytical estimate \autoref{eq:S_lin}: (blue) flux incident on the
-    critical layer, (black) flux absorbed by the critical layer, and (green)
-    flux transmitted through the critical layer. A fiducial simulation of
-    $\mathrm{Re} = 1000$ is shown.}\label{fig:nl_amps}
+    \caption{The top panel measures the incident wave amplitude $\hat{A}_i(t)$
+    (green) and the downwards propagating wave amplitude $\hat{A}_r(t)$ (red)
+    just above the forcing zone, normalized to the analytical estimate
+    \autoref{eq:uz_lin}. $A_d \neq 0$ due to reflection off the critical layer.
+    The bottom panel shows the behavior of three horizontal momentum fluxes over
+    time, in units of the analytical estimate \autoref{eq:S_lin}: (blue) flux
+    incident on the critical layer, (red) flux absorbed by the critical layer,
+    and (black) flux transmitted through the critical layer. Note that no
+    propagation time effects have been included in these plots.
+    }\label{fig:nl_f_amps}
 \end{figure}
 
-\begin{figure}[t]
-    \centering
-    \includegraphics[width=\columnwidth]{plots/nl_fluxes.png}
-    \caption{The top and bottom panels show respectively the mean flow
-    $\bar{U}_0$ (see \autoref{eq:mean_flow}) and horizontal momentum flux $S$
-    (see \autoref{eq:S_def}) as a function of $z$ at certain times $t$. The same
-    color coding is used for the two plots. The mean flow is normalized to its
-    critical value $c_{ph,x}$ while again $S$ is normalized to its analytical
-    estimate \autoref{eq:S_lin}.}\label{fig:nl_fluxes}
-\end{figure}
-
-\begin{figure}[t]
-    \centering
-    \includegraphics[width=\columnwidth]{plots/nl_front.png}
-    \caption{Shown is the position of the critical layer (black) as well as two
-    predictors of the critical layer location using the absorbed flux (see the
-    second panel of \autoref{fig:nl_amps}). The green line is generated by
-    integrating \autoref{eq:zc_anal} using $\Delta S$ from the data, while the
-    blue line is generated by using the average value of $\Delta S$ in
-    \autoref{eq:zc_sol}.}\label{fig:nl_front}
-\end{figure}
+Using these measured quantities, we may make plots of $\mathcal{R}_A^2,
+\mathcal{R}_S, \mathcal{T}_S$, which are provided in \autoref{fig:nl_f_refl}. A
+comparison between $\mathcal{R}_A^2$ and $\mathcal{R}_S$ is appropriate as $S
+\propto A^2$. While less evident in the presented plot, a longer-duration
+simulation at lower resolution confirms that the three quantities have reached
+their asymptotic values after $t \gtrsim 1750/N$. An observation may be made
+that in general $\mathcal{R}_S \geq \mathcal{R}_A^2$; this conforms to the
+intuition that reflected flux consists of the simple reflected mode and higher
+order modes as well.
 
@@
     \includegraphics[width=\columnwidth]{plots/nl_f_refl.png}
-    \caption{NL reflectivity}\label{fig:nl_refl}
+    \caption{Reflectivity and transmittivity coefficients for flux and
+    amplitude-squared as described by \autoref{eq:Ra_def} and
+    \autoref{eq:srefl_def} respectively. The coefficients seem to become
+    comparatively stable past about $t = 1750/N$, indicating that an asymptotic
+    value may have been reached.}\label{fig:nl_f_refl}
 \end{figure}
+
+\subsection{Appearance of Discrete Modes}\label{ss:modes}
+
+We briefly attribute the fluctuations in $\hat{A}_i(t)$ as observed in
+\autoref{fig:nl_f_amps} to the introduction of discrete modes by partial
+reflection off a moving critical layer. The partially reflecting boundary causes
+a muted effect of the usual normal modes calculation, so a slightly stronger
+response $\hat{A}_i > 1$ corresponds to being weakly on resonance. The
+instantaneous enhancement/attenuation of $\hat{A}_i$ when $z_c$ moves slowly can
+be computed by an overlap integral of the forcing with the modes of the finite
+domain. This is a consistent interpretation for three reasons:
+\begin{itemize}
+    \item The period of the fluctuations in $\hat{A}_i(t)$ grows longer at later
+        times. This is consistent with $\pd{z_c}{t}$ growing smaller at later
+        times, as the critical layer propagates deeper (see
+        \autoref{eq:zc_anal}).
+
+    \item The amplitude of the fluctuations in $\hat{A}_i(t)$ grows larger at
+        later times. This corresponds to the effective domain (below the
+        critical layer) growing smaller, which leads to stronger boundary
+        effects.
+
+    \item The propagation time $\delta t$, defined in \autoref{eq:dt_def}, used
+        the vertical phase velocity instead of the vertical group velocity to
+        obtain a non-oscillating reflectivity. This is consistent with the
+        fluctuations in $\hat{A}_i(t)$ being induced by internal modes of the
+        domain rather than a varying forcing strength, as the latter should
+        propagate at the \emph{group velocity} instead.
+\end{itemize}
 
@@
 
+In the previous section, we have argued for a continous train of breaking IGWs
+spontaneously forming a critical layer and strong shear flow. We have
+parameterized the width of the critical layer as well as horizontal momentum
+transport near the critical layer. In the subsequent sections, we will discuss
+the validity of these results and their application to astrophysical systems.
+
+\subsection{Convergence}\label{ss:convergence}
+
+As the primary test of convergence in our previous sections, we consider the
+convergence of the physically significant parameters of our model. In
+particular, the convergence of $\mathrm{Ri}$, shown in \autoref{fig:nl_f_ri},
+and of the reflection/transmission coefficient asymptotic values, shown in
+\autoref{fig:nl_f_refl}, are of greatest significance.
+
+To estimate convergence, we compute the median value of each of $\mathrm{Ri},
+\mathcal{R}_A^2, \mathcal{R}_S, \mathcal{T}_S$ over the last $1/4$ of simulation
+times from each simulation, where all simulations had converged to what appeared
+to be asymptotic values. Error bars are estimated with the $16$ and $84$
+percentiles. We illustrate the convergence of these averages across simulations
+in \autoref{fig:agg}.
 \begin{figure}[t]
@@
     \includegraphics[width=\columnwidth]{plots/agg.png}
-    \caption{Convergence w/ all runs}\label{fig:agg}
+    \caption{Convergence of median reflection/transmission coefficients and
+    $\mathrm{Ri}$ (critical layer width) across runs with varying viscosity,
+    parameterized with $\mathrm{Re}$ (\autoref{eq:re_def}). Error bars depict
+    $16$ and $84$ percentile values. Small horizontal displacements are made for
+    data points at identical $\mathrm{Re}$ for readability. Note that
+    simulations with larger $\mathrm{Re}$ correspond to smaller viscosity and
+    are more physically realistic, and our values seem to converge towards large
+    $\mathrm{Re}$.}\label{fig:agg}
 \end{figure}
+
+\subsection{Physical Sources of Dissipation in WDs}\label{ss:disp}
+
+The most significant linear damping in WD g-modes comes from radiative damping
+\citep{fullerI}. In \citep{wu} and \citep{fullerI}, the radiative damping rate
+is given in terms of $\omega_i = \gamma \omega_r$, where $\omega_r$ is the
+frequency of the g-mode. Typical values for $\gamma$ range from $10^{-4}$ to
+$10^{-11}$ depending on $n$ of the g-mode.
+
+We will assume this prescription directly transfers to propagating IGW, which
+results in general agreement with \citep{bukart}'s estimate of radiative damping
+rates. Then, making coarse identification $\omega_i \sim \nu k^2 \approx \nu
+k_z^2$, we find that $\mathrm{Re} \sim \frac{1}{\gamma}$. Even at $\gamma =
+10^{-4}$ however, the corresponding $\mathrm{Re}$ is far too weak to suppress
+reflection/transmission at the critical layer (e.g.\ \autoref{fig:agg}).
+
+Another source of dissipation considered in \citep{bukart} is turbulent
+convective damping. They find this damping rate to never exceed that of
+radiative damping, and so it is also too weak to suppress critical layer
+formation in our problem.
+
+Finally,
 
@@
 
-Discussion foo
+\subsection{Energy Transfer}
+
+[TODO elaborate? Need to make plots to check?]
+
+Our equations do not conserve energy, but it seems like more energy can be
+transmitted in higher modes as viscosity is decreased. Nevertheless, a
+significant fraction should still be dissipated in the critical layer since a
+significant energy cascade must happen in the critical layer.
 
@@
 We denote $x \in [0, L_x], z \in [0, L_z]$ the simulation domain and $N_x, N_z$
-the number of spectral modes in the respective dimensions. We perform direct
-numerical simulation of \autoref{eq:vol_drive} with the open-source
-pseudo-spectral code Dedalus (CITE).
+the number of spectral modes in the respective dimensions.
 
@@
     \begin{align}
-        \vec{\nabla} \cdot \vec{u} &= 0,\\
-        \pd{\Upsilon}{t} + \p{\vec{u} \cdot \vec{\nabla}} \Upsilon
+        \vec{\nabla} \cdot \vec{u}_1 &= 0,\\
+        \pd{\Upsilon}{t} + \p{\vec{u}_1 \cdot \vec{\nabla}} \Upsilon
             - \frac{u_z}{H} &= 0,\\
-        \pd{u_x}{t} + \p{\vec{u} \cdot \vec{\nabla}}u_x
+        \pd{u_{1x}}{t} + \p{\vec{u}_1 \cdot \vec{\nabla}}u_{1x}
             + \pd{\varpi_1}{x} + gH\pd{\Upsilon}{x}
             + \varpi_1 \pd{\Upsilon}{x} &= 0,\\
-        \pd{u_z}{t} + \p{\vec{u} \cdot \vec{\nabla}}u_z
+        \pd{u_z}{t} + \p{\vec{u}_1 \cdot \vec{\nabla}}u_z
             + \pd{\varpi_1}{z} + gH\pd{\Upsilon}{z}
@@
             \vec{\nabla}(\rho - \rho_0)) &= 0,\label{eq:visc_cons_mom}\\
-        \partial_t (\rho \vec{u}) + \vec{\nabla} \cdot (\rho \vec{u} \vec{u} +
+        \partial_t (\rho \vec{u}) + \vec{\nabla} \cdot (\rho \vec{u}
+        \vec{u} +
             \mathrm{diag}(\rho \varpi) - \nu \rho \vec{\nabla}\vec{u}) + \rho g

### 2019-04-30 sims/2d_3_final/strat.py

@@ -74,31 +74,31 @@ if __name__ == '__main__':
                               'N_X': 64,
                               'N_Z': 256,
                               'Re_inv': 1e-4})),
-       ('lin_1_masked',
-        get_params(overrides={'F_MULT': 0.01,
-                              'N_X': 64,
-                              'N_Z': 256,
-                              'Re_inv': 0.075})),
-       ('lin_2',
-        get_params(overrides={'F_MULT': 0.0005,
-                              'T_F': 4000,
-                              'DT': 0.1,
-                              'NL': False,
-                              'N_X': 64,
-                              'N_Z': 256,
-                              'Re_inv': 0})),
-       ('nl_1_masked',
-        get_params(overrides={'F_MULT': 1,
-                              'Re_inv': 0.7})),
-       ('nl_3_masked',
-        get_params(overrides={'F_MULT': 1,
-                              'Re_inv': 0.5})),
-       ('nl_4_masked',
-        get_params(overrides={'F_MULT': 1,
-                              'Re_inv': 0.3})),
-       ('nl_6_masked',
-        get_params(overrides={'F_MULT': 1,
-                              'Re_inv': 0.2})),
+       # ('lin_1_masked',
+       #  get_params(overrides={'F_MULT': 0.01,
+       #                        'N_X': 64,
+       #                        'N_Z': 256,
+       #                        'Re_inv': 0.075})),
+       # ('lin_2',
+       #  get_params(overrides={'F_MULT': 0.0005,
+       #                        'T_F': 4000,
+       #                        'DT': 0.1,
+       #                        'NL': False,
+       #                        'N_X': 64,
+       #                        'N_Z': 256,
+       #                        'Re_inv': 0})),
+       # ('nl_1_masked',
+       #  get_params(overrides={'F_MULT': 1,
+       #                        'Re_inv': 0.7})),
+       # ('nl_3_masked',
+       #  get_params(overrides={'F_MULT': 1,
+       #                        'Re_inv': 0.5})),
+       # ('nl_4_masked',
+       #  get_params(overrides={'F_MULT': 1,
+       #                        'Re_inv': 0.3})),
+       # ('nl_6_masked',
+       #  get_params(overrides={'F_MULT': 1,
+       #                        'Re_inv': 0.2})),
     ]
     if '-plot' in sys.argv:
         for name, params_dict in tasks:

### 2019-04-30 sims/2d_3_final/strat_helper.py

@@ -14,8 +14,6 @@ import matplotlib
 matplotlib.use('Agg')
 import matplotlib.pyplot as plt
 import numpy as np
-plt.rc('text', usetex=True)
-plt.rc('font', family='serif')
 from scipy.interpolate import interp1d
 from scipy.optimize import minimize
 
@@ -32,6 +30,11 @@ FILENAME_EXPR = '{s}/{s}_s{idx}.h5'
 Z_TOP_MULT = 1
 STRIDE = 15
 AVG_IDX = 4
+FONTSIZE = 12
+# cp snapshots_lin_0_masked/f_amps.png ../../writeups/plots/lin_amps.png; cp snapshots_lin_0_masked/fluxes.png ../../writeups/plots/lin_fluxes.png
+
+plt.rc('text', usetex=True)
+plt.rc('font', family='serif', size=FONTSIZE)
 
 def populate_globals(var_dict):
     for key, val in var_dict.items():
@@ -982,10 +985,10 @@ def plot_front(name, params):
         #          label=r'Model')
         plt.xlim(z_b, ZMAX)
         plt.ylim(-0.1, 1.1)
-        plt.legend(fontsize=6)
+        plt.legend(fontsize=FONTSIZE)
 
         plt.xlabel(r'$z(H)$')
-        plt.ylabel(r'$S / S_0$')
+        plt.ylabel(r'$\hat{S}$')
         plt.savefig('%s/fluxes.png' % snapshots_dir, dpi=400)
         plt.close()
 
@@ -1005,9 +1008,9 @@ def plot_front(name, params):
         #          'r',
         #          label=r'$A_d$',
         #          linewidth=0.7)
-        # ax1.legend(fontsize=6)
+        # ax1.legend(fontsize=FONTSIZE)
         ax1.set_xlabel(r'$t (N^{-1})$')
-        ax1.set_ylabel(r'$A/A_0$')
+        ax1.set_ylabel(r'$\hat{A}$')
         # ax2.plot(t,
         #          np.unwrap(phis_down[start_idx: ]),
         #          'r',
@@ -1064,10 +1067,10 @@ def plot_front(name, params):
         ax2.set_xlim(z_b, ZMAX)
         ax1.set_ylim(-0.2, 1.25)
         ax2.set_ylim(-0.2, 1.2)
-        ax2.legend(fontsize=6)
+        ax2.legend(fontsize=FONTSIZE)
 
[…]

### 2019-04-30 sims/2d_4_fourier/strat.py

@@ -67,30 +67,30 @@ def get_params(overrides=None):
 
 if __name__ == '__main__':
     tasks = [
-        ('yubo_nu0p5_hres',
-         get_params(overrides={'Re_inv': 0.05})),
+        # ('yubo_nu0p5_hres',
+        #  get_params(overrides={'Re_inv': 0.05})),
         ('yubo_nu0p5_shres',
          get_params(overrides={'Re_inv': 0.05,
                                'N_X': 1024,
                                'N_Z': 4096})),
-        ('yubo_nu1_hres',
-         get_params(overrides={'Re_inv': 0.1})),
-        ('yubo_nu1_vhres',
-         get_params(overrides={'Re_inv': 0.1,
-                               'N_X': 768,
-                               'N_Z': 3072})),
-        ('yubo_nu2_hres',
-         get_params(overrides={'Re_inv': 0.2})),
+        # ('yubo_nu1_hres',
+        #  get_params(overrides={'Re_inv': 0.1})),
+        # ('yubo_nu1_vhres',
+        #  get_params(overrides={'Re_inv': 0.1,
+        #                        'N_X': 768,
+        #                        'N_Z': 3072})),
+        # ('yubo_nu2_hres',
+        #  get_params(overrides={'Re_inv': 0.2})),
         # ('yubo_nu2_width',
         #  get_params(overrides={'Re_inv': 0.2,
         #                        'N_X': 256,
         #                        'NUM_SNAPSHOTS': 1500,
         #                        'STRIDE': 2,
         #                        'N_Z': 1024})),
-        ('yubo_nu3_width',
-         get_params(overrides={'Re_inv': 0.3,
-                               'N_X': 256,
-                               'N_Z': 1024})),
+        # ('yubo_nu3_width',
+        #  get_params(overrides={'Re_inv': 0.3,
+        #                        'N_X': 256,
+        #                        'N_Z': 1024})),
     ]
     if '-plot' in sys.argv:
         for name, params_dict in tasks:

### 2019-04-30 sims/2d_4_fourier/strat_helper.py

@@ -14,8 +14,6 @@ import matplotlib
 matplotlib.use('Agg')
 import matplotlib.pyplot as plt
 import numpy as np
-plt.rc('text', usetex=True)
-plt.rc('font', family='serif')
 from scipy.interpolate import interp1d
 from scipy.optimize import minimize
 
@@ -32,6 +30,11 @@ FILENAME_EXPR = '{s}/{s}_s{idx}.h5'
 Z_TOP_MULT = 1
 STRIDE = 15
 AVG_IDX = 4
+FONTSIZE = 12
+
+plt.rc('text', usetex=True)
+plt.rc('font', family='serif', size=FONTSIZE)
+# cd *shres*; for i in f*.png; do cp $i ../../../writeups/plots/nl_$i; done
 
 def populate_globals(var_dict):
     for key, val in var_dict.items():
@@ -1006,10 +1009,10 @@ def plot_front(name, params):
         #          label=r'Model')
         plt.xlim(z_b, ZMAX)
         plt.ylim(-0.1, 1.1)
-        plt.legend(fontsize=6)
+        plt.legend(fontsize=FONTSIZE)
 
         plt.xlabel(r'$z(H)$')
-        plt.ylabel(r'$S / S_0$')
+        plt.ylabel(r'$\hat{S}$')
         plt.savefig('%s/fluxes.png' % snapshots_dir, dpi=400)
         plt.close()
 
@@ -1029,9 +1032,9 @@ def plot_front(name, params):
         #          'r',
         #          label=r'$A_d$',
         #          linewidth=0.7)
-        # ax1.legend(fontsize=6)
+        # ax1.legend(fontsize=FONTSIZE)
         ax1.set_xlabel(r'$t (N^{-1})$')
-        ax1.set_ylabel(r'$A/A_0$')
+        ax1.set_ylabel(r'$\hat{A}$')
         # ax2.plot(t,
         #          np.unwrap(phis_down[start_idx: ]),
         #          'r',
@@ -1088,10 +1091,10 @@ def plot_front(name, params):
         ax2.set_xlim(z_b, ZMAX)
         ax1.set_ylim(-0.2, 1.25)
         ax2.set_ylim(-0.2, 1.2)
-        ax2.legend(fontsize=6)
+        ax2.legend(fontsize=FONTSIZE)
 
-        ax1.set_ylabel(r'$U_0 / c_{ph, x}$')
[…]

### 2019-04-30 sims/aggregate.py

@@ -7,7 +7,7 @@ import matplotlib
 matplotlib.use('Agg')
 import matplotlib.pyplot as plt
 plt.rc('text', usetex=True)
-plt.rc('font', family='serif')
+plt.rc('font', family='serif', size=12)
 import matplotlib.lines as mlines
 
 DATA = [
@@ -37,6 +37,7 @@ if __name__ == '__main__':
                  (T_med, T_min, T_max),
                  (w_med, w_min, w_max)) in DATA:
         # Reynolds number for linear mode
+        print(1024 / 10 / re_inv)
         x = (1024 / 10) / re_inv + offsets[re_inv]
         ax1.plot(x, r_med, 'ko', markersize=msize)
         ax1.errorbar(x, r_med,
@@ -66,7 +67,7 @@ if __name__ == '__main__':
                         label=r'$\left<\mathcal{R}_A(t)^2\right>$')
     ln3 = mlines.Line2D([], [], color='r', marker='*', markersize=msize,
                         label=r'$\left<\mathcal{T}_S(t)\right>$')
-    ax1.legend(handles=[ln1, ln2, ln3], fontsize=6, loc='upper left')
+    ax1.legend(handles=[ln1, ln2, ln3], fontsize=12, loc='upper left')
 
 
     ax1.set_ylabel(r'$\left<\mathcal{R}_S(t)\right>$')

### 2019-04-30 writeups/paper.tex
check in writeup w/ fft

@@
     \pd{}{t}\int\limits_0^{L_z} \rho(z) \bar{U}(z, t)\;\mathrm{d}z
-        \Delta S = 0.
+        + \Delta S = 0.
 \end{equation}
@@
 
+Finally, we may be interested in the power spectrum of the transmitted
+horizontal momentum flux. Define:
+\begin{align}
+    \tilde{S}_>(k_x, t) = \abs{
+        \int\limits_0^{L_x} e^{ik_xx}\hat{S}_>(x, z_c - \Delta z, t)\;\mathrm{d}x}
+            \label{eq:sfft_above},\\
+    \tilde{S}_<(k_x, t) = \abs{
+        \int\limits_0^{L_x} e^{ik_xx}\hat{S}_<(x, z_c + \Delta z, t)\;\mathrm{d}x}.
+            \label{eq:sfft_below}
+\end{align}
+These values are plotted in \autoref{fig:nl_fft}. Note that we expect
+$\tilde{S}_<(0, t) = S_<(t) = S_i(t) + S_d(t)$ and the same for $\tilde{S}_>$.
+This is rather accurate for $\tilde{S}_<$ but less so for $\tilde{S}_>$; this is
+since $S_>(t)$ decays very rapidly above $z_c$. Note furthermore that at later
+times, the transmitted flux increases, per expectation, and shifts towards
+higher $k_x$. Repeating this plot at higher viscosities shows significantly
+lower $\tilde{S}_>$, showing that the $k_x$ of the transmitted waves still
+changes greatly at the highest $\mathrm{Re}$ we are able to probe. Finally, note
+that the area under $\tilde{S}$ changes with time; this is okay since we are
+measuring $S$ at constant position, and different $k_x$ components of $S$ will
+translate at their separate group velocities.
+
+\begin{figure}[t]
+    \centering
+    \includegraphics[width=\columnwidth]{plots/nl_fft.png}
+    \caption{The power spectrum of the horizontal momentum flux at fixed
+    distances below and above critical layer $z_c$ at certain times $t$ for
+    simulation \texttt{0p5\_shres}. \autoref{eq:sfft_above} and
+    \autoref{eq:sfft_below} are depicted here, with $\tilde{S}_<$ in solid lines
+    and $\tilde{S}_>$ in dashed lines. Note that there is more incident flux
+    than transmitted, and that transmitted flux roughly increases over
+    time.}\label{fig:nl_fft}
+\end{figure}
+
 \subsection{Appearance of Discrete Modes}\label{ss:modes}
@@
 
-Finally,
-
-\subsection{Applicability to Astrophysical Systems}
-
-% magnetic winding, radiative diffusivity + effective Re. Might be applicable to
-% inwards focusing too
-
-\subsection{Energy Transfer}
+Finally, we consider the impact of magnetic winding. In \citep{bukart}, magnetic
+winding is used to enforce solid body rotation on the grounds that $t_A \gg
+t_{gw}$, where
+\begin{equation}
+    t_A = \int\limits_0^R \frac{\sqrt{4\pi \rho}}{B_0}\;\mathrm{d}r
+        \sim 10^2\;\mathrm{yr}\p*{\frac{10^3\;\mathrm{G}}{B}}.
+\end{equation}
+the Alfv\'en wave crossing time (evaluated for a CO WD in \citep{fullerIV})
+measures the magnetic coupling time and $t_{gw}$ measures the gravitational wave
+inspiral timescale. Before solid body rotation is attained, another relevant
+timescale is the synchronization timescale $t_s$. For a tidal torque $\tau$ and
+tidal forcing frequency $\sigma = m\p*{\Omega - \Omega_{spin}}$, we note that
+angular momentum transfer is $\pd{M_{sync}}{t} \sigma R^2 = \tau$, where
+$M_{sync}$ denotes the mass of the WD that has synchronized. Thus, the
+synchronization timescale is
+\begin{align}
+    t_{sync} &\sim \frac{M_{sync}\sigma R^2}{\tau},\nonumber\\
+        &\sim 2 \times 10^5\;\mathrm{yr}
+            \p{\frac{M}{M_{\odot}}}
+            \p{\frac{\sigma}{2\pi / (1\;\mathrm{hr})}}
+            \p{\frac{R}{R_{\oplus}}}^2
+            \p{\frac{10^{-14} GM_{\odot}^2/R_{\oplus}}{\tau}}.
+\end{align}
+A representative $\tau$ has been taken from \citep{bukart}. However, since only
+the outer $\sim 10^{-4}M_{\odot}$ need be heated for the thermodynamically
+interesting effects studied in \citep{fullerIV} and \citep{tidal_novae}, it
+seems that magnetic winding cannot absolutely rule out energetic outbursts such
+as tidal novae resulting from strong shear flows.
+
+\subsection{Applicability to Other Astrophysical Systems}
+
+[TODO flesh out]
+
+\begin{itemize}
+    \item $k_x, k_z$: In astrophysical systems, $k_{\perp} \ll k_r$. While we do
+        not explore different $k_{1x}, k_{1z}, \omega_1$ in this study, with
+        outgoing boundary conditions, there appears at first to be no $z$ length
+        scale other than $k_{1z}$, so our results would seem to be invariant
+        under rescaling of the $z$ length scale. However, true turbulence is
+        expected to be isotropic at small scales, which may couple $k_{1x},
+        k_{1z}$ in a way that $k_{1x} \ll k_{1z}$ produces different dynamics
+        than $k_{1x} \lesssim k_{1z}$ as we've studied. This is a numerically
+        difficult regime though, so we defer consideration to future work.
+
+    \item Validity of plane-parallel approximation? We're all at $\gtrsim
+        0.9R_{WD}$.
+
+    \item Solar-type stars (inner convective, outer radiative): different
+        equation of state/stratification but could be qualitatively similar.
+
+    \item Solar-type stars: In~\cite{barker_ogilvie}, inwards-propagating IGW
+        are excited that break via geometric focusing and effect
+        synchronization. They find no reflected wave despite their nonlinear
+        timescales being $10\times$ shorter than their viscous timescale.
+
+        It is not immediately clear whether our results here apply when the flux
+        is geometrically focused, but as a hypothesis we assume the convergence
+        in \autoref{fig:agg} applies under geometric focusing as a zeroth
+        approximation, perhaps as a property of the fluid motion within the
+        geometrically thin critical layer.
+
+        Associating $t_{} \sim \nu k^2 \approx \nu k_z^2$ with the visocus
+        timescale and $t_{NL} \sim \vec{u} \cdot \vec{\nabla} \sim \omega$ for
+        the nonlinear timescale, we find their $\lambda = \frac{t_{NL}}{t_L}
+        \sim \mathrm{Re}$ our Reynolds number. Our simulations indicate
+        $\mathrm{Re} \gtrsim 500$ are required to observe the correct asymptotic
+        behavior in terms of horizontal momentum flux reflection/transmission,
+        so it is possible their lack of reflection is viscosity limited.
+\end{itemize}
+
+\subsection{Heating}
 

### 2019-04-30 sims/2d_4_fourier/strat_helper.py

@@ -1129,7 +1129,7 @@ def plot_front(name, params):
         # ax1.plot(t,
         #          smooth(dSpx0),
         #          '%s-' % PLT_COLORS[color_idx],
-        #          label=r'$\Delta S_{0}|_{z=z_{c}}$',
+        #          label=r'$\Delta S_{1}|_{z=z_{c}}$',
         #          linewidth=0.7)
         # color_idx += 1
         # ax1.plot(t,
@@ -1165,7 +1165,7 @@ def plot_front(name, params):
         ax2.plot(t,
                  pos_anal,
                  'b:',
-                 label='Avg Absorbed $%.2fS_{0}$' %
+                 label='Avg Absorbed $%.2fS_{1}$' %
                     (mean_incident / flux_th),
                  linewidth=0.7)
         ax2.set_ylabel(r'$z_c$')
@@ -1327,7 +1327,7 @@ def plot_front(name, params):
                  '%s:' % color,
                    linewidth=0.5)
     ax1.set_ylabel(r'$\tilde{S}$')
-    ax1.set_xlabel(r'$k_x/k_{x0}$')
+    ax1.set_xlabel(r'$k_x/k_{x1}$')
     ax1.legend(fontsize=FONTSIZE)
     plt.savefig('%s/fft.png' % snapshots_dir, dpi=400)
     plt.close()

### 2019-05-29 sims/2d_4_fourier/strat.py
ugh

@@ -11,7 +11,7 @@ H = 1
 XMAX = 4 * H
 ZMAX = 10 * H
 
-NUM_SNAPSHOTS = 300
+NUM_SNAPSHOTS = 500
 TARGET_DISP_RAT = 0.1 # k_z * u_z / omega at base
 
 PARAMS_DEFAULT = {'XMAX': XMAX,
@@ -32,7 +32,6 @@ PARAMS_DEFAULT = {'XMAX': XMAX,
                   'SPONGE_WIDTH': 0.05 * ZMAX,
                   'SPONGE_HIGH': 0.95 * ZMAX,
                   'SPONGE_LOW': 0.03 * ZMAX,
-                  'STRIDE': 1,
 
                   'NUM_SNAPSHOTS': NUM_SNAPSHOTS,
                   'NL': True,
@@ -46,7 +45,7 @@ def get_params(overrides=None):
     OMEGA = get_omega(g, H, KX, KZ)
     VG_Z = get_vgz(g, H, KX, KZ)
 
-    PARAMS_DEFAULT['T_F'] = abs(ZMAX / VG_Z) * 2
+    PARAMS_DEFAULT['T_F'] = abs(ZMAX / VG_Z) * 3
     PARAMS_DEFAULT['OMEGA'] = OMEGA
     PARAMS_DEFAULT['S'] = PARAMS_DEFAULT['ZMAX'] / 512 * 4
     PARAMS_DEFAULT['DT'] = min(0.1 / OMEGA, 0.1)
@@ -67,30 +66,37 @@ def get_params(overrides=None):
 
 if __name__ == '__main__':
     tasks = [
-        # ('yubo_nu0p5_hres',
-        #  get_params(overrides={'Re_inv': 0.05})),
+        ('yubo_nu0p5_hres',
+         get_params(overrides={'Re_inv': 0.05})),
         ('yubo_nu0p5_shres',
          get_params(overrides={'Re_inv': 0.05,
                                'N_X': 1024,
                                'N_Z': 4096})),
-        # ('yubo_nu1_hres',
-        #  get_params(overrides={'Re_inv': 0.1})),
-        # ('yubo_nu1_vhres',
-        #  get_params(overrides={'Re_inv': 0.1,
-        #                        'N_X': 768,
-        #                        'N_Z': 3072})),
-        # ('yubo_nu2_hres',
-        #  get_params(overrides={'Re_inv': 0.2})),
+        ('yubo_nu1_hres',
+         get_params(overrides={'Re_inv': 0.1})),
[…]

### 2019-07-24 writeups/paper.tex
ugh
check in new files, stash

@@
 \maketitle
+
+% cp ../sims/agg.png plots
+% cp ../sims/2d_3_final/snapshots_lin_0_masked/f_amps.png plots/lin_amps.png
+% cp ../sims/2d_3_final/snapshots_lin_0_masked/fluxes.png plots/lin_fluxes.png
+% cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/f_amps.png plots/nl_f_amps.png
+% cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/fluxes.png plots/nl_fluxes.png
+% cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/f_refl.png plots/nl_f_refl.png
+% cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/f_ri.png plots/nl_f_ri.png
+% cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/front.png plots/nl_front.png
 
@@
 
-Finally, we may be interested in the power spectrum of the transmitted
-horizontal momentum flux. Define:
-\begin{align}
-    \tilde{S}_>(k_x, t) = \abs{
-        \int\limits_0^{L_x} e^{ik_xx}\hat{S}_>(x, z_c - \Delta z, t)\;\mathrm{d}x}
-            \label{eq:sfft_above},\\
-    \tilde{S}_<(k_x, t) = \abs{
-        \int\limits_0^{L_x} e^{ik_xx}\hat{S}_<(x, z_c + \Delta z, t)\;\mathrm{d}x}.
-            \label{eq:sfft_below}
-\end{align}
-These values are plotted in \autoref{fig:nl_fft}. Note that we expect
-$\tilde{S}_<(0, t) = S_<(t) = S_i(t) + S_d(t)$ and the same for $\tilde{S}_>$.
-This is rather accurate for $\tilde{S}_<$ but less so for $\tilde{S}_>$; this is
-since $S_>(t)$ decays very rapidly above $z_c$. Note furthermore that at later
-times, the transmitted flux increases, per expectation, and shifts towards
-higher $k_x$. Repeating this plot at higher viscosities shows significantly
-lower $\tilde{S}_>$, showing that the $k_x$ of the transmitted waves still
-changes greatly at the highest $\mathrm{Re}$ we are able to probe. Finally, note
-that the area under $\tilde{S}$ changes with time; this is okay since we are
-measuring $S$ at constant position, and different $k_x$ components of $S$ will
-translate at their separate group velocities.
-
-\begin{figure}[t]
-    \centering
-    \includegraphics[width=\columnwidth]{plots/nl_fft.png}
-    \caption{The power spectrum of the horizontal momentum flux at fixed
-    distances below and above critical layer $z_c$ at certain times $t$ for
-    simulation \texttt{0p5\_shres}. \autoref{eq:sfft_above} and
-    \autoref{eq:sfft_below} are depicted here, with $\tilde{S}_<$ in solid lines
-    and $\tilde{S}_>$ in dashed lines. Note that there is more incident flux
-    than transmitted, and that transmitted flux roughly increases over
-    time.}\label{fig:nl_fft}
-\end{figure}
+% Finally, we may be interested in the power spectrum of the transmitted
+% horizontal momentum flux. Define:
+% \begin{align}
+%     \tilde{S}_>(k_x, t) = \abs{
+%         \int\limits_0^{L_x} e^{ik_xx}\hat{S}_>(x, z_c - \Delta z, t)\;\mathrm{d}x}
+%             \label{eq:sfft_above},\\
+%     \tilde{S}_<(k_x, t) = \abs{
+%         \int\limits_0^{L_x} e^{ik_xx}\hat{S}_<(x, z_c + \Delta z, t)\;\mathrm{d}x}.
+%             \label{eq:sfft_below}
+% \end{align}
+% These values are plotted in \autoref{fig:nl_fft}. Note that we expect
+% $\tilde{S}_<(0, t) = S_<(t) = S_i(t) + S_d(t)$ and the same for $\tilde{S}_>$.
+% This is rather accurate for $\tilde{S}_<$ but less so for $\tilde{S}_>$; this is
+% since $S_>(t)$ decays very rapidly above $z_c$. Note furthermore that at later
+% times, the transmitted flux increases, per expectation, and shifts towards
+% higher $k_x$. Repeating this plot at higher viscosities shows significantly
+% lower $\tilde{S}_>$, showing that the $k_x$ of the transmitted waves still
+% changes greatly at the highest $\mathrm{Re}$ we are able to probe. Finally, note
+% that the area under $\tilde{S}$ changes with time; this is okay since we are
+% measuring $S$ at constant position, and different $k_x$ components of $S$ will
+% translate at their separate group velocities.
 
@@
 \end{itemize}
-
-\section{Discussion}\label{s:discussion}
-
-In the previous section, we have argued for a continous train of breaking IGWs
-spontaneously forming a critical layer and strong shear flow. We have
-parameterized the width of the critical layer as well as horizontal momentum
-transport near the critical layer. In the subsequent sections, we will discuss
-the validity of these results and their application to astrophysical systems.
 
@@
 
+\section{Discussion}\label{s:discussion}
+
+In the previous section, we have argued for a continous train of breaking IGWs
+spontaneously forming a critical layer and strong shear flow. We have
+parameterized the width of the critical layer as well as horizontal momentum
+transport near the critical layer. In the subsequent sections, we will discuss
+the validity of these results and their application to astrophysical systems.
+
 \subsection{Physical Sources of Dissipation in WDs}\label{ss:disp}
@@
 described in \autoref{ss:numerics}, we finally obtain the full system of
-equations as simulated in Dedalus:
+equations as simulated in Dedalus: [TODO update w/ mask]
 \begin{subequations}

### 2019-07-24 sims/2d_4_fourier/strat.py

@@ -66,14 +66,14 @@ def get_params(overrides=None):
 
 if __name__ == '__main__':
     tasks = [
-        ('yubo_nu0p5_hres',
-         get_params(overrides={'Re_inv': 0.05})),
+        # ('yubo_nu0p5_hres',
+        #  get_params(overrides={'Re_inv': 0.05})),
         ('yubo_nu0p5_shres',
          get_params(overrides={'Re_inv': 0.05,
                                'N_X': 1024,
                                'N_Z': 4096})),
-        ('yubo_nu1_hres',
-         get_params(overrides={'Re_inv': 0.1})),
+        # ('yubo_nu1_hres',
+        #  get_params(overrides={'Re_inv': 0.1})),
         ('yubo_nu1_vhres',
          get_params(overrides={'Re_inv': 0.1,
                                'N_X': 768,
@@ -111,8 +111,15 @@ if __name__ == '__main__':
             write_front(name, params_dict)
 
     elif '-front' in sys.argv:
+        out = []
         for name, params_dict in tasks:
-            plot_front(name, params_dict)
+            ret = plot_front(name, params_dict)
+            out.append('(%.2f, (%s))' %
+                       (params_dict['Re_inv'],
+                        ', '.join(['(%s)' % ', '.join(['%.4f' % i for i in j])
+                                   for j in ret])))
+        for out_line in out:
+            print(out_line)
 
     else:
         for task in tasks:

### 2019-07-24 sims/2d_4_fourier/strat_helper.py

@@ -31,6 +31,7 @@ Z_TOP_MULT = 1
 STRIDE = 15
 AVG_IDX = 4
 FONTSIZE = 12
+DPI=600
 
 plt.rc('text', usetex=True)
 plt.rc('font', family='serif', size=FONTSIZE)
@@ -598,7 +599,7 @@ def plot(name, params, stride=STRIDE):
         ux_est = uz_est * KZ / KX
 
         for t_idx, time in list(enumerate(sim_times)):
-            fig = plt.figure(dpi=400)
+            fig = plt.figure(dpi=DPI)
 
             uz_anal = get_anal_uz(params, time, x, z)
             ux_anal = get_anal_ux(params, time, x, z)
@@ -1013,7 +1014,7 @@ def plot_front(name, params):
 
         plt.xlabel(r'$z(H)$')
         plt.ylabel(r'$\hat{S}$')
-        plt.savefig('%s/fluxes.png' % snapshots_dir, dpi=400)
+        plt.savefig('%s/fluxes.png' % snapshots_dir, dpi=DPI)
         plt.close()
 
         #####################################################################
@@ -1045,7 +1046,7 @@ def plot_front(name, params):
         #          'k',
         #          label=r'$\phi_I$',
         #          linewidth=0.7)
-        plt.savefig('%s/f_amps.png' % snapshots_dir, dpi=400)
+        plt.savefig('%s/f_amps.png' % snapshots_dir, dpi=DPI)
         plt.close()
         avg_refl, avg_reflA, avg_ri, avg_trans = tuple([tuple([0, 0, 0])] * 4)
 
@@ -1096,7 +1097,7 @@ def plot_front(name, params):
         ax1.set_ylabel(r'$\bar{U} / c_{ph, x}$')
         ax2.set_ylabel(r'$\hat{S}$')
         ax2.set_xlabel(r'$z(H)$')
-        plt.savefig('%s/fluxes.png' % snapshots_dir, dpi=400)
+        plt.savefig('%s/fluxes.png' % snapshots_dir, dpi=DPI)
         plt.close()
 
         #####################################################################
@@ -1172,7 +1173,7 @@ def plot_front(name, params):
         ax2.set_xlabel(r'$t (N^{-1})$')
         ax2.set_ylim([zf, np.max(front_pos[start_idx: ])])
         ax2.legend(fontsize=FONTSIZE, loc='upper right')
-        plt.savefig('%s/front.png' % snapshots_dir, dpi=400)
[…]

### 2019-07-29 writeups/paper.tex
some first edits

@@
     \newcommand*{\z}[1]{\left\{#1\right\}}
+    \newcommand*{\bm}[1]{\boldsymbol{\mathbf{#1}}}
     \DeclareMathOperator*{\argmin}{argmin}
@@
 
-% cp ../sims/agg.png plots
-% cp ../sims/2d_3_final/snapshots_lin_0_masked/f_amps.png plots/lin_amps.png
-% cp ../sims/2d_3_final/snapshots_lin_0_masked/fluxes.png plots/lin_fluxes.png
-% cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/f_amps.png plots/nl_f_amps.png
-% cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/fluxes.png plots/nl_fluxes.png
-% cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/f_refl.png plots/nl_f_refl.png
-% cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/f_ri.png plots/nl_f_ri.png
-% cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/front.png plots/nl_front.png
+% cp ../sims/agg.png plots;
+% cp ../sims/2d_3_final/snapshots_lin_0_masked/f_amps.png plots/lin_amps.png;
+% cp ../sims/2d_3_final/snapshots_lin_0_masked/fluxes.png plots/lin_fluxes.png;
+% cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/f_amps.png plots/nl_f_amps.png;
+% cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/fluxes.png plots/nl_fluxes.png;
+% cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/f_refl.png plots/nl_f_refl.png;
+% cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/f_ri.png plots/nl_f_ri.png;
+% cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/front.png plots/nl_front.png;
 
@@
 we approximate the gravitational field as uniform. We model the background
-density stratification as $\rho_0(x, z) = \rho_0(z) = \rho_0(z=0) e^{-z/H}$ for
-some reference density $\rho_0(z=0) = \at{\rho_0(z)}_{z = 0}$.
+density stratification as $\overline{\rho}(x, z) = \overline{\rho}_0 e^{-z/H}$
+for some reference density $\overline{\rho}_0$ (we generally notate background
+quantities with overbars and perturbation quantities with primes).
 
@@
     \begin{align}
-        \vec{\nabla} \cdot \vec{u}_1 &= 0,\label{eq:nl_incomp}\\
+        \bm{\nabla} \cdot \bm{u} &= 0,\label{eq:nl_incomp}\\
         \md{\rho}{t} &= 0 ,\label{eq:nl_density}\\
-        \md{\vec{u}_1}{t} + \frac{\vec{\nabla}p}{\rho} + g\hat{z} &=
-            0.\label{eq:nl_mom}.
+        \md{\bm{u}}{t} + \frac{\bm{\nabla}P}{\rho} + g\bm{z} &=
+            0\label{eq:nl_mom}.
     \end{align}
 \end{subequations}
-$\md{}{t} = \pd{}{t} + \p{\vec{u}_1 \cdot \vec{\nabla}}$ is the Lagrangian or
-material derivative, and $\vec{u}_1, \rho, p$ denote the velocity field, density
-and pressure respectively. We denote $-g\hat{z}$ constant gravitational
+$\md{}{t} = \pd{}{t} + \p{\bm{u} \cdot \bm{\nabla}}$ is the Lagrangian or
+material derivative, and $\bm{u}, \rho, P$ denote the velocity field, density
+and pressure respectively. We denote $-g\bm{z}$ constant gravitational
 acceleration. Note that at hydrostatic equilibrium $\pd{}{t} = 0$ we have
-$\vec{\nabla}P_0 = -\rho_0 g\hat{z}$ and so $P_0 = \rho_0 gH$. A shear flow of
-form $\vec{u}_1 = u_{1x}(x, z, t)\hat{x} = u_{1x}(z)\hat{x}$ is permitted in hydrostatic
-equilibrium, but we will assume no background shear flow. Physically, this
-assumption corresponds to a non-rotating star, or going to the corotating frame
-of a rigidly rotating star.
+$\bm{\nabla}\overline{P} = -\overline{\rho} g\bm{z}$ and so $\overline{P} =
+\overline{\rho} gH$. A vertically-stratified shear flow
+$\overline{u}_x(z)\bm{x}$ is permitted in hydrostatic equilibrium, but we will
+assume no background flow. Physically, this assumption corresponds to a
+non-rotating star, or going to the corotating frame of a rigidly rotating star.
 
 In practice, it is convenient to introduce coordinate $\Upsilon = \ln
-\frac{\rho}{\rho_0}$. This both identically enforces $\rho > 0$ and avoids
-numerical issues in the $\frac{\vec{\nabla}p}{\rho}$ term if $\rho$ is small. We
-also define reduced pressure $P = \frac{p}{\rho}$. Then, we may rewrite the
-second two equations of \autoref{se:nl_orig} as
+\frac{\rho}{\overline{\rho}}$. This both identically enforces $\rho > 0$ and
+avoids numerical issues in the $\frac{\bm{\nabla}P}{\rho}$ term if $\rho$ is
+small. We also define reduced pressure $\varpi = \frac{P}{\rho}$. Then, we may
+rewrite the second two equations of \autoref{se:nl_orig} as
 \begin{subequations}\label{se:nl_upsilon}
     \begin{align}
-        \md{\Upsilon}{t} + u_{1z} \pd{\ln \rho_0}{z} &= 0
+        \md{\Upsilon}{t} + u'_z \pd{\ln \overline{\rho}}{z} &= 0
             ,\label{eq:nl_up_density} \\
-        \md{\vec{u}_1}{t} + \vec{\nabla}P + P\vec{\nabla}\Upsilon + g\hat{z} &= 0.
+        \md{\bm{u}}{t} + \bm{\nabla}\varpi
+            + \varpi\bm{\nabla}\Upsilon + g\bm{z} &= 0.
     \end{align}
@@
 Note that in the new coordinates, hydrostatic equilibrium corresponds to
-$\Upsilon_0 = 1, P_0 = gH$.
+$\overline{\Upsilon} = 1, \overline{\varpi} = gH$.
 
@@
 In the small perturbation limit, where flow velocities are small compared to the
-characteristic space and time scales $\pd{}{t} \gg \vec{u}_1 \cdot \vec{\nabla}$,
-we may linearize \autoref{se:nl_upsilon}. We ignore the advective components of the
-material derivative and consider small deviations $\Upsilon_1 = \Upsilon -
-\Upsilon_0, P_1 = P - P_0$ about hydrostatic equilibrium:
-\begin{subequations}\label{se:lin_homo}
-    \begin{align}
-        \vec{\nabla} \cdot \vec{u}_1 &= 0,\\
-        \pd{\Upsilon_1}{t} - \frac{u_{z}}{H} &= 0,\\
-        \pd{\vec{u}_1}{t} + \vec{\nabla}P_1
-            + \p{1 + \Upsilon_1} g\hat{z}
-            &= 0.
-    \end{align}
-\end{subequations}
-
-This can be solved to obtain solutions up to undetermined amplitude $A$ the
-well-known result \citep{drazin,sutherland0}:
-\begin{equation}
-    u_{1z}\p{x, z, t} = Ae^{z/2H}e^{i\p{k_{1x}x - k_{1z}z - \omega_1 t}},
+characteristic space and time scales $\pd{}{t} \gg \bm{u} \cdot
+\bm{\nabla}$, we may linearize \autoref{se:nl_upsilon}.
+%,We ignore the advective components of the material derivative and consider small
+% deviations $\Upsilon' = \Upsilon - \overline{\Upsilon}, \varpi' = \varpi -
+% \overline{\varpi}$ about hydrostatic equilibrium:
+% \begin{subequations}\label{se:lin_homo}
+%     \begin{align}
+%         \bm{\nabla} \cdot \bm{u}_1 &= 0,\\
+%         \pd{\Upsilon_1}{t} - \frac{u_{z}}{H} &= 0,\\
+%         \pd{\bm{u}_1}{t} + \bm{\nabla}P_1
+%             + \p{1 + \Upsilon_1} g\bm{z}
+%             &= 0.
+%     \end{align}
+% \end{subequations}
+The solution in this linear regime up to undetermined amplitude $A$ is
+given by \citep{drazin,sutherland0}:
+\begin{equation}
+    u'_z\p{x, z, t} = Ae^{z/2H}e^{i\p{k_{x}x + k_{z}z - \omega t}},
         \label{eq:k0z_sign}
 \end{equation}
-where $\omega_1$ satisfies dispersion relation
-\begin{equation}
-    \omega_1^2 = \frac{N^2k_{1x}^2}{k_{1x}^2 + k_{1z}^2 + \frac{1}{4H^2}},
+where $\omega$ satisfies dispersion relation
+\begin{equation}
+    \omega^2 = \frac{N^2k_{x}^2}{k_{x}^2 + k_{z}^2 + \frac{1}{4H^2}}.
         \label{eq:disp_rel}
 \end{equation}
-where
-\begin{equation}
-    N^2 \equiv g^2\p{\rd{\rho}{P} - \at{\pd{\rho}{P}}_{ad}} = \frac{g}{H},
-\end{equation}
-the Brunt-V\"ais\"al\"a frequency is constant. Our sign convention is such that
-a wave with $k_{1x}, k_{1z} > 0$ has positive vertical group velocity (see
-\autoref{eq:vg}).
-
-In the weak stratification limit $k_{1z}H \gg 1$, the solution exhibits the
+We take
+\begin{equation}
+    N^2 \equiv g^2\p{\rd{\rho}{P} - \frac{1}{c_s^2}} = \frac{g}{H},
+\end{equation}
+the \emph{Brunt-V\"ais\"al\"a frequency} to be constant and take the
+incompressible limit $c_s \to \infty$ where $c_s$ is the sound speed. Other
+dynamical quantities are simply related to $u'_z$ (see
+e.g.\ \cite{sutherland0}).
+
+In the weak stratification limit $k_{z}H \gg 1$, the solution exhibits the
 following characteristics:
@@
         \begin{align}
-            \vec{c}_{ph} &=
-                \p{k_{1x}\hat{x} - k_{1z}\hat{z}}\frac{\omega_1}
-                {k_{1x}^2 + k_{1z}^2 + \frac{1}{4H^2}},\\
-            \vec{c}_{g} &= N\frac{\p{k_{1z}^2 + \frac{1}{4H^2}}\hat{x}
-                + \p{k_{1x}k_{1z}\hat{z}}}
-                {\p{k_{1x}^2 + k_{1z}^2 + \frac{1}{4H^2}}^{3/2}}.\label{eq:vg}
-        \end{align}
-        We note $\vec{c}_{ph} \cdot \vec{c}_g = +
-        \mathcal{O}\p{\frac{1}{\p{k_{1z}H}^{2}}} \approx 0$, the usual result
-        (e.g.\  \citep{drazin,sutherland1}). Note again that $c_{gz} > 0$ for
-        $k_{1x}, k_{1z} > 0$, a consequence of our sign choice in
-        \autoref{eq:k0z_sign}.
-
-    \item The time-averaged total $x$-momentum flux carried in the $\hat{z}$
+            \bm{c}_{ph} &=
+                \p{k_{x}\bm{x} + k_{z}\bm{z}}\frac{\omega}
+                {k_{x}^2 + k_{z}^2 + \frac{1}{4H^2}},\\
+            \bm{c}_{g} &= N\frac{\p{k_{z}^2 + \frac{1}{4H^2}}\bm{x}
+                - \p{k_{x}k_{z}\bm{z}}}
+                {\p{k_{x}^2 + k_{z}^2 + \frac{1}{4H^2}}^{3/2}}.\label{eq:vg}
+        \end{align}
+        We note $\bm{c}_{ph} \cdot \bm{c}_g =
+        \mathcal{O}\p{\frac{1}{\p{k_{z}H}^{2}}} \approx 0$, the usual result
+        (e.g.~\cite{drazin,sutherland1}). Note again that $c_{gz} < 0$ for
+        $k_{x}, k_{z} > 0$, so upwards-propagating IGW have $k_z < 0$.
+
+    \item The time-averaged total $x$-momentum flux $F$ carried in the $\bm{z}$
         direction can be computed. Since the linear solution is separable as
@@
         \begin{align}
-            S &\equiv \ev{\rho u_{1x} u_{1z}}_x \equiv
-                \frac{1}{L_x}\int_0^{L_x}\limits \rho u_{1x} u_{1z}\;\mathrm{d}x,
-                    \label{eq:S_def},\\
-                &\approx -\frac{A^2}{2}\rho_0(z=0)\frac{k_{1z}}{k_{1x}},
+            F &\equiv \ev{\rho u_{x}' u_{z}'}_x \equiv
+                \frac{1}{L_x}\int_0^{L_x}\limits \rho u_{x}'u_{z}'\;\mathrm{d}x,
+                    \label{eq:F_def},\\
+                &\approx -\frac{A^2}{2}\overline{\rho}_0\frac{k_{z}}{k_{x}},
                     \label{eq:S_lin}
@@
 \begin{equation}
-    \md{\Upsilon}{t} + u_{1z}\pd{\ln \rho_0}{z}
-        = Fe^{-\frac{(z - z_0)^2}{2\sigma^2}}
-            \cos \p{k_{1x}x - \omega_1 t}.\label{eq:vol_drive}
-\end{equation}
-Using a narrow Gaussian profile excites a broad $k_z$ wavenumber spectrum, and
-only the $k_{1z}$ satisfying dispersion relation \autoref{eq:disp_rel} for the
-given $k_{1x}, \omega_1(k_{1x}, k_{1z})$ will propagate.
+    \md{\Upsilon}{t} + u_{z}'\pd{\ln \bar{\rho}}{z}
+        = Ce^{-\frac{(z - z_0)^2}{2\sigma^2}}
+            \cos \p{k_{x}x - \omega t}.\label{eq:vol_drive}
+\end{equation}
+Using a narrow Gaussian profile excites a broad $z$ power spectrum, but only the
+$k_{z}$ satisfying dispersion relation \autoref{eq:disp_rel} for the given
+$k_{x}, \omega_(k_{x}, k_{z})$ will propagate.
 
@@
 enough to be numerically resolved by the spectral code.}. This system is solved
-exactly by matching the two homogeneous solutions above and below $z_0$. We may
+exactly by matching the two plane wave solutions,above and below $z_0$. We may
 then approximately relax the solution to nonzero $\sigma$: an extra factor of
-$e^{-\frac{(k_{1z}\sigma)^2}{2}}$ arises compared to the $\delta$-function
+$e^{-\frac{(k_{z}\sigma)^2}{2}}$ arises compared to the $\delta$-function
 solution (evaluating the Fourier Transform of $e^{-\frac{(z -
-z_0)^2}{2\sigma^2}}$ at $k_z = k_{1z}$), and we obtain
-\begin{align}
-    u_{z}&(x, z, t) ={} \frac{Fgk_{1x}^2}{\omega_1^2}
-        \frac{1}{2ik_{1z}}\frac{e^{\frac{-(k_{1z}\sigma)^2}{2}}}
+z_0)^2}{2\sigma^2}}$ at $k_z$), and we obtain
+\begin{align}
+    u_{z}'&(x, z, t) ={} \frac{Cgk_{x}^2}{\omega^2}
+        \frac{1}{2ik_{z}}\frac{e^{\frac{-(k_{z}\sigma)^2}{2}}}
         {\sqrt{2\pi\sigma^2}} \times\nonumber\\
         &{}\begin{cases}
-        e^{\frac{z - z_0}{2H}}e^{i\p{k_{1x}x + k_{1z}(z - z_0) - \omega_1 t
-            + \frac{1}{2k_{1z}H}}}
+        e^{\frac{z - z_0}{2H}}e^{i\p{k_{x}x + k_{z}(z - z_0) - \omega t
+            + \frac{1}{2k_{z}H}}}
             & z > z_0\\
-        e^{\frac{z - z_0}{2H}}e^{i\p{k_{1x}x - k_{1z}(z - z_0) - \omega_1 t
-            + \frac{1}{2k_{1z}H}}}
+        e^{\frac{z - z_0}{2H}}e^{i\p{k_{x}x - k_{z}(z - z_0) - \omega t
+            + \frac{1}{2k_{z}H}}}
             & z < z_0\\
@@
 \end{align}
-The $z > z_0$ solution models an upwards-propagating IGW wavetrain inbound on
-the simulation domain from below. The solution for $u_{1x}(x, z, t)$ can be
-obtained analytically from incompressibility \autoref{eq:nl_incomp}.
+The $z > z_0$ region models an upwards-propagating IGW wavetrain excited deep
+in the atmosphere.
 
@@
 In developing the results above, we have neglected advective terms $\pd{}{t} \gg
-\vec{u}_1 \cdot \vec{\nabla}$. However, since $\abs{\vec{u}_1} \propto
+\bm{u}' \cdot \bm{\nabla}$. However, since $\abs{\bm{u}'} \propto
 e^{z/2H}$, any nonzero amplitude IGW will eventually violate the linearity
@@
 \citep{fullerI,fullerII}. We can coarsely estimate the height at which this
-happens by when the Lagrangian displacement $\vec{\xi} =
-\frac{\vec{u}_1}{-i\omega}$ of a fluid parcel satisfies
-\begin{equation}
-    \vec{\xi} \cdot \vec{k}_1 \gtrsim 1.\label{eq:nl}
-\end{equation}
-
-Another useful result quantifies the wave-induced mean flow
-\citep{eliassen_palm_cite,sutherland0}:
-\begin{equation}
-     \ev{u_{1x}} \equiv \bar{U} = \frac{1}{L_x}
-        \int\limits u_{1x}\;\mathrm{d}x = \frac{\ev{u_{1x}u_{1z}}_x}{c_{g,z}},
+happens by when the Lagrangian displacement of the perturbation $\bm{\xi}' =
+\frac{\bm{u}'}{-i\omega}$ of a fluid parcel satisfies
+\begin{equation}
+    \abs{\bm{\xi}'} \abs{\bm{k}} \gtrsim 1.\label{eq:nl}
+\end{equation}
+
+This is in agreement with the breaking location predicted by the wave-induced
+mean flow, given by \citep{eliassen_palm_cite,sutherland0}:
+\begin{equation}
+     \ev{u_{x}} \equiv \overline{U} = \frac{1}{L_x}
+        \int\limits u_{x}\;\mathrm{d}x = \frac{\ev{u_{x}u_{z}}_x}{c_{g,z}},
         \label{eq:mean_flow}
@@
 This formula is analogous to Stokes' drift for surface waves and can be derived
-by considering the propagation of $S$ (\autoref{eq:S_def}) into a medium at
-rest. Solving for $\bar{U} = c_{ph, x}$ the critical layer condition (see
-\autoref{ss:crit_layer} immediately below), we recover
-$\frac{u_{1x}k_{1x}}{\omega} = \xi_x k_{1x} \sim 1$, roughly equivalent to the
-nonlinearity criterion \autoref{eq:nl}. This condition predicts where the wave's
-self-acceleration alone is sufficient to induce critical layer absorption.
+by considering the propagation of $F$ (\autoref{eq:F_def}) into a medium at
+rest. Solving for $\overline{U} = \omega / k_x$ the critical layer condition
+(see \autoref{ss:crit_layer} immediately below), we obtain
+$\abs{\frac{u_{x}'k_{x}}{\omega}} = \abs{\xi_x k_{x}} = \abs{\xi_z k_z} \sim 1$,
+equal to the nonlinearity criterion \autoref{eq:nl}. This condition predicts
+where the wave's self-acceleration alone produces a critical layer.
 
@@
 
-A purely horizontal shear flow $\bar{U}(z) \hat{x}$ can be seen in
+A purely horizontal shear flow $\overline{U}(z) \bm{x}$ can be seen in
 \autoref{se:nl_orig} to have the effect of modifying time derivatives
 $\partial_t$ to their frequency in the comoving frame of the fluid $\partial_t -
-\bar{U}(z)\partial_x$. For a critical value $\omega_1 - \bar{U}_c k_{1x} = 0$,
-the frequency of the linear wave in the fluid's frame of reference vanishes and
-critical behavior is observed. In a linear \citep{booker_bretherton} or viscous
-\citep{hazel} theory, the incident wave has amplitude reflection and
-transmission coefficients
-\begin{align}
-    \mathrm{Ri} &\equiv \at{\frac{N^2}{\p{\pd{\bar{U}}{z}}^2}}_{z_c},
+\overline{U}(z)\partial_x$. For a critical value $\omega - \overline{U}_c k_{x}
+= 0$, the frequency of the linear wave in the fluid's frame of reference
+vanishes and critical behavior is observed. In a linear
+\citep{booker_bretherton} or viscous \citep{hazel} theory, the incident wave has
+amplitude reflection and transmission coefficients
+\begin{align}
+    \mathrm{Ri} &\equiv \at{\frac{N^2}{\p{\pd{\overline{U}}{z}}^2}}_{z_c},
         \label{eq:ri_def}\\
@@
 where we have defined local Richardson number Ri at the critical layer $z_c:
-\bar{U}(z_c) = \frac{\omega_1}{k_{1x}}$. In the $\mathrm{Ri} \gg 1$ limit,
+\overline{U}(z_c) = \frac{\omega}{k_{x}}$. In the $\mathrm{Ri} \gg 1$ limit,
 $\mathcal{R}, \mathcal{T} \ll 1$ and the incident wave is completely absorbed to
@@
 momentum flux as well, which is converted into additional horizontal momentum of
-the shear flow. Since the shear flow cannot exceed $\bar{U}_c$ the horizontal
+the shear flow. Since the shear flow cannot exceed $\overline{U}_c$ the horizontal
 phase velocity of the incident wave, the critical layer must instead propagate
@@
 \begin{equation}
-    \pd{}{t}\int\limits_0^{L_z} \rho(z) \bar{U}(z, t)\;\mathrm{d}z
+    \pd{}{t}\int\limits_0^{L_z} \rho(z) \overline{U}(z, t)\;\mathrm{d}z
         + \Delta S = 0.
@@
 We define $\Delta S$ to be the change in flux across the critical layer, or
-equivalently the negative of the absorbed flux. Treating $\bar{U}(z > z_c) =
-\bar{U}_c, \bar{U}(z < z_c) = 0$ gives us exactly
-\begin{equation}
-    \rho(z_c) \bar{U}_c\pd{z_c}{t} = \Delta S(z_c(t)).\label{eq:zc_anal}
+equivalently the negative of the absorbed flux. Treating $\overline{U}(z > z_c) =
+\overline{U}_c, \overline{U}(z < z_c) = 0$ gives us exactly
+\begin{equation}
+    \rho(z_c) \overline{U}_c\pd{z_c}{t} = \Delta S(z_c(t)).\label{eq:zc_anal}
 \end{equation}
 $\Delta S(z_c)$ denotes the change in $S$ across $z_c$. For constant $\Delta S$
-in time and $\rho \approx \rho_0$, this has analytical solution
-\begin{equation}
-    z_c(t) = -H\ln t - H\ln \frac{H\rho_0(z = 0)c_{ph, x}}{-\Delta S}
+in time and $\rho \approx \bar{\rho}$, this has analytical solution
+\begin{equation}
+    z_c(t) = -H\ln t - H\ln \frac{H\bar{\rho}_0c_{ph, x}}{-\Delta S}
         .\label{eq:zc_sol}
@@
 
-We nondimensionalize by taking $H = N = \rho_0(z=0) = 1$ in
+We nondimensionalize by taking $H = N = \bar{\rho}_0 = 1$ in
 \autoref{se:nl_orig}.
 
+% TODO physical vs simulation domain
 Denote $L_x, L_z$ to be the physical dimensions of the simulation domain. Our
@@
 the $z$ direction to damp perturbations exiting the top and bottom of the
-domain. We use a Fourier basis in the $x$ direction and either a Chebyshev or
-Fourier basis in the $z$ direction; our results were robust to either set of
-basis functions. We varied the number of $x, z$ modes (denoted $N_x, N_z$
-respectively), and used $3/2$ dealiasing \citep{boyd}.
+domain. We use a Fourier basis in both the $x$ and $z$ direction. We varied the
+number of $x, z$ modes (denoted $N_x, N_z$ respectively), and used $3/2$
+dealiasing \citep{boyd}.
 
@@
 with adaptive timesteps computed from advective Courant-Friedrichs-Lewy (CFL)
-time. Specifically, we use $\Delta t = 0.7 \min(\Delta x / u_x,\Delta z / u_{1z})$,
+time. Specifically, we use $\Delta t = 0.7 \min(\Delta x / u_x,\Delta z / u_{z})$,
 where the minimum is taken over every grid point in the domain and $\Delta
@@
 
-The physics of our simulation is fixed by four parameters: $k_{1x}, \omega_1, F,
+The physics of our simulation is fixed by four parameters: $k_{x}, \omega, C,
 \nu$. We describe our choices for these parameters below:
 \begin{itemize}
-    \item $k_{1x}$: Astrophysical IGWs in stars are generally excited by the $l
+    \item $k_{x}$: Astrophysical IGWs in stars are generally excited by the $l
         = 2$ component of the tidal potential, for which $k_{\perp} \sim
         \frac{1}{R}$ where $R$ is the radius of the star. To best emulate this,
-        we use $k_{1x} = \frac{2\pi}{L_x}$ the smallest permitted wavenumber
+        we use $k_{x} = \frac{2\pi}{L_x}$ the smallest permitted wavenumber
         permitted by periodic boundary conditions.
 
-    \item $\omega_1$: We choose $\omega_1$ by evaluating dispersion relation
-        $\omega_1\p{k_{1x}, k_{1z}}$ for some desired $k_{1z}$.
+    \item $\omega$: We choose $\omega$ by evaluating dispersion relation
+        $\omega\p{k_{x}, k_{z}}$ for some desired $k_{z}$.
 
@@
         equivalently $\frac{k_r}{k_{\perp}} \sim k_rR \gg 1$. Since $H \lesssim
-        R$, we aim to study $k_{1z}H \gg 1$. However, to satisfy this well is
+        R$, we aim to study $k_{z}H \gg 1$. However, to satisfy this well is
         computationally expensive: $L_z \gg H$ is required to give waves ample
-        room to grow within the simulation domain, but we want $k_{1z}$ to be
+        room to grow within the simulation domain, but we want $k_{z}$ to be
         sufficiently separated from the grid spacing $\sim L_z / N_z$ that
         nonlinear effects can be seen in later simulations (see
-        \autoref{ss:nl_ns}). Thus, $N_z$ sets how well we can satisfy $k_{1z}H
-        \gg 1$; we choose $k_{1z}H = 1$.
-
-    \item $F$: We first choose $F$ forcing strength such that $\vec{\xi} \cdot
-        \vec{k}_1 \ll 1$ is satisfied everywhere throughout the domain. This
-        constraints $F$ by \autoref{eq:uz_lin}.
-
-    \item $\nu$: Nonlinear effects transfer wave energy from $\vec{k}_{1}$ to
-        smaller scales. Since well-resolved simulations using spectral methods
-        have no inherent numerical viscosity, energy will accumulate at grid
-        scales in the absence of artificial dissipation. We introduce
-        dissipation parameter $\nu$ used for both artificial viscosity and artificial
-        diffusivity. We ensure our equations of motion continue to conserve
-        horizontal momentum (see \autoref{ss:strat_impl} for details). Under
-        weak forcing, we may set $\nu = 0$, since energy transfer out of the
-        parent $\vec{k}_1$ mode is negligible.
+        \autoref{ss:nl_ns}). Thus, $N_z$ sets how well we can satisfy $k_{z}H
+        \gg 1$; we choose $k_{z}H = 1$.
+
+    \item $C$: We first choose $C$ forcing strength such that $\xi_z k_z \ll 1$
+        is satisfied everywhere throughout the domain. This constrains $C$ by
+        \autoref{eq:uz_lin}.
+
+    \item $\nu$: Nonlinear effects transfer wave energy from $\bm{k}$ to
+        larger wavenumbers. Since well-resolved simulations using spectral
+        methods have no inherent numerical viscosity, energy will accumulate at
+        grid scales in the absence of artificial dissipation. We introduce
+        dissipation parameter $\nu$ used for both artificial viscosity and
+        artificial diffusivity. We ensure our equations of motion continue to
+        conserve horizontal momentum (see \autoref{ss:strat_impl} for details).
+        Under weak forcing, we may set $\nu = 0$, since energy transfer out of
+        the parent $\bm{k}$ mode is negligible.
 \end{itemize}
 
-Finally, we use initial conditions $\vec{u}(0) = \Upsilon(0) = 0, \varpi(0) =
+Finally, we use initial conditions $\bm{u}(0) = \Upsilon(0) = 0, \varpi(0) =
 1$ corresponding to hydrostatic equilibrium and no initial fluid motion.
@@
 bottom of the $z$ direction. This damps waves that reach the edge of the
-simulation domain without inducing nonphysical reflection. Specifically, these
-are implemented by introducing linear damping terms to \autoref{se:nl_upsilon}:
-\begin{align}
-    \md{\Upsilon}{t} + u_{1z} \pd{\ln \rho_0}{z} &=
-        -\Gamma(z)\Upsilon,\nonumber\\
-    \md{\vec{u}_1}{t} + \vec{\nabla}P + P\vec{\nabla}\Upsilon + g\hat{z} &=
-        - \Gamma(z)\vec{u}_1,\nonumber\\
-    \frac{1}{2\tau}\s{2 + \tanh \frac{z - z_T}{\Delta z}
-        + \tanh \frac{z_B - z}{\Delta z}} &= \Gamma(z),\label{eq:Gamma}
+simulation domain without inducing nonphysical reflection. Specifically, we
+replace material derivatives in \autoref{se:nl_upsilon} with:
+\begin{align}
+    \md{}{t} &\to \md{}{t} + \Gamma(z)\Upsilon,\nonumber\\
+    \Gamma(z) &= \frac{1}{2\tau}\s{2,+ \tanh \frac{z - z_T}{\Delta z}
+        + \tanh \frac{z_B - z}{\Delta z}},\label{eq:Gamma}
 \end{align}
@@
 We describe the results of a simulation satisfying the parameter choices in
-\autoref{ss:numerics} such that $\vec{\xi} \cdot \vec{k}_1 \ll 1$ everywhere.
+\autoref{ss:numerics} such that $\xi_z k_z \ll 1$ everywhere.
 
 During the simulation, we expect IGW of form \autoref{eq:uz_lin} to be
-excited. Incompressibility allows us to compute a $\vec{u}_1(x, z, t)$ profile
-analytically. The amplitude of the observed IGW in $\vec{u}$ relative to
-analytical solution $\vec{u}_1$  can be estimated with estimator $\hat{A}(t)$
-\begin{equation}
-    \hat{A}(t) = \frac{\int\limits_{z_b}^{z_t}\int\limits_0^{L_x}
-        \rho_0\p{k_{1x}u_xu_{1x} - k_{1z}u_{1z}u_{1z}}\;\mathrm{d}x\mathrm{d}z}
+excited. The incompressibility constraint allows us to compute a $\bm{u}(x, z,
+t)$ profile analytically. The amplitude of the observed IGW in $\bm{u}$ relative
+to analytical solution $\bm{u}_{al}$  can be estimated with estimator
+$\hat{A}_i(t)$ (subscript $i$ denotes incident wave)
+\begin{equation}
+    \hat{A}_i(t) = \frac{\int\limits_{z_b}^{z_t}\int\limits_0^{L_x}
+        \bar{\rho}\p{\bm{u} \cdot \bm{u}_{al}}\;\mathrm{d}x\mathrm{d}z}
         {\int\limits_{z_b}^{z_t}\int\limits_0^{L_x}
-        \rho_0\p{k_{1x}u_{1x}^2 - k_{1z}u_{1z}^2}\;\mathrm{d}x\mathrm{d}z}
+        \bar{\rho}u_{al}^2\;\mathrm{d}x\mathrm{d}z}
         \label{eq:ahat_def}
 \end{equation}
-If $\vec{u} = \vec{u}_1$ over $z \in [z_b, z_t], \forall t$, then it's clear
-that $\hat{A}(t) = 1$. The choice of multiplication by $\rho_0$ is such that the
-integrand is of roughly equal magnitude throughout the domain of integration (as
-$\vec{u} \propto e^{z/2H}$), while including a weighting by $k_{1x}, k_{1z}$
-ensures that reflected components with wavevector $\vec{k} = k_{1x}\hat{x} -
-k_{1z}\hat{z}$ identically vanish in the $\hat{A}$ estimator. Both of these
-features become important in later sections (see \autoref{ss:reflectivity}).
-
-For a simulation satisfying $\vec{\xi} \cdot \vec{k}_1 \ll 1$ everywhere, we
-expect $\hat{A}(t) = 1$ so long as we consider only the domain between the
-forcing and damping zones, i.e.\  $z_b \gtrsim z_0, z_t \lesssim z_T$ ($z_0,
-z_T$ are defined in \autoref{eq:vol_drive} and \autoref{eq:Gamma} respectively).
-For consistency with the nonlinear case later, we choose $z_b = z_0 + 3\sigma,
-z_t = z_b + H$, though instead using $z_t = z_T - \Delta z$ does not change the
-results appreciably. The resulting measurement of $\hat{A}(t)$ is shown in
-\autoref{fig:lin_amps}.
-\begin{figure}[t]
+If $\bm{u} = \bm{u}_{al}$ over $z \in [z_b, z_t], \forall t$, then $\hat{A}_i(t) =
+1$. The energy norm is a traditional normalization choice such that the $\bm{u}
+\propto e^{z/2H}$ is evenly weighted throughout the integration region.
+
+For a simulation satisfying $\xi_z k_z \ll 1$ everywhere, we
+expect $\hat{A}_i(t) = 1$ when integrated between the forcing and damping zones,
+i.e.\  $z_b \gtrsim z_0, z_t \lesssim z_T$ ($z_0, z_T$ are defined in
+\autoref{eq:vol_drive} and \autoref{eq:Gamma} respectively). For consistency
+with the nonlinear case later, we choose $z_b = z_0 + 3\sigma, z_t = z_b + H$
+(using $z_t = z_T - \Delta z$ instead does not change the results). The
+resulting measurement of $\hat{A}_i(t)$ is shown in \autoref{fig:lin_amps}.
+\begin{figure}
     \centering
@@
     \caption{Amplitude of excited wave over time in weak forcing simulation,
-    computed using \autoref{eq:ahat_def}. $\hat{A}(t) = 1$ corresponds to
+    computed using \autoref{eq:ahat_def}. $\hat{A}_i(t) = 1$ corresponds to
     perfect agreement with the analytical estimate. After an initial transient
     phase, we observe great agreement with \autoref{eq:uz_lin}, and in
-    particular $\hat{A}(t)$ asymptotes to a constant value, implying continuous
+    particular $\hat{A}_i(t)$ asymptotes to a constant value, implying continuous
     excitation of identical IGW.}\label{fig:lin_amps}
@@
 
-A second prediction of the analytical theory is that $S(z, t)$ should be
-spatially flat between the forcing and damping zone(\autoref{eq:S_def}). We make
-may a slightly more accurate estimate of $S_1$ the flux carried in the IGW by
-directly substituting $\vec{u}_1$ into \autoref{eq:S_def}. Calling this estimate
-$S_1$, we may then measure the agreement of our simulation with analytical
-expectation by computing
-\begin{equation}
-    \hat{S}(z, t) \equiv \frac{\int\limits_0^{L_x}\rho u_xu_z\;\mathrm{d}x}{
-        \int\limits_{0}^{L_x} \rho_0 u_{1x}u_{1z}\;\mathrm{d}x}.
-        \label{eq:shat_def}
-\end{equation}
-For our weak forcing simulation, $\hat{S} = 1$ is expected between $z_0, z_T$,
+A second prediction of the analytical theory is that the horizontal momentmum
+flux $F(z, t)$ should be spatially flat between the forcing zone where it is
+generated and the damping zone where it is dissipated (\autoref{eq:F_def}). We
+make may a slightly more accurate estimate of $F$ the flux carried in the IGW by
+directly substituting $\bm{u}_{al}$ into \autoref{eq:F_def}. Calling this
+estimate $F_{al}$, we may then measure the agreement of our simulation with
+analytical expectation by computing
+\begin{equation}
+    \hat{F}(z, t) \equiv \frac{\int\limits_0^{L_x}\rho u_xu_z\;\mathrm{d}x}{
+        \int\limits_{0}^{L_x} \bar{\rho} u_{x}u_{z}\;\mathrm{d}x}.
+        \label{eq:sbm_def}
+\end{equation}
+For our weak forcing simulation, $\hat{F} = 1$ is expected between $z_0, z_T$,
 and indeed we observe agreement with this in \autoref{fig:lin_fluxes}.
-\begin{figure}[t]
+\begin{figure}
     \centering
     \includegraphics[width=\columnwidth]{plots/lin_fluxes.png}
-    \caption{\autoref{eq:shat_def} as a function of $z$ at select times $t$. As
-    initial transients die out, $\hat{S}(z, t) = 1$ to very good agreement below
-    $z \lesssim z_T = 9.5H$. The flux excited in the forcing zone is transported
-    without loss to the top of the domain, where it is dissipated by the damping
-    layer (see \autoref{ss:damping}) without reflection.}\label{fig:lin_fluxes}
+    \caption{\autoref{eq:sbm_def} as a function of $z$ at select times $t$. As
+    initial transients die out, $\hat{F}(z, t) = 1$ to very good agreement above
+    the forcing zone $z > z_0 = 2H$ and below the damping zone $z \lesssim z_T =
+    9.5H$. The flux excited in the forcing zone is transported without loss to
+    the top of the domain, where it is dissipated by the damping layer (see
+    \autoref{ss:damping}) without reflection.}\label{fig:lin_fluxes}
 \end{figure}
@@
 To perform simulations of wave breaking phenomena, we use the same values as
-\autoref{ss:numerics} except for $F, \nu$. In particular, we choose $F$ such
-that $\vec{k}_1 \cdot \vec{\xi} = 0.1$ in the forcing zone, which implies
-$\vec{k}_1 \cdot \vec{\xi}$ is exceeded before $z_T$ the upper damping zone.
-The dissipation parameter $\nu$ was varied across the various simulations. To
-quantify $\nu$, we define Reynolds number
-\begin{equation}
-    \mathrm{Re} = \frac{LU}{\nu} \equiv \frac{\omega_1}{\nu k_{1z}^2}.
+\autoref{ss:numerics} except for $C, \nu$. In particular, we choose $C$ such
+that $\xi_z (z_0) k_z = 0.1$ in the forcing zone, which implies $\xi_z(z) k_z
+\gtrsim 1$ before $z_T$ the upper damping zone. The dissipation parameter $\nu$
+was varied across the various simulations. To quantify $\nu$, we define Reynolds
+number
+\begin{equation}
+    \mathrm{Re} = \frac{LU}{\nu} \equiv \frac{\omega}{\nu k_{z}^2}.
         \label{eq:re_def}
@@
 A table of our simulation viscosities can be found in \autoref{tab:params}.
-\begin{table*}
+\begin{table}
     \centering
     \begin{tabular}{l c c c}
-        Name & Resolution & $\mathrm{Re}$ & $z$ spectral basis\\\bottomrule
-        \texttt{0p5\_hres} & $512 \times 2048$ & $2048$ & Fourier\\
-        \texttt{0p5\_shres} & $1024 \times 4096$ & $2048$ & Fourier\\
-        \texttt{1\_hres} & $512 \times 2048$ & $1024$ & Fourier\\
-        \texttt{1\_vhres} & $768 \times 3072$ & $1024$ & Fourier\\
-        \texttt{2\_hres} & $512 \times 2048$ & $512$ & Fourier\\
-        \texttt{3\_width} & $256 \times 1024$ & $341$ & Fourier\\
-        \texttt{6\_masked} & $256 \times 1024$ & $512$ & Chebyshev\\
-        \texttt{4\_masked} & $256 \times 1024$ & $341$ & Chebyshev\\
-        \texttt{3\_masked} & $256 \times 1024$ & $205$ & Chebyshev\\
-        \texttt{1\_masked} & $256 \times 1024$ & $146$ & Chebyshev\\
+        Resolution & $\mathrm{Re}$\\\bottomrule
+        $1024 \times 4096$ & $2048$\\
+        $768 \times 3072$ & $1024$\\
+        $512 \times 2048$ & $512$\\
+        $256 \times 1024$ & $341$\\
+        $256 \times 1024$ & $205$\\
+        $256 \times 1024$ & $146$\\
     \end{tabular}
-    \caption{Table of simulation parameters. Note that $\mathrm{Re}$ is defined
+    \caption{Table of simulation resolutions. Note that $\mathrm{Re}$ is defined
     as in \autoref{eq:re_def}.}\label{tab:params}
-\end{table*}
+\end{table}
 
@@
 
-A full video of our higher-resolution run at $N_x = 1024, N_z = 4096$ can be
-found at\footnote{ http://www.princeton.edu/~lecoanet/data/breaking\_wave.mov}.
-Slices of $\bar{U}, \hat{S}$ across $z$ at various times $t$ are shown in
-\autoref{fig:nl_fluxes}. While the behavior of $\bar{U}$ seems to conform
+% TODO generate a video for the 0.1 case
+A full video of our higher-resolution run at $N_x = 1024, N_z = 4096,
+\mathrm{Re} = 2048$ is available
+online\footnote{http://www.princeton.edu/~lecoanet/data/breaking\_wave.mov}.
+We take this to be our fiducial simulation for the remainder of this paper,
+though other simulations show qualitatively identical behavior. Slices of
+$\overline{U}, \hat{F}$ across $z$ at various times $t$ are shown in
+\autoref{fig:nl_fluxes}. While the behavior of $\overline{U}$ seems to conform
 qualitatively with the predictions of \autoref{ss:crit_layer}, the behavior of
-$\hat{S}$ exhibits two salient features: (i) the incident flux seems to fluctuate
-greatly with time, and (ii) there seems to be a small transmitted feature at many
-of the later times. We will discuss further these features in
+$\hat{F}$ exhibits two salient features: (i) the incident flux seems to
+fluctuate greatly with time, and (ii) there seems to be a small transmitted
+feature at many of the later times. We will discuss further these features in
 \autoref{ss:reflectivity}, after first analyzing the propagation of the critical
 layer.
-\begin{figure}[t]
+\begin{figure}
     \centering
     \includegraphics[width=\columnwidth]{plots/nl_fluxes.png}
-    \caption{Plot of $\bar{U}(z, t)$ and $\hat{S}(z, t)$ respectively at various
-    times over $z$. $\bar{U}, \hat{S}$ follow their definitions in
-    \autoref{eq:mean_flow} and \autoref{eq:shat_def}. The values are extracted
-    from run \texttt{0p5\_shres}. $\bar{U}$ is given
-    in units of $c_{ph, x} = \frac{\omega_1}{k_{1x}}$ the critical layer flow
-    velocity (see \autoref{ss:crit_layer}). The propagation of the critical
-    layer towards lower $z$ and sharp deposition of $S$ at the critical layer
-    are evident.}\label{fig:nl_fluxes}
+    \caption{Plot of $\overline{U}(z, t)$ and $\hat{F}(z, t)$ respectively at
+    various times over $z$ in our $\mathrm{Re} = 1024$ simulation.
+    $\overline{U}, \hat{F}$ follow their definitions in \autoref{eq:mean_flow}
+    and \autoref{eq:sbm_def}. $\overline{U}$ is given in units of $c_{ph, x} =
+    \frac{\omega}{k_{x}}$ the critical layer flow velocity (see
+    \autoref{ss:crit_layer}). The propagation of the critical layer towards
+    lower $z$ and sharp deposition of $F$ at the critical layer are
+    evident.}\label{fig:nl_fluxes}
 \end{figure}
@@
 For the simulation shown in \autoref{fig:nl_fluxes}, we may analyze the location
-of the critical layer. As the simulations are very noisy, we define the location
-of the critical layer to be
-\begin{align}
-    z_{c, \min} &= \argmin_z \z{z: S(z) < 0.3S_0},\nonumber\\
-    z_{c, \max} &= \argmax_z \z{z: S(z) > 0.3S_0},\nonumber\\
+of the critical layer. As the simulations are very noisy, we measure the location
+of the critical layer using an average of where flux deposition occurs
+\begin{align}
+    z_{c, \min} &= \argmin_z \z{z: F(z) < 0.3F_{al}},\nonumber\\
+    z_{c, \max} &= \argmax_z \z{z: F(z) > 0.3F_{al}},\nonumber\\
     z_c &\equiv \frac{z_{c, \min} + z_{c, \max}}{2}.\label{eq:zc_def}
@@
 
-The evolution of $z_c$ is described by ODE \autoref{eq:zc_anal}, and in the
-limit $\Delta S(t) = \Delta S$ we obtained closed-form solution
-\autoref{eq:zc_sol}. To estimate $\Delta S(t)$ from the data, we defined:
-\begin{align}
-    S_{<}(t) &= \ev{S(z)}_{z \in [z_c - \Delta z - H, z_c - \Delta z]}
+The evolution of $z_c$ depends on the deposited flux $\Delta F(t)$ per
+\autoref{eq:zc_anal}. To estimate $\Delta F(t)$ across the critical layer $z_c$
+from the data, we defined:
+\begin{align}
+    F_{<}(t) &= \ev{F(z)}_{z \in [z_c - \Delta z - H, z_c - \Delta z]}
         ,\label{eq:sbelow_def}\\
-    S_>(t) &= \ev{\z{S(z): S(z) < 0}}_{z \in [z_c, z_c + \Delta z]},
+    F_>(t) &= \ev{\z{F(z): F(z) < 0}}_{z \in [z_c, z_c + \Delta z]},
         \label{eq:sabove_def},\\
-    \Delta S(t) &\equiv S_>(t) - S_{<}(t).\label{eq:ds_def}
+    \Delta F(t) &\equiv F_>(t) - F_{<}(t).\label{eq:ds_def}
 \end{align}
 $\ev{\dots}_{z \in [z_a, z_b]}$ denotes averaging over interval $[z_a, z_b]$.
-We average over an interval of length $H = \frac{2\pi}{k_{1z}}$ a full vertical
-wavelength. The offset $\Delta z$ is set by the observation that the critical
-layer's width is limited by $\mathrm{Ri} \lesssim 1$, translating to a critical
-layer width of $\sim \frac{1}{k_{1z}}$. Here, we found $\Delta z =
-\frac{3}{k_z}$ was necessary to be sufficiently offset from strong fluctuations
-near the critical layer.
-
-Since the transmission feature can be seen in \autoref{fig:nl_fluxes} to
-attenuate rather quickly and is indeed rather weak, it is difficult to establish
-the correct interval to average over; we use the observation that $S(z) < 0$
-only for a small interval for $z$ to select the interval length implicitly.
+Below the critical layer, we average over an interval of length $H =
+\frac{2\pi}{k_{z}}$ a full vertical wavelength. The offset $\Delta z$ is
+necessary to make a measurement of the incident flux unaffected by the
+turbulence within the critical layer. The height of the critical layer is
+limited by $\mathrm{Ri} \lesssim 1$, which bounds its vertical extent $\sim
+\frac{1}{k_{z}}$. Here, we found $\Delta z = \frac{3}{k_z}$ was necessary to be
+sufficiently offset from strong fluctuations near the critical layer.
+
+Above the critical layer, waves decay over a much smaller vertical extent as
+they have larger $z$ wavenumbers thanks to the shear flow, so we average over a
+$\Delta z$. Accurate measurement of these fluctuations above the critical layer
+is important to measuring the transmission feature accurately, seen in
+\autoref{fig:nl_fluxes} to be weak and attenuate quickly.
 
 Finally, we may plot the observed $z_c$ from data against two simple predictors:
-(i) integration of \autoref{eq:zc_anal} using the measured $\Delta S(t)$, and
-(ii) substituting the time-averaged $\ev{\Delta S}_t$ into \autoref{eq:zc_sol}.
+(i) integration of \autoref{eq:zc_anal} using the measured $\Delta F(t)$, and
+(ii) substituting the time-averaged $\ev{\Delta F}_t$ into \autoref{eq:zc_sol}.
 Since $z_c(t)$ is both stable and less well-defined at early times (when the
@@
 good agreement between the evolution of $z_c(t)$ and its estimate via $\Delta
-S(t)$ and \autoref{eq:zc_anal} are noteworthy. Further of interest is the
+F(t)$ and \autoref{eq:zc_anal} are noteworthy. Further of interest is the
 discrepancy between the time-averaged predictor and the data. The general
-agreement clearly demonstrates $\Delta S < S_1$ and that incomplete absorption
-is observed. Moreover, its significant underprediction of critical layer motion
-at early times contrasts with its comparatively satisfactory prediction at later
-times to indicate that $\Delta S(t)$ must vary significantly over time. Both of
-these observations indicate a careful characterization of reflectivity is
-necessary.
-
-\begin{figure}[t]
+agreement clearly demonstrates $\Delta F < F_{al}$ and that incomplete
+absorption is observed. Moreover, its significant underprediction of critical
+layer motion at early times contrasts with its comparatively satisfactory
+prediction at later times to indicate that $\Delta F(t)$ must vary significantly
+over time. Both of these observations indicate a careful characterization of
+reflectivity is necessary.
+
+\begin{figure}
     \centering
@@
     $z_c(t)$ from simulation data, (green) predictor of $z_c(t)$ using direct
-    integration of \autoref{eq:zc_anal} for $\Delta S(t)$ measured from
+    integration of \autoref{eq:zc_anal} for $\Delta F(t)$ measured from
     simulation data (described in \autoref{eq:ds_def}) and (blue) direct
-    substitution of time-averaged $\ev{\Delta S(t)}_t$ into
+    substitution of time-averaged $\ev{\Delta F(t)}_t$ into
     \autoref{eq:zc_sol}. Predictors use the end of the simulation as initial
@@
     The the poorer but qualitatively correct agreement of the time-averaged
-    predictor with the data shows both that $\Delta S(t) \neq S_1$ and that
-    $\Delta S(t)$ likely has some very real variation with
+    predictor with the data shows both that $\Delta F(t) \neq F_{al}$ and that
+    $\Delta F(t)$ likely has some very real variation with
     time.}\label{fig:nl_front}
@@
 vertical distance over which the local $u_x$ increases from $0.3 \times$ its
-critical value to its critical value. This can be written
-\begin{align}
-    z_{u, \min}(x, t) &= \argmin_\zeta
-        \z{z: u_x(x, \zeta, t) > 0.3\bar{U}_c},\nonumber\\
-    z_{u, \max}(x, t) &= \argmax
-        \zeta \z{z: u_x(x, \zeta, t) < \bar{U}_c},\nonumber\\
+critical value to its critical value. The value $0.3$ effectively excludes the
+self-acceleration induced by the IGW\@. This can be written
+\begin{align}
+    z_{CL, \min}(x, t) &= \argmin_\zeta
+        \z{z: u_x(x, \zeta, t) > 0.3\overline{U}_c},\nonumber\\
+    z_{CL, \max}(x, t) &= \argmax
+        \zeta \z{z: u_x(x, \zeta, t) < \overline{U}_c},\nonumber\\
     \mathrm{Ri}(t) &\equiv
-        \med_x\p{\frac{N^2 \p{z_{u, \max} - z_{u, \min}}^2}{(0.7
-            \bar{U}_c)^2}}.\label{ss:ri_med_def}
+        \med_x\p{\frac{N^2 \p{z_{CL, \max} - z_{CL, \min}}^2}{(0.7
+            \overline{U}_c)^2}}.\label{ss:ri_med_def}
 \end{align}
@@
 that as an IGW propagates into a shear flow, $\omega$ decreases and $k_z$
-increases, enhancing dissipation. Dissipation in a momentum-conserving fluid
-then corresponds to further mean flow acceleration.
-\begin{figure}[t]
+increases, enhancing dissipation and further driving mean flow acceleration.
+\begin{figure}
     \centering
@@
 We identify two instances of non-absorptive behavior: (i), the presence of a
-reflected wave with wave vector $\vec{k} = k_{1x}\hat{x} + k_{1z}\hat{z}$, and
+reflected wave with wave vector $\bm{k} = k_{x}\bm{x} + k_{z}\bm{z}$, and
 (ii) the amount of horizontal momentum flux $S$ reflected off the critical
@@
 To measure the reflected wave amplitude, we use almost the same definition
-\autoref{eq:ahat_def} except using $-k_{1z} \to +k_{1z}$; call this estimated
-$\hat{A}_r(t)$ the amplitude of the downwards-propagating reflected wave, and
-call $\hat{A}_i(t)$ the upwards-propagating incident wave. To compute
-$\hat{A}_r(t)$, we furthermore permit an arbitrary phase offset $\phi_r(t)$ at
-each time $t$, since the phase of the reflected wave is unknown, unlike that of
-the incident wave. We observe that the evolution of $\phi_r(t)$ is consistent
-with the interpretation of a reflected wave, as it is well approximated by
-$\pd{\phi_r}{t} \approx 2\pd{(k_{1z}z_c)}{t}$, consistent with a wave reflecting
-off a moving boundary at $z_c$.
+\autoref{eq:ahat_def} except using $k_{z} \to -k_{z}$; call this estimated
+$\hat{A}_r(t)$ the amplitude of the downwards-propagating reflected wave. To
+compute $\hat{A}_r(t)$, we furthermore permit an arbitrary phase offset
+$\phi_r(t)$ at each time $t$, since the phase of the reflected wave is unknown,
+unlike that of the incident wave. $\phi_r(t)$ in our simulations
+behaves in agreement with reflection off a moving boundary at $z_c$,
+well approximated by $\abs{\pd{\phi_r}{t}} \approx 2\abs{\pd{(k_{z}z_c)}{t}}$.
 
 Since reflectivity depends sensitively on accurate measurements of $\hat{A}_i,
-\hat{A}_r$, it is important \autoref{eq:ahat_def} identically not pick up
-contributions between the two $k_{1x}\hat{x} \pm k_{1z}\hat{z}$ modes, as we
-argued earlier. The integral for $\hat{A}_r$ is also performed over $z \in [z_0
-+ 3\sigma, z_0 + 3\sigma + H]$. Since in general $\hat{A}_i(t), \hat{A}_r(t)$
-are functions of time, we must measure reflectivity accounting for propagation
-time between the measurement zone and the critical layer where the wave
-reflects. For some
-\begin{equation}
-    \delta t(t) \approx \frac{z_c(t) - z_0 + 3\sigma + H/2}{c_{ph, z}},
-        \label{eq:dt_def}
-\end{equation}
-an incident wave measured at $t - \delta t$ reflects at $t$ and is re-measured
-as a reflected wave at $t + \delta t$. The choice of vertical phase velocity
-instead of group velocity is necessary to good results, and a possible
-explanation is provided in \autoref{ss:modes}. We can then define the amplitude
-reflectivity
-\begin{equation}
-    \mathcal{R}_A(t) \equiv \frac{\hat{A}_r(t - \delta t)}{\hat{A}_i(t +
-        \delta t)}.\label{eq:Ra_def}
-\end{equation}
-Since in general $\delta t$ can take on arbitrary values while $\hat{A}_i,
-\hat{A}_r$ are sampled at discrete timesteps set by our simulation output, we
-evaluate $\hat{A}_i, \hat{A}_r$ using a simple linear interpolation to
-intermediate times.
+\hat{A}_r$, we remark \autoref{eq:ahat_def} ensures orthogonality between
+$k_{x}\bm{x} \pm k_{z}\bm{z}$ modes. The integral for $\hat{A}_r$ is also
+performed over $z \in [z_0 + 3\sigma, z_0 + 3\sigma + H]$. Since
+$\hat{A}_i(t), \hat{A}_r(t)$ vary somewhat over time, we perform time averaging
+over multiple periods of the base IGW (specifically, averaged over approximately
+$8\pi/\omega$). We can then define the amplitude reflectivity
+\begin{equation}
+    \mathcal{R}_A(t) \equiv \frac{\ev{\hat{A}_r}(t)}{\ev{\hat{A}_i}(t)}
+        .\label{eq:Ra_def}
+\end{equation}
+% 7/N per timestep, 13 timestep smoothing kernel, 24/N period
+Angle brackets denote time averaging here.
 
@@
 \begin{equation}
-    S_i(t) \equiv \ev{\rho u_{1x} u_{1z}}_x\hat{A}_i^2(t).
-\end{equation}
-But then, since we already have $\Delta S(t), S_>(t)$ jump and transmitted flux
+    F_i(t) \equiv \ev{\rho u_{x} u_{z}}_x\hat{A}_i^2(t).
+\end{equation}
+But then, since we already have $\Delta F(t), F_>(t)$ jump and transmitted flux
 across the critical layer at time $t$ respectively, we can immediately write
@@
 \begin{equation}
-    S_d(t) = S_i(t - \delta t) + \Delta S(t) - S_>(t).
-\end{equation}
-$\delta t$ is still \autoref{eq:dt_def}; propagation time between $z_c$ and $z_c
-- 3/k_z - H/2$ the center averaging zone was ignored at the level of accuracy of
-the study. Then we can define flux reflectivity and transmissivity coefficients
-\begin{align}
-    \mathcal{R}_S(t) &\equiv -\frac{S_d(t)}{S_i(t - \delta t)},&
-    \mathcal{T}_S(t) &\equiv -\frac{S_>(t)}{S_i(t - \delta t)}.
+    F_d(t) = F_i(t) + \Delta F(t) - F_>(t).
+\end{equation}
+Then we can define flux reflectivity and transmissivity coefficients
+\begin{align}
+    \mathcal{R}_F(t) &\equiv -\frac{\ev{F_d}(t)}{\ev{F_i}(t)},&
+    \mathcal{T}_F(t) &\equiv -\frac{\ev{F_>}(t)}{\ev{F_i}(t)}.
         \label{eq:srefl_def}
@@
 
-The measurements of $\hat{A}_i, \hat{A}_r, S_i, \Delta S, S_d, S_>$ are given in
-\autoref{fig:nl_f_amps}. The oscillations in $\hat{A}_i, S_i$ are evident and
-neatly correspond to oscillations in $A_d, S_d$ respectively. We further analyze
-these oscillations in \autoref{ss:modes}.
-
-\begin{figure}[t]
+The measurements of $\hat{A}_i, \hat{A}_r, F_i, \Delta F, F_d, F_>$ are given in
+\autoref{fig:nl_f_amps}. The oscillations in $\hat{A}_i, F_i, A_d, F_d$ are
+apparent but the mean values seem to have temporally converged.
+
+\begin{figure}
     \centering
@@
     incident on the critical layer, (red) flux absorbed by the critical layer,
-    and (black) flux transmitted through the critical layer. Note that no
-    propagation time effects have been included in these plots.
-    }\label{fig:nl_f_amps}
+    and (black) flux transmitted through the critical
+    layer.}\label{fig:nl_f_amps}
 \end{figure}
@@
 Using these measured quantities, we may make plots of $\mathcal{R}_A^2,
-\mathcal{R}_S, \mathcal{T}_S$, which are provided in \autoref{fig:nl_f_refl}. A
-comparison between $\mathcal{R}_A^2$ and $\mathcal{R}_S$ is appropriate as $S
-\propto A^2$. While less evident in the presented plot, a longer-duration
-simulation at lower resolution confirms that the three quantities have reached
+\mathcal{R}_F, \mathcal{T}_F$, which are provided in \autoref{fig:nl_f_refl}. A
+comparison between $\mathcal{R}_A^2$ and $\mathcal{R}_F$ is appropriate as $F
+\propto A^2$. Visual inspection suggests the three quantities have reached
 their asymptotic values after $t \gtrsim 1750/N$. An observation may be made
-that in general $\mathcal{R}_S \geq \mathcal{R}_A^2$; this conforms to the
-intuition that reflected flux consists of the simple reflected mode and higher
+that in general $\mathcal{R}_F \geq \mathcal{R}_A^2$; this conforms to the
+expectation that reflected flux consists of the simple reflected mode and higher
 order modes as well.
 
-\begin{figure}[t]
+\begin{figure}
     \centering
@@
 
-% Finally, we may be interested in the power spectrum of the transmitted
-% horizontal momentum flux. Define:
-% \begin{align}
-%     \tilde{S}_>(k_x, t) = \abs{
-%         \int\limits_0^{L_x} e^{ik_xx}\hat{S}_>(x, z_c - \Delta z, t)\;\mathrm{d}x}
-%             \label{eq:sfft_above},\\
-%     \tilde{S}_<(k_x, t) = \abs{
-%         \int\limits_0^{L_x} e^{ik_xx}\hat{S}_<(x, z_c + \Delta z, t)\;\mathrm{d}x}.
-%             \label{eq:sfft_below}
-% \end{align}
-% These values are plotted in \autoref{fig:nl_fft}. Note that we expect
-% $\tilde{S}_<(0, t) = S_<(t) = S_i(t) + S_d(t)$ and the same for $\tilde{S}_>$.
-% This is rather accurate for $\tilde{S}_<$ but less so for $\tilde{S}_>$; this is
-% since $S_>(t)$ decays very rapidly above $z_c$. Note furthermore that at later
-% times, the transmitted flux increases, per expectation, and shifts towards
-% higher $k_x$. Repeating this plot at higher viscosities shows significantly
-% lower $\tilde{S}_>$, showing that the $k_x$ of the transmitted waves still
-% changes greatly at the highest $\mathrm{Re}$ we are able to probe. Finally, note
-% that the area under $\tilde{S}$ changes with time; this is okay since we are
-% measuring $S$ at constant position, and different $k_x$ components of $S$ will
-% translate at their separate group velocities.
-
-\subsection{Appearance of Discrete Modes}\label{ss:modes}
-
-We briefly attribute the fluctuations in $\hat{A}_i(t)$ as observed in
-\autoref{fig:nl_f_amps} to the introduction of discrete modes by partial
-reflection off a moving critical layer. The partially reflecting boundary causes
-a muted effect of the usual normal modes calculation, so a slightly stronger
-response $\hat{A}_i > 1$ corresponds to being weakly on resonance. The
-instantaneous enhancement/attenuation of $\hat{A}_i$ when $z_c$ moves slowly can
-be computed by an overlap integral of the forcing with the modes of the finite
-domain. This is a consistent interpretation for three reasons:
-\begin{itemize}
-    \item The period of the fluctuations in $\hat{A}_i(t)$ grows longer at later
-        times. This is consistent with $\pd{z_c}{t}$ growing smaller at later
-        times, as the critical layer propagates deeper (see
-        \autoref{eq:zc_anal}).
-
-    \item The amplitude of the fluctuations in $\hat{A}_i(t)$ grows larger at
-        later times. This corresponds to the effective domain (below the
-        critical layer) growing smaller, which leads to stronger boundary
-        effects.
-
-    \item The propagation time $\delta t$, defined in \autoref{eq:dt_def}, used
-        the vertical phase velocity instead of the vertical group velocity to
-        obtain a non-oscillating reflectivity. This is consistent with the
-        fluctuations in $\hat{A}_i(t)$ being induced by internal modes of the
-        domain rather than a varying forcing strength, as the latter should
-        propagate at the \emph{group velocity} instead.
-\end{itemize}
-
 \subsection{Convergence}\label{ss:convergence}
@@
 To estimate convergence, we compute the median value of each of $\mathrm{Ri},
-\mathcal{R}_A^2, \mathcal{R}_S, \mathcal{T}_S$ over the last $1/4$ of simulation
+\mathcal{R}_A^2, \mathcal{R}_F, \mathcal{T}_F$ over the last $1/4$ of simulation
 times from each simulation, where all simulations had converged to what appeared
@@
 in \autoref{fig:agg}.
-\begin{figure}[t]
+\begin{figure}
     \centering
@@
     parameterized with $\mathrm{Re}$ (\autoref{eq:re_def}). Error bars depict
-    $16$ and $84$ percentile values. Small horizontal displacements are made for
+    $16$ and $84$ percentile values. Fmall horizontal displacements are made for
     data points at identical $\mathrm{Re}$ for readability. Note that
@@
 Another source of dissipation considered in \citep{bukart} is turbulent
-convective damping. They find this damping rate to never exceed that of
+conbmtive damping. They find this damping rate to never exceed that of
 radiative damping, and so it is also too weak to suppress critical layer
@@
 
-    \item Solar-type stars (inner convective, outer radiative): different
+    \item Solar-type stars (inner conbmtive, outer radiative): different
         equation of state/stratification but could be qualitatively similar.
@@
         Associating $t_{} \sim \nu k^2 \approx \nu k_z^2$ with the visocus
-        timescale and $t_{NL} \sim \vec{u} \cdot \vec{\nabla} \sim \omega$ for
+        timescale and $t_{NL} \sim \bm{u} \cdot \bm{\nabla} \sim \omega$ for
         the nonlinear timescale, we find their $\lambda = \frac{t_{NL}}{t_L}
@@
 
-Numerically, the nonlinear $\frac{\vec{\nabla}P}{\rho}$ term is problematic: we
+Numerically, the nonlinear $\frac{\bm{\nabla}P}{\rho}$ term is problematic: we
 desire a system where the fluid fields are not divided by one another. We
-introduce $\varpi = \frac{P}{\rho}$ instead, then mandate $\rho_0, \varpi_0$
-background fields satisfy hydrostatic equilibrium $\vec{\nabla}\varpi_0 +
-\varpi_0 \vec{\nabla}\rho_0 + g\hat{z} = 0$. Taking isothermal stratification,
+introduce $\varpi = \frac{P}{\rho}$ instead, then mandate $\bar{\rho}, \varpi_0$
+background fields satisfy hydrostatic equilibrium $\bm{\nabla}\varpi_0 +
+\varpi_0 \bm{\nabla}\bar{\rho} + g\bm{z} = 0$. Taking isothermal stratification,
 we find $\varpi_0 = gH$. We further change variables to $\Upsilon = \ln \rho -
-\ln \rho_0$ and $\varpi_1 = \varpi - \varpi_0$ deviations from the background
+\ln \bar{\rho}$ and $\varpi_1 = \varpi - \varpi_0$ deviations from the background
 state to obtain a system of equations at most quadratic in fluid fields:
@@
     \begin{align}
-        \vec{\nabla} \cdot \vec{u}_1 &= 0,\\
-        \pd{\Upsilon}{t} + \p{\vec{u}_1 \cdot \vec{\nabla}} \Upsilon
+        \bm{\nabla} \cdot \bm{u}_1 &= 0,\\
+        \pd{\Upsilon}{t} + \p{\bm{u}_1 \cdot \bm{\nabla}} \Upsilon
             - \frac{u_z}{H} &= 0,\\
-        \pd{u_{1x}}{t} + \p{\vec{u}_1 \cdot \vec{\nabla}}u_{1x}
+        \pd{u_{1x}}{t} + \p{\bm{u}_1 \cdot \bm{\nabla}}u_{1x}
             + \pd{\varpi_1}{x} + gH\pd{\Upsilon}{x}
             + \varpi_1 \pd{\Upsilon}{x} &= 0,\\
-        \pd{u_z}{t} + \p{\vec{u}_1 \cdot \vec{\nabla}}u_z
+        \pd{u_z}{t} + \p{\bm{u}_1 \cdot \bm{\nabla}}u_z
             + \pd{\varpi_1}{z} + gH\pd{\Upsilon}{z}
@@
     \begin{align}
-        \vec{\nabla} \cdot \vec{u} &= 0,\\
-        \partial_t \rho + \vec{\nabla} \cdot (\rho \vec{u} - \nu
-            \vec{\nabla}(\rho - \rho_0)) &= 0,\label{eq:visc_cons_mom}\\
-        \partial_t (\rho \vec{u}) + \vec{\nabla} \cdot (\rho \vec{u}
-        \vec{u} +
-            \mathrm{diag}(\rho \varpi) - \nu \rho \vec{\nabla}\vec{u}) + \rho g
-            \hat{z} &= 0.
+        \bm{\nabla} \cdot \bm{u} &= 0,\\
+        \partial_t \rho + \bm{\nabla} \cdot (\rho \bm{u} - \nu
+            \bm{\nabla}(\rho - \bar{\rho})) &= 0,\label{eq:visc_cons_mom}\\
+        \partial_t (\rho \bm{u}) + \bm{\nabla} \cdot (\rho \bm{u}
+        \bm{u} +
+            \mathrm{diag}(\rho \varpi) - \nu \rho \bm{\nabla}\bm{u}) + \rho g
+            \bm{z} &= 0.
     \end{align}
@@
     \begin{align}
-        \vec{\nabla} \cdot \vec{u} &= 0,\\
-        \partial_t \Upsilon + \p{\vec{u} \cdot \vec{\nabla}} \Upsilon -
-            \frac{u_z}{H} - \nu\p{\nabla^2 \Upsilon + \p{\vec{\nabla}
-            \Upsilon} \cdot \p{\vec{\nabla}\Upsilon} - \frac{2}{H}\partial_z
+        \bm{\nabla} \cdot \bm{u} &= 0,\\
+        \partial_t \Upsilon + \p{\bm{u} \cdot \bm{\nabla}} \Upsilon -
+            \frac{u_z}{H} - \nu\p{\nabla^2 \Upsilon + \p{\bm{\nabla}
+            \Upsilon} \cdot \p{\bm{\nabla}\Upsilon} - \frac{2}{H}\partial_z
             \Upsilon + \frac{1 - e^{-\Upsilon}}{H^2}} &= 0,\\
-        \partial_t \vec{u} + \p{\vec{u} \cdot \vec{\nabla}}\vec{u} +
-            \vec{\nabla} \varpi + \varpi \vec{\nabla} \Upsilon - \nu \nabla^2
-            \vec{u} + \vec{u} \nu\p{\nabla^2 \Upsilon + \p{\vec{\nabla}
-            \Upsilon} \cdot \p{\vec{\nabla}\Upsilon} - \frac{2}{H}\partial_z
+        \partial_t \bm{u} + \p{\bm{u} \cdot \bm{\nabla}}\bm{u} +
+            \bm{\nabla} \varpi + \varpi \bm{\nabla} \Upsilon - \nu \nabla^2
+            \bm{u} + \bm{u} \nu\p{\nabla^2 \Upsilon + \p{\bm{\nabla}
+            \Upsilon} \cdot \p{\bm{\nabla}\Upsilon} - \frac{2}{H}\partial_z
             \Upsilon + \frac{1 - e^{-\Upsilon}}{H^2}}&{}\nonumber\\
-        - 2\nu \p{\p{\p{\vec{\nabla}\Upsilon} \cdot \vec{\nabla}}\vec{u} -
-            \frac{1}{H}\partial_z \vec{u}} - \frac{\varpi_1}{H} &= 0.
+        - 2\nu \p{\p{\p{\bm{\nabla}\Upsilon} \cdot \bm{\nabla}}\bm{u} -
+            \frac{1}{H}\partial_z \bm{u}} - \frac{\varpi_1}{H} &= 0.
     \end{align}
 \end{subequations}
-Hydrostatic equilibrium is still $\vec{\nabla} \varpi_0 + g\hat{z} = 0$ where
-$\rho = \rho_0, \vec{u} = 0$. Including the damping layers and forcing terms as
+Hydrostatic equilibrium is still $\bm{\nabla} \varpi_0 + g\bm{z} = 0$ where
+$\rho = \bar{\rho}, \bm{u} = 0$. Including the damping layers and forcing terms as
 described in \autoref{ss:numerics}, we finally obtain the full system of
@@
     \begin{align}
-        \vec{\nabla} \cdot \vec{u} ={}& 0,\\
+        \bm{\nabla} \cdot \bm{u} ={}& 0,\\
         \partial_t \Upsilon - \frac{u_z}{H}
-            ={}& \nu\p{\nabla^2 \Upsilon + \p{\vec{\nabla}
-            \Upsilon} \cdot \p{\vec{\nabla}\Upsilon} - \frac{2}{H}\partial_z
+            ={}& \nu\p{\nabla^2 \Upsilon + \p{\bm{\nabla}
+            \Upsilon} \cdot \p{\bm{\nabla}\Upsilon} - \frac{2}{H}\partial_z
             \Upsilon + \frac{1 - e^{-\Upsilon}}{H^2}},\nonumber\\
-            & - \p{\vec{u} \cdot \vec{\nabla}}\Upsilon
+            & - \p{\bm{u} \cdot \bm{\nabla}}\Upsilon
                 -\Gamma(z) \Upsilon
-                + \frac{F}{\rho_0(z)}e^{-\frac{(z - z_0)^2}{2\sigma^2}}
+                + \frac{F}{\bar{\rho}(z)}e^{-\frac{(z - z_0)^2}{2\sigma^2}}
                     \cos \p{k_xx - \omega t},\\
@@
             \nu \nabla^2 u_x
-            - u_x \nu\p{\nabla^2 \Upsilon + \p{\vec{\nabla} \Upsilon} \cdot
-                \p{\vec{\nabla}\Upsilon} - \frac{2}{H}\partial_z \Upsilon
+            - u_x \nu\p{\nabla^2 \Upsilon + \p{\bm{\nabla} \Upsilon} \cdot
+                \p{\bm{\nabla}\Upsilon} - \frac{2}{H}\partial_z \Upsilon
                 + \frac{1 - e^{-\Upsilon}}{H^2}}\nonumber\\
-            &+ 2\nu \p{\p{\p{\vec{\nabla}\Upsilon} \cdot \vec{\nabla}}u_x
+            &+ 2\nu \p{\p{\p{\bm{\nabla}\Upsilon} \cdot \bm{\nabla}}u_x
                 - \frac{1}{H}\partial_z u_x}
             -\Gamma(z) u_x
-                - \p{\vec{u} \cdot \vec{\nabla}}u_x
+                - \p{\bm{u} \cdot \bm{\nabla}}u_x
                 - T_1 \pd{\Upsilon}{x},\\
@@
             \nu \nabla^2 u_z
-            - u_z \nu\p{\nabla^2 \Upsilon + \p{\vec{\nabla} \Upsilon} \cdot
-                \p{\vec{\nabla}\Upsilon} - \frac{2}{H}\partial_z \Upsilon
+            - u_z \nu\p{\nabla^2 \Upsilon + \p{\bm{\nabla} \Upsilon} \cdot
+                \p{\bm{\nabla}\Upsilon} - \frac{2}{H}\partial_z \Upsilon
                 + \frac{1 - e^{-\Upsilon}}{H^2}}\nonumber\\
-            &+ 2\nu \p{\p{\p{\vec{\nabla}\Upsilon} \cdot \vec{\nabla}}u_z -
+            &+ 2\nu \p{\p{\p{\bm{\nabla}\Upsilon} \cdot \bm{\nabla}}u_z -
                 \frac{1}{H}\partial_z u_{z}}
-            -\Gamma(z) u_z - \p{\vec{u} \cdot \vec{\nabla}}u_z
+            -\Gamma(z) u_z - \p{\bm{u} \cdot \bm{\nabla}}u_z
             - T_1 \pd{\Upsilon}{z}.

### 2019-07-29 sims/2d_4_fourier/strat_helper.py

@@ -77,7 +77,10 @@ def horiz_mean(field, n_x, axis=1):
     return np.sum(field, axis=axis) / n_x
 
 def smooth(f):
-    _kernel = np.array([1, 2, 3, 3, 3, 2, 1])
+    # use much stronger smoothing since no longer using time propagation
+
+    # _kernel = np.array([1, 2, 3, 3, 3, 2, 1])
+    _kernel = np.array([1, 2, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 2, 1])
     kernel = _kernel / sum(_kernel)
     f_padded = np.concatenate(([f[0]] * (len(kernel) // 2),
                                f,
@@ -1244,9 +1247,10 @@ def plot_front(name, params):
             f_smooth = smooth(np.concatenate((f[idxs], [f[-1]])))
             return interp1d(np.concatenate((t[idxs], [t[-1]])), f_smooth)
 
-        prop_time = (smooth(front_pos[start_idx: ]) -
-                     (Z0 + 3 * S + Z_TOP_MULT * np.pi / abs(KZ)))\
-                     / abs(V_PZ)
+        # prop_time = (smooth(front_pos[start_idx: ]) -
+        #              (Z0 + 3 * S + Z_TOP_MULT * np.pi / abs(KZ)))\
+        #              / abs(V_PZ)
+        prop_time = 0 * t
 
         t_refl = np.linspace((t + prop_time)[0], (t - prop_time)[-1], len(t))
         f, ax1 = plt.subplots(1, 1, sharex=True)

### 2019-07-31 writeups/paper.tex
stash draft

@@
 $\overline{u}_x(z)\bm{x}$ is permitted in hydrostatic equilibrium, but we will
-assume no background flow. Physically, this assumption corresponds to a
-non-rotating star, or going to the corotating frame of a rigidly rotating star.
+assume no background flow, and $\bm{u} = \bm{u}'$. Physically, this assumption
+corresponds to a non-rotating star, or going to the corotating frame of a
+rigidly rotating star.
 
 In practice, it is convenient to introduce coordinate $\Upsilon = \ln
-\frac{\rho}{\overline{\rho}}$. This both identically enforces $\rho > 0$ and
-avoids numerical issues in the $\frac{\bm{\nabla}P}{\rho}$ term if $\rho$ is
-small. We also define reduced pressure $\varpi = \frac{P}{\rho}$. Then, we may
-rewrite the second two equations of \autoref{se:nl_orig} as
+\frac{\rho}{\overline{\rho}}$ (e.g.\ see \citep{lecoanet_anel}). This both
+identically enforces $\rho > 0$ and avoids numerical issues in the
+$\frac{\bm{\nabla}P}{\rho}$ term if $\rho$ is small. We also define reduced
+pressure $\varpi = \frac{P}{\rho}$. Then, we may rewrite the second two
+equations of \autoref{se:nl_orig} as
 \begin{subequations}\label{se:nl_upsilon}
@@
             ,\label{eq:nl_up_density} \\
-        \md{\bm{u}}{t} + \bm{\nabla}\varpi
+        \md{\bm{u}'}{t} + \bm{\nabla}\varpi
             + \varpi\bm{\nabla}\Upsilon + g\bm{z} &= 0.
@@
 Note that in the new coordinates, hydrostatic equilibrium corresponds to
-$\overline{\Upsilon} = 1, \overline{\varpi} = gH$.
+$\Upsilon = 1, \overline{\varpi} = gH$.
 
@@
 In the small perturbation limit, where flow velocities are small compared to the
-characteristic space and time scales $\pd{}{t} \gg \bm{u} \cdot
-\bm{\nabla}$, we may linearize \autoref{se:nl_upsilon}.
-%,We ignore the advective components of the material derivative and consider small
-% deviations $\Upsilon' = \Upsilon - \overline{\Upsilon}, \varpi' = \varpi -
-% \overline{\varpi}$ about hydrostatic equilibrium:
-% \begin{subequations}\label{se:lin_homo}
-%     \begin{align}
-%         \bm{\nabla} \cdot \bm{u}_1 &= 0,\\
-%         \pd{\Upsilon_1}{t} - \frac{u_{z}}{H} &= 0,\\
-%         \pd{\bm{u}_1}{t} + \bm{\nabla}P_1
-%             + \p{1 + \Upsilon_1} g\bm{z}
-%             &= 0.
-%     \end{align}
-% \end{subequations}
-The solution in this linear regime up to undetermined amplitude $A$ is
-given by \citep{drazin,sutherland0}:
+characteristic space and time scales $\pd{}{t} \gg \bm{u}' \cdot
+\bm{\nabla}$, we may linearize \autoref{se:nl_upsilon}. The solution in this
+linear regime up to undetermined amplitude $A$ is given by
+\citep{drazin,sutherland0}:
 \begin{equation}
@@
 \end{equation}
-the \emph{Brunt-V\"ais\"al\"a frequency} to be constant and take the
-incompressible limit $c_s \to \infty$ where $c_s$ is the sound speed. Other
-dynamical quantities are simply related to $u'_z$ (see
-e.g.\ \cite{sutherland0}).
+the \emph{Brunt-V\"ais\"al\"a frequency} to be constant in the incompressible
+limit $c_s \to \infty$ where $c_s$ is the sound speed. Other dynamical
+quantities are simply related to $u'_z$ (see e.g.\ \cite{sutherland0}).
 
@@
 
-    \item The phase and group velocities can be computed respectively:
+    \item The phase and group velocities are respectively:
         \begin{align}
@@
     \item The time-averaged total $x$-momentum flux $F$ carried in the $\bm{z}$
-        direction can be computed. Since the linear solution is separable as
-        $f(x, z, t) = f(z)e^{i(k_{1x}x - \omega_1 t)}$, $x$ averaging and time
-        averaging are equivalent, so we may write
+        direction can be computed. Since the linear solution consists of terms
+        separable as $f(x, z, t) = f(z)e^{i(k_{1x}x - \omega_1 t)}$, $x$
+        averaging and time averaging are equivalent. We may then write
         \begin{align}
@@
         \end{align}
-        where $\ev{\dots}_x$ denotes $x$-averaging.
+        where $\ev{\dots}_x$ denotes $x$-averaging. Note that indeed for an
+        upwards propagating IGW $k_z < 0, k_x > 0$, we have $F > 0$.
 \end{itemize}
@@
 As not to interfere with the incompressibility constraint, we force the system
-on the density equation. This constitutes replacing \autoref{eq:nl_up_density}
-with
-\begin{equation}
-    \md{\Upsilon}{t} + u_{z}'\pd{\ln \bar{\rho}}{z}
+on the density equation. We implement forcing with strength $C$ by replacing
+\autoref{eq:nl_up_density} with
+\begin{equation}
+    \md{\Upsilon}{t} + u_{z}'\pd{\ln \overline{\rho}}{z}
         = Ce^{-\frac{(z - z_0)^2}{2\sigma^2}}
@@
 $k_{z}$ satisfying dispersion relation \autoref{eq:disp_rel} for the given
-$k_{x}, \omega_(k_{x}, k_{z})$ will propagate.
+$k_{x}, \omega(k_{x}, k_{z})$ will propagate.
 
@@
 enough to be numerically resolved by the spectral code.}. This system is solved
-exactly by matching the two plane wave solutions,above and below $z_0$. We may
+exactly by matching the two plane wave solutions, above and below $z_0$. We may
 then approximately relax the solution to nonzero $\sigma$: an extra factor of
@@
 \begin{equation}
-    \abs{\bm{\xi}'} \abs{\bm{k}} \gtrsim 1.\label{eq:nl}
+    \abs{\bm{\xi}'} \abs{\bm{k}} \sim \xi_z k_z \gtrsim 1.\label{eq:nl}
 \end{equation}
@@
 by considering the propagation of $F$ (\autoref{eq:F_def}) into a medium at
-rest. Solving for $\overline{U} = \omega / k_x$ the critical layer condition
-(see \autoref{ss:crit_layer} immediately below), we obtain
-$\abs{\frac{u_{x}'k_{x}}{\omega}} = \abs{\xi_x k_{x}} = \abs{\xi_z k_z} \sim 1$,
-equal to the nonlinearity criterion \autoref{eq:nl}. This condition predicts
-where the wave's self-acceleration alone produces a critical layer.
+rest. Solving for $\overline{U} = \overline{U}_c \equiv \omega / k_x$ the
+critical layer condition (see \autoref{ss:crit_layer} immediately below), we
+obtain $\abs{\frac{u_{x}'k_{x}}{\omega}} = \abs{\xi_x k_{x}} = \abs{\xi_z k_z}
+\sim 1$, equal to the nonlinearity criterion \autoref{eq:nl}. This condition
+predicts where the wave's self-acceleration alone produces a critical layer.
 
@@
 \overline{U}(z)\partial_x$. For a critical value $\omega - \overline{U}_c k_{x}
-= 0$, the frequency of the linear wave in the fluid's frame of reference
-vanishes and critical behavior is observed. In a linear
+= 0$, the horizontal wavenumber of the linear wave in the fluid's frame of
+reference vanishes and critical behavior is observed. In a linear
 \citep{booker_bretherton} or viscous \citep{hazel} theory, the incident wave has
@@
 momentum flux as well, which is converted into additional horizontal momentum of
-the shear flow. Since the shear flow cannot exceed $\overline{U}_c$ the horizontal
-phase velocity of the incident wave, the critical layer must instead propagate
-downwards to accommodate the incident momentum. The total horizontal momentum of
-the shear flow then obeys conservation equation
-\begin{equation}
-    \pd{}{t}\int\limits_0^{L_z} \rho(z) \overline{U}(z, t)\;\mathrm{d}z
+the shear flow. Since the mean flow cannot exceed $\overline{U}_c$ the
+horizontal phase velocity of the incident wave, the critical layer must instead
+propagate downwards to accommodate the incident momentum. The total horizontal
+momentum of the shear flow then obeys conservation equation
+\begin{equation}
+    \pd{}{t}\int\limits \rho(z) \overline{U}(z, t)\;\mathrm{d}z
         + \Delta S = 0.
 \end{equation}
-We define $\Delta S$ to be the change in flux across the critical layer, or
-equivalently the negative of the absorbed flux. Treating $\overline{U}(z > z_c) =
-\overline{U}_c, \overline{U}(z < z_c) = 0$ gives us exactly
-\begin{equation}
-    \rho(z_c) \overline{U}_c\pd{z_c}{t} = \Delta S(z_c(t)).\label{eq:zc_anal}
-\end{equation}
-$\Delta S(z_c)$ denotes the change in $S$ across $z_c$. For constant $\Delta S$
-in time and $\rho \approx \bar{\rho}$, this has analytical solution
-\begin{equation}
-    z_c(t) = -H\ln t - H\ln \frac{H\bar{\rho}_0c_{ph, x}}{-\Delta S}
+$\Delta S$ is the total injected horizontal momentum to the flow. We next make a
+simple two-zone approximation $\overline{U}(z > z_c) = \overline{U}_c,
+\overline{U}(z < z_c) = 0$, assuming all horizontal momentum flux goes into
+critical layer propagation. Then, the conservation equation becomes simple
+differential equation
+\begin{equation}
+    \rho(z_c) \overline{U}_c\pd{z_c}{t} = \Delta S.\label{eq:zc_anal}
+\end{equation}
+For constant $\Delta S$ in time and $\rho \approx \overline{\rho}$, this has
+analytical solution
+\begin{equation}
+    z_c(t) = -H\ln t - H\ln \frac{H\overline{\rho}_0\overline{U}_c}{-\Delta S}
         .\label{eq:zc_sol}
@@
 
-We nondimensionalize by taking $H = N = \bar{\rho}_0 = 1$ in
+We nondimensionalize by taking $H = N = \overline{\rho}_0 = 1$ in
 \autoref{se:nl_orig}.
@@
 
-        Astrophysical IGWs also generally satisfy $\omega \ll N$, or
-        equivalently $\frac{k_r}{k_{\perp}} \sim k_rR \gg 1$. Since $H \lesssim
-        R$, we aim to study $k_{z}H \gg 1$. However, to satisfy this well is
-        computationally expensive: $L_z \gg H$ is required to give waves ample
-        room to grow within the simulation domain, but we want $k_{z}$ to be
-        sufficiently separated from the grid spacing $\sim L_z / N_z$ that
+        To choose $k_z$, we note astrophysical IGWs generally satisfy $\omega
+        \ll N$, or equivalently $\frac{k_r}{k_{\perp}} \sim k_rR \gg 1$. Since
+        $H \lesssim R$, we aim to study $k_{z}H \gg 1$. However, to satisfy this
+        well is computationally expensive: $L_z \gg H$ is required to give waves
+        ample room to grow within the simulation domain, but we want $k_{z}$ to
+        be sufficiently separated from the grid spacing $\sim L_z / N_z$ that
         nonlinear effects can be seen in later simulations (see
@@
         artificial diffusivity. We ensure our equations of motion continue to
-        conserve horizontal momentum (see \autoref{ss:strat_impl} for details).
-        Under weak forcing, we may set $\nu = 0$, since energy transfer out of
-        the parent $\bm{k}$ mode is negligible.
+        conserve horizontal momentum (see Appendix~\ref{se:strat_impl} for
+        details). Under weak forcing, we may set $\nu = 0$, since energy
+        transfer out of the parent $\bm{k}$ mode is negligible.
 \end{itemize}
 
-Finally, we use initial conditions $\bm{u}(0) = \Upsilon(0) = 0, \varpi(0) =
-1$ corresponding to hydrostatic equilibrium and no initial fluid motion.
+Finally, we use initial conditions $\bm{u}(x, z, 0) = \Upsilon(x, z, 0) = 0,
+\varpi(x, z, 0) = 1$ corresponding to hydrostatic equilibrium and no initial
+fluid motion.
 
@@
 important for spectral methods. Further details of our implementation of the
-fluid equations in Dedalus are described in \autoref{ss:strat_impl}.
+fluid equations in Dedalus are described in Appendix~\ref{se:strat_impl}.
 
@@
 
-During the simulation, we expect IGW of form \autoref{eq:uz_lin} to be
-excited. The incompressibility constraint allows us to compute a $\bm{u}(x, z,
-t)$ profile analytically. The amplitude of the observed IGW in $\bm{u}$ relative
-to analytical solution $\bm{u}_{al}$  can be estimated with estimator
+During the simulation, we expect IGW of form \autoref{eq:uz_lin} to be excited.
+The incompressibility constraint further yields a complete analytical solution
+for flow velocities $\bm{u}_{al}'(x, z, t)$. The amplitude of the observed IGW
+in the simulation field $\bm{u}$ relative to analytical solution $\bm{u}_{al}'$
+over some region $z \in [z_b, z_t]$ can be estimated with estimator
 $\hat{A}_i(t)$ (subscript $i$ denotes incident wave)
@@
     \hat{A}_i(t) = \frac{\int\limits_{z_b}^{z_t}\int\limits_0^{L_x}
-        \bar{\rho}\p{\bm{u} \cdot \bm{u}_{al}}\;\mathrm{d}x\mathrm{d}z}
+        \overline{\rho}\p{\bm{u} \cdot \bm{u}_{al}'}\;\mathrm{d}x\mathrm{d}z}
         {\int\limits_{z_b}^{z_t}\int\limits_0^{L_x}
-        \bar{\rho}u_{al}^2\;\mathrm{d}x\mathrm{d}z}
+        \overline{\rho}(u_{al}')^2\;\mathrm{d}x\mathrm{d}z}
         \label{eq:ahat_def}
 \end{equation}
-If $\bm{u} = \bm{u}_{al}$ over $z \in [z_b, z_t], \forall t$, then $\hat{A}_i(t) =
-1$. The energy norm is a traditional normalization choice such that the $\bm{u}
-\propto e^{z/2H}$ is evenly weighted throughout the integration region.
-
-For a simulation satisfying $\xi_z k_z \ll 1$ everywhere, we
-expect $\hat{A}_i(t) = 1$ when integrated between the forcing and damping zones,
-i.e.\  $z_b \gtrsim z_0, z_t \lesssim z_T$ ($z_0, z_T$ are defined in
-\autoref{eq:vol_drive} and \autoref{eq:Gamma} respectively). For consistency
-with the nonlinear case later, we choose $z_b = z_0 + 3\sigma, z_t = z_b + H$
-(using $z_t = z_T - \Delta z$ instead does not change the results). The
-resulting measurement of $\hat{A}_i(t)$ is shown in \autoref{fig:lin_amps}.
+If $\bm{u} = \bm{u}_{al}'$, then $\hat{A}_i(t) = 1$. The energy norm is a
+traditional normalization choice such that the overlap between $\bm{u},
+\bm{u}_{al}' \propto e^{z/2H}$ is evenly weighted throughout the integration
+region.
+
+For a simulation satisfying $\xi_z k_z \ll 1$ everywhere, we expect
+$\hat{A}_i(t) = 1$ when integrated between the forcing and damping zones,
+corresponding to choices $z_b \gtrsim z_0, z_t \lesssim z_T$ ($z_0, z_T$ are
+defined in \autoref{eq:vol_drive} and \autoref{eq:Gamma} respectively). For
+consistency with the nonlinear case later, we choose $z_b = z_0 + 3\sigma, z_t =
+z_b + H$ (using $z_t = z_T - \Delta z$ just below the damping layer instead does
+not change the results). The resulting measurement of $\hat{A}_i(t)$ is shown in
+\autoref{fig:lin_amps}.
 \begin{figure}
@@
 
-A second prediction of the analytical theory is that the horizontal momentmum
-flux $F(z, t)$ should be spatially flat between the forcing zone where it is
-generated and the damping zone where it is dissipated (\autoref{eq:F_def}). We
-make may a slightly more accurate estimate of $F$ the flux carried in the IGW by
-directly substituting $\bm{u}_{al}$ into \autoref{eq:F_def}. Calling this
-estimate $F_{al}$, we may then measure the agreement of our simulation with
-analytical expectation by computing
-\begin{equation}
-    \hat{F}(z, t) \equiv \frac{\int\limits_0^{L_x}\rho u_xu_z\;\mathrm{d}x}{
-        \int\limits_{0}^{L_x} \bar{\rho} u_{x}u_{z}\;\mathrm{d}x}.
-        \label{eq:sbm_def}
+The analytical theory also predicts that the horizontal momentmum flux $F(z, t)$
+should be spatially flat between the forcing zone where it is generated and the
+damping zone where it is dissipated (\autoref{eq:F_def}). We make may a slightly
+more accurate estimate of $F$ the flux carried in the IGW by directly
+substituting $\bm{u}_{al}'$ into \autoref{eq:F_def}. Calling this estimate
+\begin{equation}
+    F_{al}' \equiv \frac{1}{L_x}
+        \int\limits_{0}^{L_x} \overline{\rho} u_{al,x}'u_{al,z}'\;\mathrm{d}x,
+\end{equation}
+we may then measure the agreement of our simulation with analytical expectation
+by computing
+\begin{equation}
+    \hat{F}(z, t) \equiv \frac{
+        \frac{1}{L_x}\int\limits_0^{L_x}\rho u_xu_z\;\mathrm{d}x}
+        {F_{al}'}.\label{eq:sbm_def}
 \end{equation}
@@
 
+% times on fluxes.png are 054, 086, 153
+
 To perform simulations of wave breaking phenomena, we use the same values as
 \autoref{ss:numerics} except for $C, \nu$. In particular, we choose $C$ such
-that $\xi_z (z_0) k_z = 0.1$ in the forcing zone, which implies $\xi_z(z) k_z
-\gtrsim 1$ before $z_T$ the upper damping zone. The dissipation parameter $\nu$
-was varied across the various simulations. To quantify $\nu$, we define Reynolds
-number
-\begin{equation}
-    \mathrm{Re} = \frac{LU}{\nu} \equiv \frac{\omega}{\nu k_{z}^2}.
-        \label{eq:re_def}
+that $\at{\xi_z k_z}_{z_0} = 0.1$ in the forcing zone, which implies $\xi_z k_z
+$ exceeds $1$ before $z_T$ the upper damping zone. The dissipation parameter
+$\nu$ was varied across the various simulations. To quantify $\nu$, we define
+dimensionless Reynolds number
+\begin{equation}
+    \mathrm{Re}  \equiv \frac{\omega}{\nu k_{z}^2}. \label{eq:re_def}
 \end{equation}
@@
 
-% TODO generate a video for the 0.1 case
-A full video of our higher-resolution run at $N_x = 1024, N_z = 4096,
-\mathrm{Re} = 2048$ is available
-online\footnote{http://www.princeton.edu/~lecoanet/data/breaking\_wave.mov}.
-We take this to be our fiducial simulation for the remainder of this paper,
-though other simulations show qualitatively identical behavior. Slices of
+A full video of our higher-resolution run at $N_x = 768, N_z = 3072,
+\mathrm{Re} = 1024$ is available
+online\footnote{http://www.princeton.edu/~lecoanet/data/breaking\_wave.mov}. We
+take this to be our fiducial simulation for the remainder of this paper, though
+other simulations show qualitatively identical behavior. Slices of
 $\overline{U}, \hat{F}$ across $z$ at various times $t$ are shown in
@@
 \begin{align}
-    z_{c, \min} &= \argmin_z \z{z: F(z) < 0.3F_{al}},\nonumber\\
-    z_{c, \max} &= \argmax_z \z{z: F(z) > 0.3F_{al}},\nonumber\\
+    z_{c, \min} &= \argmin_z \z{z: F(z) < 0.3F_{al}'},\nonumber\\
+    z_{c, \max} &= \argmax_z \z{z: F(z) > 0.3F_{al}'},\nonumber\\
     z_c &\equiv \frac{z_{c, \min} + z_{c, \max}}{2}.\label{eq:zc_def}
@@
 discrepancy between the time-averaged predictor and the data. The general
-agreement clearly demonstrates $\Delta F < F_{al}$ and that incomplete
+agreement clearly demonstrates $\Delta F < F_{al}'$ and that incomplete
 absorption is observed. Moreover, its significant underprediction of critical
@@
     The the poorer but qualitatively correct agreement of the time-averaged
-    predictor with the data shows both that $\Delta F(t) \neq F_{al}$ and that
+    predictor with the data shows both that $\Delta F(t) \neq F_{al}'$ and that
     $\Delta F(t)$ likely has some very real variation with
@@
 $\hat{A}_i(t), \hat{A}_r(t)$ vary somewhat over time, we perform time averaging
-over multiple periods of the base IGW (specifically, averaged over approximately
-$8\pi/\omega$). We can then define the amplitude reflectivity
+over interval approximately $8\pi/\omega$, denoted by angle brackets. We can
+then define the amplitude
+reflectivity
 \begin{equation}
@@
 % 7/N per timestep, 13 timestep smoothing kernel, 24/N period
-Angle brackets denote time averaging here.
 
@@
 
-\section{Equation Implementations}\label{ss:strat_impl}
+\section{Equation Implementations}\label{se:strat_impl}
 
@@
 desire a system where the fluid fields are not divided by one another. We
-introduce $\varpi = \frac{P}{\rho}$ instead, then mandate $\bar{\rho}, \varpi_0$
+introduce $\varpi = \frac{P}{\rho}$ instead, then mandate $\overline{\rho}, \varpi_0$
 background fields satisfy hydrostatic equilibrium $\bm{\nabla}\varpi_0 +
-\varpi_0 \bm{\nabla}\bar{\rho} + g\bm{z} = 0$. Taking isothermal stratification,
+\varpi_0 \bm{\nabla}\overline{\rho} + g\bm{z} = 0$. Taking isothermal stratification,
 we find $\varpi_0 = gH$. We further change variables to $\Upsilon = \ln \rho -
-\ln \bar{\rho}$ and $\varpi_1 = \varpi - \varpi_0$ deviations from the background
+\ln \overline{\rho}$ and $\varpi_1 = \varpi - \varpi_0$ deviations from the background
 state to obtain a system of equations at most quadratic in fluid fields:
@@
         \partial_t \rho + \bm{\nabla} \cdot (\rho \bm{u} - \nu
-            \bm{\nabla}(\rho - \bar{\rho})) &= 0,\label{eq:visc_cons_mom}\\
+            \bm{\nabla}(\rho - \overline{\rho})) &= 0,\label{eq:visc_cons_mom}\\
         \partial_t (\rho \bm{u}) + \bm{\nabla} \cdot (\rho \bm{u}
@@
 Hydrostatic equilibrium is still $\bm{\nabla} \varpi_0 + g\bm{z} = 0$ where
-$\rho = \bar{\rho}, \bm{u} = 0$. Including the damping layers and forcing terms as
+$\rho = \overline{\rho}, \bm{u} = 0$. Including the damping layers and forcing terms as
 described in \autoref{ss:numerics}, we finally obtain the full system of
@@
                 -\Gamma(z) \Upsilon
-                + \frac{F}{\bar{\rho}(z)}e^{-\frac{(z - z_0)^2}{2\sigma^2}}
+                + \frac{F}{\overline{\rho}(z)}e^{-\frac{(z - z_0)^2}{2\sigma^2}}
                     \cos \p{k_xx - \omega t},\\

### 2019-07-31 sims/2d_4_fourier/plot_2d_series_movie.py

@@ -1,0 +1,96 @@
+"""
+Plot planes from joint analysis files.
+
+Usage:
+    plot_2d_series_movie.py <files>... [--output=<dir>] [--write=<num>]
+
+Options:
+    --output=<dir>  Output directory [default: ./frames]
+    --write=<num>   Write number offset [default: 0]
+
+"""
+
+import h5py
+import numpy as np
+import matplotlib
+matplotlib.use('Agg')
+import matplotlib.pyplot as plt
+plt.ioff()
+from dedalus.extras import plot_tools
+import re
+
+def main(filename, start, count, output, write_start):
+    """Save plot of specified tasks for given range of analysis writes."""
+
+    # Plot settings
+    tasks = ['ux','uz']
+    scale = 2.5
+    dpi = 300
+    title_func = lambda sim_time: r'$t = {:.3f}$'.format(sim_time)
+    # savename_func = lambda write: 'yubo_{:06}.png'.format(write+write_start)
+    def savename_func(write):
+        # regex parse out filename number, 10 writes per file
+        num_per_file = 10
+        write_num = (write - 1) % num_per_file
+        filenum = int(re.match('.*s([\d]+)\.h5', filename)[1])
+        suffix_num = (filenum - 1) * num_per_file + write_num
+        if suffix_num > 201:
+            suffix_num -= 9 # hard coded, one of the restores is misnumbered
+        print(suffix_num)
+        return 'yubo_{:06}.png'.format(suffix_num)
+    # Layout
+    nrows, ncols = 1, 2
+    image = plot_tools.Box(1, 2)
+    pad = plot_tools.Frame(0.2, 0.2, 0.1, 0.1)
+    margin = plot_tools.Frame(0.3, 0.2, 0.1, 0.1)
+
+    # Create multifigure
+    mfig = plot_tools.MultiFigure(nrows, ncols, image, pad, margin, scale)
+    fig = mfig.figure
+    # Plot writes
+    with h5py.File(filename, mode='r') as file:
+        for index in range(start, start+count):
+            for n, task in enumerate(tasks):
+                # Build subfigure axes
+                i, j = divmod(n, ncols)
+                axes = mfig.add_axes(i, j, [0, 0, 1, 1])
+                # Call 3D plotting helper, slicing in time
+                dset = file['tasks'][task]
+                if n==0: #ux
+                  clim = [-0.25, 0.25]
+                  title = r'$u_x$'
+                elif n==1: #uz
+                  clim = [-0.1, 0.1]
+                  title = r'$u_z$'
[…]

### 2019-07-31 sims/2d_4_fourier/strat.py

@@ -68,35 +68,35 @@ if __name__ == '__main__':
     tasks = [
         # ('yubo_nu0p5_hres',
         #  get_params(overrides={'Re_inv': 0.05})),
-        ('yubo_nu0p5_shres',
-         get_params(overrides={'Re_inv': 0.05,
-                               'N_X': 1024,
-                               'N_Z': 4096})),
+        # ('yubo_nu0p5_shres',
+        #  get_params(overrides={'Re_inv': 0.05,
+        #                        'N_X': 1024,
+        #                        'N_Z': 4096})),
         # ('yubo_nu1_hres',
         #  get_params(overrides={'Re_inv': 0.1})),
         ('yubo_nu1_vhres',
          get_params(overrides={'Re_inv': 0.1,
                                'N_X': 768,
                                'N_Z': 3072})),
-        ('yubo_nu2_hres',
-         get_params(overrides={'Re_inv': 0.2})),
+        # ('yubo_nu2_hres',
+        #  get_params(overrides={'Re_inv': 0.2})),
         # ('yubo_nu2_width',
         #  get_params(overrides={'Re_inv': 0.2,
         #                        'N_X': 256,
         #                        'NUM_SNAPSHOTS': 1500,
         #                        'N_Z': 1024})),
-        ('yubo_nu3_width',
-         get_params(overrides={'Re_inv': 0.3,
-                               'N_X': 256,
-                               'N_Z': 1024})),
-        ('yubo_nu5',
-         get_params(overrides={'Re_inv': 0.5,
-                               'N_X': 256,
-                               'N_Z': 1024})),
-        ('yubo_nu7',
-         get_params(overrides={'Re_inv': 0.7,
-                               'N_X': 256,
-                               'N_Z': 1024})),
+        # ('yubo_nu3_width',
+        #  get_params(overrides={'Re_inv': 0.3,
+        #                        'N_X': 256,
+        #                        'N_Z': 1024})),
+        # ('yubo_nu5',
+        #  get_params(overrides={'Re_inv': 0.5,
+        #                        'N_X': 256,
+        #                        'N_Z': 1024})),
+        # ('yubo_nu7',
+        #  get_params(overrides={'Re_inv': 0.7,
+        #                        'N_X': 256,
+        #                        'N_Z': 1024})),
     ]
     if '-plot' in sys.argv:
         for name, params_dict in tasks:

### 2019-08-01 writeups/paper.tex

@@
     ]{caption}
+    \usepackage[caption=false]{subfig} % subfigures
 
@@
 take this to be our fiducial simulation for the remainder of this paper, though
-other simulations show qualitatively identical behavior. Slices of
+other simulations show qualitatively identical behavior. Snapshots of the
+velocity field $\bm{u}$ at select times can be found in \autoref{fig:snapshots}.
+% convert yubo_000054.png -crop 2000x2000+150+250 out.png
+\begin{figure*}
+    \subfloat[$t=413.4$ At early times, the flow resembles a linear IGW lower in
+    the simulation domain but breaks down into smaller-scale features at higher
+    $z$. Some characteristic swirling motion can be seen in both $u_x, u_z$,
+    highly suggestive of Kelvin-Helmholtz instabilities.]{
+    \includegraphics[width=0.45\textwidth]{plots/yubo_000054.png}}\hfil
+    %
+    \subfloat[$t=658.5$ At a slightly later time, the mean flow in $u_x$ has
+    become much more prominent and the critical layer $z_c$ has become much more
+    definite. Small-scale fluctuations are still present in $u_z$ albeit at
+    smaller amplitudes due to being in a denser region of the fluid.]{
+    \includegraphics[width=0.45\textwidth]{plots/yubo_000086.png}}
+
+    \subfloat[$t=1171.4$ The critical layer transition is now extremely sharp,
+    and small swirls of limited vertical extent at the location of the critical
+    layer in $u_z$ suggest that the Kelvin-Helmholtz instability is responsible
+    for regulating the minimum width of this transition, a hypothesis explored
+    in \autoref{ss:khi}.]{
+    \includegraphics[width=0.45\textwidth]{plots/yubo_000153.png}}\hfil
+    %
+    \subfloat[$t=3437.8$ The end of the simulation shows very few significant
+    qualitative differences from the previous snapshot, suggesting that the
+    latter half of our simulation is temporally converged.]{
+    \includegraphics[width=0.45\textwidth]{plots/yubo_000451.png}}
+
+    \caption{Snapshots of $u_x, u_z$ in the fiducial simulation illustrating
+    distinct phases of the evolution of the flow. Note that $\bar{U}_c \approx
+    0.16$ for the parameters used.}\label{fig:snapshots}
+\end{figure*}
+
+Slices of
 $\overline{U}, \hat{F}$ across $z$ at various times $t$ are shown in
@@
 
-\subsection{Kelvin-Helmholtz Instability and Critical Layer Width}
+\subsection{Kelvin-Helmholtz Instability and Critical Layer Width}\label{ss:khi}
 

### 2019-08-01 writeups/paper.tex
stash

@@
 \overline{\rho} gH$. A vertically-stratified shear flow
-$\overline{u}_x(z)\bm{x}$ is permitted in hydrostatic equilibrium, but we will
-assume no background flow, and $\bm{u} = \bm{u}'$. Physically, this assumption
+$\overline{u}_x(z)\bm{x}$ is permitted at hydrostatic equilibrium, but we will
+assume no background flow, so $\bm{u} = \bm{u}'$. Physically, this assumption
 corresponds to a non-rotating star, or going to the corotating frame of a
@@
 Note that in the new coordinates, hydrostatic equilibrium corresponds to
-$\Upsilon = 1, \overline{\varpi} = gH$.
+$\Upsilon = 0, \overline{\varpi} = gH$.
 
@@
 
-In the weak stratification limit $k_{z}H \gg 1$, the solution exhibits the
+In the weak stratification limit $\abs{k_{z}H} \gg 1$, the solution exhibits the
 following characteristics:
@@
         \mathcal{O}\p{\frac{1}{\p{k_{z}H}^{2}}} \approx 0$, the usual result
-        (e.g.~\cite{drazin,sutherland1}). Note again that $c_{gz} < 0$ for
-        $k_{x}, k_{z} > 0$, so upwards-propagating IGW have $k_z < 0$.
+        (e.g.~\cite{drazin,sutherland1}). Note that for an upwards propagating
+        IGW $c_{g, z} > 0$, $k_z < 0, k_x > 0$.
 
@@
         direction can be computed. Since the linear solution consists of terms
-        separable as $f(x, z, t) = f(z)e^{i(k_{1x}x - \omega_1 t)}$, $x$
+        separable as $f(x, z, t) = f(z)e^{i(k_{x}x - \omega t)}$, $x$
         averaging and time averaging are equivalent. We may then write
@@
 As not to interfere with the incompressibility constraint, we force the system
-on the density equation. We implement forcing with strength $C$ by replacing
-\autoref{eq:nl_up_density} with
+on the density equation. We implement forcing with strength $C$ localized around
+$z_0$ with width $\sigma$ by replacing \autoref{eq:nl_up_density} with
 \begin{equation}
@@
 by considering the propagation of $F$ (\autoref{eq:F_def}) into a medium at
-rest. Solving for $\overline{U} = \overline{U}_c \equiv \omega / k_x$ the
-critical layer condition (see \autoref{ss:crit_layer} immediately below), we
-obtain $\abs{\frac{u_{x}'k_{x}}{\omega}} = \abs{\xi_x k_{x}} = \abs{\xi_z k_z}
-\sim 1$, equal to the nonlinearity criterion \autoref{eq:nl}. This condition
-predicts where the wave's self-acceleration alone produces a critical layer.
+rest. Solving for $\overline{U} = \omega / k_x$ the critical layer condition
+(\autoref{eq:u_c} and discussion in \autoref{ss:crit_layer}), we obtain
+$\abs{\frac{u_{x}'k_{x}}{\omega}} = \abs{\xi_x k_{x}} = \abs{\xi_z k_z} \sim 1$,
+equal to the nonlinearity criterion \autoref{eq:nl}. This condition predicts
+where the wave's self-acceleration alone produces a critical layer.
 
@@
 $\partial_t$ to their frequency in the comoving frame of the fluid $\partial_t -
-\overline{U}(z)\partial_x$. For a critical value $\omega - \overline{U}_c k_{x}
-= 0$, the horizontal wavenumber of the linear wave in the fluid's frame of
-reference vanishes and critical behavior is observed. In a linear
-\citep{booker_bretherton} or viscous \citep{hazel} theory, the incident wave has
+\overline{U}(z)\partial_x$. If anywhere $\overline{U}(z) = \overline{U}_c$ where
+\begin{equation}
+    \overline{U}_c \equiv \omega / k_x,\label{eq:u_c}
+\end{equation}
+then the horizontal wavenumber of the linear wave in the fluid frame vanishes
+and critical behavior is observed. In a linear theory, both inviscid
+\citep{booker_bretherton} and viscous \citep{hazel}, the incident wave has
 amplitude reflection and transmission coefficients
@@
         \label{eq:ri_def}\\
-    \mathcal{R} &= e^{-2\pi \sqrt{\mathrm{Ri} - \frac{1}{4}}}, &
-    \mathcal{T} &= e^{-\pi \sqrt{\mathrm{Ri} - \frac{1}{4}}},
+    \mathcal{R}_A &= e^{-2\pi \sqrt{\mathrm{Ri} - \frac{1}{4}}}, &
+    \mathcal{T}_A &= e^{-\pi \sqrt{\mathrm{Ri} - \frac{1}{4}}},
         \label{eq:crit_coeffs}
@@
     \pd{}{t}\int\limits \rho(z) \overline{U}(z, t)\;\mathrm{d}z
-        + \Delta S = 0.
-\end{equation}
-$\Delta S$ is the total injected horizontal momentum to the flow. We next make a
+        + \Delta F = 0.
+\end{equation}
+$\Delta F$ is the total injected horizontal momentum to the flow. We next make a
 simple two-zone approximation $\overline{U}(z > z_c) = \overline{U}_c,
@@
 \begin{equation}
-    \rho(z_c) \overline{U}_c\pd{z_c}{t} = \Delta S.\label{eq:zc_anal}
-\end{equation}
-For constant $\Delta S$ in time and $\rho \approx \overline{\rho}$, this has
+    \rho(z_c) \overline{U}_c\pd{z_c}{t} = \Delta F.\label{eq:zc_anal}
+\end{equation}
+If the entirety of the horizontal momentum flux carried by the IGW wavetrain is
+absorbed by the critical layer, then $\Delta F = -F'$ is constant in time.
+For constant $\Delta F$ in time and $\rho \approx \overline{\rho}$, $z_c(t)$ has
 analytical solution
 \begin{equation}
-    z_c(t) = -H\ln t - H\ln \frac{H\overline{\rho}_0\overline{U}_c}{-\Delta S}
+    z_c(t) = -H\ln t - H\ln \frac{H\overline{\rho}_0\overline{U}_c}{-\Delta F}
         .\label{eq:zc_sol}
 \end{equation}
-Note that constant $-\Delta S$ corresponds physically to the picture where
-constant $S$ carried by the IGWs, given by \autoref{eq:S_lin}, is fully absorbed
-at the critical layer, in which case $\Delta S = -S$.
 
@@
 
-We nondimensionalize by taking $H = N = \overline{\rho}_0 = 1$ in
-\autoref{se:nl_orig}.
+We nondimensionalize by taking $H = N = \overline{\rho}_0 = 1$.
 
@@
     \caption{Snapshots of $u_x, u_z$ in the fiducial simulation illustrating
-    distinct phases of the evolution of the flow. Note that $\bar{U}_c \approx
+    distinct phases of the evolution of the flow. Note that $\overline{U}_c \approx
     0.16$ for the parameters used.}\label{fig:snapshots}
@@
     $\overline{U}, \hat{F}$ follow their definitions in \autoref{eq:mean_flow}
-    and \autoref{eq:sbm_def}. $\overline{U}$ is given in units of $c_{ph, x} =
-    \frac{\omega}{k_{x}}$ the critical layer flow velocity (see
-    \autoref{ss:crit_layer}). The propagation of the critical layer towards
-    lower $z$ and sharp deposition of $F$ at the critical layer are
-    evident.}\label{fig:nl_fluxes}
+    and \autoref{eq:sbm_def}. $\overline{U}$ is given in units of
+    $\overline{U}_c = \frac{\omega}{k_{x}}$ the critical horizontal flow
+    velocity (see \autoref{ss:crit_layer}). The propagation of the critical
+    layer towards lower $z$ and sharp deposition of $F$ at the critical layer
+    are evident.}\label{fig:nl_fluxes}
 \end{figure}
@@
 The observation of reflected/transmitted waves is in accordange with previous
-studies as discussed in \autoref{ss:wave_breaking}. Many studies have shown that
-a local Richardson number $\mathrm{Ri} \sim 1/4$ corresponds to the onset of
-reflectivity. This agrees with the observed Kelvin-Helmholtz instability in our
-simulations [TODO image]. To verify that the critical layer's width is bound
-from below by Kelvin-Helmholtz instabilities, we decided to compute a
-$\mathrm{Ri}$ within the critical layer.
+studies as discussed in \autoref{ss:wave_breaking}. Studies have shown that a
+local Richardson number $\mathrm{Ri} \sim 1/4$ corresponds to the onset of
+reflectivity. This $\mathrm{Ri}$ also corresponds to the onset of the
+Kelvin-Helmholtz instability (KHI). Indeed, in our simulations, visual
+inspection suggests the KHI is present in the critical layer (see
+\autoref{fig:snapshots}). It is natural to suspect then that the shear flow
+cannot steepen any further than KHI onset. To verify this, we decided to compute
+a local $\mathrm{Ri}$ for the shear flow around the critical layer.
 
@@
 To understand the variation in $\mathrm{Ri}$ over $x$, we can also plot using
-the minimum over $x$; the maximum is significantly noisier. Both of these are
+the minimum over $x$ (the maximum is significantly noisier). Both of these are
 shown in \autoref{fig:nl_f_ri}. The quick evolution of $\mathrm{Ri}$ to its
@@
     as defined in \autoref{ss:ri_med_def}. Dotted red lines demarcate the
-    minimum and maximum of $\mathrm{Ri}$ over $x$ instead of the median. These
-    numbers effectively measure the mean and spread in width of the critical
-    layer over $x$. Note that $\mathrm{Ri} \lesssim \frac{1}{4}$ corresponds
-    to Kelvin-Helmholtz instability, so this plot suggests the critical layer
-    width roughly saturates the lower bound from the Kelvin-Helmholtz
-    instability after an initial transient phase.}\label{fig:nl_f_ri}
+    minimum and median of $\mathrm{Ri}$ over $x$. These numbers effectively
+    measure the mean and spread in width of the critical layer over $x$. Note
+    that $\mathrm{Ri} \sim \frac{1}{4}$ corresponds to the KHI, so this plot
+    suggests the shear at the critical layer does not steepen past the KHI
+    onset.}\label{fig:nl_f_ri}
 \end{figure}
@@
 reflected wave with wave vector $\bm{k} = k_{x}\bm{x} + k_{z}\bm{z}$, and
-(ii) the amount of horizontal momentum flux $S$ reflected off the critical
+(ii) the amount of horizontal momentum flux $F$ reflected off the critical
 layer.
@@
     parameterized with $\mathrm{Re}$ (\autoref{eq:re_def}). Error bars depict
-    $16$ and $84$ percentile values. Fmall horizontal displacements are made for
+    $16$ and $84$ percentile values. Small horizontal displacements are made for
     data points at identical $\mathrm{Re}$ for readability. Note that
@@
     are more physically realistic, and our values seem to converge towards large
-    $\mathrm{Re}$.}\label{fig:agg}
+    $\mathrm{Re}$. At the smallest $\mathrm{Re}$ value, $\mathrm{Ri} \approx 50$
+    is too large to fit on the plot.}\label{fig:agg}
 \end{figure}
+
+It is apparent that the Richardson number rapidly converges to $\mathrm{Ri}
+\approx 0.4$ but the reflection/transmission coefficients converge more slowly.
+This is in disagreement with \autoref{eq:crit_coeffs} calculated via the linear
+analytical theory. This tension is natural: fluid motion within the critical
+layer is turbulent, so transmission/reflection at the critical layer cannot be
+captured by any linear theory.
 
@@
     \item $k_x, k_z$: In astrophysical systems, $k_{\perp} \ll k_r$. While we do
-        not explore different $k_{1x}, k_{1z}, \omega_1$ in this study, with
+        not explore different $k_{x}, k_{z}, \omega_1$ in this study, with
         outgoing boundary conditions, there appears at first to be no $z$ length
-        scale other than $k_{1z}$, so our results would seem to be invariant
+        scale other than $k_{z}$, so our results would seem to be invariant
         under rescaling of the $z$ length scale. However, true turbulence is
-        expected to be isotropic at small scales, which may couple $k_{1x},
-        k_{1z}$ in a way that $k_{1x} \ll k_{1z}$ produces different dynamics
-        than $k_{1x} \lesssim k_{1z}$ as we've studied. This is a numerically
+        expected to be isotropic at small scales, which may couple $k_{x},
+        k_{z}$ in a way that $k_{x} \ll k_{z}$ produces different dynamics
+        than $k_{x} \lesssim k_{z}$ as we've studied. This is a numerically
         difficult regime though, so we defer consideration to future work.
@@
             - \frac{u_z}{H} &= 0,\\
-        \pd{u_{1x}}{t} + \p{\bm{u}_1 \cdot \bm{\nabla}}u_{1x}
+        \pd{u_{x}}{t} + \p{\bm{u}_1 \cdot \bm{\nabla}}u_{x}
             + \pd{\varpi_1}{x} + gH\pd{\Upsilon}{x}

### 2019-08-07 writeups/Su_IGW_break.tex
rename

--- 2019-08-01 writeups/paper.tex
@@
 
-We identify two instances of non-absorptive behavior: (i), the presence of a
-reflected wave with wave vector $\bm{k} = k_{x}\bm{x} + k_{z}\bm{z}$, and
-(ii) the amount of horizontal momentum flux $F$ reflected off the critical
-layer.
+We identify two manifestations of non-absorptive behavior: (i), the presence of
+a reflected wave with wave vector $\bm{k} = k_{x}\bm{x} - k_{z}\bm{z}$, and (ii)
+the amount of horizontal momentum flux $F$ reflected off the critical layer.
+These two quantities need not agree exactly if some reflected flux is in
+higher-order modes, which is observed. Both are of physical interest, however:
+he reflected wave amplitude may be of physical interest in setting up standing
+modes in a realistic star, while the reflected flux is important for accurately
+tracking angular momentum transfer during synchronization.
 
@@
 \autoref{fig:nl_f_amps}. The oscillations in $\hat{A}_i, F_i, A_d, F_d$ are
-apparent but the mean values seem to have temporally converged.
+significant but the mean values seem to have temporally converged.
 
@@
     t_A = \int\limits_0^R \frac{\sqrt{4\pi \rho}}{B_0}\;\mathrm{d}r
-        \sim 10^2\;\mathrm{yr}\p*{\frac{10^3\;\mathrm{G}}{B}}.
+        \sim 10^2\;\mathrm{yr}\p{\frac{10^3\;\mathrm{G}}{B}}.
 \end{equation}
@@
 timescale is the synchronization timescale $t_s$. For a tidal torque $\tau$ and
-tidal forcing frequency $\sigma = m\p*{\Omega - \Omega_{spin}}$, we note that
+tidal forcing frequency $\sigma = m\p{\Omega - \Omega_{spin}}$, we note that
 angular momentum transfer is $\pd{M_{sync}}{t} \sigma R^2 = \tau$, where
@@
 \bibliographystyle{mnras}
-\bibliography{paper}
+\bibliography{Su_IGW_break}
 
@@
 desire a system where the fluid fields are not divided by one another. We
-introduce $\varpi = \frac{P}{\rho}$ instead, then mandate $\overline{\rho}, \varpi_0$
-background fields satisfy hydrostatic equilibrium $\bm{\nabla}\varpi_0 +
-\varpi_0 \bm{\nabla}\overline{\rho} + g\bm{z} = 0$. Taking isothermal stratification,
-we find $\varpi_0 = gH$. We further change variables to $\Upsilon = \ln \rho -
-\ln \overline{\rho}$ and $\varpi_1 = \varpi - \varpi_0$ deviations from the background
-state to obtain a system of equations at most quadratic in fluid fields:
+introduce $\varpi = \frac{P}{\rho}$ instead, then mandate $\overline{\rho},
+\overline{\varpi}$ background fields satisfy hydrostatic equilibrium
+$\bm{\nabla}\overline{\varpi} + \overline{\varpi} \bm{\nabla}\overline{\rho} +
+g\bm{z} = 0$. Taking isothermal stratification, we find $\overline{\varpi} = gH$. We
+further change variables to $\Upsilon = \ln \rho - \ln \overline{\rho}$ and
+$\varpi' = \varpi - \overline{\varpi}$ deviations from the background state to obtain a
+system of equations at most quadratic in fluid fields:
 \begin{subequations}\label{se:nl_var}
     \begin{align}
-        \bm{\nabla} \cdot \bm{u}_1 &= 0,\\
-        \pd{\Upsilon}{t} + \p{\bm{u}_1 \cdot \bm{\nabla}} \Upsilon
+        \bm{\nabla} \cdot \bm{u}' &= 0,\\
+        \pd{\Upsilon}{t} + \p{\bm{u}' \cdot \bm{\nabla}} \Upsilon
             - \frac{u_z}{H} &= 0,\\
-        \pd{u_{x}}{t} + \p{\bm{u}_1 \cdot \bm{\nabla}}u_{x}
-            + \pd{\varpi_1}{x} + gH\pd{\Upsilon}{x}
-            + \varpi_1 \pd{\Upsilon}{x} &= 0,\\
-        \pd{u_z}{t} + \p{\bm{u}_1 \cdot \bm{\nabla}}u_z
-            + \pd{\varpi_1}{z} + gH\pd{\Upsilon}{z}
-            + \varpi_1 \pd{\Upsilon}{z} - \frac{\varpi_1}{H} &= 0.
+        \pd{u_{x}}{t} + \p{\bm{u}' \cdot \bm{\nabla}}u_{x}
+            + \pd{\varpi'}{x} + gH\pd{\Upsilon}{x}
+            + \varpi' \pd{\Upsilon}{x} &= 0,\\
+        \pd{u_z}{t} + \p{\bm{u}' \cdot \bm{\nabla}}u_z
+            + \pd{\varpi'}{z} + gH\pd{\Upsilon}{z}
+            + \varpi' \pd{\Upsilon}{z} - \frac{\varpi'}{H} &= 0.
     \end{align}
@@
 dissipation, artificial dissipation must be added. To ensure the dissipitive
-system conserves horizontal momentum, we begin by adding dissipitive terms to
-the flux-conservative form of the Euler fluid equations \autoref{se:nl_orig} (we
-use stress tensor $\tau_{ij} = P\delta_{ij}$):
+system conserves horizontal momentum exactly, we begin by adding dissipitive
+terms to the flux-conservative form of the Euler fluid equations
+\autoref{se:nl_orig} (we use stress tensor $\tau_{ij} = P\delta_{ij}$):
 \begin{subequations}
     \begin{align}
-        \bm{\nabla} \cdot \bm{u} &= 0,\\
-        \partial_t \rho + \bm{\nabla} \cdot (\rho \bm{u} - \nu
+        \bm{\nabla} \cdot \bm{u}' &= 0,\\
+        \partial_t \rho + \bm{\nabla} \cdot (\rho \bm{u}' - \nu
             \bm{\nabla}(\rho - \overline{\rho})) &= 0,\label{eq:visc_cons_mom}\\
-        \partial_t (\rho \bm{u}) + \bm{\nabla} \cdot (\rho \bm{u}
-        \bm{u} +
-            \mathrm{diag}(\rho \varpi) - \nu \rho \bm{\nabla}\bm{u}) + \rho g
-            \bm{z} &= 0.
+        \partial_t (\rho \bm{u}') + \bm{\nabla} \cdot (\rho \bm{u}' \bm{u}'
+            + \mathrm{diag}(\rho \varpi)
+            - \nu \rho \bm{\nabla}\bm{u})
+            + \rho g \bm{z} &= 0.
     \end{align}
@@
 numerical stability, we choose it such that hydrostatic equilibrium is not
-modified. Some algebraic manipulation to re-cast it in the form of
-\autoref{se:nl_var} gives
+modified (hence $\nu$ acts only on $\rho - \overline{\rho}$).
+% Some algebraic manipulation to re-cast it in the form of \autoref{se:nl_var}
+% gives
+% \begin{subequations}
+%     \begin{align}
+%         \bm{\nabla} \cdot \bm{u} &= 0,\\
+%         \partial_t \Upsilon + \p{\bm{u} \cdot \bm{\nabla}} \Upsilon -
+%             \frac{u_z}{H} - \nu\p{\nabla^2 \Upsilon + \p{\bm{\nabla}
+%             \Upsilon} \cdot \p{\bm{\nabla}\Upsilon} - \frac{2}{H}\partial_z
+%             \Upsilon + \frac{1 - e^{-\Upsilon}}{H^2}} &= 0,\\
+%         \partial_t \bm{u} + \p{\bm{u} \cdot \bm{\nabla}}\bm{u} +
+%             \bm{\nabla} \varpi + \varpi \bm{\nabla} \Upsilon - \nu \nabla^2
+%             \bm{u} + \bm{u} \nu\p{\nabla^2 \Upsilon + \p{\bm{\nabla}
+%             \Upsilon} \cdot \p{\bm{\nabla}\Upsilon} - \frac{2}{H}\partial_z
+%             \Upsilon + \frac{1 - e^{-\Upsilon}}{H^2}}&{}\nonumber\\
+%         - 2\nu \p{\p{\p{\bm{\nabla}\Upsilon} \cdot \bm{\nabla}}\bm{u} -
+%             \frac{1}{H}\partial_z \bm{u}} - \frac{\varpi'}{H} &= 0.
+%     \end{align}
+% \end{subequations}
+One last consideration we found necessary was masking out nonlinear terms in the
+forcing zone with a similar form to \autoref{eq:Gamma}. In the absence of this
+mask, a strong mean flow localized to the forcing zone developed. The mask used
+was
+\begin{equation}
+    \Gamma_{NL}(z) = \frac{1}{2}\s{2
+        + \tanh \frac{z - (z_0 + 8\sigma)}{\sigma}
+        - \tanh \frac{z - z_B}{\sigma}}.
+\end{equation}
+
+Including the damping layers and forcing terms as described in
+\autoref{ss:numerics}, we finally obtain the full system of equations as
+simulated in Dedalus:
 \begin{subequations}
     \begin{align}
-        \bm{\nabla} \cdot \bm{u} &= 0,\\
-        \partial_t \Upsilon + \p{\bm{u} \cdot \bm{\nabla}} \Upsilon -
-            \frac{u_z}{H} - \nu\p{\nabla^2 \Upsilon + \p{\bm{\nabla}
-            \Upsilon} \cdot \p{\bm{\nabla}\Upsilon} - \frac{2}{H}\partial_z
-            \Upsilon + \frac{1 - e^{-\Upsilon}}{H^2}} &= 0,\\
-        \partial_t \bm{u} + \p{\bm{u} \cdot \bm{\nabla}}\bm{u} +
-            \bm{\nabla} \varpi + \varpi \bm{\nabla} \Upsilon - \nu \nabla^2
-            \bm{u} + \bm{u} \nu\p{\nabla^2 \Upsilon + \p{\bm{\nabla}
-            \Upsilon} \cdot \p{\bm{\nabla}\Upsilon} - \frac{2}{H}\partial_z
-            \Upsilon + \frac{1 - e^{-\Upsilon}}{H^2}}&{}\nonumber\\
-        - 2\nu \p{\p{\p{\bm{\nabla}\Upsilon} \cdot \bm{\nabla}}\bm{u} -
-            \frac{1}{H}\partial_z \bm{u}} - \frac{\varpi_1}{H} &= 0.
-    \end{align}
-\end{subequations}
-Hydrostatic equilibrium is still $\bm{\nabla} \varpi_0 + g\bm{z} = 0$ where
-$\rho = \overline{\rho}, \bm{u} = 0$. Including the damping layers and forcing terms as
-described in \autoref{ss:numerics}, we finally obtain the full system of
-equations as simulated in Dedalus: [TODO update w/ mask]
-\begin{subequations}
-    \begin{align}
-        \bm{\nabla} \cdot \bm{u} ={}& 0,\\
+        \bm{\nabla} \cdot \bm{u}' ={}& 0,\\
         \partial_t \Upsilon - \frac{u_z}{H}
-            ={}& \nu\p{\nabla^2 \Upsilon + \p{\bm{\nabla}
-            \Upsilon} \cdot \p{\bm{\nabla}\Upsilon} - \frac{2}{H}\partial_z
-            \Upsilon + \frac{1 - e^{-\Upsilon}}{H^2}},\nonumber\\
-            & - \p{\bm{u} \cdot \bm{\nabla}}\Upsilon
-                -\Gamma(z) \Upsilon
+            ={}& -\Gamma(z) \Upsilon
                 + \frac{F}{\overline{\rho}(z)}e^{-\frac{(z - z_0)^2}{2\sigma^2}}
-                    \cos \p{k_xx - \omega t},\\
-        \pd{u_x}{t} + \pd{T}{x} + gH\pd{\Upsilon}{x} ={}&
-            \nu \nabla^2 u_x
+                    \cos \p{k_xx - \omega t},\nonumber\\
+            & + \Gamma_{NL} \bigg[-\p{\bm{u}' \cdot \bm{\nabla}}\Upsilon
+                + \nu\p{\nabla^2 \Upsilon + \p{
+                    \bm{\nabla} \Upsilon} \cdot \p{\bm{\nabla}\Upsilon}
+                    - \frac{2}{H}\partial_z \Upsilon
+                    + \frac{1 - e^{-\Upsilon}}{H^2}}\bigg],\\
+        \pd{u_x}{t} + \pd{\varpi'}{x} + gH\pd{\Upsilon}{x} ={}&
+            -\Gamma(z) u_x
+            + \Gamma_{NL}\bigg[\nu \nabla^2 u_x
             - u_x \nu\p{\nabla^2 \Upsilon + \p{\bm{\nabla} \Upsilon} \cdot
@@
                 - \frac{1}{H}\partial_z u_x}
-            -\Gamma(z) u_x
-                - \p{\bm{u} \cdot \bm{\nabla}}u_x
-                - T_1 \pd{\Upsilon}{x},\\
-        \pd{u_z}{t} + \pd{T}{z} + gH\pd{\Upsilon}{z} - \frac{T_1}{H} ={}&
-            \nu \nabla^2 u_z
+                - \p{\bm{u}' \cdot \bm{\nabla}}u_x
+                - \varpi' \pd{\Upsilon}{x}\bigg],\\
+        \pd{u_z}{t} + \pd{\varpi}{z} + gH\pd{\Upsilon}{z} - \frac{\varpi'}{H}
+            ={}& -\Gamma(z) u_z
+            +\Gamma_{NL}\bigg[\nu \nabla^2 u_z
             - u_z \nu\p{\nabla^2 \Upsilon + \p{\bm{\nabla} \Upsilon} \cdot
@@
                 \frac{1}{H}\partial_z u_{z}}
-            -\Gamma(z) u_z - \p{\bm{u} \cdot \bm{\nabla}}u_z
-            - T_1 \pd{\Upsilon}{z}.
+            - \p{\bm{u}' \cdot \bm{\nabla}}u_z
+            - \varpi' \pd{\Upsilon}{z}\bigg].
     \end{align}

### 2019-08-08 writeups/Su_IGW_break.tex
bleh

@@
     \newcommand*{\abs}[1]{\left|#1\right|}
-    \newcommand*{\ev}[1]{\langle#1\rangle}
+    \newcommand*{\ev}[1]{\left\langle#1\right\rangle}
     \newcommand*{\p}[1]{\left(#1\right)}
@@
 \begin{equation}
-    u'_z\p{x, z, t} = Ae^{z/2H}e^{i\p{k_{x}x + k_{z}z - \omega t}},
+    u'_z\p{x, z, t} = Ae^{z/2H}\cos\p{\p{k_{x}x + k_{z}z - \omega t}},
         \label{eq:k0z_sign}
@@
         direction can be computed. Since the linear solution consists of terms
-        separable as $f(x, z, t) = f(z)e^{i(k_{x}x - \omega t)}$, $x$
-        averaging and time averaging are equivalent. We may then write
+        separable as $f(x, z, t) = f(z)\cos\p{k_{x}x - \omega t}$, $x$
+        averaging and time averaging are equivalent. $x$ averaging is preferred
+        both since periodic boundary conditions pick a natural averaging length
+        and since it does not introduce extra temporal smoothing. We may then
+        write
         \begin{align}
@@
 Since we are only concerned with the behavior of the wave at high $z$ where wave
-breaking occurs, we generate waves in the most numerically convenient way. As
-long as physically represesentative IGW are generated and wave dynamics are only
-examined far from the forcing region, the exact excitation mechanism is not
-important for our conclusions.
+breaking occurs, we generate waves at low $z$ in the most numerically convenient
+way. As long as physically represesentative, upwards-propgaating IGW are
+generated and wave dynamics are only examined far from the forcing region, the
+exact excitation mechanism is not important for our conclusions.
 
@@
 on the density equation. We implement forcing with strength $C$ localized around
-$z_0$ with width $\sigma$ by replacing \autoref{eq:nl_up_density} with
+$z_0$ with small width $\sigma$ by replacing \autoref{eq:nl_up_density} with
 \begin{equation}
@@
 $k_{z}$ satisfying dispersion relation \autoref{eq:disp_rel} for the given
-$k_{x}, \omega(k_{x}, k_{z})$ will propagate.
+$k_{x}$, $\omega(k_{x}, k_{z})$ will propagate.
 
@@
         &{}\begin{cases}
-        e^{\frac{z - z_0}{2H}}e^{i\p{k_{x}x + k_{z}(z - z_0) - \omega t
-            + \frac{1}{2k_{z}H}}}
+        e^{\frac{z - z_0}{2H}}\cos\p{k_{x}x + k_{z}(z - z_0) - \omega t
+            + \frac{1}{2k_{z}H}}
             & z > z_0\\
-        e^{\frac{z - z_0}{2H}}e^{i\p{k_{x}x - k_{z}(z - z_0) - \omega t
-            + \frac{1}{2k_{z}H}}}
+        e^{\frac{z - z_0}{2H}}\cos\p{k_{x}x - k_{z}(z - z_0) - \omega t
+            + \frac{1}{2k_{z}H}}
             & z < z_0\\
@@
 
-\subsection{Wave Breaking Location}\label{ss:wave_breaking}
+\subsection{Wave Breaking Height}\label{ss:wave_breaking}
 
@@
 by considering the propagation of $F$ (\autoref{eq:F_def}) into a medium at
-rest. Solving for $\overline{U} = \omega / k_x$ the critical layer condition
-(\autoref{eq:u_c} and discussion in \autoref{ss:crit_layer}), we obtain
-$\abs{\frac{u_{x}'k_{x}}{\omega}} = \abs{\xi_x k_{x}} = \abs{\xi_z k_z} \sim 1$,
-equal to the nonlinearity criterion \autoref{eq:nl}. This condition predicts
-where the wave's self-acceleration alone produces a critical layer.
+rest. Solving for where $\overline{U} = \omega / k_x$ the critical layer
+condition is fulfilled (\autoref{eq:u_c} and discussion in
+\autoref{ss:crit_layer}), we obtain $\abs{\frac{u_{x}'k_{x}}{\omega}} =
+\abs{\xi_x k_{x}} = \abs{\xi_z k_z} \sim 1$, equal to the nonlinearity criterion
+\autoref{eq:nl}. This condition predicts where the wave's self-acceleration
+alone produces a critical layer.
 
@@
 
-% TODO physical vs simulation domain
-Denote $L_x, L_z$ to be the physical dimensions of the simulation domain. Our
-simulation domain is smaller than our full physical domain, so we adopt
-reflection-suppressing boundary conditions. We use periodic boundary conditions
-in the $x$ direction and damping layers (described in \autoref{ss:damping}) in
-the $z$ direction to damp perturbations exiting the top and bottom of the
-domain. We use a Fourier basis in both the $x$ and $z$ direction. We varied the
-number of $x, z$ modes (denoted $N_x, N_z$ respectively), and used $3/2$
-dealiasing \citep{boyd}.
-
-The geometry of our simulation domain is fixed by three parameters: $L_x, L_z$,
-and $z_0$ the forcing location. We use $z_0 = 0.2L_z$ to force sufficiently far
-from the lower damping zone and permit sufficient room for the upwards-moving
-wave to grow at $\propto e^{z/2H}$. We choose $L_z = 10H$ to give $\sim e^3$
-apmlitude growth between the damping zones. Finally, we want similar grid
-spacing $\frac{L_x}{N_x} \sim \frac{L_z}{N_z}$. For computational savings, we
-fix $N_z = 4N_x$. We use $L_x = 4H = \frac{2}{5}L_z$.
+In a realistic WD, IGW are excited very deep in the WD and propagate a
+substantial number of density scale heights before reaching the breaking zone.
+This propagation is very nearly exactly linear until just a few scale heights
+below the breaking region. In the interest of computational feasibility, we
+restrct the simulation domain to the upper WD fluid, of much smaller vertical
+extent than the full physical domain. To further decrease computational cost, we
+restrict ourselves to 2D, plane-parallel IGW rather than full 3D stellar
+g-modes. While it is well-known wave breaking is a 3D process
+(\cite{klostermeyer,winters1994}), the dynamical effect of the breaking process
+is likely to be similar in 2D.
+
+Denote $L_x, L_z$ to be the dimensions of the simulation domain. As the physical
+domain extends beyond $L_x, L_z$, we must take care to suppress reflections. We
+use periodic boundary conditions in the $x$ direction and damping layers
+(described in \autoref{ss:damping}) in the $z$ direction to damp perturbations
+incident on the top and bottom of the simulation. We use a Fourier basis in both
+the $x$ and $z$ direction. We varied the number of $x, z$ modes (denoted $N_x,
+N_z$ respectively), and used $3/2$ dealiasing \citep{boyd}.
+
+The geometry of our simulation domain is fixed by one further parameter: $z_0$
+the forcing location. We use $z_0 = 0.2L_z$ to force sufficiently far from the
+lower damping zone and permit sufficient room for the upwards-moving wave to
+grow at $\propto e^{z/2H}$. We choose $L_z = 10H$ to give $\sim e^3$ apmlitude
+growth between the damping zones. Finally, we want similar grid spacing
+$\frac{L_x}{N_x} \sim \frac{L_z}{N_z}$, guided by the intuition that turbulence
+is approximately isotropic. For computational savings, we fix $L_x = 4H < L_z$,
+a compromise between using fewer $N_x$ modes and using as large of an $L_x$ as
+possible (in stars, $L_x \sim 2\pi R$ the circumference). We then fix ratio $N_z
+/ N_x = 4$.
 
@@
     \item $\omega$: We choose $\omega$ by evaluating dispersion relation
-        $\omega\p{k_{x}, k_{z}}$ for some desired $k_{z}$.
+        $\omega\p{k_{x}, k_{z}}$ for a desired $k_{z}$.
 
@@
         $H \lesssim R$, we aim to study $k_{z}H \gg 1$. However, to satisfy this
-        well is computationally expensive: $L_z \gg H$ is required to give waves
-        ample room to grow within the simulation domain, but we want $k_{z}$ to
-        be sufficiently separated from the grid spacing $\sim L_z / N_z$ that
-        nonlinear effects can be seen in later simulations (see
+        well requires high resolution simulations: $L_z \gg H$ is required to
+        give waves ample room to grow within the simulation domain, but we want
+        $k_{z}$ to be sufficiently separated from the grid spacing $\sim L_z /
+        N_z$ that nonlinear effects can be seen in later simulations (see
         \autoref{ss:nl_ns}). Thus, $N_z$ sets how well we can satisfy $k_{z}H
-        \gg 1$; we choose $k_{z}H = 1$.
-
-    \item $C$: We first choose $C$ forcing strength such that $\xi_z k_z \ll 1$
-        is satisfied everywhere throughout the domain. This constrains $C$ by
-        \autoref{eq:uz_lin}.
+        \gg 1$; we choose $k_{z}H = 1$ as a compromise.
+
+    \item $C$: In our linear simulations, we first choose $C$ forcing strength
+        such that $\xi_z k_z \ll 1$ is satisfied everywhere in the simulation
+        domain. This constrains $C$ by \autoref{eq:uz_lin}.
 
@@
 
-To imitate an infinite fluid using a finite simulation domain, we use periodic
-boundary conditions in the $x$ direction and damping layers at the top and
-bottom of the $z$ direction. This damps waves that reach the edge of the
-simulation domain without inducing nonphysical reflection. Specifically, we
-replace material derivatives in \autoref{se:nl_upsilon} with:
+We aim to damp waves that reach the edge of the simulation domain without
+inducing nonphysical reflection. To do so, we replace material derivatives
+in \autoref{se:nl_upsilon} with:
 \begin{align}
@@
 We describe the results of a simulation satisfying the parameter choices in
-\autoref{ss:numerics} such that $\xi_z k_z \ll 1$ everywhere.
+\autoref{ss:numerics} such that $\xi_z k_z \ll 1$ everywhere in the simulation
+domain.
 
 During the simulation, we expect IGW of form \autoref{eq:uz_lin} to be excited.
-The incompressibility constraint further yields a complete analytical solution
-for flow velocities $\bm{u}_{al}'(x, z, t)$. The amplitude of the observed IGW
-in the simulation field $\bm{u}$ relative to analytical solution $\bm{u}_{al}'$
-over some region $z \in [z_b, z_t]$ can be estimated with estimator
-$\hat{A}_i(t)$ (subscript $i$ denotes incident wave)
+Invoking incompressibility, we obtain a complete analytical solution for flow
+velocities $\bm{u}_{al}'(x, z, t)$. The amplitude of the observed IGW in the
+simulation field $\bm{u}$ relative to analytical solution $\bm{u}_{al}'$ over
+some region $z \in [z_b, z_t]$ can be estimated with estimator $\hat{A}_i(t)$
+(subscript $i$ denotes incident wave)
 \begin{equation}
@@
         {\int\limits_{z_b}^{z_t}\int\limits_0^{L_x}
-        \overline{\rho}(u_{al}')^2\;\mathrm{d}x\mathrm{d}z}
+        \overline{\rho}\p{u_{al}'}^2\;\mathrm{d}x\mathrm{d}z}
         \label{eq:ahat_def}
@@
 damping zone where it is dissipated (\autoref{eq:F_def}). We make may a slightly
-more accurate estimate of $F$ the flux carried in the IGW by directly
-substituting $\bm{u}_{al}'$ into \autoref{eq:F_def}. Calling this estimate
-\begin{equation}
-    F_{al}' \equiv \frac{1}{L_x}
-        \int\limits_{0}^{L_x} \overline{\rho} u_{al,x}'u_{al,z}'\;\mathrm{d}x,
+more accurate estimate of the flux carried in the IGW by directly substituting
+$\bm{u}_{al}'$ into \autoref{eq:F_def}. Calling this estimate
+\begin{equation}
+    F_{al}'(z) \equiv \ev{\overline{\rho} u_{al,x}'u_{al,z}'}_x
 \end{equation}
@@
 \begin{equation}
-    \hat{F}(z, t) \equiv \frac{
-        \frac{1}{L_x}\int\limits_0^{L_x}\rho u_xu_z\;\mathrm{d}x}
+    \hat{F}(z, t) \equiv \frac{\ev{\rho u_xu_z}_x}
         {F_{al}'}.\label{eq:sbm_def}
 \end{equation}
-For our weak forcing simulation, $\hat{F} = 1$ is expected between $z_0, z_T$,
-and indeed we observe agreement with this in \autoref{fig:lin_fluxes}.
+For our weak forcing simulation, $\hat{F}(z) = 1$ is expected between $z_0, z_T$
+after initial transients dissipate, and indeed we observe agreement with this in
+\autoref{fig:lin_fluxes}.
 \begin{figure}
@@
     \caption{Snapshots of $u_x, u_z$ in the fiducial simulation illustrating
-    distinct phases of the evolution of the flow. Note that $\overline{U}_c \approx
-    0.16$ for the parameters used.}\label{fig:snapshots}
+    distinct phases of the evolution of the flow. Note that $\overline{U}_c
+    \approx 0.16$ for the parameters used.}\label{fig:snapshots}
 \end{figure*}
@@
 $\ev{\dots}_{z \in [z_a, z_b]}$ denotes averaging over interval $[z_a, z_b]$.
-Below the critical layer, we average over an interval of length $H =
-\frac{2\pi}{k_{z}}$ a full vertical wavelength. The offset $\Delta z$ is
+Below the critical layer, we average over an interval of length
+$\frac{2\pi}{k_{z}}$ a full vertical wavelength. The offset $\Delta z$ is
 necessary to make a measurement of the incident flux unaffected by the
-turbulence within the critical layer. The height of the critical layer is
+turbulence within the critical layer itself. The height of the critical layer is
 limited by $\mathrm{Ri} \lesssim 1$, which bounds its vertical extent $\sim
-\frac{1}{k_{z}}$. Here, we found $\Delta z = \frac{3}{k_z}$ was necessary to be
-sufficiently offset from strong fluctuations near the critical layer.
+\frac{1}{k_{z}}$. We empirically found an offset of $\Delta z = \frac{3}{k_z}$
+was necessary to be sufficiently far from strong fluctuations near the critical
+layer.
 
 Above the critical layer, waves decay over a much smaller vertical extent as
-they have larger $z$ wavenumbers thanks to the shear flow, so we average over a
-$\Delta z$. Accurate measurement of these fluctuations above the critical layer
-is important to measuring the transmission feature accurately, seen in
+they have larger $z$ wavenumbers thanks to the shear flow Doppler shift.
+Accurate measurement of these fluctuations above the critical layer is important
+to measuring the transmission feature accurately, seen in
 \autoref{fig:nl_fluxes} to be weak and attenuate quickly.
 
-Finally, we may plot the observed $z_c$ from data against two simple predictors:
+Finally, we may plot the measured $z_c$ against two semi-analytic predictors:
 (i) integration of \autoref{eq:zc_anal} using the measured $\Delta F(t)$, and
-(ii) substituting the time-averaged $\ev{\Delta F}_t$ into \autoref{eq:zc_sol}.
-Since $z_c(t)$ is both stable and less well-defined at early times (when the
-critical layer is thick and transient behavior is still strong), we instead
-integrate backwards from the end of the simulation, using $z_c(t_f)$ as initial
-condition. The resulting predictors are depicted in \autoref{fig:nl_front}. The
-good agreement between the evolution of $z_c(t)$ and its estimate via $\Delta
-F(t)$ and \autoref{eq:zc_anal} are noteworthy. Further of interest is the
-discrepancy between the time-averaged predictor and the data. The general
-agreement clearly demonstrates $\Delta F < F_{al}'$ and that incomplete
-absorption is observed. Moreover, its significant underprediction of critical
-layer motion at early times contrasts with its comparatively satisfactory
-prediction at later times to indicate that $\Delta F(t)$ must vary significantly
-over time. Both of these observations indicate a careful characterization of
-reflectivity is necessary.
+(ii) substituting the time-averaged $\ev{\Delta F}_t$ (over the entire length of
+the simulation) into \autoref{eq:zc_sol}. Since $z_c(t)$ is both stable and less
+well-defined at early times (when the critical layer is thick and transient
+behavior is still strong), we instead integrate backwards from the end of the
+simulation, using $z_c(t_f)$ as initial condition. The resulting predictors are
+depicted in \autoref{fig:nl_front}. The good agreement between the evolution of
+$z_c(t)$ and its estimate via $\Delta F(t)$ and \autoref{eq:zc_anal} are
+noteworthy.
+
+Also of interest is the behavior of the time-averaged $\ev{\Delta F}_t =
+0.71F_{al}'$ predictor, which serves mostly as a fiducial in the plot. The
+general agreement clearly demonstrates $\Delta F < F_{al}'$ and that momentum
+flux absorption is incomplete. However, its underprediction of critical layer
+velocity at early times contrasts with its overprediction at later times. This
+suggests that momentum flux absorption changes significantly over time, which is
+in line with the observations in the subsequent \autoref{ss:reflectivity}.
 
@@
 a reflected wave with wave vector $\bm{k} = k_{x}\bm{x} - k_{z}\bm{z}$, and (ii)
-the amount of horizontal momentum flux $F$ reflected off the critical layer.
-These two quantities need not agree exactly if some reflected flux is in
-higher-order modes, which is observed. Both are of physical interest, however:
-he reflected wave amplitude may be of physical interest in setting up standing
-modes in a realistic star, while the reflected flux is important for accurately
+the amount of horizontal momentum flux $F$ reflected/transmitted by the critical
+layer. The reflected wave amplitude and reflected flux need not agree exactly if
+some reflected flux is in higher-order modes, which is indeed the case in our
+simulations. Both are of physical interest, however: the reflected wave
+amplitude may be of physical interest in setting up standing modes in a
+realistic star, while the flux transfer properties is important for accurately
 tracking angular momentum transfer during synchronization.
@@
 performed over $z \in [z_0 + 3\sigma, z_0 + 3\sigma + H]$. Since
-$\hat{A}_i(t), \hat{A}_r(t)$ vary somewhat over time, we perform time averaging
-over interval approximately $8\pi/\omega$, denoted by angle brackets. We can
-then define the amplitude
-reflectivity
+$\hat{A}_i(t), \hat{A}_r(t)$ vary somewhat strongly over time, we perform time
+averaging over interval approximately $8\pi/\omega$, denoted by angle brackets.
+We can then define the amplitude reflectivity
 \begin{equation}
@@
 \begin{equation}
-    F_d(t) = F_i(t) + \Delta F(t) - F_>(t).
-\end{equation}
-Then we can define flux reflectivity and transmissivity coefficients
-\begin{align}
-    \mathcal{R}_F(t) &\equiv -\frac{\ev{F_d}(t)}{\ev{F_i}(t)},&
+    F_r(t) = F_i(t) + \Delta F(t) - F_>(t).
+\end{equation}
+Note that $F_>(t)$ is just the transmitted flux across the critical layer, since
+there is no other flux above the critical layer. We then define flux
+reflectivity and transmissivity coefficients
+\begin{align}
+    \mathcal{R}_F(t) &\equiv -\frac{\ev{F_r}(t)}{\ev{F_i}(t)},&
     \mathcal{T}_F(t) &\equiv -\frac{\ev{F_>}(t)}{\ev{F_i}(t)}.
@@
 
-The measurements of $\hat{A}_i, \hat{A}_r, F_i, \Delta F, F_d, F_>$ are given in
-\autoref{fig:nl_f_amps}. The oscillations in $\hat{A}_i, F_i, A_d, F_d$ are
+The measurements of $\hat{A}_i, \hat{A}_r, F_i, \Delta F, F_r, F_>$ are given in
+\autoref{fig:nl_f_amps}. The oscillations in $\hat{A}_i, F_i, A_d, F_r$ are
 significant but the mean values seem to have temporally converged.

### 2019-08-19 writeups/Su_IGW_break.tex
stash

@@
     \newcommand*{\bm}[1]{\boldsymbol{\mathbf{#1}}}
+    \newcommand*{\uv}[1]{\hat{\boldsymbol{\mathbf{#1}}}}
     \DeclareMathOperator*{\argmin}{argmin}
@@
 Interferometric Space Antenna (LISA) \citep{lisa}. They are also thought to
-produce interesting optical transients such as underluminous
-supernovae \citep{underlum}, Ca-rich fast transients \citep{carich}, and tidal
-novae \citep{tidal_novae}. Most importantly, they have been proposed as the likely
-progenitors of type Ia supernovae (e.g. \citep{Ia0,webbink} or more
-recently \citep{Ia1,Ia2}). While presently only a few tens of compact WD binaries
-are known \citep{lsst_wd}, \emph{Gaia} (currently gathering data) is expected to
-expand the catalog to a few hundreds \citep{lsst_wd} (results based on
+produce interesting optical transients such as underluminous supernovae
+\citep{underlum}, Ca-rich fast transients \citep{carich}, and tidal novae
+\citep{tidal_novae}. Most importantly, they have been proposed as the likely
+progenitors of type Ia supernovae (e.g. \citealt{Ia0,webbink} or more recently
+\citealt{Ia1,Ia2}). While presently only a few tens of compact WD binaries are
+known \citep{lsst_wd}, \emph{Gaia} (currently gathering data) is expected to
+expand the catalog to a few hundreds \citep{lsst_wd} \citep[results based on
 \emph{Gaia}'s second data release have already begun to
-appear \citep{gaiaDD,gaiaDD2}), and the Large Synoptic Survey Telescope (LSST,
-first light scheduled for 2020) will likely detect a few thousand
-more \citep{lsst_wd}. These observations will significantly advance the
-understanding of WD binaries and their evolution. My proposed theoretical and
-computational research is well-timed to take advantage of these new advances.
+appear][]{gaiaDD,gaiaDD2}, and the Large Synoptic Survey Telescope (LSST, first
+light scheduled for 2020) will likely detect a few thousand more
+\citep{lsst_wd}. These observations will significantly advance the understanding
+of WD binaries and their evolution.
 
@@
         \md{\rho}{t} &= 0 ,\label{eq:nl_density}\\
-        \md{\bm{u}}{t} + \frac{\bm{\nabla}P}{\rho} + g\bm{z} &=
+        \md{\bm{u}}{t} + \frac{\bm{\nabla}P}{\rho} + g\uv{z} &=
             0\label{eq:nl_mom}.
@@
 material derivative, and $\bm{u}, \rho, P$ denote the velocity field, density
-and pressure respectively. We denote $-g\bm{z}$ constant gravitational
+and pressure respectively. We denote $-g\uv{z}$ constant gravitational
 acceleration. Note that at hydrostatic equilibrium $\pd{}{t} = 0$ we have
-$\bm{\nabla}\overline{P} = -\overline{\rho} g\bm{z}$ and so $\overline{P} =
+$\bm{\nabla}\overline{P} = -\overline{\rho} g\uv{z}$ and so $\overline{P} =
 \overline{\rho} gH$. A vertically-stratified shear flow
-$\overline{u}_x(z)\bm{x}$ is permitted at hydrostatic equilibrium, but we will
+$\overline{u}_x(z)\uv{x}$ is permitted at hydrostatic equilibrium, but we will
 assume no background flow, so $\bm{u} = \bm{u}'$. Physically, this assumption
@@
 
-In practice, it is convenient to introduce coordinate $\Upsilon = \ln
-\frac{\rho}{\overline{\rho}}$ (e.g.\ see \citep{lecoanet_anel}). This both
-identically enforces $\rho > 0$ and avoids numerical issues in the
-$\frac{\bm{\nabla}P}{\rho}$ term if $\rho$ is small. We also define reduced
-pressure $\varpi = \frac{P}{\rho}$. Then, we may rewrite the second two
-equations of \autoref{se:nl_orig} as
+In practice, it is convenient to introduce the coordinate $\Upsilon = \ln
+\frac{\rho}{\overline{\rho}}$ \citep[e.g.][]{lecoanet_anel}. This both
+identically enforces $\rho > 0$ and eliminates the stiff term
+$\frac{\bm{\nabla}P}{\rho}$. We also define reduced pressure $\varpi =
+\frac{P}{\rho}$. Then, we may rewrite the second two equations of
+\autoref{se:nl_orig} as
 \begin{subequations}\label{se:nl_upsilon}
@@
         \md{\bm{u}'}{t} + \bm{\nabla}\varpi
-            + \varpi\bm{\nabla}\Upsilon + g\bm{z} &= 0.
+            + \varpi\bm{\nabla}\Upsilon + g\uv{z} &= 0.
     \end{align}
 \end{subequations}
-Note that in the new coordinates, hydrostatic equilibrium corresponds to
-$\Upsilon = 0, \overline{\varpi} = gH$.
+In the new coordinates, hydrostatic equilibrium corresponds to $\Upsilon = 0,
+\overline{\varpi} = gH$.
 
@@
 \bm{\nabla}$, we may linearize \autoref{se:nl_upsilon}. The solution in this
-linear regime up to undetermined amplitude $A$ is given by
-\citep{drazin,sutherland0}:
+linear regime is given by \citep{drazin,sutherland0}:
 \begin{equation}
@@
 \end{equation}
-where $\omega$ satisfies dispersion relation
+where $A$ is the amplitude and $\omega$ satisfies dispersion relation
 \begin{equation}
@@
 \end{equation}
-We take
+Our equations are valid in the limit of large sound speed $c_s \to \infty$, in
+which the \emph{Brunt-V\"ais\"al\"a frequency}
 \begin{equation}
@@
 \end{equation}
-the \emph{Brunt-V\"ais\"al\"a frequency} to be constant in the incompressible
-limit $c_s \to \infty$ where $c_s$ is the sound speed. Other dynamical
-quantities are simply related to $u'_z$ (see e.g.\ \cite{sutherland0}).
-
-In the weak stratification limit $\abs{k_{z}H} \gg 1$, the solution exhibits the
-following characteristics:
+is constant. Other dynamical quantities are simply related to $u'_z$ \citep[see
+e.g.][]{sutherland0}.
+
+In the short-wavelength/WKB limit $\abs{k_{z}H} \gg 1$, the solution exhibits
+the following characteristics:
 \begin{itemize}
@@
         \begin{align}
-            \bm{c}_{ph} &=
-                \p{k_{x}\bm{x} + k_{z}\bm{z}}\frac{\omega}
+            \bm{c}_{p} &=
+                \p{k_{x}\uv{x} + k_{z}\uv{z}}\frac{\omega}
                 {k_{x}^2 + k_{z}^2 + \frac{1}{4H^2}},\\
-            \bm{c}_{g} &= N\frac{\p{k_{z}^2 + \frac{1}{4H^2}}\bm{x}
-                - \p{k_{x}k_{z}\bm{z}}}
+            \bm{c}_{g} &= N\frac{\p{k_{z}^2 + \frac{1}{4H^2}}\uv{x}
+                - \p{k_{x}k_{z}\uv{z}}}
                 {\p{k_{x}^2 + k_{z}^2 + \frac{1}{4H^2}}^{3/2}}.\label{eq:vg}
         \end{align}
-        We note $\bm{c}_{ph} \cdot \bm{c}_g =
+        We note $\bm{c}_{p} \cdot \bm{c}_g =
         \mathcal{O}\p{\frac{1}{\p{k_{z}H}^{2}}} \approx 0$, the usual result
-        (e.g.~\cite{drazin,sutherland1}). Note that for an upwards propagating
-        IGW $c_{g, z} > 0$, $k_z < 0, k_x > 0$.
-
-    \item The time-averaged total $x$-momentum flux $F$ carried in the $\bm{z}$
+        \citep[e.g.][]{drazin,sutherland1}. Note that for an upwards
+        propagating IGW $c_{g, z} > 0$, $k_z < 0, k_x > 0$.
+
+    \item The time-averaged total $x$-momentum flux $F$ carried in the $\uv{z}$
         direction can be computed. Since the linear solution consists of terms
@@
 
-A purely horizontal shear flow $\overline{U}(z) \bm{x}$ can be seen in
+A purely horizontal shear flow $\overline{U}(z) \uv{x}$ can be seen in
 \autoref{se:nl_orig} to have the effect of modifying time derivatives
@@
 g-modes. While it is well-known wave breaking is a 3D process
-(\cite{klostermeyer,winters1994}), the dynamical effect of the breaking process
+\citep{klostermeyer,winters1994}, the dynamical effect of the breaking process
 is likely to be similar in 2D.
@@
 with adaptive timesteps computed from advective Courant-Friedrichs-Lewy (CFL)
-time. Specifically, we use $\Delta t = 0.7 \min(\Delta x / u_x,\Delta z / u_{z})$,
-where the minimum is taken over every grid point in the domain and $\Delta
-x,\Delta z$ are the grid spacings in the $x$ and $z$ directions respectively.
+time. Specifically, we use $\Delta t = 0.7 \min(\Delta x / u_x,\Delta z /
+u_{z})$, where the minimum is taken over every grid point in the domain and
+$\Delta x,\Delta z$ are the grid spacings in the $x$ and $z$ directions
+respectively.
 
@@
 We identify two manifestations of non-absorptive behavior: (i), the presence of
-a reflected wave with wave vector $\bm{k} = k_{x}\bm{x} - k_{z}\bm{z}$, and (ii)
+a reflected wave with wave vector $\bm{k} = k_{x}\uv{x} - k_{z}\uv{z}$, and (ii)
 the amount of horizontal momentum flux $F$ reflected/transmitted by the critical
@@
 \hat{A}_r$, we remark \autoref{eq:ahat_def} ensures orthogonality between
-$k_{x}\bm{x} \pm k_{z}\bm{z}$ modes. The integral for $\hat{A}_r$ is also
+$k_{x}\uv{x} \pm k_{z}\uv{z}$ modes. The integral for $\hat{A}_r$ is also
 performed over $z \in [z_0 + 3\sigma, z_0 + 3\sigma + H]$. Since
@@
 \end{equation}
-the Alfv\'en wave crossing time (evaluated for a CO WD in \citep{fullerIV})
+the Alfv\'en wave crossing time (evaluated for a CO WD in \citealt{fullerIV})
 measures the magnetic coupling time and $t_{gw}$ measures the gravitational wave
@@
 $\bm{\nabla}\overline{\varpi} + \overline{\varpi} \bm{\nabla}\overline{\rho} +
-g\bm{z} = 0$. Taking isothermal stratification, we find $\overline{\varpi} = gH$. We
+g\uv{z} = 0$. Taking isothermal stratification, we find $\overline{\varpi} = gH$. We
 further change variables to $\Upsilon = \ln \rho - \ln \overline{\rho}$ and
@@
             - \nu \rho \bm{\nabla}\bm{u})
-            + \rho g \bm{z} &= 0.
+            + \rho g \uv{z} &= 0.
     \end{align}
@@
                 - \varpi' \pd{\Upsilon}{x}\bigg],\\
-        \pd{u_z}{t} + \pd{\varpi}{z} + gH\pd{\Upsilon}{z} - \frac{\varpi'}{H}
+        \pd{u_z}{t} + \pd{\varpi'}{z} + gH\pd{\Upsilon}{z} - \frac{\varpi'}{H}
             ={}& -\Gamma(z) u_z

### 2019-08-20 writeups/Su_IGW_break.tex
updated writeup for Daniel revision 190820

@@
 
-We consider a two-dimensional incompressible, isothermal fluid, representative
-of degenerate matter in WDs. We furthermore neglect temperature variations and
-assume a barotropic equation of state $P(\rho, T) = P(\rho)$ as a first
-approximation. As we are interested in dynamics far from the center of the WD,
-we approximate the gravitational field as uniform. We model the background
-density stratification as $\overline{\rho}(x, z) = \overline{\rho}_0 e^{-z/H}$
-for some reference density $\overline{\rho}_0$ (we generally notate background
-quantities with overbars and perturbation quantities with primes).
+We consider a incompressible, isothermal fluid, representative of degenerate
+matter in WDs. We use barotropic equation of state $P(\rho, T) = P(\rho)$ as a
+first approximation. As we are interested in dynamics far from the center of the
+WD, we approximate the gravitational field as uniform. We model the background
+density stratification as $\overline{\rho} = \overline{\rho}_0 e^{-z/H}$ for
+some reference density $\overline{\rho}_0$ (we generally notate background
+quantities with overbars and perturbation quantities with primes). Finally, we
+consider a 2D fluid for computational feasibility: while it is well-known wave
+breaking is a 3D process \citep{klostermeyer,winters1994}, the dynamical effect
+of the breaking process is likely to be similar in 2D \citep{barker_ogilvie}.
 
@@
     \begin{align}
-        \md{\Upsilon}{t} + u'_z \pd{\ln \overline{\rho}}{z} &= 0
+        \md{\Upsilon}{t} + u_z \pd{\ln \overline{\rho}}{z} &= 0
             ,\label{eq:nl_up_density} \\
-        \md{\bm{u}'}{t} + \bm{\nabla}\varpi
-            + \varpi\bm{\nabla}\Upsilon + g\uv{z} &= 0.
+        \md{\bm{u}}{t} + \bm{\nabla}\varpi + \varpi\bm{\nabla}\Upsilon
+            - \frac{\varpi}{H}\uv{z} + g\uv{z} &= 0.
     \end{align}
@@
 
-\subsection{Analytical Properties: Linear}
+\subsection{Linear Analysis}
 
@@
 \begin{equation}
-    u'_z\p{x, z, t} = Ae^{z/2H}\cos\p{\p{k_{x}x + k_{z}z - \omega t}},
+    u'_z\p{x, z, t} = Ae^{z/2H}\cos\p{k_{x}x + k_{z}z - \omega t},
         \label{eq:k0z_sign}
@@
         \end{align}
-        We note $\bm{c}_{p} \cdot \bm{c}_g =
-        \mathcal{O}\p{\frac{1}{\p{k_{z}H}^{2}}} \approx 0$, the usual result
-        \citep[e.g.][]{drazin,sutherland1}. Note that for an upwards
-        propagating IGW $c_{g, z} > 0$, $k_z < 0, k_x > 0$.
-
-    \item The time-averaged total $x$-momentum flux $F$ carried in the $\uv{z}$
-        direction can be computed. Since the linear solution consists of terms
-        separable as $f(x, z, t) = f(z)\cos\p{k_{x}x - \omega t}$, $x$
-        averaging and time averaging are equivalent. $x$ averaging is preferred
-        both since periodic boundary conditions pick a natural averaging length
-        and since it does not introduce extra temporal smoothing. We may then
-        write
+        We note $\bm{c}_{p} \cdot \bm{c}_g = \mathcal{O}\p{ \p{k_{z}H}^{-2} }
+        \approx 0$. In the Boussinesq approximation where density stratification
+        outside of $N^2$ is completely neglected, the phase and group velocity
+        are exactly orthogonal \citep{drazin,sutherland1}. We use the convention
+        where upward propagating IGW have $c_{g, z} > 0$, $k_z < 0, k_x > 0$.
+
+    \item The averaged horizontal momentum flux $F$ is
         \begin{align}
@@
         \end{align}
-        where $\ev{\dots}_x$ denotes $x$-averaging. Note that indeed for an
-        upwards propagating IGW $k_z < 0, k_x > 0$, we have $F > 0$.
+        Thus, indeed $F > 0$ for an upward propagating IGW $c_{g, z} > 0$.
 \end{itemize}
 
-\subsection{Wave Generation: Analytical Analysis}
-
-Since we are only concerned with the behavior of the wave at high $z$ where wave
-breaking occurs, we generate waves at low $z$ in the most numerically convenient
-way. As long as physically represesentative, upwards-propgaating IGW are
-generated and wave dynamics are only examined far from the forcing region, the
-exact excitation mechanism is not important for our conclusions.
+\subsection{Wave Generation}
 
@@
 towards the surface, we use a volumetric forcing term to excite IGW near the
-bottom of the simulation domain\footnote{Interfacial forcing is numerically
-unstable when using a Chebyshev polynomial basis in the $z$ direction and
-impossible when using a Fourier basis.}. Our forcing excites both IGWs
+bottom of the simulation domain. Our forcing excites both IGWs
 propagating upwards, imitating a wave tidally excited deeper in the WD, and
@@
 In the linearized system, the effect of this forcing can be solved analytically
-up to good accuracy: we first approximate the driving term using
-$e^{\frac{-(z - z_0)^2}{2\sigma^2}} \approx \sqrt{2\pi \sigma^2}\delta(z -
-z_0)$, the $\sigma \to 0$ limit\footnote{In practice, $\sigma$ must be large
-enough to be numerically resolved by the spectral code.}. This system is solved
-exactly by matching the two plane wave solutions, above and below $z_0$. We may
-then approximately relax the solution to nonzero $\sigma$: an extra factor of
-$e^{-\frac{(k_{z}\sigma)^2}{2}}$ arises compared to the $\delta$-function
-solution (evaluating the Fourier Transform of $e^{-\frac{(z -
-z_0)^2}{2\sigma^2}}$ at $k_z$), and we obtain
+to be
 \begin{align}
@@
 \end{align}
-The $z > z_0$ region models an upwards-propagating IGW wavetrain excited deep
+The $z > z_0$ region models an upward propagating IGW wavetrain excited deep
 in the atmosphere.
@@
 
-In developing the results above, we have neglected advective terms $\pd{}{t} \gg
-\bm{u}' \cdot \bm{\nabla}$. However, since $\abs{\bm{u}'} \propto
-e^{z/2H}$, any nonzero amplitude IGW will eventually violate the linearity
-criterion. This regime is relevant in practice since the IGW may grow to
-nonlinear amplitudes before it reaches the surface of the WD
-\citep{fullerI,fullerII}. We can coarsely estimate the height at which this
-happens by when the Lagrangian displacement of the perturbation $\bm{\xi}' =
-\frac{\bm{u}'}{-i\omega}$ of a fluid parcel satisfies
-\begin{equation}
-    \abs{\bm{\xi}'} \abs{\bm{k}} \sim \xi_z k_z \gtrsim 1.\label{eq:nl}
-\end{equation}
-
-This is in agreement with the breaking location predicted by the wave-induced
-mean flow, given by \citep{eliassen_palm_cite,sutherland0}:
-\begin{equation}
-     \ev{u_{x}} \equiv \overline{U} = \frac{1}{L_x}
-        \int\limits u_{x}\;\mathrm{d}x = \frac{\ev{u_{x}u_{z}}_x}{c_{g,z}},
-        \label{eq:mean_flow}
-\end{equation}
-This formula is analogous to Stokes' drift for surface waves and can be derived
-by considering the propagation of $F$ (\autoref{eq:F_def}) into a medium at
-rest. Solving for where $\overline{U} = \omega / k_x$ the critical layer
-condition is fulfilled (\autoref{eq:u_c} and discussion in
-\autoref{ss:crit_layer}), we obtain $\abs{\frac{u_{x}'k_{x}}{\omega}} =
-\abs{\xi_x k_{x}} = \abs{\xi_z k_z} \sim 1$, equal to the nonlinearity criterion
-\autoref{eq:nl}. This condition predicts where the wave's self-acceleration
-alone produces a critical layer.
-
-\subsection{Formation and Dynamics of Critical Layer}\label{ss:crit_layer}
-
-[TODO rewrite more clearly, it's just copy pasted together right now]
-
-The general understanding of nonlinear IGW breaking is that the waves'
-interactions with the mean flow of the fluid transfer horizontal momentum from
-the waves into the fluid; such a process has been conjectured to be responsible
-for tidal synchronization in stellar binaries \citep{zahn75,gn89} as well as the
-quasi-biennial oscillation \citep{lindzen_qbo}. The details of this process have
-been laid out in a few key papers: IGWs are globally unstable to resonant
-three-wave interactions \citep{drazin}, causing energy transfer out of the IGW
-to daughter modes. These steeper daughter modes facilitate wave breaking,
-depositing horizontal momentum in the fluid's mean flow \citep{klostermeyer}.
-
-In accordance with literature and the simulation setup, we will assume momentum
-redistribution is inefficient (see \autoref{ss:disp} for analysis of this
-assumption). The mean flow will then grow in place until it reaches the
-horizontal phase velocity of the parent IGW\@. Then, a critical layer forms at
-which the frequency of the incident IGW is Doppler-shifted to zero. The
-interaction of the parent IGW with the mean flow was first studied in the
+Above, we neglected advective terms $\pd{}{t} \gg \bm{u}' \cdot \bm{\nabla}$.
+However, since $\abs{\bm{u}'} \propto e^{z/2H}$ in an infinite domain all IGWs
+eventually grow to nonlinear amplitudes and break. Nonlinear wave breaking is
+expected to be important in WDs \citep{fullerI,fullerII}. We can
+estimate the height of wave breaking using $\frac{\abs{\bm{u}'}
+\abs{\bm{k}}}{\omega} \sim 1$. This can be rewritten using the Lagrangian
+displacement $-i\omega \bm{\xi}' = \bm{u}'$:
+\begin{equation}
+    \xi_z k_z \gtrsim 1.\label{eq:nl}
+\end{equation}
+
+An equivalent estimate given by \citep{eliassen_palm_cite,sutherland0} is that
+IGWs break when the wave-induced mean flow is equal to the horizontal phase
+velocity
+\begin{equation}
+    \overline{U}(z) = \frac{\ev{u_{x}u_{z}}_x}{c_{g,z}} \sim
+        \frac{\omega}{k_x},\label{eq:mean_flow}
+\end{equation}
+where $\bar{U}(z) \equiv \ev{u_x}_x$ is the mean flow of the fluid.
+\autoref{eq:mean_flow} provides a leading-order estimate for $\bar{U}(z)$ which
+may further evolve over time (see \autoref{ss:crit_layer}); for a general mean
+flow, \autoref{eq:mean_flow} also defines the criterion for a \emph{critical
+layer}. This wave-induced mean flow is analogous to Stokes' drift for surface
+waves and can be derived by considering the propagation of $F$
+(\autoref{eq:F_def}) into a medium at rest. Evaluating \autoref{eq:mean_flow} yields
+agreement with \autoref{eq:nl}.
+
+\subsection{Critical Layer Dynamics}\label{ss:crit_layer}
+
+% critical layer does form
+% previous studies of fluid interaction @ critical layer
+
+The detailed onset of wave breaking has been laid out in a few key papers
+\citep{drazin, klostermeyer, winters1994}, where instabilities transfer momentum
+out of the IGW through daughter modes into the mean flow of the fluid. When
+viscosity is sufficiently weak (see \autoref{ss:disp} for a discussion of this
+assumption in WDs), a non-uniform shear flow and eventually a critical layer
+develop. The further evolution of the critical layer with the incident IGW is
+thought to be responsible for tidal synchronization in stellar binaries
+\citep{zahn75,gn89} as well as the quasi-biennial oscillation
+\citep{lindzen_qbo}.
+
+The behavior of an incident IGW upon a critical layer was first studied in the
 inviscid, linear regime in \citep{booker_bretherton}, which found nearly
-complete absorption. This result was reproduced with nonzero viscosity
-\citep{hazel}, but weakly nonlinear theory \citep{brown_stewartson} and fully
-nonlinear simulations \citep{winters1994} suggest that nonlinear effects can
-induce reflection.
-
-A purely horizontal shear flow $\overline{U}(z) \uv{x}$ can be seen in
-\autoref{se:nl_orig} to have the effect of modifying time derivatives
-$\partial_t$ to their frequency in the comoving frame of the fluid $\partial_t -
-\overline{U}(z)\partial_x$. If anywhere $\overline{U}(z) = \overline{U}_c$ where
-\begin{equation}
-    \overline{U}_c \equiv \omega / k_x,\label{eq:u_c}
-\end{equation}
-then the horizontal wavenumber of the linear wave in the fluid frame vanishes
-and critical behavior is observed. In a linear theory, both inviscid
-\citep{booker_bretherton} and viscous \citep{hazel}, the incident wave has
-amplitude reflection and transmission coefficients
+complete absorption of the IGW\@. In particular, the incident wave has amplitude
+reflection and transmission coefficients
 \begin{align}
@@
 \end{align}
-where we have defined local Richardson number Ri at the critical layer $z_c:
-\overline{U}(z_c) = \frac{\omega}{k_{x}}$. In the $\mathrm{Ri} \gg 1$ limit,
-$\mathcal{R}, \mathcal{T} \ll 1$ and the incident wave is completely absorbed to
-good approximation.
-
-When the fluid absorbs the incident wave, it absorbs the incident horizontal
-momentum flux as well, which is converted into additional horizontal momentum of
-the shear flow. Since the mean flow cannot exceed $\overline{U}_c$ the
-horizontal phase velocity of the incident wave, the critical layer must instead
-propagate downwards to accommodate the incident momentum. The total horizontal
-momentum of the shear flow then obeys conservation equation
+where we have defined local Richardson number Ri at the height of the critical
+layer $z_c$. In the $\mathrm{Ri} \gg 1$ limit, $\mathcal{R}, \mathcal{T} \ll 1$
+and the incident wave is completely absorbed to good approximation.
+
+This result extends to viscous fluids \citep{hazel}, but weakly nonlinear theory
+\citep{brown_stewartson} and numerical simulations \citep{winters1994} suggest
+that nonlinear effects significantly enhance reflection and transmission. These
+studies are not directly applicable to our problem where a continuous train of
+IGW are excited.
+
+Instead, we mayanalyze horizontal momentum transfer at the critical layer. Any
+incident horizontal momentum flux absorbed by the fluid must manifest as
+additional horizontal momentum of the shear flow. Since the mean flow cannot
+exceed $\overline{U}_c \equiv \omega/k_x$ the horizontal phase velocity of the
+incident wave, the critical layer must instead propagate downwards to
+accommodate the incident momentum flux. The total horizontal momentum of the
+shear flow obeys conservation equation
 \begin{equation}
     \pd{}{t}\int\limits \rho(z) \overline{U}(z, t)\;\mathrm{d}z
-        + \Delta F = 0.
-\end{equation}
-$\Delta F$ is the total injected horizontal momentum to the flow. We next make a
-simple two-zone approximation $\overline{U}(z > z_c) = \overline{U}_c,
-\overline{U}(z < z_c) = 0$, assuming all horizontal momentum flux goes into
-critical layer propagation. Then, the conservation equation becomes simple
-differential equation
+        + \Delta F = 0,
+\end{equation}
+where $\Delta F$ is the absorbed horizontal momentum. Assuming $\overline{U}(z >
+z_c) = \overline{U}_c, \overline{U}(z < z_c) = 0$, the conservation equation
+becomes differential equation for $z_c(t)$
 \begin{equation}
@@
 absorbed by the critical layer, then $\Delta F = -F'$ is constant in time.
-For constant $\Delta F$ in time and $\rho \approx \overline{\rho}$, $z_c(t)$ has
-analytical solution
+Finally, for constant $\Delta F$ in time and $\rho \approx \overline{\rho}$,
+$z_c(t)$ has analytical solution
 \begin{equation}
@@
 restrict ourselves to 2D, plane-parallel IGW rather than full 3D stellar
-g-modes. While it is well-known wave breaking is a 3D process
-\citep{klostermeyer,winters1994}, the dynamical effect of the breaking process
-is likely to be similar in 2D.
+g-modes.
 
@@
 the forcing location. We use $z_0 = 0.2L_z$ to force sufficiently far from the
-lower damping zone and permit sufficient room for the upwards-moving wave to
+lower damping zone and permit sufficient room for the upward moving wave to
 grow at $\propto e^{z/2H}$. We choose $L_z = 10H$ to give $\sim e^3$ apmlitude
@@
 modified (hence $\nu$ acts only on $\rho - \overline{\rho}$).
-% Some algebraic manipulation to re-cast it in the form of \autoref{se:nl_var}
-% gives
-% \begin{subequations}
-%     \begin{align}
-%         \bm{\nabla} \cdot \bm{u} &= 0,\\
-%         \partial_t \Upsilon + \p{\bm{u} \cdot \bm{\nabla}} \Upsilon -
-%             \frac{u_z}{H} - \nu\p{\nabla^2 \Upsilon + \p{\bm{\nabla}
-%             \Upsilon} \cdot \p{\bm{\nabla}\Upsilon} - \frac{2}{H}\partial_z
-%             \Upsilon + \frac{1 - e^{-\Upsilon}}{H^2}} &= 0,\\
-%         \partial_t \bm{u} + \p{\bm{u} \cdot \bm{\nabla}}\bm{u} +
-%             \bm{\nabla} \varpi + \varpi \bm{\nabla} \Upsilon - \nu \nabla^2
-%             \bm{u} + \bm{u} \nu\p{\nabla^2 \Upsilon + \p{\bm{\nabla}
-%             \Upsilon} \cdot \p{\bm{\nabla}\Upsilon} - \frac{2}{H}\partial_z
-%             \Upsilon + \frac{1 - e^{-\Upsilon}}{H^2}}&{}\nonumber\\
-%         - 2\nu \p{\p{\p{\bm{\nabla}\Upsilon} \cdot \bm{\nabla}}\bm{u} -
-%             \frac{1}{H}\partial_z \bm{u}} - \frac{\varpi'}{H} &= 0.
-%     \end{align}
-% \end{subequations}
+
 One last consideration we found necessary was masking out nonlinear terms in the

### 2019-08-26 writeups/Su_IGW_break.tex
small edits

@@
 
-Slices of
-$\overline{U}, \hat{F}$ across $z$ at various times $t$ are shown in
-\autoref{fig:nl_fluxes}. While the behavior of $\overline{U}$ seems to conform
-qualitatively with the predictions of \autoref{ss:crit_layer}, the behavior of
-$\hat{F}$ exhibits two salient features: (i) the incident flux seems to
-fluctuate greatly with time, and (ii) there seems to be a small transmitted
-feature at many of the later times. We will discuss further these features in
-\autoref{ss:reflectivity}, after first analyzing the propagation of the critical
-layer.
+Slices of $\overline{U}, \hat{F}$ across $z$ at the times depicted in
+\autoref{fig:snapshots} are shown in \autoref{fig:nl_fluxes}. While the behavior
+of $\overline{U}$ seems to conform qualitatively with the predictions of
+\autoref{ss:crit_layer}, the behavior of $\hat{F}$ exhibits two notable
+features: (i) the incident flux seems to fluctuate greatly with time, and (ii)
+there seems to be a small transmitted feature at many of the later times. We
+will discuss further these features in \autoref{ss:reflectivity}, after first
+analyzing the propagation of the critical layer.
 \begin{figure}
@@
     \caption{Plot of $\overline{U}(z, t)$ and $\hat{F}(z, t)$ respectively at
-    various times over $z$ in our $\mathrm{Re} = 1024$ simulation.
-    $\overline{U}, \hat{F}$ follow their definitions in \autoref{eq:mean_flow}
-    and \autoref{eq:sbm_def}. $\overline{U}$ is given in units of
-    $\overline{U}_c = \frac{\omega}{k_{x}}$ the critical horizontal flow
-    velocity (see \autoref{ss:crit_layer}). The propagation of the critical
+    the same times in \autoref{fig:snapshots} over $z$ in our $\mathrm{Re} =
+    1024$ simulation. $\overline{U}, \hat{F}$ follow their definitions in
+    \autoref{eq:mean_flow} and \autoref{eq:sbm_def}. $\overline{U}$ is given in
+    units of $\overline{U}_c = \frac{\omega}{k_{x}}$ the critical horizontal
+    flow velocity (see \autoref{ss:crit_layer}). The propagation of the critical
     layer towards lower $z$ and sharp deposition of $F$ at the critical layer
@@
     \caption{Local Richardson number of the flow at the critical layer over time
-    as defined in \autoref{ss:ri_med_def}. Dotted red lines demarcate the
-    minimum and median of $\mathrm{Ri}$ over $x$. These numbers effectively
-    measure the mean and spread in width of the critical layer over $x$. Note
-    that $\mathrm{Ri} \sim \frac{1}{4}$ corresponds to the KHI, so this plot
-    suggests the shear at the critical layer does not steepen past the KHI
-    onset.}\label{fig:nl_f_ri}
+    as defined in \autoref{ss:ri_med_def}. The dotted red and solid green lines
+    denote respectively the minimum and median of $\mathrm{Ri}$ over $x$ (the
+    maximum greatly exceeds the vertical scale of this plot). These numbers
+    effectively measure the mean and spread in width of the critical layer over
+    $x$. Note that $\mathrm{Ri} \sim \frac{1}{4}$ corresponds to the KHI, so
+    this plot suggests the shear at the critical layer does not steepen past the
+    KHI onset.}\label{fig:nl_f_ri}
 \end{figure}

### 2019-10-09 writeups/Su_IGW_break.tex
stash partial work on solve inhomo eq

--- 2019-08-20 writeups/Su_IGW_break.tex
@@
     \newcommand*{\rd}[2]{\frac{\mathrm{d}#1}{\mathrm{d}#2}}
+    \newcommand*{\rtd}[2]{\frac{\mathrm{d}^2#1}{\mathrm{d}#2^2}}
     \newcommand*{\pd}[2]{\frac{\partial#1}{\partial#2}}
+    \newcommand*{\ptd}[2]{\frac{\partial^2#1}{\partial#2^2}}
     \newcommand*{\md}[2]{\frac{\mathrm{D}#1}{\mathrm{D}#2}}
@@
 
-In the linearized system, the effect of this forcing can be solved analytically
-to be
-\begin{align}
-    u_{z}'&(x, z, t) ={} \frac{Cgk_{x}^2}{\omega^2}
+In the linearized system, the effect of this forcing can be approximated
+most clearly as two plane waves propagating away from the forcing zone
+\begin{align}
+    u_{z}'&(x, z, t) \approx{} \frac{Cgk_{x}^2}{\omega^2}
         \frac{1}{2ik_{z}}\frac{e^{\frac{-(k_{z}\sigma)^2}{2}}}
@@
 The $z > z_0$ region models an upward propagating IGW wavetrain excited deep
-in the atmosphere.
+in the atmosphere. A derivation of this result is provided in
+\autoref{s:force_solved}.
 
@@
 
+\section{Forcing Solution}\label{s:force_solved}
+
+To solve for the excited wave amplitude in the linear problem, consider the
+complexified linearized system of equations, then $\pd{}{t} \to -i\omega,
+\pd{}{x} \to ik_x$. This gives (extracting the $e^{ik_xx - i\omega t}$ mode)
+\begin{align*}
+    \pd{u_{z}'}{z} + ik_xu_x' &= 0,\\
+    -i\omega u_x' + ik_x \varpi' + gHik_x \Upsilon &= 0,\\
+    -i\omega u_{z}' + \pd{\varpi'}{z} + gH\pd{\Upsilon}{z}
+        - \frac{\varpi'}{H} &= 0,\\
+    -i\omega \Upsilon - \frac{u_{z}'}{H} &=
+        Ce^{-\frac{(z - z_0)^2}{2\sigma^2}}.
+\end{align*}
+Compare to the left hand side of \autoref{se:dedalus_eqs} below. This can be
+recast solely in terms of
+$u_{1z}$ as
+\begin{align*}
+    \omega^2\p{
+        k_x^2u_{z}' + \frac{1}{H}\pd{u_{z}'}{z} - \ptd{u_{z}'}{z}
+    } - u_{z}'N^2k_x^2 =
+    gk_x^2Ce^{-\frac{(z - z_0)^2}{2\sigma^2}}.
+\end{align*}
+% In the limit $\sigma \to 0$, the right hand side can be approximated
+% $e^{-\frac{(z - z_0)^2}{2\sigma^2}} \approx \sigma\sqrt{2\pi}\delta\p{z - z_0}$.
+% Matching this against a singularity in $\ptd{u_z'}{z}$, or a kick in
+% $\pd{u_z'}{z}$, gives \autoref{eq:uz_lin}.
+The homogeneous solutions are
+
 \section{Equation Implementations}\label{se:strat_impl}
@@
 simulated in Dedalus:
-\begin{subequations}
+\begin{subequations}\label{se:dedalus_eqs}
     \begin{align}

### 2019-10-09 writeups/Su_IGW_break.tex
stash

@@
 complexified linearized system of equations, then $\pd{}{t} \to -i\omega,
-\pd{}{x} \to ik_x$. This gives (extracting the $e^{ik_xx - i\omega t}$ mode)
+\pd{}{x} \to ik_x$. Assuming all fields have dependence $u_z(x, z, t) = u_z'(z)
+e^{ik_xx - i\omega t}$, this gives:
 \begin{align*}
-    \pd{u_{z}'}{z} + ik_xu_x' &= 0,\\
+    \rd{u_{z}'}{z} + ik_xu_x' &= 0,\\
     -i\omega u_x' + ik_x \varpi' + gHik_x \Upsilon &= 0,\\
-    -i\omega u_{z}' + \pd{\varpi'}{z} + gH\pd{\Upsilon}{z}
+    -i\omega u_{z}' + \rd{\varpi'}{z} + gH\rd{\Upsilon}{z}
         - \frac{\varpi'}{H} &= 0,\\
@@
 \end{align*}
-Compare to the left hand side of \autoref{se:dedalus_eqs} below. This can be
-recast solely in terms of
-$u_{1z}$ as
+We have replaced all partial derivatives with regular derivatives. Compare to
+the left hand side of \autoref{se:dedalus_eqs} below. This can be recast solely
+in terms of $u_{1z}$ as
 \begin{align*}
     \omega^2\p{
-        k_x^2u_{z}' + \frac{1}{H}\pd{u_{z}'}{z} - \ptd{u_{z}'}{z}
+        k_x^2u_{z}' + \frac{1}{H}\rd{u_{z}'}{z} - \rtd{u_{z}'}{z}
     } - u_{z}'N^2k_x^2 =
@@
 % $\pd{u_z'}{z}$, gives \autoref{eq:uz_lin}.
-The homogeneous solutions are
+We compute the solution to the inhomogeneous ODE by the method of variation of
+parameters. The homogeneous solutions are of form $u_{z,\pm}'(z) =
+e^{\frac{z}{2H} \pm ik_zz}$ where $k_z$ satisfies the dispersion relation
+\autoref{eq:disp_rel}. The Wronskian can be computed
+\begin{equation}
+    W \equiv \det \begin{vmatrix}
+        u_{z,+}' & u_{z,-}' \\[4pt]
+        \rd{u_{z,+}'}{z} & \rd{u_{z,-}'}{z}
+    \end{vmatrix} = -2ik_ze^{z/H}.
+\end{equation}
 

### 2019-10-09 writeups/Su_IGW_break.tex
check in exact solved inhomogeneous eq

@@
     \DeclareMathOperator*{\med}{med}
+    \DeclareMathOperator*{\erf}{erf}
 
@@
 
-In the linearized system, the effect of this forcing can be approximated
-most clearly as two plane waves propagating away from the forcing zone
-\begin{align}
-    u_{z}'&(x, z, t) \approx{} \frac{Cgk_{x}^2}{\omega^2}
-        \frac{1}{2ik_{z}}\frac{e^{\frac{-(k_{z}\sigma)^2}{2}}}
-        {\sqrt{2\pi\sigma^2}} \times\nonumber\\
+In the linearized system, the effect of this forcing can be solved exactly. If
+we approximate $k_zH \gg 1, \sigma \ll H$, the solution can be approximated as
+two plane waves propagating away from the forcing zone
+\begin{align}
+    u_{z}'&(x, z, t) \approx{} \frac{C}{2k_z}\frac{gk_x^2}{\omega^2}
+        e^{-\frac{k_z^2\sigma^2}{2}}
+        \sqrt{2\pi \sigma^2} \times\nonumber\\
         &{}\begin{cases}
-        e^{\frac{z - z_0}{2H}}\cos\p{k_{x}x + k_{z}(z - z_0) - \omega t
-            + \frac{1}{2k_{z}H}}
+        e^{\frac{z - z_0}{2H}}\sin\p{k_{x}x + k_{z}(z - z_0) - \omega t
+            + \frac{k_z\sigma^2}{2H}}
             & z > z_0\\
-        e^{\frac{z - z_0}{2H}}\cos\p{k_{x}x - k_{z}(z - z_0) - \omega t
-            + \frac{1}{2k_{z}H}}
+        e^{\frac{z - z_0}{2H}}\sin\p{k_{x}x - k_{z}(z - z_0) - \omega t
+            + \frac{k_z\sigma^2}{2H}}
             & z < z_0\\
@@
 \end{align*}
-We have replaced all partial derivatives with regular derivatives. Compare to
-the left hand side of \autoref{se:dedalus_eqs} below. This can be recast solely
-in terms of $u_{1z}$ as
+We have replaced all partial derivatives with regular derivatives. This can be
+recast solely in terms of $u_{z}'$ as
 \begin{align*}
-    \omega^2\p{
-        k_x^2u_{z}' + \frac{1}{H}\rd{u_{z}'}{z} - \rtd{u_{z}'}{z}
-    } - u_{z}'N^2k_x^2 =
-    gk_x^2Ce^{-\frac{(z - z_0)^2}{2\sigma^2}}.
+     \rtd{u_{z}'}{z} - k_x^2u_{z}' - \frac{1}{H}\rd{u_{z}'}{z}
+        + u_{z}'\frac{N^2k_x^2}{\omega^2} &=
+    -\frac{gk_x^2}{\omega^2}Ce^{-\frac{(z - z_0)^2}{2\sigma^2}}
+        .\label{eq:narrow_inhomo}
 \end{align*}
-% In the limit $\sigma \to 0$, the right hand side can be approximated
-% $e^{-\frac{(z - z_0)^2}{2\sigma^2}} \approx \sigma\sqrt{2\pi}\delta\p{z - z_0}$.
-% Matching this against a singularity in $\ptd{u_z'}{z}$, or a kick in
-% $\pd{u_z'}{z}$, gives \autoref{eq:uz_lin}.
-We compute the solution to the inhomogeneous ODE by the method of variation of
-parameters. The homogeneous solutions are of form $u_{z,\pm}'(z) =
-e^{\frac{z}{2H} \pm ik_zz}$ where $k_z$ satisfies the dispersion relation
-\autoref{eq:disp_rel}. The Wronskian can be computed
+The homogeneous solutions are of form $u_{z,\pm}'(z) = e^{\p{\frac{1}{2H} \pm
+ik_z}\p{z - z_0}}$ where $k_z$ satisfies the dispersion relation
+\autoref{eq:disp_rel}. We compute the solution to the inhomogeneous ODE by the
+method of variation of parameters. The Wronskian can be computed
 \begin{equation}
@@
 \end{equation}
+The general solution is then
+\begin{equation}
+    u_z' = -u_{z,+}\int \frac{1}{W} u_{z,-}' \p{-\frac{gk_x^2}{\omega^2}
+            Ce^{-\frac{(z - z_0)^2}{2\sigma^2}}}\;\mathrm{d}z
+        + u_{z,-}\int \frac{1}{W} u_{z,+}' \p{-\frac{gk_x^2}{\omega^2}
+            Ce^{-\frac{(z - z_0)^2}{2\sigma^2}}}\;\mathrm{d}z.
+\end{equation}
+Taking these integrals and applying boundary conditions $u_z'\p{z \to \infty} =
+u_{z,+}', u_z'\p{z \to -\infty} = u_{z,-}'$ gives exact solution
+\begin{equation}
+    u(z) = \frac{C}{2ik_z}\frac{gk_x^2}{\omega^2}
+            e^{\frac{\p{\frac{\sigma^2}{2H} \pm ik_z\sigma^2}^2}{2\sigma^2}}
+                \sqrt{2\sigma^2} \frac{\sqrt{\pi}}{2}\s{
+                    -u_+\p{1 + \erf\p{\xi}}
+                    + u_- \p{\erf\p{\xi} - 1}}
+\end{equation}
+If we are concerned with only $z$ scales significantly larger than $\sigma$,
+then we may take $\erf(\xi) \approx \Theta(\xi)$. If we further assume $k_zH \gg
+1$ and restore the $e^{ik_xx - i\omega t}$ factor, we recover the cited form
+\autoref{eq:uz_lin}
+\begin{equation}
+    u_{z}'(x, z, t) = -\frac{C}{2ik_z}\frac{gk_x^2}{\omega^2}
+        e^{-\frac{k_z^2\sigma^2}{2}}
+        \sqrt{2\pi \sigma^2} e^{ik_xx - i\omega t} \times
+    \begin{cases}
+        e^{\p{\frac{1}{2H} + ik_z}\p{z - z_0} + \frac{k_z\sigma^2}{2H}}
+            & z > z_0,\\
+        e^{\p{\frac{1}{2H} - ik_z}\p{z - z_0} - \frac{k_z\sigma^2}{2H}}
+            & z > z_0.
+    \end{cases}
+\end{equation}
+We use the plane wave approximation in our fitting routines since we generally
+fit many $\sigma$ away from $z_0$, and orthogonality with other modes is
+essential for a reliable fit.
 

### 2019-10-09 writeups/notes.tex
check in exact solved inhomogeneous eq

@@
     \DeclareMathOperator{\Res}{Res}
+    \DeclareMathOperator{\erf}{erf}
     \DeclareMathOperator{\Re}{Re}
@@
         2\rho_0\p*{z = z_0}\omega^2 k_z}e^{-k_z^2\sigma^2/2}.
+\end{equation}
+
+\subsection{Narrow-Band Forcing: Exact}
+
+We use variation of parameters. We start with PDE in $\varpi, \Upsilon$
+coordinates and transform to ODE
+\begin{align}
+    \rd{u_{z}'}{z} + ik_xu_x' &= 0,\nonumber\\
+    -i\omega u_x' + ik_x \varpi' + gHik_x \Upsilon &= 0,\nonumber\\
+    -i\omega u_{z}' + \rd{\varpi'}{z} + gH\rd{\Upsilon}{z}
+        - \frac{\varpi'}{H} &= 0,\nonumber\\
+    -i\omega \Upsilon - \frac{u_{z}'}{H} &=
+        Ce^{-\frac{(z - z_0)^2}{2\sigma^2}},\nonumber\\
+     \rtd{u_{z}'}{z} - k_x^2u_{z}' - \frac{1}{H}\rd{u_{z}'}{z}
+        + u_{z}'\frac{N^2k_x^2}{\omega^2} &=
+    -\frac{gk_x^2}{\omega^2}Ce^{-\frac{(z - z_0)^2}{2\sigma^2}}
+        .\label{eq:narrow_inhomo}
+\end{align}
+We have divided through by $e^{i(k_xx - \omega t)}$ and taken $u_z'(x, z, t)
+\Rightarrow u_z'(z)$. The inhomogeneous solutions are most easily written
+\begin{equation}
+    u_{z,\pm}'(z) = e^{\p*{\frac{1}{2H} \pm ik_z}\p*{z - z_0}},
+\end{equation}
+in which case the Wronskian is
+\begin{equation}
+    W \equiv \det \begin{vmatrix}
+        u_{z,+}' & u_{z,-}' \\[4pt]
+        \rd{u_{z,+}'}{z} & \rd{u_{z,-}'}{z}
+    \end{vmatrix} = -2ik_ze^{z/H}.
+\end{equation}
+
+The general solution to \autoref{eq:narrow_inhomo} is then given $u(z) = A_+u_+
++ A_-u_-$ where
+\begin{align}
+    A_{\pm} &= \mp \int \frac{1}{W} u_{\mp} \p*{-\frac{gk_x^2}{\omega^2}
+            Ce^{-\frac{(z - z_0)^2}{2\sigma^2}}}\;\mathrm{d}z,\nonumber\\
+        &= \mp \frac{1}{2ik_z}\frac{gk_x^2}{\omega^2}\int \exp\s*{
+            \p*{-\frac{1}{2H} \mp ik_z}z' - \frac{z'^2}{2\sigma^2}}
+            \;\mathrm{d}z',\nonumber\\
+        &= \mp \frac{1}{2ik_z}\frac{gk_x^2}{\omega^2}\int \exp\s*{
+            -\frac{z'^2 + \p*{\frac{\sigma^2}{H} \pm 2ik_z\sigma^2}z'
+            }{2\sigma^2}}\;\mathrm{d}z',\nonumber\\
+        &= \mp \frac{1}{2ik_z}\frac{gk_x^2}{\omega^2}\int \exp\s*{
+            -\frac{\p*{z' + \frac{\sigma^2}{2H} \pm ik_z\sigma^2}^2
+                - \p*{\frac{\sigma^2}{2H} \pm ik_z\sigma^2}^2
+            }{2\sigma^2}}\;\mathrm{d}z',\nonumber\\
+        &= \mp \frac{1}{2ik_z}\frac{gk_x^2}{\omega^2}
+            e^{\frac{\p*{\frac{\sigma^2}{2H} \pm ik_z\sigma^2}^2}{2\sigma^2}}
+                \sqrt{2\sigma^2} \int e^{-\xi^2}\;\mathrm{d}\xi,\nonumber\\
+        &= \mp \frac{1}{2ik_z}\frac{gk_x^2}{\omega^2}
+            e^{\frac{\p*{\frac{\sigma^2}{2H} \pm ik_z\sigma^2}^2}{2\sigma^2}}
+                \sqrt{2\sigma^2} \frac{\sqrt{\pi}}{2}\p*{1 + C +
+                    \erf\p*{\frac{z - z_0}{\sqrt{2\sigma^2}}
+                        + \frac{\sigma}{2H\sqrt{2}}
+                        \pm \frac{ik_z\sigma}{\sqrt{2}}}}
+\end{align}
+We have made $z' \equiv z - z_0, \xi = \s*{ \frac{z - z_0}{\sqrt{2\sigma^2}} +
+\frac{\sigma}{2H\sqrt{2}} \pm \frac{ik_z\sigma}{\sqrt{2}}}$, and $\erf(z) =
+\frac{2}{\sqrt{\pi}}\int\limits_0^z e^{-t^2}\;\mathrm{d}t$. $C$ is a constant of
+integration. Since we want $u(z\to \infty) = u_+, u(z \to -\infty) = u_-$, we
+can quickly see that $C_+ = 0, C_- = -2$. Thus, the specific solution can be
+written down
+\begin{equation}
+    u(z) = \frac{1}{2ik_z}\frac{gk_x^2}{\omega^2}
+            e^{\frac{\p*{\frac{\sigma^2}{2H} \pm ik_z\sigma^2}^2}{2\sigma^2}}
+                \sqrt{2\sigma^2} \frac{\sqrt{\pi}}{2}\s*{
+                    -u_+\p*{1 + \erf\p*{\xi}}
+                    + u_- \p*{\erf\p*{\xi} - 1}}
+\end{equation}
+In the limit where $H \to \infty$ and $\erf(\xi) \approx \Theta(\xi)$, we
+recover
+\begin{equation}
+    u(z) = -\frac{1}{2ik_z}\frac{gk_x^2}{\omega^2}e^{-\frac{k_z^2\sigma^2}{2}}
+        \sqrt{2\pi \sigma^2} \times
+    \begin{cases}
+        e^{\p*{\frac{1}{2H} + ik_z}\p*{z - z_0} + \frac{k_z\sigma^2}{2H}}
+            & z > z_0,\\
+        e^{\p*{\frac{1}{2H} - ik_z}\p*{z - z_0} - \frac{k_z\sigma^2}{2H}}
+            & z > z_0.
+    \end{cases}
 \end{equation}

### 2019-10-09 sims/2d_4_fourier/strat.py

@@ -74,10 +74,10 @@ if __name__ == '__main__':
         #                        'N_Z': 4096})),
         # ('yubo_nu1_hres',
         #  get_params(overrides={'Re_inv': 0.1})),
-        ('yubo_nu1_vhres',
-         get_params(overrides={'Re_inv': 0.1,
-                               'N_X': 768,
-                               'N_Z': 3072})),
+        # ('yubo_nu1_vhres',
+        #  get_params(overrides={'Re_inv': 0.1,
+        #                        'N_X': 768,
+        #                        'N_Z': 3072})),
         # ('yubo_nu2_hres',
         #  get_params(overrides={'Re_inv': 0.2})),
         # ('yubo_nu2_width',
@@ -93,10 +93,10 @@ if __name__ == '__main__':
         #  get_params(overrides={'Re_inv': 0.5,
         #                        'N_X': 256,
         #                        'N_Z': 1024})),
-        # ('yubo_nu7',
-        #  get_params(overrides={'Re_inv': 0.7,
-        #                        'N_X': 256,
-        #                        'N_Z': 1024})),
+        ('yubo_nu7',
+         get_params(overrides={'Re_inv': 0.7,
+                               'N_X': 256,
+                               'N_Z': 1024})),
     ]
     if '-plot' in sys.argv:
         for name, params_dict in tasks:

### 2019-10-09 sims/2d_4_fourier/strat_helper.py

@@ -114,7 +114,7 @@ def get_anal_uz(params, t, x, z, phi=0):
         * np.sin(params['KX'] * x
                  + params['KZ'] * (z - params['Z0'])
                  - params['OMEGA'] * t
-                 + 1 / (2 * params['KZ'] * params['H'])
+                 + (params['KZ'] * params['S'])**2 / (2 * params['H'])
                  - phi))
 
 def get_anal_ux(params, t, x, z, phi=0):
@@ -128,12 +128,12 @@ def get_anal_ux(params, t, x, z, phi=0):
             * np.sin(params['KX'] * x
                      + params['KZ'] * (z - params['Z0'])
                      - params['OMEGA'] * t
-                     + 1 / (2 * params['KZ'] * params['H'])
+                     + (params['KZ'] * params['S'])**2 / (2 * params['H'])
                      - phi)
             - np.cos(params['KX'] * x
                      + params['KZ'] * (z - params['Z0'])
                      - params['OMEGA'] * t
-                     + 1 / (2 * params['KZ'] * params['H'])
+                     + (params['KZ'] * params['S'])**2 / (2 * params['H'])
                      - phi)
                 / (2 * params['H'] * params['KX'])))
 

### 2019-11-05 writeups/Su_IGW_break.tex

@@
 
+    \newcommand*{\scinot}[2]{#1\times10^{#2}}
     \newcommand*{\rd}[2]{\frac{\mathrm{d}#1}{\mathrm{d}#2}}
@@
     \newcommand*{\z}[1]{\left\{#1\right\}}
-    \newcommand*{\bm}[1]{\boldsymbol{\mathbf{#1}}}
-    \newcommand*{\uv}[1]{\hat{\boldsymbol{\mathbf{#1}}}}
+    \newcommand*{\bm}[1]{\mathbf{#1}}
+    \newcommand*{\uv}[1]{\hat{\mathbf{#1}}}
     \DeclareMathOperator*{\argmin}{argmin}
@@
 
-\title[Internal Gravity Wave Breaking in White Dwarf Binaries]{Internal Gravity Wave Breaking in White Dwarf Binaries}
+\title[Internal Gravity Wave Breaking in White Dwarf Binaries]{Internal Gravity
+Wave Breaking in White Dwarf Binaries}
 \author[Y. Su et\ al.]{
@@
         \md{\bm{u}}{t} + \bm{\nabla}\varpi + \varpi\bm{\nabla}\Upsilon
-            - \frac{\varpi}{H}\uv{z} + g\uv{z} &= 0.
+            - \frac{\varpi}{H}\uv{z} + g\uv{z} &= 0\label{eq:nl_upsilon_u}.
     \end{align}
@@
                 \frac{1}{L_x}\int_0^{L_x}\limits \rho u_{x}'u_{z}'\;\mathrm{d}x,
-                    \label{eq:F_def},\\
+                    \label{eq:F_def}\\
                 &\approx -\frac{A^2}{2}\overline{\rho}_0\frac{k_{z}}{k_{x}},
@@
 thought to be responsible for tidal synchronization in stellar binaries
-\citep{zahn75,gn89} as well as the quasi-biennial oscillation
-\citep{lindzen_qbo}.
-
-The behavior of an incident IGW upon a critical layer was first studied in the
+\citep{zahn75,gn89} as well as terrestrial phenomena such as the quasi-biennial
+oscillation \citep{lindzen_qbo}.
+
+The behavior of an IGW incident upon a critical layer was first studied in the
 inviscid, linear regime in \citep{booker_bretherton}, which found nearly
 complete absorption of the IGW\@. In particular, the incident wave has amplitude
-reflection and transmission coefficients
+reflection and transmission coefficients that can be expressed in terms of the
+local Richardson number Ri at the height of the critical layer $z_c$:
 \begin{align}
@@
     \mathcal{R}_A &= e^{-2\pi \sqrt{\mathrm{Ri} - \frac{1}{4}}}, &
-    \mathcal{T}_A &= e^{-\pi \sqrt{\mathrm{Ri} - \frac{1}{4}}},
+    \mathcal{T}_A &= e^{-\pi \sqrt{\mathrm{Ri} - \frac{1}{4}}}.
         \label{eq:crit_coeffs}
 \end{align}
-where we have defined local Richardson number Ri at the height of the critical
-layer $z_c$. In the $\mathrm{Ri} \gg 1$ limit, $\mathcal{R}, \mathcal{T} \ll 1$
-and the incident wave is completely absorbed to good approximation.
+In the $\mathrm{Ri} \gg 1$ limit, $\mathcal{R}, \mathcal{T} \ll 1$
+and the incident wave is completely absorbed to a good approximation.
 
@@
 \citep{brown_stewartson} and numerical simulations \citep{winters1994} suggest
-that nonlinear effects significantly enhance reflection and transmission. These
-studies are not directly applicable to our problem where a continuous train of
-IGW are excited.
-
-Instead, we mayanalyze horizontal momentum transfer at the critical layer. Any
-incident horizontal momentum flux absorbed by the fluid must manifest as
-additional horizontal momentum of the shear flow. Since the mean flow cannot
-exceed $\overline{U}_c \equiv \omega/k_x$ the horizontal phase velocity of the
-incident wave, the critical layer must instead propagate downwards to
-accommodate the incident momentum flux. The total horizontal momentum of the
-shear flow obeys conservation equation
-\begin{equation}
-    \pd{}{t}\int\limits \rho(z) \overline{U}(z, t)\;\mathrm{d}z
-        + \Delta F = 0,
-\end{equation}
-where $\Delta F$ is the absorbed horizontal momentum. Assuming $\overline{U}(z >
-z_c) = \overline{U}_c, \overline{U}(z < z_c) = 0$, the conservation equation
-becomes differential equation for $z_c(t)$
-\begin{equation}
-    \rho(z_c) \overline{U}_c\pd{z_c}{t} = \Delta F.\label{eq:zc_anal}
-\end{equation}
-If the entirety of the horizontal momentum flux carried by the IGW wavetrain is
-absorbed by the critical layer, then $\Delta F = -F'$ is constant in time.
-Finally, for constant $\Delta F$ in time and $\rho \approx \overline{\rho}$,
-$z_c(t)$ has analytical solution
-\begin{equation}
-    z_c(t) = -H\ln t - H\ln \frac{H\overline{\rho}_0\overline{U}_c}{-\Delta F}
-        .\label{eq:zc_sol}
-\end{equation}
-
-\section{Internal Gravity Waves: Linear Numerical Simulation}\label{s:sim}
-
-Towards numerical simulation of IGW breaking, we first verify agreement with
-linear theory at weak forcing amplitudes. We perform direct numerical simulation
-using the pseudo-spectral code Dedalus \citep{dedalus}. In \autoref{ss:numerics}
-we discuss choices of numerical parameters, in \autoref{ss:damping} we
-separately discuss use of a damping zone at the top and bottom of the domain,
-and in \autoref{ss:lin_ns} we present the results of our ``linear'' simulations,
-where we solve the full nonlinear fluid equations for weak forcing.
+that nonlinear effects may significantly enhance reflection and transmission.
+
+Consider the horizontal momentum transfer at the critical layer. Any incident
+horizontal momentum flux absorbed by the fluid must manifest as additional
+horizontal momentum of the shear flow. Since the mean flow cannot exceed
+$\overline{U}_c \equiv \omega/k_x$, the horizontal phase velocity of the
+incident wave, the critical layer will instead propagate downward in response to
+the incident momentum flux. The horizontal momentum of the shear flow satisfies
+\begin{equation}
+    \pd{}{t}\int\limits \overline{\rho}(z) \overline{U}(z, t)\;\mathrm{d}z
+        - \Delta F = 0,
+\end{equation}
+where $\Delta F > 0$ is the absorbed horizontal momentum. Assuming
+$\overline{U}(z > z_c) = \overline{U}_c, \overline{U}(z < z_c) = 0$, this
+condition becomes
+\begin{equation}
+    \overline{\rho}(z_c) \overline{U}_c\rd{z_c}{t} = \Delta F.\label{eq:zc_anal}
+\end{equation}
+If $\Delta F$ is constant in time, $z_c(t)$ has analytical solution
+\begin{equation}
+    z_c(t) = -H\ln \frac{\overline{\rho}(t = 0)
+        + t\frac{\Delta F}{\overline{U}_c H}}{\overline{\rho}_0}
+        ,\label{eq:zc_sol}
+\end{equation}
+where $\overline{\rho}(t_0)$ is the density at the critical layer height at $t =
+0$.
+
+\section{Internal Gravity Waves: Weakly Forced Numerical
+Simulations}\label{s:sim}
 
@@
 
-We nondimensionalize by taking $H = N = \overline{\rho}_0 = 1$.
-
-In a realistic WD, IGW are excited very deep in the WD and propagate a
-substantial number of density scale heights before reaching the breaking zone.
-This propagation is very nearly exactly linear until just a few scale heights
-below the breaking region. In the interest of computational feasibility, we
-restrct the simulation domain to the upper WD fluid, of much smaller vertical
-extent than the full physical domain. To further decrease computational cost, we
-restrict ourselves to 2D, plane-parallel IGW rather than full 3D stellar
-g-modes.
-
-Denote $L_x, L_z$ to be the dimensions of the simulation domain. As the physical
-domain extends beyond $L_x, L_z$, we must take care to suppress reflections. We
-use periodic boundary conditions in the $x$ direction and damping layers
-(described in \autoref{ss:damping}) in the $z$ direction to damp perturbations
-incident on the top and bottom of the simulation. We use a Fourier basis in both
-the $x$ and $z$ direction. We varied the number of $x, z$ modes (denoted $N_x,
-N_z$ respectively), and used $3/2$ dealiasing \citep{boyd}.
+We verify the analytic theory of \autoref{s:theory} by running weakly forced
+direct numerical simulations using the pseudo-spectral code Dedalus
+\citep{dedalus,dedalus2}. IGWs are excited deep within WDs and must propagate
+across many density scale heights before reaching the surface. The waves stay
+linear until they are within a few scale heights of the breaking height. We thus
+simulate only the upper part of the WD\@. Although turbulence driven by wave
+breaking is fundamentally 3D \citep{klostermeyer}, we simplify the problem by
+restricting our simulations to two dimensions. Future work will study the
+problem in 3D spherical geometry.
+
+For numerical convenience, we simulate equations \autoref{eq:nl_incomp},
+\autoref{eq:nl_upsilon_u}, and \autoref{eq:vol_drive} in a Cartesian box with
+size $L_x, L_z$. The $x$ direction represents the azimuthal direction in a star,
+and the $z$ direction represents the radial direction. We thus choose periodic
+boundary conditions in the $x$ direction. We also use periodic boundary
+conditions in the $z$ direction, but we damp perturbations to zero near the
+top/bottom of the domain to capture their distinct behavior (described in
+\autoref{ss:damping}). We expand all variables as a Fourier series with $N_x$
+and $N_z$ modes, and use the $3/2$ dealiasing rule to avoid aliasing errors in
+the nonlinear terms \citep{boyd}.
 
 The geometry of our simulation domain is fixed by one further parameter: $z_0$
-the forcing location. We use $z_0 = 0.2L_z$ to force sufficiently far from the
-lower damping zone and permit sufficient room for the upward moving wave to
-grow at $\propto e^{z/2H}$. We choose $L_z = 10H$ to give $\sim e^3$ apmlitude
-growth between the damping zones. Finally, we want similar grid spacing
-$\frac{L_x}{N_x} \sim \frac{L_z}{N_z}$, guided by the intuition that turbulence
-is approximately isotropic. For computational savings, we fix $L_x = 4H < L_z$,
-a compromise between using fewer $N_x$ modes and using as large of an $L_x$ as
-possible (in stars, $L_x \sim 2\pi R$ the circumference). We then fix ratio $N_z
-/ N_x = 4$.
+the forcing location. We choose $L_z = 10H$ to give $\sim e^3$ amplitude growth
+between the damping zones. We force at $z_0 = 2H = 0.2L_z$ within width
+$\sigma = 0.078H$, sufficiently far from the lower damping zone and permit
+sufficient room for the upward moving wave to grow as $\propto e^{z/2H}$.
+Finally, we want similar grid spacing $\frac{L_x}{N_x} \sim \frac{L_z}{N_z}$,
+guided by the intuition that turbulence is approximately isotropic, so we use
+$L_x = 4H$ and $N_z / N_x = 4$.
 
@@
 
-The physics of our simulation is fixed by four parameters: $k_{x}, \omega, C,
+We non-dimensionalize the problem such that $H = N = \rho_0 = 1$. The physics of
+the simulation is then fixed by the four remaining parameters $k_{x}, \omega, C,
 \nu$. We describe our choices for these parameters below:
 \begin{itemize}
-    \item $k_{x}$: Astrophysical IGWs in stars are generally excited by the $l
-        = 2$ component of the tidal potential, for which $k_{\perp} \sim
-        \frac{1}{R}$ where $R$ is the radius of the star. To best emulate this,
-        we use $k_{x} = \frac{2\pi}{L_x}$ the smallest permitted wavenumber
-        permitted by periodic boundary conditions.
-
-    \item $\omega$: We choose $\omega$ by evaluating dispersion relation
-        $\omega\p{k_{x}, k_{z}}$ for a desired $k_{z}$.
-
-        To choose $k_z$, we note astrophysical IGWs generally satisfy $\omega
-        \ll N$, or equivalently $\frac{k_r}{k_{\perp}} \sim k_rR \gg 1$. Since
-        $H \lesssim R$, we aim to study $k_{z}H \gg 1$. However, to satisfy this
-        well requires high resolution simulations: $L_z \gg H$ is required to
-        give waves ample room to grow within the simulation domain, but we want
-        $k_{z}$ to be sufficiently separated from the grid spacing $\sim L_z /
-        N_z$ that nonlinear effects can be seen in later simulations (see
-        \autoref{ss:nl_ns}). Thus, $N_z$ sets how well we can satisfy $k_{z}H
-        \gg 1$; we choose $k_{z}H = 1$ as a compromise.
+    \item $k_{x}$: Tidally excited waves in stars generally have $\ell = 2$,
+        corresponding to a very small $k_\perp\sim 1/R$, where $R$ is the radius
+        of the star. We use the smallest wavenumber in our simulation,
+        $k_x=2\pi/L_x$.
+
+    \item $\omega$: We choose $\omega$ by evaluating the dispersion relation
+        $\omega(k_x, k_z)$ for a desired $k_z$. We pick $\abs{k_z H} = 2\pi$ to
+        ensure the waves are very well resolved in all of our simulations. Note
+        however that tidally forced IGWs typically have $\omega \ll N$, or
+        equivalently $k_r/k_\perp \sim k_r R \gg 1$. This requires $k_z H
+        \gtrsim 1$, which is only marginally satisfied in our simulations.
 
@@
 
-    \item $\nu$: Nonlinear effects transfer wave energy from $\bm{k}$ to
-        larger wavenumbers. Since well-resolved simulations using spectral
-        methods have no inherent numerical viscosity, energy will accumulate at
-        grid scales in the absence of artificial dissipation. We introduce
-        dissipation parameter $\nu$ used for both artificial viscosity and
-        artificial diffusivity. We ensure our equations of motion continue to
-        conserve horizontal momentum (see Appendix~\ref{se:strat_impl} for
-        details). Under weak forcing, we may set $\nu = 0$, since energy
-        transfer out of the parent $\bm{k}$ mode is negligible.
+    \item $\nu$: Nonlinear effects transfer wave energy from the injection
+        wavenumber $\bm{k}$ to larger wavenumbers. Our spectral method does not
+        have any numerical viscosity, so diffusivity must be introduced into the
+        equations to regularize the systems. We add viscosity and diffusivity to
+        the system in a way that conserves horizontal momentum (see
+        Appendix~\ref{se:strat_impl} for details). When the forcing is weak, we
+        may set $\nu=0$ because there is negligible energy transfer between
+        modes.
 \end{itemize}
@@
     \md{}{t} &\to \md{}{t} + \Gamma(z)\Upsilon,\nonumber\\
-    \Gamma(z) &= \frac{1}{2\tau}\s{2,+ \tanh \frac{z - z_T}{\Delta z}
+    \Gamma(z) &= \frac{1}{2\tau}\s{2 + \tanh \frac{z - z_T}{\Delta z}
         + \tanh \frac{z_B - z}{\Delta z}},\label{eq:Gamma}
@@
 $\tau$, negligibly affects dynamics between $z_B, z_T$ and has transition width
-governed by $\Delta z$; we use $\Delta z = 0.025L_z$. This prescription, used in
-similar studies \citep{lecoanet_damp}, has the advantage of being smooth,
-important for spectral methods. Further details of our implementation of the
-fluid equations in Dedalus are described in Appendix~\ref{se:strat_impl}.
+governed by $\Delta z$. We use $\Delta z = 0.025L_z$ and $\tau = 1/15N$. This
+prescription is similar to \citep{lecoanet_damp} and has the advantage of being
+smooth, important for spectral methods. Further details of our implementation of
+the fluid equations in Dedalus are described in Appendix~\ref{se:strat_impl}.
 
@@
 
-We describe the results of a simulation satisfying the parameter choices in
-\autoref{ss:numerics} such that $\xi_z k_z \ll 1$ everywhere in the simulation
-domain.
-
-During the simulation, we expect IGW of form \autoref{eq:uz_lin} to be excited.
-Invoking incompressibility, we obtain a complete analytical solution for flow
-velocities $\bm{u}_{al}'(x, z, t)$. The amplitude of the observed IGW in the
-simulation field $\bm{u}$ relative to analytical solution $\bm{u}_{al}'$ over
-some region $z \in [z_b, z_t]$ can be estimated with estimator $\hat{A}_i(t)$
-(subscript $i$ denotes incident wave)
+We describe the results of a simulation with $C = \scinot{1.64}{-7}$ such that
+$\xi_z k_z \ll 1$ everywhere. In particular, $\xi_z k_z \approx \scinot{5}{-5}$
+just above the forcing zone.
+
+We expect the waves to follow the solution given by \autoref{eq:uz_lin}. Along
+with incompressibility (\autoref{eq:nl_incomp}), we obtain a complete analytical
+solution for flow velocities $\bm{u}_{al}(x, z, t)$. The amplitude of the
+observed IGW in the simulation field $\bm{u}$ relative to analytical solution
+$\bm{u}_{al}$ over some region $z \in [z_b, z_t]$ can be estimated with
+estimator $\hat{A}_i(t)$ (subscript $i$ denotes incident wave)
 \begin{equation}
     \hat{A}_i(t) = \frac{\int\limits_{z_b}^{z_t}\int\limits_0^{L_x}
-        \overline{\rho}\p{\bm{u} \cdot \bm{u}_{al}'}\;\mathrm{d}x\mathrm{d}z}
+        \overline{\rho}\p{\bm{u} \cdot \bm{u}_{al}}\;\mathrm{d}x\mathrm{d}z}
         {\int\limits_{z_b}^{z_t}\int\limits_0^{L_x}
-        \overline{\rho}\p{u_{al}'}^2\;\mathrm{d}x\mathrm{d}z}
+        \overline{\rho}\abs{\bm{u}_{al}}^2\;\mathrm{d}x\mathrm{d}z}
         \label{eq:ahat_def}
 \end{equation}
-If $\bm{u} = \bm{u}_{al}'$, then $\hat{A}_i(t) = 1$. The energy norm is a
+If $\bm{u} = \bm{u}_{al}$, then $\hat{A}_i(t) = 1$. The energy norm is a
 traditional normalization choice such that the overlap between $\bm{u},
-\bm{u}_{al}' \propto e^{z/2H}$ is evenly weighted throughout the integration
+\bm{u}_{al} \propto e^{z/2H}$ is evenly weighted throughout the integration
 region.
@@
 For a simulation satisfying $\xi_z k_z \ll 1$ everywhere, we expect
-$\hat{A}_i(t) = 1$ when integrated between the forcing and damping zones,
-corresponding to choices $z_b \gtrsim z_0, z_t \lesssim z_T$ ($z_0, z_T$ are
-defined in \autoref{eq:vol_drive} and \autoref{eq:Gamma} respectively). For
-consistency with the nonlinear case later, we choose $z_b = z_0 + 3\sigma, z_t =
-z_b + H$ (using $z_t = z_T - \Delta z$ just below the damping layer instead does
-not change the results). The resulting measurement of $\hat{A}_i(t)$ is shown in
+$\hat{A}_i(t) = 1$ when integrated between the forcing and damping zones, i.e.\
+$z_b \gtrsim z_0, z_t \lesssim z_T$ ($z_0, z_T$ are defined in
+\autoref{eq:vol_drive} and \autoref{eq:Gamma} respectively). For consistency
+with the nonlinear case later, we choose $z_b = z_0 + 3\sigma, z_t = z_b + H$
+(using $z_t = z_T - \Delta z$ just below the damping layer instead does not
+change the results). The resulting measurement of $\hat{A}_i(t)$ is shown in
 \autoref{fig:lin_amps}.
@@
     phase, we observe great agreement with \autoref{eq:uz_lin}, and in
-    particular $\hat{A}_i(t)$ asymptotes to a constant value, implying continuous
-    excitation of identical IGW.}\label{fig:lin_amps}
+    particular $\hat{A}_i(t)$ asymptotes to $\approx 1$, implying continuous
+    excitation of identical IGW with the expected
+    amplitude.}\label{fig:lin_amps}
 \end{figure}
@@
 The analytical theory also predicts that the horizontal momentmum flux $F(z, t)$
-should be spatially flat between the forcing zone where it is generated and the
-damping zone where it is dissipated (\autoref{eq:F_def}). We make may a slightly
-more accurate estimate of the flux carried in the IGW by directly substituting
-$\bm{u}_{al}'$ into \autoref{eq:F_def}. Calling this estimate
-\begin{equation}
-    F_{al}'(z) \equiv \ev{\overline{\rho} u_{al,x}'u_{al,z}'}_x
+should be independent of $z$ between the forcing zone where it is generated and
+the damping zone where it is dissipated (\autoref{eq:F_def}). We make may a
+slightly more accurate estimate of the flux carried in the IGW by directly
+substituting $\bm{u}_{al}$ into \autoref{eq:F_def}. Calling this estimate
+\begin{equation}
+    F_{al}'(z) \equiv \ev{\overline{\rho} u_{al,x}u_{al,z}}_x
 \end{equation}
@@
 \end{equation}
-For our weak forcing simulation, $\hat{F}(z) = 1$ is expected between $z_0, z_T$
-after initial transients dissipate, and indeed we observe agreement with this in
-\autoref{fig:lin_fluxes}.
+\autoref{fig:lin_fluxes} shows that $\hat{F}(z)\approx 1$ between $z_0$ and
+$z_T$, as expected.
 \begin{figure}
@@
 To perform simulations of wave breaking phenomena, we use the same values as
-\autoref{ss:numerics} except for $C, \nu$. In particular, we choose $C$ such
-that $\at{\xi_z k_z}_{z_0} = 0.1$ in the forcing zone, which implies $\xi_z k_z
-$ exceeds $1$ before $z_T$ the upper damping zone. The dissipation parameter
-$\nu$ was varied across the various simulations. To quantify $\nu$, we define
-dimensionless Reynolds number
+\autoref{ss:numerics} except for $C$ and $\nu$. In particular, we choose $C$
+such that $\at{\xi_z k_z}_{z_0} = 0.1$ in the forcing zone. The linear solution
+predicts $\xi_z k_z \sim 4.25$ at the upper damping zone $z_T$. The dissipation
+parameter $\nu$ was varied across the various simulations. To quantify $\nu$, we
+define dimensionless Reynolds number
 \begin{equation}
@@
 \end{equation}
-A table of our simulation viscosities can be found in \autoref{tab:params}.
+A table of our simulations can be found in \autoref{tab:params}.
 \begin{table}
@@
     \end{tabular}
-    \caption{Table of simulation resolutions. Note that $\mathrm{Re}$ is defined
-    as in \autoref{eq:re_def}.}\label{tab:params}
+    \caption{Table of simulation resolutions.}\label{tab:params}
 \end{table}
@@
 
-A full video of our higher-resolution run at $N_x = 768, N_z = 3072,
-\mathrm{Re} = 1024$ is available
+A full video of our higher-resolution run at $N_x = 768$, $N_z = 3072$,
+$\mathrm{Re} = 1024$ is available
 online\footnote{http://www.princeton.edu/~lecoanet/data/breaking\_wave.mov}. We
 take this to be our fiducial simulation for the remainder of this paper, though
-other simulations show qualitatively identical behavior. Snapshots of the
-velocity field $\bm{u}$ at select times can be found in \autoref{fig:snapshots}.
+other simulations show qualitatively similar behavior.
+
 % convert yubo_000054.png -crop 2000x2000+150+250 out.png
+In \autoref{fig:snapshots}, we present snapshots of the velocity field $\bm{u}$
+at various interesting phases of the simulation.
+\begin{enumerate}
+    \item At early times (top left panel), the flow resembles a linear IGW lower
+        in the simulation domain but breaks down into smaller-scale features at
+        higher $z$. Some characteristic swirling motion can be seen in both
+        $u_x, u_z$, highly suggestive of Kelvin-Helmholtz instabilities.
+
+    \item At a slightly later time (top right panel), the mean flow in
+        $u_x$ has become much more prominent and the critical layer $z_c$ has
+        become much more definite. Small-scale fluctuations are still present in
+        $u_z$ but at smaller amplitudes due to being in a denser region of the
+        fluid.
+
+    \item In the bottom left panel, the critical layer transition is now
+        extremely sharp, and small swirls of limited vertical extent at the
+        location of the critical layer in $u_z$ suggest that the
+        Kelvin-Helmholtz instability is responsible for regulating the minimum
+        width of this transition. This is supported by the conclusions in
+        \autoref{ss:khi}.
+
+    \item  The end of the simulation shows very few significant qualitative
+        differences from the previous snapshot, suggesting that the latter half
+        of our simulation is temporally converged.
+\end{enumerate}
+
 \begin{figure*}
-    \subfloat[$t=413.4$ At early times, the flow resembles a linear IGW lower in
-    the simulation domain but breaks down into smaller-scale features at higher
-    $z$. Some characteristic swirling motion can be seen in both $u_x, u_z$,
-    highly suggestive of Kelvin-Helmholtz instabilities.]{
-    \includegraphics[width=0.45\textwidth]{plots/yubo_000054.png}}\hfil
+    \includegraphics[width=0.45\textwidth]{plots/yubo_000054.png}\hfil
     %
-    \subfloat[$t=658.5$ At a slightly later time, the mean flow in $u_x$ has
-    become much more prominent and the critical layer $z_c$ has become much more
-    definite. Small-scale fluctuations are still present in $u_z$ albeit at
-    smaller amplitudes due to being in a denser region of the fluid.]{
-    \includegraphics[width=0.45\textwidth]{plots/yubo_000086.png}}
-
-    \subfloat[$t=1171.4$ The critical layer transition is now extremely sharp,
-    and small swirls of limited vertical extent at the location of the critical
-    layer in $u_z$ suggest that the Kelvin-Helmholtz instability is responsible
-    for regulating the minimum width of this transition, a hypothesis explored
-    in \autoref{ss:khi}.]{
-    \includegraphics[width=0.45\textwidth]{plots/yubo_000153.png}}\hfil
+    \includegraphics[width=0.45\textwidth]{plots/yubo_000086.png}
+
+    \includegraphics[width=0.45\textwidth]{plots/yubo_000153.png}\hfil
     %
-    \subfloat[$t=3437.8$ The end of the simulation shows very few significant
-    qualitative differences from the previous snapshot, suggesting that the
-    latter half of our simulation is temporally converged.]{
-    \includegraphics[width=0.45\textwidth]{plots/yubo_000451.png}}
-
+    \includegraphics[width=0.45\textwidth]{plots/yubo_000451.png}
     \caption{Snapshots of $u_x, u_z$ in the fiducial simulation illustrating
     distinct phases of the evolution of the flow. Note that $\overline{U}_c
-    \approx 0.16$ for the parameters used.}\label{fig:snapshots}
+    \approx 0.16$ for the parameters used. The four times depicted are $t =
+    413.4$ (top left), $t = 658.5$ (top right), $t = 1171.4$ (bottom left), and
+    $t = 3437.8$ (bottom right). The four snapshots illustrate the initial
+    transient wave breaking phase, formation of a distinct critical layer,
+    steepening of the critical layer, and propagation of the critical layer
+    respectively.}\label{fig:snapshots}
 \end{figure*}
 
-Slices of $\overline{U}, \hat{F}$ across $z$ at the times depicted in
-\autoref{fig:snapshots} are shown in \autoref{fig:nl_fluxes}. While the behavior
-of $\overline{U}$ seems to conform qualitatively with the predictions of
+In \autoref{fig:nl_fluxes}, we plot $\overline{U}$ and $\hat{F}$ as a function
+of $z$ at the times depicted in \autoref{fig:snapshots}. While the behavior of
+$\overline{U}$ seems to conform qualitatively with the predictions of
 \autoref{ss:crit_layer}, the behavior of $\hat{F}$ exhibits two notable

### 2019-11-05 writeups/Su_IGW_break.tex
stash work

@@
 The geometry of our simulation domain is fixed by one further parameter: $z_0$
-the forcing location. We choose $L_z = 10H$ to give $\sim e^3$ amplitude growth
-between the damping zones. We force at $z_0 = 2H = 0.2L_z$ within width
-$\sigma = 0.078H$, sufficiently far from the lower damping zone and permit
-sufficient room for the upward moving wave to grow as $\propto e^{z/2H}$.
-Finally, we want similar grid spacing $\frac{L_x}{N_x} \sim \frac{L_z}{N_z}$,
-guided by the intuition that turbulence is approximately isotropic, so we use
-$L_x = 4H$ and $N_z / N_x = 4$.
+the forcing location. We choose $L_z = 12.5H$ to give $\sim e^3$ amplitude growth
+between the damping zones. We force at $z_0 = 3H$ within width $\sigma =
+0.078H$, sufficiently far from the lower damping zone and permit sufficient room
+for the upward moving wave to grow as $\propto e^{z/2H}$. Finally, we want
+similar grid spacing $\frac{L_x}{N_x} \sim \frac{L_z}{N_z}$, guided by the
+intuition that turbulence is approximately isotropic, so we use $L_x = 4H$ and
+$N_z / N_x = 4$.
 
@@
 \begin{equation}
-    F_{al}'(z) \equiv \ev{\overline{\rho} u_{al,x}u_{al,z}}_x
+    F_{al}(z) \equiv \ev{\overline{\rho} u_{al,x}u_{al,z}}_x
 \end{equation}
@@
 \begin{equation}
-    \hat{F}(z, t) \equiv \frac{\ev{\rho u_xu_z}_x}
-        {F_{al}'}.\label{eq:sbm_def}
-\end{equation}
-\autoref{fig:lin_fluxes} shows that $\hat{F}(z)\approx 1$ between $z_0$ and
+    F(z, t) = \frac{\ev{\rho u_xu_z}_x}.\label{eq:sbm_def}
+\end{equation}
+Where nonlinear effects are unimportant, we expect $F(z, t) = F_{al}(z)$.
+\autoref{fig:lin_fluxes} shows that $F \approx F_{al}$ between $z_0$ and
 $z_T$, as expected.
@@
     \includegraphics[width=\columnwidth]{plots/lin_fluxes.png}
-    \caption{\autoref{eq:sbm_def} as a function of $z$ at select times $t$. As
-    initial transients die out, $\hat{F}(z, t) = 1$ to very good agreement above
+    \caption{$F/F_{al}$ as a function of $z$ at select times $t$. As
+    initial transients die out, $F / F_{al} = 1$ to very good agreement above
     the forcing zone $z > z_0 = 2H$ and below the damping zone $z \lesssim z_T =
@@
 
-In \autoref{fig:nl_fluxes}, we plot $\overline{U}$ and $\hat{F}$ as a function
+In \autoref{fig:nl_fluxes}, we plot $\overline{U}$ and $F / F_{al}$ as a function
 of $z$ at the times depicted in \autoref{fig:snapshots}. While the behavior of
 $\overline{U}$ seems to conform qualitatively with the predictions of
-\autoref{ss:crit_layer}, the behavior of $\hat{F}$ exhibits two notable
+\autoref{ss:crit_layer}, the behavior of $F / F_{al}$ exhibits two notable
 features: (i) the incident flux seems to fluctuate greatly with time, and (ii)
@@
     \includegraphics[width=\columnwidth]{plots/nl_fluxes.png}
-    \caption{Plot of $\overline{U}(z, t)$ and $\hat{F}(z, t)$ respectively at
-    the same times in \autoref{fig:snapshots} over $z$ in our $\mathrm{Re} =
-    1024$ simulation. $\overline{U}, \hat{F}$ follow their definitions in
+    \caption{Plot of $\overline{U}(z, t)$ and $F(z, t) / F_{al}(z)$ respectively
+    at the same times in \autoref{fig:snapshots} over $z$ in our $\mathrm{Re} =
+    1024$ simulation. $\overline{U}, F$ follow their definitions in
     \autoref{eq:mean_flow} and \autoref{eq:sbm_def}. $\overline{U}$ is given in
@@
 \begin{align}
-    z_{c, \min} &= \argmin_z \z{z: F(z) < 0.3F_{al}'},\nonumber\\
-    z_{c, \max} &= \argmax_z \z{z: F(z) > 0.3F_{al}'},\nonumber\\
+    z_{c, \min} &= \argmin_z \z{z: F(z) < 0.3F_{al}},\nonumber\\
+    z_{c, \max} &= \argmax_z \z{z: F(z) > 0.3F_{al}},\nonumber\\
     z_c &\equiv \frac{z_{c, \min} + z_{c, \max}}{2}.\label{eq:zc_def}
@@
 Also of interest is the behavior of the time-averaged $\ev{\Delta F}_t =
-0.71F_{al}'$ predictor, which serves mostly as a fiducial in the plot. The
-general agreement clearly demonstrates $\Delta F < F_{al}'$ and that momentum
+0.71F_{al}$ predictor, which serves mostly as a fiducial in the plot. The
+general agreement clearly demonstrates $\Delta F < F_{al}$ and that momentum
 flux absorption is incomplete. However, its underprediction of critical layer
@@
     The the poorer but qualitatively correct agreement of the time-averaged
-    predictor with the data shows both that $\Delta F(t) \neq F_{al}'$ and that
+    predictor with the data shows both that $\Delta F(t) \neq F_{al}$ and that
     $\Delta F(t)$ likely has some very real variation with

### 2019-11-05 sims/2d_3_final/strat_helper.py

@@ -27,15 +27,14 @@ PLT_COLORS = ['b', 'g', 'r', 'c', 'm', 'y', 'k', 'w'];
 
 SNAPSHOTS_DIR = 'snapshots_%s'
 FILENAME_EXPR = '{s}/{s}_s{idx}.h5'
-Z_TOP_MULT = 1
 STRIDE = 15
 AVG_IDX = 4
 DPI=600
-N_X_FORCED = 64
+N_X_FORCED = 128
 N_Z_FORCED = 1024
 
 plt.rc('text', usetex=True)
-FONTSIZE = 14
+FONTSIZE = 16
 LW = 4
 plt.rc('font', family='serif', size=FONTSIZE)
 
@@ -165,6 +164,7 @@ def subtract_lins(params, state_vars, sim_times, domain):
     grid_ones = np.ones_like(x + z)
     t_t = np.array([[[t]] for t in sim_times])
 
+    Z_TOP_MULT = 1
     z_bot = get_idx(Z0 + 3 * S, z[0])
     z_top = get_idx(Z0 + 3 * S + 2 * Z_TOP_MULT * np.pi / abs(KZ), z[0])
     pos_slice = np.s_[:, z_bot:z_top]
@@ -994,24 +994,23 @@ def plot_front(name, params):
         #
         # horizontal plot showing Fpx at certain times
         #####################################################################
-        z0_cut = z0[z_b_idx: ]
+        z0_idx = get_idx(Z0, z0)
+        z0_cut = z0[z0_idx: ]
         for time in times:
-            # plt.plot(z0_cut,
-            #          S_px[time, z_b_idx: ] / flux_th,
-            plt.plot(z0,
-                     S_px[time, : ] / flux_th,
+            plt.plot(z0_cut,
+                     S_px[time, z0_idx: ] / flux_th,
                      linewidth=LW * 0.7,
                      label=r'$t=%.1f$' % sim_times[time])
         # plt.plot(z0_cut,
         #          np.exp(-k_damp * 2 * (z0_cut - Z0)),
         #          linewidth=LW * 1.5,
         #          label=r'Model')
-        plt.xlim(0, ZMAX)
+        plt.xlim(Z0, ZMAX)
         plt.ylim(-0.2, 1.1)
         plt.legend(fontsize=FONTSIZE)
 
-        plt.xlabel(r'$z(H)$')
-        plt.ylabel(r'$\hat{F}$')
+        plt.xlabel(r'$z / H$')
[…]

### 2019-11-05 sims/2d_4_fourier/strat.py

@@ -66,6 +66,13 @@ def get_params(overrides=None):
 
 if __name__ == '__main__':
     tasks = [
+       ('lin_0_masked',
+        get_params(overrides={'F_MULT': 0.0005,
+                              'T_F': 4000,
+                              'DT': 0.1,
+                              'N_X': 64,
+                              'N_Z': 256,
+                              'Re_inv': 1e-4})),
         # ('yubo_nu0p5_hres',
         #  get_params(overrides={'Re_inv': 0.05})),
         # ('yubo_nu0p5_shres',
@@ -93,10 +100,10 @@ if __name__ == '__main__':
         #  get_params(overrides={'Re_inv': 0.5,
         #                        'N_X': 256,
         #                        'N_Z': 1024})),
-        ('yubo_nu7',
-         get_params(overrides={'Re_inv': 0.7,
-                               'N_X': 256,
-                               'N_Z': 1024})),
+        # ('yubo_nu7',
+        #  get_params(overrides={'Re_inv': 0.7,
+        #                        'N_X': 256,
+        #                        'N_Z': 1024})),
     ]
     if '-plot' in sys.argv:
         for name, params_dict in tasks:

### 2019-11-05 sims/2d_4_fourier/strat_helper.py

@@ -30,11 +30,11 @@ FILENAME_EXPR = '{s}/{s}_s{idx}.h5'
 Z_TOP_MULT = 1
 STRIDE = 15
 AVG_IDX = 4
-FONTSIZE = 14
+FONTSIZE = 16
 DPI=600
 
 plt.rc('text', usetex=True)
-LW = 4
+LW = 2.5
 plt.rc('font', family='serif', size=FONTSIZE)
 
 def populate_globals(var_dict):
@@ -682,29 +682,29 @@ def plot(name, params, stride=STRIDE):
                 p = axes.plot(var_dat[t_idx],
                               z0,
                               'r-',
-                              linewidth=0.7,
+                              linewidth=LW * 0.7,
                               label='Data')
                 if var == 'uz%s' % slice_suffix:
                     p = axes.plot(
                         uz_anal[0, :],
                         z0,
                         'orange',
-                        linewidth=0.5)
+                        linewidth=LW * 0.5)
                     p = axes.plot(
                         -uz_est * np.exp((z[0] - Z0) / (2 * H)), z[0], 'g',
                         uz_est * np.exp((z[0] - Z0) / (2 * H)), z[0], 'g',
-                        linewidth=0.5)
+                        linewidth=LW * 0.5)
 
                 if var == 'ux%s' % slice_suffix:
                     p = axes.plot(
                         ux_anal[0, :],
                         z0,
                         'orange',
-                        linewidth=0.5)
+                        linewidth=LW * 0.5)
                     p = axes.plot(
                         -ux_est * np.exp((z0 - Z0) / (2 * H)), z[0], 'g',
                         ux_est * np.exp((z0 - Z0) / (2 * H)), z[0], 'g',
-                        linewidth=0.5)
+                        linewidth=LW * 0.5)
 
                 if var == 'S%s' % mean_suffix:
                     k_damp = get_k_damp(params)
@@ -716,15 +716,15 @@ def plot(name, params, stride=STRIDE):
                         z[0],
                         'orange',
                         label=r'$x_0z_0$ (Anal.)',
-                        linewidth=0.5)
+                        linewidth=LW * 0.5)
                     rho0 = RHO0 * np.exp(-z / H)
                     # compute all of the S_px cross terms (00 is model)
                     Spx01 = horiz_mean(rho0 * state_vars['ux'][t_idx] *
                                        (state_vars['uz'][t_idx] - uz_anal),
[…]

### 2019-11-05 writeups/Su_IGW_break.tex
stash, bunch of new plots again

@@
 
-% cp ../sims/agg.png plots;
-% cp ../sims/2d_3_final/snapshots_lin_0_masked/f_amps.png plots/lin_amps.png;
-% cp ../sims/2d_3_final/snapshots_lin_0_masked/fluxes.png plots/lin_fluxes.png;
-% cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/f_amps.png plots/nl_f_amps.png;
-% cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/fluxes.png plots/nl_fluxes.png;
-% cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/f_refl.png plots/nl_f_refl.png;
-% cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/f_ri.png plots/nl_f_ri.png;
-% cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/front.png plots/nl_front.png;
+% cp ../sims/agg.png plots; cp ../sims/2d_4_fourier/snapshots_lin_0_masked/f_amps.png plots/lin_amps.png; cp ../sims/2d_4_fourier/snapshots_lin_0_masked/fluxes.png plots/lin_fluxes.png; cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/f_amps.png plots/nl_f_amps.png; cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/fluxes.png plots/nl_fluxes.png; cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/f_refl.png plots/nl_f_refl.png; cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/f_ri.png plots/nl_f_ri.png; cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/front.png plots/nl_front.png;
 
@@
 
-An equivalent estimate given by \citep{eliassen_palm_cite,sutherland0} is that
-IGWs break when the wave-induced mean flow is equal to the horizontal phase
-velocity
-\begin{equation}
-    \overline{U}(z) = \frac{\ev{u_{x}u_{z}}_x}{c_{g,z}} \sim
-        \frac{\omega}{k_x},\label{eq:mean_flow}
-\end{equation}
-where $\bar{U}(z) \equiv \ev{u_x}_x$ is the mean flow of the fluid.
-\autoref{eq:mean_flow} provides a leading-order estimate for $\bar{U}(z)$ which
-may further evolve over time (see \autoref{ss:crit_layer}); for a general mean
-flow, \autoref{eq:mean_flow} also defines the criterion for a \emph{critical
-layer}. This wave-induced mean flow is analogous to Stokes' drift for surface
-waves and can be derived by considering the propagation of $F$
-(\autoref{eq:F_def}) into a medium at rest. Evaluating \autoref{eq:mean_flow} yields
-agreement with \autoref{eq:nl}.
-
-\subsection{Critical Layer Dynamics}\label{ss:crit_layer}
-
-% critical layer does form
-% previous studies of fluid interaction @ critical layer
-
-The detailed onset of wave breaking has been laid out in a few key papers
-\citep{drazin, klostermeyer, winters1994}, where instabilities transfer momentum
-out of the IGW through daughter modes into the mean flow of the fluid. When
-viscosity is sufficiently weak (see \autoref{ss:disp} for a discussion of this
-assumption in WDs), a non-uniform shear flow and eventually a critical layer
-develop. The further evolution of the critical layer with the incident IGW is
+% An equivalent estimate given by \citep{eliassen_palm_cite,sutherland0} is that
+% IGWs break when the wave-induced mean flow is equal to the horizontal phase
+% velocity
+% \begin{equation}
+%     \overline{U}(z) = \frac{\ev{u_{x}u_{z}}_x}{c_{g,z}} \sim
+%         \frac{\omega}{k_x},\label{eq:rip}
+% \end{equation}
+% where \overline{U}(z) \equiv \ev{u_x}_x$ is the mean flow of the fluid.
+% \autoref{eq:rip} provides a leading-order estimate for \overline{U}(z)$ which
+% may further evolve over time (see \autoref{ss:crit_layer}); for a general mean
+% flow, \autoref{eq:rip} also defines the criterion for a \emph{critical
+% layer}. This wave-induced mean flow is analogous to Stokes' drift for surface
+% waves and can be derived by considering the propagation of $F$
+% (\autoref{eq:F_def}) into a medium at rest. Evaluating \autoref{eq:rip}
+% yields agreement with \autoref{eq:nl}.
+
+\citealt{drazin, klostermeyer, winters1994} describe the onset of wave breaking.
+At intermediate amplitudes, wave breaking occurs via triadic resonances which
+transfer energy from the ``parent’’ internal gravity wave (in our problem, the
+tidally forced wave) to ``daughter’’ internal gravity waves on smaller length
+scales that efficiently damp. The horizontal momentum flux decreases from $F$ to
+$0$ over this breaking region. The lost flux is deposited into mean flow
+\begin{equation}
+    \overline{U}(z, t) \equiv \ev*{u_x}_x.\label{eq:mean_flow}
+\end{equation}
+As the mean flow grows, a \emph{critical layer} may form, discussed below.
+
+\subsection{Critical Layers}\label{ss:crit_layer}
+
+% TODO stash here
+\begin{equation}
+    \overline{U}_c \equiv \omega/k_x,
+\end{equation}
+The further evolution of the critical layer with the incident IGW is
 thought to be responsible for tidal synchronization in stellar binaries
@@
 horizontal momentum flux absorbed by the fluid must manifest as additional
-horizontal momentum of the shear flow. Since the mean flow cannot exceed
-$\overline{U}_c \equiv \omega/k_x$, the horizontal phase velocity of the
-incident wave, the critical layer will instead propagate downward in response to
-the incident momentum flux. The horizontal momentum of the shear flow satisfies
+horizontal momentum of the shear flow. Since $\overline{U}$ cannot exceed
+$\overline{U}_c$, the critical layer will instead propagate downward in response
+to the incident momentum flux. The horizontal momentum of the shear flow
+satisfies
 \begin{equation}
@@
 \begin{equation}
-    z_c(t) = -H\ln \frac{\overline{\rho}(t = 0)
-        + t\frac{\Delta F}{\overline{U}_c H}}{\overline{\rho}_0}
-        ,\label{eq:zc_sol}
-\end{equation}
-where $\overline{\rho}(t_0)$ is the density at the critical layer height at $t =
-0$.
-
-\section{Internal Gravity Waves: Weakly Forced Numerical
-Simulations}\label{s:sim}
-
-\subsection{Numerical Setup}\label{ss:numerics}
+    z_c(t) = -H\ln \s{e^{-\frac{z_c(t = 0)}{H}} +
+        \frac{t\Delta F}{\overline{U}_c H\overline{\rho}_0}},\label{eq:zc_sol}
+\end{equation}
+where $z_c(t = 0)$ is the initial critical layer height.
+
+\section{Numerical Simulation Setup}\label{ss:numerics}
 
@@
 problem in 3D spherical geometry.
+
+\subsection{Parameter Choices}
 
@@
 
-    \item $C$: In our linear simulations, we first choose $C$ forcing strength
-        such that $\xi_z k_z \ll 1$ is satisfied everywhere in the simulation
-        domain. This constrains $C$ by \autoref{eq:uz_lin}.
+    \item $C$: In our weak forcing simulations, we first choose $C$ forcing
+        strength such that $\xi_z k_z \ll 1$ is satisfied everywhere in the
+        simulation domain. This constrains $C$ by \autoref{eq:uz_lin}.
 
@@
 \end{align}
-where $z_B = 0.05L_z, z_T = 0.95L_z$ are the boundaries of the damping zone.
+where $z_B = 1.5H, z_T = 10.5H$ are the boundaries of the damping zone.
 This strongly damps perturbations below $z_B$ and above $z_T$ with damping time
 $\tau$, negligibly affects dynamics between $z_B, z_T$ and has transition width
-governed by $\Delta z$. We use $\Delta z = 0.025L_z$ and $\tau = 1/15N$. This
+governed by $\Delta z$. We use $\Delta z = 0.25H$ and $\tau = 1/15N$. This
 prescription is similar to \citep{lecoanet_damp} and has the advantage of being
@@
 
-\subsection{Simulation Results}\label{ss:lin_ns}
+\section{Weakly Forced Numerical Simulation}\label{s:weak_sim}
 
@@
 we may then measure the agreement of our simulation with analytical expectation
-by computing
-\begin{equation}
-    F(z, t) = \frac{\ev{\rho u_xu_z}_x}.\label{eq:sbm_def}
-\end{equation}
-Where nonlinear effects are unimportant, we expect $F(z, t) = F_{al}(z)$.
-\autoref{fig:lin_fluxes} shows that $F \approx F_{al}$ between $z_0$ and
-$z_T$, as expected.
+by computing $F$ from the simulation via \autoref{eq:F_def}. Where nonlinear
+effects are unimportant, we expect $F(z, t) = F_{al}(z)$.
+\autoref{fig:lin_fluxes} shows that $F \approx F_{al}$ between $z_0$ and $z_T$,
+as expected.
 \begin{figure}
@@
 
-\section{Internal Gravity Waves: Nonlinear Simulation}\label{s:sim}
+\section{Wave Breaking Simulations}\label{s:sim}
 
@@
     1024$ simulation. $\overline{U}, F$ follow their definitions in
-    \autoref{eq:mean_flow} and \autoref{eq:sbm_def}. $\overline{U}$ is given in
+    \autoref{eq:mean_flow} and \autoref{eq:F_def}. $\overline{U}$ is given in
     units of $\overline{U}_c = \frac{\omega}{k_{x}}$ the critical horizontal
@@
 For the simulation shown in \autoref{fig:nl_fluxes}, we may analyze the location
-of the critical layer. As the simulations are very noisy, we measure the location
-of the critical layer using an average of where flux deposition occurs
-\begin{align}
-    z_{c, \min} &= \argmin_z \z{z: F(z) < 0.3F_{al}},\nonumber\\
-    z_{c, \max} &= \argmax_z \z{z: F(z) > 0.3F_{al}},\nonumber\\
+of the critical layer. As $F$ fluctuates significantly in the simulation, we
+measure the location of the critical layer using an average of where flux
+deposition occurs
+\begin{align}
+    z_{c, \min} &= \argmin_z \z{z: F(z) > 0.3F_{al}},\nonumber\\
+    z_{c, \max} &= \argmax_z \z{z: F(z) < 0.3F_{al}},\nonumber\\
     z_c &\equiv \frac{z_{c, \min} + z_{c, \max}}{2}.\label{eq:zc_def}
@@
 This was found to a relatively stable estimator of the critical layer location.
-Other estimators were used and do not significantly change the results of the
+Using other estimators does not significantly change the results of the
 analysis.
@@
 The evolution of $z_c$ depends on the deposited flux $\Delta F(t)$ per
-\autoref{eq:zc_anal}. To estimate $\Delta F(t)$ across the critical layer $z_c$
+\autoref{eq:zc_anal}. To measure $\Delta F(t)$ across the critical layer $z_c$
 from the data, we defined:
@@
     F_>(t) &= \ev{\z{F(z): F(z) < 0}}_{z \in [z_c, z_c + \Delta z]},
-        \label{eq:sabove_def},\\
-    \Delta F(t) &\equiv F_>(t) - F_{<}(t).\label{eq:ds_def}
+        \label{eq:sabove_def}\\
+    \Delta F(t) &\equiv F_<(t) - F_{>}(t).\label{eq:ds_def}
 \end{align}
@@
 to measuring the transmission feature accurately, seen in
-\autoref{fig:nl_fluxes} to be weak and attenuate quickly.
+\autoref{fig:nl_fluxes} to be weak and attenuate quickly. We average only where
+$F < 0$ since the decay scale of the trasmitted flux varies in time (see
+\autoref{fig:nl_fluxes} again), so this was found to be an adequate way of
+measuring only the transmitted flux.
 
@@
 To measure the reflected horizontal momentum flux, we recall from our linear
-simulations described in \autoref{ss:lin_ns} that we are able to accurately
+simulations described in \autoref{s:weak_sim} that we are able to accurately
 predict the incident flux from the incident wave amplitude. Thus, we write for
@@
 \begin{equation}
-    F_r(t) = F_i(t) + \Delta F(t) - F_>(t).
+    F_r(t) = F_i(t) - \Delta F(t) - F_>(t).
 \end{equation}
@@
 \end{equation}
-the Alfv\'en wave crossing time (evaluated for a CO WD in \citealt{fullerIV})
+the Alfv\'en wave crossing time (evaluated for a CO WD in \citealp{fullerIV})
 measures the magnetic coupling time and $t_{gw}$ measures the gravitational wave

### 2019-11-05 sims/2d_4_fourier/strat_helper.py

@@ -158,11 +158,13 @@ def subtract_lins(params, state_vars, sim_times, domain):
     global N_X, N_Z
     xstride = N_X // 256
     zstride = N_Z // 1024
+    xscale = 1 if xstride == 0 else int(1 / xstride)
+    zscale = 1 if zstride == 0 else int(1 / zstride)
     N_X = 256
     N_Z = 1024
 
-    x = domain.grid(0, scales=1 / xstride)
-    z = domain.grid(1, scales=1 / zstride)
+    x = domain.grid(0, scales=xscale)
+    z = domain.grid(1, scales=zscale)
     x_t = np.array([x] * len(sim_times)) # all-time
     z_t = np.array([z] * len(sim_times))
     grid_ones = np.ones_like(x + z)
@@ -173,8 +175,8 @@ def subtract_lins(params, state_vars, sim_times, domain):
     pos_slice = np.s_[:, z_bot:z_top]
     time_slice = np.s_[:, :, z_bot:z_top]
 
-    dx = domain.grid_spacing(0, scales=1 / xstride)
-    dz = domain.grid_spacing(1, scales=1 / zstride)
+    dx = domain.grid_spacing(0, scales=xscale)
+    dz = domain.grid_spacing(1, scales=zscale)
     dxdz = np.outer(dx, dz)[pos_slice]
     norm = np.exp(-z/H)[pos_slice]
 
@@ -340,12 +342,14 @@ def add_nl_problem(problem):
 def get_solver(params):
     ''' sets up solver '''
     populate_globals(params)
+    N_X_FORCED = N_X if N_X > 256 else 256
+    N_Z_FORCED = N_Z if N_Z > 1024 else 1024
     x_basis = de.Fourier('x',
-                         N_X,
+                         N_X_FORCED,
                          interval=(0, XMAX),
                          dealias=3/2)
     z_basis = de.Fourier('z',
-                         N_Z,
+                         N_Z_FORCED,
                          interval=(-1, ZMAX + 1.5),
                          dealias=3/2)
     domain = de.Domain([x_basis, z_basis], np.float64)
@@ -453,13 +457,15 @@ def load(name, params, dyn_vars, stride, start=0):
     global N_X, N_Z
     xstride = N_X // 256
     zstride = N_Z // 1024
+    xscale = 1 if xstride == 0 else int(1 / xstride)
+    zscale = 1 if zstride == 0 else int(1 / zstride)
     N_X = 256
[…]

### 2019-11-07 writeups/Su_IGW_break.tex
okay actual draft for Daniel

@@
 \citep{tidal_novae}. Most importantly, they have been proposed as the likely
-progenitors of type Ia supernovae (e.g. \citealt{Ia0,webbink} or more recently
-\citealt{Ia1,Ia2}). While presently only a few tens of compact WD binaries are
+progenitors of type Ia supernovae (e.g. \citealp{Ia0,webbink} or more recently
+\citealp{Ia1,Ia2}). While presently only a few tens of compact WD binaries are
 known \citep{lsst_wd}, \emph{Gaia} (currently gathering data) is expected to
@@
 
-\citealt{drazin, klostermeyer, winters1994} describe the onset of wave breaking.
+\citealp{drazin, klostermeyer, winters1994} describe the onset of wave breaking.
 At intermediate amplitudes, wave breaking occurs via triadic resonances which
@@
 scales that efficiently damp. The horizontal momentum flux decreases from $F$ to
-$0$ over this breaking region. The lost flux is deposited into mean flow
-\begin{equation}
-    \overline{U}(z, t) \equiv \ev*{u_x}_x.\label{eq:mean_flow}
+$0$ over this breaking region. The lost flux is deposited into a horizontal mean
+flow
+\begin{equation}
+    \overline{U}(z, t) \equiv \ev{u_x}_x.\label{eq:mean_flow}
 \end{equation}
@@
 
-% TODO stash here
-\begin{equation}
-    \overline{U}_c \equiv \omega/k_x,
-\end{equation}
-The further evolution of the critical layer with the incident IGW is
-thought to be responsible for tidal synchronization in stellar binaries
-\citep{zahn75,gn89} as well as terrestrial phenomena such as the quasi-biennial
-oscillation \citep{lindzen_qbo}.
+A horizontal shear flow $\overline{U}(z, t)\hat{x}$ enters the fluid equations
+via the Lagrangian derivative, which can be decomposed as
+\begin{equation}
+    \md{}{t} = \pd{}{t} + \overline{U} \pd{}{x} + \p{\bm{u}' \cdot \vec{\nabla}},
+\end{equation}
+where $\bm{u}'$ is the velocity field \emph{without} the shear flow, for
+clarity. Thus, $\overline{U}$ has the effect of Doppler shifting the time
+derivative into the frame comoving with the mean flow. If $\overline{U}$ is
+roughly constant, then the behavior of a small plane wave perturbation satisfies
+modified dispersion relation
+\begin{equation}
+    \p{\omega - \overline{U}k_x}^2 =
+        \frac{N^2k_{x}^2}{k_{x}^2 + k_{z}^2 + \frac{1}{4H^2}}.
+        \label{eq:disp_rel_U}
+\end{equation}
+This is just \autoref{eq:disp_rel} with $\omega \to \omega - \overline{U}k_x$.
+
+It is apparent that if $\overline{U} = \overline{U}_c$, where
+\begin{equation}
+    \overline{U}_c \equiv \frac{\omega}{k_x},
+\end{equation}
+then the dispersion relation is singular and the linear solution breaks down.
+Physically, this corresponds to the Doppler-shifted frequency of the IGW being
+zero. Anywhere $\overline{U} = \overline{U}_c$ is called a \emph{critical
+layer}.
 
 The behavior of an IGW incident upon a critical layer was first studied in the
-inviscid, linear regime in \citep{booker_bretherton}, which found nearly
+inviscid, linear regime in \citealp{booker_bretherton}, which found nearly
 complete absorption of the IGW\@. In particular, the incident wave has amplitude
@@
 that nonlinear effects may significantly enhance reflection and transmission.
+Understanding the interaction of IGW and critical layers has important
+applications: it is thought to be responsible for tidal synchronization in
+stellar binaries \citep{zahn75,gn89} as well as terrestrial phenomena such as
+the quasi-biennial oscillation \citep{lindzen_qbo}.
 
@@
 \begin{align}
-    \md{}{t} &\to \md{}{t} + \Gamma(z)\Upsilon,\nonumber\\
+    \md{}{t} &\to \md{}{t} + \Gamma(z)\Upsilon,\\
     \Gamma(z) &= \frac{1}{2\tau}\s{2 + \tanh \frac{z - z_T}{\Delta z}
@@
     particular $\hat{A}_i(t)$ asymptotes to $\approx 1$, implying continuous
-    excitation of identical IGW with the expected
-    amplitude.}\label{fig:lin_amps}
+    excitation of identical IGW with the expected amplitude. The small
+    suppression may be due to timestepping errors, as a relatively large fixed
+    step size $\Delta t = 0.1/N$ was used for this
+    simulation.}\label{fig:lin_amps}
 \end{figure}
@@
 
-\section{Wave Breaking Simulations}\label{s:sim}
+\section{Numerical Simulations of Wave Breaking}\label{s:sim}
 
@@
 In \autoref{fig:nl_fluxes}, we plot $\overline{U}$ and $F / F_{al}$ as a function
-of $z$ at the times depicted in \autoref{fig:snapshots}. While the behavior of
-$\overline{U}$ seems to conform qualitatively with the predictions of
-\autoref{ss:crit_layer}, the behavior of $F / F_{al}$ exhibits two notable
-features: (i) the incident flux seems to fluctuate greatly with time, and (ii)
-there seems to be a small transmitted feature at many of the later times. We
-will discuss further these features in \autoref{ss:reflectivity}, after first
-analyzing the propagation of the critical layer.
+of $z$ at the times depicted in \autoref{fig:snapshots}.
+At each time, $\overline{U}$ is close to zero below the critical layer, but then
+sharply increases to $\overline{U}_c$ at the critical layer. Above the critical
+layer, $\overline{U}$ varies slightly due to momentum transport within the
+spun-up layer. This is in agreement with the predictions
+of\autoref{ss:crit_layer}.
+
+Similarly, $F \lesssim F_{al}$ below the critical layer, and then decreases to
+about zero above the critical layer. However, two notable deviations can be
+observed: (i) the incident flux on the critical layer seems to fluctuate
+somewhat temporally, and (ii) there seems to be a small transmitted feature at
+many of the later times. We will discuss further these features in
+\autoref{ss:reflectivity}.
 \begin{figure}
@@
 
-For the simulation shown in \autoref{fig:nl_fluxes}, we may analyze the location
-of the critical layer. As $F$ fluctuates significantly in the simulation, we
-measure the location of the critical layer using an average of where flux
-deposition occurs
-\begin{align}
-    z_{c, \min} &= \argmin_z \z{z: F(z) > 0.3F_{al}},\nonumber\\
-    z_{c, \max} &= \argmax_z \z{z: F(z) < 0.3F_{al}},\nonumber\\
+For our fiducial simulation, we may analyze the location of the critical layer
+and compare to \autoref{eq:zc_anal}. As $F$ fluctuates significantly in the
+simulation, we measure the location of the critical layer using an average of
+where flux deposition occurs
+\begin{align}
+    z_{c, \min} &\equiv \argmin_z \z{z: F(z) > 0.3F_{al}},\\
+    z_{c, \max} &\equiv \argmax_z \z{z: F(z) < 0.3F_{al}},\\
     z_c &\equiv \frac{z_{c, \min} + z_{c, \max}}{2}.\label{eq:zc_def}
@@
 
-The evolution of $z_c$ depends on the deposited flux $\Delta F(t)$ per
-\autoref{eq:zc_anal}. To measure $\Delta F(t)$ across the critical layer $z_c$
-from the data, we defined:
-\begin{align}
-    F_{<}(t) &= \ev{F(z)}_{z \in [z_c - \Delta z - H, z_c - \Delta z]}
+We also need to measure $\Delta F(t)$ across the critical layer $z_c$ from the
+data, so we define
+\begin{align}
+    F_{<}(t) &\equiv \ev{F(z)}_{z \in [z_c - \Delta z - H, z_c - \Delta z]}
         ,\label{eq:sbelow_def}\\
-    F_>(t) &= \ev{\z{F(z): F(z) < 0}}_{z \in [z_c, z_c + \Delta z]},
+    F_>(t) &\equiv \ev{\z{F(z): F(z) < 0}}_{z \in [z_c, z_c + \Delta z]},
         \label{eq:sabove_def}\\
@@
 turbulence within the critical layer itself. The height of the critical layer is
-limited by $\mathrm{Ri} \lesssim 1$, which bounds its vertical extent $\sim
-\frac{1}{k_{z}}$. We empirically found an offset of $\Delta z = \frac{3}{k_z}$
-was necessary to be sufficiently far from strong fluctuations near the critical
-layer.
+limited by $\mathrm{Ri} \lesssim 1$ (see \autoref{ss:khi}), which bounds its
+vertical extent $\sim \frac{1}{k_{z}}$. We empirically found an offset of
+$\Delta z = \frac{3}{k_z}$ was necessary to be sufficiently far from strong
+fluctuations near the critical layer.
 
@@
 \autoref{fig:nl_fluxes} again), so this was found to be an adequate way of
-measuring only the transmitted flux.
+measuring the transmitted flux before it dissipates.
 
@@
 (ii) substituting the time-averaged $\ev{\Delta F}_t$ (over the entire length of
-the simulation) into \autoref{eq:zc_sol}. Since $z_c(t)$ is both stable and less
-well-defined at early times (when the critical layer is thick and transient
+the simulation) into \autoref{eq:zc_sol}. Since $z_c(t)$ is both unstable and
+less well-defined at early times (when the critical layer is thick and transient
 behavior is still strong), we instead integrate backwards from the end of the
@@
 velocity at early times contrasts with its overprediction at later times. This
-suggests that momentum flux absorption changes significantly over time, which is
-in line with the observations in the subsequent \autoref{ss:reflectivity}.
+suggests that momentum flux absorption changes significantly over time, in
+agreement with \autoref{ss:reflectivity}.
 
@@
 vertical distance over which the local $u_x$ increases from $0.3 \times$ its
-critical value to its critical value. The value $0.3$ effectively excludes the
-self-acceleration induced by the IGW\@. This can be written
-\begin{align}
-    z_{CL, \min}(x, t) &= \argmin_\zeta
-        \z{z: u_x(x, \zeta, t) > 0.3\overline{U}_c},\nonumber\\
-    z_{CL, \max}(x, t) &= \argmax
-        \zeta \z{z: u_x(x, \zeta, t) < \overline{U}_c},\nonumber\\
+critical value to its critical value. The value $0.3$ is necessary to exclude
+the small mean flow generated in the weakly nonlinear regime well below the
+critical layer. This is written
+\begin{align}
+    z_{CL, \min}(x, t) &\equiv \argmin_z
+        \z{z\mid u_x(x, z, t) > 0.3\overline{U}_c},\\
+    z_{CL, \max}(x, t) &\equiv \argmax_z
+        \z{z\mid u_x(x, z, t) < \overline{U}_c},\\
     \mathrm{Ri}(t) &\equiv
         \med_x\p{\frac{N^2 \p{z_{CL, \max} - z_{CL, \min}}^2}{(0.7
-            \overline{U}_c)^2}}.\label{ss:ri_med_def}
+            \overline{U}_c)^2}}.\label{eq:ri_med_def}
 \end{align}
@@
     \caption{Local Richardson number of the flow at the critical layer over time
-    as defined in \autoref{ss:ri_med_def}. The dotted red and solid green lines
+    as defined in \autoref{eq:ri_med_def}. The dotted red and solid green lines
     denote respectively the minimum and median of $\mathrm{Ri}$ over $x$ (the
@@
     t_A = \int\limits_0^R \frac{\sqrt{4\pi \rho}}{B_0}\;\mathrm{d}r
-        \sim 10^2\;\mathrm{yr}\p{\frac{10^3\;\mathrm{G}}{B}}.
-\end{equation}
-the Alfv\'en wave crossing time (evaluated for a CO WD in \citealp{fullerIV})
+        \sim 10^2\;\mathrm{yr}\p{\frac{10^3\;\mathrm{G}}{B}},
+\end{equation}
+the Alfv\'en wave crossing time (evaluated for a CO WD in \citealp{fullerIV}),
 measures the magnetic coupling time and $t_{gw}$ measures the gravitational wave
@@
 \begin{align}
-    t_{sync} &\sim \frac{M_{sync}\sigma R^2}{\tau},\nonumber\\
+    t_{sync} &\sim \frac{M_{sync}\sigma R^2}{\tau},\\
         &\sim 2 \times 10^5\;\mathrm{yr}
@@
 
-    \item Solar-type stars (inner conbmtive, outer radiative): different
+    \item Solar-type stars (inner convective, outer radiative): different
         equation of state/stratification but could be qualitatively similar.
@@
                 + \frac{F}{\overline{\rho}(z)}e^{-\frac{(z - z_0)^2}{2\sigma^2}}
-                    \cos \p{k_xx - \omega t},\nonumber\\
+                    \cos \p{k_xx - \omega t}\nonumber\\
             & + \Gamma_{NL} \bigg[-\p{\bm{u}' \cdot \bm{\nabla}}\Upsilon

### 2019-11-07 sims/2d_4_fourier/strat_helper.py
wtf ok actual last

@@ -1090,7 +1090,7 @@ def plot_front(name, params):
                      u0_avg / u_c,
                      '%s-' % color,
                      linewidth=LW * 0.5,
-                     label=r'$t=%.1f$' % sim_times[time])
+                     label=r'$t=%.1f/N$' % sim_times[time])
             # S_px sliced at time, just one for comparison
             # if time == times[len(times) // 2]:
             # #     ax2.plot(z0_cut,
@@ -1104,7 +1104,7 @@ def plot_front(name, params):
                      S_px_avg / flux_th,
                      '%s-' % color,
                      linewidth=LW * 0.5,
-                     label=r'$t=%.1f$' % sim_times[time])
+                     label=r'$t=%.1f/N$' % sim_times[time])
         # overlay analytical flux including viscous dissipation
         # # ax2.plot(z0_cut,
         # ax2.plot(z0,
@@ -1116,7 +1116,7 @@ def plot_front(name, params):
         ax2.set_xlim(z_b, ZMAX)
         ax1.set_ylim(-0.2, 1.25)
         ax2.set_ylim(-0.2, 1.2)
-        ax2.legend(fontsize=FONTSIZE)
+        ax2.legend(fontsize=FONTSIZE - 4)
 
         ax1.set_ylabel(r'$\overline{U} / \overline{U}_c$')
         ax2.set_ylabel(r'$F / F_{al}$')

### 2019-11-07 sims/aggregate.py

@@ -63,12 +63,11 @@ if __name__ == '__main__':
                         label=r'$\left<\mathcal{R}_A(t)^2\right>$')
     ln3 = mlines.Line2D([], [], color='r', marker='*', markersize=msize,
                         label=r'$\left<\mathcal{T}_S(t)\right>$')
-    ax1.legend(handles=[ln1, ln2, ln3], fontsize=12, loc='upper left')
+    ax1.legend(handles=[ln1, ln2, ln3], fontsize=16, loc='upper left')
 
 
     # ax1.set_ylabel(r'$\left<\mathcal{R}_S(t)\right>$')
     ax3.set_ylabel('Ri')
     ax3.set_xlabel('Re')
-    ax = plt.gca()
     plt.savefig('agg.png', dpi=600)
 

### 2019-11-13 writeups/Su_IGW_break.tex
wtf ok actual last
updated a ton of plots, first real draft

@@
         usenatbib,
+        referee,
     ]{mnras}
@@
     In sufficiently compact white dwarf binaries, dynamical tides raise a train
-    of internal gravity waves that propagate towards the surface. We perform 2D
-    numerical simulations of these waves undergoing nonlinear wave breaking in
-    an incompressible, isothermal atmosphere. After an initial transient phase,
-    we find that these waves induce a sharp transition between a non-rotating
-    core and synchronously rotating envelope. We find evidence that the width of
-    this transition layer is bound from below by the Kelvin-Helmholtz
-    Instability. We provide analytical formulae for absorption and reflection of
-    incident waves off the critical layer up to prefactors of order unity. These
-    prefactors converge to constant values when artificial dissipation is
-    decreased. We provide dimensionless criteria necessary to resolving momentum
-    transfer within the critical layer. Finally, we speculate on the application
-    of our model to tidal synchronization and heating in astrophysical systems.
+    of internal gravity waves that propagate towards the surface and dissipate
+    via nonlinear wave breaking. We perform 2D hydrodynamical simulations of
+    this wave breaking in an incompressible, isothermal atmosphere. After an
+    initial transient phase, we find that these waves generate a sharp
+    transition, a critical layer, between the non-rotating core and
+    synchronously rotating envelope. We find evidence that indefinite steepening
+    of this critical layer is suppressed by the Kelvin-Helmholtz instability. We
+    study the absorption and reflection of incident waves off the critical layer
+    and provide analytical formulae describing its evolution. We infer
+    necessary conditions to resolving momentum transfer within the critical
+    layer. Finally, we speculate on the application of our model to tidal
+    synchronization and heating in astrophysical systems.
 \end{abstract}
@@
 \section{Introduction}
-
-[Below copied from proposal, will rewrite]
 
@@
 
-% Usual WD binaries introduction, from proposals
-
-% IGW tidal dissipation has been thoroughly studied in many astrophysical
-% binary systems (Barker Ogilvie, Goldreich Nicholson, Zahn?). Its effect in WDs
-% (Fuller Lai) resembles early solar type stars but has not been studied
-% numerically. While the linear IGW solution is an exact solution of the
-% nonlinear system (Drazin) but is globally unstable (Drazin), the IVP must be
-% studied numerically to determine how the linear solution breaks down and
-% deposits energy/angular momentum.
-
-% From numerical studies, IGWs are known to be anti-diffusive (Lecoanet mean
-% flow papers, Lindzen) and sharpen mean flows (m = 0). Mean flows produce
-% reflection, known both analytically (Booker & Bretherton) and numerically
-% (Winters d'Asaro). IGW self-interaction via its wave-induced mean flow is
-% known to play a crucial role in IGW wavebreaking (Sutherland x2).
-
-% We study IGW breaking in the plane-parallel approximation in 2D. Known to
-% describe both wave steepening (klostermeyer 1991) and captures mean flow
-% absorption (Booker & Bretherton). While the details of breaking are presently
-% understood to be 3D (Winters & d'Asaro), we argue in (section) that our
-% results should be qualitatively similar to 3D (both 2D and 3D exhibit KHI, and
-% since we are able to resolve KHI, the spinup in 3D should be similar?).
-
-In \autoref{s:equations}, we will describe the system of equations we will
-use to analyze IGW breaking. In \autoref{s:theory}, we discuss relevant
-analytical results. In \autoref{s:sim} we present the results of numerical
-simulations. Finally, in \autoref{s:discussion} we discuss the results of the
-preceeding section.
+In this paper, we perform numerical simulations of tidally-excited IGWs
+undergoing wave breaking in an idealized WD fluid. The primary aim of the paper
+is to thoroughly characterize the angular momentum transfer within a WD by
+purely hydrodynamical effects. As a first approximation, we analyze the problem
+in a 2D Cartesian geometry, and consider IGW propagating into a fluid initially
+at rest. We find that, after an initial transient phase, the fluid develops two
+distinct zones: (i) a lower zone that has no horizontal mean flow, and (ii) an
+upper zone with significant horizontal mean flow.
+
+These two zones are separated by a \emph{critical layer}. The majority of our
+paper is dedicated to characterizing the behavior of the critical layer when
+interacting with a continuous train of IGW excited from the bottom of the fluid.
+IGW are generally \emph{anti}-diffusive, in that they steepen shear flows
+\citep{lindzen_qbo,lecoanet_meanflow}, and act to narrow the critical layer. We
+find this steepening is suppressed by the Kelvin-Helmholtz instability and
+turbulence within the very narrow critical layer. By careful accounting of the
+momentum flux budget about the critical layer, we are able to model the
+propagation of the critical layer, reflection of the incident IGW, and other
+momentum redistribution behaviors. Our results are numerically converged
+for sufficiently large resolutions.
+
+In \autoref{s:equations}, we will describe the system of equations we will use
+to analyze IGW breaking. In \autoref{s:theory}, we review existing understanding
+of wave breaking and present new analytical results. In \autoref{s:numerics}, we
+describe our numerical setup, which in \autoref{s:weak_sim} we validate in the
+weak forcing limit against linear theory. In \autoref{s:sim}, we present the
+results of simulation of IGW breaking and our characterization of the critical
+layer. Finally, we summarize and conclude in \autoref{s:discussion}.
 
@@
 some reference density $\overline{\rho}_0$ (we generally notate background
-quantities with overbars and perturbation quantities with primes). Finally, we
-consider a 2D fluid for computational feasibility: while it is well-known wave
-breaking is a 3D process \citep{klostermeyer,winters1994}, the dynamical effect
-of the breaking process is likely to be similar in 2D \citep{barker_ogilvie}.
+quantities with overbars and perturbation quantities with primes). We
+consider a 2D fluid with for computational feasibility. While
+it is well-known wave breaking is a 3D process \citep{klostermeyer,winters1994},
+the dynamical effect of the breaking process is likely to be similar in 2D
+\citep{barker_ogilvie}. Finally, we simplify by considering simple
+plane-parallel IGW\@. Since wave breaking generally occurs near the surface of
+WDs \citep[$\gtrsim 0.9R_{WD}$][]{fullerI}, the curvature of the wave front is
+negligible.
 
@@
 In the small perturbation limit, where flow velocities are small compared to the
-characteristic space and time scales $\pd{}{t} \gg \bm{u}' \cdot
+characteristic space and time scales $\pd{}{t} \gg \bm{u} \cdot
 \bm{\nabla}$, we may linearize \autoref{se:nl_upsilon}. The solution in this
@@
 \begin{equation}
-    u'_z\p{x, z, t} = Ae^{z/2H}\cos\p{k_{x}x + k_{z}z - \omega t},
+    u_z\p{x, z, t} = Ae^{z/2H}\cos\p{k_{x}x + k_{z}z - \omega t},
         \label{eq:k0z_sign}
@@
         \begin{align}
-            F &\equiv \ev{\rho u_{x}' u_{z}'}_x \equiv
-                \frac{1}{L_x}\int_0^{L_x}\limits \rho u_{x}'u_{z}'\;\mathrm{d}x,
+            F &\equiv \ev{\rho u_{x} u_{z}}_x \equiv
+                \frac{1}{L_x}\int_0^{L_x}\limits \rho u_{x}u_{z}\;\mathrm{d}x,
                     \label{eq:F_def}\\
@@
 propagating upwards, imitating a wave tidally excited deeper in the WD, and
-downwards, which are damped away by the damping layers described in
-\autoref{ss:damping}.
+downwards, which are not physically meaningful, but are damped away by the
+damping layers described in \autoref{ss:damping}.
 
@@
 \begin{equation}
-    \md{\Upsilon}{t} + u_{z}'\pd{\ln \overline{\rho}}{z}
+    \md{\Upsilon}{t} + u_{z}\pd{\ln \overline{\rho}}{z}
         = Ce^{-\frac{(z - z_0)^2}{2\sigma^2}}
@@
 
-In the linearized system, the effect of this forcing can be solved exactly. If
-we approximate $k_zH \gg 1, \sigma \ll H$, the solution can be approximated as
-two plane waves propagating away from the forcing zone
-\begin{align}
-    u_{z}'&(x, z, t) \approx{} \frac{C}{2k_z}\frac{gk_x^2}{\omega^2}
+In the linearized system, the effect of this forcing can be solved exactly
+(\autoref{s:force_solved}). If we approximate $k_zH \gg 1, \sigma \ll H$, the
+solution can be approximated as two plane waves propagating away from the
+forcing zone
+\begin{align}
+    u_{z}&(x, z, t) \approx{} \frac{C}{2k_z}\frac{gk_x^2}{\omega^2}
         e^{-\frac{k_z^2\sigma^2}{2}}
@@
 The $z > z_0$ region models an upward propagating IGW wavetrain excited deep
-in the atmosphere. A derivation of this result is provided in
-\autoref{s:force_solved}.
+in the atmosphere.
 
@@
 
-Above, we neglected advective terms $\pd{}{t} \gg \bm{u}' \cdot \bm{\nabla}$.
-However, since $\abs{\bm{u}'} \propto e^{z/2H}$ in an infinite domain all IGWs
+Above, we neglected advective terms $\pd{}{t} \gg \bm{u} \cdot \bm{\nabla}$.
+However, since $\abs{\bm{u}} \propto e^{z/2H}$ in an infinite domain all IGWs
 eventually grow to nonlinear amplitudes and break. Nonlinear wave breaking is
 expected to be important in WDs \citep{fullerI,fullerII}. We can
-estimate the height of wave breaking using $\frac{\abs{\bm{u}'}
+estimate the height of wave breaking using $\frac{\abs{\bm{u}}
 \abs{\bm{k}}}{\omega} \sim 1$. This can be rewritten using the Lagrangian
-displacement $-i\omega \bm{\xi}' = \bm{u}'$:
+displacement $-i\omega \bm{\xi} = \bm{u}$:
 \begin{equation}
@@
 This is just \autoref{eq:disp_rel} with $\omega \to \omega - \overline{U}k_x$.
-
 It is apparent that if $\overline{U} = \overline{U}_c$, where
@@
     \pd{}{t}\int\limits \overline{\rho}(z) \overline{U}(z, t)\;\mathrm{d}z
-        - \Delta F = 0,
-\end{equation}
-where $\Delta F > 0$ is the absorbed horizontal momentum. Assuming
-$\overline{U}(z > z_c) = \overline{U}_c, \overline{U}(z < z_c) = 0$, this
-condition becomes
-\begin{equation}
-    \overline{\rho}(z_c) \overline{U}_c\rd{z_c}{t} = \Delta F.\label{eq:zc_anal}
-\end{equation}
-If $\Delta F$ is constant in time, $z_c(t)$ has analytical solution
+        - F_a(t) = 0,
+\end{equation}
+where $F_a(t) > 0$ is the absorbed horizontal momentum. Assuming $\overline{U}(z
+> z_c) = \overline{U}_c, \overline{U}(z < z_c) = 0$, this condition becomes
+\begin{equation}
+    \overline{\rho}(z_c) \overline{U}_c\rd{z_c}{t} = F_a(t).\label{eq:zc_anal}
+\end{equation}
+If $F_a$ is constant in time, $z_c(t)$ has analytical solution
 \begin{equation}
     z_c(t) = -H\ln \s{e^{-\frac{z_c(t = 0)}{H}} +
-        \frac{t\Delta F}{\overline{U}_c H\overline{\rho}_0}},\label{eq:zc_sol}
+        \frac{tF_a}{\overline{U}_c H\overline{\rho}_0}},\label{eq:zc_sol}
 \end{equation}
@@
 
-\section{Numerical Simulation Setup}\label{ss:numerics}
+\section{Numerical Simulation Setup}\label{s:numerics}
 
@@
 
-For numerical convenience, we simulate equations \autoref{eq:nl_incomp},
-\autoref{eq:nl_upsilon_u}, and \autoref{eq:vol_drive} in a Cartesian box with
-size $L_x, L_z$. The $x$ direction represents the azimuthal direction in a star,
-and the $z$ direction represents the radial direction. We thus choose periodic
-boundary conditions in the $x$ direction. We also use periodic boundary
-conditions in the $z$ direction, but we damp perturbations to zero near the
-top/bottom of the domain to capture their distinct behavior (described in
-\autoref{ss:damping}). We expand all variables as a Fourier series with $N_x$
-and $N_z$ modes, and use the $3/2$ dealiasing rule to avoid aliasing errors in
-the nonlinear terms \citep{boyd}.
+We simulate equations \autoref{eq:nl_incomp}, \autoref{eq:nl_upsilon_u}, and
+\autoref{eq:vol_drive} in a Cartesian box with size $L_x, L_z$. The $x$
+direction represents the azimuthal direction in a star, and the $z$ direction
+represents the radial direction. We thus choose periodic boundary conditions in
+the $x$ direction. We also use periodic boundary conditions in the $z$
+direction, but we damp perturbations to zero near the top/bottom of the domain
+to capture their distinct behavior (described in \autoref{ss:damping}). We
+expand all variables as Fourier series with $N_x$ and $N_z$ modes, and use the
+$3/2$ dealiasing rule to avoid aliasing errors in the nonlinear terms
+\citep{boyd}.
 
@@
         have any numerical viscosity, so diffusivity must be introduced into the
-        equations to regularize the systems. We add viscosity and diffusivity to
-        the system in a way that conserves horizontal momentum (see
-        Appendix~\ref{se:strat_impl} for details). When the forcing is weak, we
-        may set $\nu=0$ because there is negligible energy transfer between
-        modes.
+        equations to regularize the systems at large wavenumbers. We add
+        viscosity and diffusivity to the system in a way that conserves
+        horizontal momentum (see Appendix~\ref{se:strat_impl} for details).
+
+        To quantify $\nu$, we define the dimensionless Reynolds number
+        \begin{equation}
+            \mathrm{Re}  \equiv \frac{\omega}{\nu k_{z}^2}. \label{eq:re_def}
+        \end{equation}
+        When the forcing is weak, we may set $\nu \approx 0$, or $\mathrm{Re}
+        \gg 1$ because there is negligible energy transfer between modes. Note
+        that $\nu = 0$ exactly changes the order of the PDE system, as the
+        highest-order derivatives appear only in the dissipitive terms, so a
+        nonzero $\nu$ must be chosen even in linear simulations.
 \end{itemize}
@@
 \begin{align}
-    \md{}{t} &\to \md{}{t} + \Gamma(z)\Upsilon,\\
+    \md{}{t} &\to \md{}{t} + \Gamma(z),\\
     \Gamma(z) &= \frac{1}{2\tau}\s{2 + \tanh \frac{z - z_T}{\Delta z}
@@
 $\xi_z k_z \ll 1$ everywhere. In particular, $\xi_z k_z \approx \scinot{5}{-5}$
-just above the forcing zone.
+just above the forcing zone. A nonzero $\nu$ corresponding to $\mathrm{Re} =
+10^7$ was used.
 
@@
 $\bm{u}_{al}$ over some region $z \in [z_b, z_t]$ can be estimated with
-estimator $\hat{A}_i(t)$ (subscript $i$ denotes incident wave)
-\begin{equation}
-    \hat{A}_i(t) = \frac{\int\limits_{z_b}^{z_t}\int\limits_0^{L_x}
+\begin{equation}
+    A_i(t) = \frac{\int\limits_{z_b}^{z_t}\int\limits_0^{L_x}
         \overline{\rho}\p{\bm{u} \cdot \bm{u}_{al}}\;\mathrm{d}x\mathrm{d}z}
@@
 \end{equation}
-If $\bm{u} = \bm{u}_{al}$, then $\hat{A}_i(t) = 1$. The energy norm is a
-traditional normalization choice such that the overlap between $\bm{u},
+The subscript $i$ denotes the incident wave. If $\bm{u} = \bm{u}_{al}$, then
+$A_i(t) = 1$. The normalization is chosen such that the overlap between $\bm{u},
 \bm{u}_{al} \propto e^{z/2H}$ is evenly weighted throughout the integration
@@
 For a simulation satisfying $\xi_z k_z \ll 1$ everywhere, we expect
-$\hat{A}_i(t) = 1$ when integrated between the forcing and damping zones, i.e.\
+$A_i(t) = 1$ when integrated between the forcing and damping zones, i.e.\
 $z_b \gtrsim z_0, z_t \lesssim z_T$ ($z_0, z_T$ are defined in
@@
 (using $z_t = z_T - \Delta z$ just below the damping layer instead does not
-change the results). The resulting measurement of $\hat{A}_i(t)$ is shown in
+change the results). The resulting measurement of $A_i(t)$ is shown in
 \autoref{fig:lin_amps}.
@@
     \caption{Amplitude of excited wave over time in weak forcing simulation,
-    computed using \autoref{eq:ahat_def}. $\hat{A}_i(t) = 1$ corresponds to
+    computed using \autoref{eq:ahat_def}. $A_i(t) = 1$ corresponds to
     perfect agreement with the analytical estimate. After an initial transient
-    phase, we observe great agreement with \autoref{eq:uz_lin}, and in
-    particular $\hat{A}_i(t)$ asymptotes to $\approx 1$, implying continuous
+    phase, we observe $A_i(t)$ asymptotes to $\approx 1$, implying continuous
     excitation of identical IGW with the expected amplitude. The small
@@
 
-% times on fluxes.png are 054, 086, 153
+% times on fluxes.png are 054, 086, 153, 451
 
 To perform simulations of wave breaking phenomena, we use the same values as
-\autoref{ss:numerics} except for $C$ and $\nu$. In particular, we choose $C$
-such that $\at{\xi_z k_z}_{z_0} = 0.1$ in the forcing zone. The linear solution
-predicts $\xi_z k_z \sim 4.25$ at the upper damping zone $z_T$. The dissipation
-parameter $\nu$ was varied across the various simulations. To quantify $\nu$, we
-define dimensionless Reynolds number
-\begin{equation}
-    \mathrm{Re}  \equiv \frac{\omega}{\nu k_{z}^2}. \label{eq:re_def}
-\end{equation}
-A table of our simulations can be found in \autoref{tab:params}.
+described in \autoref{s:numerics} except for $C$ and $\nu$. In particular, we
+choose $C$ such that $\at{\xi_z k_z}_{z_0} = 0.1$ in the forcing zone. The
+linear solution predicts $\xi_z k_z \sim 4.25$ at the upper damping zone $z_T$.
+$\mathrm{Re}$ was chosen to be as large as possible across the various
+resolutions. A table of our simulations can be found in \autoref{tab:params}.
 \begin{table}
@@
     \end{tabular}
-    \caption{Table of simulation resolutions.}\label{tab:params}
+    \caption{Table of simulation resolutions with wave
+    breaking.}\label{tab:params}
 \end{table}
@@
 
-A full video of our higher-resolution run at $N_x = 768$, $N_z = 3072$,
-$\mathrm{Re} = 1024$ is available
+A full video of our simulation with $N_x = 768$, $N_z = 3072$, $\mathrm{Re} =
+1024$ is available
 online\footnote{http://www.princeton.edu/~lecoanet/data/breaking\_wave.mov}. We
@@
 % convert yubo_000054.png -crop 2000x2000+150+250 out.png
-In \autoref{fig:snapshots}, we present snapshots of the velocity field $\bm{u}$
-at various interesting phases of the simulation.
+In \autoref{fig:snapshots}, we present snapshots of $u_x$ and $\Upsilon$ at
+various interesting phases of the simulation. The flow evolves through distinct
+stages:
 \begin{enumerate}
@@
         in the simulation domain but breaks down into smaller-scale features at
-        higher $z$. Some characteristic swirling motion can be seen in both
-        $u_x, u_z$, highly suggestive of Kelvin-Helmholtz instabilities.
+        higher $z$. Some characteristic swirling motion can be seen in the
+        advected scalar $\Upsilon$, indicating Kelvin-Helmholtz instabilities.
 
@@
         width of this transition. This is supported by the conclusions in
-        \autoref{ss:khi}.
+        \autoref{s:khi}.
 
@@
     distinct phases of the evolution of the flow. Note that $\overline{U}_c
-    \approx 0.16$ for the parameters used. The four times depicted are $t =
-    413.4$ (top left), $t = 658.5$ (top right), $t = 1171.4$ (bottom left), and
-    $t = 3437.8$ (bottom right). The four snapshots illustrate the initial
-    transient wave breaking phase, formation of a distinct critical layer,
-    steepening of the critical layer, and propagation of the critical layer
-    respectively.}\label{fig:snapshots}
+    \approx 0.16HN$ for the parameters used. The four times depicted are $t =
+    413.4/N$ (top left), $t = 658.5/N$ (top right), $t = 1171.4/N$ (bottom
+    left), and $t = 3437.8/N$ (bottom right). The four snapshots illustrate (i)
+    the initial transient wave breaking phase, (ii) formation of a distinct
+    critical layer, (iii) steepening of the critical layer, and (iv) propagation
+    of the critical layer respectively.}\label{fig:snapshots}
 \end{figure*}
@@
 spun-up layer. This is in agreement with the predictions
-of\autoref{ss:crit_layer}.
+of \autoref{ss:crit_layer}.
 
@@
 observed: (i) the incident flux on the critical layer seems to fluctuate
-somewhat temporally, and (ii) there seems to be a small transmitted feature at
-many of the later times. We will discuss further these features in
-\autoref{ss:reflectivity}.
+somewhat temporally, and (ii) there seems to be a small negative flux above the
+critical layer at later times. These are addressed in the following subsection.
 \begin{figure}
@@
     1024$ simulation. $\overline{U}, F$ follow their definitions in
-    \autoref{eq:mean_flow} and \autoref{eq:F_def}. $\overline{U}$ is given in
-    units of $\overline{U}_c = \frac{\omega}{k_{x}}$ the critical horizontal
-    flow velocity (see \autoref{ss:crit_layer}). The propagation of the critical
-    layer towards lower $z$ and sharp deposition of $F$ at the critical layer
-    are evident.}\label{fig:nl_fluxes}
+    \autoref{eq:mean_flow} and \autoref{eq:F_def}. The two distinct zones of
+    mean flow separated by a critical layer, the propagation of this critical
+    layer towards lower $z$, and the sharp deposition of $F$ at the critical
+    layer are evident.}\label{fig:nl_fluxes}
 \end{figure}
 
-\subsection{Propagating Critical Layer}
-
-For our fiducial simulation, we may analyze the location of the critical layer
-and compare to \autoref{eq:zc_anal}. As $F$ fluctuates significantly in the
-simulation, we measure the location of the critical layer using an average of
-where flux deposition occurs
-\begin{align}
-    z_{c, \min} &\equiv \argmin_z \z{z: F(z) > 0.3F_{al}},\\
-    z_{c, \max} &\equiv \argmax_z \z{z: F(z) < 0.3F_{al}},\\
-    z_c &\equiv \frac{z_{c, \min} + z_{c, \max}}{2}.\label{eq:zc_def}
-\end{align}
-This was found to a relatively stable estimator of the critical layer location.
-Using other estimators does not significantly change the results of the
-analysis.
-
-We also need to measure $\Delta F(t)$ across the critical layer $z_c$ from the
-data, so we define
-\begin{align}
-    F_{<}(t) &\equiv \ev{F(z)}_{z \in [z_c - \Delta z - H, z_c - \Delta z]}
-        ,\label{eq:sbelow_def}\\
-    F_>(t) &\equiv \ev{\z{F(z): F(z) < 0}}_{z \in [z_c, z_c + \Delta z]},
-        \label{eq:sabove_def}\\
-    \Delta F(t) &\equiv F_<(t) - F_{>}(t).\label{eq:ds_def}
-\end{align}
-$\ev{\dots}_{z \in [z_a, z_b]}$ denotes averaging over interval $[z_a, z_b]$.
-Below the critical layer, we average over an interval of length
+\subsection{Flux Budget}\label{ss:flux_budget}
+
+The propagation of the critical layer $z_c(t)$ is driven by the absorbed flux at
+$z_c$, following \autoref{eq:zc_anal}. When analyzing the simulation, one must
+first compute $z_c$ before analyzing the flux budget in its vicinity, but
+conceptually $z_c$ is a consequence of flux redistribution at the critical
+layer. We proceed in conceptual order and first describe how we analyze the flux
+transfer at $z_c$.
+
+In general, the flux budget at the critical layer can be decomposed as
+\begin{equation}
+    F_i(t) = F_a(t) + F_r(t) + F_s(t),
+\end{equation}
+where $F_i > 0$ is the incident flux, $F_a > 0$ is the absorbed flux, $F_r > 0$
+is the reflected flux, and $F_s < 0$ is some flux above the critical layer,
+likely responsible for momentum redistribution within the synchronized outer
+zone of the WD\@. Careful accounting of $F_s$ turns out to be important to
+obtain the correct critical layer propagation. A more specific physical
+interpretation of $F_s$ is unclear; it is somewhat tempting but unfounded to
+identify $F_s$ with some transmitted flux.
+
+After measuring $z_c$ at each time (see \autoref{ss:cl_prop}), we can then
+determine each of $F_i$, $F_a$, $F_r$, $F_s$ as follows:
+\begin{align}
+    F_i(t) &= \ev{\rho u_{x} u_{z}}_xA_i^2(t),\\
+    F_i(t) - F_r(t) &= \ev{F(z)}_{z \in [z_c - \Delta z - H, z_c - \Delta z]}
+        ,\label{eq:fr_def}\\
+    F_s(t) &= \ev{\z{F(z): F(z) < 0}}_{z \in [z_c, z_c + \Delta z]},
+        \label{eq:fs_def}\\
+    F_a(t) &= F_i - F_r - F_s.\label{eq:fa_def}
+\end{align}
+Here, $\ev{\dots}_{z \in [z_a, z_b]}$ denotes an average over the interval
+$[z_a, z_b]$. Below the critical layer, we average over an interval of length
 $\frac{2\pi}{k_{z}}$ a full vertical wavelength. The offset $\Delta z$ is
@@
 turbulence within the critical layer itself. The height of the critical layer is
-limited by $\mathrm{Ri} \lesssim 1$ (see \autoref{ss:khi}), which bounds its
+limited by $\mathrm{Ri} \lesssim 1$ (see \autoref{s:khi}), which bounds its
 vertical extent $\sim \frac{1}{k_{z}}$. We empirically found an offset of
@@
 
-Above the critical layer, waves decay over a much smaller vertical extent as
-they have larger $z$ wavenumbers thanks to the shear flow Doppler shift.
-Accurate measurement of these fluctuations above the critical layer is important
-to measuring the transmission feature accurately, seen in
-\autoref{fig:nl_fluxes} to be weak and attenuate quickly. We average only where
-$F < 0$ since the decay scale of the trasmitted flux varies in time (see
-\autoref{fig:nl_fluxes} again), so this was found to be an adequate way of
-measuring the transmitted flux before it dissipates.
+Above the critical layer, we observe that the $F_s$ feature has varying width
+(see \autoref{fig:nl_fluxes} again) but contributes significantly to the total
+flux budget. We average only where $F < 0$ to be robust to such width
+variations, and this was found to be a sufficiently accurate way of measuring
+$F_s$.
+
+\subsection{Critical Layer Propagation}\label{ss:cl_prop}
+
+With a careful determination of $F_a$, we can next make predictions for the
+propagation of $z_c(t)$ and compare to the observed propagation.
+In principle, $z_c$ is the location where the incident flux significantly
+attenuates. In the simulations, shear turbulence causes $F$ to have significant
+spatial and temporal fluctuations that propagate to large temporal fluctuations
+in $z_c(t)$. To minimize these spurious fluctuations, we measure the location of
+the critical layer using a spatial average of where flux deposition occurs
+\begin{align}
+    z_{c, \min}(t) &\equiv \argmin_z \z{z: F(z, t) > 0.3F_{al}},\\
+    z_{c, \max}(t) &\equiv \argmax_z \z{z: F(z, t) < 0.3F_{al}},\\
+    z_c(t) &\equiv \frac{z_{c, \min}(t) + z_{c, \max}(t)}{2}.\label{eq:zc_def}
+\end{align}
+Measuring $z_c$ in other ways does not significantly change the results of the
+analysis.
 
 Finally, we may plot the measured $z_c$ against two semi-analytic predictors:
-(i) integration of \autoref{eq:zc_anal} using the measured $\Delta F(t)$, and
-(ii) substituting the time-averaged $\ev{\Delta F}_t$ (over the entire length of
+(i) integration of \autoref{eq:zc_anal} using the measured $F_a(t)$, and
+(ii) substituting the time-averaged $\ev{F_a}_t$ (over the entire length of
 the simulation) into \autoref{eq:zc_sol}. Since $z_c(t)$ is both unstable and
 less well-defined at early times (when the critical layer is thick and transient
-behavior is still strong), we instead integrate backwards from the end of the
-simulation, using $z_c(t_f)$ as initial condition. The resulting predictors are
-depicted in \autoref{fig:nl_front}. The good agreement between the evolution of
-$z_c(t)$ and its estimate via $\Delta F(t)$ and \autoref{eq:zc_anal} are
-noteworthy.
-
-Also of interest is the behavior of the time-averaged $\ev{\Delta F}_t =
-0.71F_{al}$ predictor, which serves mostly as a fiducial in the plot. The
-general agreement clearly demonstrates $\Delta F < F_{al}$ and that momentum
-flux absorption is incomplete. However, its underprediction of critical layer
-velocity at early times contrasts with its overprediction at later times. This
-suggests that momentum flux absorption changes significantly over time, in
-agreement with \autoref{ss:reflectivity}.
-
+behavior is still strong), we solve \autoref{eq:zc_anal} by integrating
+backwards from the end of the simulation, using $z_c(t_f)$ as initial condition.
+The resulting predictors are depicted in \autoref{fig:nl_front}. The good
+agreement between the evolution of $z_c(t)$ and its estimate via $F_a(t)$ and
+\autoref{eq:zc_anal} are noteworthy.
+
+The behavior of the behavior of the fiducial, time-averaged $\ev{F_a}_t =
+0.71F_{al}$ predictor provides interesting insights. The general agreement
+clearly demonstrates $F_a < F_{al}$ and that momentum flux absorption is
+not complete. However, its underprediction of critical layer velocity at early
+times contrasts with its overprediction at later times. This suggests that
+momentum flux absorption changes significantly over time, in agreement with
+detailed analysis in \autoref{ss:reflectivity}.
 \begin{figure}
@@
     $z_c(t)$ from simulation data, (green) predictor of $z_c(t)$ using direct
-    integration of \autoref{eq:zc_anal} for $\Delta F(t)$ measured from
-    simulation data (described in \autoref{eq:ds_def}) and (blue) direct
-    substitution of time-averaged $\ev{\Delta F(t)}_t$ into
+    integration of \autoref{eq:zc_anal} for $F_a(t)$ measured from
+    simulation data (described in \autoref{eq:fa_def}) and (blue) direct
+    substitution of time-averaged $\ev{F_a(t)}_t$ into
     \autoref{eq:zc_sol}. Predictors use the end of the simulation as initial
@@
     The the poorer but qualitatively correct agreement of the time-averaged
-    predictor with the data shows both that $\Delta F(t) \neq F_{al}$ and that
-    $\Delta F(t)$ likely has some very real variation with
-    time.}\label{fig:nl_front}
+    predictor with the data shows both that $F_a(t) \neq F_{al}$ and that
+    $F_a(t)$ likely varies in time.}\label{fig:nl_front}
 \end{figure}
 
-\subsection{Kelvin-Helmholtz Instability and Critical Layer Width}\label{ss:khi}
-
-The observation of reflected/transmitted waves is in accordange with previous
+\subsection{Non-absorption at Critical Layer}\label{ss:reflectivity}
+
+To further understand the behavior at the critical layer, we compare two
+reflective behaviors observed in the simulation: (i), the presence of a
+reflected wave with wave vector $\bm{k}_r = k_{x}\uv{x} - k_{z}\uv{z}$, and (ii)
+$F_r$. The reflected wave amplitude and reflected flux need not agree exactly if
+some reflected flux is in higher-order modes, which is indeed the case in our
+simulations. Both are of physical interest, however: the reflected wave
+amplitude may be of physical interest in setting up standing modes in a
+realistic star, while the flux transfer properties is important for accurately
+tracking angular momentum transfer during synchronization.
+
+To measure the reflected wave amplitude $A_r(t)$, we use \autoref{eq:ahat_def}
+with $\bm{u}_{al}$ computed with $\bm{k}_r$ with one key change: since the phase
+of the reflected wave is unknown, unlike that of the incident wave, we fit for
+an arbitrary horizontal offset $\delta x(t)$ at each time $t$ (not constrained
+to be on-grid), picking $\delta x$ that maximizes $A_r$. To be explicit, we
+compute
+\begin{equation}
+    A_r(t) = \max_{\delta x}\frac{\int\limits_{z_b}^{z_t}\int\limits_0^{L_x}
+        \overline{\rho}\p{\bm{u} \cdot \at{\bm{u}_{al,
+        \bm{k}_r}}_{x = x + \delta x}}\;\mathrm{d}x\mathrm{d}z}
+        {\int\limits_{z_b}^{z_t}\int\limits_0^{L_x}
+        \overline{\rho}\abs{\bm{u}_{al}}^2\;\mathrm{d}x\mathrm{d}z}
+        \label{eq:ar_def}
+\end{equation}
+In our simulations, the phase offset $\phi_r(t) \equiv k_x \delta x(t) $ behaves
+in agreement with reflection off a moving boundary at $z_c$, well approximated
+by $\abs{\pd{\phi_r}{t}} \approx 2\abs{\pd{(k_{z}z_c)}{t}}$.
+
+Since reflectivity depends sensitively on accurate measurements of $A_i,
+A_r$, we remark \autoref{eq:ahat_def} and \autoref{eq:ar_def} ensure
+orthogonality between $k_{x}\uv{x} \pm k_{z}\uv{z}$ modes. The integral for
+$A_r$ is also performed over $z \in [z_0 + 3\sigma, z_0 + 3\sigma + H]$. Since
+$A_i(t), A_r(t)$ vary somewhat strongly over time, we perform time averaging
+over interval approximately $8\pi/\omega$, denoted by angle brackets. We can
+then define the amplitude reflectivity
+\begin{equation}
+    \mathcal{R}_A(t) \equiv \frac{\ev{A_r}(t)}{\ev{A_i}(t)}
+        .\label{eq:Ra_def}
+\end{equation}
+% 7/N per timestep, 13 timestep smoothing kernel, 24/N period
+
+To compare to the behavior of the flux at the critical layer, we compare
+$\mathcal{R}_A^2(t)$ to the ratios of $F_r$ and $F_s$ to $F_i$. We define
+shorthand
+\begin{align}
+    \hat{F}_r &\equiv \frac{\ev{F_r}(t)}{\ev{F_i}(t)},
+        \label{eq:srefl_def1}\\
+    \hat{F}_s &\equiv -\frac{\ev{F_s}(t)}{\ev{F_i}(t)}.
+        \label{eq:srefl_def}
+\end{align}
+\autoref{fig:nl_f_amps} demonstrates the relations between these quantities. All
+of these quantities vary significantly temporally but their mean values appear
+to converge towards the end of the simulation. A significant fluctuation in
+$A_i(t)$ is observed. We were unable to identify the cause of this unintended
+fluctuation, but it doees not affect $\mathcal{R}_A$ thanks to the time
+averaging used.
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8\columnwidth]{plots/nl_f_amps.png}
+    \caption{The top panel shows the incident wave amplitude $A_i(t)$ (green)
+    and the reflected wave amplitude $A_r(t)$ (red) just above the forcing zone.
+    The bottom panel shows the behavior of the four components of the horizontal
+    momentum flux budget over time, in units of the analytical estimate
+    \autoref{eq:S_lin}: (green) flux incident on the critical layer, (blue) flux
+    absorbed by the critical layer, (red) flux reflected at the critical layer,
+    and (black) flux inside the synchronized layer.}\label{fig:nl_f_amps}
+\end{figure}
+
+Using these measured quantities, we may make plots of $\mathcal{R}_A^2$,
+$\hat{F}_r$, $\hat{F}_s$, which are provided in \autoref{fig:nl_f_refl}. A
+comparison between $\mathcal{R}_A^2$ and $\hat{F}_r$ is appropriate as $F
+\propto A^2$. Visual inspection suggests the three quantities have reached their
+asymptotic values after $t \gtrsim 1750/N$. An observation may be made that in
+general $\hat{F}_r \geq \mathcal{R}_A^2$; this conforms to the expectation
+that reflected flux consists of the simple reflected mode and higher order modes
+as well.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\columnwidth]{plots/nl_f_refl.png}
+    \caption{Behavior of $\mathcal{R}_A^2$, $\hat{F}_r$, and $\hat{F}_s$ as
+    described by Equations~\ref{eq:Ra_def},~\ref{eq:srefl_def1},
+    and~\ref{eq:srefl_def}. The coefficients seem to become comparatively stable
+    past about $t = 1750/N$, indicating that an asymptotic value may have been
+    reached.}\label{fig:nl_f_refl}
+\end{figure}
+
+\subsection{Resolution Study}\label{ss:convergence}
+
+As the primary test of whether our simulations have obtained sufficient
+resolution, we consider the convergence of $\mathcal{R}_A^2$, $\hat{F}_r$,
+$\hat{F}_s$. Furthermore, we are able to compare our results against
+\autoref{eq:crit_coeffs}, which appear to suggest that the behavior of incident
+waves on the critical layer depends only on the local Richardson number (see
+\autoref{s:khi} for details on our measurement of $\mathrm{Ri}$).
+
+For each simulation in \autoref{tab:params}, we compute the median value of each
+of $\mathcal{R}_A^2$, $\hat{F}_r$, $\hat{F}_s$, and $\mathrm{Ri}$ over the last
+$1/4$ of simulation times from each simulation, where all simulations temporally
+converged to what appeared to be asymptotic values. The bars show the temporal
+variation of each measurement between the $16\%$ and $84\%$ range. We illustrate
+the behavior of these averages at different $\mathrm{Re}$ in \autoref{fig:agg}.
+\begin{figure}
+    \centering
+    \includegraphics[width=\columnwidth]{plots/agg.png}
+    \caption{Convergence of median $\hat{F}_r$, $\mathcal{R}_A^2$, $\hat{F}_s$,
+    and $\mathrm{Ri}$ across runs with varying resolution and viscosity as given
+    in \autoref{tab:params}. The bars show the temporal variation of each
+    measurement between the $16\%$ and $84\%$ range. Small horizontal
+    displacements are made for data points at identical $\mathrm{Re}$ for
+    readability. Note that simulations with larger $\mathrm{Re}$ and resolution
+    correspond to smaller viscosity and are more physically realistic. Our
+    measurements seem to converge at large $\mathrm{Re}$. At the smallest
+    $\mathrm{Re}$ value, $\mathrm{Ri} \approx 50$ is too large to fit on the
+    plot.}\label{fig:agg}
+\end{figure}
+
+It is apparent that the Richardson number rapidly converges to $\mathrm{Ri}
+\approx 0.4$ but the non-absorption measures converge more slowly. This is in
+disagreement with \autoref{eq:crit_coeffs} calculated via the linear analytical
+theory. This tension is natural: fluid motion within the critical layer is
+turbulent, so transmission/reflection at the critical layer cannot be captured
+by any linear calculation.
+
+\section{Summary and Discussion}\label{s:discussion}
+
+In this paper, we have performed numerical simulations of breaking IGWs in a
+2D idealized WD fluid. We observe spontaneous formation of a critical layer that
+separates a synchronized upper layer of fluid and a lower layer of fluid with no
+mean horizontal flow. This critical layer then propagates downwards as a
+continuous train of IGW are incident on it (see \autoref{fig:snapshots} for
+snapshots from a fiducial simulation). Our primary conclusions regarding the
+evolution of the critical layer are as follows:
+\begin{itemize}
+    \item The critical layer propagates downward by partially absorbing incident
+        flux supplied by the continuous train of IGW, eventually spinning up the
+        entire fluid (see \autoref{fig:nl_fluxes} for the time evolution of the
+        mean flow and momentum flux in the fluid).
+
+    \item \autoref{fig:nl_front} shows that the height of the critical layer
+        $z_c(t)$ can be estimated from \autoref{eq:zc_anal}. The critical layer
+        only absorbs $\sim 70\%$ of the incident flux in most simulations (see
+        \autoref{fig:agg} for convergence at high resolutions).
+
+        Note this absorbed fraction of incident momentum flux we
+        measure differs from both analytical \citep{booker_bretherton} and
+        numerical \citep{barker_ogilvie} results in the existing literature. The
+        former result is a linear calculation and is superceded by our nonlinear
+        simulation; we speculate on our discrepancy with the latter in
+        \autoref{ss:other} below.
+
+    \item We compute the asymptotic amplitude of reflected waves off the
+        critical layer as well as the average reflected momentum flux in
+        \autoref{fig:nl_f_refl}. Also illustrated is the asymptotic value of an
+        extra momentum flux just above the critical layer, likely responsible
+        for momentum redistribution within the critical layer. Our simulations
+        are run sufficiently long to be temporally converged.
+
+    \item Finally, the largest flow variations are found within the turbulent
+        critical layer, whose minimum width is set by the Kelvin-Helmholtz
+        instability.
+\end{itemize}
+In the subsequent sections, we will discuss the validity of these results and
+their application to astrophysical systems.
+
+\subsection{Physical Sources of Dissipation in WDs}\label{ss:disp}
+
+The most significant linear damping in WD g-modes comes from radiative damping
+\citep{fullerI}. In \citep{wu} and \citep{fullerI}, the radiative damping rate
+is given in terms of $\omega_i = \gamma \omega_r$, where $\omega_r$ is the
+frequency of the g-mode. Typical values for $\gamma$ range from $10^{-4}$ to
+$10^{-11}$ depending on $n$ of the g-mode.
+
+We will assume this prescription directly transfers to propagating IGW, which
+results in general agreement with \citep{bukart}'s estimate of radiative damping
+rates. Then, making coarse identification $\omega_i \sim \nu k^2 \approx \nu
+k_z^2$, we find that $\mathrm{Re} \sim \frac{1}{\gamma}$. Even at $\gamma =
+10^{-4}$ however, the corresponding $\mathrm{Re}$ is far too weak to suppress
+reflection/transmission at the critical layer (e.g.\ \autoref{fig:agg}).
+
+Another source of dissipation considered in \citep{bukart} is turbulent
+conbmtive damping. They find this damping rate to never exceed that of
+radiative damping, and so it is also too weak to suppress critical layer
+formation in our problem.
+
+Finally, we consider the impact of magnetic winding. In \citep{bukart}, magnetic
+winding is used to enforce solid body rotation on the grounds that $t_A \gg
+t_{gw}$, where
+\begin{equation}
+    t_A = \int\limits_0^R \frac{\sqrt{4\pi \rho}}{B_0}\;\mathrm{d}r
+        \sim 10^2\;\mathrm{yr}\p{\frac{10^3\;\mathrm{G}}{B}},
+\end{equation}
+the Alfv\'en wave crossing time (evaluated for a CO WD in \citealp{fullerIV}),
+measures the magnetic coupling time and $t_{gw}$ measures the gravitational wave
+inspiral timescale. Before solid body rotation is attained, another relevant
+timescale is the synchronization timescale $t_s$. For a tidal torque $\tau$ and
+tidal forcing frequency $\sigma = m\p{\Omega - \Omega_{spin}}$, we note that
+angular momentum transfer is $\pd{M_{sync}}{t} \sigma R^2 = \tau$, where
+$M_{sync}$ denotes the mass of the WD that has synchronized. Thus, the
+synchronization timescale is
+\begin{align}
+    t_{sync} &\sim \frac{M_{sync}\sigma R^2}{\tau},\\
+        &\sim 20\;\mathrm{yr}
+            \p{\frac{M_{sync}}{10^{-4}M_{\odot}}}
+            \p{\frac{\sigma}{2\pi / (1\;\mathrm{hr})}}
+            \p{\frac{R}{R_{\oplus}}}^2
+            \p{\frac{10^{-14} GM_{\odot}^2/R_{\oplus}}{\tau}}.
+\end{align}
+A representative $\tau$ has been taken from \citep{bukart}. Only $M_{sync} \sim
+10^{-4}M_{\odot}$ need be heated for the thermodynamically interesting effects
+studied in \citep{fullerIV} and \citep{tidal_novae}, it seems that strong shear
+flows may exist even in the presence of magnetic coupling.
+
+\subsection{Applicability to Astrophysical Systems}\label{ss:other}
+
+Below, we speculate on the applicability of our results to general astrophysical
+problems of interest:
+\begin{itemize}
+    \item In astrophysical systems, $k_{\perp} \ll k_r$, while in our study,
+        $k_x \simeq k_z$. True turbulence is expected to be isotropic at small
+        scales, which may introduce different behavior for different ratios of
+        $k_z/k_x$. However, $k_x \ll k_z$ is a numerically difficult regime; we
+        defer its exploration to future work.
+
+    \item In~\cite{barker_ogilvie}, inwards-propagating IGW are excited that
+        break via geometric focusing. They find no reflected wave despite their
+        nonlinear timescales being $10\times$ shorter than their viscous
+        timescale, in contrast with the reflective behavior observed in our
+        simulations. It is plausible the discrepancy arises due to the differing
+        causes of wave breaking (geometric focusing vs.\ density
+        stratification), but their simulations are also at significantly higher
+        artificial viscosities than ours, as we show below.
+
+        Associating $t_{L} \sim \nu k^2 \approx \nu k_z^2$ with the visocus
+        timescale and $t_{NL} \sim \bm{u} \cdot \bm{\nabla} \sim \omega$ for the
+        nonlinear timescale, we find their $\lambda = \frac{t_{NL}}{t_L} \sim
+        \mathrm{Re}$ our Reynolds number. Our simulations indicate $\mathrm{Re}
+        \gtrsim 500$ are required to observe the correct asymptotic behavior in
+        terms of horizontal momentum flux reflection/transmission, so it is
+        possible their lack of reflection is resolution limited.
+
+    \item Our results give some tentative understanding of where tidal heating
+        is expected to occur. While we use a barotropic equation of state and do
+        not track heat deposition, we can recall that the strongest small-scale
+        variations in fluid flow are localized within the critical layer. Thus,
+        we posit that a substantial fraction of the energy carried by the IGW is
+        deposited in the very narrow critical layer. Then, as the critical layer
+        propagates downwards, the entire star is heated. This tidal heating
+        model differs flom that used in \citep{tidal_novae}, and its effects
+        will be studied in future work.
+\end{itemize}
+
+\section{Acknowledgements}\label{s:ack}
+
+This work has been supported in part by NASA FINESST grant
+19-ASTRO19-0041.%chktex 8
+
+\bibliographystyle{mnras}
+\bibliography{Su_IGW_break}
+
+\clearpage
+\onecolumn
+\appendix
+
+\section{Forcing Solution}\label{s:force_solved}
+
+To solve for the excited wave amplitude in the linear problem, consider the
+complexified linearized system of equations, then $\pd{}{t} \to -i\omega,
+\pd{}{x} \to ik_x$. Assuming all fields have dependence $u_z(x, z, t) =
+\tilde{u_z}(z) e^{ik_xx - i\omega t}$ and dropping the tildes for convenience,
+this gives:
+\begin{align*}
+    \rd{u_{z}}{z} + ik_xu_x &= 0,\\
+    -i\omega u_x + ik_x \varpi + gHik_x \Upsilon &= 0,\\
+    -i\omega u_{z} + \rd{\varpi}{z} + gH\rd{\Upsilon}{z}
+        - \frac{\varpi}{H} &= 0,\\
+    -i\omega \Upsilon - \frac{u_{z}}{H} &=
+        Ce^{-\frac{(z - z_0)^2}{2\sigma^2}}.
+\end{align*}
+We have replaced all partial derivatives with regular derivatives. This can be
+recast solely in terms of $u_{z}$ as
+\begin{align*}
+     \rtd{u_{z}}{z} - k_x^2u_{z} - \frac{1}{H}\rd{u_{z}}{z}
+        + u_{z}\frac{N^2k_x^2}{\omega^2} &=
+    -\frac{gk_x^2}{\omega^2}Ce^{-\frac{(z - z_0)^2}{2\sigma^2}}
+        .\label{eq:narrow_inhomo}
+\end{align*}
+The homogeneous solutions are of form $u_{z,\pm}(z) = e^{\p{\frac{1}{2H} \pm
+ik_z}\p{z - z_0}}$ where $k_z$ satisfies the dispersion relation
+\autoref{eq:disp_rel}. We compute the solution to the inhomogeneous ODE by the
+method of variation of parameters. The Wronskian can be computed
+\begin{equation}
+    W \equiv \det \begin{vmatrix}
+        u_{z,+} & u_{z,-} \\[4pt]
+        \rd{u_{z,+}}{z} & \rd{u_{z,-}}{z}
+    \end{vmatrix} = -2ik_ze^{z/H}.
+\end{equation}
+The general solution is then
+\begin{equation}
+    u_z = -u_{z,+}\int \frac{1}{W} u_{z,-} \p{-\frac{gk_x^2}{\omega^2}
+            Ce^{-\frac{(z - z_0)^2}{2\sigma^2}}}\;\mathrm{d}z
+        + u_{z,-}\int \frac{1}{W} u_{z,+} \p{-\frac{gk_x^2}{\omega^2}
+            Ce^{-\frac{(z - z_0)^2}{2\sigma^2}}}\;\mathrm{d}z.
+\end{equation}
+Taking these integrals and applying boundary conditions $u_z\p{z \to \infty} =
+u_{z,+}, u_z\p{z \to -\infty} = u_{z,-}$ gives exact solution
+\begin{equation}
+    u(z) = \frac{C}{2ik_z}\frac{gk_x^2}{\omega^2}
+            e^{\frac{\p{\frac{\sigma^2}{2H} \pm ik_z\sigma^2}^2}{2\sigma^2}}
+                \sqrt{2\sigma^2} \frac{\sqrt{\pi}}{2}\s{
+                    -u_+\p{1 + \erf\p{\xi}}
+                    + u_- \p{\erf\p{\xi} - 1}}
+\end{equation}
+If we are concerned with only $z$ scales significantly larger than $\sigma$,
+then we may take $\erf(\xi) \approx \Theta(\xi)$. If we further assume $k_zH \gg
+1$ and restore the $e^{ik_xx - i\omega t}$ factor, we recover the cited form
+\autoref{eq:uz_lin}
+\begin{equation}
+    u_{z}(x, z, t) = -\frac{C}{2ik_z}\frac{gk_x^2}{\omega^2}
+        e^{-\frac{k_z^2\sigma^2}{2}}
+        \sqrt{2\pi \sigma^2} e^{ik_xx - i\omega t} \times
+    \begin{cases}
+        e^{\p{\frac{1}{2H} + ik_z}\p{z - z_0} + \frac{k_z\sigma^2}{2H}}
+            & z > z_0,\\
+        e^{\p{\frac{1}{2H} - ik_z}\p{z - z_0} - \frac{k_z\sigma^2}{2H}}
+            & z > z_0.
+    \end{cases}
+\end{equation}
+We use the plane wave approximation in our fitting routines since we generally
+fit many $\sigma$ away from $z_0$, and orthogonality with other modes is
+essential for a reliable fit.
+
+\section{Kelvin-Helmholtz Instability and Critical Layer Width}\label{s:khi}
+
+The observation of reflected/transmitted waves is in accordance with previous
 studies as discussed in \autoref{ss:wave_breaking}. Studies have shown that a
@@
 \autoref{fig:snapshots}). It is natural to suspect then that the shear flow
-cannot steepen any further than KHI onset. To verify this, we decided to compute
-a local $\mathrm{Ri}$ for the shear flow around the critical layer.
+cannot steepen any further than KHI onset. To verify this, we compute a local
+$\mathrm{Ri}$ for the shear flow around the critical layer.
 
@@
 the minimum over $x$ (the maximum is significantly noisier). Both of these are
-shown in \autoref{fig:nl_f_ri}. The quick evolution of $\mathrm{Ri}$ to its
-saturated value is reflective of the fact that IGW are \emph{anti-diffusive}.
-This property is a simple consequence of the IGW dispersion relation
-\autoref{eq:disp_rel}, which is approximately $\omega k_z \approx Nk_x$ such
-that as an IGW propagates into a shear flow, $\omega$ decreases and $k_z$
-increases, enhancing dissipation and further driving mean flow acceleration.
+shown in \autoref{fig:nl_f_ri}. The critical layer quickly narrows to its
+minimum width, driven by the anti-diffusivity of IGWs.
 \begin{figure}
     \centering
-    \includegraphics[width=\columnwidth]{plots/nl_f_ri.png}
+    \includegraphics[width=0.6\columnwidth]{plots/nl_f_ri.png}
     \caption{Local Richardson number of the flow at the critical layer over time
-    as defined in \autoref{eq:ri_med_def}. The dotted red and solid green lines
-    denote respectively the minimum and median of $\mathrm{Ri}$ over $x$ (the
-    maximum greatly exceeds the vertical scale of this plot). These numbers
-    effectively measure the mean and spread in width of the critical layer over
-    $x$. Note that $\mathrm{Ri} \sim \frac{1}{4}$ corresponds to the KHI, so
-    this plot suggests the shear at the critical layer does not steepen past the
-    KHI onset.}\label{fig:nl_f_ri}
+    as defined in \autoref{eq:ri_med_def}. The red and solid green lines denote
+    respectively the minimum and median of $\mathrm{Ri}$ over $x$ (the maximum
+    greatly exceeds the vertical scale of this plot). These numbers effectively
+    measure the mean and spread in width of the critical layer over $x$. Note
+    that $\mathrm{Ri} \sim \frac{1}{4}$ corresponds to the KHI, so this plot
+    suggests the shear at the critical layer does not steepen past the KHI
+    onset.}\label{fig:nl_f_ri}
 \end{figure}
-
-\subsection{Non-absorption at Critical Layer}\label{ss:reflectivity}
-
-We identify two manifestations of non-absorptive behavior: (i), the presence of
-a reflected wave with wave vector $\bm{k} = k_{x}\uv{x} - k_{z}\uv{z}$, and (ii)
-the amount of horizontal momentum flux $F$ reflected/transmitted by the critical
-layer. The reflected wave amplitude and reflected flux need not agree exactly if
-some reflected flux is in higher-order modes, which is indeed the case in our
-simulations. Both are of physical interest, however: the reflected wave
-amplitude may be of physical interest in setting up standing modes in a
-realistic star, while the flux transfer properties is important for accurately
-tracking angular momentum transfer during synchronization.
-
-To measure the reflected wave amplitude, we use almost the same definition
-\autoref{eq:ahat_def} except using $k_{z} \to -k_{z}$; call this estimated
-$\hat{A}_r(t)$ the amplitude of the downwards-propagating reflected wave. To
-compute $\hat{A}_r(t)$, we furthermore permit an arbitrary phase offset
-$\phi_r(t)$ at each time $t$, since the phase of the reflected wave is unknown,
-unlike that of the incident wave. $\phi_r(t)$ in our simulations
-behaves in agreement with reflection off a moving boundary at $z_c$,
-well approximated by $\abs{\pd{\phi_r}{t}} \approx 2\abs{\pd{(k_{z}z_c)}{t}}$.
-
-Since reflectivity depends sensitively on accurate measurements of $\hat{A}_i,
-\hat{A}_r$, we remark \autoref{eq:ahat_def} ensures orthogonality between
-$k_{x}\uv{x} \pm k_{z}\uv{z}$ modes. The integral for $\hat{A}_r$ is also
-performed over $z \in [z_0 + 3\sigma, z_0 + 3\sigma + H]$. Since
-$\hat{A}_i(t), \hat{A}_r(t)$ vary somewhat strongly over time, we perform time
-averaging over interval approximately $8\pi/\omega$, denoted by angle brackets.
-We can then define the amplitude reflectivity
-\begin{equation}
-    \mathcal{R}_A(t) \equiv \frac{\ev{\hat{A}_r}(t)}{\ev{\hat{A}_i}(t)}
-        .\label{eq:Ra_def}
-\end{equation}
-% 7/N per timestep, 13 timestep smoothing kernel, 24/N period
-
-To measure the reflected horizontal momentum flux, we recall from our linear
-simulations described in \autoref{s:weak_sim} that we are able to accurately
-predict the incident flux from the incident wave amplitude. Thus, we write for
-incident flux
-\begin{equation}
-    F_i(t) \equiv \ev{\rho u_{x} u_{z}}_x\hat{A}_i^2(t).
-\end{equation}
-But then, since we already have $\Delta F(t), F_>(t)$ jump and transmitted flux
-across the critical layer at time $t$ respectively, we can immediately write
-down the reflected flux
-\begin{equation}
-    F_r(t) = F_i(t) - \Delta F(t) - F_>(t).
-\end{equation}
-Note that $F_>(t)$ is just the transmitted flux across the critical layer, since
-there is no other flux above the critical layer. We then define flux
-reflectivity and transmissivity coefficients
-\begin{align}
-    \mathcal{R}_F(t) &\equiv -\frac{\ev{F_r}(t)}{\ev{F_i}(t)},&
-    \mathcal{T}_F(t) &\equiv -\frac{\ev{F_>}(t)}{\ev{F_i}(t)}.
-        \label{eq:srefl_def}
-\end{align}
-
-The measurements of $\hat{A}_i, \hat{A}_r, F_i, \Delta F, F_r, F_>$ are given in
-\autoref{fig:nl_f_amps}. The oscillations in $\hat{A}_i, F_i, A_d, F_r$ are
-significant but the mean values seem to have temporally converged.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\columnwidth]{plots/nl_f_amps.png}
-    \caption{The top panel measures the incident wave amplitude $\hat{A}_i(t)$
-    (green) and the downwards propagating wave amplitude $\hat{A}_r(t)$ (red)
-    just above the forcing zone, normalized to the analytical estimate
-    \autoref{eq:uz_lin}. $A_d \neq 0$ due to reflection off the critical layer.
-    The bottom panel shows the behavior of three horizontal momentum fluxes over
-    time, in units of the analytical estimate \autoref{eq:S_lin}: (blue) flux
-    incident on the critical layer, (red) flux absorbed by the critical layer,
-    and (black) flux transmitted through the critical
-    layer.}\label{fig:nl_f_amps}
-\end{figure}
-
-Using these measured quantities, we may make plots of $\mathcal{R}_A^2,
-\mathcal{R}_F, \mathcal{T}_F$, which are provided in \autoref{fig:nl_f_refl}. A
-comparison between $\mathcal{R}_A^2$ and $\mathcal{R}_F$ is appropriate as $F
-\propto A^2$. Visual inspection suggests the three quantities have reached
-their asymptotic values after $t \gtrsim 1750/N$. An observation may be made
-that in general $\mathcal{R}_F \geq \mathcal{R}_A^2$; this conforms to the
-expectation that reflected flux consists of the simple reflected mode and higher
-order modes as well.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\columnwidth]{plots/nl_f_refl.png}
-    \caption{Reflectivity and transmittivity coefficients for flux and
-    amplitude-squared as described by \autoref{eq:Ra_def} and
-    \autoref{eq:srefl_def} respectively. The coefficients seem to become
-    comparatively stable past about $t = 1750/N$, indicating that an asymptotic
-    value may have been reached.}\label{fig:nl_f_refl}
-\end{figure}
-
-\subsection{Convergence}\label{ss:convergence}
-
-As the primary test of convergence in our previous sections, we consider the
-convergence of the physically significant parameters of our model. In
-particular, the convergence of $\mathrm{Ri}$, shown in \autoref{fig:nl_f_ri},
-and of the reflection/transmission coefficient asymptotic values, shown in
-\autoref{fig:nl_f_refl}, are of greatest significance.
-
-To estimate convergence, we compute the median value of each of $\mathrm{Ri},
-\mathcal{R}_A^2, \mathcal{R}_F, \mathcal{T}_F$ over the last $1/4$ of simulation
-times from each simulation, where all simulations had converged to what appeared
-to be asymptotic values. Error bars are estimated with the $16$ and $84$
-percentiles. We illustrate the convergence of these averages across simulations
-in \autoref{fig:agg}.
-\begin{figure}
-    \centering
-    \includegraphics[width=\columnwidth]{plots/agg.png}
-    \caption{Convergence of median reflection/transmission coefficients and
-    $\mathrm{Ri}$ (critical layer width) across runs with varying viscosity,
-    parameterized with $\mathrm{Re}$ (\autoref{eq:re_def}). Error bars depict
-    $16$ and $84$ percentile values. Small horizontal displacements are made for
-    data points at identical $\mathrm{Re}$ for readability. Note that
-    simulations with larger $\mathrm{Re}$ correspond to smaller viscosity and
-    are more physically realistic, and our values seem to converge towards large
-    $\mathrm{Re}$. At the smallest $\mathrm{Re}$ value, $\mathrm{Ri} \approx 50$
-    is too large to fit on the plot.}\label{fig:agg}
-\end{figure}
-
-It is apparent that the Richardson number rapidly converges to $\mathrm{Ri}
-\approx 0.4$ but the reflection/transmission coefficients converge more slowly.
-This is in disagreement with \autoref{eq:crit_coeffs} calculated via the linear
-analytical theory. This tension is natural: fluid motion within the critical
-layer is turbulent, so transmission/reflection at the critical layer cannot be
-captured by any linear theory.
-
-\section{Discussion}\label{s:discussion}
-
-In the previous section, we have argued for a continous train of breaking IGWs
-spontaneously forming a critical layer and strong shear flow. We have
-parameterized the width of the critical layer as well as horizontal momentum
-transport near the critical layer. In the subsequent sections, we will discuss
-the validity of these results and their application to astrophysical systems.
-
-\subsection{Physical Sources of Dissipation in WDs}\label{ss:disp}
-
-The most significant linear damping in WD g-modes comes from radiative damping
-\citep{fullerI}. In \citep{wu} and \citep{fullerI}, the radiative damping rate
-is given in terms of $\omega_i = \gamma \omega_r$, where $\omega_r$ is the
-frequency of the g-mode. Typical values for $\gamma$ range from $10^{-4}$ to
-$10^{-11}$ depending on $n$ of the g-mode.
-
-We will assume this prescription directly transfers to propagating IGW, which
-results in general agreement with \citep{bukart}'s estimate of radiative damping
-rates. Then, making coarse identification $\omega_i \sim \nu k^2 \approx \nu
-k_z^2$, we find that $\mathrm{Re} \sim \frac{1}{\gamma}$. Even at $\gamma =
-10^{-4}$ however, the corresponding $\mathrm{Re}$ is far too weak to suppress
-reflection/transmission at the critical layer (e.g.\ \autoref{fig:agg}).
-
-Another source of dissipation considered in \citep{bukart} is turbulent
-conbmtive damping. They find this damping rate to never exceed that of
-radiative damping, and so it is also too weak to suppress critical layer
-formation in our problem.
-
-Finally, we consider the impact of magnetic winding. In \citep{bukart}, magnetic
-winding is used to enforce solid body rotation on the grounds that $t_A \gg
-t_{gw}$, where
-\begin{equation}
-    t_A = \int\limits_0^R \frac{\sqrt{4\pi \rho}}{B_0}\;\mathrm{d}r
-        \sim 10^2\;\mathrm{yr}\p{\frac{10^3\;\mathrm{G}}{B}},
-\end{equation}
-the Alfv\'en wave crossing time (evaluated for a CO WD in \citealp{fullerIV}),
-measures the magnetic coupling time and $t_{gw}$ measures the gravitational wave
-inspiral timescale. Before solid body rotation is attained, another relevant
-timescale is the synchronization timescale $t_s$. For a tidal torque $\tau$ and
-tidal forcing frequency $\sigma = m\p{\Omega - \Omega_{spin}}$, we note that
-angular momentum transfer is $\pd{M_{sync}}{t} \sigma R^2 = \tau$, where
-$M_{sync}$ denotes the mass of the WD that has synchronized. Thus, the
-synchronization timescale is
-\begin{align}
-    t_{sync} &\sim \frac{M_{sync}\sigma R^2}{\tau},\\
-        &\sim 2 \times 10^5\;\mathrm{yr}
-            \p{\frac{M}{M_{\odot}}}
-            \p{\frac{\sigma}{2\pi / (1\;\mathrm{hr})}}
-            \p{\frac{R}{R_{\oplus}}}^2
-            \p{\frac{10^{-14} GM_{\odot}^2/R_{\oplus}}{\tau}}.
-\end{align}
-A representative $\tau$ has been taken from \citep{bukart}. However, since only
-the outer $\sim 10^{-4}M_{\odot}$ need be heated for the thermodynamically
-interesting effects studied in \citep{fullerIV} and \citep{tidal_novae}, it
-seems that magnetic winding cannot absolutely rule out energetic outbursts such
-as tidal novae resulting from strong shear flows.
-
-\subsection{Applicability to Other Astrophysical Systems}
-
-[TODO flesh out]
-
-\begin{itemize}
-    \item $k_x, k_z$: In astrophysical systems, $k_{\perp} \ll k_r$. While we do
-        not explore different $k_{x}, k_{z}, \omega_1$ in this study, with
-        outgoing boundary conditions, there appears at first to be no $z$ length
-        scale other than $k_{z}$, so our results would seem to be invariant
-        under rescaling of the $z$ length scale. However, true turbulence is
-        expected to be isotropic at small scales, which may couple $k_{x},
-        k_{z}$ in a way that $k_{x} \ll k_{z}$ produces different dynamics
-        than $k_{x} \lesssim k_{z}$ as we've studied. This is a numerically
-        difficult regime though, so we defer consideration to future work.
-
-    \item Validity of plane-parallel approximation? We're all at $\gtrsim
-        0.9R_{WD}$.
-
-    \item Solar-type stars (inner convective, outer radiative): different
-        equation of state/stratification but could be qualitatively similar.
-
-    \item Solar-type stars: In~\cite{barker_ogilvie}, inwards-propagating IGW
-        are excited that break via geometric focusing and effect
-        synchronization. They find no reflected wave despite their nonlinear
-        timescales being $10\times$ shorter than their viscous timescale.
-
-        It is not immediately clear whether our results here apply when the flux
-        is geometrically focused, but as a hypothesis we assume the convergence
-        in \autoref{fig:agg} applies under geometric focusing as a zeroth
-        approximation, perhaps as a property of the fluid motion within the
-        geometrically thin critical layer.
-
-        Associating $t_{} \sim \nu k^2 \approx \nu k_z^2$ with the visocus
-        timescale and $t_{NL} \sim \bm{u} \cdot \bm{\nabla} \sim \omega$ for
-        the nonlinear timescale, we find their $\lambda = \frac{t_{NL}}{t_L}
-        \sim \mathrm{Re}$ our Reynolds number. Our simulations indicate
-        $\mathrm{Re} \gtrsim 500$ are required to observe the correct asymptotic
-        behavior in terms of horizontal momentum flux reflection/transmission,
-        so it is possible their lack of reflection is viscosity limited.
-\end{itemize}
-
-\subsection{Heating}
-
-[TODO elaborate? Need to make plots to check?]
-
-Our equations do not conserve energy, but it seems like more energy can be
-transmitted in higher modes as viscosity is decreased. Nevertheless, a
-significant fraction should still be dissipated in the critical layer since a
-significant energy cascade must happen in the critical layer.
-
-\section{Acknowledgements}\label{s:ack}
-
-\bibliographystyle{mnras}
-\bibliography{Su_IGW_break}
-
-\clearpage
-\onecolumn
-\appendix
-
-\section{Forcing Solution}\label{s:force_solved}
-
-To solve for the excited wave amplitude in the linear problem, consider the
-complexified linearized system of equations, then $\pd{}{t} \to -i\omega,
-\pd{}{x} \to ik_x$. Assuming all fields have dependence $u_z(x, z, t) = u_z'(z)
-e^{ik_xx - i\omega t}$, this gives:
-\begin{align*}
-    \rd{u_{z}'}{z} + ik_xu_x' &= 0,\\
-    -i\omega u_x' + ik_x \varpi' + gHik_x \Upsilon &= 0,\\
-    -i\omega u_{z}' + \rd{\varpi'}{z} + gH\rd{\Upsilon}{z}
-        - \frac{\varpi'}{H} &= 0,\\
-    -i\omega \Upsilon - \frac{u_{z}'}{H} &=
-        Ce^{-\frac{(z - z_0)^2}{2\sigma^2}}.
-\end{align*}
-We have replaced all partial derivatives with regular derivatives. This can be
-recast solely in terms of $u_{z}'$ as
-\begin{align*}
-     \rtd{u_{z}'}{z} - k_x^2u_{z}' - \frac{1}{H}\rd{u_{z}'}{z}
-        + u_{z}'\frac{N^2k_x^2}{\omega^2} &=
-    -\frac{gk_x^2}{\omega^2}Ce^{-\frac{(z - z_0)^2}{2\sigma^2}}
-        .\label{eq:narrow_inhomo}
-\end{align*}
-The homogeneous solutions are of form $u_{z,\pm}'(z) = e^{\p{\frac{1}{2H} \pm
-ik_z}\p{z - z_0}}$ where $k_z$ satisfies the dispersion relation
-\autoref{eq:disp_rel}. We compute the solution to the inhomogeneous ODE by the
-method of variation of parameters. The Wronskian can be computed
-\begin{equation}
-    W \equiv \det \begin{vmatrix}
-        u_{z,+}' & u_{z,-}' \\[4pt]
-        \rd{u_{z,+}'}{z} & \rd{u_{z,-}'}{z}
-    \end{vmatrix} = -2ik_ze^{z/H}.
-\end{equation}
-The general solution is then
-\begin{equation}
-    u_z' = -u_{z,+}\int \frac{1}{W} u_{z,-}' \p{-\frac{gk_x^2}{\omega^2}
-            Ce^{-\frac{(z - z_0)^2}{2\sigma^2}}}\;\mathrm{d}z
-        + u_{z,-}\int \frac{1}{W} u_{z,+}' \p{-\frac{gk_x^2}{\omega^2}
-            Ce^{-\frac{(z - z_0)^2}{2\sigma^2}}}\;\mathrm{d}z.
-\end{equation}
-Taking these integrals and applying boundary conditions $u_z'\p{z \to \infty} =
-u_{z,+}', u_z'\p{z \to -\infty} = u_{z,-}'$ gives exact solution
-\begin{equation}
-    u(z) = \frac{C}{2ik_z}\frac{gk_x^2}{\omega^2}
-            e^{\frac{\p{\frac{\sigma^2}{2H} \pm ik_z\sigma^2}^2}{2\sigma^2}}
-                \sqrt{2\sigma^2} \frac{\sqrt{\pi}}{2}\s{
-                    -u_+\p{1 + \erf\p{\xi}}
-                    + u_- \p{\erf\p{\xi} - 1}}
-\end{equation}
-If we are concerned with only $z$ scales significantly larger than $\sigma$,
-then we may take $\erf(\xi) \approx \Theta(\xi)$. If we further assume $k_zH \gg
-1$ and restore the $e^{ik_xx - i\omega t}$ factor, we recover the cited form
-\autoref{eq:uz_lin}
-\begin{equation}
-    u_{z}'(x, z, t) = -\frac{C}{2ik_z}\frac{gk_x^2}{\omega^2}
-        e^{-\frac{k_z^2\sigma^2}{2}}
-        \sqrt{2\pi \sigma^2} e^{ik_xx - i\omega t} \times
-    \begin{cases}
-        e^{\p{\frac{1}{2H} + ik_z}\p{z - z_0} + \frac{k_z\sigma^2}{2H}}
-            & z > z_0,\\
-        e^{\p{\frac{1}{2H} - ik_z}\p{z - z_0} - \frac{k_z\sigma^2}{2H}}
-            & z > z_0.
-    \end{cases}
-\end{equation}
-We use the plane wave approximation in our fitting routines since we generally
-fit many $\sigma$ away from $z_0$, and orthogonality with other modes is
-essential for a reliable fit.
 
@@
 further change variables to $\Upsilon = \ln \rho - \ln \overline{\rho}$ and
-$\varpi' = \varpi - \overline{\varpi}$ deviations from the background state to obtain a
-system of equations at most quadratic in fluid fields:
+$\varpi' = \varpi - \overline{\varpi}$ deviations from the background state to
+obtain a system of equations at most quadratic in fluid fields:
 \begin{subequations}\label{se:nl_var}
     \begin{align}
-        \bm{\nabla} \cdot \bm{u}' &= 0,\\
-        \pd{\Upsilon}{t} + \p{\bm{u}' \cdot \bm{\nabla}} \Upsilon
+        \bm{\nabla} \cdot \bm{u} &= 0,\\
+        \pd{\Upsilon}{t} + \p{\bm{u} \cdot \bm{\nabla}} \Upsilon
             - \frac{u_z}{H} &= 0,\\
-        \pd{u_{x}}{t} + \p{\bm{u}' \cdot \bm{\nabla}}u_{x}
+        \pd{u_{x}}{t} + \p{\bm{u} \cdot \bm{\nabla}}u_{x}
             + \pd{\varpi'}{x} + gH\pd{\Upsilon}{x}
             + \varpi' \pd{\Upsilon}{x} &= 0,\\
-        \pd{u_z}{t} + \p{\bm{u}' \cdot \bm{\nabla}}u_z
+        \pd{u_z}{t} + \p{\bm{u} \cdot \bm{\nabla}}u_z
             + \pd{\varpi'}{z} + gH\pd{\Upsilon}{z}
@@
     \begin{align}
-        \bm{\nabla} \cdot \bm{u}' &= 0,\\
-        \partial_t \rho + \bm{\nabla} \cdot (\rho \bm{u}' - \nu
+        \bm{\nabla} \cdot \bm{u} &= 0,\\
+        \partial_t \rho + \bm{\nabla} \cdot (\rho \bm{u} - \nu
             \bm{\nabla}(\rho - \overline{\rho})) &= 0,\label{eq:visc_cons_mom}\\
-        \partial_t (\rho \bm{u}') + \bm{\nabla} \cdot (\rho \bm{u}' \bm{u}'
+        \partial_t (\rho \bm{u}) + \bm{\nabla} \cdot (\rho \bm{u} \bm{u}
             + \mathrm{diag}(\rho \varpi)
@@
 Including the damping layers and forcing terms as described in
-\autoref{ss:numerics}, we finally obtain the full system of equations as
+\autoref{s:numerics}, we finally obtain the full system of equations as
 simulated in Dedalus:
@@
     \begin{align}
-        \bm{\nabla} \cdot \bm{u}' ={}& 0,\\
+        \bm{\nabla} \cdot \bm{u} ={}& 0,\\
         \partial_t \Upsilon - \frac{u_z}{H}
@@
                     \cos \p{k_xx - \omega t}\nonumber\\
-            & + \Gamma_{NL} \bigg[-\p{\bm{u}' \cdot \bm{\nabla}}\Upsilon
+            & + \Gamma_{NL} \bigg[-\p{\bm{u} \cdot \bm{\nabla}}\Upsilon
                 + \nu\p{\nabla^2 \Upsilon + \p{
@@
                 - \frac{1}{H}\partial_z u_x}
-                - \p{\bm{u}' \cdot \bm{\nabla}}u_x
+                - \p{\bm{u} \cdot \bm{\nabla}}u_x
                 - \varpi' \pd{\Upsilon}{x}\bigg],\\
@@
                 \frac{1}{H}\partial_z u_{z}}
-            - \p{\bm{u}' \cdot \bm{\nabla}}u_z
+            - \p{\bm{u} \cdot \bm{\nabla}}u_z
             - \varpi' \pd{\Upsilon}{z}\bigg].

### 2019-11-13 sims/2d_4_fourier/strat.py

@@ -66,13 +66,13 @@ def get_params(overrides=None):
 
 if __name__ == '__main__':
     tasks = [
-       ('lin_0_masked',
-        get_params(overrides={'F_MULT': 0.0005,
-                              'T_F': 4000,
-                              'DT': 0.1,
-                              'N_X': 64,
-                              'N_Z': 256,
-                              'Re_inv': 1e-4})),
+        ('lin_0_masked',
+         get_params(overrides={'F_MULT': 0.0005,
+                               'T_F': 4000,
+                               'DT': 0.1,
+                               'N_X': 64,
+                               'N_Z': 256,
+                               'Re_inv': 1e-4})),
         # ('yubo_nu0p5_hres',
         #  get_params(overrides={'Re_inv': 0.05})),
         # ('yubo_nu0p5_shres',

### 2019-11-13 sims/2d_4_fourier/strat_helper.py

@@ -1052,7 +1052,7 @@ def plot_front(name, params):
         #          linewidth=LW * 0.7)
         # ax1.legend(fontsize=FONTSIZE)
         ax1.set_xlabel(r'$Nt$')
-        ax1.set_ylabel(r'$\hat{A}_i$')
+        ax1.set_ylabel(r'$A_i$')
         # ax2.plot(t,
         #          np.unwrap(phis_down[start_idx: ]),
         #          'r',
@@ -1077,7 +1077,7 @@ def plot_front(name, params):
         f, (ax1, ax2) = plt.subplots(2, 1, sharex=True)
         # z0_cut = z0[z_b_idx: ]
 
-        for time, color in zip(times, PLT_COLORS):
+        for time, color in zip(times[::-1], PLT_COLORS):
             u0_avg = np.mean(u0[time - AVG_IDX: time + AVG_IDX,
                                 # z_b_idx: ], axis=0)
                                 : ], axis=0)
@@ -1192,7 +1192,7 @@ def plot_front(name, params):
         ax2.plot(t,
                  pos_anal,
                  'b',
-                 label="Avg Absorbed $%.2fF_{al}'$" %
+                 label="$F_a = %.2fF_{al}$" %
                     (mean_incident / flux_th),
                  linewidth=LW * 0.7)
         ax2.set_ylabel(r'$z_c$')
@@ -1208,18 +1208,18 @@ def plot_front(name, params):
         #
         # convolved amplitudes over time
         #####################################################################
-        f, (ax1, ax2) = plt.subplots(2, 1, sharex=True)
+        f, (ax1, ax2) = plt.subplots(2, 1, figsize=(6, 9), sharex=True)
         ax1.plot(t,
                 smooth(amps[start_idx::]),
                 'g',
-                label=r'$\hat{A}_i(t)$',
+                label=r'$A_i(t)$',
                 linewidth=LW * 0.7)
         ax1.plot(t,
                  smooth(amps_down[start_idx::]),
                  'r',
-                 label=r'$\hat{A}_r(t)$',
+                 label=r'$A_r(t)$',
                  linewidth=LW * 0.7)
-        ax1.set_ylabel(r'$\hat{A}$')
+        ax1.set_ylabel(r'$A$')
         ax1.set_ylim(bottom=0)
[…]

### 2019-12-22 writeups/Su_IGW_break.tex
check in draft 191222

@@
     \newcommand*{\ptd}[2]{\frac{\partial^2#1}{\partial#2^2}}
+    % inline
+    \newcommand*{\mdil}[2]{\mathrm{D}#1/\mathrm{D}#2}
+    \newcommand*{\pdil}[2]{\partial#1/\partial#2}
+    \newcommand*{\rdil}[2]{\mathrm{d}#1/\mathrm{d}#2}
     \newcommand*{\md}[2]{\frac{\mathrm{D}#1}{\mathrm{D}#2}}
@@
 
-\title[Internal Gravity Wave Breaking in White Dwarf Binaries]{Internal Gravity
-Wave Breaking in White Dwarf Binaries}
+\title[Internal Gravity Wave Breaking]{Physics of Tidal Dissipation:
+Hydrodynamical Simulations of Internal Gravity Wave Breaking in Stellar
+Envelopes and Massive Stellar Binaries}
 \author[Y. Su et\ al.]{
@@
 \pagerange{\pageref{firstpage}--\pageref{lastpage}}
-\renewcommand*{\sectionautorefname}{Section}
 \maketitle
 
-% cp ../sims/agg.png plots; cp ../sims/2d_4_fourier/snapshots_lin_0_masked/f_amps.png plots/lin_amps.png; cp ../sims/2d_4_fourier/snapshots_lin_0_masked/fluxes.png plots/lin_fluxes.png; cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/f_amps.png plots/nl_f_amps.png; cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/fluxes.png plots/nl_fluxes.png; cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/f_refl.png plots/nl_f_refl.png; cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/f_ri.png plots/nl_f_ri.png; cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/front.png plots/nl_front.png;
+% cp ../sims/agg.png plots; cp ../sims/2d_4_fourier/snapshots_lin_0_masked/f_amps.png plots/lin_amps.png; cp ../sims/2d_4_fourier/snapshots_lin_0_masked/fluxes.png plots/lin_fluxes.png; cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/f_amps2.png plots/nl_f_amps2.png; cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/f_amps.png plots/nl_f_amps.png; cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/fluxes.png plots/nl_fluxes.png; cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/f_refl.png plots/nl_f_refl.png; cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/f_ri.png plots/nl_f_ri.png; cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/front.png plots/nl_front.png;
 
@@
 
-\section{Introduction}
+\section{Introduction}\label{s:intro}
 
@@
 \citep{tidal_novae}. Most importantly, they have been proposed as the likely
-progenitors of type Ia supernovae (e.g. \citealp{Ia0,webbink} or more recently
-\citealp{Ia1,Ia2}). While presently only a few tens of compact WD binaries are
+progenitors of type Ia supernovae (e.g. \citet{Ia0,webbink} or more recently
+\citet{Ia1,Ia2}). While presently only a few tens of compact WD binaries are
 known \citep{lsst_wd}, \emph{Gaia} (currently gathering data) is expected to
@@
 
-These two zones are separated by a \emph{critical layer}. The majority of our
-paper is dedicated to characterizing the behavior of the critical layer when
+These two zones are separated by a \emph{critical layer}. The interaction of IGW
+and critical layers plays an important role in the tidal synchronization process
+of stellar binaries \citep{zahn75,gn89} as well as in terrestrial phenomena such
+as the quasi-biennial oscillation \citep{lindzen_qbo}. The majority of our paper
+is dedicated to characterizing the behavior of the critical layer when
 interacting with a continuous train of IGW excited from the bottom of the fluid.
@@
 propagation of the critical layer, reflection of the incident IGW, and other
-momentum redistribution behaviors. Our results are numerically converged
-for sufficiently large resolutions.
-
-In \autoref{s:equations}, we will describe the system of equations we will use
-to analyze IGW breaking. In \autoref{s:theory}, we review existing understanding
-of wave breaking and present new analytical results. In \autoref{s:numerics}, we
-describe our numerical setup, which in \autoref{s:weak_sim} we validate in the
-weak forcing limit against linear theory. In \autoref{s:sim}, we present the
-results of simulation of IGW breaking and our characterization of the critical
-layer. Finally, we summarize and conclude in \autoref{s:discussion}.
-
-\section{Problem Description}\label{s:equations}
-
-We consider a incompressible, isothermal fluid, representative of degenerate
-matter in WDs. We use barotropic equation of state $P(\rho, T) = P(\rho)$ as a
-first approximation. As we are interested in dynamics far from the center of the
-WD, we approximate the gravitational field as uniform. We model the background
-density stratification as $\overline{\rho} = \overline{\rho}_0 e^{-z/H}$ for
-some reference density $\overline{\rho}_0$ (we generally notate background
-quantities with overbars and perturbation quantities with primes). We
-consider a 2D fluid with for computational feasibility. While
-it is well-known wave breaking is a 3D process \citep{klostermeyer,winters1994},
-the dynamical effect of the breaking process is likely to be similar in 2D
-\citep{barker_ogilvie}. Finally, we simplify by considering simple
-plane-parallel IGW\@. Since wave breaking generally occurs near the surface of
-WDs \citep[$\gtrsim 0.9R_{WD}$][]{fullerI}, the curvature of the wave front is
-negligible.
+momentum redistribution behaviors. Our results are numerically converged for
+sufficiently large resolutions.
+
+In Section~\ref{s:equations}, we will describe the system of equations we will
+use to analyze IGW breaking. In Section~\ref{s:theory}, we review existing
+understanding of wave breaking and present new analytical results. In
+Section~\ref{s:numerics}, we describe our numerical setup, which in
+Section~\ref{s:weak_sim} we validate in the weak forcing limit against linear
+theory. In Section~\ref{s:sim}, we present the results of simulation of IGW
+breaking and our characterization of the critical layer. Finally, we summarize
+and conclude in Section~\ref{s:discussion}.
+
+\section{Problem Setup}\label{s:equations}
+
+We consider a stratified, incompressible, isothermal fluid representing a
+stellar envelope or atmosphere. We use barotropic equation of state $P(\rho, T)
+= P(\rho)$ as a first approximation. We study dynamics in 2D, so that fluid
+variables depend only on the Cartesian coordinates $x$ and $z$. While it is
+well-known wave breaking is a 3D process \citep{klostermeyer,winters1994}, the
+dynamical effect of the breaking process is likely to be similar in 2D
+\citep{barker_ogilvie}. We approximate the gravitational field as uniform,
+pointing in the $(-\uv{z})$ direction. The plane-parallel approximation is
+justified since wave breaking generally occurs near the stellar surface. The
+background density stratification is given by
+\begin{equation}
+    \overline{\rho} = \overline{\rho}_0 e^{-z/H},
+\end{equation}
+with $\overline{\rho}_0$ some reference density (we generally denote background
+quantities with overbars and perturbation quantities with primes).
 
@@
         \md{\bm{u}}{t} + \frac{\bm{\nabla}P}{\rho} + g\uv{z} &=
-            0\label{eq:nl_mom}.
+            0\label{eq:nl_mom},
     \end{align}
 \end{subequations}
-$\md{}{t} = \pd{}{t} + \p{\bm{u} \cdot \bm{\nabla}}$ is the Lagrangian or
-material derivative, and $\bm{u}, \rho, P$ denote the velocity field, density
-and pressure respectively. We denote $-g\uv{z}$ constant gravitational
-acceleration. Note that at hydrostatic equilibrium $\pd{}{t} = 0$ we have
-$\bm{\nabla}\overline{P} = -\overline{\rho} g\uv{z}$ and so $\overline{P} =
-\overline{\rho} gH$. A vertically-stratified shear flow
-$\overline{u}_x(z)\uv{x}$ is permitted at hydrostatic equilibrium, but we will
-assume no background flow, so $\bm{u} = \bm{u}'$. Physically, this assumption
-corresponds to a non-rotating star, or going to the corotating frame of a
-rigidly rotating star.
-
-In practice, it is convenient to introduce the coordinate $\Upsilon = \ln
-\frac{\rho}{\overline{\rho}}$ \citep[e.g.][]{lecoanet_anel}. This both
-identically enforces $\rho > 0$ and eliminates the stiff term
-$\frac{\bm{\nabla}P}{\rho}$. We also define reduced pressure $\varpi =
-\frac{P}{\rho}$. Then, we may rewrite the second two equations of
-\autoref{se:nl_orig} as
+where $\mdil{}{t} = \pdil{}{t}\,+\,\p{\bm{u} \cdot \bm{\nabla}}$ is the
+Lagrangian or material derivative, and $\bm{u}, \rho, P$ denote the velocity
+field, density and pressure respectively. We denote $-g\uv{z}$ constant
+gravitational acceleration. Note that hydrostatic equilibrium ($\pdil{}{t} = 0$)
+of the background implies $\bm{\nabla}\overline{P} = -\overline{\rho} g\uv{z}$,
+so that $\overline{P} = \overline{\rho} gH$. A vertically-stratified shear flow
+$\overline{u}_x(z)\uv{x}$ is permitted in hydrostatic equilibrium, but we will
+assume no such background flow, so $\bm{u} = \bm{u}'$. Physically, this
+assumption corresponds to a non-rotating star, or going to the corotating frame
+of a rigidly rotating star.
+
+For convenience, we introduce the dimensionless density variable $\Upsilon$ and
+the reduced pressure $\varpi$ \citep[e.g.][]{lecoanet_anel} via
+\begin{align}
+    \Upsilon &\equiv \ln \frac{\rho}{\bar{\rho}},\\
+    \varpi &\equiv \frac{P}{\rho}.
+\end{align}
+These variables automatically enforce $\rho > 0$ and eliminate the stiff term
+$\vec{\nabla} P / \rho$ in the Euler equation. In terms of $\Upsilon$ and
+$\varpi$, the second two equations in~\eqref{se:nl_orig} become
 \begin{subequations}\label{se:nl_upsilon}
@@
 \end{subequations}
-In the new coordinates, hydrostatic equilibrium corresponds to $\Upsilon = 0,
-\overline{\varpi} = gH$.
+Hydrostatic equilibrium corresponds to $\Upsilon = 0, \overline{\varpi} = gH$.
 
@@
 
-\subsection{Linear Analysis}
-
-In the small perturbation limit, where flow velocities are small compared to the
-characteristic space and time scales $\pd{}{t} \gg \bm{u} \cdot
-\bm{\nabla}$, we may linearize \autoref{se:nl_upsilon}. The solution in this
-linear regime is given by \citep{drazin,sutherland0}:
-\begin{equation}
-    u_z\p{x, z, t} = Ae^{z/2H}\cos\p{k_{x}x + k_{z}z - \omega t},
-        \label{eq:k0z_sign}
-\end{equation}
-where $A$ is the amplitude and $\omega$ satisfies dispersion relation
-\begin{equation}
-    \omega^2 = \frac{N^2k_{x}^2}{k_{x}^2 + k_{z}^2 + \frac{1}{4H^2}}.
+\subsection{Linear Analysis}\label{ss:lin_analysis}
+
+In the small perturbation limit, we may linearize
+Eqs.~\eqref{se:nl_upsilon}. The solution in this linear regime is given by
+\citep{drazin,sutherland0}
+\begin{equation}
+    u_z'\p{x, z, t} = Ae^{z/2H}\cos\p{k_{x}x + k_{z}z - \omega t},
+        \label{eq:lin_sol}
+\end{equation}
+where $A$ is a constant amplitude, and the frequency $\omega$ and the wave
+number $\p{k_x, k_z}$ satisfy the dispersion relation
+\begin{equation}
+    \omega^2 = \frac{N^2k_{x}^2}{k_{x}^2 + k_{z}^2 + \p{2H}^{-2}}.
         \label{eq:disp_rel}
 \end{equation}
-Our equations are valid in the limit of large sound speed $c_s \to \infty$, in
-which the \emph{Brunt-V\"ais\"al\"a frequency}
+Our equations are valid in the limit of large sound speed ($c_s \to \infty$), in
+which the \emph{Brunt-V\"ais\"al\"a frequency} $N$ is given by
 \begin{equation}
@@
 \end{equation}
-is constant. Other dynamical quantities are simply related to $u'_z$ \citep[see
-e.g.][]{sutherland0}.
-
-In the short-wavelength/WKB limit $\abs{k_{z}H} \gg 1$, the solution exhibits
+is constant. Other dynamical quantities are simply related to $u'_z$.
+
+In the short-wavelength/WKB limit ($\abs{k_{z}H} \gg 1$), the solution exhibits
 the following characteristics:
-\begin{itemize}
+\begin{enumerate}
     \item The amplitude of the wave grows with $z$ as $e^{z/2H}$. Thus, the
-        linear approximation is always violated for sufficiently large $z$.
-
-    \item The phase and group velocities are respectively:
+        linear approximation always breaks down for sufficiently large $z$.
+
+    \item The phase and group velocities are given by:
         \begin{align}
@@
                 \p{k_{x}\uv{x} + k_{z}\uv{z}}\frac{\omega}
-                {k_{x}^2 + k_{z}^2 + \frac{1}{4H^2}},\\
-            \bm{c}_{g} &= N\frac{\p{k_{z}^2 + \frac{1}{4H^2}}\uv{x}
+                {k_{x}^2 + k_{z}^2 + \p{2H}^{-2}},\\
+            \bm{c}_{g} &= N\frac{\s{k_{z}^2 + \p{2H}^{-2}}\uv{x}
                 - \p{k_{x}k_{z}\uv{z}}}
-                {\p{k_{x}^2 + k_{z}^2 + \frac{1}{4H^2}}^{3/2}}.\label{eq:vg}
-        \end{align}
-        We note $\bm{c}_{p} \cdot \bm{c}_g = \mathcal{O}\p{ \p{k_{z}H}^{-2} }
-        \approx 0$. In the Boussinesq approximation where density stratification
-        outside of $N^2$ is completely neglected, the phase and group velocity
-        are exactly orthogonal \citep{drazin,sutherland1}. We use the convention
+                {\s{k_{x}^2 + k_{z}^2 + \p{2H}^{-2}}^{3/2}}.\label{eq:vg}
+        \end{align}
+        We note $\bm{c}_{p} \cdot \bm{c}_g = \mathcal{O}\s{\p{k_{z}H}^{-2}}
+        \approx 0$. In the Boussinesq approximation where terms
+        $\mathcal{O}\p{H^{-2}}$ are ignored, the phase and group velocities are
+        exactly orthogonal \citep{drazin,sutherland1}. We use the convention
         where upward propagating IGW have $c_{g, z} > 0$, $k_z < 0, k_x > 0$.
 
-    \item The averaged horizontal momentum flux $F$ is
-        \begin{align}
-            F &\equiv \ev{\rho u_{x} u_{z}}_x \equiv
-                \frac{1}{L_x}\int_0^{L_x}\limits \rho u_{x}u_{z}\;\mathrm{d}x,
-                    \label{eq:F_def}\\
-                &\approx -\frac{A^2}{2}\overline{\rho}_0\frac{k_{z}}{k_{x}},
+    \item The averaged horizontal momentum flux $F$ (in the $+\uv{z}$
+        direction) carried by the IGW is defined by
+        \begin{equation}
+            F(z, t) \equiv \ev{\rho u_{x}' u_{z}'}_x \equiv
+                \frac{1}{L_x}\int_0^{L_x}\limits \rho u_{x}'u_{z}'\;\mathrm{d}x.
+                    \label{eq:F_def}
+        \end{equation}
+        The notation $\ev{\dots}_x$ denotes averaging over the $x$ domain. For
+        the linear solution (Eq.~\eqref{eq:lin_sol}), this evaluates to
+        \begin{equation}
+            F \approx -\frac{A^2}{2}\overline{\rho}_0\frac{k_{z}}{k_{x}},
                     \label{eq:S_lin}
-        \end{align}
-        Thus, indeed $F > 0$ for an upward propagating IGW $c_{g, z} > 0$.
-\end{itemize}
+        \end{equation}
+        Thus, indeed $F > 0$ for an upward propagating IGW ($c_{g, z} > 0$).
+\end{enumerate}
 
@@
 
-To model continuous excitation of IGWs deep in the WD interior propagating
+To model continuous excitation of IGWs deep in the stellar envelope propagating
 towards the surface, we use a volumetric forcing term to excite IGW near the
 bottom of the simulation domain. Our forcing excites both IGWs
-propagating upwards, imitating a wave tidally excited deeper in the WD, and
-downwards, which are not physically meaningful, but are damped away by the
-damping layers described in \autoref{ss:damping}.
+propagating upwards, imitating a wave tidally excited deeper in the star, and
+downwards, which are not physically meaningful, but are dissipated by a
+damping zone described in Section~\ref{ss:damping}.
 
@@
 on the density equation. We implement forcing with strength $C$ localized around
-$z_0$ with small width $\sigma$ by replacing \autoref{eq:nl_up_density} with
+$z_0$ with small width $\sigma$ by replacing Eq.~\eqref{eq:nl_up_density}
+with
 \begin{equation}
@@
 Using a narrow Gaussian profile excites a broad $z$ power spectrum, but only the
-$k_{z}$ satisfying dispersion relation \autoref{eq:disp_rel} for the given
-$k_{x}$, $\omega(k_{x}, k_{z})$ will propagate.
+$k_{z}$ satisfying dispersion relation (Eq.~\eqref{eq:disp_rel}) for the given
+$k_{x}$ and $\omega$ will propagate.
 
 In the linearized system, the effect of this forcing can be solved exactly
-(\autoref{s:force_solved}). If we approximate $k_zH \gg 1, \sigma \ll H$, the
-solution can be approximated as two plane waves propagating away from the
-forcing zone
+(see Appendix~\ref{s:force_solved}). If we approximate $\abs{k_zH} \gg 1, \sigma
+\ll H$, the solution can be approximated as two plane waves propagating away
+from the forcing zone
 \begin{align}
     u_{z}&(x, z, t) \approx{} \frac{C}{2k_z}\frac{gk_x^2}{\omega^2}
-        e^{-\frac{k_z^2\sigma^2}{2}}
-        \sqrt{2\pi \sigma^2} \times\nonumber\\
-        &{}\begin{cases}
+        \exp\p{-\frac{k_z^2\sigma^2}{2}}
+        \sqrt{2\pi \sigma^2} \nonumber\\
+        &{}\times\begin{cases}
         e^{\frac{z - z_0}{2H}}\sin\p{k_{x}x + k_{z}(z - z_0) - \omega t
             + \frac{k_z\sigma^2}{2H}}
-            & z > z_0\\
+            & \text{for }z > z_0,\\[5pt]
         e^{\frac{z - z_0}{2H}}\sin\p{k_{x}x - k_{z}(z - z_0) - \omega t
             + \frac{k_z\sigma^2}{2H}}
-            & z < z_0\\
-    \end{cases}.\label{eq:uz_lin}
-\end{align}
-The $z > z_0$ region models an upward propagating IGW wavetrain excited deep
-in the atmosphere.
+            & \text{for }z < z_0.\\
+    \end{cases}\label{eq:uz_lin}
+\end{align}
+The $z > z_0$ region models an upward propagating IGW wavetrain. The $x$
+component of the velocity can be obtained by the incompressibility constraint
+(Eq.~\eqref{eq:nl_incomp}).
 
@@
 
-Above, we neglected advective terms $\pd{}{t} \gg \bm{u} \cdot \bm{\nabla}$.
-However, since $\abs{\bm{u}} \propto e^{z/2H}$ in an infinite domain all IGWs
-eventually grow to nonlinear amplitudes and break. Nonlinear wave breaking is
-expected to be important in WDs \citep{fullerI,fullerII}. We can
-estimate the height of wave breaking using $\frac{\abs{\bm{u}}
-\abs{\bm{k}}}{\omega} \sim 1$. This can be rewritten using the Lagrangian
-displacement $-i\omega \bm{\xi} = \bm{u}$:
-\begin{equation}
-    \xi_z k_z \gtrsim 1.\label{eq:nl}
-\end{equation}
-
-% An equivalent estimate given by \citep{eliassen_palm_cite,sutherland0} is that
-% IGWs break when the wave-induced mean flow is equal to the horizontal phase
-% velocity
-% \begin{equation}
-%     \overline{U}(z) = \frac{\ev{u_{x}u_{z}}_x}{c_{g,z}} \sim
-%         \frac{\omega}{k_x},\label{eq:rip}
-% \end{equation}
-% where \overline{U}(z) \equiv \ev{u_x}_x$ is the mean flow of the fluid.
-% \autoref{eq:rip} provides a leading-order estimate for \overline{U}(z)$ which
-% may further evolve over time (see \autoref{ss:crit_layer}); for a general mean
-% flow, \autoref{eq:rip} also defines the criterion for a \emph{critical
-% layer}. This wave-induced mean flow is analogous to Stokes' drift for surface
-% waves and can be derived by considering the propagation of $F$
-% (\autoref{eq:F_def}) into a medium at rest. Evaluating \autoref{eq:rip}
-% yields agreement with \autoref{eq:nl}.
-
-\citealp{drazin, klostermeyer, winters1994} describe the onset of wave breaking.
-At intermediate amplitudes, wave breaking occurs via triadic resonances which
-transfer energy from the ``parent’’ internal gravity wave (in our problem, the
-tidally forced wave) to ``daughter’’ internal gravity waves on smaller length
-scales that efficiently damp. The horizontal momentum flux decreases from $F$ to
-$0$ over this breaking region. The lost flux is deposited into a horizontal mean
-flow
+As the upward propagating IGW grows in amplitude ($\abs{\bm{u}} \propto
+e^{z/2H}$), it is expected to break due to nonlinear effects. We can estimate
+the height of wave breaking as when $\abs{\bm{u}} \sim \omega / \abs{\bm{k}}$.
+This can be rewritten using the Lagrangian displacement $\bm{\xi} = \bm{u} /
+\p{-i\omega}$:
+\begin{equation}
+    \abs{\xi_z k_z} \gtrsim 1.\label{eq:nl}
+\end{equation}
+
+\citet{drazin, klostermeyer, winters1994} describe the onset of wave breaking in
+some detail. At intermediate amplitudes, wave breaking occurs via triadic
+resonances which transfer energy from the ``parent’’ IGW to ``daughter’’ waves
+on smaller length scales that efficiently damp. The horizontal momentum flux
+decreases from $F$ to $0$ over this breaking region. The lost flux is deposited
+into a horizontal mean flow
 \begin{equation}
@@
 \end{equation}
-As the mean flow grows, a \emph{critical layer} may form, discussed below.
+As the mean flow grows, a \emph{critical layer} may form, as discussed below.
 
@@
 
-A horizontal shear flow $\overline{U}(z, t)\hat{x}$ enters the fluid equations
+A horizontal shear flow $\overline{U}(z, t)\uv{x}$ enters the fluid equations
 via the Lagrangian derivative, which can be decomposed as
@@
 \end{equation}
-where $\bm{u}'$ is the velocity field \emph{without} the shear flow, for
-clarity. Thus, $\overline{U}$ has the effect of Doppler shifting the time
-derivative into the frame comoving with the mean flow. If $\overline{U}$ is
-roughly constant, then the behavior of a small plane wave perturbation satisfies
-modified dispersion relation
+where $\bm{u}'$ is the velocity field \emph{without} the shear flow. Thus,
+$\overline{U}$ has the effect of Doppler shifting the time derivative into the
+frame comoving with the mean flow. If $\overline{U}$ is roughly constant, then
+the behavior of a linear plane-wave perturbation satisfies the modified
+dispersion relation
 \begin{equation}
     \p{\omega - \overline{U}k_x}^2 =
-        \frac{N^2k_{x}^2}{k_{x}^2 + k_{z}^2 + \frac{1}{4H^2}}.
+        \frac{N^2k_{x}^2}{k_{x}^2 + k_{z}^2 + \p{2H}^{-2}}.
         \label{eq:disp_rel_U}
 \end{equation}
-This is just \autoref{eq:disp_rel} with $\omega \to \omega - \overline{U}k_x$.
-It is apparent that if $\overline{U} = \overline{U}_c$, where
-\begin{equation}
-    \overline{U}_c \equiv \frac{\omega}{k_x},
+This is just Eq.~\eqref{eq:disp_rel} with $\omega \to \omega - \overline{U}
+k_x$. It is apparent that if $\overline{U} = \overline{U}_c$, where
+\begin{equation}
+    \overline{U}_c \equiv \frac{\omega}{k_x},\label{eq:u_crit}
 \end{equation}
@@
 The behavior of an IGW incident upon a critical layer was first studied in the
-inviscid, linear regime in \citealp{booker_bretherton}, which found nearly
-complete absorption of the IGW\@. In particular, the incident wave has amplitude
-reflection and transmission coefficients that can be expressed in terms of the
-local Richardson number Ri at the height of the critical layer $z_c$:
-\begin{align}
-    \mathrm{Ri} &\equiv \at{\frac{N^2}{\p{\pd{\overline{U}}{z}}^2}}_{z_c},
-        \label{eq:ri_def}\\
-    \mathcal{R}_A &= e^{-2\pi \sqrt{\mathrm{Ri} - \frac{1}{4}}}, &
-    \mathcal{T}_A &= e^{-\pi \sqrt{\mathrm{Ri} - \frac{1}{4}}}.
+inviscid, linear regime in \citep{booker_bretherton}, which found nearly
+complete absorption of the IGW\@. The amplitude
+reflection and transmission coefficients are given by
+\begin{align}
+    \mathcal{R} &= \exp\p{-2\pi \sqrt{\mathrm{Ri} - \frac{1}{4}}}, &
+    \mathcal{T} &= \exp\p{-\pi \sqrt{\mathrm{Ri} - \frac{1}{4}}},
         \label{eq:crit_coeffs}
 \end{align}
-In the $\mathrm{Ri} \gg 1$ limit, $\mathcal{R}, \mathcal{T} \ll 1$
-and the incident wave is completely absorbed to a good approximation.
-
-This result extends to viscous fluids \citep{hazel}, but weakly nonlinear theory
-\citep{brown_stewartson} and numerical simulations \citep{winters1994} suggest
-that nonlinear effects may significantly enhance reflection and transmission.
-Understanding the interaction of IGW and critical layers has important
-applications: it is thought to be responsible for tidal synchronization in
-stellar binaries \citep{zahn75,gn89} as well as terrestrial phenomena such as
-the quasi-biennial oscillation \citep{lindzen_qbo}.
-
-Consider the horizontal momentum transfer at the critical layer. Any incident
-horizontal momentum flux absorbed by the fluid must manifest as additional
-horizontal momentum of the shear flow. Since $\overline{U}$ cannot exceed
-$\overline{U}_c$, the critical layer will instead propagate downward in response
-to the incident momentum flux. The horizontal momentum of the shear flow
-satisfies
+where $\mathrm{Ri}$ is the local Richardson number evaluated at the critical
+layer height $z_c$:
+\begin{equation}
+    \mathrm{Ri} \equiv \at{\frac{N^2}{\p{\pdil{\overline{U}}{z}}^2}}_{z_c}.
+        \label{eq:ri_def}
+\end{equation}
+In the $\mathrm{Ri} \gg 1$ limit, $\mathcal{R}, \mathcal{T} \ll 1$ and the
+incident wave is almost completely absorbed. This result also applies to viscous
+fluids \citep{hazel}. However, weakly nonlinear theory \citep{brown_stewartson}
+and numerical simulations \citep{winters1994} suggest that nonlinear effects may
+significantly enhance reflection and transmission.
+
+Instead, we consider the horizontal momentum transfer at the critical layer. Any
+incident horizontal momentum flux $F_a(t)$ absorbed by the fluid must manifest
+as additional horizontal momentum of the shear flow. Since $\overline{U}$ cannot
+exceed $\overline{U}_c$, the critical layer will instead propagate downward in
+response to the incident momentum flux. The horizontal momentum of the shear
+flow satisfies
 \begin{equation}
     \pd{}{t}\int\limits \overline{\rho}(z) \overline{U}(z, t)\;\mathrm{d}z
-        - F_a(t) = 0,
-\end{equation}
-where $F_a(t) > 0$ is the absorbed horizontal momentum. Assuming $\overline{U}(z
-> z_c) = \overline{U}_c, \overline{U}(z < z_c) = 0$, this condition becomes
-\begin{equation}
-    \overline{\rho}(z_c) \overline{U}_c\rd{z_c}{t} = F_a(t).\label{eq:zc_anal}
-\end{equation}
-If $F_a$ is constant in time, $z_c(t)$ has analytical solution
-\begin{equation}
-    z_c(t) = -H\ln \s{e^{-\frac{z_c(t = 0)}{H}} +
+        - F_a(t) = 0.
+\end{equation}
+Assuming $\overline{U}(z > z_c) \approx \overline{U}_c$ and $\overline{U}(z <
+z_c) \approx 0$ are both approximately constant, this condition becomes
+\begin{equation}
+    -\overline{\rho}(z_c) \overline{U}_c\rd{z_c}{t} = F_a(t).\label{eq:zc_anal}
+\end{equation}
+If $F_a$ is constant in time, $z_c(t)$ has as analytical solution
+\begin{equation}
+    z_c(t) = -H\ln \s{\exp\p{-\frac{z_c(t = 0)}{H}} +
         \frac{tF_a}{\overline{U}_c H\overline{\rho}_0}},\label{eq:zc_sol}
@@
 
-We verify the analytic theory of \autoref{s:theory} by running weakly forced
-direct numerical simulations using the pseudo-spectral code Dedalus
-\citep{dedalus,dedalus2}. IGWs are excited deep within WDs and must propagate
-across many density scale heights before reaching the surface. The waves stay
-linear until they are within a few scale heights of the breaking height. We thus
-simulate only the upper part of the WD\@. Although turbulence driven by wave
-breaking is fundamentally 3D \citep{klostermeyer}, we simplify the problem by
-restricting our simulations to two dimensions. Future work will study the
-problem in 3D spherical geometry.
-
-\subsection{Parameter Choices}
-
-We simulate equations \autoref{eq:nl_incomp}, \autoref{eq:nl_upsilon_u}, and
-\autoref{eq:vol_drive} in a Cartesian box with size $L_x, L_z$. The $x$
-direction represents the azimuthal direction in a star, and the $z$ direction
-represents the radial direction. We thus choose periodic boundary conditions in
-the $x$ direction. We also use periodic boundary conditions in the $z$
-direction, but we damp perturbations to zero near the top/bottom of the domain
-to capture their distinct behavior (described in \autoref{ss:damping}). We
-expand all variables as Fourier series with $N_x$ and $N_z$ modes, and use the
-$3/2$ dealiasing rule to avoid aliasing errors in the nonlinear terms
-\citep{boyd}.
-
-The geometry of our simulation domain is fixed by one further parameter: $z_0$
-the forcing location. We choose $L_z = 12.5H$ to give $\sim e^3$ amplitude growth
-between the damping zones. We force at $z_0 = 3H$ within width $\sigma =
-0.078H$, sufficiently far from the lower damping zone and permit sufficient room
-for the upward moving wave to grow as $\propto e^{z/2H}$. Finally, we want
-similar grid spacing $\frac{L_x}{N_x} \sim \frac{L_z}{N_z}$, guided by the
-intuition that turbulence is approximately isotropic, so we use $L_x = 4H$ and
-$N_z / N_x = 4$.
+We use the pseudo-spectral code Dedalus \citep{dedalus,dedalus2} to simulate
+the excitation and propagation of IGWs (Section~\ref{s:weak_sim}) as well as
+their nonlinear breaking and the formation of a critical layer
+(Section~\ref{s:sim}).
+
+\subsection{Parameter Choices}\label{ss:params}
+
+We solve Eqs.~\eqref{eq:nl_incomp},~\eqref{eq:nl_upsilon_u},
+and~\eqref{eq:vol_drive} in a Cartesian box with size $L_x, L_z$. We choose
+periodic boundary conditions in both the $x$ and $z$ direction. We damp
+perturbations to zero near the top/bottom of the domain using damping zones
+(see Section~\ref{ss:damping}). We expand all variables as Fourier series
+with $N_x$ and $N_z$ modes, and use the $3/2$ dealiasing rule to avoid aliasing
+errors in the nonlinear terms \citep{boyd}.
+
+The geometry of our simulation domain is fixed by one further parameter, $z_0$,
+the forcing location. We choose $L_z = 12.5H$, and the lower and upper damping
+zones are located at $z < 1.5H$ and $z > 10.5H$ respectively. The forcing (see
+Eq.~\eqref{eq:vol_drive}) is at $z_0 = 3H$ with width $\sigma = 0.078H$,
+sufficiently far from the lower damping zone and permitting sufficient room for
+the upward propagating wave to grow as $\propto e^{z/2H}$. Finally, we want
+similar grid spacing in the $x$ and $z$ directions (i.e.\ $L_x / N_x \sim L_z
+/N_z$), guided by the intuition that turbulence generated by wave breaking is
+approximately isotropic, so we use $L_x = 4H$ and $N_z / N_x = 4$.
 
@@
 time. Specifically, we use $\Delta t = 0.7 \min(\Delta x / u_x,\Delta z /
-u_{z})$, where the minimum is taken over every grid point in the domain and
-$\Delta x,\Delta z$ are the grid spacings in the $x$ and $z$ directions
-respectively.
+u_{z})$, where the minimum is taken over every grid point in the domain, and
+$\Delta x \equiv L_x / N_x$ and $\Delta z \equiv L_z / N_z$ are the grid
+spacings in the $x$ and $z$ directions respectively.
 
 We non-dimensionalize the problem such that $H = N = \rho_0 = 1$. The physics of
-the simulation is then fixed by the four remaining parameters $k_{x}, \omega, C,
-\nu$. We describe our choices for these parameters below:
-\begin{itemize}
+the simulation is then fixed by the four remaining parameters $k_{x}$, $\omega$,
+$C$, and the viscosity $\nu$. We describe our choices for these parameters
+below:
+\begin{enumerate}
     \item $k_{x}$: Tidally excited waves in stars generally have $\ell = 2$,
-        corresponding to a very small $k_\perp\sim 1/R$, where $R$ is the radius
-        of the star. We use the smallest wavenumber in our simulation,
-        $k_x=2\pi/L_x$.
+        corresponding to a horizontal wavenumber $k_\perp\sim 1/R$, where $R$ is
+        the radius of the star. We use the smallest wavenumber in our
+        simulation, $k_x=2\pi/L_x$.
 
     \item $\omega$: We choose $\omega$ by evaluating the dispersion relation
-        $\omega(k_x, k_z)$ for a desired $k_z$. We pick $\abs{k_z H} = 2\pi$ to
-        ensure the waves are very well resolved in all of our simulations. Note
-        however that tidally forced IGWs typically have $\omega \ll N$, or
-        equivalently $k_r/k_\perp \sim k_r R \gg 1$. This requires $k_z H
-        \gtrsim 1$, which is only marginally satisfied in our simulations.
-
-    \item $C$: In our weak forcing simulations, we first choose $C$ forcing
-        strength such that $\xi_z k_z \ll 1$ is satisfied everywhere in the
-        simulation domain. This constrains $C$ by \autoref{eq:uz_lin}.
+        $\omega(k_x, k_z)$ for a desired $k_z$ (see Eq.~\eqref{eq:disp_rel}). We
+        pick $\abs{k_z H} = 2\pi$ to ensure the waves are very well resolved in
+        all of our simulations. Note however that tidally forced IGWs typically
+        have $\omega \ll N$, or equivalently $k_r/k_\perp \sim k_r R \gg 1$.
+        This requires $\abs{k_z H} \gtrsim 1$, which is only marginally
+        satisfied in our simulations.
+
+    \item $C$: In our weak forcing simulations, we first choose the forcing
+        strength $C$ (see Eq.~\eqref{eq:vol_drive}) such that $\abs{\xi_z k_z}
+        \ll 1$ is satisfied everywhere in the simulation domain. This constrains
+        $C$ by Eq.~\eqref{eq:uz_lin}.
 
@@
         viscosity and diffusivity to the system in a way that conserves
-        horizontal momentum (see Appendix~\ref{se:strat_impl} for details).
-
-        To quantify $\nu$, we define the dimensionless Reynolds number
+        horizontal momentum (see Appendix~\ref{se:strat_impl} for details). To
+        quantify $\nu$, we define the dimensionless Reynolds number
         \begin{equation}
@@
         When the forcing is weak, we may set $\nu \approx 0$, or $\mathrm{Re}
-        \gg 1$ because there is negligible energy transfer between modes. Note
-        that $\nu = 0$ exactly changes the order of the PDE system, as the
+        \gg 1$ because there is negligible energy transfer into higher-order
+        modes. Note that $\nu = 0$ changes the order of the PDE system, as the
         highest-order derivatives appear only in the dissipitive terms, so a
-        nonzero $\nu$ must be chosen even in linear simulations.
-\end{itemize}
-
-Finally, we use initial conditions $\bm{u}(x, z, 0) = \Upsilon(x, z, 0) = 0,
-\varpi(x, z, 0) = 1$ corresponding to hydrostatic equilibrium and no initial
+        nonzero $\nu$ must be chosen even in weakly forced simulations.
+\end{enumerate}
+
+Finally, we use initial conditions $\bm{u}(x, z, 0) = \Upsilon(x, z, 0) = 0$ and
+$\varpi(x, z, 0) = 1$, corresponding to hydrostatic equilibrium and no initial
 fluid motion.
@@
 
-We aim to damp waves that reach the edge of the simulation domain without
-inducing nonphysical reflection. To do so, we replace material derivatives
-in \autoref{se:nl_upsilon} with:
+We aim to damp disturbances that reach the vertical boundaries of the simulation
+domain without inducing nonphysical reflection. To do so, we replace material
+derivatives in Eq.~\eqref{se:nl_upsilon} with:
 \begin{align}
@@
 \end{align}
-where $z_B = 1.5H, z_T = 10.5H$ are the boundaries of the damping zone.
-This strongly damps perturbations below $z_B$ and above $z_T$ with damping time
-$\tau$, negligibly affects dynamics between $z_B, z_T$ and has transition width
-governed by $\Delta z$. We use $\Delta z = 0.25H$ and $\tau = 1/15N$. This
-prescription is similar to \citep{lecoanet_damp} and has the advantage of being
-smooth, important for spectral methods. Further details of our implementation of
-the fluid equations in Dedalus are described in Appendix~\ref{se:strat_impl}.
+where $z_B = 1.5H$ and $z_T = 10.5H$ are the boundaries of the lower and upper
+damping zones respectively. This damps perturbations below $z_B$ and
+above $z_T$ with damping time $\tau$ and negligibly affects the dynamics between
+$z_B$ and $z_T$. We choose the transition width $\Delta z = 0.25H$ and damping
+time $\tau = 1 / (15N)$. This prescription is similar to \citet{lecoanet_damp}
+and has the advantage of being smooth, important for spectral methods. Further
+details of our implementation of the fluid equations in Dedalus are described in
+Appendix~\ref{se:strat_impl}.
 
@@
 
-We describe the results of a simulation with $C = \scinot{1.64}{-7}$ such that
-$\xi_z k_z \ll 1$ everywhere. In particular, $\xi_z k_z \approx \scinot{5}{-5}$
-just above the forcing zone. A nonzero $\nu$ corresponding to $\mathrm{Re} =
-10^7$ was used.
-
-We expect the waves to follow the solution given by \autoref{eq:uz_lin}. Along
-with incompressibility (\autoref{eq:nl_incomp}), we obtain a complete analytical
-solution for flow velocities $\bm{u}_{al}(x, z, t)$. The amplitude of the
-observed IGW in the simulation field $\bm{u}$ relative to analytical solution
-$\bm{u}_{al}$ over some region $z \in [z_b, z_t]$ can be estimated with
+To test our numerical code and implementation, we carry out a simulation in the
+weakly forced regime with $C = \scinot{1.64}{-7}$. According to the linear
+solution (Eq.~\eqref{eq:uz_lin}), this generates IGW with $\abs{\xi_z k_z}
+\approx \scinot{5}{-5}$ just above the forcing zone. This amplitude is
+sufficiently small to ensure $\abs{\xi_z k_z} \ll 1$ in the entire simulation
+domain. We include a nonzero $\nu$ corresponding to $\mathrm{Re} = 10^7$.
+
+We expect the waves to follow the analytical solution given by
+Eq.~\eqref{eq:uz_lin} and the corresponding $u_x(x, z, t)$; we denote this
+solution $\bm{u}_{al}(x, z, t)$. The amplitude of the observed IGW in the
+simulation field $\bm{u}$ relative to analytical solution $\bm{u}_{al}$ over
+some region $z \in [z_b, z_t]$ can be estimated from
 \begin{equation}
@@
         {\int\limits_{z_b}^{z_t}\int\limits_0^{L_x}
-        \overline{\rho}\abs{\bm{u}_{al}}^2\;\mathrm{d}x\mathrm{d}z}
+        \overline{\rho}\abs{\bm{u}_{al}}^2\;\mathrm{d}x\mathrm{d}z}.
         \label{eq:ahat_def}
@@
 The subscript $i$ denotes the incident wave. If $\bm{u} = \bm{u}_{al}$, then
-$A_i(t) = 1$. The normalization is chosen such that the overlap between $\bm{u},
-\bm{u}_{al} \propto e^{z/2H}$ is evenly weighted throughout the integration
-region.
-
-For a simulation satisfying $\xi_z k_z \ll 1$ everywhere, we expect
+$A_i(t) = 1$. The normalization in Eq.~\eqref{eq:ahat_def} is chosen such that
+the overlap between $\bm{u}, \bm{u}_{al} \propto e^{z/2H}$ is evenly weighted
+throughout the integration region.
+
+For the weakly forced simulation, we expect
 $A_i(t) = 1$ when integrated between the forcing and damping zones, i.e.\
-$z_b \gtrsim z_0, z_t \lesssim z_T$ ($z_0, z_T$ are defined in
-\autoref{eq:vol_drive} and \autoref{eq:Gamma} respectively). For consistency
-with the nonlinear case later, we choose $z_b = z_0 + 3\sigma, z_t = z_b + H$
-(using $z_t = z_T - \Delta z$ just below the damping layer instead does not
-change the results). The resulting measurement of $A_i(t)$ is shown in
-\autoref{fig:lin_amps}.
+$z_b \gtrsim z_0$ and $z_t \lesssim z_T$ ($z_0, z_T$ are defined in
+Eq.~\eqref{eq:vol_drive} and Eq.~\eqref{eq:Gamma} respectively). For consistency
+with the nonlinear case later, we choose $z_b = z_0 + 3\sigma$ and $z_t = z_b +
+H$. Note that using a larger integration domain by choosing $z_t = z_T - \Delta
+z$ just below the upper damping zone instead does not change the measured $A_i$.
+The resulting measurement of $A_i(t)$ is shown in Fig.~\ref{fig:lin_amps}, and
+indeed $A_i \approx 1$ after the initial transient.
 \begin{figure}
     \centering
-    \includegraphics[width=\columnwidth]{plots/lin_amps.png}
-    \caption{Amplitude of excited wave over time in weak forcing simulation,
-    computed using \autoref{eq:ahat_def}. $A_i(t) = 1$ corresponds to
-    perfect agreement with the analytical estimate. After an initial transient
-    phase, we observe $A_i(t)$ asymptotes to $\approx 1$, implying continuous
-    excitation of identical IGW with the expected amplitude. The small
-    suppression may be due to timestepping errors, as a relatively large fixed
-    step size $\Delta t = 0.1/N$ was used for this
-    simulation.}\label{fig:lin_amps}
+    \includegraphics[width=0.9\columnwidth]{plots/lin_amps.png}
+    \caption{Amplitude of the excited IGW over time (in units of $N^{-1}$) in
+    the weakly forced simulation, computed using Eq.~\eqref{eq:ahat_def}.
+    $A_i(t) = 1$ corresponds to perfect agreement with the analytical estimate.
+    After an initial transient phase, we observe $A_i(t)$ asymptotes to $\approx
+    1$, implying continuous excitation of identical IGW with the expected
+    amplitude. The small deviation of $A_i(t)$ from unity may be due to
+    timestepping errors, as a relatively large fixed step size $\Delta t =
+    0.1/N$ was used for this simulation.}\label{fig:lin_amps}
 \end{figure}
 
-The analytical theory also predicts that the horizontal momentmum flux $F(z, t)$
-should be independent of $z$ between the forcing zone where it is generated and
-the damping zone where it is dissipated (\autoref{eq:F_def}). We make may a
-slightly more accurate estimate of the flux carried in the IGW by directly
-substituting $\bm{u}_{al}$ into \autoref{eq:F_def}. Calling this estimate
-\begin{equation}
-    F_{al}(z) \equiv \ev{\overline{\rho} u_{al,x}u_{al,z}}_x
-\end{equation}
-we may then measure the agreement of our simulation with analytical expectation
-by computing $F$ from the simulation via \autoref{eq:F_def}. Where nonlinear
-effects are unimportant, we expect $F(z, t) = F_{al}(z)$.
-\autoref{fig:lin_fluxes} shows that $F \approx F_{al}$ between $z_0$ and $z_T$,
-as expected.
+The analytical theory (Section~\ref{ss:lin_analysis}) also predicts that the
+horizontal momentmum flux $F(z, t)$ is independent of $z$ between the forcing
+zone where the wave is generated and the damping zone where it is dissipated.
+The expected horizontal momentum flux carried by the excited IGW in the linear
+theory can be computed by simply evaluating Eq.~\eqref{eq:F_def} for
+$\bm{u}_{al}$ and is a constant:
+\begin{equation}
+    F_{al} \equiv \ev{\rho u_{al, x} u_{al, z}}_x.\label{eq:F_al}
+\end{equation}
+Denote the momentum flux measured in the simulation by $F(z, t)$
+(use Eq.~\eqref{eq:F_def} with velocities taken from the simulation), then we
+expect $F(z, t) = F_{al}$ between $z_0$ and $z_T$. Fig.~\ref{fig:lin_fluxes}
+shows agreement with this prediction.
 \begin{figure}
     \centering
-    \includegraphics[width=\columnwidth]{plots/lin_fluxes.png}
-    \caption{$F/F_{al}$ as a function of $z$ at select times $t$. As
-    initial transients die out, $F / F_{al} = 1$ to very good agreement above
-    the forcing zone $z > z_0 = 2H$ and below the damping zone $z \lesssim z_T =
-    9.5H$. The flux excited in the forcing zone is transported without loss to
-    the top of the domain, where it is dissipated by the damping layer (see
-    \autoref{ss:damping}) without reflection.}\label{fig:lin_fluxes}
+    \includegraphics[width=0.9\columnwidth]{plots/lin_fluxes.png}
+    \caption{$F/F_{al}$ as a function of $z$ at select times $t$ (in units of
+    $N^{-1}$). As the initial transient dies out, $F / F_{al} \approx 1$ to a
+    good approximation above the forcing zone $z > z_0 = 2H$ and below the
+    damping zone $z \lesssim z_T = 9.5H$. The horizonal momentum flux excited in
+    the forcing zone is transported without loss to the top of the domain, where
+    it is dissipated by the damping zone (see Section~\ref{ss:damping}) without
+    reflection.}\label{fig:lin_fluxes}
 \end{figure}
 
+% GOT HERE
+
 \section{Numerical Simulations of Wave Breaking}\label{s:sim}
@@
 
-To perform simulations of wave breaking phenomena, we use the same values as
-described in \autoref{s:numerics} except for $C$ and $\nu$. In particular, we
-choose $C$ such that $\at{\xi_z k_z}_{z_0} = 0.1$ in the forcing zone. The
-linear solution predicts $\xi_z k_z \sim 4.25$ at the upper damping zone $z_T$.
-$\mathrm{Re}$ was chosen to be as large as possible across the various
-resolutions. A table of our simulations can be found in \autoref{tab:params}.
+To perform simulations of wave breaking phenomena, we use the same setup as
+described in Section~\ref{s:numerics} and Section~\ref{s:weak_sim} except for
+different values of $C$ and $\nu$. In particular, we choose $C$ such that
+$\abs{\xi_z k_z} = 0.1$ in the forcing zone ($z = z_0$). The linear solution
+predicts $\abs{\xi_z k_z} \sim 4.25$ at the upper damping zone $z_T$. We choose
+the viscosity $\nu$ so that $\mathrm{Re}$ is as large as possible across the
+various resolutions. A table of our simulations can be found in
+Table~\ref{tab:params}.
 \begin{table}
@@
 % convert yubo_000054.png -crop 2000x2000+150+250 out.png
-In \autoref{fig:snapshots}, we present snapshots of $u_x$ and $\Upsilon$ at
-various interesting phases of the simulation. The flow evolves through distinct
+In Fig.~\ref{fig:snapshots}, we present snapshots of $u_x$ and $\Upsilon$ at
+various phases of the simulation. The flow evolves through several distinct
 stages:
@@
     \item At a slightly later time (top right panel), the mean flow in
-        $u_x$ has become much more prominent and the critical layer $z_c$ has
+        $u_x$ becomes much more prominent and the critical layer $z_c$ has
         become much more definite. Small-scale fluctuations are still present in
-        $u_z$ but at smaller amplitudes due to being in a denser region of the
+        $u_x$ but at smaller amplitudes due to being in a denser region of the
         fluid.
 
-    \item In the bottom left panel, the critical layer transition is now
-        extremely sharp, and small swirls of limited vertical extent at the
-        location of the critical layer in $u_z$ suggest that the
-        Kelvin-Helmholtz instability is responsible for regulating the minimum
-        width of this transition. This is supported by the conclusions in
-        \autoref{s:khi}.
-
-    \item  The end of the simulation shows very few significant qualitative
-        differences from the previous snapshot, suggesting that the latter half
-        of our simulation is temporally converged.
+    \item In the bottom left panel, the critical layer transition becomes very
+        sharp, and small swirls of limited vertical extent in $\Upsilon$ at the
+        location of the critical layer suggest that the Kelvin-Helmholtz
+        instability is responsible for regulating the width of this transition.
+        More discussion can be found in Section~\ref{ss:khi}.
+
+    \item At the end of the simulation (bottom right panel), the flow shows very
+        few significant qualitative differences from the previous snapshot,
+        suggesting that the latter phase of the simulation has reached a steady
+        state.
 \end{enumerate}
@@
     \includegraphics[width=0.45\textwidth]{plots/yubo_000451.png}
-    \caption{Snapshots of $u_x, u_z$ in the fiducial simulation illustrating
-    distinct phases of the evolution of the flow. Note that $\overline{U}_c
-    \approx 0.16HN$ for the parameters used. The four times depicted are $t =
-    413.4/N$ (top left), $t = 658.5/N$ (top right), $t = 1171.4/N$ (bottom
-    left), and $t = 3437.8/N$ (bottom right). The four snapshots illustrate (i)
-    the initial transient wave breaking phase, (ii) formation of a distinct
-    critical layer, (iii) steepening of the critical layer, and (iv) propagation
-    of the critical layer respectively.}\label{fig:snapshots}
+    \caption{Snapshots of $u_x$ (in units of $HN$) and $\Upsilon \equiv
+    \ln\p{\rho / \bar{\rho}}$ in the fiducial simulation illustrating distinct
+    phases of the evolution of the flow. Note that $\overline{U}_c \approx
+    0.16HN$ (see Eq.~\eqref{eq:u_crit}) for the parameters used. The four times
+    depicted are $t = 413.4/N$ (top left), $t = 658.5/N$ (top right), $t =
+    1171.4/N$ (bottom left), and $t = 3437.8/N$ (bottom right). The four
+    snapshots illustrate (i) the initial transient wave breaking phase, (ii)
+    formation of a distinct critical layer, (iii) steepening of the critical
+    layer, and (iv) downward advance of the critical
+    layer.}\label{fig:snapshots}
 \end{figure*}
 
-In \autoref{fig:nl_fluxes}, we plot $\overline{U}$ and $F / F_{al}$ as a function
-of $z$ at the times depicted in \autoref{fig:snapshots}.
-At each time, $\overline{U}$ is close to zero below the critical layer, but then
-sharply increases to $\overline{U}_c$ at the critical layer. Above the critical
-layer, $\overline{U}$ varies slightly due to momentum transport within the
-spun-up layer. This is in agreement with the predictions
-of \autoref{ss:crit_layer}.
+In Fig.~\ref{fig:nl_fluxes}, we plot the mean horizontal flow velocity
+$\overline{U}$ (Eq.~\eqref{eq:mean_flow}) and the dimensionless momentum flux $F
+/ F_{al}$ (Eqs.~\eqref{eq:F_def} and~\eqref{eq:F_al}) as a function of $z$ at
+the times depicted in Fig.~\ref{fig:snapshots}. At each time, $\overline{U}$ is
+close to zero below the critical layer, but then sharply increases to
+$\overline{U}_c$ at the critical layer (i.e.\ the flow is ``spun-up''). Above
+the critical layer, $\overline{U}$ varies slightly due to momentum transport
+within the spun-up layer. This agrees with the expectation discussed in
+Section~\ref{ss:crit_layer}.
 
 Similarly, $F \lesssim F_{al}$ below the critical layer, and then decreases to
-about zero above the critical layer. However, two notable deviations can be
-observed: (i) the incident flux on the critical layer seems to fluctuate
-somewhat temporally, and (ii) there seems to be a small negative flux above the
-critical layer at later times. These are addressed in the following subsection.
+about zero above the critical layer. However, two notable deviations from the
+discussion in Section~\ref{ss:crit_layer} can be observed: (i) the incident flux
+on the critical layer fluctuates somewhat temporally, and (ii) there is a small
+negative flux just above the critical layer at later times. These are addressed
+in subsequent sections.
 \begin{figure}
     \centering
-    \includegraphics[width=\columnwidth]{plots/nl_fluxes.png}
-    \caption{Plot of $\overline{U}(z, t)$ and $F(z, t) / F_{al}(z)$ respectively
-    at the same times in \autoref{fig:snapshots} over $z$ in our $\mathrm{Re} =
-    1024$ simulation. $\overline{U}, F$ follow their definitions in
-    \autoref{eq:mean_flow} and \autoref{eq:F_def}. The two distinct zones of
-    mean flow separated by a critical layer, the propagation of this critical
-    layer towards lower $z$, and the sharp deposition of $F$ at the critical
-    layer are evident.}\label{fig:nl_fluxes}
+    \includegraphics[width=0.9\columnwidth]{plots/nl_fluxes.png}
+    \caption{The mean horizontal flow velocity $\overline{U}(z, t)$
+    (Eq.~\eqref{eq:mean_flow}) and the dimensionless momentum flux $F(z, t) /
+    F_{al}(z)$ (Eqs.~\eqref{eq:F_def} and~\eqref{eq:F_al}) are plotted at the
+    same times as in Fig.~\ref{fig:snapshots} as a function of $z$ in our
+    fiducial simulation. The two distinct zones of mean flow are separated by a
+    critical layer. The propagation of this critical layer towards lower $z$ and
+    the sharp deposition of $F$ at the critical layer are
+    evident.}\label{fig:nl_fluxes}
 \end{figure}
 
+\subsection{Kelvin-Helmholtz Instability and Critical Layer Width}\label{ss:khi}
+
+The formation of the critical layer is associated with a strong shear flow. What
+is the width of this layer? Inspection of Fig.~\ref{fig:snapshots} suggests the
+presents of the Kelvin-Helmholtz Instability (KHI) in the critical layer. In a
+stratified medium, KHI occurs when the Richardson number (Eq.~\eqref{eq:ri_def})
+satisfies $\mathrm{Ri} \lesssim 1/4$ \citep[e.g.][]{shu1991physics}. It is
+natural to suspect that the shear flow cannot steepen further than the onset of
+KHI\@. To test this, we compute the local $\mathrm{Ri}$ for the shear flow
+around the critical layer.
+
+To avoid noisiness, $\mathrm{Ri}$ is measured as follows: we first assign an
+$\mathrm{Ri}_x(x, t)$ for every $x$ in the critical layer, then take the median
+as $\mathrm{Ri}$ for the entire layer. $\mathrm{Ri}_x$ is computed using the
+vertical distance over which the local $u_x$ increases from $0.3\bar{U}_c$ to
+$\bar{U}_c$ (see Eq.~\eqref{eq:u_crit}). The value $0.3$ is necessary to exclude
+the small mean flow generated in the weakly nonlinear regime far below the
+critical layer. This procedure can be written:
+\begin{align}
+    z_{CL, \min}(x, t) &\equiv \argmin_z
+        \z{z\mid u_x(x, z, t) > 0.3\overline{U}_c},\\
+    z_{CL, \max}(x, t) &\equiv \argmax_z
+        \z{z\mid u_x(x, z, t) < \overline{U}_c},\\
+    \mathrm{Ri}_x(x, t) &\equiv
+        \p{\frac{N^2 \p{z_{CL, \max} - z_{CL, \min}}^2}{(0.7
+            \overline{U}_c)^2}},\\
+    \mathrm{Ri}(t) &\equiv \med_x\mathrm{Ri}\p{x, t}.\label{eq:ri_med_def}
+\end{align}
+To understand the variation in $\mathrm{Ri}$ over $x$, we can also compute
+$\min\limits_x \mathrm{Ri}_x(x, t)$ (the maximum is very noisy). Both of these
+are shown in Fig.~\ref{fig:nl_f_ri}. The critical layer quickly narrows to its
+minimum width, driven by the anti-diffusivity of IGWs.
+\begin{figure}
+    \centering
+    \includegraphics[width=0.9\columnwidth]{plots/nl_f_ri.png}
+    \caption{Local Richardson number (Eq.~\eqref{eq:ri_med_def}) of the flow at
+    the critical layer over time (in units of $N^{-1}$) in our fiducial
+    simulation. The red and solid green lines denote respectively the minimum
+    and median of $\mathrm{Ri}_x(x, t)$. These numbers measure the mean and
+    spread in width of the critical layer over $x$. Note that $\mathrm{Ri} \sim
+    \frac{1}{4}$ corresponds to the KHI, so this plot suggests the shear at the
+    critical layer does not steepen past the onset of the
+    KHI.}\label{fig:nl_f_ri}
+\end{figure}
+
 \subsection{Flux Budget}\label{ss:flux_budget}
 
-The propagation of the critical layer $z_c(t)$ is driven by the absorbed flux at
-$z_c$, following \autoref{eq:zc_anal}. When analyzing the simulation, one must
-first compute $z_c$ before analyzing the flux budget in its vicinity, but
-conceptually $z_c$ is a consequence of flux redistribution at the critical
-layer. We proceed in conceptual order and first describe how we analyze the flux
-transfer at $z_c$.
+Knowing now the width of the critical layer, we next seek its location. The
+downward propagation of the critical layer location $z_c(t)$ is driven by the
+absorption of horizontal momentum flux at $z_c$, following
+Eq.~\eqref{eq:zc_anal}. When analyzing the simulation, one must first compute
+$z_c$ before examining the flux budget in its vicinity, but conceptually $z_c$
+is a consequence of flux redistribution at the critical layer. Here, we proceed
+in conceptual order and first describe how we analyze the flux transfer at
+$z_c$.
 
@@
 \begin{equation}
-    F_i(t) = F_a(t) + F_r(t) + F_s(t),
-\end{equation}
-where $F_i > 0$ is the incident flux, $F_a > 0$ is the absorbed flux, $F_r > 0$
-is the reflected flux, and $F_s < 0$ is some flux above the critical layer,
-likely responsible for momentum redistribution within the synchronized outer
-zone of the WD\@. Careful accounting of $F_s$ turns out to be important to
-obtain the correct critical layer propagation. A more specific physical
+    F_i(t) = F_a(t) + F_r(t) + F_s(t),\label{eq:f_budget}
+\end{equation}
+where $F_i$ is the incident flux, $F_a$ is the absorbed flux, $F_r$ is the
+reflected flux, and $F_s$ is the ``residual'' flux above the critical layer,
+likely responsible for momentum redistribution within the synchronized upper
+layer. Careful accounting of $F_s$ turns out to be important to obtain the
+correct $F_a$ and resulting critical layer propagation. A more specific physical
 interpretation of $F_s$ is unclear; it is somewhat tempting but unfounded to
-identify $F_s$ with some transmitted flux.
-
-After measuring $z_c$ at each time (see \autoref{ss:cl_prop}), we can then
-determine each of $F_i$, $F_a$, $F_r$, $F_s$ as follows:
-\begin{align}
-    F_i(t) &= \ev{\rho u_{x} u_{z}}_xA_i^2(t),\\
-    F_i(t) - F_r(t) &= \ev{F(z)}_{z \in [z_c - \Delta z - H, z_c - \Delta z]}
+identify $F_s$ with the transmitted flux. In the simulation, we find $F_s < 0$.
+
+After measuring $z_c$ (see Section~\ref{ss:cl_prop}) and $F(z)$
+(Eq.~\eqref{eq:F_def}) at each time step, we can then determine each of $F_i$,
+$F_a$, $F_r$, $F_s$ as follows:
+\begin{align}
+    F_i(t) &= F_{al}A_i^2(t),\\
+    F_i(t) - F_r(t) &= \ev{F(z, t)}_{z \in [z_c - \Delta z - H, z_c - \Delta z]}
         ,\label{eq:fr_def}\\
-    F_s(t) &= \ev{\z{F(z): F(z) < 0}}_{z \in [z_c, z_c + \Delta z]},
+    F_s(t) &= \ev{\z{F(z, t): F(z, t) < 0}}_{z \in [z_c, z_c + \Delta z]},
         \label{eq:fs_def}\\
@@
 \end{align}
-Here, $\ev{\dots}_{z \in [z_a, z_b]}$ denotes an average over the interval
-$[z_a, z_b]$. Below the critical layer, we average over an interval of length
-$\frac{2\pi}{k_{z}}$ a full vertical wavelength. The offset $\Delta z$ is
-necessary to make a measurement of the incident flux unaffected by the
-turbulence within the critical layer itself. The height of the critical layer is
-limited by $\mathrm{Ri} \lesssim 1$ (see \autoref{s:khi}), which bounds its
-vertical extent $\sim \frac{1}{k_{z}}$. We empirically found an offset of
-$\Delta z = \frac{3}{k_z}$ was necessary to be sufficiently far from strong
+Here, $\ev{\dots}_{z \in [z_a, z_b]}$ denotes a vertical average over the
+interval $[z_a, z_b]$. Below the critical layer, we average over an interval of
+length $H$, also the vertical wavelength. The offset $\Delta z$ is necessary to
+make the measurement of the incident flux unaffected by the turbulence within
+the critical layer itself. The width of the critical layer is limited by
+$\mathrm{Ri} \lesssim 1$ (see Section~\ref{ss:khi}), which bounds its vertical
+extent $\sim \frac{1}{\abs{k_{z}}}$. We empirically found an offset of $\Delta z
+= \frac{3}{\abs{k_z}}$ was necessary to be sufficiently far from strong
 fluctuations near the critical layer.
@@
 Above the critical layer, we observe that the $F_s$ feature has varying width
-(see \autoref{fig:nl_fluxes} again) but contributes significantly to the total
-flux budget. We average only where $F < 0$ to be robust to such width
-variations, and this was found to be a sufficiently accurate way of measuring
-$F_s$.
+(compare e.g.\ the $t = 1171.4/N$ and $t = 3437.8/N$ lines in the bottom panel
+of Fig.~\ref{fig:nl_fluxes}) but contributes significantly to the total
+flux budget. We average only where $F < 0$ so that $F_s$ is robust to such width
+variations. We find that this is a sufficiently accurate way of measuring $F_s$
+and determining $F_a$.
+
+Fig.~\ref{fig:nl_f_amps2} depicts the four components of this flux decomposition.
+\begin{figure}
+    \centering
+    \includegraphics[width=0.9\columnwidth]{plots/nl_f_amps2.png}
+    \caption{Momentum flux decomposition calculated from the simulation. Plotted
+    are the four components of the horizontal momentum flux budget over time
+    (see Eq.~\eqref{eq:f_budget}), in units of the analytical estimate for the
+    incident wave flux $F_{al}$ (Eq.~\eqref{eq:S_lin}): $F_i$, the flux incident
+    on the critical layer (green); $F_a$, the flux absorbed by the critical
+    layer (blue); $F_r$, the flux reflected at the critical layer (red); and
+    $F_s$, the flux inside the synchronized layer
+    (black).}\label{fig:nl_f_amps2}
+\end{figure}
 
@@
 With a careful determination of $F_a$, we can next make predictions for the
-propagation of $z_c(t)$ and compare to the observed propagation.
+propagation of $z_c(t)$ and compare to the measured propagation in the
+simulation.
 In principle, $z_c$ is the location where the incident flux significantly
-attenuates. In the simulations, shear turbulence causes $F$ to have significant
-spatial and temporal fluctuations that propagate to large temporal fluctuations
+attenuates. In the simulation, shear turbulence causes $F$ to have significant
+spatial and temporal fluctuations that translate to large temporal fluctuations
 in $z_c(t)$. To minimize these spurious fluctuations, we measure the location of
-the critical layer using a spatial average of where flux deposition occurs
+the critical layer using a spatial average of where flux deposition occurs:
 \begin{align}
@@
 
-Finally, we may plot the measured $z_c$ against two semi-analytic predictors:
-(i) integration of \autoref{eq:zc_anal} using the measured $F_a(t)$, and
-(ii) substituting the time-averaged $\ev{F_a}_t$ (over the entire length of
-the simulation) into \autoref{eq:zc_sol}. Since $z_c(t)$ is both unstable and
-less well-defined at early times (when the critical layer is thick and transient
-behavior is still strong), we solve \autoref{eq:zc_anal} by integrating
-backwards from the end of the simulation, using $z_c(t_f)$ as initial condition.
-The resulting predictors are depicted in \autoref{fig:nl_front}. The good
-agreement between the evolution of $z_c(t)$ and its estimate via $F_a(t)$ and
-\autoref{eq:zc_anal} are noteworthy.
-
-The behavior of the behavior of the fiducial, time-averaged $\ev{F_a}_t =
-0.71F_{al}$ predictor provides interesting insights. The general agreement
-clearly demonstrates $F_a < F_{al}$ and that momentum flux absorption is
-not complete. However, its underprediction of critical layer velocity at early
-times contrasts with its overprediction at later times. This suggests that
-momentum flux absorption changes significantly over time, in agreement with
-detailed analysis in \autoref{ss:reflectivity}.
+Finally, in Fig.~\ref{fig:nl_front} we plot the numerically measured $z_c$
+against two semi-analytic predictors: (i) integration of Eq.~\eqref{eq:zc_anal}
+using the measured $F_a(t)$, and (ii) substituting the time-averaged
+$\ev{F_a}_t$ (over the entire length of the simulation) into
+Eq.~\eqref{eq:zc_sol}. Since $z_c(t)$ is less well-defined at early times (when
+the critical layer is thick and transient behavior is strong), we solve
+Eq.~\eqref{eq:zc_anal} by integrating backwards from the end of the simulation
+($t = t_f$), using $z_c(t_f)$ as the initial condition. From
+Fig.~\ref{fig:nl_front}, we see that the agreement between the measured $z_c(t)$
+and its estimate via $F_a(t)$ and Eq.~\eqref{eq:zc_anal} is excellent.
+
+By time-averaging the numerically measured $F_a$, we find $\ev{F_a}_t \approx
+0.71F_{al}$, which is used for the second predictor in Fig.~\ref{fig:nl_front}.
+Note that $F_a < F_{al}$, so momentum flux absorption at the critical layer is
+incomplete.
+% Moreover, careful comparison of the two predicted curves in
+% Fig.~\ref{fig:nl_front} shows that there is . This suggests that
+% momentum flux absorption changes significantly over time, in agreement with
+% detailed analysis in Section~\ref{ss:reflectivity}.
 \begin{figure}
     \centering
-    \includegraphics[width=\columnwidth]{plots/nl_front.png}
-    \caption{Propagation of the critical layer over time. Shown are (black)
+    \includegraphics[width=0.9\columnwidth]{plots/nl_front.png}
+    \caption{Propagation of the critical layer over time. Shown are: (black)
     $z_c(t)$ from simulation data, (green) predictor of $z_c(t)$ using direct
-    integration of \autoref{eq:zc_anal} for $F_a(t)$ measured from
-    simulation data (described in \autoref{eq:fa_def}) and (blue) direct
+    integration of Eq.~\eqref{eq:zc_anal} for $F_a(t)$ measured from
+    simulation data (described in Eq.~\eqref{eq:fa_def}), and (blue) direct
     substitution of time-averaged $\ev{F_a(t)}_t$ into
-    \autoref{eq:zc_sol}. Predictors use the end of the simulation as initial
+    Eq.~\eqref{eq:zc_sol}. Predictors use the end of the simulation as initial
     conditions and integrate backwards, as $z_c$ is less well-defined at early
     times. The agreement of the directly-integrated predictor with the data
-    shows \autoref{eq:zc_anal} is a good description of the evolution of $z_c$.
+    shows Eq.~\eqref{eq:zc_anal} is a good description of the evolution of $z_c$.
     The the poorer but qualitatively correct agreement of the time-averaged
-    predictor with the data shows both that $F_a(t) \neq F_{al}$ and that
-    $F_a(t)$ likely varies in time.}\label{fig:nl_front}
+    predictor with the data shows both that $F_a(t) <
+    F_{al}$.}\label{fig:nl_front}
 \end{figure}
@@
 To further understand the behavior at the critical layer, we compare two
-reflective behaviors observed in the simulation: (i), the presence of a
-reflected wave with wave vector $\bm{k}_r = k_{x}\uv{x} - k_{z}\uv{z}$, and (ii)
-$F_r$. The reflected wave amplitude and reflected flux need not agree exactly if
-some reflected flux is in higher-order modes, which is indeed the case in our
-simulations. Both are of physical interest, however: the reflected wave
-amplitude may be of physical interest in setting up standing modes in a
-realistic star, while the flux transfer properties is important for accurately
-tracking angular momentum transfer during synchronization.
-
-To measure the reflected wave amplitude $A_r(t)$, we use \autoref{eq:ahat_def}
-with $\bm{u}_{al}$ computed with $\bm{k}_r$ with one key change: since the phase
-of the reflected wave is unknown, unlike that of the incident wave, we fit for
-an arbitrary horizontal offset $\delta x(t)$ at each time $t$ (not constrained
-to be on-grid), picking $\delta x$ that maximizes $A_r$. To be explicit, we
-compute
+reflective behaviors observed in the simulation: (i) the presence of a reflected
+wave with wave vector $\bm{k}_r = k_{x}\uv{x} - k_{z}\uv{z}$, and (ii) the
+reflected flux $F_r$. The reflected wave amplitude and flux need not agree
+exactly if some reflected flux is in higher-order modes, which is indeed the
+case in our simulations. Both are of physical interest, however: the reflected
+wave amplitude is essential for setting up standing modes in a realistic star,
+while the flux is important for accurately tracking angular momentum transfer
+during synchronization.
+
+To measure the reflected wave amplitude $A_r(t)$, we use an approach similar to
+the calculation of $A_i(t)$ (Eq.~\eqref{eq:ahat_def}):
 \begin{equation}
@@
         {\int\limits_{z_b}^{z_t}\int\limits_0^{L_x}
-        \overline{\rho}\abs{\bm{u}_{al}}^2\;\mathrm{d}x\mathrm{d}z}
+        \overline{\rho}\abs{\bm{u}_{al}}^2\;\mathrm{d}x\mathrm{d}z},
         \label{eq:ar_def}
 \end{equation}
-In our simulations, the phase offset $\phi_r(t) \equiv k_x \delta x(t) $ behaves
-in agreement with reflection off a moving boundary at $z_c$, well approximated
-by $\abs{\pd{\phi_r}{t}} \approx 2\abs{\pd{(k_{z}z_c)}{t}}$.
-
-Since reflectivity depends sensitively on accurate measurements of $A_i,
-A_r$, we remark \autoref{eq:ahat_def} and \autoref{eq:ar_def} ensure
-orthogonality between $k_{x}\uv{x} \pm k_{z}\uv{z}$ modes. The integral for
-$A_r$ is also performed over $z \in [z_0 + 3\sigma, z_0 + 3\sigma + H]$. Since
-$A_i(t), A_r(t)$ vary somewhat strongly over time, we perform time averaging
-over interval approximately $8\pi/\omega$, denoted by angle brackets. We can
-then define the amplitude reflectivity
+where $z_b = z_0 + 3\sigma$ and $z_t = z_b + H$ as before. The primary
+difference from Eq.~\eqref{eq:ahat_def} is the introduction of free parameter
+$\delta x$, the horizontal phase offset of the reflected wave. Since $\delta x$
+is unknown \emph{a priori}, we choose $\delta x \in [0, 2\pi]$ that maximizes
+$A_r(t)$. In our simulation, the phase offset $\phi_r(t) \equiv k_x \delta x(t)
+$ is consistent with reflection off a moving boundary at $z_c$, well
+approximated by $\abs{\pdil{\phi_r}{t}} \simeq 2\abs{\pdil{(k_{z}z_c)}{t}}$.
+Finally, note that $\bm{u}_{al}$ and $\bm{u}_{al, \bm{k}_r}$ are orthogonal
+under the integral used for Eqs.~\eqref{eq:ahat_def} and~\eqref{eq:ar_def}, so
+$A_i$ and $A_r$ are independent.
+
+Fig.~\ref{fig:nl_f_amps} illustrates the behaviors of these two quantities. Both
+of them vary significantly in time but their mean values appear to converge
+towards the end of the simulation. A modest ($\sim20\%$), long-term fluctuation
+in $A_i(t)$ is observed, likely of numerical origin.
+\begin{figure}
+    \centering
+    \includegraphics[width=0.9\columnwidth]{plots/nl_f_amps.png}
+    \caption{Wave amplitudes observed in the simulation. The
+    top panel shows the incident wave amplitude $A_i(t)$ (green) and the
+    reflected wave amplitude $A_r(t)$ (red) just above the forcing
+    zone. The modest fluctuation in $A_i(t)$ is likely of numerical origin but
+    has no bearing on our results.}\label{fig:nl_f_amps}
+\end{figure}
+
+Since $A_i(t), A_r(t)$ vary somewhat over time, we perform time averaging over
+interval of approximately $8\pi/\omega$, denoted by angle brackets. We can then
+define the amplitude reflectivity
 \begin{equation}
@@
 
-To compare to the behavior of the flux at the critical layer, we compare
-$\mathcal{R}_A^2(t)$ to the ratios of $F_r$ and $F_s$ to $F_i$. We define
-shorthand
-\begin{align}
-    \hat{F}_r &\equiv \frac{\ev{F_r}(t)}{\ev{F_i}(t)},
-        \label{eq:srefl_def1}\\
-    \hat{F}_s &\equiv -\frac{\ev{F_s}(t)}{\ev{F_i}(t)}.
-        \label{eq:srefl_def}
-\end{align}
-\autoref{fig:nl_f_amps} demonstrates the relations between these quantities. All
-of these quantities vary significantly temporally but their mean values appear
-to converge towards the end of the simulation. A significant fluctuation in
-$A_i(t)$ is observed. We were unable to identify the cause of this unintended
-fluctuation, but it doees not affect $\mathcal{R}_A$ thanks to the time
-averaging used.
+This reflectivity can be compared to the flux redistribution at the critical
+layer by calculating $\mathcal{R}_A^2(t)$ and the ratios of $F_r$ and $-F_s$ (as
+$F_s < 0$) to $F_i$. A comparison between $\mathcal{R}_A^2$ and $\hat{F}_r$ is
+appropriate as $F \propto A^2$ (Eq.~\eqref{eq:S_lin}). We define \begin{align}
+\hat{F}_r &\equiv \frac{\ev{F_r}(t)}{\ev{F_i}(t)}, \label{eq:srefl_def1}\\
+\hat{F}_s &\equiv -\frac{\ev{F_s}(t)}{\ev{F_i}(t)}. \label{eq:srefl_def}
+\end{align} Fig.~\ref{fig:nl_f_refl} shows $\mathcal{R}_A^2$, $\hat{F}_r$,
+$\hat{F}_s$ using the results of Fig.~\ref{fig:nl_f_amps2} and
+Fig.~\ref{fig:nl_f_amps} as inputs. Visual inspection suggests the three
+quantities have reached their asymptotic values (in the time-averaged sense) for
+$t \gtrsim 2500/N$. The aforementioned fluctuations in $A_i$ do not affect our
+reflectivity results thanks to the time averaging used in
+Eqs.~\eqref{eq:Ra_def}--\eqref{eq:srefl_def}.% chktex 2
+We see that in general $\hat{F}_r \gtrsim \mathcal{R}_A^2$, conforming with the
+expectation that the reflected flux consists of the simple reflected mode and
+higher order modes as well.
 \begin{figure}
     \centering
-    \includegraphics[width=0.8\columnwidth]{plots/nl_f_amps.png}
-    \caption{The top panel shows the incident wave amplitude $A_i(t)$ (green)
-    and the reflected wave amplitude $A_r(t)$ (red) just above the forcing zone.
-    The bottom panel shows the behavior of the four components of the horizontal
-    momentum flux budget over time, in units of the analytical estimate
-    \autoref{eq:S_lin}: (green) flux incident on the critical layer, (blue) flux
-    absorbed by the critical layer, (red) flux reflected at the critical layer,
-    and (black) flux inside the synchronized layer.}\label{fig:nl_f_amps}
+    \includegraphics[width=0.9\columnwidth]{plots/nl_f_refl.png}
+    \caption{Behavior of dimensionless reflectivities $\mathcal{R}_A^2$,
+    $\hat{F}_r$, and residual flux $\hat{F}_s$
+    (Eqs.~\eqref{eq:Ra_def}--\eqref{eq:srefl_def}) as a function of time (in
+    units of $N^{-1}$) in the simulation. These quantities seem to become
+    comparatively stable past about $t = 2500/N$, indicating that an asymptotic
+    value may have been reached. That $\hat{F}_r \gtrsim \mathcal{R}_A^2$
+    implies a substantial fraction of reflected flux is in higher-order modes
+    than the reflected IGW.}\label{fig:nl_f_refl}
 \end{figure}
 
-Using these measured quantities, we may make plots of $\mathcal{R}_A^2$,
-$\hat{F}_r$, $\hat{F}_s$, which are provided in \autoref{fig:nl_f_refl}. A
-comparison between $\mathcal{R}_A^2$ and $\hat{F}_r$ is appropriate as $F
-\propto A^2$. Visual inspection suggests the three quantities have reached their
-asymptotic values after $t \gtrsim 1750/N$. An observation may be made that in
-general $\hat{F}_r \geq \mathcal{R}_A^2$; this conforms to the expectation
-that reflected flux consists of the simple reflected mode and higher order modes
-as well.
+\subsection{Resolution Study}\label{ss:convergence}
+
+As the primary test of whether our simulations have reached sufficient
+resolution, we consider the convergence of $\mathcal{R}_A^2$, $\hat{F}_r$,
+$\hat{F}_s$. Furthermore, we compare our results against
+Eqs.~\eqref{eq:crit_coeffs}, which suggest that wave reflection at the critical
+layer depends only on the local Richardson number (see Section~\ref{ss:khi} for
+details on our measurement of $\mathrm{Ri}$).
+
+For each simulation in Tab.~\ref{tab:params}, we compute the median values of
+$\mathcal{R}_A^2$, $\hat{F}_r$, $\hat{F}_s$, and $\mathrm{Ri}$
+(Eqs.~\eqref{eq:Ra_def}--\eqref{eq:srefl_def} and~\eqref{eq:ri_def}
+respectively) over the last $1/4$ of the simulation time from each simulation,
+when these quantities have converged to their asymptotic values. These results
+are shown in Fig.~\ref{fig:agg}.
 
@@
     \centering
-    \includegraphics[width=\columnwidth]{plots/nl_f_refl.png}
-    \caption{Behavior of $\mathcal{R}_A^2$, $\hat{F}_r$, and $\hat{F}_s$ as
-    described by Equations~\ref{eq:Ra_def},~\ref{eq:srefl_def1},
-    and~\ref{eq:srefl_def}. The coefficients seem to become comparatively stable
-    past about $t = 1750/N$, indicating that an asymptotic value may have been
-    reached.}\label{fig:nl_f_refl}
-\end{figure}
-
-\subsection{Resolution Study}\label{ss:convergence}
-
-As the primary test of whether our simulations have obtained sufficient
-resolution, we consider the convergence of $\mathcal{R}_A^2$, $\hat{F}_r$,
-$\hat{F}_s$. Furthermore, we are able to compare our results against
-\autoref{eq:crit_coeffs}, which appear to suggest that the behavior of incident
-waves on the critical layer depends only on the local Richardson number (see
-\autoref{s:khi} for details on our measurement of $\mathrm{Ri}$).
-
-For each simulation in \autoref{tab:params}, we compute the median value of each
-of $\mathcal{R}_A^2$, $\hat{F}_r$, $\hat{F}_s$, and $\mathrm{Ri}$ over the last
-$1/4$ of simulation times from each simulation, where all simulations temporally
-converged to what appeared to be asymptotic values. The bars show the temporal
-variation of each measurement between the $16\%$ and $84\%$ range. We illustrate
-the behavior of these averages at different $\mathrm{Re}$ in \autoref{fig:agg}.
-\begin{figure}
-    \centering
-    \includegraphics[width=\columnwidth]{plots/agg.png}
-    \caption{Convergence of median $\hat{F}_r$, $\mathcal{R}_A^2$, $\hat{F}_s$,
-    and $\mathrm{Ri}$ across runs with varying resolution and viscosity as given
-    in \autoref{tab:params}. The bars show the temporal variation of each
-    measurement between the $16\%$ and $84\%$ range. Small horizontal
-    displacements are made for data points at identical $\mathrm{Re}$ for
-    readability. Note that simulations with larger $\mathrm{Re}$ and resolution
-    correspond to smaller viscosity and are more physically realistic. Our
-    measurements seem to converge at large $\mathrm{Re}$. At the smallest
+    \includegraphics[width=0.9\columnwidth]{plots/agg.png}
+    \caption{Convergence of the median $\hat{F}_r$, $\mathcal{R}_A^2$,
+    $\hat{F}_s$, and $\mathrm{Ri}$ (Eqs.~\eqref{eq:Ra_def}--\eqref{eq:srefl_def}
+    and~\eqref{eq:ri_def} respectively) across runs with varying resolution and
+    viscosity as given in Tab.~\ref{tab:params}. Vertical bars show the temporal
+    variation of each measurement between the $16\%$ and $84\%$ range. Small
+    horizontal displacements are made for data points at identical $\mathrm{Re}$
+    for readability. Note that simulations with larger $\mathrm{Re}$ correspond
+    to smaller viscosity and are more physically realistic.  At the smallest
     $\mathrm{Re}$ value, $\mathrm{Ri} \approx 50$ is too large to fit on the
@@
 It is apparent that the Richardson number rapidly converges to $\mathrm{Ri}
-\approx 0.4$ but the non-absorption measures converge more slowly. This is in
-disagreement with \autoref{eq:crit_coeffs} calculated via the linear analytical
-theory. This tension is natural: fluid motion within the critical layer is
-turbulent, so transmission/reflection at the critical layer cannot be captured
-by any linear calculation.
+\approx 0.4$ but the reflectivity measure converges more slowly. This disagrees
+with Eq.~\eqref{eq:crit_coeffs}, which is based on the linear theory. This
+tension is natural: fluid motion within the critical layer is turbulent, so
+reflecton at the critical layer cannot be captured by the linear theory.
 
@@
 
-In this paper, we have performed numerical simulations of breaking IGWs in a
-2D idealized WD fluid. We observe spontaneous formation of a critical layer that
-separates a synchronized upper layer of fluid and a lower layer of fluid with no
-mean horizontal flow. This critical layer then propagates downwards as a
-continuous train of IGW are incident on it (see \autoref{fig:snapshots} for
-snapshots from a fiducial simulation). Our primary conclusions regarding the
-evolution of the critical layer are as follows:
-\begin{itemize}
-    \item The critical layer propagates downward by partially absorbing incident
-        flux supplied by the continuous train of IGW, eventually spinning up the
-        entire fluid (see \autoref{fig:nl_fluxes} for the time evolution of the
-        mean flow and momentum flux in the fluid).
-
-    \item \autoref{fig:nl_front} shows that the height of the critical layer
-        $z_c(t)$ can be estimated from \autoref{eq:zc_anal}. The critical layer
-        only absorbs $\sim 70\%$ of the incident flux in most simulations (see
-        \autoref{fig:agg} for convergence at high resolutions).
-
-        Note this absorbed fraction of incident momentum flux we
-        measure differs from both analytical \citep{booker_bretherton} and
-        numerical \citep{barker_ogilvie} results in the existing literature. The
-        former result is a linear calculation and is superceded by our nonlinear
-        simulation; we speculate on our discrepancy with the latter in
-        \autoref{ss:other} below.
-
-    \item We compute the asymptotic amplitude of reflected waves off the
-        critical layer as well as the average reflected momentum flux in
-        \autoref{fig:nl_f_refl}. Also illustrated is the asymptotic value of an
-        extra momentum flux just above the critical layer, likely responsible
-        for momentum redistribution within the critical layer. Our simulations
-        are run sufficiently long to be temporally converged.
-
-    \item Finally, the largest flow variations are found within the turbulent
-        critical layer, whose minimum width is set by the Kelvin-Helmholtz
-        instability.
-\end{itemize}
+\subsection{Key Results}
+
+In this paper, we have performed numerical simulations of nonlinear breaking of
+IGWs in a stratified isothermal atmosphere. Such a setup represents the
+plane-parallel idealization of the outer envelope of massive stars, WDs, or
+planets. Our simulations use the spectral code Dedalus \citep{dedalus,dedalus2},
+and are carried out in 2D. We observe spontaneous formation of a critical layer
+that separates a synchronized upper layer of fluid and a lower layer of fluid
+with no mean horizontal flow. This critical layer then propagates downwards as a
+continuous train of IGW are incident on it (see Fig.~\ref{fig:snapshots} for
+snapshots from our fiducial simulation). Our primary conclusions regarding the
+evolution of the critical layer are as
+follows:
+\begin{enumerate}
+    \item The width of the turbulent critical layer is determined by requiring
+        the local Richardson number (Eq.~\eqref{eq:ri_def}) $\mathrm{Ri} \gtrsim
+        0.5$, as shown in Fig.~\ref{fig:nl_f_ri}.
+
+    \item The location of the critical layer $z_c(t)$ can be predicted by
+        careful measurement of the absorbed flux at the critical layer
+        (see Eq.~\eqref{eq:zc_anal}) as shown in Fig.~\ref{fig:nl_front}.
+
+    \item Flux absorption at the critical layer is incomplete. The
+        critical layer only absorbs $\sim 70\%$ of the incident flux in most
+        simulations (see Fig.~\ref{fig:agg} for convergence at high
+        resolutions). The reflected flux is carried away from the critical layer
+        as both a lowest-order reflected wave and waves with larger $z$
+        wavenumbers.
+\end{enumerate}
 In the subsequent sections, we will discuss the validity of these results and
@@
 10^{-4}$ however, the corresponding $\mathrm{Re}$ is far too weak to suppress
-reflection/transmission at the critical layer (e.g.\ \autoref{fig:agg}).
+reflection/transmission at the critical layer (Fig.~\ref{fig:agg}).
 
@@
 \end{equation}
-the Alfv\'en wave crossing time (evaluated for a CO WD in \citealp{fullerIV}),
+the Alfv\'en wave crossing time (evaluated for a CO WD in \citet{fullerIV}),
 measures the magnetic coupling time and $t_{gw}$ measures the gravitational wave
@@
 problems of interest:
-\begin{itemize}
+\begin{enumerate}
     \item In astrophysical systems, $k_{\perp} \ll k_r$, while in our study,
@@
         will be studied in future work.
-\end{itemize}
+\end{enumerate}
 
@@
 
-To solve for the excited wave amplitude in the linear problem, consider the
-complexified linearized system of equations, then $\pd{}{t} \to -i\omega,
-\pd{}{x} \to ik_x$. Assuming all fields have dependence $u_z(x, z, t) =
-\tilde{u_z}(z) e^{ik_xx - i\omega t}$ and dropping the tildes for convenience,
-this gives:
+To solve for the linear excited IGW amplitude due to bulk forcing
+(see Eq.~\eqref{eq:vol_drive}), we consider the
+complexified linearized system of equations, with all dynamical variables having
+dependence $u_z(x, z, t) = \tilde{u_z}(z) e^{ik_xx - i\omega t}$. Thus,
+$\pdil{}{t} \to -i\omega t$, $\pd{}{x} \to ik_x$, and the dynamical fluid
+equations become (see Eqs.~\eqref{se:nl_orig} and~\eqref{se:nl_upsilon}):
 \begin{align*}
@@
     -i\omega \Upsilon - \frac{u_{z}}{H} &=
-        Ce^{-\frac{(z - z_0)^2}{2\sigma^2}}.
+        C\exp\s{-\frac{(z - z_0)^2}{2\sigma^2}} \equiv \mathcal{C}(z).
 \end{align*}
-We have replaced all partial derivatives with regular derivatives. This can be
-recast solely in terms of $u_{z}$ as
+These can be recast solely in terms of $u_{z}$ as
 \begin{align*}
@@
         + u_{z}\frac{N^2k_x^2}{\omega^2} &=
-    -\frac{gk_x^2}{\omega^2}Ce^{-\frac{(z - z_0)^2}{2\sigma^2}}
+    -\frac{gk_x^2}{\omega^2}\mathcal{C}(z).
         .\label{eq:narrow_inhomo}
 \end{align*}
-The homogeneous solutions are of form $u_{z,\pm}(z) = e^{\p{\frac{1}{2H} \pm
+The homogeneous solutions are of form $u_{z,\pm}(z) = \exp\s{\p{\frac{1}{2H} \pm
 ik_z}\p{z - z_0}}$ where $k_z$ satisfies the dispersion relation
-\autoref{eq:disp_rel}. We compute the solution to the inhomogeneous ODE by the
-method of variation of parameters. The Wronskian can be computed
+(Eq.~\eqref{eq:disp_rel}). We compute the solution to the inhomogeneous ODE by
+the method of variation of parameters. The Wronskian is
 \begin{equation}
@@
         u_{z,+} & u_{z,-} \\[4pt]
-        \rd{u_{z,+}}{z} & \rd{u_{z,-}}{z}
+        \rdil{u_{z,+}}{z} & \rdil{u_{z,-}}{z}
     \end{vmatrix} = -2ik_ze^{z/H}.
@@
     u_z = -u_{z,+}\int \frac{1}{W} u_{z,-} \p{-\frac{gk_x^2}{\omega^2}
-            Ce^{-\frac{(z - z_0)^2}{2\sigma^2}}}\;\mathrm{d}z
+            \mathcal{C}(z)}\;\mathrm{d}z
         + u_{z,-}\int \frac{1}{W} u_{z,+} \p{-\frac{gk_x^2}{\omega^2}
-            Ce^{-\frac{(z - z_0)^2}{2\sigma^2}}}\;\mathrm{d}z.
-\end{equation}
-Taking these integrals and applying boundary conditions $u_z\p{z \to \infty} =
-u_{z,+}, u_z\p{z \to -\infty} = u_{z,-}$ gives exact solution
-\begin{equation}
-    u(z) = \frac{C}{2ik_z}\frac{gk_x^2}{\omega^2}
-            e^{\frac{\p{\frac{\sigma^2}{2H} \pm ik_z\sigma^2}^2}{2\sigma^2}}
-                \sqrt{2\sigma^2} \frac{\sqrt{\pi}}{2}\s{
-                    -u_+\p{1 + \erf\p{\xi}}
-                    + u_- \p{\erf\p{\xi} - 1}}
-\end{equation}
-If we are concerned with only $z$ scales significantly larger than $\sigma$,
-then we may take $\erf(\xi) \approx \Theta(\xi)$. If we further assume $k_zH \gg
-1$ and restore the $e^{ik_xx - i\omega t}$ factor, we recover the cited form
-\autoref{eq:uz_lin}
+            \mathcal{C}(z)}\;\mathrm{d}z.
+\end{equation}
+Taking these integrals and applying the boundary conditions $u_z\p{z \to \infty}
+= u_{z,+}$, $u_z\p{z \to -\infty} = u_{z,-}$ give the exact solution:
+\begin{align}
+    u(z) &= \frac{\sqrt{\pi}C\sigma}{2^{3/2}ik_z}\frac{gk_x^2}{\omega^2}
+            \exp\s{\frac{\p{\frac{\sigma^2}{2H} \pm ik_z\sigma^2}^2}{2\sigma^2}}
+                \s{-u_{z,+}\p{1 + \erf\p{\xi_+}}
+                    + u_{z,-} \p{\erf\p{\xi_-} - 1}},\\
+    \xi_{\pm} &\equiv \s{\frac{z - z_0}{\sqrt{2\sigma^2}} +
+        \frac{\sigma}{2^{3/2}H} \pm \frac{ik_z\sigma}{\sqrt{2}}}
+\end{align}
+Here, the error function is defined $\erf(z) \equiv
+\p{2/\sqrt{\pi}}\int\limits_0^z \exp\p{-t^2}\;\mathrm{d}t$. If we are concerned
+with only $z$ scales significantly larger than $\sigma$, then we may take
+$\erf(\xi_{\pm}) \approx \Theta(z - z_0)$ (the Heaviside step function). If we
+further assume $\abs{k_zH} \gg 1$ and restore the $e^{ik_xx - i\omega t}$
+factor, we recover Eq.~\eqref{eq:uz_lin} in the main text
 \begin{equation}
     u_{z}(x, z, t) = -\frac{C}{2ik_z}\frac{gk_x^2}{\omega^2}
-        e^{-\frac{k_z^2\sigma^2}{2}}
+        e^{-k_z^2\sigma^2 / 2}
         \sqrt{2\pi \sigma^2} e^{ik_xx - i\omega t} \times
     \begin{cases}
-        e^{\p{\frac{1}{2H} + ik_z}\p{z - z_0} + \frac{k_z\sigma^2}{2H}}
-            & z > z_0,\\
-        e^{\p{\frac{1}{2H} - ik_z}\p{z - z_0} - \frac{k_z\sigma^2}{2H}}
-            & z > z_0.
+        \exp\s{\p{\frac{1}{2H} + ik_z}\p{z - z_0} + i\frac{k_z\sigma^2}{2H}}
+            & \text{for }z > z_0,\\[5pt]
+        \exp\s{\p{\frac{1}{2H} - ik_z}\p{z - z_0} - i\frac{k_z\sigma^2}{2H}}
+            & \text{for }z > z_0.
     \end{cases}
 \end{equation}
-We use the plane wave approximation in our fitting routines since we generally
-fit many $\sigma$ away from $z_0$, and orthogonality with other modes is
-essential for a reliable fit.
-
-\section{Kelvin-Helmholtz Instability and Critical Layer Width}\label{s:khi}
-
-The observation of reflected/transmitted waves is in accordance with previous
-studies as discussed in \autoref{ss:wave_breaking}. Studies have shown that a
-local Richardson number $\mathrm{Ri} \sim 1/4$ corresponds to the onset of
-reflectivity. This $\mathrm{Ri}$ also corresponds to the onset of the
-Kelvin-Helmholtz instability (KHI). Indeed, in our simulations, visual
-inspection suggests the KHI is present in the critical layer (see
-\autoref{fig:snapshots}). It is natural to suspect then that the shear flow
-cannot steepen any further than KHI onset. To verify this, we compute a local
-$\mathrm{Ri}$ for the shear flow around the critical layer.
-
-Since fluid instabilities are local, $\mathrm{Ri}$ must be measured in the
-immediate vicinity of a fluid parcel. Thus, we first assign an $\mathrm{Ri}$ for
-every $x$ in the critical layer, then take the median as $\mathrm{Ri}$ for the
-entire layer. To avoid noisiness, the local $\mathrm{Ri}$ is computed using the
-vertical distance over which the local $u_x$ increases from $0.3 \times$ its
-critical value to its critical value. The value $0.3$ is necessary to exclude
-the small mean flow generated in the weakly nonlinear regime well below the
-critical layer. This is written
-\begin{align}
-    z_{CL, \min}(x, t) &\equiv \argmin_z
-        \z{z\mid u_x(x, z, t) > 0.3\overline{U}_c},\\
-    z_{CL, \max}(x, t) &\equiv \argmax_z
-        \z{z\mid u_x(x, z, t) < \overline{U}_c},\\
-    \mathrm{Ri}(t) &\equiv
-        \med_x\p{\frac{N^2 \p{z_{CL, \max} - z_{CL, \min}}^2}{(0.7
-            \overline{U}_c)^2}}.\label{eq:ri_med_def}
-\end{align}
-To understand the variation in $\mathrm{Ri}$ over $x$, we can also plot using
-the minimum over $x$ (the maximum is significantly noisier). Both of these are
-shown in \autoref{fig:nl_f_ri}. The critical layer quickly narrows to its
-minimum width, driven by the anti-diffusivity of IGWs.
-\begin{figure}
-    \centering
-    \includegraphics[width=0.6\columnwidth]{plots/nl_f_ri.png}
-    \caption{Local Richardson number of the flow at the critical layer over time
-    as defined in \autoref{eq:ri_med_def}. The red and solid green lines denote
-    respectively the minimum and median of $\mathrm{Ri}$ over $x$ (the maximum
-    greatly exceeds the vertical scale of this plot). These numbers effectively
-    measure the mean and spread in width of the critical layer over $x$. Note
-    that $\mathrm{Ri} \sim \frac{1}{4}$ corresponds to the KHI, so this plot
-    suggests the shear at the critical layer does not steepen past the KHI
-    onset.}\label{fig:nl_f_ri}
-\end{figure}
+Note that in the main text, this approximate form is used to compute
+$\bm{u}_{al}$, as it is easier to work with and sufficiently accurate in the
+regions of interest (many $\sigma$ away from $z_0$).
 
@@
 
-We denote $x \in [0, L_x], z \in [0, L_z]$ the simulation domain and $N_x, N_z$
-the number of spectral modes in the respective dimensions.
-
-Numerically, the nonlinear $\frac{\bm{\nabla}P}{\rho}$ term is problematic: we
-desire a system where the fluid fields are not divided by one another. We
-introduce $\varpi = \frac{P}{\rho}$ instead, then mandate $\overline{\rho},
-\overline{\varpi}$ background fields satisfy hydrostatic equilibrium
-$\bm{\nabla}\overline{\varpi} + \overline{\varpi} \bm{\nabla}\overline{\rho} +
-g\uv{z} = 0$. Taking isothermal stratification, we find $\overline{\varpi} = gH$. We
-further change variables to $\Upsilon = \ln \rho - \ln \overline{\rho}$ and
-$\varpi' = \varpi - \overline{\varpi}$ deviations from the background state to
-obtain a system of equations at most quadratic in fluid fields:
-\begin{subequations}\label{se:nl_var}
-    \begin{align}
-        \bm{\nabla} \cdot \bm{u} &= 0,\\
-        \pd{\Upsilon}{t} + \p{\bm{u} \cdot \bm{\nabla}} \Upsilon
-            - \frac{u_z}{H} &= 0,\\
-        \pd{u_{x}}{t} + \p{\bm{u} \cdot \bm{\nabla}}u_{x}
-            + \pd{\varpi'}{x} + gH\pd{\Upsilon}{x}
-            + \varpi' \pd{\Upsilon}{x} &= 0,\\
-        \pd{u_z}{t} + \p{\bm{u} \cdot \bm{\nabla}}u_z
-            + \pd{\varpi'}{z} + gH\pd{\Upsilon}{z}
-            + \varpi' \pd{\Upsilon}{z} - \frac{\varpi'}{H} &= 0.
-    \end{align}
-\end{subequations}
-It bears noting that these equations are exactly equivalent to the original
-Euler equations and hence conserve horizontal momentum.
-
-\subsection{Artificial Dissipation}
-
-The nonlinear terms in the above equations will transfer energy from lower
+The system of equations we wish to simulate consists of
+Eqs.~\eqref{eq:nl_incomp},~\eqref{eq:nl_upsilon_u}, and~\eqref{eq:vol_drive}.
+% can be explicitly written in
+% component form and partial derivatives below:
+% \begin{subequations}\label{se:nl_var}
+%     \begin{align}
+%         \bm{\nabla} \cdot \bm{u} &= 0,\\
+%         \pd{\Upsilon}{t} + \p{\bm{u} \cdot \bm{\nabla}} \Upsilon
+%             - \frac{u_z}{H} &= 0,\\
+%         \pd{u_{x}}{t} + \p{\bm{u} \cdot \bm{\nabla}}u_{x}
+%             + \pd{\varpi'}{x} + gH\pd{\Upsilon}{x}
+%             + \varpi' \pd{\Upsilon}{x} &= 0,\\
+%         \pd{u_z}{t} + \p{\bm{u} \cdot \bm{\nabla}}u_z
+%             + \pd{\varpi'}{z} + gH\pd{\Upsilon}{z}
+%             + \varpi' \pd{\Upsilon}{z} - \frac{\varpi'}{H} &= 0.
+%     \end{align}
+% \end{subequations}
+The nonlinear terms in the these equations will transfer energy from lower
 wavenumbers to higher wavenumbers. Since spectral codes have no numerical
@@
 terms to the flux-conservative form of the Euler fluid equations
-\autoref{se:nl_orig} (we use stress tensor $\tau_{ij} = P\delta_{ij}$):
+(equivalent to Eqs.~\ref{se:nl_orig}):
 \begin{subequations}
@@
 \end{subequations}
-The same $\nu$ is used for both the diffusive and viscous term, though this is
-not required. Since the dissipation is not physical and is purely used for
-numerical stability, we choose it such that hydrostatic equilibrium is not
-modified (hence $\nu$ acts only on $\rho - \overline{\rho}$).
-
-One last consideration we found necessary was masking out nonlinear terms in the
-forcing zone with a similar form to \autoref{eq:Gamma}. In the absence of this
-mask, a strong mean flow localized to the forcing zone developed. The mask used
-was
+The same viscosity $\nu$ is used for both the diffusive and viscous terms,
+although this is not required. Since the dissipation is not physical and is
+purely used for numerical stability, we choose it such that hydrostatic
+equilibrium is not modified (hence $\nu$ acts only on $\rho - \overline{\rho}$).
+
+It is necessary to mask out nonlinear terms in the forcing zone using a form
+similar to Eq.~\eqref{eq:Gamma}. In the absence of this mask, a nonphysical mean
+flow localized to the forcing zone develops. We use mask
 \begin{equation}
@@
 
-Including the damping layers and forcing terms as described in
-\autoref{s:numerics}, we finally obtain the full system of equations as
-simulated in Dedalus:
+Including the damping zones and forcing terms as described in
+Section~\ref{s:numerics}, and again making change of variables to $\Upsilon,
+\varpi$, we finally obtain the full system of equations as simulated in Dedalus:
 \begin{subequations}\label{se:dedalus_eqs}

### 2019-12-22 sims/2d_4_fourier/strat_helper.py

@@ -9,7 +9,7 @@ logger = logging.getLogger()
 import os
 from collections import defaultdict
 
-import h5py
+# import h5py
 import matplotlib
 matplotlib.use('Agg')
 import matplotlib.pyplot as plt
@@ -17,12 +17,12 @@ import numpy as np
 from scipy.interpolate import interp1d
 from scipy.optimize import minimize
 
-from dedalus import public as de
-from dedalus.tools import post
-from dedalus.extras.flow_tools import CFL, GlobalFlowProperty
-from dedalus.extras.plot_tools import quad_mesh, pad_limits
-from mpi4py import MPI
-CW = MPI.COMM_WORLD
+# from dedalus import public as de
+# from dedalus.tools import post
+# from dedalus.extras.flow_tools import CFL, GlobalFlowProperty
+# from dedalus.extras.plot_tools import quad_mesh, pad_limits
+# from mpi4py import MPI
+# CW = MPI.COMM_WORLD
 PLT_COLORS = ['b', 'g', 'r', 'c', 'm', 'y', 'k', 'w'];
 
 SNAPSHOTS_DIR = 'snapshots_%s'
@@ -34,7 +34,7 @@ FONTSIZE = 16
 DPI=600
 
 plt.rc('text', usetex=True)
-LW = 2.5
+LW = 3.5
 plt.rc('font', family='serif', size=FONTSIZE)
 
 def populate_globals(var_dict):
@@ -1019,17 +1019,17 @@ def plot_front(name, params):
             plt.plot(z0_cut,
                      S_px[time, z_0_idx: ] / flux_th,
                      linewidth=LW * 0.7,
-                     label=r'$t=%.1f$' % sim_times[time])
+                     label=r'$t=%.1f/N$' % sim_times[time])
         # plt.plot(z0_cut,
         #          np.exp(-k_damp * 2 * (z0_cut - Z0)),
         #          linewidth=LW * 1.5,
         #          label=r'Model')
         plt.xlim(Z0, ZMAX)
-        plt.ylim(-0.2, 1.1)
+        plt.ylim(-0.2, 1.2)
         plt.legend(fontsize=FONTSIZE)
 
-        plt.xlabel(r'$z / H$')
-        plt.ylabel(r'$F / F_{al}$')
+        plt.xlabel(r'$z / H$', fontsize=int(1.5 * FONTSIZE))
+        plt.ylabel(r'$F / F_{al}$', fontsize=int(1.5 * FONTSIZE))
         plt.tight_layout()
[…]

### 2019-12-22 sims/2d_4_fourier/testo/plot_2d_series_movie.py

@@ -1,0 +1,99 @@
+"""
+Plot planes from joint analysis files.
+
+Usage:
+    plot_2d_series_movie.py <files>... [--output=<dir>] [--write=<num>]
+
+Options:
+    --output=<dir>  Output directory [default: ./frames]
+    --write=<num>   Write number offset [default: 0]
+
+"""
+
+import h5py
+import numpy as np
+import matplotlib
+matplotlib.use('Agg')
+import matplotlib.pyplot as plt
+plt.ioff()
+from dedalus.extras import plot_tools
+import re
+
+plt.rc('text', usetex=True)
+plt.rc('font', family='serif', size=12)
+
+def main(filename, start, count, output, write_start):
+    """Save plot of specified tasks for given range of analysis writes."""
+
+    # Plot settings
+    tasks = ['ux','uz']
+    scale = 2.5
+    dpi = 300
+    title_func = lambda sim_time: r'$t = {:.3f}$'.format(sim_time)
+    # savename_func = lambda write: 'yubo_{:06}.png'.format(write+write_start)
+    def savename_func(write):
+        # regex parse out filename number, 10 writes per file
+        num_per_file = 10
+        write_num = (write - 1) % num_per_file
+        filenum = int(re.match('.*s([\d]+)\.h5', filename)[1])
+        suffix_num = (filenum - 1) * num_per_file + write_num
+        if suffix_num > 201:
+            suffix_num -= 9 # hard coded, one of the restores is misnumbered
+        print(suffix_num)
+        return 'yubo_{:06}.png'.format(suffix_num)
+    # Layout
+    nrows, ncols = 1, 2
+    image = plot_tools.Box(1, 2)
+    pad = plot_tools.Frame(0.2, 0.2, 0.1, 0.1)
+    margin = plot_tools.Frame(0.3, 0.2, 0.1, 0.1)
+
+    # Create multifigure
+    mfig = plot_tools.MultiFigure(nrows, ncols, image, pad, margin, scale)
+    fig = mfig.figure
+    # Plot writes
+    with h5py.File(filename, mode='r') as file:
+        for index in range(start, start+count):
+            for n, task in enumerate(tasks):
+                # Build subfigure axes
+                i, j = divmod(n, ncols)
+                axes = mfig.add_axes(i, j, [0, 0, 1, 1])
+                # Call 3D plotting helper, slicing in time
+                dset = file['tasks'][task]
+                if n==0: #ux
[…]

### 2019-12-22 sims/aggregate.py

@@ -56,10 +56,11 @@ if __name__ == '__main__':
         ax3.errorbar(x, w_med,
                      [[w_med - w_min], [w_max - w_med]],
                      ecolor='k', linewidth=lwidth)
-        ax3.set_ylim([0, 1.5])
-        ax3.set_yticks([0, 0.5, 1])
-        ax3.set_xticks([100, 1000])
         offsets[re_inv] += offset_mult * 3
+    ax3.set_xlim([100, 2200])
+    ax3.set_ylim([0, 1.5])
+    ax3.set_yticks([0, 0.5, 1])
+    ax3.set_xticks([100, 1000])
 
     ln1 = mlines.Line2D([], [], color='b', marker='o', markersize=msize,
                         label=r'$\left<\mathcal{R}_A(t)^2\right>$')
@@ -71,6 +72,6 @@ if __name__ == '__main__':
 
     ax3.set_xlabel('Re')
     ax3.set_ylabel('Ri')
-    f.subplots_adjust(hspace=0)
+    f.subplots_adjust(hspace=0.1)
     plt.savefig('agg.png', dpi=400)
     plt.clf()

### 2020-01-22 writeups/Su_IGW_break.tex
Latest comments from Dong

@@
 
-\title[Internal Gravity Wave Breaking]{Physics of Tidal Dissipation:
-Hydrodynamical Simulations of Internal Gravity Wave Breaking in Stellar
-Envelopes and Massive Stellar Binaries}
+\title[Physics of Tidal Dissipation]{Physics of Tidal Dissipation in Early-Type
+Stars and White Dwarfs: Hydrodynamical Simulations of Internal Gravity Wave
+Breaking in Isothermal Stellar Envelopes}
 \author[Y. Su et\ al.]{
@@
 Daniel Lecoanet,$^2$
-Dong Lai,$^1$
+Dong Lai,$^{1,3}$
 \\
@@
 NJ 08544, USA
+\\
+$^3$ Tsung-Dao Lee Institute, 800 Dongchuan Road, Shanghai, 200240, China
 }
@@
 
-\pubyear{2019}
+\pubyear{2020}
 
@@
 \begin{abstract}
-    In sufficiently compact white dwarf binaries, dynamical tides raise a train
-    of internal gravity waves that propagate towards the surface and dissipate
-    via nonlinear wave breaking. We perform 2D hydrodynamical simulations of
-    this wave breaking in an incompressible, isothermal atmosphere. After an
-    initial transient phase, we find that these waves generate a sharp
-    transition, a critical layer, between the non-rotating core and
-    synchronously rotating envelope. We find evidence that indefinite steepening
-    of this critical layer is suppressed by the Kelvin-Helmholtz instability. We
-    study the absorption and reflection of incident waves off the critical layer
-    and provide analytical formulae describing its evolution. We infer
-    necessary conditions to resolving momentum transfer within the critical
-    layer. Finally, we speculate on the application of our model to tidal
-    synchronization and heating in astrophysical systems.
+    In binaries composed of early-type stars and white dwarfs, the dominant
+    tidal process involve the excitation of internal gravity waves (IGWs) that
+    propagate towards the stellar surface and their dissipation via nonlinear
+    wave breaking. We perform 2D hydrodynamical simulations of this wave
+    breaking process in a stratified isothermal atmosphere. We find that, after
+    an initial transient phase, the dissipation of the IGWs naturally generate a
+    sharp critical layer, separating the lower stationary region (with no mean
+    flow) and the upper ``synchronized'' envelope (with the mean flow speed
+    equal to the horizontal wave phase velocity). We find that the indefinite
+    steepening of this critical layer is suppressed by the Kelvin-Helmholtz
+    instability, and the width of the layer is determined by the Richardson
+    criterion. We study the absorption and reflection of incident waves off the
+    critical layer and provide analytical formulae describing its evolution. The
+    result of this study is important for characterizing the evolution of
+    tidally heated white dwarfs and other binary stars.
 \end{abstract}
@@
 
+% so that \cite[see][for q]{QQ} gives (see QQ for q) without comma after QQ
+\setcitestyle{notesep={ }}
+
 \section{Introduction}\label{s:intro}
 
-Compact white dwarf (WD) binary systems, with orbital periods in the range of
-minutes to hours, are important for a range of astrophysical problems. They are
-the most important sources of gravitational waves (GWs) for the Laser
-Interferometric Space Antenna (LISA) \citep{lisa}. They are also thought to
-produce interesting optical transients such as underluminous supernovae
-\citep{underlum}, Ca-rich fast transients \citep{carich}, and tidal novae
-\citep{tidal_novae}. Most importantly, they have been proposed as the likely
-progenitors of type Ia supernovae (e.g. \citet{Ia0,webbink} or more recently
-\citet{Ia1,Ia2}). While presently only a few tens of compact WD binaries are
-known \citep{lsst_wd}, \emph{Gaia} (currently gathering data) is expected to
-expand the catalog to a few hundreds \citep{lsst_wd} \citep[results based on
-\emph{Gaia}'s second data release have already begun to
-appear][]{gaiaDD,gaiaDD2}, and the Large Synoptic Survey Telescope (LSST, first
-light scheduled for 2020) will likely detect a few thousand more
-\citep{lsst_wd}. These observations will significantly advance the understanding
-of WD binaries and their evolution.
-
-In spite of the broad importance of WD binaries, the evolution of these systems
-prior to their final mergers is not well understood. Much of this uncertainty
-comes from our imprecise understanding of tidal interactions, which play an
-important role during a compact WD binary's inspiral \citep{fullerII}. Previous
-studies have shown that these interactions manifest as tidal excitation of
-internal gravity waves (IGW), waves in the WD fluid restored by the buoyancy
-force due to density stratification \citep{fullerI}. As these waves propagate
-outwards towards the WD surface, they grow in amplitude until they break, as do
-ocean waves on a shore, and transfer both energy and angular momentum from the
-binary orbit to the outer envelope of the WD \citep{fullerI,fullerII}.
-
-Previous works have found that the dissipation of IGW can generate significantly
-more energy than thermal radiation from the isolated WD surface and is thus a
-major contributor to the WD energy budget \citep{fullerII,fullerIV}. However,
-these works parameterized the wave breaking process in an ad hoc manner. The
-details of dissipation, namely the location and spatial extent of the wave
-breaking, affect the observable outcome: dissipation near the surface of the WD
-can be efficiently radiated away and simply brightens the WD, while dissipation
-deep in the WD envelope causes an energy buildup that results in energetic
-flares \citep{tidal_novae}. Works in other fields based on numerical simulations
-show that strongly nonlinear wave breaking behaves differently than predictions
-based in linear and weakly nonlinear theory \citep{winters1994,barker_ogilvie}.
-Such fully nonlinear numerical simulations have not been performed for WDs.
-
-In this paper, we perform numerical simulations of tidally-excited IGWs
-undergoing wave breaking in an idealized WD fluid. The primary aim of the paper
-is to thoroughly characterize the angular momentum transfer within a WD by
-purely hydrodynamical effects. As a first approximation, we analyze the problem
-in a 2D Cartesian geometry, and consider IGW propagating into a fluid initially
-at rest. We find that, after an initial transient phase, the fluid develops two
-distinct zones: (i) a lower zone that has no horizontal mean flow, and (ii) an
-upper zone with significant horizontal mean flow.
-
-These two zones are separated by a \emph{critical layer}. The interaction of IGW
-and critical layers plays an important role in the tidal synchronization process
-of stellar binaries \citep{zahn75,gn89} as well as in terrestrial phenomena such
-as the quasi-biennial oscillation \citep{lindzen_qbo}. The majority of our paper
-is dedicated to characterizing the behavior of the critical layer when
-interacting with a continuous train of IGW excited from the bottom of the fluid.
-IGW are generally \emph{anti}-diffusive, in that they steepen shear flows
-\citep{lindzen_qbo,lecoanet_meanflow}, and act to narrow the critical layer. We
-find this steepening is suppressed by the Kelvin-Helmholtz instability and
-turbulence within the very narrow critical layer. By careful accounting of the
-momentum flux budget about the critical layer, we are able to model the
-propagation of the critical layer, reflection of the incident IGW, and other
-momentum redistribution behaviors. Our results are numerically converged for
-sufficiently large resolutions.
-
-In Section~\ref{s:equations}, we will describe the system of equations we will
-use to analyze IGW breaking. In Section~\ref{s:theory}, we review existing
-understanding of wave breaking and present new analytical results. In
-Section~\ref{s:numerics}, we describe our numerical setup, which in
-Section~\ref{s:weak_sim} we validate in the weak forcing limit against linear
-theory. In Section~\ref{s:sim}, we present the results of simulation of IGW
-breaking and our characterization of the critical layer. Finally, we summarize
-and conclude in Section~\ref{s:discussion}.
+The physical processes responsible for tidal evolution in close binaries often
+involve the excitation and dissipation of internal waves, going beyond the
+``weak friction'' of equilibrium tides \citep[see][for a
+review]{ogilvie2014tidal}. In particular, internal gravity waves (IGWs), arising
+from buoyancy of stratified stellar fluid, play an important role in several
+types of binary systems. In solar-type stars with radiative cores and convective
+envelopes, IGWs are excited by tidal forcing at the radiative-convective
+boundary and propagate inward; as the wave amplitude grows due to geometric
+focusing, nonlinear effects can lead to efficient damping of the wave
+\citep{goodman1998dynamical,barker_ogilvie,essick2015orbital}. In massive
+main-sequence stars, with convective cores and radiative envelopes, IGWs are
+similarly excited at the convective-radiative interface but travel toward the
+stellar surface; nonlinearity develops as the wave amplitude grows, leading to
+efficient dissipation \citep{zahn75,zahn77}. As the outgoing wave deposites its
+angular momentum to the stellar surface layer, a critical layer may form and the
+star synchronize from outside-in \citep{gn89}.
+
+Tidal dissipation can also play an important role in compact double white dwarf
+(WD) binary systems (with orbital periods in the range of minutes to hours).
+Such binaries may produce a variety of exotic astrophysical systems and
+phenomena, ranging from isolated sdB/sdO stars,
+%(likely created by the merger of two He WDs),
+R CrB stars,
+%(likely created by the merger of a CO and He WD),
+AM CVn binaries,
+%(possibly created by stable mass transfer from a He WD onto a CO WD),
+high-mass neutron stars and magnetars (created by the accretion-induced
+collapse of merging WDs),
+%CO WDs or accreting O-Ne-Mg WDs),
+and various optical transients (underluminous supernovae, Ca-rich fast
+transients, and type Ia supernovae)
+\citep[e.g.][]{livio2018progenitors,gaiaDD2,toloza2019understanding,
+underlum,carich,Ia1,Ia2}. The outcomes of WD mergers depend on the WD masses
+and composition, but tidal dissipation can strongly affect the pre-merger
+conditions of the WDs. Tidal dessipation may also incluence the evoluton of
+eccentric WD-massive black hole binaries prior to the eventual tidal disruption
+of the WD \citep{vick2017tidal}.
+
+Previous studies have identified nonlinear dissipation of IGWs as the key tidal
+process in compact WD binaries \citep{fullerII,fullerIV,tidal_novae,bukart}:
+IGWs are tidally excited mainly at the compositon transitions of the WD
+envelope; as these waves propagate outwards towards the WD surface, they grow
+in amplitude until they break, and transfer both energy and angular momentum
+from the binary orbit to the outer envelope of the WD\@.
+%Fuller \& Lai (20xx,20xx) found that the dissipation of IGW can generate
+%significantly more energy than thermal radiation from the isolated WD surface
+%and is thus a major contributor to the WD energy budget.
+However, these previous works parameterized the wave breaking process in an ad
+hoc manner. The details of dissipation, namely the location and spatial extent
+of the wave breaking, affect the observable outcome: Dissipation near the
+surface of the WD can be efficiently radiated away and simply brightens the WD,
+while dissipation deep in the WD envelope causes an energy buildup that results
+in energetic flares \citep{tidal_novae}. An important goal of this paper is to
+elucidate the details of the nonlinear IGW breaking process; the result of this
+``microphysics'' study will help determine the observational manifestation
+tidally heated binary WDs.
+
+In this paper, we perform numerical simulations of IGW breaking in a
+plane-parallel stratified atmosphere (a stellar enevelope). We use the
+pseudo-spectral code Dedalus \citep{dedalus,dedalus2} and a 2D Cartesian
+geometry, and consider IGWs propagating into an isothermal fluid initially at
+rest. We find that, after an initial transient phase, a \emph{critical layer}
+naturally develops, separating a lower zone that has no horizontal mean flow
+and an upper zone with mean flow at the horizontal phase velocity of the IGW\@.
+The major part of our paper is dedicated to characterizing the behavior of the
+critical layer when interacting with a continuous train of IGW excited from the
+bottom of the fluid. IGWs are generally \emph{anti}-diffusive, in that they
+steepen shear flows \citep{lindzen_qbo,lecoanet_meanflow} and act to narrow the
+critical layer. We find this steepening is suppressed by the Kelvin-Helmholtz
+instability and turbulence within the narrow critical layer. By careful
+accounting of the momentum flux budget about the critical layer, we are able to
+model the reflection and absorption of the incident IGW, and the slow downward
+propagation of the critical layer.
+
+While the motivation of our study is to understand tidal dissipation
+in WD and mass star binaries, the IGW breaking process studied in this paper
+is also quite relevant to the circulation dynamics of planetary atmospheres
+\citep[see e.g.][]{lindzen1981turbulence,holton1983influence}.
+
+This paper is organized as follows. In Section~\ref{s:equations} we present the
+system of equations used in our simulations. In Section~\ref{s:theory}, we
+review the existing understanding of wave breaking and present new analytical
+results on IGWs around a critical layer. In Section~\ref{s:numerics} we describe
+our numerical setup and in Section~\ref{s:weak_sim} we validate our method in
+the weak forcing limit against linear theory. In Section~\ref{s:sim}, we present
+the results of simulations of IGW breaking and our characterization of the
+critical layer. We summarize and conclude in Section~\ref{s:discussion}.
 
@@
         \begin{equation}
-            \mathrm{Re}  \equiv \frac{\omega}{\nu k_{z}^2}. \label{eq:re_def}
-        \end{equation}
-        When the forcing is weak, we may set $\nu \approx 0$, or $\mathrm{Re}
-        \gg 1$ because there is negligible energy transfer into higher-order
-        modes. Note that $\nu = 0$ changes the order of the PDE system, as the
-        highest-order derivatives appear only in the dissipitive terms, so a
-        nonzero $\nu$ must be chosen even in weakly forced simulations.
+            \mathrm{Re} \equiv \frac{\omega}{\nu k_{z}^2}
+                = \frac{\omega}{\nu}\p{\frac{H}{2\pi}}^2.\label{eq:re_def}
+        \end{equation}
+        We use $\mathrm{Re} \gg 1$ in our simulations\footenote{This condition
+        is always satisfied in stars. For example, in WDs, the dominant linear
+        dissipation mechanism of g-modes is radiative damping, with damping rate
+        ranging from $10^{-11}$--$10^-4$ of the mode frequency \citep{fullerI}.
+        This corresponds to a small effective viscosity or $\mathrm{Re} \gg
+        1$.}. Note that $\nu = 0$ changes the order of the PDE system, as the
+        highest-order derivatives appear only in the dissipitive terms, so only
+        nonzero $\nu$ can be used even in weakly forced simulations.
 \end{enumerate}
@@
 
-\subsection{Physical Sources of Dissipation in WDs}\label{ss:disp}
-
-The most significant linear damping in WD g-modes comes from radiative damping
-\citep{fullerI}. In \citep{wu} and \citep{fullerI}, the radiative damping rate
-is given in terms of $\omega_i = \gamma \omega_r$, where $\omega_r$ is the
-frequency of the g-mode. Typical values for $\gamma$ range from $10^{-4}$ to
-$10^{-11}$ depending on $n$ of the g-mode.
-
-We will assume this prescription directly transfers to propagating IGW, which
-results in general agreement with \citep{bukart}'s estimate of radiative damping
-rates. Then, making coarse identification $\omega_i \sim \nu k^2 \approx \nu
-k_z^2$, we find that $\mathrm{Re} \sim \frac{1}{\gamma}$. Even at $\gamma =
-10^{-4}$ however, the corresponding $\mathrm{Re}$ is far too weak to suppress
-reflection/transmission at the critical layer (Fig.~\ref{fig:agg}).
-
-Another source of dissipation considered in \citep{bukart} is turbulent
-conbmtive damping. They find this damping rate to never exceed that of
-radiative damping, and so it is also too weak to suppress critical layer
-formation in our problem.
-
-Finally, we consider the impact of magnetic winding. In \citep{bukart}, magnetic
-winding is used to enforce solid body rotation on the grounds that $t_A \gg
-t_{gw}$, where
-\begin{equation}
-    t_A = \int\limits_0^R \frac{\sqrt{4\pi \rho}}{B_0}\;\mathrm{d}r
-        \sim 10^2\;\mathrm{yr}\p{\frac{10^3\;\mathrm{G}}{B}},
-\end{equation}
-the Alfv\'en wave crossing time (evaluated for a CO WD in \citet{fullerIV}),
-measures the magnetic coupling time and $t_{gw}$ measures the gravitational wave
-inspiral timescale. Before solid body rotation is attained, another relevant
-timescale is the synchronization timescale $t_s$. For a tidal torque $\tau$ and
-tidal forcing frequency $\sigma = m\p{\Omega - \Omega_{spin}}$, we note that
-angular momentum transfer is $\pd{M_{sync}}{t} \sigma R^2 = \tau$, where
-$M_{sync}$ denotes the mass of the WD that has synchronized. Thus, the
-synchronization timescale is
-\begin{align}
-    t_{sync} &\sim \frac{M_{sync}\sigma R^2}{\tau},\\
-        &\sim 20\;\mathrm{yr}
-            \p{\frac{M_{sync}}{10^{-4}M_{\odot}}}
-            \p{\frac{\sigma}{2\pi / (1\;\mathrm{hr})}}
-            \p{\frac{R}{R_{\oplus}}}^2
-            \p{\frac{10^{-14} GM_{\odot}^2/R_{\oplus}}{\tau}}.
-\end{align}
-A representative $\tau$ has been taken from \citep{bukart}. Only $M_{sync} \sim
-10^{-4}M_{\odot}$ need be heated for the thermodynamically interesting effects
-studied in \citep{fullerIV} and \citep{tidal_novae}, it seems that strong shear
-flows may exist even in the presence of magnetic coupling.
-
-\subsection{Applicability to Astrophysical Systems}\label{ss:other}
-
-Below, we speculate on the applicability of our results to general astrophysical
-problems of interest:
-\begin{enumerate}
-    \item In astrophysical systems, $k_{\perp} \ll k_r$, while in our study,
-        $k_x \simeq k_z$. True turbulence is expected to be isotropic at small
-        scales, which may introduce different behavior for different ratios of
-        $k_z/k_x$. However, $k_x \ll k_z$ is a numerically difficult regime; we
-        defer its exploration to future work.
-
-    \item In~\cite{barker_ogilvie}, inwards-propagating IGW are excited that
-        break via geometric focusing. They find no reflected wave despite their
-        nonlinear timescales being $10\times$ shorter than their viscous
-        timescale, in contrast with the reflective behavior observed in our
-        simulations. It is plausible the discrepancy arises due to the differing
-        causes of wave breaking (geometric focusing vs.\ density
-        stratification), but their simulations are also at significantly higher
-        artificial viscosities than ours, as we show below.
-
-        Associating $t_{L} \sim \nu k^2 \approx \nu k_z^2$ with the visocus
-        timescale and $t_{NL} \sim \bm{u} \cdot \bm{\nabla} \sim \omega$ for the
-        nonlinear timescale, we find their $\lambda = \frac{t_{NL}}{t_L} \sim
-        \mathrm{Re}$ our Reynolds number. Our simulations indicate $\mathrm{Re}
-        \gtrsim 500$ are required to observe the correct asymptotic behavior in
-        terms of horizontal momentum flux reflection/transmission, so it is
-        possible their lack of reflection is resolution limited.
-
-    \item Our results give some tentative understanding of where tidal heating
-        is expected to occur. While we use a barotropic equation of state and do
-        not track heat deposition, we can recall that the strongest small-scale
-        variations in fluid flow are localized within the critical layer. Thus,
-        we posit that a substantial fraction of the energy carried by the IGW is
-        deposited in the very narrow critical layer. Then, as the critical layer
-        propagates downwards, the entire star is heated. This tidal heating
-        model differs flom that used in \citep{tidal_novae}, and its effects
-        will be studied in future work.
-\end{enumerate}
+\subsection{Discussion}
+
+In this paper, we have studied the nonlinear behavior of IGWs with $\abs{k_x /
+k_z} \sim 1$ in a plane-parallel atmospheres. On the other hand, tidally
+excited IGWs in binary stars have horizontal wave number $k_\perp \sim 1/R$
+(where $R$ is the stellar radius) much smaller than the radial wave number
+$k_r$. Strictly speaking, our results may not directly apply to IGWs with
+$\abs{k_x/k_z}\ll 1$. However the qualitative behavior is likely to be similar,
+as the turbulence driving the critical layer dynamics occurs at scales
+significantly smaller than either $1/k_x$ or $1/\abs{k_z}$. Simulating IGWs with
+$k_x \ll k_z$ is more numerically challenging and we defer its exploration to
+future work.
+
+It is interesting to compare our work with that of \citet{barker_ogilvie}, who
+studied inward-propagating IGWs in solar-type stars and their nonlinear breaking
+due to geometric focusing. In their numerical simulation in a 2D polar geometry,
+they find no evidence for reflected waves, contrary to our result. It is
+plausible the discrepancy arises due to the differing causes of wave breaking
+(geometric focusing vs.\ density rarefaction). Additionally, their nonlinearity
+parameter $\lambda \sim 1/\mathrm{Re}$ where $\mathrm{Re}$ is our Reynolds
+number (Eq.~\eqref{eq:re_def}). For $\lambda \sim 10^{-1}$ as explored in
+\citet{barker_ogilvie}, our results also predict no reflected wave.
+
+Regardless of the limitations inherent in our simulations (e.g.\ plane-parallel
+geometry), our results shed light on the physical mechanism of tidal heating in
+close binaries. In particular, our simulations indicate that energy dissipation
+occurs in a narrow critical layer. The star heats up from outside-in as the
+critical layer propagates inwards. This tidal heating profile differs flom that
+used by \citet{tidal_novae}. We plan to study this issue in a future work.
 

### 2020-02-05 writeups/Su_IGW_break.tex
updated draft

@@
         usenatbib,
-        referee,
+        % referee,
     ]{mnras}
@@
 Stars and White Dwarfs: Hydrodynamical Simulations of Internal Gravity Wave
-Breaking in Isothermal Stellar Envelopes}
+Breaking in Stellar Envelopes}
 \author[Y. Su et\ al.]{
@@
 Daniel Lecoanet,$^2$
-Dong Lai,$^{1,3}$
+Dong Lai$^{1,3}$
 \\
@@
 \\
-$^3$ Tsung-Dao Lee Institute, 800 Dongchuan Road, Shanghai, 200240, China
+$^3$ Tsung-Dao Lee Institute, Shanghai Jiao Tong University, Shanghai, 200240, China
 }
@@
     In binaries composed of early-type stars and white dwarfs, the dominant
-    tidal process involve the excitation of internal gravity waves (IGWs) that
+    tidal process involves the excitation of internal gravity waves (IGWs) that
     propagate towards the stellar surface and their dissipation via nonlinear
@@
     sharp critical layer, separating the lower stationary region (with no mean
-    flow) and the upper ``synchronized'' envelope (with the mean flow speed
-    equal to the horizontal wave phase velocity). We find that the indefinite
-    steepening of this critical layer is suppressed by the Kelvin-Helmholtz
-    instability, and the width of the layer is determined by the Richardson
+    flow) and the upper ``synchronized'' region (with the mean flow velocity
+    equal to the horizontal wave phase speed). We find that the indefinite
+    steepening of this critical layer is suppressed by Kelvin-Helmholtz
+    instabilities, and the width of the layer is determined by the Richardson
     criterion. We study the absorption and reflection of incident waves off the
@@
 focusing, nonlinear effects can lead to efficient damping of the wave
-\citep{goodman1998dynamical,barker_ogilvie,essick2015orbital}. In massive
+\citep{goodman1998dynamical,barker_ogilvie,essick2015orbital}. In early-type
 main-sequence stars, with convective cores and radiative envelopes, IGWs are
@@
 angular momentum to the stellar surface layer, a critical layer may form and the
-star synchronize from outside-in \citep{gn89}.
+star is expected to synchronize from outside-in \citep{gn89}.
 
@@
 Such binaries may produce a variety of exotic astrophysical systems and
-phenomena, ranging from isolated sdB/sdO stars,
-%(likely created by the merger of two He WDs),
-R CrB stars,
-%(likely created by the merger of a CO and He WD),
-AM CVn binaries,
-%(possibly created by stable mass transfer from a He WD onto a CO WD),
-high-mass neutron stars and magnetars (created by the accretion-induced
-collapse of merging WDs),
-%CO WDs or accreting O-Ne-Mg WDs),
-and various optical transients (underluminous supernovae, Ca-rich fast
-transients, and type Ia supernovae)
-\citep[e.g.][]{livio2018progenitors,gaiaDD2,toloza2019understanding,
-underlum,carich,Ia1,Ia2}. The outcomes of WD mergers depend on the WD masses
-and composition, but tidal dissipation can strongly affect the pre-merger
-conditions of the WDs. Tidal dessipation may also incluence the evoluton of
-eccentric WD-massive black hole binaries prior to the eventual tidal disruption
-of the WD \citep{vick2017tidal}.
-
-Previous studies have identified nonlinear dissipation of IGWs as the key tidal
+phenomena, ranging from isolated sdB/sdO stars, R CrB stars, AM CVn binaries,
+high-mass neutron stars and magnetars (created by the accretion-induced collapse
+of merging WDs), and various optical transients (underluminous supernovae,
+Ca-rich fast transients, and type Ia supernovae)
+\citep[e.g.][]{livio2018progenitors,toloza2019understanding}. The outcomes of WD
+mergers depend on the WD masses and composition, but tidal dissipation can
+strongly affect the pre-merger conditions of the WDs and therefore the merger
+outcomes. Tidal dessipation may also influence the evoluton of eccentric
+WD-massive black hole binaries prior to the eventual tidal disruption of the WD
+\citep{vick2017tidal}.
+
+Recent studies have identified nonlinear dissipation of IGWs as the key tidal
 process in compact WD binaries \citep{fullerII,fullerIV,tidal_novae,bukart}:
 IGWs are tidally excited mainly at the compositon transitions of the WD
-envelope; as these waves propagate outwards towards the WD surface, they grow
-in amplitude until they break, and transfer both energy and angular momentum
-from the binary orbit to the outer envelope of the WD\@.
-%Fuller \& Lai (20xx,20xx) found that the dissipation of IGW can generate
-%significantly more energy than thermal radiation from the isolated WD surface
-%and is thus a major contributor to the WD energy budget.
-However, these previous works parameterized the wave breaking process in an ad
-hoc manner. The details of dissipation, namely the location and spatial extent
-of the wave breaking, affect the observable outcome: Dissipation near the
-surface of the WD can be efficiently radiated away and simply brightens the WD,
-while dissipation deep in the WD envelope causes an energy buildup that results
-in energetic flares \citep{tidal_novae}. An important goal of this paper is to
-elucidate the details of the nonlinear IGW breaking process; the result of this
-``microphysics'' study will help determine the observational manifestation
-tidally heated binary WDs.
+envelope; as these waves propagate outwards towards the WD surface, they grow in
+amplitude until they break, and transfer both energy and angular momentum from
+the binary orbit to the outer envelope of the WD\@. However, these previous
+works parameterized the wave breaking process in an ad hoc manner. The details
+of dissipation, namely the location and spatial extent of the wave breaking,
+affect the observable outcomes: dissipation near the surface of the WD can be
+efficiently radiated away and simply brightens the WD, while dissipation deep in
+the WD envelope causes an energy buildup that results in energetic flares
+\citep{tidal_novae}. An important goal of this paper is to elucidate the details
+of the nonlinear IGW breaking process; the result of this ``microphysics'' study
+will help determine the thermal evolution and and the observational
+manifestations of tidally heated binary WDs.
 
 In this paper, we perform numerical simulations of IGW breaking in a
-plane-parallel stratified atmosphere (a stellar enevelope). We use the
+plane-parallel stratified atmosphere (stellar enevelope). We use the
 pseudo-spectral code Dedalus \citep{dedalus,dedalus2} and a 2D Cartesian
@@
 critical layer when interacting with a continuous train of IGW excited from the
-bottom of the fluid. IGWs are generally \emph{anti}-diffusive, in that they
+bottom of the atmosphere. IGWs are generally \emph{anti}-diffusive, in that they
 steepen shear flows \citep{lindzen_qbo,lecoanet_meanflow} and act to narrow the
@@
 
-While the motivation of our study is to understand tidal dissipation
-in WD and mass star binaries, the IGW breaking process studied in this paper
-is also quite relevant to the circulation dynamics of planetary atmospheres
+While the motivation of our study is to understand tidal dissipation in WD and
+early-type stellar binaries, the IGW breaking process studied in this paper is
+also quite relevant to the circulation dynamics of planetary atmospheres
 \citep[see e.g.][]{lindzen1981turbulence,holton1983influence}.
@@
 our numerical setup and in Section~\ref{s:weak_sim} we validate our method in
-the weak forcing limit against linear theory. In Section~\ref{s:sim}, we present
+the weak-forcing limit against linear theory. In Section~\ref{s:sim}, we present
 the results of simulations of IGW breaking and our characterization of the
@@
 
-\section{Problem Setup}\label{s:equations}
+\section{Problem Setup and Equations}\label{s:equations}
 
@@
 \end{equation}
-with $\overline{\rho}_0$ some reference density (we generally denote background
-quantities with overbars and perturbation quantities with primes).
+with $\overline{\rho}_0$ some reference density. We generally denote background
+quantities with overbars and perturbation quantities with primes.
 
@@
 
-In the small perturbation limit, we may linearize
-Eqs.~\eqref{se:nl_upsilon}. The solution in this linear regime is given by
-\citep{drazin,sutherland0}
+In the small perturbation limit, we may linearize Eq.~\eqref{se:nl_upsilon}. The
+solution in this linear regime is given by \citep{drazin,sutherland0}
 \begin{equation}
@@
 Our equations are valid in the limit of large sound speed ($c_s \to \infty$), in
-which the \emph{Brunt-V\"ais\"al\"a frequency} $N$ is given by
+which the \emph{Brunt-V\"ais\"al\"a frequency}, $N$, is given by
 \begin{equation}
@@
         We note $\bm{c}_{p} \cdot \bm{c}_g = \mathcal{O}\s{\p{k_{z}H}^{-2}}
-        \approx 0$. In the Boussinesq approximation where terms
+        \approx 0$. In the Boussinesq approximation where terms of order
         $\mathcal{O}\p{H^{-2}}$ are ignored, the phase and group velocities are
@@
         The notation $\ev{\dots}_x$ denotes averaging over the $x$ domain. For
-        the linear solution (Eq.~\eqref{eq:lin_sol}), this evaluates to
+        the linear solution (Eq.~\ref{eq:lin_sol}), this evaluates to
         \begin{equation}
@@
 propagating upwards, imitating a wave tidally excited deeper in the star, and
-downwards, which are not physically meaningful, but are dissipated by a
-damping zone described in Section~\ref{ss:damping}.
+downwards, which are not physically relevant in binaries. In our simulations,
+these downward propagating waves are dissipated by a damping zone described in
+Section~\ref{ss:damping}.
 
@@
 on the density equation. We implement forcing with strength $C$ localized around
-$z_0$ with small width $\sigma$ by replacing Eq.~\eqref{eq:nl_up_density}
+height $z_0$ with small width $\sigma$ by replacing Eq.~\eqref{eq:nl_up_density}
 with
@@
 Using a narrow Gaussian profile excites a broad $z$ power spectrum, but only the
-$k_{z}$ satisfying dispersion relation (Eq.~\eqref{eq:disp_rel}) for the given
+$k_{z}$ satisfying the dispersion relation (Eq.~\ref{eq:disp_rel}) for the given
 $k_{x}$ and $\omega$ will propagate.
@@
 \end{align}
-The $z > z_0$ region models an upward propagating IGW wavetrain. The $x$
+The $z > z_0$ region contains an upward propagating IGW wavetrain. The $x$
 component of the velocity can be obtained by the incompressibility constraint
-(Eq.~\eqref{eq:nl_incomp}).
+(Eq.~\ref{eq:nl_incomp}).
 
@@
 e^{z/2H}$), it is expected to break due to nonlinear effects. We can estimate
-the height of wave breaking as when $\abs{\bm{u}} \sim \omega / \abs{\bm{k}}$.
-This can be rewritten using the Lagrangian displacement $\bm{\xi} = \bm{u} /
-\p{-i\omega}$:
-\begin{equation}
-    \abs{\xi_z k_z} \gtrsim 1.\label{eq:nl}
+the height of wave breaking using the condition $\abs{\bm{u}} \sim \omega /
+\abs{\bm{k}}$. This can be rewritten using the Lagrangian displacement $\bm{\xi}
+= \bm{u} / \p{-i\omega}$:
+\begin{equation}
+    \abs{\xi_z k_z} \sim 1.\label{eq:nl}
 \end{equation}
@@
 some detail. At intermediate amplitudes, wave breaking occurs via triadic
-resonances which transfer energy from the ``parent’’ IGW to ``daughter’’ waves
-on smaller length scales that efficiently damp. The horizontal momentum flux
+resonances, transfering energy from the ``parent’’ IGW to ``daughter’’ waves on
+smaller length scales that efficiently damp. The horizontal momentum flux
 decreases from $F$ to $0$ over this breaking region. The lost flux is deposited
@@
 The behavior of an IGW incident upon a critical layer was first studied in the
-inviscid, linear regime in \citep{booker_bretherton}, which found nearly
+inviscid, linear regime in \citet{booker_bretherton}, which found nearly
 complete absorption of the IGW\@. The amplitude
@@
 
-Instead, we consider the horizontal momentum transfer at the critical layer. Any
-incident horizontal momentum flux $F_a(t)$ absorbed by the fluid must manifest
-as additional horizontal momentum of the shear flow. Since $\overline{U}$ cannot
-exceed $\overline{U}_c$, the critical layer will instead propagate downward in
-response to the incident momentum flux. The horizontal momentum of the shear
-flow satisfies
+Consider now the long-term evolution of the critical layer due to continuous
+horizontal momentum transfer by IGWs. Any incident horizontal momentum flux
+$F_a(t)$ absorbed by the fluid must manifest as additional horizontal momentum
+of the shear flow. Since $\overline{U}$ cannot exceed $\overline{U}_c$, the
+critical layer will instead propagate downward in response to the incident
+momentum flux. The horizontal momentum of the shear flow satisfies
 \begin{equation}
@@
 Assuming $\overline{U}(z > z_c) \approx \overline{U}_c$ and $\overline{U}(z <
-z_c) \approx 0$ are both approximately constant, this condition becomes
+z_c) \approx 0$, this condition becomes
 \begin{equation}
@@
 \end{equation}
-If $F_a$ is constant in time, $z_c(t)$ has as analytical solution
+If $F_a$ is constant in time, the height of the critical layer $z_c(t)$ has
+analytical solution:
 \begin{equation}
@@
         \end{equation}
-        We use $\mathrm{Re} \gg 1$ in our simulations\footenote{This condition
+        We use $\mathrm{Re} \gg 1$ in our simulations\footnote{This condition
         is always satisfied in stars. For example, in WDs, the dominant linear
@@
 IGWs in a stratified isothermal atmosphere. Such a setup represents the
-plane-parallel idealization of the outer envelope of massive stars, WDs, or
-planets. Our simulations use the spectral code Dedalus \citep{dedalus,dedalus2},
-and are carried out in 2D. We observe spontaneous formation of a critical layer
-that separates a synchronized upper layer of fluid and a lower layer of fluid
-with no mean horizontal flow. This critical layer then propagates downwards as a
-continuous train of IGW are incident on it (see Fig.~\ref{fig:snapshots} for
-snapshots from our fiducial simulation). Our primary conclusions regarding the
-evolution of the critical layer are as
-follows:
+plane-parallel idealization of the outer stellar envelope. Our simulations use
+the spectral code Dedalus \citep{dedalus,dedalus2}, and are carried out in 2D.
+We observe spontaneous formation of a critical layer that separates a
+``synchronized'' upper layer of fluid and a lower layer with no mean
+horizontal flow. This critical layer then propagates downwards as incident IGWs
+break and deposit horizontal momentum to the fluid (see Fig.~\ref{fig:snapshots}
+for snapshots from our fiducial simulation). Our primary conclusions regarding
+the evolution of the critical layer are as follows:
 \begin{enumerate}
@@
         the local Richardson number (Eq.~\eqref{eq:ri_def}) $\mathrm{Ri} \gtrsim
-        0.5$, as shown in Fig.~\ref{fig:nl_f_ri}.
+        0.5$ (see Fig.~\ref{fig:nl_f_ri}).
 
     \item The location of the critical layer $z_c(t)$ can be predicted by
-        careful measurement of the absorbed flux at the critical layer
-        (see Eq.~\eqref{eq:zc_anal}) as shown in Fig.~\ref{fig:nl_front}.
-
-    \item Flux absorption at the critical layer is incomplete. The
-        critical layer only absorbs $\sim 70\%$ of the incident flux in most
-        simulations (see Fig.~\ref{fig:agg} for convergence at high
-        resolutions). The reflected flux is carried away from the critical layer
-        as both a lowest-order reflected wave and waves with larger $z$
-        wavenumbers.
+        careful measurement of the absorbed horizontal momentum flux at the
+        critical layer (see Eq.~\eqref{eq:zc_anal} and
+        Fig.~\ref{fig:nl_front}).
+
+    \item The absorption of IGW momentum flux at the critical layer is
+        incomplete. The critical layer only absorbs $\sim 70\%$ of the incident
+        flux in most simulations (see Fig.~\ref{fig:agg}). The reflected flux is
+        carried away from the critical layer as both lowest-order reflected
+        waves and waves with larger $z$ wavenumbers.
 \end{enumerate}
-In the subsequent sections, we will discuss the validity of these results and
-their application to astrophysical systems.
 
@@
 In this paper, we have studied the nonlinear behavior of IGWs with $\abs{k_x /
-k_z} \sim 1$ in a plane-parallel atmospheres. On the other hand, tidally
-excited IGWs in binary stars have horizontal wave number $k_\perp \sim 1/R$
-(where $R$ is the stellar radius) much smaller than the radial wave number
-$k_r$. Strictly speaking, our results may not directly apply to IGWs with
-$\abs{k_x/k_z}\ll 1$. However the qualitative behavior is likely to be similar,
-as the turbulence driving the critical layer dynamics occurs at scales
-significantly smaller than either $1/k_x$ or $1/\abs{k_z}$. Simulating IGWs with
-$k_x \ll k_z$ is more numerically challenging and we defer its exploration to
-future work.
+k_z} \sim 1$ in a plane-parallel geometry. Tidally excited IGWs in binary stars
+have horizontal wave number $k_\perp \sim 1/R$ (where $R$ is the stellar radius)
+much smaller than the radial wave number $k_r$. Strictly speaking, our results
+may not directly apply to IGWs with $\abs{k_x/k_z}\ll 1$. However, the
+qualitative behavior is likely to be similar, as the turbulence driving the
+critical layer dynamics occurs at scales significantly smaller than either
+$1/k_x$ or $1/\abs{k_z}$. Simulating IGWs with $k_x \ll \abs{k_z}$ is more
+challenging numerically and we defer its exploration to future work.
 
@@
 studied inward-propagating IGWs in solar-type stars and their nonlinear breaking
-due to geometric focusing. In their numerical simulation in a 2D polar geometry,
-they find no evidence for reflected waves, contrary to our result. It is
-plausible the discrepancy arises due to the differing causes of wave breaking
-(geometric focusing vs.\ density rarefaction). Additionally, their nonlinearity
-parameter $\lambda \sim 1/\mathrm{Re}$ where $\mathrm{Re}$ is our Reynolds
-number (Eq.~\eqref{eq:re_def}). For $\lambda \sim 10^{-1}$ as explored in
-\citet{barker_ogilvie}, our results also predict no reflected wave.
+due to geometric focusing. In their numerical simulations in a 2D polar
+geometry, they found no evidence for reflected waves, contrary to our result. It
+is plausible that the discrepancy arises due to the differing causes of wave
+breaking (geometric focusing vs.\ density rarefaction). Note that their
+nonlinearity parameter $\lambda \sim 1/\mathrm{Re}$ where $\mathrm{Re}$ is the
+Reynolds number (Eq.~\eqref{eq:re_def}). For $\lambda \sim 10^{-1}$ as explored
+in \citet{barker_ogilvie}, our results also show no reflected wave.
 
@@
 
-This work has been supported in part by NASA FINESST grant
-19-ASTRO19-0041.%chktex 8
+This work has been supported in part by the NSF grant AST-17152. YS is supported
+by the NASA FINESST grant 19-ASTRO19-0041.%chktex 8
 

### 2020-02-19 writeups/Su_IGW_break.tex
stash WIP Daniel review comments

@@
+% TODO
+% - Figure 1: Could you include a tick mark at t=0? Does the curve start at t=0?
+% Section 6.2 needs more narrative discussion of the results.
+We should add another paragraph saying that the typical Richardson number of the critical layer is between 0.25 and 0.5, characteristic of Kelvin-Helmholtz instabilities.
+This suggestions the critical layer width is self-consistently determined by the ``turbulence’’ in our simulations.
+The layer width does not change with resolution.
+We should also say that here we calculate Ri using the background buoyancy frequency, because the fluctuations do not change N^2 significantly; thus, Ri~0.5 can be used to solve for the critical layer width in a star, as we know N^2 and U_c (corresponding to the tidal frequency Omega).
+% rewrite eq_fadef in explicit integrals
+% Investigate a smarter time averaging for the "average over everything
+% predictor"?
     \documentclass[
@@
     \newcommand*{\uv}[1]{\hat{\mathbf{#1}}}
-    \DeclareMathOperator*{\argmin}{argmin}
-    \DeclareMathOperator*{\argmax}{argmax}
     \DeclareMathOperator*{\med}{med}
@@
 \\
-$^3$ Tsung-Dao Lee Institute, Shanghai Jiao Tong University, Shanghai, 200240, China
+$^3$ Department of Astrophysical Sciences, Princeton University, Princeton NJ
+08544, USA
+\\
+$^4$ Tsung-Dao Lee Institute, Shanghai Jiao Tong University, Shanghai, 200240, China
 }
@@
     In binaries composed of early-type stars and white dwarfs, the dominant
-    tidal process involves the excitation of internal gravity waves (IGWs) that
-    propagate towards the stellar surface and their dissipation via nonlinear
-    wave breaking. We perform 2D hydrodynamical simulations of this wave
-    breaking process in a stratified isothermal atmosphere. We find that, after
-    an initial transient phase, the dissipation of the IGWs naturally generate a
-    sharp critical layer, separating the lower stationary region (with no mean
-    flow) and the upper ``synchronized'' region (with the mean flow velocity
-    equal to the horizontal wave phase speed). We find that the indefinite
-    steepening of this critical layer is suppressed by Kelvin-Helmholtz
-    instabilities, and the width of the layer is determined by the Richardson
-    criterion. We study the absorption and reflection of incident waves off the
-    critical layer and provide analytical formulae describing its evolution. The
-    result of this study is important for characterizing the evolution of
-    tidally heated white dwarfs and other binary stars.
+    tidal process involves the excitation of internal gravity waves (IGWs),
+    which propagate towards the stellar surface, and their dissipation via
+    nonlinear wave breaking. We perform 2D hydrodynamical simulations of this
+    wave breaking process in a stratified isothermal atmosphere. We find that,
+    after an initial transient phase, the dissipation of the IGWs naturally
+    generate a sharp critical layer, separating the lower stationary region
+    (with no mean flow) and the upper ``synchronized'' region (with the mean
+    flow velocity equal to the horizontal wave phase speed). While the critical
+    layer is steepened by these waves, it is simultaneously broadened by
+    Kelvin-Helmholtz instabilities such that, in steady state, the critical
+    layer width is determined by the Richardson criterion. We study the
+    absorption and reflection of incident waves off the critical layer and
+    provide analytical formulae describing its evolution. The result of this
+    study is important for characterizing the evolution of tidally heated white
+    dwarfs and other binary stars.
 \end{abstract}
@@
 stellar surface; nonlinearity develops as the wave amplitude grows, leading to
-efficient dissipation \citep{zahn75,zahn77}. As the outgoing wave deposites its
+efficient dissipation \citep{zahn75,zahn77}. As the outgoing wave deposits its
 angular momentum to the stellar surface layer, a critical layer may form and the
@@
 In this paper, we perform numerical simulations of IGW breaking in a
-plane-parallel stratified atmosphere (stellar enevelope). We use the
-pseudo-spectral code Dedalus \citep{dedalus,dedalus2} and a 2D Cartesian
+plane-parallel stratified atmosphere (a simple model for a stellar envelope). We
+use the pseudo-spectral code Dedalus \citep{dedalus,dedalus2} and a 2D Cartesian
 geometry, and consider IGWs propagating into an isothermal fluid initially at
 rest. We find that, after an initial transient phase, a \emph{critical layer}
-naturally develops, separating a lower zone that has no horizontal mean flow
-and an upper zone with mean flow at the horizontal phase velocity of the IGW\@.
-The major part of our paper is dedicated to characterizing the behavior of the
+naturally develops, separating a lower zone that has no horizontal mean flow and
+an upper zone with mean flow at the horizontal phase velocity of the IGW\@. The
+major part of our paper is dedicated to characterizing the behavior of the
 critical layer when interacting with a continuous train of IGW excited from the
@@
 steepen shear flows \citep{lindzen_qbo,lecoanet_meanflow} and act to narrow the
-critical layer. We find this steepening is suppressed by the Kelvin-Helmholtz
-instability and turbulence within the narrow critical layer. By careful
-accounting of the momentum flux budget about the critical layer, we are able to
-model the reflection and absorption of the incident IGW, and the slow downward
-propagation of the critical layer.
+critical layer. We find this steepening is counter-balanced by the
+Kelvin-Helmholtz instability and turbulence within the narrow critical layer. By
+careful accounting of the momentum flux budget about the critical layer, we are
+able to model the reflection and absorption of the incident IGW, and the slow
+downward propagation of the critical layer.
 
@@
 also quite relevant to the circulation dynamics of planetary atmospheres
-\citep[see e.g.][]{lindzen1981turbulence,holton1983influence}.
+\citep[see e.g.][]{lindzen1981turbulence,holton1983influence,baldwin2001quasi}.
 
@@
 system of equations used in our simulations. In Section~\ref{s:theory}, we
-review the existing understanding of wave breaking and present new analytical
-results on IGWs around a critical layer. In Section~\ref{s:numerics} we describe
-our numerical setup and in Section~\ref{s:weak_sim} we validate our method in
-the weak-forcing limit against linear theory. In Section~\ref{s:sim}, we present
-the results of simulations of IGW breaking and our characterization of the
-critical layer. We summarize and conclude in Section~\ref{s:discussion}.
+review the existing understanding of wave breaking and present analytical
+results characterizing IGW behavior near a critical layer. In
+Section~\ref{s:numerics} we describe our numerical setup and in
+Section~\ref{s:weak_sim} we validate our method in the weak-forcing limit
+against linear theory. In Section~\ref{s:sim}, we present the results of
+simulations of IGW breaking and our characterization of the critical layer. We
+summarize and conclude in Section~\ref{s:discussion}.
 
@@
 We consider a stratified, incompressible, isothermal fluid representing a
-stellar envelope or atmosphere. We use barotropic equation of state $P(\rho, T)
-= P(\rho)$ as a first approximation. We study dynamics in 2D, so that fluid
+stellar envelope or atmosphere. We approximate the equation of state as
+baratropic, i.e.\ $P(\rho, T) = P(\rho)$. We study dynamics in 2D, so that fluid
 variables depend only on the Cartesian coordinates $x$ and $z$. While it is
@@
 \end{equation}
-with $\overline{\rho}_0$ some reference density. We generally denote background
-quantities with overbars and perturbation quantities with primes.
+with $\overline{\rho}_0$ some reference density. We denote background quantities
+with overbars and perturbation quantities with primes.
 
@@
 Lagrangian or material derivative, and $\bm{u}, \rho, P$ denote the velocity
-field, density and pressure respectively. We denote $-g\uv{z}$ constant
-gravitational acceleration. Note that hydrostatic equilibrium ($\pdil{}{t} = 0$)
-of the background implies $\bm{\nabla}\overline{P} = -\overline{\rho} g\uv{z}$,
-so that $\overline{P} = \overline{\rho} gH$. A vertically-stratified shear flow
-$\overline{u}_x(z)\uv{x}$ is permitted in hydrostatic equilibrium, but we will
-assume no such background flow, so $\bm{u} = \bm{u}'$. Physically, this
-assumption corresponds to a non-rotating star, or going to the corotating frame
-of a rigidly rotating star.
+field, density and pressure respectively. The constant gravitational
+acceleration is $-g\uv{z}$. For this isothermal background, hydrostatic
+equilibrium implies $P = \overline{\rho} g H$. We assume there is initially no
+background flow, so $\bm{u} = \bm{u}'$. Physically, this assumption corresponds
+to a non-rotating star.
 
@@
 These variables automatically enforce $\rho > 0$ and eliminate the stiff term
-$\vec{\nabla} P / \rho$ in the Euler equation. In terms of $\Upsilon$ and
+$\bm{\nabla} P / \rho$ in the Euler equation. In terms of $\Upsilon$ and
 $\varpi$, the second two equations in~\eqref{se:nl_orig} become
@@
 \end{equation}
-is constant. Other dynamical quantities are simply related to $u'_z$.
+and is constant. Other dynamical quantities are simply related to $u'_z$.
 
@@
         \end{equation}
-        The notation $\ev{\dots}_x$ denotes averaging over the $x$ domain. For
-        the linear solution (Eq.~\ref{eq:lin_sol}), this evaluates to
+        The notation $\ev{\dots}_x$ denotes averaging over the $x$ direction.
+        For the linear solution (Eq.~\ref{eq:lin_sol}), this evaluates to
         \begin{equation}
@@
 In the linearized system, the effect of this forcing can be solved exactly
-(see Appendix~\ref{s:force_solved}). If we approximate $\abs{k_zH} \gg 1, \sigma
+(see Appendix~\ref{s:force_solved}). In the limit $\abs{k_zH} \gg 1, \sigma
 \ll H$, the solution can be approximated as two plane waves propagating away
@@
 the height of wave breaking using the condition $\abs{\bm{u}} \sim \omega /
-\abs{\bm{k}}$. This can be rewritten using the Lagrangian displacement $\bm{\xi}
-= \bm{u} / \p{-i\omega}$:
+\abs{\bm{k}}$. This can be rewritten using the Lagrangian displacement
+$\boldsymbol{\xi} = \bm{u} / \p{-i\omega}$:
 \begin{equation}
@@
 some detail. At intermediate amplitudes, wave breaking occurs via triadic
-resonances, transfering energy from the ``parent’’ IGW to ``daughter’’ waves on
+resonances, transfering energy from the ``parent'' IGW to ``daughter'' waves on
 smaller length scales that efficiently damp. The horizontal momentum flux
@@
 \begin{equation}
-    \md{}{t} = \pd{}{t} + \overline{U} \pd{}{x} + \p{\bm{u}' \cdot \vec{\nabla}},
+    \md{}{t} = \pd{}{t} + \overline{U} \pd{}{x} + \p{\bm{u}' \cdot \bm{\nabla}},
 \end{equation}
@@
 and~\eqref{eq:vol_drive} in a Cartesian box with size $L_x, L_z$. We choose
-periodic boundary conditions in both the $x$ and $z$ direction. We damp
-perturbations to zero near the top/bottom of the domain using damping zones
+periodic boundary conditions in both the $x$ and $z$ direction. To mimic the
+absence of physical boundaries at the top/bottom of the simulation domain, we damp
+perturbations to zero near the top/bottom using damping zones
 (see Section~\ref{ss:damping}). We expand all variables as Fourier series
@@
 
-    \item $C$: In our weak forcing simulations, we first choose the forcing
-        strength $C$ (see Eq.~\eqref{eq:vol_drive}) such that $\abs{\xi_z k_z}
-        \ll 1$ is satisfied everywhere in the simulation domain. This constrains
-        $C$ by Eq.~\eqref{eq:uz_lin}.
+    \item $C$: In our weak forcing simulations (Section~\ref{s:weak_sim}),
+        we first choose the forcing strength $C$ (see Eq.~\eqref{eq:vol_drive})
+        to be sufficiently weak such that $\abs{\xi_z k_z} \ll 1$ is satisfied
+        everywhere in the simulation domain. This constrains $C$ by
+        Eq.~\eqref{eq:uz_lin}. In our wave breaking simulations
+        (Section~\ref{s:sim}), we choose larger $C$.
 
@@
         viscosity and diffusivity to the system in a way that conserves
-        horizontal momentum (see Appendix~\ref{se:strat_impl} for details). To
-        quantify $\nu$, we define the dimensionless Reynolds number
+        horizontal momentum (see Appendix~\ref{se:strat_impl} for details). We
+        define the dimensionless Reynolds number
         \begin{equation}
@@
         dissipation mechanism of g-modes is radiative damping, with damping rate
-        ranging from $10^{-11}$--$10^-4$ of the mode frequency \citep{fullerI}.
-        This corresponds to a small effective viscosity or $\mathrm{Re} \gg
-        1$.}. Note that $\nu = 0$ changes the order of the PDE system, as the
-        highest-order derivatives appear only in the dissipitive terms, so only
-        nonzero $\nu$ can be used even in weakly forced simulations.
+        ranging from $10^{-11}$--$10^{-4}$ of the mode frequency
+        \citep{fullerI}. This corresponds to a small effective viscosity or
+        $\mathrm{Re} \gg 1$.}.
 \end{enumerate}
@@
 solution (Eq.~\eqref{eq:uz_lin}), this generates IGW with $\abs{\xi_z k_z}
-\approx \scinot{5}{-5}$ just above the forcing zone. This amplitude is
-sufficiently small to ensure $\abs{\xi_z k_z} \ll 1$ in the entire simulation
-domain. We include a nonzero $\nu$ corresponding to $\mathrm{Re} = 10^7$.
+\approx \scinot{5}{-5}$ just above the forcing zone. The IGW grows to $\abs{\xi_z
+k_z} \approx \scinot{7.4}{-3}$ at the upper damping zone and satisfies
+$\abs{\xi_z k_z} \ll 1$ in the entire simulation domain. We include a nonzero
+$\nu$ corresponding to $\mathrm{Re} = 10^7$.
 
@@
 \end{equation}
-Denote the momentum flux measured in the simulation by $F(z, t)$
-(use Eq.~\eqref{eq:F_def} with velocities taken from the simulation), then we
-expect $F(z, t) = F_{al}$ between $z_0$ and $z_T$. Fig.~\ref{fig:lin_fluxes}
-shows agreement with this prediction.
+Denoting the momentum flux measured in the simulation by $F(z, t)$
+(use Eq.~\eqref{eq:F_def} with velocities taken from the simulation), we expect
+$F(z, t) = F_{al}$ between $z_0$ and $z_T$. Fig.~\ref{fig:lin_fluxes} shows
+agreement with this prediction.
 \begin{figure}
@@
     \includegraphics[width=0.9\columnwidth]{plots/lin_fluxes.png}
-    \caption{$F/F_{al}$ as a function of $z$ at select times $t$ (in units of
+    \caption{$F/F_{al}$ plotted at select times $t$ (in units of
     $N^{-1}$). As the initial transient dies out, $F / F_{al} \approx 1$ to a
@@
     \end{tabular}
-    \caption{Table of simulation resolutions with wave
+    \caption{Spectral resolutions and Reynolds numbers of simulations of wave
     breaking.}\label{tab:params}
@@
     (Eq.~\eqref{eq:mean_flow}) and the dimensionless momentum flux $F(z, t) /
-    F_{al}(z)$ (Eqs.~\eqref{eq:F_def} and~\eqref{eq:F_al}) are plotted at the
-    same times as in Fig.~\ref{fig:snapshots} as a function of $z$ in our
-    fiducial simulation. The two distinct zones of mean flow are separated by a
-    critical layer. The propagation of this critical layer towards lower $z$ and
-    the sharp deposition of $F$ at the critical layer are
-    evident.}\label{fig:nl_fluxes}
+    F_{al}$ (Eqs.~\eqref{eq:F_def} and~\eqref{eq:F_al}) in our fiducial
+    simulation plotted at the same times as in Fig.~\ref{fig:snapshots}. The two
+    distinct zones of mean flow are separated by a critical layer. The critical
+    layer propagates toward lower $z$ due to momentum transport
+    ($\pdil{F}{z}$).}\label{fig:nl_fluxes}
 \end{figure}
@@
 
-To avoid noisiness, $\mathrm{Ri}$ is measured as follows: we first assign an
-$\mathrm{Ri}_x(x, t)$ for every $x$ in the critical layer, then take the median
-as $\mathrm{Ri}$ for the entire layer. $\mathrm{Ri}_x$ is computed using the
-vertical distance over which the local $u_x$ increases from $0.3\bar{U}_c$ to
-$\bar{U}_c$ (see Eq.~\eqref{eq:u_crit}). The value $0.3$ is necessary to exclude
-the small mean flow generated in the weakly nonlinear regime far below the
+It is difficult to accurately measure the Richardson number, as it depends on
+the derivative of the velocity. We measure $\mathrm{Ri}$ as follows: we first
+assign an $\mathrm{Ri}_x(x, t)$ for every $x$ in the critical layer, then take
+the median of $\mathrm{Ri}$ for the entire layer. $\mathrm{Ri}_x$ is computed
+using the vertical distance over which the local $u_x$ increases from
+$0.3\bar{U}_c$ to $\bar{U}_c$ (see Eq.~\eqref{eq:u_crit}). The value $0.3$ is
+necessary to exclude the small mean flow generated in the weakly nonlinear
+regime far below the
 critical layer. This procedure can be written:
 \begin{align}
-    z_{CL, \min}(x, t) &\equiv \argmin_z
+    z_{CL, \min}(x, t) &\equiv \min_z
         \z{z\mid u_x(x, z, t) > 0.3\overline{U}_c},\\
-    z_{CL, \max}(x, t) &\equiv \argmax_z
+    z_{CL, \max}(x, t) &\equiv \max_z
         \z{z\mid u_x(x, z, t) < \overline{U}_c},\\
@@
 \end{align}
-To understand the variation in $\mathrm{Ri}$ over $x$, we can also compute
-$\min\limits_x \mathrm{Ri}_x(x, t)$ (the maximum is very noisy). Both of these
-are shown in Fig.~\ref{fig:nl_f_ri}. The critical layer quickly narrows to its
-minimum width, driven by the anti-diffusivity of IGWs.
+To understand the variation in $\mathrm{Ri}$ over $x$, we also compute
+$\min\limits_x \mathrm{Ri}_x(x, t)$ (the maximum is very noisy). Both are shown
+in Fig.~\ref{fig:nl_f_ri}. The critical layer quickly narrows to its minimum
+width, driven by the anti-diffusivity of IGWs.
 \begin{figure}
@@
 
-Knowing now the width of the critical layer, we next seek its location. The
-downward propagation of the critical layer location $z_c(t)$ is driven by the
-absorption of horizontal momentum flux at $z_c$, following
-Eq.~\eqref{eq:zc_anal}. When analyzing the simulation, one must first compute
-$z_c$ before examining the flux budget in its vicinity, but conceptually $z_c$
-is a consequence of flux redistribution at the critical layer. Here, we proceed
-in conceptual order and first describe how we analyze the flux transfer at
-$z_c$.
-
-In general, the flux budget at the critical layer can be decomposed as
+The downward propagation of the critical layer location $z_c(t)$ is driven by
+the absorption of horizontal momentum flux at $z_c$, following
+Eq.~\eqref{eq:zc_anal}. The flux budget at the critical layer can be decomposed
+as
 \begin{equation}
@@
 where $F_i$ is the incident flux, $F_a$ is the absorbed flux, $F_r$ is the
-reflected flux, and $F_s$ is the ``residual'' flux above the critical layer,
-likely responsible for momentum redistribution within the synchronized upper
+reflected flux, and $F_s$ is some ``redistribution'' flux above the critical
+layer, responsible for momentum redistribution within the synchronized upper
 layer. Careful accounting of $F_s$ turns out to be important to obtain the
@@
 interpretation of $F_s$ is unclear; it is somewhat tempting but unfounded to
-identify $F_s$ with the transmitted flux. In the simulation, we find $F_s < 0$.
+identify $F_s$ with the transmitted flux. In these simulations, we find $F_s <
+0$, corresponding to net momentum transport \emph{into} the critical layer from
+the synchronized layer above it.
 
 After measuring $z_c$ (see Section~\ref{ss:cl_prop}) and $F(z)$
-(Eq.~\eqref{eq:F_def}) at each time step, we can then determine each of $F_i$,
-$F_a$, $F_r$, $F_s$ as follows:
+(Eq.~\eqref{eq:F_def}) at each time step, we determine each of $F_i$, $F_a$,
+$F_r$, $F_s$ as follows:
 \begin{align}
     F_i(t) &= F_{al}A_i^2(t),\\
-    F_i(t) - F_r(t) &= \ev{F(z, t)}_{z \in [z_c - \Delta z - H, z_c - \Delta z]}
+    F_r(t) &= F_i(t) - \ev{F(z, t)}_{z \in [z_c - \Delta z - H, z_c - \Delta z]}
         ,\label{eq:fr_def}\\
@@
 \end{align}
-Here, $\ev{\dots}_{z \in [z_a, z_b]}$ denotes a vertical average over the
-interval $[z_a, z_b]$. Below the critical layer, we average over an interval of
-length $H$, also the vertical wavelength. The offset $\Delta z$ is necessary to
-make the measurement of the incident flux unaffected by the turbulence within
-the critical layer itself. The width of the critical layer is limited by
-$\mathrm{Ri} \lesssim 1$ (see Section~\ref{ss:khi}), which bounds its vertical
-extent $\sim \frac{1}{\abs{k_{z}}}$. We empirically found an offset of $\Delta z
-= \frac{3}{\abs{k_z}}$ was necessary to be sufficiently far from strong
+Fig.~\ref{fig:nl_f_amps2} depicts the four components of this flux
+decomposition.
+% Here, $\ev{\dots}_{z \in [z_a, z_b]}$ denotes a vertical average over the
+% interval $[z_a, z_b]$.
+Below the critical layer, we average over an interval of length $H$, also the
+vertical wavelength. The offset $\Delta z$ is necessary to make the measurement
+of the incident flux unaffected by the turbulence within the critical layer
+itself. The width of the critical layer is limited by $\mathrm{Ri} \lesssim 1$
+(see Section~\ref{ss:khi}), which bounds its vertical extent $\sim
+\frac{1}{\abs{k_{z}}}$. We empirically found an offset of $\Delta z =
+\frac{3}{\abs{k_z}}$ was necessary to be sufficiently far from strong
 fluctuations near the critical layer.
@@
 flux budget. We average only where $F < 0$ so that $F_s$ is robust to such width
-variations. We find that this is a sufficiently accurate way of measuring $F_s$
-and determining $F_a$.
-
-Fig.~\ref{fig:nl_f_amps2} depicts the four components of this flux decomposition.
+variations. We find that this is an accurate way of measuring $F_s$ and
+determining $F_a$.
+
 \begin{figure}
@@
 
-With a careful determination of $F_a$, we can next make predictions for the
+With a careful determination of $F_a$, we can make predictions for the
 propagation of $z_c(t)$ and compare to the measured propagation in the
-simulation.
-In principle, $z_c$ is the location where the incident flux significantly
-attenuates. In the simulation, shear turbulence causes $F$ to have significant
-spatial and temporal fluctuations that translate to large temporal fluctuations
-in $z_c(t)$. To minimize these spurious fluctuations, we measure the location of
-the critical layer using a spatial average of where flux deposition occurs:
-\begin{align}
-    z_{c, \min}(t) &\equiv \argmin_z \z{z: F(z, t) > 0.3F_{al}},\\
-    z_{c, \max}(t) &\equiv \argmax_z \z{z: F(z, t) < 0.3F_{al}},\\
+simulation. In principle, $z_c$ is the location where the incident flux
+significantly attenuates. In the simulation, shear turbulence causes $F$ to have
+significant spatial and temporal fluctuations that translate to large temporal
+fluctuations in $z_c(t)$. To minimize these spurious fluctuations, we measure
+the location of the critical layer using a spatial average of where flux
+deposition occurs:
+\begin{align}
+    z_{c, \min}(t) &\equiv \min_z \z{z: F(z, t) > 0.3F_{al}},\\
+    z_{c, \max}(t) &\equiv \max_z \z{z: F(z, t) < 0.3F_{al}},\\
     z_c(t) &\equiv \frac{z_{c, \min}(t) + z_{c, \max}(t)}{2}.\label{eq:zc_def}
@@
 
-Finally, in Fig.~\ref{fig:nl_front} we plot the numerically measured $z_c$
-against two semi-analytic predictors: (i) integration of Eq.~\eqref{eq:zc_anal}
-using the measured $F_a(t)$, and (ii) substituting the time-averaged
-$\ev{F_a}_t$ (over the entire length of the simulation) into
-Eq.~\eqref{eq:zc_sol}. Since $z_c(t)$ is less well-defined at early times (when
-the critical layer is thick and transient behavior is strong), we solve
-Eq.~\eqref{eq:zc_anal} by integrating backwards from the end of the simulation
-($t = t_f$), using $z_c(t_f)$ as the initial condition. From
-Fig.~\ref{fig:nl_front}, we see that the agreement between the measured $z_c(t)$
-and its estimate via $F_a(t)$ and Eq.~\eqref{eq:zc_anal} is excellent.
+In Fig.~\ref{fig:nl_front} we plot the numerically measured $z_c$ against two
+semi-analytic models: (i) integration of Eq.~\eqref{eq:zc_anal} using the
+measured $F_a(t)$, and (ii) substituting the time-averaged $\ev{F_a}_t$ (over
+the entire length of the simulation) into Eq.~\eqref{eq:zc_sol}. Since $z_c(t)$
+is less well-defined at early times (when the critical layer is thick and
+transient behavior is strong), we solve Eq.~\eqref{eq:zc_anal} by integrating
+backwards from the end of the simulation ($t = t_f$), using $z_c(t_f)$ as the
+initial condition. From Fig.~\ref{fig:nl_front}, we see that the agreement
+between the measured $z_c(t)$ and its estimate via $F_a(t)$ and
+Eq.~\eqref{eq:zc_anal} is excellent.
 
 By time-averaging the numerically measured $F_a$, we find $\ev{F_a}_t \approx
-0.71F_{al}$, which is used for the second predictor in Fig.~\ref{fig:nl_front}.
+0.71F_{al}$, which is used for the second model in Fig.~\ref{fig:nl_front}.
 Note that $F_a < F_{al}$, so momentum flux absorption at the critical layer is
-incomplete.
-% Moreover, careful comparison of the two predicted curves in
-% Fig.~\ref{fig:nl_front} shows that there is . This suggests that
-% momentum flux absorption changes significantly over time, in agreement with
-% detailed analysis in Section~\ref{ss:reflectivity}.
+incomplete. This is due to reflection of waves off the critical layer, which
+carry momentum downward.
 \begin{figure}
@@
     \caption{Propagation of the critical layer over time. Shown are: (black)
-    $z_c(t)$ from simulation data, (green) predictor of $z_c(t)$ using direct
-    integration of Eq.~\eqref{eq:zc_anal} for $F_a(t)$ measured from
-    simulation data (described in Eq.~\eqref{eq:fa_def}), and (blue) direct
-    substitution of time-averaged $\ev{F_a(t)}_t$ into
-    Eq.~\eqref{eq:zc_sol}. Predictors use the end of the simulation as initial
-    conditions and integrate backwards, as $z_c$ is less well-defined at early
-    times. The agreement of the directly-integrated predictor with the data
-    shows Eq.~\eqref{eq:zc_anal} is a good description of the evolution of $z_c$.
-    The the poorer but qualitatively correct agreement of the time-averaged
-    predictor with the data shows both that $F_a(t) <
-    F_{al}$.}\label{fig:nl_front}
+    $z_c(t)$ from simulation data, (green) model for $z_c(t)$ using direct
+    integration of Eq.~\eqref{eq:zc_anal} for $F_a(t)$ measured from simulation
+    data (described in Eq.~\eqref{eq:fa_def}), and (blue) direct substitution of
+    time-averaged $\ev{F_a(t)}_t$ into Eq.~\eqref{eq:zc_sol}. The models use the
+    end of the simulation as initial conditions and integrate backwards, as
+    $z_c$ is less well-defined at early times. The agreement of the
+    directly-integrated model with the data shows Eq.~\eqref{eq:zc_anal} is a
+    good description of the evolution of $z_c$. The the poorer but qualitatively
+    correct agreement of the time-averaged model with the data shows both that
+    $F_a(t) < F_{al}$.}\label{fig:nl_front}
 \end{figure}
@@
 reflective behaviors observed in the simulation: (i) the presence of a reflected
-wave with wave vector $\bm{k}_r = k_{x}\uv{x} - k_{z}\uv{z}$, and (ii) the
-reflected flux $F_r$. The reflected wave amplitude and flux need not agree
-exactly if some reflected flux is in higher-order modes, which is indeed the
-case in our simulations. Both are of physical interest, however: the reflected
-wave amplitude is essential for setting up standing modes in a realistic star,
-while the flux is important for accurately tracking angular momentum transfer
-during synchronization.
+wave with the same frequency as the incident wave (i.e.\ with wave vector
+$\bm{k}_r = k_{x}\uv{x} - k_{z}\uv{z}$), and (ii) the reflected flux $F_r$. The
+reflected wave amplitude and flux need not agree exactly if some reflected flux
+is in higher-order modes, which is indeed the case in our simulations. Both are
+of physical interest, however: the reflected wave amplitude is essential for
+setting up standing modes in a realistic star, while the flux is important for
+accurately tracking angular momentum transfer during synchronization.
 
@@
 $A_r(t)$. In our simulation, the phase offset $\phi_r(t) \equiv k_x \delta x(t)
-$ is consistent with reflection off a moving boundary at $z_c$, well
-approximated by $\abs{\pdil{\phi_r}{t}} \simeq 2\abs{\pdil{(k_{z}z_c)}{t}}$.
-Finally, note that $\bm{u}_{al}$ and $\bm{u}_{al, \bm{k}_r}$ are orthogonal
-under the integral used for Eqs.~\eqref{eq:ahat_def} and~\eqref{eq:ar_def}, so
-$A_i$ and $A_r$ are independent.
-
-Fig.~\ref{fig:nl_f_amps} illustrates the behaviors of these two quantities. Both
-of them vary significantly in time but their mean values appear to converge
-towards the end of the simulation. A modest ($\sim20\%$), long-term fluctuation
-in $A_i(t)$ is observed, likely of numerical origin.
+$ is consistent with reflection off a moving boundary at $z_c$,
+i.e.\ $\abs{\pdil{\phi_r}{t}} \simeq 2\abs{\pdil{(k_{z}z_c)}{t}}$.
+
+Fig.~\ref{fig:nl_f_amps} illustrates the behaviors of $A_i$ and $A_r$. Both vary
+significantly in time but their mean values appear to converge towards the end
+of the simulation.
 \begin{figure}
@@
     \includegraphics[width=0.9\columnwidth]{plots/nl_f_amps.png}
-    \caption{Wave amplitudes observed in the simulation. The
-    top panel shows the incident wave amplitude $A_i(t)$ (green) and the
-    reflected wave amplitude $A_r(t)$ (red) just above the forcing
-    zone. The modest fluctuation in $A_i(t)$ is likely of numerical origin but
-    has no bearing on our results.}\label{fig:nl_f_amps}
+    \caption{The incident wave amplitude $A_i(t)$ (green) and the reflected wave
+    amplitude $A_r(t)$ (red) just above the forcing zone.}\label{fig:nl_f_amps}
 \end{figure}
@@
 Since $A_i(t), A_r(t)$ vary somewhat over time, we perform time averaging over
-interval of approximately $8\pi/\omega$, denoted by angle brackets. We can then
-define the amplitude reflectivity
+interval of four wave periods, denoted by angle brackets. We can then define the
+amplitude reflectivity
 \begin{equation}
@@
 
-This reflectivity can be compared to the flux redistribution at the critical
-layer by calculating $\mathcal{R}_A^2(t)$ and the ratios of $F_r$ and $-F_s$ (as
-$F_s < 0$) to $F_i$. A comparison between $\mathcal{R}_A^2$ and $\hat{F}_r$ is
-appropriate as $F \propto A^2$ (Eq.~\eqref{eq:S_lin}). We define \begin{align}
+We compare the square of the reflectivity to the ratios of $F_r$ and $-F_s$ to
+$F_i$, as $F \propto A^2$ (Eq.~\eqref{eq:S_lin}). We define
+\begin{align}
     \hat{F}_r &\equiv \frac{\ev{F_r}(t)}{\ev{F_i}(t)}, \label{eq:srefl_def1}\\
     \hat{F}_s &\equiv -\frac{\ev{F_s}(t)}{\ev{F_i}(t)}. \label{eq:srefl_def}
-\end{align} Fig.~\ref{fig:nl_f_refl} shows $\mathcal{R}_A^2$, $\hat{F}_r$,
-$\hat{F}_s$ using the results of Fig.~\ref{fig:nl_f_amps2} and
-Fig.~\ref{fig:nl_f_amps} as inputs. Visual inspection suggests the three
-quantities have reached their asymptotic values (in the time-averaged sense) for
-$t \gtrsim 2500/N$. The aforementioned fluctuations in $A_i$ do not affect our
+\end{align}
+Fig.~\ref{fig:nl_f_refl} shows $\mathcal{R}_A^2$, $\hat{F}_r$, and $\hat{F}_s$
+as functions of time. The three quantities appear to be roughly stationary for
+$t \gtrsim 2500/N$. Modest fluctuations ($\sim 20\%$) in $A_i$ do not affect our
 reflectivity results thanks to the time averaging used in
@@
     \includegraphics[width=0.9\columnwidth]{plots/nl_f_refl.png}
-    \caption{Behavior of dimensionless reflectivities $\mathcal{R}_A^2$,
-    $\hat{F}_r$, and residual flux $\hat{F}_s$
+    \caption{$\mathcal{R}_A^2$, $\hat{F}_r$, and redistribution flux $\hat{F}_s$
     (Eqs.~\eqref{eq:Ra_def}--\eqref{eq:srefl_def}) as a function of time (in
-    units of $N^{-1}$) in the simulation. These quantities seem to become
-    comparatively stable past about $t = 2500/N$, indicating that an asymptotic
-    value may have been reached. That $\hat{F}_r \gtrsim \mathcal{R}_A^2$
-    implies a substantial fraction of reflected flux is in higher-order modes
-    than the reflected IGW.}\label{fig:nl_f_refl}
+    units of $N^{-1}$). These quantities seem to become comparatively stable
+    past about $t = 2500/N$, indicating that an asymptotic value may have been
+    reached. That $\hat{F}_r \gtrsim \mathcal{R}_A^2$ implies a substantial
+    fraction of reflected flux is in higher-order modes than the reflected
+    IGW.}\label{fig:nl_f_refl}
 \end{figure}
@@
 
-As the primary test of whether our simulations have reached sufficient
-resolution, we consider the convergence of $\mathcal{R}_A^2$, $\hat{F}_r$,
-$\hat{F}_s$. Furthermore, we compare our results against
-Eqs.~\eqref{eq:crit_coeffs}, which suggest that wave reflection at the critical
-layer depends only on the local Richardson number (see Section~\ref{ss:khi} for
-details on our measurement of $\mathrm{Ri}$).
+Although throughout this paper we focused on our fiducial simulation with
+$\mathrm{Re} = 1024$ and resolution $N_x=768$, $N_z=3072$, we also ran a suite
+of simulations varying the resolution and corresponding Reynolds number
+(Tab.~\ref{tab:params}). We find that our global, quantitative measurements in
+the simulations ($\mathcal{R}_A^2$, $F_r$, $F_s$, and $\mathrm{Ri}$) are very
+similar for our highest Reynolds numbers (1024 and 2048).
 
@@
 (Eqs.~\eqref{eq:Ra_def}--\eqref{eq:srefl_def} and~\eqref{eq:ri_def}
-respectively) over the last $1/4$ of the simulation time from each simulation,
-when these quantities have converged to their asymptotic values. These results
-are shown in Fig.~\ref{fig:agg}.
+respectively) over the last $1/4$ of the simulation time, when these quantities
+have reached their asymptotic values. These results are shown in
+Fig.~\ref{fig:agg}.
 
@@
     $\hat{F}_s$, and $\mathrm{Ri}$ (Eqs.~\eqref{eq:Ra_def}--\eqref{eq:srefl_def}
-    and~\eqref{eq:ri_def} respectively) across runs with varying resolution and
-    viscosity as given in Tab.~\ref{tab:params}. Vertical bars show the temporal
-    variation of each measurement between the $16\%$ and $84\%$ range. Small
-    horizontal displacements are made for data points at identical $\mathrm{Re}$
-    for readability. Note that simulations with larger $\mathrm{Re}$ correspond
-    to smaller viscosity and are more physically realistic. At the smallest
-    $\mathrm{Re}$ value, $\mathrm{Ri} \approx 50$ is too large to fit on the
-    plot.}\label{fig:agg}
+    and~\eqref{eq:ri_def} respectively) in simulations with varying resolution
+    and viscosity as given in Tab.~\ref{tab:params}. Vertical bars show the
+    temporal variation of each measurement between the $16\%$ and $84\%$ range.
+    Small horizontal displacements are made for data points at identical
+    $\mathrm{Re}$ for readability. Note that simulations with larger
+    $\mathrm{Re}$ correspond to smaller viscosity and are more physically
+    realistic. At the smallest $\mathrm{Re}$ value, $\mathrm{Ri} \approx 50$ is
+    too large to fit on the plot.}\label{fig:agg}
 \end{figure}
 
-It is apparent that the Richardson number rapidly converges to $\mathrm{Ri}
-\approx 0.4$ but the reflectivity measure converges more slowly. This disagrees
-with Eq.~\eqref{eq:crit_coeffs}, which is based on the linear theory. This
-tension is natural: fluid motion within the critical layer is turbulent, so
-reflecton at the critical layer cannot be captured by the linear theory.
+As the simulation resolution increases and the viscosity decreases, we find that
+the Richardson number decreases, while the reflection and redistribution fluxes
+increase. The Richardson number is roughly constant for $\mathrm{Re} > 200$ with
+a value of $\mathrm{Ri} \sim 0.4$. The behavior of the fluxes is more
+complicated. While the fraction of reflected and redistributed flux is similar
+for our simulations with $\mathrm{Re} = 1024$ and $2048$, higher resolution
+simulations would be required to determine these flux fractions in the limit
+$\mathrm{Re}\rightarrow\infty$.
+
+Nevertheless, the difference in behavior of $\mathrm{Ri}$ and the flux
+reflectivity as $\mathrm{Re}$ is varied is in tension with
+Eq.~\eqref{eq:crit_coeffs}. This tension is natural: Eq.~\eqref{eq:crit_coeffs}
+is derived from a linear theory, while fluid motion within the critical layer is
+turbulent, so reflecton at the critical layer cannot be captured by the linear
+theory.
 
@@
     \item The width of the turbulent critical layer is determined by requiring
-        the local Richardson number (Eq.~\eqref{eq:ri_def}) $\mathrm{Ri} \gtrsim
+        the local Richardson number (Eq.~\eqref{eq:ri_def}) $\mathrm{Ri} \sim
         0.5$ (see Fig.~\ref{fig:nl_f_ri}).
@@
         incomplete. The critical layer only absorbs $\sim 70\%$ of the incident
-        flux in most simulations (see Fig.~\ref{fig:agg}). The reflected flux is
-        carried away from the critical layer as both lowest-order reflected
-        waves and waves with larger $z$ wavenumbers.
+        flux in our highest resolution simulations (see Fig.~\ref{fig:agg}). The
+        reflected flux is carried away from the critical layer as both
+        lowest-order reflected waves and waves with larger $z$ wavenumbers.
 \end{enumerate}
@@
 In this paper, we have studied the nonlinear behavior of IGWs with $\abs{k_x /
-k_z} \sim 1$ in a plane-parallel geometry. Tidally excited IGWs in binary stars
-have horizontal wave number $k_\perp \sim 1/R$ (where $R$ is the stellar radius)
-much smaller than the radial wave number $k_r$. Strictly speaking, our results
-may not directly apply to IGWs with $\abs{k_x/k_z}\ll 1$. However, the
-qualitative behavior is likely to be similar, as the turbulence driving the
-critical layer dynamics occurs at scales significantly smaller than either
-$1/k_x$ or $1/\abs{k_z}$. Simulating IGWs with $k_x \ll \abs{k_z}$ is more
-challenging numerically and we defer its exploration to future work.
+k_z} \sim 1 / (2\pi)$ in a plane-parallel geometry. Tidally excited IGWs in
+binary stars have horizontal wave number $k_\perp \sim 1/R$ (where $R$ is the
+stellar radius) much smaller than the radial wave number $k_r$. While our
+simulations do not have $\abs{k_x/k_z}\ll 1$, the qualitative behavior is likely
+to be similar, as the turbulence driving the critical layer dynamics occurs at
+scales significantly smaller than either $1/k_x$ or $1/\abs{k_z}$. Simulating
+IGWs with $k_x \ll \abs{k_z}$ is more challenging numerically and we defer its
+exploration to future work.
 
@@
 due to geometric focusing. In their numerical simulations in a 2D polar
-geometry, they found no evidence for reflected waves, contrary to our result. It
-is plausible that the discrepancy arises due to the differing causes of wave
-breaking (geometric focusing vs.\ density rarefaction). Note that their
-nonlinearity parameter $\lambda \sim 1/\mathrm{Re}$ where $\mathrm{Re}$ is the
-Reynolds number (Eq.~\eqref{eq:re_def}). For $\lambda \sim 10^{-1}$ as explored
-in \citet{barker_ogilvie}, our results also show no reflected wave.
+geometry, they found no evidence for reflected waves, contrary to our result.
+% It is plausible that the discrepancy arises due to the differing causes of wave
+% breaking (geometric focusing vs.\ density rarefaction). Note that their
+% nonlinearity parameter $\lambda \sim 1/\mathrm{Re}$ where $\mathrm{Re}$ is the
+% Reynolds number (Eq.~\eqref{eq:re_def}). For $\lambda \sim 10^{-1}$ as explored
+% in \citet{barker_ogilvie}, our results also show no reflected wave.
+Note, however, that their simulations were run with substantially higher
+viscosity, i.e.\ lower resolution (their $\lambda \sim 1/\mathrm{Re}$, so they
+studied $\mathrm{Re} \sim 10$), than the simulations presented here. We also
+find at low Reynolds number that there is negligible wave reflection. It is
+possible that reflection and flux redistribution are also important processes
+for geometrically breaking waves, but was absent from the low-Reynolds number
+simulations of \citet{barker_ogilvie}.
 
@@
 by the NASA FINESST grant 19-ASTRO19-0041.%chktex 8
+DL is supported by the Princeton Center for Theoretical Sciences and Lyman
+Spitzer Jr fellowships. Computations were conducted with support by the NASA
+High End Computing (HEC) Program through the NASA Advanced Supercomputing (NAS)
+Division at Ames Research Center on Pleiades with allocation GID s1647.
 
@@
 
-To solve for the linear excited IGW amplitude due to bulk forcing
-(see Eq.~\eqref{eq:vol_drive}), we consider the
-complexified linearized system of equations, with all dynamical variables having
-dependence $u_z(x, z, t) = \tilde{u_z}(z) e^{ik_xx - i\omega t}$. Thus,
-$\pdil{}{t} \to -i\omega t$, $\pd{}{x} \to ik_x$, and the dynamical fluid
+To solve for the linear excited IGW amplitude due to bulk forcing (see
+Eq.~\eqref{eq:vol_drive}), we consider the linearized system of equations, with
+all dynamical variables having dependence $u_z(x, z, t) = \tilde{u_z}(z)
+e^{ik_xx - i\omega t}$. Thus, $\pdil{}{t} \to -i\omega t$, $\pd{}{x} \to ik_x$,
+and the dynamical fluid
 equations become (see Eqs.~\eqref{se:nl_orig} and~\eqref{se:nl_upsilon}):
@@
 
-\section{Equation Implementations}\label{se:strat_impl}
+\section{Equation Implementation}\label{se:strat_impl}
 
@@
 wavenumbers to higher wavenumbers. Since spectral codes have no numerical
-dissipation, artificial dissipation must be added. To ensure the dissipitive
-system conserves horizontal momentum exactly, we begin by adding dissipitive
+diffusion, explicit diffusion must be added. To ensure the non-ideal
+system conserves horizontal momentum exactly, we begin by adding diffusion
 terms to the flux-conservative form of the Euler fluid equations
@@
 \end{subequations}
-The same viscosity $\nu$ is used for both the diffusive and viscous terms,
-although this is not required. Since the dissipation is not physical and is
-purely used for numerical stability, we choose it such that hydrostatic
-equilibrium is not modified (hence $\nu$ acts only on $\rho - \overline{\rho}$).
+For simplicity, we use the same diffusivity $\nu$ for both the momentum and mass
+diffusivities. Although mass diffusivity is not physical, we incclude it for
+numerical stability. We choose the mass diffusion term to conserve mass, and not
+to affect the background density profile.
 
@@
 similar to Eq.~\eqref{eq:Gamma}. In the absence of this mask, a nonphysical mean
-flow localized to the forcing zone develops. We use mask
+flow localized to the forcing zone develops. We use the mask
 \begin{equation}

### 2020-02-19 sims/2d_4_fourier/strat.py

@@ -5,9 +5,12 @@ import numpy as np
 import sys
 from strat_helper import *
 
-# comment out when venv not working
-from mpi4py import MPI
-CW = MPI.COMM_WORLD
+# Not needed for plotting after pkl is generated
+try:
+    from mpi4py import MPI
+    CW = MPI.COMM_WORLD
+except ModuleNotFoundError:
+    print('Not loading mpi packages')
 
 H = 1
 XMAX = 4 * H

### 2020-02-19 sims/2d_4_fourier/strat_helper.py

@@ -17,14 +17,17 @@ from scipy.interpolate import interp1d
 from scipy.optimize import minimize
 PLT_COLORS = ['b', 'g', 'r', 'c', 'm', 'y', 'k', 'w'];
 
-# comment out when venv not working
-import h5py
-from dedalus import public as de
-from dedalus.tools import post
-from dedalus.extras.flow_tools import CFL, GlobalFlowProperty
-from dedalus.extras.plot_tools import quad_mesh, pad_limits
-from mpi4py import MPI
-CW = MPI.COMM_WORLD
+# Not needed for plotting after pkl is generated
+try:
+    import h5py
+    from dedalus import public as de
+    from dedalus.tools import post
+    from dedalus.extras.flow_tools import CFL, GlobalFlowProperty
+    from dedalus.extras.plot_tools import quad_mesh, pad_limits
+    from mpi4py import MPI
+    CW = MPI.COMM_WORLD
+except ModuleNotFoundError:
+    print('Not loading dedalus and h5 packages')
 
 SNAPSHOTS_DIR = 'snapshots_%s'
 FILENAME_EXPR = '{s}/{s}_s{idx}.h5'
@@ -516,7 +519,9 @@ def load(name, params, dyn_vars, stride, start=0):
         state_vars[key] = np.array(state_vars[key])
 
     state_vars['ux_z'] = np.gradient(state_vars['ux'], axis=2)
-
+    state_vars['U_z'] = np.gradient(state_vars['U'], axis=2)
+    state_vars['W_z'] = np.gradient(state_vars['W'], axis=2)
+    state_vars['W'] += params['g'] * params['H'] # oops wrong gauge choice
     state_vars['S'] = RHO0 * np.exp(-z/ H) * (
         (state_vars['ux'] * state_vars['uz']) +
         (NU * np.exp(state_vars['U']) * state_vars['ux_z']))
@@ -885,6 +890,8 @@ def write_front(name, params, stride=1):
         for x_idx in range(N_X):
             ux_slice = state_vars['ux'][t_idx, x_idx]
             w_slice = state_vars['W'][t_idx, x_idx]
+            Uz_slice = state_vars['U_z'][t_idx, x_idx]
+            Wz_slice = state_vars['W_z'][t_idx, x_idx]
             def interp_idx(f, idx, val):
                 df_grid = f[idx] - f[idx - 1]
[…]

### 2020-02-19 writeups/Su_IGW_break.tex
small

@@
 Yubo Su,$^1$,
-Daniel Lecoanet,$^2$
-Dong Lai$^{1,3}$
+Daniel Lecoanet,$^{2,3}$
+Dong Lai$^{1,4}$
 \\
@@
     nonlinear wave breaking. We perform 2D hydrodynamical simulations of this
-    wave breaking process in a stratified isothermal atmosphere. We find that,
+    wave breaking process in a stratified, isothermal atmosphere. We find that,
     after an initial transient phase, the dissipation of the IGWs naturally

### 2020-02-20 writeups/Su_IGW_break.tex
better movie plotting

@@
 We should also say that here we calculate Ri using the background buoyancy frequency, because the fluctuations do not change N^2 significantly; thus, Ri~0.5 can be used to solve for the critical layer width in a star, as we know N^2 and U_c (corresponding to the tidal frequency Omega).
-% rewrite eq_fadef in explicit integrals
 % Investigate a smarter time averaging for the "average over everything
@@
 the forcing location. We choose $L_z = 12.5H$, and the lower and upper damping
-zones are located at $z < 1.5H$ and $z > 10.5H$ respectively. The forcing (see
-Eq.~\eqref{eq:vol_drive}) is at $z_0 = 3H$ with width $\sigma = 0.078H$,
-sufficiently far from the lower damping zone and permitting sufficient room for
-the upward propagating wave to grow as $\propto e^{z/2H}$. Finally, we want
-similar grid spacing in the $x$ and $z$ directions (i.e.\ $L_x / N_x \sim L_z
-/N_z$), guided by the intuition that turbulence generated by wave breaking is
-approximately isotropic, so we use $L_x = 4H$ and $N_z / N_x = 4$.
+zones are located at $z < 0.3H$ and $z > 9.5H$ respectively ($z \in [-1,
+11.5]$). The forcing (see Eq.~\eqref{eq:vol_drive}) is at $z_0 = 2H$ with width
+$\sigma = 0.078H$, sufficiently far from the lower damping zone and permitting
+sufficient room for the upward propagating wave to grow as $\propto e^{z/2H}$.
+Finally, we want similar grid spacing in the $x$ and $z$ directions (i.e.\ $L_x
+/ N_x \sim L_z /N_z$), guided by the intuition that turbulence generated by wave
+breaking is approximately isotropic, so we use $L_x = 4H$ and $N_z / N_x = 4$.
 
@@
     F_i(t) &= F_{al}A_i^2(t),\\
-    F_r(t) &= F_i(t) - \ev{F(z, t)}_{z \in [z_c - \Delta z - H, z_c - \Delta z]}
+    F_r(t) &= F_i(t) - \frac{1}{H}
+        \int\limits_{z_c - \Delta z - H}^{z_c - \Delta z}F(z, t)\;\mathrm{d}z
+        % \ev{F(z, t)}_{z \in [z_c - \Delta z - H, z_c - \Delta z]}
         ,\label{eq:fr_def}\\
-    F_s(t) &= \ev{\z{F(z, t): F(z, t) < 0}}_{z \in [z_c, z_c + \Delta z]},
+    F_s(t) &=
+        \frac{1}{\Delta z}\int\limits_{z_c}^{z_c + \Delta z}
+            \min(F(z, t), 0)\;\mathrm{d}z,
+        % \ev{\z{F(z, t): F(z, t) < 0}}_{z \in [z_c, z_c + \Delta z]},
         \label{eq:fs_def}\\

### 2020-02-20 dedalus/dedalus/extras/plot_tools.py

@@ -132,6 +132,7 @@ def plot_bot(dset, image_axes, data_slices, image_scales=(0,0), clim=None, even_
         ticks=ticker.MaxNLocator(nbins=5))
     cbar.outline.set_visible(False)
     caxes.xaxis.set_ticks_position('top')
+    caxes.set_xticklabels(caxes.get_xticklabels(), fontsize=12)
 
     # Labels
     if title is None:
@@ -139,7 +140,7 @@ def plot_bot(dset, image_axes, data_slices, image_scales=(0,0), clim=None, even_
             title = dset.attrs['name']
         except KeyError:
             title = dset.name
-    caxes.set_xlabel(title, fontsize=14)
+    caxes.set_xlabel(title, fontsize=14, labelpad=10)
     caxes.xaxis.set_label_position('top')
     if isinstance(xscale, str):
         paxes.set_xlabel(xscale)

### 2020-02-20 sims/2d_4_fourier/plot_2d_series_movie.py

@@ -1,7 +1,16 @@
 """
-rm -f frames/*.png && python plot_2d_series_movie.py testo/*.h5
-ffmpeg -framerate 12 -pattern_type glob -i 'yubo_*.png' test.mp4
+Plot planes from joint analysis files.
+
+Usage:
+    plot_2d_series_movie.py <files>... [--output=<dir>] [--write=<num>]
+
+Options:
+    --output=<dir>  Output directory [default: ./frames]
+    --write=<num>   Write number offset [default: 0]
 """
+
+# rm -f frames/*.png && python plot_2d_series_movie.py --output frames snapshots_yubo_nu1_vhres/*.h5
+# ffmpeg -framerate 12 -pattern_type glob -i 'yubo_*.png' test.mp4
 
 import h5py
 import numpy as np
@@ -13,7 +22,19 @@ from dedalus.extras import plot_tools
 import re
 
 plt.rc('text', usetex=True)
-plt.rc('font', family='serif', size=12)
+plt.rc('font', family='serif', size=16)
+
+# critical horizontal velocity
+UX_C = 0.15394489478851475
+SIGMA = 0.078
+Z0 = 2
+Z_BOT = 0.3
+Z_TOP = 9.5
+
+# save two versions of these indicies, one version to put in paper (times are
+# labeled)
+PLOT_IDXS = [231, 50, 86, 451]
+PLOT_PREFIXES = ['(i)', '(ii)', '(iii)', '(iv)']
 
 def main(filename, start, count, output, write_start):
     """Save plot of specified tasks for given range of analysis writes."""
@@ -22,51 +43,83 @@ def main(filename, start, count, output, write_start):
     tasks = ['ux','uz']
     scale = 2.5
     dpi = 300
-    title_func = lambda sim_time: r'$t = {:.3f}$'.format(sim_time)
     # savename_func = lambda write: 'yubo_{:06}.png'.format(write+write_start)
-    def savename_func(write):
-        # regex parse out filename number, 10 writes per file
-        num_per_file = 10
-        write_num = (write - 1) % num_per_file
-        filenum = int(re.match('.*s([\d]+)\.h5', filename)[1])
-        suffix_num = (filenum - 1) * num_per_file + write_num
-        if suffix_num > 201:
-            suffix_num -= 9 # hard coded, one of the restores is misnumbered
-        print(suffix_num)
-        return 'yubo_{:06}.png'.format(suffix_num)
[…]

### 2020-02-20 sims/2d_4_fourier/strat.py

@@ -120,7 +120,7 @@ if __name__ == '__main__':
 
     elif '-write' in sys.argv:
         for name, params_dict in tasks:
-            write_front(name, params_dict)
+            write_front(name, params_dict, start=0)
 
     elif '-front' in sys.argv:
         out = []

### 2020-02-20 sims/2d_4_fourier/strat_helper.py

@@ -512,7 +512,8 @@ def load(name, params, dyn_vars, stride, start=0):
             start = (((simlen - start - 1) // stride) + 1) * stride\
                 + start - simlen
         filename = FILENAME_EXPR.format(s=snapshots_dir, idx=i)
-    print('Loaded total %d times' % len(total_sim_times))
+    print('Loaded total %d times' % len(total_sim_times),
+          total_sim_times[0], total_sim_times[-1])
 
     # cast to np arrays
     for key in state_vars.keys():
@@ -826,7 +827,7 @@ def plot(name, params, stride=STRIDE):
             logger.info('Saved %s/%s' % (snapshots_dir, savefig))
             plt.close()
 
-def write_front(name, params, stride=1):
+def write_front(name, params, stride=1, start=10):
     ''' few plots for front, defined where flux drops below 1/2 of theory '''
     populate_globals(params)
     # HACK HACK coerce N_X, N_Z to be loadable on exo15c
@@ -843,7 +844,7 @@ def write_front(name, params, stride=1):
     flux_threshold = flux_th * 0.3
 
     sim_times, domain, state_vars = load(
-        name, params, dyn_vars, stride=stride, start=10)
+        name, params, dyn_vars, stride=stride, start=start)
     x = domain.grid(0, scales=1)
     z = domain.grid(1, scales=1)
     dz = domain.grid_spacing(1, scales=1)[0]
@@ -999,7 +1000,7 @@ def get_dS_front(params, S_px, sim_times, z0):
         S_aboves.append(S_above)
     return np.array(dSpx), np.array(front_pos), front_idxs, np.array(S_aboves)
 
-def plot_front(name, params):
+def plot_front(name, params, start_time=None):
     populate_globals(params)
     N = np.sqrt(g / H)
     u_c = OMEGA / KX
@@ -1028,7 +1029,10 @@ def plot_front(name, params):
         Spx11 = np.array(Spx11) / flux_th
 
     tf = sim_times[-1]
-    start_idx = get_idx(200, sim_times)
+    if start_time is not None:
+        start_idx = get_idx(start_time, sim_times)
+    else:
+        start_idx = 0
     t = sim_times[start_idx: ]
 
     S_px0 = amps**2 * flux_th
[…]

### 2020-02-20 writeups/Su_IGW_break.tex
everything but new movie stuff is checked in

@@
-% TODO
-% - Figure 1: Could you include a tick mark at t=0? Does the curve start at t=0?
-% Section 6.2 needs more narrative discussion of the results.
-We should add another paragraph saying that the typical Richardson number of the critical layer is between 0.25 and 0.5, characteristic of Kelvin-Helmholtz instabilities.
-This suggestions the critical layer width is self-consistently determined by the ``turbulence’’ in our simulations.
-The layer width does not change with resolution.
-We should also say that here we calculate Ri using the background buoyancy frequency, because the fluctuations do not change N^2 significantly; thus, Ri~0.5 can be used to solve for the critical layer width in a star, as we know N^2 and U_c (corresponding to the tidal frequency Omega).
-% Investigate a smarter time averaging for the "average over everything
-% predictor"?
     \documentclass[
@@
 1024$ is available
-online\footnote{http://www.princeton.edu/~lecoanet/data/breaking\_wave.mov}. We
+online\footnote{http://www.princeton.edu/\~lecoanet/data/breaking\_wave.mov}. We
 take this to be our fiducial simulation for the remainder of this paper, though
@@
 \begin{align}
-    z_{CL, \min}(x, t) &\equiv \min_z
+    z_{CL, \min}(x, t) &\equiv \min
         \z{z\mid u_x(x, z, t) > 0.3\overline{U}_c},\\
-    z_{CL, \max}(x, t) &\equiv \max_z
+    z_{CL, \max}(x, t) &\equiv \max
         \z{z\mid u_x(x, z, t) < \overline{U}_c},\\
@@
 \end{align}
-To understand the variation in $\mathrm{Ri}$ over $x$, we also compute
-$\min\limits_x \mathrm{Ri}_x(x, t)$ (the maximum is very noisy). Both are shown
-in Fig.~\ref{fig:nl_f_ri}. The critical layer quickly narrows to its minimum
-width, driven by the anti-diffusivity of IGWs.
+We use the background buoyancy frequency to compute $\mathrm{Ri}$, as
+fluctuations do not change $N^2$ significantly ($\sim 1\%$). To understand the
+variation in $\mathrm{Ri}$ over $x$, we also compute $\min\limits_x
+\mathrm{Ri}_x(x, t)$ (the maximum is very noisy). Both are shown in
+Fig.~\ref{fig:nl_f_ri}. The Richardson number at the critical layer quickly
+decreases to a typical value between $0.25$ and $0.5$, characteristic of the
+onset of the KHI\@.
 \begin{figure}
@@
 
+The steadiness of $\mathrm{Ri}$ at this characteristic value suggests that the
+critical layer width is self-consistently determined by the ``turbulence'' in
+our simulations. The width does not change significantly with resolution in our
+resolved simulations (see Fig.~\ref{fig:agg}). Thus, in stars, $\mathrm{Ri} \sim
+0.5$ can be used to calculate the critical layer width, where $N^2$ and
+$\overline{U}_c$ (corresponding to the tidal frequency) are known.
+
 \subsection{Flux Budget}\label{ss:flux_budget}
@@
 
-In Fig.~\ref{fig:nl_front} we plot the numerically measured $z_c$ against two
-semi-analytic models: (i) integration of Eq.~\eqref{eq:zc_anal} using the
-measured $F_a(t)$, and (ii) substituting the time-averaged $\ev{F_a}_t$ (over
-the entire length of the simulation) into Eq.~\eqref{eq:zc_sol}. Since $z_c(t)$
-is less well-defined at early times (when the critical layer is thick and
-transient behavior is strong), we solve Eq.~\eqref{eq:zc_anal} by integrating
-backwards from the end of the simulation ($t = t_f$), using $z_c(t_f)$ as the
-initial condition. From Fig.~\ref{fig:nl_front}, we see that the agreement
-between the measured $z_c(t)$ and its estimate via $F_a(t)$ and
-Eq.~\eqref{eq:zc_anal} is excellent.
+In Fig.~\ref{fig:nl_front} we plot the numerically measured $z_c$ against
+numerical integration of Eq.~\eqref{eq:zc_anal} using the measured $F_a(t)$.
+Since the critical layer is still forming at early times, we solve
+Eq.~\eqref{eq:zc_anal} by integrating backwards from the end of the simulation
+($t = t_f$), using $z_c(t_f)$ as the initial condition. From
+Fig.~\ref{fig:nl_front}, we see that the agreement between the measured $z_c(t)$
+and its estimate via $F_a(t)$ is excellent.
 
 By time-averaging the numerically measured $F_a$, we find $\ev{F_a}_t \approx
-0.71F_{al}$, which is used for the second model in Fig.~\ref{fig:nl_front}.
-Note that $F_a < F_{al}$, so momentum flux absorption at the critical layer is
-incomplete. This is due to reflection of waves off the critical layer, which
-carry momentum downward.
+0.71F_{al}$. Note that $F_a < F_{al}$, so momentum flux absorption at the
+critical layer is incomplete. This is due to reflection of waves off the
+critical layer, which carry momentum downward.
 \begin{figure}
@@
     \caption{Propagation of the critical layer over time. Shown are: (black)
-    $z_c(t)$ from simulation data, (green) model for $z_c(t)$ using direct
+    $z_c(t)$ from simulation data, and (red) model for $z_c(t)$ using direct
     integration of Eq.~\eqref{eq:zc_anal} for $F_a(t)$ measured from simulation
-    data (described in Eq.~\eqref{eq:fa_def}), and (blue) direct substitution of
-    time-averaged $\ev{F_a(t)}_t$ into Eq.~\eqref{eq:zc_sol}. The models use the
-    end of the simulation as initial conditions and integrate backwards, as
-    $z_c$ is less well-defined at early times. The agreement of the
-    directly-integrated model with the data shows Eq.~\eqref{eq:zc_anal} is a
-    good description of the evolution of $z_c$. The the poorer but qualitatively
-    correct agreement of the time-averaged model with the data shows both that
-    $F_a(t) < F_{al}$.}\label{fig:nl_front}
+    data (described in Eq.~\eqref{eq:fa_def}). The model uses the end of the
+    simulation as its initial condition and integrates backwards, as the
+    critical layer is still forming at earlier times. The agreement of the model
+    with the simulation shows Eq.~\eqref{eq:zc_anal} is a good description of
+    the evolution of $z_c$.}\label{fig:nl_front}
 \end{figure}

### 2020-02-20 sims/2d_4_fourier/plot_2d_series_movie.py

@@ -33,7 +33,7 @@ Z_TOP = 9.5
 
 # save two versions of these indicies, one version to put in paper (times are
 # labeled)
-PLOT_IDXS = [231, 50, 86, 451]
+PLOT_IDXS = [40, 89, 134, 451]
 PLOT_PREFIXES = ['(i)', '(ii)', '(iii)', '(iv)']
 
 def main(filename, start, count, output, write_start):
@@ -112,7 +112,7 @@ def main(filename, start, count, output, write_start):
             try:
                 fig_idx = PLOT_IDXS.index(suffix_num)
                 annotation.remove()
-                print('Adding extra annotation')
+                print('Adding extra annotation', suffix_num)
                 overlay = '%s %s' % (PLOT_PREFIXES[fig_idx], time_str)
                 paxes0.text(0.1, 10.8, overlay, fontsize=14)
             except ValueError:

### 2020-02-20 sims/2d_4_fourier/strat_helper.py

@@ -1043,7 +1043,9 @@ def plot_front(name, params, start_time=None):
     dSpx, front_pos, front_idxs, S_aboves =\
         get_dS_front(params, S_px, sim_times, z0)
 
-    times = get_times([1/15, 1/7, 3/10, 1], sim_times, start_idx)
+    times = get_times([9/100, 1/5, 3/10, 1], sim_times, start_idx)
+    times[0] + 1 # this fraction thing is killing me
+    print(times)
     fig = plt.figure()
     if 'lin' in name:
         #####################################################################
@@ -1068,6 +1070,9 @@ def plot_front(name, params, start_time=None):
 
         plt.xlabel(r'$z / H$', fontsize=int(1.5 * FONTSIZE))
         plt.ylabel(r'$F / F_{al}$', fontsize=int(1.5 * FONTSIZE))
+        axes = plt.gca()
+        axes.xaxis.set_ticks_position('both')
+        axes.yaxis.set_ticks_position('both')
         plt.tight_layout()
         plt.savefig('%s/fluxes.png' % snapshots_dir, dpi=DPI)
         plt.close()
@@ -1102,7 +1107,10 @@ def plot_front(name, params, start_time=None):
         #          label=r'$\phi_I$',
         #          linewidth=LW * 0.7)
         ax1.set_xlim(left=0)
-        print(t[0], t[-1])
+        ax1.set_ylim(0.96, 1.025)
+        axes = plt.gca()
+        axes.xaxis.set_ticks_position('both')
+        axes.yaxis.set_ticks_position('both')
         plt.tight_layout()
         plt.savefig('%s/f_amps.png' % snapshots_dir, dpi=DPI)
         plt.close()
@@ -1130,7 +1138,7 @@ def plot_front(name, params, start_time=None):
                      u0_avg / u_c,
                      '%s-' % color,
                      linewidth=LW * 0.5,
-                     label=r'$t=%.1f/N$' % sim_times[time])
+                     label=r'$t=%d/N$' % sim_times[time])
             # S_px sliced at time, just one for comparison
             # if time == times[len(times) // 2]:
             # #     ax2.plot(z0_cut,
@@ -1144,7 +1152,7 @@ def plot_front(name, params, start_time=None):
                      S_px_avg / flux_th,
[…]

### 2020-02-21 writeups/Su_IGW_break.tex
everything is done except for movie plots now

@@
 % cp ../sims/agg.png plots; cp ../sims/2d_4_fourier/snapshots_lin_0_masked/f_amps.png plots/lin_amps.png; cp ../sims/2d_4_fourier/snapshots_lin_0_masked/fluxes.png plots/lin_fluxes.png; cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/f_amps2.png plots/nl_f_amps2.png; cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/f_amps.png plots/nl_f_amps.png; cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/fluxes.png plots/nl_fluxes.png; cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/f_refl.png plots/nl_f_refl.png; cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/f_ri.png plots/nl_f_ri.png; cp ../sims/2d_4_fourier/snapshots_yubo_nu1_vhres/front.png plots/nl_front.png;
+% cp ../sims/2d_4_fourier/frames2/*labeled.png plots
 
@@
 
-\subsection{Wave Generation}
+\subsection{Wave Generation}\label{ss:forcing}
 
@@
 The geometry of our simulation domain is fixed by one further parameter, $z_0$,
-the forcing location. We choose $L_z = 12.5H$, and the lower and upper damping
-zones are located at $z < 0.3H$ and $z > 9.5H$ respectively ($z \in [-1,
-11.5]$). The forcing (see Eq.~\eqref{eq:vol_drive}) is at $z_0 = 2H$ with width
-$\sigma = 0.078H$, sufficiently far from the lower damping zone and permitting
-sufficient room for the upward propagating wave to grow as $\propto e^{z/2H}$.
-Finally, we want similar grid spacing in the $x$ and $z$ directions (i.e.\ $L_x
-/ N_x \sim L_z /N_z$), guided by the intuition that turbulence generated by wave
-breaking is approximately isotropic, so we use $L_x = 4H$ and $N_z / N_x = 4$.
+the forcing location. We choose $L_z = 12.5H$ ($z$ runs from $-H$ to $11.5H$),
+and the lower and upper damping zones are active for $z < 0.3H$ and $z > 9.5H$
+respectively. The forcing (see Eq.~\eqref{eq:vol_drive}) is at $z_0 = 2H$ with
+width $\sigma = 0.078H$, sufficiently far from the lower damping zone and
+permitting sufficient room for the upward propagating wave to grow as $\propto
+e^{z/2H}$. Finally, we want similar grid spacing in the $x$ and $z$ directions
+(i.e.\ $L_x / N_x \sim L_z /N_z$), guided by the intuition that turbulence
+generated by wave breaking is approximately isotropic, so we use $L_x = 4H$ and
+$N_z / N_x = 4$.
 
@@
 \end{align}
-where $z_B = 1.5H$ and $z_T = 10.5H$ are the boundaries of the lower and upper
+where $z_B = 0.3H$ and $z_T = 9.5H$ are the boundaries of the lower and upper
 damping zones respectively. This damps perturbations below $z_B$ and
@@
 \begin{figure*}
-    \includegraphics[width=0.45\textwidth]{plots/yubo_000054.png}\hfil
+    \includegraphics[width=0.45\textwidth]{plots/yubo_000040_labeled.png}\hfil
     %
-    \includegraphics[width=0.45\textwidth]{plots/yubo_000086.png}
-
-    \includegraphics[width=0.45\textwidth]{plots/yubo_000153.png}\hfil
+    \includegraphics[width=0.45\textwidth]{plots/yubo_000089_labeled.png}
+
+    \includegraphics[width=0.45\textwidth]{plots/yubo_000134_labeled.png}\hfil
     %
-    \includegraphics[width=0.45\textwidth]{plots/yubo_000451.png}
-    \caption{Snapshots of $u_x$ (in units of $HN$) and $\Upsilon \equiv
-    \ln\p{\rho / \bar{\rho}}$ in the fiducial simulation illustrating distinct
-    phases of the evolution of the flow. Note that $\overline{U}_c \approx
-    0.16HN$ (see Eq.~\eqref{eq:u_crit}) for the parameters used. The four times
-    depicted are $t = 413.4/N$ (top left), $t = 658.5/N$ (top right), $t =
-    1171.4/N$ (bottom left), and $t = 3437.8/N$ (bottom right). The four
-    snapshots illustrate (i) the initial transient wave breaking phase, (ii)
-    formation of a distinct critical layer, (iii) steepening of the critical
-    layer, and (iv) downward advance of the critical
+    \includegraphics[width=0.45\textwidth]{plots/yubo_000451_labeled.png}
+    \caption{Snapshots of $u_x$ and $\Upsilon$ in the fiducial simulation
+    illustrating distinct phases of the evolution of the flow. Damping layers
+    are shaded in light grey (see Section~\ref{ss:damping}), while the forcing
+    zone (see Section~\ref{ss:forcing}) is shaded in light green (boundaries are
+    at $z_0 \pm 3\sigma$). The four panels illustrate (i) the initial transient
+    wave breaking phase, (ii) formation of a distinct critical layer, (iii)
+    steepening of the critical layer, and (iv) downward advance of the critical
     layer.}\label{fig:snapshots}
@@
 \mathrm{Ri}_x(x, t)$ (the maximum is very noisy). Both are shown in
-Fig.~\ref{fig:nl_f_ri}. The Richardson number at the critical layer quickly
-decreases to a typical value between $0.25$ and $0.5$, characteristic of the
-onset of the KHI\@.
+Fig.~\ref{fig:nl_f_ri}. Absorbtion of incident IGWs quickly decreases the
+Richardson number to between $0.25$ and $0.5$, characteristic of the onset of
+the KHI\@.
+
+This result suggests that the critical layer width is regulated by the
+competition between steepening induced by IGW breaking and broadening due to
+shear instability. This width does not vary significantly with resolution in our
+resolved simulations (see Fig.~\ref{fig:agg}). $\mathrm{Ri} \sim 0.5$ can be
+used to calculate the critical layer width in stars, where $N^2$ and
+$\overline{U}_c$ (corresponding to the tidal frequency) are known.
 \begin{figure}
@@
 \end{figure}
-
-The steadiness of $\mathrm{Ri}$ at this characteristic value suggests that the
-critical layer width is self-consistently determined by the ``turbulence'' in
-our simulations. The width does not change significantly with resolution in our
-resolved simulations (see Fig.~\ref{fig:agg}). Thus, in stars, $\mathrm{Ri} \sim
-0.5$ can be used to calculate the critical layer width, where $N^2$ and
-$\overline{U}_c$ (corresponding to the tidal frequency) are known.
 
@@
 geometry, they found no evidence for reflected waves, contrary to our result.
-% It is plausible that the discrepancy arises due to the differing causes of wave
-% breaking (geometric focusing vs.\ density rarefaction). Note that their
-% nonlinearity parameter $\lambda \sim 1/\mathrm{Re}$ where $\mathrm{Re}$ is the
-% Reynolds number (Eq.~\eqref{eq:re_def}). For $\lambda \sim 10^{-1}$ as explored
-% in \citet{barker_ogilvie}, our results also show no reflected wave.
-Note, however, that their simulations were run with substantially higher
-viscosity, i.e.\ lower resolution (their $\lambda \sim 1/\mathrm{Re}$, so they
-studied $\mathrm{Re} \sim 10$), than the simulations presented here. We also
-find at low Reynolds number that there is negligible wave reflection. It is
-possible that reflection and flux redistribution are also important processes
-for geometrically breaking waves, but was absent from the low-Reynolds number
-simulations of \citet{barker_ogilvie}.
+Note that their simulations were run with substantially higher viscosity, or
+lower resolution, than explored here, and their effective Reynolds number (equal
+to $1/\lambda$ in their notation) is of order $10$. We also find at low Reynolds
+numbers that there is negligible wave reflection.
 

### 2020-02-21 writeups/Su_IGW_break.tex

@@
     \includegraphics[width=0.45\textwidth]{plots/yubo_000451_labeled.png}
-    \caption{Snapshots of $u_x$ and $\Upsilon$ in the fiducial simulation
-    illustrating distinct phases of the evolution of the flow. Damping layers
-    are shaded in light grey (see Section~\ref{ss:damping}), while the forcing
-    zone (see Section~\ref{ss:forcing}) is shaded in light green (boundaries are
-    at $z_0 \pm 3\sigma$). The four panels illustrate (i) the initial transient
-    wave breaking phase, (ii) formation of a distinct critical layer, (iii)
+    \caption{Snapshots of $u_x$ and $\Upsilon \equiv \ln \p{\rho / \bar{\rho}}$
+    in the fiducial simulation illustrating distinct phases of the evolution of
+    the flow. Damping layers are shaded in light grey (see
+    Section~\ref{ss:damping}), while the forcing zone (see
+    Section~\ref{ss:forcing}) is shaded in light green (boundaries are at $z_0
+    \pm 3\sigma$). The four panels illustrate (i) the initial transient wave
+    breaking phase, (ii) formation of a distinct critical layer, (iii)
     steepening of the critical layer, and (iv) downward advance of the critical
@@
 shear instability. This width does not vary significantly with resolution in our
-resolved simulations (see Fig.~\ref{fig:agg}). $\mathrm{Ri} \sim 0.5$ can be
-used to calculate the critical layer width in stars, where $N^2$ and
+resolved simulations (see Fig.~\ref{fig:agg}). As such, $\mathrm{Ri} \sim 0.5$
+can be used to calculate the critical layer width in stars, where $N^2$ and
 $\overline{U}_c$ (corresponding to the tidal frequency) are known.

### 2020-02-23 writeups/Su_IGW_break.tex
fix tilde at last

@@
 Daniel Lecoanet,$^{2,3}$
-Dong Lai$^{1,4}$
+Dong Lai$^{1,4, 5}$
 \\
@@
 \\
-$^4$ Tsung-Dao Lee Institute, Shanghai Jiao Tong University, Shanghai, 200240, China
+$^4$ Tsung-Dao Lee Institute, Shanghai Jiao Tong University, Shanghai, 200240,
+China
+\\
+$^5$ Department of Astronomy and Miller Institute for Basic Research In Science,
+UC Berkeley, Berkeley, CA 94720, USA
 }
@@
     flow velocity equal to the horizontal wave phase speed). While the critical
-    layer is steepened by these waves, it is simultaneously broadened by
-    Kelvin-Helmholtz instabilities such that, in steady state, the critical
-    layer width is determined by the Richardson criterion. We study the
+    layer is steepened by absorption of these waves, it is simultaneously
+    broadened by Kelvin-Helmholtz instabilities such that, in steady state, the
+    critical layer width is determined by the Richardson criterion. We study the
     absorption and reflection of incident waves off the critical layer and
-    provide analytical formulae describing its evolution. The result of this
-    study is important for characterizing the evolution of tidally heated white
-    dwarfs and other binary stars.
+    provide analytical formulae describing its long-term evolution. The result
+    of this study is important for characterizing the evolution of tidally
+    heated white dwarfs and other binary stars.
 \end{abstract}
@@
 of the nonlinear IGW breaking process; the result of this ``microphysics'' study
-will help determine the thermal evolution and and the observational
-manifestations of tidally heated binary WDs.
+will help determine the thermal evolution and the observational manifestations
+of tidally heated binary WDs.
 
@@
 1024$ is available
-online\footnote{http://www.princeton.edu/\~lecoanet/data/breaking\_wave.mov}. We
+online\footnote{\url{http://www.princeton.edu/\~lecoanet/data/breaking\_wave.mov}}. We
 take this to be our fiducial simulation for the remainder of this paper, though
@@
 stellar radius) much smaller than the radial wave number $k_r$. While our
-simulations do not have $\abs{k_x/k_z}\ll 1$, the qualitative behavior is likely
+simulations do not satisfy $\abs{k_x/k_z}\ll 1$, the qualitative behavior is likely
 to be similar, as the turbulence driving the critical layer dynamics occurs at

### 2020-02-25 writeups/Su_IGW_break.tex
typo in abstract

@@
     after an initial transient phase, the dissipation of the IGWs naturally
-    generate a sharp critical layer, separating the lower stationary region
+    generates a sharp critical layer, separating the lower stationary region
     (with no mean flow) and the upper ``synchronized'' region (with the mean

### 2020-04-04 writeups/Su_IGW_break.tex
responded to half of ref comments

@@
         newtxmath,
-        ae, aecompl,
+        ae,
+        aecompl,
         graphicx,
         booktabs,
+        xcolor,
+        soul,
     }
@@
     \DeclareMathOperator*{\erf}{erf}
+
+    \colorlet{Corr}{red}
 
@@
 \begin{abstract}
-    In binaries composed of early-type stars and white dwarfs, the dominant
-    tidal process involves the excitation of internal gravity waves (IGWs),
-    which propagate towards the stellar surface, and their dissipation via
-    nonlinear wave breaking. We perform 2D hydrodynamical simulations of this
-    wave breaking process in a stratified, isothermal atmosphere. We find that,
-    after an initial transient phase, the dissipation of the IGWs naturally
-    generates a sharp critical layer, separating the lower stationary region
-    (with no mean flow) and the upper ``synchronized'' region (with the mean
-    flow velocity equal to the horizontal wave phase speed). While the critical
-    layer is steepened by absorption of these waves, it is simultaneously
-    broadened by Kelvin-Helmholtz instabilities such that, in steady state, the
-    critical layer width is determined by the Richardson criterion. We study the
-    absorption and reflection of incident waves off the critical layer and
-    provide analytical formulae describing its long-term evolution. The result
-    of this study is important for characterizing the evolution of tidally
-    heated white dwarfs and other binary stars.
+    In binaries composed of \textcolor{Corr}{either early-type stars or white
+    dwarfs}, the dominant tidal process involves the excitation of internal
+    gravity waves (IGWs), which propagate towards the stellar surface, and their
+    dissipation via nonlinear wave breaking. We perform 2D hydrodynamical
+    simulations of this wave breaking process in a stratified, isothermal
+    atmosphere. We find that, after an initial transient phase, the dissipation
+    of the IGWs naturally generates a sharp critical layer, separating the lower
+    stationary region (with no mean flow) and the upper ``synchronized'' region
+    (with the mean flow velocity equal to the horizontal wave phase speed).
+    While the critical layer is steepened by absorption of these waves, it is
+    simultaneously broadened by Kelvin-Helmholtz instabilities such that, in
+    steady state, the critical layer width is determined by the Richardson
+    criterion. We study the absorption and reflection of incident waves off the
+    critical layer and provide analytical formulae describing its long-term
+    evolution. The result of this study is important for characterizing the
+    evolution of tidally heated white dwarfs and other binary stars.
 \end{abstract}
@@
 stellar envelope or atmosphere. We approximate the equation of state as
-baratropic, i.e.\ $P(\rho, T) = P(\rho)$. We study dynamics in 2D, so that fluid
-variables depend only on the Cartesian coordinates $x$ and $z$. While it is
-well-known wave breaking is a 3D process \citep{klostermeyer,winters1994}, the
-dynamical effect of the breaking process is likely to be similar in 2D
-\citep{barker_ogilvie}. We approximate the gravitational field as uniform,
-pointing in the $(-\uv{z})$ direction. The plane-parallel approximation is
-justified since wave breaking generally occurs near the stellar surface. The
-background density stratification is given by
+\textcolor{Corr}{barotropic}, i.e.\ $P(\rho, T) = P(\rho)$. We study dynamics in
+2D, so that fluid variables depend only on the Cartesian coordinates $x$ and
+$z$. \textcolor{Corr}{While it is well known that waves break differently in two
+versus three dimensions} \citep{klostermeyer,winters1994}, the dynamical effect
+of the breaking process is likely to be similar in 2D \citep{barker_ogilvie}. We
+approximate the gravitational field as uniform, pointing in the $(-\uv{z})$
+direction. The plane-parallel approximation is justified since wave breaking
+generally occurs near the stellar surface. The background density stratification
+is given by
 \begin{equation}
@@
 acceleration is $-g\uv{z}$. For this isothermal background, hydrostatic
-equilibrium implies $P = \overline{\rho} g H$. We assume there is initially no
-background flow, so $\bm{u} = \bm{u}'$. Physically, this assumption corresponds
-to a non-rotating star.
+equilibrium implies \textcolor{Corr}{$\overline{P}(z) = \overline{\rho}(z) g
+H$}. We assume there is initially no background flow, so $\bm{u} = \bm{u}'$.
+Physically, this assumption corresponds to a non-rotating star.
 
@@
 
-In the small perturbation limit, we may linearize Eq.~\eqref{se:nl_upsilon}. The
-solution in this linear regime is given by \citep{drazin,sutherland0}
+In the small perturbation limit, we may linearize Eq.~\eqref{se:nl_upsilon}.
+\textcolor{Corr}{These equations admit the canonical IGW solution}, given by
+\citep{drazin,sutherland0}
 \begin{equation}
@@
         \end{align}
+        \textcolor{Corr}{The additional $\p{2H}^{-2}$ term in the denominator
+        accounts for the growing amplitude of the IGW in the $z$ direction (the
+        $z$ wavenumber is effectively $k_z - i / (2H)$).}
         We note $\bm{c}_{p} \cdot \bm{c}_g = \mathcal{O}\s{\p{k_{z}H}^{-2}}
@@
 Consider now the long-term evolution of the critical layer due to continuous
-horizontal momentum transfer by IGWs. Any incident horizontal momentum flux
-$F_a(t)$ absorbed by the fluid must manifest as additional horizontal momentum
-of the shear flow. Since $\overline{U}$ cannot exceed $\overline{U}_c$, the
-critical layer will instead propagate downward in response to the incident
-momentum flux. The horizontal momentum of the shear flow satisfies
+horizontal momentum transfer by IGWs. \textcolor{Corr}{Any incident horizontal
+momentum flux absorbed by the fluid, denoted $F_a(t)$, must manifest as
+additional horizontal momentum of the shear flow}.
+\textcolor{Corr}{Additionally, as the mean flow $\overline{U}$ cannot grow
+efficiently above $\overline{U}_c$ (due to the breakdown of the linear
+solution), we assume $\overline{U}$ saturates at $\overline{U}_c$, which holds
+to good accuracy (see Fig.~\ref{fig:nl_fluxes}). Instead, the critical layer
+must propagate downward in response to the incident momentum flux.} The
+horizontal momentum of the shear flow satisfies
 \begin{equation}
@@
 periodic boundary conditions in both the $x$ and $z$ direction. To mimic the
-absence of physical boundaries at the top/bottom of the simulation domain, we damp
-perturbations to zero near the top/bottom using damping zones
-(see Section~\ref{ss:damping}). We expand all variables as Fourier series
-with $N_x$ and $N_z$ modes, and use the $3/2$ dealiasing rule to avoid aliasing
-errors in the nonlinear terms \citep{boyd}.
-
-The geometry of our simulation domain is fixed by one further parameter, $z_0$,
-the forcing location. We choose $L_z = 12.5H$ ($z$ runs from $-H$ to $11.5H$),
-and the lower and upper damping zones are active for $z < 0.3H$ and $z > 9.5H$
-respectively. The forcing (see Eq.~\eqref{eq:vol_drive}) is at $z_0 = 2H$ with
-width $\sigma = 0.078H$, sufficiently far from the lower damping zone and
-permitting sufficient room for the upward propagating wave to grow as $\propto
-e^{z/2H}$. Finally, we want similar grid spacing in the $x$ and $z$ directions
-(i.e.\ $L_x / N_x \sim L_z /N_z$), guided by the intuition that turbulence
-generated by wave breaking is approximately isotropic, so we use $L_x = 4H$ and
-$N_z / N_x = 4$.
+absence of physical boundaries at the top/bottom of the simulation domain, we
+damp perturbations to zero near the top/bottom using damping zones (see
+Section~\ref{ss:damping}). We expand all variables as Fourier series with $N_x$
+and $N_z$ modes, and use the $3/2$ dealiasing rule to avoid aliasing errors in
+the nonlinear terms \citep{boyd}. We choose $L_z = 12.5H$ ($z$ runs from $-H$ to
+$11.5H$), and the lower and upper damping zones are active for $z < 0.3H$ and $z
+> 9.5H$ respectively. The forcing (see Eq.~\eqref{eq:vol_drive}) is centered at
+$z_0 = 2H$ with width $\sigma = 0.078H$, sufficiently far from the lower damping
+zone and permitting sufficient room for the upward propagating wave to grow as
+$\propto e^{z/2H}$. Finally, we want similar grid spacing in the $x$ and $z$
+directions (i.e.\ $L_x / N_x \sim L_z /N_z$), guided by the intuition that
+turbulence generated by wave breaking is approximately isotropic, so we use $L_x
+= 4H$ and $N_z / N_x = 4$.
 
@@
 above $z_T$ with damping time $\tau$ and negligibly affects the dynamics between
-$z_B$ and $z_T$. We choose the transition width $\Delta z = 0.25H$ and damping
-time $\tau = 1 / (15N)$. This prescription is similar to \citet{lecoanet_damp}
-and has the advantage of being smooth, important for spectral methods. Further
-details of our implementation of the fluid equations in Dedalus are described in
-Appendix~\ref{se:strat_impl}.
+$z_B$ and $z_T$. \textcolor{Corr}{Most importantly, horizontal momentum is still
+conserved between $z_B$ and $z_T$}. We choose the transition width $\Delta z =
+0.25H$ and damping time $\tau = 1 / (15N)$. This prescription is similar to
+\citet{lecoanet_damp} and has the advantage of being smooth, important for
+spectral methods. Further details of our implementation of the fluid equations
+in Dedalus are described in Appendix~\ref{se:strat_impl}.
 
@@
 The subscript $i$ denotes the incident wave. If $\bm{u} = \bm{u}_{al}$, then
-$A_i(t) = 1$. The normalization in Eq.~\eqref{eq:ahat_def} is chosen such that
-the overlap between $\bm{u}, \bm{u}_{al} \propto e^{z/2H}$ is evenly weighted
-throughout the integration region.
+$A_i(t) = 1$. \textcolor{Corr}{The factor of $\bar{\rho}$ inside the integrands
+in Eq.~\eqref{eq:ahat_def} corrects for the $\propto e^{z/2H}$ growth of
+$\bm{u}_{al}$; without it, $A_i(t)$ would be dominated by the contribution near
+$z_t$.}
 
@@
 
-% GOT HERE
-
 \section{Numerical Simulations of Wave Breaking}\label{s:sim}
@@
 $\abs{\xi_z k_z} = 0.1$ in the forcing zone ($z = z_0$). The linear solution
-predicts $\abs{\xi_z k_z} \sim 4.25$ at the upper damping zone $z_T$. We choose
-the viscosity $\nu$ so that $\mathrm{Re}$ is as large as possible across the
-various resolutions. A table of our simulations can be found in
-Table~\ref{tab:params}.
+predicts $\abs{\xi_z k_z} \sim 4.25$ at the upper damping zone $z_T$.
+\textcolor{Corr}{We choose the viscosity $\nu$ for each resolution to be as
+small as possible while still resolving the shortest spatial scales of the wave
+breaking.} A table of our simulations can be found in Table~\ref{tab:params}.
 \begin{table}
@@
 In Fig.~\ref{fig:snapshots}, we present snapshots of $u_x$ and $\Upsilon$ at
-various phases of the simulation. The flow evolves through several distinct
-stages:
+various phases of the simulation. \textcolor{Corr}{Note that $\Upsilon \lesssim
+0.1$, so the density stratification does not deviate significantly from
+equilibrium.} The flow evolves through several distinct stages:
 \begin{enumerate}
@@
 
-    \item At the end of the simulation (bottom right panel), the flow shows very
-        few significant qualitative differences from the previous snapshot,
-        suggesting that the latter phase of the simulation has reached a steady
-        state.
+    \item At the end of the simulation (bottom right panel), the
+        \textcolor{Corr}{critical layer has advanced downwards, but otherwise}
+        the flow shows very few significant qualitative differences from the
+        previous snapshot. This suggests that the latter phase of the simulation
+        has reached a steady state.
+
+        \textcolor{Corr}{Notably, the horizontal banded structure of $u_x$ in
+        the synchronized fluid is mostly unchanged (also visible in the top
+        panel of Fig.~\ref{fig:nl_fluxes}), suggesting that momentum
+        redistribution and mixing within the synchronized fluid are negligible.}
 \end{enumerate}
@@
 is the width of this layer? Inspection of Fig.~\ref{fig:snapshots} suggests the
-presents of the Kelvin-Helmholtz Instability (KHI) in the critical layer. In a
-stratified medium, KHI occurs when the Richardson number (Eq.~\eqref{eq:ri_def})
-satisfies $\mathrm{Ri} \lesssim 1/4$ \citep[e.g.][]{shu1991physics}. It is
-natural to suspect that the shear flow cannot steepen further than the onset of
-KHI\@. To test this, we compute the local $\mathrm{Ri}$ for the shear flow
-around the critical layer.
+\textcolor{Corr}{presence} of the Kelvin-Helmholtz Instability (KHI) in the
+critical layer. In a stratified medium, KHI occurs when the Richardson number
+(Eq.~\eqref{eq:ri_def}) satisfies $\mathrm{Ri} \lesssim 1/4$
+\citep[e.g.][]{shu1991physics}. It is natural to suspect that the shear flow
+cannot steepen further than the onset of KHI\@. To test this, we compute the
+local $\mathrm{Ri}$ for the shear flow around the critical layer.
 
@@
 \mathrm{Ri}_x(x, t)$ (the maximum is very noisy). Both are shown in
-Fig.~\ref{fig:nl_f_ri}. Absorbtion of incident IGWs quickly decreases the
-Richardson number to between $0.25$ and $0.5$, characteristic of the onset of
-the KHI\@.
+Fig.~\ref{fig:nl_f_ri}. \textcolor{Corr}{Absorption} of incident IGWs quickly
+decreases the Richardson number to between $0.25$ and $0.5$, characteristic of
+the onset of the KHI\@.
 

### 2020-04-04 writeups/Su_IGW_break.tex
check in comments, add typo fixes

@@
 strongly affect the pre-merger conditions of the WDs and therefore the merger
-outcomes. Tidal dessipation may also influence the evoluton of eccentric
+outcomes. Tidal dissipation may also influence the evolution of eccentric
 WD-massive black hole binaries prior to the eventual tidal disruption of the WD
@@
 process in compact WD binaries \citep{fullerII,fullerIV,tidal_novae,bukart}:
-IGWs are tidally excited mainly at the compositon transitions of the WD
+IGWs are tidally excited mainly at the composition transitions of the WD
 envelope; as these waves propagate outwards towards the WD surface, they grow in
@@
 some detail. At intermediate amplitudes, wave breaking occurs via triadic
-resonances, transfering energy from the ``parent'' IGW to ``daughter'' waves on
+resonances, transferring energy from the ``parent'' IGW to ``daughter'' waves on
 smaller length scales that efficiently damp. The horizontal momentum flux
@@
 The analytical theory (Section~\ref{ss:lin_analysis}) also predicts that the
-horizontal momentmum flux $F(z, t)$ is independent of $z$ between the forcing
+horizontal momentum flux $F(z, t)$ is independent of $z$ between the forcing
 zone where the wave is generated and the damping zone where it is dissipated.
@@
     good approximation above the forcing zone $z > z_0 = 2H$ and below the
-    damping zone $z \lesssim z_T = 9.5H$. The horizonal momentum flux excited in
-    the forcing zone is transported without loss to the top of the domain, where
-    it is dissipated by the damping zone (see Section~\ref{ss:damping}) without
-    reflection.}\label{fig:lin_fluxes}
+    damping zone $z \lesssim z_T = 9.5H$. The horizontal momentum flux excited
+    in the forcing zone is transported without loss to the top of the domain,
+    where it is dissipated by the damping zone (see Section~\ref{ss:damping})
+    without reflection.}\label{fig:lin_fluxes}
 \end{figure}
@@
 is derived from a linear theory, while fluid motion within the critical layer is
-turbulent, so reflecton at the critical layer cannot be captured by the linear
+turbulent, so reflection at the critical layer cannot be captured by the linear
 theory.
@@
 k_z} \sim 1 / (2\pi)$ in a plane-parallel geometry. Tidally excited IGWs in
-binary stars have horizontal wave number $k_\perp \sim 1/R$ (where $R$ is the
-stellar radius) much smaller than the radial wave number $k_r$. While our
+binary stars have horizontal wavenumber $k_\perp \sim 1/R$ (where $R$ is the
+stellar radius) much smaller than the radial wavenumber $k_r$. While our
 simulations do not satisfy $\abs{k_x/k_z}\ll 1$, the qualitative behavior is likely
@@
 For simplicity, we use the same diffusivity $\nu$ for both the momentum and mass
-diffusivities. Although mass diffusivity is not physical, we incclude it for
+diffusivities. Although mass diffusivity is not physical, we include it for
 numerical stability. We choose the mass diffusion term to conserve mass, and not

### 2020-04-05 writeups/Su_IGW_break.tex
check in colorblind plots, need to update captions

@@
     amplitude. The small deviation of $A_i(t)$ from unity may be due to
-    timestepping errors, as a relatively large fixed step size $\Delta t =
-    0.1/N$ was used for this simulation.}\label{fig:lin_amps}
+    \textcolor{Corr}{truncation error in our implicit timestepping scheme}, as a
+    relatively large fixed step size $\Delta t = 0.1/N$ was used for this
+    simulation.}\label{fig:lin_amps}
 \end{figure}
@@
     in the fiducial simulation illustrating distinct phases of the evolution of
-    the flow. Damping layers are shaded in light grey (see
-    Section~\ref{ss:damping}), while the forcing zone (see
+    the flow. \textcolor{Corr}{See the online PDF for a color version.} Damping
+    layers \textcolor{Corr}{at the top and bottom of the simulation domain} are
+    shaded in light grey (see Section~\ref{ss:damping}), while the forcing zone
+    \textcolor{Corr}{in the lower middle portion of the simulation domain} (see
     Section~\ref{ss:forcing}) is shaded in light green (boundaries are at $z_0

### 2020-04-05 sims/2d_4_fourier/strat.py

@@ -71,13 +71,13 @@ def get_params(overrides=None):
 
 if __name__ == '__main__':
     tasks = [
-        # ('lin_0_masked',
-        #  get_params(overrides={'F_MULT': 0.0005,
-        #                        'T_F': 4000,
-        #                        'DT': 0.1,
-        #                        'N_X': 64,
-        #                        'N_Z': 256,
-        #                        'Re_inv': 1e-4})),
+        ('lin_0_masked',
+         get_params(overrides={'F_MULT': 0.0005,
+                               'T_F': 4000,
+                               'DT': 0.1,
+                               'N_X': 64,
+                               'N_Z': 256,
+                               'Re_inv': 1e-4})),
         # ('yubo_nu0p5_hres',
         #  get_params(overrides={'Re_inv': 0.05})),
         # ('yubo_nu0p5_shres',

### 2020-04-05 sims/2d_4_fourier/strat_helper.py

@@ -15,7 +15,8 @@ import matplotlib.pyplot as plt
 import numpy as np
 from scipy.interpolate import interp1d
 from scipy.optimize import minimize
-PLT_COLORS = ['b', 'g', 'r', 'c', 'm', 'y', 'k', 'w'];
+# PLT_COLORS = ['b', 'g', 'r', 'c', 'm', 'y', 'k', 'w'];
+PLT_STYLES = ['k-', 'b--', 'r-.', 'g:'];
 
 # Not needed for plotting after pkl is generated
 try:
@@ -1055,10 +1056,12 @@ def plot_front(name, params, start_time=None):
         #####################################################################
         z_0_idx = get_idx(Z0, z0)
         z0_cut = z0[z_0_idx: ]
-        for time in times:
+        for idx, (time, style) in enumerate(zip(times, PLT_STYLES[::-1])):
             plt.plot(z0_cut,
                      S_px[time, z_0_idx: ] / flux_th,
-                     linewidth=LW * 0.7,
+                     style,
+                     linewidth=LW * 0.8,
+                     alpha=0.7,
                      label=r'$t=%.1f/N$' % sim_times[time])
         # plt.plot(z0_cut,
         #          np.exp(-k_damp * 2 * (z0_cut - Z0)),
@@ -1125,7 +1128,7 @@ def plot_front(name, params, start_time=None):
         f, (ax1, ax2) = plt.subplots(2, 1, figsize=(6, 8), sharex=True)
         # z0_cut = z0[z_b_idx: ]
 
-        for time, color in zip(times[::-1], PLT_COLORS):
+        for idx, (time, style) in enumerate(zip(times, PLT_STYLES[::-1])):
             u0_avg = np.mean(u0[time - AVG_IDX: time + AVG_IDX,
                                 # z_b_idx: ], axis=0)
                                 : ], axis=0)
@@ -1136,8 +1139,9 @@ def plot_front(name, params, start_time=None):
             # ax1.plot(z0_cut,
             ax1.plot(z0,
                      u0_avg / u_c,
-                     '%s-' % color,
-                     linewidth=LW * 0.5,
+                     style,
+                     alpha=0.7,
+                     linewidth=LW * 0.8,
                      label=r'$t=%d/N$' % sim_times[time])
             # S_px sliced at time, just one for comparison
             # if time == times[len(times) // 2]:
[…]

### 2020-04-05 sims/aggregate.py

@@ -41,11 +41,11 @@ if __name__ == '__main__':
         ax1.errorbar(x, r_med,
                      [[r_med - r_min], [r_max - r_med]],
                      ecolor='k', linewidth=lwidth)
-        ax1.plot(x + 2 * offset_mult, T_med, 'ro', markersize=msize)
+        ax1.plot(x + 2 * offset_mult, T_med, 'r*', markersize=msize)
         ax1.errorbar(x + 2 * offset_mult, T_med,
                      [[T_med - T_min], [T_max - T_med]],
                      ecolor='r', linewidth=lwidth)
-        ax1.plot(x + offset_mult, rA_med**2, 'bo', markersize=msize)
+        ax1.plot(x + offset_mult, rA_med**2, 'bx', markersize=msize)
         ax1.errorbar(x + offset_mult, rA_med**2,
                      [[rA_med**2 - rA_min**2], [rA_max**2 - rA_med**2]],
                      ecolor='b', linewidth=lwidth)
@@ -65,12 +65,12 @@ if __name__ == '__main__':
     ax3.set_yticks([0, 0.5, 1])
     ax3.set_xticks([100, 1000])
 
-    ln1 = mlines.Line2D([], [], color='b', marker='o', markersize=msize,
-                        label=r'$\left<\mathcal{R}_A(t)^2\right>$')
     ln2 = mlines.Line2D([], [], color='k', marker='o', markersize=msize,
-                        label=r'$\left<\hat{F}_r\right>$')
+                        ls='', label=r'$\left<\hat{F}_r\right>$')
+    ln1 = mlines.Line2D([], [], color='b', marker='x', markersize=msize,
+                        ls='', label=r'$\left<\mathcal{R}_A(t)^2\right>$')
     ln3 = mlines.Line2D([], [], color='r', marker='*', markersize=msize,
-                        label=r'$\left<\hat{F}_s\right>$')
+                        ls='', label=r'$\left<\hat{F}_s\right>$')
     ax1.legend(handles=[ln1, ln2, ln3], fontsize=14, loc='upper left')
 
     ax3.set_xlabel('Re')

### 2020-04-08 writeups/Su_IGW_break.tex
try add revisions

@@
 
-We consider a stratified, incompressible, isothermal fluid representing a
-stellar envelope or atmosphere. We approximate the equation of state as
-\textcolor{Corr}{barotropic}, i.e.\ $P(\rho, T) = P(\rho)$. We study dynamics in
-2D, so that fluid variables depend only on the Cartesian coordinates $x$ and
-$z$. \textcolor{Corr}{While it is well known that waves break differently in two
-versus three dimensions} \citep{klostermeyer,winters1994}, the dynamical effect
-of the breaking process is likely to be similar in 2D \citep{barker_ogilvie}. We
-approximate the gravitational field as uniform, pointing in the $(-\uv{z})$
-direction. The plane-parallel approximation is justified since wave breaking
-generally occurs near the stellar surface. The background density stratification
-is given by
+We consider a incompressible, isothermally stratified fluid representing a
+stellar envelope or atmosphere. We study dynamics in 2D, so that fluid variables
+depend only on the Cartesian coordinates $x$ and $z$. \textcolor{Corr}{While it
+is well known that waves break differently in 2D versus 3D} \citep{klostermeyer,
+winters1994}, the dynamical effect of the breaking process is likely to be
+similar in 2D \citep{barker_ogilvie}. We approximate the gravitational field as
+uniform, pointing in the $(-\uv{z})$ direction. The plane-parallel approximation
+is justified since wave breaking generally occurs near the stellar surface. The
+background density stratification is given by
 \begin{equation}
@@
 
-The Euler equations for an incompressible, barotropic fluid in a uniform
-gravitational field are
+The Euler equations for an incompressible fluid in a uniform gravitational field
+are\footnote{\textcolor{Corr}{These equations arise by enforcing fluid
+incompressibility ($\mdil{\rho}{t} = 0$ or the adiabatic index $\gamma \to
+\infty$) in the Euler equations. They also can be derived from the commonly used
+``sound-proof'' \emph{anelastic equations} \citep[see e.g.][]{ogura1962scale,
+sutherland1} in the limit where (i) the density stratification is weak ($H \gg
+L$ for perturbation wavelength $L$) and (ii) the fluid is incompressible (note
+the anelastic equations are defined for arbitrary $\gamma$). Under condition
+(i), the anelastic approximation reduces to the Boussinesq approximation
+\citep{klein2009asymptotics}. Our equations permit a weak but nonzero density
+stratification and can be interpreted as ``locally Boussinesq''. Condition (ii)
+is motivated by WD fluid properties and, more generally, simplicity, as the
+dynamical effect of wave breaking is likely to be similar for different sets of
+fluid equations.}}
 \begin{subequations}\label{se:nl_orig}
@@
 field, density and pressure respectively. The constant gravitational
-acceleration is $-g\uv{z}$. For this isothermal background, hydrostatic
-equilibrium implies \textcolor{Corr}{$\overline{P}(z) = \overline{\rho}(z) g
-H$}. We assume there is initially no background flow, so $\bm{u} = \bm{u}'$.
-Physically, this assumption corresponds to a non-rotating star.
+acceleration is $(-g\uv{z})$.
+
+For this isothermal background, hydrostatic equilibrium implies
+\textcolor{Corr}{$\overline{P}(z) = \overline{\rho}(z) g H$}. We assume there is
+initially no background flow, so $\bm{u} = \bm{u}'$. Physically, this assumption
+corresponds to a non-rotating star.
 
@@
 In the small perturbation limit, we may linearize Eq.~\eqref{se:nl_upsilon}.
-\textcolor{Corr}{These equations admit the canonical IGW solution}, given by
+\textcolor{Corr}{The resulting equations admit the canonical IGW solution}
 \citep{drazin,sutherland0}
@@
         \textcolor{Corr}{The additional $\p{2H}^{-2}$ term in the denominator
-        accounts for the growing amplitude of the IGW in the $z$ direction (the
-        $z$ wavenumber is effectively $k_z - i / (2H)$).}
-        We note $\bm{c}_{p} \cdot \bm{c}_g = \mathcal{O}\s{\p{k_{z}H}^{-2}}
-        \approx 0$. In the Boussinesq approximation where terms of order
+        accounts for the growing amplitude of the IGW in the $z$ direction (as
+        the $z$ wavenumber is effectively $k_z - i / (2H)$).} We note
+        $\bm{c}_{p} \cdot \bm{c}_g = \mathcal{O}\s{\p{k_{z}H}^{-2}} \approx 0$.
+        In the Boussinesq approximation where terms of order
         $\mathcal{O}\p{H^{-2}}$ are ignored, the phase and group velocities are
@@
 solution), we assume $\overline{U}$ saturates at $\overline{U}_c$, which holds
-to good accuracy (see Fig.~\ref{fig:nl_fluxes}). Instead, the critical layer
-must propagate downward in response to the incident momentum flux.} The
+to good accuracy (see Fig.~\ref{fig:nl_fluxes}). In this case, the critical
+layer must propagate downward in response to the incident momentum flux.} The
 horizontal momentum of the shear flow satisfies
@@
 explicitly. A third-order, four-stage DIRK-ERK scheme \citep{ascher} is used
-with adaptive timesteps computed from advective Courant-Friedrichs-Lewy (CFL)
-time. Specifically, we use $\Delta t = 0.7 \min(\Delta x / u_x,\Delta z /
-u_{z})$, where the minimum is taken over every grid point in the domain, and
-$\Delta x \equiv L_x / N_x$ and $\Delta z \equiv L_z / N_z$ are the grid
-spacings in the $x$ and $z$ directions respectively.
+with adaptive timesteps computed from \textcolor{Corr}{the minimum of $0.1 / N$
+and the advective Courant-Friedrichs-Lewy (CFL) time. The CFL time is given by}
+$\Delta t = 0.7 \min(\Delta x / u_x,\Delta z / u_{z})$, where the minimum is
+taken over every grid point in the domain, and $\Delta x \equiv L_x / N_x$ and
+$\Delta z \equiv L_z / N_z$ are the grid spacings in the $x$ and $z$ directions
+respectively.
 
@@
 above $z_T$ with damping time $\tau$ and negligibly affects the dynamics between
-$z_B$ and $z_T$. \textcolor{Corr}{Most importantly, horizontal momentum is still
-conserved between $z_B$ and $z_T$}. We choose the transition width $\Delta z =
-0.25H$ and damping time $\tau = 1 / (15N)$. This prescription is similar to
-\citet{lecoanet_damp} and has the advantage of being smooth, important for
-spectral methods. Further details of our implementation of the fluid equations
-in Dedalus are described in Appendix~\ref{se:strat_impl}.
+$z_B$ and $z_T$. \textcolor{Corr}{Most importantly, horizontal momentum remains
+conserved between $z_B$ and $z_T$, and outgoing boundary conditions are imposed
+at $z_B, z_T$}. We choose the transition width $\Delta z = 0.25H$ and damping
+time $\tau = 1 / (15N)$. This prescription is similar to \citet{lecoanet_damp}
+and has the advantage of being smooth, important for spectral methods. Further
+details of our implementation of the fluid equations in Dedalus are described in
+Appendix~\ref{se:strat_impl}.
 
@@
 Eq.~\eqref{eq:uz_lin} and the corresponding $u_x(x, z, t)$; we denote this
-solution $\bm{u}_{al}(x, z, t)$. The amplitude of the observed IGW in the
-simulation field $\bm{u}$ relative to analytical solution $\bm{u}_{al}$ over
+analytical solution $\bm{u}_{an}(x, z, t)$. The amplitude of the observed IGW in the
+simulation field $\bm{u}$ relative to $\bm{u}_{an}$ over
 some region $z \in [z_b, z_t]$ can be estimated from
@@
     A_i(t) = \frac{\int\limits_{z_b}^{z_t}\int\limits_0^{L_x}
-        \overline{\rho}\p{\bm{u} \cdot \bm{u}_{al}}\;\mathrm{d}x\mathrm{d}z}
+        \overline{\rho}\p{\bm{u} \cdot \bm{u}_{an}}\;\mathrm{d}x\mathrm{d}z}
         {\int\limits_{z_b}^{z_t}\int\limits_0^{L_x}
-        \overline{\rho}\abs{\bm{u}_{al}}^2\;\mathrm{d}x\mathrm{d}z}.
+        \overline{\rho}\abs{\bm{u}_{an}}^2\;\mathrm{d}x\mathrm{d}z}.
         \label{eq:ahat_def}
 \end{equation}
-The subscript $i$ denotes the incident wave. If $\bm{u} = \bm{u}_{al}$, then
+The subscript $i$ denotes the incident wave. If $\bm{u} = \bm{u}_{an}$, then
 $A_i(t) = 1$. \textcolor{Corr}{The factor of $\bar{\rho}$ inside the integrands
 in Eq.~\eqref{eq:ahat_def} corrects for the $\propto e^{z/2H}$ growth of
-$\bm{u}_{al}$; without it, $A_i(t)$ would be dominated by the contribution near
+$\bm{u}_{an}$; without it, $A_i(t)$ would be dominated by the contribution near
 $z_t$.}
@@
 theory can be computed by simply evaluating Eq.~\eqref{eq:F_def} for
-$\bm{u}_{al}$ and is a constant:
-\begin{equation}
-    F_{al} \equiv \ev{\rho u_{al, x} u_{al, z}}_x.\label{eq:F_al}
+$\bm{u}_{an}$ and is a constant:
+\begin{equation}
+    F_{an} \equiv \ev{\rho u_{an, x} u_{an, z}}_x.\label{eq:F_al}
 \end{equation}
@@
 (use Eq.~\eqref{eq:F_def} with velocities taken from the simulation), we expect
-$F(z, t) = F_{al}$ between $z_0$ and $z_T$. Fig.~\ref{fig:lin_fluxes} shows
+$F(z, t) = F_{an}$ between $z_0$ and $z_T$. Fig.~\ref{fig:lin_fluxes} shows
 agreement with this prediction.
@@
     \includegraphics[width=0.9\columnwidth]{plots/lin_fluxes.png}
-    \caption{$F/F_{al}$ plotted at select times $t$ (in units of
-    $N^{-1}$). As the initial transient dies out, $F / F_{al} \approx 1$ to a
+    \caption{$F/F_{an}$ plotted at select times $t$ (in units of
+    $N^{-1}$). As the initial transient dies out, $F / F_{an} \approx 1$ to a
     good approximation above the forcing zone $z > z_0 = 2H$ and below the
@@
         previous snapshot. This suggests that the latter phase of the simulation
-        has reached a steady state.
-
-        \textcolor{Corr}{Notably, the horizontal banded structure of $u_x$ in
-        the synchronized fluid is mostly unchanged (also visible in the top
-        panel of Fig.~\ref{fig:nl_fluxes}), suggesting that momentum
-        redistribution and mixing within the synchronized fluid are negligible.}
+        has reached a steady state. \textcolor{Corr}{Notably, the horizontal
+        banded structure of $u_x$ in the upper, synchronized fluid does not
+        continue to evolve (also visible in the top panel of
+        Fig.~\ref{fig:nl_fluxes}), suggesting that momentum redistribution and
+        mixing within the synchronized fluid are negligible.}
 \end{enumerate}
@@
 $\overline{U}$ (Eq.~\eqref{eq:mean_flow}) and the dimensionless momentum flux $F
-/ F_{al}$ (Eqs.~\eqref{eq:F_def} and~\eqref{eq:F_al}) as a function of $z$ at
+/ F_{an}$ (Eqs.~\eqref{eq:F_def} and~\eqref{eq:F_al}) as a function of $z$ at
 the times depicted in Fig.~\ref{fig:snapshots}. At each time, $\overline{U}$ is
@@
 
-Similarly, $F \lesssim F_{al}$ below the critical layer, and then decreases to
+Similarly, $F \lesssim F_{an}$ below the critical layer, and then decreases to
 about zero above the critical layer. However, two notable deviations from the
@@
     (Eq.~\eqref{eq:mean_flow}) and the dimensionless momentum flux $F(z, t) /
-    F_{al}$ (Eqs.~\eqref{eq:F_def} and~\eqref{eq:F_al}) in our fiducial
+    F_{an}$ (Eqs.~\eqref{eq:F_def} and~\eqref{eq:F_al}) in our fiducial
     simulation plotted at the same times as in Fig.~\ref{fig:snapshots}. The two
@@
     the critical layer over time (in units of $N^{-1}$) in our fiducial
-    simulation. The red and solid green lines denote respectively the minimum
-    and median of $\mathrm{Ri}_x(x, t)$. These numbers measure the mean and
-    spread in width of the critical layer over $x$. Note that $\mathrm{Ri} \sim
-    \frac{1}{4}$ corresponds to the KHI, so this plot suggests the shear at the
-    critical layer does not steepen past the onset of the
+    simulation. The solid red and dotted black lines denote respectively the
+    minimum and median of $\mathrm{Ri}_x(x, t)$. These numbers measure the mean
+    and spread in width of the critical layer over $x$. Note that $\mathrm{Ri}
+    \sim \frac{1}{4}$ corresponds to the KHI, so this plot suggests the shear at
+    the critical layer does not steepen past the onset of the
     KHI.}\label{fig:nl_f_ri}
@@
 \begin{align}
-    F_i(t) &= F_{al}A_i^2(t),\\
+    F_i(t) &= F_{an}A_i^2(t),\\
     F_r(t) &= F_i(t) - \frac{1}{H}
@@
     (see Eq.~\eqref{eq:f_budget}), in units of the analytical estimate for the
-    incident wave flux $F_{al}$ (Eq.~\eqref{eq:S_lin}): $F_i$, the flux incident
-    on the critical layer (green); $F_a$, the flux absorbed by the critical
-    layer (blue); $F_r$, the flux reflected at the critical layer (red); and
-    $F_s$, the flux inside the synchronized layer
-    (black).}\label{fig:nl_f_amps2}
+    incident wave flux $F_{an}$ (Eq.~\eqref{eq:S_lin}): $F_i$, the flux incident
+    on the critical layer; $F_a$, the flux absorbed by the critical layer;
+    $F_r$, the flux reflected at the critical layer; and $F_s$, the flux inside
+    the synchronized layer.}\label{fig:nl_f_amps2}
 \end{figure}
@@
 \begin{align}
-    z_{c, \min}(t) &\equiv \min_z \z{z: F(z, t) > 0.3F_{al}},\\
-    z_{c, \max}(t) &\equiv \max_z \z{z: F(z, t) < 0.3F_{al}},\\
+    z_{c, \min}(t) &\equiv \min_z \z{z: F(z, t) > 0.3F_{an}},\\
+    z_{c, \max}(t) &\equiv \max_z \z{z: F(z, t) < 0.3F_{an}},\\
     z_c(t) &\equiv \frac{z_{c, \min}(t) + z_{c, \max}(t)}{2}.\label{eq:zc_def}
@@
 By time-averaging the numerically measured $F_a$, we find $\ev{F_a}_t \approx
-0.71F_{al}$. Note that $F_a < F_{al}$, so momentum flux absorption at the
+0.71F_{an}$. Note that $F_a < F_{an}$, so momentum flux absorption at the
 critical layer is incomplete. This is due to reflection of waves off the
@@
     \includegraphics[width=0.9\columnwidth]{plots/nl_front.png}
-    \caption{Propagation of the critical layer over time. Shown are: (black)
-    $z_c(t)$ from simulation data, and (red) model for $z_c(t)$ using direct
+    \caption{Propagation of the critical layer over time. Shown are: (solid black)
+    $z_c(t)$ from simulation data, and (dashed red) model for $z_c(t)$ using direct
     integration of Eq.~\eqref{eq:zc_anal} for $F_a(t)$ measured from simulation
@@
     A_r(t) = \max_{\delta x}\frac{\int\limits_{z_b}^{z_t}\int\limits_0^{L_x}
-        \overline{\rho}\p{\bm{u} \cdot \at{\bm{u}_{al,
+        \overline{\rho}\p{\bm{u} \cdot \at{\bm{u}_{an,
         \bm{k}_r}}_{x = x + \delta x}}\;\mathrm{d}x\mathrm{d}z}
         {\int\limits_{z_b}^{z_t}\int\limits_0^{L_x}
-        \overline{\rho}\abs{\bm{u}_{al}}^2\;\mathrm{d}x\mathrm{d}z},
+        \overline{\rho}\abs{\bm{u}_{an}}^2\;\mathrm{d}x\mathrm{d}z},
         \label{eq:ar_def}
@@
     \includegraphics[width=0.9\columnwidth]{plots/nl_f_amps.png}
-    \caption{The incident wave amplitude $A_i(t)$ (green) and the reflected wave
-    amplitude $A_r(t)$ (red) just above the forcing zone.}\label{fig:nl_f_amps}
+    \caption{The incident wave amplitude $A_i(t)$ (solid black) and the
+    reflected wave amplitude $A_r(t)$ (dashed red) just above the forcing
+    zone.}\label{fig:nl_f_amps}
 \end{figure}
@@
     \caption{$\mathcal{R}_A^2$, $\hat{F}_r$, and redistribution flux $\hat{F}_s$
-    (Eqs.~\eqref{eq:Ra_def}--\eqref{eq:srefl_def}) as a function of time (in
+    [Eqs.~(\ref{eq:Ra_def}--\ref{eq:srefl_def})] as a function of time (in
     units of $N^{-1}$). These quantities seem to become comparatively stable
@@
 $\mathcal{R}_A^2$, $\hat{F}_r$, $\hat{F}_s$, and $\mathrm{Ri}$
-(Eqs.~\eqref{eq:Ra_def}--\eqref{eq:srefl_def} and~\eqref{eq:ri_def}
-respectively) over the last $1/4$ of the simulation time, when these quantities
+[Eqs.~(\ref{eq:Ra_def}--\ref{eq:srefl_def}) and~\eqref{eq:ri_def}
+respectively] over the last $1/4$ of the simulation time, when these quantities
 have reached their asymptotic values. These results are shown in
@@
 Note that in the main text, this approximate form is used to compute
-$\bm{u}_{al}$, as it is easier to work with and sufficiently accurate in the
+$\bm{u}_{an}$, as it is easier to work with and sufficiently accurate in the
 regions of interest (many $\sigma$ away from $z_0$).

### 2020-04-08 sims/2d_4_fourier/strat_helper.py

@@ -1072,7 +1072,7 @@ def plot_front(name, params, start_time=None):
         plt.legend(fontsize=FONTSIZE)
 
         plt.xlabel(r'$z / H$', fontsize=int(1.5 * FONTSIZE))
-        plt.ylabel(r'$F / F_{al}$', fontsize=int(1.5 * FONTSIZE))
+        plt.ylabel(r'$F / F_{an}$', fontsize=int(1.5 * FONTSIZE))
         axes = plt.gca()
         axes.xaxis.set_ticks_position('both')
         axes.yaxis.set_ticks_position('both')
@@ -1173,7 +1173,7 @@ def plot_front(name, params, start_time=None):
 
         ax1.set_ylabel(r'$\overline{U} / \overline{U}_c$',
                        fontsize=int(1.5 * FONTSIZE))
-        ax2.set_ylabel(r'$F / F_{al}$', fontsize=int(1.5 * FONTSIZE))
+        ax2.set_ylabel(r'$F / F_{an}$', fontsize=int(1.5 * FONTSIZE))
         ax2.set_xlabel(r'$z / H$', fontsize=int(1.5 * FONTSIZE))
         for ax in [ax1, ax2]:
             ax.xaxis.set_ticks_position('both')
@@ -1235,7 +1235,7 @@ def plot_front(name, params, start_time=None):
                  linewidth=LW * 0.5)
         ax2.plot(t,
                  front_pos_intg_S,
-                 PLT_STYLES[1],
+                 'r--',
                  alpha=0.7,
                  label='Model (Eq. 23)',
                  linewidth=LW * 0.8)
@@ -1251,7 +1251,7 @@ def plot_front(name, params, start_time=None):
         # ax2.plot(t,
         #          pos_anal,
         #          'g',
-        #          label="Model (Eq. 24, $F_a = %.2fF_{al}$)" %
+        #          label="Model (Eq. 24, $F_a = %.2fF_{an}$)" %
         #             (mean_incident / flux_th),
         #          linewidth=LW * 0.7)
         ax2.set_ylabel(r'$z_c$', fontsize=int(1.5 * FONTSIZE))
@@ -1279,7 +1279,7 @@ def plot_front(name, params, start_time=None):
                  linewidth=LW * 0.7)
         ax1.plot(t,
                  smooth(amps_down[start_idx::]),
-                 PLT_STYLES[1],
[…]

### 2020-04-13 writeups/Su_IGW_break.tex
added daniel appendix

@@
 The Euler equations for an incompressible fluid in a uniform gravitational field
-are\footnote{\textcolor{Corr}{These equations arise by enforcing fluid
-incompressibility ($\mdil{\rho}{t} = 0$ or the adiabatic index $\gamma \to
-\infty$) in the Euler equations. They also can be derived from the commonly used
-``sound-proof'' \emph{anelastic equations} \citep[see e.g.][]{ogura1962scale,
-sutherland1} in the limit where (i) the density stratification is weak ($H \gg
-L$ for perturbation wavelength $L$) and (ii) the fluid is incompressible (note
-the anelastic equations are defined for arbitrary $\gamma$). Under condition
-(i), the anelastic approximation reduces to the Boussinesq approximation
-\citep{klein2009asymptotics}. Our equations permit a weak but nonzero density
-stratification and can be interpreted as ``locally Boussinesq''. Condition (ii)
-is motivated by WD fluid properties and, more generally, simplicity, as the
-dynamical effect of wave breaking is likely to be similar for different sets of
-fluid equations.}}
+are
+% \footnote{\textcolor{Corr}{These equations arise by enforcing fluid
+% incompressibility ($\mdil{\rho}{t} = 0$ or the adiabatic index $\gamma \to
+% \infty$) in the Euler equations. They also can be derived from the commonly used
+% ``sound-proof'' \emph{anelastic equations} \citep[see e.g.][]{ogura1962scale,
+% sutherland1} in the limit where (i) the density stratification is weak ($H \gg
+% L$ for perturbation wavelength $L$) and (ii) the fluid is incompressible (note
+% the anelastic equations are defined for arbitrary $\gamma$). Under condition
+% (i), the anelastic approximation reduces to the Boussinesq approximation
+% \citep{klein2009asymptotics}. Our equations permit a weak but nonzero density
+% stratification and can be interpreted as ``locally Boussinesq''. Condition (ii)
+% filters out sound waves. The dynamical effect of wave breaking is likely to be
+% similar for different sets of fluid equations.}}
 \begin{subequations}\label{se:nl_orig}
@@
 field, density and pressure respectively. The constant gravitational
-acceleration is $(-g\uv{z})$.
+acceleration is $(-g\uv{z})$. \textcolor{Corr}{Note that these equations
+conserve the same wave energy as the commonly used anelastic equations
+\citep{ogura1962scale,anel_part1} and thus give the same wave amplitude
+growth. Appendix~\ref{s:equation_deriv} provides a derivation of these
+equations and justification for using them.}
 
@@
 \appendix
+
+\section{Derivation of Fluid Equations}\label{s:equation_deriv}
+
+We aim to model wave dynamics over multiple density and pressure scaleheights.
+Start with the compressible Euler equations.
+\begin{align}
+    \partial_t \rho + \p{\bm{u}\cdot\bm{\nabla}} \rho
+            + \rho \bm{\nabla}\cdot\bm{u} =\, &0, \\
+    \partial_t s + \p{\bm{u}\cdot\bm{\nabla}} s =\, &0, \\
+    \partial_t \bm{u} + \p{\bm{u}\cdot\bm{\nabla}}\bm{u} +
+        \frac{1}{\rho}\bm{\nabla} p =& \, \bm{g},
+\end{align}
+where $s$ is the entropy. The pressure is calculated from the entropy and
+density using the equation of state
+\begin{equation}
+    \frac{s}{c_p} = \log \frac{p^{1/\gamma}}{\rho},
+\end{equation}
+where $c_p$ is the specific heat at constant pressure, and $\gamma$ is the ratio
+of specific heats. For computational ease, it is convenient to filter the fast
+sound waves from these equations. One approach is to assume pressure
+perturbations are small \citep[yielding the ``pseudo-incompressible''
+equations,][]{anel_part2}, or that all thermodynamic perturbations are small
+\citep[yielding the ``anelastic'' equations,][]{anel_part1}. In these
+approximations, one of the thermodynamic equations is replaced by a constraint
+equation: $\bm{\nabla}\cdot \p{p_0^{1/\gamma} \bm{u}}=0$ for
+pseudo-incompressible; $\bm{\nabla}\cdot\p{\rho_0 \bm{u}} =0$ for anelastic.
+Here $p_0$ and $\rho_0$ are the background density and pressure profiles. The
+pressure in the momentum equation can be interpreted as a Lagrange multiplier
+which enforces the constraint \citep{anel_part2}. Upon linearization, both
+approximations conserve a wave energy
+\begin{equation}
+    E_w = \frac{1}{2}\rho_0 \abs{\bm{u}}^2
+            + \frac{1}{2}\frac{g^2}{N^2} \frac{(\rho')^2}{\rho_0}
+        = \frac{1}{2}\rho_0 \abs{\bm{u}}^2
+            + \frac{1}{2}\rho_0 N^2 \frac{(s')^2}{\abs{\bm{\nabla} s_0}^2},
+            \label{eq:ewave}
+\end{equation}
+where $\rho'$ and $s'$ represent the density and entropy perturbations.
+
+Rather than assume thermodynamic perturbations are small, we instead filter
+sound waves by taking the limit $\gamma\rightarrow \infty$. Then the entropy and
+log density are proportional to each other, and the entropy equation becomes
+\begin{equation}
+    \partial_t \log \rho + \p{\bm{u}\cdot \bm{\nabla}} \log \rho = 0.
+\end{equation}
+Together with mass conservation, this implies
+\begin{equation}
+    \bm{\nabla} \cdot \bm{u} = 0.
+\end{equation}
+We solve these equations together with the normal momentum equation.
+
+Although these equations are non-standard, they have various desirable
+properties. They conserve mass and momentum, and the linearized equations
+conserve the wave energy $E_w$ [Eq.~\eqref{eq:ewave} above] similar to the
+pseudo-incompressible equations and anelastic equations. Our equations also
+satisfy the same linear dispersion relation as the fully compressible equations
+in the limit of large sound speed \citep[this is also true for the anelastic
+equations, but not pseudo-incompressible,][]{anel_part1, anel_part2}. Thus, the
+vertical propagation of internal gravity waves is similar to the
+pseudo-incompressible equations and anelastic equations.
+
+In this work we are interested in waves which reach large amplitudes and break.
+For breaking waves, \citet{achatz_igw_pi} suggests that the anelastic equations
+may miss important effects. Although the pseudo-incompressible equations may
+capture wave-breaking more accurately, they are more complicated, and do not
+satisfy the correct dispersion relation to order $(k_z H)^{-2}$
+\citep{anel_part2}. In the absence of a clear choice to study this wave breaking
+problem, we have elected to use these simple equations derived in the
+$\gamma\rightarrow\infty$ limit.
 

### 2020-04-13 writeups/Su_IGW_break.tex
filter out

@@
 where $c_p$ is the specific heat at constant pressure, and $\gamma$ is the ratio
-of specific heats. For computational ease, it is convenient to filter the fast
-sound waves from these equations. One approach is to assume pressure
+of specific heats. For computational ease, it is convenient to filter out the
+fast sound waves from these equations. One approach is to assume pressure
 perturbations are small \citep[yielding the ``pseudo-incompressible''
@@
 
-Rather than assume thermodynamic perturbations are small, we instead filter
+Rather than assume thermodynamic perturbations are small, we instead filter out
 sound waves by taking the limit $\gamma\rightarrow \infty$. Then the entropy and

### 2020-05-06 writeups/Su_IGW_break.tex
check in accepted version

@@
 
-\date{Accepted XXX\@. Received YYY\@; in original form ZZZ}
+\date{Accepted 2020 May 06. Received 2020 April 13; in original form 2020
+February 26}
 
@@
 \begin{abstract}
-    In binaries composed of \textcolor{Corr}{either early-type stars or white
-    dwarfs}, the dominant tidal process involves the excitation of internal
+    In binaries composed of either early-type stars or white
+    dwarfs, the dominant tidal process involves the excitation of internal
     gravity waves (IGWs), which propagate towards the stellar surface, and their
@@
 stellar envelope or atmosphere. We study dynamics in 2D, so that fluid variables
-depend only on the Cartesian coordinates $x$ and $z$. \textcolor{Corr}{While it
-is well known that waves break differently in 2D versus 3D} \citep{klostermeyer,
+depend only on the Cartesian coordinates $x$ and $z$. While it
+is well known that waves break differently in 2D versus 3D \citep{klostermeyer,
 winters1994}, the dynamical effect of the breaking process is likely to be
@@
 are
-% \footnote{\textcolor{Corr}{These equations arise by enforcing fluid
+% \footnote{These equations arise by enforcing fluid
 % incompressibility ($\mdil{\rho}{t} = 0$ or the adiabatic index $\gamma \to
@@
 % filters out sound waves. The dynamical effect of wave breaking is likely to be
-% similar for different sets of fluid equations.}}
+% similar for different sets of fluid equations.}
 \begin{subequations}\label{se:nl_orig}
@@
 field, density and pressure respectively. The constant gravitational
-acceleration is $(-g\uv{z})$. \textcolor{Corr}{Note that these equations
+acceleration is $(-g\uv{z})$. Note that these equations
 conserve the same wave energy as the commonly used anelastic equations
@@
 growth. Appendix~\ref{s:equation_deriv} provides a derivation of these
-equations and justification for using them.}
+equations and justification for using them.
 
 For this isothermal background, hydrostatic equilibrium implies
-\textcolor{Corr}{$\overline{P}(z) = \overline{\rho}(z) g H$}. We assume there is
+$\overline{P}(z) = \overline{\rho}(z) g H$. We assume there is
 initially no background flow, so $\bm{u} = \bm{u}'$. Physically, this assumption
@@
 In the small perturbation limit, we may linearize Eq.~\eqref{se:nl_upsilon}.
-\textcolor{Corr}{The resulting equations admit the canonical IGW solution}
+The resulting equations admit the canonical IGW solution
 \citep{drazin,sutherland0}
@@
         \end{align}
-        \textcolor{Corr}{The additional $\p{2H}^{-2}$ term in the denominator
+        The additional $\p{2H}^{-2}$ term in the denominator
         accounts for the growing amplitude of the IGW in the $z$ direction (as
-        the $z$ wavenumber is effectively $k_z - i / (2H)$).} We note
+        the $z$ wavenumber is effectively $k_z - i / (2H)$). We note
         $\bm{c}_{p} \cdot \bm{c}_g = \mathcal{O}\s{\p{k_{z}H}^{-2}} \approx 0$.
@@
 Consider now the long-term evolution of the critical layer due to continuous
-horizontal momentum transfer by IGWs. \textcolor{Corr}{Any incident horizontal
+horizontal momentum transfer by IGWs. Any incident horizontal
 momentum flux absorbed by the fluid, denoted $F_a(t)$, must manifest as
-additional horizontal momentum of the shear flow}.
-\textcolor{Corr}{Additionally, as the mean flow $\overline{U}$ cannot grow
+additional horizontal momentum of the shear flow.
+Additionally, as the mean flow $\overline{U}$ cannot grow
 efficiently above $\overline{U}_c$ (due to the breakdown of the linear
@@
 to good accuracy (see Fig.~\ref{fig:nl_fluxes}). In this case, the critical
-layer must propagate downward in response to the incident momentum flux.} The
+layer must propagate downward in response to the incident momentum flux. The
 horizontal momentum of the shear flow satisfies
@@
 explicitly. A third-order, four-stage DIRK-ERK scheme \citep{ascher} is used
-with adaptive timesteps computed from \textcolor{Corr}{the minimum of $0.1 / N$
-and the advective Courant-Friedrichs-Lewy (CFL) time. The CFL time is given by}
+with adaptive timesteps computed from the minimum of $0.1 / N$
+and the advective Courant-Friedrichs-Lewy (CFL) time. The CFL time is given by
 $\Delta t = 0.7 \min(\Delta x / u_x,\Delta z / u_{z})$, where the minimum is
@@
 above $z_T$ with damping time $\tau$ and negligibly affects the dynamics between
-$z_B$ and $z_T$. \textcolor{Corr}{Most importantly, horizontal momentum remains
+$z_B$ and $z_T$. Most importantly, horizontal momentum remains
 conserved between $z_B$ and $z_T$, and outgoing boundary conditions are imposed
-at $z_B, z_T$}. We choose the transition width $\Delta z = 0.25H$ and damping
+at $z_B, z_T$. We choose the transition width $\Delta z = 0.25H$ and damping
 time $\tau = 1 / (15N)$. This prescription is similar to \citet{lecoanet_damp}
@@
 The subscript $i$ denotes the incident wave. If $\bm{u} = \bm{u}_{an}$, then
-$A_i(t) = 1$. \textcolor{Corr}{The factor of $\bar{\rho}$ inside the integrands
+$A_i(t) = 1$. The factor of $\bar{\rho}$ inside the integrands
 in Eq.~\eqref{eq:ahat_def} corrects for the $\propto e^{z/2H}$ growth of
 $\bm{u}_{an}$; without it, $A_i(t)$ would be dominated by the contribution near
-$z_t$.}
+$z_t$.
 
@@
     amplitude. The small deviation of $A_i(t)$ from unity may be due to
-    \textcolor{Corr}{truncation error in our implicit timestepping scheme}, as a
+    truncation error in our implicit timestepping scheme, as a
     relatively large fixed step size $\Delta t = 0.1/N$ was used for this
@@
 predicts $\abs{\xi_z k_z} \sim 4.25$ at the upper damping zone $z_T$.
-\textcolor{Corr}{We choose the viscosity $\nu$ for each resolution to be as
+We choose the viscosity $\nu$ for each resolution to be as
 small as possible while still resolving the shortest spatial scales of the wave
-breaking.} A table of our simulations can be found in Table~\ref{tab:params}.
+breaking. A table of our simulations can be found in Table~\ref{tab:params}.
 \begin{table}
@@
 In Fig.~\ref{fig:snapshots}, we present snapshots of $u_x$ and $\Upsilon$ at
-various phases of the simulation. \textcolor{Corr}{Note that $\Upsilon \lesssim
+various phases of the simulation. Note that $\Upsilon \lesssim
 0.1$, so the density stratification does not deviate significantly from
-equilibrium.} The flow evolves through several distinct stages:
+equilibrium. The flow evolves through several distinct stages:
 \begin{enumerate}
@@
     \item At the end of the simulation (bottom right panel), the
-        \textcolor{Corr}{critical layer has advanced downwards, but otherwise}
+        critical layer has advanced downwards, but otherwise
         the flow shows very few significant qualitative differences from the
         previous snapshot. This suggests that the latter phase of the simulation
-        has reached a steady state. \textcolor{Corr}{Notably, the horizontal
+        has reached a steady state. Notably, the horizontal
         banded structure of $u_x$ in the upper, synchronized fluid does not
@@
         Fig.~\ref{fig:nl_fluxes}), suggesting that momentum redistribution and
-        mixing within the synchronized fluid are negligible.}
+        mixing within the synchronized fluid are negligible.
 \end{enumerate}
@@
     in the fiducial simulation illustrating distinct phases of the evolution of
-    the flow. \textcolor{Corr}{See the online PDF for a color version.} Damping
-    layers \textcolor{Corr}{at the top and bottom of the simulation domain} are
+    the flow. See the online PDF for a color version. Damping
+    layers at the top and bottom of the simulation domain are
     shaded in light grey (see Section~\ref{ss:damping}), while the forcing zone
-    \textcolor{Corr}{in the lower middle portion of the simulation domain} (see
+    in the lower middle portion of the simulation domain (see
     Section~\ref{ss:forcing}) is shaded in light green (boundaries are at $z_0
@@
 is the width of this layer? Inspection of Fig.~\ref{fig:snapshots} suggests the
-\textcolor{Corr}{presence} of the Kelvin-Helmholtz Instability (KHI) in the
+presence of the Kelvin-Helmholtz Instability (KHI) in the
 critical layer. In a stratified medium, KHI occurs when the Richardson number
@@
 \mathrm{Ri}_x(x, t)$ (the maximum is very noisy). Both are shown in
-Fig.~\ref{fig:nl_f_ri}. \textcolor{Corr}{Absorption} of incident IGWs quickly
+Fig.~\ref{fig:nl_f_ri}. Absorption of incident IGWs quickly
 decreases the Richardson number to between $0.25$ and $0.5$, characteristic of

### 2020-05-23 writeups/Su_IGW_break.tex
use MNRAS video link

@@
 1024$ is available
-online\footnote{\url{http://www.princeton.edu/\~lecoanet/data/breaking\_wave.mov}}. We
-take this to be our fiducial simulation for the remainder of this paper, though
-other simulations show qualitatively similar behavior.
+online\footnote{\url{https://academic.oup.com/mnras/article-abstract/495/1/1239/5835700#supplementary-data}}.
+We take this to be our fiducial simulation for the remainder of this paper,
+though other simulations show qualitatively similar behavior.
 

### 2020-05-23 writeups/Su_IGW_break.tex
fix typo

--- 2020-05-06 writeups/Su_IGW_break.tex
@@
 1024$ is available
-online\footnote{\url{http://www.princeton.edu/\~lecoanet/data/breaking\_wave.mov}}. We
-take this to be our fiducial simulation for the remainder of this paper, though
-other simulations show qualitatively similar behavior.
+online\footnote{\url{https://academic.oup.com/mnras/article-abstract/495/1/1239/5835700\#supplementary-data}}.
+We take this to be our fiducial simulation for the remainder of this paper,
+though other simulations show qualitatively similar behavior.
 
